### 2018-01-29 paper.tex
initial commit

%%%%%%%% ICML 2018 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2018}

\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{listings}
\usepackage{amsthm}
% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfig} 
\usepackage{fancyvrb}

\usepackage{caption}
\usepackage{subcaption}

\fvset{fontsize=\footnotesize}

\usepackage{amssymb}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{tabularx}

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text,decorations.pathreplacing}
%\usetikzlibrary{decorations.pathmorphing}
\usepackage{dsfont}
\usepackage{amsmath}
\usepackage{hyperref}
\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
\DeclareMathOperator{\argmin}{argmin} % no space, limits underneath in displays

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.

\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em E}}
\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2018}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Inducing Domain Specific Languages for Bayesian Program Learning}

\begin{document}

\twocolumn[
\icmltitle{Inducing Domain Specific Languages for Bayesian Program Learning}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2018
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Aeiau Zzzz}{equal,to}
\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
\icmlauthor{Cieua Vvvvv}{goo}
\icmlauthor{Iaesut Saoeu}{ed}
\icmlauthor{Fiuea Rrrr}{to}
\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
\icmlauthor{Aaoeu Iasoh}{goo}
\icmlauthor{Buiui Eueu}{ed}
\icmlauthor{Aeuia Zzzz}{ed}
\icmlauthor{Bieea C.~Yyyy}{to,goo}
\icmlauthor{Teoau Xxxx}{ed}
\icmlauthor{Eee Pppp}{ed}
\end{icmlauthorlist}

\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}

\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
This document provides a basic paper template and submission guidelines.
Abstracts must be a single paragraph, ideally between 4--6 sentences long.
Gross violations will trigger corrections at the camera-ready phase.
\end{abstract}

\section{Introduction}

Imagine an agent faced with a suite of new problems totally different
from anything it has seen before. It has at its disposal a basic set
of primitives it can compose to build solutions to these problems, but
it is no idea what kinds of primitives are appropriate for which
problems nor does it know the higher-level domain-specific language in
which solutions are best expressed.

Our algorithm accomplishes the following:
\begin{itemize}
\item Learns from relatively small amounts of data and with weak supervision. We do not use ground truth programs -- weak supervision. Our DSL learner does not need huge amounts of data -- tens of examples suffice.
  \item Jointly infers a \emph{generative model} along with a \emph{recognition model}. The \emph{generative model} is a probabilistic context-sensitive grammar over programs, and includes a DSL. The \emph{recognition model} is a neural network that guides the agents use of the DSL.
\end{itemize}
The generative model and the recognition model bootstrap off of each other:
\begin{itemize}
\item The recognition model works by upweighting the probability of program components likely to be useful for a given problem. By learning a DSL, the recognition model gets more power because it can upweight new more powerful primitives that are better suited for the domain.
\item The generative model (DSL) is learned from programs that the agent has found so far. Because the recognition model speeds up search, it generates more training data for the generative model.
\end{itemize}

\section{Experiments}

\begin{figure}
  \includegraphics[width = 9cm]{figures/polynomial.png}
  \includegraphics[width = 9cm]{figures/text.png}
  \includegraphics[width = \columnwidth]{figures/circuit.png} 
\end{figure}

\section{Model}

\begin{figure}
  \begin{tikzpicture}

  \node[latent] at (1,3) (d){$\mathcal{D}$};
  \node[latent] at (2.5,3) (t){$\theta$};
  \node[latent] at (1,1.5) (z){$z$};
  \node[latent] at (2.5,1.5) (tx){$\theta_x$};
  \node[obs] at (1,0) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(tx)(z)(x)}{$N$};
  \draw [->,dashed] (x.east) to[out = 0,in = -90] node[fill = white]{$q(x)$} (tx.south);
  \draw [->,dashed] (tx.west) -- (z.east);
  \end{tikzpicture}
  \caption{DSL $\mathcal{D}$ generates programs $z$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). We observe program outputs $x$. A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta_x = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}
  \end{figure}

\section{Implementation}

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,[\text{var}:\alpha]+\mathcal{E},\beta)$
   \STATE \textbf{return} $\lambda \text{var}.$ body
   \ENDIF
   %   \ELSE
   \STATE\begin{align*}     \text{primitives} \gets\{p | &p:\alpha\to\cdots\to\beta \in \mathcal{D}\cup\mathcal{E}\\     &\text{if canUnify}(\tau,\beta) \}     \end{align*}

   \STATE Sample $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ and w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Let $e:\alpha_1\to\alpha_2\to\cdots\to \alpha_K\to\beta$. Unify $\tau$ with $\beta$.
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} $e(a_1, a_2, \cdots, a_K)$
 %  \UNTIL{$noChange$ is $true$}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[tb]
   \caption{DSL Learner}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
     \ENDFOR
     \STATE $\mathcal{D}_i,\theta_i\gets $induceGrammar$(\{\mathcal{F}_x\}_{x\in X})$
     \STATE Define $Q_x(z) \propto \begin{cases}
       \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
       0&x\not \in \mathcal{F}_x
     \end{cases}$
     \STATE $q_i\gets \argmin_q \sum_{x\in X}\text{KL}(Q_x(\cdot )||\probability[\cdot |\mathcal{D}_i,q(x)])$
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[tb]
   \caption{Grammar Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$, AIC coefficient $a$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $\log \probability[\mathcal{D}]\stackrel{+}{ = } -\lambda\sum_{p\in \mathcal{D}} \text{size}(p)$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - a|\mathcal{D}|$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ a subtree of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \IF{$\text{score}(\mathcal{D}') > \text{score}(\mathcal{D})$}
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ELSE
     \STATE\textbf{return} $\mathcal{D},\theta^*(\mathcal{D})$
     \ENDIF
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

\section{Estimating the grammar parameters}

I justify this estimator by proving that it maximizes a lower bound on the log likelihood of the data. Writing $L$ for the log likelihood, $\theta$ for the parameters of the grammar, $N$ for the number of random choices, $A$ to range over the alternative choices for a random variable, $c(x)$ to mean the number of times that primitive $x$ was used, and $a(x) = \sum_A \indicator [x\in A]$ to mean the number of times that primitive $x$ could have been used:
\begin{align}
  L& = \sum_x c(x)\log \theta_x - \sum_A \log \sum_{x\in A}\theta_x\\
  & = \sum_x c(x)\log \theta_x - N\expect_A \log \sum_{x\in A}\theta_x\\
  &\geq\sum_x c(x)\log \theta_x - N \log \expect_A \sum_{x\in A}\theta_x\text{, Jensen's inequality}\\
  & = \sum_x c(x)\log \theta_x - N \log \frac{1}{N}\sum_A \sum_{x} \indicator [x\in A]\theta_x\\
  & \stackrel{+}{=} \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x.
\end{align}
Differentiate with respect to $\theta_x$ and set to zero:
\begin{align}
  \frac{c(x)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
\end{align}
This equality holds if $\theta_x = c(x)/a(x)$:
\begin{align}
  \frac{c(x)}{\theta_x} &= a(x).\\
N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
\end{align}
If this equality holds then $\theta_x \propto c(x)/a(x)$:
\begin{align}
  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}.
\end{align}

Now what we are actually after is the parameters that maximize the joint log probability of the data+parameters, which I will write $J$:
\begin{align}
  J& = L + \log \text{D}(\theta|\alpha)\\
  &\stackrel{+}{\geq } \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x  + \sum_x(\alpha_x - 1)\log \theta_x\\
  & = \sum_x (c(x) + \alpha_x - 1)\log \theta_x -  N \log \sum_x a(x)\theta_x
\end{align}
So you add the pseudocounts to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$).

\bibliography{example_paper}
\bibliographystyle{icml2018}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018. It was modified from a version from Dan Roy in
% 2017, which was based on a version from Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2018-01-29 paper.tex
lambda calculus and intro

@@
 from anything it has seen before. It has at its disposal a basic set
-of primitives it can compose to build solutions to these problems, but
+of primitive actions it can compose to build solutions to these problems, but
 it is no idea what kinds of primitives are appropriate for which
@@
 which solutions are best expressed.
+How can our agent get off the ground?
+
+The AI and machine learning literature contains two broad takes on this problem.
+The first take is that the agent should come up with a better representation of the space of solutions,
+for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
+The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill}.
 
@@
 
-\section{Experiments}
-
-\begin{figure}
-  \includegraphics[width = 9cm]{figures/polynomial.png}
-  \includegraphics[width = 9cm]{figures/text.png}
-  \includegraphics[width = \columnwidth]{figures/circuit.png} 
-\end{figure}
-
-\section{Model}
-
 \begin{figure}
@@
   \draw [->,dashed] (x.east) to[out = 0,in = -90] node[fill = white]{$q(x)$} (tx.south);
+  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
   \draw [->,dashed] (tx.west) -- (z.east);
@@
 \end{figure}
+
+\section{Program Representation}
+We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
+A $\lambda$-calculus expression is either:
+\\\noindent A \emph{primitive}, like the number 5 or
+  the function \texttt{sum}.
+\\\noindent A \emph{variable}, like $x$, $y$, $z$
+\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function.
+$\lambda$-abstractions have a variable and a body.
+The body is a $\lambda$-calculus expression.
+Abstractions are written as $\lambda \text{var}.
+\text{body}$.
+\\\noindent An \emph{application} of a function to an argument.
+Both the function and the argument are $\lambda$-calculus expressions.
+The application of the function $f$ to the argument $x$ is written as $f\; x$.
+
+For example, the function which squares the logarithm of a number is
+$\lambda x.\text{\texttt{square}} (\text{\texttt{log} } x)$, and the identity function $f(x) = x$ is $\lambda x.x$. The
+$\lambda$-calculus serves as a spartan but expressive Turing complete
+program representation, and distills the essential features of functional languages like Lisp.
+
+However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
+(i.e., $5 \; (\lambda x.x)$).
+We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
+The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
+Hindley-Milner has two important features:
+Feature 1: It supports \emph{parametric polymorphism}: meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
+For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
+A function that returns the first element of a list has the type $\texttt{list}(\alpha)\to\alpha$.
+Type variables are not the same has variables introduced by $\lambda$-abstractions.
+Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
+A detailed exposition of Hindley-Milner is beyond the scope of this work.
+
+%% \begin{figure}
+%%   \begin{align}
+%%     \text{primitive types} & 
+%%     \end{align}
+%%   \end{figure}
+
+\section{Experiments}
+
+\begin{figure}
+  \includegraphics[width = 9cm]{figures/polynomial.png}
+  \includegraphics[width = 9cm]{figures/text.png}
+  \includegraphics[width = \columnwidth]{figures/circuit.png} 
+\end{figure}
+
+\section{Model}
 
@@
 
-\bibliography{example_paper}
+\bibliography{main}
 \bibliographystyle{icml2018}

### 2018-01-29 paper.tex
more writing

@@
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+\newcommand{\system}{EC2.0~}
 
@@
 it is no idea what kinds of primitives are appropriate for which
-problems nor does it know the higher-level domain-specific language in
+problems nor does it know the higher-level vocabulary in
 which solutions are best expressed.
@@
 for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
-The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill}.
+The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
+Our contribution is a general algorithm for fusing these two takes on the problem:
+we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
+alongside a bottom-up discriminative model that regresses from problems to solutions.
+We evaluate our algorithm on four domains:
+building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style programming problems.
+We show that \system can construct a set of basis primitives suitable for discovering solutions in each of these domains
 
-Our algorithm accomplishes the following:
-\begin{itemize}
-\item Learns from relatively small amounts of data and with weak supervision.
-We do not use ground truth programs -- weak supervision.
-Our DSL learner does not need huge amounts of data -- tens of examples suffice.
-  \item Jointly infers a \emph{generative model} along with a \emph{recognition model}.
-The \emph{generative model} is a probabilistic context-sensitive grammar over programs, and includes a DSL.
-The \emph{recognition model} is a neural network that guides the agents use of the DSL.
-\end{itemize}
-The generative model and the recognition model bootstrap off of each other:
-\begin{itemize}
-\item The recognition model works by upweighting the probability of program components likely to be useful for a given problem.
-By learning a DSL, the recognition model gets more power because it can upweight new more powerful primitives that are better suited for the domain.
-\item The generative model (DSL) is learned from programs that the agent has found so far. Because the recognition model speeds up search, it generates more training data for the generative model.
-\end{itemize}
+We cast these problems as \emph{Bayesian Program
+  Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
+where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
+A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
+We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
+define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
+In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
+where the likelihood $\probability[x|p]$ is domain-dependent.
+The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
+Alongside this generative model,
+we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
+
+Our key observation is that the generative and recognition models can bootstrap off of each other,
+greatly increasing the tractability of BPL.
 
@@
   \node[latent] at (2.5,3) (t){$\theta$};
-  \node[latent] at (1,1.5) (z){$z$};
+  \node[latent] at (1,1.5) (z){$p$};
   \node[latent] at (2.5,1.5) (tx){$\theta_x$};
@@
   \end{tikzpicture}
-  \caption{DSL $\mathcal{D}$ generates programs $z$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
+  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
 We observe program outputs $x$.
@@
 Solid arrows correspond to the top-down generative model.
-Dashed arrows correspond to the bottom-up recognition model.}
+Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
 \end{figure}

### 2018-01-30 paper.tex
losses

@@
 
+\subsection{Boolean circuits}
+pedagogical example; easy domain
 \begin{figure}
-  \includegraphics[width = 9cm]{figures/polynomial.png}
-  \includegraphics[width = 9cm]{figures/text.png}
   \includegraphics[width = \columnwidth]{figures/circuit.png} 
@@
 
+\subsection{Symbolic Regression}
+Here, we show how to use \system to infer programs with continuous, real-valued parameters.
+
+Demonstrates how to integrate continuous parameters into a program using our framework.
+$p:\mathbb{R}\to\mathbb{R}$
+\begin{align*}
+&\log   \probability\left[\{(x_i,y_i)\}_{i\leq N}|p\right]\stackrel{+}{ = }\log \int \mathrm{d}C\; p_C(C)\prod_{i\leq N}\text{Normal}(p(x_i)|y_i)\\
+\text{\emph{BIC:}} &\approx \sum_{i\leq N}\log \text{Normal}(p(x_i)|y_i) - \frac{|C|\log N}{2}
+  \end{align}
+\begin{figure}
+      \includegraphics[width = 9cm]{figures/polynomial.png}
+\end{figure}
+\subsection{String editing}
+\begin{figure}
+
+  \includegraphics[width = 9cm]{figures/text.png}
+
+\end{figure}
+
+\subsection{List problems}
+
 \section{Model}
 
-\section{Implementation}
+\begin{align}
+  \mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
+  \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_p Q_x(p)\log \probability [p|\mathcal{D},q(x)]\right]\nonumber\\
+    \mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log \probability[p|\mathcal{D},q(x)]\right],\text{ $p$ evaluates to $x$}\nonumber
+  \end{align}
 
@@
   \STATE var $\gets$ an unused variable name
-  \STATE body $\sim$ sample$(\mathcal{D},\theta,[\text{var}:\alpha]+\mathcal{E},\beta)$
+  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
    \STATE \textbf{return} $\lambda \text{var}.$ body
@@
    %   \ELSE
-   \STATE\begin{align*}     \text{primitives} \gets\{p | &p:\alpha\to\cdots\to\beta \in \mathcal{D}\cup\mathcal{E}\\     &\text{if canUnify}(\tau,\beta) \}     \end{align*}
+   \STATE $\text{primitives} \gets\{p | &p:\alpha\to\cdots\to\beta \in \mathcal{D}\cup\mathcal{E}$
+   \STATE \hspace{2.5cm}$\text{if canUnify}(\tau,\beta) \} $
 
-   \STATE Sample $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ and w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
+   \STATE Sample $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
+   \STATE \hspace{3.35cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
    \STATE Let $e:\alpha_1\to\alpha_2\to\cdots\to \alpha_K\to\beta$. Unify $\tau$ with $\beta$.
@@
  \ENDFOR
- \STATE \textbf{return} $e(a_1, a_2, \cdots, a_K)$
+ \STATE \textbf{return} $e\;a_1\; a_2\; \cdots\; a_K$
  %  \UNTIL{$noChange$ is $true$}
@@
 
-\section{Estimating the grammar parameters}
+\section{Estimating $\theta$}
 
@@
 Writing $L$ for the log likelihood, $\theta$ for the parameters of the grammar, $N$ for the number of random choices, $A$ to range over the alternative choices for a random variable, $c(x)$ to mean the number of times that primitive $x$ was used, and $a(x) = \sum_A \indicator [x\in A]$ to mean the number of times that primitive $x$ could have been used:
-\begin{align}
-  L& = \sum_x c(x)\log \theta_x - \sum_A \log \sum_{x\in A}\theta_x\\
-  & = \sum_x c(x)\log \theta_x - N\expect_A \log \sum_{x\in A}\theta_x\\
-  &\geq\sum_x c(x)\log \theta_x - N \log \expect_A \sum_{x\in A}\theta_x\text{, Jensen's inequality}\\
-  & = \sum_x c(x)\log \theta_x - N \log \frac{1}{N}\sum_A \sum_{x} \indicator [x\in A]\theta_x\\
-  & \stackrel{+}{=} \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x.
+\begin{align*}
+  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
+  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
+  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
 \end{align}
-Differentiate with respect to $\theta_x$ and set to zero:
+where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.
+
+Differentiate with respect to $\theta_e$ and set to zero 
 \begin{align}

### 2018-01-30 paper.tex
symbolic regression

@@
 \subsection{Symbolic Regression}
-Here, we show how to use \system to infer programs with continuous, real-valued parameters.
+We show how to use \system to infer programs containing both discrete
+structure and continuous parameters. The high-level idea is to synthesize programs with unspecified-real-valued parameters, and to fit those parameters using gradient descent.
+Concretely, we ask the algorithm to
+solve a set of 1000 symbolic regression problems, each a polynomial of
+degree 0, 1, or 2, where our observations $x$ take the form of $N$
+input/output examples, which we write as $x = \left\{(i_n,o_n)
+\right\}_{n\leq N}$. For example, one task is to infer a program
+calculating $3x + 2$, and the observations are the input-output
+examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.
 
-Demonstrates how to integrate continuous parameters into a program using our framework.
-$p:\mathbb{R}\to\mathbb{R}$
+We initially equip our DSL learner with addition and multiplication,
+along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
+We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
 \begin{align*}
-&\log   \probability\left[\{(x_i,y_i)\}_{i\leq N}|p\right]\stackrel{+}{ = }\log \int \mathrm{d}C\; p_C(C)\prod_{i\leq N}\text{Normal}(p(x_i)|y_i)\\
-\text{\emph{BIC:}} &\approx \sum_{i\leq N}\log \text{Normal}(p(x_i)|y_i) - \frac{|C|\log N}{2}
+\log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P_{\vec{\mathcal{R}}}(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
+%\text{\emph{BIC:}} &\approx 
 \end{align}
+where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$.
+We approximate this marginal using the BIC~\cite{Bishop:2006:PRM:1162264}:
+\begin{align*}
+  \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
+\end{align}
+where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.
+
+\begin{tabular}{ll}
+  \toprule
+  Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
+  & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
+  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
+  \midrule
+  Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
+  \\\midrule
+  \begin{tabular}{l}
+    Subset of \\Learned DSL
+    \end{tabular}&
+  \begin{tabular}{ll}
+    $\lambda x. \mathcal{R}\times x + \mathcal{R}$ &linear\\
+    $\lambda x. \mathcal{R} + x$ &increment\\
+    $\lambda x. x\times (\text{linear } x)$&$\text{quadratic}_0$\\
+    $\lambda x. \text{increment }(\text{quadratic}_0\; x)$&quadratic\\
+  \end{tabular}
+
+  \\\bottomrule
+  \end{tabular}
+
 \begin{figure}
@@
 
-I justify this estimator by proving that it maximizes a lower bound on the log likelihood of the data.
-Writing $L$ for the log likelihood, $\theta$ for the parameters of the grammar, $N$ for the number of random choices, $A$ to range over the alternative choices for a random variable, $c(x)$ to mean the number of times that primitive $x$ was used, and $a(x) = \sum_A \indicator [x\in A]$ to mean the number of times that primitive $x$ could have been used:
+We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.\ref{programGenerativeModel}.
+Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
 \begin{align*}

### 2018-01-30 paper.tex
neural network figure

@@
   \node[latent] at (1,1.5) (z){$p$};
-  \node[latent] at (2.5,1.5) (tx){$\theta_x$};
-  \node[obs] at (1,0) (x) {$x$};
+  \node[latent] at (3.5,1.5) (tx){$\theta_x$};
+  \node[obs] at (1,-0.5) (x) {$x$};
   \edge {z}{x};
   \edge {d,t}{z};
-  \plate {}{(tx)(z)(x)}{$N$};
-  \draw [->,dashed] (x.east) to[out = 0,in = -90] node[fill = white]{$q(x)$} (tx.south);
+  \plate {}{(tx)(z)(x)}{$x\in X$};
+  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn)[fill = white]{
+          \includegraphics[width = 1.5cm]{figures/MLP.png} 
+  } (tx.south);
+  \node at (nn.south) {$q(x)$};
   \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
@@
 \end{figure}
+
+\section{The \system Algorithm}
 
@@
 
+What DSL does \system learn?
+The learned DSL contains templates for quadratic and linear functions,
+which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain.
+Examining the programs themselves,
+one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
+for example, it represents the polynomial $8x^2+8x$
+
 \begin{tabular}{ll}
@@
   \end{tabular}
-
   \\\bottomrule

### 2018-01-31 paper.tex
more intro

@@
 
+\subsection{Motivation and Overview}
+
+\system takes as input a set of \emph{tasks}, written $X$, each of which is a program induction problem.
+It has at its disposal a \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
+Its goal is to solve each of the tasks by writing a program,
+and also to infer a DSL $\mathcal{D}$ that distills the commonalities across all of the programs that solve the tasks.
+
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}.
+From Fig.~\ref{graphicalModel}, the MAP probability of $(\mathcal{D},\theta)$ is
+\begin{equation}
+  \probability[\mathcal{D},\theta|X]\propto \probability[\mathcal{D},\theta] \prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\label{probabilityDSL}
+\end{equation}
+If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{probabilityDSL}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
+Through this lens we now take as our goal to maximize Eq.~\ref{probabilityDSL}.
+But even \emph{evaluating} Eq.~\ref{probabilityDSL} is intractable because it involves summing over the infinite set of all possible programs.
+In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
+
+With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{probabilityDSL},
+which we call $J$:
+\begin{equation}
+J = \probability[\mathcal{D},\theta]\prod_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
+\end{equation}
+This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:
+
+\noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+is the set of programs discovered by \system where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
+
+We maximize $J$ by alternatingly maximizing it w.r.t. the DSL and the frontiers:
+\\\noindent \textbf{Program Synthesis: Maximizing $J$ w.r.t. the frontiers.} Here we
+want to find new programs to add to  the frontiers so that $J$ increases the most.
+Adding new programs to the frontiers means attempting to synthesize new programs for each of the tasks.
+So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
+explains how \system synthesizes new programs and given a DSL.
+\\\noindent \textbf{DSL Induction: Maximizing $J$ w.r.t.
+the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed and so we can evaluate $J$.
+Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
+Section~\ref{grammarInductionSection} explains this step of the algorithm.
+
+In practice, synthesizing programs is extremely difficult because
+of how large the search space is. Learning the DSL eases the difficulty of synthesis
+by exposing a domain-specific basis for constructing programs.
+Another complementary means of meeting search is to learn a
+bottom-up \emph{recognition model}
+
+\subsection{Synthesizing Programs from a DSL}\label{explorationSection}
+
+Describe the enumerator, depth first search on the random choices made by the generative model with iterative deepening
+
+Limitation: has a hard time learning arbitrary constants. Solution: solve those constants using specialized solvers.
+
+\subsection{Learning a Neural Recognition Model}
+The key idea here is that if the DSL is suitably fit to the domain then
+we can get away with a simple, low-capacity model.
+
+\begin{align}
+  \mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
+  \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_p Q_x(p)\log \probability [p|\mathcal{D},q(x)]\right]\nonumber\\
+    \mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log \probability[p|\mathcal{D},q(x)]\right],\text{ $p$ evaluates to $x$}\nonumber
+\end{align}
+
+\subsection{Inducing the DSL from the Frontiers}\label{grammarInductionSection}
+
+\begin{algorithm}[tb]
+   \caption{Grammar Induction Algorithm}
+   \label{grammarInductionAlgorithm}
+   \begin{algorithmic}
+     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
+     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$, AIC coefficient $a$
+     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+     \STATE Define $\log \probability[\mathcal{D}]\stackrel{+}{ = } -\lambda\sum_{p\in \mathcal{D}} \text{size}(p)$
+     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
+     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
+     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - a|\mathcal{D}|$
+     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
+     \WHILE {true}
+     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ a subtree of }z\}$
+     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
+     \IF{$\text{score}(\mathcal{D}') > \text{score}(\mathcal{D})$}
+     \STATE $\mathcal{D}\gets\mathcal{D}'$
+     \ELSE
+     \STATE\textbf{return} $\mathcal{D},\theta^*(\mathcal{D})$
+     \ENDIF
+     \ENDWHILE
+   \end{algorithmic}
+\end{algorithm}
 \section{Program Representation}
@@
 
-\begin{align}
-  \mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
-  \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_p Q_x(p)\log \probability [p|\mathcal{D},q(x)]\right]\nonumber\\
-    \mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log \probability[p|\mathcal{D},q(x)]\right],\text{ $p$ evaluates to $x$}\nonumber
-  \end{align}
-
 \begin{algorithm}[tb]
@@
  \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[tb]
-   \caption{Grammar Induction Algorithm}
-   \label{grammarInductionAlgorithm}
-   \begin{algorithmic}
-     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$, AIC coefficient $a$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-     \STATE Define $\log \probability[\mathcal{D}]\stackrel{+}{ = } -\lambda\sum_{p\in \mathcal{D}} \text{size}(p)$
-     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
-     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
-     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - a|\mathcal{D}|$
-     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-     \WHILE {true}
-     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ a subtree of }z\}$
-     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
-     \IF{$\text{score}(\mathcal{D}') > \text{score}(\mathcal{D})$}
-     \STATE $\mathcal{D}\gets\mathcal{D}'$
-     \ELSE
-     \STATE\textbf{return} $\mathcal{D},\theta^*(\mathcal{D})$
-     \ENDIF
-     \ENDWHILE
 \end{algorithmic}

### 2018-01-31 paper.tex
more work on the intro

@@
 
-\subsection{Motivation and Overview}
+The goal of \system is to both induce a DSL and find good programs solving each of the tasks.
+Our strategy is to
+alternate between 3 different steps: (1) improving the DSL, (2) searching for programs that solve the tasks, and (3) learning a
+better neural recognition model, which we use to accelerate the search over programs.
+The key observation here is that each of these three steps can bootstrap off of each other:
+
+\begin{itemize}
+\item Improving the DSL: We induce a DSL from the programs we have found so far which solve the tasks;
+  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
+\item Searching for programs: Searching for programs is done using the neural recognition model --
+  so the recognition model bootstraps the search process.
+  \item Learning a recognition model: The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+As the DSL improves and we find more programs, the recognition model gets both more data to train on and better data.
+\end{itemize}
+
+Section~\ref{mathematicalFraming} frames this 3-step procedure as
+a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
+Section~\ref{grammarInductionSection} explains how \system induces a DSL from programs;
+Section~\ref{explorationSection} explains how we search for programs that solve the tasks; and Section~\ref{recognitionSection}
+explains how we train a neural network to accelerate the search over programs.
+
+\subsection{Probabilistic Framing}\label{mathematicalFraming}
 
@@
 
-\subsection{Learning a Neural Recognition Model}
+\subsection{Learning a Neural Recognition Model}\label{recognitionSection}
 The key idea here is that if the DSL is suitably fit to the domain then

### 2018-01-31 paper.tex
enumeration section

@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}.
-From Fig.~\ref{graphicalModel}, the MAP probability of $(\mathcal{D},\theta)$ is
-\begin{equation}
-  \probability[\mathcal{D},\theta|X]\propto \probability[\mathcal{D},\theta] \prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\label{probabilityDSL}
-\end{equation}
-If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{probabilityDSL}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
-Through this lens we now take as our goal to maximize Eq.~\ref{probabilityDSL}.
-But even \emph{evaluating} Eq.~\ref{probabilityDSL} is intractable because it involves summing over the infinite set of all possible programs.
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}. We wish to maximize the MAP probability of $\mathcal{D}$:
+\begin{equation*}
+  \probability[\mathcal{D}|X]\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]
+\end{equation*}
+In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
+Conventional NLP approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
+\begin{align}
+  A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
+&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
+  \end{align}
+
+If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{AIC}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
+Through this lens we now take as our goal to maximize Eq.~\ref{AIC}.
+But even \emph{evaluating} Eq.~\ref{AIC} is intractable because it involves summing over the infinite set of all possible programs.
 In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
-
-With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{probabilityDSL},
+With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{AIC},
 which we call $J$:
 \begin{equation}
-J = \probability[\mathcal{D},\theta]\prod_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
+J = \log \probability[\mathcal{D},\theta] + \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
 \end{equation}
@@
 \noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is the set of programs discovered by \system where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
+is a set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
 
 We maximize $J$ by alternatingly maximizing it w.r.t. the DSL and the frontiers:
-\\\noindent \textbf{Program Synthesis: Maximizing $J$ w.r.t. the frontiers.} Here we
-want to find new programs to add to  the frontiers so that $J$ increases the most.
-Adding new programs to the frontiers means attempting to synthesize new programs for each of the tasks.
-So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
-explains how \system synthesizes new programs and given a DSL.
-\\\noindent \textbf{DSL Induction: Maximizing $J$ w.r.t.
+%So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
+%explains how \system synthesizes new programs and given a DSL.
+\\\noindent \textbf{DSL Induction: Maxing $J$ w.r.t.
 the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed and so we can evaluate $J$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
-Section~\ref{grammarInductionSection} explains this step of the algorithm.
+\\\noindent \textbf{Program Search: Maxing $J$ w.r.t. the frontiers.} Here we
+want to find new programs to add to  the frontiers so that $J$ increases the most.
+Adding new programs to the frontiers means searching for new programs $p$ for task $x$
+where $\probability[x,p|\mathcal{D},\theta]$ is large.
 
-In practice, synthesizing programs is extremely difficult because
-of how large the search space is. Learning the DSL eases the difficulty of synthesis
-by exposing a domain-specific basis for constructing programs.
-Another complementary means of meeting search is to learn a
-bottom-up \emph{recognition model}
+Searching for programs is extremely difficult because
+of how large the search space is. We ease the difficulty of the search by learning a neural recognition model:
 
-\subsection{Synthesizing Programs from a DSL}\label{explorationSection}
+\textbf{Neural recognition model: tractably maxing $J$ w.r.t. the
+  frontiers.}  Here we train a neural network, $q$, to predict a
+distribution over programs conditioned on a task. The objective of $q$
+is to assign high probability to programs $p$ where
+$\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
+frontier $\mathcal{F}_x$ by searching for programs maximizing
+$q(p|x)$.
+The network $q$ exploits the structure of the DSL $\mathcal{D}$:
+rather than directly predicting a distribution over $p$ conditioned on $x$,
+it predicts a weight vector, $\theta_x$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
+This approach implements an amortized
+inference scheme~\cite{ritchie2016deep} for the generative model in
+Fig.~\ref{graphicalModel}.
 
-Describe the enumerator, depth first search on the random choices made by the generative model with iterative deepening
+%% Learning the DSL eases the difficulty of synthesis
+%% by exposing a domain-specific basis for constructing programs.
+%% Another complementary means of meeting search is to learn a
+%% bottom-up \emph{recognition model}
 
-Limitation: has a hard time learning arbitrary constants. Solution: solve those constants using specialized solvers.
+\subsection{Searching for Programs}\label{explorationSection}
+
+Now our goal is to search for programs that solve the tasks.  In this
+work we use the simple search strategy of enumerating programs from
+the DSL  in decreasing order of their probability,
+and then checking if an enumerated program $p$ assigns positive
+probability to a task ($\probability[x|p] > 0$); if so, we include $p$ in
+the frontier $\mathcal{F}_x$.
+
+To make this concrete we need to define what programs actually are and
+what form $\probability[p |\mathcal{D},\theta]$ takes.
+In this work, we represent programs as \emph{polymorphicly-type $\lambda$-calculus expressions}.
+$\lambda$-calculus is a formalism for expressing functional programs.
+It includes variables, function application, and the ability to create new functions using ...
+
+TODO: summarize lambda calculus and types in one paragraph
+
+\noindent\textbf{Definition: $\mathcal{D}$.}
+A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
+\\\noindent\textbf{Definition: $\theta$.}
+A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
+one number for each DSL primitive $e:\tau\in \mathcal{D}$, written $\theta_e$,
+and a weight controlling the probability of a variable occurring in a program, written $\theta_{\text{var}}$.
+
+Algorithm~\ref{programGenerativeModel} is a procedure for drawing
+samples from $\probability[p|\mathcal{D},\theta]$.  In practice, we
+enumerate programs rather than sampling them.  Enumeration proceeds by
+a depth-first search over the random choices made by
+Algorithm~\ref{programGenerativeModel}; we wrap the depth-first search
+in iterative deepening to (approximately) build $\lambda$-calculus
+expressions in order of their probability.
+
+\begin{algorithm}[tb]
+   \caption{Generative model over programs}
+   \label{programGenerativeModel}
+   \begin{algorithmic}
+     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
+  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
+  \STATE \textbf{Output:} a program whose type unifies with $\tau$
+  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
+\STATE
+     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
+  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
+  \STATE \textbf{Output:} a program whose type unifies with $\tau$
+  \IF{$\tau = \alpha\to\beta$}
+  \STATE var $\gets$ an unused variable name
+  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
+   \STATE \textbf{return} $\lambda \text{var}.$ body
+   \ENDIF
+   %   \ELSE
+   \STATE $\text{primitives} \gets\{p | &p:\alpha\to\cdots\to\beta \in \mathcal{D}\cup\mathcal{E}$
+   \STATE \hspace{2.5cm}$\text{if canUnify}(\tau,\beta) \} $
+
+   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
+   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
+   \STATE Let $e:\alpha_1\to\alpha_2\to\cdots\to \alpha_K\to\beta$. Unify $\tau$ with $\beta$.
+%   \STATE unify$(\tau,\beta)$
+   \FOR{$k=1$ {\bfseries to} $K$}
+ \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
+ \ENDFOR
+ \STATE \textbf{return} $e\;a_1\; a_2\; \cdots\; a_K$
+ %  \UNTIL{$noChange$ is $true$}
+\end{algorithmic}
+\end{algorithm}
+
+Why enumerate, when the program synthesis community has invented many
+sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
+We have two reasons:% for using the simple strategy of enumerating programs:
+\begin{itemize}
+\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
+  the space of of programs.
+\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the
+  search algorithm is very simple.  
+\end{itemize}
+A main drawback of an enumerative search algorithm is that we have no
+efficient means of solving for arbitrary constants that might occur in the
+program. In Section~\ref{regressionSection},
+we will show how to find programs with real-valued constants
+by automatically differentiating through the program and setting the constants using gradient descent.
 
@@
      \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$, AIC coefficient $a$
+     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
@@
      \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
-     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - a|\mathcal{D}|$
+     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
      \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
@@
 
-\subsection{Symbolic Regression}
+\subsection{Symbolic Regression}\label{regressionSection}
 We show how to use \system to infer programs containing both discrete
@@
 \section{Model}
-
-\begin{algorithm}[tb]
-   \caption{Generative model over programs}
-   \label{programGenerativeModel}
-   \begin{algorithmic}
-     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
-  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
-  \STATE \textbf{Output:} a program whose type unifies with $\tau$
-  \IF{$\tau = \alpha\to\beta$}
-  \STATE var $\gets$ an unused variable name
-  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
-   \STATE \textbf{return} $\lambda \text{var}.$ body
-   \ENDIF
-   %   \ELSE
-   \STATE $\text{primitives} \gets\{p | &p:\alpha\to\cdots\to\beta \in \mathcal{D}\cup\mathcal{E}$
-   \STATE \hspace{2.5cm}$\text{if canUnify}(\tau,\beta) \} $
-
-   \STATE Sample $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
-   \STATE \hspace{3.35cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
-   \STATE Let $e:\alpha_1\to\alpha_2\to\cdots\to \alpha_K\to\beta$. Unify $\tau$ with $\beta$.
-%   \STATE unify$(\tau,\beta)$
-   \FOR{$k=1$ {\bfseries to} $K$}
- \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
- \ENDFOR
- \STATE \textbf{return} $e\;a_1\; a_2\; \cdots\; a_K$
- %  \UNTIL{$noChange$ is $true$}
-\end{algorithmic}
-\end{algorithm}
 

### 2018-01-31 paper.tex
reorder summary according to exposition

@@
 We evaluate our algorithm on four domains:
-building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style programming problems.
+building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
 We show that \system can construct a set of basis primitives suitable for discovering solutions in each of these domains
@@
 The goal of \system is to both induce a DSL and find good programs solving each of the tasks.
-Our strategy is to
-alternate between 3 different steps: (1) improving the DSL, (2) searching for programs that solve the tasks, and (3) learning a
-better neural recognition model, which we use to accelerate the search over programs.
+Our strategy is to iterate through three steps: (1) searching for programs that
+solve the tasks, (2) learning a better neural recognition model -- which we use
+to accelerate the search over programs -- and (3) improving the DSL.
 The key observation here is that each of these three steps can bootstrap off of each other:
@@
 \begin{itemize}
-\item Improving the DSL: We induce a DSL from the programs we have found so far which solve the tasks;
-  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
-\item Searching for programs: Searching for programs is done using the neural recognition model --
+  \item Searching for programs: Program search uses a distribution determined by the neural recognition model --
   so the recognition model bootstraps the search process.
@@
 As the DSL improves and we find more programs, the recognition model gets both more data to train on and better data.
+  \item Improving the DSL: We induce a DSL from the programs we have found so far which solve the tasks;
+  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
 \end{itemize}
@@
 a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
-Section~\ref{grammarInductionSection} explains how \system induces a DSL from programs;
-Section~\ref{explorationSection} explains how we search for programs that solve the tasks; and Section~\ref{recognitionSection}
-explains how we train a neural network to accelerate the search over programs.
+Section~\ref{explorationSection} explains how we search for programs that solve the tasks;
+Section~\ref{recognitionSection} explains how we train a neural network to accelerate the search over programs; and
+Section~\ref{grammarInductionSection} explains how \system induces a DSL from programs.
 

### 2018-01-31 paper.tex
more reordering according to exposition
merge PDF

@@
   \node[latent] at (1,1.5) (z){$p$};
-  \node[latent] at (3.5,1.5) (tx){$\theta_x$};
+  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
   \node[obs] at (1,-0.5) (x) {$x$};
@@
 We observe program outputs $x$.
-A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta_x = q(x)$).
+A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta^{(x)} = q(x)$).
 Solid arrows correspond to the top-down generative model.
@@
 %explains how \system synthesizes new programs and given a DSL.
-\\\noindent \textbf{DSL Induction: Maxing $J$ w.r.t.
-the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed and so we can evaluate $J$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
 \\\noindent \textbf{Program Search: Maxing $J$ w.r.t. the frontiers.} Here we
@@
 where $\probability[x,p|\mathcal{D},\theta]$ is large.
+\\\noindent \textbf{DSL Induction: Maxing $J$ w.r.t.
+the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed and so we can evaluate $J$.
+Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
 
@@
 rather than directly predicting a distribution over $p$ conditioned on $x$,
-it predicts a weight vector, $\theta_x$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
+it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
 This approach implements an amortized
@@
 by automatically differentiating through the program and setting the constants using gradient descent.
+In Section~\ref{textSection}
+we will show that the bottom-up neural recognition model can learn
+which discrete constants should be included in a program.
 
 \subsection{Learning a Neural Recognition Model}\label{recognitionSection}
-The key idea here is that if the DSL is suitably fit to the domain then
-we can get away with a simple, low-capacity model.
 
+The purpose of the recognition model is to accelerate the search over
+programs.  It does this by learning to predict which programs both 
+assign high likelihood to a task, and at the same time 
+have high prior probability under the prior $\probability[\cdot |\mathcal{D},\theta]$.
+
+The recognition model $q$ is a neural network that predicts,
+for each task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.
+Together with the DSL, this defines a distribution over programs,
+$\probability[p|\mathcal{D},\theta = q(x)]$.
+We abbreviate this distribution as $q(p|x)$.
+The crucial aspect of this framing is that the neural network
+can leverage the structure of the DSL,
+and is \emph{not} responsible for
+generating programs wholesale.
+We will show that this lets us get away with a simple, low-capacity neural network.
+
+We want a recognition model that closely approximates the true posteriors over programs, and so aim to minimize the following KL-divergence:
+\begin{equation*}
+  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]
+\end{equation*}
+which is equivalent to maximizing
+$$
+  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
+$$
+  where the expectation is taken over tasks. One could take this expectation
+  over the empirical distribution of the observations,
+  like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
+  We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the \system objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
   \begin{align}
   \mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
-  \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_p Q_x(p)\log \probability [p|\mathcal{D},q(x)]\right]\nonumber\\
-    \mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log \probability[p|\mathcal{D},q(x)]\right],\text{ $p$ evaluates to $x$}\nonumber
+  \mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\\
+  \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+    \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
   \end{align}
+
+  Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
+  the current DSL, running them to get their outputs,
+  and then training $q$ to regress from the outputs to the program.
+  If these programs map inputs to outputs,
+  then we need some way of sampling these inputs as well.
+  Our solution to this problem is to sample the inputs
+  from the empirical observed distribution of inputs in $X$.
 
@@
 \end{figure}
-\subsection{String editing}
+\subsection{String editing}\label{textSection}
 \begin{figure}
@@
 \begin{align}
-  \frac{c(x)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
+  \frac{c(x)}{\theta^{(x)}} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
 \end{align}
-This equality holds if $\theta_x = c(x)/a(x)$:
+This equality holds if $\theta^{(x)} = c(x)/a(x)$:
 \begin{align}

### 2018-01-31 paper.tex
started on list functions

@@
 %\text{\emph{BIC:}} &\approx 
-\end{align}
+\end{align*}
 where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$.
@@
   \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
-\end{align}
+\end{align*}
 where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.
@@
 
-\subsection{List problems}
+\subsection{List functions}
+Our list function domain consists of tasks which are solved by functions
+that take as input an integer or a list of integers, and have as output
+either a Boolean, an integer, or a list of integers. Examples of these
+functions are in Table~\ref{listexamples}. For each function, we create a
+task $x$ by generating 15 input/output examples used for testing whether a
+program produces the correct output. Supplying many examples reduces
+ambiguity in the task's function, ensuring solutions achieve the desired
+concept.
+
+\begin{table}
+\centering
+\begin{tabular}{| l | l | l |}
+  \hline
+  \emph{name} & \emph{input} & \emph{output} \\
+  \hline
+  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
+  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
+  len & [3\, 5\, 12\, 1] & 4 \\
+  range & 3 & [1\, 2\, 3] \\
+  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
+  has-4 & [4\, 5\, 7\, 4] & \texttt{true} \\
+  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
+  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+  \hline
+\end{tabular}
+\caption{Examples from the domain of list functions.}
+\label{listexamples}
+\end{table}
+
+We supply \system with the DSL outlined in Table~\ref{listdsl}.
+
+\begin{table}
+\centering
+\begin{tabular}{| l | l |}
+  \hline
+  \emph{name} & \emph{type} \\
+  \hline
+    empty & \texttt{tlist($\alpha$)} \\
+    singleton & \texttt{arrow($\alpha$, tlist($\alpha$))} \\
+    range & \texttt{int $\rightarrow$ tlist(int)} \\
+    concat & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
+    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\beta$)} \\
+    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\beta$} \\
+    \hline
+    true & \texttt{bool} \\
+    not & \texttt{bool $\rightarrow$ bool} \\
+    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
+    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
+    \hline
+    0, $\ldots$, 9 & \texttt{int} \\
+    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
+    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
+    negate & \texttt{int $\rightarrow$ int} \\
+    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
+    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
+    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
+    is-prime & \texttt{int $\rightarrow$ bool} \\
+    is-square & \texttt{int $\rightarrow$ bool} \\
+    sort & \texttt{tlist(int) $\rightarrow$ tlist(int)} \\
+    \hline
+    sum & \texttt{tlist(int) $\rightarrow$ int} \\
+    reverse & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
+    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
+    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
+    index & \texttt{int $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\alpha$} \\
+    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
+    slice & \texttt{int $\rightarrow$ int $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
+  \hline
+\end{tabular}
+\caption{DSL for the domain of list function.}
+\label{listdsl}
+\end{table}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figures/list.eps}
+\end{figure}
+
+We found that using a less sophisticated but equally-capable DSL made common
+patterns, such as summation, unlikely and unlearnable in a small enumeration
+bound.
 

### 2018-01-31 paper.tex
progress on DSL induction section

@@
    %   \ELSE
-   \STATE $\text{primitives} \gets\{p | &p:\alpha\to\cdots\to\beta \in \mathcal{D}\cup\mathcal{E}$
-   \STATE \hspace{2.5cm}$\text{if canUnify}(\tau,\beta) \} $
+   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau')) \} $
 
@@
  \STATE \textbf{return} $e\;a_1\; a_2\; \cdots\; a_K$
- %  \UNTIL{$noChange$ is $true$}
+ \STATE
+ \STATE \textbf{function} yield$(\tau)$:
+ \STATE \textbf{Input:} type $\tau$
+ \STATE \textbf{Output:} the return type of $\tau$
+ \STATE \textbf{if} $\tau = \alpha\to \beta$ \textbf{then return }yield$(\beta)$ \textbf{else return }$\tau$
 \end{algorithmic}
@@
 
+The purpose of the DSL is to
+offer a set of abstractions
+that allow an agent to easily express solutions to the tasks at hand.
+In the \system algorithm we infer the DSL from a collection of frontiers.
+Intuitively, we want the algorithm to
+look at the programs in the frontiers and
+generalize beyond them; 
+not only so the DSL can better express the current solutions,
+but also so that the DSL might expose new abstractions
+which will later used to
+discover even better programs.
+
+Exact maximization of $J$ (Eq.\ref{lowerBound}) w.r.t. $(\mathcal{D},\theta)$
+is intractable,
+so we take a more heuristic approach.
+The strategy is to search locally through the space of DSLs,
+proposing small local changes to $\mathcal{D}$ until $J$ fails to increase.
+The search moves work by introducing new
+$\lambda$-expressions into the DSL.
+We propose these new expressions by extracting subexpressions from
+programs already in the frontier.
+These extracted subexpressions
+are fragments of the original programs, and can introduce new variables (Figure~\ref{fragmentExample}),
+which then become new functions in the DSL.
+
+Closely related to Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
+but context-sensitive
+
+We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
+\begin{align*}
+  \probability[\mathcal{D}]&\propto\exp\left(\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
+  P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)
+\end{align*}
+where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
+and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
+Algorithm~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
+
+Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
+Although this problem may seem 
+very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+$\mathcal{D}$
+ is actually context-sensitive
+due to the presence of variables in the programs and also due to the
+polymorphic typing system.
+In the Appendix we derive an efficient MAP estimator for $\theta$.
+
 \begin{algorithm}[tb]
-   \caption{Grammar Induction Algorithm}
+   \caption{DSL Induction Algorithm}
    \label{grammarInductionAlgorithm}
@@
      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-     \STATE Define $\log \probability[\mathcal{D}]\stackrel{+}{ = } -\lambda\sum_{p\in \mathcal{D}} \text{size}(p)$
      \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
@@
      \WHILE {true}
-     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ a subtree of }z\}$
+     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
      \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
-     \IF{$\text{score}(\mathcal{D}') > \text{score}(\mathcal{D})$}
+     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
      \STATE $\mathcal{D}\gets\mathcal{D}'$
-     \ELSE
-     \STATE\textbf{return} $\mathcal{D},\theta^*(\mathcal{D})$
-     \ENDIF
      \ENDWHILE
@@
 \end{algorithm}
+
+\begin{figure*}
+  \begin{tabular}{lll}
+    \toprule
+    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
+    Boolean circuits&
+    \begin{tabular}{l}
+      \texttt{(lambda x (nand x x))}\\
+      \texttt{(lambda x (lambda y (nand x (nand y y))}
+    \end{tabular}&
+    \texttt{(nand z z)}\\\midrule
+    String editing&
+    \begin{tabular}{l}
+      \texttt{(lambda s (+ ',' (index 0 (split ',' s))))}\\
+      \texttt{(lambda s (index 0 (split ' ' s)))}
+    \end{tabular}&
+    \texttt{(index 0 (split z s))}\\\bottomrule
+  \end{tabular}
+  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
+    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} in the right column).}
+  \label{fragmentExample}
+  \end{figure*}
 \section{Program Representation}

### 2018-02-01 paper.tex
moved stuff that is going to be in the appendix to the bottom of the document
minor edits

@@
 \noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is a set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
+is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
 
@@
 what form $\probability[p |\mathcal{D},\theta]$ takes.
-In this work, we represent programs as \emph{polymorphicly-type $\lambda$-calculus expressions}.
+In this work, we represent programs as \emph{polymorphicly-typed $\lambda$-calculus expressions}.
 $\lambda$-calculus is a formalism for expressing functional programs.
@@
    \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
-   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau')) \} $
+   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
 
@@
 \begin{itemize}
+\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the  search algorithm is very simple.  
 \item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
   the space of of programs.
-\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the
-  search algorithm is very simple.  
 \end{itemize}
@@
 polymorphic typing system.
-In the Appendix we derive an efficient MAP estimator for $\theta$.
+In the Appendix we derive a tractable MAP estimator for $\theta$.
 
@@
   \end{figure*}
-\section{Program Representation}
-We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
-A $\lambda$-calculus expression is either:
-\\\noindent A \emph{primitive}, like the number 5 or
-  the function \texttt{sum}.
-\\\noindent A \emph{variable}, like $x$, $y$, $z$
-\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function.
-$\lambda$-abstractions have a variable and a body.
-The body is a $\lambda$-calculus expression.
-Abstractions are written as $\lambda \text{var}.
-\text{body}$.
-\\\noindent An \emph{application} of a function to an argument.
-Both the function and the argument are $\lambda$-calculus expressions.
-The application of the function $f$ to the argument $x$ is written as $f\; x$.
-
-For example, the function which squares the logarithm of a number is
-$\lambda x.\text{\texttt{square}} (\text{\texttt{log} } x)$, and the identity function $f(x) = x$ is $\lambda x.x$. The
-$\lambda$-calculus serves as a spartan but expressive Turing complete
-program representation, and distills the essential features of functional languages like Lisp.
-
-However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
-(i.e., $5 \; (\lambda x.x)$).
-We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
-The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
-Hindley-Milner has two important features:
-Feature 1: It supports \emph{parametric polymorphism}: meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
-For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
-A function that returns the first element of a list has the type $\texttt{list}(\alpha)\to\alpha$.
-Type variables are not the same has variables introduced by $\lambda$-abstractions.
-Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
-A detailed exposition of Hindley-Milner is beyond the scope of this work.
-
-%% \begin{figure}
-%%   \begin{align}
-%%     \text{primitive types} & 
-%%     \end{align}
-%%   \end{figure}
 
@@
 
+\pagebreak
+
+\section{Program Representation}
+We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
+A $\lambda$-calculus expression is either:
+\\\noindent A \emph{primitive}, like the number 5 or
+  the function \texttt{sum}.
+\\\noindent A \emph{variable}, like $x$, $y$, $z$
+\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function.
+$\lambda$-abstractions have a variable and a body.
+The body is a $\lambda$-calculus expression.
+Abstractions are written as $\lambda \text{var}.
+\text{body}$.
+\\\noindent An \emph{application} of a function to an argument.
+Both the function and the argument are $\lambda$-calculus expressions.
+The application of the function $f$ to the argument $x$ is written as $f\; x$.
+
+For example, the function which squares the logarithm of a number is
+$\lambda x.\text{\texttt{square}} (\text{\texttt{log} } x)$, and the identity function $f(x) = x$ is $\lambda x.x$. The
+$\lambda$-calculus serves as a spartan but expressive Turing complete
+program representation, and distills the essential features of functional languages like Lisp.
+
+However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
+(i.e., $5 \; (\lambda x.x)$).
+We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
+The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
+Hindley-Milner has two important features:
+Feature 1: It supports \emph{parametric polymorphism}: meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
+For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
+A function that returns the first element of a list has the type $\texttt{list}(\alpha)\to\alpha$.
+Type variables are not the same has variables introduced by $\lambda$-abstractions.
+Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
+A detailed exposition of Hindley-Milner is beyond the scope of this work.
+
+%% \begin{figure}
+%%   \begin{align}
+%%     \text{primitive types} & 
+%%     \end{align}
+%%   \end{figure}
+
 \section{Estimating $\theta$}

### 2018-02-01 paper.tex
minor edits
proposed the name Helmholtz hacker
better list function DSL table

@@
 
-\newcommand{\system}{EC2.0~}
+\newcommand{\system}{\textsc{HelmholtzHacker}~}
 
@@
 In the Appendix we derive a tractable MAP estimator for $\theta$.
-
 \begin{algorithm}[tb]
@@
 \end{figure*}
+
+\noindent\textbf{Putting it all together.}
+Algorithm~\ref{mainAlgorithm} describes how we combine the
+program search, recognition model training, and DSL induction.
+
+\begin{algorithm}[tb]
+   \caption{The \system Algorithm}
+   \label{mainAlgorithm}
+   \begin{algorithmic}
+     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
+     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
+     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
+     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
+     \FOR{$i=1$ {\bfseries to} $I$}
+     \FOR{$x:\tau\in X$}
+     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
+     \ENDFOR
+     \STATE $\mathcal{D}_i,\theta_i\gets $induceDSL$(\{\mathcal{F}_x\}_{x\in X})$
+     %% \STATE Define $Q_x(z) \propto \begin{cases}
+     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
+     %%   0&x\not \in \mathcal{F}_x
+     %% \end{cases}
+
+     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
+      \ENDFOR
+ \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
+\end{algorithmic}
+\end{algorithm}
 
@@
   \hline
-    empty & \texttt{tlist($\alpha$)} \\
-    singleton & \texttt{arrow($\alpha$, tlist($\alpha$))} \\
-    range & \texttt{int $\rightarrow$ tlist(int)} \\
-    concat & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
-    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\beta$)} \\
-    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\beta$} \\
+    empty & \texttt{[$\alpha$]} \\
+    singleton & \texttt{$\alpha$ $\rightarrow$ [$\alpha$]} \\
+    range & \texttt{int $\rightarrow$ [int]} \\
+    concat & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
+    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$]} \\
+    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$} \\
     \hline
@@
     is-square & \texttt{int $\rightarrow$ bool} \\
-    sort & \texttt{tlist(int) $\rightarrow$ tlist(int)} \\
+    sort & \texttt{[int] $\rightarrow$ [int]} \\
     \hline
-    sum & \texttt{tlist(int) $\rightarrow$ int} \\
-    reverse & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
-    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
-    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
-    index & \texttt{int $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\alpha$} \\
-    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
-    slice & \texttt{int $\rightarrow$ int $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
+    sum & \texttt{[int] $\rightarrow$ int} \\
+    reverse & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$]} \\
+    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
+    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
+    index & \texttt{int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$} \\
+    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
+    slice & \texttt{int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
   \hline
@@
 \section{Model}
-
-\begin{algorithm}[tb]
-   \caption{DSL Learner}
-   \label{mainAlgorithm}
-   \begin{algorithmic}
-     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-     \STATE \textbf{Hyperparameters:} Frontier size $F$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
-     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
-     \FOR{$i=1$ {\bfseries to} $I$}
-     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
-     \ENDFOR
-     \STATE $\mathcal{D}_i,\theta_i\gets $induceGrammar$(\{\mathcal{F}_x\}_{x\in X})$
-     \STATE Define $Q_x(z) \propto \begin{cases}
-       \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-       0&x\not \in \mathcal{F}_x
-     \end{cases}$
-     \STATE $q_i\gets \argmin_q \sum_{x\in X}\text{KL}(Q_x(\cdot )||\probability[\cdot |\mathcal{D}_i,q(x)])$
-      \ENDFOR
- \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
-\end{algorithmic}
-\end{algorithm}
 

### 2018-02-01 paper.tex
tiny dsl table change

@@
     singleton & \texttt{$\alpha$ $\rightarrow$ [$\alpha$]} \\
-    range & \texttt{int $\rightarrow$ [int]} \\
     concat & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
@@
     is-square & \texttt{int $\rightarrow$ bool} \\
+    range & \texttt{int $\rightarrow$ [int]} \\
     sort & \texttt{[int] $\rightarrow$ [int]} \\

### 2018-02-02 paper.tex
new figure explaining how the different components bootstrap off of each other

@@
 %\usetikzlibrary{arrows.meta}
-\usetikzlibrary{positioning}
-%\usetikzlibrary{decorations.text,decorations.pathreplacing}
+%\usetikzlibrary{positioning}
+%\usetikzlibrary{decorations.text}
 %\usetikzlibrary{decorations.pathmorphing}
@@
   \draw [->,dashed] (tx.west) -- (z.east);
+
+  \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](DSL) at (6,4) {DSL: $\mathcal{D}$};
+    \node[align = center](synthesis) at (4,2) {Search for \\programs: $p$};
+    \node[ align = center](recognitionModel) at (6,0) {Recognition \\model: $q$};
+
+    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
+    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
+    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
+    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
+    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
+  \end{scope}
   \end{tikzpicture}
@@
 
-\noindent\textbf{Putting it all together.}
+\subsection{Implementing the \system}
+
 Algorithm~\ref{mainAlgorithm} describes how we combine the
 program search, recognition model training, and DSL induction.
+We additionally make the following implementation details:
+(1) On the first iteration we do \emph{not} train the recognition model on samples from
+the generative model because, on the first iteration, we have not yet learned a
+new DSL, so we instead trained the network to only maximize $\mathcal{L}_{\text{AE}}$;
+(2) Because the frontiers can grow very large,
+we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
+with the highest $\probability[x,p|\mathcal{D},\theta]$;
+(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
+by only summing over the top $K$
+programs in $\mathcal{F}_x$.
+We found that $K = 2$ sufficed.
 
@@
      \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
-     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
+     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
+     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
      \FOR{$i=1$ {\bfseries to} $I$}
-     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
-     \ENDFOR
-     \STATE $\mathcal{D}_i,\theta_i\gets $induceDSL$(\{\mathcal{F}_x\}_{x\in X})$
+%     \FOR{$x:\tau\in X$}
+     \STATE  $\mathcal{F}^{(\theta)}_x\gets \{z| z\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|z] > 0\}$
+ %    \ENDFOR
+     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
+     \STATE  $\mathcal{F}^{(q)}_x\gets\{z|z\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|z] > 0\}$
+     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{(\theta)}_x\cup\mathcal{F}^{(q)}_x\}_{x\in X})$
      %% \STATE Define $Q_x(z) \propto \begin{cases}
@@
      %% \end{cases}
-
-     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
       \ENDFOR
- \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
+ \STATE \textbf{return} $\mathcal{D},\theta,q$
 \end{algorithmic}

### 2018-02-02 paper.tex
got rid of first paragraph

@@
 
-Imagine an agent faced with a suite of new problems totally different
-from anything it has seen before. It has at its disposal a basic set
-of primitive actions it can compose to build solutions to these problems, but
-it is no idea what kinds of primitives are appropriate for which
-problems nor does it know the higher-level vocabulary in
-which solutions are best expressed.
-How can our agent get off the ground?
+%% Imagine an agent faced with a suite of new problems totally different
+%% from anything it has seen before. It has at its disposal a basic set
+%% of primitive actions it can compose to build solutions to these problems, but
+%% it is no idea what kinds of primitives are appropriate for which
+%% problems nor does it know the higher-level vocabulary in
+%% which solutions are best expressed.
+%% How can our agent get off the ground?
 

### 2018-02-02 paper.tex
edits

--- 2018-02-02 paper.tex
@@
 
-Imagine an agent faced with a suite of new problems totally different
-from anything it has seen before. It has at its disposal a basic set
-of primitive actions it can compose to build solutions to these problems, but
-it is no idea what kinds of primitives are appropriate for which
-problems nor does it know the higher-level vocabulary in
-which solutions are best expressed.
-How can our agent get off the ground?
+Imagine an agent faced with a new domain of unfamiliar tasks: think
+playing new kinds of videogames, authoring Python code, or navigating
+through mazes.  Our agent has at its disposal a basis of primitive
+actions it can compose to propose solutions to these tasks, but it is
+no idea what kinds of primitives are appropriate for which tasks
+nor does it know the higher-level vocabulary in which solutions are
+best expressed.  How can our agent learn to solve problems in this new
+domain?
 
@@
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
-We show that \system can construct a set of basis primitives suitable for discovering solutions in each of these domains
+Because we want a domain-general
+approach,
+we model solutions to tasks as
+programs,
+which frames our problem as one of program induction.
+
+Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}: solid lines):
+A DSL $\mathcal{D}$
+probabilistically generates a latent program $p$
+which generates the agent's observations, $x$.
+In tandem we learn a learn a neural network $q$ that regresses
+from observations to programs.
 
@@
 we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
-
-Our key observation is that the generative and recognition models can bootstrap off of each other,
-greatly increasing the tractability of BPL.
 
@@
   \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
-We observe program outputs $x$.
-A neural network $q(\cdot )$ called the \emph{recognition model} regresses from program outputs to a distribution over programs ($\theta^{(x)} = q(x)$).
+We observe tasks $x$, which are program inputs/outputs.
+A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs ($\theta^{(x)} = q(x)$).
 Solid arrows correspond to the top-down generative model.
@@
 
-The goal of \system is to both induce a DSL and find good programs solving each of the tasks.
+Our goal in this work is to both induce a DSL and find good programs solving each of the tasks.
 Our strategy is to iterate through three steps: (1) searching for programs that
@@
 to accelerate the search over programs -- and (3) improving the DSL.
-The key observation here is that each of these three steps can bootstrap off of each other:
+The key observation here is that each of these three steps can bootstrap off each other:
 
@@
 \end{itemize}
+
+The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
+learning a generative model and training a recognition model on samples from the generative model.
+We apply this same idea to the problem of program induction,
+and so we call our algorithm the \system.
 

### 2018-02-02 paper.tex
lambda calculus description

@@
 A DSL $\mathcal{D}$
-probabilistically generates a latent program $p$
-which generates the agent's observations, $x$.
+samples a latent program $p$
+which determines the agent's observations, $x$.
+This makes our work in instance of
+ \emph{Bayesian Program
+  Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
 In tandem we learn a learn a neural network $q$ that regresses
 from observations to programs.
+We will show that the interaction of
+the generative and recognition models is important for solving these program induction problems.
 
-We cast these problems as \emph{Bayesian Program
-  Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
-where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
-A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
-We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
-define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
-In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
-where the likelihood $\probability[x|p]$ is domain-dependent.
-The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
-Alongside this generative model,
-we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
+%% We cast these problems as \emph{Bayesian Program
+%%   Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
+%% where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
+%% A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
+%% We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
+%% define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
+%% In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
+%% where the likelihood $\probability[x|p]$ is domain-dependent.
+%% The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
+%% Alongside this generative model,
+%% we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
 
@@
 
-  \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](DSL) at (6,4) {DSL: $\mathcal{D}$};
-    \node[align = center](synthesis) at (4,2) {Search for \\programs: $p$};
-    \node[ align = center](recognitionModel) at (6,0) {Recognition \\model: $q$};
-
-    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
-    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
-    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
-    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
-    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
-  \end{scope}
   \end{tikzpicture}
@@
 \begin{itemize}
-  \item Searching for programs: Program search uses a distribution determined by the neural recognition model --
-  so the recognition model bootstraps the search process.
+\item Searching for programs: Our program search  is informed by both the DSL and the recognition model.
+When these improve we find better programs.
   \item Learning a recognition model: The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
-As the DSL improves and we find more programs, the recognition model gets both more data to train on and better data.
+As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
   \item Improving the DSL: We induce a DSL from the programs we have found so far which solve the tasks;
@@
 and so we call our algorithm the \system.
+Figure~\ref{feeding} diagrams how these three steps feed off of each other.
+\begin{figure}[b]\centering
+  \begin{tikzpicture}
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](DSL) at (6,4) {DSL: $\mathcal{D}$};
+    \node[align = center](synthesis) at (4,2) {Search for \\programs: $p$};
+    \node[ align = center](recognitionModel) at (6,0) {Recognition \\model: $q$};
+
+    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
+    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
+    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
+    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
+    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
+  \end{scope}
+    \end{tikzpicture}
+\caption{Our algorithm alternatingly solves for the programs, DSL, and recognition model. Each of these steps bootstraps off of the others.}  \label{feeding}
+  \end{figure}
 
@@
 
-\noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
 is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
@@
 what form $\probability[p |\mathcal{D},\theta]$ takes.
-In this work, we represent programs as \emph{polymorphicly-typed $\lambda$-calculus expressions}.
-$\lambda$-calculus is a formalism for expressing functional programs.
-It includes variables, function application, and the ability to create new functions using ...
+In this work, we represent programs as $\lambda$-calculus expressions.
+$\lambda$-calculus is a formalism for expressing functional programs
+that closely resembles Lisp.
+$\lambda$-calculus includes variables, function application, and the ability to create new functions.
+Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
+Our programs are all strongly typed.
+We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is used in functional languages like OCaml.
+Variables that range over types are always written using lowercase Greek letters
+and we write $\alpha\to \beta$ to mean a function that takes as input type $\alpha$
+and returns something of type $\beta$.
+We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
+has the type $\tau$.
+For example, to describe the type of the identity function
+we would say \texttt{(lambda (x) x)}$:\alpha\to \alpha$.
+With this notation in hand we can now define DSLs:
 
-TODO: summarize lambda calculus and types in one paragraph
-
-\noindent\textbf{Definition: $\mathcal{D}$.}
+\noindent\textbf{Definition.}
 A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-\\\noindent\textbf{Definition: $\theta$.}
+\\\noindent\textbf{Definition.}
 A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:

### 2018-02-02 paper.tex
edits

@@
 Figure~\ref{feeding} diagrams how these three steps feed off of each other.
-\begin{figure}[b]\centering
+\begin{figure}\centering
   \begin{tikzpicture}
@@
   \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
-   \STATE \textbf{return} $\lambda \text{var}.$ body
+   \STATE \textbf{return} \texttt{(lambda (}var\texttt{) }body\texttt{)}
    \ENDIF
@@
    \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
-   \STATE Let $e:\alpha_1\to\alpha_2\to\cdots\to \alpha_K\to\beta$. Unify $\tau$ with $\beta$.
+   \STATE Unify $\tau$ with yield$(\tau')$.
+   \STATE $\left\{\alpha_k \right\}_{k\leq K}\gets\text{argTypes}(\tau')$
 %   \STATE unify$(\tau,\beta)$
@@
  \ENDFOR
- \STATE \textbf{return} $e\;a_1\; a_2\; \cdots\; a_K$
- \STATE
- \STATE \textbf{function} yield$(\tau)$:
- \STATE \textbf{Input:} type $\tau$
- \STATE \textbf{Output:} the return type of $\tau$
- \STATE \textbf{if} $\tau = \alpha\to \beta$ \textbf{then return }yield$(\beta)$ \textbf{else return }$\tau$
+ \STATE \textbf{return} \texttt{(}$e\;a_1\; a_2\; \cdots\; a_K$\texttt{)}
+ \STATE\textbf{where:}
+ \STATE  yield$(\tau) = \begin{cases}
+\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
+\tau   &\text{ otherwise.}
+ \end{cases}$ 
+ \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
+[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
+[]   &\text{ otherwise.}
+ \end{cases}$
 \end{algorithmic}
@@
 The crucial aspect of this framing is that the neural network
-can leverage the structure of the DSL,
-and is \emph{not} responsible for
+leverages the structure of the learned DSL,
+and so is \emph{not} responsible for
 generating programs wholesale.
-We will show that this lets us get away with a simple, low-capacity neural network.
 
@@
 but also so that the DSL might expose new abstractions
-which will later used to
+which will later be used to
 discover even better programs.
@@
 We found that $K = 2$ sufficed.
+(4) For added robustness, at each iteration we enumerate programs from both the generative model and the recognition model.
 

### 2018-02-02 paper.tex
ec references and start of circuit section

@@
  \emph{Bayesian Program
-  Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
+  Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
 In tandem we learn a learn a neural network $q$ that regresses
@@
 
+TODO: talk explicitly about what ideas we are borrowing from EC
 %% We cast these problems as \emph{Bayesian Program
@@
 \subsection{Boolean circuits}
-pedagogical example; easy domain
+As a toy warm-up domain, we consider the problem of learning
+to build the Boolean circuits out of logic gates.
+Now, if we equipped our agent with a full repertoire of
+circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
+and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
+Instead, we gave \system only a NAND gate,
+and tasked it with building a set of 500 random circuits made out of
+AND, NOT, and OR gates.
+This experimental setup was introduced in the original EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
+
+Figure~\ref{circuitLearningCurve}
+
 \begin{figure}
   \includegraphics[width = \columnwidth]{figures/circuit.png}
+  \caption{\system learning to build Boolean circuits. We start out with only NAND gates and learn DSL that contains ...}\label{circuitLearningCurve}
 \end{figure}

### 2018-02-03 paper.tex
text primitives
started writing circuit section
vectorize mlp; changes to intro; small edits

@@
 
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}
+
+  \node[latent] at (1,3) (d){$\mathcal{D}$};
+  \node[latent] at (2.5,3) (t){$\theta$};
+  \node[latent] at (1,1.5) (z){$p$};
+  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+  \node[obs] at (1,-0.5) (x) {$x$};
+  \edge {z}{x};
+  \edge {d,t}{z};
+  \plate {}{(tx)(z)(x)}{$x\in X$};
+  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn){} (tx.south);
+  \draw [->,dashed] (tx.west) -- (z.east);
+
+  \node at (nn) {
+    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle,width=2cm] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}
+  };
+  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
+
+  \end{tikzpicture}
+  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
+We observe tasks $x$, which are program inputs/outputs.
+A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs ($\theta^{(x)} = q(x)$).
+Solid arrows correspond to the top-down generative model.
+Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
+\end{figure}
+
 Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}: solid lines):
 A DSL $\mathcal{D}$
-samples a latent program $p$
-which determines the agent's observations, $x$.
+determines a distribution with production probabilities $\theta$,
+from which a latent program $p$ is sampled,
+which finally determines the agent's observations, $x$.
 This makes our work in instance of
@@
   Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
-In tandem we learn a learn a neural network $q$ that regresses
-from observations to programs.
-We will show that the interaction of
-the generative and recognition models is important for solving these program induction problems.
+In tandem we learn a neural network $q$ that regresses from observations to
+programs by proposing a distribution over programs $\theta^{(x)} = q(x)$.
+We will show that the interaction of the generative and recognition models
+is important for solving these program induction problems.
 
-TODO: talk explicitly about what ideas we are borrowing from EC
+TODO: talk about existing works as special cases of, and points of inspiration for, \system:
+\begin{itemize}
+  \item Maximum-likelihood estimation of a PCFG \citep{DBLP:conf/icml/LiangJK10} with
+    production probabilities $\theta$, a fixed the DSL $\mathcal{D}$ and no neural network
+    $q$;
+    % although some of the Liang paper actually learns subexpressions and therefore DSL
+  \item The EC algorithm \citep{Dechter:2013:BLV:2540128.2540316} has no recognition model
+    $q$;
+    % because this is a primary inspiration, might be worth further mention:
+    % different program representation (routing vs. substitution; simple vs polymorphic types);
+    % different grammar induction (sequitur vs. fragment grammar)
+  \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
+    DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
+    network $q$.
+    % although RobustFill has very different enumeration
+\end{itemize}
+
 %% We cast these problems as \emph{Bayesian Program
@@
 %% we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
-
-\begin{figure}
-  \begin{tikzpicture}
-
-  \node[latent] at (1,3) (d){$\mathcal{D}$};
-  \node[latent] at (2.5,3) (t){$\theta$};
-  \node[latent] at (1,1.5) (z){$p$};
-  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-  \node[obs] at (1,-0.5) (x) {$x$};
-  \edge {z}{x};
-  \edge {d,t}{z};
-  \plate {}{(tx)(z)(x)}{$x\in X$};
-  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn)[fill = white]{
-          \includegraphics[width = 1.5cm]{figures/MLP.png} 
-  } (tx.south);
-  \node at (nn.south) {$q(x)$};
-  \draw [->,dashed] (d.west) to[out = 180,in = 180] (z.west);
-  \draw [->,dashed] (tx.west) -- (z.east);
-
-  \end{tikzpicture}
-  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
-We observe tasks $x$, which are program inputs/outputs.
-A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs ($\theta^{(x)} = q(x)$).
-Solid arrows correspond to the top-down generative model.
-Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
-\end{figure}
 
@@
 In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
-Conventional NLP approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
+Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
 \begin{align}
@@
 
-We maximize $J$ by alternatingly maximizing it w.r.t. the DSL and the frontiers:
+We maximize $J$ by alternating maximization w.r.t. the DSL and the frontiers:
 %So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
@@
 This experimental setup was introduced in the original EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
-
-Figure~\ref{circuitLearningCurve}
+For this domain our recognition model is a simple multilayer perceptron
+whose inputs are the input/output examples.
+Figure~\ref{circuitLearningCurve} shows our algorithm learning to build circuits; Table~\ref{circuitExperimentTable}
 
@@
 \end{figure}
+
+\begin{table}
+\begin{tabular}{ll}
+  \toprule
+  Primitives&\texttt{nand}
+
+\\\midrule
+  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
+  \midrule
+  Likelihood $\probability[x|p]$& $\prod_n \indicator[p(i_n) = o_n]$
+  \\\midrule
+  \begin{tabular}{l}
+    Subset of \\Learned DSL
+    \end{tabular}&
+
+  \\\bottomrule
+\end{tabular}
+\caption{}\label{circuitExperimentTable}
+\end{table}
 
@@
 \subsection{String editing}\label{textSection}
+
+\begin{tabular}{ll}
+  \toprule
+  Primitives&\begin{tabular}{l}
+    \texttt{0}, \texttt{+1},\texttt{-1}\\
+    \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
+    \texttt{append}, \texttt{chr->str}\\
+    \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
+    \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}
+    \end{tabular}
+
+\\\midrule
+  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
+  \midrule
+  Likelihood $\probability[x|p]$& $\prod_n \indicator[p(i_n) = o_n]$
+  \\\midrule
+  \begin{tabular}{l}
+    Subset of \\Learned DSL
+    \end{tabular}&
+
+  \\\bottomrule
+  \end{tabular}
 \begin{figure}

### 2018-02-05 paper.tex
citation changes

@@
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
-Because we want a domain-general
-approach,
-we model solutions to tasks as
+%Because we want a domain-general
+%approach,
+We model solutions to tasks as
 programs,
@@
 We observe tasks $x$, which are program inputs/outputs.
-A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs ($\theta^{(x)} = q(x)$).
-Solid arrows correspond to the top-down generative model.
-Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
+A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs
+    ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
 \end{figure}
@@
 \begin{itemize}
-  \item Maximum-likelihood estimation of a PCFG \citep{DBLP:conf/icml/LiangJK10} with
-    production probabilities $\theta$, a fixed the DSL $\mathcal{D}$ and no neural network
-    $q$;
-    % although some of the Liang paper actually learns subexpressions and therefore DSL
-  \item The EC algorithm \citep{Dechter:2013:BLV:2540128.2540316} has no recognition model
+  \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
+  \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
     $q$;
@@
 This approach implements an amortized
-inference scheme~\cite{ritchie2016deep} for the generative model in
+inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
 Fig.~\ref{graphicalModel}.

### 2018-02-05 paper.tex
improved overview figure

@@
     \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](DSL) at (6,4) {DSL: $\mathcal{D}$};
-    \node[align = center](synthesis) at (4,2) {Search for \\programs: $p$};
-    \node[ align = center](recognitionModel) at (6,0) {Recognition \\model: $q$};
+    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
+    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
 
-    \draw [->,thick] (DSL.west) to[out = 180,in = 90] node[above,rotate = 45]{{\footnotesize Makes tractable}} (synthesis.north);
-    \draw [->,thick] (synthesis.east) to[out = 0,in = -90] node[above,rotate = 70]{{\footnotesize Trains}} (DSL.south);
-    \draw [->,thick] (synthesis.east) to[out = 0,in = 90] node[below,rotate = -60]{{\footnotesize Trains}} (recognitionModel.north);
-    \draw [->,thick] (recognitionModel.west) to[out = 180,in = -90] node[below,rotate = -70]{{\footnotesize Makes tractable}} (synthesis.south);
-    \draw [->,thick] (DSL.east) to[out = 0,in = 0] node[above,rotate = -90]{{\footnotesize Trains}} (recognitionModel.east);
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45]{{\footnotesize Trains}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45]{{\footnotesize Trains}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above]{{\footnotesize Trains}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.east);
   \end{scope}
     \end{tikzpicture}
-\caption{Our algorithm alternatingly solves for the programs, DSL, and recognition model. Each of these steps bootstraps off of the others.}  \label{feeding}
+\caption{Our algorithm solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others.}  \label{feeding}
   \end{figure}

### 2018-02-05 paper.tex
introduction rewrite

@@
 
-Imagine an agent faced with a new domain of unfamiliar tasks: think
-playing new kinds of videogames, authoring Python code, or navigating
-through mazes.  Our agent has at its disposal a basis of primitive
-actions it can compose to propose solutions to these tasks, but it is
-no idea what kinds of primitives are appropriate for which tasks
-nor does it know the higher-level vocabulary in which solutions are
-best expressed.  How can our agent learn to solve problems in this new
-domain?
+Automatically inducing programs from examples is a long-standing goal
+of artificial intelligence.  Recent work in program induction show
+that deep networks can be very successful at learning to write certain
+kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
+DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
+techniques (e.g., Metagol:~\cite{muggleton2015meta} and
+FlashFill:~\cite{gulwani2011automating}).
+However, the success of these approaches -- both symbolic and neural --
+relies critically upon a hand-engineered \emph{Domain Specific Language} (DSL).
+The DSL is an inventory of restricted programming primitives,
+crafted by an engineer,
+which imparts domain specific knowledge about the space of programs.
+To what extent can we dispense with hand-engineered DSLs?
 
-The AI and machine learning literature contains two broad takes on this problem.
-The first take is that the agent should come up with a better representation of the space of solutions,
-for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
-The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
-Our contribution is a general algorithm for fusing these two takes on the problem:
-we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
-alongside a bottom-up discriminative model that regresses from problems to solutions.
-We evaluate our algorithm on four domains:
-building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
-%Because we want a domain-general
-%approach,
-We model solutions to tasks as
-programs,
-which frames our problem as one of program induction.
+This work proposes \emph{learning} DSLs.
+We consider the problem of solving many related program induction tasks,
+starting with a Lisp-like representation.
+We introduce an algorithm, called the \system,
+which creates its own DSL by discovering and then reusing
+new useful programming idioms and subroutines.
+
+%% Imagine an agent faced with a new domain of unfamiliar tasks: think
+%% playing new kinds of videogames, authoring Python code, or navigating
+%% through mazes.  Our agent has at its disposal a basis of primitive
+%% actions it can compose to propose solutions to these tasks, but it is
+%% no idea what kinds of primitives are appropriate for which tasks
+%% nor does it know the higher-level vocabulary in which solutions are
+%% best expressed.  How can our agent learn to solve problems in this new
+%% domain?
+
+%% The AI and machine learning literature contains two broad takes on this problem.
+%% The first take is that the agent should come up with a better representation of the space of solutions,
+%% for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
+%% The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
+%% Our contribution is a general algorithm for fusing these two takes on the problem:
+%% we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
+%% alongside a bottom-up discriminative model that regresses from problems to solutions.
+%% We evaluate our algorithm on four domains:
+%% building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
+%% %Because we want a domain-general
+%% %approach,
+%% We model solutions to tasks as
+%% programs,
+%% which frames our problem as one of program induction.
 
@@
 
-  \node[latent] at (1,3) (d){$\mathcal{D}$};
-  \node[latent] at (2.5,3) (t){$\theta$};
-  \node[latent] at (1,1.5) (z){$p$};
-  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-  \node[obs] at (1,-0.5) (x) {$x$};
+  \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+  \node[latent] at (1.5,3) (t){$\theta$};
+  \node[latent] at (1,1.75) (z){$p$};
+%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+  \node[obs] at (1,0) (x) {$x$};
   \edge {z}{x};
   \edge {d,t}{z};
-  \plate {}{(tx)(z)(x)}{$x\in X$};
-  \draw [->,dashed] (x.east) to[out = 0,in = -90] node(nn){} (tx.south);
-  \draw [->,dashed] (tx.west) -- (z.east);
+  \plate {}{(z)(x)}{$x\in X$};
+  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative \\model};
+
+  \begin{scope}[shift = {(-1,0)}]  
+  \node[latent] at (3.5,3) (dx){$\mathcal{D}$};
+  \node[latent] at (3.5,1.75) (zp){$p$};
+  \node[latent] at (5.5,1.75) (tx){$\theta^{(x)}$};
+  \node[obs] at (3.5,0) (xp) {$x$};
+  \plate {}{(zp)(xp)(tx)}{$x\in X$};
+  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
+  \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
+  \draw [->,red] (tx.west) -- (zp.east);
+  \draw [->,red] (dx.south) -- (zp.north);
 
@@
 
+  \end{scope}
+
   \end{tikzpicture}
-  \caption{DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
-We observe tasks $x$, which are program inputs/outputs.
-A neural network $q(\cdot )$ called the \emph{recognition model} regresses from $x$ to a distribution over programs
-    ($\theta^{(x)} = q(x)$). Solid arrows correspond to the top-down generative model. Dashed arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
+  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
+Observe tasks $x$ (program inputs/outputs).
+(b)  neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to a distribution over programs
+    ($\theta^{(x)} = q(x)$). Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
 \end{figure}
 
-Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}: solid lines):
-A DSL $\mathcal{D}$
-determines a distribution with production probabilities $\theta$,
-from which a latent program $p$ is sampled,
-which finally determines the agent's observations, $x$.
+Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}(a): black lines):
+A DSL $\mathcal{D}$, equipped with a real-valued weight vector $\theta$,
+determine a distribution over programs
+from which a latent program is sampled,
+which finally determines the agent's observations.
+Our goal will be both to infer the programs that best explain our observations,
+and at the same time induce a DSL
+that distills the commonalities across all of the programs that solve the tasks.
+Inducing the DSL means learning the symbolic structure of a latent programming language,
+and so we are confronted with two distinct structure learning problems:
+the structure of the programs and the structure of the DSL.
+
+Inference in the generative model in Figure~\ref{graphicalModel}(a) is
+difficult because of the vastness of the combinatorial space of all
+programs. To make inference tractable our model learns a bottom-up
+\emph{recognition model} alongside the top-down generative model
+(Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
+network that regresses from observations to a distribution over
+programs.  The neural network $q$ implements an amortized
+inference/compiled inference
+scheme~\cite{le2016inference,ritchie2016deep} for the generative
+model.
+
+How should we perform inference for the model
+in~\ref{graphicalModel}(a)?  An intuitive strategy is to exploit the
+conditional independence of the programs conditioned on the DSL, and
+so to alternate between estimating the DSL ($\mathcal{D},\theta$) and
+estimating a posterior distribution over the programs by searching for
+programs written using the current DSL.  In practice searching the
+space of programs is difficult, so we use the recognition model to
+learn how to efficiently search for programs in the current DSL.
+
 This makes our work in instance of
@@
   Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
-In tandem we learn a neural network $q$ that regresses from observations to
-programs by proposing a distribution over programs $\theta^{(x)} = q(x)$.
-We will show that the interaction of the generative and recognition models
-is important for solving these program induction problems.
+ Alongside this generative model, our algorithm learns a bottom-up recognition model (Figure~\ref{graphicalModel}(b)).
+ The recognition model is a neural network $q$ that regresses from programming tasks $x$
+ to a distribution over programs.
 

### 2018-02-05 paper.tex
more introductions

@@
 How should we perform inference for the model
-in~\ref{graphicalModel}(a)?  An intuitive strategy is to exploit the
+in~\ref{graphicalModel}(a)?  We take the intuitive strategy of exploiting the
 conditional independence of the programs conditioned on the DSL, and
-so to alternate between estimating the DSL ($\mathcal{D},\theta$) and
-estimating a posterior distribution over the programs by searching for
-programs written using the current DSL.  In practice searching the
+so alternate between estimating the DSL ($\mathcal{D},\theta$) and
+estimating a posterior distribution over $p$ by searching for
+programs written using $\mathcal{D}$.  In practice searching the
 space of programs is difficult, so we use the recognition model to
-learn how to efficiently search for programs in the current DSL.
+learn how to efficiently search for programs.
+
+We apply \system to four domains:
+ building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
+ For each of these we deliberately provide an impoverished
+ set of programming primitives,
+ and show that our algorithm discovers
+ its own domain-specific vocabulary for expressing solutions in each of these domains (Table~\ref{initialExampleDSL}).
+ \begin{table}
+    \begin{tabular}{ll}
+   \toprule
+   Domain&Subset of induced DSL\\\midrule
+   \\\bottomrule
+ \end{tabular}
+   \end{table}
 

### 2018-02-05 paper.tex
minor edits; briefly explain unification

@@
 learning a generative model and training a recognition model on samples from the generative model.
-We apply this same idea to the problem of program induction,
-and so we call our algorithm the \system.
+We apply this idea to program induction, motivating why we call our algorithm the \system.
 Figure~\ref{feeding} diagrams how these three steps feed off of each other.
@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}. We wish to maximize the MAP probability of $\mathcal{D}$:
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}. We wish to maximize the posterior probability of $\mathcal{D}$:
 \begin{equation*}
@@
 
-We maximize $J$ by alternating maximization w.r.t. the DSL and the frontiers:
+We maximize $J$ by alternating maximization w.r.t. the frontiers and the DSL:
 %So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
@@
 \\\noindent \textbf{DSL Induction: Maxing $J$ w.r.t.
-the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed and so we can evaluate $J$.
+the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $J$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
@@
 Searching for programs is extremely difficult because
-of how large the search space is. We ease the difficulty of the search by learning a neural recognition model:
+of the large search space. We ease this difficulty by learning a neural recognition model:
 
@@
 $\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
-frontier $\mathcal{F}_x$ by searching for programs maximizing
+the frontier of a task $\mathcal{F}_x$ by searching for programs that maximize
 $q(p|x)$.
@@
 inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
-Fig.~\ref{graphicalModel}.
+Figure~\ref{graphicalModel}.
 
@@
 what form $\probability[p |\mathcal{D},\theta]$ takes.
-In this work, we represent programs as $\lambda$-calculus expressions.
+We represent programs as $\lambda$-calculus expressions.
 $\lambda$-calculus is a formalism for expressing functional programs
-that closely resembles Lisp.
+that closely resembles the Lisp programming language.
 $\lambda$-calculus includes variables, function application, and the ability to create new functions.
@@
 we would say \texttt{(lambda (x) x)}$:\alpha\to \alpha$.
+We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
+$p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
+a type $\alpha$ with $\tau$ is to introduce constraints on the type
+variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
+See Section~\ref{programrepresentation} for more detail on program representation.
 With this notation in hand we can now define DSLs:
@@
 
-\section{Program Representation}
+\section{Program Representation}\label{programrepresentation}
 We choose to represent programs using $\lambda$-calculus~\cite{pierce}.

### 2018-02-05 paper.tex
related work

@@
 However, the success of these approaches -- both symbolic and neural --
-relies critically upon a hand-engineered \emph{Domain Specific Language} (DSL).
+relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
 The DSL is an inventory of restricted programming primitives,
-crafted by an engineer,
-which imparts domain specific knowledge about the space of programs.
+that imparts domain specific knowledge about the space of programs.
 To what extent can we dispense with hand-engineered DSLs?
@@
 A DSL $\mathcal{D}$, equipped with a real-valued weight vector $\theta$,
-determine a distribution over programs
+give a distribution over programs
 from which a latent program is sampled,
-which finally determines the agent's observations.
+which then determines the agent's observations.
 Our goal will be both to infer the programs that best explain our observations,
@@
 estimating a posterior distribution over $p$ by searching for
-programs written using $\mathcal{D}$.  In practice searching the
-space of programs is difficult, so we use the recognition model to
-learn how to efficiently search for programs.
+programs written using $\mathcal{D}$.  In practice searching for
+programs is difficult, so we use the recognition model to
+learn how to search efficiently.
 
@@
    \toprule
-   Domain&Subset of induced DSL\\\midrule
+   Domain&Part of the learned DSL\\\midrule
+   Circuits&\texttt{(nand (nand a a) (nand b b))}\\
+   &\hspace{0.5cm} \emph{(an OR gate)}\\
+   Regression& \texttt{(+ (* real x) real)} \\
+      &\hspace{0.5cm} \emph{(a linear function of x)}\\
+   Strings& \texttt{(join '' (split c s))}\\
+   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
+   Lists&
    \\\bottomrule
     \end{tabular}
+    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful fragments of code.}\label{initialExampleDSL}
  \end{table}
 
-This makes our work in instance of
+ \subsection{Related Work}
+ Our work is far from the first to learn to learn programs,
+ an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:
+
+ \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
+ focused on engineering neural networks that regress from
+ input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to a lesioned version of
+ \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned.
+
+ \noindent \emph{Predicate Invention in ILP:} The Inductive Logic Programming (ILP) community
+ has long worked on the problem of inventing new primitive predicates~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}, which is analogous to our model with $\theta$ held fixed and $q$ ablated.
+
+ \noindent \emph{Inventing new subroutines for program induction:}
+ Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting.
+We find this work inspiring and motivating,
+ and go beyond it along two dimensions: (1) we propose more sophisticated algorithms for
+ inducing reusable structure, based on a formalism called \emph{Fragment Grammars}~\cite{tim};
+ and (2) we show how to combine these techniques with bottom-up neural recognition models.
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.
+
+Our work is an instance of
  \emph{Bayesian Program
   Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
- Alongside this generative model, our algorithm learns a bottom-up recognition model (Figure~\ref{graphicalModel}(b)).
- The recognition model is a neural network $q$ that regresses from programming tasks $x$
- to a distribution over programs.
-
-TODO: talk about existing works as special cases of, and points of inspiration for, \system:
-\begin{itemize}
-  \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
-  \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
-    $q$;
-    % because this is a primary inspiration, might be worth further mention:
-    % different program representation (routing vs. substitution; simple vs polymorphic types);
-    % different grammar induction (sequitur vs. fragment grammar)
-  \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
-    DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
-    network $q$.
-    % although RobustFill has very different enumeration
-\end{itemize}
+
+%% \begin{itemize}
+%%   \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
+%%   \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
+%%     $q$;
+%%     % because this is a primary inspiration, might be worth further mention:
+%%     % different program representation (routing vs. substitution; simple vs polymorphic types);
+%%     % different grammar induction (sequitur vs. fragment grammar)
+%%   \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
+%%     DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
+%%     network $q$.
+%%     % although RobustFill has very different enumeration
+%% \end{itemize}
 
@@
 
-Our goal in this work is to both induce a DSL and find good programs solving each of the tasks.
+Our goal  is to induce a DSL and find good programs solving each of the tasks.
 Our strategy is to iterate through three steps: (1) searching for programs that
@@
 
-\begin{itemize}
-\item Searching for programs: Our program search  is informed by both the DSL and the recognition model.
+\begin{description}
+\item[ Searching for programs:] Our program search  is informed by both the DSL and the recognition model.
 When these improve we find better programs.
-  \item Learning a recognition model: The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+  \item[ Learning a recognition model:] The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-  \item Improving the DSL: We induce a DSL from the programs we have found so far which solve the tasks;
+  \item[Improving the DSL:] We induce a DSL from the programs we have found so far which solve the tasks;
   as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
-\end{itemize}
+\end{description}
 
@@
 Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL $\mathcal{D}$ that distills the commonalities across all of the programs that solve the tasks.
+and also to infer a DSL $\mathcal{D}$.
 

### 2018-02-05 paper.tex
add lists DSL example

@@
    &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
-   Lists&
+   Lists& \texttt{(map (lambda (x) (+ x a)) l)}\\
+   &\hspace{0.5cm} \emph{(add a to every element of list l)}
    \\\bottomrule

### 2018-02-05 paper.tex
attempt to redo the lower bound section

@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}. We wish to maximize the posterior probability of $\mathcal{D}$:
-\begin{equation*}
-  \probability[\mathcal{D}|X]\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]
-\end{equation*}
-In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
-Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}(a), finding the $\mathcal{D}^*,\theta^*$ solving:
+\begin{align}\label{intractableObjectives}
+  \nonumber\probability[X|\mathcal{D},\theta] &= \prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  \nonumber\probability[\mathcal{D}|X]&\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\probability[X|\mathcal{D},\theta]\\
+  \mathcal{D}^* &= \argmax_{\mathcal{D}}\probability[\mathcal{D}|X]\\
+\nonumber  \theta^*& =\argmax_\theta P(\theta |\mathcal{D}^*)\probability[X|\mathcal{D}^*,\theta]
+\end{align}
+The above equations summarize the problem from the point of view of an ideal Bayesian learner.
+However, Eq.~\ref{intractableObjectives}
+is wildly intractable because evaluating $\probability[X|\mathcal{D},\theta]$ involves
+summing over the  infinite set of all possible programs.
+In practice we will only ever be able to sum over a finite set of programs.
+So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
+\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
+
+With the frontiers in hand, we can define the following intuitive lower bound on the likelihood of the tasks, which we call $\mathcal{L}$:
 \begin{align}
-  A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
-&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
+  \probability[X|\mathcal{D},\theta]\geq J\triangleq\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
 \end{align}
-
-If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{AIC}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
-Through this lens we now take as our goal to maximize Eq.~\ref{AIC}.
-But even \emph{evaluating} Eq.~\ref{AIC} is intractable because it involves summing over the infinite set of all possible programs.
+along with a corresponding lower bound on the marginal likelihood of a DSL, which we call $$
+\begin{align}
+  \probability[\mathcal{D}|X]\geq 
+  \end{align}
+
+If we had a $(\mathcal{D},\theta)$ solving Eq.~\ref{intractableObjectives}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
+Through this lens we now take as our goal to solve Eq.~\ref{intractableObjectives}.
+But even \emph{evaluating} Eq.~\ref{intractableObjectives} is intractable because it involves summing over the infinite set of all possible programs, as an ideal Bayesian learner would.
+In practice, we must instead marginalize over
+some finite set of programs.
+
 In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
+
+Now in theory we would like to sum over the entire infinite space
+of all programs -- but this is of course impossible.
+
+%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
+%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
+%% \begin{align}
+%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
+%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
+%%   \end{align}
+
 With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{AIC},
@@
 This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:
-
-\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
 

### 2018-02-05 paper.tex
moved to stuff to supplement

@@
 
-\pagebreak
-
-\section{Program Representation}\label{programrepresentation}
-We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
-A $\lambda$-calculus expression is either:
-\\\noindent A \emph{primitive}, like the number 5 or
-  the function \texttt{sum}.
-\\\noindent A \emph{variable}, like $x$, $y$, $z$
-\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function.
-$\lambda$-abstractions have a variable and a body.
-The body is a $\lambda$-calculus expression.
-Abstractions are written as $\lambda \text{var}.
-\text{body}$.
-\\\noindent An \emph{application} of a function to an argument.
-Both the function and the argument are $\lambda$-calculus expressions.
-The application of the function $f$ to the argument $x$ is written as $f\; x$.
-
-For example, the function which squares the logarithm of a number is
-$\lambda x.\text{\texttt{square}} (\text{\texttt{log} } x)$, and the identity function $f(x) = x$ is $\lambda x.x$. The
-$\lambda$-calculus serves as a spartan but expressive Turing complete
-program representation, and distills the essential features of functional languages like Lisp.
-
-However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
-(i.e., $5 \; (\lambda x.x)$).
-We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
-The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
-Hindley-Milner has two important features:
-Feature 1: It supports \emph{parametric polymorphism}: meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
-For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
-A function that returns the first element of a list has the type $\texttt{list}(\alpha)\to\alpha$.
-Type variables are not the same has variables introduced by $\lambda$-abstractions.
-Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
-A detailed exposition of Hindley-Milner is beyond the scope of this work.
-
-%% \begin{figure}
-%%   \begin{align}
-%%     \text{primitive types} & 
-%%     \end{align}
-%%   \end{figure}
-
-\section{Estimating $\theta$}
-
-We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.\ref{programGenerativeModel}.
-Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
-\begin{align*}
-  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
-  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
-  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
-\end{align}
-where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.
-
-Differentiate with respect to $\theta_e$ and set to zero 
-\begin{align}
-  \frac{c(x)}{\theta^{(x)}} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
-\end{align}
-This equality holds if $\theta^{(x)} = c(x)/a(x)$:
-\begin{align}
-  \frac{c(x)}{\theta_x} &= a(x).\\
-N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
-\end{align}
-If this equality holds then $\theta_x \propto c(x)/a(x)$:
-\begin{align}
-  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}.
-\end{align}
-
-Now what we are actually after is the parameters that maximize the joint log probability of the data+parameters, which I will write $J$:
-\begin{align}
-  J& = L + \log \text{D}(\theta|\alpha)\\
-  &\stackrel{+}{\geq } \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x  + \sum_x(\alpha_x - 1)\log \theta_x\\
-  & = \sum_x (c(x) + \alpha_x - 1)\log \theta_x -  N \log \sum_x a(x)\theta_x
-\end{align}
-So you add the pseudocounts to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$).
-
 \bibliography{main}

### 2018-02-05 supplement.tex
moved to stuff to supplement

--- 2018-02-05 paper.tex
@@
 \twocolumn[
-\icmltitle{Inducing Domain Specific Languages for Bayesian Program Learning}
+\icmltitle{Supplement to: Inducing Domain Specific Languages for Bayesian Program Learning}
 
@@
 \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
-\begin{abstract}
-This document provides a basic paper template and submission guidelines.
-Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-Gross violations will trigger corrections at the camera-ready phase.
-\end{abstract}
-
-\section{Introduction}
-
-Automatically inducing programs from examples is a long-standing goal
-of artificial intelligence.  Recent work in program induction show
-that deep networks can be very successful at learning to write certain
-kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
-DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
-techniques (e.g., Metagol:~\cite{muggleton2015meta} and 
-FlashFill:~\cite{gulwani2011automating}).
-However, the success of these approaches -- both symbolic and neural --
-relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
-The DSL is an inventory of restricted programming primitives,
-that imparts domain specific knowledge about the space of programs.
-To what extent can we dispense with hand-engineered DSLs?
-
-This work proposes \emph{learning} DSLs.
-We consider the problem of solving many related program induction tasks,
-starting with a Lisp-like representation.
-We introduce an algorithm, called the \system,
-which creates its own DSL by discovering and then reusing
-new useful programming idioms and subroutines.
-
-%% Imagine an agent faced with a new domain of unfamiliar tasks: think
-%% playing new kinds of videogames, authoring Python code, or navigating
-%% through mazes.  Our agent has at its disposal a basis of primitive
-%% actions it can compose to propose solutions to these tasks, but it is
-%% no idea what kinds of primitives are appropriate for which tasks
-%% nor does it know the higher-level vocabulary in which solutions are
-%% best expressed.  How can our agent learn to solve problems in this new
-%% domain?
-
-%% The AI and machine learning literature contains two broad takes on this problem.
-%% The first take is that the agent should come up with a better representation of the space of solutions,
-%% for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
-%% The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
-%% Our contribution is a general algorithm for fusing these two takes on the problem:
-%% we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
-%% alongside a bottom-up discriminative model that regresses from problems to solutions.
-%% We evaluate our algorithm on four domains:
-%% building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
-%% %Because we want a domain-general
-%% %approach,
-%% We model solutions to tasks as
-%% programs,
-%% which frames our problem as one of program induction.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}
-
-  \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-  \node[latent] at (1.5,3) (t){$\theta$};
-  \node[latent] at (1,1.75) (z){$p$};
-%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-  \node[obs] at (1,0) (x) {$x$};
-  \edge {z}{x};
-  \edge {d,t}{z};
-  \plate {}{(z)(x)}{$x\in X$};
-  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative \\model};
-
-  \begin{scope}[shift = {(-1,0)}]  
-  \node[latent] at (3.5,3) (dx){$\mathcal{D}$};
-  \node[latent] at (3.5,1.75) (zp){$p$};
-  \node[latent] at (5.5,1.75) (tx){$\theta^{(x)}$};
-  \node[obs] at (3.5,0) (xp) {$x$};
-  \plate {}{(zp)(xp)(tx)}{$x\in X$};
-  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
-  \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
-  \draw [->,red] (tx.west) -- (zp.east);
-  \draw [->,red] (dx.south) -- (zp.north);
-
-  \node at (nn) {
-    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-      \node[rectangle,width=2cm] at (1,1) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (I-\name) at (0,-\y) {};
-      \foreach \name / \y in {1,...,3}
-          \node[neuron] (H-\name) at (1,-\y-0.5) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (O-\name) at (2,-\y) {};
-      \foreach \source in {1,...,4}
-          \foreach \dest in {1,...,3}
-              \draw [-latex] (I-\source) -- (H-\dest);
-      \foreach \source in {1,...,3}
-          \foreach \dest in {1,...,4}
-              \draw [-latex] (H-\source) -- (O-\dest);
-    \end{tikzpicture}
-  };
-  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
-
-  \end{scope}
-
-  \end{tikzpicture}
-  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
-Observe tasks $x$ (program inputs/outputs).
-(b)  neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to a distribution over programs
-    ($\theta^{(x)} = q(x)$). Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
-\end{figure}
-
-Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}(a): black lines):
-A DSL $\mathcal{D}$, equipped with a real-valued weight vector $\theta$,
-give a distribution over programs
-from which a latent program is sampled,
-which then determines the agent's observations.
-Our goal will be both to infer the programs that best explain our observations,
-and at the same time induce a DSL
-that distills the commonalities across all of the programs that solve the tasks.
-Inducing the DSL means learning the symbolic structure of a latent programming language,
-and so we are confronted with two distinct structure learning problems:
-the structure of the programs and the structure of the DSL.
-
-Inference in the generative model in Figure~\ref{graphicalModel}(a) is
-difficult because of the vastness of the combinatorial space of all
-programs. To make inference tractable our model learns a bottom-up
-\emph{recognition model} alongside the top-down generative model
-(Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
-network that regresses from observations to a distribution over
-programs.  The neural network $q$ implements an amortized
-inference/compiled inference
-scheme~\cite{le2016inference,ritchie2016deep} for the generative
-model.
-
-How should we perform inference for the model
-in~\ref{graphicalModel}(a)?  We take the intuitive strategy of exploiting the
-conditional independence of the programs conditioned on the DSL, and
-so alternate between estimating the DSL ($\mathcal{D},\theta$) and
-estimating a posterior distribution over $p$ by searching for
-programs written using $\mathcal{D}$.  In practice searching for
-programs is difficult, so we use the recognition model to
-learn how to search efficiently.
-
-We apply \system to four domains:
- building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
- For each of these we deliberately provide an impoverished
- set of programming primitives,
- and show that our algorithm discovers
- its own domain-specific vocabulary for expressing solutions in each of these domains (Table~\ref{initialExampleDSL}).
- \begin{table}
-    \begin{tabular}{ll}
-   \toprule
-   Domain&Part of the learned DSL\\\midrule
-   Circuits&\texttt{(nand (nand a a) (nand b b))}\\
-   &\hspace{0.5cm} \emph{(an OR gate)}\\
-   Regression& \texttt{(+ (* real x) real)} \\
-      &\hspace{0.5cm} \emph{(a linear function of x)}\\
-   Strings& \texttt{(join '' (split c s))}\\
-   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
-   Lists& \texttt{(map (lambda (x) (+ x a)) l)}\\
-   &\hspace{0.5cm} \emph{(add a to every element of list l)}
-   \\\bottomrule
-    \end{tabular}
-    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful fragments of code.}\label{initialExampleDSL}
- \end{table}
-
- \subsection{Related Work}
- Our work is far from the first to learn to learn programs,
- an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:
-
- \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
- focused on engineering neural networks that regress from
- input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to a lesioned version of
- \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned.
-
- \noindent \emph{Predicate Invention in ILP:} The Inductive Logic Programming (ILP) community
- has long worked on the problem of inventing new primitive predicates~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}, which is analogous to our model with $\theta$ held fixed and $q$ ablated.
-
- \noindent \emph{Inventing new subroutines for program induction:}
- Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting.
-We find this work inspiring and motivating,
- and go beyond it along two dimensions: (1) we propose more sophisticated algorithms for
- inducing reusable structure, based on a formalism called \emph{Fragment Grammars}~\cite{tim};
- and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.
-
-Our work is an instance of
- \emph{Bayesian Program
-  Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
-
-%% \begin{itemize}
-%%   \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
-%%   \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
-%%     $q$;
-%%     % because this is a primary inspiration, might be worth further mention:
-%%     % different program representation (routing vs. substitution; simple vs polymorphic types);
-%%     % different grammar induction (sequitur vs. fragment grammar)
-%%   \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
-%%     DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
-%%     network $q$.
-%%     % although RobustFill has very different enumeration
-%% \end{itemize}
-
-%% We cast these problems as \emph{Bayesian Program
-%%   Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
-%% where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
-%% A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
-%% We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
-%% define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
-%% In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
-%% where the likelihood $\probability[x|p]$ is domain-dependent.
-%% The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
-%% Alongside this generative model,
-%% we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
-
-\section{The \system Algorithm}
-
-Our goal  is to induce a DSL and find good programs solving each of the tasks.
-Our strategy is to iterate through three steps: (1) searching for programs that
-solve the tasks, (2) learning a better neural recognition model -- which we use
-to accelerate the search over programs -- and (3) improving the DSL.
-The key observation here is that each of these three steps can bootstrap off each other:
-
-\begin{description}
-\item[ Searching for programs:] Our program search  is informed by both the DSL and the recognition model.
-When these improve we find better programs.
-  \item[ Learning a recognition model:] The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
-As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-  \item[Improving the DSL:] We induce a DSL from the programs we have found so far which solve the tasks;
-  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
-\end{description}
-
-The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
-learning a generative model and training a recognition model on samples from the generative model.
-We apply this idea to program induction, motivating why we call our algorithm the \system.
-Figure~\ref{feeding} diagrams how these three steps feed off of each other.
-\begin{figure}\centering
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
-    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45]{{\footnotesize Trains}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45]{{\footnotesize Trains}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above]{{\footnotesize Trains}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
-\caption{Our algorithm solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others.}  \label{feeding}
-  \end{figure}
-
-Section~\ref{mathematicalFraming} frames this 3-step procedure as
-a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
-Section~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Section~\ref{recognitionSection} explains how we train a neural network to accelerate the search over programs; and
-Section~\ref{grammarInductionSection} explains how \system induces a DSL from programs.
-
-\subsection{Probabilistic Framing}\label{mathematicalFraming}
-
-\system takes as input a set of \emph{tasks}, written $X$, each of which is a program induction problem.
-It has at its disposal a \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
-Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL $\mathcal{D}$.
-
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}. We wish to maximize the posterior probability of $\mathcal{D}$:
-\begin{equation*}
-  \probability[\mathcal{D}|X]\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]
-\end{equation*}
-In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
-Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
-\begin{align}
-  A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
-&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
-  \end{align}
-
-If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{AIC}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
-Through this lens we now take as our goal to maximize Eq.~\ref{AIC}.
-But even \emph{evaluating} Eq.~\ref{AIC} is intractable because it involves summing over the infinite set of all possible programs.
-In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
-With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{AIC},
-which we call $J$:
-\begin{equation}
-J = \log \probability[\mathcal{D},\theta] + \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
-\end{equation}
-This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:
-
-\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
-
-We maximize $J$ by alternating maximization w.r.t. the frontiers and the DSL:
-%So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
-%explains how \system synthesizes new programs and given a DSL.
-\\\noindent \textbf{Program Search: Maxing $J$ w.r.t. the frontiers.} Here we
-want to find new programs to add to  the frontiers so that $J$ increases the most.
-Adding new programs to the frontiers means searching for new programs $p$ for task $x$
-where $\probability[x,p|\mathcal{D},\theta]$ is large.
-\\\noindent \textbf{DSL Induction: Maxing $J$ w.r.t.
-the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $J$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
-
-Searching for programs is extremely difficult because
-of the large search space. We ease this difficulty by learning a neural recognition model:
-
-\textbf{Neural recognition model: tractably maxing $J$ w.r.t. the
-  frontiers.}  Here we train a neural network, $q$, to predict a
-distribution over programs conditioned on a task. The objective of $q$
-is to assign high probability to programs $p$ where
-$\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
-the frontier of a task $\mathcal{F}_x$ by searching for programs that maximize
-$q(p|x)$.
-The network $q$ exploits the structure of the DSL $\mathcal{D}$:
-rather than directly predicting a distribution over $p$ conditioned on $x$,
-it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
-This approach implements an amortized
-inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
-Figure~\ref{graphicalModel}.
-
-%% Learning the DSL eases the difficulty of synthesis
-%% by exposing a domain-specific basis for constructing programs.
-%% Another complementary means of meeting search is to learn a
-%% bottom-up \emph{recognition model}
-
-\subsection{Searching for Programs}\label{explorationSection}
-
-Now our goal is to search for programs that solve the tasks.  In this
-work we use the simple search strategy of enumerating programs from
-the DSL  in decreasing order of their probability,
-and then checking if an enumerated program $p$ assigns positive
-probability to a task ($\probability[x|p] > 0$); if so, we include $p$ in
-the frontier $\mathcal{F}_x$.
-
-To make this concrete we need to define what programs actually are and
-what form $\probability[p |\mathcal{D},\theta]$ takes.
-We represent programs as $\lambda$-calculus expressions.
-$\lambda$-calculus is a formalism for expressing functional programs
-that closely resembles the Lisp programming language.
-$\lambda$-calculus includes variables, function application, and the ability to create new functions.
-Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
-Our programs are all strongly typed.
-We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is used in functional languages like OCaml.
-Variables that range over types are always written using lowercase Greek letters
-and we write $\alpha\to \beta$ to mean a function that takes as input type $\alpha$
-and returns something of type $\beta$.
-We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
-has the type $\tau$.
-For example, to describe the type of the identity function
-we would say \texttt{(lambda (x) x)}$:\alpha\to \alpha$.
-We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
-$p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
-a type $\alpha$ with $\tau$ is to introduce constraints on the type
-variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
-See Section~\ref{programrepresentation} for more detail on program representation.
-With this notation in hand we can now define DSLs:
-
-\noindent\textbf{Definition.}
-A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-\\\noindent\textbf{Definition.}
-A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL primitive $e:\tau\in \mathcal{D}$, written $\theta_e$,
-and a weight controlling the probability of a variable occurring in a program, written $\theta_{\text{var}}$.
-
-Algorithm~\ref{programGenerativeModel} is a procedure for drawing
-samples from $\probability[p|\mathcal{D},\theta]$.  In practice, we
-enumerate programs rather than sampling them.  Enumeration proceeds by
-a depth-first search over the random choices made by
-Algorithm~\ref{programGenerativeModel}; we wrap the depth-first search
-in iterative deepening to (approximately) build $\lambda$-calculus
-expressions in order of their probability.
-
-\begin{algorithm}[tb]
-   \caption{Generative model over programs}
-   \label{programGenerativeModel}
-   \begin{algorithmic}
-     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
-  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
-  \STATE \textbf{Output:} a program whose type unifies with $\tau$
-  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
-\STATE
-     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
-  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
-  \STATE \textbf{Output:} a program whose type unifies with $\tau$
-  \IF{$\tau = \alpha\to\beta$}
-  \STATE var $\gets$ an unused variable name
-  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
-   \STATE \textbf{return} \texttt{(lambda (}var\texttt{) }body\texttt{)}
-   \ENDIF
-   %   \ELSE
-   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
-   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
-
-   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
-   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
-   \STATE Unify $\tau$ with yield$(\tau')$.
-   \STATE $\left\{\alpha_k \right\}_{k\leq K}\gets\text{argTypes}(\tau')$
-%   \STATE unify$(\tau,\beta)$
-   \FOR{$k=1$ {\bfseries to} $K$}
- \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
- \ENDFOR
- \STATE \textbf{return} \texttt{(}$e\;a_1\; a_2\; \cdots\; a_K$\texttt{)}
- \STATE\textbf{where:}
- \STATE  yield$(\tau) = \begin{cases}
-\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
-\tau   &\text{ otherwise.}
- \end{cases}$ 
- \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
-[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
-[]   &\text{ otherwise.}
- \end{cases}$
-\end{algorithmic}
-\end{algorithm}
-
-Why enumerate, when the program synthesis community has invented many
-sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
-We have two reasons:% for using the simple strategy of enumerating programs:
-\begin{itemize}
-\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
-\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-  the space of of programs.
-\end{itemize}
-A main drawback of an enumerative search algorithm is that we have no
-efficient means of solving for arbitrary constants that might occur in the
-program. In Section~\ref{regressionSection},
-we will show how to find programs with real-valued constants
-by automatically differentiating through the program and setting the constants using gradient descent.
-In Section~\ref{textSection}
-we will show that the bottom-up neural recognition model can learn
-which discrete constants should be included in a program.
-
-\subsection{Learning a Neural Recognition Model}\label{recognitionSection}
-
-The purpose of the recognition model is to accelerate the search over
-programs.  It does this by learning to predict which programs both 
-assign high likelihood to a task, and at the same time 
-have high prior probability under the prior $\probability[\cdot |\mathcal{D},\theta]$.
-
-The recognition model $q$ is a neural network that predicts,
-for each task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.
-Together with the DSL, this defines a distribution over programs,
-$\probability[p|\mathcal{D},\theta = q(x)]$.
-We abbreviate this distribution as $q(p|x)$.
-The crucial aspect of this framing is that the neural network
-leverages the structure of the learned DSL,
-and so is \emph{not} responsible for
-generating programs wholesale.
-
-We want a recognition model that closely approximates the true posteriors over programs, and so aim to minimize the following KL-divergence:
-\begin{equation*}
-  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]
-\end{equation*}
-which is equivalent to maximizing
-$$
-  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
-$$
-  where the expectation is taken over tasks. One could take this expectation
-  over the empirical distribution of the observations,
-  like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
-  We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the \system objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
-  \begin{align}
-  \mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
-  \mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\\
-  \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-    \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
-  \end{align}
-
-  Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
-  the current DSL, running them to get their outputs,
-  and then training $q$ to regress from the outputs to the program.
-  If these programs map inputs to outputs,
-  then we need some way of sampling these inputs as well.
-  Our solution to this problem is to sample the inputs
-  from the empirical observed distribution of inputs in $X$.
-
-\subsection{Inducing the DSL from the Frontiers}\label{grammarInductionSection}
-
-The purpose of the DSL is to
-offer a set of abstractions
-that allow an agent to easily express solutions to the tasks at hand.
-In the \system algorithm we infer the DSL from a collection of frontiers.
-Intuitively, we want the algorithm to
-look at the programs in the frontiers and
-generalize beyond them; 
-not only so the DSL can better express the current solutions,
-but also so that the DSL might expose new abstractions
-which will later be used to
-discover even better programs.
-
-Exact maximization of $J$ (Eq.\ref{lowerBound}) w.r.t. $(\mathcal{D},\theta)$
-is intractable,
-so we take a more heuristic approach.
-The strategy is to search locally through the space of DSLs,
-proposing small local changes to $\mathcal{D}$ until $J$ fails to increase.
-The search moves work by introducing new
-$\lambda$-expressions into the DSL.
-We propose these new expressions by extracting subexpressions from
-programs already in the frontier.
-These extracted subexpressions
-are fragments of the original programs, and can introduce new variables (Figure~\ref{fragmentExample}),
-which then become new functions in the DSL.
-
-Closely related to Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
-but context-sensitive
-
-We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
-\begin{align*}
-  \probability[\mathcal{D}]&\propto\exp\left(\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
-  P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)
-\end{align*}
-where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
-and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-Algorithm~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
-
-Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
-Although this problem may seem 
-very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-$\mathcal{D}$
- is actually context-sensitive
-due to the presence of variables in the programs and also due to the
-polymorphic typing system.
-In the Appendix we derive a tractable MAP estimator for $\theta$.
-\begin{algorithm}[tb]
-   \caption{DSL Induction Algorithm}
-   \label{grammarInductionAlgorithm}
-   \begin{algorithmic}
-     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
-     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
-     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
-     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-     \WHILE {true}
-     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
-     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
-     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
-     \STATE $\mathcal{D}\gets\mathcal{D}'$
-     \ENDWHILE
-   \end{algorithmic}
-\end{algorithm}
-
-\begin{figure*}
-  \begin{tabular}{lll}
-    \toprule
-    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
-    Boolean circuits&
-    \begin{tabular}{l}
-      \texttt{(lambda x (nand x x))}\\
-      \texttt{(lambda x (lambda y (nand x (nand y y))}
-    \end{tabular}&
-    \texttt{(nand z z)}\\\midrule
-    String editing&
-    \begin{tabular}{l}
-      \texttt{(lambda s (+ ',' (index 0 (split ',' s))))}\\
-      \texttt{(lambda s (index 0 (split ' ' s)))}
-    \end{tabular}&
-    \texttt{(index 0 (split z s))}\\\bottomrule
-  \end{tabular}
-  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
-    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} in the right column).}
-  \label{fragmentExample}
-\end{figure*}
-
-\subsection{Implementing the \system}
-
-Algorithm~\ref{mainAlgorithm} describes how we combine the
-program search, recognition model training, and DSL induction.
-We additionally make the following implementation details:
-(1) On the first iteration we do \emph{not} train the recognition model on samples from
-the generative model because, on the first iteration, we have not yet learned a
-new DSL, so we instead trained the network to only maximize $\mathcal{L}_{\text{AE}}$;
-(2) Because the frontiers can grow very large,
-we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
-with the highest $\probability[x,p|\mathcal{D},\theta]$;
-(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
-by only summing over the top $K$
-programs in $\mathcal{F}_x$.
-We found that $K = 2$ sufficed.
-(4) For added robustness, at each iteration we enumerate programs from both the generative model and the recognition model.
-
-\begin{algorithm}[tb]
-   \caption{The \system Algorithm}
-   \label{mainAlgorithm}
-   \begin{algorithmic}
-     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-     \FOR{$i=1$ {\bfseries to} $I$}
-%     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}^{(\theta)}_x\gets \{z| z\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|z] > 0\}$
- %    \ENDFOR
-     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
-     \STATE  $\mathcal{F}^{(q)}_x\gets\{z|z\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|z] > 0\}$
-     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{(\theta)}_x\cup\mathcal{F}^{(q)}_x\}_{x\in X})$
-     %% \STATE Define $Q_x(z) \propto \begin{cases}
-     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-     %%   0&x\not \in \mathcal{F}_x
-     %% \end{cases}
-      \ENDFOR
- \STATE \textbf{return} $\mathcal{D},\theta,q$
-\end{algorithmic}
-\end{algorithm}
-
-\section{Experiments}
-
-\subsection{Boolean circuits}
-As a toy warm-up domain, we consider the problem of learning
-to build the Boolean circuits out of logic gates.
-Now, if we equipped our agent with a full repertoire of
-circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
-and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
-Instead, we gave \system only a NAND gate,
-and tasked it with building a set of 500 random circuits made out of
-AND, NOT, and OR gates.
-This experimental setup was introduced in the original EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
-For this domain our recognition model is a simple multilayer perceptron
-whose inputs are the input/output examples.
-Figure~\ref{circuitLearningCurve} shows our algorithm learning to build circuits; Table~\ref{circuitExperimentTable}
-
-\begin{figure}
-  \includegraphics[width = \columnwidth]{figures/circuit.png}
-  \caption{\system learning to build Boolean circuits. We start out with only NAND gates and learn DSL that contains ...}\label{circuitLearningCurve}
-\end{figure}
-
-\begin{table}
-\begin{tabular}{ll}
-  \toprule
-  Primitives&\texttt{nand}
-
-\\\midrule
-  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  \midrule
-  Likelihood $\probability[x|p]$& $\prod_n \indicator[p(i_n) = o_n]$
-  \\\midrule
-  \begin{tabular}{l}
-    Subset of \\Learned DSL
-    \end{tabular}&
-
-  \\\bottomrule
-\end{tabular}
-\caption{}\label{circuitExperimentTable}
-\end{table}
-
-\subsection{Symbolic Regression}\label{regressionSection}
-We show how to use \system to infer programs containing both discrete
-structure and continuous parameters. The high-level idea is to synthesize programs with unspecified-real-valued parameters, and to fit those parameters using gradient descent.
-Concretely, we ask the algorithm to
-solve a set of 1000 symbolic regression problems, each a polynomial of
-degree 0, 1, or 2, where our observations $x$ take the form of $N$
-input/output examples, which we write as $x = \left\{(i_n,o_n)
-\right\}_{n\leq N}$. For example, one task is to infer a program
-calculating $3x + 2$, and the observations are the input-output
-examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.
-
-We initially equip our DSL learner with addition and multiplication,
-along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
-We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
-\begin{align*}
-\log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P_{\vec{\mathcal{R}}}(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
-%\text{\emph{BIC:}} &\approx 
-\end{align*}
-where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$.
-We approximate this marginal using the BIC~\cite{Bishop:2006:PRM:1162264}:
-\begin{align*}
-  \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
-\end{align*}
-where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.
-
-What DSL does \system learn?
-The learned DSL contains templates for quadratic and linear functions,
-which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain.
-Examining the programs themselves,
-one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
-for example, it represents the polynomial $8x^2+8x$
-
-\begin{tabular}{ll}
-  \toprule
-  Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
-  & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
-  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  \midrule
-  Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
-  \\\midrule
-  \begin{tabular}{l}
-    Subset of \\Learned DSL
-    \end{tabular}&
-  \begin{tabular}{ll}
-    $\lambda x. \mathcal{R}\times x + \mathcal{R}$ &linear\\
-    $\lambda x. \mathcal{R} + x$ &increment\\
-    $\lambda x. x\times (\text{linear } x)$&$\text{quadratic}_0$\\
-    $\lambda x. \text{increment }(\text{quadratic}_0\; x)$&quadratic\\
-  \end{tabular}
-  \\\bottomrule
-  \end{tabular}
-
-\begin{figure}
-      \includegraphics[width = 9cm]{figures/polynomial.png}
-\end{figure}
-\subsection{String editing}\label{textSection}
-
-\begin{tabular}{ll}
-  \toprule
-  Primitives&\begin{tabular}{l}
-    \texttt{0}, \texttt{+1},\texttt{-1}\\
-    \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
-    \texttt{append}, \texttt{chr->str}\\
-    \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
-    \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}
-    \end{tabular}
-
-\\\midrule
-  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  \midrule
-  Likelihood $\probability[x|p]$& $\prod_n \indicator[p(i_n) = o_n]$
-  \\\midrule
-  \begin{tabular}{l}
-    Subset of \\Learned DSL
-    \end{tabular}&
-
-  \\\bottomrule
-  \end{tabular}
-\begin{figure}
-
-  \includegraphics[width = 9cm]{figures/text.png}
-
-\end{figure}
-
-\subsection{List functions}
-Our list function domain consists of tasks which are solved by functions
-that take as input an integer or a list of integers, and have as output
-either a Boolean, an integer, or a list of integers. Examples of these
-functions are in Table~\ref{listexamples}. For each function, we create a
-task $x$ by generating 15 input/output examples used for testing whether a
-program produces the correct output. Supplying many examples reduces
-ambiguity in the task's function, ensuring solutions achieve the desired
-concept.
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l | l |}
-  \hline
-  \emph{name} & \emph{input} & \emph{output} \\
-  \hline
-  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
-  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
-  len & [3\, 5\, 12\, 1] & 4 \\
-  range & 3 & [1\, 2\, 3] \\
-  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
-  has-4 & [4\, 5\, 7\, 4] & \texttt{true} \\
-  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
-  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-  \hline
-\end{tabular}
-\caption{Examples from the domain of list functions.}
-\label{listexamples}
-\end{table}
-
-We supply \system with the DSL outlined in Table~\ref{listdsl}.
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l |}
-  \hline
-  \emph{name} & \emph{type} \\
-  \hline
-    empty & \texttt{[$\alpha$]} \\
-    singleton & \texttt{$\alpha$ $\rightarrow$ [$\alpha$]} \\
-    concat & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
-    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$]} \\
-    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$} \\
-    \hline
-    true & \texttt{bool} \\
-    not & \texttt{bool $\rightarrow$ bool} \\
-    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
-    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
-    \hline
-    0, $\ldots$, 9 & \texttt{int} \\
-    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
-    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
-    negate & \texttt{int $\rightarrow$ int} \\
-    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
-    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
-    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
-    is-prime & \texttt{int $\rightarrow$ bool} \\
-    is-square & \texttt{int $\rightarrow$ bool} \\
-    range & \texttt{int $\rightarrow$ [int]} \\
-    sort & \texttt{[int] $\rightarrow$ [int]} \\
-    \hline
-    sum & \texttt{[int] $\rightarrow$ int} \\
-    reverse & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$]} \\
-    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
-    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
-    index & \texttt{int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$} \\
-    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
-    slice & \texttt{int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
-  \hline
-\end{tabular}
-\caption{DSL for the domain of list function.}
-\label{listdsl}
-\end{table}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figures/list.eps}
-\end{figure}
-
-We found that using a less sophisticated but equally-capable DSL made common
-patterns, such as summation, unlikely and unlearnable in a small enumeration
-bound.
-
-\section{Model}
-
-\pagebreak
 

### 2018-02-05 paper.tex
edits in grammar induction section

@@
 leverages the structure of the learned DSL,
-and so is \emph{not} responsible for
+so it is \emph{not} responsible for
 generating programs wholesale.
 
-We want a recognition model that closely approximates the true posteriors over programs, and so aim to minimize the following KL-divergence:
+We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence:
 \begin{equation*}
@@
 which is equivalent to maximizing
-$$
+\begin{equation*}
   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
-$$
+\end{equation*}
 where the expectation is taken over tasks. One could take this expectation
@@
 which then become new functions in the DSL.
-
-Closely related to Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
-but context-sensitive
+Our model of subexpressions is loosely related to Fragment
+Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
+but context-sensitive.
 
@@
 
-Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
+To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
 Although this problem may seem 
@@
 for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-$\mathcal{D}$
- is actually context-sensitive
-due to the presence of variables in the programs and also due to the
-polymorphic typing system.
+a DSL in \system may be context-sensitive due to the presence of variables
+in the programs and also due to the polymorphic typing system.
 In the Appendix we derive a tractable MAP estimator for $\theta$.
@@
 
-Algorithm~\ref{mainAlgorithm} describes how we combine the
-program search, recognition model training, and DSL induction.
-We additionally make the following implementation details:
-(1) On the first iteration we do \emph{not} train the recognition model on samples from
-the generative model because, on the first iteration, we have not yet learned a
-new DSL, so we instead trained the network to only maximize $\mathcal{L}_{\text{AE}}$;
+Algorithm~\ref{mainAlgorithm} describes how we combine program search,
+recognition model training, and DSL induction.
+We note the following implementation details:
+(1) on the first iteration, we do \emph{not} train the recognition model on samples from
+the generative model because a DSL has not yet been learned --
+we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
 (2) Because the frontiers can grow very large,
 we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
-with the highest $\probability[x,p|\mathcal{D},\theta]$;
+with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
 (3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
 by only summing over the top $K$
-programs in $\mathcal{F}_x$.
-We found that $K = 2$ sufficed.
+programs in $\mathcal{F}_x$ -- we found that $K = 2$ sufficed.
 (4) For added robustness, at each iteration we enumerate programs from both the generative model and the recognition model.

### 2018-02-06 paper.tex
merge - not sure how I messed this up :/

@@
 \newcommand{\system}{\textsc{HelmholtzHacker}~}
-
+\newcommand{\lowerBound}{\mathscr{L}}
 % Use the following line for the initial blind version submitted for review:
@@
 \usepackage{microtype}      % microtypography
-
+\usepackage{mathrsfs}
 \usepackage{listings}
@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}(a), finding the $\mathcal{D}^*,\theta^*$ solving:
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}(a).
+Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*,\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
-  \nonumber\probability[X|\mathcal{D},\theta] &= \prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  \nonumber\probability[\mathcal{D}|X]&\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\probability[X|\mathcal{D},\theta]\\
-  \mathcal{D}^* &= \argmax_{\mathcal{D}}\probability[\mathcal{D}|X]\\
-\nonumber  \theta^*& =\argmax_\theta P(\theta |\mathcal{D}^*)\probability[X|\mathcal{D}^*,\theta]
+\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
+\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
 \end{align}
@@
 However, Eq.~\ref{intractableObjectives}
-is wildly intractable because evaluating $\probability[X|\mathcal{D},\theta]$ involves
+is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
 summing over the  infinite set of all possible programs.
@@
 
-With the frontiers in hand, we can define the following intuitive lower bound on the likelihood of the tasks, which we call $\mathcal{L}$:
+Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
 \begin{align}
-  \probability[X|\mathcal{D},\theta]\geq J\triangleq\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
-\end{align}
-along with a corresponding lower bound on the marginal likelihood of a DSL, which we call $$
-\begin{align}
-  \probability[\mathcal{D}|X]\geq 
+ J\geq \lowerBound\triangleq\probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
 \end{align}
 
-If we had a $(\mathcal{D},\theta)$ solving Eq.~\ref{intractableObjectives}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
-Through this lens we now take as our goal to solve Eq.~\ref{intractableObjectives}.
-But even \emph{evaluating} Eq.~\ref{intractableObjectives} is intractable because it involves summing over the infinite set of all possible programs, as an ideal Bayesian learner would.
-In practice, we must instead marginalize over
-some finite set of programs.
+%% If we had a $(\mathcal{D},\theta)$ solving Eq.~\ref{intractableObjectives}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
+%% Through this lens we now take as our goal to solve Eq.~\ref{intractableObjectives}.
+%% But even \emph{evaluating} Eq.~\ref{intractableObjectives} is intractable because it involves summing over the infinite set of all possible programs, as an ideal Bayesian learner would.
+%% In practice, we must instead marginalize over
+%% some finite set of programs.
 
-In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
+%% In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
 
-Now in theory we would like to sum over the entire infinite space
-of all programs -- but this is of course impossible.
+%% Now in theory we would like to sum over the entire infinite space
+%% of all programs -- but this is of course impossible.
 
@@
 %%   \end{align}
-
-With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{AIC},
-which we call $J$:
-\begin{equation}
-J = \log \probability[\mathcal{D},\theta] + \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
-\end{equation}
-This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:
-
-We maximize $J$ by alternating maximization w.r.t. the frontiers and the DSL:
+Mechanistically, \system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability.
+We maximize $\lowerBound$ by alternatingly maximizing w.r.t. the frontiers and the DSL:
 %So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
 %explains how \system synthesizes new programs and given a DSL.
-\\\noindent \textbf{Program Search: Maxing $J$ w.r.t. the frontiers.} Here we
-want to find new programs to add to  the frontiers so that $J$ increases the most.
+\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
+want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
 Adding new programs to the frontiers means searching for new programs $p$ for task $x$
 where $\probability[x,p|\mathcal{D},\theta]$ is large.
-\\\noindent \textbf{DSL Induction: Maxing $J$ w.r.t.
-the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $J$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $J$.
+\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.
+the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$.
+Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
+Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
 
@@
 
-\textbf{Neural recognition model: tractably maxing $J$ w.r.t. the
+\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
   frontiers.}  Here we train a neural network, $q$, to predict a
@@
 $\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
-the frontier of a task $\mathcal{F}_x$ by searching for programs that maximize
+the frontier $\mathcal{F}_x$ by searching for programs that maximize
 $q(p|x)$.
@@
 it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
-This approach implements an amortized
-inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
-Figure~\ref{graphicalModel}.
+%% This approach implements an amortized
+%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
+%% Figure~\ref{graphicalModel}.
+
+Our lower bound $\lowerBound$ is unconventional,
+and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
+For the interested reader, we explain in the supplement that maximizing an ELBO-style bound
+leads to a pathological behavior that would cause the algorithm to easily become trapped in local optima.
 
@@
 variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
-See Section~\ref{programrepresentation} for more detail on program representation.
+See Supplement for more detail on program representation.
 With this notation in hand we can now define DSLs:
@@
 
-Exact maximization of $J$ (Eq.\ref{lowerBound}) w.r.t. $(\mathcal{D},\theta)$
+Exact maximization of $\lowerBound$ (Eq.\ref{lowerBound}) w.r.t. $(\mathcal{D},\theta)$
 is intractable,
@@
 The strategy is to search locally through the space of DSLs,
-proposing small local changes to $\mathcal{D}$ until $J$ fails to increase.
+proposing small local changes to $\mathcal{D}$ until $\lowerBound$ fails to increase.
 The search moves work by introducing new
@@
 which then become new functions in the DSL.
-Our model of subexpressions is loosely related to Fragment
-Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
-but context-sensitive.
+
+%% Our model of subexpressions is loosely related to Fragment
+%% Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
+%% but context-sensitive.
+
+The idea of storing and reusing
+fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.
+Somehow I need to clarify that this is not like an adapter grammar, and that previous work has done adapter grammar like things...
+In fact this might be a baseline, because with the right parameter settings our algorithm can revert to an adapter grammar like phase.
 
@@
 \begin{align*}
-  \probability[\mathcal{D}]&\propto\exp\left(\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
+  \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
   P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)

### 2018-02-06 paper.tex
symbolic regression writing and figure

@@
 We show how to use \system to infer programs containing both discrete
-structure and continuous parameters. The high-level idea is to synthesize programs with unspecified-real-valued parameters, and to fit those parameters using gradient descent.
+structure and continuous parameters. The high-level idea is to synthesize programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
+and optimizing using gradient descent.
 Concretely, we ask the algorithm to
-solve a set of 1000 symbolic regression problems, each a polynomial of
-degree 0, 1, or 2, where our observations $x$ take the form of $N$
-input/output examples, which we write as $x = \left\{(i_n,o_n)
-\right\}_{n\leq N}$. For example, one task is to infer a program
-calculating $3x + 2$, and the observations are the input-output
-examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.
-
-We initially equip our DSL learner with addition and multiplication,
+solve a set of 400 symbolic regression problems, each a polynomial of
+degree 0 to 4, where our observations $x$ are 10
+input/output examples.%, which we write as $x = \left\{(i_n,o_n)
+%\right\}_{n\leq N}$. %% For example, one task is to infer a program
+%% calculating $3x + 2$, and the observations are the input-output
+%% examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.
+
+We initially give our system only addition and multiplication,
 along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
-We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
-\begin{align*}
-\log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P_{\vec{\mathcal{R}}}(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
-%\text{\emph{BIC:}} &\approx 
-\end{align*}
-where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$.
-We approximate this marginal using the BIC~\cite{Bishop:2006:PRM:1162264}:
-\begin{align*}
-  \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
-\end{align*}
-where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.
+We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples,
+and penalize extra continuous degrees of freedom using the BIC~\cite{Bishop:2006:PRM:1162264}:
+%% \begin{align*}
+%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
+%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
+%% \end{align*}
+
+%% %% and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
+%% %% \begin{align*}
+%% %% \log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
+%% %% %\text{\emph{BIC:}} &\approx 
+%% %% \end{align*}
+%% where $\mathcal{N}(\cdot |\cdot )$ is the normal density and
+%% $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over
+%% $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~:
+
+%% where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$
+%% found by performing gradient ascent on the likelihood of the
+%% observations w.r.t. $\vec{\mathcal{R}}$.
 
 What DSL does \system learn?
-The learned DSL contains templates for quadratic and linear functions,
-which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain.
+The learned DSL contains templates for polynomials of different orders,
+which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Table~\ref{regressionDSL}).
 Examining the programs themselves,
 one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
-for example, it represents the polynomial $8x^2+8x$
-
-\begin{tabular}{ll}
+for example, it represents the linear function $-x+2$ with the program 
+\texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
+This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood's model BIC penalty.
+\begin{table}
+\begin{tabular}{l}
   \toprule
-  Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
-  & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
-  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  \midrule
-  Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
-  \\\midrule
-  \begin{tabular}{l}
-    Subset of \\Learned DSL
-    \end{tabular}&
-  \begin{tabular}{ll}
-    $\lambda x. \mathcal{R}\times x + \mathcal{R}$ &linear\\
-    $\lambda x. \mathcal{R} + x$ &increment\\
-    $\lambda x. x\times (\text{linear } x)$&$\text{quadratic}_0$\\
-    $\lambda x. \text{increment }(\text{quadratic}_0\; x)$&quadratic\\
-  \end{tabular}
+  %% Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
+  %% & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
+  %% Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
+  %% \midrule
+  %% Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
+  %% \\\midrule
+    $f_0 = $\texttt{(+ }$\mathcal{R}$\texttt{)}\\
+    $f_1($\texttt{x}$) =$ \texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)} \\
+      $f_2($\texttt{x}$) =$ \texttt{(+ }$\mathcal{R}$\texttt{ ($f_1$ x))} \\
+      $f_3($\texttt{x}$) = $\texttt{(* x (}$f_2$\texttt{ x))}\\
+  $f_4($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_3$\texttt{ x)))}\\
+  $f_5($\texttt{x}$) =$ \texttt{(* x (}$f_4$\texttt{ x))}\\
+  $f_6($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_4$\texttt{ x))} (\emph{4th order polynomial})
   \\\bottomrule
 \end{tabular}
-
-\begin{figure}
-      \includegraphics[width = 9cm]{figures/polynomial.png}
+\caption{Some induced DSL primitives for symbolic regression.
+System starts out with addition, multiplication, and real numbers, and learns functions for building 4-th order polynomials.}\label{regressionDSL}
+\end{table}
+
+\begin{figure}\centering
+  \includegraphics[width = 7cm]{figures/symbolicRegression.png}
+  \caption{DSL learning curve for symbolic regression.}\label{regressionLearningCurve}
 \end{figure}

### 2018-02-06 paper.tex
miscellaneous improvements

@@
 \begin{abstract}
-This document provides a basic paper template and submission guidelines.
-Abstracts must be a single paragraph, ideally between 4--6 sentences long.
-Gross violations will trigger corrections at the camera-ready phase.
+
 \end{abstract}
@@
 A DSL $\mathcal{D}$, equipped with a real-valued weight vector $\theta$,
-give a distribution over programs
-from which a latent program is sampled,
-which then determines the agent's observations.
+defines a distribution over programs.
+A program is sampled from $(\mathcal{D},\theta)$,
+and then the program is executed to determines the agent's observations.
 Our goal will be both to infer the programs that best explain our observations,
@@
 that distills the commonalities across all of the programs that solve the tasks.
-Inducing the DSL means learning the symbolic structure of a latent programming language,
-and so we are confronted with two distinct structure learning problems:
-the structure of the programs and the structure of the DSL.
+%% Inducing the DSL means learning the symbolic structure of a latent programming language,
+%% and so we are confronted with two distinct structure learning problems:
+%% the structure of the programs and the structure of the DSL.
 
 Inference in the generative model in Figure~\ref{graphicalModel}(a) is
-difficult because of the vastness of the combinatorial space of all
-programs. To make inference tractable our model learns a bottom-up
-\emph{recognition model} alongside the top-down generative model
+difficult because the programs are unobserved, and so we must solve a
+difficult search problem to find any programs at all. To make
+inference tractable our model learns a bottom-up \emph{recognition
+  model} alongside the top-down generative model
 (Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
@@
 generating programs wholesale.
+This taps into the intuition that programming is primarily a top-down activity:
+as human programmers, we first decide what kind of
+programming constructs we might need to use,
+and then we figure out how to assemble them in to the desired program.
 
@@
 
+The neural architecture of $q$ depends upon the domain;
+for domains with sequential structure like strings and lists,
+we use an RNN,
+and for domains with a fixed-dimensionality inputs/outputs we use a multilayer perceptron.
+
 \subsection{Inducing the DSL from the Frontiers}\label{grammarInductionSection}
@@
 
-Exact maximization of $\lowerBound$ (Eq.\ref{lowerBound}) w.r.t. $(\mathcal{D},\theta)$
-is intractable,
-so we take a more heuristic approach.
-The strategy is to search locally through the space of DSLs,
-proposing small local changes to $\mathcal{D}$ until $\lowerBound$ fails to increase.
+Recall from Section~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
+We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
+\begin{align}
+\nonumber  \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
+&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
+  \end{align}
+
+We induce a DSL by searching locally through the space of DSLs,
+proposing small local changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
 The search moves work by introducing new
@@
 which then become new functions in the DSL.
-
-%% Our model of subexpressions is loosely related to Fragment
-%% Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing},
-%% but context-sensitive.
-
 The idea of storing and reusing
 fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.
-Somehow I need to clarify that this is not like an adapter grammar, and that previous work has done adapter grammar like things...
-In fact this might be a baseline, because with the right parameter settings our algorithm can revert to an adapter grammar like phase.
 
@@
 \begin{figure*}
-  \begin{tabular}{lll}
+\centering  \begin{tabular}{lll}
     \toprule
     Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
-    Boolean circuits&
+    Circuits&
     \begin{tabular}{l}
@@
     \texttt{(nand z z)}\\\midrule
-    String editing&
+    Strings&
     \begin{tabular}{l}
@@
   \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
-    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} in the right column).}
+    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} and \texttt{s} in the right column).}
   \label{fragmentExample}
@@
 %     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}^{(\theta)}_x\gets \{z| z\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|z] > 0\}$
+     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$
  %    \ENDFOR
-     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
-     \STATE  $\mathcal{F}^{(q)}_x\gets\{z|z\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|z] > 0\}$
-     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{(\theta)}_x\cup\mathcal{F}^{(q)}_x\}_{x\in X})$
+     \STATE $q(\cdot )\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
+     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$
+     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$
      %% \STATE Define $Q_x(z) \propto \begin{cases}
@@
 \subsection{Boolean circuits}
-As a toy warm-up domain, we consider the problem of learning
+As a toy domain, we consider the problem of learning
 to build the Boolean circuits out of logic gates.
@@
 and tasked it with building a set of 500 random circuits made out of
-AND, NOT, and OR gates.
-This experimental setup was introduced in the original EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
+AND, NOT, and OR gates,
+as well as 2 or 4 input multiplexers.
+This experimental setup is an extension of one introduced in the original EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
 For this domain our recognition model is a simple multilayer perceptron
@@
 \end{figure}
-
-\begin{table}
-\begin{tabular}{ll}
-  \toprule
-  Primitives&\texttt{nand}
-
-\\\midrule
-  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  \midrule
-  Likelihood $\probability[x|p]$& $\prod_n \indicator[p(i_n) = o_n]$
-  \\\midrule
-  \begin{tabular}{l}
-    Subset of \\Learned DSL
-    \end{tabular}&
-
-  \\\bottomrule
-\end{tabular}
-\caption{}\label{circuitExperimentTable}
-\end{table}
 
@@
 \caption{Some induced DSL primitives for symbolic regression.
-System starts out with addition, multiplication, and real numbers, and learns functions for building 4-th order polynomials.}\label{regressionDSL}
+System starts out with addition, multiplication, and real numbers, and learns functions for building 4th order polynomials.}\label{regressionDSL}
 \end{table}
@@
 \subsection{String editing}\label{textSection}
+\begin{table}
 
-\begin{tabular}{ll}
+  \centering
+  \begin{tabular}{l}
   \toprule
-  Primitives&\begin{tabular}{l}
     \texttt{0}, \texttt{+1},\texttt{-1}\\
     \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
-    \texttt{append}, \texttt{chr->str}\\
+    \texttt{append}, \texttt{chr->str}, \texttt{trim}\\
     \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
     \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}  
-    \end{tabular}
-
-\\\midrule
-  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  \midrule
-  Likelihood $\probability[x|p]$& $\prod_n \indicator[p(i_n) = o_n]$
-  \\\midrule
-  \begin{tabular}{l}
-    Subset of \\Learned DSL
-    \end{tabular}&
-
   \\\bottomrule
   \end{tabular}
+  \caption{Initial set of primitives provided to the system for learning string editing programs}
+  \end{table}
 \begin{figure}

### 2018-02-06 paper.tex
recompiled document
added a caption
writing for string editing section

@@
 \end{figure}
+
 \subsection{String editing}\label{textSection}
+Synthesizing programs that manipulate strings is a classic problem in the
+programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
+and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
+However, this prior work presumes a ready-made DSL,
+expertly crafted to suit string editing.
+We show in this experiment that we can instead start out with a
+more restricted set of primitives -- close to the string routines provided by Python --
+and automatically recover many of the higher-level building blocks that have made these
+other system successful.
+Table~\ref{textPrimitives} lists the primitive functions provided to \system.
+Absent are any routines for searching for characters, extracting substrings,
+or replacing one character with another.
+Instead, we provide array/string indexing primitives (\texttt{slice}, \texttt{index}),
+basic arithmetic, and functions for splitting a string into a list (\texttt{split})
+or turning a list into a string (\texttt{join}).
+
+We automatically generated 600 string editing problems,
+along with input/output examples (Figure~\ref{exampleTextProblem}).
+
 \begin{table}
@@
     \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
-    \texttt{append}, \texttt{chr->str}, \texttt{trim}\\
+    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{index}\\
     \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
@@
   \end{tabular}
-  \caption{Initial set of primitives provided to the system for learning string editing programs}
+  \caption{Initial set of primitives provided to the system for learning string editing programs}\label{textPrimitives}
   \end{table}
@@
 
-  \includegraphics[width = 9cm]{figures/text.png}
-
+  \includegraphics[width = 7cm]{figures/text.png}
+  \caption{Learning curves for string editing programs}\label{textLearningCurve}
+
+\end{figure}
+
+\begin{figure}
+  \begin{tabular}{ll}\toprule
+    Input&Output\\\midrule
+    kIDY eMSz,zON& 	eM\\
+KCQ njp,yzhq 	&nj\\
+zeChf zzZ,dtFP 	&zz\\
+plbKi fmLcO,ykkOR 	&fm    \\\bottomrule
+  \end{tabular}
+  \caption{Example string edit problem (left) and the program that our system recovers (right)}\label{exampleTextProblem}
   \end{figure}

### 2018-02-06 paper.tex
a single sentence
small edits

@@
 
-  \begin{scope}[shift = {(-1,0)}]  
+  \begin{scope}[shift = {(0.7,0)}]  
   \node[latent] at (3.5,3) (dx){$\mathcal{D}$};
@@
 Our work is an instance of
- \emph{Bayesian Program
-  Learning} (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
+ Bayesian Program
+ Learning (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
+Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
+ and our contribution here is a general scheme for getting BPL off the ground with less hand-engineering of the DSL.
 
@@
 
-Searching for programs is extremely difficult because
-of the large search space. We ease this difficulty by learning a neural recognition model:
+Searching for programs is hard because
+of the large combinatorial search space. We ease this difficulty by learning a neural recognition model:
 
@@
 it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
+This taps into the intuition that programming is primarily a top-down activity:
+as human programmers, we first decide what kind of
+programming constructs we might need to use,
+and then we figure out how to assemble them in to the desired program.
 %% This approach implements an amortized
@@
 generating programs wholesale.
-This taps into the intuition that programming is primarily a top-down activity:
-as human programmers, we first decide what kind of
-programming constructs we might need to use,
-and then we figure out how to assemble them in to the desired program.
 

### 2018-02-06 paper.tex
example invented primitives for text
at its introduction

@@
 
-\newcommand{\system}{\textsc{HelmholtzHacker}~}
+\newcommand{\system}{\textsc{DreamHacker}~}
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 
-This work proposes \emph{learning} DSLs.
-We consider the problem of solving many related program induction tasks,
-starting with a Lisp-like representation.
-We introduce an algorithm, called the \system,
-which creates its own DSL by discovering and then reusing
-new useful programming idioms and subroutines.
+This work proposes learning DSLs.
+We consider the setting where we have a collection of related programming tasks,
+each specified by a set of input/output examples.
+Our algorithm, called \system,
+does two things:
+it finds programs that solve each of the tasks,
+and at the same time creates its own DSL by discovering and reusing
+new domain-specific subroutines.
+The idea is that induced DSL
+should distill the commonalities across the programs that solve the tasks,
+and so can be used to solve other related program learning problems.
+
+%% starting with a Lisp-like representation.
+%% We introduce an algorithm, called the \system,
+%% which creates its own DSL by discovering and then reusing
+%% new useful programming idioms and subroutines.
 
@@
 
-Our algorithm is best understood as inference in a hierarchical Bayesian model (Figure~\ref{graphicalModel}(a): black lines):
-A DSL $\mathcal{D}$, equipped with a real-valued weight vector $\theta$,
-defines a distribution over programs.
-A program is sampled from $(\mathcal{D},\theta)$,
-and then the program is executed to determines the agent's observations.
-Our goal will be both to infer the programs that best explain our observations,
-and at the same time induce a DSL
-that distills the commonalities across all of the programs that solve the tasks.
+Viewed as an inference problem,
+our goal is to infer the DSL, written $\mathcal{D}$,
+along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
+A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
+We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together
+$(\mathcal{D},\theta)$ define a distribution over programs.
+Figure~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
+
 %% Inducing the DSL means learning the symbolic structure of a latent programming language,
@@
   \centering
+  \begin{tabular}{c}
+    String editing primitives\\
     \begin{tabular}{l}
@@
   \\\bottomrule
+    \end{tabular}\\
+\\    Subset of learned DSL\\
+    \begin{tabular}{l}
+      \toprule
+      $f_0 = $\texttt{(+ (chr->str c))}\\
+      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s)) (chr->str s))}\\
+      $f_2($\texttt{s,c}$) = $\texttt{(}$f_0$\texttt{ c) s (}$f_1$\texttt{ c s (1+ 0))} (\emph{Extracts substring delimited by character})
+    %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
+    %% $f_1 = $\texttt{(join "")}
+  \\\bottomrule
+    \end{tabular}
   \end{tabular}

### 2018-02-06 paper.tex
more improvements to the introduction

@@
 each specified by a set of input/output examples.
+We do not assume that the tasks are annotated with
+ground-truth programs.
 Our algorithm, called \system,
@@
 and at the same time creates its own DSL by discovering and reusing
-new domain-specific subroutines.
-The idea is that induced DSL
+domain-specific subroutines.
+The learned DSL
 should distill the commonalities across the programs that solve the tasks,
@@
 Observe tasks $x$ (program inputs/outputs).
-(b)  neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to a distribution over programs
+(b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs
     ($\theta^{(x)} = q(x)$). Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
@@
 Inference in the generative model in Figure~\ref{graphicalModel}(a) is
-difficult because the programs are unobserved, and so we must solve a
-difficult search problem to find any programs at all. To make
+complicated by the fact that the programs are unobserved latent variables,
+and so we must solve a difficult search problem to find any programs at all. To make
 inference tractable our model learns a bottom-up \emph{recognition
@@
 (Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
-network that regresses from observations to a distribution over
+network that regresses from input/output pairs to a distribution over
 programs.  The neural network $q$ implements an amortized
@@
 scheme~\cite{le2016inference,ritchie2016deep} for the generative
-model.
+model by speeding up the search for programs.
 
@@
 
-The neural architecture of $q$ depends upon the domain;
-for domains with sequential structure like strings and lists,
-we use an RNN,
+The neural architecture of $q$ depends upon the domain.
+For domains with sequential structure like strings and lists,
+we use an RNN to encode each input/output example,
 and for domains with a fixed-dimensionality inputs/outputs we use a multilayer perceptron.
+We MaxPool across all of the input/output examples. See supplement for more details on the neural architecture.
 

### 2018-02-06 paper.tex
recompiled document

@@
 \usepackage{graphicx}
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage{booktabs} % for professional tables
@@
 \usepackage{caption}
-\usepackage{subcaption}
+%\usepackage{subcaption}
 
@@
 \DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
-\DeclareMathOperator{\argmin}{argmin} % no space, limits underneath in displays
 
@@
 The DSL is an inventory of restricted programming primitives,
-that imparts domain specific knowledge about the space of programs.
+and imparts domain specific knowledge about the space of programs.
 To what extent can we dispense with hand-engineered DSLs?

### 2018-02-06 paper.tex
temporarily disabled hyperlinks

@@
 % \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
-\usepackage{hyperref}
+%\usepackage{hyperref}
 
@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage{icml2018}
-
-\usepackage{hyperref}       % hyperlinks
+\usepackage[nohyperref]{icml2018}
+
+%\usepackage{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
@@
 \usepackage{amsmath}
-\usepackage{hyperref}
+
 \DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
@@
 Inference in the generative model in Figure~\ref{graphicalModel}(a) is
-complicated by the fact that the programs are unobserved latent variables,
+complicated by the fact that the programs are unobserved,
 and so we must solve a difficult search problem to find any programs at all. To make
-inference tractable our model learns a bottom-up \emph{recognition
+search tractable we learn a bottom-up \emph{recognition
   model} alongside the top-down generative model

### 2018-02-06 paper.tex
improved text DSL figure

@@
 
-This work proposes learning DSLs.
+This work proposes \emph{learning} the DSL.
 We consider the setting where we have a collection of related programming tasks,
@@
 estimating a posterior distribution over $p$ by searching for
-programs written using $\mathcal{D}$.  In practice searching for
+programs.  In practice searching for
 programs is difficult, so we use the recognition model to
@@
 or replacing one character with another.
-Instead, we provide array/string indexing primitives (\texttt{slice}, \texttt{index}),
+Instead, we provide array/string indexing primitives (\texttt{slice}, \texttt{nth}),
 basic arithmetic, and functions for splitting a string into a list (\texttt{split})
@@
     \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
-    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{index}\\
+    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{nth}\\
     \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
@@
       \toprule
-      $f_0 = $\texttt{(+ (chr->str c))}\\
-      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s)) (chr->str s))}\\
-      $f_2($\texttt{s,c}$) = $\texttt{(}$f_0$\texttt{ c) s (}$f_1$\texttt{ c s (1+ 0))} (\emph{Extracts substring delimited by character})
+      $f_0($\texttt{c}$) = $\texttt{(+ (chr->str c))}\\
+      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s))}\\\hspace{2.55cm}\texttt{ (chr->str c))}\\
+      $f_2($\texttt{s,c}$) = $\texttt{((}$f_0$\texttt{ c) (}$f_1$\texttt{ (1+ 0) c s))} \\
+      \hspace{1cm}($f_2$: \emph{Extract substring delimited by character})
     %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
@@
   \end{tabular}
-  \caption{Initial set of primitives provided to the system for learning string editing programs}\label{textPrimitives}
+  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
   \end{table}

### 2018-02-06 paper.tex
title change

@@
 \twocolumn[
-\icmltitle{Inducing Domain Specific Languages for Bayesian Program Learning}
+\icmltitle{\system: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
 \texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
-This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood's model BIC penalty.
+This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
 \begin{table}

### 2018-02-07 paper.tex
progress on wake sleep

@@
 
-\newcommand{\system}{\textsc{DreamHacker}~}
+\newcommand{\system}{\textsc{DreamCoder}~}
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 \twocolumn[
-\icmltitle{\system: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+\icmltitle{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
 Our algorithm, called \system,
-does two things:
-it finds programs that solve each of the tasks,
-and at the same time creates its own DSL by discovering and reusing
-domain-specific subroutines.
-The learned DSL
-should distill the commonalities across the programs that solve the tasks,
-and so can be used to solve other related program learning problems.
+does three things:
+it finds programs that solve each of the tasks;
+it creates its own DSL by discovering and reusing
+domain-specific subroutines;
+and it learns a neural network that
+efficiently searches for programs in the DSL.
+The purpose of the DSL
+is to distill the commonalities across the programs that solve the tasks,
+and it can be used to solve other related program learning problems.
 
@@
 How should we perform inference for the model
-in~\ref{graphicalModel}(a)?  We take the intuitive strategy of exploiting the
+in~\ref{graphicalModel}(a)?
+We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
+which alternates between ``waking'', or training a generative model (our $\mathcal{D},\theta$),
+and  ``sleeping'', or training a recognition model on samples from the generative model (our $q$).
+For us, a wake cycle means using the recognition model to search for programs that solve the tasks,
+while a sleep cycle means updating the DSL and the recognition model
+to more closely match the distribution of tasks.
+This wake/sleep style of inference is the reason why we call our approach \system.
+
+We take the intuitive strategy of exploiting the
 conditional independence of the programs conditioned on the DSL, and
@@
     \end{tabular}
-    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful fragments of code.}\label{initialExampleDSL}
+    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
  \end{table}
-
- \subsection{Related Work}
- Our work is far from the first to learn to learn programs,
- an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:
-
- \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
- focused on engineering neural networks that regress from
- input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to a lesioned version of
- \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned.
-
- \noindent \emph{Predicate Invention in ILP:} The Inductive Logic Programming (ILP) community
- has long worked on the problem of inventing new primitive predicates~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}, which is analogous to our model with $\theta$ held fixed and $q$ ablated.
-
- \noindent \emph{Inventing new subroutines for program induction:}
- Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting.
-We find this work inspiring and motivating,
- and go beyond it along two dimensions: (1) we propose more sophisticated algorithms for
- inducing reusable structure, based on a formalism called \emph{Fragment Grammars}~\cite{tim};
- and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.
-
-Our work is an instance of
- Bayesian Program
- Learning (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
-Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
- and our contribution here is a general scheme for getting BPL off the ground with less hand-engineering of the DSL.
 
@@
 by automatically differentiating through the program and setting the constants using gradient descent.
-In Section~\ref{textSection}
-we will show that the bottom-up neural recognition model can learn
-which discrete constants should be included in a program.
+%% In Section~\ref{textSection}
+%% we will show that the bottom-up neural recognition model can learn
+%% which discrete constants should be included in a program.
 
@@
 assign high likelihood to a task, and at the same time 
-have high prior probability under the prior $\probability[\cdot |\mathcal{D},\theta]$.
+have high probability under $(\mathcal{D},\theta)$.
 
@@
 
-\section{Model}
+\pagebreak
+ \subsection{Related Work}
+ Our work is far from the first to learn to learn programs,
+ an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:
+
+ \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
+ focused on engineering neural networks that regress from
+ input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to a lesioned version of
+ \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned.
+
+ \noindent \emph{Predicate Invention in ILP:} The Inductive Logic Programming (ILP) community
+ has long worked on the problem of inventing new primitive predicates~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}, which is analogous to our model with $\theta$ held fixed and $q$ ablated.
+
+ \noindent \emph{Inventing new subroutines for program induction:}
+ Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting.
+We find this work inspiring and motivating,
+ and go beyond it along two dimensions: (1) we propose more sophisticated algorithms for
+ inducing reusable structure, based on a formalism called \emph{Fragment Grammars}~\cite{tim};
+ and (2) we show how to combine these techniques with bottom-up neural recognition models.
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.
+
+Our work is an instance of
+ Bayesian Program
+ Learning (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
+Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
+ and our contribution here is a general scheme for getting BPL off the ground with less hand-engineering of the DSL.
 

### 2018-02-07 paper.tex
figure change

@@
 search tractable we learn a bottom-up \emph{recognition
-  model} alongside the top-down generative model
+  model}, written $q(\cdot )$
 (Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
 network that regresses from input/output pairs to a distribution over
-programs.  The neural network $q$ implements an amortized
-inference/compiled inference
+programs likely to explain the input/outputs. $q(\cdot )$ implements an amortized
+inference
 scheme~\cite{le2016inference,ritchie2016deep} for the generative
-model by speeding up the search for programs.
+model.
 
@@
 
-We take the intuitive strategy of exploiting the
-conditional independence of the programs conditioned on the DSL, and
-so alternate between estimating the DSL ($\mathcal{D},\theta$) and
-estimating a posterior distribution over $p$ by searching for
-programs.  In practice searching for
-programs is difficult, so we use the recognition model to
-learn how to search efficiently.
+%% We take the intuitive strategy of exploiting the
+%% conditional independence of the programs conditioned on the DSL, and
+%% so alternate between estimating the DSL ($\mathcal{D},\theta$) and
+%% estimating a posterior distribution over $p$ by searching for
+%% programs.  In practice searching for
+%% programs is difficult, so we use the recognition model to
+%% learn how to search efficiently.
 
@@
 
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.west);
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Provides}\\ \footnotesize{inductive bias}} (synthesis.west);
     \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.east);

### 2018-02-07 paper.tex
wake sleep improvements

@@
 We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
-which alternates between ``waking'', or training a generative model (our $\mathcal{D},\theta$),
-and  ``sleeping'', or training a recognition model on samples from the generative model (our $q$).
+which alternates between ``waking'', which draws samples from a recognition model (our $q$);
+and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
 For us, a wake cycle means using the recognition model to search for programs that solve the tasks,
@@
 
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45]{{\footnotesize Trains}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45]{{\footnotesize Trains}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above]{{\footnotesize Trains}} (recognitionModel.west);
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Wake)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.west);
 

### 2018-02-07 paper.tex
progress on describing baseline comparisons

@@
 Our goal  is to induce a DSL and find good programs solving each of the tasks.
-Our strategy is to iterate through three steps: (1) searching for programs that
-solve the tasks, (2) learning a better neural recognition model -- which we use
+Our strategy is to iterate through three steps: (1) a \emph{wake} phase where we search for programs, and then a pair of \emph{sleep} phases: (2) learning a neural recognition model -- which we use
 to accelerate the search over programs -- and (3) improving the DSL.
-The key observation here is that each of these three steps can bootstrap off each other:
-
-\begin{description}
-\item[ Searching for programs:]  Our program search  is informed by both the DSL and the recognition model.
+The key observation here is that these three steps bootstrap off each other:
+
+\noindent \textbf{Searching for programs:}  Our program search  is informed by both the DSL and the recognition model.
 When these improve we find better programs.
-  \item[ Learning a recognition model:] The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+
+\noindent\textbf{Learning a recognition model:} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-  \item[Improving the DSL:] We induce a DSL from the programs we have found so far which solve the tasks;
+
+\noindent\textbf{Improving the DSL:} We induce a DSL from the programs we have found so far which solve the tasks;
   as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
-\end{description}
-
-The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
-learning a generative model and training a recognition model on samples from the generative model.
-We apply this idea to program induction, motivating why we call our algorithm the \system.
+
+%% The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
+%% learning a generative model and training a recognition model on samples from the generative model.
+%% We apply this idea to program induction, motivating why we call our algorithm the \system.
 Figure~\ref{feeding} diagrams how these three steps feed off of each other.
@@
     \end{tikzpicture}
-\caption{Our algorithm solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others.}  \label{feeding}
+\caption{Our algorithm solves for programs, the DSL, and a recognition model.
+Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
   \end{figure}
@@
 is to assign high probability to programs $p$ where
-$\probability[x,p|\mathcal{D},\theta]$ is large.  With $q$ in hand we can find programs for
+$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
+in the frontiers will most increase $\lowerBound$.  With $q$ in hand we can find programs for
 the frontier $\mathcal{F}_x$ by searching for programs that maximize
@@
 sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
-We have two reasons:% for using the simple strategy of enumerating programs:
-\begin{itemize}
-\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
-\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
+We have two reasons:
+(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
+(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
   the space of of programs.
-\end{itemize}
+
 A main drawback of an enumerative search algorithm is that we have no
@@
 as well as 2 or 4 input multiplexers.
-This experimental setup is an extension of one introduced in the original EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
+This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
 For this domain our recognition model is a simple multilayer perceptron
@@
 
-\begin{figure}
-  \includegraphics[width = \columnwidth]{figures/circuit.png}
-  \caption{\system learning to build Boolean circuits. We start out with only NAND gates and learn DSL that contains ...}\label{circuitLearningCurve}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width = \columnwidth]{figures/circuit.png}
+%%   \caption{\system learning to build Boolean circuits. We start out with only NAND gates and learn DSL that contains ...}\label{circuitLearningCurve}
+%% \end{figure}
 
@@
 
+The neural recognition model accelerates inference,
+but given enough iterations of our algorithm,
+a lesioned version without the recognition model eventually comes close to catching up (red curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
+We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}):
+No NN, which lesions $q$;
+RF/DC,
+which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
+a $q$ only from samples from the fixed generative model --
+this is how RobustFill and DeepCoder are trained;
+PCFG, which lesions $q$, fixes $\mathcal{D}$, and learns only $\theta$ (so it is analogous to learning the parameters of a PCFG);
+and Enum, which does no learning and just enumerates a large frontier.
 \begin{figure}\centering
-  \includegraphics[width = 7cm]{figures/symbolicRegression.png}
-  \caption{DSL learning curve for symbolic regression.}\label{regressionLearningCurve}
+  \includegraphics[width = 7cm]{figures/regression.eps}
+  \caption{Learning curves for symbolic regression.
+Green: \system.
+Red: no neural network.
+Notice how the model without the neural network eventually almost catches up.}\label{regressionLearningCurve}
 \end{figure}
+
+\begin{table}
+\begin{tabular}{lllll}
+  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
+  98\% solved&94\% \\
+  $2.71\pm 0.07$ sec& $2.84\pm0.07$
+  \\\bottomrule
+\end{tabular}\caption{\% held out tasks solved and average solve time for symbolic regression}\label{regressionComparison}
+\end{table}
 
@@
 
-We automatically generated  600 string editing problems,
-along with input/output examples (Figure~\ref{exampleTextProblem}).
+We automatically generated  600 string editing along with input/output examples (Figure~\ref{exampleTextProblem}), split into a 75/25 train/test split.
+
+\begin{table}
+\begin{tabular}{lllll}
+  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
+  99\% solved\\
+  $0.88\pm 0.04$ sec
+  \\\bottomrule
+\end{tabular}\caption{\% held out tasks solved and average solve time for string edit problems}
+\end{table}
 
@@
   \end{table}
-\begin{figure}
-
-  \includegraphics[width = 7cm]{figures/text.png}
-  \caption{Learning curves for string editing programs}\label{textLearningCurve}
-
-\end{figure}
+%% \begin{figure}
+
+%%   \includegraphics[width = 7cm]{figures/text.png}
+%%   \caption{Learning curves for string editing programs}\label{textLearningCurve}
+
+%% \end{figure}
 

### 2018-02-07 supplement.tex
more improvements to the introduction
progress on describing baseline comparisons

@@
 Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
-A detailed exposition of Hindley-Milner is beyond the scope of this work.
+Our generative model over programs performs Hindley-Milner type inference during sampling;
+\texttt{unify} in the generative model uses the machinery of Hindley-Milner to
+ensure that the generated programs have valid polymorphic types.
+A satisfactory exposition of Hindley-Milner is beyond the scope of this paper,
+but~\cite{pierce} offers a nice overview of lambda calculus and typing systems like Hindley-Milner.
 
@@
 %%   \end{figure}
+
+\section{Neural recognition model architecture}
+
+The neural recognition model regresses from an observation (set of input/output pairs: $\left\{(i_n,o_n) \right\}_{n\leq N}$) to a $|\mathcal{D}| + 1$ dimensional vector.
+Each input/output pair is processed by an identical encoder network;
+the outputs of the encoders are max pooled and passed to an MLP with 1 hidden layer, 16 hidden units, and a ReLU activation:
+\begin{align}
+  q(x) = \text{MLP}\left(\text{MaxPool}\left(\left\{\text{encoder}\left(i_n,o_n \right) \right\}_{n\leq N} \right) \right)
+\end{align}
+For the Boolean circuits and symbolic regression domains,
+the inputs and outputs have a fixed dimensionality,
+and the set of possible inputs is prespecified.
+For these fixed-dimensionality domains our encoder is an MLP with 1 hidden layer having 16  units and a ReLU activation function.
+The input to the encoder is
+the sequence of program outputs $\left\{o_n \right\}_{n\leq N}$.
+
+For the string editing and list domains,
+the inputs and outputs are sequences.
+Our encoder for these domains is a bidirectional GRU with 16 hidden units that reads each input/output pair; we concatenate the input and output along with a special delimiter
+symbol between them.
+We MaxPool the final hidden unit activations in the GRU along both passes of the bidirectional GRU.
+
+\section{Why not the ELBO bound?}
 
@@
 
-Now what we are actually after is the parameters that maximize the joint log probability of the data+parameters, which I will write $J$:
+Now what we are actually after is the parameters that maximize the joint log probability of the data and parameters $\theta$, which we  write $J$:
 \begin{align}
@@
 \end{align}
-So you add the pseudocounts to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$).
+We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$).
 

### 2018-02-07 paper.tex
more wake/sleep

@@
 inference
-scheme~\cite{le2016inference,ritchie2016deep} for the generative
+scheme~\cite{le2016inference} for the generative %ritchie2016deep
 model.
@@
 to accelerate the search over programs -- and (3) improving the DSL.
-The key observation here is that these three steps bootstrap off each other:
+The key observation here is that these three steps bootstrap off each other (Figure~\ref{feeding}):
 
@@
 %% We apply this idea to program induction, motivating why we call our algorithm the \system.
-Figure~\ref{feeding} diagrams how these three steps feed off of each other.
+
 \begin{figure}\centering
@@
 %%   \end{align}
-Mechanistically, \system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability.
-We maximize $\lowerBound$ by alternatingly maximizing w.r.t. the frontiers and the DSL:
+\system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
+alternatingly maximizing w.r.t. the frontiers (``wake'') and the DSL (``sleep''):
 %So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
@@
 Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by learning a neural recognition model:
+of the large combinatorial search space. We ease this difficulty by learning a neural recognition model
+during a second ``sleep'' phase:
 
@@
 
-\subsection{Searching for Programs}\label{explorationSection}
+\subsection{Wake Phase: Searching for Programs}\label{explorationSection}
 
@@
 
-\noindent\textbf{Definition.}
+\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
 A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-\\\noindent\textbf{Definition.}
+%\\\noindent\textbf{Definition.}
 A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
 one number for each DSL primitive $e:\tau\in \mathcal{D}$, written $\theta_e$,
-and a weight controlling the probability of a variable occurring in a program, written $\theta_{\text{var}}$.
+and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
 
@@
 Algorithm~\ref{programGenerativeModel}; we wrap the depth-first search
-in iterative deepening to (approximately) build $\lambda$-calculus
+in iterative deepening to build $\lambda$-calculus
 expressions in order of their probability.
@@
 
-\subsection{Learning a Neural Recognition Model}\label{recognitionSection}
+\subsection{Sleep Phase I: Learning a Recognition Model}\label{recognitionSection}
 
@@
 
-\subsection{Inducing the DSL from the Frontiers}\label{grammarInductionSection}
+\subsection{Sleep Phase II: Inducing the DSL}\label{grammarInductionSection}
 
@@
 Intuitively, we want the algorithm to
-look at the programs in the frontiers and
-generalize beyond them; 
-not only so the DSL can better express the current solutions,
-but also so that the DSL might expose new abstractions
+look at  the frontiers and
+generalize beyond them, 
+both so the DSL can better express the current solutions,
+and  also so that the DSL might expose new abstractions
 which will later be used to
-discover even better programs.
+discover more programs.
 

### 2018-02-07 paper.tex
more sleep

@@
     \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Wake)}} (recognitionModel.150);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.150);
     \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.west);

### 2018-02-07 paper.tex
justification of Helmholtz objective

@@
 like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
-We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the \system objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
+We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
 \begin{align}
@@
 \end{align}
+The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
+all of our experiments train \system on only a few hundred tasks, which is too little for
+a high-capacity $q$.
+Once we have bootstrapped a $(\mathcal{D},\theta)$,
+we can draw unlimited samples from $(\mathcal{D},\theta)$,
+and train $q$ on those samples.
 

### 2018-02-07 paper.tex
at its two recognition model section

@@
 
-We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence:
-\begin{equation*}
-  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]
-\end{equation*}
+We want a recognition model that closely approximates the true posteriors over programs.
+We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]$,
 which is equivalent to maximizing
@@
 where the expectation is taken over tasks. One could take this expectation
-over the empirical distribution of the observations,
+over the observed empirical distribution of tasks,
 like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.

### 2018-02-07 paper.tex
baseline data for symbolic regression
symbolic regression data

@@
   \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
-  98\% solved&94\% \\
-  $2.71\pm 0.07$ sec& $2.84\pm0.07$
+  98\% solved&94\%&0.7\%&0\%&2\% \\
+  $2.71$ sec& $2.84$  &$3.6$ sec&--&\\
+  $9.3$ nats &$11$ nats&$2.3$ nats&--&$4.6$ nats
   \\\bottomrule
-\end{tabular}\caption{\% held out tasks solved and average solve time for symbolic regression}\label{regressionComparison}
+\end{tabular}\caption{Symbolic regression.
+Top: \% held out tasks solved w/ 5 second timeout.
+Middle: average solve time.
+Bottom: average description length $-\log \probability[x|p]$.
+``No NN'' solves almost as many tasks but
+using worse programs w/ extraneous degrees of freedom, increasing  description length.}\label{regressionComparison}
 \end{table}

### 2018-02-07 paper.tex
included the last bit of symbolic regression data

@@
   98\% solved&94\%&0.7\%&0\%&2\% \\
-  $2.71$ sec& $2.84$  &$3.6$ sec&--&\\
+  $2.7$ sec& $2.8$ sec  &$3.6$ sec&--&$2.1$ sec\\
   $9.3$ nats &$11$ nats&$2.3$ nats&--&$4.6$ nats

### 2018-02-07 paper.tex
changes to list section

@@
 
-\begin{table}
-
+\begin{table}[t]
   \centering
@@
 \subsection{List functions}
-Our list function domain consists of tasks which are solved by functions
-that take as input an integer or a list of integers, and have as output
-either a Boolean, an integer, or a list of integers. Examples of these
-functions are in Table~\ref{listexamples}. For each function, we create a
-task $x$ by generating 15 input/output examples used for testing whether a
-program produces the correct output. Supplying many examples reduces
-ambiguity in the task's function, ensuring solutions achieve the desired
-concept.
+Our list function domain consists of hand-designed concepts from which we
+sample tasks. All tasks are solved by functions that take as input an
+integer or a list of integers, and have as output either a Boolean, an
+integer, a list of Booleans, or a list of integers.  Examples of these
+tasks are in Table~\ref{listexamples}.
 
@@
   has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
-  has-4 & [4\, 5\, 7\, 4] & \texttt{true} \\
+  has-head-in-tail & [4\, 5\, 7\, 4] & \texttt{true} \\
   repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
@@
 \end{tabular}
-\caption{Examples from the domain of list functions.}
+\caption{Examples of tasks from our list function domain. These are simpler
+  than most other tasks in our dataset which are too complex to warrant
+  description here.}
 \label{listexamples}
@@
 
-We supply \system with the DSL outlined in Table~\ref{listdsl}.
-
-\begin{table}
-\centering
-\begin{tabular}{| l | l |}
-  \hline
-  \emph{name} & \emph{type} \\
-  \hline
-    empty & \texttt{[$\alpha$]} \\
-    singleton & \texttt{$\alpha$ $\rightarrow$ [$\alpha$]} \\
-    concat & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
-    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$]} \\
-    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$} \\
-    \hline
-    true & \texttt{bool} \\
-    not & \texttt{bool $\rightarrow$ bool} \\
-    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
-    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
-    \hline
-    0, $\ldots$, 9 & \texttt{int} \\
-    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
-    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
-    negate & \texttt{int $\rightarrow$ int} \\
-    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
-    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
-    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
-    is-prime & \texttt{int $\rightarrow$ bool} \\
-    is-square & \texttt{int $\rightarrow$ bool} \\
-    range & \texttt{int $\rightarrow$ [int]} \\
-    sort & \texttt{[int] $\rightarrow$ [int]} \\
-    \hline
-    sum & \texttt{[int] $\rightarrow$ int} \\
-    reverse & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$]} \\
-    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
-    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
-    index & \texttt{int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$} \\
-    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
-    slice & \texttt{int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
-  \hline
-\end{tabular}
-\caption{DSL for the domain of list function.}
-\label{listdsl}
-\end{table}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figures/list.eps}
-\end{figure}
-
-We found that using a less sophisticated but equally-capable DSL made common
-patterns, such as summation, unlikely and unlearnable in a small enumeration
-bound.
+We take approaches from two different initial DSLs, \emph{base} and
+\emph{rich}, and assess their performance. The base DSL starts with only
+low-level primitives such as \texttt{mapi}, \texttt{reducei}, and
+\texttt{if}, whereas the rich DSL includes some common structure that can be
+built from these such as \texttt{all}, \texttt{filter}, and \texttt{slice}.
+Both are theoretically capable of solving all tasks supplied by our dataset.
+See Supplement for more detail on these DSLs.
+
+% TODO: table of list results, show some induced primitives
 

### 2018-02-07 supplement.tex
changes to list section

@@
   & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
-\end{align}
+\end{align*}
 where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.
@@
 
+\vfill\null\columnbreak
+
+\section{Initial domain-specific languages}
+\subsection{List functions}
+DSL for the list function experiment domain is shown below. The lower
+section with seven primitives designates those given to the \emph{rich} DSL
+variant and not the \emph{base} variant. The rich variant is not supplied
+the \texttt{if} primitive, because its primary intention is to enable the
+learning of the \texttt{filter}, \texttt{index}, and \texttt{slice}
+concepts.
+
+\begin{center}
+\tt
+\begin{tabular}{| l | l |}
+  \hline
+  \textrm{\emph{name}} & \textrm{\emph{type}} \\
+  \hline
+    empty & [$\alpha$] \\
+    singleton & $\alpha$ $\rightarrow$ [$\alpha$] \\
+    concat & [$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
+    mapi & (int $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) \\
+    & \qquad\qquad$\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$] \\
+    reducei & ($\beta$ $\rightarrow$ int $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$)\\
+    & \qquad\qquad$\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$ \\
+    \hline
+    true & bool \\
+    not & bool $\rightarrow$ bool \\
+    and & bool $\rightarrow$ bool $\rightarrow$ bool \\
+    or & bool $\rightarrow$ bool $\rightarrow$ bool \\
+    if & bool $\rightarrow$ $\alpha$ $\rightarrow$ $\alpha$ $\rightarrow$ $\alpha$ \\
+    \hline
+    0 $\cdots$ 5 & int \\
+    + & int $\rightarrow$ int $\rightarrow$ int \\
+    * & int $\rightarrow$ int $\rightarrow$ int \\
+    negate & int $\rightarrow$ int \\
+    mod & int $\rightarrow$ int $\rightarrow$ int \\
+    eq? & int $\rightarrow$ int $\rightarrow$ bool \\
+    gt? & int $\rightarrow$ int $\rightarrow$ bool \\
+    is-prime & int $\rightarrow$ bool \\
+    is-square & int $\rightarrow$ bool \\
+    range & int $\rightarrow$ [int] \\
+    sort & [int] $\rightarrow$ [int] \\
+    \hline
+    sum & [int] $\rightarrow$ int \\
+    reverse & [$\alpha$] $\rightarrow$ [$\alpha$] \\
+    all & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool \\
+    any & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool \\
+    index & int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$ \\
+    filter & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
+    slice & int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
+  \hline
+\end{tabular}
+\end{center}
+
 \bibliography{main}

### 2018-02-07 paper.tex
conclusion

@@
 This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
-\begin{table}
+\begin{table}\centering
 \begin{tabular}{l}
@@
  \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
- focused on engineering neural networks that regress from
- input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to a lesioned version of
- \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned.
+ focused on creating neural networks that regress from
+ input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to the lesioned version of
+ \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned (RF/DC baseline).
+ These neural networks are typically trained with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
+ In our work we consider a weakly-supervised setting where ground truth programs are not provided and
+ the agent must learn from a few hundred examples.
 
@@
  and our contribution here is a general scheme for getting BPL off the ground with less hand-engineering of the DSL.
+
+ \section{Conclusion}
+ We have contributed a model, \system, that learns to program by
+ bootstrapping a domain-specific language for expressing the solutions
+ to the tasks at hand.  Alongside this DSL, we train a bottom-up
+ neural recognition model that learns how to efficiently deploy the
+ DSL and hone in on the desired programs.  We believe that this
+ integration of top-down symbolic representation and bottom-up neural
+ networks -- both of them learned -- is a useful general approach for
+ getting program induction off the ground.
+
+\pagebreak
 

### 2018-02-07 paper.tex
more stuff for text domain

@@
 and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
-For us, a wake cycle means using the recognition model to search for programs that solve the tasks,
-while a sleep cycle means updating the DSL and the recognition model
+For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
+The wake cycle can be thought of as a kind of active problem solving,
+guided by the current generative and recognition models.
+A sleep cycle means updating these models so that the DSL and  recognition model
 to more closely match the distribution of tasks.
@@
  and show that our algorithm discovers
- its own domain-specific vocabulary for expressing solutions in each of these domains (Table~\ref{initialExampleDSL}).
+ its own domain-specific vocabulary for expressing solutions the domain (Table~\ref{initialExampleDSL}).
  \begin{table}
@@
 Instead, we gave \system only a NAND gate,
-and tasked it with building a set of 500 random circuits made out of
+and tasked it with building a set of 250 random circuits made out of
 AND, NOT, and OR gates,
@@
 This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
-For this domain our recognition model is a simple multilayer perceptron
-whose inputs are the input/output examples.
-Figure~\ref{circuitLearningCurve} shows our algorithm learning to build circuits; Table~\ref{circuitExperimentTable}
+%% For this domain our recognition model is a simple multilayer perceptron 
+%% whose inputs are the input/output examples.
+\system learns to build these
+circuits by inventing new subroutines that correspond to
+logic gates and multiplexers (Figure~\ref{circuitTables}, top).
+We compare  with four baselines on 250 held-out circuits:
+No NN, which lesions $q$;
+RF/DC,
+which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
+a $q$ only from samples from the fixed generative model --
+this is how RobustFill and DeepCoder are trained;
+PCFG, which lesions $q$, fixes $\mathcal{D}$, and learns only $\theta$ (so it is analogous to learning the parameters of a PCFG);
+and Enum, which does no learning and just enumerates a large frontier.
+\begin{figure}
+\begin{tabular}{lllll}
+  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
+  84\% solved&76\%& \\
+  $0.9$ sec&  1.1
+  \\\bottomrule
+\end{tabular}
+\end{figure}
 
@@
 but given enough iterations of our algorithm,
-a lesioned version without the recognition model eventually comes close to catching up (red curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
-We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}):
-No NN, which lesions $q$;
-RF/DC,
-which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
-a $q$ only from samples from the fixed generative model --
-this is how RobustFill and DeepCoder are trained;
-PCFG, which lesions $q$, fixes $\mathcal{D}$, and learns only $\theta$ (so it is analogous to learning the parameters of a PCFG);
-and Enum, which does no learning and just enumerates a large frontier.
+a lesioned version without the recognition model eventually comes close to catching up (green curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
+We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}).
 \begin{figure}\centering
@@
   \caption{Learning curves for symbolic regression.
-Green: \system.
-Red: no neural network.
-Notice how the model without the neural network eventually almost catches up.}\label{regressionLearningCurve}
+Red: \system.
+Green: no neural network.
+Notice how the model without the neural network eventually solves almost all the tasks, but does so using programs with lower likelihood (see Table~\ref{regressionComparison}).}\label{regressionLearningCurve}
 \end{figure}
@@
 We show in this experiment that we can instead start out with a
-more restricted set of primitives -- close to the string routines provided by Python --
+less specialized set of primitives -- close to the string routines provided by Python --
 and automatically recover many of the higher-level building blocks that have made these
@@
 
-We automatically generated  600 string editing along with input/output examples (Figure~\ref{exampleTextProblem}), split into a 75/25 train/test split.
+We automatically generated  600 string editing along with input/output examples (Figure~\ref{exampleTextProblem}), split  75/25 train/test.
 
@@
       $f_2($\texttt{s,c}$) = $\texttt{((}$f_0$\texttt{ c) (}$f_1$\texttt{ (1+ 0) c s))} \\
-      \hspace{1cm}($f_2$: \emph{Extract substring delimited by character})
+      \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
+      $f_3($\texttt{c,s}$) = $\texttt{(nth (1+ 0) (split c (nth 0 s)))}\\
+      \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
     %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
@@
 \begin{figure}
-  \begin{tabular}{ll}\toprule
+  \centering\begin{tabular}{ll}\toprule
     Input&Output\\\midrule
-    kIDY eMSz,zON& 	eM\\
-KCQ njp,yzhq 	&nj\\
-zeChf zzZ,dtFP 	&zz\\
-plbKi fmLcO,ykkOR 	&fm    \\\bottomrule
+cqbHT$>$aIKd$<$szW& 	AIKD\\
+YBW$>$TIg$<$WSho 	&TIG\\
+DkI$>$DwKI$<$xaPoc& 	DWKI\\
+KfvTh$>$Dbl$<$OIJso 	&DBL
+    \\\bottomrule
   \end{tabular}
-  \caption{Example string edit problem (left) and the program that our system recovers (right)}\label{exampleTextProblem}
+
+  \vspace{0.5cm}
+
+  \begin{minipage}{7.5cm}
+ $f($\texttt{s}$) = $   \texttt{(}$f_3$\texttt{ (split '>' (upper s)) '<')}
+    \end{minipage}
+  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Table~\ref{textPrimitives}}\label{exampleTextProblem}
   \end{figure}

### 2018-02-07 paper.tex
through a random circuit into the paper

@@
 \usepackage{tikz}
+\usepackage{circuitikz}
 \usetikzlibrary{fit,bayesnet}
@@
 
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Provides}\\ \footnotesize{inductive bias}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white]{{\footnotesize Makes tractable}} (synthesis.east);
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
   \end{scope}
@@
 as well as 2 or 4 input multiplexers.
+\system observes the input/output truth table of the circuits.
 This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
@@
   \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
-  84\% solved&76\%& \\
-  $0.9$ sec&  1.1
+  84\% solved&76\%& 63\%&62\%&62\%\\
+  $0.9$ sec&  1.1&1.2&1.3&1.3
   \\\bottomrule
 \end{tabular}
+\caption{Top: \% held-out Boolean circuits solved by \system and baselines w/ 5sec timeout. Bottom: average search time.}
+\end{figure}
+
+\begin{figure}
+  \begin{circuitikz} \draw
+(0,2) node[and port] (myand1) {}
+(0,0) node[and port] (myand2) {}
+(2,1) node[xnor port] (myxnor) {}
+(myand1.out) -- (myxnor.in 1)
+(myand2.out) -- (myxnor.in 2);
+\end{circuitikz}
   \end{figure}

### 2018-02-07 paper.tex
shrink font size of code

@@
 \newcommand{\lowerBound}{\mathscr{L}}
+\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
 % Use the following line for the initial blind version submitted for review:
@@
    Domain&Part of the learned DSL\\\midrule
-   Circuits&\texttt{(nand (nand a a) (nand b b))}\\
+   Circuits&\code{(nand (nand a a) (nand b b))}\\
    &\hspace{0.5cm} \emph{(an OR gate)}\\
-   Regression& \texttt{(+ (* real x) real)} \\
+   Regression& \code{(+ (* real x) real)} \\
       &\hspace{0.5cm} \emph{(a linear function of x)}\\
-   Strings& \texttt{(join '' (split c s))}\\
+   Strings& \code{(join '' (split c s))}\\
    &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
-   Lists& \texttt{(map (lambda (x) (+ x a)) l)}\\
+   Lists& \code{(map (lambda (x) (+ x a)) l)}\\
    &\hspace{0.5cm} \emph{(add a to every element of list l)}
@@
 For example, to describe the type of the identity function
-we would say \texttt{(lambda (x) x)}$:\alpha\to \alpha$.
+we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
 We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
@@
   \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
-   \STATE \textbf{return} \texttt{(lambda (}var\texttt{) }body\texttt{)}
+   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
    \ENDIF
@@
  \ENDFOR
- \STATE \textbf{return} \texttt{(}$e\;a_1\; a_2\; \cdots\; a_K$\texttt{)}
+ \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
  \STATE\textbf{where:}
@@
     \begin{tabular}{l}
-      \texttt{(lambda x (nand x x))}\\
-      \texttt{(lambda x (lambda y (nand x (nand y y))}
+      \code{(lambda x (nand x x))}\\
+      \code{(lambda x (lambda y (nand x (nand y y))}
     \end{tabular}&
-    \texttt{(nand z z)}\\\midrule
+    \code{(nand z z)}\\\midrule
     Strings&
     \begin{tabular}{l}
-      \texttt{(lambda s (+ ',' (index 0 (split ',' s))))}\\
-      \texttt{(lambda s (index 0 (split ' ' s)))}
+      \code{(lambda s (+ ',' (index 0 (split ',' s))))}\\
+      \code{(lambda s (index 0 (split ' ' s)))}
     \end{tabular}&
-    \texttt{(index 0 (split z s))}\\\bottomrule
+    \code{(index 0 (split z s))}\\\bottomrule
   \end{tabular}
   \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
-    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} and \texttt{s} in the right column).} 
+    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\code{z} and \code{s} in the right column).} 
   \label{fragmentExample}
@@
 for example, it represents the linear function $-x+2$ with the program 
-\texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
+\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
 This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
@@
   %% \\\midrule
-    $f_0 = $\texttt{(+ }$\mathcal{R}$\texttt{)}\\
-    $f_1($\texttt{x}$) =$ \texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)} \\
-      $f_2($\texttt{x}$) =$ \texttt{(+ }$\mathcal{R}$\texttt{ ($f_1$ x))} \\
-      $f_3($\texttt{x}$) = $\texttt{(* x (}$f_2$\texttt{ x))}\\
-  $f_4($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_3$\texttt{ x)))}\\
-  $f_5($\texttt{x}$) =$ \texttt{(* x (}$f_4$\texttt{ x))}\\
-  $f_6($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_4$\texttt{ x))} (\emph{4th order polynomial})
+    $f_0 = $\code{(+ }$\mathcal{R}$\code{)}\\
+    $f_1($\code{x}$) =$ \code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} \\
+      $f_2($\code{x}$) =$ \code{(+ }$\mathcal{R}$\code{ ($f_1$ x))} \\
+      $f_3($\code{x}$) = $\code{(* x (}$f_2$\code{ x))}\\
+  $f_4($\code{x}$) = $\code{(}$f_0$\code{ (}$f_3$\code{ x)))}\\
+  $f_5($\code{x}$) =$ \code{(* x (}$f_4$\code{ x))}\\
+  $f_6($\code{x}$) = $\code{(}$f_0$\code{ (}$f_4$\code{ x))} (\emph{4th order polynomial})
   \\\bottomrule
@@
 or replacing one character with another.
-Instead, we provide array/string indexing primitives (\texttt{slice}, \texttt{nth}),
-basic arithmetic, and functions for splitting a string into a list (\texttt{split})
-or turning a list into a string (\texttt{join}).
+Instead, we provide array/string indexing primitives (\code{slice}, \code{nth}),
+basic arithmetic, and functions for splitting a string into a list (\code{split})
+or turning a list into a string (\code{join}).
 
@@
     \toprule
-    \texttt{0}, \texttt{+1},\texttt{-1}\\
-    \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
-    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{nth}\\
-    \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
-    \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}  
+    \code{0}, \code{+1},\code{-1}\\
+    \code{split}, \code{join}, \code{map}, \code{length}, \code{slice}\\
+    \code{append}, \code{chr->str}, \code{trim}, \code{nth}\\
+    \code{lowercase}, \code{uppercase}, \code{capitalize}\\
+    \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  
   \\\bottomrule
@@
       \toprule
-      $f_0($\texttt{c}$) = $\texttt{(+ (chr->str c))}\\
-      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s))}\\\hspace{2.55cm}\texttt{ (chr->str c))}\\
-      $f_2($\texttt{s,c}$) = $\texttt{((}$f_0$\texttt{ c) (}$f_1$\texttt{ (1+ 0) c s))} \\
+      $f_0($\code{c}$) = $\code{(+ (chr->str c))}\\
+      $f_1($\code{i,c,s}$) = $      \code{(+ (nth i (split c s))}\\\hspace{2.55cm}\code{ (chr->str c))}\\
+      $f_2($\code{s,c}$) = $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
       \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
-      $f_3($\texttt{c,s}$) = $\texttt{(nth (1+ 0) (split c (nth 0 s)))}\\
+      $f_3($\code{c,s}$) = $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
       \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
-    %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
-    %% $f_1 = $\texttt{(join "")}
+    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
+    %% $f_1 = $\code{(join "")}
   \\\bottomrule
@@
   \begin{minipage}{7.5cm}
- $f($\texttt{s}$) = $   \texttt{(}$f_3$\texttt{ (split '>' (upper s)) '<')}
+ $f($\code{s}$) = $   \code{(}$f_3$\code{ (split '>' (upper s)) '<')}
     \end{minipage}
@@
   range & 3 & [1\, 2\, 3] \\
-  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
-  has-head-in-tail & [4\, 5\, 7\, 4] & \texttt{true} \\
+  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
+  has-head-in-tail & [4\, 5\, 7\, 4] & \code{true} \\
   repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
@@
 \emph{rich}, and assess their performance. The base DSL starts with only
-low-level primitives such as \texttt{mapi}, \texttt{reducei}, and
-\texttt{if}, whereas the rich DSL includes some common structure that can be
-built from these such as \texttt{all}, \texttt{filter}, and \texttt{slice}.
+low-level primitives such as \code{mapi}, \code{reducei}, and
+\code{if}, whereas the rich DSL includes some common structure that can be
+built from these such as \code{all}, \code{filter}, and \code{slice}.
 Both are theoretically capable of solving all tasks supplied by our dataset.

### 2018-02-08 paper.tex
circuit figure

@@
 circuits by inventing new subroutines that correspond to
-logic gates and multiplexers (Figure~\ref{circuitTables}, top).
+logic gates and multiplexers (Figure~\ref{multiplexer}, top).
 We compare  with four baselines on 250 held-out circuits:
@@
 
-\begin{figure}
+\begin{figure}\centering
   \begin{circuitikz} \draw
-(0,2) node[and port] (myand1) {}
-(0,0) node[and port] (myand2) {}
-(2,1) node[xnor port] (myxnor) {}
-(myand1.out) -- (myxnor.in 1)
-(myand2.out) -- (myxnor.in 2);
+    (0,0) node[port] (x) {$x$}
+    (x)++(0,-0.3) node[port] (y) {$y$}
+    (y)++(0,-0.5) node[port] (z) {$z$}
+
+    (x)++(1.7,-0.15) node[nand port, scale = 0.5] (xy) {}
+    (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
+    (xz) ++ (1,0.2) node[nand port, scale = 0.5] (zxz) {}
+    (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
+    (x.out) -- (xy.in 1)
+    (y.out) -- (xy.in 2)
+
+    (x.out) -- (xz.in 1)
+    (z.out) -- (xz.in 2)
+
+    (xz.out) -- (zxz.in 2)
+    (z.out) |- (zxz.in 1)
+
+    (zxz.out) -- (m.in 2)
+    (xy.out) |- (m.in 1)
+
+    (m)++(1.2,0) node [port] (l) {\textsc{MUX}$(x,y,z)$}
+    ;
+%% (0,0) node[and port] (myand2) {}
+%% (2,1) node[xnor port] (myxnor) {}
+%% (myand1.out) -- (myxnor.in 1)
+%% (myand2.out) -- (myxnor.in 2);
   \end{circuitikz}
+\caption{\system discovers the above multiplexer circuit, which it uses to make more complicated circuits like the one below.}\label{multiplexer}
   \end{figure}

### 2018-02-08 paper.tex
small edits

@@
 is to distill the commonalities across the programs that solve the tasks,
-and it can be used to solve other related program learning problems.
+letting the agent solve other related program learning problems.
 
@@
 A sleep cycle means updating these models so that the DSL and  recognition model
-to more closely match the distribution of tasks.
+more closely match the distribution of tasks.
 This wake/sleep style of inference is the reason why we call our approach \system.
@@
 A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL primitive $e:\tau\in \mathcal{D}$, written $\theta_e$,
+one number for each DSL primitive $e\in \mathcal{D}$, written $\theta_e$,
 and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
@@
     \toprule
-    \code{0}, \code{+1},\code{-1}\\
-    \code{split}, \code{join}, \code{map}, \code{length}, \code{slice}\\
+    \code{0}, \code{+1},\code{-1},     \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  \\
+    \code{split}, \code{join}, \code{map}, 
     \code{append}, \code{chr->str}, \code{trim}, \code{nth}\\
-    \code{lowercase}, \code{uppercase}, \code{capitalize}\\
-    \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  
+    \code{length}, \code{slice},    \code{lower}, \code{upper}, \code{capitalize}
   \\\bottomrule

### 2018-02-08 paper.tex
re- factored experiments

@@
 \section{Experiments}
+We first present our experimental results qualitatively,
+illustrating the DSLs that  \system
+learns in the domains of Boolean circuits, symbolic regression, string editing, and list functions.
+We then show quantitative results (Section~\ref{quantitative})
+comparing \system with lesioned versions
+of our model that are analogous to prior work.
 
@@
 to build the Boolean circuits out of logic gates.
-Now, if we equipped our agent with a full repertoire of
+If we gave our agent the full repertoire of
 circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
@@
 logic gates and multiplexers (Figure~\ref{multiplexer}, top).
-We compare  with four baselines on 250 held-out circuits:
-No NN, which lesions $q$;
-RF/DC,
-which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
-a $q$ only from samples from the fixed generative model --
-this is how RobustFill and DeepCoder are trained;
-PCFG, which lesions $q$, fixes $\mathcal{D}$, and learns only $\theta$ (so it is analogous to learning the parameters of a PCFG);
-and Enum, which does no learning and just enumerates a large frontier.
-\begin{figure}
-\begin{tabular}{lllll}
-  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
-  84\% solved&76\%& 63\%&62\%&62\%\\
-  $0.9$ sec&  1.1&1.2&1.3&1.3
-  \\\bottomrule
-\end{tabular}
-\caption{Top: \% held-out Boolean circuits solved by \system and baselines w/ 5sec timeout. Bottom: average search time.}
-\end{figure}
 
@@
 
-    (x.out) -- (xz.in 1)
+    (x.out) -| (xz.in 1)
     (z.out) -- (xz.in 2)
@@
 
-\begin{table}
-\begin{tabular}{lllll}
-  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
-  98\% solved&94\%&0.7\%&0\%&2\% \\
-  $2.7$ sec& $2.8$ sec  &$3.6$ sec&--&$2.1$ sec\\
-  $9.3$ nats &$11$ nats&$2.3$ nats&--&$4.6$ nats
-  \\\bottomrule
-\end{tabular}\caption{Symbolic regression.
-Top: \% held out tasks solved w/ 5 second timeout.
-Middle: average solve time.
-Bottom: average description length $-\log \probability[x|p]$.
-``No NN'' solves almost as many tasks but
-using worse programs w/ extraneous degrees of freedom, increasing  description length.}\label{regressionComparison}
-\end{table}
-
 \subsection{String editing}\label{textSection}
@@
 We automatically generated  600 string editing along with input/output examples (Figure~\ref{exampleTextProblem}), split  75/25 train/test.
-
-\begin{table}
-\begin{tabular}{lllll}
-  \toprule \system&No NN  & RF/DC & PCFG & Enum\\\midrule
-  99\% solved\\
-  $0.88\pm 0.04$ sec
-  \\\bottomrule
-\end{tabular}\caption{\% held out tasks solved and average solve time for string edit problems}
-\end{table}
 
@@
 
+\subsection{Quantitative Results}\label{quantitative}
+
+We compare  with four baselines on  held-out tasks:
+\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
+\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
+a recognition model only from samples from the fixed generative model.
+This is equivalent to our algorithm with $\lambda = \infty$ (see Section~\ref{grammarInductionSection})
+and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Section~\ref{recognitionSection}).
+We call this baseline RF/DC because
+this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
+\\\noindent \textbf{PCFG}, which lesions the recognition model, fixes $\mathcal{D}$, and learns only $\theta$.
+This is like learning the parameters of a PCFG.
+\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
+This is equivalent to the first wake cycle of our \system.
+
+For each domain,
+we are interested both in how many tasks the
+agent can solve and how quickly can find those solutions.
+For symbolic regression, we also care about the quality
+of the solution, as measured by the likelihood model $\probability[x|p]$.
+Concretely, this cashes out as measuring how many
+real numbers were used in the programs: did the model correctly explain a linear function
+using two numbers, or did it introduce extraneous parameters?
+Table~\ref{baselineComparisons}
+compares our model against these baselines.
+
+\begin{table}
+%% This radically shrinks the table
+\tabcolsep=4pt
+\renewcommand{\arraystretch}{0.5}
+\begin{tabular}{llllll}
+  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum
+
+  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
+  \% solved &
+  84\% &76\%& 63\%&62\%&62\%\\
+  Solve time&  $0.9$s&  1.1&1.2&1.3&1.3
+
+  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
+  \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
+  Solve time&  $2.7$s& $2.8$s  &$3.6$s&--&$2.1$s\\
+  MDL (nats)&   $9.3$ &$11.0$&$2.3$&--&$4.6$
+
+  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
+  \% solved&99\% \\
+  Solve time&  0.9s
+  \\\bottomrule
+  \end{tabular}
+\caption{\% solved: fraction solved with 5 sec timeout.
+Solve time: averaged over solved tasks.
+MDL: $-\expect\left[\probability[x|p] \right]$.
+For domains other than symbolic regression MDL is 0 nats.
+For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}
 \pagebreak
- \subsection{Related Work}
+ \section{Related Work}
  Our work is far from the first to learn to learn programs,
@@
  \section{Conclusion}
- We have contributed a model, \system, that learns to program by
- bootstrapping a domain-specific language for expressing the solutions
- to the tasks at hand.  Alongside this DSL, we train a bottom-up
+ We contributed a model, \system, that learns to program by
+ bootstrapping a domain-specific language.
+ The  DSL learns to expose new domain-specific primitives
+  Alongside this DSL, we train a bottom-up
  neural recognition model that learns how to efficiently deploy the

### 2018-02-08 paper.tex
jo\begin{figure}

@@
 more closely match the distribution of tasks.
+An important difference from the Helmholtz machine
+is that our setup has two distinct sleep phases: a phase where we update the generative model
+$(\mathcal{D},\theta)$, which we call \textbf{G-Sleep},
+and a second phase where we update the recognition model $q$,
+which we call \textbf{R-Sleep}.
 This wake/sleep style of inference is the reason why we call our approach \system.
@@
 Our goal  is to induce a DSL and find good programs solving each of the tasks.
-Our strategy is to iterate through three steps: (1) a \emph{wake} phase where we search for programs, and then a pair of \emph{sleep} phases: (2) learning a neural recognition model -- which we use
+Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
 to accelerate the search over programs -- and (3) improving the DSL.
 The key observation here is that these three steps bootstrap off each other (Figure~\ref{feeding}):
-
-\noindent \textbf{Searching for programs:}  Our program search  is informed by both the DSL and the recognition model.
+\\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
 When these improve we find better programs.
-
-\noindent\textbf{Learning a recognition model:} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+\\\noindent\textbf{R-Sleep: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-
-\noindent\textbf{Improving the DSL:} We induce a DSL from the programs we have found so far which solve the tasks;
+\\\noindent\textbf{G-Sleep: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
   as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
@@
 
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.west);
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (G-Sleep)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (R-Sleep)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (R-Sleep)}} (recognitionModel.west);
 
@@
 Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
+\end{figure}
+
+\begin{figure}
+\begin{tikzpicture}
+
+  \begin{scope}[shift = {(0.7,0)}]  
+    \node[obs] at (3,3) (dx){$\mathcal{D}$};
+  \node[latent] at (3.5,1.75) (zp){$p$};
+  \node[obs] at (4,3) (tx){$\theta^{(x)}$};
+  \node[obs] at (3.5,0.5) (xp) {$x$};
+  \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
+  \draw [->,red] (xp.east) to[out = 0,in = 20] node(nn){} (tx.north);
+  \draw [->,red] (tx.south) -- (zp.north);
+  \draw [->] (dx.south) -- (zp.north);
+
+  \node at (nn) {
+    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle,width=2cm] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}
+  };
+  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
+  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
+  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
+
+  \end{scope}
+
+    \begin{scope}[shift = {(5,0)}]  
+      \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+      \draw [->] (p2.south) -- (x2.north);
+      \draw [->] (dt.south) -- (p2.north);
+
+      \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+      \draw [->] (p1.south) -- (x1.north);
+      \draw [->] (dt.south) -- (p1.north);
+
+      \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+      \draw [->] (p3.south) -- (x3.north);
+      \draw [->] (dt.south) -- (p3.north);
+
+      \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+      \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+
+  \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+  \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+  \draw [->] (zp.south) -- (xp.north);
+  \plate {}{(zp)(xp)}{$x\in X$};
+
+  \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
+  \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {R-Sleep: Train $q$\\training data: cyan $(x,p)$};
+    \end{scope}
+
+    \begin{scope}[shift = {(5,-6)}]  
+      \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+  \node[latent] at (1.5,3) (t){$\theta$};
+  \node[obs] at (1,1.75) (z){$p$};
+%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+  \node[obs] at (1,0.5) (x) {$x$};
+  \edge {z}{x};
+  \edge {d,t}{z};
+  \plate {}{(z)(x)}{$x\in X$};
+  \node[align = center] at ([yshift = -1cm]x.south) {G-Sleep: Induce $(\mathcal{D},\theta})$};
+    \end{scope}
+
+  \end{tikzpicture}
   \end{figure}
@@
 
-\subsection{Sleep Phase I: Learning a Recognition Model}\label{recognitionSection}
+\subsection{R-Sleep: Learning a Recognition Model}\label{recognitionSection}
 
@@
 
-\subsection{Sleep Phase II: Inducing the DSL}\label{grammarInductionSection}
+\subsection{G-Sleep: Inducing the DSL}\label{grammarInductionSection}
 

### 2018-02-08 paper.tex
included held out baseline data for text - the neural network helps!

@@
 and tasked it with building a set of 250 random circuits made out of
-AND, NOT, and OR gates,
-as well as 2 or 4 input multiplexers.
+AND, NOT, and OR gates, along with a multiplexer.
 \system observes the input/output truth table of the circuits.
@@
   \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
-  \% solved&99\% \\
-  Solve time&  0.9s
+  \% solved&99\% &82\% \\
+  Solve time&  0.9s&2.2s
   \\\bottomrule

### 2018-02-08 paper.tex
conclusion

@@
 
-The neural recognition model accelerates inference,
-but given enough iterations of our algorithm,
-a lesioned version without the recognition model eventually comes close to catching up (green curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
-We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}).
-\begin{figure}\centering
-  \includegraphics[width = 7cm]{figures/regression.eps}
-  \caption{Learning curves for symbolic regression.
-Red: \system.
-Green: no neural network.
-Notice how the model without the neural network eventually solves almost all the tasks, but does so using programs with lower likelihood (see Table~\ref{regressionComparison}).}\label{regressionLearningCurve}
-\end{figure}
+%% The neural recognition model accelerates inference,
+%% but given enough iterations of our algorithm,
+%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
+%% We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}).
 
@@
 or replacing one character with another.
-Instead, we provide array/string indexing primitives (\code{slice}, \code{nth}),
+Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
 basic arithmetic, and functions for splitting a string into a list (\code{split})
@@
 
-We automatically generated  600 string editing along with input/output examples (Figure~\ref{exampleTextProblem}), split  75/25 train/test.
+We automatically generated  600 string editing tasks along with input/output examples (Figure~\ref{exampleTextProblem}). %, split  75/25 train/test.
 
@@
 we are interested both in how many tasks the
-agent can solve and how quickly can find those solutions.
+agent can solve and how quickly it can find those solutions.
 For symbolic regression, we also care about the quality
@@
 Concretely, this cashes out as measuring how many
-real numbers were used in the programs: did the model correctly explain a linear function
+real numbers were used in the programs: e.g. did the model correctly explain a linear function
 using two numbers, or did it introduce extraneous parameters?
@@
 compares our model against these baselines.
-
+We see that our full model consistently
+beats the baselines,
+and that the recognition model 
+consistently increases the number of solved held-out tasks.
+Lesioning the recognition model also slows down the convergence of the algorithm,
+taking more wake/sleep cycles to reach a given number of tasks solved (Figure~\ref{learningCurves}).
+This data supports our view of the recognition model as a way of accelerating the search over programs.
 \begin{table}
@@
 For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}
+
+\begin{figure}\centering
+  \includegraphics[width = 7cm]{figures/regression.eps}
+  \caption{Learning curves for \system both with and without the recognition model (No NN in the above plots).
+Red: \system.
+Green: no neural network.
+Notice how the model without the neural network eventually solves almost all the tasks, but does so using programs with lower likelihood (see Table~\ref{regressionComparison}).}\label{regressionLearningCurve}
+\end{figure}
+
 \pagebreak
@@
 We find this work inspiring and motivating,
- and go beyond it along two dimensions: (1) we propose more sophisticated algorithms for
- inducing reusable structure, based on a formalism called \emph{Fragment Grammars}~\cite{tim};
+ and go beyond it along two dimensions: (1) we propose a new algorithm for
+ inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
@@
  We contributed a model, \system, that learns to program by
- bootstrapping a domain-specific language.
- The  DSL learns to expose new domain-specific primitives
+ bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
   Alongside this DSL, we train a bottom-up

### 2018-02-08 paper.tex
learning curve figure
added circuit figure

@@
   \end{circuitikz}
-\caption{\system discovers the above multiplexer circuit, which it uses to make more complicated circuits like the one below.}\label{multiplexer}
+  \begin{circuitikz} \draw
+    (0,0) node[port] (a) {$a$}
+    (a)++(0,-0.7) node[port] (b) {$b$}
+
+    (a)++(1,-0) node[nand port, scale = 0.5] (ab1) {}
+    (b)++(1,-0) node[nand port, scale = 0.5] (ab2) {}
+    (ab1)++(1,-0.35) node[nand port, scale = 0.5] (o) {}
+
+    (a.out) -| (ab1.in 1)
+    (a.out) -| (ab1.in 2)
+    (b.out) -| (ab2.in 1)
+    (b.out) -| (ab2.in 2)
+
+    (ab1.out) -- (o.in 1)
+    (ab2.out) -- (o.in 2)
+
+    (o)++(0.8,0) node [port] (l) {\textsc{OR}$(a,b)$}
+    ;
+  \end{circuitikz}
+  \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
+    (0,0) node[port] (w) {$w$}
+    (w)++(0,-0.5) node[port] (x) {$x$}
+    (x)++(0,-0.5) node[port] (y) {$y$}
+    (y)++(0,-0.5) node[port] (z) {$z$}
+
+    (z.out) -| (m.south)
+    (y.out) -| (m.west)
+    (x.out) -| (m.north)
+
+    (y)++(2.5,0.13) node[or port, scale = 0.5] (z) {}
+    (m.east) -- (z.in 2)
+    (w.east) -| (z.in 1)
+    ;
+  \end{circuitikz}
+\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right), which it was unable to build without discovering these new primitives.}\label{multiplexer}
   \end{figure}
@@
   \includegraphics[width = 7cm]{figures/regression.eps}
+  \includegraphics[width = 7cm]{figures/textLearningCurve.eps}
+  \includegraphics[width = 7cm]{figures/circuitLearningCurve.eps}  
   \caption{Learning curves for \system both with and without the recognition model (No NN in the above plots).

### 2018-02-08 paper.tex
text baseline data
supplement improvements

@@
 this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-\\\noindent \textbf{PCFG}, which lesions the recognition model, fixes $\mathcal{D}$, and learns only $\theta$.
-This is like learning the parameters of a PCFG.
+\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
+This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
+and is like learning the parameters of a PCFG while not learning any of the structure.
 \\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
-This is equivalent to the first wake cycle of our \system.
+This is equivalent to the first wake cycle of \system.
 
@@
   \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
-  \% solved&99\% &82\% \\
-  Solve time&  0.9s&2.2s
+  \% solved&99\% &82\% &9\%&24\%&18\%\\
+  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
   \\\bottomrule

### 2018-02-08 supplement.tex
supplement improvements

@@
 
-\newcommand{\system}{\textsc{HelmholtzHacker}~}
+\newcommand{\system}{\textsc{DreamCoder}~}
 
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Inducing Domain Specific Languages for Bayesian Program Learning}
+\icmltitlerunning{Bootstrapping Domain Specific Languages for Bayesian Program Learning}
 
@@
 \twocolumn[
-\icmltitle{Supplement to: Inducing Domain Specific Languages for Bayesian Program Learning}
+\icmltitle{Supplement: Bootstrapping Domain Specific Languages for Bayesian Program Learning}
 
@@
 
+If we were to maximize the ELBO bound to perform inference in
+our generative model,
+then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along an unimportant regularizing term on the DSL):
+\begin{align}
+  &  \sum_{x\in X}\expect_{p\sim Q_x}\left[\log \probability[p|\mathcal{D}^*,\theta^*] \right] \label{EM}\\
+  & Q_x(p)\triangleq \probability[p|x,\mathcal{D},\theta]
+\end{align}
+where $(\mathcal{D},\theta)$ is our current estimate of the generative
+model.  These equations fall out of an EM-style derivation, and one
+could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
+using importance sampling (so the expectation in~\ref{EM} is taken
+over $q$ and we reweight using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.
+
+How the recognition model is worked into this is unimportant to the reason why we do not maximize a bound of this form:
+this bound takes an expectation over the \emph{previous} sleep cycle's
+posterior over programs,
+and so the approximate posterior $Q_x$ at the next iteration
+ends up being very close to previous approximate posterior.
+Intuitively,
+we want the DSL induction to be a function \emph{only} of the programs that we have found,
+and \emph{not} be a function of how the previous generative model weighted them.
+In practice,
+we found that maximizing EM-style bounds, like the ELBO,
+leads to a kind of hysteresis effect,
+where the next generative model too closely matches the previous one,
+causing the algorithm to easily become trapped in local optima.
+
 \section{Estimating $\theta$}
@@
 \begin{align}
-  \frac{c(x)}{\theta^{(x)}} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
+  \frac{c(x)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
 \end{align}
-This equality holds if $\theta^{(x)} = c(x)/a(x)$:
+This equality holds if $\theta_x = c(x)/a(x)$:
 \begin{align}
@@
 We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$).
+
+When is this lower bound tight? This lower bound is tight whenever all
+of the types of the expressions in the DSL are not polymorphic, in which case our DSL is equivalent to a PCFG
+and this estimator is equivalent to the inside/outside algorithm.
+Polymorphism introduces context-sensitivity to the DSL,
+and exactly maximizing the likelihood with respect to $\theta$
+becomes intractable,
+so for domains with polymorphic types we use this estimator.
 

### 2018-02-08 paper.tex
updates for list domain

@@
 Our list function domain consists of hand-designed concepts from which we
-sample tasks. All tasks are solved by functions that take as input an
-integer or a list of integers, and have as output either a Boolean, an
-integer, a list of Booleans, or a list of integers.  Examples of these
-tasks are in Table~\ref{listexamples}.
+sample tasks, so there is no ``ground truth'' formal DSL from which these
+tasks are generated. Additionally, few of these tasks are concatenative ---
+many of them represent distinct operations. We found that this domain was
+difficult to tackle without starting from a relatively rich DSL, and even
+then it was still hard to solve all tasks. All tasks are solved by functions
+that take as input an integer or a list of integers, and have as output
+either a Boolean, an integer, a list of Booleans, or a list of integers.
+Examples of these tasks are in Table~\ref{listexamples}.
 
@@
 
-% TODO: table of list results, show some induced primitives
+\begin{table}[t]
+  \centering
+  Subset of learned DSL\\
+  \begin{tabular}{l}
+    \toprule
+    $(f_0$ \code{i $\ell$}$) = $\code{(singleton (index i $\ell$))}\\
+    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
+    $(f_1$ \code{i $\ell$}$) = $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
+    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
+    $(f_2$ \code{n $\ell$}$) = $\code{(any (lambda (x) (= n x)) $\ell$)}\\
+    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
+    $(f_3$ \code{i $\ell$}$) = $\code{(index (negate i) (sort $\ell$))}\\
+    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
+    $(f_4$ \code{n $\ell$}$) = $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
+    $(f_5$ \code{k $\ell$}$) = $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
+    $(f_6$ \code{k n $\ell$}$) = $\code{($f_4$ n ($f_5$ k $\ell$)}\\
+    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
+  \bottomrule
+  \end{tabular}
+  \caption{Some induced primitives from the list DSL}\label{listinduced}
+\end{table}
 
@@
 This data supports our view of the recognition model as a way of accelerating the search over programs.
+
 \begin{table}
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
+  \\\midrule\\\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
   \% solved &
   84\% &76\%& 63\%&62\%&62\%\\
-  Solve time&  $0.9$s&  1.1&1.2&1.3&1.3
-
-  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
+  Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s
+
+  \\\midrule\\\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
   \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
-  Solve time&  $2.7$s& $2.8$s  &$3.6$s&--&$2.1$s\\
-  MDL (nats)&   $9.3$ &$11.0$&$2.3$&--&$4.6$
-
-  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
+  Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
+  MDL (nats)&   9.3 &11.0&2.3&--&4.6
+
+  \\\midrule\\\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
   \% solved&99\% &82\% &9\%&24\%&18\%\\
   Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
+
+  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
+  \% solved&52\% &52\% &40\%&44\%&42\%\\
+  Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
+
+  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
+  \% solved&86\% &84\% &60\%&74\%&70\%\\
+  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
   \\\bottomrule

### 2018-02-08 paper.tex
tiny figure improvements
more writing for the experiment section

@@
   \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
-  \draw [->,red] (xp.east) to[out = 0,in = 20] node(nn){} (tx.north);
-  \draw [->,red] (tx.south) -- (zp.north);
-  \draw [->] (dx.south) -- (zp.north);
+  \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
+  \draw [->,red] (tx) -- (zp);
+  \draw [->] (dx) -- (zp);
 
@@
   };
-  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
-  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
-  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
+  \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
 
@@
 \subsection{Boolean circuits}
-As a toy domain, we consider the problem of learning
-to build the Boolean circuits out of logic gates.
+We consider the problem of learning
+to build  Boolean circuits out of logic gates.
 If we gave our agent the full repertoire of
@@
 AND, NOT, and OR gates, along with a multiplexer.
-\system observes the input/output truth table of the circuits.
+\system observes the input/output truth table of the circuits,
+and we define the likelihood (where $x$ is a truth table and $p$ a program implementing a circuit):
+\begin{align}
+  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]\nonumber
+  \end{align}
+
 This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
@@
 \system learns to build these
-circuits by inventing new subroutines that correspond to
-logic gates and multiplexers (Figure~\ref{multiplexer}, top).
+circuits by creating new subroutines that correspond to
+logic gates and multiplexers (Figure~\ref{multiplexer}).
+We will show data showing that the neural recognition model
+assists in building these circuits (Section~\ref{quantitative}),
+suggesting that an agent can glean something about the structure of the circuit just by looking at its truth table.
 
@@
 Concretely, we ask the algorithm to
-solve a set of 400 symbolic regression problems, each a polynomial of
-degree 0 to 4, where our observations $x$ are 10
+solve a set of 200 symbolic regression problems, each a polynomial of
+degree 1 to 4, where our observations $x$ are 10
 input/output examples.%, which we write as $x = \left\{(i_n,o_n)
@@
 We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples,
-and penalize extra continuous degrees of freedom using the BIC~\cite{Bishop:2006:PRM:1162264}:
+and penalize continuous degrees of freedom using the BIC~\cite{Bishop:2006:PRM:1162264}.
 %% \begin{align*}
@@
 other system successful.
-Table~\ref{textPrimitives} lists the primitive functions provided to \system.
+Table~\ref{textPrimitives} (top) lists the primitive functions provided to \system.
 Absent are any routines for searching for characters, extracting substrings,
@@
 or turning a list into a string (\code{join}).
-
 We automatically generated  600 string editing tasks along with input/output examples (Figure~\ref{exampleTextProblem}). %, split  75/25 train/test.
+At first, \system cannot find any programs that
+work at all for most of the tasks.
+After a few wake/sleep iterations,
+it assembles a DSL (Table~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
+all of the tasks.
 

### 2018-02-08 paper.tex
reorganized wake/sleep diagrams
add list figure before more results come in

@@
 \begin{figure}
+\centering
 \begin{tikzpicture}
-
-  \begin{scope}[shift = {(0.7,0)}]  
+  \begin{scope}[shift={(2.5,0)}]
     \node[obs] at (3,3) (dx){$\mathcal{D}$};
@@
     \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-  \node[obs] at (3.5,0.5) (xp) {$x$};
+    \node[obs] at (3.5,0.7) (xp) {$x$};
     \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
@@
     \draw [->] (dx) -- (zp);
-
     \node at (nn) {
@@
     \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-
   \end{scope}
 
-    \begin{scope}[shift = {(5,0)}]  
+  \begin{scope}[shift={(0.7,-4.2)}]
     \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
@@
 
-    \begin{scope}[shift = {(5,-6)}]  
+  \begin{scope}[shift={(7.7,-4)}]
     \node[latent] at (0.5,3) (d){$\mathcal{D}$};
@@
 \end{tabular}
-\caption{Examples of tasks from our list function domain. These are simpler
-  than most other tasks in our dataset which are too complex to warrant
-  description here.}
+\caption{Examples of simple tasks from our list function domain. We used
+  15 input/output examples for each task to reduce ambiguity in the task's
+  function, ensuring solutions achieve the desired concept.}
 \label{listexamples}
@@
 \code{if}, whereas the rich DSL includes some common structure that can be
-built from these such as \code{all}, \code{filter}, and \code{slice}.
+built from these such as \code{all}, \code{filter}, \code{slice}, and
+\code{index}.
 Both are theoretically capable of solving all tasks supplied by our dataset.
@@
   \includegraphics[width = 7cm]{figures/regression.eps}
+  \includegraphics[width = 7cm]{figures/list.eps}
   \includegraphics[width = 7cm]{figures/textLearningCurve.eps}

### 2018-02-08 paper.tex
another learning curve for strings

@@
    \STATE Unify $\tau$ with yield$(\tau')$.
-   \STATE $\left\{\alpha_k \right\}_{k\leq K}\gets\text{argTypes}(\tau')$
+   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{argTypes}(\tau')$ 
 %   \STATE unify$(\tau,\beta)$
@@
 For symbolic regression, we also care about the quality
-of the solution, as measured by the likelihood model $\probability[x|p]$.
-Concretely, this cashes out as measuring how many
-real numbers were used in the programs: e.g. did the model correctly explain a linear function
+of the solution, as measured by the likelihood model $\probability[x|p]$,
+which measures how many
+real numbers were used in the program: e.g. did the model correctly explain a linear function
 using two numbers, or did it introduce extraneous parameters?
@@
   \includegraphics[width = 7cm]{figures/circuitLearningCurve.eps}  
-  \caption{Learning curves for \system both with and without the recognition model (No NN in the above plots).
-Red: \system.
-Green: no neural network.
-Notice how the model without the neural network eventually solves almost all the tasks, but does so using programs with lower likelihood (see Table~\ref{regressionComparison}).}\label{regressionLearningCurve}
+  \caption{Learning curves for \system both with (blue) and without (red) the recognition model as the frontier size is varied (solid/dashed/dotted lines).}\label{learningCurves} 
 \end{figure}

### 2018-02-08 paper.tex
renamed sleep phases

@@
 is that our setup has two distinct sleep phases: a phase where we update the generative model
-$(\mathcal{D},\theta)$, which we call \textbf{G-Sleep},
+$(\mathcal{D},\theta)$, which we call \textbf{Sleep-G},
 and a second phase where we update the recognition model $q$,
-which we call \textbf{R-Sleep}.
+which we call \textbf{Sleep-R}.
 This wake/sleep style of inference is the reason why we call our approach \system.
@@
 When these improve we find better programs.
-\\\noindent\textbf{R-Sleep: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+\\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-\\\noindent\textbf{G-Sleep: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
+\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
   as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
@@
 
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (G-Sleep)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (R-Sleep)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (R-Sleep)}} (recognitionModel.west);
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
 
@@
     \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {R-Sleep: Train $q$\\training data: cyan $(x,p)$};
+    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
   \end{scope}
@@
     \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {G-Sleep: Induce $(\mathcal{D},\theta})$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
   \end{scope}
@@
 
-\subsection{R-Sleep: Learning a Recognition Model}\label{recognitionSection}
+\subsection{Sleep-R: Learning a Recognition Model}\label{recognitionSection}
 
@@
 
-\subsection{G-Sleep: Inducing the DSL}\label{grammarInductionSection}
+\subsection{Sleep-G: Inducing the DSL}\label{grammarInductionSection}
 

### 2018-02-08 paper.tex
miscellaneous edits

@@
 \begin{abstract}
+  Successful  approaches to program induction require a
+  hand-engineered  domain-specific languaget (DSL),
+  constraining the space of allowed programs and imparting
+  prior knowledge of the domain.
+  We contribute a program learning algorithm
+  that learns the DSL called \system,
+which infers
+  a DSL 
+  and trains a neural network to
+  efficiently search for programs in the learned DSL.
+  We use our model to build circuits,
+  edit strings, do symbolic regression,
+  and synthesize Lisp-style functions on lists,
+  in each case showing that
+  \system learns a domain-specific vocabulary for expressing programs.
 
@@
 DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
-techniques (e.g., Metagol:~\cite{muggleton2015meta} and 
-FlashFill:~\cite{gulwani2011automating}).
+techniques (e.g., Metagol:~\cite{muggleton2015meta},
+FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
 However, the success of these approaches -- both symbolic and neural --
@@
 The purpose of the DSL
-is to distill the commonalities across the programs that solve the tasks,
+is to expose the commonalities across the programs that solve tasks within a domain,
 letting the agent solve other related program learning problems.
@@
 Observe tasks $x$ (program inputs/outputs).
-(b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs
-    ($\theta^{(x)} = q(x)$). Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
+(b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
+    $\theta^{(x)}$. Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
 \end{figure}
@@
 For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
-The wake cycle can be thought of as a kind of active problem solving,
+The wake cycle can be thought of active problem solving,
 guided by the current generative and recognition models.
@@
 \system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
-alternatingly maximizing w.r.t. the frontiers (``wake'') and the DSL (``sleep''):
+alternatingly maximizing w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
 %So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
@@
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-Adding new programs to the frontiers means searching for new programs $p$ for task $x$
-where $\probability[x,p|\mathcal{D},\theta]$ is large.
+$\lowerBound$ most increases by searching for programs where $\probability[x,p|\mathcal{D},\theta]$
+is large.
+%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
+%% where  is large.
 \\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.
@@
 of the large combinatorial search space. We ease this difficulty by learning a neural recognition model
-during a second ``sleep'' phase:
+during the Sleep-R phase:
 
@@
 $\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
-in the frontiers will most increase $\lowerBound$.  With $q$ in hand we can find programs for
-the frontier $\mathcal{F}_x$ by searching for programs that maximize
-$q(p|x)$.
-The network $q$ exploits the structure of the DSL $\mathcal{D}$:
+in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
+%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
+%% $q(p|x)$.
+The recognition model exploits the structure of the DSL: %$\mathcal{D}$:
 rather than directly predicting a distribution over $p$ conditioned on $x$,
-it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
+it predicts a distribution over components of the DSL.
 This taps into the intuition that programming is primarily a top-down activity:
@@
 and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
-For the interested reader, we explain in the supplement that maximizing an ELBO-style bound
-leads to a pathological behavior that would cause the algorithm to easily become trapped in local optima.
+We explain in the supplement that maximizing an ELBO-style bound
+leads to a pathological behavior that  causes the model to easily become trapped in local optima.
 
@@
 generating programs wholesale.
+We share this aspect with the DeepCoder system~\cite{balog2016deepcoder}.
 
@@
 Although this problem may seem 
-very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
+very similar to estimating the parameters of a probabilistic context free grammar,
 for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
@@
 in the programs and also due to the polymorphic typing system.
-In the Appendix we derive a tractable MAP estimator for $\theta$.
+In the Supplement we derive a tractable MAP estimator for $\theta$.
 \begin{algorithm}[tb]
@@
 We note the following implementation details:
-(1) on the first iteration, we do \emph{not} train the recognition model on samples from
+(1) We interleave wake cycles with each of the sleep cycles;
+(2) on the first iteration, we do \emph{not} train the recognition model on samples from
 the generative model because a DSL has not yet been learned --
 we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
-(2) Because the frontiers can grow very large,
+(3) Because the frontiers can grow very large,
 we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
 with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
-(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
+(4) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
 by only summing over the top $K$
 programs in $\mathcal{F}_x$ -- we found that $K = 2$ sufficed.
-(4) For added robustness, at each iteration we enumerate programs from both the generative model and the recognition model.
+(5) For added robustness, we enumerate programs from both the generative model and the recognition model.
 
@@
 assists in building these circuits (Section~\ref{quantitative}),
-suggesting that an agent can glean something about the structure of the circuit just by looking at its truth table.
+suggesting that an agent can glean something about the structure of a circuit just by looking at its truth table.
 
@@
 \code{index}.
-Both are theoretically capable of solving all tasks supplied by our dataset.
+Both are capable of solving all tasks supplied by our dataset.
 See Supplement for more detail on these DSLs.
@@
   \% solved &
-  84\% &76\%& 63\%&62\%&62\%\\
+  \textbf{84\%} &76\%& 63\%&62\%&62\%\\
   Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s
@@
   \\\midrule\\\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
-  \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
+  \% solved&   \textbf{ 98\% }&94\%&0.7\%&0\%&2\% \\
   Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
@@
   \\\midrule\\\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
-  \% solved&99\% &82\% &9\%&24\%&18\%\\
+  \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
   Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
@@
   \\\midrule\\\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
-  \% solved&52\% &52\% &40\%&44\%&42\%\\
+  \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
   Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
@@
   \\\midrule\\\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
-  \% solved&86\% &84\% &60\%&74\%&70\%\\
+  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
   Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
@@
  focused on creating neural networks that regress from
- input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to the lesioned version of
+ input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is closest to the lesioned version of
  \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned (RF/DC baseline).
@@
 We find this work inspiring and motivating,
- and go beyond it along two dimensions: (1) we propose a new algorithm for
+ and extend it along two dimensions: (1) we propose a new algorithm for
  inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
@@
  neural recognition model that learns how to efficiently deploy the
- DSL and hone in on the desired programs.  We believe that this
+ DSL and hone in on the right programs.  We believe that this
  integration of top-down symbolic representation and bottom-up neural

### 2018-02-08 paper.tex
updated list stuff

@@
 tasks are generated. Additionally, few of these tasks are concatenative ---
-many of them represent distinct operations. We found that this domain was
-difficult to tackle without starting from a relatively rich DSL, and even
-then it was still hard to solve all tasks. All tasks are solved by functions
-that take as input an integer or a list of integers, and have as output
-either a Boolean, an integer, a list of Booleans, or a list of integers.
-Examples of these tasks are in Table~\ref{listexamples}.
+many of them represent distinct operations.
+Tasks in this domain are solved by functions that take as input an integer
+or a list of integers, and have as output either a Boolean, an integer, a
+list of Booleans, or a list of integers.  There are 225 tasks in total, some
+of which are shown in Table~\ref{listexamples}.
+
+We take approaches from two different initial DSLs, \emph{base} and
+\emph{rich}, and assess their performance. The base DSL starts with only
+low-level primitives such as \code{mapi}, \code{reducei}, and
+\code{if}, whereas the rich DSL includes some common structure that can be
+built from these such as \code{all}, \code{filter}, \code{slice}, and
+\code{index}.
+Both are capable of solving all tasks supplied by our dataset.
+See Supplement for more detail on these DSLs.
+
+We found that this domain was difficult to tackle without starting from a
+relatively rich DSL, and even then it was still hard to solve all tasks. Our
+algorithm, like human learners, requires a spectrum of problems ranging from
+easy to hard. Without this curriculum, \system cannot get off the ground.
+The behavior with the base DSL illustrates this, because the initial set of
+primitives is not paired with a sufficiently large and graded curriculum.
 
@@
 \centering
-\begin{tabular}{| l | l | l |}
-  \hline
-  \emph{name} & \emph{input} & \emph{output} \\
-  \hline
+\begin{tabular}{lll}
+  \toprule
+  Name & Input & Output \\\midrule
   add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
@@
   drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-  \hline
+  \bottomrule
 \end{tabular}
-\caption{Examples of simple tasks from our list function domain. We used
-  15 input/output examples for each task to reduce ambiguity in the task's
-  function, ensuring solutions achieve the desired concept.}
+\caption{Simple tasks from our list function domain. We used 15 input/output
+  examples for each task to reduce ambiguity in the task's function,
+  ensuring solutions achieve the desired concept.}
 \label{listexamples}
@@
 
-We take approaches from two different initial DSLs, \emph{base} and
-\emph{rich}, and assess their performance. The base DSL starts with only
-low-level primitives such as \code{mapi}, \code{reducei}, and
-\code{if}, whereas the rich DSL includes some common structure that can be
-built from these such as \code{all}, \code{filter}, \code{slice}, and
-\code{index}.
-Both are capable of solving all tasks supplied by our dataset.
-See Supplement for more detail on these DSLs.
-
 \begin{table}[t]
   \centering
-  Subset of learned DSL\\
   \begin{tabular}{l}
@@
   \end{tabular}
-  \caption{Some induced primitives from the list DSL}\label{listinduced}
+  \caption{Some induced primitives for list functions.}\label{listinduced}
 \end{table}

### 2018-02-08 paper.tex
smaller learning curve figures
reworded slightly confusing sentence in abstract

@@
   Successful approaches to program induction require a
-  hand-engineered  domain-specific languaget (DSL),
+  hand-engineered domain-specific language (DSL),
   constraining the space of allowed programs and imparting
@@
   We contribute a program learning algorithm
-  that learns the DSL called \system,
-which infers
-  a DSL 
+  called \system
+  that infers a DSL
   and trains a neural network to
@@
 \begin{figure}\centering
-\includegraphics[width = 7cm]{figures/regression.eps}
+\includegraphics[width = 3.5cm]{figures/regression.eps} 
   \includegraphics[width = 7cm]{figures/list.eps}
-  \includegraphics[width = 7cm]{figures/textLearningCurve.eps}
-  \includegraphics[width = 7cm]{figures/circuitLearningCurve.eps}  
+  \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        
+  \includegraphics[width = 3.5cm]{figures/circuitLearningCurve.eps}  
   \caption{Learning curves for \system both with (blue) and without (red) the recognition model as the frontier size is varied (solid/dashed/dotted lines).}\label{learningCurves} 
@@
 
-\pagebreak
  \section{Related Work}

### 2018-02-08 paper.tex
update list graph
aggressive editing - we need to kill half page

@@
   \end{tikzpicture}
-  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}).
+  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Alg.~\ref{programGenerativeModel}).
 Observe tasks $x$ (program inputs/outputs).
@@
 $(\mathcal{D},\theta)$ define a distribution over programs.
-Figure~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
+Fig.~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
 
@@
 
-Inference in the generative model in Figure~\ref{graphicalModel}(a) is
+Inference in the generative model in Fig.~\ref{graphicalModel}(a) is
 complicated by the fact that the programs are unobserved,
@@
   model}, written $q(\cdot )$
-(Figure~\ref{graphicalModel}b).  The recognition model $q$ is a neural
+(Fig.~\ref{graphicalModel}b).  The recognition model $q$ is a neural
 network that regresses from input/output pairs to a distribution over
@@
  and show that our algorithm discovers
- its own domain-specific vocabulary for expressing solutions the domain (Table~\ref{initialExampleDSL}).
+ its own domain-specific vocabulary for expressing solutions the domain (Tbl.~\ref{initialExampleDSL}).
  \begin{table}
@@
 to accelerate the search over programs -- and (3) improving the DSL.
-The key observation here is that these three steps bootstrap off each other (Figure~\ref{feeding}):
+The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
 \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
@@
 
-\begin{figure}
-\centering
-\begin{tikzpicture}
-  \begin{scope}[shift={(2.5,0)}]
-    \node[obs] at (3,3) (dx){$\mathcal{D}$};
-    \node[latent] at (3.5,1.75) (zp){$p$};
-    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-    \node[obs] at (3.5,0.7) (xp) {$x$};
-    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
-    \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
-    \draw [->,red] (tx) -- (zp);
-    \draw [->] (dx) -- (zp);
-    \node at (nn) {
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle,width=2cm] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-  \end{scope}
-
-  \begin{scope}[shift={(0.7,-4.2)}]
-    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-    \draw [->] (p2.south) -- (x2.north);
-    \draw [->] (dt.south) -- (p2.north);
-
-    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-    \draw [->] (p1.south) -- (x1.north);
-    \draw [->] (dt.south) -- (p1.north);
-
-    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-    \draw [->] (p3.south) -- (x3.north);
-    \draw [->] (dt.south) -- (p3.north);
-
-    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
-
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-    \draw [->] (zp.south) -- (xp.north);
-    \plate {}{(zp)(xp)}{$x\in X$};
-
-    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
-  \end{scope}
-
-  \begin{scope}[shift={(7.7,-4)}]
-    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-    \node[latent] at (1.5,3) (t){$\theta$};
-    \node[obs] at (1,1.75) (z){$p$};
-    % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-    \node[obs] at (1,0.5) (x) {$x$};
-    \edge {z}{x};
-    \edge {d,t}{z};
-    \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
-  \end{scope}
-
-\end{tikzpicture}
-\end{figure}
-
-Section~\ref{mathematicalFraming} frames this 3-step procedure as
+%% \begin{figure}
+%% \centering
+%% \begin{tikzpicture}
+%%   \begin{scope}[shift={(2.5,0)}]
+%%     \node[obs] at (3,3) (dx){$\mathcal{D}$};
+%%     \node[latent] at (3.5,1.75) (zp){$p$};
+%%     \node[obs] at (4,3) (tx){$\theta^{(x)}$};
+%%     \node[obs] at (3.5,0.7) (xp) {$x$};
+%%     \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
+%%     \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
+%%     \draw [->,red] (tx) -- (zp);
+%%     \draw [->] (dx) -- (zp);
+%%     \node at (nn) {
+%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+%%         \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%         \node[rectangle,width=2cm] at (1,1) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (I-\name) at (0,-\y) {};
+%%         \foreach \name / \y in {1,...,3}
+%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (O-\name) at (2,-\y) {};
+%%         \foreach \source in {1,...,4}
+%%             \foreach \dest in {1,...,3}
+%%                 \draw [-latex] (I-\source) -- (H-\dest);
+%%         \foreach \source in {1,...,3}
+%%             \foreach \dest in {1,...,4}
+%%                 \draw [-latex] (H-\source) -- (O-\dest);
+%%       \end{tikzpicture}
+%%     };
+%%     \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
+%%   \end{scope}
+
+%%   \begin{scope}[shift={(0.7,-4.2)}]
+%%     \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+%%     \draw [->] (p2.south) -- (x2.north);
+%%     \draw [->] (dt.south) -- (p2.north);
+
+%%     \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+%%     \draw [->] (p1.south) -- (x1.north);
+%%     \draw [->] (dt.south) -- (p1.north);
+
+%%     \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+%%     \draw [->] (p3.south) -- (x3.north);
+%%     \draw [->] (dt.south) -- (p3.north);
+
+%%     \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+%%     \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+
+%%     \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+%%     \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+%%     \draw [->] (zp.south) -- (xp.north);
+%%     \plate {}{(zp)(xp)}{$x\in X$};
+
+%%     \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
+%%     \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
+%%   \end{scope}
+
+%%   \begin{scope}[shift={(7.7,-4)}]
+%%     \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+%%     \node[latent] at (1.5,3) (t){$\theta$};
+%%     \node[obs] at (1,1.75) (z){$p$};
+%%     % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+%%     \node[obs] at (1,0.5) (x) {$x$};
+%%     \edge {z}{x};
+%%     \edge {d,t}{z};
+%%     \plate {}{(z)(x)}{$x\in X$};
+%%     \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
+%%   \end{scope}
+
+%% \end{tikzpicture}
+%% \end{figure}
+
+Sec.~\ref{mathematicalFraming} frames this 3-step procedure as
 a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
-Section~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Section~\ref{recognitionSection} explains how we train a neural network to accelerate the search over programs; and
-Section~\ref{grammarInductionSection} explains how \system induces a DSL from programs.
+Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
+Sec.~\ref{recognitionSection} how we train a neural network to accelerate the search over programs; and
+Sec.~\ref{grammarInductionSection}  how \system induces a DSL from programs.
 
@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}(a).
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}(a).
 Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*,\theta^*$ solving:
@@
 %%   \end{align}
-\system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
+\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
 alternatingly maximizing w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
-%So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
+%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
 %explains how \system synthesizes new programs and given a DSL.
@@
 This taps into the intuition that programming is primarily a top-down activity:
-as human programmers, we first decide what kind of
+as human programmers, we often first decide what kind of
 programming constructs we might need to use,
@@
 %% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
-%% Figure~\ref{graphicalModel}.
+%% Fig.~\ref{graphicalModel}.
 
@@
 See Supplement for more detail on program representation.
-With this notation in hand we can now define DSLs:
+With this notation in hand we now define DSLs:
 
@@
 
-Algorithm~\ref{programGenerativeModel} is a procedure for drawing
+Alg.~\ref{programGenerativeModel} is a procedure for drawing
 samples from $\probability[p|\mathcal{D},\theta]$.  In practice, we
@@
 a depth-first search over the random choices made by
-Algorithm~\ref{programGenerativeModel}; we wrap the depth-first search
+Alg.~\ref{programGenerativeModel}; we wrap the depth-first search
 in iterative deepening to build $\lambda$-calculus
@@
 efficient means of solving for arbitrary constants that might occur in the
-program. In Section~\ref{regressionSection},
+program. In Sec.~\ref{regressionSection},
 we will show how to find programs with real-valued constants
 by automatically differentiating through the program and setting the constants using gradient descent.
-%% In Section~\ref{textSection}
+%% In Sec.~\ref{textSection}
 %% we will show that the bottom-up neural recognition model can learn
@@
 generating programs wholesale.
-We share this aspect with the DeepCoder system~\cite{balog2016deepcoder}.
+We share this aspect with DeepCoder~\cite{balog2016deepcoder}.
 
@@
 We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]$,
-which is equivalent to maximizing
+equivalently maximizing
 \begin{equation*}
@@
 and for domains with a fixed-dimensionality inputs/outputs we use a multilayer perceptron.
-We MaxPool across all of the input/output examples. See supplement for more details on the neural architecture.
+We MaxPool across all of the input/output examples. See supplement for details.
 
@@
 
-Recall from Section~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
+Recall from Sec.~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
 We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
@@
 These extracted subexpressions
-are fragments of the original programs, and can introduce new variables (Figure~\ref{fragmentExample}),
+are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
 which then become new functions in the DSL.
@@
 and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-Algorithm~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
+Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
 
@@
 
-Algorithm~\ref{mainAlgorithm} describes how we combine program search,
+Alg.~\ref{mainAlgorithm} describes how we combine program search,
 recognition model training, and DSL induction.
@@
 learns in the domains of Boolean circuits, symbolic regression, string editing, and list functions.
-We then show quantitative results (Section~\ref{quantitative})
+We then show quantitative results (Sec.~\ref{quantitative})
 comparing \system with lesioned versions
@@
 \system observes the input/output truth table of the circuits,
-and we define the likelihood (where $x$ is a truth table and $p$ a program implementing a circuit):
-\begin{align}
-  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]\nonumber
-  \end{align}
+and we define the likelihood 
+$  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]$,
+where $x$ is a truth table and $p$ a program implementing a circuit.
 
@@
 circuits by creating new subroutines that correspond to
-logic gates and multiplexers (Figure~\ref{multiplexer}).
+logic gates and multiplexers (Fig.~\ref{multiplexer}).
 We will show data showing that the neural recognition model
-assists in building these circuits (Section~\ref{quantitative}),
+assists in building these circuits (Sec.~\ref{quantitative}),
 suggesting that an agent can glean something about the structure of a circuit just by looking at its truth table.
@@
 We show how to use \system to infer programs containing both discrete
-structure and continuous parameters. The high-level idea is to synthesize programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
+structure and continuous parameters. The high-level idea is to write programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
 and optimizing using gradient descent.
-Concretely, we ask the algorithm to
+We have the algorithm 
 solve a set of 200 symbolic regression problems, each a polynomial of
@@
 The learned DSL contains templates for polynomials of different orders,
-which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Table~\ref{regressionDSL}).
+which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Tbl.~\ref{regressionDSL}).
 Examining the programs themselves,
@@
 for example, it represents the linear function $-x+2$ with the program 
-\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
-This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
+\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}.
+This falls out of our Bayesian framing -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
 \begin{table}\centering
@@
 %% but given enough iterations of our algorithm,
-%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Figure~\ref{regressionLearningCurve}; No NN in Table~\ref{regressionComparison}).
-%% We compare our model on held-out tasks against 4 baselines (Table~\ref{regressionComparison}).
+%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Fig.~\ref{regressionLearningCurve}; No NN in Tbl.~\ref{regressionComparison}).
+%% We compare our model on held-out tasks against 4 baselines (Tbl.~\ref{regressionComparison}).
 
@@
 less specialized set of primitives -- close to the string routines provided by Python --
-and automatically recover many of the higher-level building blocks that have made these
+and recover many of the higher-level building blocks that have made these
 other system successful.
-Table~\ref{textPrimitives} (top) lists the primitive functions provided to \system.
+Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \system.
 Absent are any routines for searching for characters, extracting substrings,
@@
 or turning a list into a string (\code{join}).
-We automatically generated  600 string editing tasks along with input/output examples (Figure~\ref{exampleTextProblem}). %, split  75/25 train/test.
+We automatically generated  600 string editing tasks along with input/output examples (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
 At first, \system cannot find any programs that
 work at all for most of the tasks.
-After a few wake/sleep iterations,
-it assembles a DSL (Table~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
+After three wake/sleep iterations,
+it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
 all of the tasks.
@@
   \end{table}
-%% \begin{figure}
-
-%%   \includegraphics[width = 7cm]{figures/text.png}
-%%   \caption{Learning curves for string editing programs}\label{textLearningCurve}
-
-%% \end{figure}
 
@@
 YBW$>$TIg$<$WSho 	&TIG\\
-DkI$>$DwKI$<$xaPoc& 	DWKI\\
+%DkI$>$DwKI$<$xaPoc& 	DWKI\\
 KfvTh$>$Dbl$<$OIJso 	&DBL
@@
 
-  \vspace{0.5cm}
+  \vspace{0.25cm}
 
@@
     \end{minipage}
-  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Table~\ref{textPrimitives}}\label{exampleTextProblem}
+  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}
   \end{figure}
@@
 list of Booleans, or a list of integers.  There are 225 tasks in total, some
-of which are shown in Table~\ref{listexamples}.
+of which are shown in Tbl.~\ref{listexamples}.
 
@@
 Both are capable of solving all tasks supplied by our dataset.
-See Supplement for more detail on these DSLs.
+See Supplement for details on these DSLs.
 
@@
 a recognition model only from samples from the fixed generative model.
-This is equivalent to our algorithm with $\lambda = \infty$ (see Section~\ref{grammarInductionSection})
-and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Section~\ref{recognitionSection}).
+This is equivalent to our algorithm with $\lambda = \infty$ (see Sec.~\ref{grammarInductionSection})
+and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Sec.~\ref{recognitionSection}).
 We call this baseline RF/DC because
@@
 using two numbers, or did it introduce extraneous parameters?
-Table~\ref{baselineComparisons}
+Tbl.~\ref{baselineComparisons}
 compares our model against these baselines.
-We see that our full model consistently
+Our full model consistently
 beats the baselines,
@@
 Lesioning the recognition model also slows down the convergence of the algorithm,
-taking more wake/sleep cycles to reach a given number of tasks solved (Figure~\ref{learningCurves}).
-This data supports our view of the recognition model as a way of accelerating the search over programs.
+taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
+This supports our view of the recognition model as a way of accelerating the search over programs.
 
@@
 
-  \\\midrule\\\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
   \% solved &
@@
 
-  \\\midrule\\\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
   \% solved&   \textbf{ 98\% }&94\%&0.7\%&0\%&2\% \\
@@
 
-  \\\midrule\\\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
   \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
@@
 
-  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
   \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
@@
 
-  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
   \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
@@
 \includegraphics[width = 3.5cm]{figures/regression.eps} 
-  \includegraphics[width = 7cm]{figures/list.eps}
+  \includegraphics[width = 3.5cm]{figures/list.eps}
   \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        

### 2018-02-09 paper.tex
more editing
small change to section title
removed the pseudocode for grammar induction
shortened bibliography

@@
 and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
+%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
 
@@
 In the Supplement we derive a tractable MAP estimator for $\theta$.
-\begin{algorithm}[tb]
-   \caption{DSL Induction Algorithm}
-   \label{grammarInductionAlgorithm}
-   \begin{algorithmic}
-     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
-     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
-     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
-     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-     \WHILE {true}
-     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
-     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
-     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
-     \STATE $\mathcal{D}\gets\mathcal{D}'$
-     \ENDWHILE
-   \end{algorithmic}
-\end{algorithm}
+%% \begin{algorithm}[tb]
+%%    \caption{DSL Induction Algorithm}
+%%    \label{grammarInductionAlgorithm}
+%%    \begin{algorithmic}
+%%      \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
+%%      \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
+%%      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+%%      \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
+%%      \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
+%%      \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
+%%      \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
+%%      \WHILE {true}
+%%      \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
+%%      \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
+%%      \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
+%%      \STATE $\mathcal{D}\gets\mathcal{D}'$
+%%      \ENDWHILE
+%%    \end{algorithmic}
+%% \end{algorithm}
 
@@
 
-\subsection{Implementing the \system}
+\subsection{Implementing \system}
 
@@
 \\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
-This is equivalent to the first wake cycle of \system.
+This is equivalent to our first wake cycle.
 
@@
 Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
- and our contribution here is a general scheme for getting BPL off the ground with less hand-engineering of the DSL.
+ and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.
 
@@
  getting program induction off the ground.
-
-\pagebreak
 

### 2018-02-09 paper.tex
another symbolic regression experiment curve
fixed our use of system at end of sentences

@@
 \newcommand{\system}{\textsc{DreamCoder}~}
+\newcommand{\endingSystem}{\textsc{DreamCoder}}
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 which we call \textbf{Sleep-R}.
-This wake/sleep style of inference is the reason why we call our approach \system.
+This wake/sleep style of inference is the reason why we call our approach \endingSystem.
 
@@
     \end{tikzpicture}
-\caption{Our algorithm solves for programs, the DSL, and a recognition model.
+  \caption{\system solves for programs, the DSL, and a recognition model.
 Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
@@
 
-The neural architecture of $q$ depends upon the domain.
+The architecture of $q$ depends upon the domain.
 For domains with sequential structure like strings and lists,
 we use an RNN to encode each input/output example,
-and for domains with a fixed-dimensionality inputs/outputs we use a multilayer perceptron.
+and for domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
 We MaxPool across all of the input/output examples. See supplement for details.
@@
 We first present our experimental results qualitatively,
-illustrating the DSLs that  \system
-learns in the domains of Boolean circuits, symbolic regression, string editing, and list functions.
+illustrating the DSLs we
+learn.
 We then show quantitative results (Sec.~\ref{quantitative})
@@
 \end{tabular}
-\caption{Some induced DSL primitives for symbolic regression.
-System starts out with addition, multiplication, and real numbers, and learns functions for building 4th order polynomials.}\label{regressionDSL}
+\caption{Some induced DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns to build 4th order polynomials.}\label{regressionDSL}
 \end{table}
@@
 other system successful.
-Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \system.
+Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \endingSystem.
 Absent are any routines for searching for characters, extracting substrings,
@@
     \code{split}, \code{join}, \code{map}, 
-    \code{append}, \code{chr->str}, \code{trim}, \code{nth}\\
+    \code{+}, \code{chr->str}, \code{trim}, \code{nth}\\
     \code{length}, \code{slice},    \code{lower}, \code{upper}, \code{capitalize}
@@
   \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
-  \% solved&   \textbf{ 98\% }&94\%&0.7\%&0\%&2\% \\
+  \% solved&   \textbf{98\% }&94\%&0.7\%&0\%&2\% \\
   Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
@@
   \end{tabular}
-\caption{\% solved: fraction solved with 5 sec timeout.
+\caption{\% solved: fraction solved w/ 5 sec timeout.
 Solve time: averaged over solved tasks.
@@
 
- \section{Conclusion}
+ \section{Contribution and Outlook}
  We contributed a model, \system, that learns to program by

### 2018-02-09 paper.tex
LUCAS I edited your section, feel free to change things back

@@
 \subsection{List functions}
-Our list function domain consists of hand-designed concepts from which we
-sample tasks, so there is no ``ground truth'' formal DSL from which these
-tasks are generated. Additionally, few of these tasks are concatenative ---
-many of them represent distinct operations.
-Tasks in this domain are solved by functions that take as input an integer
-or a list of integers, and have as output either a Boolean, an integer, a
-list of Booleans, or a list of integers.  There are 225 tasks in total, some
-of which are shown in Tbl.~\ref{listexamples}.
-
-We take approaches from two different initial DSLs, \emph{base} and
-\emph{rich}, and assess their performance. The base DSL starts with only
+Synthesizing programs that manipulate data structures is
+a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
+%with applications to computer aided programming~\cite{solar2008program}.
+We consider this problem within the context of
+learning functions that manipulate lists.
+We created 225 Lisp-style list manipulation tasks,
+each with 15 input/output examples (Tbl.~\ref{listexamples}).
+Our data set is challenging along two dimensions:
+many of the functions are very complicated,
+and the agent must learn to solve these complicated problems from only 225 tasks.
+
+%% Our list function domain consists of hand-designed concepts from which we
+%% sample tasks, so there is no ``ground truth'' formal DSL from which these
+%% tasks are generated. Additionally, few of these tasks are concatenative ---
+%% many of them represent distinct operations.
+%% Tasks in this domain are solved by functions that take as input an integer
+%% or a list of integers, and have as output either a Boolean, an integer, a
+%% list of Booleans, or a list of integers.  There are 225 tasks in total, some
+%% of which are shown in Tbl.~\ref{listexamples}.
+
+We evaluate \system starting from two different initial DSLs, \emph{base} and
+\emph{rich}. The base DSL starts with only
 low-level primitives such as \code{mapi}, \code{reducei}, and
@@
 
-We found that this domain was difficult to tackle without starting from a
-relatively rich DSL, and even then it was still hard to solve all tasks. Our
+We found  this domain  difficult to tackle without starting from 
+the rich DSL. Our
 algorithm, like human learners, requires a spectrum of problems ranging from
-easy to hard. Without this curriculum, \system cannot get off the ground.
-The behavior with the base DSL illustrates this, because the initial set of
-primitives is not paired with a sufficiently large and graded curriculum.
+easy to hard. Without this curriculum, we cannot get off the ground.
+%The behavior with the base DSL illustrates this phenomena  because the initial set of
+%primitives is not paired with a sufficiently large and graded curriculum.
 
@@
 \end{tabular}
-\caption{Simple tasks from our list function domain. We used 15 input/output
-  examples for each task to reduce ambiguity in the task's function,
-  ensuring solutions achieve the desired concept.}
+\caption{Sample tasks from our list function domain. %% We used 15 input/output
+  %% examples for each task to reduce ambiguity in the task's function,
+  %% ensuring solutions achieve the desired concept.
+}
 \label{listexamples}

### 2018-02-09 paper.tex
LUCAS more edits to your section, of course so free to change them back
fixed bug in previous commit

@@
 \newcommand{\system}{\textsc{DreamCoder}~}
-\newcommand{\endingSystem}{\textsc{DreamCoder}}
+\newcommand{\systemEnding}{\textsc{DreamCoder}}
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 ground-truth programs.
-Our algorithm, called \system,
+Our algorithm, called \systemEnding,
 does three things:
@@
 which we call \textbf{Sleep-R}.
-This wake/sleep style of inference is the reason why we call our approach \endingSystem.
+This wake/sleep style of inference is the reason why we call our approach \systemEnding.
 
@@
 other system successful.
-Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \endingSystem.
+Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
 Absent are any routines for searching for characters, extracting substrings,
@@
     \end{minipage}
-  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}
+  \caption{Example string edit task (top) and program  \system writes (bottom).
+$f_3$ is a new DSL subroutine discovered by \systemEnding, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}
   \end{figure}
@@
 algorithm, like human learners, requires a spectrum of problems ranging from
-easy to hard. Without this curriculum, we cannot get off the ground.
+easy to hard.
+When starting with the base DSL,
+we found that there were not enough steppingstones in the curriculum to
+get \system off the ground.
+
 %The behavior with the base DSL illustrates this phenomena  because the initial set of
@@
  \section{Contribution and Outlook}
- We contributed a model, \system, that learns to program by
+ We contributed a model, \systemEnding, that learns to program by
  bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.

### 2018-02-09 paper.tex
put in Josh 's introduction

@@
 Automatically inducing programs from examples is a long-standing goal
-of artificial intelligence.  Recent work in program induction show
-that deep networks can be very successful at learning to write certain
-kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
-DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
-techniques (e.g., Metagol:~\cite{muggleton2015meta},
-FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
-However, the success of these approaches -- both symbolic and neural --
-relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
-The DSL is an inventory of restricted programming primitives,
-and imparts domain specific knowledge about the space of programs.
-To what extent can we dispense with hand-engineered DSLs?
-
-This work proposes \emph{learning} the DSL.
-We consider the setting where we have a collection of related programming tasks,
-each specified by a set of input/output examples.
-We do not assume that the tasks are annotated with
-ground-truth programs.
-Our algorithm, called \systemEnding,
-does three things:
-it finds programs that solve each of the tasks;
-it creates its own DSL by discovering and reusing
-domain-specific subroutines;
-and it learns a neural network that
-efficiently searches for programs in the DSL.
-The purpose of the DSL
-is to expose the commonalities across the programs that solve tasks within a domain,
-letting the agent solve other related program learning problems.
+of artificial intelligence. Recent work has successfully used
+symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
+FlashFill:~\cite{gulwani2011automating}), neural networks trained from
+a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill},
+DeepCoder:~\cite{balog2016deepcoder}), and hybrids of neural and
+symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}) to
+synthesize programs for task domains such as string transformations,
+list processing, and robot navigation and planning. However, all these
+approaches -- symbolic, neural and neural-symbolic -- rely upon a
+hand-engineered \emph{Domain Specific Language} (DSL). The DSL is an
+inventory of restricted programming primitives, encoding domain
+specific knowledge about the space of programs to be considered. In
+practice
+we are often given only one or a few examples of each
+program to be induced, and thus success often hinges on having a good
+DSL, to
+provide a crucial inductive bias for what would otherwise be an
+unconstrained search through the space of all computable functions.
+Here we ask, to what extent can we dispense with such highly
+hand-engineered DSLs?
+
+We propose \emph{learning} the DSL. We consider the setting where we
+have a collection of related programming tasks, each specified by a
+set of input/output examples. We do not assume that the tasks are
+annotated with ground-truth programs. We typically will not try to
+learn a DSL completely from scratch, but rather to start from a weaker
+or more general set of primitives in a Lisp-like representation and
+construct a richer, more powerful and better-tuned domain-specific
+language.
+
+Our algorithm is called \system because it is based on a
+novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
+between ``wake'' and ``sleep'' phases to achieve three goals: finding
+programs that solve each of the tasks; creating a DSL by discovering
+and reusing domain-specific subroutines; and learning a neural network
+that efficiently guides search for programs in the DSL. The DSL
+distills commonalities across programs that solve the tasks, letting
+the agent solve other related program learning problems. The neural
+network ensures that searching for program solutions remains tractable
+even as the DSL (and hence the search space for programs) expands.
+
+%% Automatically inducing programs from examples is a long-standing goal
+%% of artificial intelligence.  Recent work in program induction show
+%% that deep networks can be very successful at learning to write certain
+%% kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
+%% DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
+%% techniques (e.g., Metagol:~\cite{muggleton2015meta},
+%% FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
+%% However, the success of these approaches -- both symbolic and neural --
+%% relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
+%% The DSL is an inventory of restricted programming primitives,
+%% and imparts domain specific knowledge about the space of programs.
+%% To what extent can we dispense with hand-engineered DSLs?
+
+%% This work proposes \emph{learning} the DSL.
+%% We consider the setting where we have a collection of related programming tasks,
+%% each specified by a set of input/output examples.
+%% We do not assume that the tasks are annotated with
+%% ground-truth programs.
+%% Our algorithm, called \systemEnding,
+%% does three things:
+%% it finds programs that solve each of the tasks;
+%% it creates its own DSL by discovering and reusing
+%% domain-specific subroutines;
+%% and it learns a neural network that
+%% efficiently searches for programs in the DSL.
+%% The purpose of the DSL
+%% is to expose the commonalities across the programs that solve tasks within a domain,
+%% letting the agent solve other related program learning problems.
 

### 2018-02-09 paper.tex
stitching things together with the Josh section

@@
   edit strings, do symbolic regression,
-  and synthesize Lisp-style functions on lists,
+  and synthesize functions on lists,
   in each case showing that
@@
 %  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-  \node[obs] at (1,0) (x) {$x$};
+  \node[obs] at (1,0.5) (x) {$x$};
   \edge {z}{x};
@@
   \plate {}{(z)(x)}{$x\in X$};
-  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative \\model};
+  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};
 
   \begin{scope}[shift = {(0.7,0)}]  
-  \node[latent] at (3.5,3) (dx){$\mathcal{D}$};
-  \node[latent] at (3.5,1.75) (zp){$p$};
-  \node[latent] at (5.5,1.75) (tx){$\theta^{(x)}$};
-  \node[obs] at (3.5,0) (xp) {$x$};
+  \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
+  \node[latent] at (3.5,1.9) (zp){$p$};
+  \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
+  \node[obs] at (3.5,0.5) (xp) {$x$};
   \plate {}{(zp)(xp)(tx)}{$x\in X$};
@@
   \node at (nn) {
-    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
@@
   \end{tikzpicture}
-  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Alg.~\ref{programGenerativeModel}).
+  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$.
 Observe tasks $x$ (program inputs/outputs).
 (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
-    $\theta^{(x)}$. Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
+    $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
 \end{figure}
@@
 Fig.~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
-
-%% Inducing the DSL means learning the symbolic structure of a latent programming language,
-%% and so we are confronted with two distinct structure learning problems:
-%% the structure of the programs and the structure of the DSL.
-
-Inference in the generative model in Fig.~\ref{graphicalModel}(a) is
-complicated by the fact that the programs are unobserved,
-and so we must solve a difficult search problem to find any programs at all. To make
+Inference in this model is
+difficult because the programs are unobserved,
+and so we must solve a hard search problem to recover them. To make
 search tractable we learn a bottom-up \emph{recognition
@@
 
-How should we perform inference for the model
-in~\ref{graphicalModel}(a)?
-We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
+To perform inference, we take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
 which alternates between ``waking'', which draws samples from a recognition model (our $q$);
@@
 For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
-The wake cycle can be thought of active problem solving,
-guided by the current generative and recognition models.
 A sleep cycle means  updating these models so that the DSL and  recognition model
@@
  and show that our algorithm discovers
- its own domain-specific vocabulary for expressing solutions the domain (Tbl.~\ref{initialExampleDSL}).
+ its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
  \begin{table}
@@
 
-Our goal  is to induce a DSL and find good programs solving each of the tasks.
+Our goal  is to induce a DSL while finding programs solving each of the tasks.
 Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use

### 2018-02-09 paper.tex
more edits

@@
 inventory of restricted programming primitives, encoding domain
-specific knowledge about the space of programs to be considered. In
+specific knowledge about the space of programs. In
 practice
-we are often given only one or a few examples of each
+we are often given only a few examples for each
 program to be induced, and thus success often hinges on having a good
@@
 or more general set of primitives in a Lisp-like representation and
-construct a richer, more powerful and better-tuned domain-specific
-language.
+construct a richer, more powerful and better-tuned DSL.
 
@@
 %% which frames our problem as one of program induction.
+
+Viewed as an inference problem,
+our goal is to infer the DSL, written $\mathcal{D}$,
+along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
+A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
+We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together
+$(\mathcal{D},\theta)$ define a distribution over programs.
+Fig.~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
+Inference in this model is
+difficult because the programs are unobserved,
+and so we must solve a hard search problem to recover them. To make
+search tractable we learn a bottom-up \emph{recognition
+  model}, written $q(\cdot )$
+(Fig.~\ref{graphicalModel}b).  The recognition model $q$ is a neural
+network that regresses from input/output pairs to a distribution over
+programs likely to explain the input/outputs. $q(\cdot )$ implements an amortized
+inference
+scheme~\cite{le2016inference} for the generative %ritchie2016deep
+model.
 
@@
 
-Viewed as an inference problem,
-our goal is to infer the DSL, written $\mathcal{D}$,
-along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
-A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
-We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together
-$(\mathcal{D},\theta)$ define a distribution over programs.
-Fig.~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
-Inference in this model is
-difficult because the programs are unobserved,
-and so we must solve a hard search problem to recover them. To make
-search tractable we learn a bottom-up \emph{recognition
-  model}, written $q(\cdot )$
-(Fig.~\ref{graphicalModel}b).  The recognition model $q$ is a neural
-network that regresses from input/output pairs to a distribution over
-programs likely to explain the input/outputs. $q(\cdot )$ implements an amortized
-inference
-scheme~\cite{le2016inference} for the generative %ritchie2016deep
-model.
-
-To perform inference, we take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
+To perform inference, we took inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
 which alternates between ``waking'', which draws samples from a recognition model (our $q$);
@@
 and a second phase where we update the recognition model $q$,
-which we call \textbf{Sleep-R}.
-This wake/sleep style of inference is the reason why we call our approach \systemEnding.
+ called \textbf{Sleep-R}.
 

### 2018-02-09 paper.tex
tiny edits to intro

@@
 Viewed as an inference problem,
-our goal is to infer the DSL, written $\mathcal{D}$,
+our goal is to infer the DSL (written $\mathcal{D}$)
 along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
@@
 $(\mathcal{D},\theta)$ define a distribution over programs.
-Fig.~\ref{graphicalModel}(a) the diagrams this inference problem as a hierarchical Bayesian model.
+Fig.~\ref{graphicalModel}(a) diagrams this inference problem as a hierarchical Bayesian model.
 Inference in this model is
@@
 search tractable we learn a bottom-up \emph{recognition
-  model}, written $q(\cdot )$
-(Fig.~\ref{graphicalModel}b).  The recognition model $q$ is a neural
+  model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
+The recognition model $q$ is a neural
 network that regresses from input/output pairs to a distribution over

### 2018-02-09 paper.tex
only three more lines to kill

@@
 of the solution, as measured by the likelihood model $\probability[x|p]$,
-which measures how many
-real numbers were used in the program: e.g. did the model correctly explain a linear function
+e.g. did the agent correctly explain a linear function
 using two numbers, or did it introduce extraneous parameters?
@@
  In our work we consider a weakly-supervised setting where ground truth programs are not provided and
- the agent must learn from a few hundred examples.
-
- \noindent \emph{Predicate Invention in ILP:} The Inductive Logic Programming (ILP) community
- has long worked on the problem of inventing new primitive predicates~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}, which is analogous to our model with $\theta$ held fixed and $q$ ablated.
+ the agent must learn from a few hundred tasks.
 
@@
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}.
 

### 2018-02-09 paper.tex
8 pages!

@@
 we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
-(3) Because the frontiers can grow very large,
-we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
-with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
-(4) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
+%% (3) Because the frontiers can grow very large,
+%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
+%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
+(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
 by only summing over the top $K$
-programs in $\mathcal{F}_x$ -- we found that $K = 2$ sufficed.
-(5) For added robustness, we enumerate programs from both the generative model and the recognition model.
+programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
+(4) For added robustness, we enumerate programs from both the generative model and the recognition model.
 
@@
 or turning a list into a string (\code{join}).
-We automatically generated  600 string editing tasks along with input/output examples (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
+We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
 At first, \system cannot find any programs that
@@
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}.
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14}.
 

### 2018-02-09 paper.tex
miscellaneous edits

@@
   in each case showing that
-  \system learns a domain-specific vocabulary for expressing programs.
+  \system learns a domain-specific vocabulary for expressing solutions to the problems in the domain.
 
@@
 programs that solve each of the tasks; creating a DSL by discovering
-and reusing domain-specific subroutines; and learning a neural network
+and reusing domain-specific subroutines; and training a neural network
 that efficiently guides search for programs in the DSL. The DSL
@@
 
-To perform inference, we took inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
+To perform inference, we take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
 which alternates between ``waking'', which draws samples from a recognition model (our $q$);
@@
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by searching for programs where $\probability[x,p|\mathcal{D},\theta]$
+$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
 is large.
@@
 Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by learning a neural recognition model
+of the large combinatorial search space. We ease this difficulty by training a neural recognition model
 during the Sleep-R phase:
@@
 and then training $q$ to regress from the outputs to the program.
-If these programs map inputs to outputs,
-then we need some way of sampling these inputs as well.
+Since these programs map inputs to outputs,
+we need some way of sampling these inputs as well.
 Our solution to this problem is to sample the inputs
@@
 %% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
-(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
+(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
 by only summing over the top $K$

### 2018-02-09 paper.tex
add difficult list tasks

@@
   Name & Input & Output \\\midrule
-  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
   append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
   len & [3\, 5\, 12\, 1] & 4 \\
-  range & 3 & [1\, 2\, 3] \\
   has-2 & [4\, 5\, 7\, 4] & \code{false} \\
-  has-head-in-tail & [4\, 5\, 7\, 4] & \code{true} \\
   repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
   drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+  count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+  rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
+  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
+  keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
   \bottomrule

### 2018-02-09 supplement.tex
initial DSL primitives and types

@@
 
-\section{Why not the ELBO bound?}
+\section{DSL Induction}
 
-If we were to maximize the ELBO bound to perform inference in
-our generative model,
-then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along an unimportant regularizing term on the DSL):
-\begin{align}
-  &  \sum_{x\in X}\expect_{p\sim Q_x}\left[\log \probability[p|\mathcal{D}^*,\theta^*] \right] \label{EM}\\
-  & Q_x(p)\triangleq \probability[p|x,\mathcal{D},\theta]
-\end{align}
-where $(\mathcal{D},\theta)$ is our current estimate of the generative
-model.  These equations fall out of an EM-style derivation, and one
-could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
-using importance sampling (so the expectation in~\ref{EM} is taken
-over $q$ and we reweight using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.
+\subsection{Structure learning}
 
-How the recognition model is worked into this is unimportant to the reason why we do not maximize a bound of this form:
-this bound takes an expectation over the \emph{previous} sleep cycle's
-posterior over programs,
-and so the approximate posterior $Q_x$ at the next iteration
-ends up being very close to previous approximate posterior.
-Intuitively,
-we want the DSL induction to be a function \emph{only} of the programs that we have found,
-and \emph{not} be a function of how the previous generative model weighted them.
-In practice,
-we found that maximizing EM-style bounds, like the ELBO,
-leads to a kind of hysteresis effect,
-where the next generative model too closely matches the previous one,
-causing the algorithm to easily become trapped in local optima.
+We use Algorithm~\ref{grammarInductionAlgorithm} to search for the structure of the DSL that best explains the frontiers.
 
-\section{Estimating $\theta$}
+\begin{algorithm}[tb]
+   \caption{DSL Induction Algorithm}
+   \label{grammarInductionAlgorithm}
+   \begin{algorithmic}
+     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
+     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
+     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
+     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
+     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
+     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
+     \WHILE {true}
+     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
+     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
+     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
+     \STATE $\mathcal{D}\gets\mathcal{D}'$
+     \ENDWHILE
+   \end{algorithmic}
+\end{algorithm}
 
-We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.\ref{programGenerativeModel}.
+\subsection{Estimating $\theta$}
+
+We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.~1 of the main paper.
 Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
@@
 
+\subsection{Boolean circuits}
+\begin{center}
+\tt
+\begin{tabular}{| l | l |}
+  \hline
+  \textrm{\emph{name}} & \textrm{\emph{type}} \\
+  \hline
+    nand & $\text{bool}\to \text{bool}\to \text{bool}$ \\  \hline
+\end{tabular}
+\end{center}
+
+\subsection{Symbolic regression}
+\begin{center}
+\tt
+\begin{tabular}{| l | l |}
+  \hline
+  \textrm{\emph{name}} & \textrm{\emph{type}} \\
+  \hline
+  + & $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$ \\
+  *&  $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
+  real&$\mathbb{R}$\\\hline
+\end{tabular}
+\end{center}
+
+\subsection{String editing}
+\begin{center}
+\tt
+\begin{tabular}{| l | l |}
+  \hline
+  \textrm{\emph{name}} & \textrm{\emph{type}} \\
+  \hline
+  0&$\mathbb{Z}$\\
+  +1&$\mathbb{Z}\to \mathbb{Z}$\\
+  -1&$\mathbb{Z}\to \mathbb{Z}$\\
+  split&$\text{char}\to \text{string}\to \left[\text{string} \right]$\\
+  join&$\text{string}\to \left[\text{string} \right]\to \text{string}$\\
+  map&$(\text{string}\to \text{string})\to \left[\text{string} \right]\to \left[\text{string} \right]$\\
+  nth&$\mathbb{Z}\to \left[\text{string} \right]\to \text{string}$\\
+  slice&$\mathbb{Z}\to \mathbb{Z}\to \text{string}\to \text{string}$\\
+  length&$\text{string}\to \mathbb{Z}$\\
+  chr->str&$\text{char}\to \text{string}$\\
+  trim&$\text{string}\to \text{string}$\\
+  upper&$\text{string}\to \text{string}$\\
+  lower&$\text{string}\to \text{string}$\\
+  capitalize&$\text{string}\to \text{string}$\\\hline
+\end{tabular}
+\end{center}
+We also provide the constant empty string which has type string and also provide
+constant characters which have the type char.
+
+\section{Why not the ELBO bound?}
+
+If we were to maximize the ELBO bound to perform inference in
+our generative model,
+then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along an unimportant regularizing term on the DSL):
+\begin{align}
+  &  \sum_{x\in X}\expect_{p\sim Q_x}\left[\log \probability[p|\mathcal{D}^*,\theta^*] \right] \label{EM}\\
+  & Q_x(p)\triangleq \probability[p|x,\mathcal{D},\theta]
+\end{align}
+where $(\mathcal{D},\theta)$ is our current estimate of the generative
+model.  These equations fall out of an EM-style derivation, and one
+could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
+using importance sampling (so the expectation in~\ref{EM} is taken
+over $q$ and we reweight using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.
+
+How the recognition model is worked into this is unimportant to the reason why we do not maximize a bound of this form:
+this bound takes an expectation over the \emph{previous} sleep cycle's
+posterior over programs,
+and so the approximate posterior $Q_x$ at the next iteration
+ends up being very close to previous approximate posterior.
+Intuitively,
+we want the DSL induction to be a function \emph{only} of the programs that we have found,
+and \emph{not} be a function of how the previous generative model weighted them.
+In practice,
+we found that maximizing EM-style bounds, like the ELBO,
+leads to a kind of hysteresis effect,
+where the next generative model too closely matches the previous one,
+causing the algorithm to easily become trapped in local optima.
+
 \bibliography{main}

### 2018-02-09 paper.tex
initial DSL primitives and types
restyled induced primitives

@@
 In the Supplement we derive a tractable MAP estimator for $\theta$.
-%% \begin{algorithm}[tb]
-%%    \caption{DSL Induction Algorithm}
-%%    \label{grammarInductionAlgorithm}
-%%    \begin{algorithmic}
-%%      \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-%%      \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-%%      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-%%      \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
-%%      \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
-%%      \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
-%%      \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-%%      \WHILE {true}
-%%      \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
-%%      \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
-%%      \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
-%%      \STATE $\mathcal{D}\gets\mathcal{D}'$
-%%      \ENDWHILE
-%%    \end{algorithmic}
-%% \end{algorithm}
 
@@
   %% \\\midrule
-    $f_0 = $\code{(+ }$\mathcal{R}$\code{)}\\
-    $f_1($\code{x}$) =$ \code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} \\
-      $f_2($\code{x}$) =$ \code{(+ }$\mathcal{R}$\code{ ($f_1$ x))} \\
-      $f_3($\code{x}$) = $\code{(* x (}$f_2$\code{ x))}\\
-  $f_4($\code{x}$) = $\code{(}$f_0$\code{ (}$f_3$\code{ x)))}\\
-  $f_5($\code{x}$) =$ \code{(* x (}$f_4$\code{ x))}\\
-  $f_6($\code{x}$) = $\code{(}$f_0$\code{ (}$f_4$\code{ x))} (\emph{4th order polynomial})
+  $f_0\,=\,$\code{(+ $\mathcal{R}$)}\\
+  $f_1($\code{x}$)\,=\,$\code{(* (* (+ x $\mathcal{R}$) $\mathcal{R}$) x)} \\
+  $f_2($\code{x}$)\,=\,$\code{(+ $\mathcal{R}$ ($f_1$ x))} \\
+  $f_3($\code{x}$)\,=\,$\code{(* x ($f_2$ x))}\\
+  $f_4($\code{x}$)\,=\,$\code{($f_0$ ($f_3$ x))}\\
+  $f_5($\code{x}$)\,=\,$\code{(* x ($f_4$ x))}\\
+  $f_6($\code{x}$)\,=\,$\code{($f_0$ ($f_4$ x))} (\emph{4th order polynomial})
   \\\bottomrule
@@
       \toprule
-      $f_0($\code{c}$) = $\code{(+ (chr->str c))}\\
-      $f_1($\code{i,c,s}$) = $      \code{(+ (nth i (split c s))}\\\hspace{2.55cm}\code{ (chr->str c))}\\
-      $f_2($\code{s,c}$) = $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
+      $f_0($\code{c}$) \,=\, $\code{(+ (chr->str c))}\\
+      $f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ (nth i (split c s))}\\
+        \phantom{$f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ }}\code{(chr->str c))}\\
+      $f_2($\code{s$,$c}$) \,=\, $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
         \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
-      $f_3($\code{c,s}$) = $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
+      $f_3($\code{c$,$s}$) \,=\, $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
         \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
@@
   \begin{minipage}{7.5cm}
- $f($\code{s}$) = $   \code{(}$f_3$\code{ (split '>' (upper s)) '<')}
+    $f($\code{s}$) \,=\, $\code{(}$f_3$\code{ (split '>' (upper s)) '<')}
   \end{minipage}
-  \caption{Example string edit task (top) and program  \system writes (bottom).
-$f_3$ is a new DSL subroutine discovered by \systemEnding, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}
+  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}.
   \end{figure}
@@
     \toprule
-    $(f_0$ \code{i $\ell$}$) = $\code{(singleton (index i $\ell$))}\\
+    $f_0($\code{i$,\ell$}$) \,=\, $\code{(singleton (index i $\ell$))}\\
     \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
-    $(f_1$ \code{i $\ell$}$) = $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
+    $f_1($\code{i$,\ell$}$) \,=\, $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
     \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
-    $(f_2$ \code{n $\ell$}$) = $\code{(any (lambda (x) (= n x)) $\ell$)}\\
+    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any (lambda (x) (= n x)) $\ell$)}\\
     \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
-    $(f_3$ \code{i $\ell$}$) = $\code{(index (negate i) (sort $\ell$))}\\
+    $f_3($\code{i$,\ell$}$) \,=\, $\code{(index (negate i) (sort $\ell$))}\\
     \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
-    $(f_4$ \code{n $\ell$}$) = $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
-    $(f_5$ \code{k $\ell$}$) = $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
-    $(f_6$ \code{k n $\ell$}$) = $\code{($f_4$ n ($f_5$ k $\ell$)}\\
+    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
+    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
+    $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
     \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\

### 2018-02-09 supplement.tex
moved the pretty diagram to the supplement

@@
 \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+
+\section{An illustration of the wake-sleep cycle}
+Below we diagram the wake/sleep cycles employed by our algorithm. At each stage of the cycle,
+we have shaded the observed variables in gray and left the unobserved variables white.
+Black lines correspond to a connection from the top-down
+generative model, while red lines correspond to connections from the bottom-up recognition model.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+  \begin{scope}[shift={(2.5,0)}]
+    \node[obs] at (3,3) (dx){$\mathcal{D}$};
+    \node[latent] at (3.5,1.75) (zp){$p$};
+    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
+    \node[obs] at (3.5,0.7) (xp) {$x$};
+    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
+    \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
+    \draw [->,red] (tx) -- (zp);
+    \draw [->] (dx) -- (zp);
+    \node at (nn) {
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle,width=2cm] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
+  \end{scope}
+
+  \begin{scope}[shift={(0.7,-4.2)}]
+    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+    \draw [->] (p2.south) -- (x2.north);
+    \draw [->] (dt.south) -- (p2.north);
+
+    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+    \draw [->] (p1.south) -- (x1.north);
+    \draw [->] (dt.south) -- (p1.north);
+
+    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+    \draw [->] (p3.south) -- (x3.north);
+    \draw [->] (dt.south) -- (p3.north);
+
+    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+    \draw [->] (zp.south) -- (xp.north);
+    \plate {}{(zp)(xp)}{$x\in X$};
+
+    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
+    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
+  \end{scope}
+
+  \begin{scope}[shift={(7.7,-4)}]
+    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+    \node[latent] at (1.5,3) (t){$\theta$};
+    \node[obs] at (1,1.75) (z){$p$};
+    % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+    \node[obs] at (1,0.5) (x) {$x$};
+    \edge {z}{x};
+    \edge {d,t}{z};
+    \plate {}{(z)(x)}{$x\in X$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
+  \end{scope}
+
+\end{tikzpicture}
+\end{figure}
 

### 2018-02-09 supplement.tex
reorganize DSL listings

@@
 
-\vfill\null\columnbreak
-
 \section{Initial domain-specific languages}
-\subsection{List functions}
-DSL for the list function experiment domain is shown below. The lower
-section with seven primitives designates those given to the \emph{rich} DSL
-variant and not the \emph{base} variant. The rich variant is not supplied
-the \texttt{if} primitive, because its primary intention is to enable the
-learning of the \texttt{filter}, \texttt{index}, and \texttt{slice}
-concepts.
-
-\begin{center}
-\tt
-\begin{tabular}{| l | l |}
-  \hline
-  \textrm{\emph{name}} & \textrm{\emph{type}} \\
-  \hline
-    empty & [$\alpha$] \\
-    singleton & $\alpha$ $\rightarrow$ [$\alpha$] \\
-    concat & [$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
-    mapi & (int $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) \\
-    & \qquad\qquad$\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$] \\
-    reducei & ($\beta$ $\rightarrow$ int $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$)\\
-    & \qquad\qquad$\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$ \\
-    \hline
-    true & bool \\
-    not & bool $\rightarrow$ bool \\
-    and & bool $\rightarrow$ bool $\rightarrow$ bool \\
-    or & bool $\rightarrow$ bool $\rightarrow$ bool \\
-    if & bool $\rightarrow$ $\alpha$ $\rightarrow$ $\alpha$ $\rightarrow$ $\alpha$ \\
-    \hline
-    0 $\cdots$ 5 & int \\
-    + & int $\rightarrow$ int $\rightarrow$ int \\
-    * & int $\rightarrow$ int $\rightarrow$ int \\
-    negate & int $\rightarrow$ int \\
-    mod & int $\rightarrow$ int $\rightarrow$ int \\
-    eq? & int $\rightarrow$ int $\rightarrow$ bool \\
-    gt? & int $\rightarrow$ int $\rightarrow$ bool \\
-    is-prime & int $\rightarrow$ bool \\
-    is-square & int $\rightarrow$ bool \\
-    range & int $\rightarrow$ [int] \\
-    sort & [int] $\rightarrow$ [int] \\
-    \hline
-    sum & [int] $\rightarrow$ int \\
-    reverse & [$\alpha$] $\rightarrow$ [$\alpha$] \\
-    all & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool \\
-    any & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool \\
-    index & int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$ \\
-    filter & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
-    slice & int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
-  \hline
-\end{tabular}
-\end{center}
-
 \subsection{Boolean circuits}
@@
   join&$\text{string}\to \left[\text{string} \right]\to \text{string}$\\
-  map&$(\text{string}\to \text{string})\to \left[\text{string} \right]\to \left[\text{string} \right]$\\
+  map&$(\text{string}\to \text{string})$\\
+  &$\qquad\quad\to\left[\text{string} \right]\to \left[\text{string} \right]$\\
   nth&$\mathbb{Z}\to \left[\text{string} \right]\to \text{string}$\\
@@
 constant characters which have the type char.
+
+\subsection{List functions}
+\begin{center}
+\tt
+\begin{tabular}{| l | l |}
+  \hline
+  \textrm{\emph{name}} & \textrm{\emph{type}} \\
+  \hline
+    empty & [$\alpha$] \\
+    singleton & $\alpha \to$ [$\alpha$] \\
+    concat & [$\alpha$] $\to$ [$\alpha$] $\to$ [$\alpha$] \\
+    mapi & $(\mathbb{Z} \to \alpha \to \beta)$ \\
+    & \qquad\qquad$\to$ [$\alpha$] $\to$ [$\beta$] \\
+    reducei & $(\beta \to \mathbb{Z} \to \alpha \to \beta)$\\
+    & \qquad\qquad$\to \beta \to$ [$\alpha$] $\to \beta$ \\
+    \hline
+    if & bool $\to \alpha \to \alpha \to \alpha$ \\
+    \hline
+    true & bool \\
+    not & bool $\to$ bool \\
+    and & bool $\to$ bool $\to$ bool \\
+    or & bool $\to$ bool $\to$ bool \\
+    0 $\cdots$ 5 & $\mathbb{Z}$ \\
+    + & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
+    * & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
+    negate & $\mathbb{Z} \to \mathbb{Z}$ \\
+    mod & $\mathbb{Z} \to \mathbb{Z} \to \mathbb{Z}$ \\
+    eq? & $\mathbb{Z} \to \mathbb{Z} \to$ bool \\
+    gt? & $\mathbb{Z} \to \mathbb{Z} \to$ bool \\
+    is-prime & $\mathbb{Z} \to$ bool \\
+    is-square & $\mathbb{Z} \to$ bool \\
+    range & $\mathbb{Z} \to$ [$\mathbb{Z}$] \\
+    sort & [$\mathbb{Z}$] $\to$ [$\mathbb{Z}$] \\
+    \hline
+    sum & [$\mathbb{Z}$] $\to \mathbb{Z}$ \\
+    reverse & [$\alpha$] $\to$ [$\alpha$] \\
+    all & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
+    any & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
+    index & $\mathbb{Z} \to$ [$\alpha$] $\to \alpha$ \\
+    filter & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ [$\alpha$] \\
+    slice & $\mathbb{Z} \to \mathbb{Z} \to$ [$\alpha$] $\to$ [$\alpha$] \\
+  \hline
+\end{tabular}
+\end{center}
+The lower section with seven primitives designates those given to the
+\emph{rich} DSL variant and not the \emph{base} variant. The rich variant is
+not supplied the \texttt{if} primitive, because its primary intention is to
+enable the learning of the \texttt{filter}, \texttt{index}, and
+\texttt{slice} concepts. Greek letters represent polymorphic types.
 

### 2018-02-09 supplement.tex
(supplement) make string editing sentence less awkward
hyper parameters

@@
 
+\vfill
+
+\section{Hyperparameters \& Implementation details}
+We set $\alpha = 10$ and $\lambda = 1$ for all experiments except for string
+editing. For string editing we found that more strongly regularizing
+the DSL was necessary,
+and so we set $\lambda = 5$.
+We believe this is because
+we have many more tasks for string editing than we do for the other domains.
+The frontier size used for Boolean circuits was $5000$,
+for symbolic regression we used $500$,
+for string editing we used $10^5$,
+and for list functions we used $10^4$.
+
+Because the frontiers can become very large in later iterations of the algorithm,
+we only keep around the top $10^4$ programs in the frontier $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$.
+
 \subsection{List functions}

### 2018-02-09 paper.tex
no more errors

@@
       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-      \node[rectangle,width=2cm] at (1,1) {};
+      \node[rectangle] at (1,1) {};
       \foreach \name / \y in {1,...,4}
@@
 
-%% \begin{figure}
-%% \centering
-%% \begin{tikzpicture}
-%%   \begin{scope}[shift={(2.5,0)}]
-%%     \node[obs] at (3,3) (dx){$\mathcal{D}$};
-%%     \node[latent] at (3.5,1.75) (zp){$p$};
-%%     \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-%%     \node[obs] at (3.5,0.7) (xp) {$x$};
-%%     \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
-%%     \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
-%%     \draw [->,red] (tx) -- (zp);
-%%     \draw [->] (dx) -- (zp);
-%%     \node at (nn) {
-%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-%%         \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-%%         \node[rectangle,width=2cm] at (1,1) {};
-%%         \foreach \name / \y in {1,...,4}
-%%             \node[neuron] (I-\name) at (0,-\y) {};
-%%         \foreach \name / \y in {1,...,3}
-%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
-%%         \foreach \name / \y in {1,...,4}
-%%             \node[neuron] (O-\name) at (2,-\y) {};
-%%         \foreach \source in {1,...,4}
-%%             \foreach \dest in {1,...,3}
-%%                 \draw [-latex] (I-\source) -- (H-\dest);
-%%         \foreach \source in {1,...,3}
-%%             \foreach \dest in {1,...,4}
-%%                 \draw [-latex] (H-\source) -- (O-\dest);
-%%       \end{tikzpicture}
-%%     };
-%%     \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-%%   \end{scope}
-
-%%   \begin{scope}[shift={(0.7,-4.2)}]
-%%     \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-%%     \draw [->] (p2.south) -- (x2.north);
-%%     \draw [->] (dt.south) -- (p2.north);
-
-%%     \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-%%     \draw [->] (p1.south) -- (x1.north);
-%%     \draw [->] (dt.south) -- (p1.north);
-
-%%     \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-%%     \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-%%     \draw [->] (p3.south) -- (x3.north);
-%%     \draw [->] (dt.south) -- (p3.north);
-
-%%     \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-%%     \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
-
-%%     \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-%%     \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-%%     \draw [->] (zp.south) -- (xp.north);
-%%     \plate {}{(zp)(xp)}{$x\in X$};
-
-%%     \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
-%%     \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
-%%   \end{scope}
-
-%%   \begin{scope}[shift={(7.7,-4)}]
-%%     \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-%%     \node[latent] at (1.5,3) (t){$\theta$};
-%%     \node[obs] at (1,1.75) (z){$p$};
-%%     % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-%%     \node[obs] at (1,0.5) (x) {$x$};
-%%     \edge {z}{x};
-%%     \edge {d,t}{z};
-%%     \plate {}{(z)(x)}{$x\in X$};
-%%     \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
-%%   \end{scope}
-
-%% \end{tikzpicture}
-%% \end{figure}
-
 Sec.~\ref{mathematicalFraming} frames this 3-step procedure as
@@
    %   \ELSE
-   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
-   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
+   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
+     \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
 
    \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
-   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
+   \STATE \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
    \STATE Unify $\tau$ with yield$(\tau')$.
@@
  \STATE\textbf{where:}
- \STATE  yield$(\tau) = \begin{cases}
-\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
+ \STATE \quad yield$(\tau) = \begin{cases}
+    \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     \tau   &\text{ otherwise.}
  \end{cases}$ 
- \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
-[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
+ \STATE \quad argTypes$(\tau) = \begin{cases}
+    [\alpha] + \text{argTypes}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
     []   &\text{ otherwise.}
@@
   \begin{circuitikz} \draw
-    (0,0) node[port] (x) {$x$}
-    (x)++(0,-0.3) node[port] (y) {$y$}
-    (y)++(0,-0.5) node[port] (z) {$z$}
-
-    (x)++(1.7,-0.15) node[nand port, scale = 0.5] (xy) {}
+    (0,0) node (x) {$x$}
+    (x)++(0,-0.3) node (y) {$y$}
+    (y)++(0,-0.5) node (z) {$z$}
+
+    (x)++(1.7,-0.14) node[nand port, scale = 0.5] (xy) {}
     (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
@@
     (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
-    (x.out) -- (xy.in 1)
-    (y.out) -- (xy.in 2)
-
-    (x.out) -| (xz.in 1)
-    (z.out) -- (xz.in 2)
+    (x) -- (xy.in 1)
+    (y) -- (xy.in 2)
+
+    (x) -| (xz.in 1)
+    (z.east) -- (xz.in 2)
 
     (xz.out) -- (zxz.in 2)
-    (z.out) |- (zxz.in 1)
+    (z.east) |- (zxz.in 1)
 
@@
 
-    (m)++(1.2,0) node [port] (l) {\textsc{MUX}$(x,y,z)$}
+    (m)++(1.2,0) node (l) {\textsc{MUX}$(x,y,z)$}
     ;
@@
   \begin{circuitikz} \draw
-    (0,0) node[port] (a) {$a$}
-    (a)++(0,-0.7) node[port] (b) {$b$}
+    (0,0) node (a) {$a$}
+    (a)++(0,-0.7) node (b) {$b$}
 
@@
 
-    (a.out) -| (ab1.in 1)
-    (a.out) -| (ab1.in 2)
-    (b.out) -| (ab2.in 1)
-    (b.out) -| (ab2.in 2)
+    (a.east) -| (ab1.in 1)
+    (a.east) -| (ab1.in 2)
+    (b.east) -| (ab2.in 1)
+    (b.east) -| (ab2.in 2)
 
@@
 
-    (o)++(0.8,0) node [port] (l) {\textsc{OR}$(a,b)$}
+    (o)++(0.8,0) node (l) {\textsc{OR}$(a,b)$}
     ;
@@
   \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
-    (0,0) node[port] (w) {$w$}
-    (w)++(0,-0.5) node[port] (x) {$x$}
-    (x)++(0,-0.5) node[port] (y) {$y$}
-    (y)++(0,-0.5) node[port] (z) {$z$}
-
-    (z.out) -| (m.south)
-    (y.out) -| (m.west)
-    (x.out) -| (m.north)
+    (0,0) node (w) {$w$}
+    (w)++(0,-0.5) node (x) {$x$}
+    (x)++(0,-0.5) node (y) {$y$}
+    (y)++(0,-0.5) node (z) {$z$}
+
+    (z.east) -| (m.south)
+    (y.east) -| (m.west)
+    (x.east) -| (m.north)
 

### 2018-02-09 paper.tex
tiny edit in abstract
hyphenate domain-specific. CHANGES TITLE. Kevin make sure this is alright

@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Inducing Domain Specific Languages for Bayesian Program Learning}
+\icmltitlerunning{Inducing Domain-Specific Languages for Bayesian Program Learning}
 
@@
 \twocolumn[
-\icmltitle{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+\icmltitle{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
   in each case showing that
-  \system learns a domain-specific vocabulary for expressing solutions to the problems in the domain.
+  \system learns a domain-specific vocabulary for expressing solutions to problems in the domain.
 
@@
 approaches -- symbolic, neural and neural-symbolic -- rely upon a
-hand-engineered \emph{Domain Specific Language} (DSL). The DSL is an
-inventory of restricted programming primitives, encoding domain
-specific knowledge about the space of programs. In
+hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
+inventory of restricted programming primitives, encoding domain-specific
+knowledge about the space of programs. In
 practice
@@
 along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
-A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
+A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
 We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together

### 2018-02-09 paper.tex
edits to introduction

@@
 program to be induced, and thus success often hinges on having a good
-DSL, to
+DSL to
 provide a crucial inductive bias for what would otherwise be an
@@
 learn a DSL completely from scratch, but rather to start from a weaker
-or more general set of primitives in a Lisp-like representation and
-construct a richer, more powerful and better-tuned DSL.
+or more general set of primitives and
+construct a richer, more powerful, and better-tuned DSL.
 
@@
 between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve each of the tasks; creating a DSL by discovering
+programs that solve tasks; creating a DSL by discovering
 and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL. The DSL
-distills commonalities across programs that solve the tasks, letting
-the agent solve other related program learning problems. The neural
+that efficiently guides search for programs in the DSL. The learned DSL
+distills commonalities across programs that solve tasks, helping
+the agent solve related program learning problems. The neural
 network ensures that searching for program solutions remains tractable
@@
 A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
-We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together
+We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
 $(\mathcal{D},\theta)$ define a distribution over programs.
@@
   \end{tikzpicture}
-  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$.
-Observe tasks $x$ (program inputs/outputs).
+  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$.
+From each program we observe a task $x$ (program inputs/outputs).
 (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
@@
 and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
-For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
+For us, a \textbf{Wake} cycle means using the recognition model to search for programs that solve tasks.
 A sleep cycle means  updating these models so that the DSL and  recognition model
@@
  its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
- \begin{table}
+ \begin{table}[b]
     \begin{tabular}{ll}

### 2018-02-09 paper.tex
edits to 2.0

@@
    &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
-   Lists& \code{(map (lambda (x) (+ x a)) l)}\\
-   &\hspace{0.5cm} \emph{(add a to every element of list l)}
+   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
+   &\hspace{0.5cm} \emph{(add k to every element of list l)}
    \\\bottomrule
@@
 \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-When these improve we find better programs.
+  When these improve, we find better programs.
 \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
-  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
+\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we found which solve the tasks.
+  As we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
 
@@
 a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
-Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Sec.~\ref{recognitionSection} how we train a neural network to accelerate the search over programs; and
-Sec.~\ref{grammarInductionSection}  how \system induces a DSL from programs.
+Sec.~\ref{explorationSection} explains how we search for programs that solve tasks;
+Sec.~\ref{recognitionSection} explains how we train a neural network to accelerate program search; and
+Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
 

### 2018-02-09 paper.tex
edits to 2.1

@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}(a).
-Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*,\theta^*$ solving:
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed by Fig.~\ref{graphicalModel}(a).
+Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  \nonumber
+    J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
   \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
-\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
+  \nonumber
+    \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
 \end{align}
@@
 \begin{align}
- J\geq \lowerBound\triangleq\probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
+ J\geq \lowerBound\triangleq\probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
 \end{align}
@@
 \system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
-alternatingly maximizing w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
+alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
 %So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
@@
 $\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
-in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
+in the frontiers will most likely increase $\lowerBound$.  %% With $q$ in hand we can find programs for
 %% the frontier $\mathcal{F}_x$ by searching for programs that maximize
 %% $q(p|x)$.
-The recognition model exploits the structure of the DSL: %$\mathcal{D}$:
-rather than directly predicting a distribution over $p$ conditioned on $x$,
-it predicts a distribution over components of the DSL.
+Rather than directly predicting a distribution over $p$ conditioned on $x$,
+the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
 This taps into the intuition that programming is primarily a top-down activity:

### 2018-02-09 paper.tex
capitalize refs to Supplement

@@
 and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
-We explain in the supplement that maximizing an ELBO-style bound
+We explain in the Supplement that maximizing an ELBO-style bound
 leads to a pathological behavior that  causes the model to easily become trapped in local optima.
@@
 and for domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
-We MaxPool across all of the input/output examples. See supplement for details.
+We MaxPool across all of the input/output examples. See Supplement for details.
 

### 2018-02-09 paper.tex
edits to 2.2

@@
 Our programs are all strongly typed.
-We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is used in functional languages like OCaml.
-Variables that range over types are always written using lowercase Greek letters
-and we write $\alpha\to \beta$ to mean a function that takes as input type $\alpha$
+We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
+used in functional programming languages like OCaml.
+Type variables are always written using lowercase Greek letters
+and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
 and returns something of type $\beta$.
@@
 (2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-  the space of of programs.
-
-A main drawback of an enumerative search algorithm is that we have no
+  the space of programs.
+
+A drawback of using an enumerative search algorithm is that we have no
 efficient means of solving for arbitrary constants that might occur in the

### 2018-02-09 paper.tex
edits to 2.3

@@
 
-The purpose of the recognition model is to accelerate the search over
-programs.  It does this by learning to predict which programs both 
-assign high likelihood to a task, and at the same time 
-have high probability under $(\mathcal{D},\theta)$.
+The purpose of the recognition model is to accelerate the search for
+programs.  It does this by learning to predict programs which are probable
+under $(\mathcal{D},\theta)$ while also assigning high likelihood for a task
+according to $\probability[x|p]$.
 
@@
 We want a recognition model that closely approximates the true posteriors over programs.
-We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]$,
-equivalently maximizing
+We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
+or equivalently maximizing
 \begin{equation*}
@@
 like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
-We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
+We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
 \begin{align}
@@
 all of our experiments train \system on only a few hundred tasks, which is too little for
-a high-capacity $q$.
-Once we have bootstrapped a $(\mathcal{D},\theta)$,
-we can draw unlimited samples from $(\mathcal{D},\theta)$,
+a high-capacity neural network $q$.
+Once we bootstrap a $(\mathcal{D},\theta)$,
+we can draw unlimited samples from $(\mathcal{D},\theta)$
 and train $q$ on those samples.
@@
 For domains with sequential structure like strings and lists,
-we use an RNN to encode each input/output example,
-and for domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
+we use an RNN to encode each input/output example.
+For domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
 We MaxPool across all of the input/output examples. See Supplement for details.

### 2018-02-09 paper.tex
edits to 2.4

@@
 \begin{align}
-\nonumber  \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
-&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
+  \nonumber
+    \log \probability[\mathcal{D}] + \argmax_{\theta}&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  &+ \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
   \end{align}
@@
   \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
-  P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)
+  \probability[\theta|\mathcal{D}]& = \text{Dir}(\theta|\alpha)
 \end{align*}
@@
 Although this problem may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
+very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
 for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},

### 2018-02-09 paper.tex
edits to 2.5

@@
 We note the following implementation details:
-(1) We interleave wake cycles with each of the sleep cycles;
-(2) on the first iteration, we do \emph{not} train the recognition model on samples from
+(1) We perform a wake cycle before each of the sleep cycles.
+(2) On the first iteration, we do \emph{not} train the recognition model on samples from
 the generative model because a DSL has not yet been learned --

### 2018-02-09 paper.tex
edits to 3.1

@@
 logic gates and multiplexers (Fig.~\ref{multiplexer}).
-We will show data showing that the neural recognition model
+We will show that the neural recognition model
 assists in building these circuits (Sec.~\ref{quantitative}),
@@
   \end{circuitikz}
-\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right), which it was unable to build without discovering these new primitives.}\label{multiplexer}
+\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right) that it was unable to build without discovering these new primitives.}\label{multiplexer}
   \end{figure}

### 2018-02-09 paper.tex
edits to 3.2

@@
 Examining the programs themselves,
-one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
-for example, it represents the linear function $-x+2$ with the program 
-\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}.
-This falls out of our Bayesian framing -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
+one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom.
+For example, it represents the linear function $-x+2$ with the program
+\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
+which has five continuous degrees of freedom.
+This arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
 \begin{table}\centering

### 2018-02-09 paper.tex
back to 8 pages
edits to 3.3

@@
   \end{tabular}
-  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
+  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
     These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\code{z} and \code{s} in the right column).} 
@@
 We show in this experiment that we can instead start out with a
-less specialized set of primitives -- close to the string routines provided by Python --
+less specialized set of primitives -- close to the string methods provided by Python --
 and recover many of the higher-level building blocks that have made these
@@
 We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
-At first, \system cannot find any programs that
-work at all for most of the tasks.
+At first, \system cannot find programs for most of the tasks.
 After three wake/sleep iterations,
@@
 
-  \begin{minipage}{7.5cm}
+  \begin{minipage}{\linewidth}
+    \centering
     $f($\code{s}$) \,=\, $\code{(}$f_3$\code{ (split '>' (upper s)) '<')}
   \end{minipage}
-  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}.
+  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
   \end{figure}

### 2018-02-09 paper.tex
edits to 3.5

@@
 compares our model against these baselines.
-Our full model consistently
-beats the baselines,
-and that the recognition model 
-consistently increases the number of solved held-out tasks.
-Lesioning the recognition model also slows down the convergence of the algorithm,
+Our full model consistently beats the baselines.
+Lesioning the recognition model slows down the convergence of the algorithm,
 taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).

### 2018-02-09 paper.tex
edits to related work
add "from examples" to conclusion

@@
 \section{Related Work}
- Our work is far from the first to learn to learn programs,
+ Our work is far from the first for learning to learn programs,
  an idea that goes back to Solomonoff~\cite{solomonoff1989system}:
@@
  \section{Contribution and Outlook}
- We contributed a model, \systemEnding, that learns to program by
+ We contributed a model, \systemEnding, that learns to program from examples by
  bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.

### 2018-02-09 supplement.tex
no more errors
edits to supplement

@@
 %\usepackage{epsfig} % less modern
-\usepackage{subfig} 
+%\usepackage{subfig}
 \usepackage{fancyvrb}
@@
 \usepackage{hyperref}
-\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
+%\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
 \DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
@@
         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle,width=2cm] at (1,1) {};
+        \node[rectangle] at (1,1) {};
         \foreach \name / \y in {1,...,4}
@@
     \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta})$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
   \end{scope}
@@
 A $\lambda$-calculus expression is either:
-\\\noindent A \emph{primitive}, like the number 5 or
-  the function \texttt{sum}.
-\\\noindent A \emph{variable}, like $x$, $y$, $z$
-\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function.
+\begin{itemize}
+  \item[--] A \emph{primitive}, like the number 5 or the function \texttt{sum}.
+  \item[--] A \emph{variable}, like $x$, $y$, or $z$.
+  \item[--] A $\lambda$\emph{-abstraction}, which creates a new function.
 $\lambda$-abstractions have a variable and a body.
@@
 Abstractions are written as $\lambda \text{var}.
-\text{body}$.
-\\\noindent An \emph{application} of a function to an argument.
+\text{body}$ or in Lisp syntax as \mbox{\texttt{(lambda (\textrm{var}) \textrm{body})}}.
+  \item[--] An \emph{application} of a function to an argument.
 Both the function and the argument are $\lambda$-calculus expressions.
-The application of the function $f$ to the argument $x$ is written as $f\; x$.
+The application of the function $f$ to the argument $x$ is written as $f\; x$ or as \texttt{($f$ x)}.
+\end{itemize}
 
 For example, the function which squares the logarithm of a number is
-$\lambda x.\text{\texttt{square}} (\text{\texttt{log} } x)$, and the identity function $f(x) = x$ is $\lambda x.x$. The
+$\lambda x.\text{\tt(square (log $x$))}$, and the identity function $f(x) = x$ is $\lambda x.x$. The
 $\lambda$-calculus serves as a spartan but expressive Turing complete
-program representation, and distills the essential features of functional languages like Lisp.
+program representation, and distills the essential features of functional
+programming languages like Lisp.
 
@@
 Hindley-Milner has two important features:
-Feature 1: It supports \emph{parametric polymorphism}: meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
+Feature 1: It supports \emph{parametric polymorphism}, meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
 For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
-A function that returns the first element of a list has the type $\texttt{list}(\alpha)\to\alpha$.
-Type variables are not the same has variables introduced by $\lambda$-abstractions.
+A function that returns the first element of a list has the type $[\alpha]\to\alpha$.
+Type variables are not the same as variables introduced by $\lambda$-abstractions.
 Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
-Our generative model over programs performs Hindley-Milner type inference during sampling;
-\texttt{unify} in the generative model uses the machinery of Hindley-Milner to
+Our generative model over programs performs Hindley-Milner type inference during sampling:
+\emph{Unify} in the generative model uses the machinery of Hindley-Milner to
 ensure that the generated programs have valid polymorphic types.
@@
 
-We use Algorithm~\ref{grammarInductionAlgorithm} to search for the structure of the DSL that best explains the frontiers.
+We use Alg.~\ref{grammarInductionAlgorithm} to search for the structure of the DSL that best explains the frontiers.
 
+\setcounter{algorithm}{2} % there are two algorithms in the main paper.
 \begin{algorithm}[tb]
@@
      \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
-     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
+     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - \|\theta\|_0$
      \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
@@
 
-We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.~1 of the main paper.
+We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$; $c(p)= \sum_{e\in \mathcal{D}}c(e,p)$ to mean the total number of primitives used in program $p$; $R(p)$ to mean the sequence of types input to sample in Alg.~1.
 Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
@@
 \end{align*}
-where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.
+where  $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.
 
-Differentiate with respect to $\theta_e$ and set to zero 
-\begin{align}
-  \frac{c(x)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
+Differentiate with respect to $\theta_e$ and set to zero:
+%\begin{align*}
+%  \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}}
+%\end{align*}
+%or more succinctly,
+\begin{align*}
+  \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
 \end{align}
+where $N=c(p)$ and $a(x) = r(x, p)$. From here forward, we similarly let
+$c(x)=c(x,p)$ for notational convenience.
+
 This equality holds if $\theta_x = c(x)/a(x)$:
@@
 \begin{align}
-  J& = L + \log \text{D}(\theta|\alpha)\\
+  J& = L + \log \text{Dir}(\theta|\alpha)\\
   &\stackrel{+}{\geq } \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x  + \sum_x(\alpha_x - 1)\log \theta_x\\
@@
 
-When is this lower bound tight? This lower bound is tight whenever all
+This lower bound is tight whenever all
 of the types of the expressions in the DSL are not polymorphic, in which case our DSL is equivalent to a PCFG
@@
 
-\section{Hyperparameters \& Implementation details}
-We set $\alpha = 10$ and $\lambda = 1$ for all experiments except for string
-editing. For string editing we found that more strongly regularizing
-the DSL was necessary,
-and so we set $\lambda = 5$.
-We believe this is because
-we have many more tasks for string editing than we do for the other domains.
-The frontier size used for Boolean circuits was $5000$,
-for symbolic regression we used $500$,
-for string editing we used $10^5$,
-and for list functions we used $10^4$.
-
-Because the frontiers can become very large in later iterations of the algorithm,
-we only keep around the top $10^4$ programs in the frontier $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$.
-
 \subsection{List functions}
@@
 
+\section{Hyperparameters \& Implementation details}
+We set $\alpha = 10$ and $\lambda = 1$ for all experiments except for string
+editing. For string editing we found that more strongly regularizing
+the DSL was necessary,
+and so we set $\lambda = 5$.
+We believe this is because
+we have many more tasks for string editing than we do for the other domains.
+The frontier size used for Boolean circuits was $5000$,
+for symbolic regression we used $500$,
+for string editing we used $10^5$,
+and for list functions we used $10^4$.
+
+Because the frontiers can become very large in later iterations of the algorithm,
+we only keep around the top $10^4$ programs in the frontier $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$.
+
 \section{Why not the ELBO bound?}
@@
 could replace $Q_x(p)$ with the recognition model $q(p|x)$, either
-using importance sampling (so the expectation in~\ref{EM} is taken
-over $q$ and we reweight using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.
+using importance sampling (so the expectation in Eq.~\ref{EM} is taken
+over $q$ and we reweigh using $Q_x$) or by directly using $q$ as our approximate posterior over the program that solves task $x$.
 
@@
 we want the DSL induction to be a function \emph{only} of the programs that we have found,
-and \emph{not} be a function of how the previous generative model weighted them.
+and \emph{not} be a function of how the previous generative model weighed them.
 In practice,

### 2018-02-09 supplement.tex
updates to list supplement

@@
 %or more succinctly,
-\begin{align*}
+\begin{align}
   \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
@@
 \subsection{List functions}
+The lower section with seven primitives designates those given to the
+\emph{rich} DSL variant and not the \emph{base} variant. The rich variant is
+not supplied the \texttt{if} primitive, because its primary intention is to
+enable the learning of the lower five concepts.
+Greek letters represent polymorphic types.
 \begin{center}
@@
     not & bool $\to$ bool \\
-    and & bool $\to$ bool $\to$ bool \\
-    or & bool $\to$ bool $\to$ bool \\
+%UNUSED+UNNECESSARY:    and & bool $\to$ bool $\to$ bool \\
+%UNUSED+UNNECESSARY:    or & bool $\to$ bool $\to$ bool \\
     0 $\cdots$ 5 & $\mathbb{Z}$ \\
@@
 \end{center}
-The lower section with seven primitives designates those given to the
-\emph{rich} DSL variant and not the \emph{base} variant. The rich variant is
-not supplied the \texttt{if} primitive, because its primary intention is to
-enable the learning of the \texttt{filter}, \texttt{index}, and
-\texttt{slice} concepts. Greek letters represent polymorphic types.
+
+\vfill
 

### 2018-02-09 paper.tex
miscellaneous edits

--- 2018-02-09 paper.tex
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Inducing Domain-Specific Languages for Bayesian Program Learning}
+\icmltitlerunning{Inducing Domain Specific Languages for Bayesian Program Learning}
 
@@
 \twocolumn[
-\icmltitle{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
+\icmltitle{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
   in each case showing that
-  \system learns a domain-specific vocabulary for expressing solutions to problems in the domain.
+  \system learns a domain-specific vocabulary for expressing solutions to the problems in the domain.
 
@@
 approaches -- symbolic, neural and neural-symbolic -- rely upon a
-hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
-inventory of restricted programming primitives, encoding domain-specific
-knowledge about the space of programs. In
+hand-engineered \emph{Domain Specific Language} (DSL). The DSL is an
+inventory of restricted programming primitives, encoding domain
+specific knowledge about the space of programs. In
 practice
@@
 program to be induced, and thus success often hinges on having a good
-DSL to
+DSL, to
 provide a crucial inductive bias for what would otherwise be an
@@
 learn a DSL completely from scratch, but rather to start from a weaker
-or more general set of primitives and
-construct a richer, more powerful, and better-tuned DSL.
+or more general set of primitives in a Lisp-like representation and
+construct a richer, more powerful and better-tuned DSL.
 
@@
 between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve tasks; creating a DSL by discovering
+programs that solve each of the tasks; creating a DSL by discovering
 and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL. The learned DSL
-distills commonalities across programs that solve tasks, helping
-the agent solve related program learning problems. The neural
+that efficiently guides search for programs in the DSL. The DSL
+distills commonalities across programs that solve the tasks, letting
+the agent solve other related program learning problems. The neural
 network ensures that searching for program solutions remains tractable
@@
 along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
-A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
-We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
+A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
+We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together
 $(\mathcal{D},\theta)$ define a distribution over programs.
@@
       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-      \node[rectangle] at (1,1) {};
+      \node[rectangle,width=2cm] at (1,1) {};
       \foreach \name / \y in {1,...,4}
@@
   \end{tikzpicture}
-  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$.
-From each program we observe a task $x$ (program inputs/outputs).
+  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$.
+Observe tasks $x$ (program inputs/outputs).
 (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
@@
 and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
-For us, a \textbf{Wake} cycle means using the recognition model to search for programs that solve tasks.
+For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
 A sleep cycle means  updating these models so that the DSL and  recognition model
@@
  its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
- \begin{table}[b]
+ \begin{table}
     \begin{tabular}{ll}
@@
    &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
-   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
-   &\hspace{0.5cm} \emph{(add k to every element of list l)}
+   Lists& \code{(map (lambda (x) (+ x a)) l)}\\
+   &\hspace{0.5cm} \emph{(add a to every element of list l)}
    \\\bottomrule
@@
 \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-  When these improve, we find better programs.
+When these improve we find better programs.
 \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we found which solve the tasks.
-  As we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
+\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
+  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
 
@@
 a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
-Sec.~\ref{explorationSection} explains how we search for programs that solve tasks;
-Sec.~\ref{recognitionSection} explains how we train a neural network to accelerate program search; and
-Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
+Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
+Sec.~\ref{recognitionSection} how we train a neural network to accelerate the search over programs; and
+Sec.~\ref{grammarInductionSection}  how \system induces a DSL from programs.
 
@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed by Fig.~\ref{graphicalModel}(a).
-Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
+We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}(a).
+Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*,\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
-  \nonumber
-    J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
   \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
-  \nonumber
-    \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
+\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
 \end{align}
@@
 \begin{align}
- J\geq \lowerBound\triangleq\probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
+ J\geq \lowerBound\triangleq\probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
 \end{align}
@@
 \system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
+alternatingly maximizing w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
 %So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
@@
 $\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
-in the frontiers will most likely increase $\lowerBound$.  %% With $q$ in hand we can find programs for
+in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
 %% the frontier $\mathcal{F}_x$ by searching for programs that maximize
 %% $q(p|x)$.
-Rather than directly predicting a distribution over $p$ conditioned on $x$,
-the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
+The recognition model exploits the structure of the DSL: %$\mathcal{D}$:
+rather than directly predicting a distribution over $p$ conditioned on $x$,
+it predicts a distribution over components of the DSL.
 This taps into the intuition that programming is primarily a top-down activity:
@@
 and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
-We explain in the Supplement that maximizing an ELBO-style bound
+We explain in the supplement that maximizing an ELBO-style bound
 leads to a pathological behavior that  causes the model to easily become trapped in local optima.
@@
 Our programs are all strongly typed.
-We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
-used in functional programming languages like OCaml.
-Type variables are always written using lowercase Greek letters
-and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
+We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is used in functional languages like OCaml.
+Variables that range over types are always written using lowercase Greek letters
+and we write $\alpha\to \beta$ to mean a function that takes as input type $\alpha$
 and returns something of type $\beta$.
@@
    %   \ELSE
-   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
-     \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
+   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
 
    \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
-   \STATE \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
+   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
    \STATE Unify $\tau$ with yield$(\tau')$.
@@
  \STATE\textbf{where:}
- \STATE \quad yield$(\tau) = \begin{cases}
-    \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+ \STATE  yield$(\tau) = \begin{cases}
+\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
 \tau   &\text{ otherwise.}
  \end{cases}$ 
- \STATE \quad argTypes$(\tau) = \begin{cases}
-    [\alpha] + \text{argTypes}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+ \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
+[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
 []   &\text{ otherwise.}
@@
 (2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-  the space of programs.
-
-A drawback of using an enumerative search algorithm is that we have no
+  the space of of programs.
+
+A main drawback of an enumerative search algorithm is that we have no
 efficient means of solving for arbitrary constants that might occur in the
@@
 
-The purpose of the recognition model is to accelerate the search for
-programs.  It does this by learning to predict programs which are probable
-under $(\mathcal{D},\theta)$ while also assigning high likelihood for a task
-according to $\probability[x|p]$.
+The purpose of the recognition model is to accelerate the search over
+programs.  It does this by learning to predict which programs both 
+assign high likelihood to a task, and at the same time 
+have high probability under $(\mathcal{D},\theta)$.
 
@@
 We want a recognition model that closely approximates the true posteriors over programs.
-We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
-or equivalently maximizing
+We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]||q(p|x) \right) \right]$,
+equivalently maximizing
 \begin{equation*}
@@
 like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
-We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
+We found it useful to maximize both an autoencoder-style objective (written $\mathcal{L}_{\text{AE}}$) and a Helmholtz-style objective ($\mathcal{L}_{\text{HM}}$), giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
 \begin{align}
@@
 all of our experiments train \system on only a few hundred tasks, which is too little for
-a high-capacity neural network $q$.
-Once we bootstrap a $(\mathcal{D},\theta)$,
-we can draw unlimited samples from $(\mathcal{D},\theta)$
+a high-capacity $q$.
+Once we have bootstrapped a $(\mathcal{D},\theta)$,
+we can draw unlimited samples from $(\mathcal{D},\theta)$,
 and train $q$ on those samples.
@@
 For domains with sequential structure like strings and lists,
-we use an RNN to encode each input/output example.
-For domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
-We MaxPool across all of the input/output examples. See Supplement for details.
+we use an RNN to encode each input/output example,
+and for domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
+We MaxPool across all of the input/output examples. See supplement for details.
 
@@
 \begin{align}
-  \nonumber
-    \log \probability[\mathcal{D}] + \argmax_{\theta}&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  &+ \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
+\nonumber  \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
+&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
   \end{align}
@@
   \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
-  \probability[\theta|\mathcal{D}]& = \text{Dir}(\theta|\alpha)
+  P(\theta|\mathcal{D})& = \text{Dir}(\theta|\alpha)
 \end{align*}
@@
 Although this problem may seem 
-very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
+very similar to estimating the parameters of a probabilistic context free grammar,
 for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
@@
   \end{tabular}
-  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
+  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
     These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\code{z} and \code{s} in the right column).} 
@@
 We note the following implementation details:
-(1) We perform a wake cycle before each of the sleep cycles.
-(2) On the first iteration, we do \emph{not} train the recognition model on samples from
+(1) We interleave wake cycles with each of the sleep cycles;
+(2) on the first iteration, we do \emph{not} train the recognition model on samples from
 the generative model because a DSL has not yet been learned --
@@
 logic gates and multiplexers (Fig.~\ref{multiplexer}).
-We will show that the neural recognition model
+We will show data showing that the neural recognition model
 assists in building these circuits (Sec.~\ref{quantitative}),
@@
   \begin{circuitikz} \draw
-    (0,0) node (x) {$x$}
-    (x)++(0,-0.3) node (y) {$y$}
-    (y)++(0,-0.5) node (z) {$z$}
-
-    (x)++(1.7,-0.14) node[nand port, scale = 0.5] (xy) {}
+    (0,0) node[port] (x) {$x$}
+    (x)++(0,-0.3) node[port] (y) {$y$}
+    (y)++(0,-0.5) node[port] (z) {$z$}
+
+    (x)++(1.7,-0.15) node[nand port, scale = 0.5] (xy) {}
     (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
@@
     (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
-    (x) -- (xy.in 1)
-    (y) -- (xy.in 2)
-
-    (x) -| (xz.in 1)
-    (z.east) -- (xz.in 2)
+    (x.out) -- (xy.in 1)
+    (y.out) -- (xy.in 2)
+
+    (x.out) -| (xz.in 1)
+    (z.out) -- (xz.in 2)
 
     (xz.out) -- (zxz.in 2)
-    (z.east) |- (zxz.in 1)
+    (z.out) |- (zxz.in 1)
 
@@
 
-    (m)++(1.2,0) node (l) {\textsc{MUX}$(x,y,z)$}
+    (m)++(1.2,0) node [port] (l) {\textsc{MUX}$(x,y,z)$}
     ;
@@
   \begin{circuitikz} \draw
-    (0,0) node (a) {$a$}
-    (a)++(0,-0.7) node (b) {$b$}
+    (0,0) node[port] (a) {$a$}
+    (a)++(0,-0.7) node[port] (b) {$b$}
 
@@
 
-    (a.east) -| (ab1.in 1)
-    (a.east) -| (ab1.in 2)
-    (b.east) -| (ab2.in 1)
-    (b.east) -| (ab2.in 2)
+    (a.out) -| (ab1.in 1)
+    (a.out) -| (ab1.in 2)
+    (b.out) -| (ab2.in 1)
+    (b.out) -| (ab2.in 2)
 
@@
 
-    (o)++(0.8,0) node (l) {\textsc{OR}$(a,b)$}
+    (o)++(0.8,0) node [port] (l) {\textsc{OR}$(a,b)$}
     ;
@@
   \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
-    (0,0) node (w) {$w$}
-    (w)++(0,-0.5) node (x) {$x$}
-    (x)++(0,-0.5) node (y) {$y$}
-    (y)++(0,-0.5) node (z) {$z$}
-
-    (z.east) -| (m.south)
-    (y.east) -| (m.west)
-    (x.east) -| (m.north)
+    (0,0) node[port] (w) {$w$}
+    (w)++(0,-0.5) node[port] (x) {$x$}
+    (x)++(0,-0.5) node[port] (y) {$y$}
+    (y)++(0,-0.5) node[port] (z) {$z$}
+
+    (z.out) -| (m.south)
+    (y.out) -| (m.west)
+    (x.out) -| (m.north)
 
@@
   \end{circuitikz}
-\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right) that it was unable to build without discovering these new primitives.}\label{multiplexer}
+\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right), which it was unable to build without discovering these new primitives.}\label{multiplexer}
   \end{figure}
@@
 Examining the programs themselves,
-one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom.
-For example, it represents the linear function $-x+2$ with the program
-\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
-which has five continuous degrees of freedom.
-This arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
+one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
+for example, it represents the linear function $-x+2$ with the program 
+\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}.
+This falls out of our Bayesian framing -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
 \begin{table}\centering
@@
 We show in this experiment that we can instead start out with a
-less specialized set of primitives -- close to the string methods provided by Python --
+less specialized set of primitives -- close to the string routines provided by Python --
 and recover many of the higher-level building blocks that have made these
@@
 We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
-At first, \system cannot find programs for most of the tasks.
+At first, \system cannot find any programs that
+work at all for most of the tasks.
 After three wake/sleep iterations,
@@
 
-  \begin{minipage}{\linewidth}
-    \centering
+  \begin{minipage}{7.5cm}
     $f($\code{s}$) \,=\, $\code{(}$f_3$\code{ (split '>' (upper s)) '<')}
   \end{minipage}
-  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}.
   \end{figure}

### 2018-02-09 paper.tex
changed name of ours-nn

@@
 practice
-we are often given only a few examples for each
+we often have only a few input/output examples for each
 program to be induced, and thus success often hinges on having a good
-DSL, to
-provide a crucial inductive bias for what would otherwise be an
+DSL, which
+provides a crucial inductive bias for what would otherwise be an
 unconstrained search through the space of all computable functions.
@@
   \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}.
-  \end{figure}
+\vspace{-0.5cm}  \end{figure}
 
@@
 We compare  with four baselines on  held-out tasks:
-\\\noindent \textbf{Ours--NN}, which lesions the recognition model.
-\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
+\\\noindent \textbf{Ours (no $q$)}, which lesions the recognition model, $q$.
+\\\noindent \textbf{RF/DC}, which holds $(\mathcal{D},\theta)$ fixed and learns
 a recognition model only from samples from the fixed generative model.
@@
 this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
+We can not compare directly with these systems,
+because they are engineered for one specific domain, and 
+do not have publicly available code and datasets.
 \\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
@@
 \begin{tabular}{llllll}
-  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum
+  \toprule& Ours& \begin{tabular}{c}
+    Ours\\(no $q$)
+    \end{tabular}& RF/DC & PCFG & Enum
 

### 2018-02-09 paper.tex
improved experiments section

@@
 Our full model consistently
-beats the baselines,
-and that the recognition model 
-consistently increases the number of solved held-out tasks.
-Lesioning the recognition model also slows down the convergence of the algorithm,
+improves on the baselines,
+often dramatically (string editing and symbolic regression).
+The recognition model 
+consistently increases the number of solved held-out tasks,
+and lesioning it also slows down the convergence of the algorithm,
 taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).

### 2018-02-09 paper.tex
small changes to abstract and title

@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Inducing Domain Specific Languages for Bayesian Program Learning}
+\icmltitlerunning{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
 \twocolumn[
-\icmltitle{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+\icmltitle{\systemEnding: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
   and synthesize functions on lists,
-  in each case showing that
-  \system learns a domain-specific vocabulary for expressing solutions to the problems in the domain.
+  in each case showing the model
+   learns a domain-specific vocabulary for expressing solutions to the problems in the domain.
 

### 2018-02-09 paper.tex
Made titles consistent

@@
 \begin{abstract}
-  Successful approaches to program induction require a
-  hand-engineered domain-specific language (DSL),
-  constraining the space of allowed programs and imparting
-  prior knowledge of the domain.
-  We contribute a program learning algorithm
-  called \system
-  that infers a DSL
-  and trains a neural network to
-  efficiently search for programs in the learned DSL.
-  We use our model to build circuits,
-  edit strings, do symbolic regression,
-  and synthesize functions on lists,
-  in each case showing the model
-   learns a domain-specific vocabulary for expressing solutions to the problems in the domain.
+  Successful approaches to program induction require a hand-engineered
+  domain-specific language (DSL), constraining the space of allowed
+  programs and imparting prior knowledge of the domain.  We contribute
+  a program induction algorithm called \system that learns a DSL and
+  trains a neural network to efficiently search for programs in the
+  learned DSL.  We use our model to build circuits, edit strings, do
+  symbolic regression, and synthesize functions on lists, in each case
+  showing the model learns a domain-specific vocabulary for expressing
+  solutions to the problems in the domain.
 

### 2018-02-09 supplement.tex
Made titles consistent

@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Bootstrapping Domain Specific Languages for Bayesian Program Learning}
+\icmltitlerunning{\textsc{DreamCoder}: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
 \twocolumn[
-\icmltitle{Supplement: Bootstrapping Domain Specific Languages for Bayesian Program Learning}
+\icmltitle{Supplement: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
 

### 2018-02-09 paper.tex
miscellaneous edits

@@
 Here we ask, to what extent can we dispense with such highly
-hand-engineered DSLs?
+hand-engineered domain-specific languages?
 
@@
 distills commonalities across programs that solve the tasks, letting
-the agent solve other related program learning problems. The neural
-network ensures that searching for program solutions remains tractable
+the agent solve other related program induction problems. The neural
+network ensures that searching for programs remains tractable
 even as the DSL (and hence the search space for programs) expands.
@@
 \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-When these improve we find better programs.
+When these improve we find more programs solving the tasks.
 \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
@@
 (2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-  the space of of programs.
+  the space of  programs.
 
@@
 the current DSL, running them to get their outputs,
-and then training $q$ to regress from the outputs to the program.
+and then training $q$ to regress from the input/outputs to the program.
 Since these programs map inputs to outputs,
-we need some way of sampling these inputs as well.
-Our solution to this problem is to sample the inputs
+we need to sample the inputs as well.
+Our solution is to sample the inputs
 from the empirical observed distribution of inputs in $X$.
@@
 We first present our experimental results qualitatively,
-illustrating the DSLs we
-learn.
+illustrating the learned DSLs.
 We then show quantitative results (Sec.~\ref{quantitative})
@@
 We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples,
-and penalize continuous degrees of freedom using the BIC~\cite{Bishop:2006:PRM:1162264}.
+and penalize the use of continuous parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
 %% \begin{align*}
@@
 \end{tabular}
-\caption{Some induced DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns to build 4th order polynomials.}\label{regressionDSL}
+\caption{Some learned DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns to build 4th order polynomials.}\label{regressionDSL}
 \end{table}
@@
   \end{tabular}
-  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
+  \caption{Top: provided string editing primitives. Bottom: some learned DSL primitives.}\label{textPrimitives}
   \end{table}
@@
 \end{tabular}
-\caption{Sample tasks from our list function domain. %% We used 15 input/output
+\caption{Sample tasks from our list function domain %% We used 15 input/output
   %% examples for each task to reduce ambiguity in the task's function,
@@
   \end{tabular}
-  \caption{Some induced primitives for list functions.}\label{listinduced}
+  \caption{Some learned DSL primitives for list functions}\label{listinduced}
 \end{table}

### 2018-02-09 supplement.tex
INCOMPLETE EM description of estimating grammar parameters

@@
 
-We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$; $c(p)= \sum_{e\in \mathcal{D}}c(e,p)$ to mean the total number of primitives used in program $p$; $R(p)$ to mean the sequence of types input to sample in Alg.~1.
+We use an EM algorithm to estimate the continuous parameters of the DSL, e.g. $\theta$.
+Suppressing dependencies on $\mathcal{D}$, the EM updates are
+\begin{align}
+\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
+  Q_x(p)&\propto \probability[x|p]\probability[p|\theta]
+  \end{align}
+In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$.
+We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$; $c(p)= \sum_{e\in \mathcal{D}}c(e,p)$ to mean the total number of primitives used in program $p$; $R(p)$ to mean the sequence of types input to sample in Alg.~1 of the main paper.
 Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
@@
 \end{align*}
-where  $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.
+where we have defined
+$$r(e:\tau',p)\triangleq\sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$$
 
-Differentiate with respect to $\theta_e$ and set to zero:
+Returning to equation~\ref{maximizeStep}:
+\begin{align*}
+\log P(\theta) +& \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\geq \\
+\log   \text{Dir}(\theta|\alpha) +& \sum_x\sum_{e\in \mathcal{D}}\expect_{Q_x}\left[ c(e,p)\right]\log \theta_e -\\
+&\sum_x  \expect_{Q_x}\left[c(p)\log \sum_{e\in \mathcal{D}}r(e,p)\theta_e \right]
+  \end{align*}
+
+Differentiate with respect to $\theta_e$:
+\begin{align}
+  & \frac{\alpha + \sum_x\expect_{Q_x}\left[ c(e,p)\right]}{\theta_e} -
+\theta_e\sum_x  \expect_{Q_x}\left[c(p)\frac{r(p,e)}{\sum_{e'\in \mathcal{D}} r(e',p)\theta_{e'}} \right]  
+  \end{align}
+and set to zero:
 %\begin{align*}
@@
 
-This equality holds if $\theta_x = c(x)/a(x)$:
-\begin{align}
-  \frac{c(x)}{\theta_x} &= a(x).\\
-N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
-\end{align}
+%% This equality holds if $\theta_x = c(x)/a(x)$:
+%% \begin{align}
+%%   \frac{c(x)}{\theta_x} &= a(x).\\
+%% N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
+%% \end{align}
 If this equality holds then $\theta_x \propto c(x)/a(x)$:

### 2018-02-09 supplement.tex
example text problems

@@
 \end{figure}
+
+\section{Example program induction tasks}
+
+\subsection{String editing}
+
+We automatically generated 600 string editing problems split 75/25 train/test.
+We show representative problems:
+\pagebreak
+
+double
+
+\begin{tabular}{ll}
+        \toprule Input&Output\\\midrule
+vlxS & vlxSvlxS \\
+ JkVC & JkVCJkVC \\
+ dbxD & dbxDdbxD \\
+ gQgB & gQgBgQgB
+\\\bottomrule
+\end{tabular}
+
+Extract string delimited by ',' (inclusive),' '
+\begin{tabular}{ll}
+        \toprule Input&Output\\\midrule
+LsR,uhMZI wIL & ,uhMZI  \\
+ tMY,KdF hhT & ,KdF  \\
+ UUPdC,UAd kCgtq & ,UAd  \\
+ gQgGN,VMJ FjrXZ & ,VMJ 
+\\\bottomrule
+\end{tabular}
+
+Apply first 2 characters to input delimited by '.'
+\begin{tabular}{ll}
+        \toprule Input&Output\\\midrule
+OhCQN.DlH.LdkJ.pLh & OhDlLdpL \\
+ CZr.mveZr.dvjP.YMJpg & CZmvdvYM \\
+ bpub.VkkZ.dYdPr & bpVkdY \\
+ zjNA.wFwY.WMMqh.OFdbr & zjwFWMOF
+\\\bottomrule
+\end{tabular}
+
+Apply first 2 characters to string delimited by ' ','.'
+\begin{tabular}{ll}
+        \toprule Input&Output\\\midrule
+fNA YWyL.JDD & YW \\
+ YWozt TPW.nczz & TP \\
+ RJwiT dpW.uFU & dp \\
+ FTp gNMo.JsZB & gN
+\\\bottomrule
+\end{tabular}
+
+%% Apply double to input delimited by '.'
+%% \begin{tabular}{ll}
+%%         \toprule Input&Output\\\midrule
+%% FWhb.jYMdi & FWhbFWhbjYMdijYMdi \\
+%%  JZB.WEwgH & JZBJZBWEwgHWEwgH \\
+%%  SgA.Kxo.RyF & SgASgAKxoKxoRyFRyF \\
+%%  NMApj.OaPBs.uZm.iCZy & NMApjNMApjOaPBsOaPBsuZmuZmiCZyiCZy
+%% \\\bottomrule
+%% \end{tabular}
+
+Apply strip to string delimited by ',','.'
+\begin{tabular}{ll}
+        \toprule Input&Output\\\midrule
+iNEP,  GigW .Xlm & GigW \\
+ xhpH,  QvwnW .mfK & QvwnW \\
+ weWIC,IQjhWa .fGKz & IQjhWa \\
+ xZUm, FpdE.ETcr & FpdE
+\\\bottomrule
+\end{tabular}
+
+drop first character
+
+\begin{tabular}{ll}
+        \toprule Input&Output\\\midrule
+WYPvm & YPvm \\
+ terHk & erHk \\
+ asWM & sWM \\
+ tBdT & BdT
+\\\bottomrule
+\end{tabular}
 

### 2018-02-09 paper.tex
tiny edit in abstract

@@
   showing the model learns a domain-specific vocabulary for expressing
-  solutions to the problems in the domain.
+  solutions to problems in the domain.
 

### 2018-02-09 paper.tex
edits to introduction

@@
 program to be induced, and thus success often hinges on having a good
-DSL, which
+DSL which
 provides a crucial inductive bias for what would otherwise be an
@@
 between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve each of the tasks; creating a DSL by discovering
+programs that solve tasks; creating a DSL by discovering
 and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL. The DSL
-distills commonalities across programs that solve the tasks, letting
-the agent solve other related program induction problems. The neural
+that efficiently guides search for programs in the DSL. The learned DSL
+distills commonalities across programs that solve tasks, helping
+the agent solve related program induction problems. The neural
 network ensures that searching for programs remains tractable
@@
 A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
-We equip $\mathcal{D}$ with a real-valued weight vector, $\theta$, and together
+We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
 $(\mathcal{D},\theta)$ define a distribution over programs.
@@
 and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
-For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
+For us, a \textbf{Wake} cycle means using the recognition model to search for programs that solve the tasks.
 A sleep cycle means  updating these models so that the DSL and  recognition model

### 2018-02-09 paper.tex
edits to 2.0

@@
 \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-When these improve we find more programs solving the tasks.
+When these improve, we find more programs solving the tasks.
 \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
-  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
+\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we found which solve the tasks.
+  As we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
 
@@
 Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Sec.~\ref{recognitionSection} how we train a neural network to accelerate the search over programs; and
-Sec.~\ref{grammarInductionSection}  how \system induces a DSL from programs.
+Sec.~\ref{recognitionSection} explains how we train a neural network to accelerate program search; and
+Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
 

### 2018-02-09 paper.tex
edits to 2.1

@@
 
-We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Fig.~\ref{graphicalModel}(a).
-Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*,\theta^*$ solving:
+We frame this problem as maximum a posteriori (MAP) inference in the
+generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]P(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]\probability(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
   \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\

### 2018-02-09 paper.tex
edits to 2.2

@@
 
-A main drawback of an enumerative search algorithm is that we have no
+A drawback of using an enumerative search algorithm is that we have no
 efficient means of solving for arbitrary constants that might occur in the

### 2018-02-09 paper.tex
hyphenate domain-specific. CHANGES TITLE

@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+\icmltitlerunning{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
 \twocolumn[
-\icmltitle{\systemEnding: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+\icmltitle{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
 
@@
 approaches -- symbolic, neural and neural-symbolic -- rely upon a
-hand-engineered \emph{Domain Specific Language} (DSL). The DSL is an
+hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
 inventory of restricted programming primitives, encoding domain
@@
 along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
-A DSL $\mathcal{D}$ is a collection of domain specific subroutines.
+A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
 We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together

### 2018-02-09 paper.tex
back to 8 pages
edits to 3.3

@@
 We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
-At first, \system cannot find any programs that
-work at all for most of the tasks.
+At first, \system cannot programs for most of the tasks.
 After three wake/sleep iterations,
@@
 
-  \begin{minipage}{7.5cm}
+  \begin{minipage}{\linewidth}
+    \centering
     $f($\code{s}$) \,=\, $\code{(}$f_3$\code{ (split '>' (upper s)) '<')}
   \end{minipage}
-  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}}\label{exampleTextProblem}.
-\vspace{-0.5cm}  \end{figure}
+  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+\end{figure}
 

### 2018-02-09 paper.tex
no more errors

@@
    %   \ELSE
-   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
-   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
-
-   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
-   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
+   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
+      \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
+
+   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ \\
+      \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
    \STATE Unify $\tau$ with yield$(\tau')$.
@@
  \STATE\textbf{where:}
- \STATE  yield$(\tau) = \begin{cases}
-\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
+ \STATE \quad yield$(\tau) = \begin{cases}
+   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
    \tau   &\text{ otherwise.}
  \end{cases}$ 
- \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
-[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
+ \STATE \quad argTypes$(\tau) = \begin{cases}
+   [\alpha] + \text{argTypes}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
    []   &\text{ otherwise.}

### 2018-02-09 paper.tex
hyphenate subtle "domain specific"
edits in response to Josh comments

@@
 hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
-inventory of restricted programming primitives, encoding domain
-specific knowledge about the space of programs. In
+inventory of restricted programming primitives, encoding domain-specific
+knowledge about the space of programs. In
 practice
@@
      %     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$
- %    \ENDFOR
-     \STATE $q(\cdot )\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
-     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$
-     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$
+     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Wake})}
+     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Sleep-R}})
+     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Wake)}}
+
+     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$   \footnotesize{\textbf{(Sleep-G)}}
      %% \STATE Define $Q_x(z) \propto \begin{cases}
@@
 We compare  with four baselines on  held-out tasks:
-\\\noindent \textbf{Ours (no $q$)}, which lesions the recognition model, $q$.
+\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model, $q$.
 \\\noindent \textbf{RF/DC}, which holds $(\mathcal{D},\theta)$ fixed and learns
@@
   \toprule& Ours& \begin{tabular}{c}
-    Ours\\(no $q$)
+    Ours\\(no NN)
     \end{tabular}& RF/DC & PCFG & Enum

### 2018-02-09 supplement.tex
simplified derivation of parameter estimator

@@
 
-We use an EM algorithm to estimate the continuous parameters of the DSL, e.g. $\theta$.
-Suppressing dependencies on $\mathcal{D}$, the EM updates are
-\begin{align}
-\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
-  Q_x(p)&\propto \probability[x|p]\probability[p|\theta]
-  \end{align}
-In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$.
+%% We use an EM algorithm to estimate the continuous parameters of the DSL, e.g. $\theta$.
+%% Suppressing dependencies on $\mathcal{D}$, the EM updates are
+%% \begin{align}
+%% \label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
+%%   Q_x(p)&\propto \probability[x|p]\probability[p|\theta]
+%%   \end{align}
+%% In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$.
 We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$; $c(p)= \sum_{e\in \mathcal{D}}c(e,p)$ to mean the total number of primitives used in program $p$; $R(p)$ to mean the sequence of types input to sample in Alg.~1 of the main paper.
@@
 
-Returning to equation~\ref{maximizeStep}:
+%% Returning to equation~\ref{maximizeStep}:
+%% \begin{align*}
+%% \log P(\theta) +& \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\geq \\
+%% \log   \text{Dir}(\theta|\alpha) +& \sum_x\sum_{e\in \mathcal{D}}\expect_{Q_x}\left[ c(e,p)\right]\log \theta_e -\\
+%% &\sum_x  \expect_{Q_x}\left[c(p)\log \sum_{e\in \mathcal{D}}r(e,p)\theta_e \right]
+%%   \end{align*}
+
+Differentiate with respect to $\theta_e$ and set to zero:
+%% \begin{align}
+%%   & \frac{\alpha + \sum_x\expect_{Q_x}\left[ c(e,p)\right]}{\theta_e} -
+%% \theta_e\sum_x  \expect_{Q_x}\left[c(p)\frac{r(p,e)}{\sum_{e'\in \mathcal{D}} r(e',p)\theta_{e'}} \right]  
+%%   \end{align}
+
 \begin{align*}
-\log P(\theta) +& \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\geq \\
-\log   \text{Dir}(\theta|\alpha) +& \sum_x\sum_{e\in \mathcal{D}}\expect_{Q_x}\left[ c(e,p)\right]\log \theta_e -\\
-&\sum_x  \expect_{Q_x}\left[c(p)\log \sum_{e\in \mathcal{D}}r(e,p)\theta_e \right]
+ \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}}
 \end{align*}
-
-Differentiate with respect to $\theta_e$:
-\begin{align}
-  & \frac{\alpha + \sum_x\expect_{Q_x}\left[ c(e,p)\right]}{\theta_e} -
-\theta_e\sum_x  \expect_{Q_x}\left[c(p)\frac{r(p,e)}{\sum_{e'\in \mathcal{D}} r(e',p)\theta_{e'}} \right]  
-  \end{align}
-and set to zero:
-%\begin{align*}
-%  \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}}
-%\end{align*}
 %or more succinctly,
@@
 \end{align}
-We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$).
+We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$),
+and so we estimate $\theta_e$ as the ratio of the sum of the number of times that $e$ was used, plus the pseudocounts $\alpha$, with the number of times that it could have been used, $a(e)$.
 

### 2018-02-09 paper.tex
simplified derivation of parameter estimator
another text problem example

@@
 Alg.~\ref{programGenerativeModel} is a procedure for drawing
-samples from $\probability[p|\mathcal{D},\theta]$.  In practice, we
+samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
 enumerate programs rather than sampling them.  Enumeration proceeds by
@@
 
-\subsection{Sleep-G: Inducing the DSL}\label{grammarInductionSection}
+\subsection{Sleep-G: Learning a DSL (Generative Model)}\label{grammarInductionSection}
 
@@
 (2) On the first iteration, we do \emph{not} train the recognition model on samples from
-the generative model because a DSL has not yet been learned --
+the generative model because a generative model has not yet been learned --
 we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
@@
 
-     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$   \footnotesize{\textbf{(Sleep-G)}}
+     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Sleep-G)}}
      %% \STATE Define $Q_x(z) \propto \begin{cases}
@@
 \begin{figure}
-  \centering\begin{tabular}{ll}\toprule
-    Input&Output\\\midrule
-cqbHT$>$aIKd$<$szW& 	AIKD\\
-YBW$>$TIg$<$WSho 	&TIG\\
+  \centering\begin{tabular}{ll|ll}\toprule
+    Input&Output&Input&Output\\\midrule
+cqbHT$>$aIKd$<$W& 	AIKD&BgZQ..PBBnD & BgZQ. \\
+YBW$>$TIg$<$WS 	&TIG& qqfgg..mNIx & qqfgg. 
 %DkI$>$DwKI$<$xaPoc& 	DWKI\\
-KfvTh$>$Dbl$<$OIJso 	&DBL
+%KfvTh$>$Dbl$<$OIJso 	&DBL
+\\\midrule
+\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_3$\code{ (split '>' } }&
+\multicolumn{2}{l}{$f(\text{\code{s}}) = $\code{(}$f_1$\code{ s '.' 0)}}
+\\
+\multicolumn{2}{l|}{\hspace{1.1cm}\code{(upper s)) '<')}}
 \\\bottomrule
@@
 
-  \vspace{0.25cm}
-
-  \begin{minipage}{\linewidth}
-    \centering
-    $f($\code{s}$) \,=\, $\code{(}$f_3$\code{ (split '>' (upper s)) '<')}
-  \end{minipage}
-  \caption{A string edit task and the program \system writes for it. $f_3$ is a discovered DSL subroutine, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+  %% \vspace{0.25cm}
+
+  %% \begin{minipage}{\linewidth}
+  %%   \centering
+  %%   $f($\code{s}$) \,=\, $
+  %% \end{minipage}
+
+  \caption{Two string edit tasks (top) and the programs \system writes for them (bottom). $f_3$ are $f_1$  discovered DSL subroutines, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
 \end{figure}

### 2018-02-09 paper.tex
more informative caption

@@
 FlashFill:~\cite{gulwani2011automating}), neural networks trained from
-a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill},
-DeepCoder:~\cite{balog2016deepcoder}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}) to
+a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
+symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
 synthesize programs for task domains such as string transformations,
@@
 
-Viewed as an inference problem,
+Viewed as a probabilistic inference problem,
 our goal is to infer the DSL (written $\mathcal{D}$)
@@
 network that regresses from input/output pairs to a distribution over
-programs likely to explain the input/outputs. $q(\cdot )$ implements an amortized
+programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
 inference
-scheme~\cite{le2016inference} for the generative %ritchie2016deep
-model.
+scheme~\cite{le2016inference}.
 
@@
 \end{tabular}
-\caption{Some learned DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns to build 4th order polynomials.}\label{regressionDSL}
+\caption{Some learned DSL primitives for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
 \end{table}
@@
   \end{tabular}
-\caption{\% solved: fraction solved w/ 5 sec timeout.
-Solve time: averaged over solved tasks.
+\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved tasks.
+  RF/DC: trained like RobustFill/DeepCoder.
+  PCFG: model w/o structure learning.
+  Enum: model w/o learning.
   MDL: $-\expect\left[\probability[x|p] \right]$.

### 2018-02-09 supplement.tex
another text problem example
reorganize text examples

@@
 \end{figure}
-
-\section{Example program induction tasks}
-
-\subsection{String editing}
-
-We automatically generated 600 string editing problems split 75/25 train/test.
-We show representative problems:
-\pagebreak
-
-double
-
-\begin{tabular}{ll}
-        \toprule Input&Output\\\midrule
-vlxS & vlxSvlxS \\
- JkVC & JkVCJkVC \\
- dbxD & dbxDdbxD \\
- gQgB & gQgBgQgB
-\\\bottomrule
-\end{tabular}
-
-Extract string delimited by ',' (inclusive),' '
-\begin{tabular}{ll}
-        \toprule Input&Output\\\midrule
-LsR,uhMZI wIL & ,uhMZI  \\
- tMY,KdF hhT & ,KdF  \\
- UUPdC,UAd kCgtq & ,UAd  \\
- gQgGN,VMJ FjrXZ & ,VMJ 
-\\\bottomrule
-\end{tabular}
-
-Apply first 2 characters to input delimited by '.'
-\begin{tabular}{ll}
-        \toprule Input&Output\\\midrule
-OhCQN.DlH.LdkJ.pLh & OhDlLdpL \\
- CZr.mveZr.dvjP.YMJpg & CZmvdvYM \\
- bpub.VkkZ.dYdPr & bpVkdY \\
- zjNA.wFwY.WMMqh.OFdbr & zjwFWMOF
-\\\bottomrule
-\end{tabular}
-
-Apply first 2 characters to string delimited by ' ','.'
-\begin{tabular}{ll}
-        \toprule Input&Output\\\midrule
-fNA YWyL.JDD & YW \\
- YWozt TPW.nczz & TP \\
- RJwiT dpW.uFU & dp \\
- FTp gNMo.JsZB & gN
-\\\bottomrule
-\end{tabular}
-
-%% Apply double to input delimited by '.'
-%% \begin{tabular}{ll}
-%%         \toprule Input&Output\\\midrule
-%% FWhb.jYMdi & FWhbFWhbjYMdijYMdi \\
-%%  JZB.WEwgH & JZBJZBWEwgHWEwgH \\
-%%  SgA.Kxo.RyF & SgASgAKxoKxoRyFRyF \\
-%%  NMApj.OaPBs.uZm.iCZy & NMApjNMApjOaPBsOaPBsuZmuZmiCZyiCZy
-%% \\\bottomrule
-%% \end{tabular}
-
-Apply strip to string delimited by ',','.'
-\begin{tabular}{ll}
-        \toprule Input&Output\\\midrule
-iNEP,  GigW .Xlm & GigW \\
- xhpH,  QvwnW .mfK & QvwnW \\
- weWIC,IQjhWa .fGKz & IQjhWa \\
- xZUm, FpdE.ETcr & FpdE
-\\\bottomrule
-\end{tabular}
-
-drop first character
-
-\begin{tabular}{ll}
-        \toprule Input&Output\\\midrule
-WYPvm & YPvm \\
- terHk & erHk \\
- asWM & sWM \\
- tBdT & BdT
-\\\bottomrule
-\end{tabular}
 
@@
 
+\section{Example program induction tasks}
+
+\subsection{String editing}
+
+We automatically generated 600 string editing problems split 75/25 train/test.
+The following are representative problems:
+\pagebreak
+
+\begin{center}
+\begin{tabular}{|l|l|}
+  \hline Input&Output\\
+  \hline\multicolumn{2}{|c|}{Double}\\
+  \hline
+vlxS & vlxSvlxS \\
+ JkVC & JkVCJkVC \\
+ dbxD & dbxDdbxD \\
+ gQgB & gQgBgQgB \\
+  \hline\multicolumn{2}{|c|}{Extract string delimited by ',' (inclusive),' '}\\
+  \hline
+LsR,uhMZI wIL & ,uhMZI  \\
+ tMY,KdF hhT & ,KdF  \\
+ UUPdC,UAd kCgtq & ,UAd  \\
+ gQgGN,VMJ FjrXZ & ,VMJ \\
+  \hline\multicolumn{2}{|c|}{Apply first 2 characters to input delimited by '.'}\\
+  \hline
+OhCQN.DlH.LdkJ.pLh & OhDlLdpL \\
+ CZr.mveZr.dvjP.YMJpg & CZmvdvYM \\
+ bpub.VkkZ.dYdPr & bpVkdY \\
+ zjNA.wFwY.WMMqh.OFdbr & zjwFWMOF \\
+  \hline\multicolumn{2}{|c|}{Apply first 2 characters to string delimited by ' ','.'}\\
+  \hline
+fNA YWyL.JDD & YW \\
+ YWozt TPW.nczz & TP \\
+ RJwiT dpW.uFU & dp \\
+ FTp gNMo.JsZB & gN \\
+%  \hline\multicolumn{2}{|c|}{Apply double to input delimited by '.'}\\
+%  \hline
+%FWhb.jYMdi & FWhbFWhbjYMdijYMdi \\
+% JZB.WEwgH & JZBJZBWEwgHWEwgH \\
+% SgA.Kxo.RyF & SgASgAKxoKxoRyFRyF \\
+% NMApj.OaPBs.uZm.iCZy & NMApjNMApjOaPBsOaPBsuZmuZmiCZyiCZy \\
+  \hline\multicolumn{2}{|c|}{Apply strip to string delimited by ',','.'}\\
+  \hline
+iNEP,  GigW .Xlm & GigW \\
+ xhpH,  QvwnW .mfK & QvwnW \\
+ weWIC,IQjhWa .fGKz & IQjhWa \\
+ xZUm, FpdE.ETcr & FpdE \\
+  \hline\multicolumn{2}{|c|}{Drop first character}\\
+  \hline
+WYPvm & YPvm \\
+ terHk & erHk \\
+ asWM & sWM \\
+ tBdT & BdT \\
+\hline
+\end{tabular}
+\end{center}
+
 \bibliography{main}

### 2018-02-09 supplement.tex
tiny structure change for text examples
edits re. ELBO

@@
 our generative model,
-then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along an unimportant regularizing term on the DSL):
+then, during DSL induction, we would seek a new $(\mathcal{D}^*,\theta^*)$ maximizing the following lower bound on the likelihood (along with an unimportant regularizing term on the DSL):
 \begin{align}
@@
 
-How the recognition model is worked into this is unimportant to the reason why we do not maximize a bound of this form:
-this bound takes an expectation over the \emph{previous} sleep cycle's
+We do not maximize a bound of this form because it
+takes an expectation over the \emph{previous} sleep cycle's
 posterior over programs,
-and so the approximate posterior $Q_x$ at the next iteration
+so the approximate posterior $Q_x$ at the next iteration
 ends up being very close to previous approximate posterior.
@@
 The following are representative problems:
-\pagebreak
+
+\vfill
 

### 2018-02-09 paper.tex
improved caption
edits re. hyperparameters

@@
 We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
-\begin{align*}
+\begin{align}
   \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
   \probability[\theta|\mathcal{D}]& = \text{Dir}(\theta|\alpha)
-\end{align*}
+\end{align}
 where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
@@
   \end{tabular}
-\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved tasks.
-  RF/DC: trained like RobustFill/DeepCoder.
-  PCFG: model w/o structure learning.
-  Enum: model w/o learning.
-  MDL: $-\expect\left[\probability[x|p] \right]$.
-For domains other than symbolic regression MDL is 0 nats.
-For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}
+\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
+  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
+  structure learning.  Enum: model w/o any learning.  MDL:
+  $-\expect\left[\probability[x|p] \right]$. For domains other than
+  symbolic regression MDL is 0 nats. For symbolic regression MDL is a
+  proxy for \# of continuous parameters.}\label{baselineComparisons} \end{table}
 

### 2018-02-09 supplement.tex
edits re. hyperparameters

@@
 \section{Hyperparameters \& Implementation details}
-We set $\alpha = 10$ and $\lambda = 1$ for all experiments except for string
-editing. For string editing we found that more strongly regularizing
+We set structure penalty $\lambda = 1$ (Eq.~5 of the main paper) and
+smoothness parameter $\alpha = 10$ (Eq.~6 of the main paper)
+for all experiments except for string editing.
+For string editing we found that more strongly regularizing
 the DSL was necessary,
-and so we set $\lambda = 5$.
+so we set $\lambda = 5$.
 We believe this is because

### 2018-02-09 paper.tex
changed a single word

@@
 \newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
+\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
 % Use the following line for the initial blind version submitted for review:
@@
 
-\subsection{Boolean circuits}
+\subsection{Boolean Circuits}
 We consider the problem of learning
@@
 
-\subsection{String editing}\label{textSection}
+\subsection{String Editing}\label{textSection}
 Synthesizing programs that manipulate strings is a classic problem in the
@@
     \toprule
-    \code{0}, \code{+1},\code{-1},     \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  \\
+    \code{0}, \code{+1},\code{-1},\codechar{},\codechar{,},\codechar{ },\codechar{<},\codechar{>}  \\
     \code{split}, \code{join}, \code{map}, 
@@
         \phantom{$f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ }}\code{(chr->str c))}\\
-      $f_2($\code{s$,$c}$) \,=\, $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
+      $f_2($\code{s$,$c}$) \,=\, $\code{((}$f_0$\code{ c) (}$f_1$\code{ (+1 0) c s))} \\
         \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
-      $f_3($\code{c$,$s}$) \,=\, $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
+      $f_3($\code{c$,$s}$) \,=\, $\code{(nth (+1 0) (split c (nth 0 s)))}\\
         \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
@@
 \\\midrule
-\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_3$\code{ (split '>' } }&
-\multicolumn{2}{l}{$f(\text{\code{s}}) = $\code{(}$f_1$\code{ s '.' 0)}}
+\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_3$\code{ (split }\codechar{>}\code{ } }&
+\multicolumn{2}{l}{$f(\text{\code{s}}) = $\code{(}$f_1$\code{ s }\codechar{.}\code{ 0)}}
 \\
-\multicolumn{2}{l|}{\hspace{1.1cm}\code{(upper s)) '<')}}
+\multicolumn{2}{l|}{\hspace{1.1cm}\code{(upper s)) }\codechar{<}\code{)}}
 \\\bottomrule
@@
 
-\subsection{List functions}
+\subsection{List Functions}
 Synthesizing programs that manipulate data structures is
@@
 improves on the baselines,
-often dramatically (string editing and symbolic regression).
+sometimes dramatically (string editing and symbolic regression).
 The recognition model 

### 2018-02-09 supplement.tex
capitalize titles

@@
 
-\section{An illustration of the wake-sleep cycle}
+\section{An Illustration of the Wake-Sleep Cycle}
 Below we diagram the wake/sleep cycles employed by our algorithm. At each stage of the cycle,
@@
 
-\section{Neural recognition model architecture}
+\section{Neural Recognition Model Architecture}
 
@@
 
-\subsection{Structure learning}
+\subsection{Structure Learning}
 
@@
 
-\section{Initial domain-specific languages}
-\subsection{Boolean circuits}
+\section{Initial Domain-Specific Languages}
+\subsection{Boolean Circuits}
 \begin{center}
@@
 
-\subsection{Symbolic regression}
+\subsection{Symbolic Regression}
 \begin{center}
@@
 
-\subsection{String editing}
+\subsection{String Editing}
 \begin{center}
@@
 
-\subsection{List functions}
+\subsection{List Functions}
 The lower section with seven primitives designates those given to the
@@
 
-\section{Hyperparameters \& Implementation details}
+\section{Hyperparameters \& Implementation Details}
 We set structure penalty $\lambda = 1$ (Eq.~5 of the main paper) and
@@
 
-\section{Why not the ELBO bound?}
+\section{Why not the ELBO Bound?}
 
@@
 
-\section{Example program induction tasks}
+\section{Example Program Induction Tasks}
 
-\subsection{String editing}
+\subsection{String Editing}
 
@@
 The following are representative problems:
-
-\vfill
 

### 2018-02-09 supplement.tex
consistent list type notation

@@
   \hline
-    empty & [$\alpha$] \\
-    singleton & $\alpha \to$ [$\alpha$] \\
-    concat & [$\alpha$] $\to$ [$\alpha$] $\to$ [$\alpha$] \\
+    empty & $\left[\alpha\right]$ \\
+    singleton & $\alpha \to \left[\alpha\right]$ \\
+    concat & $\left[\alpha\right] \to \left[\alpha\right] \to \left[\alpha\right]$ \\
     mapi & $(\mathbb{Z} \to \alpha \to \beta)$ \\
-    & \qquad\qquad$\to$ [$\alpha$] $\to$ [$\beta$] \\
+    & \qquad\qquad$\to \left[\alpha\right] \to \left[\beta\right]$ \\
     reducei & $(\beta \to \mathbb{Z} \to \alpha \to \beta)$\\
-    & \qquad\qquad$\to \beta \to$ [$\alpha$] $\to \beta$ \\
+    & \qquad\qquad$\to \beta \to \left[\alpha\right] \to \beta$ \\
     \hline
@@
     range & $\mathbb{Z} \to$ [$\mathbb{Z}$] \\
-    sort & [$\mathbb{Z}$] $\to$ [$\mathbb{Z}$] \\
+    sort & $\left[\mathbb{Z}\right] \to \left[\mathbb{Z}\right]$ \\
     \hline
-    sum & [$\mathbb{Z}$] $\to \mathbb{Z}$ \\
-    reverse & [$\alpha$] $\to$ [$\alpha$] \\
-    all & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
-    any & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ bool \\
-    index & $\mathbb{Z} \to$ [$\alpha$] $\to \alpha$ \\
-    filter & $(\alpha \to$ bool$) \to$ [$\alpha$] $\to$ [$\alpha$] \\
-    slice & $\mathbb{Z} \to \mathbb{Z} \to$ [$\alpha$] $\to$ [$\alpha$] \\
+    sum & $\left[\mathbb{Z}\right] \to \mathbb{Z}$ \\
+    reverse & $\left[\alpha\right] \to \left[\alpha\right]$ \\
+    all & $(\alpha \to$ bool$) \to \left[\alpha\right] \to$ bool \\
+    any & $(\alpha \to$ bool$) \to \left[\alpha\right] \to$ bool \\
+    index & $\mathbb{Z} \to \left[\alpha\right] \to \alpha$ \\
+    filter & $(\alpha \to$ bool$) \to \left[\alpha\right] \to \left[\alpha\right]$ \\
+    slice & $\mathbb{Z} \to \mathbb{Z} \to \left[\alpha\right] \to \left[\alpha\right]$ \\
   \hline

### 2018-02-09 supplement.tex
edits re. estimating theta

@@
 
-Differentiate with respect to $\theta_e$ and set to zero:
+Differentiating with respect to $\theta_e$ and setting to zero:
 %% \begin{align}
@@
 %%   \end{align}
-
+%%
 \begin{align*}
- \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}}
-\end{align*}
-%or more succinctly,
-\begin{align}
+  \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}} \\
   \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
@@
 %% \end{align}
-If this equality holds then $\theta_x \propto c(x)/a(x)$:
+If the above equality holds then $\theta_x \propto c(x)/a(x)$:
 \begin{align}
-  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}.
+  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}
 \end{align}
@@
 \begin{align}
-  J& = L + \log \text{Dir}(\theta|\alpha)\\
-  &\stackrel{+}{\geq } \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x  + \sum_x(\alpha_x - 1)\log \theta_x\\
+  \nonumber J& = L + \log \text{Dir}(\theta|\alpha)\\
+  \nonumber&\stackrel{+}{\geq } \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x \\
+  \nonumber&\qquad\quad+ \sum_x(\alpha_x - 1)\log \theta_x\\
   & = \sum_x (c(x) + \alpha_x - 1)\log \theta_x -  N \log \sum_x a(x)\theta_x
 \end{align}
-We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} ($c(x)$), but not to the \emph{possible counts} ($a(x)$),
-and so we estimate $\theta_e$ as the ratio of the sum of the number of times that $e$ was used, plus the pseudocounts $\alpha$, with the number of times that it could have been used, $a(e)$.
+We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} $c(x)$, but not to the \emph{possible counts} $a(x)$,
+and so we estimate $\theta_e$ as the ratio of the sum of the number of times that $e$ was used plus the pseudocounts $\alpha$, with the number of times that it could have been used, $a(e)$.
 

### 2018-02-09 supplement.tex
better placement of DSL induction algorithm
use p to refer to programs
made text chars more consistent

@@
 \setcounter{algorithm}{2} % there are two algorithms in the main paper.
-\begin{algorithm}[tb]
+\begin{algorithm}[hb]
    \caption{DSL Induction Algorithm}
@@
      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
+     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[p|\mathcal{D},\theta]$
      \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
@@
      \WHILE {true}
-     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
+     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\text{ subexpression of }p\}$
      \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
@@
   length&$\text{string}\to \mathbb{Z}$\\
+  ""&string\\
+  ","\quad" "&char\\
+  "<"\quad">" &char \\
   chr->str&$\text{char}\to \text{string}$\\
@@
 \end{center}
-We also provide the constant empty string which has type string and also provide
-constant characters which have the type char.
 

### 2018-02-09 paper.tex
got it down to eight pages using negative vertical space
made text chars more consistent
Josh conclusion

--- 2018-02-09 paper.tex
@@
 Viewed as a probabilistic inference problem,
-our goal is to infer the DSL (written $\mathcal{D}$)
-along with a program (written $p$) for each set of inputs/outputs (\emph{tasks}, written $x$).
+our goal is to jointly infer a single DSL (written $\mathcal{D}$)
+and many programs (written $p$), one for each set of inputs/outputs (\emph{tasks}, written $x$).
 A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
@@
 and optimizing using gradient descent.
-We have the algorithm 
-solve a set of 200 symbolic regression problems, each a polynomial of
+We task our model
+with solving 200 symbolic regression problems, each a polynomial of
 degree 1 to 4, where our observations $x$ are 10
@@
 
-What DSL does \system learn?
-The learned DSL contains templates for polynomials of different orders,
-which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Tbl.~\ref{regressionDSL}).
-Examining the programs themselves,
-one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom.
+\system learns a DSL containing templates for polynomials of different orders,
+which lets the algorithm quickly find the functions that are most appropriate to this domain (Tbl.~\ref{regressionDSL}).
+We also found that the algorithm discovers programs that minimize the number of continuous degrees of freedom.
 For example, it represents the linear function $-x+2$ with the program
 \code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
-which has five continuous degrees of freedom.
-This arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
+which has five continuous parameters.
+This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
 \begin{table}\centering
@@
   symbolic regression MDL is 0 nats. For symbolic regression MDL is a
-  proxy for \# of continuous parameters.}\label{baselineComparisons} \end{table}
+  proxy for \# of continuous parameters.}\vspace{-0.5cm}\label{baselineComparisons} \end{table}
 
@@
  and extend it along two dimensions: (1) we propose a new algorithm for
- inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
+ inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
@@
  \section{Contribution and Outlook}
- We contributed a model, \systemEnding, that learns to program from examples by
+ We contributed a model, \systemEnding, that learns to program by
  bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
@@
  neural recognition model that learns how to efficiently deploy the
- DSL and hone in on the right programs.  We believe that this
- integration of top-down symbolic representation and bottom-up neural
- networks -- both of them learned -- is a useful general approach for
- getting program induction off the ground.
+ DSL and hone in on the right programs.
+ %% We believe that this
+ %% integration of top-down symbolic representation and bottom-up neural
+ %% networks -- both of them learned -- is a useful general approach for
+ %% getting program induction off the ground. 
+Many directions remain open for future work. Two immediate goals are to
+integrate more sophisticated neural recognition models~\cite{devlin2017robustfill} and
+program synthesizers~\cite{solar2008program}, which may improve performance in some
+domains over the simple, generic methods used here, and to explore DSL
+meta-learning: Can we find a \emph{single} universal primitive set that could
+effectively bootstrap DSLs for new domains, including the four domains
+considered but many others?
 

### 2018-02-09 paper.tex
small cleanups to notation and conclusion

@@
 We frame this problem as maximum a posteriori (MAP) inference in the
-generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
+generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]\probability(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
   \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
@@
 \begin{align}
- J\geq \lowerBound\triangleq\probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
+ J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
 \end{align}
@@
 domains over the simple, generic methods used here, and to explore DSL
-meta-learning: Can we find a \emph{single} universal primitive set that could
+meta-learning: Can we find a single universal primitive set that could
 effectively bootstrap DSLs for new domains, including the four domains

### 2018-02-09 paper.tex
re- factored conclusion

@@
 
- \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
+ \noindent \emph{Deep Learning:} Much recent work in the ML community has
  focused on creating neural networks that regress from
- input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is closest to the lesioned version of
- \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned (RF/DC baseline).
+ input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. %This family of work is closest to our RF/DC baseline.
  These neural networks are typically trained with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
- In our work we consider a weakly-supervised setting where ground truth programs are not provided and
+ Our work  considers a weakly-supervised regime where ground truth programs are not provided and
  the agent must learn from a few hundred tasks.
@@
  \section{Contribution and Outlook}
- We contributed a model, \systemEnding, that learns to program by
+ We contribute a model, \systemEnding, that learns to program by
  bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
@@
  DSL and hone in on the right programs.
- %% We believe that this
- %% integration of top-down symbolic representation and bottom-up neural
- %% networks -- both of them learned -- is a useful general approach for
- %% getting program induction off the ground. 
+ We believe that this
+ integration of top-down symbolic representation and bottom-up neural
+ networks -- both of them learned -- is a useful general approach for
+ program induction.
 Many directions remain open for future work. Two immediate goals are to
@@
 program synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the simple, generic methods used here, and to explore DSL
+domains over the generic methods used here, and to explore DSL
 meta-learning: Can we find a single universal primitive set that could
 effectively bootstrap DSLs for new domains, including the four domains
-considered but many others?
+considered but also many others?
 

### 2018-02-09 paper.tex
short and related work and baselines

@@
   trains a neural network to efficiently search for programs in the
-  learned DSL.  We use our model to build circuits, edit strings, do
-  symbolic regression, and synthesize functions on lists, in each case
+  learned DSL.  We use our model to build circuits, edit strings, solve
+  symbolic regression problems, and synthesize functions on lists, in each case
   showing the model learns a domain-specific vocabulary for expressing
@@
 program to be induced, and thus success often hinges on having a good
-DSL which
-provides a crucial inductive bias for what would otherwise be an
+DSL, by providing a crucial inductive bias for what would otherwise be an
 unconstrained search through the space of all computable functions.
@@
 We compare  with four baselines on  held-out tasks:
-\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model, $q$.
-\\\noindent \textbf{RF/DC}, which holds $(\mathcal{D},\theta)$ fixed and learns
+\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
+\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
 a recognition model only from samples from the fixed generative model.
-This is equivalent to our algorithm with $\lambda = \infty$ (see Sec.~\ref{grammarInductionSection})
-and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Sec.~\ref{recognitionSection}).
+This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
+and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
 We call this baseline RF/DC because
@@
  Bayesian Program
- Learning (BPL; see~\citep{lake2013one,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
+ Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
 Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),

### 2018-02-09 paper.tex
fix capitalization

@@
 
- \noindent \emph{Deep Learning:} Much recent work in the ML community has
+ \noindent \emph{Deep learning:} Much recent work in the ML community has
  focused on creating neural networks that regress from

### 2018-02-09 paper.tex
change experiment ordering in abstract to match exposition

@@
   trains a neural network to efficiently search for programs in the
-  learned DSL.  We use our model to build circuits, edit strings, solve
-  symbolic regression problems, and synthesize functions on lists, in each case
+  learned DSL.  We use our model to build circuits, solve
+  symbolic regression problems, edit strings, and synthesize functions on lists, in each case
   showing the model learns a domain-specific vocabulary for expressing

### 2018-02-09 paper.tex
replace weird comma

@@
 program to be induced, and thus success often hinges on having a good
-DSL, by providing a crucial inductive bias for what would otherwise be an
+DSL that provides a crucial inductive bias for what would otherwise be an
 unconstrained search through the space of all computable functions.

### 2018-02-09 supplement.tex
putting DSL 's in supplement

@@
  FTp gNMo.JsZB & gN \\
-%  \hline\multicolumn{2}{|c|}{Apply double to input delimited by '.'}\\
-%  \hline
-%FWhb.jYMdi & FWhbFWhbjYMdijYMdi \\
-% JZB.WEwgH & JZBJZBWEwgHWEwgH \\
-% SgA.Kxo.RyF & SgASgAKxoKxoRyFRyF \\
-% NMApj.OaPBs.uZm.iCZy & NMApjNMApjOaPBsOaPBsuZmuZmiCZyiCZy \\
   \hline\multicolumn{2}{|c|}{Apply strip to string delimited by ',','.'}\\
@@
 \end{center}
+
+\section{Learned DSLs}
+
+Here we present representative DSLs learned by
+our model. DSL primitives discovered by the algorithm are prefixed with \code{#}.
+Variables are prefixed with \$, and we use De Bruijn indices to model variables~\cite{pierce}.
+We abbreviate lambda as \textbackslash.
+
+\pagebreak
+\onecolumn
+\subsection{Boolean circuits}
+
+\begin{verbatim}
+#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2))))))
+#(\ (\ (nand (nand $0 $1) (nand $0 $1))))
+#(\ (\ (nand (nand $0 $0) (nand $1 $1))))
+#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 $3))))))
+#(\ (\ (\ (nand (nand $0 $0) (nand $1 (nand $2 $1))))))
+#(\ (\ (#(\ (\ (nand (nand $0 $1) (nand $0 $1)))) $0 (nand $1 $0))))
+#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0))))))
+#(\ (\ (#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0)))))) $0 $1 (nand $1 $1))))
+#(\ (\ (nand $0 (nand $1 $0))))
+#(\ (#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2)))))) (nand $0 $0)))
+#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $3 $2)))))))
+\end{verbatim}
+
+\subsection{Symbolic regression}
+\begin{verbatim}
+#(\ (* (+ REAL $0) REAL))
+#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0)))
+#(\ (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))
+#(\ (#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0))) (+ $0 $0)))
+#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0)))
+#(\ (* (#(\ (+ $0 REAL)) $0) (#(\ (+ (* $0 (* $0 REAL)) REAL)) $0)))
+#(\ (+ (* $0 (* $0 REAL)) REAL))
+#(* REAL)
+#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0))
+#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0))))
+#(\ (+ $0 REAL))
+\end{verbatim}
+
+\subsection{String editing}
+\begin{verbatim}
+#(\ (\ (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0)))))
+#(\ (\ (join (chr->str $0) (map (\ (strip $0)) $1))))
+#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0))))
+#(\ (#(join empty) (split '.' $0)))
+#(#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2))))) (decr 0))
+#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (slice $2 (incr $2) $0)))))
+#(#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (incr 0))
+#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2)))))
+#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1))))))
+#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr $0)))
+#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (slice $0 (incr (incr $0)))) $2 $0)))))
+#(\ (\ (slice $0 (incr $0) (nth (incr $0) $1))))
+#(\ (\ (#(\ (\ (slice $0 (len $1) $1))) (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1) (incr 0))))
+#(\ (\ (caseUpper (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0))))))
+#(\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (nth 0 $1)) $0)))
+#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr (decr $0))))
+#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1))))
+#(\ (\ (#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0)))) (split '.' (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1)) '.')))
+#(\ (\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1))))) $1 $2 0)) $0))))
+#(\ (\ (slice $0 (incr (incr $0)) (#(nth (incr 0)) $1))))
+#(join empty)
+#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1)))))
+#(\ (\ (nth 0 (split $0 (#(nth (incr 0)) $1)))))
+#(\ (slice $0 (incr (incr $0))))
+#(nth (incr 0))
+#(\ (join empty (map (\ (caseCapitalize $0)) $0)))
+#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2)))))
+#(\ (\ (slice $0 (len $1) $1)))
+#(\ (\ (\ (join (chr->str $0) (map $1 $2)))))
+#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1))))
+\end{verbatim}
+
+\subsection{List functions}
 

### 2018-02-09 paper.tex
Josh conclusion

@@
  \section{Contribution and Outlook}
- We contribute a model, \systemEnding, that learns to program by
- bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
-  Alongside this DSL, we train a bottom-up
- neural recognition model that learns how to efficiently deploy the
- DSL and hone in on the right programs.
- We believe that this
- integration of top-down symbolic representation and bottom-up neural
- networks -- both of them learned -- is a useful general approach for
- program induction.
-Many directions remain open for future work. Two immediate goals are to
-integrate more sophisticated neural recognition models~\cite{devlin2017robustfill} and
-program synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the generic methods used here, and to explore DSL
-meta-learning: Can we find a single universal primitive set that could
-effectively bootstrap DSLs for new domains, including the four domains
-considered but also many others?
-
+\system learns to program by bootstrapping a DSL with new
+domain-specific primitives that the algorithm itself discovers,
+together with a neural recognition model that learns how to
+efficiently deploy the DSL on new tasks. We believe this integration
+of top-down symbolic representations and bottom-up neural networks --
+both of them learned -- could help make program induction systems more
+generally useful for AI. Many directions remain open. Two immediate
+goals 
+are to integrate more sophisticated neural recognition
+models~\cite{devlin2017robustfill} and program
+synthesizers~\cite{solar2008program}, which may improve performance in
+some domains over the generic methods used here, and to
+explore DSL meta-learning: Can we find a \emph{single} universal
+primitive set that could effectively bootstrap DSLs for new domains,
+including the four domains considered but many others?
 \bibliography{main}

### 2018-02-09 paper.tex
conclusion edits

@@
  \section{Contribution and Outlook}
-\system learns to program by bootstrapping a DSL with new
+We contribute a model, \systemEnding, that learns to program by bootstrapping a DSL with new
 domain-specific primitives that the algorithm itself discovers,

### 2018-02-09 supplement.tex
induced list DSL

@@
 \subsection{List functions}
+\begin{verbatim}
+#(\ (\ (index $0 (sort $1))))
+#(\ (\ (slice $0 (+ $1 4))))
+#(mapi (\ (\ (eq? $0 $2))))
+#(\ (mapi (\ (\ (gt? $0 $2)))))
+#(\ (any (\ (eq? $0 $1))))
+#(\ (mapi (\ (\ (+ $0 $2)))))
+#(\ (\ (mapi (\ (\ (index $2 $3))) $1)))
+#(\ (mapi (\ (\ (mod $0 $2)))))
+#(\ (mapi (\ (\ (* $0 $2)))))
+#(\ (\ (singleton (index $0 $1))))
+#(\ (\ (slice $0 (sum $1) $1)))
+#(slice 0)
+#(\ (\ (index (negate $0) (sort $1))))
+#(\ (++ (singleton $0)))
+#(\ (\ (++ $0 (singleton $1))))
+#(\ (\ (++ $0 (singleton (index $1 $0)))))
+#(\ (\ (++ (singleton (index $0 $1)) $1)))
+#(\ (\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) $1 $2)))))
+#(\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))
+#(#(\ (mapi (\ (\ (gt? $0 $2))))) 3 empty)
+#(all (\ (is-square $0)))
+#(\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) 1 $1))))
+#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1))))
+#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2)))
+#(filter (\ (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0))))
+#(\ (\ (\ (++ (#(slice 0) $0 $1) (#(\ (\ (slice $0 (sum $1) $1))) $1 $2)))))
+#(\ (filter (\ (#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1)))) $1 (singleton $0)))))
+#(filter (\ (not (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0)))))
+#(\ (++ $0 (++ $0 $0)))
+#(\ (++ $0 (#(\ (++ $0 (++ $0 $0))) $0)))
+#(index 0)
+
+\end{verbatim}
 

### 2018-02-09 paper.tex
small changes

@@
 a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
+symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
 synthesize programs for task domains such as string transformations,
@@
 \\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.
-the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$.
+the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
@@
 and then checking if an enumerated program $p$ assigns positive
-probability to a task ($\probability[x|p] > 0$); if so, we include $p$ in
-the frontier $\mathcal{F}_x$.
+probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
+into the frontier $\mathcal{F}_x$.
 

### 2018-02-09 supplement.tex
clean up style of induced DSLs

--- 2018-02-09 supplement.tex
@@
 \end{center}
+
+\section{Learned DSLs}
+
+Here we present representative DSLs learned by
+our model. DSL primitives discovered by the algorithm are prefixed with \code{#}.
+Variables are prefixed with \$, and we use De Bruijn indices to model variables~\cite{pierce}.
+We abbreviate lambda as \textbackslash.
+
+\pagebreak
+\onecolumn
+\subsection{Boolean circuits}
+
+\begin{lstlisting}[breaklines=true]
+#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2))))))
+#(\ (\ (nand (nand $0 $1) (nand $0 $1))))
+#(\ (\ (nand (nand $0 $0) (nand $1 $1))))
+#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 $3))))))
+#(\ (\ (\ (nand (nand $0 $0) (nand $1 (nand $2 $1))))))
+#(\ (\ (#(\ (\ (nand (nand $0 $1) (nand $0 $1)))) $0 (nand $1 $0))))
+#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0))))))
+#(\ (\ (#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0)))))) $0 $1 (nand $1 $1))))
+#(\ (\ (nand $0 (nand $1 $0))))
+#(\ (#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2)))))) (nand $0 $0)))
+#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $3 $2)))))))
+\end{lstlisting}
+
+\subsection{Symbolic regression}
+\begin{lstlisting}[breaklines=true]
+#(\ (* (+ REAL $0) REAL))
+#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0)))
+#(\ (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))
+#(\ (#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0))) (+ $0 $0)))
+#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0)))
+#(\ (* (#(\ (+ $0 REAL)) $0) (#(\ (+ (* $0 (* $0 REAL)) REAL)) $0)))
+#(\ (+ (* $0 (* $0 REAL)) REAL))
+#(* REAL)
+#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0))
+#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0))))
+#(\ (+ $0 REAL))
+\end{lstlisting}
+
+\subsection{String editing}
+\begin{lstlisting}[breaklines=true]
+#(\ (\ (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0)))))
+#(\ (\ (join (chr->str $0) (map (\ (strip $0)) $1))))
+#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0))))
+#(\ (#(join empty) (split '.' $0)))
+#(#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2))))) (decr 0))
+#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (slice $2 (incr $2) $0)))))
+#(#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (incr 0))
+#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2)))))
+#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1))))))
+#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr $0)))
+#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (slice $0 (incr (incr $0)))) $2 $0)))))
+#(\ (\ (slice $0 (incr $0) (nth (incr $0) $1))))
+#(\ (\ (#(\ (\ (slice $0 (len $1) $1))) (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1) (incr 0))))
+#(\ (\ (caseUpper (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0))))))
+#(\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (nth 0 $1)) $0)))
+#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr (decr $0))))
+#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1))))
+#(\ (\ (#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0)))) (split '.' (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1)) '.')))
+#(\ (\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1))))) $1 $2 0)) $0))))
+#(\ (\ (slice $0 (incr (incr $0)) (#(nth (incr 0)) $1))))
+#(join empty)
+#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1)))))
+#(\ (\ (nth 0 (split $0 (#(nth (incr 0)) $1)))))
+#(\ (slice $0 (incr (incr $0))))
+#(nth (incr 0))
+#(\ (join empty (map (\ (caseCapitalize $0)) $0)))
+#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2)))))
+#(\ (\ (slice $0 (len $1) $1)))
+#(\ (\ (\ (join (chr->str $0) (map $1 $2)))))
+#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1))))
+\end{lstlisting}
+
+\subsection{List functions}
+\begin{lstlisting}[breaklines=true]
+#(\ (\ (index $0 (sort $1))))
+#(\ (\ (slice $0 (+ $1 4))))
+#(mapi (\ (\ (eq? $0 $2))))
+#(\ (mapi (\ (\ (gt? $0 $2)))))
+#(\ (any (\ (eq? $0 $1))))
+#(\ (mapi (\ (\ (+ $0 $2)))))
+#(\ (\ (mapi (\ (\ (index $2 $3))) $1)))
+#(\ (mapi (\ (\ (mod $0 $2)))))
+#(\ (mapi (\ (\ (* $0 $2)))))
+#(\ (\ (singleton (index $0 $1))))
+#(\ (\ (slice $0 (sum $1) $1)))
+#(slice 0)
+#(\ (\ (index (negate $0) (sort $1))))
+#(\ (++ (singleton $0)))
+#(\ (\ (++ $0 (singleton $1))))
+#(\ (\ (++ $0 (singleton (index $1 $0)))))
+#(\ (\ (++ (singleton (index $0 $1)) $1)))
+#(\ (\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) $1 $2)))))
+#(\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))
+#(#(\ (mapi (\ (\ (gt? $0 $2))))) 3 empty)
+#(all (\ (is-square $0)))
+#(\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) 1 $1))))
+#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1))))
+#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2)))
+#(filter (\ (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0))))
+#(\ (\ (\ (++ (#(slice 0) $0 $1) (#(\ (\ (slice $0 (sum $1) $1))) $1 $2)))))
+#(\ (filter (\ (#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1)))) $1 (singleton $0)))))
+#(filter (\ (not (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0)))))
+#(\ (++ $0 (++ $0 $0)))
+#(\ (++ $0 (#(\ (++ $0 (++ $0 $0))) $0)))
+#(index 0)
+\end{lstlisting}
 

### 2018-02-09 paper.tex
fixed weird citation bug

@@
 very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+for which we have effective approaches like the Inside/Outside algorithm~(Lafferty,~2000), %\cite{international2000derivation},
 a DSL in \system may be context-sensitive due to the presence of variables

### 2018-02-09 supplement.tex
fix latex errors
change listings style for consistency
fixed weird citation bug

--- 2018-02-09 supplement.tex
@@
 \usepackage{listings}
+\lstset{%
+  basicstyle=\footnotesize\ttfamily,
+  breaklines=true}
 \usepackage{wrapfig}
@@
   \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
-\end{align}
+\end{align*}
 where $N=c(p)$ and $a(x) = r(x, p)$. From here forward, we similarly let
@@
 
-\section{Example Program Induction Tasks}
-
-\subsection{String Editing}
+\section{Example String Editing Tasks}
 
@@
 
+\vfill
+
 \section{Learned DSLs}
@@
 Here we present representative DSLs learned by
-our model. DSL primitives discovered by the algorithm are prefixed with \code{#}.
-Variables are prefixed with \$, and we use De Bruijn indices to model variables~\cite{pierce}.
-We abbreviate lambda as \textbackslash.
+our model. DSL primitives discovered by the algorithm are prefixed with
+\lstinline!#!.
+Variables are prefixed with \lstinline!$!, and we use De Bruijn indices to model variables~\cite{pierce}.
+We abbreviate lambda as \lstinline!\!.
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 #(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2))))))
@@
 #(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $3 $2)))))))
-\end{verbatim}
+\end{lstlisting}
 
 \subsection{Symbolic regression}
-\begin{verbatim}
+\begin{lstlisting}
 #(\ (* (+ REAL $0) REAL))
@@
 #(\ (+ $0 REAL))
-\end{verbatim}
+\end{lstlisting}
 
 \subsection{String editing}
-\begin{verbatim}
+\begin{lstlisting}
 #(\ (\ (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0)))))
@@
 #(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1))))
-\end{verbatim}
+\end{lstlisting}
 
 \subsection{List functions}
-\begin{verbatim}
+\begin{lstlisting}
 #(\ (\ (index $0 (sort $1))))
@@
 #(index 0)
-
-\end{verbatim}
+\end{lstlisting}
 

### 2018-02-09 supplement.tex
better spacing around table

--- 2018-02-09 supplement.tex
@@
 \usepackage{listings}
+\lstset{%
+  basicstyle=\ttfamily,
+  breaklines=true}
 \usepackage{wrapfig}

### 2018-02-09 paper.tex
fixed a typo in an equation
changed a section heading to be more consistent
Josh introduction

@@
 scheme~\cite{le2016inference}.
+ The neural recognition model and the
+generative model embodied in the DSL jointly train each other, as they
+iteratively learn to solve more programming tasks.
 
@@
 
-To perform inference, we take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
-which alternates between ``waking'', which draws samples from a recognition model (our $q$);
-and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
-For us, a \textbf{Wake} cycle means using the recognition model to search for programs that solve the tasks.
-A sleep cycle means  updating these models so that the DSL and  recognition model
-more closely match the distribution of tasks.
-An important difference from the Helmholtz machine
-is that our setup has two distinct sleep phases: a phase where we update the generative model
-$(\mathcal{D},\theta)$, which we call \textbf{Sleep-G},
-and a second phase where we update the recognition model $q$,
- called \textbf{Sleep-R}.
+%% To perform inference, we take inspiration from the Helmholtz machine
+%% and the Wake/Sleep algorithm~\cite{hinton1995wake}, which alternates
+%% between ``waking'', which draws samples from a recognition model (our
+%% $q$); and ``sleeping'', which trains a recognition model on samples
+%% from the generative model (our $\mathcal{D},\theta$).  For us, a
+%% \textbf{Wake} cycle means using the recognition model to search for
+%% programs that solve the tasks.  A sleep cycle means updating these
+%% models so that the DSL and recognition model more closely match the
+%% distribution of tasks.  An important difference from the Helmholtz
+%% machine is that our setup has two distinct sleep phases: a phase where
+%% we update the generative model $(\mathcal{D},\theta)$, which we call
+%% \textbf{Sleep-G}, and a second phase where we update the recognition
+%% model $q$, called \textbf{Sleep-R}.
 
@@
 
-Our goal  is to induce a DSL while finding programs solving each of the tasks.
-Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
-to accelerate the search over programs -- and (3) improving the DSL.
-The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
-\\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-When these improve, we find more programs solving the tasks.
-\\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+ Our goal is to induce a DSL while finding programs solving each of the
+tasks.  We take inspiration from the Wake/Sleep
+algorithm~\cite{hinton1995wake}, as well as the
+Exploration-Compression algorithm for bootstrap
+learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Wake/Sleep
+alternates between training a generative model on samples from a
+recognition model (our $q$), and training a recognition model on
+samples from the generative model (our $\mathcal{D},\theta$).  EC
+alternates between exploring the space of solutions to a set of tasks,
+and compressing those solutions to suggest new search primitives
+for the next exploration stage.
+We combine these ideas into an inference strategy that iterates
+through three steps: a \textbf{Wake} cycle uses the current DSL and
+recognition model to search for programs that solve the tasks.  The
+\textbf{Sleep-G} and \textbf{Sleep-R} cycles update the DSL and the
+recognition model, respectively.  Crucially, these steps bootstrap 
+each off other (Fig.~\ref{feeding}):
+
+\\\noindent \textbf{Wake: Searching for programs.}  Our program search
+is informed by both the DSL and the recognition model. When these
+improve, we find more programs solving the tasks.
+\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce the DSL
+from the programs found in the wake phase, aiming
+to maximally compress (or, raise the prior probability of) these
+programs.  As we solve more tasks, we hone in on richer DSLs that
+more closely match the domain.
+\\\noindent \textbf{Sleep-R: Learning a neural recognition model.}  We
+update the recognition model training on two data sources: samples
+from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
+programs found by the search procedure during waking. As the DSL
+improves and as search finds more programs, the recognition model gets
+both more data to train on, and better data.
+
+%% Our goal  is to induce a DSL while finding programs solving each of the tasks.
+%% Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
+%% to accelerate the search over programs -- and (3) improving the DSL.
+%% The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
+%% \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
+%% When these improve, we find more programs solving the tasks.
+%% \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
 As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce a DSL from the programs we found which solve the tasks.
-  As we solve more tasks, we can hone in on richer DSLs that more closely match the domain.
 
@@
 %% Fig.~\ref{graphicalModel}.
-
-Our lower bound $\lowerBound$ is unconventional,
-and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
-We explain in the Supplement that maximizing an ELBO-style bound
-leads to a pathological behavior that  causes the model to easily become trapped in local optima.
 
@@
 \mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
+  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
 \end{align}
@@
 
-\subsection{Sleep-G: Learning a DSL (Generative Model)}\label{grammarInductionSection}
+\subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
 

### 2018-02-09 supplement.tex
remove some redundant tex
change listings font size
Josh introduction

--- 2018-02-09 supplement.tex
@@
 
+Our lower bound $\lowerBound$ is unconventional,
+and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
+We explain in the Supplement that maximizing an ELBO-style bound
+leads to a pathological behavior that  causes the model to easily become trapped in local optima.
+
 If we were to maximize the ELBO bound to perform inference in

### 2018-02-11 paper.tex
added neuroscience citations

@@
   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
-The recognition model $q$ is a neural
+The recognition model $q(\cdot )$ is a neural
 network that regresses from input/output pairs to a distribution over
@@
 algorithm~\cite{hinton1995wake}, as well as the
-Exploration-Compression algorithm for bootstrap
+Exploration-Compression algorithm (EC) for bootstrap
 learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Wake/Sleep
@@
 recognition model, respectively.  Crucially, these steps bootstrap 
-each off other (Fig.~\ref{feeding}):
+off each other (Fig.~\ref{feeding}):
 
@@
 programs.  As we solve more tasks, we hone in on richer DSLs that
-more closely match the domain.
+more closely match the domain.\footnote{This is loosely biologically
+    inspired by the formation of abstractions during sleep memory
+    consolidation~\cite{DUDAI201520}}
 \\\noindent \textbf{Sleep-R: Learning a neural recognition model.}  We
-update the recognition model training on two data sources: samples
+update the recognition model by training on two data sources: samples
 from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
@@
 improves and as search finds more programs, the recognition model gets
-both more data to train on, and better data.
+both more data to train on, and better data.\footnote{These two sources are also
+    loosely biologically inspired by the distinct episodic replay and
+    hallucination components of dream sleep~\cite{fosse2003dreaming}}
 
@@
 Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Sec.~\ref{recognitionSection} explains how we train a neural network to accelerate program search; and
+Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
 Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
@@
 taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
-This supports our view of the recognition model as a way of accelerating the search over programs.
+This supports the view of the recognition model as a way of accelerating the search over programs.
 
@@
 synthesizers~\cite{solar2008program}, which may improve performance in
-some domains over the generic methods used here, and to
+some domains over the generic methods used here.
+Another direction is to
 explore DSL meta-learning: Can we find a \emph{single} universal
 primitive set that could effectively bootstrap DSLs for new domains,
-including the four domains considered but many others?
+including the four domains considered,  but also many others?
 \bibliography{main}

### 2018-02-14 supplement.tex
correct derivation of estimator of DSL parameters

@@
 
-%% We use an EM algorithm to estimate the continuous parameters of the DSL, e.g. $\theta$.
-%% Suppressing dependencies on $\mathcal{D}$, the EM updates are
-%% \begin{align}
-%% \label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
-%%   Q_x(p)&\propto \probability[x|p]\probability[p|\theta]
-%%   \end{align}
-%% In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$.
+We use an EM algorithm to estimate the continuous parameters of the DSL, e.g. $\theta$.
+Suppressing dependencies on $\mathcal{D}$, the EM updates are
+\begin{align}
+\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
+  Q_x(p)&\propto \probability[x|p]\probability[p|\theta]
+  \end{align}
+In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$,
+making our approach an instance of Generalized EM.
+
 We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$; $c(p)= \sum_{e\in \mathcal{D}}c(e,p)$ to mean the total number of primitives used in program $p$; $R(p)$ to mean the sequence of types input to sample in Alg.~1 of the main paper.
-Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
+Jensen's inequality gives a lower bound on the likelihood:
 \begin{align*}
-  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
-  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
-  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
+  &\sum_x\expect_{Q_x}\left[  \log \probability[p|\theta] \right] =\\
+  &\sum_{e\in \mathcal{D}} \log \theta_e \sum_x\expect\left[c(e,p_x) \right] -
+  \sum_\tau\expect\left[\sum_x c(\tau,p_x)  \right]\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+ =   &\sum_e C(e)\log \theta_e  - \beta\sum_\tau\frac{\expect\left[\sum_x c(\tau,p_x)  \right]}{\beta}\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+ \geq    &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{\expect\left[\sum_x c(\tau,p_x)  \right]}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+     =     &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{R(\tau)}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  
+  %% &\geq\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \frac{1}{c(p)}\sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
+  %% & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log\frac{1}{c(p)} \sum_{e\in \mathcal{D}} r(e,p)\theta_e
 \end{align*}
 where we have defined
-$$r(e:\tau',p)\triangleq\sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$$
+\begin{align*}
+  C(e)&\triangleq  \sum_x\expect\left[c(e,p_x) \right]\\
+  R(\tau)&\triangleq \expect\left[\sum_x c(\tau,p_x)  \right]\\
+  \beta&\triangleq\sum_\tau \expect\left[\sum_x c(\tau,p_x)  \right]
+\end{align*}
+Crucially it was defining $\beta$ that let us use Jensen's inequality. 
+Recalling from the main paper that $P(\theta)\triangleq\text{Dir}(\alpha)$,
+we have the following lower bound on M-step objective:
+\begin{align}
+\sum_e (C(e) + \alpha)\log \theta_e  - \beta\log \sum_\tau\frac{R(\tau)}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e    
+\end{align}
+Differentiate with respect to $\theta_e$, where $e:\tau$, and set to zero to obtain:
+\begin{align}
+  &  \frac{C(e) + \alpha}{\theta_e}\propto\sum_{\tau'}\indicator\left[\text{unify}(\tau,\tau') \right] R(\tau')\\
+&  \theta_e\propto\frac{C(e) + \alpha}{\sum_{\tau'}\indicator\left[\text{unify}(\tau,\tau') \right] R(\tau')}
+\end{align}
+The above is our estimator for $\theta_e$.
+Despite the convoluted derivation, the above estimator has an intuitive interpretation.
+The quantity $C(e)$ is the expected number of times that we used $e$.
+The quantity $\sum_{\tau'}\indicator\left[\text{unify}(\tau,\tau') \right] R(\tau')$
+is the expected number of times that we \emph{could have} used $e$.
+The hyperparameter $\alpha$ acts as pseudocounts that are
+added to the number of times that we used each primitive,
+and are not added to the number of times that we could have used each primitive.
 
-%% Returning to equation~\ref{maximizeStep}:
-%% \begin{align*}
-%% \log P(\theta) +& \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\geq \\
-%% \log   \text{Dir}(\theta|\alpha) +& \sum_x\sum_{e\in \mathcal{D}}\expect_{Q_x}\left[ c(e,p)\right]\log \theta_e -\\
-%% &\sum_x  \expect_{Q_x}\left[c(p)\log \sum_{e\in \mathcal{D}}r(e,p)\theta_e \right]
-%%   \end{align*}
-
-Differentiating with respect to $\theta_e$ and setting to zero:
-%% \begin{align}
-%%   & \frac{\alpha + \sum_x\expect_{Q_x}\left[ c(e,p)\right]}{\theta_e} -
-%% \theta_e\sum_x  \expect_{Q_x}\left[c(p)\frac{r(p,e)}{\sum_{e'\in \mathcal{D}} r(e',p)\theta_{e'}} \right]  
-%%   \end{align}
-%%
-\begin{align*}
-  \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}} \\
-  \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
-\end{align*}
-where $N=c(p)$ and $a(x) = r(x, p)$. From here forward, we similarly let
-$c(x)=c(x,p)$ for notational convenience.
-
-%% This equality holds if $\theta_x = c(x)/a(x)$:
-%% \begin{align}
-%%   \frac{c(x)}{\theta_x} &= a(x).\\
-%% N\frac{a(x)}{\sum_y a(y)\theta_y}& = N\frac{a(x)}{\sum_y c(y)}   = N\frac{a(x)}{N} = a(x).
-%% \end{align}
-If the above equality holds then $\theta_x \propto c(x)/a(x)$:
-\begin{align}
-  \theta_x = \frac{c(x)}{a(x)}\times \underbrace{\frac{\sum_y a(y)\theta_y}{N}}_{\text{Independent of $x$}}
-\end{align}
-
-Now what we are actually after is the parameters that maximize the joint log probability of the data and parameters $\theta$, which we  write $J$:
-\begin{align}
-  \nonumber J& = L + \log \text{Dir}(\theta|\alpha)\\
-  \nonumber&\stackrel{+}{\geq } \sum_x c(x)\log \theta_x - N \log \sum_x a(x)\theta_x \\
-  \nonumber&\qquad\quad+ \sum_x(\alpha_x - 1)\log \theta_x\\
-  & = \sum_x (c(x) + \alpha_x - 1)\log \theta_x -  N \log \sum_x a(x)\theta_x
-\end{align}
-We identify the parameters of the prior over $\theta$, $\alpha$, as  pseudocounts that are added to the \emph{counts} $c(x)$, but not to the \emph{possible counts} $a(x)$,
-and so we estimate $\theta_e$ as the ratio of the sum of the number of times that $e$ was used plus the pseudocounts $\alpha$, with the number of times that it could have been used, $a(e)$.
+We are only maximizing a lower bound on the log posterior; when is this lower bound tight?
+Whenever the DSL does not contain polymorphic types,
 

### 2018-05-12 nips_2018.tex
NIPS

--- 2018-02-11 paper.tex
@@
-%%%%%%%% ICML 2018 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
 \documentclass{article}
 
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
-\usepackage{graphicx}
-%\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
-
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
-%\usepackage{hyperref}
-
-% Attempt to make hyperref and algorithmic work together better:
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-\newcommand{\system}{\textsc{DreamCoder}~}
-\newcommand{\systemEnding}{\textsc{DreamCoder}}
-\newcommand{\lowerBound}{\mathscr{L}}
-\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
-\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
-% Use the following line for the initial blind version submitted for review:
-\usepackage[nohyperref]{icml2018}
-
-%\usepackage{hyperref}       % hyperlinks
+% if you need to pass options to natbib, use, e.g.:
+% \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading nips_2018
+
+% ready for submission
+\usepackage{nips_2018}
+
+% to compile a preprint version, e.g., for submission to arXiv, add
+% add the [preprint] option:
+% \usepackage[preprint]{nips_2018}
+
+% to compile a camera-ready version, add the [final] option, e.g.:
+% \usepackage[final]{nips_2018}
+
+% to avoid loading the natbib package, add option nonatbib:
+% \usepackage[nonatbib]{nips_2018}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
@@
 \usepackage{microtype}      % microtypography
+
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+\newcommand{\system}{\textsc{ECC}~}
+\newcommand{\systemEnding}{\textsc{ECC}}
+\newcommand{\lowerBound}{\mathscr{L}}
+\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
+\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
+
 \usepackage{mathrsfs}
@@
 \usepackage{amsthm}
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-%\usepackage{epsfig} % less modern
+
 \usepackage{subfig} 
@@
 \usepackage{caption}
-%\usepackage{subcaption}
-
-\fvset{fontsize=\footnotesize}
-
 \usepackage{amssymb}
@@
 \usetikzlibrary{fit,bayesnet}
-%\usetikzlibrary{arrows.meta}
-%\usetikzlibrary{positioning}
-%\usetikzlibrary{decorations.text}
-%\usetikzlibrary{decorations.pathmorphing}
 \usepackage{dsfont}
@@
 
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2018}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
+\title{Learning Reusable Components for Neurally--Guided Bayesian Program Learning}
+
+% The \author macro works with any number of authors. There are two
+% commands used to separate the names and addresses of multiple
+% authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to
+% break the lines. Using \AND forces a line break at that point. So,
+% if LaTeX puts 3 of 4 authors names on the first line, and the last
+% on the second line, try using \AND instead of \And before the third
+% author name.
+
+\author{
+  David S.~Hippocampus\thanks{Use footnote for providing further
+    information about author (webpage, alternative
+    address)---\emph{not} for acknowledging funding agencies.} \\
+  Department of Computer Science\\
+  Cranberry-Lemon University\\
+  Pittsburgh, PA 15213 \\
+  \texttt{hippo@cs.cranberry-lemon.edu} \\
+  %% examples of more authors
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \AND
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+}
 
 \begin{document}
-
-\twocolumn[
-\icmltitle{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2018
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Aeiau Zzzz}{equal,to}
-\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
-\icmlauthor{Cieua Vvvvv}{goo}
-\icmlauthor{Iaesut Saoeu}{ed}
-\icmlauthor{Fiuea Rrrr}{to}
-\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
-\icmlauthor{Aaoeu Iasoh}{goo}
-\icmlauthor{Buiui Eueu}{ed}
-\icmlauthor{Aeuia Zzzz}{ed}
-\icmlauthor{Bieea C.~Yyyy}{to,goo}
-\icmlauthor{Teoau Xxxx}{ed}
-\icmlauthor{Eee Pppp}{ed}
-\end{icmlauthorlist}
-
-\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
-\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
-\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
-
-\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
-\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Machine Learning, ICML}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+% \nipsfinalcopy is no longer used
+
+\maketitle
 
@@
   solutions to problems in the domain.
-
 \end{abstract}
@@
 $\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
-in the frontiers will most likely increase $\lowerBound$.  %% With $q$ in hand we can find programs for
+in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
 %% the frontier $\mathcal{F}_x$ by searching for programs that maximize
@@
 For example, it represents the linear function $-x+2$ with the program
-\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
+\code{(* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
 which has five continuous parameters.
@@
 We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
-At first, \system cannot programs for most of the tasks.
+At first, \system cannot find any correct programs for most of the tasks.
 After three wake/sleep iterations,
@@
 including the four domains considered,  but also many others?
-\bibliography{main}
-\bibliographystyle{icml2018}
-
+
+\bibliographystyle{unsrt}
+{\small \bibliography{main}}
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018. It was modified from a version from Dan Roy in
-% 2017, which was based on a version from Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2018-05-12 paper.tex
minor edit
NIPS

@@
 \begin{icmlauthorlist}
-\icmlauthor{Aeiau Zzzz}{equal,to}
-\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
-\icmlauthor{Cieua Vvvvv}{goo}
-\icmlauthor{Iaesut Saoeu}{ed}
-\icmlauthor{Fiuea Rrrr}{to}
-\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
-\icmlauthor{Aaoeu Iasoh}{goo}
-\icmlauthor{Buiui Eueu}{ed}
-\icmlauthor{Aeuia Zzzz}{ed}
-\icmlauthor{Bieea C.~Yyyy}{to,goo}
-\icmlauthor{Teoau Xxxx}{ed}
-\icmlauthor{Eee Pppp}{ed}
+\icmlauthor{Kevin Ellis}{equal,MIT}
+\icmlauthor{Lucas Morales}{equal,MIT}
+\icmlauthor{Armando Solar-Lezama}{MIT}
+\icmlauthor{Joshua B. Tenenbaum}{MIT}
 \end{icmlauthorlist}
 
-\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
-\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
-\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
-
-\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
-\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
+%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
+\icmlaffiliation{MIT}{MIT}
+%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
+
+\icmlcorrespondingauthor{Kevin Ellis}{ellisk@mit.edu}
+%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
 

### 2018-05-12 nips_2018.tex
intro

@@
   programs and imparting prior knowledge of the domain.  We contribute
-  a program induction algorithm called \system that learns a DSL and
-  trains a neural network to efficiently search for programs in the
-  learned DSL.  We use our model to build circuits, solve
-  symbolic regression problems, edit strings, and synthesize functions on lists, in each case
-  showing the model learns a domain-specific vocabulary for expressing
-  solutions to problems in the domain.
+  a program induction algorithm called \system that learns a DSL while
+  jointly training a neural network to efficiently search for programs
+  in the learned DSL.  We use our model to solve symbolic regression
+  problems, edit strings, and synthesize functions on lists, showing
+  how the model learns a domain-specific library of program components
+  for expressing solutions to problems in the domain.
 \end{abstract}
@@
 approaches -- symbolic, neural and neural-symbolic -- rely upon a
-hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
+hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
 inventory of restricted programming primitives, encoding domain-specific
@@
 
-We propose \emph{learning} the DSL. We consider the setting where we
-have a collection of related programming tasks, each specified by a
-set of input/output examples. We do not assume that the tasks are
-annotated with ground-truth programs. We typically will not try to
-learn a DSL completely from scratch, but rather to start from a weaker
-or more general set of primitives and
-construct a richer, more powerful, and better-tuned DSL.
-
-Our algorithm is called \system because it is based on a
-novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
-between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve tasks; creating a DSL by discovering
-and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL. The learned DSL
+We propose \emph{learning} the DSL by inducing a library of domain--specific
+subroutines. We consider the setting where we have a collection of
+related programming tasks, each specified by a set of input/output
+examples. %% We do not assume that the tasks are annotated with
+%% ground-truth programs.
+Start from a weaker or more
+general library of primitives, we give an algorithm for constructing a richer, more powerful,
+and better-tuned DSL.
+Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding)
+because it iterates between three different phases: an
+\textbf{Explore} phase uses the DSL to explore the space of programs,
+searching for ones that solve the tasks; a \textbf{Compress} phase
+modifies the DSL by discovering regularities in the programs found by
+the previous Explore phase; and a \textbf{Compile} phase, which
+improves the program search procedure by training a neural network to
+write programs in the current DSL, in the spirit of ``amortized'' or
+``compiled'' inference~\cite{le2016inference}.
+The learned DSL
 distills commonalities across programs that solve tasks, helping
@@
 even as the DSL (and hence the search space for programs) expands.
+
+Because any model may be encoded as a (deterministic or probabilistic) program,
+we carefully delineate the scope of program learning problems considered here.
+We think of \system as learning to solve the kinds of problems that humans
+can solve relatively quickly -- once they acquire the relevant domain expertise.
+These correspond to short programs -- once you have the right DSL.
+Even with the right DSL, program search may be intractable, so
+we amortize the cost of program search by training a neural network to
+assist the search procedure.
 
@@
 
+%% \begin{figure}
+%%   \centering
+
+%%   \begin{tikzpicture}
+
+%%   \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+%%   \node[latent] at (1.5,3) (t){$\theta$};
+%%   \node[latent] at (1,1.75) (z){$p$};
+%% %  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+%%   \node[obs] at (1,0.5) (x) {$x$};
+%%   \edge {z}{x};
+%%   \edge {d,t}{z};
+%%   \plate {}{(z)(x)}{$x\in X$};
+%%   \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};
+
+%%   \begin{scope}[shift = {(0.7,0)}]  
+%%   \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
+%%   \node[latent] at (3.5,1.9) (zp){$p$};
+%%   \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
+%%   \node[obs] at (3.5,0.5) (xp) {$x$};
+%%   \plate {}{(zp)(xp)(tx)}{$x\in X$};
+%%   \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
+%%   \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
+%%   \draw [->,red] (tx.west) -- (zp.east);
+%%   \draw [->,red] (dx.south) -- (zp.north);
+
+%%   \node at (nn) {
+%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
+%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%       \node[rectangle] at (1,1) {};
+%%       \foreach \name / \y in {1,...,4}
+%%           \node[neuron] (I-\name) at (0,-\y) {};
+%%       \foreach \name / \y in {1,...,3}
+%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%       \foreach \name / \y in {1,...,4}
+%%           \node[neuron] (O-\name) at (2,-\y) {};
+%%       \foreach \source in {1,...,4}
+%%           \foreach \dest in {1,...,3}
+%%               \draw [-latex] (I-\source) -- (H-\dest);
+%%       \foreach \source in {1,...,3}
+%%           \foreach \dest in {1,...,4}
+%%               \draw [-latex] (H-\source) -- (O-\dest);
+%%     \end{tikzpicture}
+%%   };
+%%   \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
+
+%%   \end{scope}
+
+%%   \end{tikzpicture}
+%%   \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$.
+From each program we observe a task $x$ (program inputs/outputs).
+(b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
+%%     $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
+%% \end{figure}
+
+%% To perform inference, we take inspiration from the Helmholtz machine
+%% and the Wake/Sleep algorithm~\cite{hinton1995wake}, which alternates
+%% between ``waking'', which draws samples from a recognition model (our
+%% $q$); and ``sleeping'', which trains a recognition model on samples
+%% from the generative model (our $\mathcal{D},\theta$).  For us, a
+%% \textbf{Wake} cycle means using the recognition model to search for
+%% programs that solve the tasks.  A sleep cycle means updating these
+%% models so that the DSL and recognition model more closely match the
+%% distribution of tasks.  An important difference from the Helmholtz
+%% machine is that our setup has two distinct sleep phases: a phase where
+%% we update the generative model $(\mathcal{D},\theta)$, which we call
+%% \textbf{Sleep-G}, and a second phase where we update the recognition
+%% model $q$, called \textbf{Sleep-R}.
+
+%% We take the intuitive strategy of exploiting the
+%% conditional independence of the programs conditioned on the DSL, and
+%% so alternate between estimating the DSL ($\mathcal{D},\theta$) and
+%% estimating a posterior distribution over $p$ by searching for
+%% programs.  In practice searching for
+%% programs is difficult, so we use the recognition model to
+%% learn how to search efficiently.
+
+We apply \system to three domains:
+ symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
+ For each of these we provide a generic
+ set of programming primitives in a Lisp-like language,
+ including conditionals, variables, and higher-order functions.
+ Our algorithm then discovers
+ its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+
+ \begin{wrapfig}
+    \begin{tabular}{ll}
+   \toprule
+   Domain&Part of the learned DSL\\\midrule
+   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
+   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
+   Regression& \code{(+ (* real x) real)} \\
+      &\hspace{0.5cm} \emph{(a linear function of x)}\\
+   Strings& \code{(join '' (split c s))}\\
+   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
+   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
+   &\hspace{0.5cm} \emph{(add k to every element of list l)}
+   \\\bottomrule
+    \end{tabular}
+    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
+ \end{wrapfig}
+
+%% \begin{itemize}
+%%   \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
+%%   \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
+%%     $q$;
+%%     % because this is a primary inspiration, might be worth further mention:
+%%     % different program representation (routing vs. substitution; simple vs polymorphic types);
+%%     % different grammar induction (sequitur vs. fragment grammar)
+%%   \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
+%%     DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
+%%     network $q$.
+%%     % although RobustFill has very different enumeration
+%% \end{itemize}
+
+%% We cast these problems as \emph{Bayesian Program
+%%   Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
+%% where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
+%% A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
+%% We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
+%% define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
+%% In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
+%% where the likelihood $\probability[x|p]$ is domain-dependent.
+%% The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
+%% Alongside this generative model,
+%% we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
+
+ \section{The \system Algorithm}
+
+ Our goal is to induce a DSL while finding programs solving each of the
+tasks.  We take inspiration primarily from the
+Exploration-Compression algorithm for bootstrap
+learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
+alternates between exploring the space of solutions to a set of tasks,
+and compressing those solutions to suggest new search primitives
+for the next exploration stage.
+We extend these ideas into an inference strategy that iterates
+through three steps: an \textbf{Explore} cycle uses the current DSL and
+recognition model to search for programs that solve the tasks.  The
+\textbf{Compress} and \textbf{Compile} cycles update the DSL and the
+recognition model, respectively.  Crucially, these steps bootstrap 
+off each other (Fig.~\ref{feeding}):
+
+\\\noindent \textbf{Wake: Searching for programs.}  Our program search
+is informed by both the DSL and the recognition model. When these
+improve, we find more programs solving the tasks.
+\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce the DSL
+from the programs found in the wake phase, aiming
+to maximally compress (or, raise the prior probability of) these
+programs.  As we solve more tasks, we hone in on richer DSLs that
+more closely match the domain.\footnote{This is loosely biologically
+    inspired by the formation of abstractions during sleep memory
+    consolidation~\cite{DUDAI201520}}
+\\\noindent \textbf{Sleep-R: Learning a neural recognition model.}  We
+update the recognition model by training on two data sources: samples
+from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
+programs found by the search procedure during waking. As the DSL
+improves and as search finds more programs, the recognition model gets
+both more data to train on, and better data.\footnote{These two sources are also
+    loosely biologically inspired by the distinct episodic replay and
+    hallucination components of dream sleep~\cite{fosse2003dreaming}}
+
+%% Our goal  is to induce a DSL while finding programs solving each of the tasks.
+%% Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
+%% to accelerate the search over programs -- and (3) improving the DSL.
+%% The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
+%% \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
+%% When these improve, we find more programs solving the tasks.
+%% \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
+As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
+
+%% The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
+%% learning a generative model and training a recognition model on samples from the generative model.
+%% We apply this idea to program induction, motivating why we call our algorithm the \system.
+
+\begin{wrapfig}\centering
+  \begin{tikzpicture}
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs};
+    \node[align = center](DSL) at (3,1) {DSL};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};
+
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compression)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+  \caption{\system solves for programs, the DSL, and a neural network called the \emph{recognition model}.
+Each of these steps bootstrap off of the others in an iterative inference algorithm.}  \label{feeding}
+\end{wrapfigure}
+
+Sec.~\ref{mathematicalFraming} frames this 3-step procedure as
+a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
+Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
+Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
+Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
+
+\subsection{Probabilistic Framing}\label{mathematicalFraming}
 Viewed as a probabilistic inference problem,
@@
 iteratively learn to solve more programming tasks.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}
-
-  \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-  \node[latent] at (1.5,3) (t){$\theta$};
-  \node[latent] at (1,1.75) (z){$p$};
-%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-  \node[obs] at (1,0.5) (x) {$x$};
-  \edge {z}{x};
-  \edge {d,t}{z};
-  \plate {}{(z)(x)}{$x\in X$};
-  \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};
-
-  \begin{scope}[shift = {(0.7,0)}]  
-  \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
-  \node[latent] at (3.5,1.9) (zp){$p$};
-  \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
-  \node[obs] at (3.5,0.5) (xp) {$x$};
-  \plate {}{(zp)(xp)(tx)}{$x\in X$};
-  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
-  \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
-  \draw [->,red] (tx.west) -- (zp.east);
-  \draw [->,red] (dx.south) -- (zp.north);
-
-  \node at (nn) {
-    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
-      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-      \node[rectangle] at (1,1) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (I-\name) at (0,-\y) {};
-      \foreach \name / \y in {1,...,3}
-          \node[neuron] (H-\name) at (1,-\y-0.5) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (O-\name) at (2,-\y) {};
-      \foreach \source in {1,...,4}
-          \foreach \dest in {1,...,3}
-              \draw [-latex] (I-\source) -- (H-\dest);
-      \foreach \source in {1,...,3}
-          \foreach \dest in {1,...,4}
-              \draw [-latex] (H-\source) -- (O-\dest);
-    \end{tikzpicture}
-  };
-  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
-
-  \end{scope}
-
-  \end{tikzpicture}
-  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$.
-From each program we observe a task $x$ (program inputs/outputs).
-(b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
-    $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
-\end{figure}
-
-%% To perform inference, we take inspiration from the Helmholtz machine
-%% and the Wake/Sleep algorithm~\cite{hinton1995wake}, which alternates
-%% between ``waking'', which draws samples from a recognition model (our
-%% $q$); and ``sleeping'', which trains a recognition model on samples
-%% from the generative model (our $\mathcal{D},\theta$).  For us, a
-%% \textbf{Wake} cycle means using the recognition model to search for
-%% programs that solve the tasks.  A sleep cycle means updating these
-%% models so that the DSL and recognition model more closely match the
-%% distribution of tasks.  An important difference from the Helmholtz
-%% machine is that our setup has two distinct sleep phases: a phase where
-%% we update the generative model $(\mathcal{D},\theta)$, which we call
-%% \textbf{Sleep-G}, and a second phase where we update the recognition
-%% model $q$, called \textbf{Sleep-R}.
-
-%% We take the intuitive strategy of exploiting the
-%% conditional independence of the programs conditioned on the DSL, and
-%% so alternate between estimating the DSL ($\mathcal{D},\theta$) and
-%% estimating a posterior distribution over $p$ by searching for
-%% programs.  In practice searching for
-%% programs is difficult, so we use the recognition model to
-%% learn how to search efficiently.
-
-We apply \system to four domains:
- building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
- For each of these we deliberately provide an impoverished
- set of programming primitives,
- and show that our algorithm discovers
- its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
- \begin{table}[b]
-    \begin{tabular}{ll}
-   \toprule
-   Domain&Part of the learned DSL\\\midrule
-   Circuits&\code{(nand (nand a a) (nand b b))}\\
-   &\hspace{0.5cm} \emph{(an OR gate)}\\
-   Regression& \code{(+ (* real x) real)} \\
-      &\hspace{0.5cm} \emph{(a linear function of x)}\\
-   Strings& \code{(join '' (split c s))}\\
-   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
-   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
-   &\hspace{0.5cm} \emph{(add k to every element of list l)}
-   \\\bottomrule
-    \end{tabular}
-    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
- \end{table}
-
-%% \begin{itemize}
-%%   \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
-%%   \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
-%%     $q$;
-%%     % because this is a primary inspiration, might be worth further mention:
-%%     % different program representation (routing vs. substitution; simple vs polymorphic types);
-%%     % different grammar induction (sequitur vs. fragment grammar)
-%%   \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
-%%     DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
-%%     network $q$.
-%%     % although RobustFill has very different enumeration
-%% \end{itemize}
-
-%% We cast these problems as \emph{Bayesian Program
-%%   Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
-%% where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
-%% A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
-%% We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
-%% define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
-%% In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
-%% where the likelihood $\probability[x|p]$ is domain-dependent.
-%% The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
-%% Alongside this generative model,
-%% we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
-
- \section{The \system Algorithm}
-
- Our goal is to induce a DSL while finding programs solving each of the
-tasks.  We take inspiration from the Wake/Sleep
-algorithm~\cite{hinton1995wake}, as well as the
-Exploration-Compression algorithm (EC) for bootstrap
-learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Wake/Sleep
-alternates between training a generative model on samples from a
-recognition model (our $q$), and training a recognition model on
-samples from the generative model (our $\mathcal{D},\theta$).  EC
-alternates between exploring the space of solutions to a set of tasks,
-and compressing those solutions to suggest new search primitives
-for the next exploration stage.
-We combine these ideas into an inference strategy that iterates
-through three steps: a \textbf{Wake} cycle uses the current DSL and
-recognition model to search for programs that solve the tasks.  The
-\textbf{Sleep-G} and \textbf{Sleep-R} cycles update the DSL and the
-recognition model, respectively.  Crucially, these steps bootstrap 
-off each other (Fig.~\ref{feeding}):
-
-\\\noindent \textbf{Wake: Searching for programs.}  Our program search
-is informed by both the DSL and the recognition model. When these
-improve, we find more programs solving the tasks.
-\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce the DSL
-from the programs found in the wake phase, aiming
-to maximally compress (or, raise the prior probability of) these
-programs.  As we solve more tasks, we hone in on richer DSLs that
-more closely match the domain.\footnote{This is loosely biologically
-    inspired by the formation of abstractions during sleep memory
-    consolidation~\cite{DUDAI201520}}
-\\\noindent \textbf{Sleep-R: Learning a neural recognition model.}  We
-update the recognition model by training on two data sources: samples
-from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
-programs found by the search procedure during waking. As the DSL
-improves and as search finds more programs, the recognition model gets
-both more data to train on, and better data.\footnote{These two sources are also
-    loosely biologically inspired by the distinct episodic replay and
-    hallucination components of dream sleep~\cite{fosse2003dreaming}}
-
-%% Our goal  is to induce a DSL while finding programs solving each of the tasks.
-%% Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
-%% to accelerate the search over programs -- and (3) improving the DSL.
-%% The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
-%% \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-%% When these improve, we find more programs solving the tasks.
-%% \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
-As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-
-%% The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
-%% learning a generative model and training a recognition model on samples from the generative model.
-%% We apply this idea to program induction, motivating why we call our algorithm the \system.
-
-\begin{figure}\centering
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
-    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
-  \caption{\system solves for programs, the DSL, and a recognition model.
-Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
-\end{figure}
-
-Sec.~\ref{mathematicalFraming} frames this 3-step procedure as
-a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
-Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
-Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
-
-\subsection{Probabilistic Framing}\label{mathematicalFraming}
 
@@
 We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
-\begin{align}
-  \nonumber
-    \log \probability[\mathcal{D}] + \argmax_{\theta}&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  &+ \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
-  \end{align}
-
+\begin{equation}
+      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
+  \end{equation}
 We induce a DSL by searching locally through the space of DSLs,
@@
 We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
-\begin{align}
-  \probability[\mathcal{D}]&\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\\
-  \probability[\theta|\mathcal{D}]& = \text{Dir}(\theta|\alpha)
-\end{align}
+\begin{equation}
+    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\hspace{1cm}  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)
+  \end{equation}
+
 where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
@@
 of our model that are analogous to prior work.
-
-\subsection{Boolean Circuits}
-We consider the problem of learning
-to build  Boolean circuits out of logic gates.
-If we gave our agent the full repertoire of
-circuit primitives -- XOR gates, multiplexers, etc. -- this problem would be easy,
-and we could use sophisticated symbolic search algorithms like SMT solvers to synthesize circuits.
-Instead, we gave \system only a NAND gate,
-and tasked it with building a set of 250 random circuits made out of
-AND, NOT, and OR gates, along with a multiplexer.
-\system observes the input/output truth table of the circuits,
-and we define the likelihood 
-$  \probability[x|p]\triangleq\indicator\left[p\text{ has  the same truth table as }x \right]$,
-where $x$ is a truth table and $p$ a program implementing a circuit.
-
-This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
-%% For this domain our recognition model is a simple multilayer perceptron 
-%% whose inputs are the input/output examples.
-\system learns to build these
-circuits by creating new subroutines that correspond to
-logic gates and multiplexers (Fig.~\ref{multiplexer}).
-We will show that the neural recognition model
-assists in building these circuits (Sec.~\ref{quantitative}),
-suggesting that an agent can glean something about the structure of a circuit just by looking at its truth table.
-
-\begin{figure}\centering
-  \begin{circuitikz} \draw
-    (0,0) node (x) {$x$}
-    (x)++(0,-0.3) node (y) {$y$}
-    (y)++(0,-0.5) node (z) {$z$}
-
-    (x)++(1.7,-0.14) node[nand port, scale = 0.5] (xy) {}
-    (xy) ++ (0,-1) node[nand port, scale = 0.5] (xz) {}
-    (xz) ++ (1,0.2) node[nand port, scale = 0.5] (zxz) {}
-    (zxz) ++ (1,0.2) node[nand port, scale = 0.5] (m) {}
-    (x) -- (xy.in 1)
-    (y) -- (xy.in 2)
-
-    (x) -| (xz.in 1)
-    (z.east) -- (xz.in 2)
-
-    (xz.out) -- (zxz.in 2)
-    (z.east) |- (zxz.in 1)
-
-    (zxz.out) -- (m.in 2)
-    (xy.out) |- (m.in 1)
-
-    (m)++(1.2,0) node (l) {\textsc{MUX}$(x,y,z)$}
-    ;
-%% (0,0) node[and port] (myand2) {}
-%% (2,1) node[xnor port] (myxnor) {}
-%% (myand1.out) -- (myxnor.in 1)
-%% (myand2.out) -- (myxnor.in 2);
-  \end{circuitikz}
-  \begin{circuitikz} \draw
-    (0,0) node (a) {$a$}
-    (a)++(0,-0.7) node (b) {$b$}
-
-    (a)++(1,-0) node[nand port, scale = 0.5] (ab1) {}
-    (b)++(1,-0) node[nand port, scale = 0.5] (ab2) {}
-    (ab1)++(1,-0.35) node[nand port, scale = 0.5] (o) {}
-
-    (a.east) -| (ab1.in 1)
-    (a.east) -| (ab1.in 2)
-    (b.east) -| (ab2.in 1)
-    (b.east) -| (ab2.in 2)
-
-    (ab1.out) -- (o.in 1)
-    (ab2.out) -- (o.in 2)
-
-    (o)++(0.8,0) node (l) {\textsc{OR}$(a,b)$}
-    ;
-  \end{circuitikz}
-  \begin{circuitikz}     \node(m)[draw] at (1,-1) {MUX}; \draw
-    (0,0) node (w) {$w$}
-    (w)++(0,-0.5) node (x) {$x$}
-    (x)++(0,-0.5) node (y) {$y$}
-    (y)++(0,-0.5) node (z) {$z$}
-
-    (z.east) -| (m.south)
-    (y.east) -| (m.west)
-    (x.east) -| (m.north)
-
-    (y)++(2.5,0.13) node[or port, scale = 0.5] (z) {}
-    (m.east) -- (z.in 2)
-    (w.east) -| (z.in 1)
-    ;
-  \end{circuitikz}
-\caption{\system discovers multiplexers (above) and OR gates (bottom left), which it uses to make more complicated circuits (bottom right) that it was unable to build without discovering these new primitives.}\label{multiplexer}
-  \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width = \columnwidth]{figures/circuit.png}
-%%   \caption{\system learning to build Boolean circuits. We start out with only NAND gates and learn DSL that contains ...}\label{circuitLearningCurve}
-%% \end{figure}
 

### 2018-05-12 nips_2018.tex
shortened and got rid of a lot of wake/sleep

@@
 %% ground-truth programs.
-Start from a weaker or more
+Starting from a weaker or more
 general library of primitives, we give an algorithm for constructing a richer, more powerful,
@@
 ``compiled'' inference~\cite{le2016inference}.
+We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
 The learned DSL
@@
 the agent solve related program induction problems. The neural
-network ensures that searching for programs remains tractable
+recognition model ensures that searching for programs remains tractable
 even as the DSL (and hence the search space for programs) expands.
@@
  its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
-
- \begin{wrapfig}
+\begin{wrapfigure}{r}{0.6\textwidth}
     \begin{tabular}{ll}
@@
     \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
- \end{wrapfig}
+ \end{wrapfigure}
 
@@
 
-\\\noindent \textbf{Wake: Searching for programs.}  Our program search
+\\\noindent \textbf{Exploration: Searching for programs.}  Our program search
 is informed by both the DSL and the recognition model. When these
 improve, we find more programs solving the tasks.
-\\\noindent\textbf{Sleep-G: Improving the DSL.} We induce the DSL
-from the programs found in the wake phase, aiming
+\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
+from the programs found in the exploration phase, aiming
 to maximally compress (or, raise the prior probability of) these
 programs.  As we solve more tasks, we hone in on richer DSLs that
-more closely match the domain.\footnote{This is loosely biologically
-    inspired by the formation of abstractions during sleep memory
-    consolidation~\cite{DUDAI201520}}
-\\\noindent \textbf{Sleep-R: Learning a neural recognition model.}  We
+more closely match the domain.
+\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
 update the recognition model by training on two data sources: samples
 from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
-programs found by the search procedure during waking. As the DSL
+programs found by the search procedure during exploration. As the DSL
 improves and as search finds more programs, the recognition model gets
-both more data to train on, and better data.\footnote{These two sources are also
-    loosely biologically inspired by the distinct episodic replay and
-    hallucination components of dream sleep~\cite{fosse2003dreaming}}
-
-%% Our goal  is to induce a DSL while finding programs solving each of the tasks.
-%% Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
-%% to accelerate the search over programs -- and (3) improving the DSL.
-%% The key observation is that these steps bootstrap off each other (Fig.~\ref{feeding}):
-%% \\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
-%% When these improve, we find more programs solving the tasks.
-%% \\\noindent\textbf{Sleep-R: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure.
-As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
-
-%% The Helmholtz machine~\cite{dayan1995helmholtz} introduced the idea of alternating between
-%% learning a generative model and training a recognition model on samples from the generative model.
-%% We apply this idea to program induction, motivating why we call our algorithm the \system.
-
-\begin{wrapfig}\centering
+both more data to train on, and better data.
+\begin{wrapfigure}{r}{0.6\textwidth}
   \begin{tikzpicture}
@@
 
-Sec.~\ref{mathematicalFraming} frames this 3-step procedure as
-a means of maximizing a lower bound on the posterior probability of the DSL given the tasks.
+Sec.~\ref{mathematicalFraming} frames this 3-step procedure as probabilistic inference.
 Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
@@
 \subsection{Probabilistic Framing}\label{mathematicalFraming}
-Viewed as a probabilistic inference problem,
-our goal is to jointly infer a single DSL (written $\mathcal{D}$)
-and many programs (written $p$), one for each set of inputs/outputs (\emph{tasks}, written $x$).
-A DSL $\mathcal{D}$ is a collection of domain-specific subroutines.
-We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
-$(\mathcal{D},\theta)$ define a distribution over programs.
-Fig.~\ref{graphicalModel}(a) diagrams this inference problem as a hierarchical Bayesian model.
-Inference in this model is
-difficult because the programs are unobserved,
-and so we must solve a hard search problem to recover them. To make
-search tractable we learn a bottom-up \emph{recognition
-  model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
-The recognition model $q(\cdot )$ is a neural
-network that regresses from input/output pairs to a distribution over
-programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
-inference
-scheme~\cite{le2016inference}.
- The neural recognition model and the
-generative model embodied in the DSL jointly train each other, as they
-iteratively learn to solve more programming tasks.
 
@@
 Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL $\mathcal{D}$.
-
-We frame this problem as maximum a posteriori (MAP) inference in the
-generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
+and also to infer a DSL, written $\mathcal{D}$.
+We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
+$(\mathcal{D},\theta)$ define a generative model over programs.
+We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
+Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
-\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
+  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta],\quad
+  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta, \quad
+  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
 \end{align}
+
+%% Inference in this model is
+%% difficult because the programs are unobserved,
+%% and so we must solve a hard search problem to recover them. To make
+%% search tractable we learn a bottom-up \emph{recognition
+%%   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
+%% The recognition model $q(\cdot )$ is a neural
+%% network that regresses from input/output pairs to a distribution over
+%% programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
+%% inference
+%% scheme~\cite{le2016inference}.
+%%  The neural recognition model and the
+%% generative model embodied in the DSL jointly train each other, as they
+%% iteratively learn to solve more programming tasks.
+
 The above equations summarize the problem from the point of view of an ideal Bayesian learner.
@@
 is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
-summing over the  infinite set of all possible programs.
+summing over the  infinite set of all programs.
 In practice we will only ever be able to sum over a finite set of programs.
@@
 
-Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
+Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
 \begin{align}
@@
 %%   \end{align}
-\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
-%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
-%explains how \system synthesizes new programs and given a DSL.
+\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
+alternating maximization w.r.t. the frontiers (Exploration) and the DSL (Compression):
 \\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
@@
 of the large combinatorial search space. We ease this difficulty by training a neural recognition model
-during the Sleep-R phase:
+during the compilation phase:
 
@@
 %% $q(p|x)$.
-Rather than directly predicting a distribution over $p$ conditioned on $x$,
-the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
-This taps into the intuition that programming is primarily a top-down activity:
-as human programmers, we often first decide what kind of
-programming constructs we might need to use,
-and then we figure out how to assemble them in to the desired program.
+%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
+%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
+%% This taps into the intuition that programming is primarily a top-down activity:
+%% as human programmers, we often first decide what kind of
+%% programming constructs we might need to use,
+%% and then we figure out how to assemble them in to the desired program.
 %% This approach implements an amortized
@@
 
-\subsection{Wake Phase: Searching for Programs}\label{explorationSection}
+\subsection{Exploration: Searching for Programs}\label{explorationSection}
 
@@
 
-\subsection{Sleep-R: Learning a Recognition Model}\label{recognitionSection}
+\subsection{Compilation: Learning a Recognition Model}\label{recognitionSection}
 
 The purpose of the recognition model is to accelerate the search for
-programs.  It does this by learning to predict programs which are probable
-under $(\mathcal{D},\theta)$ while also assigning high likelihood for a task
-according to $\probability[x|p]$.
-
-The recognition model $q$ is a neural network that predicts,
-for each task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.
-Together with the DSL, this defines a distribution over programs,
-$\probability[p|\mathcal{D},\theta = q(x)]$.
-We abbreviate this distribution as $q(p|x)$.
-The crucial aspect of this framing is that the neural network
-leverages the structure of the learned DSL,
-so it is \emph{not} responsible for
-generating programs wholesale.
-We share this aspect with DeepCoder~\cite{balog2016deepcoder}.
+programs.  It does this by learning to predict programs which are
+probable under $(\mathcal{D},\theta)$ while also assigning high
+likelihood for a task according to $\probability[x|p]$.  Concretely,
+the recognition model $q$ is a neural network that predicts, for each
+task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
+\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
+a distribution over programs, $\probability[p|\mathcal{D},\theta =
+  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
+aspect of this framing is that the neural network leverages the
+structure of the learned DSL, so it is \emph{not} responsible for
+generating programs wholesale.  We share this aspect with
+DeepCoder~\cite{balog2016deepcoder}.
 
@@
 like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
-We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}}$:
+We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
 \begin{align}
-\mathcal{L}_{\text{RM}}& = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}\\
-\mathcal{L}_{\text{HM}}& = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\\
-\mathcal{L}_{\text{AE}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]\nonumber
+\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\qquad
+\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
 \end{align}
@@
 
-\subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
+\subsection{Compression: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
 
@@
 very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
-for which we have effective approaches like the Inside/Outside algorithm~(Lafferty,~2000), %\cite{international2000derivation},
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
 a DSL in \system may be context-sensitive due to the presence of variables
@@
 We note the following implementation details:
-(1) We perform a wake cycle before each of the sleep cycles.
+(1) We perform an exploration cycle before each compression and compilation  cycle.
 (2) On the first iteration, we do \emph{not} train the recognition model on samples from
@@
 programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
-(4) For added robustness, we enumerate programs from both the generative model and the recognition model.
+(4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.
 
@@
      \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
+     \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
      \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
@@
      %     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Wake})}
-     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Sleep-R}})
-     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Wake)}}
-
-     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Sleep-G)}}
+     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
+     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Compile}})
+     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}
+
+     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
      %% \STATE Define $Q_x(z) \propto \begin{cases}

### 2018-05-12 nips_2018.tex
DSL fragment example

@@
 \usepackage{microtype}      % microtypography
+\usepackage{stmaryrd}
 
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
-
+\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
 \newcommand{\system}{\textsc{ECC}~}
@@
 \\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
+is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
 
@@
 Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by training a neural recognition model
-during the compilation phase:
+of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
+during the compilation phase: $q$ is trained to approximate the
+posterior over programs, $q(p|x)\propto\probability[x,p|\mathcal{D},\theta}]$,
+  thus amortizing the cost of finding programs with high posterior probability.
 
@@
 
-\subsection{Compilation: Learning a Recognition Model}\label{recognitionSection}
+\subsection{Compilation: Learning a Neural Recognition Model}\label{recognitionSection}
 
@@
 generating programs wholesale.  We share this aspect with
-DeepCoder~\cite{balog2016deepcoder}.
+DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.
 
@@
 We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
-\begin{align}
-\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right],\text{ $p$ evaluates to $x$}\nonumber\qquad
+\begin{align}\nonumber
+\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\text{ s.t. $x = \sem{p}$}\quad
 \mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
@@
 \begin{figure*}
-\centering  \begin{tabular}{lll}
+\centering  \begin{tabular}{ll}
     \toprule
-    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
-    Circuits&
+    Example programs in frontiers&Proposed subexpression\\\midrule
     \begin{tabular}{l}
-      \code{(lambda x (nand x x))}\\
-      \code{(lambda x (lambda y (nand x (nand y y))}
+      \code{(lambda (a b) (fold b (cons "," a) (lambda (x z) (cons x z))))}\\
+      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
     \end{tabular}&
-    \code{(nand z z)}\\\midrule
-    Strings&
-    \begin{tabular}{l}
-      \code{(lambda s (+ ',' (index 0 (split ',' s))))}\\
-      \code{(lambda s (index 0 (split ' ' s)))}
-    \end{tabular}&
-    \code{(index 0 (split z s))}\\\bottomrule
+    \code{(fold a b (lambda (x z) (cons x z)))}\\\bottomrule
   \end{tabular}
   \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
-    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\code{z} and \code{s} in the right column).} 
+    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}).
+Here, the proposed subexpression appends two lists.}
   \label{fragmentExample}

### 2018-05-12 nips_2018.tex
more writing

@@
 We task our model
-with solving 200 symbolic regression problems, each a polynomial of
-degree 1 to 4, where our observations $x$ are 10
-input/output examples.%, which we write as $x = \left\{(i_n,o_n)
-%\right\}_{n\leq N}$. %% For example, one task is to infer a program
-%% calculating $3x + 2$, and the observations are the input-output
-%% examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.
-
-We initially give our system only addition and multiplication,
+with solving 200 symbolic regression problems, each either a polynomial of
+degree 1 to 4 or a rational function.
+The recognition model is a convolutional network that observes an image
+of the target function's graph (Fig.~\ref{functions}) --
+visually, different kinds of polynomials and rational functions
+produce different kinds of graphs,
+and so the recognition model must learn to
+look at a graph and predict what kind of function best explains it.
+\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}
+  \includegraphics[width = \functionSize]{figures/functions/0.png}
+  \includegraphics[width = \functionSize]{figures/functions/6.png}
+  \includegraphics[width = \functionSize]{figures/functions/207.png}
+  \includegraphics[width = \functionSize]{figures/functions/160.png}
+  \includegraphics[width = \functionSize]{figures/functions/303.png}\\  
+  \includegraphics[width = \functionSize]{figures/functions/402.png}
+  \includegraphics[width = \functionSize]{figures/functions/415.png}
+  \includegraphics[width = \functionSize]{figures/functions/426.png}
+  \includegraphics[width = \functionSize]{figures/functions/597.png}
+  \includegraphics[width = \functionSize]{figures/functions/591.png}
+  \caption{Recognition model input for symbolic regression.
+While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
+  \end{wrapfigure}
+
+We initially give our system addition, multiplication, and division,
 along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
@@
   \centering
-  \begin{tabular}{c}
-    String editing primitives\\
     \begin{tabular}{l}
       \toprule
-    \code{0}, \code{+1},\code{-1},\codechar{},\codechar{,},\codechar{ },\codechar{<},\codechar{>}  \\
-    \code{split}, \code{join}, \code{map}, 
-    \code{+}, \code{chr->str}, \code{trim}, \code{nth}\\
-    \code{length}, \code{slice},    \code{lower}, \code{upper}, \code{capitalize}
-  \\\bottomrule
-    \end{tabular}\\
-\\    Subset of learned DSL\\
-    \begin{tabular}{l}
-      \toprule
-      $f_0($\code{c}$) \,=\, $\code{(+ (chr->str c))}\\
-      $f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ (nth i (split c s))}\\
-        \phantom{$f_1($\code{i$,$c$,$s}$) \,=\, $\code{(+ }}\code{(chr->str c))}\\
-      $f_2($\code{s$,$c}$) \,=\, $\code{((}$f_0$\code{ c) (}$f_1$\code{ (+1 0) c s))} \\
-        \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
-      $f_3($\code{c$,$s}$) \,=\, $\code{(nth (+1 0) (split c (nth 0 s)))}\\
-        \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
-    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
-    %% $f_1 = $\code{(join "")}
+      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
+      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
+        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
+        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
   \\\bottomrule
     \end{tabular}
-  \end{tabular}
-  \caption{Top: provided string editing primitives. Bottom: some learned DSL primitives.}\label{textPrimitives}
+  \caption{Some learned DSL primitives for string editing}\label{textPrimitives}
   \end{table}

### 2018-05-12 nips_2018.tex
text edit examples

@@
 \begin{figure}
-  \centering\begin{tabular}{ll|ll}\toprule
-    Input&Output&Input&Output\\\midrule
-cqbHT$>$aIKd$<$W& 	AIKD&BgZQ..PBBnD & BgZQ. \\
-YBW$>$TIg$<$WS 	&TIG& qqfgg..mNIx & qqfgg. 
+  \centering\begin{tabular}{ll|lll}\toprule
+    Input&Output&Input 1&Input 2&Output\\\midrule
+Temple Annalisa Haven 185&TAH1&Launa&Withers&Launa Withers\\
+Lara Gregori Bradford&LGB& Rudolf&Akiyama&Rudolf Akiyama
 %DkI$>$DwKI$<$xaPoc& 	DWKI\\
@@
 \\\midrule
-\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_3$\code{ (split }\codechar{>}\code{ } }&
-\multicolumn{2}{l}{$f(\text{\code{s}}) = $\code{(}$f_1$\code{ s }\codechar{.}\code{ 0)}}
+\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
+\multicolumn{2}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
 \\
-\multicolumn{2}{l|}{\hspace{1.1cm}\code{(upper s)) }\codechar{<}\code{)}}
-\\\bottomrule
+\bottomrule
   \end{tabular}
@@
 
-  \caption{Two string edit tasks (top) and the programs \system writes for them (bottom). $f_3$ are $f_1$  discovered DSL subroutines, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+  \caption{Two string edit tasks (top) and the programs \system writes for them (bottom).
+$f_0$ and $f_2$ are   subroutines written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
 \end{figure}

### 2018-05-12 nips_2018.tex
listed primitives

@@
 
-The architecture of $q$ depends upon the domain.
-For domains with sequential structure like strings and lists,
-we use an RNN to encode each input/output example.
-For domains with fixed-dimensionality inputs/outputs we use a multilayer perceptron.
-We MaxPool across all of the input/output examples. See Supplement for details.
+%% The architecture of $q$ depends upon the domain.  For domains with
+%% sequential structure like strings and lists, we use an RNN to encode
+%% each input/output example.  For domains with fixed-dimensionality
+%% inputs/outputs we use a multilayer perceptron.  We MaxPool across all
+%% of the input/output examples. See Supplement for details.
 
@@
 \section{Experiments}
-We first present our experimental results qualitatively,
-illustrating the learned DSLs.
-We then show quantitative results (Sec.~\ref{quantitative})
-comparing \system with lesioned versions
-of our model that are analogous to prior work.
+%% We first present our experimental results qualitatively,
+%% illustrating the learned DSLs.
+%% We then show quantitative results (Sec.~\ref{quantitative})
+%% comparing \system with lesioned versions
+%% of our model that are analogous to prior work.
 
 \subsection{Symbolic Regression}\label{regressionSection}
-We show how to use \system to infer programs containing both discrete
-structure and continuous parameters. The high-level idea is to write programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
-and optimizing using gradient descent.
-We task our model
-with solving 200 symbolic regression problems, each either a polynomial of
-degree 1 to 4 or a rational function.
-The recognition model is a convolutional network that observes an image
-of the target function's graph (Fig.~\ref{functions}) --
-visually, different kinds of polynomials and rational functions
-produce different kinds of graphs,
-and so the recognition model must learn to
-look at a graph and predict what kind of function best explains it.
 \begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}
@@
 While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
-  \end{wrapfigure}
+  \end{wrapfigure}We show how to use \system to infer programs containing both discrete
+structure and continuous parameters. The high-level idea is to write programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
+and optimizing using gradient descent.
+We task our model
+with solving 200 symbolic regression problems, each either a polynomial of
+degree 1 to 4 or a rational function.
+The recognition model is a convolutional network that observes an image
+of the target function's graph (Fig.~\ref{functions}) --
+visually, different kinds of polynomials and rational functions
+produce different kinds of graphs,
+and so the recognition model can learn to
+look at a graph and predict what kind of function best explains it.
 
@@
 
-\subsection{String Editing}\label{textSection}
+\subsection{Programs that manipulate sequences}\label{sequences}
+We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection})  problems.
+For both these domains we initially provide the system with a generic
+minimal set of programming primitives,
+and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
+The initial set of primitives includes \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+
+\subsubsection{String Editing}\label{textSection}
 Synthesizing programs that manipulate strings is a classic problem in the
@@
 expertly crafted to suit string editing.
-We show in this experiment that we can instead start out with a
-less specialized set of primitives -- close to the string methods provided by Python --
+We show \system can instead start out with generic primitives
 and recover many of the higher-level building blocks that have made these
 other system successful.
-Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
-Absent are any routines for searching for characters, extracting substrings,
-or replacing one character with another.
-Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
-basic arithmetic, and functions for splitting a string into a list (\code{split})
-or turning a list into a string (\code{join}).
 We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
 At first, \system cannot find any correct programs for most of the tasks.
-After three wake/sleep iterations,
-it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
+It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
 all of the tasks.
@@
 
-\subsection{List Functions}
+\subsubsection{List Functions}\label{listSection}
 Synthesizing programs that manipulate data structures is

### 2018-05-12 nips_2018.tex
NIPS.pdf

@@
 We induce a DSL by searching locally through the space of DSLs,
-proposing small local changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
+proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
 The search moves work by introducing new
@@
 We propose these new expressions by extracting subexpressions from
-programs already in the frontier.
-These extracted subexpressions
+programs already in the frontiers.
+These subexpressions
 are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
@@
 \subsection{Programs that manipulate sequences}\label{sequences}
-We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection})  problems.
+We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
 For both these domains we initially provide the system with a generic
@@
 and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
-The initial set of primitives includes \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
 
@@
 expertly crafted to suit string editing.
-We show \system can instead start out with generic primitives
+We show \system can instead start out with generic Lisp primitives
 and recover many of the higher-level building blocks that have made these
@@
  \section{Contribution and Outlook}
-We contribute a model, \systemEnding, that learns to program by bootstrapping a DSL with new
+We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
 domain-specific primitives that the algorithm itself discovers,

### 2018-05-13 nips_2018.tex
small fixed to table
wrgpfig

@@
 samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
-enumerate programs rather than sampling them.  Enumeration proceeds by
-a depth-first search over the random choices made by
-Alg.~\ref{programGenerativeModel}; we wrap the depth-first search
-in iterative deepening to build $\lambda$-calculus
-expressions in order of their probability.
-
-\begin{algorithm}[tb]
+enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sampling them.
+\begin{wrapfigure}{L}{0.5\textwidth}
+  \begin{minipage}{0.49\textwidth}    
+\begin{algorithm}[H]
    \caption{Generative model over programs}
@@
    \begin{algorithmic}
-     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
-  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
-  \STATE \textbf{Output:} a program whose type unifies with $\tau$
-  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
-\STATE
+%%      \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
+%%   \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
+%%   \STATE \textbf{Output:} a program whose type unifies with $\tau$
+%%   \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
+%% \STATE
      \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
-  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
+  \STATE {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$
   \STATE \textbf{Output:} a program whose type unifies with $\tau$
@@
    \STATE Unify $\tau$ with yield$(\tau')$.
-   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{argTypes}(\tau')$ 
+   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
 %   \STATE unify$(\tau,\beta)$
@@
  \STATE\textbf{where:}
- \STATE \quad yield$(\tau) = \begin{cases}
+ \STATE yield$(\tau) = \begin{cases}
    \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
@@
  \end{cases}$ 
- \STATE \quad argTypes$(\tau) = \begin{cases}
-   [\alpha] + \text{argTypes}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+ \STATE  args$(\tau) = \begin{cases}
+   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
    []   &\text{ otherwise.}
@@
 \end{algorithm}
+    \end{minipage}
+  \end{wrapfigure}
 
@@
 We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
-or equivalently maximizing
-\begin{equation*}
-  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
-\end{equation*}
+or equivalently maximizing $  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]$,
+%% \begin{equation*}
+%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
+%% \end{equation*}
 where the expectation is taken over tasks. One could take this expectation
@@
     \begin{tabular}{l}
-      \code{(lambda (a b) (fold b (cons "," a) (lambda (x z) (cons x z))))}\\
+      \code{(lambda (a b) (fold b (cons "," a)}\\
+      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
       \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
     \end{tabular}&
-    \code{(fold a b (lambda (x z) (cons x z)))}\\\bottomrule
+    \begin{tabular}{l}
+      \code{(fold a b (lambda (x z) }\\
+      \phantom{\code{(fold a b }}\code{(cons x z)))}
+      \end{tabular}
+\\\bottomrule
   \end{tabular}
@@
 
+Alg.~\ref{mainAlgorithm} describes how we combine program search,
+recognition model training, and DSL induction.
+\begin{comment}
 \subsection{Implementing \system}
@@
 (4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.
-
-\begin{algorithm}[tb]
+\end{comment}
+\begin{wrapfigure}{L}{0.6\textwidth}
+  \begin{minipage}{0.6\textwidth}    
+\begin{algorithm}[H]
    \caption{The \system Algorithm}
@@
      \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
-     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Compile}})
+     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
      \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}
@@
 \subsection{Symbolic Regression}\label{regressionSection}
-\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}
+\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
   \includegraphics[width = \functionSize]{figures/functions/0.png}
@@
   \includegraphics[width = \functionSize]{figures/functions/160.png}
-  \includegraphics[width = \functionSize]{figures/functions/303.png}\\  
-  \includegraphics[width = \functionSize]{figures/functions/402.png}
+  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
+  \includegraphics[width = \functionSize]{figures/functions/432.png}
   \includegraphics[width = \functionSize]{figures/functions/415.png}
@@
 While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
-  \end{wrapfigure}We show how to use \system to infer programs containing both discrete
-structure and continuous parameters. The high-level idea is to write programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
-and optimizing using gradient descent.
-We task our model
-with solving 200 symbolic regression problems, each either a polynomial of
-degree 1 to 4 or a rational function.
-The recognition model is a convolutional network that observes an image
-of the target function's graph (Fig.~\ref{functions}) --
-visually, different kinds of polynomials and rational functions
-produce different kinds of graphs,
-and so the recognition model can learn to
-look at a graph and predict what kind of function best explains it.
-
-We initially give our system addition, multiplication, and division,
-along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
-We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples,
+\end{wrapfigure}
+As a simple first example of how our algorithm works, we apply \system
+to symbolic regression problems.  For each of these problems, the
+agent observes points along the curve of a function, and must write a
+program that fits those points.  We initially equip our learner with
+addition, multiplication, and division, and task it with solving
+200 symbolic regression problems, each either a polynomial of degree
+1--4 or a rational function.  Our recognition model here is a
+convolutional network that observes an image of the target function's
+graph (Fig.~\ref{functions}) -- visually, different kinds of
+polynomials and rational functions produce different kinds of graphs,
+and so the recognition model can learn to look at a graph and predict
+what kind of function best explains it.  A key difficulty, however, is
+that these problems are best solved with programs containing real
+numbers.  Our solution to this difficulty is to allow the system to
+write programs with real-valued parameters, and then fit those
+parameters by automatically differentiating through the programs the
+system writes and use gradient descent.
+We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
 and penalize the use of continuous parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
@@
 
-\system learns a DSL containing templates for polynomials of different orders,
+\system learns a DSL containing templates for polynomials of different orders, as well as
+
 which lets the algorithm quickly find the functions that are most appropriate to this domain (Tbl.~\ref{regressionDSL}).
@@
 This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
-\begin{table}\centering
+\begin{wrapfigure}{l}{0.5\textwidth}\centering
 \begin{tabular}{l}
   \toprule
-  %% Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
-  %% & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
-  %% Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
-  %% \midrule
-  %% Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
-  %% \\\midrule
   $f_0\,=\,$\code{(+ $\mathcal{R}$)}\\
@@
 \caption{Some learned DSL primitives for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
-\end{table}
+\end{wrapfigure}
 
@@
 \multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
-\multicolumn{2}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
+\multicolumn{3}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
 \\

### 2018-05-13 nips_2018.tex
string example structure

@@
       &\hspace{0.5cm} \emph{(a linear function of x)}\\
-   Strings& \code{(join '' (split c s))}\\
-   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
+   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
+   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
    Lists& \code{(map (lambda (x) (+ x k)) l)}\\
@@
 parameters by automatically differentiating through the programs the
-system writes and use gradient descent.
+system writes and use gradient descent to fit the parameters.
 We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
-and penalize the use of continuous parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
+and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
 %% \begin{align*}
@@
 and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
-The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
 
@@
 other system successful.
-We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
+We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
 At first, \system cannot find any correct programs for most of the tasks.

### 2018-05-13 nips_2018.tex
shortened: we might actually make it within the page limit...

@@
 \begin{align}\label{intractableObjectives}
-  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta],\quad
-  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta, \quad
+\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \qquad
   \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
@@
 
-Alg.~\ref{programGenerativeModel} is a procedure for drawing
-samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
-enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sampling them.
 \begin{wrapfigure}{L}{0.5\textwidth}
@@
 \end{algorithm}
-    \end{minipage}
-  \end{wrapfigure}
+\vspace{-1.8cm}    \end{minipage}
+  \end{wrapfigure}Alg.~\ref{programGenerativeModel} is a procedure for drawing
+samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
+enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.
 
@@
 \begin{align}\nonumber
-\mathcal{L}_{\text{HM}} = \expect_{p\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\text{ s.t. $x = \sem{p}$}\quad
+\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
 \mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
@@
   \end{equation}
-
 where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-$\lambda$ is a hyperparameter that acts as a regularizer on the size of the DSL,
+$\lambda$ acts as a regularizer on the size of the DSL,
 and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
@@
 
-To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
+\begin{wrapfigure}{R}{0.6\textwidth}\vspace{-0.7cm}
+  \begin{minipage}{0.6\textwidth}    
+    \begin{algorithm}[H]
+      \caption{The \system Algorithm}
+      \label{mainAlgorithm}
+      \begin{algorithmic}
+        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
+        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
+        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
+        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
+        \FOR{$i=1$ {\bfseries to} $I$}
+        %     \FOR{$x:\tau\in X$}
+        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
+        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
+        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}
+
+        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
+        %% \STATE Define $Q_x(z) \propto \begin{cases}
+        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
+        %%   0&x\not \in \mathcal{F}_x
+        %% \end{cases}
+        \ENDFOR
+        \STATE \textbf{return} $\mathcal{D},\theta,q$
+      \end{algorithmic}
+    \end{algorithm}
+  \end{minipage}
+\vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
 Although this problem may seem 
@@
 for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-a DSL in \system may be context-sensitive due to the presence of variables
+our DSLs are context-sensitive due to the presence of variables
 in the programs and also due to the polymorphic typing system.
@@
 \end{figure*}
-
 Alg.~\ref{mainAlgorithm} describes how we combine program search,
@@
 \end{comment}
-\begin{wrapfigure}{L}{0.6\textwidth}
-  \begin{minipage}{0.6\textwidth}    
-\begin{algorithm}[H]
-   \caption{The \system Algorithm}
-   \label{mainAlgorithm}
-   \begin{algorithmic}
-     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-     \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-     \FOR{$i=1$ {\bfseries to} $I$}
-     %     \FOR{$x:\tau\in X$}
-     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
-     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}
-
-     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
-     %% \STATE Define $Q_x(z) \propto \begin{cases}
-     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-     %%   0&x\not \in \mathcal{F}_x
-     %% \end{cases}
-      \ENDFOR
- \STATE \textbf{return} $\mathcal{D},\theta,q$
-\end{algorithmic}
-\end{algorithm}
-    \end{minipage}
-  \end{wrapfigure}
 

### 2018-05-13 nips_2018.tex
we can do sygus

@@
 and better-tuned DSL.
-Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding)
-because it iterates between three different phases: an
-\textbf{Explore} phase uses the DSL to explore the space of programs,
-searching for ones that solve the tasks; a \textbf{Compress} phase
+Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
+and iterates between three different steps: an
+\textbf{Explore} step uses the DSL to explore the space of programs,
+searching for ones that solve the tasks; a \textbf{Compress} step
 modifies the DSL by discovering regularities in the programs found by
-the previous Explore phase; and a \textbf{Compile} phase, which
+the previous Explore step; and a \textbf{Compile} step, which
 improves the program search procedure by training a neural network to
@@
  symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
- For each of these we provide a generic
- set of programming primitives in a Lisp-like language,
- including conditionals, variables, and higher-order functions.
+ For each of these we initially provide a generic
+ set of programming primitives in a Lisp-like language.
+% including conditionals, variables, and higher-order functions.
  Our algorithm then discovers
@@
     \end{tabular}
-    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
+    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}\vspace{-1cm}
  \end{wrapfigure}
@@
 We extend these ideas into an inference strategy that iterates
-through three steps: an \textbf{Explore} cycle uses the current DSL and
+through three steps: an \textbf{Explore} step uses the current DSL and
 recognition model to search for programs that solve the tasks.  The
-\textbf{Compress} and \textbf{Compile} cycles update the DSL and the
+\textbf{Compress} and \textbf{Compile} steps update the DSL and the
 recognition model, respectively.  Crucially, these steps bootstrap 
@@
 
+\begin{wrapfigure}{r}{0.5\textwidth}\vspace{-0.5}
+  \begin{tikzpicture}
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs};
+    \node[align = center](DSL) at (3,1) {DSL};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};
+
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compression)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+  \caption{\system solves for programs, the DSL, and a neural network (recognition model). Each of these steps iteratively bootstraps off of the others.}  \label{feeding}\vspace{-0.4cm}
+\end{wrapfigure}%
 \\\noindent \textbf{Exploration: Searching for programs.}  Our program search
@@
 to maximally compress (or, raise the prior probability of) these
-programs.  As we solve more tasks, we hone in on richer DSLs that
+programs.  As we solve more tasks, we hone in on DSLs that
 more closely match the domain.
@@
 improves and as search finds more programs, the recognition model gets
-both more data to train on, and better data.
-\begin{wrapfigure}{r}{0.6\textwidth}
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs};
-    \node[align = center](DSL) at (3,1) {DSL};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compression)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
-  \caption{\system solves for programs, the DSL, and a neural network called the \emph{recognition model}.
-Each of these steps bootstrap off of the others in an iterative inference algorithm.}  \label{feeding}
-\end{wrapfigure}
-
+more data to train on, and better data.
+
+\begin{comment}
 Sec.~\ref{mathematicalFraming} frames this 3-step procedure as probabilistic inference.
@@
 Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
-
-\subsection{Probabilistic Framing}\label{mathematicalFraming}
+  \end{comment}
+\subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}
 
@@
 during the compilation phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\propto\probability[x,p|\mathcal{D},\theta}]$,
+posterior over programs, $q(p|x)\propto\probability[p|x,\mathcal{D},\theta}]=\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
   thus amortizing the cost of finding programs with high posterior probability.
@@
 
-Now our goal is to search for programs that solve the tasks.  In this
-work we use the simple search strategy of enumerating programs from
+Now our goal is to search for programs solving the tasks.  We use the simple search strategy of enumerating programs from
 the DSL  in decreasing order of their probability,
@@
 
-The purpose of the recognition model is to accelerate the search for
+The purpose of the recognition model is to amortize the cost of searching for
 programs.  It does this by learning to predict programs which are
@@
 where the expectation is taken over tasks. One could take this expectation
-over the observed empirical distribution of tasks,
+over the empirical distribution of tasks,
 like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
@@
 \vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
-Although this problem may seem 
-very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
+Although this  may seem 
+very similar to estimating the parameters of a probabilistic context free grammar,
 for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
@@
 In the Supplement we derive a tractable MAP estimator for $\theta$.
-
+Putting it all together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+recognition model training, and DSL induction.
 \begin{figure*}
@@
 \end{figure*}
-Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
+
 \begin{comment}
@@
 
-\section{Experiments}
+%\section{Experiments}
 %% We first present our experimental results qualitatively,
@@
 
-\subsection{Symbolic Regression}\label{regressionSection}
+\section{Symbolic Regression Experiments}\label{regressionSection}
 \begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
@@
 
-\system learns a DSL containing templates for polynomials of different orders, as well as
-
-which lets the algorithm quickly find the functions that are most appropriate to this domain (Tbl.~\ref{regressionDSL}).
-We also found that the algorithm discovers programs that minimize the number of continuous degrees of freedom.
+\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
+The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
 For example, it represents the linear function $-x+2$ with the program
-\code{(* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{)} which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
+\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
 which has five continuous parameters.
 This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
-\begin{wrapfigure}{l}{0.5\textwidth}\centering
+\begin{wrapfigure}{l}{0.4\textwidth}\centering
 \begin{tabular}{l}
   \toprule
-  $f_0\,=\,$\code{(+ $\mathcal{R}$)}\\
-  $f_1($\code{x}$)\,=\,$\code{(* (* (+ x $\mathcal{R}$) $\mathcal{R}$) x)} \\
-  $f_2($\code{x}$)\,=\,$\code{(+ $\mathcal{R}$ ($f_1$ x))} \\
-  $f_3($\code{x}$)\,=\,$\code{(* x ($f_2$ x))}\\
-  $f_4($\code{x}$)\,=\,$\code{($f_0$ ($f_3$ x))}\\
-  $f_5($\code{x}$)\,=\,$\code{(* x ($f_4$ x))}\\
-  $f_6($\code{x}$)\,=\,$\code{($f_0$ ($f_4$ x))} (\emph{4th order polynomial})
+  $f_0($\code{x}$)\,=\,$\code{(* real (+ x real))}\\
+  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
   \\\bottomrule
 \end{tabular}
-\caption{Some learned DSL primitives for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
+\caption{Some learned subroutines for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
 \end{wrapfigure}
@@
 
-\subsection{Programs that manipulate sequences}\label{sequences}
+\section{Sequence manipulating programs}\label{sequences}
 We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
@@
 
-\subsubsection{String Editing}\label{textSection}
+\subsection{String Editing}\label{textSection}
 Synthesizing programs that manipulate strings is a classic problem in the
@@
 
+How well does the  learned DSL generalized to real text-editing scenarios?
+We tested, but did not train, our model on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+our system solves 32/108 of the problems with an average search time of 11 minutes.
+After learning,
+our system solves 80/108, and does so much faster,
+solving them in an average of 29 seconds.
+
 \begin{table}[t]
@@
     \end{tabular}
-  \caption{Some learned DSL primitives for string editing}\label{textPrimitives}
+  \caption{Some string editing learned subroutines }\label{textPrimitives}
   \end{table}
@@
 
-\subsubsection{List Functions}\label{listSection}
+\subsection{List Functions}\label{listSection}
 Synthesizing programs that manipulate data structures is
@@
   \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        
-  \includegraphics[width = 3.5cm]{figures/circuitLearningCurve.eps}  
   \caption{Learning curves for \system both with (blue) and without (red) the recognition model as the frontier size is varied (solid/dashed/dotted lines).}\label{learningCurves} 

### 2018-05-13 nips_2018.tex
writing

@@
 
-\title{Learning Reusable Components for Neurally--Guided Bayesian Program Learning}
+\title{Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning}
 
@@
 The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
-For example, it represents the linear function $-x+2$ with the program
+For example, it learns to represent linear functions with the program
 \code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
@@
   $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
+  $f_2($\code{x}$)\,=\,$\code{(/ (/ x real) x)}
   \\\bottomrule
 \end{tabular}
-\caption{Some learned subroutines for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
+\caption{Some learned subroutines for symbolic regression.
+The system starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
 \end{wrapfigure}
@@
 How well does the  learned DSL generalized to real text-editing scenarios?
-We tested, but did not train, our model on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
-our system solves 32/108 of the problems with an average search time of 11 minutes.
+We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+\system solves 32/108 of the problems with an average search time of 11 minutes.
 After learning,
-our system solves 80/108, and does so much faster,
+it solves 80/108, and does so much faster,
 solving them in an average of 29 seconds.
+As of the 2017 SyGuS competition,
+the best-performing algorithm solves 86/108 problems,
+and does so with a
+\emph{different hand-engineered DSL for each problem}.
+Here  we learned a single DSL
+that applied generically to
+all of the tasks,
+and perform comparably to the best
+prior work.
 
@@
 and the agent must learn to solve these complicated problems from only 225 tasks.
+Our data set primarily consists of arithmetic operations upon sequences,
+and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
+\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
+
+A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
+Although we provided \system with some of the standard higher-order functions,
+like \code{fold} and \code{unfold},
+we did not include others.
+When trained on these list functions,
+our system rediscovers
+and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}
 
@@
 
-We evaluate \system starting from two different initial DSLs, \emph{base} and
-\emph{rich}. The base DSL starts with only
-low-level primitives such as \code{mapi}, \code{reducei}, and
-\code{if}, whereas the rich DSL includes some common structure that can be
-built from these such as \code{all}, \code{filter}, \code{slice}, and
-\code{index}.
-Both are capable of solving all tasks supplied by our dataset.
-See Supplement for details on these DSLs.
-
-We found  this domain  difficult to tackle without starting from 
-the rich DSL. Our
-algorithm, like human learners, requires a spectrum of problems ranging from
-easy to hard.
-When starting with the base DSL,
-we found that there were not enough steppingstones in the curriculum to
-get \system off the ground.
+%% We evaluate \system starting from two different initial DSLs, \emph{base} and
+%% \emph{rich}. The base DSL starts with only
+%% low-level primitives such as \code{mapi}, \code{reducei}, and
+%% \code{if}, whereas the rich DSL includes some common structure that can be
+%% built from these such as \code{all}, \code{filter}, \code{slice}, and
+%% \code{index}.
+%% Both are capable of solving all tasks supplied by our dataset.
+%% See Supplement for details on these DSLs.
+
+%% We found  this domain  difficult to tackle without starting from 
+%% the rich DSL. Our
+%% algorithm, like human learners, requires a spectrum of problems ranging from
+%% easy to hard.
+%% When starting with the base DSL,
+%% we found that there were not enough steppingstones in the curriculum to
+%% get \system off the ground.
 
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
-  \% solved &
-  \textbf{84\%} &76\%& 63\%&62\%&62\%\\
-  Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s
+  %% \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
+  %% \% solved &
+  %% \textbf{84\%} &76\%& 63\%&62\%&62\%\\
+  %% Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s
 
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
-  \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
-  Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
-
-  \\\midrule\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
+  %% \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
+  %% \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
+  %% Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
+
+  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
   \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\

### 2018-05-14 nips_2018.tex
editing

@@
 
-Now our goal is to search for programs solving the tasks.  We use the simple search strategy of enumerating programs from
+Now our goal is to search for programs solving the tasks.  We use the simple approach of enumerating programs from
 the DSL  in decreasing order of their probability,
-and then checking if an enumerated program $p$ assigns positive
+and then checking if a program $p$ assigns positive
 probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
@@
 probable under $(\mathcal{D},\theta)$ while also assigning high
-likelihood for a task according to $\probability[x|p]$.  Concretely,
+likelihood to a task according to $\probability[x|p]$.  Concretely,
 the recognition model $q$ is a neural network that predicts, for each
@@
 that allow an agent to easily express solutions to the tasks at hand.
-In the \system algorithm we infer the DSL from a collection of frontiers.
+%In the \system algorithm we infer the DSL from a collection of frontiers.
 Intuitively, we want the algorithm to
@@
 discover more programs.
-
-Recall from Sec.~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
+ Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
 We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
@@
 \end{wrapfigure}
-As a simple first example of how our algorithm works, we apply \system
-to symbolic regression problems.  For each of these problems, the
+As a simple first example, we apply \system
+to symbolic regression problems.  Here, the
 agent observes points along the curve of a function, and must write a
@@
 addition, multiplication, and division, and task it with solving
-200 symbolic regression problems, each either a polynomial of degree
-1--4 or a rational function.  Our recognition model here is a
+100 symbolic regression problems, each either a polynomial of degree
+1--4 or a rational function.  The recognition model is a
 convolutional network that observes an image of the target function's
@@
 We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
-For both these domains we initially provide the system with a generic
-minimal set of programming primitives,
-and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
-The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
+and initially provide the system with a generic
+set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
 
@@
 and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
-However, this prior work presumes a ready-made DSL,
+This prior work presumes a ready-made DSL,
 expertly crafted to suit string editing.
@@
 other system successful.
-We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
+We automatically generated 109 string editing tasks  %% with 4 input/output examples each
+(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
 At first, \system cannot find any correct programs for most of the tasks.

### 2018-05-15 nips_2018.tex
put list functions first

@@
 
-We define a prior distribution over DSLs which penalizes the sizes of the $\lambda$-calculus expressions in the DSL, and put a Dirichlet prior over the weight vector:
-\begin{equation}
-    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)\hspace{1cm}  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)
-  \end{equation}
-where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-$\lambda$ acts as a regularizer on the size of the DSL,
-and $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
+To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+and place a symmetric Dirichlet prior over the weight vector $\theta$.
+%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
+ %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
 %Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
@@
 In the Supplement we derive a tractable MAP estimator for $\theta$.
-Putting it all together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
 recognition model training, and DSL induction.
@@
 
+\section{Sequence manipulating programs}\label{sequences}
+We apply \system to list processing (Section~\ref{textSection}) and text editing (Section~\ref{listSection}).
+For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
+and initially provide the system with a generic
+set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+
+\subsection{List Functions}\label{listSection}
+Synthesizing programs that manipulate data structures is
+a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
+%with applications to computer aided programming~\cite{solar2008program}.
+We consider this problem within the context of
+learning functions that manipulate lists.
+We created 225 Lisp-style list manipulation tasks,
+each with 15 input/output examples (Tbl.~\ref{listexamples}).
+Our data set is challenging along two dimensions:
+many of the functions are very complicated,
+and the agent must learn to solve these complicated problems from only 225 tasks.
+Our data set primarily consists of arithmetic operations upon sequences,
+and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
+\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
+
+A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
+Although we provided \system with some of the standard higher-order functions,
+like \code{fold} and \code{unfold},
+we did not include others.
+When trained on these list functions,
+our system rediscovers
+and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}
+
+%% Our list function domain consists of hand-designed concepts from which we
+%% sample tasks, so there is no ``ground truth'' formal DSL from which these
+%% tasks are generated. Additionally, few of these tasks are concatenative ---
+%% many of them represent distinct operations.
+%% Tasks in this domain are solved by functions that take as input an integer
+%% or a list of integers, and have as output either a Boolean, an integer, a
+%% list of Booleans, or a list of integers.  There are 225 tasks in total, some
+%% of which are shown in Tbl.~\ref{listexamples}.
+
+%% We evaluate \system starting from two different initial DSLs, \emph{base} and
+%% \emph{rich}. The base DSL starts with only
+%% low-level primitives such as \code{mapi}, \code{reducei}, and
+%% \code{if}, whereas the rich DSL includes some common structure that can be
+%% built from these such as \code{all}, \code{filter}, \code{slice}, and
+%% \code{index}.
+%% Both are capable of solving all tasks supplied by our dataset.
+%% See Supplement for details on these DSLs.
+
+%% We found  this domain  difficult to tackle without starting from 
+%% the rich DSL. Our
+%% algorithm, like human learners, requires a spectrum of problems ranging from
+%% easy to hard.
+%% When starting with the base DSL,
+%% we found that there were not enough steppingstones in the curriculum to
+%% get \system off the ground.
+
+%The behavior with the base DSL illustrates this phenomena  because the initial set of
+%primitives is not paired with a sufficiently large and graded curriculum.
+
+\begin{comment}
+\begin{table}
+\centering
+\begin{tabular}{lll}
+  \toprule
+  Name & Input & Output \\\midrule
+  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
+  len & [3\, 5\, 12\, 1] & 4 \\
+  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
+  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
+  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+  count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+  rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
+  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
+  keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
+  \bottomrule
+\end{tabular}
+\caption{Sample tasks from our list function domain %% We used 15 input/output
+  %% examples for each task to reduce ambiguity in the task's function,
+  %% ensuring solutions achieve the desired concept.
+}
+\label{listexamples}
+\end{table}
+
+\begin{table}[t]
+  \centering
+  \begin{tabular}{l}
+    \toprule
+    $f_0($\code{i$,\ell$}$) \,=\, $\code{(singleton (index i $\ell$))}\\
+    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
+    $f_1($\code{i$,\ell$}$) \,=\, $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
+    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
+    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any (lambda (x) (= n x)) $\ell$)}\\
+    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
+    $f_3($\code{i$,\ell$}$) \,=\, $\code{(index (negate i) (sort $\ell$))}\\
+    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
+    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
+    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
+    $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
+    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
+  \bottomrule
+  \end{tabular}
+  \caption{Some learned DSL primitives for list functions}\label{listinduced}
+\end{table}
+
+  \end{comment}
+
+\subsection{String Editing}\label{textSection}
+Synthesizing programs that manipulate strings is a classic problem in the
+programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
+and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
+This prior work presumes a ready-made DSL,
+expertly crafted to suit string editing.
+We show \system can instead start out with generic Lisp primitives
+and recover many of the higher-level building blocks that have made these
+other system successful.
+We automatically generated 109 string editing tasks  %% with 4 input/output examples each
+(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
+At first, \system cannot find any correct programs for most of the tasks.
+It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
+all of the tasks.
+
+How well does the  learned DSL generalized to real text-editing scenarios?
+We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+\system solves 32/108 of the problems with an average search time of 11 minutes.
+After learning,
+it solves 80/108, and does so much faster,
+solving them in an average of 29 seconds.
+As of the 2017 SyGuS competition,
+the best-performing algorithm solves 86/108 problems,
+and does so with a
+\emph{different hand-engineered DSL for each problem}.
+Here  we learned a single DSL
+that applied generically to
+all of the tasks,
+and perform comparably to the best
+prior work.
+
+\begin{table}[t]
+  \centering
+    \begin{tabular}{l}
+      \toprule
+      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
+      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
+        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
+        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
+  \\\bottomrule
+    \end{tabular}
+  \caption{Some string editing learned subroutines }\label{textPrimitives}
+  \end{table}
+
+\begin{figure}
+  \centering\begin{tabular}{ll|lll}\toprule
+    Input&Output&Input 1&Input 2&Output\\\midrule
+Temple Annalisa Haven 185&TAH1&Launa&Withers&Launa Withers\\
+Lara Gregori Bradford&LGB& Rudolf&Akiyama&Rudolf Akiyama
+%DkI$>$DwKI$<$xaPoc& 	DWKI\\
+%KfvTh$>$Dbl$<$OIJso 	&DBL
+\\\midrule
+\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
+\multicolumn{3}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
+\\
+\bottomrule
+  \end{tabular}
+
+  %% \vspace{0.25cm}
+
+  %% \begin{minipage}{\linewidth}
+  %%   \centering
+  %%   $f($\code{s}$) \,=\, $
+  %% \end{minipage}
+
+  \caption{Two string edit tasks (top) and the programs \system writes for them (bottom).
+$f_0$ and $f_2$ are   subroutines written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+\end{figure}
+
 \section{Symbolic Regression Experiments}\label{regressionSection}
@@
 \end{wrapfigure}
-As a simple first example, we apply \system
+We apply \system
 to symbolic regression problems.  Here, the
@@
 
-\section{Sequence manipulating programs}\label{sequences}
-We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
-For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
-and initially provide the system with a generic
-set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
-
-\subsection{String Editing}\label{textSection}
-Synthesizing programs that manipulate strings is a classic problem in the
-programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
-and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
-This prior work presumes a ready-made DSL,
-expertly crafted to suit string editing.
-We show \system can instead start out with generic Lisp primitives
-and recover many of the higher-level building blocks that have made these
-other system successful.
-We automatically generated 109 string editing tasks  %% with 4 input/output examples each
-(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
-At first, \system cannot find any correct programs for most of the tasks.
-It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
-all of the tasks.
-
-How well does the  learned DSL generalized to real text-editing scenarios?
-We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
-\system solves 32/108 of the problems with an average search time of 11 minutes.
-After learning,
-it solves 80/108, and does so much faster,
-solving them in an average of 29 seconds.
-As of the 2017 SyGuS competition,
-the best-performing algorithm solves 86/108 problems,
-and does so with a
-\emph{different hand-engineered DSL for each problem}.
-Here  we learned a single DSL
-that applied generically to
-all of the tasks,
-and perform comparably to the best
-prior work.
-
-\begin{table}[t]
-  \centering
-    \begin{tabular}{l}
-      \toprule
-      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
-      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
-        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
-        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
-  \\\bottomrule
-    \end{tabular}
-  \caption{Some string editing learned subroutines }\label{textPrimitives}
-  \end{table}
-
-\begin{figure}
-  \centering\begin{tabular}{ll|lll}\toprule
-    Input&Output&Input 1&Input 2&Output\\\midrule
-Temple Annalisa Haven 185&TAH1&Launa&Withers&Launa Withers\\
-Lara Gregori Bradford&LGB& Rudolf&Akiyama&Rudolf Akiyama
-%DkI$>$DwKI$<$xaPoc& 	DWKI\\
-%KfvTh$>$Dbl$<$OIJso 	&DBL
-\\\midrule
-\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
-\multicolumn{3}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
-\\
-\bottomrule
-  \end{tabular}
-
-  %% \vspace{0.25cm}
-
-  %% \begin{minipage}{\linewidth}
-  %%   \centering
-  %%   $f($\code{s}$) \,=\, $
-  %% \end{minipage}
-
-  \caption{Two string edit tasks (top) and the programs \system writes for them (bottom).
-$f_0$ and $f_2$ are   subroutines written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
-\end{figure}
-
-\subsection{List Functions}\label{listSection}
-Synthesizing programs that manipulate data structures is
-a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
-%with applications to computer aided programming~\cite{solar2008program}.
-We consider this problem within the context of
-learning functions that manipulate lists.
-We created 225 Lisp-style list manipulation tasks,
-each with 15 input/output examples (Tbl.~\ref{listexamples}).
-Our data set is challenging along two dimensions:
-many of the functions are very complicated,
-and the agent must learn to solve these complicated problems from only 225 tasks.
-Our data set primarily consists of arithmetic operations upon sequences,
-and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
-\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
-
-A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
-Although we provided \system with some of the standard higher-order functions,
-like \code{fold} and \code{unfold},
-we did not include others.
-When trained on these list functions,
-our system rediscovers
-and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}
-
-%% Our list function domain consists of hand-designed concepts from which we
-%% sample tasks, so there is no ``ground truth'' formal DSL from which these
-%% tasks are generated. Additionally, few of these tasks are concatenative ---
-%% many of them represent distinct operations.
-%% Tasks in this domain are solved by functions that take as input an integer
-%% or a list of integers, and have as output either a Boolean, an integer, a
-%% list of Booleans, or a list of integers.  There are 225 tasks in total, some
-%% of which are shown in Tbl.~\ref{listexamples}.
-
-%% We evaluate \system starting from two different initial DSLs, \emph{base} and
-%% \emph{rich}. The base DSL starts with only
-%% low-level primitives such as \code{mapi}, \code{reducei}, and
-%% \code{if}, whereas the rich DSL includes some common structure that can be
-%% built from these such as \code{all}, \code{filter}, \code{slice}, and
-%% \code{index}.
-%% Both are capable of solving all tasks supplied by our dataset.
-%% See Supplement for details on these DSLs.
-
-%% We found  this domain  difficult to tackle without starting from 
-%% the rich DSL. Our
-%% algorithm, like human learners, requires a spectrum of problems ranging from
-%% easy to hard.
-%% When starting with the base DSL,
-%% we found that there were not enough steppingstones in the curriculum to
-%% get \system off the ground.
-
-%The behavior with the base DSL illustrates this phenomena  because the initial set of
-%primitives is not paired with a sufficiently large and graded curriculum.
-
-\begin{table}
-\centering
-\begin{tabular}{lll}
-  \toprule
-  Name & Input & Output \\\midrule
-  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
-  len & [3\, 5\, 12\, 1] & 4 \\
-  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
-  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
-  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-  count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-  rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
-  keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
-  \bottomrule
-\end{tabular}
-\caption{Sample tasks from our list function domain %% We used 15 input/output
-  %% examples for each task to reduce ambiguity in the task's function,
-  %% ensuring solutions achieve the desired concept.
-}
-\label{listexamples}
-\end{table}
-
-\begin{table}[t]
-  \centering
-  \begin{tabular}{l}
-    \toprule
-    $f_0($\code{i$,\ell$}$) \,=\, $\code{(singleton (index i $\ell$))}\\
-    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
-    $f_1($\code{i$,\ell$}$) \,=\, $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
-    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
-    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any (lambda (x) (= n x)) $\ell$)}\\
-    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
-    $f_3($\code{i$,\ell$}$) \,=\, $\code{(index (negate i) (sort $\ell$))}\\
-    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
-    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
-    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
-    $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
-    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
-  \bottomrule
-  \end{tabular}
-  \caption{Some learned DSL primitives for list functions}\label{listinduced}
-\end{table}
-
 \subsection{Quantitative Results}\label{quantitative}

### 2018-05-15 nips_2018.tex
reverse example

@@
    &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
-   Lists& \code{(map (lambda (x) (+ x k)) l)}\\
-   &\hspace{0.5cm} \emph{(add k to every element of list l)}
+   Lists& \code{(fold nil (lambda (a b) (cons a b)))}\\
+   &\hspace{0.5cm} \emph{(reverses a list)}
    \\\bottomrule

### 2018-05-15 nips_2018.tex
shorter

@@
 \\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
-This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
-and is like learning the parameters of a PCFG while not learning any of the structure.
-\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
-This is equivalent to our first wake cycle.
+This is equivalent to \system with $q(x) = \theta$ and $\lambda = \infty$,
+and is like learning the parameters of a PCFG while not learning its structure.
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first exploration cycle.
 
@@
 agent can solve and how quickly it can find those solutions.
-For symbolic regression, we also care about the quality
-of the solution, as measured by the likelihood model $\probability[x|p]$,
-e.g. did the agent correctly explain a linear function
-using two numbers, or did it introduce extraneous parameters?
+%% For symbolic regression, we also care about the quality
+%% of the solution, as measured by the likelihood model $\probability[x|p]$,
+%% e.g. did the agent correctly explain a linear function
+%% using two numbers, or did it introduce extraneous parameters?
 Tbl.~\ref{baselineComparisons}
@@
 and lesioning it also slows down the convergence of the algorithm,
-taking more wake/sleep cycles to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
-This supports the view of the recognition model as a way of accelerating the search over programs.
-
-\begin{table}
+taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
+This supports a view of the recognition model as a way of amortizing the cost of searching for programs.
+
+\begin{wraptable}{r}{0.6\textwidth}
 %% This radically shrinks the table
@@
   Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
-  MDL (nats)&   9.3 &11.0&2.3&--&4.6
+  %% MDL (nats)&   9.3 &11.0&2.3&--&4.6
 
@@
   tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
-  structure learning.  Enum: model w/o any learning.  MDL:
-  $-\expect\left[\probability[x|p] \right]$. For domains other than
-  symbolic regression MDL is 0 nats. For symbolic regression MDL is a
-  proxy for \# of continuous parameters.}\vspace{-0.5cm}\label{baselineComparisons} \end{table}
+  structure learning.  Enum: model w/o any learning.  %% MDL:
+  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
+  %% symbolic regression MDL is 0 nats.
+  %% For symbolic regression MDL is a
+  %% proxy for \# of continuous parameters.
+}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
 

### 2018-05-15 nips_2018.tex
First geom stuff insertion

@@
  \section{Contribution and Outlook}
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+  \begin{center}
+    \includegraphics[width = 0.4\textwidth]{figures/geomCompiled.eps} 
+  \end{center}
+  \caption{
+  Future: dreams.
+  Top left: hand crafted traning targets. Top right: examples of
+  discovered compiled new programs. Bottom: compiled programm across iterations
+  to highlight structure emergence.
+  }\label{geomCompiled} 
+\end{wrapfigure}
+
 We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new

### 2018-05-15 nips_2018.tex
rational data

@@
 other system successful.
+An obstacle here, however,
+is that
+our enumerative search procedure
+has no means of generating string constants,
+and so we incorporate  string-valued parameters as a primitive,
+  defining $\probability[x|p]$ by marginalizing out the
+  values of the string via dynamic programming.
+  In  Section~\ref{regressionSection},
+  we will use a similar trick to synthesize programs containing real numbers using gradient descent.
+
 We automatically generated 109 string editing tasks  %% with 4 input/output examples each
@@
       \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (cdr (if (= c x) s a))))}\\
         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
       $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
-        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
+        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
+        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
+        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
+        %(lambda (lambda (fold $0 $0 (lambda (lambda (if (char-eq? $1 $3) empty (cons $1 $0)))))))
   \\\bottomrule
@@
 \begin{figure}
-  \centering\begin{tabular}{ll|lll}\toprule
-    Input&Output&Input 1&Input 2&Output\\\midrule
-Temple Annalisa Haven 185&TAH1&Launa&Withers&Launa Withers\\
-Lara Gregori Bradford&LGB& Rudolf&Akiyama&Rudolf Akiyama
+  \centering\begin{tabular}{r|r}\toprule
+Temple Annalisa Haven 185 $\longrightarrow$ TAH1&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
+Lara Gregori Bradford $\longrightarrow$ LGB& Andrew Cencici $\longrightarrow$ Dr. Andrew
 %DkI$>$DwKI$<$xaPoc& 	DWKI\\
@@
 \\\midrule
-\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
-\multicolumn{3}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
+\multicolumn{1}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
+\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
 \\
@@
 
-\section{Symbolic Regression Experiments}\label{regressionSection}
-\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
-  \includegraphics[width = \functionSize]{figures/functions/0.png}
-  \includegraphics[width = \functionSize]{figures/functions/6.png}
-  \includegraphics[width = \functionSize]{figures/functions/207.png}
-  \includegraphics[width = \functionSize]{figures/functions/160.png}
-  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
-  \includegraphics[width = \functionSize]{figures/functions/432.png}
-  \includegraphics[width = \functionSize]{figures/functions/415.png}
-  \includegraphics[width = \functionSize]{figures/functions/426.png}
-  \includegraphics[width = \functionSize]{figures/functions/597.png}
-  \includegraphics[width = \functionSize]{figures/functions/591.png}
-  \caption{Recognition model input for symbolic regression.
-While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
-\end{wrapfigure}
+\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
 We apply \system
@@
 and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
+
+\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
+  \includegraphics[width = \functionSize]{figures/functions/0.png}
+  \includegraphics[width = \functionSize]{figures/functions/6.png}
+  \includegraphics[width = \functionSize]{figures/functions/207.png}
+  \includegraphics[width = \functionSize]{figures/functions/160.png}
+  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
+  \includegraphics[width = \functionSize]{figures/functions/432.png}
+  \includegraphics[width = \functionSize]{figures/functions/415.png}
+  \includegraphics[width = \functionSize]{figures/functions/426.png}
+  \includegraphics[width = \functionSize]{figures/functions/597.png}
+  \includegraphics[width = \functionSize]{figures/functions/591.png}
+  \caption{Recognition model input for symbolic regression.
+While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
+\end{wrapfigure}
 %% \begin{align*}
@@
 This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
-\begin{wrapfigure}{l}{0.4\textwidth}\centering
+\begin{wrapfigure}{l}{0.35\textwidth}\centering
 \begin{tabular}{l}
   \toprule
-  $f_0($\code{x}$)\,=\,$\code{(* real (+ x real))}\\
-  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
-  $f_2($\code{x}$)\,=\,$\code{(/ (/ x real) x)}
+  $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
+  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
+  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
+  \hspace{0.5cm}\emph{($f_2$: quadratics)}\\
+  $f_3($\code{x}$)\,=\,$\code{(/ (}$f_2$\code{ x) (}$f_0$\code{ x))}\\
+    \hspace{0.5cm}\emph{($f_3$: ratio of polynomials)}
   \\\bottomrule
@@
 \caption{Some learned subroutines for symbolic regression.
-The system starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
+System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
 \end{wrapfigure}
@@
   \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
-  \% solved&   \textbf{98\% }&94\%&0.7\%&0\%&2\% \\
-  Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
+  \% solved&   \textbf{84\% }&75\%&38\%&38\%&37\% \\
+  Solve time&  24s& 40s  &31s&55s&29s\\
   %% MDL (nats)&   9.3 &11.0&2.3&--&4.6
@@
 \begin{figure}\centering
-\includegraphics[width = 3.5cm]{figures/regression.eps} 
+\includegraphics[width = 3.5cm]{figures/rationalCurve.eps} 
   \includegraphics[width = 3.5cm]{figures/list.eps}

### 2018-05-16 nips_2018.tex
cut a bunch of stuff - premature optimization

@@
 during the compilation phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\propto\probability[p|x,\mathcal{D},\theta}]=\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
+posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta}]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
   thus amortizing the cost of finding programs with high posterior probability.
@@
 used in functional programming languages like OCaml.
-Type variables are always written using lowercase Greek letters
-and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
-and returns something of type $\beta$.
-We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
-has the type $\tau$.
-For example, to describe the type of the identity function
-we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
-We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
-$p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
-a type $\alpha$ with $\tau$ is to introduce constraints on the type
-variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
-See Supplement for more detail on program representation.
-With this notation in hand we now define DSLs:
+%% Type variables are always written using lowercase Greek letters
+%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
+%% and returns something of type $\beta$.
+%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
+%% has the type $\tau$.
+%% For example, to describe the type of the identity function
+%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
+%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
+%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
+%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
+%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
+%% See Supplement for more detail on program representation.
+We now define DSLs:
 
@@
 A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL primitive $e\in \mathcal{D}$, written $\theta_e$,
+one number for each DSL primitive $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of primitive $e$ occurring in a program,
 and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
 
-\begin{wrapfigure}{L}{0.5\textwidth}
-  \begin{minipage}{0.49\textwidth}    
-\begin{algorithm}[H]
-   \caption{Generative model over programs}
-   \label{programGenerativeModel}
-   \begin{algorithmic}
-%%      \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
-%%   \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
-%%   \STATE \textbf{Output:} a program whose type unifies with $\tau$
-%%   \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
-%% \STATE
-     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
-  \STATE {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$
-  \STATE \textbf{Output:} a program whose type unifies with $\tau$
-  \IF{$\tau = \alpha\to\beta$}
-  \STATE var $\gets$ an unused variable name
-  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
-   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
-   \ENDIF
-   %   \ELSE
-   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
-      \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
-
-   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ \\
-      \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
-   \STATE Unify $\tau$ with yield$(\tau')$.
-   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
-%   \STATE unify$(\tau,\beta)$
-   \FOR{$k=1$ {\bfseries to} $K$}
- \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
- \ENDFOR
- \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
- \STATE\textbf{where:}
- \STATE yield$(\tau) = \begin{cases}
-   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-   \tau   &\text{ otherwise.}
- \end{cases}$ 
- \STATE  args$(\tau) = \begin{cases}
-   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-   []   &\text{ otherwise.}
- \end{cases}$
-\end{algorithmic}
-\end{algorithm}
-\vspace{-1.8cm}    \end{minipage}
-  \end{wrapfigure}Alg.~\ref{programGenerativeModel} is a procedure for drawing
-samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
-enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.
+Together with its weight vector,
+a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
+In the supplement, we define this distribution  
+by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$. Care must be taken to ensure that programs are well-typed
+and that variable scoping rules are obeyed.
+With this distribution in hand, we search for programs by enumerating
+$\lambda$-calculus expressions in decreasing order of their probability under
+$(\mathcal{D},\theta)$.
 
@@
   \end{minipage}
-\vspace{-0.7cm}\end{wrapfigure}To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
-Although this  may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-our DSLs are context-sensitive due to the presence of variables
-in the programs and also due to the polymorphic typing system.
-In the Supplement we derive a tractable MAP estimator for $\theta$.
+\vspace{-0.7cm}\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
+%% Although this  may seem 
+%% very similar to estimating the parameters of a probabilistic context free grammar,
+%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+%% our DSLs are context-sensitive due to the presence of variables
+%% in the programs and also due to the polymorphic typing system.
+%% In the Supplement we derive a tractable MAP estimator for $\theta$.
 Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,

### 2018-05-16 supplement.tex
cut a bunch of stuff - premature optimization

@@
 \end{figure}
+
+\section{Generative model over the programs}
+\begin{wrapfigure}{L}{0.5\textwidth}
+  \begin{minipage}{0.49\textwidth}    
+\begin{algorithm}[H]
+   \caption{Generative model over programs}
+   \label{programGenerativeModel}
+   \begin{algorithmic}
+%%      \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
+%%   \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
+%%   \STATE \textbf{Output:} a program whose type unifies with $\tau$
+%%   \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
+%% \STATE
+     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
+  \STATE {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$
+  \STATE \textbf{Output:} a program whose type unifies with $\tau$
+  \IF{$\tau = \alpha\to\beta$}
+  \STATE var $\gets$ an unused variable name
+  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
+   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
+   \ENDIF
+   %   \ELSE
+   \STATE $\text{primitives} \gets\{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$ \\
+      \hspace*{6.9em}$\text{if }\tau\text{ can unify with yield}(\tau') \} $
+
+   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$ \\
+      \hspace*{8.8em}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
+   \STATE Unify $\tau$ with yield$(\tau')$.
+   \STATE $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
+%   \STATE unify$(\tau,\beta)$
+   \FOR{$k=1$ {\bfseries to} $K$}
+ \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
+ \ENDFOR
+ \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
+ \STATE\textbf{where:}
+ \STATE yield$(\tau) = \begin{cases}
+   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+   \tau   &\text{ otherwise.}
+ \end{cases}$ 
+ \STATE  args$(\tau) = \begin{cases}
+   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+   []   &\text{ otherwise.}
+ \end{cases}$
+\end{algorithmic}
+\end{algorithm}
+\vspace{-1.8cm}    \end{minipage}
+  \end{wrapfigure}Alg.~\ref{programGenerativeModel} is a procedure for drawing
+samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
+enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.
 

### 2018-05-16 nips_2018.tex
example task program and DSL at beginning

@@
 
+\begin{tabular}{r|c|}\cmidrule{2-2}
+  Task&
+  \begin{tabular}{rcl}
+  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
+ Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
+ Pushmeet Kohli&$\longrightarrow$&???
+  \end{tabular}\\\cmidrule{2-2}
+  Program&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
+  DSL&\begin{tabular}{l}
+    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y)))}\\
+    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
+    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
+    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
+        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
+    \end{tabular}\\\cmidrule{2-2}
+  \end{tabular}
+
 Automatically inducing programs from examples is a long-standing goal
@@
 
-%% Automatically inducing programs from examples is a long-standing goal
-%% of artificial intelligence.  Recent work in program induction show
-%% that deep networks can be very successful at learning to write certain
-%% kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
-%% DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
-%% techniques (e.g., Metagol:~\cite{muggleton2015meta},
-%% FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
-%% However, the success of these approaches -- both symbolic and neural --
-%% relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
-%% The DSL is an inventory of restricted programming primitives,
-%% and imparts domain specific knowledge about the space of programs.
-%% To what extent can we dispense with hand-engineered DSLs?
-
-%% This work proposes \emph{learning} the DSL.
-%% We consider the setting where we have a collection of related programming tasks,
-%% each specified by a set of input/output examples.
-%% We do not assume that the tasks are annotated with
-%% ground-truth programs.
-%% Our algorithm, called \systemEnding,
-%% does three things:
-%% it finds programs that solve each of the tasks;
-%% it creates its own DSL by discovering and reusing
-%% domain-specific subroutines;
-%% and it learns a neural network that
-%% efficiently searches for programs in the DSL.
-%% The purpose of the DSL
-%% is to expose the commonalities across the programs that solve tasks within a domain,
-%% letting the agent solve other related program learning problems.
-
-%% starting with a Lisp-like representation.
-%% We introduce an algorithm, called the \system,
-%% which creates its own DSL by discovering and then reusing
-%% new useful programming idioms and subroutines.
-
-%% Imagine an agent faced with a new domain of unfamiliar tasks: think
-%% playing new kinds of videogames, authoring Python code, or navigating
-%% through mazes.  Our agent has at its disposal a basis of primitive
-%% actions it can compose to propose solutions to these tasks, but it is
-%% no idea what kinds of primitives are appropriate for which tasks
-%% nor does it know the higher-level vocabulary in which solutions are
-%% best expressed.  How can our agent learn to solve problems in this new
-%% domain?
-
-%% The AI and machine learning literature contains two broad takes on this problem.
-%% The first take is that the agent should come up with a better representation of the space of solutions,
-%% for example, by inventing new primitive actions: see \emph{options} in reinforcement learning~\cite{stolle2002learning}, the EC algorithm in program synthesis~\cite{Dechter:2013:BLV:2540128.2540316}, or predicate invention in inductive logic programming~\cite{muggleton2015meta}.
-%% The second take is that the agent should learn a discriminative model mapping problems to a distribution over solutions: for example, policy gradient methods in reinforcement learning or neural models of program synthesis~\cite{devlin2017robustfill,balog2016deepcoder}.
-%% Our contribution is a general algorithm for fusing these two takes on the problem:
-%% we propose jointly inducing a representation language, called a \emph{Domain Specific Language} (DSL),
-%% alongside a bottom-up discriminative model that regresses from problems to solutions.
-%% We evaluate our algorithm on four domains:
-%% building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and functions on lists.
-%% %Because we want a domain-general
-%% %approach,
-%% We model solutions to tasks as
-%% programs,
-%% which frames our problem as one of program induction.
-
-%% \begin{figure}
-%%   \centering
-
-%%   \begin{tikzpicture}
-
-%%   \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-%%   \node[latent] at (1.5,3) (t){$\theta$};
-%%   \node[latent] at (1,1.75) (z){$p$};
-%% %  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-%%   \node[obs] at (1,0.5) (x) {$x$};
-%%   \edge {z}{x};
-%%   \edge {d,t}{z};
-%%   \plate {}{(z)(x)}{$x\in X$};
-%%   \node[align = center] at ([yshift = -1.3cm]x.south) {(a) \\generative model};
-
-%%   \begin{scope}[shift = {(0.7,0)}]  
-%%   \node[obs] at (3.5,3) (dx){$\mathcal{D}$};
-%%   \node[latent] at (3.5,1.9) (zp){$p$};
-%%   \node[latent] at (5,1.9) (tx){$\theta^{(x)}$};
-%%   \node[obs] at (3.5,0.5) (xp) {$x$};
-%%   \plate {}{(zp)(xp)(tx)}{$x\in X$};
-%%   \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.south) {(b) \\recognition model};
-%%   \draw [->,red] (xp.east) to[out = 0,in = -90] node(nn){} (tx.south);
-%%   \draw [->,red] (tx.west) -- (zp.east);
-%%   \draw [->,red] (dx.south) -- (zp.north);
-
-%%   \node at (nn) {
-%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.15,shift={+(-1,2.5)}}]
-%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-%%       \node[rectangle] at (1,1) {};
-%%       \foreach \name / \y in {1,...,4}
-%%           \node[neuron] (I-\name) at (0,-\y) {};
-%%       \foreach \name / \y in {1,...,3}
-%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
-%%       \foreach \name / \y in {1,...,4}
-%%           \node[neuron] (O-\name) at (2,-\y) {};
-%%       \foreach \source in {1,...,4}
-%%           \foreach \dest in {1,...,3}
-%%               \draw [-latex] (I-\source) -- (H-\dest);
-%%       \foreach \source in {1,...,3}
-%%           \foreach \dest in {1,...,4}
-%%               \draw [-latex] (H-\source) -- (O-\dest);
-%%     \end{tikzpicture}
-%%   };
-%%   \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
-
-%%   \end{scope}
-
-%%   \end{tikzpicture}
-%%   \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives according to probabilities $\theta$.
-From each program we observe a task $x$ (program inputs/outputs).
-(b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
-%%     $\theta^{(x)}$. Black/red arrows correspond to the generative/recognition models.}\label{graphicalModel}
-%% \end{figure}
-
-%% To perform inference, we take inspiration from the Helmholtz machine
-%% and the Wake/Sleep algorithm~\cite{hinton1995wake}, which alternates
-%% between ``waking'', which draws samples from a recognition model (our
-%% $q$); and ``sleeping'', which trains a recognition model on samples
-%% from the generative model (our $\mathcal{D},\theta$).  For us, a
-%% \textbf{Wake} cycle means using the recognition model to search for
-%% programs that solve the tasks.  A sleep cycle means updating these
-%% models so that the DSL and recognition model more closely match the
-%% distribution of tasks.  An important difference from the Helmholtz
-%% machine is that our setup has two distinct sleep phases: a phase where
-%% we update the generative model $(\mathcal{D},\theta)$, which we call
-%% \textbf{Sleep-G}, and a second phase where we update the recognition
-%% model $q$, called \textbf{Sleep-R}.
-
-%% We take the intuitive strategy of exploiting the
-%% conditional independence of the programs conditioned on the DSL, and
-%% so alternate between estimating the DSL ($\mathcal{D},\theta$) and
-%% estimating a posterior distribution over $p$ by searching for
-%% programs.  In practice searching for
-%% programs is difficult, so we use the recognition model to
-%% learn how to search efficiently.
-
 We apply \system to three domains:
@@
 We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
-used in functional programming languages like OCaml.
+used in functional programming languages like OCaml and Haskell.
 %% Type variables are always written using lowercase Greek letters

### 2018-05-16 nips_2018.tex
visual updates to initial figure, tiny edits

@@
 
-\begin{tabular}{r|c|}\cmidrule{2-2}
-  Task&
-  \begin{tabular}{rcl}
-  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
+\begin{figure*}[h]
+  \centering
+  \begin{tabular}{rc}\toprule
+    \sc{Task}& \begin{tabular}{rcl}
+      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
       Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
       Pushmeet Kohli&$\longrightarrow$&???
-  \end{tabular}\\\cmidrule{2-2}
-  Program&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
-  DSL&\begin{tabular}{l}
+    \end{tabular}
+      \\\midrule
+    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
+      \\\midrule
+    \sc{DSL}& \begin{tabular}{l}
       $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y)))}\\
@@
       \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-    \end{tabular}\\\cmidrule{2-2}
     \end{tabular}
+      \\\bottomrule
+  \end{tabular}\right.
+\end{figure*}
 
@@
 can solve relatively quickly -- once they acquire the relevant domain expertise.
-These correspond to short programs -- once you have the right DSL.
-Even with the right DSL, program search may be intractable, so
+These correspond to short programs if you have an expressive DSL.
+Even with such an expressive DSL, program search may be intractable, so
 we amortize the cost of program search by training a neural network to
@@
     \end{tabular}
-    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}\vspace{-1cm}
+    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
+    \vspace{-4em}
  \end{wrapfigure}
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}\vspace{-0.5}
+\begin{wrapfigure}{r}{0.5\textwidth}
   \begin{tikzpicture}

### 2018-05-16 nips_2018.tex
rename fold to foldr

@@
     \sc{DSL}& \begin{tabular}{l}
-      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y)))}\\
+      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y)))}\\
       \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a)}\\
       \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
@@
    &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
-   Lists& \code{(fold nil (lambda (a b) (cons a b)))}\\
+   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
    &\hspace{0.5cm} \emph{(reverses a list)}
@@
     \begin{tabular}{l}
-      \code{(lambda (a b) (fold b (cons "," a)}\\
+      \code{(lambda (a b) (foldr b (cons "," a)}\\
       \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
-      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
+      \code{(lambda (a b) (foldr a b (lambda (x z) (cons x z))))}
     \end{tabular}&
     \begin{tabular}{l}
-      \code{(fold a b (lambda (x z) }\\
-      \phantom{\code{(fold a b }}\code{(cons x z)))}
+      \code{(foldr a b (lambda (x z) }\\
+      \phantom{\code{(foldr a b }}\code{(cons x z)))}
       \end{tabular}
@@
 and initially provide the system with a generic
-set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+set of list processing primitives: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
 
@@
 Although we provided \system with some of the standard higher-order functions,
-like \code{fold} and \code{unfold},
+like \code{foldr} and \code{unfold},
 we did not include others.
@@
       \toprule
-      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
+      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y))))))}\\
       \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (cdr (if (= c x) s a))))}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (cdr (if (= c x) s a))))}\\
         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
@@
         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
-        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
+        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}

### 2018-05-16 nips_2018.tex
reorder experiment presentation

@@
   jointly training a neural network to efficiently search for programs
-  in the learned DSL.  We use our model to solve symbolic regression
-  problems, edit strings, and synthesize functions on lists, showing
-  how the model learns a domain-specific library of program components
-  for expressing solutions to problems in the domain.
+  in the learned DSL.  We use our model to synthesize functions on lists,
+  edit strings, and solve symbolic regression problems, showing how the
+  model learns a domain-specific library of program components for
+  expressing solutions to problems in the domain.
 \end{abstract}
@@
       \\\bottomrule
-  \end{tabular}\right.
+  \end{tabular}
 \end{figure*}
@@
 We apply \system to three domains:
- symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
+Lisp-style functions on lists;
+FlashFill-style~\cite{gulwani2011automating} string processing; and
+symbolic regression;
 For each of these we initially provide a generic
- set of programming primitives in a Lisp-like language.
+set of programming primitives in a Lisp-like language. % TODO: we don't actually give lisp primitives for regression..
 % including conditionals, variables, and higher-order functions.
@@
    %% &\hspace{0.5cm} \emph{(an OR gate)}\\
+   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
+   &\hspace{0.5cm} \emph{(reverses a list)}
+   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
+   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
    Regression& \code{(+ (* real x) real)} \\
       &\hspace{0.5cm} \emph{(a linear function of x)}\\
-   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
-   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
-   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
-   &\hspace{0.5cm} \emph{(reverses a list)}
    \\\bottomrule
@@
 \section{Sequence manipulating programs}\label{sequences}
-We apply \system to list processing (Section~\ref{textSection}) and text editing (Section~\ref{listSection}).
+We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
 For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
@@
 learning functions that manipulate lists.
-We created 225 Lisp-style list manipulation tasks,
+We created 244 Lisp-style list manipulation tasks,
 each with 15 input/output examples (Tbl.~\ref{listexamples}).
@@
 many of the functions are very complicated,
-and the agent must learn to solve these complicated problems from only 225 tasks.
+and the agent must learn to solve these complicated problems from only 244 tasks.
 Our data set primarily consists of arithmetic operations upon sequences,
 and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
-\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
+\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}, \code{2},
+\code{3}, \code{4}, \code{5}.
 
@@
 
+  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
+  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
+  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
+
+  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
+  \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
+  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
+
+  %% \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
+  %% \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
+  %% Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
+
   \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
-  \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
-  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
-
-  %% \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
-  %% \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
-  %% Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
-
-  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
-  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
-  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
   \\\bottomrule

### 2018-05-16 nips_2018.tex
name introduction

@@
 
-\begin{figure*}[h]
-  \centering
-  \begin{tabular}{rc}\toprule
-    \sc{Task}& \begin{tabular}{rcl}
-      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
-      Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
+Imagine that you are asked to edit some text,
+and told that you should change the text ``Nancy FreeHafer'' to ``Dr. Nancy''.
+From this example,
+you likely infer that ``Pushmeet Kohli'' should be changed to ``Dr. Pushmeet'',
+drawing upon your prior knowledge of text,
+like that words are separated by spaces and that
+one commonly prepends titles like ``Dr.''
+In this work,
+we consider the problem of
+building agents that solve few-shot learning tasks like these,
+and also the problem of acquiring
+the prior knowledge necessary to quickly
+solve these tasks (Figure~\ref{firstPageFigure}).
+We think of
+solutions to these tasks as being best represented by programs,
+and so our problem can be stated as follows:
+how should an agent learn to write programs?
+We take inspiration from two sources:
+(1) Good software engineers compose libraries
+of reusable subroutines
+that are shared across
+related programming tasks.
+Returning to Figure~\ref{firstPageFigure},
+a good string processing library should support appending strings and splitting on spaces --
+exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
+(2) Skilled human programmers
+can quickly recognize what kinds of programming idioms and library routines would be useful for solving
+the task at hand, even if they cannot instantly work out the details.
+\begin{figure}[h]
+  \begin{minipage}[c]{0.68\textwidth}
+\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
+  \sc{Task}&
+  \begin{tabular}{rcl}
+  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
+% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
  Pushmeet Kohli&$\longrightarrow$&???
-    \end{tabular}
-      \\\midrule
-    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
-      \\\midrule
-    \sc{DSL}& \begin{tabular}{l}
-      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y)))}\\
+  \end{tabular}\\\cmidrule{2-2}
+  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
+  \begin{tabular}{c}
+    \sc{Library}\\
+    \sc{(DSL)}
+    \end{tabular}&\begin{tabular}{l}
+    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
+    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
     \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a)}\\
+    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
     \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
         \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
+    \end{tabular}\\\cmidrule{2-2}
   \end{tabular}
-      \\\bottomrule
-  \end{tabular}
-\end{figure*}
-
-Automatically inducing programs from examples is a long-standing goal
-of artificial intelligence. Recent work has successfully used
-symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
-FlashFill:~\cite{gulwani2011automating}), neural networks trained from
-a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
-synthesize programs for task domains such as string transformations,
-list processing, and robot navigation and planning. However, all these
-approaches -- symbolic, neural and neural-symbolic -- rely upon a
-hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
-inventory of restricted programming primitives, encoding domain-specific
-knowledge about the space of programs. In
-practice
-we often have only a few input/output examples for each
-program to be induced, and thus success often hinges on having a good
-DSL that provides a crucial inductive bias for what would otherwise be an
-unconstrained search through the space of all computable functions.
-Here we ask, to what extent can we dispense with such highly
-hand-engineered domain-specific languages?
-
-We propose \emph{learning} the DSL by inducing a library of domain--specific
-subroutines. We consider the setting where we have a collection of
-related programming tasks, each specified by a set of input/output
-examples. %% We do not assume that the tasks are annotated with
-%% ground-truth programs.
-Starting from a weaker or more
-general library of primitives, we give an algorithm for constructing a richer, more powerful,
-and better-tuned DSL.
+  \end{minipage}%
+  \begin{minipage}[c]{0.31\textwidth}
+      \caption{\textbf{Task}: Few-shot learning problem.
+Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}.
+Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
+    \end{minipage}
+\end{figure}
+We combine these two ideas into
+a new algorithm called \systemEnding,
+which takes as input a collection of
+programming \textbf{tasks},
+and then jointly
+solves three problems:
+(1) Writing programs that solve the tasks;
+(2) Composing a library of
+domain-specific
+subroutines -- which allow the agent to
+more compactly write programs in the domain,
+and (3) Training a neural network
+to recognize
+which library components are
+useful for which kinds of tasks.
+Together, the library and neural net
+encode the domain specific knowledge needed to
+quickly write programs.
+
+We call the learned library of subroutines a \textbf{Domain-Specific
+  Language} (\textbf{DSL}), a term widely used in the program
+synthesis community~\cite{polozov2015flashmeta}.
+Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
+both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
+FlashFill:~\cite{gulwani2011automating}),
+neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
+symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
+The contribution of this work is
+an algorithm for learning DSLs while jointly training a neural net to
+search for programs in the DSL.
+
+%% synthesize programs for task domains such as string transformations,
+%% list processing, and robot navigation and planning. However, all these
+%% approaches -- symbolic, neural and neural-symbolic -- rely upon a
+%% hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
+%% inventory of restricted programming primitives, encoding domain-specific
+%% knowledge about the space of programs. In
+%% practice
+%% we often have only a few input/output examples for each
+%% program to be induced, and thus success often hinges on having a good
+%% DSL that provides a crucial inductive bias for what would otherwise be an
+%% unconstrained search through the space of all computable functions.
+%% Here we ask, to what extent can we dispense with such highly
+%% hand-engineered domain-specific languages?
+
+%% We propose \emph{learning} the DSL by inducing a library of domain--specific
+%% subroutines. We consider the setting where we have a collection of
+%% related programming tasks, each specified by a set of input/output
+%% examples. %% We do not assume that the tasks are annotated with
+%% %% ground-truth programs.
+%% Starting from a weaker or more
+%% general library of primitives, we give an algorithm for constructing a richer, more powerful,
+%% and better-tuned DSL.
+
 Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
-and iterates between three different steps: an
+because it iterates between three different steps: an
 \textbf{Explore} step uses the DSL to explore the space of programs,
@@
 
-We apply \system to three domains:
-Lisp-style functions on lists;
-FlashFill-style~\cite{gulwani2011automating} string processing; and
-symbolic regression;
+We apply \system to four domains:
+list processing; FlashFill-style~\cite{gulwani2011automating} string editing;  symbolic regression; and turtle graphics~\cite{turtle}.
  For each of these we initially provide a generic
-set of programming primitives in a Lisp-like language. % TODO: we don't actually give lisp primitives for regression..
-% including conditionals, variables, and higher-order functions.
+ set of programming primitives.
 Our algorithm then discovers
@@
    Domain&Part of the learned DSL\\\midrule
-   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
-   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
    Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
-   &\hspace{0.5cm} \emph{(reverses a list)}
+   &\hspace{0.5cm} \emph{(reverses a list)}\\
    Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
@@
   \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
-  \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
-  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
+  \% solved&\textbf{75\%} &\% &33\%&0\%&30\%\\
+  Solve time&  29s&s&80s&--&
 

### 2018-05-16 nips_2018.tex
better introduction figures

@@
 
+Because any model may be encoded as a (deterministic or probabilistic) program,
+we carefully delineate the scope of program learning problems considered here.
+We think of \system as learning to solve the kinds of problems that humans
+can solve relatively quickly -- once they acquire the relevant domain expertise.
+These correspond to short programs -- if you have an expressive DSL.
+Even with a good DSL, program search may be intractable, so
+we amortize the cost of program search by training a neural network to
+assist the search procedure.
+
 Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
@@
 searching for ones that solve the tasks; a \textbf{Compress} step
-modifies the DSL by discovering regularities in the programs found by
+modifies the structure of the DSL by discovering regularities across programs found by
 the previous Explore step; and a \textbf{Compile} step, which
@@
 
-Because any model may be encoded as a (deterministic or probabilistic) program,
-we carefully delineate the scope of program learning problems considered here.
-We think of \system as learning to solve the kinds of problems that humans
-can solve relatively quickly -- once they acquire the relevant domain expertise.
-These correspond to short programs if you have an expressive DSL.
-Even with such an expressive DSL, program search may be intractable, so
-we amortize the cost of program search by training a neural network to
-assist the search procedure.
-
 We apply \system to four domains:
@@
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
-\begin{wrapfigure}{r}{0.6\textwidth}
-    \begin{tabular}{ll}
+\begin{figure}
+  \begin{tabular}{lcl}
     \toprule
-   Domain&Part of the learned DSL\\\midrule
-   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
-   &\hspace{0.5cm} \emph{(reverses a list)}\\
-   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
-   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
-   Regression& \code{(+ (* real x) real)} \\
-      &\hspace{0.5cm} \emph{(a linear function of x)}\\
+    Domain&Example Task&Part of the learned DSL\\\midrule
+    Lists&
+    \begin{tabular}{rcl}
+      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
+      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
+    \end{tabular}
+    &
+    \begin{tabular}{l}
+      \code{(foldr nil (lambda (a b) (cons a b)))}\\
+          \hspace{0.5cm} \emph{(appends lists)}
+      \end{tabular}
+
+    \\    Strings&
+    \begin{tabular}{rcl}
+      Temple Anna H& $\longrightarrow$& TAH\\
+      Lara Gregori&$\longrightarrow$&LG
+      \end{tabular}
+    &
+    \begin{tabular}{l}
+      \code{(map (lambda (x) (if (= x a) b x)))}\\
+      \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
+      \end{tabular}
+
+    \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
+      \code{(+ (* real x) real)} \\
+          \hspace{0.5cm} \emph{(a linear function of x)}
+      \end{tabular}
+
     \\\bottomrule
@@
   \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
-    \vspace{-4em}
- \end{wrapfigure}
+%  \vspace{-4em}
+\end{figure}
 
@@
 \textbf{Compress} and \textbf{Compile} steps update the DSL and the
-recognition model, respectively.  Crucially, these steps bootstrap 
-off each other (Fig.~\ref{feeding}):
-
+recognition model, respectively.  Crucially, these steps synergistically bootstrap 
+off each other:
+\begin{comment}
   \begin{wrapfigure}{r}{0.5\textwidth}
@@
   \caption{\system solves for programs, the DSL, and a neural network (recognition model). Each of these steps iteratively bootstraps off of the others.}  \label{feeding}\vspace{-0.4cm}
-\end{wrapfigure}%
+\end{wrapfigure}
+  \end{comment}
 \\\noindent \textbf{Exploration: Searching for programs.}  Our program search
 is informed by both the DSL and the recognition model. When these
-improve, we find more programs solving the tasks.
+improve, we can solve more tasks.
 \\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
@@
 
-\system takes as input a set of \emph{tasks}, written $X$, each of which is a program induction problem.
-It has at its disposal a \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
+\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
+It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.\footnote{For example, for string editing,
+  the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
+and 0 otherwise.}
 Its goal is to solve each of the tasks by writing a program,

### 2018-05-16 nips_2018.tex
recognition model section

@@
 We think of
-solutions to these tasks as being best represented by programs,
+solutions to these tasks as being well represented by programs,
 and so our problem can be stated as follows:
@@
 \textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
-  frontiers.}  Here we train a neural network, $q$, to predict a
-distribution over programs conditioned on a task. The objective of $q$
-is to assign high probability to programs $p$ where
+  frontiers.}  Here we train %% a neural network, $q$, to predict a
+%% distribution over programs conditioned on a task. The objective of $q$
+%% is
+$q(p|x)$ to assign high probability to programs $p$ where
 $\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
@@
 A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL primitive $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of primitive $e$ occurring in a program,
+one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
 and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
@@
 
-The purpose of the recognition model is to amortize the cost of searching for
-programs.  It does this by learning to predict programs which are
-probable under $(\mathcal{D},\theta)$ while also assigning high
-likelihood to a task according to $\probability[x|p]$.  Concretely,
-the recognition model $q$ is a neural network that predicts, for each
+The purpose of training the recognition model is to amortize the cost of searching for
+programs.  It does this by learning to predict, for each task,
+programs with high likelihood according to 
+$\probability[x|p]$ while also being
+probable under the prior $(\mathcal{D},\theta)$.  Concretely,
+the recognition model $q$ predicts, for each
 task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
@@
 
-We want a recognition model that closely approximates the true posteriors over programs.
-We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
-or equivalently maximizing $  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]$,
+How should we get the data to train $q$?
+This is nonobvious  because \system is only weakly supervised (i.e., learns only from tasks and not from (program, task) pairs).
+One approach is to sample programs from the DSL,
+run them to get their input/outputs,
+and then train $q$ to predict the program from the input/outputs.
+This approach is like how a Helmholtz machine
+trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
+The advantage of ``Helmholtz machine'' training is that
+we can draw unlimited samples from the DSL,
+training on a large amount of data.
+Another approach is
+ self-supervised learning,
+training $q$ on the (program, task)
+pairs discovered by the algorithm so far.
+The advantage of self-supervised learning
+is that the training data is much higher quality,
+because we are training on the actual tasks.
+Due to these complementary advantages,
+we train on both these sources of data.
+
+Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
+ equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
 %% \begin{equation*}
@@
 %% \end{equation*}
-where the expectation is taken over tasks. One could take this expectation
-over the empirical distribution of tasks,
-like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
-We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
+ where the expectation is taken over tasks.
+Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
+ The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
 \begin{align}\nonumber
 \mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
-\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
   \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
 \end{align}
+\begin{comment}
   The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
@@
 and train $q$ on those samples.
-
+  \end{comment}
 Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from

### 2018-05-16 nips_2018.tex
different introduction

@@
 
-Imagine that you are asked to edit some text,
-and told that you should change the text ``Nancy FreeHafer'' to ``Dr. Nancy''.
-From this example,
-you likely infer that ``Pushmeet Kohli'' should be changed to ``Dr. Pushmeet'',
-drawing upon your prior knowledge of text,
-like that words are separated by spaces and that
-one commonly prepends titles like ``Dr.''
+Imagine that you are asked to write a program to edit some text, and
+told that you should change the text ``Nancy FreeHafer'' to
+``Dr. Nancy''.  From this example, you likely infer that ``Jane
+Goodall'' should be ``Dr. Jane'', drawing upon your prior knowledge of
+text, like that words are separated by spaces. For few-shot program learning problems
+like these, the program synthesis and AI communities have had great
+success engineering program learning systems,
+not just for text editing~\cite{gulwani2011automating} but also for graphics
+programs~\cite{ellis2017learning}, planning
+procedures~\cite{devlin2017neural}, and many others.  However, the
+success of these systems hinges upon having a carefully
+hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
+impart prior knowledge of a domain
+by providing a restricted set
+of finely-tuned programming primitives: for
+%that are finely tuned to the problems in the domain: for
+text editing,
+these are primitives like appending and splitting on characters.
 In this work,
 we consider the problem of
-building agents that solve few-shot learning tasks like these,
+building agents that solve program learning tasks,
 and also the problem of acquiring
@@
 solve these tasks (Figure~\ref{firstPageFigure}).
-We think of
-solutions to these tasks as being well represented by programs,
-and so our problem can be stated as follows:
-how should an agent learn to write programs?
-We take inspiration from two sources:
-(1) Good software engineers compose libraries
-of reusable subroutines
-that are shared across
-related programming tasks.
-Returning to Figure~\ref{firstPageFigure},
-a good string processing library should support appending strings and splitting on spaces --
-exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
-(2) Skilled human programmers
-can quickly recognize what kinds of programming idioms and library routines would be useful for solving
-the task at hand, even if they cannot instantly work out the details.
+Our solution is an algorithm that learns a DSL
+while jointly training a neural network to write programs in that learned DSL.
 \begin{figure}[h]
@@
 % Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
- Pushmeet Kohli&$\longrightarrow$&???
+ Jane Goodall&$\longrightarrow$&???
   \end{tabular}\\\cmidrule{2-2}
@@
 \end{figure}
+
+%% We think of
+%% solutions to these tasks as being well represented by programs,
+%% and so our problem can be stated as follows:
+%% how should an agent learn to write programs?
+We take inspiration from two sources:
+(1) Good software engineers compose libraries
+of reusable subroutines
+that are shared across
+related programming tasks.
+Returning to Figure~\ref{firstPageFigure},
+a good string processing library should support appending strings and splitting on spaces --
+exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
+(2) Skilled human programmers
+can quickly recognize what kinds of programming idioms and library routines would be useful for solving
+the task at hand, even if they cannot instantly work out the details.
 We combine these two ideas into
-a new algorithm called \systemEnding,
+an algorithm called \systemEnding,
 which takes as input a collection of
@@
 
-We call the learned library of subroutines a \textbf{Domain-Specific
-  Language} (\textbf{DSL}), a term widely used in the program
-synthesis community~\cite{polozov2015flashmeta}.
-Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
-both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
-FlashFill:~\cite{gulwani2011automating}),
-neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
-The contribution of this work is
-an algorithm for learning DSLs while jointly training a neural net to
-search for programs in the DSL.
-
 %% synthesize programs for task domains such as string transformations,
@@
 
-Because any model may be encoded as a (deterministic or probabilistic) program,
-we carefully delineate the scope of program learning problems considered here.
-We think of \system as learning to solve the kinds of problems that humans
-can solve relatively quickly -- once they acquire the relevant domain expertise.
-These correspond to short programs -- if you have an expressive DSL.
-Even with a good DSL, program search may be intractable, so
-we amortize the cost of program search by training a neural network to
-assist the search procedure.
-
 Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
 because it iterates between three different steps: an
-\textbf{Explore} step uses the DSL to explore the space of programs,
+\textbf{Explore} step uses the DSL and neural network to explore the space of programs,
 searching for ones that solve the tasks; a \textbf{Compress} step
@@
 
+Because any model may be encoded as a (deterministic or probabilistic) program,
+we carefully delineate the scope of problems considered here.
+We think of \system as learning to solve the kinds of problems that humans
+can solve relatively quickly -- once they acquire the relevant domain expertise.
+These correspond to short programs -- if you have an expressive DSL.
+Even with a good DSL, program search may be intractable, so
+we amortize the cost of program search by training a neural network to
+assist the search procedure.
+
 We apply \system to four domains:
-list processing; FlashFill-style~\cite{gulwani2011automating} string editing;  symbolic regression; and turtle graphics~\cite{turtle}.
+list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
  For each of these we initially provide a generic
@@
 Our algorithm then discovers
-its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
 \begin{figure}
@@
 \end{figure}
+
+Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
+both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
+FlashFill:~\cite{gulwani2011automating}),
+neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
+symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
+A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
+which also learns a library of subroutines.
+We were inspired by EC,
+and go beyond it by giving a new algorithm for learning DSLs, as well as a way of combining DSL learning with
+neurally guided program search.
+In the experiments section,
+we compared directly with EC,
+showing empirical improvements.
+The contribution of this work is
+then
+an effective algorithm for learning DSLs which also trains a neural net to
+search for programs in the DSL.
 
@@
 $\lambda$-calculus is a formalism for expressing functional programs
-that closely resembles the Lisp programming language.
-$\lambda$-calculus includes variables, function application, and the ability to create new functions.
+that closely resembles Lisp,
+including variables, function application, and the ability to create new functions.
 Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.

### 2018-05-17 nips_2018.tex
Tweaking figure - feedback: is it too much?
big table at the beginning

@@
 ``Dr. Nancy''.  From this example, you likely infer that ``Jane
-Goodall'' should be ``Dr. Jane'', drawing upon your prior knowledge of
+Goodall'' goes to ``Dr. Jane'', drawing upon your prior knowledge of
 text, like that words are separated by spaces. For few-shot program learning problems
-like these, the program synthesis and AI communities have had great
-success engineering program learning systems,
+like these, the program synthesis and AI communities have had many successes,
 not just for text editing~\cite{gulwani2011automating} but also for graphics
@@
 Our solution is an algorithm that learns a DSL
-while jointly training a neural network to write programs in that learned DSL.
+while jointly training a neural network to help write programs in the learned DSL.
 \begin{figure}[h]
@@
   \end{tabular}\\\cmidrule{2-2}
-  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
-  \begin{tabular}{c}
-    \sc{Library}\\
-    \sc{(DSL)}
-    \end{tabular}&\begin{tabular}{l}
-    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
+  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
+    \sc{DSL}
+&\begin{tabular}{l}
+    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y)}\\
     \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
     \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
+    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s ($\lambda$ (x a)}\\
     \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
@@
       \caption{\textbf{Task}: Few-shot learning problem.
-Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}.
+Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}.
 Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
@@
 (1) Writing programs that solve the tasks;
-(2) Composing a library of
+(2) Composing a library (DSL) of
 domain-specific
@@
 to recognize
-which library components are
+which DSL components are
 useful for which kinds of tasks.
-Together, the library and neural net
+Together, the DSL and neural net
 encode the domain specific knowledge needed to
@@
     \begin{tabular}{l}
-      \code{(foldr nil (lambda (a b) (cons a b)))}\\
+      \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
           \hspace{0.5cm} \emph{(appends lists)}
@@
     \begin{tabular}{l}
-      \code{(map (lambda (x) (if (= x a) b x)))}\\
+      \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
       \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
@@
 \end{figure}
+
+\begin{tabular}{ccc}\toprule
+  \begin{tabular}{c}
+    \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
+      \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
+      \code{$f(x) = $}\code{($f_0$ nil)}\\\\
+      \code{[99\, 4\, 2]}$\to $\code{99}\\
+      \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
+      \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
+      \code{($f_1$ )}
+  \end{tabular}&
+  \begin{tabular}{c}
+    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
+    +83 973-831$\to $83.973.831\\
+    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} ($f_0$ \codechar{.} \codechar{ } (cdr s)))}    \\\\
+Temple Anna H $\to $TAH\\
+Lara Gregori$\to $LG\\
+$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\
+
+  \end{tabular}&
+
+  \begin{tabular}{c}
+    \includegraphics[width = 1cm]{figures/functions/4.png}\\
+    \code{$f($x$) = $($f_1$ -0.8 1.6)}\\
+    \includegraphics[width = 1cm]{figures/functions/112.png}\\
+    \code{$f($x$) = $($f_0$ -0.8 1.6)}\\
+
+    \end{tabular}
+
+  \\\midrule&
+
+  \begin{tabular}{l}
+          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
+      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
+        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
+        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
+    \end{tabular}&
+
+  \begin{tabular}{l}
+    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\quad
+  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
+  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
+    \end{tabular}
+
+      \\\bottomrule
+
+  \end{tabular}
 
@@
 then
-an effective algorithm for learning DSLs which also trains a neural net to
+an effective algorithm for learning DSLs which jointly trains a neural net to
 search for programs in the DSL.
@@
 \textbf{Compress} and \textbf{Compile} steps update the DSL and the
-recognition model, respectively.  Crucially, these steps synergistically bootstrap 
-off each other:
+recognition model, respectively.  Crucially, these steps bootstrap 
+each other:
 \begin{comment}
@@
 \system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.\footnote{For example, for string editing,
-  the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
-and 0 otherwise.}
+It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.%% \footnote{For example, for string editing,
+%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
+%% and 0 otherwise.}
 Its goal is to solve each of the tasks by writing a program,
@@
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
+$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
 is large.
@@
 $q(p|x)$ to assign high probability to programs $p$ where
-$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
+$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
 in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
@@
 In the supplement, we define this distribution  
-by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$. Care must be taken to ensure that programs are well-typed
-and that variable scoping rules are obeyed.
-With this distribution in hand, we search for programs by enumerating
-$\lambda$-calculus expressions in decreasing order of their probability under
-$(\mathcal{D},\theta)$.
+by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$. %%  Care must be taken to ensure that programs are well-typed
+%% and that variable scoping rules are obeyed.
+%% With this distribution in hand, we search for programs by enumerating
+%% $\lambda$-calculus expressions in decreasing order of their probability under
+%% $(\mathcal{D},\theta)$.
 
@@
 
-  A drawback of using an enumerative search algorithm is that we have no
-efficient means of solving for arbitrary constants that might occur in the
+  However, a drawback of   enumerative search  is that we have no
+efficient means of solving for arbitrary constants that might occur in a
 program. In Sec.~\ref{regressionSection},
@@
 $\lambda$-expressions into the DSL.
-We propose these new expressions by extracting subexpressions from
-programs already in the frontiers.
-These subexpressions
-are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
-which then become new functions in the DSL.
-The idea of storing and reusing
-fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.
+We propose these new expressions by extracting fragments of
+programs already in the frontiers (Fig.~\ref{fragmentExample}).
+An important point here is that we are \emph{not} simply adding
+subexpressions of programs to $\mathcal{D}$.  Instead, we are
+extracting fragments that unify with programs in the frontiers.  This
+idea of storing and reusing fragments of expressions comes from
+Fragment Grammars~\cite{tim} and Tree-Substitution
+Grammars~\cite{cohn2010inducing}, and is closely related to the idea
+of antiunification~\cite{henderson2013cumulative}.
 
@@
 recognition model training, and DSL induction.
-\begin{figure*}
+\begin{figure*}[b]
 \centering  \begin{tabular}{ll}
     \toprule
-    Example programs in frontiers&Proposed subexpression\\\midrule
+    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
     \begin{tabular}{l}
@@
       \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
-      \code{(lambda (a b) (foldr a b (lambda (x z) (cons x z))))}
+      \code{(lambda (a b) (foldr a b }\\
+      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}
     \end{tabular}&
@@
       \phantom{\code{(foldr a b }}\code{(cons x z)))}
+    \end{tabular}\\\midrule
+    \begin{tabular}{l}
+      \code{(lambda (s) (map (lambda (x)}\\
+      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
+      \code{(lambda (s) (map (lambda (x) }\\
+      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
+      \end{tabular}&
+    \begin{tabular}{l}      \code{(lambda (s) (map (lambda (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
       \end{tabular}
@@
   \end{tabular}
-  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
-    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}).
-Here, the proposed subexpression appends two lists.}
+  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
+    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
   \label{fragmentExample}
@@
 \subsection{String Editing}\label{textSection}
-Synthesizing programs that manipulate strings is a classic problem in the
+Synthesizing programs that edit strings is a classic problem in the
 programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
 and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
-This prior work presumes a ready-made DSL,
-expertly crafted to suit string editing.
+This prior work presumes a hand-engineered DSL.
 We show \system can instead start out with generic Lisp primitives
 and recover many of the higher-level building blocks that have made these
-other system successful.
-An obstacle here, however,
-is that
-our enumerative search procedure
-has no means of generating string constants,
-and so we incorporate  string-valued parameters as a primitive,
-  defining $\probability[x|p]$ by marginalizing out the
-  values of the string via dynamic programming.
-  In  Section~\ref{regressionSection},
-  we will use a similar trick to synthesize programs containing real numbers using gradient descent.
+other systems successful.
+
+Because our enumerative search procedure has no means of generating
+string constants, we have the enumerator propose programs with string-valued holes --
+e.g., in Fig.~\ref{firstPageFigure},
+we enumerate $\text{\code{(}}f_0\code{ string (}f_2\text{\code{ s ' '))}}$ --
+and define $\probability[x|p]$ by marginalizing out the values of the
+strings via dynamic programming.  In Section~\ref{regressionSection},
+we will use a similar trick to synthesize programs containing real
+numbers, but using gradient descent instead of dynamic programming.
 
 We automatically generated 109 string editing tasks  %% with 4 input/output examples each
-(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
+(Fig.~\ref{exampleTextProblem}).
 At first, \system cannot find any correct programs for most of the tasks.
-It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
+After three iterations, it assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
 all of the tasks.
@@
 We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
-\system solves 32/108 of the problems with an average search time of 11 minutes.
+\system solves 3.7\% of the problems with an average search time of 235 seconds.
 After learning,
-it solves 80/108, and does so much faster,
+it solves 74.1\%, and does so much faster,
 solving them in an average of 29 seconds.
 As of the 2017 SyGuS competition,
-the best-performing algorithm solves 86/108 problems,
-and does so with a
-\emph{different hand-engineered DSL for each problem}.
+the best-performing algorithm solves 79.6\% problems.
+But, SyGuS comes with a
+different hand-engineered DSL \emph{for each text editing problem}.
 Here  we learned a single DSL
@@
       \toprule
-      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y))))))}\\
-      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
+      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map (lambda (x) (if (= x a) b x)) s)}\\
+      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
       $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (cdr (if (= c x) s a))))}\\
@@
       $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
-        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
+        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
         $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
@@
 
-\begin{figure}
-  \centering\begin{tabular}{r|r}\toprule
-Temple Annalisa Haven 185 $\longrightarrow$ TAH1&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
-Lara Gregori Bradford $\longrightarrow$ LGB& Andrew Cencici $\longrightarrow$ Dr. Andrew
+\begin{wrapfigure}{r}{0.6\textwidth}
+  \centering\begin{tabular}{c}\toprule
++106 769-858-438 $\longrightarrow$106.769.858.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
++83 973-757-831 $\longrightarrow$83.973.757.831%& Andrew Cencici $\longrightarrow$ Dr. Andrew
 %DkI$>$DwKI$<$xaPoc& 	DWKI\\
@@
 \\\midrule
-\multicolumn{1}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
-\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
+%(lambda (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' SPACE (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' '-' (cdr $0))))
+\multicolumn{1}{c}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  '.' ' ' ($f_0$ '.' '-' (cdr s)))}}
+%\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
 \\
@@
 
-  \caption{Two string edit tasks (top) and the programs \system writes for them (bottom).
-$f_0$ and $f_2$ are   subroutines written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
-\end{figure}
+  \caption{String edit task (top) and the program \system writes for it (bottom). $f_0$ is a subroutine written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+\end{wrapfigure}
 
@@
 do not have publicly available code and datasets.
-\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
-This is equivalent to \system with $q(x) = \theta$ and $\lambda = \infty$,
-and is like learning the parameters of a PCFG while not learning its structure.
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first exploration cycle.
+\\\noindent \textbf{EC}, an implementation of the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, publicly available at...
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first exploration step.
 
@@
   \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
-  \% solved&\textbf{75\%} &\% &33\%&0\%&30\%\\
-  Solve time&  29s&s&80s&--&
+  \% solved&\textbf{75\%} &\% &33\%&0\%&4\%\\
+  Solve time&  29s&s&80s&--&235s
 
@@
   \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        
-  \caption{Learning curves for \system both with (blue) and without (red) the recognition model as the frontier size is varied (solid/dashed/dotted lines).}\label{learningCurves} 
+  \caption{Learning curves for \system both with (blue) and without (red) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
 \end{figure}
@@
   \caption{
-  Future: dreams.
   Top left: hand crafted traning targets. Top right: examples of

### 2018-05-17 nips_2018.tex
improved the big table

@@
 
-\begin{tabular}{ccc}\toprule
+\begin{figure}[hb]
+\makebox[\textwidth][c]{
+\scriptsize
+\renewcommand\code\texttt
+\renewcommand\codechar[1]{\texttt{"#1"}}
+\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
+  Lists&Strings&Symbolic Regression\\\midrule
   \begin{tabular}{c}
-    \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
-      \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
-      \code{$f(x) = $}\code{($f_0$ nil)}\\\\
+      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
+      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
+      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
       \code{[99\, 4\, 2]}$\to $\code{99}\\
@@
       \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
-      \code{($f_1$ )}
+      \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
+      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
+      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
+      \code{$f(\ell) = $}\code{($f_1$ $\ell$ k)}
   \end{tabular}&
@@
     +83 973-831$\to $83.973.831\\
-    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} ($f_0$ \codechar{.} \codechar{ } (cdr s)))}    \\\\
+    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} ($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
 Temple Anna H $\to $TAH\\
@@
   \begin{tabular}{c}
-    \includegraphics[width = 1cm]{figures/functions/4.png}\\
+    \includegraphics[width = 3em]{figures/functions/4.png}\\
     \code{$f($x$) = $($f_1$ -0.8 1.6)}\\
-    \includegraphics[width = 1cm]{figures/functions/112.png}\\
+    \includegraphics[width = 3em]{figures/functions/112.png}\\
     \code{$f($x$) = $($f_0$ -0.8 1.6)}\\
@@
 
-  \\\midrule&
+  \\\midrule
+
+  \begin{tabular}{l}
+    % $f_0($\code{r}$,\ell) \,=\, $\code{(fold $\ell$ r ($\lambda$ (x a)}\\
+    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(fold}}\code{(cons (index (length a) $\ell$) a)))}\\
+      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
+    $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a) (if (> a x) a x)))}\\
+      \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
+    $f_1(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
+    \phantom{$f_1(\ell,$\code{k}$) \,=\, $\code{(foldr}}
+    \code{($\lambda$ (x a) (if a a (= k x))))}\\
+      \hspace{0.5cm}($f_1$: \emph{Whether $\ell$ contains $k$})\\
+  \end{tabular}&
 
@@
         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s empty$?$ car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
@@
   \begin{tabular}{l}
-    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\quad
+    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
   $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
@@
 
-  \end{tabular}
+\end{tabular}}
+\end{figure}
 

### 2018-05-17 nips_2018.tex
regression figures are now better

@@
 
-Imagine that you are asked to write a program to edit some text, and
+Imagine you are asked  to edit some text, and
 told that you should change the text ``Nancy FreeHafer'' to
@@
 Goodall'' goes to ``Dr. Jane'', drawing upon your prior knowledge of
-text, like that words are separated by spaces. For few-shot program learning problems
-like these, the program synthesis and AI communities have had many successes,
+text, like that words are separated by spaces. Few-shot learning problems
+like these
+are commonplace in both human and machine learning,
+
+, the program synthesis and AI communities have had many successes,
 not just for text editing~\cite{gulwani2011automating} but also for graphics
@@
     \begin{tabular}{l}
-      \code{(lambda (a b) (foldr b (cons "," a)}\\
-      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
-      \code{(lambda (a b) (foldr a b }\\
-      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}
+      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
+      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
+      \code{($\lambda$ (a b) (foldr a b }\\
+      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
     \end{tabular}&
     \begin{tabular}{l}
-      \code{(foldr a b (lambda (x z) }\\
+      \code{(foldr a b ($\lambda$ (x z) }\\
       \phantom{\code{(foldr a b }}\code{(cons x z)))}
@@
     \begin{tabular}{l}
-      \code{(lambda (s) (map (lambda (x)}\\
+      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
       \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
-      \code{(lambda (s) (map (lambda (x) }\\
+      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
       \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
       \end{tabular}&
-    \begin{tabular}{l}      \code{(lambda (s) (map (lambda (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
+    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
       \end{tabular}
@@
     \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
-    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any (lambda (x) (= n x)) $\ell$)}\\
+    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any ($\lambda$ (x) (= n x)) $\ell$)}\\
     \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
@@
     \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
-    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
-    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
+    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (mod x n)) $\ell$)}\\
+    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (+ x k)) $\ell$)}\\
     $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
@@
 As of the 2017 SyGuS competition,
-the best-performing algorithm solves 79.6\% problems.
+the best-performing algorithm solves 79.6\% of the problems.
 But, SyGuS comes with a
@@
       \toprule
-      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map (lambda (x) (if (= x a) b x)) s)}\\
+      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
       \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (cdr (if (= c x) s a))))}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
-        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
+        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-        %(lambda (lambda (fold $0 $0 (lambda (lambda (if (char-eq? $1 $3) empty (cons $1 $0)))))))
+        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
   \\\bottomrule
@@
 
-\begin{wrapfigure}{r}{0.6\textwidth}
-  \centering\begin{tabular}{c}\toprule
-+106 769-858-438 $\longrightarrow$106.769.858.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
-+83 973-757-831 $\longrightarrow$83.973.757.831%& Andrew Cencici $\longrightarrow$ Dr. Andrew
-%DkI$>$DwKI$<$xaPoc& 	DWKI\\
-%KfvTh$>$Dbl$<$OIJso 	&DBL
-\\\midrule
-%(lambda (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' SPACE (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' '-' (cdr $0))))
-\multicolumn{1}{c}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  '.' ' ' ($f_0$ '.' '-' (cdr s)))}}
-%\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
-\\
-\bottomrule
-  \end{tabular}
-
-  %% \vspace{0.25cm}
-
-  %% \begin{minipage}{\linewidth}
-  %%   \centering
-  %%   $f($\code{s}$) \,=\, $
-  %% \end{minipage}
-
-  \caption{String edit task (top) and the program \system writes for it (bottom). $f_0$ is a subroutine written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
-\end{wrapfigure}
+%% \begin{wrapfigure}{r}{0.6\textwidth}
+%%   \centering\begin{tabular}{c}\toprule
+%% +106 769-858-438 $\longrightarrow$106.769.858.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
+%% +83 973-757-831 $\longrightarrow$83.973.757.831%& Andrew Cencici $\longrightarrow$ Dr. Andrew
+%% %DkI$>$DwKI$<$xaPoc& 	DWKI\\
+%% %KfvTh$>$Dbl$<$OIJso 	&DBL
+%% \\\midrule
+%% %(lambda (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' SPACE (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' '-' (cdr $0))))
+%% \multicolumn{1}{c}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  '.' ' ' ($f_0$ '.' '-' (cdr s)))}}
+%% %\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
+%% \\
+%% \bottomrule
+%%   \end{tabular}
+
+%%   %% \vspace{0.25cm}
+
+%%   %% \begin{minipage}{\linewidth}
+%%   %%   \centering
+%%   %%   $f($\code{s}$) \,=\, $
+%%   %% \end{minipage}
+
+%%   \caption{String edit task (top) and the program \system writes for it (bottom). $f_0$ is a subroutine written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
+%% \end{wrapfigure}
 
@@
 
-\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
-  \includegraphics[width = \functionSize]{figures/functions/0.png}
+\begin{wrapfigure}{r}{6.4cm} \newcommand{\functionSize}{1.5cm}\vspace{-0.5cm}\centering
+%  \includegraphics[width = \functionSize]{figures/functions/0.png}
   \includegraphics[width = \functionSize]{figures/functions/6.png}
-  \includegraphics[width = \functionSize]{figures/functions/207.png}
+  \includegraphics[width = \functionSize]{figures/functions/116.png}
   \includegraphics[width = \functionSize]{figures/functions/160.png}
-  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
-  \includegraphics[width = \functionSize]{figures/functions/432.png}
-  \includegraphics[width = \functionSize]{figures/functions/415.png}
-  \includegraphics[width = \functionSize]{figures/functions/426.png}
-  \includegraphics[width = \functionSize]{figures/functions/597.png}
-  \includegraphics[width = \functionSize]{figures/functions/591.png}
+  \includegraphics[width = \functionSize]{figures/functions/149.png}
+  %% \includegraphics[width = \functionSize]{figures/functions/432.png}
+  %% \includegraphics[width = \functionSize]{figures/functions/415.png}
+  %% \includegraphics[width = \functionSize]{figures/functions/426.png}
+  %% \includegraphics[width = \functionSize]{figures/functions/597.png}
+  %% \includegraphics[width = \functionSize]{figures/functions/591.png}
   \caption{Recognition model input for symbolic regression.
@@
   \% solved&   \textbf{84\% }&75\%&38\%&38\%&37\% \\
-  Solve time&  24s& 40s  &31s&55s&29s\\
+  Solve time&  24s& 40s  &31s&55s&29s
   %% MDL (nats)&   9.3 &11.0&2.3&--&4.6

### 2018-05-17 nips_2018.tex
improved big table

@@
 &\begin{tabular}{l}
-    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y)}\\
-    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
+    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y) (cons x y)))}\\
     \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
@@
 its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
-\begin{figure}
-  \begin{tabular}{lcl}
-    \toprule
-    Domain&Example Task&Part of the learned DSL\\\midrule
-    Lists&
-    \begin{tabular}{rcl}
-      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
-      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
-    \end{tabular}
-    &
-    \begin{tabular}{l}
-      \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
-          \hspace{0.5cm} \emph{(appends lists)}
-      \end{tabular}
-
-    \\    Strings&
-    \begin{tabular}{rcl}
-      Temple Anna H& $\longrightarrow$& TAH\\
-      Lara Gregori&$\longrightarrow$&LG
-      \end{tabular}
-    &
-    \begin{tabular}{l}
-      \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
-      \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
-      \end{tabular}
-
-    \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
-      \code{(+ (* real x) real)} \\
-          \hspace{0.5cm} \emph{(a linear function of x)}
-      \end{tabular}
-
-    \\\bottomrule
-  \end{tabular}
-  \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
-%  \vspace{-4em}
-\end{figure}
+%% \begin{figure}
+%%   \begin{tabular}{lcl}
+%%     \toprule
+%%     Domain&Example Task&Part of the learned DSL\\\midrule
+%%     Lists&
+%%     \begin{tabular}{rcl}
+%%       \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
+%%       \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
+%%     \end{tabular}
+%%     &
+%%     \begin{tabular}{l}
+%%       \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
+%%           \hspace{0.5cm} \emph{(appends lists)}
+%%       \end{tabular}
+
+%%     \\    Strings&
+%%     \begin{tabular}{rcl}
+%%       Temple Anna H& $\longrightarrow$& TAH\\
+%%       Lara Gregori&$\longrightarrow$&LG
+%%       \end{tabular}
+%%     &
+%%     \begin{tabular}{l}
+%%       \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
+%%       \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
+%%       \end{tabular}
+
+%%     \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
+%%       \code{(+ (* real x) real)} \\
+%%           \hspace{0.5cm} \emph{(a linear function of x)}
+%%       \end{tabular}
+
+%%     \\\bottomrule
+%%   \end{tabular}
+%%   \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
+%% %  \vspace{-4em}
+%% \end{figure}
 
@@
 \begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
-  Lists&Strings&Symbolic Regression\\\midrule
+  {\normalsize Lists}&{\normalsize Strings}&{\normalsize Symbolic Regression}\\\midrule
   \begin{tabular}{c}
@@
     +83 973-831$\to $83.973.831\\
-    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} ($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
+    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
+    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
 Temple Anna H $\to $TAH\\
@@
 
-  \begin{tabular}{c}
-    \includegraphics[width = 3em]{figures/functions/4.png}\\
-    \code{$f($x$) = $($f_1$ -0.8 1.6)}\\
-    \includegraphics[width = 3em]{figures/functions/112.png}\\
-    \code{$f($x$) = $($f_0$ -0.8 1.6)}\\
+  \begin{tabular}{cc}
+    \includegraphics[width = 3em]{figures/functions/4.png}&
+    \includegraphics[width = 3em]{figures/functions/146}\\
+    \code{$f($x$) = $($f_1$ x)}&    \code{$f($x$) = $($f_6$ x)}\\
+    \includegraphics[width = 3em]{figures/functions/112.png}&
+        \includegraphics[width = 3em]{figures/functions/92.png}
+    \\
+    \code{$f($x$) = $($f_4$ x)}&    \code{$f($x$) = $($f_3$ x)}\\
 
@@
       % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
-    $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a) (if (> a x) a x)))}\\
+    $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
       \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
     $f_1(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
-    \phantom{$f_1(\ell,$\code{k}$) \,=\, $\code{(foldr}}
+    \phantom{$f_1(\ell,$\code{k}}
     \code{($\lambda$ (x a) (if a a (= k x))))}\\
@@
   \begin{tabular}{l}
-          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
+          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
       \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
-        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s empty$?$ car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
+      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
+        \hspace{0.5cm}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s empty$?$ car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
@@
     $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
+    $f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}\\
+    $f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}\\
+    \hspace{0.5cm}\emph{($f_4$: 4th order polynomial)}\\
+        $f_5($\code{x}$)\,=\,$\code{(/ real x)}\\
+    $f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))} \\
+    \hspace{0.5cm}\emph{($f_6$: rational function)}\\
+
     \end{tabular}
@@
 \end{tabular}}
+\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them.
+Bottom: Subset of learned DSL.
+Notice that learned DSL primitives can call each other ($f_1$/$f_2$ in Strings \& Regression)}
 \end{figure}

### 2018-05-17 nips_2018.tex
more interesting list inventions in big figure
more spacing for list inventions in big figure
fix list example induced programs
introduction edits

@@
 like these
-are commonplace in both human and machine learning,
-
-, the program synthesis and AI communities have had many successes,
-not just for text editing~\cite{gulwani2011automating} but also for graphics
-programs~\cite{ellis2017learning}, planning
+are commonplace in both human and machine learning.
+We cast these few-shot learning problems as each being a program synthesis problem.
+The program synthesis and AI communities have built many successful program synthesis algorithms,
+spanning text editing (e.g., FlashFill~\cite{gulwani2011automating}), motor programs~\cite{lake2015human}, procedural graphics~\cite{ellis2017learning}, planning
 procedures~\cite{devlin2017neural}, and many others.  However, the
-success of these systems hinges upon having a carefully
+success of these systems hinges upon a carefully
 hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
@@
       % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
-      \code{[99\, 4\, 2]}$\to $\code{99}\\
-      \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
-      \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
-      \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
+      % \code{[99\, 4\, 2]}$\to $\code{99}\\
+      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
+      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
+      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
+      \code{[1]}$\to $\code{[1\, 3]}\\
+      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
+      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
       \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
       \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
-      \code{$f(\ell) = $}\code{($f_1$ $\ell$ k)}
+      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
   \end{tabular}&
@@
       % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
-    $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
-      \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
-    $f_1(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
+    % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
+      % \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
+    %NOTE: this is the actual invention, but I removed a redundant lambda
+    % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
+    $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
+    \vspace{2ex}
+      \hspace{0.5cm}($f_0$: \emph{Concatenate }\code{r}\emph{ with $\ell$})\\
+    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k empty) $\ell$)}\\
+    \vspace{2ex}
+      \hspace{0.5cm}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
+    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
     \phantom{$f_1(\ell,$\code{k}}
     \code{($\lambda$ (x a) (if a a (= k x))))}\\
-      \hspace{0.5cm}($f_1$: \emph{Whether $\ell$ contains $k$})\\
+      \hspace{0.5cm}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
   \end{tabular}&

### 2018-05-17 nips_2018.tex
consistent domain titles: "list processing" and "text editing"

@@
 Returning to Figure~\ref{firstPageFigure},
-a good string processing library should support appending strings and splitting on spaces --
+a good text editing library should support appending strings and splitting on spaces --
 exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
@@
 We apply \system to four domains:
-list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
- For each of these we initially provide a generic
- set of programming primitives.
-Our algorithm then discovers
-its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+list processing;
+text editing (in the style of FlashFill~\cite{gulwani2011automating});
+and symbolic regression.
+For each of these we initially provide a generic set of programming primitives.
+Our algorithm then discovers its own DSL for expressing solutions in the
+domain (Tbl.~\ref{initialExampleDSL}).
 %% \begin{figure}
@@
 \begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
-  {\normalsize Lists}&{\normalsize Strings}&{\normalsize Symbolic Regression}\\\midrule
+  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
   \begin{tabular}{c}
@@
 \end{tabular}}
-\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them.
-Bottom: Subset of learned DSL.
-Notice that learned DSL primitives can call each other ($f_1$/$f_2$ in Strings \& Regression)}
+\caption{Top: Example tasks from each domain, each followed by the programs
+  \system discovers for them. Bottom: Subset of learned DSL. Notice that
+  learned DSL primitives can call each other
+  ($f_0$/$f_1$ in Lists, $f_1$/$f_2$ in Text \& Regression)}
 \end{figure}
@@
 We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
-For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
-and initially provide the system with a generic
-set of list processing primitives: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
-
-\subsection{List Functions}\label{listSection}
-Synthesizing programs that manipulate data structures is
-a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
+For both these domains we use a bidirectional GRU~\cite{cho2014learning} for
+the recognition model, and initially provide the system with a generic set
+of list processing primitives:
+\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
+\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
+\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+
+\subsection{List Processing}\label{listSection}
+Synthesizing programs that manipulate data structures is a widely studied
+problem in the programming languages community~\cite{feser2015synthesizing}.
 %with applications to computer aided programming~\cite{solar2008program}.
-We consider this problem within the context of
-learning functions that manipulate lists.
-We created 244 Lisp-style list manipulation tasks,
-each with 15 input/output examples (Tbl.~\ref{listexamples}).
-Our data set is challenging along two dimensions:
-many of the functions are very complicated,
-and the agent must learn to solve these complicated problems from only 244 tasks.
-Our data set primarily consists of arithmetic operations upon sequences,
-and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
+We consider this problem within the context of learning functions that
+manipulate lists.
+We created 236 Lisp-style list manipulation tasks, each with 15 input/output
+examples (Tbl.~\ref{listexamples}).
+Our data set is challenging along two dimensions: many of the functions are
+very complicated, and the agent must learn to solve these complicated
+problems from only 236 tasks.
+Our data set primarily consists of arithmetic operations upon sequences, and
+so, in addition to the Lisp primitives provided for the text editing
+experiments, we additionally start the system out with the following
+primitives:
 \code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}, \code{2},
@@
 
-A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
-Although we provided \system with some of the standard higher-order functions,
-like \code{foldr} and \code{unfold},
-we did not include others.
-When trained on these list functions,
-our system rediscovers
-and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}
+A complete repertoire of higher-order functions is a staple of functional
+programming standard libraries.
+Although we provided \system with some of the standard higher-order
+functions, like \code{foldr} and \code{unfold}, we did not include others.
+When trained on these list functions, our system rediscovers and then reuses
+the higher-order function \code{filter} \textbf{Lucas: not sure if it
+actually does this! It would be cool if we can write something like this
+though.}
 
@@
 
-\subsection{String Editing}\label{textSection}
-Synthesizing programs that edit strings is a classic problem in the
+\subsection{Text Editing}\label{textSection}
+Synthesizing programs that edit text is a classic problem in the
 programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
-and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
+and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
 This prior work presumes a hand-engineered DSL.
@@
 
-We automatically generated 109 string editing tasks  %% with 4 input/output examples each
+We automatically generated 109 text editing tasks  %% with 4 input/output examples each
 (Fig.~\ref{exampleTextProblem}).
@@
     \end{tabular}
-  \caption{Some string editing learned subroutines }\label{textPrimitives}
+  \caption{Some text editing learned subroutines }\label{textPrimitives}
   \end{table}
@@
 improves on the baselines,
-sometimes dramatically (string editing and symbolic regression).
+sometimes dramatically (text editing and symbolic regression).
 The recognition model 
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{List Processing}}\\\midrule
   \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
+  \\\midrule\multicolumn{6}{c}{\emph{Text Editing}}\\\midrule
   \% solved&\textbf{75\%} &\% &33\%&0\%&4\%\\

### 2018-05-17 nips_2018.tex
change experiment-related section titles

@@
 
-\section{Sequence manipulating programs}\label{sequences}
+\section{Sequence Manipulation: Domains from generic primitives}\label{sequences}
 We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
@@
 
-\subsection{Quantitative Results}\label{quantitative}
+\section{Quantitative Results}\label{quantitative}
 

### 2018-05-17 nips_2018.tex
move figures around in alg section

@@
   \end{equation}
-We induce a DSL by searching locally through the space of DSLs,
-proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
-The search moves work by introducing new
-$\lambda$-expressions into the DSL.
-We propose these new expressions by extracting fragments of
-programs already in the frontiers (Fig.~\ref{fragmentExample}).
-An important point here is that we are \emph{not} simply adding
-subexpressions of programs to $\mathcal{D}$.  Instead, we are
-extracting fragments that unify with programs in the frontiers.  This
-idea of storing and reusing fragments of expressions comes from
-Fragment Grammars~\cite{tim} and Tree-Substitution
-Grammars~\cite{cohn2010inducing}, and is closely related to the idea
-of antiunification~\cite{henderson2013cumulative}.
-
-To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp\left(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) \right)$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-and place a symmetric Dirichlet prior over the weight vector $\theta$.
-%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
- %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
-
-\begin{wrapfigure}{R}{0.6\textwidth}\vspace{-0.7cm}
+
+\begin{wrapfigure}{R}{0.6\textwidth}
   \begin{minipage}{0.6\textwidth}    
@@
   \end{minipage}
-\vspace{-0.7cm}\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
-%% Although this  may seem 
-%% very similar to estimating the parameters of a probabilistic context free grammar,
-%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-%% our DSLs are context-sensitive due to the presence of variables
-%% in the programs and also due to the polymorphic typing system.
-%% In the Supplement we derive a tractable MAP estimator for $\theta$.
-Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
-\begin{figure*}[b]
+\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
+
+We induce a DSL by searching locally through the space of DSLs,
+proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
+The search moves work by introducing new
+$\lambda$-expressions into the DSL.
+We propose these new expressions by extracting fragments of
+programs already in the frontiers (Fig.~\ref{fragmentExample}).
+An important point here is that we are \emph{not} simply adding
+subexpressions of programs to $\mathcal{D}$.  Instead, we are
+extracting fragments that unify with programs in the frontiers.  This
+idea of storing and reusing fragments of expressions comes from
+Fragment Grammars~\cite{tim} and Tree-Substitution
+Grammars~\cite{cohn2010inducing}, and is closely related to the idea
+of antiunification~\cite{henderson2013cumulative}.
+
+\begin{figure}[b]
 \centering  \begin{tabular}{ll}
@@
   \label{fragmentExample}
-\end{figure*}
+\end{figure}
+
+To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+and place a symmetric Dirichlet prior over the weight vector $\theta$.
+%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
+ %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
+%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
+
+%% Although this  may seem 
+%% very similar to estimating the parameters of a probabilistic context free grammar,
+%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+%% our DSLs are context-sensitive due to the presence of variables
+%% in the programs and also due to the polymorphic typing system.
+%% In the Supplement we derive a tractable MAP estimator for $\theta$.
+Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+recognition model training, and DSL induction.
 

### 2018-05-17 nips_2018.tex
consistent is-nil/nil usage

@@
       \hspace{0.5cm}($f_0$: \emph{Concatenate }\code{r}\emph{ with $\ell$})\\
-    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k empty) $\ell$)}\\
+    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
     \vspace{2ex}
       \hspace{0.5cm}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
-    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (empty$?$ $\ell$)}\\
+    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
     \phantom{$f_1(\ell,$\code{k}}
@@
         \hspace{0.5cm}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s empty$?$ car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
@@
         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
+      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
@@
         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
+        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
   \\\bottomrule

### 2018-05-17 nips_2018.tex
s/fold/foldr/
cocosea

@@
 \newcommand{\sem}[1]{\llbracket #1 \rrbracket}
-\newcommand{\system}{\textsc{ECC}~}
-\newcommand{\systemEnding}{\textsc{ECC}}
+\newcommand{\system}{\textsc{CoCoSea}~}
+\newcommand{\systemEnding}{\textsc{CoCoSea}}
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 We cast these few-shot learning problems as each being a program synthesis problem.
-The program synthesis and AI communities have built many successful program synthesis algorithms,
+The programming languages and AI communities have built many successful program synthesis algorithms,
 spanning text editing (e.g., FlashFill~\cite{gulwani2011automating}), motor programs~\cite{lake2015human}, procedural graphics~\cite{ellis2017learning}, planning
@@
 &\begin{tabular}{l}
-    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y) (cons x y)))}\\
+    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b ($\lambda$ (x y) (cons x y)))}\\
     \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s ($\lambda$ (x a)}\\
+    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\
     \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
@@
 the task at hand, even if they cannot instantly work out the details.
-We combine these two ideas into
-an algorithm called \systemEnding,
+We weave these ideas together into
+an algorithm called \system (\textbf{Co}mpress/\textbf{Co}mpile/\textbf{Sea}rch),
 which takes as input a collection of
@@
 solves three problems:
-(1) Writing programs that solve the tasks;
+(1) Searching for programs that solve the tasks;
 (2) Composing a library (DSL) of
@@
 
-Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
-because it iterates between three different steps: an
-\textbf{Explore} step uses the DSL and neural network to explore the space of programs,
+%% Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
+%% because it
+\system iterates through three different steps: a
+\textbf{Search} step uses the DSL and neural network to explore the space of programs,
 searching for ones that solve the tasks; a \textbf{Compress} step
-modifies the structure of the DSL by discovering regularities across programs found by
-the previous Explore step; and a \textbf{Compile} step, which
-improves the program search procedure by training a neural network to
+modifies the structure of the DSL by discovering regularities across programs found during search; and a \textbf{Compile} step, which
+improves the search procedure by training a neural network to
 write programs in the current DSL, in the spirit of ``amortized'' or
@@
 distills commonalities across programs that solve tasks, helping
-the agent solve related program induction problems. The neural
+the agent solve related programming problems. The neural
 recognition model ensures that searching for programs remains tractable
 even as the DSL (and hence the search space for programs) expands.
-
-Because any model may be encoded as a (deterministic or probabilistic) program,
-we carefully delineate the scope of problems considered here.
+%% Because any model may be encoded as a (deterministic or probabilistic) program,
+%% we carefully delineate the scope of problems considered here.
 We think of \system as learning to solve the kinds of problems that humans
@@
 Even with a good DSL, program search may be intractable, so
-we amortize the cost of program search by training a neural network to
+we amortize the cost of search by training a neural network to
 assist the search procedure.
 
-We apply \system to four domains:
-list processing;
-text editing (in the style of FlashFill~\cite{gulwani2011automating});
-and symbolic regression.
-For each of these we initially provide a generic set of programming primitives.
-Our algorithm then discovers its own DSL for expressing solutions in the
-domain (Tbl.~\ref{initialExampleDSL}).
-%% \begin{figure}
-%%   \begin{tabular}{lcl}
-%%     \toprule
-%%     Domain&Example Task&Part of the learned DSL\\\midrule
-%%     Lists&
-%%     \begin{tabular}{rcl}
-%%       \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
-%%       \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
-%%     \end{tabular}
-%%     &
-%%     \begin{tabular}{l}
-%%       \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
-%%           \hspace{0.5cm} \emph{(appends lists)}
-%%       \end{tabular}
-
-%%     \\    Strings&
-%%     \begin{tabular}{rcl}
-%%       Temple Anna H& $\longrightarrow$& TAH\\
-%%       Lara Gregori&$\longrightarrow$&LG
-%%       \end{tabular}
-%%     &
-%%     \begin{tabular}{l}
-%%       \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
-%%       \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
-%%       \end{tabular}
-
-%%     \\Regression&\includegraphics[width = 1cm]{figures/functions/476.png}&\begin{tabular}{l}
-%%       \code{(+ (* real x) real)} \\
-%%           \hspace{0.5cm} \emph{(a linear function of x)}
-%%       \end{tabular}
-
-%%     \\\bottomrule
-%%   \end{tabular}
-%%   \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
-%% %  \vspace{-4em}
-%% \end{figure}
-
+We apply \system to three domains:
+list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
+ For each of these we initially provide a generic
+ set of programming primitives.
+Our algorithm then constructs
+its own DSL for expressing solutions in the domain (Fig.~\ref{initialExampleDSL}).
 \begin{figure}[hb]
@@
   \begin{tabular}{l}
-    % $f_0($\code{r}$,\ell) \,=\, $\code{(fold $\ell$ r ($\lambda$ (x a)}\\
-    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(fold}}\code{(cons (index (length a) $\ell$) a)))}\\
+    % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
+    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
       % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
@@
 \end{tabular}}
-\caption{Top: Example tasks from each domain, each followed by the programs
-  \system discovers for them. Bottom: Subset of learned DSL. Notice that
-  learned DSL primitives can call each other
-  ($f_0$/$f_1$ in Lists, $f_1$/$f_2$ in Text \& Regression)}
+\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them.
+Bottom: Subset of learned DSL.
+Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
 \end{figure}
@@
 We were inspired by EC,
-and go beyond it by giving a new algorithm for learning DSLs, as well as a way of combining DSL learning with
+and go beyond it by giving a new algorithm that learns DSLs, as well as a way of combining DSL learning with
 neurally guided program search.
@@
 showing empirical improvements.
-The contribution of this work is
-then
-an effective algorithm for learning DSLs which jointly trains a neural net to
+
+The new contribution of this work is
+thus an  algorithm for learning DSLs which jointly trains a neural net to
 search for programs in the DSL.
@@
 We extend these ideas into an inference strategy that iterates
-through three steps: an \textbf{Explore} step uses the current DSL and
+through three steps: a \textbf{Search} step uses the current DSL and
 recognition model to search for programs that solve the tasks.  The
@@
         %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Explore})}
+        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Search})}
         \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Explore)}}
+        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Search)}}
 
@@
         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-        %($\lambda$ ($\lambda$ (fold $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
+        %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
   \\\bottomrule
@@
 and lesioning it also slows down the convergence of the algorithm,
+
 taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).

### 2018-05-17 nips_2018.tex
got rid of exploration terminology

@@
   \end{comment}
-\\\noindent \textbf{Exploration: Searching for programs.}  Our program search
+\\\noindent \textbf{Search: Solving tasks.}  Our program search
 is informed by both the DSL and the recognition model. When these
@@
 \\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
-from the programs found in the exploration phase, aiming
+from the programs found in the search phase, aiming
 to maximally compress (or, raise the prior probability of) these
@@
 from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
-programs found by the search procedure during exploration. As the DSL
+programs found by the search procedure during search. As the DSL
 improves and as search finds more programs, the recognition model gets
@@
 \system does approximate MAP inference  by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t. the frontiers (Exploration) and the DSL (Compression):
+alternating maximization w.r.t. the frontiers (Search) and the DSL (Compression):
 \\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
@@
 
-\subsection{Exploration: Searching for Programs}\label{explorationSection}
+\subsection{Searching for Programs}\label{explorationSection}
 
@@
 We note the following implementation details:
-(1) We perform an exploration cycle before each compression and compilation  cycle.
+(1) We perform an search cycle before each compression and compilation  cycle.
 (2) On the first iteration, we do \emph{not} train the recognition model on samples from
@@
 \\\noindent \textbf{EC}, an implementation of the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, publicly available at...
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first exploration step.
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
 

### 2018-05-17 nips_2018.tex
placeholder list tables

@@
 How should we get the data to train $q$?
-This is nonobvious  because \system is only weakly supervised (i.e., learns only from tasks and not from (program, task) pairs).
+This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
 One approach is to sample programs from the DSL,
@@
 and then train $q$ to predict the program from the input/outputs.
-This approach is like how a Helmholtz machine
+This is like how a Helmholtz machine
 trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
@@
 training $q$ on the (program, task)
-pairs discovered by the algorithm so far.
+pairs discovered by search.
 The advantage of self-supervised learning
@@
 An important point here is that we are \emph{not} simply adding
-subexpressions of programs to $\mathcal{D}$.  Instead, we are
+subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
 extracting fragments that unify with programs in the frontiers.  This
@@
 
-\begin{comment}
-\begin{table}
-\centering
+\begin{figure*}
+\makebox[\textwidth][c]{
+\begin{minipage}{0.6\textwidth}
   \begin{tabular}{lll}
@@
     Name & Input & Output \\\midrule
-  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
-  len & [3\, 5\, 12\, 1] & 4 \\
     has-2 & [4\, 5\, 7\, 4] & \code{false} \\
@@
     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
     keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
@@
   \end{tabular}
-\caption{Sample tasks from our list function domain %% We used 15 input/output
-  %% examples for each task to reduce ambiguity in the task's function,
-  %% ensuring solutions achieve the desired concept.
-}
+  \captionof{table}{Sample tasks from our list function domain}
   \label{listexamples}
-\end{table}
-
-\begin{table}[t]
+\end{minipage}
+\hfill
+\begin{minipage}{0.6\textwidth}
   \centering
@@
     \toprule
-    $f_0($\code{i$,\ell$}$) \,=\, $\code{(singleton (index i $\ell$))}\\
-    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
-    $f_1($\code{i$,\ell$}$) \,=\, $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
-    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
-    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any ($\lambda$ (x) (= n x)) $\ell$)}\\
-    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
-    $f_3($\code{i$,\ell$}$) \,=\, $\code{(index (negate i) (sort $\ell$))}\\
-    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
-    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (mod x n)) $\ell$)}\\
-    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (+ x k)) $\ell$)}\\
-    $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
-    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
+    $f_0($\code{i$,\ell$}$) \,=\, $\code{(TODO)}\\
+    \hspace{0.5cm}($f_0$: \emph{TODO})\\
   \bottomrule
   \end{tabular}
-  \caption{Some learned DSL primitives for list functions}\label{listinduced}
-\end{table}
-
-  \end{comment}
+  \captionof{table}{Some learned DSL primitives for list functions}\label{listinduced}
+\end{minipage}
+  }
+\end{figure*}
 
@@
 
-We automatically generated 109 text editing tasks  %% with 4 input/output examples each
-(Fig.~\ref{exampleTextProblem}).
+We automatically generated 109 text editing tasks with 4 input/output examples each
+%(Fig.~\ref{exampleTextProblem}).
 At first, \system cannot find any correct programs for most of the tasks.
-After three iterations, it assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
+After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
 all of the tasks.
@@
 
-\begin{table}[t]
-  \centering
-    \begin{tabular}{l}
-      \toprule
-      $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
-      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
-        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
-        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
-        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
-        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-        %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
-  \\\bottomrule
-    \end{tabular}
-  \caption{Some text editing learned subroutines }\label{textPrimitives}
-  \end{table}
+%% \begin{table}[t]
+%%   \centering
+%%     \begin{tabular}{l}
+%%       \toprule
+%%       $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
+%%       \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
+%%       $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
+%%         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
+%%       $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
+%%         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
+%%         $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
+%%         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
+%%         %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
+%%   \\\bottomrule
+%%     \end{tabular}
+%%   \caption{Some text editing learned subroutines }\label{textPrimitives}
+%%   \end{table}
 
@@
 \begin{tabular}{llllll}
-  \toprule& Ours& \begin{tabular}{c}
+  \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
     Ours\\(no NN)
-    \end{tabular}& RF/DC & PCFG & Enum
+    \end{tabular}}& RF/DC & PCFG & Enum
 
@@
 
- \noindent \emph{Deep learning:} Much recent work in the ML community has
+ \noindent \textbf{Deep learning:} Much recent work in the ML community has
  focused on creating neural networks that regress from
  input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. %This family of work is closest to our RF/DC baseline.
- These neural networks are typically trained with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
+\systemEnding's recognition model draws heavily from this line of work, particularly from~\cite{menon2013machine}.
+We see these prior works as operating in a different regime: typically, they train with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
  Our work   considers a weakly-supervised regime where ground truth programs are not provided and
- the agent must learn from a few hundred tasks.
-
- \noindent \emph{Inventing new subroutines for program induction:}
+ the agent must learn from at most a few hundred tasks,
+ which is facilitated by our ``Helmholtz machine'' style recognition model training (Sec.~\ref{recognitionSection}).
+
+ \noindent \textbf{Inventing new subroutines for program induction:}
  Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting.
@@
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14}.
-
- Our work is an instance of
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
+
+\noindent\textbf{Bayesian Program
+ Learning:} Our work is an instance of
  Bayesian Program
@@
 primitive set that could effectively bootstrap DSLs for new domains,
-including the four domains considered,  but also many others?
+including the three domains considered,  but also many others?
 

### 2018-05-17 nips_2018.tex
random edits

@@
 
-\begin{figure*}
-\makebox[\textwidth][c]{
-\begin{minipage}{0.6\textwidth}
+\begin{comment}
+\begin{table}
+\centering
 \begin{tabular}{lll}
@@
   Name & Input & Output \\\midrule
+  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
+  len & [3\, 5\, 12\, 1] & 4 \\
   has-2 & [4\, 5\, 7\, 4] & \code{false} \\
@@
   rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
+  pow-3 & [10\, 4\, 13] & [1000\, 64\, 2197] \\
   keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
@@
 \end{tabular}
-  \captionof{table}{Sample tasks from our list function domain}
+\caption{Sample tasks from our list function domain %% We used 15 input/output
+  %% examples for each task to reduce ambiguity in the task's function,
+  %% ensuring solutions achieve the desired concept.
+}
 \label{listexamples}
-\end{minipage}
-\hfill
-\begin{minipage}{0.6\textwidth}
+\end{table}
+
+\begin{table}[t]
   \centering
@@
     \toprule
-    $f_0($\code{i$,\ell$}$) \,=\, $\code{(TODO)}\\
-    \hspace{0.5cm}($f_0$: \emph{TODO})\\
+    $f_0($\code{i$,\ell$}$) \,=\, $\code{(singleton (index i $\ell$))}\\
+    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
+    $f_1($\code{i$,\ell$}$) \,=\, $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
+    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
+    $f_2($\code{n$,\ell$}$) \,=\, $\code{(any ($\lambda$ (x) (= n x)) $\ell$)}\\
+    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
+    $f_3($\code{i$,\ell$}$) \,=\, $\code{(index (negate i) (sort $\ell$))}\\
+    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
+    $f_4($\code{n$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (mod x n)) $\ell$)}\\
+    $f_5($\code{k$,\ell$}$) \,=\, $\code{(mapi ($\lambda$ (i x) (+ x k)) $\ell$)}\\
+    $f_6($\code{k$,$n$,\ell$}$) \,=\, $\code{($f_4$ n ($f_5$ k $\ell$)}\\
+    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
   \bottomrule
   \end{tabular}
-  \captionof{table}{Some learned DSL primitives for list functions}\label{listinduced}
-\end{minipage}
-  }
-\end{figure*}
+  \caption{Some learned DSL primitives for list functions}\label{listinduced}
+\end{table}
+
+  \end{comment}
 

### 2018-05-17 nips_2018.tex
improvements to figures

--- 2018-05-17 nips_2018.tex
@@
 Our algorithm then constructs
-its own DSL for expressing solutions in the domain (Fig.~\ref{initialExampleDSL}).
-\begin{figure}[hb]
+its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+\begin{table}[hb]
 \makebox[\textwidth][c]{
@@
   \end{tabular}
-
   \\\bottomrule 
-
-\end{tabular}}
+\end{tabular}}\\\\
 \caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them.
@@
 Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
-\end{figure}
+\end{table}
 
@@
 We propose these new expressions by extracting fragments of
-programs already in the frontiers (Fig.~\ref{fragmentExample}).
+programs already in the frontiers (Tbl.~\ref{fragmentExample}).
 An important point here is that we are \emph{not} simply adding
@@
 
-\begin{figure}[b]
+\begin{table}[b]
 \centering  \begin{tabular}{ll}
@@
   \label{fragmentExample}
-\end{figure}
+\end{table}
 
@@
 
-\begin{wrapfigure}{r}{6.4cm} \newcommand{\functionSize}{1.5cm}\vspace{-0.5cm}\centering
+%% \begin{align*}
+%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
+%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
+%% \end{align*}
+
+%% %% and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
+%% %% \begin{align*}
+%% %% \log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
+%% %% %\text{\emph{BIC:}} &\approx 
+%% %% \end{align*}
+%% where $\mathcal{N}(\cdot |\cdot )$ is the normal density and
+%% $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over
+%% $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~:
+
+%% where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$
+%% found by performing gradient ascent on the likelihood of the
+%% observations w.r.t. $\vec{\mathcal{R}}$.
+
+\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Tbl.~\ref{regressionDSL}).
+The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
+For example, it learns to represent linear functions with the program
+\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
+which has five continuous parameters.
+This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
+
+%% The neural recognition model accelerates inference,
+%% but given enough iterations of our algorithm,
+%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Fig.~\ref{regressionLearningCurve}; No NN in Tbl.~\ref{regressionComparison}).
+%% We compare our model on held-out tasks against 4 baselines (Tbl.~\ref{regressionComparison}).
+
+\begin{figure*}[h]
+\makebox[\textwidth][c]{
+\begin{minipage}{0.55\textwidth} \newcommand{\functionSize}{1.5cm}\centering
 %  \includegraphics[width = \functionSize]{figures/functions/0.png}
@@
 While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
-\end{wrapfigure}
-%% \begin{align*}
-%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
-%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
-%% \end{align*}
-
-%% %% and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
-%% %% \begin{align*}
-%% %% \log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
-%% %% %\text{\emph{BIC:}} &\approx 
-%% %% \end{align*}
-%% where $\mathcal{N}(\cdot |\cdot )$ is the normal density and
-%% $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over
-%% $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~:
-
-%% where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$
-%% found by performing gradient ascent on the likelihood of the
-%% observations w.r.t. $\vec{\mathcal{R}}$.
-
-\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
-The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
-For example, it learns to represent linear functions with the program
-\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
-which has five continuous parameters.
-This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
-\begin{wrapfigure}{l}{0.35\textwidth}\centering
-\begin{tabular}{l}
+\end{minipage}
+\hspace{1em}
+\begin{minipage}{0.5\textwidth}
+  \centering\begin{tabular}{l}
   \toprule
@@
 \end{tabular}
-\caption{Some learned subroutines for symbolic regression.
+  \captionof{table}{Some learned subroutines for symbolic regression.
 System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
-\end{wrapfigure}
-
-%% The neural recognition model accelerates inference,
-%% but given enough iterations of our algorithm,
-%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Fig.~\ref{regressionLearningCurve}; No NN in Tbl.~\ref{regressionComparison}).
-%% We compare our model on held-out tasks against 4 baselines (Tbl.~\ref{regressionComparison}).
+\end{minipage}
+}
+\end{figure*}
 
@@
 
-\begin{wrapfigure}{r}{0.4\textwidth}
-  \begin{center}
-    \includegraphics[width = 0.4\textwidth]{figures/geomCompiled.eps} 
-  \end{center}
-  \caption{
-  Top left: hand crafted traning targets. Top right: examples of
-  discovered compiled new programs. Bottom: compiled programm across iterations
-  to highlight structure emergence.
-  }\label{geomCompiled} 
-\end{wrapfigure}
-
+\parbox{0.58\textwidth}{
 We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
@@
 including the three domains considered,  but also many others?
+}\hfill\begin{minipage}{0.4\textwidth}\centering
+  \includegraphics[width=0.8\textwidth]{figures/geomCompiled.eps} 
+  \captionof{figure}{
+  Top left: hand crafted traning targets. Top right: examples of
+  discovered compiled new programs. Bottom: compiled programm across iterations
+  to highlight structure emergence.
+  }\label{geomCompiled}
+\end{minipage}
 

### 2018-05-17 nips_2018.tex
writing

@@
 After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
-all of the tasks.
+all of the tasks. The learned DSL contains \textbf{X} new functions, listed in the supplement.
 
@@
 
-\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Tbl.~\ref{regressionDSL}).
-The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
+\system learns a DSL containing 13 new functions,
+most of which are templates for polynomials of different orders or ratios of polynomials.
+The algorithm also learns to find programs that minimize the number of continuous degrees of freedom.
 For example, it learns to represent linear functions with the program
-\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
+\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
 which has five continuous parameters.
-This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
-
-%% The neural recognition model accelerates inference,
-%% but given enough iterations of our algorithm,
-%% a lesioned version without the recognition model eventually comes close to catching up (green curves in Fig.~\ref{regressionLearningCurve}; No NN in Tbl.~\ref{regressionComparison}).
-%% We compare our model on held-out tasks against 4 baselines (Tbl.~\ref{regressionComparison}).
-
-\begin{figure*}[h]
-\makebox[\textwidth][c]{
-\begin{minipage}{0.55\textwidth} \newcommand{\functionSize}{1.5cm}\centering
-%  \includegraphics[width = \functionSize]{figures/functions/0.png}
+This phenomenon arises from our Bayesian framing -- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
+\begin{wrapfigure}{r}{0.45\textwidth} \newcommand{\functionSize}{1.5cm}\centering
   \includegraphics[width = \functionSize]{figures/functions/6.png}
@@
   \includegraphics[width = \functionSize]{figures/functions/149.png}
-  %% \includegraphics[width = \functionSize]{figures/functions/432.png}
-  %% \includegraphics[width = \functionSize]{figures/functions/415.png}
-  %% \includegraphics[width = \functionSize]{figures/functions/426.png}
-  %% \includegraphics[width = \functionSize]{figures/functions/597.png}
-  %% \includegraphics[width = \functionSize]{figures/functions/591.png}
   \caption{Recognition model input for symbolic regression.
 While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
-\end{minipage}
-\hspace{1em}
-\begin{minipage}{0.5\textwidth}
-  \centering\begin{tabular}{l}
-  \toprule
-  $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
-  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
-  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
-  \hspace{0.5cm}\emph{($f_2$: quadratics)}\\
-  $f_3($\code{x}$)\,=\,$\code{(/ (}$f_2$\code{ x) (}$f_0$\code{ x))}\\
-    \hspace{0.5cm}\emph{($f_3$: ratio of polynomials)}
-  \\\bottomrule
-\end{tabular}
-  \captionof{table}{Some learned subroutines for symbolic regression.
-System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
-\end{minipage}
-}
-\end{figure*}
+    \end{wrapfigure}
 
@@
 do not have publicly available code and datasets.
-\\\noindent \textbf{EC}, an implementation of the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, publicly available at...
+\\\noindent \textbf{EC}, which lesions the recognition model and modifies  the DSL learner to model the library learning technique of  EC~\cite{Dechter:2013:BLV:2540128.2540316}.
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
@@
 \renewcommand{\arraystretch}{0.5}
-\begin{tabular}{llllll}
+\begin{tabular}{lllllll}
   \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
     Ours\\(no NN)
-    \end{tabular}}& RF/DC & PCFG & Enum
+    \end{tabular}}& EC&RF/DC & PCFG & Enum
 
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{List Processing}}\\\midrule
-  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
-  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
-
-  \\\midrule\multicolumn{6}{c}{\emph{Text Editing}}\\\midrule
-  \% solved&\textbf{75\%} &\% &33\%&0\%&4\%\\
-  Solve time&  29s&s&80s&--&235s
+  \\\midrule\multicolumn{7}{c}{\emph{List Processing}}\\\midrule
+  \% solved&\textbf{86\%} &84\% &&60\%&74\%&70\%\\
+  Solve time&  0.8s&0.7s&&1.0s&1.0s&1.1s
+
+  \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
+  \% solved&\textbf{75\%} &43\% &&33\%&0\%&4\%\\
+  Solve time&  29s&65s&&80s&--&235s
 
@@
 
-  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
-  \% solved&   \textbf{84\% }&75\%&38\%&38\%&37\% \\
-  Solve time&  24s& 40s  &31s&55s&29s
+  \\\midrule\multicolumn{7}{c}{\emph{Symbolic Regression}}\\\midrule
+  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
+  Solve time&  \textbf{24}s& 40s  &28s&31s&55s&29s
   %% MDL (nats)&   9.3 &11.0&2.3&--&4.6
@@
   \end{tabular}
-\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
+\caption{\% solved before timeout. Solve time: averaged over solved
   tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o

### 2018-05-17 nips_2018.tex
updates to list section

@@
 We consider this problem within the context of learning functions that
-manipulate lists.
-We created 236 Lisp-style list manipulation tasks, each with 15 input/output
-examples (Tbl.~\ref{listexamples}).
-Our data set is challenging along two dimensions: many of the functions are
-very complicated, and the agent must learn to solve these complicated
-problems from only 236 tasks.
-Our data set primarily consists of arithmetic operations upon sequences, and
-so, in addition to the Lisp primitives provided for the text editing
-experiments, we additionally start the system out with the following
-primitives:
-\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}, \code{2},
-\code{3}, \code{4}, \code{5}.
-
-A complete repertoire of higher-order functions is a staple of functional
-programming standard libraries.
-Although we provided \system with some of the standard higher-order
-functions, like \code{foldr} and \code{unfold}, we did not include others.
-When trained on these list functions, our system rediscovers and then reuses
-the higher-order function \code{filter} \textbf{Lucas: not sure if it
-actually does this! It would be cool if we can write something like this
-though.}
-
-%% Our list function domain consists of hand-designed concepts from which we
-%% sample tasks, so there is no ``ground truth'' formal DSL from which these
-%% tasks are generated. Additionally, few of these tasks are concatenative ---
-%% many of them represent distinct operations.
-%% Tasks in this domain are solved by functions that take as input an integer
-%% or a list of integers, and have as output either a Boolean, an integer, a
-%% list of Booleans, or a list of integers.  There are 225 tasks in total, some
-%% of which are shown in Tbl.~\ref{listexamples}.
-
-%% We evaluate \system starting from two different initial DSLs, \emph{base} and
-%% \emph{rich}. The base DSL starts with only
-%% low-level primitives such as \code{mapi}, \code{reducei}, and
-%% \code{if}, whereas the rich DSL includes some common structure that can be
-%% built from these such as \code{all}, \code{filter}, \code{slice}, and
-%% \code{index}.
-%% Both are capable of solving all tasks supplied by our dataset.
-%% See Supplement for details on these DSLs.
-
-%% We found  this domain  difficult to tackle without starting from 
-%% the rich DSL. Our
-%% algorithm, like human learners, requires a spectrum of problems ranging from
-%% easy to hard.
-%% When starting with the base DSL,
-%% we found that there were not enough steppingstones in the curriculum to
-%% get \system off the ground.
-
-%The behavior with the base DSL illustrates this phenomena  because the initial set of
-%primitives is not paired with a sufficiently large and graded curriculum.
-
-\begin{figure*}
-\makebox[\textwidth][c]{
-\begin{minipage}{0.6\textwidth}
+manipulate lists with some involvement of numerical arithmetic.
+
+\begin{wrapfigure}{r}{0.55\textwidth}\centering
   \begin{tabular}{lll}
@@
     Name & Input & Output \\\midrule
-    has-2 & [4\, 5\, 7\, 4] & \code{false} \\
     repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
     drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
     count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
     keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
+    product & [7\, 1\, 6\, 2] & 84 \\
     \bottomrule
   \end{tabular}
-  \captionof{table}{Sample tasks from our list function domain}
+  \captionof{table}{Some tasks in our list function domain}
   \label{listexamples}
-\end{minipage}
-\hfill
-\begin{minipage}{0.6\textwidth}
-  \centering
-  \begin{tabular}{l}
-    \toprule
-    $f_0($\code{i$,\ell$}$) \,=\, $\code{(TODO)}\\
-    \hspace{0.5cm}($f_0$: \emph{TODO})\\
-  \bottomrule
-  \end{tabular}
-  \captionof{table}{Some learned DSL primitives for list functions}\label{listinduced}
-\end{minipage}
-  }
-\end{figure*}
+\end{wrapfigure}
+We created 236 human-interpretable list manipulation tasks, each with 15
+input/output examples (Tbl.~\ref{listexamples}).
+Our data set is challenging in three major ways: many of the tasks require
+complex solutions, the tasks were not generated from some latent DSL, and
+the agent must learn to solve these complicated problems from only 236
+tasks.
+Our data set assumes arithmetic operations as well as sequence operations,
+so we additionally provide our system with the following arithmetic
+primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
+\code{is-prime}.
 
@@
 This prior work presumes a hand-engineered DSL.
-We show \system can instead start out with generic Lisp primitives
-and recover many of the higher-level building blocks that have made these
-other systems successful.
+We show \system can instead start out with generic sequence manipulation
+primitives and recover many of the higher-level building blocks that have
+made these other text editing systems successful.
 

### 2018-05-17 nips_2018.tex
table improvements

@@
   \end{tabular}
-  \\\bottomrule 
-\end{tabular}}\\\\
+  \\\bottomrule\\
+\end{tabular}}
 \caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them.
@@
       \end{tabular}
-\\\bottomrule
+\\\bottomrule\\
   \end{tabular}
@@
 At first, \system cannot find any correct programs for most of the tasks.
-After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
+After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Tbl.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
 all of the tasks. The learned DSL contains \textbf{X} new functions, listed in the supplement.

### 2018-05-17 supplement.tex
supplement

@@
-%%%%%%%% ICML 2018 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
 \documentclass{article}
 
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
-\usepackage{graphicx}
-%\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
+% if you need to pass options to natbib, use, e.g.:
+% \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading nips_2018
 
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
-\usepackage{hyperref}
+% ready for submission
+\usepackage{nips_2018}
 
-% Attempt to make hyperref and algorithmic work together better:
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
+% to compile a preprint version, e.g., for submission to arXiv, add
+% add the [preprint] option:
+% \usepackage[preprint]{nips_2018}
 
-\newcommand{\system}{\textsc{DreamCoder}~}
+% to compile a camera-ready version, add the [final] option, e.g.:
+% \usepackage[final]{nips_2018}
 
-% Use the following line for the initial blind version submitted for review:
-\usepackage{icml2018}
+% to avoid loading the natbib package, add option nonatbib:
+% \usepackage[nonatbib]{nips_2018}
 
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
 \usepackage{hyperref}       % hyperlinks
@@
 \usepackage{microtype}      % microtypography
+\usepackage{stmaryrd}
 
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
+\newcommand{\system}{\textsc{CoCoSea}~}
+\newcommand{\systemEnding}{\textsc{CoCoSea}}
+\newcommand{\lowerBound}{\mathscr{L}}
+\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
+\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
+
+\usepackage{mathrsfs}
 \usepackage{listings}
 \usepackage{amsthm}
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-%\usepackage{epsfig} % less modern
-%\usepackage{subfig}
+
+\usepackage{subfig} 
 \usepackage{fancyvrb}
@@
 \usepackage{caption}
-\usepackage{subcaption}
-
-\fvset{fontsize=\footnotesize}
-
 \usepackage{amssymb}
 \usepackage{listings}
-\lstset{%
-  basicstyle=\footnotesize\ttfamily,
-  breaklines=true}
 \usepackage{wrapfig}
@@
 \usepackage{tikz}
+\usepackage{circuitikz}
 \usetikzlibrary{fit,bayesnet}
-%\usetikzlibrary{arrows.meta}
-%\usetikzlibrary{positioning}
-%\usetikzlibrary{decorations.text}
-%\usetikzlibrary{decorations.pathmorphing}
 \usepackage{dsfont}
 \usepackage{amsmath}
-\usepackage{hyperref}
-%\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
+
+\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
 \DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
-\DeclareMathOperator{\argmin}{argmin} % no space, limits underneath in displays
 
@@
 
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2018}
+\title{Supplement to: Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning}
 
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{\textsc{DreamCoder}: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+% The \author macro works with any number of authors. There are two
+% commands used to separate the names and addresses of multiple
+% authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to
+% break the lines. Using \AND forces a line break at that point. So,
+% if LaTeX puts 3 of 4 authors names on the first line, and the last
+% on the second line, try using \AND instead of \And before the third
+% author name.
+
+\author{
+  David S.~Hippocampus\thanks{Use footnote for providing further
+    information about author (webpage, alternative
+    address)---\emph{not} for acknowledging funding agencies.} \\
+  Department of Computer Science\\
+  Cranberry-Lemon University\\
+  Pittsburgh, PA 15213 \\
+  \texttt{hippo@cs.cranberry-lemon.edu} \\
+  %% examples of more authors
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \AND
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+  %% \And
+  %% Coauthor \\
+  %% Affiliation \\
+  %% Address \\
+  %% \texttt{email} \\
+}
 
 \begin{document}
+% \nipsfinalcopy is no longer used
 
-\twocolumn[
-\icmltitle{Supplement: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}
+\maketitle
 
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2018
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Aeiau Zzzz}{equal,to}
-\icmlauthor{Bauiu C.~Yyyy}{equal,to,goo}
-\icmlauthor{Cieua Vvvvv}{goo}
-\icmlauthor{Iaesut Saoeu}{ed}
-\icmlauthor{Fiuea Rrrr}{to}
-\icmlauthor{Tateu H.~Yasehe}{ed,to,goo}
-\icmlauthor{Aaoeu Iasoh}{goo}
-\icmlauthor{Buiui Eueu}{ed}
-\icmlauthor{Aeuia Zzzz}{ed}
-\icmlauthor{Bieea C.~Yyyy}{to,goo}
-\icmlauthor{Teoau Xxxx}{ed}
-\icmlauthor{Eee Pppp}{ed}
-\end{icmlauthorlist}
-
-\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
-\icmlaffiliation{goo}{Googol ShallowMind, New London, Michigan, USA}
-\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
-
-\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
-\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Machine Learning, ICML}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
-\section{An Illustration of the Wake-Sleep Cycle}
-Below we diagram the wake/sleep cycles employed by our algorithm. At each stage of the cycle,
+\section{An Illustration of the 3 iterations of our algorithm}
+Below we diagram the iterations employed by our algorithm. At each stage of the algorithm,
 we have shaded the observed variables in gray and left the unobserved variables white.
@@
     \node[obs] at (3.5,0.7) (xp) {$x$};
-    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
+    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Search: Infer $p$};
     \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
@@
     \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
+    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
   \end{scope}
@@
     \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
   \end{scope}
@@
 \end{algorithm}
-\vspace{-1.8cm}    \end{minipage}
+    \end{minipage}
   \end{wrapfigure}Alg.~\ref{programGenerativeModel} is a procedure for drawing
@@
 Each input/output pair is processed by an identical encoder network;
-the outputs of the encoders are max pooled and passed to an MLP with 1 hidden layer, 16 hidden units, and a ReLU activation:
+the outputs of the encoders are average and passed to an MLP with 1 hidden layer, 32 hidden units, and a ReLU activation:
 \begin{align}
-  q(x) = \text{MLP}\left(\text{MaxPool}\left(\left\{\text{encoder}\left(i_n,o_n \right) \right\}_{n\leq N} \right) \right)
+  q(x) = \text{MLP}\left(\text{Average}\left(\left\{\text{encoder}\left(i_n,o_n \right) \right\}_{n\leq N} \right) \right)
 \end{align}
-For the Boolean circuits and symbolic regression domains,
-the inputs and outputs have a fixed dimensionality,
-and the set of possible inputs is prespecified.
-For these fixed-dimensionality domains our encoder is an MLP with 1 hidden layer having 16  units and a ReLU activation function.
-The input to the encoder is
-the sequence of program outputs $\left\{o_n \right\}_{n\leq N}$.
 
@@
 the inputs and outputs are sequences.
-Our encoder for these domains is a bidirectional GRU with 16 hidden units that reads each input/output pair; we concatenate the input and output along with a special delimiter
+Our encoder for these domains is a bidirectional GRU with 64 hidden units that reads each input/output pair; we concatenate the input and output along with a special delimiter
 symbol between them.
 We MaxPool the final hidden unit activations in the GRU along both passes of the bidirectional GRU.
+
+For symbolic regression,
+the input/outputs are densely sampled points along the curve of the function.
+We rendered these points to a graph,
+and pass the image of the graph to a convolutional network,
+which acts as the encoder.
 
@@
      \WHILE {true}
-     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\text{ subexpression of }p\}$
+     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\text{ a fragment  of }p\}$
      \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
@@
 We are only maximizing a lower bound on the log posterior; when is this lower bound tight?
-Whenever the DSL does not contain polymorphic types,
-
 This lower bound is tight whenever all
@@
 
+\begin{comment}
 \section{Initial Domain-Specific Languages}
@@
 \end{center}
-
+  \end{comment}
 \vfill
@@
 smoothness parameter $\alpha = 10$ (Eq.~6 of the main paper)
-for all experiments except for string editing.
-For string editing we found that more strongly regularizing
-the DSL was necessary,
-so we set $\lambda = 5$.
-We believe this is because
-we have many more tasks for string editing than we do for the other domains.
-The frontier size used for Boolean circuits was $5000$,
-for symbolic regression we used $500$,
-for string editing we used $10^5$,
-and for list functions we used $10^4$.
+for all experiments.
+For list processing and text editing we used a search timeout of two hours;
+because the symbolic regression problems are easier,
+we used a timeout of only five minutes for these.
 
@@
 and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
-We explain in the Supplement that maximizing an ELBO-style bound
+Surprisingly, maximizing an ELBO-style bound
 leads to a pathological behavior that  causes the model to easily become trapped in local optima.
@@
 We do not maximize a bound of this form because it
-takes an expectation over the \emph{previous} sleep cycle's
+takes an expectation over the \emph{previous} iteration's
 posterior over programs,
@@
 
+\begin{comment}
+b
 \section{Example String Editing Tasks}
@@
 \onecolumn
-\subsection{Boolean circuits}
-
-\begin{lstlisting}
-#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2))))))
-#(\ (\ (nand (nand $0 $1) (nand $0 $1))))
-#(\ (\ (nand (nand $0 $0) (nand $1 $1))))
-#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 $3))))))
-#(\ (\ (\ (nand (nand $0 $0) (nand $1 (nand $2 $1))))))
-#(\ (\ (#(\ (\ (nand (nand $0 $1) (nand $0 $1)))) $0 (nand $1 $0))))
-#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0))))))
-#(\ (\ (#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0)))))) $0 $1 (nand $1 $1))))
-#(\ (\ (nand $0 (nand $1 $0))))
-#(\ (#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2)))))) (nand $0 $0)))
-#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $3 $2)))))))
-\end{lstlisting}
-
 \subsection{Symbolic regression}
 \begin{lstlisting}
-#(\ (* (+ REAL $0) REAL))
-#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0)))
-#(\ (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))
-#(\ (#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0))) (+ $0 $0)))
-#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0)))
-#(\ (* (#(\ (+ $0 REAL)) $0) (#(\ (+ (* $0 (* $0 REAL)) REAL)) $0)))
-#(\ (+ (* $0 (* $0 REAL)) REAL))
-#(* REAL)
-#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0))
-#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0))))
-#(\ (+ $0 REAL))
+#(\ (#(\ (+. $0 REAL)) (*. $0 REAL)))
+#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0)))
+#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0)))
+#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0)))
+#(/. REAL)
+#(\ (/. (/. REAL $0) $0))
+#(\ (+. $0 REAL))
+#(\ (#(\ (+. $0 REAL)) (*. $0 (#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0))) $0))))
+#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0))
+#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1))))
+#(\ (#(\ (+. $0 REAL)) (#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0))) $0)))
+#(#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1)))) (#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0)) REAL))
+#(\ (/. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) (#(\ (+. $0 REAL)) $0)))
 \end{lstlisting}
@@
 \begin{lstlisting}
-#(\ (\ (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0)))))
-#(\ (\ (join (chr->str $0) (map (\ (strip $0)) $1))))
-#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0))))
-#(\ (#(join empty) (split '.' $0)))
-#(#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2))))) (decr 0))
-#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (slice $2 (incr $2) $0)))))
-#(#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (incr 0))
-#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2)))))
-#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1))))))
-#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr $0)))
-#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (slice $0 (incr (incr $0)))) $2 $0)))))
-#(\ (\ (slice $0 (incr $0) (nth (incr $0) $1))))
-#(\ (\ (#(\ (\ (slice $0 (len $1) $1))) (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1) (incr 0))))
-#(\ (\ (caseUpper (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0))))))
-#(\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (nth 0 $1)) $0)))
-#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr (decr $0))))
-#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1))))
-#(\ (\ (#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0)))) (split '.' (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1)) '.')))
-#(\ (\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1))))) $1 $2 0)) $0))))
-#(\ (\ (slice $0 (incr (incr $0)) (#(nth (incr 0)) $1))))
-#(join empty)
-#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1)))))
-#(\ (\ (nth 0 (split $0 (#(nth (incr 0)) $1)))))
-#(\ (slice $0 (incr (incr $0))))
-#(nth (incr 0))
-#(\ (join empty (map (\ (caseCapitalize $0)) $0)))
-#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2)))))
-#(\ (\ (slice $0 (len $1) $1)))
-#(\ (\ (\ (join (chr->str $0) (map $1 $2)))))
-#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1))))
+#(\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
+#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0))))))
+#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0)))))))
+#(\ (\ (#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0)))))) (cons $0 $1))))
+#(\ (\ (fold $0 $0 (\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
+#(\ (#(\ (\ (\ (cons (car $0) (cons $1 $2))))) (#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty) '.' $0) '.'))
+#(\ (\ (map (\ (if (char-eq? $1 $0) $2 $0)))))
+#(\ (map (\ (index $0 $1))))
+#(\ (unfold $0 (\ (empty? $0)) (\ (car $0)) (\ (#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0))))))) SPACE $0))))
+#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty)
 \end{lstlisting}

### 2018-05-17 supplement.tex
First geom addition to supplement

@@
 \maketitle
+
+\section{Learning abstract geometry}
+The brifly presented geometry project is a recent addition to the set of
+problems adressed with our method. The original DSL, while inspired by the
+Logo~\cite{abelson1974logo} in the sense that ones controls a moving on-off
+point, was designed to be congitively plausible in the sense that some
+abstractions are baked in for curves and a few structures --- it was first
+described in~\cite{sablemeyer2017geom}.
+
+It is to be thought as a turtle with as internal state a position, a facing
+direction, a speed, an acceleration, an angular speed and an angular
+acceleration.
+
+The DSL is the following:
+
+\begin{center}
+\tt
+\begin{tabular}{rr}
+  \toprule
+  \textrm{\emph{name}}    & \textrm{\emph{type}} \\
+  \midrule
+  Concat    & \text{prog}      \to \text{prog}      \to \text{prog} \\
+  Repeat    & var option        \to \text{prog}      \to \text{prog} \\
+  Embed     & \text{prog}      \to \text{prog}                       \\
+  Define    & \text{var}        \to \text{prog}                       \\
+  Turn      & \text{var option} \to \text{prog}                       \\
+  Integrate & \text{var option} \to \text{bool}       \to              \\
+            & \text{var option} \to \text{var option} \to              \\
+            & \text{prog}                                             \\
+  \midrule
+  True      & bool                                                     \\
+  False     & bool                                                     \\
+  \midrule
+  Nothing   & \text{var option}                                        \\
+  Just      & \text{var}        \to \text{var option}                  \\
+  \midrule
+  Unit      & var                                                      \\
+  Name      & var                                                      \\
+  Next      & \text{var}        \to \text{var}                         \\
+  Prev      & \text{var}        \to \text{var}                         \\
+  Double    & \text{var}        \to \text{var}                         \\
+  Half      & \text{var}        \to \text{var}                         \\
+  Opposite  & \text{var}        \to \text{var}                         \\
+  \bottomrule
+\end{tabular}
+\end{center}
+
+In this context programs to be learned are directly resulting shapes, since they
+don't take an input, and all tasks will be unique in the sense that each program
+ignores input and produces a single output.
+
+Some elements of the semantics are common, the others are as follow.
+\texttt{Repeat} takes a \texttt{variable} and a \texttt{prog} and repeats
+said \texttt{prog} $n$ times where $n$ is the evaluation of the
+\texttt{variable} --- if not set it is defaulted to two. \texttt{Embed} of a
+\texttt{prog} means that said \texttt{prog} will be executed and then returns to
+the current state --- leaving what has been drawn in the meantime on the canvas.
+\texttt{Integrate} is the main instruction and the only one that draws anything:
+\texttt{Integrate(t,p,a,c)} takes a time \texttt{var}, a pen \texttt{bool}, an
+acceleration \texttt{var} and an angular speed \texttt{var}, and it moves the
+turtle according to these parameters --- with or without actually drawing
+depending on the pen variable \texttt{p}.
+
+\begin{wrapfigure}{r}{0.25\textwidth}
+  \begin{center}
+    \includegraphics[width = 0.25\textwidth]{figures/DSLGeom.eps}
+  \end{center}\label{dslgeom}\caption{Some defaults for \texttt{Integrate}}
+\end{wrapfigure}
+
+The default values of these parameters are set such that \texttt{Integrate(nothing,
+true, nothing, nothing)} draws a unit segment, \texttt{Integrate(nothing, true,
+nothing, Just(Unit))} draws a circle of unit length, and
+\texttt{Integrate(nothing, true, Just(Unit), Just(unit))} draws the first spire
+of a spiral.
+
+The training was done on the following set of tasks
+
+\begin{figure}[h]
+\centering
+\includegraphics[width = \textwidth]{figures/geomAllTasks.png}
+  \caption{The set of tasks for the geometry domain}
+\end{figure}
 

### 2018-05-17 nips_2018.tex
induced text table fix
latest for Josh

@@
   in the learned DSL.  We use our model to synthesize functions on lists,
-  edit strings, and solve symbolic regression problems, showing how the
+  edit text, and solve symbolic regression problems, showing how the
   model learns a domain-specific library of program components for
@@
       \end{tabular}&
-    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
+    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x a) b x))) s)}
       \end{tabular}
@@
 
-We compare  with four baselines on  held-out tasks:
+We compare with ablations of our model on held out tasks.
+Each of these ablations is chosen to approximate
+a prior approach in the program learning literature:
 \\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.

### 2018-05-18 nips_2018.tex
greek lettering for proposed expressions
bunches of changes

@@
 \usepackage{amsmath}
+\usetikzlibrary{trees}
 
@@
   \scriptsize
+  \tabcolsep=4pt
 \renewcommand\code\texttt
 \renewcommand\codechar[1]{\texttt{"#1"}}
+\newcommand{\helpSize}{0.25cm}
 \begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
-  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
+  {\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
+  {  \tabcolsep=4pt \begin{tabular}{cc}
   \begin{tabular}{c}
-      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
-      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
-      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
-      % \code{[99\, 4\, 2]}$\to $\code{99}\\
-      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
-      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
-      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
-      \code{[1]}$\to $\code{[1\, 3]}\\
-      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
-      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
-      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
-      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
+      \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
+      \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
+      \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
+      \code{[2\, 7\, 8\, 1]}$\to $\code{8}\\
+      \code{[3\, 19\, 14]}$\to $\code{19}\\
+      \code{$f(\ell) = $}\code{($f_0$ $\ell$)}
+      %% \code{[1]}$\to $\code{[1\, 3]}\\
+      %% \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
+      %% \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
+  \end{tabular}&
+  \begin{tabular}{c}
+          \code{[7\, 3]}$\to $\code{\#f}\\
+          \code{[9\, 4\, 4]}$\to $\code{\#t}\\
+          \code{[]}$\to $\code{\#f}\\
+          \code{[4\, 7\, 3]}$\to $\code{\#t}\\
       \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
-  \end{tabular}&
+    \end{tabular}
+  \end{tabular}}
+  &
   \begin{tabular}{c}
@@
     % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
-      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
+      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
     % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
-      % \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
+      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
     %NOTE: this is the actual invention, but I removed a redundant lambda
@@
     \vspace{2ex}
-      \hspace{0.5cm}($f_0$: \emph{Concatenate }\code{r}\emph{ with $\ell$})\\
+      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
     $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
     \vspace{2ex}
-      \hspace{0.5cm}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
+      \hspace{\helpSize}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
     $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
@@
     \code{($\lambda$ (x a) (if a a (= k x))))}\\
-      \hspace{0.5cm}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
+      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
   \end{tabular}&
@@
           $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
-      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
+      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
       $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
-        \hspace{0.5cm}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
+        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
       $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
-        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
+        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
     \end{tabular}&
@@
     $f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}\\
-    \hspace{0.5cm}\emph{($f_4$: 4th order polynomial)}\\
+    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
         $f_5($\code{x}$)\,=\,$\code{(/ real x)}\\
     $f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))} \\
-    \hspace{0.5cm}\emph{($f_6$: rational function)}\\
+    \hspace{\helpSize}\emph{($f_6$: rational function)}\\
 
@@
 \end{tabular}}
-\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them.
-Bottom: Subset of learned DSL.
+\caption{Top: Tasks from each domain, each followed by the programs \system discovers for them.
+Bottom: Several examples from learned DSL.
 Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
@@
 and train $q$ on those samples.
-  \end{comment}
+
 Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
@@
 from the empirical observed distribution of inputs in $X$.
+  \end{comment}
 
@@
         %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Search})}
+        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
+        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
+        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
         \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Search)}}
+        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
+        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
+        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}
 
@@
 
-\begin{table}[b]
-\centering  \begin{tabular}{ll}
+\begin{figure}[b!]\tabcolsep=2pt
+\centering  \begin{minipage}[c]{0.2\textwidth}
+  \begin{tikzpicture}[scale=0.7]
+    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
+    %%     \footnotesize Program:\\
+    %%     \code{($\lambda$ (x) (+ (- x) 1))}
+    %% \end{tabular}};
+    %\node at (,0.5) {\code{cons}};
+    \node [rotate=90] at (-2.3,-0.5) {\small program};
+
+          \node(l1) at (0,0) {};
+  \node[color=red](p1) at (-1,-1) {\code{+}};
+  \node[color=red](n1) at (0.7,-0.9) {\code{1}};
+  \node(x1) at (0,-1) {\code{1}};
+  \draw[color=red] (l1.south) -- (p1.north);
+  \draw[color=red] (l1.south) -- (n1.north);
+  \draw[color=red] (-0.5,-0.5) -- (x1.north);
+
+  \node(t) at (-0.5,0.5) {};
+  \draw (l1.south) -- (t.south);
+  \node(c) at (-1.5,-0.2) {\code{cons}};
+  \draw (t.south) -- (c.north);
+
+%    \draw  (l1.south) -- (-0.5,0.5);
+
+  %% \node(c) at (-0.5,-1.5) {\code{-}};
+  %% \node(z) at (0.5,-1.5) {\code{x}};
+
+  %% \draw (0,-1) -- (c.north);
+  %% \draw (0,-1) -- (z.north);
+
+  \begin{scope}[shift={(0,-2.5)}]
+      \node(l1) at (0,0) {};
+  \node[color=red](p1) at (-1,-1) {\code{+}};
+  \node[color=red](n1) at (0.7,-0.9) {\code{1}};
+  %\node(x1) at (0,-1) {};
+  \draw[color=red] (l1.south) -- (p1.north);
+  \draw[color=red] (l1.south) -- (n1.north);
+  \draw[color=red] (-0.5,-0.5) -- (0,-1);
+
+  \node(c) at (-0.5,-1.5) {\code{car}};
+  \node(z) at (0.5,-1.5) {\code{z}};
+
+  \draw (0,-1) -- (c.north);
+  \draw (0,-1) -- (z.north);
+
+  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+  \end{scope}
+
+\begin{scope}[shift={(0,-5)}]
+  \node[red](p1) at (-1,-1) {\code{+}};
+  \node[red](n1) at (0.8,-0.7) {\code{1}};
+  \node[red](a) at (0,-1) {\code{ }};
+  %\node(x1) at (0,-1) {};
+  \draw[red] (0,0) -- (p1.north);
+  \draw[red] (0,0) -- (n1.north);
+  \draw[red] (-0.55,-0.4) -- (a.north);
+  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
+
+  \end{scope}
+
+\end{tikzpicture}
+  \end{minipage}
+\hspace{0.1cm}\begin{tabular}{ll}
     \toprule
@@
     \begin{tabular}{l}
-      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
-      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
-      \code{($\lambda$ (a b) (foldr a b }\\
-      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
+      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
+      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
+      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
+      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
     \end{tabular}&
     \begin{tabular}{l}
-      \code{(foldr a b ($\lambda$ (x z) }\\
-      \phantom{\code{(foldr a b }}\code{(cons x z)))}
+      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
+      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
     \end{tabular}\\\midrule
@@
       \end{tabular}&
-    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x a) b x))) s)}
+    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
       \end{tabular}
@@
 \end{tabular}
-  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
-    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
+\caption{\textbf{Left:} syntax trees of two programs sharing common structure, highlighted in {\color{red} red},
+  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
+  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}
+
   \label{fragmentExample}
-\end{table}
+\end{figure}
 
@@
 After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Tbl.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
-all of the tasks. The learned DSL contains \textbf{X} new functions, listed in the supplement.
+all of the tasks. The learned DSL contains 10 new functions, listed in the supplement.
 
@@
 a prior approach in the program learning literature:
-\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
-\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
-a recognition model only from samples from the fixed generative model.
-This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
-and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
-We call this baseline RF/DC because
-this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-We can not compare directly with these systems,
-because they are engineered for one specific domain, and 
-do not have publicly available code and datasets.
-\\\noindent \textbf{EC}, which lesions the recognition model and modifies  the DSL learner to model the library learning technique of  EC~\cite{Dechter:2013:BLV:2540128.2540316}.
+\\\noindent \textbf{No NN}, which lesions the recognition model.
+\\\noindent \textbf{NPS}, which does not learn the DSL,
+instead learning the recognition model
+from samples drawn from the fixed DSL.
+We call this NPS (Neural Program Synthesis)
+because this setup is closest to how
+RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
+%% We can not compare directly with these systems,
+%% because they are engineered for one specific domain, and 
+%% do not have publicly available code and datasets.
+\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
+only add subexpressions of programs in the frontiers to the DSL.
+This is how prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
@@
 Tbl.~\ref{baselineComparisons}
-compares our model against these baselines.
+compares our model against these alternatives.
 Our full model consistently
@@
 and lesioning it also slows down the convergence of the algorithm,
-
 taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
@@
 \begin{tabular}{lllllll}
-  \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
-    Ours\\(no NN)
-    \end{tabular}}& EC&RF/DC & PCFG & Enum
+  \toprule& Ours& 
+    No NN
+& SE&NPS & PCFG & Enum
 
@@
 \caption{\% solved before timeout. Solve time: averaged over solved
-  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
+  tasks.  SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
   structure learning.  Enum: model w/o any learning.  %% MDL:

### 2018-05-18 nips_2018.tex
Alignment and stuff
Some work on the table

@@
   \newcommand{\helpSize}{0.25cm}
-\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
+  \begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
+    \toprule
     {\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
-  {  \tabcolsep=4pt \begin{tabular}{cc}
+    {\tabcolsep=4pt
+      \begin{tabular}{cc}
         \begin{tabular}{c}
-      \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
-      \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
-      \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
-      \code{[2\, 7\, 8\, 1]}$\to $\code{8}\\
-      \code{[3\, 19\, 14]}$\to $\code{19}\\
+          \code{[7\, 0\, 2]}$\to $\code{[7\, 0\, 2\, 4]}         \\
+          \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2\, 4]} \\
+          \code{$f(x) = $}\code{($f_0$ nil $x$)}             \\
+          ~                                                  \\
+          ~                                                  \\
+          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
+          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
           \code{$f(\ell) = $}\code{($f_0$ $\ell$)}
-      %% \code{[1]}$\to $\code{[1\, 3]}\\
-      %% \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
-      %% \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
-  \end{tabular}&
-  \begin{tabular}{c}
-          \code{[7\, 3]}$\to $\code{\#f}\\
-          \code{[9\, 4\, 4]}$\to $\code{\#t}\\
-          \code{[]}$\to $\code{\#f}\\
-          \code{[4\, 7\, 3]}$\to $\code{\#t}\\
+        \end{tabular}
+        &
+        \hspace{-0.3cm}\begin{tabular}{c}
+          \code{[7\, 3]}$\to $\code{False}                              \\
+          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
+          \hspace{-0.3cm}\code{[9\, 4\, 4]}$\to $\code{True\phantom{e}} \\
+          \hspace{0.3cm}\code{[4]}$\to $\code{True\phantom{e}}                        \\
+          \hspace{-0.3cm}\code{[4\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
           \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
         \end{tabular}
-  \end{tabular}}
+      \end{tabular}
+    }
     &
-  \begin{tabular}{c}
+    \hspace{-0.5cm}\begin{tabular}{c}
       +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
@@
       $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
-    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
+      \hspace{1.25cm}\code{($f_0$ \codechar{.} \codechar{ }}\\
+      \hspace{1.4cm}\code{(cdr s)))}\\
+      ~\\
       Temple Anna H $\to $TAH\\
@@
       $f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\
-
-  \end{tabular}&
-
+    \end{tabular}
+    &
     \begin{tabular}{cc}
@@
       \code{$f($x$) = $($f_1$ x)}&    \code{$f($x$) = $($f_6$ x)}\\
+      ~\\
       \includegraphics[width = 3em]{figures/functions/112.png}&
@@
     \end{tabular}
-
-  \\\midrule
-
-  \begin{tabular}{l}
+    ~\\
+    \midrule
+    \hspace{1cm}\begin{tabular}{l}
       % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
@@
 
-  \begin{tabular}{l}
+  \hspace{0.5cm}\begin{tabular}{l}
     $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\

### 2018-05-18 nips_2018.tex
data update and different figures

@@
 How well does the  learned DSL generalized to real text-editing scenarios?
-We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+We tested, but did not train, our system on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
 \system solves 3.7\% of the problems with an average search time of 235 seconds.
@@
 it solves 74.1\%, and does so much faster,
-solving them in an average of 29 seconds.
+solving them in an average of 38 seconds.
 As of the 2017 SyGuS competition,
@@
   \includegraphics[width = \functionSize]{figures/functions/6.png}
-  \includegraphics[width = \functionSize]{figures/functions/116.png}
+  \includegraphics[width = \functionSize]{figures/functions/48.png}
+  \includegraphics[width = \functionSize]{figures/functions/102.png}
+  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
+  \includegraphics[width = \functionSize]{figures/functions/181.png}
   \includegraphics[width = \functionSize]{figures/functions/160.png}
   \includegraphics[width = \functionSize]{figures/functions/149.png}
+    \includegraphics[width = \functionSize]{figures/functions/187.png}
   \caption{Recognition model input for symbolic regression.
-While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
+DSL learns subroutines for polynomials (top row) and polynomials (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}
     \end{wrapfigure}

### 2018-05-18 nips_2018.tex
updated super figure
Almost midnight commit

--- 2018-05-18 nips_2018.tex
@@
 including the three domains considered,  but also many others?
-}\hfill\begin{minipage}{0.4\textwidth}\centering
-  \includegraphics[width=0.8\textwidth]{figures/geomCompiled.eps} 
+}\hfill
+\begin{minipage}{0.4\textwidth}
+  \centering
+  \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
   \captionof{figure}{
-  Top left: hand crafted traning targets. Top right: examples of
-  discovered compiled new programs. Bottom: compiled programm across iterations
-  to highlight structure emergence.
-  }\label{geomCompiled}
+  Top left: a few example of the shapes used as targets. Top right: a few
+  examples of discovered new shapes. Bottom: learning curve across iterations in
+  percent in the middle, and on both side average of the compiled new shapes
+  before any traning and at the end of the last iteration. See supplementary
+  material for more details.}
+\label{geomCompiled}
 \end{minipage}

### 2018-05-18 supplement.tex
Almost midnight commit

@@
 The brifly presented geometry project is a recent addition to the set of
-problems adressed with our method. The original DSL, while inspired by the
-Logo~\cite{abelson1974logo} in the sense that ones controls a moving on-off
-point, was designed to be congitively plausible in the sense that some
-abstractions are baked in for curves and a few structures --- it was first
-described in~\cite{sablemeyer2017geom}.
+problems adressed with our method.
+
+The original DSL, while inspired by the Logo~\cite{abelson1974logo} in the sense
+that ones controls a moving on-off point, was designed to be congitively
+plausible in the sense that some abstractions are baked in for curves and a few
+structures --- it was first described in~\cite{sablemeyer2017geom}.
 
@@
 \tt
-\begin{tabular}{rr}
+\begin{tabular}{rl}
   \toprule
@@
 
-In this context programs to be learned are directly resulting shapes, since they
-don't take an input, and all tasks will be unique in the sense that each program
-ignores input and produces a single output.
+In this context programs to be learned correspond to a single shape since they
+don't take an input. All tasks will ignore input and produces a single output.
+This differs from the previous examples where each target could be evaluated on
+various input.
 
@@
 the current state --- leaving what has been drawn in the meantime on the canvas.
+
+\begin{wrapfigure}{r}{0.25\textwidth}
+  \begin{center}
+    \includegraphics[width = 0.25\textwidth]{figures/DSLGeom.eps}
+  \end{center}\label{dslgeom}\caption{Top: Some defaults for \texttt{Integrate}.
+  Bottom, the result of the \texttt{Cross} example bellow.}
+\end{wrapfigure}
+
 \texttt{Integrate} is the main instruction and the only one that draws anything:
@@
 
-\begin{wrapfigure}{r}{0.25\textwidth}
-  \begin{center}
-    \includegraphics[width = 0.25\textwidth]{figures/DSLGeom.eps}
-  \end{center}\label{dslgeom}\caption{Some defaults for \texttt{Integrate}}
-\end{wrapfigure}
-
 The default values of these parameters are set such that \texttt{Integrate(nothing,
@@
 \texttt{Integrate(nothing, true, Just(Unit), Just(unit))} draws the first spire
-of a spiral.
+of a spiral. Playing with the first arguments decides on the duration during
+which the arguments are integrated.
 
-The training was done on the following set of tasks
+In the \texttt{var} type, most are self explicit, while \texttt{Opposite(v)}
+takes the opposite of \texttt{v}. The behaviour of \texttt{Name} was originally
+designed to handle arbitrary variables in a call by name fashion but was latter
+reduced to a single storage location, which proves to be enough for the
+targeted shapes. One can therefore store elements in that placeholder using
+\texttt{Define} and retrieve it through \texttt{Name}.
+
+Because of these defaults, the following program draws a cross:
+
+\[
+  \text{\texttt{Cross = Repeat(Double(Double(Unit)), Concat(Embed(Integrate), Turn(None)))}}
+\]
+
+Describing highly regular complex shapes in this language is easy to do as a
+human but quickly escapes the reach of naïve enumeration search. By using a
+curriculum of shapes our approach compresses the search in the corresponding
+directions --- another way to say this is that upon being given a dataset of
+shapes, it picks up the simple ones and abstract them as building blocks for
+latter staged of search.
+
+For example, the simplest way to draw a square in this language is already of
+length 12. Placing several around, for example to draw a grid is out of reach of
+the initial search. However by first abstracting as a primitive the segment ---
+thus reducing the length by four ---, then assuming that after a segment it
+often needs to draw something else, then abstracting the first arguments of
+\texttt{Repeat} --- further reducing the length by two --- as well as the one of
+\texttt{Turn} and finally making the square a primitive on its own once it
+starts using it often enough, the length of this particular shape drops.
+
+\begin{table}[htbp]
+  \begin{center}
+    \begin{tabular}{rcl}
+      \toprule
+      New Primitive & Type & Definition \\
+      \midrule
+      Segment = $f_0$ & \texttt{prog} & \texttt{Concat(Nothing, True, Nothing, Nothing)} \\
+      Right-angle = $f_1$ & \texttt{prog} & \texttt{Turn(Nothing)} \\
+      RepeatTwice = $f_2$ & \texttt{prog} $\to$ \texttt{prog} & $\lambda p.$ \texttt{Repeat(nothing,} $p$) \\
+      AddToUnitSegment = $f_3$ & \texttt{prog} $\to$ \texttt{prog} & $\lambda p.$ \texttt{Concat(}$f_0, p$) \\
+      Square = $f_4$ & \texttt{prog} & $f_3(f_3(f_2(f_1)))$ \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{How our method compresses the square step by step. On the example
+  given in the main article this was produced by the compressor after the second
+  search phase and leads to the second jump in success, the first one being the
+  abstraction of the Segment. Names are \emph{not} produced by the compressor
+  and are here as indication to help the reader.}
+\end{table}
+
+The underlying hypothesis is that the new primitive distort the space of search
+toward something that looks more like what human actually produce and moves away
+from semantically valid but meaningless programs --- in a sense, learns to care
+about what \emph{matters} rather than what is \emph{true} in a very
+pragmatic-like way.
+
+In Figure \ref{alltasksgeom} are listed all the shapes used for this project
+without particular order.
 
@@
 \caption{The set of tasks for the geometry domain}
+\label{alltasksgeom}
 \end{figure}
+
+Since this is a first step in broader project of program induction for abstract
+geometry the likelyhood is currently all-or-none --- ongoing work moves this to
+a neural net based distance function to abstract away from noise in the shapes.
+
+\newpage
 
@@
 \centering
-\begin{tikzpicture}
-  \begin{scope}[shift={(2.5,0)}]
-    \node[obs] at (3,3) (dx){$\mathcal{D}$};
-    \node[latent] at (3.5,1.75) (zp){$p$};
-    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-    \node[obs] at (3.5,0.7) (xp) {$x$};
-    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Search: Infer $p$};
-    \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
-    \draw [->,red] (tx) -- (zp);
-    \draw [->] (dx) -- (zp);
-    \node at (nn) {
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-  \end{scope}
+  \includegraphics{figures/iterations.eps}
+%\begin{tikzpicture}
+  %\begin{scope}[shift={(2.5,0)}]
+    %\node[obs] at (3,3) (dx){$\mathcal{D}$};
+    %\node[latent] at (3.5,1.75) (zp){$p$};
+    %\node[obs] at (4,3) (tx){$\theta^{(x)}$};
+    %\node[obs] at (3.5,0.7) (xp) {$x$};
+    %\node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Search: Infer $p$};
+    %\draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
+    %\draw [->,red] (tx) -- (zp);
+    %\draw [->] (dx) -- (zp);
+    %\node at (nn) {
+      %\begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        %\tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        %\fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        %\node[rectangle] at (1,1) {};
+        %\foreach \name / \y in {1,...,4}
+            %\node[neuron] (I-\name) at (0,-\y) {};
+        %\foreach \name / \y in {1,...,3}
+            %\node[neuron] (H-\name) at (1,-\y-0.5) {};
+        %\foreach \name / \y in {1,...,4}
+            %\node[neuron] (O-\name) at (2,-\y) {};
+        %\foreach \source in {1,...,4}
+            %\foreach \dest in {1,...,3}
+                %\draw [-latex] (I-\source) -- (H-\dest);
+        %\foreach \source in {1,...,3}
+            %\foreach \dest in {1,...,4}
+                %\draw [-latex] (H-\source) -- (O-\dest);
+      %\end{tikzpicture}
+    %};
+    %\node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
+  %\end{scope}
 
-  \begin{scope}[shift={(0.7,-4.2)}]
-    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-    \draw [->] (p2.south) -- (x2.north);
-    \draw [->] (dt.south) -- (p2.north);
+  %\begin{scope}[shift={(0.7,-4.2)}]
+    %\node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+    %\draw [->] (p2.south) -- (x2.north);
+    %\draw [->] (dt.south) -- (p2.north);
 
-    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-    \draw [->] (p1.south) -- (x1.north);
-    \draw [->] (dt.south) -- (p1.north);
+    %\node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+    %\draw [->] (p1.south) -- (x1.north);
+    %\draw [->] (dt.south) -- (p1.north);
 
-    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-    \draw [->] (p3.south) -- (x3.north);
-    \draw [->] (dt.south) -- (p3.north);
+    %\node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+    %\draw [->] (p3.south) -- (x3.north);
+    %\draw [->] (dt.south) -- (p3.north);
 
-    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+    %\node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+    %\node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
 
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-    \draw [->] (zp.south) -- (xp.north);
-    \plate {}{(zp)(xp)}{$x\in X$};
+    %\node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+    %\node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+    %\draw [->] (zp.south) -- (xp.north);
+    %\plate {}{(zp)(xp)}{$x\in X$};
 
-    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
-  \end{scope}
+    %\draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
+    %rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
+    %\node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
+  %\end{scope}
 
-  \begin{scope}[shift={(7.7,-4)}]
-    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-    \node[latent] at (1.5,3) (t){$\theta$};
-    \node[obs] at (1,1.75) (z){$p$};
-    % \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-    \node[obs] at (1,0.5) (x) {$x$};
-    \edge {z}{x};
-    \edge {d,t}{z};
-    \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
-  \end{scope}
+  %\begin{scope}[shift={(7.7,-4)}]
+    %\node[latent] at (0.5,3) (d){$\mathcal{D}$};
+    %\node[latent] at (1.5,3) (t){$\theta$};
+    %\node[obs] at (1,1.75) (z){$p$};
+     %\node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+    %\node[obs] at (1,0.5) (x) {$x$};
+    %\edge {z}{x};
+    %\edge {d,t}{z};
+    %\plate {}{(z)(x)}{$x\in X$};
+    %\node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
+  %\end{scope}
 
-\end{tikzpicture}
+%\end{tikzpicture}
 \end{figure}
@@
 \section{Generative model over the programs}
+
+Alg.~\ref{programGenerativeModel} is a procedure for drawing
+samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
+enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.
+
 \begin{wrapfigure}{L}{0.5\textwidth}
@@
     \end{minipage}
-  \end{wrapfigure}Alg.~\ref{programGenerativeModel} is a procedure for drawing
-samples from the generative model $(\mathcal{D},\theta)$.  In practice, we
-enumerate programs in order of their probability under  Alg.~\ref{programGenerativeModel} rather than sample them.
+  \end{wrapfigure}
 

### 2018-05-18 nips_2018.tex

--- 2018-05-18 nips_2018.tex
@@
         \begin{tabular}{c}
-          \code{[7\, 0\, 2]}$\to $\code{[7\, 0\, 2\, 4]}         \\
-          \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2\, 4]} \\
-          \code{$f(x) = $}\code{($f_0$ nil $x$)}             \\
+          \code{[7\, 0\, 2]}$\to $\code{[7\, 0\, 2\, 9]}         \\
+          \code{[9\, 3]}$\to $\code{[9\, 3\, 9]} \\
+          \code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}             \\
           ~                                                  \\
@@
           \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
-          \code{$f(\ell) = $}\code{($f_0$ $\ell$)}
+          \code{$f(\ell) = $}\code{($f_2$ $\ell$)}
         \end{tabular}
@@
           \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
-          \hspace{-0.3cm}\code{[9\, 4\, 4]}$\to $\code{True\phantom{e}} \\
-          \hspace{0.3cm}\code{[4]}$\to $\code{True\phantom{e}}                        \\
-          \hspace{-0.3cm}\code{[4\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
-          \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
+          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
+          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
+          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
+          \code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}
         \end{tabular}
@@
       $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
-      \vspace{2ex}
       \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
       $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
-      \vspace{2ex}
-      \hspace{\helpSize}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
-      $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
-      \phantom{$f_1(\ell,$\code{k}}
+      \hspace{\helpSize}($f_1$: \emph{Append the number }\code{k}\emph{ to $\ell$})\\
+      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
+      $f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\
+      \phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}\\
+      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
+      $f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
+      \phantom{$f_1(\ell,$}
       \code{($\lambda$ (x a) (if a a (= k x))))}\\

### 2018-05-18 nips_2018.tex
writing

@@
 
-\system learns a DSL containing 13 new functions,
-most of which are templates for polynomials of different orders or ratios of polynomials.
-The algorithm also learns to find programs that minimize the number of continuous degrees of freedom.
-For example, it learns to represent linear functions with the program
-\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
-which has five continuous parameters.
-This phenomenon arises from our Bayesian framing -- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
-\begin{wrapfigure}{r}{0.45\textwidth} \newcommand{\functionSize}{1.5cm}\centering
+\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
   \includegraphics[width = \functionSize]{figures/functions/6.png}
@@
   \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top row) and polynomials (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}
-    \end{wrapfigure}
+DSL learns subroutines for polynomials (top row) and polynomials (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
+    \end{wrapfigure}\system learns a DSL containing 13 new functions,
+most of which are templates for polynomials of different orders or ratios of polynomials.
+It also learns to find programs that minimize the number of continuous degrees of freedom.
+For example, it learns to represent linear functions with the program
+\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
+which has five continuous parameters.
+This phenomenon arises from our Bayesian framing -- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
 
@@
 We compare with ablations of our model on held out tasks.
-Each of these ablations is chosen to approximate
-a prior approach in the program learning literature:
+The purpose of this ablation study is to
+both to examine the role of each component of \systemEnding,
+as well as to compare with
+prior approaches in the literature:
+A head-to-head
+comparison of program synthesizers is complicated by the fact that
+each system, including ours, makes idiosyncratic 
+assumptions about the space of programs and the statement of tasks.
+Nevertheless, much prior work can be modeled as
+special cases of our setup. We compare with the following baselines:
 \\\noindent \textbf{No NN}, which lesions the recognition model.
@@
 RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-%% We can not compare directly with these systems,
-%% because they are engineered for one specific domain, and 
-%% do not have publicly available code and datasets.
 \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
@@
 agent can solve and how quickly it can find those solutions.
-%% For symbolic regression, we also care about the quality
-%% of the solution, as measured by the likelihood model $\probability[x|p]$,
-%% e.g. did the agent correctly explain a linear function
-%% using two numbers, or did it introduce extraneous parameters?
 Tbl.~\ref{baselineComparisons}
@@
   \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
-  \% solved&\textbf{75\%} &43\% &&33\%&0\%&4\%\\
-  Solve time&  29s&65s&&80s&--&235s
+  \% solved&\textbf{74\%} &43\% &&33\%&0\%&4\%\\
+  Solve time&  38s&65s&&80s&--&235s
 

### 2018-05-18 supplement.tex
More tweaking the sections 2 to 4
add list dataset to supplement

@@
 
-\begin{wrapfigure}{L}{0.5\textwidth}
-  \begin{minipage}{0.49\textwidth}    
+\begin{figure}
+  \centering
+  \begin{minipage}{0.5\textwidth}    
     \begin{algorithm}[H]
@@
   \end{minipage}
-  \end{wrapfigure}
+\end{figure}
 
@@
 
+\section{List Processing Data Set}
+Each list processing tasks we created is in described in
+Tbl~\ref{listdataset}.
+
+\begin{table}[h]
+  \centering\footnotesize
+  \begin{tabular}{|l|l|}\hline
+    add-k for k $\in\{0..5\}$&
+kth-largest for k $\in\{1..5\}$\\
+    append-index-k for k $\in\{1..5\}$&
+kth-smallest for k $\in\{1..5\}$\\
+    append-k for k $\in\{0..5\}$&
+last\\
+    bool-identify-geq-k for k $\in\{0..5\}$&
+len\\
+    bool-identify-is-mod-k for k $\in\{1..5\}$&
+max\\
+    bool-identify-is-prime&
+min\\
+    bool-identify-k for k $\in\{0..5\}$&
+modulo-k for k $\in\{1..5\}$\\
+    caesar-cipher-k-modulo-n&
+mult-k for k $\in\{0..5\}$\\
+    \qquad for k $\in\{0..5\}$ and n $\in\{1..5\}$&
+odds\\
+    count-head-in-tail&
+pop\\
+    count-k for k $\in\{0..5\}$&
+pow-k for k $\in\{1..5\}$\\
+    drop-k for k $\in\{0..5\}$&
+prepend-index-k for k $\in\{1..5\}$\\
+    dup&
+prepend-k for k $\in\{0..5\}$\\
+    empty&
+product\\
+    evens&
+range\\
+    fibonacci&
+remove-empty-lists\\
+    has-head-in-tail&
+remove-eq-k for k $\in\{0..3\}$\\
+    has-k for k $\in\{0..5\}$&
+remove-gt-k for k $\in\{0..3\}$\\
+    head&
+remove-index-k for k $\in\{1..5\}$\\
+    index-head&
+remove-mod-head\\
+    index-k for k $\in\{1..5\}$&
+remove-mod-k for k $\in\{2..5\}$\\
+    is-evens&
+repeat\\
+    is-mod-k for k $\in\{1..5\}$&
+repeat-k for k $\in\{1..5\}$\\
+    is-odds&
+repeat-many\\
+    is-primes&
+replace-all-with-index-k for k $\in\{1..5\}$\\
+    is-squares&
+reverse\\
+    keep-eq-k for k $\in\{0..3\}$&
+rotate-k for k $\in\{1..5\}$\\
+    keep-gt-k for k $\in\{0..3\}$&
+slice-k-n for k $\in\{1..5\}$ and n $\in\{1..5\}$\\
+    keep-mod-head&
+sort\\
+    keep-mod-k for k $\in\{1..5\}$&
+sum\\
+    keep-primes&
+tail\\
+    keep-squares&
+take-k for k $\in\{1..5\}$\\
+\hline
+  \end{tabular}
+  \normalsize
+  ~\\~\\\caption{Our list processing data set}\label{listdataset}
+\end{table}
+
 \section{Learned DSLs}
@@
 \onecolumn
-\subsection{Symbolic regression}
-\begin{lstlisting}
-#(\ (#(\ (+. $0 REAL)) (*. $0 REAL)))
-#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0)))
-#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0)))
-#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0)))
-#(/. REAL)
-#(\ (/. (/. REAL $0) $0))
-#(\ (+. $0 REAL))
-#(\ (#(\ (+. $0 REAL)) (*. $0 (#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0))) $0))))
-#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0))
-#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1))))
-#(\ (#(\ (+. $0 REAL)) (#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0))) $0)))
-#(#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1)))) (#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0)) REAL))
-#(\ (/. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) (#(\ (+. $0 REAL)) $0)))
-\end{lstlisting}
-
-\subsection{String editing}
-\begin{lstlisting}
-#(\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
-#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0))))))
-#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0)))))))
-#(\ (\ (#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0)))))) (cons $0 $1))))
-#(\ (\ (fold $0 $0 (\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
-#(\ (#(\ (\ (\ (cons (car $0) (cons $1 $2))))) (#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty) '.' $0) '.'))
-#(\ (\ (map (\ (if (char-eq? $1 $0) $2 $0)))))
-#(\ (map (\ (index $0 $1))))
-#(\ (unfold $0 (\ (empty? $0)) (\ (car $0)) (\ (#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0))))))) SPACE $0))))
-#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty)
-\end{lstlisting}
 
@@
 
+\subsection{Text editing}
+\begin{lstlisting}
+#(\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
+#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0))))))
+#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0)))))))
+#(\ (\ (#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0)))))) (cons $0 $1))))
+#(\ (\ (fold $0 $0 (\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
+#(\ (#(\ (\ (\ (cons (car $0) (cons $1 $2))))) (#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty) '.' $0) '.'))
+#(\ (\ (map (\ (if (char-eq? $1 $0) $2 $0)))))
+#(\ (map (\ (index $0 $1))))
+#(\ (unfold $0 (\ (empty? $0)) (\ (car $0)) (\ (#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0))))))) SPACE $0))))
+#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty)
+\end{lstlisting}
+
+\subsection{Symbolic regression}
+\begin{lstlisting}
+#(\ (#(\ (+. $0 REAL)) (*. $0 REAL)))
+#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0)))
+#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0)))
+#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0)))
+#(/. REAL)
+#(\ (/. (/. REAL $0) $0))
+#(\ (+. $0 REAL))
+#(\ (#(\ (+. $0 REAL)) (*. $0 (#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0))) $0))))
+#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0))
+#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1))))
+#(\ (#(\ (+. $0 REAL)) (#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0))) $0)))
+#(#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1)))) (#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0)) REAL))
+#(\ (/. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) (#(\ (+. $0 REAL)) $0)))
+\end{lstlisting}
+
 \bibliography{main}

### 2018-05-18 supplement.tex
beef up end of supplement

@@
   \end{comment}
-\vfill
 
@@
 
-Here we present representative DSLs learned by
-our model. DSL primitives discovered by the algorithm are prefixed with
-\lstinline!#!.
-Variables are prefixed with \lstinline!$!, and we use De Bruijn indices to model variables~\cite{pierce}.
-We abbreviate lambda as \lstinline!\!.
+Here we present representative DSLs learned by our model. DSL primitives
+discovered by the algorithm are prefixed with \lstinline!#!.
+Variables are prefixed with \lstinline!$!, and we adopt De Bruijn indices to
+model bound variables~\cite{pierce}.
 
@@
 
+\lstset{escapechar=@}
 \subsection{List functions}
 \begin{lstlisting}
-#(\ (\ (index $0 (sort $1))))
-#(\ (\ (slice $0 (+ $1 4))))
-#(mapi (\ (\ (eq? $0 $2))))
-#(\ (mapi (\ (\ (gt? $0 $2)))))
-#(\ (any (\ (eq? $0 $1))))
-#(\ (mapi (\ (\ (+ $0 $2)))))
-#(\ (\ (mapi (\ (\ (index $2 $3))) $1)))
-#(\ (mapi (\ (\ (mod $0 $2)))))
-#(\ (mapi (\ (\ (* $0 $2)))))
-#(\ (\ (singleton (index $0 $1))))
-#(\ (\ (slice $0 (sum $1) $1)))
-#(slice 0)
-#(\ (\ (index (negate $0) (sort $1))))
-#(\ (++ (singleton $0)))
-#(\ (\ (++ $0 (singleton $1))))
-#(\ (\ (++ $0 (singleton (index $1 $0)))))
-#(\ (\ (++ (singleton (index $0 $1)) $1)))
-#(\ (\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) $1 $2)))))
-#(\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))
-#(#(\ (mapi (\ (\ (gt? $0 $2))))) 3 empty)
-#(all (\ (is-square $0)))
-#(\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) 1 $1))))
-#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1))))
-#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2)))
-#(filter (\ (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0))))
-#(\ (\ (\ (++ (#(slice 0) $0 $1) (#(\ (\ (slice $0 (sum $1) $1))) $1 $2)))))
-#(\ (filter (\ (#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1)))) $1 (singleton $0)))))
-#(filter (\ (not (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0)))))
-#(\ (++ $0 (++ $0 $0)))
-#(\ (++ $0 (#(\ (++ $0 (++ $0 $0))) $0)))
-#(index 0)
+#(+ 1 1)
+#(@$\lambda$@ (cdr (cdr $0)))
+#(@$\lambda$@ (fold $0 1 (@$\lambda$@ (@$\lambda$@ (* $0 $1)))))
+#(@$\lambda$@ (cons (car $0) empty))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 (empty? $0) (@$\lambda$@ (@$\lambda$@ (if $0 $0 (eq? $3 $1)))))))
+#(@$\lambda$@ (map (@$\lambda$@ (eq? $1 $0))))
+#(@$\lambda$@ (* $0 (* $0 $0)))
+#(+ 1 #(+ 1 1))
+#(@$\lambda$@ (map (@$\lambda$@ (gt? $0 $1))))
+#(@$\lambda$@ (fold $0 empty (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0)))))
+#(@$\lambda$@ (map (@$\lambda$@ (eq? $0 (length (range $0)))) $0))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $1)) (range $1))))
+#(@$\lambda$@ (cdr (#(@$\lambda$@ (cdr (cdr $0))) $0)))
+#(@$\lambda$@ (map (@$\lambda$@ (index 1 $1))))
+#(@$\lambda$@ (fold $0 empty (@$\lambda$@ (@$\lambda$@ (cons $1 (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (cons (car $0) $1)))
+#(+ 1 #(+ 1 #(+ 1 1)))
+#(@$\lambda$@ (map (@$\lambda$@ (gt? 1 (mod $0 $1)))))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (mod (+ $0 $1) $2)))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 $0) $0))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons $0 empty) $1)))
+#(+ #(+ 1 #(+ 1 1)) #(+ 1 1))
+#(@$\lambda$@ (map (@$\lambda$@ (+ #(+ 1 #(+ 1 1)) (+ $1 $0)))))
+#(@$\lambda$@ (map (@$\lambda$@ (+ $0 $1))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 (empty? $0) (@$\lambda$@ (@$\lambda$@ (gt? $1 (#(@$\lambda$@ (* $0 (* $0 $0))) $3)))))))
+#(@$\lambda$@ (fold $0 0 (@$\lambda$@ (@$\lambda$@ (+ $0 (#(@$\lambda$@ (fold $0 1 (@$\lambda$@ (@$\lambda$@ (* $0 $1))))) (range $1)))))))
+#(@$\lambda$@ (#(@$\lambda$@ (cdr (#(@$\lambda$@ (cdr (cdr $0))) $0))) (cdr $0)))
+#(@$\lambda$@ (#(@$\lambda$@ (fold $0 empty (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0)))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (cons (car $0) empty)) $0) $1)))
+#(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (#(@$\lambda$@ (fold $0 empty (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0))) (map (@$\lambda$@ (- $1 $0)) $1)))))
+#(@$\lambda$@ (@$\lambda$@ (is-square (#(@$\lambda$@ (fold $0 1 (@$\lambda$@ (@$\lambda$@ (* $0 $1))))) (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (mod (+ $0 $1) $2))))) (length (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 $0) $0)) $0)) $1 $0)))))
+#(@$\lambda$@ (@$\lambda$@ (fold (cdr $0) $1 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (cons (car $0) empty)) $0) $1))) (cdr $0) $0))))))
+#(@$\lambda$@ (empty?
+(#(@$\lambda$@ (#(@$\lambda$@ (fold $0 empty (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0))) (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (mod (+ $0 $1) $2))))) #(+ 1 1) 1 $0))))
+#(@$\lambda$@ (@$\lambda$@ (gt?
+(#(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (#(@$\lambda$@ (fold $0 empty (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0))) (map (@$\lambda$@ (- $1 $0)) $1))))) $0 $1) 1)))
 \end{lstlisting}
@@
 \begin{lstlisting}
-#(\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
-#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0))))))
-#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0)))))))
-#(\ (\ (#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0)))))) (cons $0 $1))))
-#(\ (\ (fold $0 $0 (\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
-#(\ (#(\ (\ (\ (cons (car $0) (cons $1 $2))))) (#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty) '.' $0) '.'))
-#(\ (\ (map (\ (if (char-eq? $1 $0) $2 $0)))))
-#(\ (map (\ (index $0 $1))))
-#(\ (unfold $0 (\ (empty? $0)) (\ (car $0)) (\ (#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0))))))) SPACE $0))))
-#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty)
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq? $1 $3) $2 $0)))))))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons $0 $1))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) empty) '.' $0) '.'))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (if (char-eq? $1 $0) $2 $0)))))
+#(@$\lambda$@ (map (@$\lambda$@ (index $0 $1))))
+#(@$\lambda$@ (unfold $0 (@$\lambda$@ (empty?
+$0)) (@$\lambda$@ (car $0)) (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq?
+$1 $3) $2 $0))))))) SPACE $0))))
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) empty)
 \end{lstlisting}
@@
 \begin{lstlisting}
-#(\ (#(\ (+. $0 REAL)) (*. $0 REAL)))
-#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0)))
-#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0)))
-#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0)))
-#(/. REAL)
-#(\ (/. (/. REAL $0) $0))
-#(\ (+. $0 REAL))
-#(\ (#(\ (+. $0 REAL)) (*. $0 (#(\ (#(\ (+. $0 REAL)) (*. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) $0))) $0))))
-#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0))
-#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1))))
-#(\ (#(\ (+. $0 REAL)) (#(\ (#(/. REAL) (#(\ (+. $0 REAL)) $0))) $0)))
-#(#(\ (\ (#(/. REAL) (/. (#(\ (+. $0 REAL)) $0) $1)))) (#(\ (/. (#(\ (/. (/. REAL $0) $0)) $0) $0)) REAL))
-#(\ (/. (#(\ (#(\ (#(\ (+. $0 REAL)) (*. $0 REAL))) (*. (#(\ (+. $0 REAL)) $0) $0))) $0) (#(\ (+. $0 REAL)) $0)))
+#(@$\lambda$@ (/. (/. REAL $0) $0))
+#(@$\lambda$@ (+. $0 REAL))
+#(@$\lambda$@ (#(@$\lambda$@ (+.
+$0 REAL)) (*.
+$0 (#(@$\lambda$@ (#(@$\lambda$@ (+.
+$0 REAL)) (*.
+(#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+.
+$0 REAL)) (*.
+$0 REAL))) (*.
+(#(@$\lambda$@ (+.
+$0 REAL)) $0) $0))) $0) $0))) $0))))
+#(@$\lambda$@ (/. (#(@$\lambda$@ (/. (/. REAL $0) $0)) $0) $0))
+#(@$\lambda$@ (@$\lambda$@ (#(/. REAL) (/. (#(@$\lambda$@ (+. $0 REAL)) $0) $1))))
+#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (#(@$\lambda$@ (#(/. REAL) (#(@$\lambda$@ (+. $0 REAL)) $0))) $0)))
+#(#(@$\lambda$@ (@$\lambda$@ (#(/. REAL) (/. (#(@$\lambda$@ (+. $0 REAL)) $0) $1)))) (#(@$\lambda$@ (/. (#(@$\lambda$@ (/. (/.  REAL $0) $0)) $0) $0)) REAL))
+#(@$\lambda$@ (/. (#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0))) $0) (#(@$\lambda$@ (+. $0 REAL)) $0)))
 \end{lstlisting}

### 2018-05-18 nips_2018.tex
add list times
add list dataset to supplement
add lists curve
finish list section

@@
 We consider this problem within the context of learning functions that
-manipulate lists with some involvement of numerical arithmetic.
+manipulate lists with some involvement of integer arithmetic.
 
@@
   \end{tabular}
-  \captionof{table}{Some tasks in our list function domain}
+  \captionof{table}{Some tasks in our list function domain. See the
+  supplement for the complete data set.}
   \label{listexamples}
@@
 \code{is-prime}.
+
+We experimented with \system by randomly splitting the data set into two
+equal-sized parts: one for training, and one held out for testing.
+We found that the recognition model provided no benefit for the training
+tasks. However, it yielded faster search times on held out tasks allowing
+more tasks to be solved before timing out.
+Several inventions were produced to yield a more expressive DSL for the
+domain (left of Tbl.~\ref{initialExampleDSL}, right of
+Fig.~\ref{fragmentExample}). See the supplement for a complete list of
+inventions.
 
@@
   \\\midrule\multicolumn{7}{c}{\emph{List Processing}}\\\midrule
-  \% solved&\textbf{86\%} &84\% &&60\%&74\%&70\%\\
-  Solve time&  0.8s&0.7s&&1.0s&1.0s&1.1s
+  \% solved&\textbf{79\%} &76\% &71\%&35\%&62\%&37\%\\
+  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s
+  %NOTE: "Ours" result was without helmholtz
 
@@
 \begin{figure}\centering
-\includegraphics[width = 3.5cm]{figures/rationalCurve.eps} 
   \includegraphics[width = 3.5cm]{figures/list.eps}
   \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        
+  \includegraphics[width = 3.5cm]{figures/rationalCurve.eps} 
   \caption{Learning curves for \system both with (blue) and without (red) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 

### 2018-05-18 supplement.tex
wrap lines for supplement inventions
consistent use of foldr/is-nil/nil

--- 2018-05-18 supplement.tex
@@
 
-Here we present representative DSLs learned by
-our model. DSL primitives discovered by the algorithm are prefixed with
-\lstinline!#!.
-Variables are prefixed with \lstinline!$!, and we use De Bruijn indices to model variables~\cite{pierce}.
-We abbreviate lambda as \lstinline!\!.
+Here we present representative DSLs learned by our model. DSL primitives
+discovered by the algorithm are prefixed with \lstinline!#!.
+Variables are prefixed with \lstinline!$!, and we adopt De Bruijn indices to
+model bound variables~\cite{pierce}.
 
-\pagebreak
 \onecolumn
 
-\subsection{List functions}
+\lstset{
+  basicstyle=\footnotesize,
+  escapechar=@,
+  breaklines=true,
+  breakatwhitespace=true,
+  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
+}
+
+\subsection{List processing}
 \begin{lstlisting}
-#(\ (\ (index $0 (sort $1))))
-#(\ (\ (slice $0 (+ $1 4))))
-#(mapi (\ (\ (eq? $0 $2))))
-#(\ (mapi (\ (\ (gt? $0 $2)))))
-#(\ (any (\ (eq? $0 $1))))
-#(\ (mapi (\ (\ (+ $0 $2)))))
-#(\ (\ (mapi (\ (\ (index $2 $3))) $1)))
-#(\ (mapi (\ (\ (mod $0 $2)))))
-#(\ (mapi (\ (\ (* $0 $2)))))
-#(\ (\ (singleton (index $0 $1))))
-#(\ (\ (slice $0 (sum $1) $1)))
-#(slice 0)
-#(\ (\ (index (negate $0) (sort $1))))
-#(\ (++ (singleton $0)))
-#(\ (\ (++ $0 (singleton $1))))
-#(\ (\ (++ $0 (singleton (index $1 $0)))))
-#(\ (\ (++ (singleton (index $0 $1)) $1)))
-#(\ (\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) $1 $2)))))
-#(\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))
-#(#(\ (mapi (\ (\ (gt? $0 $2))))) 3 empty)
-#(all (\ (is-square $0)))
-#(\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) 1 $1))))
-#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1))))
-#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2)))
-#(filter (\ (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0))))
-#(\ (\ (\ (++ (#(slice 0) $0 $1) (#(\ (\ (slice $0 (sum $1) $1))) $1 $2)))))
-#(\ (filter (\ (#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1)))) $1 (singleton $0)))))
-#(filter (\ (not (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0)))))
-#(\ (++ $0 (++ $0 $0)))
-#(\ (++ $0 (#(\ (++ $0 (++ $0 $0))) $0)))
-#(index 0)
+#(+ 1 1)
+#(@$\lambda$@ (cdr (cdr $0)))
+#(@$\lambda$@ (foldr $0 1 (@$\lambda$@ (@$\lambda$@ (* $0 $1)))))
+#(@$\lambda$@ (cons (car $0) nil))
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 (is-nil $0) (@$\lambda$@ (@$\lambda$@ (if $0 $0 (eq? $3 $1)))))))
+#(@$\lambda$@ (map (@$\lambda$@ (eq? $1 $0))))
+#(@$\lambda$@ (* $0 (* $0 $0)))
+#(+ 1 #(+ 1 1))
+#(@$\lambda$@ (map (@$\lambda$@ (gt? $0 $1))))
+#(@$\lambda$@ (foldr $0 nil (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0)))))
+#(@$\lambda$@ (map (@$\lambda$@ (eq? $0 (length (range $0)))) $0))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $1)) (range $1))))
+#(@$\lambda$@ (cdr (#(@$\lambda$@ (cdr (cdr $0))) $0)))
+#(@$\lambda$@ (map (@$\lambda$@ (index 1 $1))))
+#(@$\lambda$@ (foldr $0 nil (@$\lambda$@ (@$\lambda$@ (cons $1 (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (cons (car $0) $1)))
+#(+ 1 #(+ 1 #(+ 1 1)))
+#(@$\lambda$@ (map (@$\lambda$@ (gt? 1 (mod $0 $1)))))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (mod (+ $0 $1) $2)))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 $0) $0))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons $0 nil) $1)))
+#(+ #(+ 1 #(+ 1 1)) #(+ 1 1))
+#(@$\lambda$@ (map (@$\lambda$@ (+ #(+ 1 #(+ 1 1)) (+ $1 $0)))))
+#(@$\lambda$@ (map (@$\lambda$@ (+ $0 $1))))
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 (is-nil $0) (@$\lambda$@ (@$\lambda$@ (gt? $1 (#(@$\lambda$@ (* $0 (* $0 $0))) $3)))))))
+#(@$\lambda$@ (foldr $0 0 (@$\lambda$@ (@$\lambda$@ (+ $0 (#(@$\lambda$@ (foldr $0 1 (@$\lambda$@ (@$\lambda$@ (* $0 $1))))) (range $1)))))))
+#(@$\lambda$@ (#(@$\lambda$@ (cdr (#(@$\lambda$@ (cdr (cdr $0))) $0))) (cdr $0)))
+#(@$\lambda$@ (#(@$\lambda$@ (foldr $0 nil (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0)))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (cons (car $0) nil)) $0) $1)))
+#(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (#(@$\lambda$@ (foldr $0 nil (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0))) (map (@$\lambda$@ (- $1 $0)) $1)))))
+#(@$\lambda$@ (@$\lambda$@ (is-square (#(@$\lambda$@ (foldr $0 1 (@$\lambda$@ (@$\lambda$@ (* $0 $1))))) (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (mod (+ $0 $1) $2))))) (length (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 $0) $0)) $0)) $1 $0)))))
+#(@$\lambda$@ (@$\lambda$@ (foldr (cdr $0) $1 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (#(@$\lambda$@ (cons (car $0) nil)) $0) $1))) (cdr $0) $0))))))
+#(@$\lambda$@ (is-nil (#(@$\lambda$@ (#(@$\lambda$@ (foldr $0 nil (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0))) (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (mod (+ $0 $1) $2))))) #(+ 1 1) 1 $0))))
+#(@$\lambda$@ (@$\lambda$@ (gt?
+(#(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (#(@$\lambda$@ (foldr $0 nil (@$\lambda$@ (@$\lambda$@ (if (is-square $1) (cons $1 $0) $0))))) (map (@$\lambda$@ (* $0 (+ $0 $0))) $0))) (map (@$\lambda$@ (- $1 $0)) $1))))) $0 $1) 1)))
 \end{lstlisting}
@@
 \begin{lstlisting}
-#(\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $3) empty (cons $1 $0)))))))
-#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0))))))
-#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0)))))))
-#(\ (\ (#(\ (\ (fold $0 $1 (\ (\ (cons $1 $0)))))) (cons $0 $1))))
-#(\ (\ (fold $0 $0 (\ (\ (fold $0 $0 (\ (\ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
-#(\ (#(\ (\ (\ (cons (car $0) (cons $1 $2))))) (#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty) '.' $0) '.'))
-#(\ (\ (map (\ (if (char-eq? $1 $0) $2 $0)))))
-#(\ (map (\ (index $0 $1))))
-#(\ (unfold $0 (\ (empty? $0)) (\ (car $0)) (\ (#(\ (\ (fold $0 $0 (\ (\ (cdr (if (char-eq? $1 $3) $2 $0))))))) SPACE $0))))
-#(#(\ (\ (\ (cons (car $0) (cons $1 $2))))) empty)
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq? $1 $3) $2 $0)))))))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons $0 $1))))
+#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) nil) '.' $0) '.'))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (if (char-eq? $1 $0) $2 $0)))))
+#(@$\lambda$@ (map (@$\lambda$@ (index $0 $1))))
+#(@$\lambda$@ (unfold $0 (@$\lambda$@ (is-nil $0)) (@$\lambda$@ (car $0)) (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq?
+$1 $3) $2 $0))))))) SPACE $0))))
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) nil)
 \end{lstlisting}

### 2018-05-18 nips_2018.tex
updated supplement and minor edits

@@
 
-\section{Sequence Manipulation: Domains from generic primitives}\label{sequences}
+\section{Programs that manipulate sequences}\label{sequences}
 We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
@@
 input/output examples (Tbl.~\ref{listexamples}).
-Our data set is challenging in three major ways: many of the tasks require
-complex solutions, the tasks were not generated from some latent DSL, and
+Our data set is interesting in three major ways: many of the tasks require
+complex solutions; the tasks were not generated from some latent DSL, and
 the agent must learn to solve these complicated problems from only 236
@@
 
-We experimented with \system by randomly splitting the data set into two
-equal-sized parts: one for training, and one held out for testing.
-We found that the recognition model provided no benefit for the training
-tasks. However, it yielded faster search times on held out tasks allowing
+We evaluated \system on  random 50/50 test/train split.
+Interestingly, we found that the recognition model provided little benefit for the training
+tasks. However, it yielded faster search times on held out tasks, allowing
 more tasks to be solved before timing out.
-Several inventions were produced to yield a more expressive DSL for the
+The system composed 38 new subroutines, yielding a more expressive DSL more closely matching the
 domain (left of Tbl.~\ref{initialExampleDSL}, right of
 Fig.~\ref{fragmentExample}). See the supplement for a complete list of
-inventions.
+DSL primitives discovered by \systemEnding.
 
@@
 At first, \system cannot find any correct programs for most of the tasks.
-After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Tbl.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
-all of the tasks. The learned DSL contains 10 new functions, listed in the supplement.
+After three iterations, it assembles a DSL containing 12 new functions (center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
+all of the tasks.
 
@@
 We compare with ablations of our model on held out tasks.
-The purpose of this ablation study is to
+The purpose of this ablation study is 
 both to examine the role of each component of \systemEnding,
@@
   \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
-  \% solved&\textbf{74\%} &43\% &&33\%&0\%&4\%\\
-  Solve time&  38s&65s&&80s&--&235s
+  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
+  Solve time&  38s&65s&38s&80s&--&235s
 

### 2018-05-18 supplement.tex
updated supplement and minor edits

@@
 \begin{lstlisting}
-#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
-#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
-#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq? $1 $3) $2 $0)))))))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons $0 $1))))
-#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
+#(+ 1)
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq? $1 $3) $2 $0)))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons $0 $1))))
 #(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (cons (car $0) (cons $1 $2))))) nil) '.' $0) '.'))
+#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (if (char-eq? $1 $5) (cdr $2) $0)))))))))
 #(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (if (char-eq? $1 $0) $2 $0)))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $1 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 STRING))
 #(@$\lambda$@ (map (@$\lambda$@ (index $0 $1))))
-#(@$\lambda$@ (unfold $0 (@$\lambda$@ (is-nil $0)) (@$\lambda$@ (car $0)) (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (foldr $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq?
+#(@$\lambda$@ (unfold $0 (@$\lambda$@ (nil?
+$0)) (@$\lambda$@ (car $0)) (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $0 $0 (@$\lambda$@ (@$\lambda$@ (cdr (if (char-eq?
 $1 $3) $2 $0))))))) SPACE $0))))

### 2018-05-18 nips_2018.tex
text editing figure

@@
         \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
+        $f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}\\
+      \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
   \end{tabular}&
@@
 
-Because our enumerative search procedure has no means of generating
-string constants, we have the enumerator propose programs with string-valued holes --
-e.g., in Fig.~\ref{firstPageFigure},
-we enumerate $\text{\code{(}}f_0\code{ string (}f_2\text{\code{ s ' '))}}$ --
-and define $\probability[x|p]$ by marginalizing out the values of the
-strings via dynamic programming.  In Section~\ref{regressionSection},
-we will use a similar trick to synthesize programs containing real
-numbers, but using gradient descent instead of dynamic programming.
-
-We automatically generated 109 text editing tasks with 4 input/output examples each
-%(Fig.~\ref{exampleTextProblem}).
-At first, \system cannot find any correct programs for most of the tasks.
-After three iterations, it assembles a DSL containing 12 new functions (center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
-all of the tasks.
+Because our enumerative search procedure cannot generate string
+constants, we instead enumerate programs with string-valued
+parameters.  For example, to learn a program that prepends ``Dr.'', we
+enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
+learned appending primitive (Fig.~\ref{initialExampleDSL}) -- and then
+define $\probability[x|p]$ by approximately marginalizing out the
+string parameters via a simple dynamic program.
+In Sec.~\ref{regressionSection}, we will use a similar trick to
+synthesize programs containing real numbers, but using gradient
+descent instead of dynamic programming.
+
+We trained our system on a corpus of 109 automatically generated text editing tasks, with 4 input/output examples each.
+After three iterations, it assembles a DSL containing a dozen new functions (center of Fig.~\ref{initialExampleDSL}) that let it solve 
+all of the training tasks.
 
 How well does the  learned DSL generalized to real text-editing scenarios?
-We tested, but did not train, our system on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
 \system solves 3.7\% of the problems with an average search time of 235 seconds.
@@
 it solves 74.1\%, and does so much faster,
-solving them in an average of 38 seconds.
+solving them in an average of 29 seconds.
 As of the 2017 SyGuS competition,
@@
 But, SyGuS comes with a
-different hand-engineered DSL \emph{for each text editing problem}.
+different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants \emph{for each task}.
+Our system does not use this assistance.}
 Here  we learned a single DSL
@@
   \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
-  Solve time&  38s&65s&38s&80s&--&235s
+  Solve time&  29s&65s&38s&80s&--&235s
 

### 2018-05-18 nips_2018.tex
Fiddling and manually merging

@@
 
-%% \begin{table}[t]
-%%   \centering
-%%     \begin{tabular}{l}
-%%       \toprule
-%%       $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
-%%       \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
-%%       $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
-%%         \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-%%       $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
-%%         \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words. See Fig.~\ref{initialExampleDSL}})\\
-%%         $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (if (= c x) nil (cons x a))))}\\
-%%         \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-%%         %($\lambda$ ($\lambda$ (foldr $0 $0 ($\lambda$ ($\lambda$ (if (char-eq? $1 $3) nil (cons $1 $0)))))))
-%%   \\\bottomrule
-%%     \end{tabular}
-%%   \caption{Some text editing learned subroutines }\label{textPrimitives}
-%%   \end{table}
-
-%% \begin{wrapfigure}{r}{0.6\textwidth}
-%%   \centering\begin{tabular}{c}\toprule
-%% +106 769-858-438 $\longrightarrow$106.769.858.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
-%% +83 973-757-831 $\longrightarrow$83.973.757.831%& Andrew Cencici $\longrightarrow$ Dr. Andrew
-%% %DkI$>$DwKI$<$xaPoc& 	DWKI\\
-%% %KfvTh$>$Dbl$<$OIJso 	&DBL
-%% \\\midrule
-%% %(lambda (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' SPACE (#(lambda (lambda (map (lambda (if (char-eq? $0 $1) $2 $0))))) '.' '-' (cdr $0))))
-%% \multicolumn{1}{c}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  '.' ' ' ($f_0$ '.' '-' (cdr s)))}}
-%% %\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
-%% \\
-%% \bottomrule
-%%   \end{tabular}
-
-%%   %% \vspace{0.25cm}
-
-%%   %% \begin{minipage}{\linewidth}
-%%   %%   \centering
-%%   %%   $f($\code{s}$) \,=\, $
-%%   %% \end{minipage}
-
-%%   \caption{String edit task (top) and the program \system writes for it (bottom). $f_0$ is a subroutine written by \systemEnding, defined in Tbl.~\ref{textPrimitives}.}\label{exampleTextProblem}
-%% \end{wrapfigure}
-
 \section{Symbolic Regression: Programs from visual input}\label{regressionSection}
@@
 
-%% \begin{align*}
-%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
-%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
-%% \end{align*}
-
-%% %% and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
-%% %% \begin{align*}
-%% %% \log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
-%% %% %\text{\emph{BIC:}} &\approx 
-%% %% \end{align*}
-%% where $\mathcal{N}(\cdot |\cdot )$ is the normal density and
-%% $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over
-%% $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~:
-
-%% where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$
-%% found by performing gradient ascent on the likelihood of the
-%% observations w.r.t. $\vec{\mathcal{R}}$.
-
 \begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
@@
   \includegraphics[width = \functionSize]{figures/functions/116.png}\\
-  \includegraphics[width = \functionSize]{figures/functions/181.png}
+  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
   \includegraphics[width = \functionSize]{figures/functions/160.png}
@@
 DSL learns subroutines for polynomials (top row) and polynomials (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
-    \end{wrapfigure}\system learns a DSL containing 13 new functions,
+\end{wrapfigure}
+\system learns a DSL containing 13 new functions,
 most of which are templates for polynomials of different orders or ratios of polynomials.
@@
 prior approaches in the literature:
-A head-to-head
+a head-to-head
 comparison of program synthesizers is complicated by the fact that
@@
 assumptions about the space of programs and the statement of tasks.
+\begin{wraptable}{r}{0.55\textwidth}
+\vspace{-5pt}
+%% This radically shrinks the table
+\tabcolsep=2.5pt
+\renewcommand{\arraystretch}{0.5}
+\begin{tabular}{lcccccc}
+  \toprule& Ours& 
+    No NN & SE&NPS & PCFG & Enum \\
+  \midrule
+  \multicolumn{7}{c}{\emph{List Processing}}\\
+  \midrule
+  \% solved&\textbf{79\%} &76\% &71\%&35\%&62\%&37\%\\
+  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
+  \midrule
+  \multicolumn{7}{c}{\emph{Text Editing}}\\
+  \midrule
+  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
+  Solve time& 38s&65s&38s&80s&--&235s\\
+  \midrule
+  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
+  \midrule
+  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
+  Solve time&  \textbf{24}s& 40s  &28s&31s&55s&29s\\
+  \bottomrule
+  \end{tabular}
+\caption{\% solved before timeout. Solve time: averaged over solved
+  tasks.  SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
+  structure learning.  Enum: model w/o any learning.  %% MDL:
+  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
+  %% symbolic regression MDL is 0 nats.
+  %% For symbolic regression MDL is a
+  %% proxy for \# of continuous parameters.
+}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
 Nevertheless, much prior work can be modeled as
-special cases of our setup. We compare with the following baselines:
+special cases of our setup. 
+We compare with the following baselines:
 \\\noindent \textbf{No NN}, which lesions the recognition model.
@@
 
- \section{Contribution and Outlook}
+\section{Discussion}
 
 \parbox{0.58\textwidth}{
-We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
-domain-specific primitives that the algorithm itself discovers,
-together with a neural recognition model that learns how to
-efficiently deploy the DSL on new tasks. We believe this integration
-of top-down symbolic representations and bottom-up neural networks --
-both of them learned -- could help make program induction systems more
-generally useful for AI. Many directions remain open. Two immediate
-goals 
-are to integrate more sophisticated neural recognition
+
+\paragraph{Outlook.}
+We contribute an algorithm, \systemEnding, that learns to program by
+bootstrapping a DSL with new domain-specific primitives that the algorithm
+itself discovers, together with a neural recognition model that learns how to
+efficiently deploy the DSL on new tasks. We believe this integration of top-down
+symbolic representations and bottom-up neural networks --- both of them learned
+--- helps make program induction systems more generally useful for AI. Many
+directions remain open.
+
+\paragraph{Future.}
+Two immediate goals are to integrate more sophisticated neural recognition
 models~\cite{devlin2017robustfill} and program
-synthesizers~\cite{solar2008program}, which may improve performance in
-some domains over the generic methods used here.
-Another direction is to
-explore DSL meta-learning: Can we find a \emph{single} universal
-primitive set that could effectively bootstrap DSLs for new domains,
-including the three domains considered,  but also many others?
+synthesizers~\cite{solar2008program}, which may improve performance in some
+domains over the generic methods used here. Figure \ref{geomCompiled} gives an
+example of a promising domain for this, abstract geometry, see supplementary material for details.  Another direction is to explore DSL
+meta-learning: can we find a \emph{single} universal primitive set that could
+effectively bootstrap DSLs for new domains, including the three domains
+considered,  but also many others?
 }\hfill
@@
   percent in the middle, and on both side average of the compiled new shapes
-  before any traning and at the end of the last iteration. See supplementary
-  material for more details.}
+  before any traning and at the end of the last iteration.}
 \label{geomCompiled}

### 2018-05-18 nips_2018.tex
Josh introduction

@@
 
-Imagine you are asked  to edit some text, and
-told that you should change the text ``Nancy FreeHafer'' to
-``Dr. Nancy''.  From this example, you likely infer that ``Jane
-Goodall'' goes to ``Dr. Jane'', drawing upon your prior knowledge of
-text, like that words are separated by spaces. Few-shot learning problems
-like these
-are commonplace in both human and machine learning.
-We cast these few-shot learning problems as each being a program synthesis problem.
-The programming languages and AI communities have built many successful program synthesis algorithms,
-spanning text editing (e.g., FlashFill~\cite{gulwani2011automating}), motor programs~\cite{lake2015human}, procedural graphics~\cite{ellis2017learning}, planning
-procedures~\cite{devlin2017neural}, and many others.  However, the
-success of these systems hinges upon a carefully
-hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
-impart prior knowledge of a domain
-by providing a restricted set
-of finely-tuned programming primitives: for
-%that are finely tuned to the problems in the domain: for
-text editing,
-these are primitives like appending and splitting on characters.
-In this work,
-we consider the problem of
-building agents that solve program learning tasks,
-and also the problem of acquiring
-the prior knowledge necessary to quickly
-solve these tasks (Figure~\ref{firstPageFigure}).
-Our solution is an algorithm that learns a DSL
-while jointly training a neural network to help write programs in the learned DSL.
-\begin{figure}[h]
-  \begin{minipage}[c]{0.68\textwidth}
-\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
-  \sc{Task}&
-  \begin{tabular}{rcl}
-  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
-% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
- Jane Goodall&$\longrightarrow$&???
-  \end{tabular}\\\cmidrule{2-2}
-  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
-    \sc{DSL}
-&\begin{tabular}{l}
-    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b ($\lambda$ (x y) (cons x y)))}\\
-    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
-    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\
-    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
-        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
-    \end{tabular}\\\cmidrule{2-2}
-  \end{tabular}
-  \end{minipage}%
-  \begin{minipage}[c]{0.31\textwidth}
-      \caption{\textbf{Task}: Few-shot learning problem.
-Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}.
-Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
-    \end{minipage}
-\end{figure}
-
-%% We think of
-%% solutions to these tasks as being well represented by programs,
-%% and so our problem can be stated as follows:
-%% how should an agent learn to write programs?
-We take inspiration from two sources:
-(1) Good software engineers compose libraries
-of reusable subroutines
-that are shared across
-related programming tasks.
-Returning to Figure~\ref{firstPageFigure},
-a good text editing library should support appending strings and splitting on spaces --
-exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
-(2) Skilled human programmers
-can quickly recognize what kinds of programming idioms and library routines would be useful for solving
-the task at hand, even if they cannot instantly work out the details.
-We weave these ideas together into
-an algorithm called \system (\textbf{Co}mpress/\textbf{Co}mpile/\textbf{Sea}rch),
-which takes as input a collection of
-programming \textbf{tasks},
-and then jointly
-solves three problems:
-(1) Searching for programs that solve the tasks;
-(2) Composing a library (DSL) of
-domain-specific
-subroutines -- which allow the agent to
-more compactly write programs in the domain,
-and (3) Training a neural network
-to recognize
-which DSL components are
-useful for which kinds of tasks.
-Together, the DSL and neural net
-encode the domain specific knowledge needed to
-quickly write programs.
-
-%% synthesize programs for task domains such as string transformations,
-%% list processing, and robot navigation and planning. However, all these
-%% approaches -- symbolic, neural and neural-symbolic -- rely upon a
-%% hand-engineered \emph{Domain-Specific Language} (DSL). DSLs contain an
-%% inventory of restricted programming primitives, encoding domain-specific
-%% knowledge about the space of programs. In
-%% practice
-%% we often have only a few input/output examples for each
-%% program to be induced, and thus success often hinges on having a good
-%% DSL that provides a crucial inductive bias for what would otherwise be an
-%% unconstrained search through the space of all computable functions.
-%% Here we ask, to what extent can we dispense with such highly
-%% hand-engineered domain-specific languages?
-
-%% We propose \emph{learning} the DSL by inducing a library of domain--specific
-%% subroutines. We consider the setting where we have a collection of
-%% related programming tasks, each specified by a set of input/output
-%% examples. %% We do not assume that the tasks are annotated with
-%% %% ground-truth programs.
-%% Starting from a weaker or more
-%% general library of primitives, we give an algorithm for constructing a richer, more powerful,
-%% and better-tuned DSL.
-
-%% Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
-%% because it
-\system iterates through three different steps: a
-\textbf{Search} step uses the DSL and neural network to explore the space of programs,
-searching for ones that solve the tasks; a \textbf{Compress} step
-modifies the structure of the DSL by discovering regularities across programs found during search; and a \textbf{Compile} step, which
-improves the search procedure by training a neural network to
-write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}.
-We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
-The learned DSL
-distills commonalities across programs that solve tasks, helping
-the agent solve related programming problems. The neural
-recognition model ensures that searching for programs remains tractable
-even as the DSL (and hence the search space for programs) expands.
-%% Because any model may be encoded as a (deterministic or probabilistic) program,
-%% we carefully delineate the scope of problems considered here.
-We think of \system as learning to solve the kinds of problems that humans
-can solve relatively quickly -- once they acquire the relevant domain expertise.
-These correspond to short programs -- if you have an expressive DSL.
-Even with a good DSL, program search may be intractable, so
-we amortize the cost of search by training a neural network to
-assist the search procedure.
-
-We apply \system to three domains:
-list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
- For each of these we initially provide a generic
- set of programming primitives.
-Our algorithm then constructs
-its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
-\begin{table}[hb]
+Much of everyday human thinking and learning can be understood in
+terms of program induction: constructing a procedure that maps inputs
+to desired outputs, based on observing example input-output pairs.
+People can induce programs flexibly across many different domains, and
+remarkably, often from just one or a few examples.  For instance, if
+shown that a text-editing program should map ``Jane Morris Goodall''
+to ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey''
+to ``R. E. Leakey''; if instead the first input mapped to
+``Dr. Jane'', ``Goodall, Jane'', or ''Morris'', we might have guessed
+the latter should map to ``Dr. Richard'', ``Leakey, Richard'', or
+``Erskine'', respectively.
+
+The FlashFill system ~\cite{gulwani2011automating} developed by
+Microsoft researchers and now embedded in Excel solves problems such
+as these and is probably the best known practical program-induction
+algorithm, but researchers in programming languages and AI have built
+successful program induction algorithms for many applications, 
+such as handwriting recognition and generation ~\cite{lake2015human},
+procedural graphics ~\cite{ellis2017learning}, question
+answering~\cite{johnson2017clevr} and robot motion
+planning~\cite{devlin2017neural}, to name just a few.
+These systems work in different ways, but most hinge upon having a
+carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
+especially true for systems such as FlashFill that aim to induce a
+wide range of programs very quickly, in a few seconds or less.  DSLs
+constrain the search over programs with strong prior knowledge in the
+form of a restricted set of programming primitives tuned to the needs
+of the domain: for text editing, these are operations like appending
+strings and splitting on characters.
+
+In this work, we consider the problem of building agents that learn to
+solve program induction tasks, and also the problem of acquiring the
+prior knowledge necessary to quickly solve these tasks in a new domain.
+Representative problems in three domains are shown in 
+Table~\ref{initialExampleDSL}.  Our solution is an algorithm that
+grows or boostraps a DSL while jointly training a neural network to help write
+programs in the increasingly rich DSL.  
+%
+% (JBT: MAYBE INSERT FIGURE FROM ICML PAPER?)
+%
+
+\begin{table}[t!]
   \makebox[\textwidth][c]{
@@
 
+Because any computable learning problem can in principle be cast as
+program induction, it is important to delimit our focus.  In contrast
+to computer assisted programming~\cite{solar2008program} or genetic
+programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
+engineering, to learn to synthesize large bodies of code, or to learn
+complex programs starting from scratch.  Ours is a basic AI goal:
+capturing the human ability to learn to think flexibly and efficiently
+in new domains -- to learn what you need to know about a domain so you
+don't have to solve new problems starting from scratch.  We are
+focused on the kinds of problems that humans can solve relatively
+quickly, once they acquire the relevant domain expertise.  These
+correspond to tasks solved by short programs -- if you have an
+expressive DSL.  Even with a good DSL, program search may be
+intractable;
+%
+% , and adding new library routines to the DSL only broadens the
+% search space;
+%
+so we amortize the cost of search by training a
+neural network to assist the search procedure.
+
+Our algorithm takes inspiration from several ways that skilled human
+programmers have learned to code: Skilled coders build libraries of
+reusable subroutines that are shared across related programming tasks,
+and can be composed to generate increasingly complex and powerful
+subroutines.  In text editing, a good library should support routines
+for splitting on characters, but also specialize these routines to
+split on particular characters such as spaces or commas that are
+frequently used to delimit substrings across tasks.  Skilled coders
+also learn to recognize what kinds of programming idioms and library
+routines would be useful for solving the task at hand, even if they
+cannot instantly work out the details.  In text editing, one might
+learn that if outputs are consistently shorter than inputs, removing
+characters is likely to be part of the solution; if every output
+contains a constant substring (e.g., ``Dr.''), inserting or appending
+that constant string is likely to be a subroutine.
+
+Our \system (\textbf{(S)}search/\textbf{(C})ompress/\textbf{(C})ompile)
+algorithm incorporates these insights by iterating through three steps. 
+The \textbf{Search} step takes a given set of \textbf{tasks}, typically
+several hundred, and searches for compact programs that solve these tasks
+guided by the current DSL and neural network.
+The \textbf{Compress} step grows the library (or DSL) of
+domain-specific subroutines which allow the agent to more compactly
+write programs in the domain; it modifies the structure of the DSL by
+discovering regularities across programs found during search, compressing
+them to distill out common code fragments across successful programs.
+The \textbf{Compile} step improves the search procedure by training a neural network to
+write programs in the current DSL, in the spirit of ``amortized'' or
+``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
+%% and others too on amortized inference.)
+
+The learned DSL effectively encodes a prior on programs likely to
+solve tasks in the domain, while the neural net looks at the example
+input-output pairs for a specific task and produces a ``posterior''
+for programs likely to solve that specific task.  The neural network
+thus functions as a \textbf{recognition model} supporting a form of
+approximate Bayesian program induction, jointly trained with a
+\textbf{generative model} for programs encoded in the DSL, in the
+spirit of the Helmholtz machine ~\cite{hinton1995wake}). The
+recognition model ensures that searching for programs remains
+tractable even as the DSL (and hence the search space for programs)
+expands.
+
+We apply \system to three domains:
+list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
+ For each of these we initially provide a generic
+ set of programming primitives.
+Our algorithm then constructs
+its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+
 Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
@@
  \section{The \system Algorithm}
-
- Our goal is to induce a DSL while finding programs solving each of the
-tasks.  We take inspiration primarily from the
-Exploration-Compression algorithm for bootstrap
-learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
-alternates between exploring the space of solutions to a set of tasks,
-and compressing those solutions to suggest new search primitives
-for the next exploration stage.
-We extend these ideas into an inference strategy that iterates
-through three steps: a \textbf{Search} step uses the current DSL and
-recognition model to search for programs that solve the tasks.  The
-\textbf{Compress} and \textbf{Compile} steps update the DSL and the
-recognition model, respectively.  Crucially, these steps bootstrap 
-each other:
-\begin{comment}
-  \begin{wrapfigure}{r}{0.5\textwidth}
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs};
-    \node[align = center](DSL) at (3,1) {DSL};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compression)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compilation)}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
-  \caption{\system solves for programs, the DSL, and a neural network (recognition model). Each of these steps iteratively bootstraps off of the others.}  \label{feeding}\vspace{-0.4cm}
-\end{wrapfigure}
-  \end{comment}
-\\\noindent \textbf{Search: Solving tasks.}  Our program search
-is informed by both the DSL and the recognition model. When these
-improve, we can solve more tasks.
-\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
-from the programs found in the search phase, aiming
-to maximally compress (or, raise the prior probability of) these
-programs.  As we solve more tasks, we hone in on DSLs that
-more closely match the domain.
-\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
-update the recognition model by training on two data sources: samples
-from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
-programs found by the search procedure during search. As the DSL
-improves and as search finds more programs, the recognition model gets
-more data to train on, and better data.
-
-\begin{comment}
-Sec.~\ref{mathematicalFraming} frames this 3-step procedure as probabilistic inference.
-Sec.~\ref{explorationSection} explains how we search for programs that solve the tasks;
-Sec.~\ref{recognitionSection} explains how we train a neural network to  search for programs; and
-Sec.~\ref{grammarInductionSection} explains how we induce a DSL from programs.
-  \end{comment}
+ We first mathematically describe our 3-step algorithm as
+ an inference procedure for a hierarchical Bayesian model (Section~\ref{mathematicalFraming}),
+  and then describe each step algorithmically in detail (Section~\ref{explorationSection}-\ref{grammarInductionSection}).
+
  \subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}
@@
 \begin{figure}\centering
-  \includegraphics[width = 3.5cm]{figures/list.eps}
-  \includegraphics[width =  3.5cm]{figures/textLearningCurve.eps}        
-  \includegraphics[width = 3.5cm]{figures/rationalCurve.eps} 
+  \includegraphics[width = 4.5cm]{figures/list.eps}
+  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}        
+  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}   
   \caption{Learning curves for \system both with (blue) and without (red) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 

### 2018-05-18 nips_2018.tex
Changed colours here and there

--- 2018-05-18 nips_2018.tex
@@
 \newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
+
+\usepackage{xcolor}
+\definecolor{pop1}{HTML}{1b9e77}
+\definecolor{pop2}{HTML}{7570b3}
+\newcommand{\pop}[1]{\textcolor{pop1}{#1}}
+\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
 
@@
     \scriptsize
-  \tabcolsep=4pt
+  \tabcolsep=3pt
   \renewcommand\code\texttt
@@
   \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
+  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
     \toprule
-    {\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
-    {\tabcolsep=4pt
+    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
+    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=4pt
       \begin{tabular}{cc}
@@
           \code{[9\, 3]}$\to $\code{[9\, 3\, 9]} \\
-          \code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}             \\
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}}       \\
           ~                                                  \\
@@
           \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
-          \code{$f(\ell) = $}\code{($f_2$ $\ell$)}
+          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
         \end{tabular}
@@
           \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
-          \code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}
+          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
         \end{tabular}
@@
       +83 973-831$\to $83.973.831\\
-      $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
+      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
       \hspace{1.25cm}\code{($f_0$ \codechar{.} \codechar{ }}\\
-      \hspace{1.4cm}\code{(cdr s)))}\\
+      \hspace{1.4cm}\code{(cdr s)))}}\\
       ~\\
@@
       Lara Gregori$\to $LG\\
-      $f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\
+      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
     \end{tabular}
@@
       \includegraphics[width = 3em]{figures/functions/146}\\
-      \code{$f($x$) = $($f_1$ x)}&    \code{$f($x$) = $($f_6$ x)}\\
+      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
       ~\\
@@
       \\
-      \code{$f($x$) = $($f_4$ x)}&    \code{$f($x$) = $($f_3$ x)}\\
+      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\
 
@@
     \midrule
-    \hspace{1cm}\begin{tabular}{l}
+    \rotatebox[origin=c]{90}{\normalsize DSL}&
+    \hspace{0cm}\begin{tabular}{l}
       % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
@@
           \node(l1) at (0,0) {};
-  \node[color=red](p1) at (-1,-1) {\code{+}};
-  \node[color=red](n1) at (0.7,-0.9) {\code{1}};
+  \node[color=pop1](p1) at (-1,-1) {\code{+}};
+  \node[color=pop1](n1) at (0.7,-0.9) {\code{1}};
   \node(x1) at (0,-1) {\code{1}};
-  \draw[color=red] (l1.south) -- (p1.north);
-  \draw[color=red] (l1.south) -- (n1.north);
-  \draw[color=red] (-0.5,-0.5) -- (x1.north);
+  \draw[color=pop1] (l1.south) -- (p1.north);
+  \draw[color=pop1] (l1.south) -- (n1.north);
+  \draw[color=pop1] (-0.5,-0.5) -- (x1.north);
 
@@
       \node(l1) at (0,0) {};
-  \node[color=red](p1) at (-1,-1) {\code{+}};
-  \node[color=red](n1) at (0.7,-0.9) {\code{1}};
+  \node[color=pop1](p1) at (-1,-1) {\code{+}};
+  \node[color=pop1](n1) at (0.7,-0.9) {\code{1}};
   %\node(x1) at (0,-1) {};
-  \draw[color=red] (l1.south) -- (p1.north);
-  \draw[color=red] (l1.south) -- (n1.north);
-  \draw[color=red] (-0.5,-0.5) -- (0,-1);
+  \draw[color=pop1] (l1.south) -- (p1.north);
+  \draw[color=pop1] (l1.south) -- (n1.north);
+  \draw[color=pop1] (-0.5,-0.5) -- (0,-1);
 
@@
 \begin{scope}[shift={(0,-5)}]
-  \node[red](p1) at (-1,-1) {\code{+}};
-  \node[red](n1) at (0.8,-0.7) {\code{1}};
-  \node[red](a) at (0,-1) {\code{ }};
+  \node[pop1](p1) at (-1,-1) {\code{+}};
+  \node[pop1](n1) at (0.8,-0.7) {\code{1}};
+  \node[pop1](a) at (0,-1) {\code{ }};
   %\node(x1) at (0,-1) {};
-  \draw[red] (0,0) -- (p1.north);
-  \draw[red] (0,0) -- (n1.north);
-  \draw[red] (-0.55,-0.4) -- (a.north);
+  \draw[pop1] (0,0) -- (p1.north);
+  \draw[pop1] (0,0) -- (n1.north);
+  \draw[pop1] (-0.55,-0.4) -- (a.north);
   \node [rotate=90] at (-2.3,-0.7) {\small fragment};

### 2018-05-18 nips_2018.tex
Rephrased "future" part of conclusion

@@
 synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the generic methods used here. Figure \ref{geomCompiled} gives an
-example of a promising domain for this, abstract geometry, see supplementary material for details.  Another direction is to explore DSL
-meta-learning: can we find a \emph{single} universal primitive set that could
-effectively bootstrap DSLs for new domains, including the three domains
-considered,  but also many others?
+domains over the generic methods used here.
+We are in the process of applying our algorithm to generative programs and
+prototyped this with a turtle-like domain: Figure \ref{geomCompiled} gives some
+preliminary results for our graphics programsdomain --- see supplementary material for more
+details.  Another direction is to explore DSL meta-learning: can we find a
+\emph{single} universal primitive set that could effectively bootstrap DSLs for
+new domains, including the three domains considered,  but also many others?
 }\hfill

### 2018-05-18 nips_2018.tex
contributions

@@
 which also learns a library of subroutines.
-We were inspired by EC,
-and go beyond it by giving a new algorithm that learns DSLs, as well as a way of combining DSL learning with
-neurally guided program search.
-In the experiments section,
-we compared directly with EC,
-showing empirical improvements.
-
-The new contribution of this work is
-thus an  algorithm for learning DSLs which jointly trains a neural net to
-search for programs in the DSL.
+We were motivated by EC, and go beyond it and other prior work through the following contributions:
+(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language;
+(2) We give an algorithm for learning DSLs,
+built on a formalism known as Fragment Grammars~\cite{tim};
+and (3) We give a hierarchical Bayesian framing of
+the problem that
+allows a joint inference of the DSL and recognition model.
 
@@
 
-\begin{wraptable}{r}{0.6\textwidth}
-%% This radically shrinks the table
-\tabcolsep=4pt
-\renewcommand{\arraystretch}{0.5}
-\begin{tabular}{lllllll}
-  \toprule& Ours& 
-    No NN
-& SE&NPS & PCFG & Enum
-
-  %% \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
-  %% \% solved &
-  %% \textbf{84\%} &76\%& 63\%&62\%&62\%\\
-  %% Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s
-
-  \\\midrule\multicolumn{7}{c}{\emph{List Processing}}\\\midrule
-  \% solved&\textbf{79\%} &76\% &71\%&35\%&62\%&37\%\\
-  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s
-  %NOTE: "Ours" result was without helmholtz
-
-  \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
-  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
-  Solve time&  29s&65s&38s&80s&--&235s
-
-  %% \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
-  %% \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
-  %% Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s
-
-  \\\midrule\multicolumn{7}{c}{\emph{Symbolic Regression}}\\\midrule
-  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
-  Solve time&  \textbf{24}s& 40s  &28s&31s&55s&29s
-  %% MDL (nats)&   9.3 &11.0&2.3&--&4.6
-
-  \\\bottomrule
-  \end{tabular}
-\caption{\% solved before timeout. Solve time: averaged over solved
-  tasks.  SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
-  structure learning.  Enum: model w/o any learning.  %% MDL:
-  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
-  %% symbolic regression MDL is 0 nats.
-  %% For symbolic regression MDL is a
-  %% proxy for \# of continuous parameters.
-}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
-
 \begin{figure}\centering

### 2018-05-18 nips_2018.tex
itemize contributions

@@
 We were motivated by EC, and go beyond it and other prior work through the following contributions:
-(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language;
-(2) We give an algorithm for learning DSLs,
-built on a formalism known as Fragment Grammars~\cite{tim};
-and (3) We give a hierarchical Bayesian framing of
-the problem that
-allows a joint inference of the DSL and recognition model.
+
+\begin{itemize}
+  \setlength\itemsep{0}
+  \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
+  \item We give an algorithm for learning DSLs, built on a formalism known
+    as Fragment Grammars~\cite{tim}; and 
+  \item We give a hierarchical Bayesian framing of the problem that allows a
+    joint inference of the DSL and recognition model.
+\end{itemize}
 

### 2018-05-18 nips_2018.tex
Changed colours for figures
8 pages!

--- 2018-05-18 nips_2018.tex
@@
 \definecolor{pop1}{HTML}{1b9e77}
-\definecolor{pop2}{HTML}{7570b3}
+\definecolor{pop2}{HTML}{d95f02}
 \newcommand{\pop}[1]{\textcolor{pop1}{#1}}
@@
       +83 973-831$\to $83.973.831\\
-      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
-      \hspace{1.25cm}\code{($f_0$ \codechar{.} \codechar{ }}\\
-      \hspace{1.4cm}\code{(cdr s)))}}\\
+      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,}
+      }}\\
+      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
+      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
       ~\\
@@
 which also learns a library of subroutines.
-We were motivated by EC, and go beyond it and other prior work through the following contributions:
-(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language;
+We find EC motivating, and go beyond it and other prior work through the following contributions:%% \begin{itemize}
+%%   \setlength\itemsep{0}
+%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
+%%   \item We give an algorithm for learning DSLs, built on a formalism known
+%%     as Fragment Grammars~\cite{tim}; and 
+%%   \item We give a hierarchical Bayesian framing of the problem that allows a
+%%     joint inference of the DSL and recognition model.
+%% \end{itemize}
+\noindent\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
 (2) We give an algorithm for learning DSLs,
@@
 the problem that
-allows a joint inference of the DSL and recognition model.
+allows joint inference of the DSL and neural recognition model.
 
@@
 We consider this problem within the context of learning functions that
-manipulate lists with some involvement of integer arithmetic.
+manipulate lists, and which also perform arithmetic operations upon lists of numbers.
 
@@
 all of the training tasks.
-
-How well does the  learned DSL generalized to real text-editing scenarios?
+But, how well does the  learned DSL generalized to real text-editing scenarios?
 We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
@@
   \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
-  Solve time& 38s&65s&38s&80s&--&235s\\
+  Solve time& 29s&65s&38s&80s&--&235s\\
   \midrule
@@
   \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
-  Solve time&  \textbf{24}s& 40s  &28s&31s&55s&29s\\
+  Solve time&  24s& 40s  &28s&31s&55s&29s\\
   \bottomrule
@@
 }\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
-Nevertheless, much prior work can be modeled as
-special cases of our setup. 
+Nevertheless, much prior work can be modeled within our setup. 
 We compare with the following baselines:
@@
 \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-only add subexpressions of programs in the frontiers to the DSL.
+only add \textbf{s}ub\textbf{e}xpressions of programs in the frontiers to the DSL.
 This is how prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
+\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
+but instead learns the parameters of the DSL ($\theta$).
+and is like learning the parameters of a PCFG while not learning any of the structure.
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
 
-For each domain,
-we are interested both in how many tasks the
-agent can solve and how quickly it can find those solutions.
-Tbl.~\ref{baselineComparisons}
-compares our model against these alternatives.
-Our full model consistently
-improves on the baselines,
-sometimes dramatically (text editing and symbolic regression).
-The recognition model 
-consistently increases the number of solved held-out tasks,
-and lesioning it also slows down the convergence of the algorithm,
-taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
-This supports a view of the recognition model as a way of amortizing the cost of searching for programs.
+%For each domain,
+%% We are interested both in how many tasks the
+%% agent can solve and how quickly it can find those solutions.
+%% Tbl.~\ref{baselineComparisons}
+%% compares our model against these alternatives.
+%% We consistently
+%% improve on the baselines,
+%% and find that lesioning the recognition model
+%% and lesioning it also slows down the convergence of the algorithm,
+%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
+%% This supports a view of the recognition model as a way of amortizing the cost of search.
 
@@
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}   
-  \caption{Learning curves for \system both with (blue) and without (red) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
+  \caption{Learning curves for \system both with (\popp{orange}) and without
+  (\pop{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
 \end{figure}
@@
  the agent must learn from at most a few hundred tasks,
- which is facilitated by our ``Helmholtz machine'' style recognition model training (Sec.~\ref{recognitionSection}).
+ which is facilitated by our ``Helmholtz machine'' style recognition model.%(Sec.~\ref{recognitionSection}).
 
@@
 
-\paragraph{Outlook.}
+%\paragraph{Outlook.}
 We contribute an algorithm, \systemEnding, that learns to program by
@@
 directions remain open.
-
-\paragraph{Future.}
+%\paragraph{Future.}
 Two immediate goals are to integrate more sophisticated neural recognition
@@
 prototyped this with a turtle-like domain: Figure \ref{geomCompiled} gives some
-preliminary results for our graphics programsdomain --- see supplementary material for more
+preliminary results for turtle graphics  --- see supplementary material for more
 details.  Another direction is to explore DSL meta-learning: can we find a
@@
   \captionof{figure}{
-  Top left: a few example of the shapes used as targets. Top right: a few
-  examples of discovered new shapes. Bottom: learning curve across iterations in
-  percent in the middle, and on both side average of the compiled new shapes
-  before any traning and at the end of the last iteration.}
+  Top left: Example training tasks. Top right: samples from the learned DSL.
+ Bottom: learning curve  (middle), and on both sides the average sample from the DSL
+  before any training (left) and after training (right). See Sec.~1 of supplement.}
 \label{geomCompiled}

### 2018-05-18 nips_2018.tex
Manual merge

--- 2018-05-18 nips_2018.tex
@@
 \newcommand{\sem}[1]{\llbracket #1 \rrbracket}
-\newcommand{\system}{\textsc{CoCoSea}~}
-\newcommand{\systemEnding}{\textsc{CoCoSea}}
+
+\newcommand{\system}{\textsc{SCC}~}
+\newcommand{\systemEnding}{\textsc{SCC}}
+
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 \usepackage{xcolor}
-\definecolor{pop1}{HTML}{1b9e77}
-\definecolor{pop2}{HTML}{d95f02}
+\definecolor{pop1}{HTML}{1f78b4}
+\definecolor{pop2}{HTML}{33a02c}
+\definecolor{pop3}{HTML}{d95f02}
 \newcommand{\pop}[1]{\textcolor{pop1}{#1}}
 \newcommand{\popp}[1]{\textcolor{pop2}{#1}}
+\newcommand{\tree}[1]{\textcolor{pop3}{#1}}
 
@@
 
-\title{Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning}
+\title{Search, Compress, Compile: Library Learning in Neurally-Guided Bayesian Program Learning}
 
@@
     \midrule
-    \rotatebox[origin=c]{90}{\normalsize DSL}&
+    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
     \hspace{0cm}\begin{tabular}{l}
@@
       % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
-      $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
+      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
       \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
-      $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
+      \popp{$f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}}\\
       \hspace{\helpSize}($f_1$: \emph{Append the number }\code{k}\emph{ to $\ell$})\\
       %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
-      $f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\
-      \phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}\\
+      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
+      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
       \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
-      $f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
+      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
       \phantom{$f_1(\ell,$}
-      \code{($\lambda$ (x a) (if a a (= k x))))}\\
+      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
       \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
@@
   \hspace{0.5cm}\begin{tabular}{l}
-    $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
+    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
+    (x)}}\\
+    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
       \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
-      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
+      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
+      a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}}\\
         \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
+      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
+      }}\\
+      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
         \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
-        $f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}\\
+        \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
       \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
@@
   \begin{tabular}{l}
-    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
-    $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
-    $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
-    $f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}\\
-    $f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}\\
+    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
+    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
+    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
+    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
+    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
     \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
-        $f_5($\code{x}$)\,=\,$\code{(/ real x)}\\
-    $f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))} \\
+    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
+    \popp{$f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))}}\\
     \hspace{\helpSize}\emph{($f_6$: rational function)}\\
@@
           \node(l1) at (0,0) {};
-  \node[color=pop1](p1) at (-1,-1) {\code{+}};
-  \node[color=pop1](n1) at (0.7,-0.9) {\code{1}};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
   \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop1] (l1.south) -- (p1.north);
-  \draw[color=pop1] (l1.south) -- (n1.north);
-  \draw[color=pop1] (-0.5,-0.5) -- (x1.north);
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
 
@@
       \node(l1) at (0,0) {};
-  \node[color=pop1](p1) at (-1,-1) {\code{+}};
-  \node[color=pop1](n1) at (0.7,-0.9) {\code{1}};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
   %\node(x1) at (0,-1) {};
-  \draw[color=pop1] (l1.south) -- (p1.north);
-  \draw[color=pop1] (l1.south) -- (n1.north);
-  \draw[color=pop1] (-0.5,-0.5) -- (0,-1);
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
 
@@
 \begin{scope}[shift={(0,-5)}]
-  \node[pop1](p1) at (-1,-1) {\code{+}};
-  \node[pop1](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop1](a) at (0,-1) {\code{ }};
+  \node[pop3](p1) at (-1,-1) {\code{+}};
+  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
+  \node[pop3](a) at (0,-1) {\code{ }};
   %\node(x1) at (0,-1) {};
-  \draw[pop1] (0,0) -- (p1.north);
-  \draw[pop1] (0,0) -- (n1.north);
-  \draw[pop1] (-0.55,-0.4) -- (a.north);
+  \draw[pop3] (0,0) -- (p1.north);
+  \draw[pop3] (0,0) -- (n1.north);
+  \draw[pop3] (-0.55,-0.4) -- (a.north);
   \node [rotate=90] at (-2.3,-0.7) {\small fragment};
@@
 that these problems are best solved with programs containing real
-numbers.  Our solution to this difficulty is to allow the system to
-write programs with real-valued parameters, and then fit those
+numbers.  Our solution to this difficulty is to enumerate
+ programs with real-valued parameters, and then fit those
 parameters by automatically differentiating through the programs the
@@
   \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top row) and polynomials (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
+DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
 \end{wrapfigure}
@@
 For example, it learns to represent linear functions with the program
-\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
+\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
 which has five continuous parameters.
@@
   \midrule
-  \% solved&\textbf{79\%} &76\% &71\%&35\%&62\%&37\%\\
+  \% solved&\textbf{79\%}\footnote{A last-minute bug caused this  to run without ``Helmholtz machine'' style training ($\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{SS}}$).
+We are rerunning this experiment and expect this number to be higher.} &
+  76\% &71\%&35\%&62\%&37\%\\
   Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
@@
 \caption{\% solved before timeout. Solve time: averaged over solved
-  tasks.  SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
-  structure learning.  Enum: model w/o any learning.  %% MDL:
+  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
+  %structure learning.  Enum: model w/o any learning.  %% MDL:
   %% $-\expect\left[\probability[x|p] \right]$. For domains other than
@@
 Nevertheless, much prior work can be modeled within our setup. 
-We compare with the following baselines:
-\\\noindent \textbf{No NN}, which lesions the recognition model.
+We compare with the following ablations (Table~\ref{baselineComparisons}):
+\\\noindent \textbf{No NN:} lesions the recognition model.
 \\\noindent \textbf{NPS}, which does not learn the DSL,
@@
 We call this NPS (Neural Program Synthesis)
-because this setup is closest to how
+because this is closest to how
 RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
@@
 \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-but instead learns the parameters of the DSL ($\theta$).
-and is like learning the parameters of a PCFG while not learning any of the structure.
+but instead learns the parameters of the DSL ($\theta$) -- learning the parameters of a PCFG while not learning any of the structure.
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.

### 2018-05-18 nips_2018.tex
colors

--- 2018-05-18 nips_2018.tex
@@
 \definecolor{pop1}{HTML}{1b9e77}
-\definecolor{pop2}{HTML}{d95f02}
+%\definecolor{pop2}{HTML}{d95f02}
 \newcommand{\pop}[1]{\textcolor{pop1}{#1}}
-\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
+%\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
 

### 2018-05-18 nips_2018.tex
Prepare merge

--- 2018-05-18 nips_2018.tex
@@
 \definecolor{pop3}{HTML}{d95f02}
+\definecolor{orange}{HTML}{d95f02}
+\definecolor{teal}{HTML}{1b9e77}
 \newcommand{\pop}[1]{\textcolor{pop1}{#1}}
@@
 \newcommand{\tree}[1]{\textcolor{pop3}{#1}}
+\newcommand{\orange}[1]{\textcolor{orange}{#1}}
+\newcommand{\teal}[1]{\textcolor{teal}{#1}}
 
@@
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}   
-  \caption{Learning curves for \system both with (\popp{orange}) and without
-  (\pop{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
+  \caption{Learning curves for \system both with (\orange{orange}) and without
+  (\teal{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
 \end{figure}
@@
 prototyped this with a turtle-like domain: Figure \ref{geomCompiled} gives some
-preliminary results for turtle graphics  --- see supplementary material for more
+preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
 details.  Another direction is to explore DSL meta-learning: can we find a
@@
   Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: learning curve  (middle), and on both sides the average sample from the DSL
-  before any training (left) and after training (right). See Sec.~1 of supplement.}
+ Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
+  before any training (left) and after last iteration (right).}
 \label{geomCompiled}

### 2018-05-18 nips_2018.tex

@@
 But, SyGuS comes with a
-different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants \emph{for each task}.
-Our system does not use this assistance.}
+different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
+For these experiments, our system did not use this assistance.}
 Here  we learned a single DSL
@@
   \midrule
-  \% solved&\textbf{79\%}\footnote{A last-minute bug caused this  to run without ``Helmholtz machine'' style training ($\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{SS}}$).
-We are rerunning this experiment and expect this number to be higher.} &
+  \% solved&\textbf{79\%}\footnotemark & 
   76\% &71\%&35\%&62\%&37\%\\
@@
 \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-only add \textbf{s}ub\textbf{e}xpressions of programs in the frontiers to the DSL.
-This is how prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
+only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
+This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
 \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
@@
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
-
+\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments without training the recognition model on ``Helmholtz machine'' style data.
+We are rerunning this experiment and anticipate improved quantitative results for list processing -- notice the recognition model helps much more for text editing and symbolic regression.} &
 %For each domain,

### 2018-05-18 nips_2018.tex
embarrassing footnote

@@
 \definecolor{pop3}{HTML}{d95f02}
-\definecolor{orange}{HTML}{d95f02}
-\definecolor{teal}{HTML}{1b9e77}
 \newcommand{\pop}[1]{\textcolor{pop1}{#1}}
@@
 \newcommand{\tree}[1]{\textcolor{pop3}{#1}}
-\newcommand{\orange}[1]{\textcolor{orange}{#1}}
-\newcommand{\teal}[1]{\textcolor{teal}{#1}}
 
@@
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}   
-  \caption{Learning curves for \system both with (\orange{orange}) and without
-  (\teal{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
+  \caption{Learning curves for \system both with (\popp{orange}) and without
+  (\pop{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
 \end{figure}
@@
 prototyped this with a turtle-like domain: Figure \ref{geomCompiled} gives some
-preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
+preliminary results for turtle graphics  --- see supplementary material for more
 details.  Another direction is to explore DSL meta-learning: can we find a
@@
   Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
-  before any training (left) and after last iteration (right).}
+ Bottom: learning curve  (middle), and on both sides the average sample from the DSL
+  before any training (left) and after training (right). See Sec.~1 of supplement.}
 \label{geomCompiled}

### 2018-05-18 supplement.tex
Fiddling and manually merging
Added examples of new primitives

@@
 
+\section{Program Representation}\label{programrepresentation}
+We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
+A $\lambda$-calculus expression is either:
+\begin{itemize}
+  \item[--] A \emph{primitive}, like the number 5 or the function \texttt{sum}.
+  \item[--] A \emph{variable}, like $x$, $y$, or $z$.
+  \item[--] A $\lambda$\emph{-abstraction}, which creates a new function.
+$\lambda$-abstractions have a variable and a body.
+The body is a $\lambda$-calculus expression.
+Abstractions are written as $\lambda \text{var}.
+\text{body}$ or in Lisp syntax as \mbox{\texttt{(lambda (\textrm{var}) \textrm{body})}}.
+  \item[--] An \emph{application} of a function to an argument.
+Both the function and the argument are $\lambda$-calculus expressions.
+The application of the function $f$ to the argument $x$ is written as $f\; x$ or as \texttt{($f$ x)}.
+\end{itemize}
+
+For example, the function which squares the logarithm of a number is
+$\lambda x.\text{\tt(square (log $x$))}$, and the identity function $f(x) = x$ is $\lambda x.x$. The
+$\lambda$-calculus serves as a spartan but expressive Turing complete
+program representation, and distills the essential features of functional
+programming languages like Lisp.
+
+However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
+(i.e., $5 \; (\lambda x.x)$).
+We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
+The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
+Hindley-Milner has two important features:
+Feature 1: It supports \emph{parametric polymorphism}, meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
+For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
+A function that returns the first element of a list has the type $[\alpha]\to\alpha$.
+Type variables are not the same as variables introduced by $\lambda$-abstractions.
+Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
+Our generative model over programs performs Hindley-Milner type inference during sampling:
+\emph{Unify} in the generative model uses the machinery of Hindley-Milner to
+ensure that the generated programs have valid polymorphic types.
+A satisfactory exposition of Hindley-Milner is beyond the scope of this paper,
+but~\cite{pierce} offers a nice overview of lambda calculus and typing systems like Hindley-Milner.
+
+%% \begin{figure}
+%%   \begin{align}
+%%     \text{primitive types} & 
+%%     \end{align}
+%%   \end{figure}
+
 \section{Generative model over the programs}
@@
 
-\begin{figure}
+\begin{figure}[H]
   \centering
@@
 
-\section{Program Representation}\label{programrepresentation}
-We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
-A $\lambda$-calculus expression is either:
-\begin{itemize}
-  \item[--] A \emph{primitive}, like the number 5 or the function \texttt{sum}.
-  \item[--] A \emph{variable}, like $x$, $y$, or $z$.
-  \item[--] A $\lambda$\emph{-abstraction}, which creates a new function.
-$\lambda$-abstractions have a variable and a body.
-The body is a $\lambda$-calculus expression.
-Abstractions are written as $\lambda \text{var}.
-\text{body}$ or in Lisp syntax as \mbox{\texttt{(lambda (\textrm{var}) \textrm{body})}}.
-  \item[--] An \emph{application} of a function to an argument.
-Both the function and the argument are $\lambda$-calculus expressions.
-The application of the function $f$ to the argument $x$ is written as $f\; x$ or as \texttt{($f$ x)}.
-\end{itemize}
-
-For example, the function which squares the logarithm of a number is
-$\lambda x.\text{\tt(square (log $x$))}$, and the identity function $f(x) = x$ is $\lambda x.x$. The
-$\lambda$-calculus serves as a spartan but expressive Turing complete
-program representation, and distills the essential features of functional
-programming languages like Lisp.
-
-However, many $\lambda$-calculus expressions correspond to ill-typed programs, such as the program that takes the logarithm of the Boolean \texttt{true} (i.e., \texttt{log true}) or which applies the number five to the identity function
-(i.e., $5 \; (\lambda x.x)$).
-We use a well-established typing system for $\lambda$-calculus called \emph{Hindley-Milner typing}~\cite{pierce}, which is used in programming languages like OCaml.
-The purpose of the typing system is to ensure that our programs never call a function with a type it is not expecting (like trying to take the logarithm of \texttt{true}).
-Hindley-Milner has two important features:
-Feature 1: It supports \emph{parametric polymorphism}, meaning that types can have variables in them, called \emph{type variables}. Lowercase Greek letters are conventionally used for  type variables.
-For example, the type of the identity function is $\alpha\to\alpha$, meaning it takes something of type $\alpha$ and return something of type $\alpha$.
-A function that returns the first element of a list has the type $[\alpha]\to\alpha$.
-Type variables are not the same as variables introduced by $\lambda$-abstractions.
-Feature 2: Remarkably, there is a  simple algorithm for automatically inferring the polymorphic Hindley-Milner type of a $\lambda$-calculus expression~\cite{damas1982principal}.
-Our generative model over programs performs Hindley-Milner type inference during sampling:
-\emph{Unify} in the generative model uses the machinery of Hindley-Milner to
-ensure that the generated programs have valid polymorphic types.
-A satisfactory exposition of Hindley-Milner is beyond the scope of this paper,
-but~\cite{pierce} offers a nice overview of lambda calculus and typing systems like Hindley-Milner.
-
-%% \begin{figure}
-%%   \begin{align}
-%%     \text{primitive types} & 
-%%     \end{align}
-%%   \end{figure}
-
 \section{Neural Recognition Model Architecture}
@@
 
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{0.675\textwidth}    
+    \begin{algorithm}[H]
     \setcounter{algorithm}{2} % there are two algorithms in the main paper.
-\begin{algorithm}[hb]
        \caption{DSL Induction Algorithm}
@@
     \end{algorithm}
+  \end{minipage}
+\end{figure}
 
@@
 
+\subsection{Geometry}
+\begin{lstlisting}
+#(var_half var_name)
+#(var_double var_name)
+#(var_next #(var_half var_name))
+#(concat (turn (just #(var_half var_name))))
+#(@$\lambda$@ (integrate nothing $0 nothing nothing))
+#(@$\lambda$@ (integrate $0 true nothing (just var_name)))
+#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1)))))
+#(integrate (just #(var_half var_name)) true nothing nothing)
+#(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)
+#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))
+#(@$\lambda$@ (@$\lambda$@ (integrate (just #(var_double var_name)) true $0 (just $1))))
+#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))
+#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)))
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2)))))
+#(#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) nothing)
+#(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name))))
+#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0)))
+#(concat #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name)))))
+#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (embed $0)))
+#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false)))))
+#(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn $0) nothing nothing))
+#(@$\lambda$@ (#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (integrate $0 false nothing nothing)))
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit))))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (embed (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))))
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit)))))) nothing)
+#(@$\lambda$@ (concat (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true) (turn nothing)) (integrate (just $0) true nothing nothing)))
+#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0)))
+#(#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) nothing)
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (integrate nothing true nothing (just #(var_half var_name))) nothing)
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (integrate nothing $0 nothing $1) $2 nothing))))
+#(@$\lambda$@ (@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) $1)))))
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false) (just var_name) nothing)
+#(#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) (just #(var_half var_name)))
+#(#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name)))))
+#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) (just (var_double $0))))
+#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (define #(var_half var_name)) (just #(var_half var_name)) $0)))
+#(#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing))
+#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (concat (define $0) #(integrate (just #(var_half var_name)) true nothing nothing))))
+#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit)))))) nothing (just (var_half $0)) false)))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 #(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name))))))
+#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (embed $0))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just $0))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) $0)))
+#(embed (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 #(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name)))))) nothing))
+#(@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (embed (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))) $0)))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (integrate nothing $0 nothing $1) $2 nothing)))) (just #(var_half var_name)) (just $0) true))
+#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) $0)))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) nothing)))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn $0) nothing nothing)) (just #(var_half var_name)))))
+#(#(@$\lambda$@ (@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) $1))))) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false) true)
+#(repeat nothing (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit)))))) nothing (just (var_half $0)) false))) #(var_half var_name)))
+#(#(@$\lambda$@ (@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) $1))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing) false)
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (embed (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))) nothing) (just $0) nothing))
+#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) $0))) (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) $0)))
+#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) #(#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) nothing) nothing $0)))
+#(repeat nothing (concat #(#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)) #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name))))))
+#(@$\lambda$@ (repeat (just $0) #(repeat nothing (concat #(#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)) #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name))))))))
+#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (embed $0))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 #(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name)))))) (just $0))))
+\end{lstlisting}
+
 \bibliography{main}

### 2018-05-18 nips_2018.tex
better list figure
Tweaking

--- 2018-05-18 nips_2018.tex
@@
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
-\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
-
-\newcommand{\system}{\textsc{SCC}~}
+\newcommand{\sem}[1]{\llbracket{} #1 \rrbracket{}}
+
+\newcommand{\system}{\textsc{SCC} }
 \newcommand{\systemEnding}{\textsc{SCC}}
@@
 \newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
-\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
+\newcommand{\codechar}[1]{{\footnotesize{\texttt{``#1''}}}}
 
@@
   jointly training a neural network to efficiently search for programs
-  in the learned DSL.  We use our model to synthesize functions on lists,
+  in the learned DSL.\@ We use our model to synthesize functions on lists,
   edit text, and solve symbolic regression problems, showing how the
@@
 
-The FlashFill system ~\cite{gulwani2011automating} developed by
+The FlashFill system~\cite{gulwani2011automating} developed by
 Microsoft researchers and now embedded in Excel solves problems such
@@
 successful program induction algorithms for many applications, 
-such as handwriting recognition and generation ~\cite{lake2015human},
-procedural graphics ~\cite{ellis2017learning}, question
+such as handwriting recognition and generation~\cite{lake2015human},
+procedural graphics~\cite{ellis2017learning}, question
 answering~\cite{johnson2017clevr} and robot motion
@@
 
-In this work, we consider the problem of building agents that learn to
-solve program induction tasks, and also the problem of acquiring the
-prior knowledge necessary to quickly solve these tasks in a new domain.
-Representative problems in three domains are shown in 
-Table~\ref{initialExampleDSL}.  Our solution is an algorithm that
-grows or boostraps a DSL while jointly training a neural network to help write
-programs in the increasingly rich DSL.  
+In this work, we consider the problem of building agents that learn to solve
+program induction tasks, and also the problem of acquiring the prior knowledge
+necessary to quickly solve these tasks in a new domain.  Representative problems
+in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
+algorithm that grows or boostraps a DSL while jointly training a neural network
+to help write programs in the increasingly rich DSL.
 %
@@
           \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}}       \\
-          ~                                                  \\
-          ~                                                  \\
+          \\
+          \\
           \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
@@
 
-Our \system (\textbf{(S)}search/\textbf{(C})ompress/\textbf{(C})ompile)
+Our \system (\textbf{S}earch/\textbf{C}ompress/\textbf{C}ompile)
 algorithm incorporates these insights by iterating through three steps. 

### 2018-05-18 nips_2018.tex
Follows (some) chktex recommandations

@@
 capturing the human ability to learn to think flexibly and efficiently
-in new domains -- to learn what you need to know about a domain so you
+in new domains --- to learn what you need to know about a domain so you
 don't have to solve new problems starting from scratch.  We are
@@
 quickly, once they acquire the relevant domain expertise.  These
-correspond to tasks solved by short programs -- if you have an
+correspond to tasks solved by short programs --- if you have an
 expressive DSL.  Even with a good DSL, program search may be
@@
 \textbf{generative model} for programs encoded in the DSL, in the
-spirit of the Helmholtz machine ~\cite{hinton1995wake}). The
+spirit of the Helmholtz machine~\cite{hinton1995wake}). The
 recognition model ensures that searching for programs remains
@@
 \system does approximate MAP inference  by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t. the frontiers (Search) and the DSL (Compression):
-\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
+alternating maximization w.r.t.\ the frontiers (Search) and the DSL (Compression):
+\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
@@
 %% where  is large.
-\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.
-the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
+\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
@@
 during the compilation phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta}]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
+posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
   thus amortizing the cost of finding programs with high posterior probability.
@@
   from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
-  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}
-
-  \label{fragmentExample}
+  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
 \end{figure}
@@
   \captionof{table}{Some tasks in our list function domain. See the
-  supplement for the complete data set.}
-  \label{listexamples}
+  supplement for the complete data set.}\label{listexamples}
 \end{wrapfigure}
@@
 enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
-learned appending primitive (Fig.~\ref{initialExampleDSL}) -- and then
+learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
 define $\probability[x|p]$ by approximately marginalizing out the
@@
 convolutional network that observes an image of the target function's
-graph (Fig.~\ref{functions}) -- visually, different kinds of
+graph (Fig.~\ref{functions}) --- visually, different kinds of
 polynomials and rational functions produce different kinds of graphs,
@@
 which has five continuous parameters.
-This phenomenon arises from our Bayesian framing -- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
+This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
 
@@
 \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-but instead learns the parameters of the DSL ($\theta$) -- learning the parameters of a PCFG while not learning any of the structure.
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
+but instead learns the parameters of the DSL ($\theta$) --- learning the parameters of a PCFG while not learning any of the structure.
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
 \footnotetext[2]{Due to a last-minute bug we ran the list processing experiments without training the recognition model on ``Helmholtz machine'' style data.
@@
 \begin{figure}\centering
-  \includegraphics[width = 4.5cm]{figures/list.eps}
+  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps}
   \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
-  \caption{Learning curves for \system both with (\orange{orange}) and without
-  (\teal{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
+  \caption{Learning curves for \system both with (\orange{in orange}) and without
+  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves}
 \end{figure}
@@
 We are in the process of applying our algorithm to generative programs and
-prototyped this with a turtle-like domain: Figure \ref{geomCompiled} gives some
+prototyped this with a turtle-like domain: Figure~\ref{geomCompiled} gives some
 preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more

### 2018-05-18 nips_2018.tex
writing

--- 2018-05-18 nips_2018.tex
@@
   \end{tabular}
-\caption{\% solved before timeout. Solve time: averaged over solved
+\caption{\% held-out test tasks solved. Solve time: averaged over solved
   tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
@@
 \begin{figure}\centering
-  \includegraphics[width = 4.5cm]{figures/list.eps}
+  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps}
   \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}        
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}   
-  \caption{Learning curves for \system both with (\popp{orange}) and without
-  (\pop{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
+  \caption{Learning curves for \system both with (\orange{orange}) and without
+  (\teal{teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves} 
 \end{figure}

### 2018-05-18 nips_2018.tex
--- vs -- and table

--- 2018-05-18 nips_2018.tex
@@
 Nevertheless, much prior work can be modeled within our setup. 
-We compare with the following ablations (Table~\ref{baselineComparisons}):
+We compare with the following ablations (Tbl~\ref{baselineComparisons};
+Fig~\ref{learningCurves}):
 \\\noindent \textbf{No NN:} lesions the recognition model.
@@
 \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-but instead learns the parameters of the DSL ($\theta$) --- learning the parameters of a PCFG while not learning any of the structure.
+but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
-\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments without training the recognition model on ``Helmholtz machine'' style data.
-We are rerunning this experiment and anticipate improved quantitative results for list processing -- notice the recognition model helps much more for text editing and symbolic regression.} &
+\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments
+without training the recognition model on ``Helmholtz machine'' style data. We
+are rerunning this experiment and anticipate improved quantitative results for
+list processing --- notice the recognition model helps much more for text editing and symbolic regression.} &
 %For each domain,

### 2018-05-18 supplement.tex
More changes

@@
 
-The original DSL, while inspired by the Logo~\cite{abelson1974logo} in the sense
+The original DSL, while inspired by the Logo language~(\cite{abelson1974logo}) in the sense
 that ones controls a moving on-off point, was designed to be congitively
@@
 a neural net based distance function to abstract away from noise in the shapes.
+
+The result presented in the main article desribe a sample of tasks and compiled
+new shapes on the top row. On the bottom row is a condesed description of a run
+where are displayed the mean of typical compiled programs on both sides, once
+before any training and once after the last iteration --- note how the
+probability weighting shaped the output space to add structure. In the middle is
+the learning rate measured by holding some tasks out, training on all the other,
+and with each iteration measuring the success rate on the unseen tasks. In the
+given example the split was $25\%$ test and $73\%$ train.
 
@@
 \end{comment}
+
 \vfill

### 2018-05-18 supplement.tex
More dreams in supp

@@
 
-\begin{figure}[h]
+\begin{figure}[htp]
 \centering
@@
 \end{figure}
+
+\begin{figure}[htp]
+  \begin{center}
+  \includegraphics[width = 0.75\textwidth]{figures/moreDreams.png}
+  \end{center}
+\caption{Another example of compiled figures generated by our method.}\label{moreDreams}
+\end{figure}
+
+In Figure~\ref{moreDreams} is another montage with several compiled shapes generated by our
+method after some training. While not all are that regular, often offer high
+level structure in the latter run while no structure is to be found at the
+begining.
 

### 2018-05-18 supplement.tex

@@
 \newcommand{\sem}[1]{\llbracket #1 \rrbracket}
-\newcommand{\system}{\textsc{CoCoSea}~}
-\newcommand{\systemEnding}{\textsc{CoCoSea}}
+\newcommand{\system}{\textsc{SCC}~}
+\newcommand{\systemEnding}{\textsc{SCC}}
 \newcommand{\lowerBound}{\mathscr{L}}
@@
 
-\section{Learning abstract geometry}
-The brifly presented geometry project is a recent addition to the set of
-problems adressed with our method.
-
-The original DSL, while inspired by the Logo language~(\cite{abelson1974logo}) in the sense
-that ones controls a moving on-off point, was designed to be congitively
-plausible in the sense that some abstractions are baked in for curves and a few
-structures --- it was first described in~\cite{sablemeyer2017geom}.
-
-It is to be thought as a turtle with as internal state a position, a facing
-direction, a speed, an acceleration, an angular speed and an angular
-acceleration.
+\section{Learning Generative Graphics Programs}
+
+A natural extension of our work is to consider the problem of learning generative models:
+here, we would learn programs that generate (either deterministically or probabilistically)
+structures like images or words.
+As a first step in this direction,
+we apply \system to synthesizing graphics programs.
+
+A natural starting DSL for graphics programming is the
+Logo language~(\cite{abelson1974logo}), also sometimes called \textbf{turtle graphics}.
+These programs control a pen (sometimes called a ``turtle''),
+and can do things like pick the pen up,
+move the pen forward,
+rotate the pen,
+or trace out a programmatically specified arc.
+We take turtle graphics primitives from prior work~\cite{sablemeyer2017geom}.
+In our setting, we will encapsulate turtle graphics primitives inside of $\lambda$-calculus,
+and seek to infer graphics programs from images: thus the task is to look at an image,
+and write the program that would have drawn it.
+
+%% The brifly presented geometry project is a recent addition to the set
+%% of problems adressed with our method.
+
+%% The original DSL, while inspired by the in the sense that ones
+%% controls a moving on-off point, was designed to be congitively
+%% plausible in the sense that some abstractions are baked in for curves
+%% and a few structures ---
+
+%% It is to be thought as a turtle with as internal state a position, a
+%% facing direction, a speed, an acceleration, an angular speed and an
+%% angular acceleration.
 
@@
 
-In this context programs to be learned correspond to a single shape since they
-don't take an input. All tasks will ignore input and produces a single output.
-This differs from the previous examples where each target could be evaluated on
-various input.
+%% In this context programs to be learned correspond to a single shape since they
+%% don't take an input. All tasks will ignore input and produces a single output.
+%% This differs from the previous examples where each target could be evaluated on
+%% various input.
 

### 2018-05-18 supplement.tex
supplemental edits

@@
 
-\begin{figure}[htp]
+\begin{figure}[h]
 \centering
@@
 \end{figure}
-
-\begin{figure}[htp]
-  \begin{center}
-  \includegraphics[width = 0.75\textwidth]{figures/moreDreams.png}
-  \end{center}
-\caption{Another example of compiled figures generated by our method.}\label{moreDreams}
-\end{figure}
-
-In Figure~\ref{moreDreams} is another montage with several compiled shapes generated by our
-method after some training. While not all are that regular, often offer high
-level structure in the latter run while no structure is to be found at the
-begining.
 

### 2018-05-18 nips_2018.tex
Updated colors and compiling

@@
 \definecolor{pop1}{HTML}{1f78b4}
-\definecolor{pop2}{HTML}{33a02c}
+\definecolor{pop2}{HTML}{297F23}
 \definecolor{pop3}{HTML}{d95f02}
@@
 \end{tabular}
-\caption{\textbf{Left:} syntax trees of two programs sharing common structure, highlighted in {\color{red} red},
+\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
+highlighted in {\orange{orange}},
   from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,

### 2018-06-06 nampi.tex
nampi

--- 2018-05-18 nips_2018.tex
@@
+%%%%%%%% ICML 2018 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+
 \documentclass{article}
 
-% if you need to pass options to natbib, use, e.g.:
-% \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading nips_2018
+% Recommended, but optional, packages for figures and better typesetting:
+\usepackage{microtype}
+\usepackage{graphicx}
+%\usepackage{subfigure}
+\usepackage{booktabs} % for professional tables
 
-% ready for submission
-\usepackage{nips_2018}
+% hyperref makes hyperlinks in the resulting PDF.
+% If your build breaks (sometimes temporarily if a hyperlink spans a page)
+% please comment out the following usepackage line and replace
+% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
+%\usepackage{hyperref}
 
-% to compile a preprint version, e.g., for submission to arXiv, add
-% add the [preprint] option:
-% \usepackage[preprint]{nips_2018}
+% Attempt to make hyperref and algorithmic work together better:
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
 
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2018}
-
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2018}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{stmaryrd}
-
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-\newcommand{\sem}[1]{\llbracket{} #1 \rrbracket{}}
-
-\newcommand{\system}{\textsc{SCC} }
-\newcommand{\systemEnding}{\textsc{SCC}}
-
+\newcommand{\system}{\textsc{DreamCoder}~}
+\newcommand{\systemEnding}{\textsc{DreamCoder}}
 \newcommand{\lowerBound}{\mathscr{L}}
 \newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
-\newcommand{\codechar}[1]{{\footnotesize{\texttt{``#1''}}}}
+\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
+% Use the following line for the initial blind version submitted for review:
+\usepackage[nohyperref]{icml2018}
 
@@
 
+%\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
 \usepackage{mathrsfs}
@@
 \usepackage{amsthm}
-
+% use Times
+\usepackage{times}
+% For figures
+\usepackage{graphicx} % more modern
+%\usepackage{epsfig} % less modern
 \usepackage{subfig} 
@@
 \usepackage{caption}
+%\usepackage{subcaption}
+
+\fvset{fontsize=\footnotesize}
+
 \usepackage{amssymb}
@@
 \usetikzlibrary{fit,bayesnet}
+%\usetikzlibrary{arrows.meta}
+%\usetikzlibrary{positioning}
+%\usetikzlibrary{decorations.text}
+%\usetikzlibrary{decorations.pathmorphing}
 \usepackage{dsfont}
 \usepackage{amsmath}
-\usetikzlibrary{trees}
 
@@
 
-\title{Search, Compress, Compile: Library Learning in Neurally-Guided Bayesian Program Learning}
+% If accepted, instead use the following line for the camera-ready submission:
+%\usepackage[accepted]{icml2018}
 
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
-
-\author{
-  David S.~Hippocampus\thanks{Use footnote for providing further
-    information about author (webpage, alternative
-    address)---\emph{not} for acknowledging funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  %% examples of more authors
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+\icmltitlerunning{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
 
 \begin{document}
-% \nipsfinalcopy is no longer used
 
-\maketitle
+\twocolumn[
+\icmltitle{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
+
+% It is OKAY to include author information, even for blind
+% submissions: the style file will automatically remove it for you
+% unless you've provided the [accepted] option to the icml2018
+% package.
+
+% List of affiliations: The first argument should be a (short)
+% identifier you will use later to specify author affiliations
+% Academic affiliations should list Department, University, City, Region, Country
+% Industry affiliations should list Company, City, Region, Country
+
+% You can specify symbols, otherwise they are numbered in order.
+% Ideally, you should not use this facility. Affiliations will be numbered
+% in order of appearance and this is the preferred way.
+\icmlsetsymbol{equal}{*}
+
+\begin{icmlauthorlist}
+\icmlauthor{Kevin Ellis}{MIT}
+\icmlauthor{Lucas Morales}{MIT}
+\icmlauthor{Matthias}{MIT}
+\icmlauthor{Armando Solar-Lezama}{MIT}
+\icmlauthor{Joshua B. Tenenbaum}{MIT}
+\end{icmlauthorlist}
+
+%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
+\icmlaffiliation{MIT}{MIT}
+%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
+
+\icmlcorrespondingauthor{Kevin Ellis}{ellisk@mit.edu}
+%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
+
+% You may provide any keywords that you
+% find helpful for describing your paper; these are used to populate
+% the "keywords" metadata in the PDF but will not be shown in the document
+\icmlkeywords{Machine Learning, ICML}
+
+\vskip 0.3in
+]
+
+% this must go after the closing bracket ] following \twocolumn[ ...
+
+% This command actually creates the footnote in the first column
+% listing the affiliations and the copyright notice.
+% The command takes one argument, which is text to display at the start of the footnote.
+% The \icmlEqualContribution command is standard text for equal contribution.
+% Remove it (just {}) if you do not need this facility.
+
+%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
 
@@
 %
-
-\begin{table}[t!]
+\begin{table*} %[t!]
   \makebox[\textwidth][c]{
     \scriptsize
-  \tabcolsep=3pt
+  \tabcolsep=4pt
   \renewcommand\code\texttt
@@
     &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=4pt
+    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
       \begin{tabular}{cc}
         \begin{tabular}{c}
-          \code{[7\, 0\, 2]}$\to $\code{[7\, 0\, 2\, 9]}         \\
-          \code{[9\, 3]}$\to $\code{[9\, 3\, 9]} \\
-          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}}       \\
+          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
+          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
+          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
+          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
           \\
@@
       +83 973-831$\to $83.973.831\\
-      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,}
+      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
       }}\\
@@
       \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
-      \popp{$f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}}\\
-      \hspace{\helpSize}($f_1$: \emph{Append the number }\code{k}\emph{ to $\ell$})\\
+      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
+      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
+      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
       %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
@@
       \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
-      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
+      \hspace{\helpSize}($f_3$: \emph{Whether $\ell$ contains }\code{k})\\
     \end{tabular}&
@@
       \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
-      a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}}\\
+      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
         \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
@@
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
-\end{table}
-
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ in List Functions)}\label{initialExampleDSL}
+\end{table*}
 Because any computable learning problem can in principle be cast as
@@
 
-Our \system (\textbf{S}earch/\textbf{C}ompress/\textbf{C}ompile)
+Our \system
 algorithm incorporates these insights by iterating through three steps. 
@@
 
+\begin{comment}
 Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
@@
 allows joint inference of the DSL and neural recognition model.
+  \end{comment}
 
@@
 \section{The \system Algorithm}
- We first mathematically describe our 3-step algorithm as
- an inference procedure for a hierarchical Bayesian model (Section~\ref{mathematicalFraming}),
-  and then describe each step algorithmically in detail (Section~\ref{explorationSection}-\ref{grammarInductionSection}).
+\begin{figure}\centering
+  \begin{tikzpicture}
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs};
+    \node[align = center](DSL) at (3,1) {DSL};
+    \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
 
- \subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
 
-\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.%% \footnote{For example, for string editing,
-%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
-%% and 0 otherwise.}
-Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL, written $\mathcal{D}$.
-We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
-$(\mathcal{D},\theta)$ define a generative model over programs.
-We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
-Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
-\begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \qquad
-  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
-\end{align}
-
-%% Inference in this model is
-%% difficult because the programs are unobserved,
-%% and so we must solve a hard search problem to recover them. To make
-%% search tractable we learn a bottom-up \emph{recognition
-%%   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
-%% The recognition model $q(\cdot )$ is a neural
-%% network that regresses from input/output pairs to a distribution over
-%% programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
-%% inference
-%% scheme~\cite{le2016inference}.
-%%  The neural recognition model and the
-%% generative model embodied in the DSL jointly train each other, as they
-%% iteratively learn to solve more programming tasks.
-
-The above equations summarize the problem from the point of view of an ideal Bayesian learner.
-However, Eq.~\ref{intractableObjectives}
-is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
-summing over the  infinite set of all programs.
-In practice we will only ever be able to sum over a finite set of programs.
-So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
-\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
-
-Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
-\begin{align}
- J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
-\end{align}
-
-%% If we had a $(\mathcal{D},\theta)$ solving Eq.~\ref{intractableObjectives}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
-%% Through this lens we now take as our goal to solve Eq.~\ref{intractableObjectives}.
-%% But even \emph{evaluating} Eq.~\ref{intractableObjectives} is intractable because it involves summing over the infinite set of all possible programs, as an ideal Bayesian learner would.
-%% In practice, we must instead marginalize over
-%% some finite set of programs.
-
-%% In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
-
-%% Now in theory we would like to sum over the entire infinite space
-%% of all programs -- but this is of course impossible.
-
-%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
-%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
-%% \begin{align}
-%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
-%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
-%%   \end{align}
-\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t.\ the frontiers (Search) and the DSL (Compression):
-\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
-want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
-is large.
-%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
-%% where  is large.
-\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
-Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
-
-Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
-during the compilation phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
-  thus amortizing the cost of finding programs with high posterior probability.
-
-\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
-  frontiers.}  Here we train %% a neural network, $q$, to predict a
-%% distribution over programs conditioned on a task. The objective of $q$
-%% is
-$q(p|x)$ to assign high probability to programs $p$ where
-$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
-in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
-%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
-%% $q(p|x)$.
-%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
-%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
-%% This taps into the intuition that programming is primarily a top-down activity:
-%% as human programmers, we often first decide what kind of
-%% programming constructs we might need to use,
-%% and then we figure out how to assemble them in to the desired program.
-%% This approach implements an amortized
-%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
-%% Fig.~\ref{graphicalModel}.
-
-%% Learning the DSL eases the difficulty of synthesis
-%% by exposing a domain-specific basis for constructing programs.
-%% Another complementary means of meeting search is to learn a
-%% bottom-up \emph{recognition model}
-
-\subsection{Searching for Programs}\label{explorationSection}
-
-Now our goal is to search for programs solving the tasks.  We use the simple approach of enumerating programs from
-the DSL  in decreasing order of their probability,
-and then checking if a program $p$ assigns positive
-probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
-into the frontier $\mathcal{F}_x$.
-
-To make this concrete we need to define what programs actually are and
-what form $\probability[p |\mathcal{D},\theta]$ takes.
-We represent programs as $\lambda$-calculus expressions.
-$\lambda$-calculus is a formalism for expressing functional programs
-that closely resembles Lisp,
-including variables, function application, and the ability to create new functions.
-Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
-Our programs are all strongly typed.
-We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
-used in functional programming languages like OCaml and Haskell.
-%% Type variables are always written using lowercase Greek letters
-%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
-%% and returns something of type $\beta$.
-%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
-%% has the type $\tau$.
-%% For example, to describe the type of the identity function
-%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
-%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
-%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
-%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
-%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
-%% See Supplement for more detail on program representation.
-We now define DSLs:
-
-\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
-A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-%\\\noindent\textbf{Definition.}
-A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
-and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
-
-Together with its weight vector,
-a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
-In the supplement, we define this distribution  
-by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$. %%  Care must be taken to ensure that programs are well-typed
-%% and that variable scoping rules are obeyed.
-%% With this distribution in hand, we search for programs by enumerating
-%% $\lambda$-calculus expressions in decreasing order of their probability under
-%% $(\mathcal{D},\theta)$.
-
-Why enumerate, when the program synthesis community has invented many
-sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
-We have two reasons:
-(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
-(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-  the space of  programs.
-
-  However, a drawback of   enumerative search  is that we have no
-efficient means of solving for arbitrary constants that might occur in a
-program. In Sec.~\ref{regressionSection},
-we will show how to find programs with real-valued constants
-by automatically differentiating through the program and setting the constants using gradient descent.
-%% In Sec.~\ref{textSection}
-%% we will show that the bottom-up neural recognition model can learn
-%% which discrete constants should be included in a program.
-
-\subsection{Compilation: Learning a Neural Recognition Model}\label{recognitionSection}
-
-The purpose of training the recognition model is to amortize the cost of searching for
-programs.  It does this by learning to predict, for each task,
-programs with high likelihood according to 
-$\probability[x|p]$ while also being
-probable under the prior $(\mathcal{D},\theta)$.  Concretely,
-the recognition model $q$ predicts, for each
-task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
-\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
-a distribution over programs, $\probability[p|\mathcal{D},\theta =
-  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
-aspect of this framing is that the neural network leverages the
-structure of the learned DSL, so it is \emph{not} responsible for
-generating programs wholesale.  We share this aspect with
-DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.
-
-How should we get the data to train $q$?
-This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
-One approach is to sample programs from the DSL,
-run them to get their input/outputs,
-and then train $q$ to predict the program from the input/outputs.
-This is like how a Helmholtz machine
-trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
-The advantage of ``Helmholtz machine'' training is that
-we can draw unlimited samples from the DSL,
-training on a large amount of data.
-Another approach is
- self-supervised learning,
-training $q$ on the (program, task)
-pairs discovered by search.
-The advantage of self-supervised learning
-is that the training data is much higher quality,
-because we are training on the actual tasks.
-Due to these complementary advantages,
-we train on both these sources of data.
-
-Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
- equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
-%% \begin{equation*}
-%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
-%% \end{equation*}
- where the expectation is taken over tasks.
-Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
- The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
-\begin{align}\nonumber
-\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
-\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
-\end{align}
-\begin{comment}
-  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
-all of our experiments train \system on only a few hundred tasks, which is too little for
-a high-capacity neural network $q$.
-Once we bootstrap a $(\mathcal{D},\theta)$,
-we can draw unlimited samples from $(\mathcal{D},\theta)$
-and train $q$ on those samples.
-
-Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
-the current DSL, running them to get their outputs,
-and then training $q$ to regress from the input/outputs to the program.
-Since these programs map inputs to outputs,
-we need to sample the inputs as well.
-Our solution is to sample the inputs
-from the empirical observed distribution of inputs in $X$.
-  \end{comment}
-
-%% The architecture of $q$ depends upon the domain.  For domains with
-%% sequential structure like strings and lists, we use an RNN to encode
-%% each input/output example.  For domains with fixed-dimensionality
-%% inputs/outputs we use a multilayer perceptron.  We MaxPool across all
-%% of the input/output examples. See Supplement for details.
-
-\subsection{Compression: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
-
-The purpose of the DSL is to
-offer a set of abstractions
-that allow an agent to easily express solutions to the tasks at hand.
-%In the \system algorithm we infer the DSL from a collection of frontiers.
-Intuitively, we want the algorithm to
-look at  the frontiers and
-generalize beyond them, 
-both so the DSL can better express the current solutions,
-and  also so that the DSL might expose new abstractions
-which will later be used to
-discover more programs.
- Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
-We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
-\begin{equation}
-      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
-  \end{equation}
-
-\begin{wrapfigure}{R}{0.6\textwidth}
-  \begin{minipage}{0.6\textwidth}    
-    \begin{algorithm}[H]
-      \caption{The \system Algorithm}
-      \label{mainAlgorithm}
-      \begin{algorithmic}
-        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-        \FOR{$i=1$ {\bfseries to} $I$}
-        %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
-        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
-        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
-        %% \STATE Define $Q_x(z) \propto \begin{cases}
-        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-        %%   0&x\not \in \mathcal{F}_x
-        %% \end{cases}
-        \ENDFOR
-        \STATE \textbf{return} $\mathcal{D},\theta,q$
-      \end{algorithmic}
-    \end{algorithm}
-  \end{minipage}
-\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
-
-We induce a DSL by searching locally through the space of DSLs,
-proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
-The search moves work by introducing new
-$\lambda$-expressions into the DSL.
-We propose these new expressions by extracting fragments of
-programs already in the frontiers (Tbl.~\ref{fragmentExample}).
-An important point here is that we are \emph{not} simply adding
-subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
-extracting fragments that unify with programs in the frontiers.  This
-idea of storing and reusing fragments of expressions comes from
-Fragment Grammars~\cite{tim} and Tree-Substitution
-Grammars~\cite{cohn2010inducing}, and is closely related to the idea
-of antiunification~\cite{henderson2013cumulative}.
-
-\begin{figure}[b!]\tabcolsep=2pt
-\centering  \begin{minipage}[c]{0.2\textwidth}
-  \begin{tikzpicture}[scale=0.7]
-    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
-    %%     \footnotesize Program:\\
-    %%     \code{($\lambda$ (x) (+ (- x) 1))}
-    %% \end{tabular}};
-    %\node at (,0.5) {\code{cons}};
-    \node [rotate=90] at (-2.3,-0.5) {\small program};
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-%    \draw  (l1.south) -- (-0.5,0.5);
-
-  %% \node(c) at (-0.5,-1.5) {\code{-}};
-  %% \node(z) at (0.5,-1.5) {\code{x}};
-
-  %% \draw (0,-1) -- (c.north);
-  %% \draw (0,-1) -- (z.north);
-
-  \begin{scope}[shift={(0,-2.5)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
   \end{scope}
-
-\begin{scope}[shift={(0,-5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  %\node(x1) at (0,-1) {};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.55,-0.4) -- (a.north);
-  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
-
-  \end{scope}
-
     \end{tikzpicture}
-  \end{minipage}
-\hspace{0.1cm}\begin{tabular}{ll}
-    \toprule
-    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
-    \begin{tabular}{l}
-      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
-      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
-    \end{tabular}&
-    \begin{tabular}{l}
-      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
-    \end{tabular}\\\midrule
-    \begin{tabular}{l}
-      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
-      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
-      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
-      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
-      \end{tabular}&
-    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
-      \end{tabular}
-\\\bottomrule\\
-\end{tabular}
-\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
-highlighted in {\orange{orange}},
-  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
-  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
+  \caption{\system solves for programs, the DSL, and a recognition model.
+Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
 \end{figure}
-
-To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-and place a symmetric Dirichlet prior over the weight vector $\theta$.
-%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
- %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
-
-%% Although this  may seem 
-%% very similar to estimating the parameters of a probabilistic context free grammar,
-%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-%% our DSLs are context-sensitive due to the presence of variables
-%% in the programs and also due to the polymorphic typing system.
-%% In the Supplement we derive a tractable MAP estimator for $\theta$.
-Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
-
-\begin{comment}
-\subsection{Implementing \system}
-
-Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
-We note the following implementation details:
-(1) We perform an search cycle before each compression and compilation  cycle.
-(2) On the first iteration, we do \emph{not} train the recognition model on samples from
-the generative model because a generative model has not yet been learned --
-we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
-%% (3) Because the frontiers can grow very large,
-%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
-%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
-(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
-by only summing over the top $K$
-programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
-(4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.
-\end{comment}
-
-%\section{Experiments}
-%% We first present our experimental results qualitatively,
-%% illustrating the learned DSLs.
-%% We then show quantitative results (Sec.~\ref{quantitative})
-%% comparing \system with lesioned versions
-%% of our model that are analogous to prior work.
 
@@
 
-\bibliographystyle{unsrt}
-{\small \bibliography{main}}
+\bibliography{main}
+\bibliographystyle{icml2018}
 \end{document}

### 2018-06-07 nampi.tex
learning from scratch as well as an introduction rewrite

@@
 \usepackage{xcolor}
-\definecolor{pop1}{HTML}{1f78b4}
+\definecolor{pop1}{HTML}{1F78b4}
 \definecolor{pop2}{HTML}{297F23}
-\definecolor{pop3}{HTML}{d95f02}
-\definecolor{orange}{HTML}{d95f02}
+\definecolor{pop3}{HTML}{d95F02}
+\definecolor{orange}{HTML}{d95F02}
 \definecolor{teal}{HTML}{1b9e77}
@@
 
-\begin{abstract}
-  Successful approaches to program induction require a hand-engineered
-  domain-specific language (DSL), constraining the space of allowed
-  programs and imparting prior knowledge of the domain.  We contribute
-  a program induction algorithm called \system that learns a DSL while
-  jointly training a neural network to efficiently search for programs
-  in the learned DSL.\@ We use our model to synthesize functions on lists,
-  edit text, and solve symbolic regression problems, showing how the
-  model learns a domain-specific library of program components for
-  expressing solutions to problems in the domain.
-\end{abstract}
+%% \begin{abstract}
+%%   Successful approaches to program induction require a hand-engineered
+%%   domain-specific language (DSL), constraining the space of allowed
+%%   programs and imparting prior knowledge of the domain.  We contribute
+%%   a program induction algorithm called \system that learns a DSL while
+%%   jointly training a neural network to efficiently search for programs
+%%   in the learned DSL.\@ We use our model to synthesize functions on lists,
+%%   edit text, and solve symbolic regression problems, showing how the
+%%   model learns a domain-specific library of program components for
+%%   expressing solutions to problems in the domain.
+%% \end{abstract}
 
-\section{Introduction}
+\section{Introduction: The Centrality of Domain-Specific Languages}
 
@@
 planning~\cite{devlin2017neural}, to name just a few.
-These systems work in different ways, but most hinge upon having a
+These systems work in different ways, but most hinge upon  a
 carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
@@
 \end{tabular}}
-\caption{Top: Tasks from each domain, each followed by the programs \system discovers for them.
+\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ in List Functions)}\label{initialExampleDSL}
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
 \end{table*}
@@
 %
-so we amortize the cost of search by training a
-neural network to assist the search procedure.
+so we train a
+neural network to guide the search procedure.
 
-Our algorithm takes inspiration from several ways that skilled human
+\section{The \system Algorithm}
+
+We take inspiration from several ways that skilled human
 programmers have learned to code: Skilled coders build libraries of
@@
 for splitting on characters, but also specialize these routines to
-split on particular characters such as spaces or commas that are
-frequently used to delimit substrings across tasks.  Skilled coders
+split on characters such as spaces or commas that are
+frequently used to delimit substrings.  Skilled coders
 also learn to recognize what kinds of programming idioms and library
@@
 learn that if outputs are consistently shorter than inputs, removing
-characters is likely to be part of the solution; if every output
-contains a constant substring (e.g., ``Dr.''), inserting or appending
-that constant string is likely to be a subroutine.
+characters is likely to be part of the solution.%% ; if every output
+%% contains a constant substring (e.g., ``Dr.''), inserting or appending
+%% that constant string is likely to be a subroutine.
 
-Our \system
-algorithm incorporates these insights by iterating through three steps. 
-The \textbf{Search} step takes a given set of \textbf{tasks}, typically
-several hundred, and searches for compact programs that solve these tasks
-guided by the current DSL and neural network.
-The \textbf{Compress} step grows the library (or DSL) of
-domain-specific subroutines which allow the agent to more compactly
-write programs in the domain; it modifies the structure of the DSL by
-discovering regularities across programs found during search, compressing
-them to distill out common code fragments across successful programs.
-The \textbf{Compile} step improves the search procedure by training a neural network to
-write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
-%% and others too on amortized inference.)
+Our algorithm is called \system because it is based on a
+novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
+between ``wake'' and ``sleep'' phases to achieve three goals: finding
+programs that solve tasks; creating a DSL by discovering
+and reusing domain-specific subroutines; and training a neural network
+that efficiently guides search for programs in the DSL. %% The learned DSL
+%% distills commonalities across programs that solve tasks, helping
+%% the agent solve related program induction problems. The neural
+%% network ensures that searching for programs remains tractable
+%% even as the DSL (and hence the search space for programs) expands.
 
@@
 \textbf{generative model} for programs encoded in the DSL, in the
-spirit of the Helmholtz machine~\cite{hinton1995wake}). The
+spirit of the Helmholtz machine~\cite{hinton1995wake}. The
 recognition model ensures that searching for programs remains
@@
 
-We apply \system to three domains:
-list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
- For each of these we initially provide a generic
- set of programming primitives.
-Our algorithm then constructs
-its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+%% Our \system
+%% algorithm incorporates these insights by iterating
+Concretely, our algorithm iterates through three steps. 
+The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
+several hundred, and searches for compact programs that solve these tasks
+guided by the current DSL and neural network.
+The \textbf{Sleep-G} step grows the DSL (or \textbf{G}enerative model) 
+which allows the agent to more compactly
+write programs in the domain; it modifies the structure of the DSL by
+discovering regularities across programs found during waking, compressing
+them to distill out common code fragments across successful programs.
+The \textbf{Sleep-R} step improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
+write programs in the current DSL, in the spirit of ``amortized'' or
+``compiled'' inference~\cite{le2016inference}. %%  (CITE ,
+%% and others too on amortized inference.)
+Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
+\begin{figure}[h]\centering
+  \begin{tikzpicture}
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs};
+    \node[align = center](DSL) at (3,1) {DSL};
+    \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
 
-\begin{comment}
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+  \caption{Algorithm solves for programs, the DSL, and a recognition model. These steps bootstrap off each other in a Helmholtz-machine inspired wake/sleep inference strategy.}  \label{feeding}
+\end{figure}
+
+%% We apply \system to three domains:
+%% list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
+%%  For each of these we initially provide a generic
+%%  set of programming primitives.
+%% Our algorithm then constructs
+%% its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
+
 Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
@@
 which also learns a library of subroutines.
-We find EC motivating, and go beyond it and other prior work through the following contributions:%% \begin{itemize}
+We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
 %%   \setlength\itemsep{0}
@@
 %% \end{itemize}
-\noindent\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
+\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
 (2) We give an algorithm for learning DSLs,
@@
 allows joint inference of the DSL and neural recognition model.
-  \end{comment}
 
@@
 
-\section{The \system Algorithm}
-\begin{figure}\centering
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs};
-    \node[align = center](DSL) at (3,1) {DSL};
-    \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
-  \caption{\system solves for programs, the DSL, and a recognition model.
-Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
-\end{figure}
-
 \section{Programs that manipulate sequences}\label{sequences}
@@
 
-\begin{wrapfigure}{r}{0.55\textwidth}\centering
+\begin{figure}\centering
   \begin{tabular}{lll}
@@
   \end{tabular}
-  \captionof{table}{Some tasks in our list function domain. See the
-  supplement for the complete data set.}\label{listexamples}
-\end{wrapfigure}
+  \captionof{table}{Some tasks in our list function domain.}\label{listexamples}
+\end{figure}
 We created 236 human-interpretable list manipulation tasks, each with 15
 input/output examples (Tbl.~\ref{listexamples}).
-Our data set is interesting in three major ways: many of the tasks require
-complex solutions; the tasks were not generated from some latent DSL, and
-the agent must learn to solve these complicated problems from only 236
-tasks.
+%% Our data set is interesting in three major ways: many of the tasks
+%% require complex solutions; the tasks were not generated from some
+%% latent DSL, and the agent must learn to solve these complicated
+%% problems from only 236 tasks.
 Our data set assumes arithmetic operations as well as sequence operations,
@@
 
-We evaluated \system on  random 50/50 test/train split.
-Interestingly, we found that the recognition model provided little benefit for the training
-tasks. However, it yielded faster search times on held out tasks, allowing
-more tasks to be solved before timing out.
-The system composed 38 new subroutines, yielding a more expressive DSL more closely matching the
-domain (left of Tbl.~\ref{initialExampleDSL}, right of
+We evaluated \system on random 50/50 test/train split.  The system
+composed 38 new subroutines, and rediscovered the higher-order
+function \code{filter}, yielding a more expressive DSL more closely
+matching the domain (left of Tbl.~\ref{initialExampleDSL}, right of
 Fig.~\ref{fragmentExample}). See the supplement for a complete list of
@@
 
-Because our enumerative search procedure cannot generate string
-constants, we instead enumerate programs with string-valued
-parameters.  For example, to learn a program that prepends ``Dr.'', we
-enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
-learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
-define $\probability[x|p]$ by approximately marginalizing out the
-string parameters via a simple dynamic program.
-In Sec.~\ref{regressionSection}, we will use a similar trick to
-synthesize programs containing real numbers, but using gradient
-descent instead of dynamic programming.
+%% Because our enumerative search procedure cannot generate string %
+%% constants, we instead enumerate programs with string-valued
+%% parameters.  For example, to learn a program that prepends ``Dr.'', we
+%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
+%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
+%% define $\probability[x|p]$ by approximately marginalizing out the
+%% string parameters via a simple dynamic program.
+%% In Sec.~\ref{regressionSection}, we will use a similar trick to
+%% synthesize programs containing real numbers, but using gradient
+%% descent instead of dynamic programming.
 
@@
 and so the recognition model can learn to look at a graph and predict
-what kind of function best explains it.  A key difficulty, however, is
-that these problems are best solved with programs containing real
-numbers.  Our solution to this difficulty is to enumerate
- programs with real-valued parameters, and then fit those
-parameters by automatically differentiating through the programs the
-system writes and use gradient descent to fit the parameters.
-We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
-and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
+what kind of function best explains it.  %% A key difficulty, however, is
+%% that these problems are best solved with programs containing real
+%% numbers.  Our solution to this difficulty is to enumerate
+%%  programs with real-valued parameters, and then fit those
+%% parameters by automatically differentiating through the programs the
+%% system writes and use gradient descent to fit the parameters.
+%% We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
+%% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
 
-\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
+\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.5cm}\centering
   \includegraphics[width = \functionSize]{figures/functions/6.png}
@@
   \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
-\end{wrapfigure}
+DSL learns subroutines for polynomials (top row) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (above) and predict which learned subroutines best explain the observation.}\label{functions}
+\end{figure}
 \system learns a DSL containing 13 new functions,
@@
 This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
+
+\section{Learning from Scratch}
+A long-standing dream within the program induction community
+is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
+and then learning to solve a wide swath of
+induction problems~\cite{solomonoff1964formal}.
+All existing systems,
+including ours,
+fall far short of this dream,
+and it is unclear (and we believe unlikely)
+that this dream could ever be realized in its fullness.
+How far can our system push in this direction?
+
+``Learning from scratch'' is subjective, but we believe a reasonable
+starting point is the set of primitives provided in McCarthy's 1959
+Lisp implementation~\cite{mccarthy1960recursive}: these include
+conditionals, \code{lambda}'s, recursion, arithmetic, and the basic
+Lisp operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
+A sensible and modest goal is to start with
+1959 Lisp primitives,
+and then recover a DSL that
+more closely resembles modern functional languages like Haskell and OCaml.
+Recall (Section~\ref{sequences})
+that we initially provided our system with modern functional programming routines like
+\code{map} and \code{fold}.
+
+We ran the following experiment: \system was provided with a subset of
+the 1959 Lisp primitives, and tasked with solving 22 functional
+programming exercises.  After running for 93 hours on 64 CPUs, our
+algorithm solves these exercises, along the way assembling a DSL
+containing 9 subroutines, recovering a modern repertoire of
+functional programming idioms and subroutines, including \code{map},
+\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
+and simple arithmetic operations like incrementing, decrementing, and
+building lists of natural numbers between an interval.
+We did not use the recognition model for this experiment:
+a bottom-up pattern recognizer is of little use
+for acquiring this abstract knowledge
+from less than two dozen problems.
+
+To be clear, we believe that program learners should \emph{not}
+start from scratch,
+but instead should start from
+a rich, domain-agnostic
+basis like those embodied in the standard libraries of modern programming languages.
+What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
+and can in principle recover many of the amenities of modern programming systems,
+provided it is given enough computational power and a suitable
+spectrum of programming exercises.
 
@@
 assumptions about the space of programs and the statement of tasks.
-\begin{wraptable}{r}{0.55\textwidth}
-\vspace{-5pt}
-%% This radically shrinks the table
-\tabcolsep=2.5pt
-\renewcommand{\arraystretch}{0.5}
-\begin{tabular}{lcccccc}
-  \toprule& Ours& 
-    No NN & SE&NPS & PCFG & Enum \\
-  \midrule
-  \multicolumn{7}{c}{\emph{List Processing}}\\
-  \midrule
-  \% solved&\textbf{79\%}\footnotemark & 
-  76\% &71\%&35\%&62\%&37\%\\
-  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
-  \midrule
-  \multicolumn{7}{c}{\emph{Text Editing}}\\
-  \midrule
-  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
-  Solve time& 29s&65s&38s&80s&--&235s\\
-  \midrule
-  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
-  \midrule
-  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
-  Solve time&  24s& 40s  &28s&31s&55s&29s\\
-  \bottomrule
-  \end{tabular}
-\caption{\% held-out test tasks solved. Solve time: averaged over solved
-  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
-  %structure learning.  Enum: model w/o any learning.  %% MDL:
-  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
-  %% symbolic regression MDL is 0 nats.
-  %% For symbolic regression MDL is a
-  %% proxy for \# of continuous parameters.
-}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
+
 Nevertheless, much prior work can be modeled within our setup. 
@@
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
-\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments
-without training the recognition model on ``Helmholtz machine'' style data. We
-are rerunning this experiment and anticipate improved quantitative results for
-list processing --- notice the recognition model helps much more for text editing and symbolic regression.} &
 %For each domain,

### 2018-06-07 nampi.tex
almost 4 pages

@@
 
-\section{Introduction: The Centrality of Domain-Specific Languages}
+\section{Introduction}
 
@@
 
-\section{The \system Algorithm}
+\section{Overview of \system}
 
@@
 and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL. %% The learned DSL
-%% distills commonalities across programs that solve tasks, helping
-%% the agent solve related program induction problems. The neural
-%% network ensures that searching for programs remains tractable
-%% even as the DSL (and hence the search space for programs) expands.
-
+that efficiently guides search for programs in the DSL.
 The learned DSL effectively encodes a prior on programs likely to
@@
 %% and others too on amortized inference.)
-Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
-\begin{figure}[h]\centering
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs};
-    \node[align = center](DSL) at (3,1) {DSL};
-    \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
+%Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
+%% \begin{figure}[h]\centering
+%%   \begin{tikzpicture}
+%%     \begin{scope}[shift = {(1,-1)}]
+%%     \node[align = center](synthesis) at (6,4) {Search for \\programs};
+%%     \node[align = center](DSL) at (3,1) {DSL};
+%%     \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
 
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+%%     \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+%%     \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+%%     \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
 
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
-  \caption{Algorithm solves for programs, the DSL, and a recognition model. These steps bootstrap off each other in a Helmholtz-machine inspired wake/sleep inference strategy.}  \label{feeding}
-\end{figure}
+%%     \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+%%     \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+%%   \end{scope}
+%%     \end{tikzpicture}
+%%   \caption{Algorithm solves for programs, the DSL, and a recognition model. These steps bootstrap off each other in a Helmholtz-machine inspired wake/sleep inference strategy.}  \label{feeding}
+%% \end{figure}
 
@@
 
-%% \begin{itemize}
-%%   \item Fixed DSL, no recognition model, learn parameters $\theta$ of an inductive bias over programs\citep{menon2013machine,singh2015predicting,learningToRank}
-%%   \item The EC algorithm and related work \citep{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14} has no recognition model
-%%     $q$;
-%%     % because this is a primary inspiration, might be worth further mention:
-%%     % different program representation (routing vs. substitution; simple vs polymorphic types);
-%%     % different grammar induction (sequitur vs. fragment grammar)
-%%   \item DeepCoder and RobustFill \citep{balog2016deepcoder,devlin2017robustfill} both fix the
-%%     DSL $\mathcal{D}$ and does not use $\theta$ in favor of solely relying on the neural
-%%     network $q$.
-%%     % although RobustFill has very different enumeration
-%% \end{itemize}
+\section{The \system Algorithm}
 
-%% We cast these problems as \emph{Bayesian Program
-%%   Learning} (BPL; see~\citep{lake2013one,ellis2016sampling,DBLP:conf/icml/LiangJK10}),
-%% where the goal is to infer from an observation $x$ a posterior distribution over programs, $\probability[p|x]$.
-%% A DSL $\mathcal{D}$ specifies the vocabulary in which programs $p$ are written.
-%% We equip our DSLs with a \emph{weight vector} $\theta$; together, $(\mathcal{D},\theta)$
-%% define a probabilistic generative model over programs, $\probability[p|\mathcal{D},\theta]$.
-%% In this BPL setting, $\probability[p|x]\propto \probability[x|p]\probability[p|\mathcal{D},\theta]$,
-%% where the likelihood $\probability[x|p]$ is domain-dependent.
-%% The solid lines in Fig.~\ref{graphicalModel} the diagram this generative model.
-%% Alongside this generative model,
-%% we infer a bottom-up recognition model, $q(x)$, which is a neural network that regresses from observations to a distribution over programs.
+\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
+It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
+%% \footnote{For example, for string editing,
+%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
+%% and 0 otherwise.}
+Its goal is to solve each of the tasks by writing a program,
+and also to infer a DSL, written $\mathcal{D}$,
+which is a generative model over programs, written $\probability[p |\mathcal{D}]$.
+We frame our goal as maximum a posteriori (MAP) inference of $\mathcal{D}$ given $X$:
+$$
+\argmax_{\mathcal{D}}\probability[\mathcal{D}]\prod_{x\in X}\sum_p\probability[x|p]\probability[p|\mathcal{D}]
+$$
+Even evaluating the above objective is intractable due to the sum over all programs,
+and so we instead marginalize over finite sets of programs, called \textbf{frontiers}, one for each task $x\in X$, written $\mathcal{F}_x$, giving the following lower bound on the marginal, which we call $\lowerBound$:
+\begin{equation}\label{lowerBound}
+  \lowerBound  \triangleq\probability[\mathcal{D}]\prod_{x\in X}\sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D}]
+\end{equation}
+We alternate maximization w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ (\textbf{Wake}) and $\mathcal{D}$  (\textbf{Sleep-G}):
+
+\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the programs.} Here $\mathcal{D}$ is fixed and we
+want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
+$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D}]$
+is large.
+Here we use a simple and generic enumerative program synthesis algorithm.
+%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
+%% where  is large.
+\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
+Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\mathcal{D}$.
+To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.
+
+Searching for programs is hard because
+of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
+during the \textbf{Sleep-R} phase: $q$ is trained to approximate the
+posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$,
+  thus amortizing the cost of finding programs with high posterior probability.
+
+\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
+  frontiers.}  Here we train %% a neural network, $q$, to predict a
+%% distribution over programs conditioned on a task. The objective of $q$
+%% is
+$q(p|x)$ to assign high probability to programs $p$ where
+$\probability[x|p]\probability[p|\mathcal{D}]$ is large, because including those programs
+in the frontiers will most increase $\lowerBound$.
+Concretely, the loss for $q(\cdot |x)$ is:
+$$
+\expect_X\left[ \text{KL}\left(\probability[\cdot |x,\mathcal{D}]\; ||\; q(\cdot |x) \right) \right]
+$$
+which minimizes the KL between the distribution predicted by $q$ and the true posterior over programs solving task $x$.
 
@@
 For both these domains we use a bidirectional GRU~\cite{cho2014learning} for
-the recognition model, and initially provide the system with a generic set
-of list processing primitives:
+the recognition model, and initially provide the system with generic list processing primitives:
 \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
@@
 We consider this problem within the context of learning functions that
-manipulate lists, and which also perform arithmetic operations upon lists of numbers.
-
+manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
 \begin{figure}\centering
@@
   \end{tabular}
-  \captionof{table}{Some tasks in our list function domain.}\label{listexamples}
+  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}
 \end{figure}
 We created 236 human-interpretable list manipulation tasks, each with 15
-input/output examples (Tbl.~\ref{listexamples}).
+input/output examples,
 %% Our data set is interesting in three major ways: many of the tasks
@@
 %% problems from only 236 tasks.
-Our data set assumes arithmetic operations as well as sequence operations,
-so we additionally provide our system with the following arithmetic
-primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
-\code{is-prime}.
-
-We evaluated \system on random 50/50 test/train split.  The system
+%% Our data set assumes arithmetic operations as well as sequence operations,
+%% so we additionally provide our system with the following arithmetic
+%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
+%% \code{is-prime}.
+evaluating  on a 50/50 test/train split.  The system
 composed 38 new subroutines, and rediscovered the higher-order
 function \code{filter}, yielding a more expressive DSL more closely
-matching the domain (left of Tbl.~\ref{initialExampleDSL}, right of
-Fig.~\ref{fragmentExample}). See the supplement for a complete list of
-DSL primitives discovered by \systemEnding.
+matching the domain (Tbl.~\ref{initialExampleDSL}, left).
 
@@
 Synthesizing programs that edit text is a classic problem in the
-programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
+programming languages and AI literatures~\cite{lau2001programming},
 and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
@@
 
-We trained our system on a corpus of 109 automatically generated text editing tasks, with 4 input/output examples each.
-After three iterations, it assembles a DSL containing a dozen new functions (center of Fig.~\ref{initialExampleDSL}) that let it solve 
-all of the training tasks.
+We trained our system on 109 synthetic tasks, with 4 input/output examples each.
+After three iterations, it assembles a DSL containing a dozen new functions (Fig.~\ref{initialExampleDSL}, center) solving
+all the training tasks.
 But, how well does the  learned DSL generalized to real text-editing scenarios?
@@
 But, SyGuS comes with a
-different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
+different hand-engineered DSL \emph{for each text editing problem}.%\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
 For these experiments, our system did not use this assistance.}
@@
 addition, multiplication, and division, and task it with solving
-100 symbolic regression problems, each either a polynomial of degree
-1--4 or a rational function.  The recognition model is a
+100 symbolic regression problems, each either a polynomial or rational function.  The recognition model is a
 convolutional network that observes an image of the target function's
@@
 and so the recognition model can learn to look at a graph and predict
-what kind of function best explains it.  %% A key difficulty, however, is
+what kind of function best explains it.
+These programs can contain real numbers,
+which we set with gradient descent,
+and penalize using real numbers by
+incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
+%% A key difficulty, however, is
 %% that these problems are best solved with programs containing real
@@
 For example, it learns to represent linear functions with the program
-\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
+\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the discovered DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
 which has five continuous parameters.
-This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.
+This phenomenon arises from our Bayesian framing.
 
@@
 
-``Learning from scratch'' is subjective, but we believe a reasonable
-starting point is the set of primitives provided in McCarthy's 1959
-Lisp implementation~\cite{mccarthy1960recursive}: these include
-conditionals, \code{lambda}'s, recursion, arithmetic, and the basic
-Lisp operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
+``Learning from scratch'' is subjective, but a reasonable
+starting point is the set of primitives provided in 1959
+Lisp~\cite{mccarthy1960recursive}: these include
+conditionals, recursion, arithmetic, and the 
+list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
 A sensible and modest goal is to start with
@@
 more closely resembles modern functional languages like Haskell and OCaml.
-Recall (Section~\ref{sequences})
+Recall (Sec.~\ref{sequences})
 that we initially provided our system with modern functional programming routines like
@@
 
+\section{Learning Generative Models}
+
+We apply \system to learning generative models for images (Figure~\ref{geomCompiled}) and text (Figure~\ref{regularExpressions}).
+For images, we learn programs in a LOGO/Turtle-like representation,
+and the task is to look at an image and explain it in terms of a graphics program.
+For text, we learn probabilistic regular expressions -- a simple probabilistic program for which inference is always tractable -- and the task is to
+infer a regex from a set of strings.
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{figures/geomCompiled.eps} 
+  \caption{
+  Top left: Example training tasks. Top right: samples from the learned DSL.
+ Bottom: \% holdout testing tasks solved (middle); on the sides are averaged samples from the DSL
+  before any training (left) and after last iteration (right).}\label{geomCompiled}
+
+  \end{figure}
+
 \section{Quantitative Results}\label{quantitative}
 
-We compare with ablations of our model on held out tasks.
-The purpose of this ablation study is 
-both to examine the role of each component of \systemEnding,
-as well as to compare with
-prior approaches in the literature:
-a head-to-head
-comparison of program synthesizers is complicated by the fact that
-each system, including ours, makes idiosyncratic 
-assumptions about the space of programs and the statement of tasks.
+%% We compare with ablations of our model on held out tasks.
+%% The purpose of this ablation study is 
+%% both to examine the role of each component of \systemEnding,
+%% as well as to compare with
+%% prior approaches in the literature:
+%% a head-to-head
+%% comparison of program synthesizers is complicated by the fact that
+%% each system, including ours, makes idiosyncratic 
+%% assumptions about the space of programs and the statement of tasks.
 
-Nevertheless, much prior work can be modeled within our setup. 
-We compare with the following ablations (Tbl~\ref{baselineComparisons};
-Fig~\ref{learningCurves}):
-\\\noindent \textbf{No NN:} lesions the recognition model.
-\\\noindent \textbf{NPS}, which does not learn the DSL,
-instead learning the recognition model
-from samples drawn from the fixed DSL.
-We call this NPS (Neural Program Synthesis)
-because this is closest to how
-RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
+%% Nevertheless, much prior work can be modeled within our setup. 
+%% We compare with the following ablations (Tbl~\ref{baselineComparisons};
+%% Fig~\ref{learningCurves}):
+%% \\\noindent \textbf{No NN:} lesions the recognition model.
+%% \\\noindent \textbf{NPS}, which does not learn the DSL,
+%% instead learning the recognition model
+%% from samples drawn from the fixed DSL.
+%% We call this NPS (Neural Program Synthesis)
+%% because this is closest to how
+%% RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
+%% \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
+%% only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
 This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
-\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
+%% \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
+%% but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
+%% \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
 %For each domain,
@@
 
-\begin{figure}\centering
-  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} 
-  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}
+\begin{figure*}\centering
+  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
+  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
   \caption{Learning curves for \system both with (\orange{in orange}) and without
-  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves}
-\end{figure}
+  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time, averaged only over those tasks that are solved.}\label{learningCurves}
+\end{figure*}
 
+\begin{comment}
 \section{Related Work}
@@
  and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.
-
+  \end{comment} 
 \section{Discussion}
-
-\parbox{0.58\textwidth}{
 
@@
 domains over the generic methods used here.
-We are in the process of applying our algorithm to generative programs and
-prototyped this with a turtle-like domain: Figure~\ref{geomCompiled} gives some
-preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
-details.  Another direction is to explore DSL meta-learning: can we find a
+Another direction is to explore DSL meta-learning: can we find a
 \emph{single} universal primitive set that could effectively bootstrap DSLs for
 new domains, including the three domains considered,  but also many others?
-}\hfill
-\begin{minipage}{0.4\textwidth}
-  \centering
-  \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
-  \captionof{figure}{
-  Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
-  before any training (left) and after last iteration (right).}
-\label{geomCompiled}
-\end{minipage}
 

### 2018-06-07 nips_2018.tex
fixed typos
nampi
almost 4 pages

@@
         \begin{tabular}{c}
-          \code{[7\, 0\, 2]}$\to $\code{[7\, 0\, 2\, 9]}         \\
-          \code{[9\, 3]}$\to $\code{[9\, 3\, 9]} \\
-          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}}       \\
+          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
+          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
+          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
+          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
           \\
@@
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ in List Functions)}\label{initialExampleDSL}
 \end{table}
@@
 \system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.%% \footnote{For example, for string editing,
+It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
+%% \footnote{For example, for string editing,
 %%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,

### 2018-06-07 nampi.tex
writing

@@
 
-The FlashFill system~\cite{gulwani2011automating} developed by
-Microsoft researchers and now embedded in Excel solves problems such
-as these and is probably the best known practical program-induction
-algorithm, but researchers in programming languages and AI have built
-successful program induction algorithms for many applications, 
-such as handwriting recognition and generation~\cite{lake2015human},
-procedural graphics~\cite{ellis2017learning}, question
+The FlashFill system~\cite{gulwani2011automating} embedded in
+Microsoft Excel solves problems such as these and is probably the best
+known practical program-induction algorithm, but researchers in
+programming languages and AI have built successful program induction
+algorithms for many applications, such as handwriting recognition and
+generation~\cite{lake2015human}, procedural
+graphics~\cite{ellis2017learning}, question
 answering~\cite{johnson2017clevr} and robot motion
-planning~\cite{devlin2017neural}, to name just a few.
-These systems work in different ways, but most hinge upon  a
-carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
-especially true for systems such as FlashFill that aim to induce a
-wide range of programs very quickly, in a few seconds or less.  DSLs
-constrain the search over programs with strong prior knowledge in the
-form of a restricted set of programming primitives tuned to the needs
-of the domain: for text editing, these are operations like appending
-strings and splitting on characters.
+planning~\cite{devlin2017neural}.  These systems
+work in different ways, but most hinge upon a carefully engineered
+\textbf{Domain Specific Language (DSL)}.  This is especially true for
+systems such as FlashFill that aim to induce a wide range of programs
+very quickly, in a few seconds or less.  DSLs constrain the search
+over programs with strong prior knowledge in the form of a restricted
+set of programming primitives tuned to the needs of the domain: for
+text editing, these are operations like appending and
+splitting on characters.
 
@@
           \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
-          \\
           \\
@@
       % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
-      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
-      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
+      %% \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
+      %% \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
       \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
@@
         \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
-        \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
-      \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
+      %%   \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
+      %% \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
   \end{tabular}&
@@
 programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
-engineering, to learn to synthesize large bodies of code, or to learn
-complex programs starting from scratch.  Ours is a basic AI goal:
+engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
 capturing the human ability to learn to think flexibly and efficiently
@@
 don't have to solve new problems starting from scratch.  We are
-focused on the kinds of problems that humans can solve relatively
+focused on the kinds of problems that people solve relatively
 quickly, once they acquire the relevant domain expertise.  These
 correspond to tasks solved by short programs --- if you have an
-expressive DSL.  Even with a good DSL, program search may be
-intractable;
+expressive DSL.  %% Even with a good DSL, program search may be
+%% intractable;
 %
@@
 %
-so we train a
-neural network to guide the search procedure.
+%% so we train a
+%% neural network to guide the search procedure.
 
@@
 for splitting on characters, but also specialize these routines to
-split on characters such as spaces or commas that are
-frequently used to delimit substrings.  Skilled coders
+split on frequent delimiters such as spaces or commas. Skilled coders
 also learn to recognize what kinds of programming idioms and library
@@
 built on a formalism known as Fragment Grammars~\cite{tim};
-and (3) We give a hierarchical Bayesian framing of
-the problem that
-allows joint inference of the DSL and neural recognition model.
+and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.
 
@@
 \section{Programs that manipulate sequences}\label{sequences}
-We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
-For both these domains we use a bidirectional GRU~\cite{cho2014learning} for
-the recognition model, and initially provide the system with generic list processing primitives:
+We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}),
+using a bidirectional GRU~\cite{cho2014learning} for
+the recognition model, and initially providing the system with generic sequence manipulation primitives:
 \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
@@
     repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
-    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+%    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+%    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
     keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
@@
 We created 236 human-interpretable list manipulation tasks, each with 15
-input/output examples,
+input/output examples.
 %% Our data set is interesting in three major ways: many of the tasks
@@
 %% \code{is-prime}.
-evaluating  on a 50/50 test/train split.  The system
+%% evaluating  on a 50/50 test/train split.
+In solving these tasks, the system
 composed 38 new subroutines, and rediscovered the higher-order
-function \code{filter}, yielding a more expressive DSL more closely
-matching the domain (Tbl.~\ref{initialExampleDSL}, left).
+function \code{filter} (Tbl.~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
+%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
 
@@
 This prior work presumes a hand-engineered DSL.
-We show \system can instead start out with generic sequence manipulation
+Here we will instead start out with generic sequence manipulation
 primitives and recover many of the higher-level building blocks that have
 made these other text editing systems successful.
-
 %% Because our enumerative search procedure cannot generate string %
@@
 100 symbolic regression problems, each either a polynomial or rational function.  The recognition model is a
-convolutional network that observes an image of the target function's
+convnet that observes an image of the target function's
 graph (Fig.~\ref{functions}) --- visually, different kinds of
 polynomials and rational functions produce different kinds of graphs,
-and so the recognition model can learn to look at a graph and predict
+and so the recognition model can  look at a graph and predict
 what kind of function best explains it.
@@
 which we set with gradient descent,
-and penalize using real numbers by
+and penalize continuous parameters by
 incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
@@
 %% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
-
-\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.5cm}\centering
+\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
+  \begin{minipage}[c]{0.3\columnwidth}
   \includegraphics[width = \functionSize]{figures/functions/6.png}
-  \includegraphics[width = \functionSize]{figures/functions/48.png}
+  \includegraphics[width = \functionSize]{figures/functions/48.png}\\
   \includegraphics[width = \functionSize]{figures/functions/102.png}
   \includegraphics[width = \functionSize]{figures/functions/116.png}\\
-  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
+  \vspace{1pt}%\includegraphics[width = \functionSize]{figures/functions/181.png}
+  %\includegraphics[width = \functionSize]{figures/functions/160.png}
   \includegraphics[width = \functionSize]{figures/functions/160.png}
     \includegraphics[width = \functionSize]{figures/functions/149.png}
-    \includegraphics[width = \functionSize]{figures/functions/187.png}
+  \end{minipage}
+  \begin{minipage}[c]{0.69\columnwidth}    
     \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top row) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (above) and predict which learned subroutines best explain the observation.}\label{functions}
+DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}
+        \end{minipage}
 \end{figure}
-\system learns a DSL containing 13 new functions,
-most of which are templates for polynomials of different orders or ratios of polynomials.
-It also learns to find programs that minimize the number of continuous degrees of freedom.
-For example, it learns to represent linear functions with the program
-\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the discovered DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
+We learn a DSL containing 13 new functions,
+mainly templates for polynomials of different orders or ratios of polynomials.
+It also learns to find programs minimizing the number of continuous parameters ---
+learning to represent linear functions with 
+\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
 which has five continuous parameters.
@@
 that this dream could ever be realized in its fullness.
-How far can our system push in this direction?
-
+How far can we in this direction?
 ``Learning from scratch'' is subjective, but a reasonable
@@
 list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
-A sensible and modest goal is to start with
-1959 Lisp primitives,
+A  basic first goal is to start with
+these primitives,
 and then recover a DSL that
@@
 Recall (Sec.~\ref{sequences})
-that we initially provided our system with modern functional programming routines like
+that we initially provided our system with functional programming routines like
 \code{map} and \code{fold}.
@@
 algorithm solves these exercises, along the way assembling a DSL
-containing 9 subroutines, recovering a modern repertoire of
+with a modern repertoire of
 functional programming idioms and subroutines, including \code{map},
 \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
-and simple arithmetic operations like incrementing, decrementing, and
+and simple arithmetic operations like 
 building lists of natural numbers between an interval.
-We did not use the recognition model for this experiment:
-a bottom-up pattern recognizer is of little use
-for acquiring this abstract knowledge
-from less than two dozen problems.
+%% We did not use the recognition model for this experiment:
+%% a bottom-up pattern recognizer is of little use
+%% for acquiring this abstract knowledge
+%% from less than two dozen problems.
 
@@
 \section{Quantitative Results}\label{quantitative}
+We evaluate  on held-out testing tasks,
+measuring how many
+tasks are solved and how long it takes to solve them~\ref{learningCurves}.
+Prior to any learning,
+the system cannot find solutions for most of the tasks,
+and those it does solve take a long time;
+with more wake/sleep iterations,
+we converge upon DSLs and recognition models that
+more closely match the domains.
 
@@
   \caption{Learning curves for \system both with (\orange{in orange}) and without
-  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time, averaged only over those tasks that are solved.}\label{learningCurves}
+  (\teal{in teal}) the recognition model.
+Solid lines: \% holdout testing tasks solved w/ 10m timeout.
+Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}
 \end{figure*}
@@
 symbolic representations and bottom-up neural networks --- both of them learned
---- helps make program induction systems more generally useful for AI. Many
+--- can help make program induction systems more generally useful for AI. Many
 directions remain open.

### 2018-06-07 nampi.tex
shortened

@@
 
-\section{Overview of \system}
+%\section{Overview of \system}
 
@@
 reusable subroutines that are shared across related programming tasks,
-and can be composed to generate increasingly complex and powerful
+and can be composed to generate increasingly complex
 subroutines.  In text editing, a good library should support routines
@@
 learn that if outputs are consistently shorter than inputs, removing
-characters is likely to be part of the solution.%% ; if every output
+characters is likely part of the solution.%% ; if every output
 %% contains a constant substring (e.g., ``Dr.''), inserting or appending
@@
 The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
-several hundred, and searches for compact programs that solve these tasks
+several hundred, and searches for compact programs that solve these tasks,
 guided by the current DSL and neural network.
@@
 
-Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
+\textbf{Related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
 both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},

### 2018-06-07 nampi.tex
shortened

@@
 to desired outputs, based on observing example input-output pairs.
-People can induce programs flexibly across many different domains, and
-remarkably, often from just one or a few examples.  For instance, if
+People can induce programs flexibly across many different domains, often from just one or a few examples.  For instance, if
 shown that a text-editing program should map ``Jane Morris Goodall''
@@
 to ``R. E. Leakey''; if instead the first input mapped to
-``Dr. Jane'', ``Goodall, Jane'', or ''Morris'', we might have guessed
-the latter should map to ``Dr. Richard'', ``Leakey, Richard'', or
-``Erskine'', respectively.
+``Dr. Jane'' or ``Goodall, Jane'', we might guess
+the latter should map to ``Dr. Richard'' or ``Leakey, Richard'',
+ respectively.
 
@@
 known practical program-induction algorithm, but researchers in
-programming languages and AI have built successful program induction
-algorithms for many applications, such as handwriting recognition and
+programming languages and AI have had successes in many domains, such as handwriting recognition and
 generation~\cite{lake2015human}, procedural
@@
 \textbf{Domain Specific Language (DSL)}.  This is especially true for
-systems such as FlashFill that aim to induce a wide range of programs
+systems like FlashFill that induce a wide range of programs
 very quickly, in a few seconds or less.  DSLs constrain the search
 over programs with strong prior knowledge in the form of a restricted
-set of programming primitives tuned to the needs of the domain: for
+set of programming primitives finely tuned to the domain: for
 text editing, these are operations like appending and
@@
 \end{table*}
-Because any computable learning problem can in principle be cast as
+Because any learning problem can in principle be cast as
 program induction, it is important to delimit our focus.  In contrast
@@
 don't have to solve new problems starting from scratch.  We are
-focused on the kinds of problems that people solve relatively
+focused on problems that people solve relatively
 quickly, once they acquire the relevant domain expertise.  These
@@
 recognition model ensures that searching for programs remains
-tractable even as the DSL (and hence the search space for programs)
+tractable even as the DSL %% (and hence the search space for programs)
 expands.
@@
 Here we will instead start out with generic sequence manipulation
-primitives and recover many of the higher-level building blocks that have
-made these other text editing systems successful.
+primitives.%%  and recover many of the higher-level building blocks that have
+%% made these other text editing systems successful.
 %% Because our enumerative search procedure cannot generate string %
@@
 
-To be clear, we believe that program learners should \emph{not}
+We believe that program learners should \emph{not}
 start from scratch,
@@
 a rich, domain-agnostic
-basis like those embodied in the standard libraries of modern programming languages.
+basis like those embodied in the standard libraries of modern  languages.
 What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
@@
 provided it is given enough computational power and a suitable
-spectrum of programming exercises.
+spectrum of tasks.% programming exercises.
 
@@
 measuring how many
-tasks are solved and how long it takes to solve them~\ref{learningCurves}.
+tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
 Prior to any learning,
@@
 with more wake/sleep iterations,
-we converge upon DSLs and recognition models that
-more closely match the domains.
+we converge upon DSLs and recognition models 
+more closely matching the domain.
 

### 2018-06-07 nampi.tex
shortened and random edits

@@
 and it is unclear (and we believe unlikely)
-that this dream could ever be realized in its fullness.
-How far can we in this direction?
+that this dream could ever be fully realized.
+How far can we push in this direction?
 ``Learning from scratch'' is subjective, but a reasonable
@@
 efficiently deploy the DSL on new tasks. We believe this integration of top-down
-symbolic representations and bottom-up neural networks --- both of them learned
+symbolic representations and bottom-up neural nets --- both of them learned
 --- can help make program induction systems more generally useful for AI. Many
@@
 domains over the generic methods used here.
-Another direction is to explore DSL meta-learning: can we find a
+Another direction is DSL meta-learning: can we find a
 \emph{single} universal primitive set that could effectively bootstrap DSLs for
-new domains, including the three domains considered,  but also many others?
+new domains, including the domains considered here,  but also many others?
 

### 2018-06-08 nampi.tex
added unformatted regex figure
shortened

@@
 \newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
+\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
 \newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
@@
 \definecolor{pop1}{HTML}{1F78b4}
-\definecolor{pop2}{HTML}{297F23}
+\definecolor{pop2}{HTML}{164C13}
 \definecolor{pop3}{HTML}{d95F02}
@@
 \icmlauthor{Matthias}{MIT}
+\icmlauthor{Maxwell Nye}{MIT}
 \icmlauthor{Armando Solar-Lezama}{MIT}
@@
 %\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
 
@@
 
-\section{Programs that manipulate sequences}\label{sequences}
-We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}),
+\section{Experiments}
+
+\subsection{Programs that manipulate sequences}\label{sequences}
+We apply \system to list processing %% (Section~\ref{listSection})
+and text editing, %% (Section~\ref{textSection})
 using a bidirectional GRU~\cite{cho2014learning} for
@@
 
-\subsection{List Processing}\label{listSection}
-Synthesizing programs that manipulate data structures is a widely studied
+%\subsubsection{List Processing}\label{listSection}
+\textbf{List Processing:} Synthesizing programs that manipulate data structures is a widely studied
 problem in the programming languages community~\cite{feser2015synthesizing}.
@@
 manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
-\begin{figure}\centering
+We created 236 human-interpretable list manipulation tasks, each with 15
+input/output examples.
+%% Our data set is interesting in three major ways: many of the tasks
+%% require complex solutions; the tasks were not generated from some
+%% latent DSL, and the agent must learn to solve these complicated
+%% problems from only 236 tasks.
+%% Our data set assumes arithmetic operations as well as sequence operations,
+%% so we additionally provide our system with the following arithmetic
+%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
+%% \code{is-prime}.
+%% evaluating  on a 50/50 test/train split.
+In solving these tasks, the system
+composed 38 new subroutines, and rediscovered the higher-order
+function \code{filter} ($f_1$ in Table~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
+%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
+\begin{figure}[b]\centering
   \begin{tabular}{lll}
@@
 \end{figure}
-We created 236 human-interpretable list manipulation tasks, each with 15
-input/output examples.
-%% Our data set is interesting in three major ways: many of the tasks
-%% require complex solutions; the tasks were not generated from some
-%% latent DSL, and the agent must learn to solve these complicated
-%% problems from only 236 tasks.
-%% Our data set assumes arithmetic operations as well as sequence operations,
-%% so we additionally provide our system with the following arithmetic
-%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
-%% \code{is-prime}.
-%% evaluating  on a 50/50 test/train split.
-In solving these tasks, the system
-composed 38 new subroutines, and rediscovered the higher-order
-function \code{filter} (Tbl.~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
-%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
 
-\subsection{Text Editing}\label{textSection}
-Synthesizing programs that edit text is a classic problem in the
-programming languages and AI literatures~\cite{lau2001programming},
-and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
-This prior work presumes a hand-engineered DSL.
-Here we will instead start out with generic sequence manipulation
-primitives.%%  and recover many of the higher-level building blocks that have
-%% made these other text editing systems successful.
+%\subsubsection{Text Editing}\label{textSection}
+\textbf{Text Editing:} Synthesizing programs that edit text is a classic problem in the
+programming languages and AI literatures~\cite{gulwani2011automating,lau2001programming}.
+This prior work uses hand-engineered DSLs.
+Here, we instead start out with generic sequence manipulation
+primitives and recover many of the higher-level building blocks that 
+have made these other systems successful.
 %% Because our enumerative search procedure cannot generate string %
@@
 
-We trained our system on 109 synthetic tasks, with 4 input/output examples each.
+We trained our system on 109 automatically-generated text editing tasks, with 4 input/output examples each.
 After three iterations, it assembles a DSL containing a dozen new functions (Fig.~\ref{initialExampleDSL}, center) solving
@@
 But, SyGuS comes with a
-different hand-engineered DSL \emph{for each text editing problem}.%\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
+different hand-engineered DSL \emph{for each text editing problem}.
+%\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
 For these experiments, our system did not use this assistance.}
@@
 
-\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
+\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
 We apply \system
@@
 mainly templates for polynomials of different orders or ratios of polynomials.
-It also learns to find programs minimizing the number of continuous parameters ---
+The model also learns to find programs minimizing the number of continuous parameters ---
 learning to represent linear functions with 
@@
 
-\section{Learning from Scratch}
+\subsection{Learning from Scratch}
 A long-standing dream within the program induction community
@@
 fall far short of this dream,
-and it is unclear (and we believe unlikely)
+%and it is unclear (and we believe unlikely)
+and we believe it is unlikely
 that this dream could ever be fully realized.
-How far can we push in this direction?
+%How far can we push in this direction?
 ``Learning from scratch'' is subjective, but a reasonable
@@
 
-\section{Learning Generative Models}
+\subsection{Learning Generative Models}
 
-We apply \system to learning generative models for images (Figure~\ref{geomCompiled}) and text (Figure~\ref{regularExpressions}).
-For images, we learn programs in a LOGO/Turtle-like representation,
+We apply \system to learning generative models for images and text (Fig.~\ref{geomCompiled}-\ref{regularExpressions}).
+For images, we learn programs controlling a simulated ``pen,''
 and the task is to look at an image and explain it in terms of a graphics program.
 For text, we learn probabilistic regular expressions -- a simple probabilistic program for which inference is always tractable -- and the task is to
-infer a regex from a set of strings.
+infer a regex from a collection of strings.
 \begin{figure}[t]
@@
   \end{figure}
+\begin{figure}\centering
+\tiny
+  \begin{tabular}{lll}
+\multicolumn{3}{c}{Tasks and synthesized programs:}\\
+cut&F &Moss Side \\ 
+control&CL &Burnage \\
+  control&F & City Centre\\
+cut& PCFL& Brooklands \\
+    \hline
+\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash u\textbackslash w*)*} \\
+    \bottomrule
+  \end{tabular}
+  \quad
+    \begin{tabular}{l}
+\multicolumn{1}{c}{DSL:}\\
+$f_1 = $ \scode{\textbackslash u\textbackslash w*}\\
+$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}\\
+$f_3(x) =  f_2(\scode{space}) = $\scode{( \textbar\textbackslash u\textbackslash w*)*}\\
+\hline
+$f_4(x) = $ \scode{($x$*$x$)} (equivalent to regex `plus')\\
+$f_5 = f_4(\scode{\textbackslash l}) = \scode{\textbackslash l*\textbackslash l}$\\
+    \bottomrule
+  \end{tabular}
+\\
+\begin{tabular}{ccc}
+\multicolumn{3}{c}{samples from synthesized programs:}\\
+ \scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash u\textbackslash w*)*} \\
+\hline
+ya&DQDF & Vr DR \\ 
+fcbadcvoaobvjf&F &BeF lKQ \\
+mgs&F& W\\
+piljnl& KI& kqBfZ 0 \\
+    \bottomrule
+  \end{tabular}
+            \centering 
+            \includegraphics[width=.25\textwidth]{figures/nampiregexjoshcurve.png}   
+\caption{regex stuff}\label{regularExpressions}
+\end{figure}
 
-\section{Quantitative Results}\label{quantitative}
+\subsection{Quantitative Results}\label{quantitative}
 We evaluate  on held-out testing tasks,

### 2018-06-08 nampi.tex
shortened things

@@
 
+\usepackage{dblfloatfix}    % To enable figures at the bottom of page
 % Recommended, but optional, packages for figures and better typesetting:
@@
     \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
-    \popp{$f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))}}\\
+    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
     \hspace{\helpSize}\emph{($f_6$: rational function)}\\
@@
   \\\bottomrule\\
-\end{tabular}}
+\end{tabular}}\vspace{-0.5cm}
 \caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}\vspace{-0.5cm}
 \end{table*}
@@
 \end{equation}
-We alternate maximization w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ (\textbf{Wake}) and $\mathcal{D}$  (\textbf{Sleep-G}):
+We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ %% (\textbf{Wake})
+and $\mathcal{D}$:  %% (\textbf{Sleep-G})
 
-\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the programs.} Here $\mathcal{D}$ is fixed and we
+\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $\mathcal{D}$ is fixed and we
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
@@
 and text editing, %% (Section~\ref{textSection})
-using a bidirectional GRU~\cite{cho2014learning} for
+using a GRU~\cite{cho2014learning} for
 the recognition model, and initially providing the system with generic sequence manipulation primitives:
@@
 \begin{figure}[b]\centering
-  \begin{tabular}{lll}
+\vspace{-0.5cm}  \begin{tabular}{lll}
     \toprule
     Name & Input & Output \\\midrule
-    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
+    repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
 %    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
@@
 %    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
+    keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
     product & [7\, 1\, 6\, 2] & 84 \\
@@
   \end{tabular}
-  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}
+  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
 \end{figure}
@@
 addition, multiplication, and division, and task it with solving
-100 symbolic regression problems, each either a polynomial or rational function.  The recognition model is a
+100 %% symbolic regression
+problems, each either a polynomial or rational function.  The recognition model is a
 convnet that observes an image of the target function's
@@
 polynomials and rational functions produce different kinds of graphs,
-and so the recognition model can  look at a graph and predict
+and so the convnet can  look at a graph and predict
 what kind of function best explains it.
@@
 \begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
-  \begin{minipage}[c]{0.3\columnwidth}
+  \begin{minipage}[c]{0.3\columnwidth}%\vspace{-0.3cm}
   \includegraphics[width = \functionSize]{figures/functions/6.png}
@@
     \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}
+DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}\vspace{-1cm}
         \end{minipage}
@@
 
-We ran the following experiment: \system was provided with a subset of
-the 1959 Lisp primitives, and tasked with solving 22 functional
+We ran the following experiment: \system was given a subset of
+the 1959 Lisp primitives, and tasked with solving 22
 programming exercises.  After running for 93 hours on 64 CPUs, our
@@
 \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
-and simple arithmetic operations like 
+and  arithmetic operations like 
 building lists of natural numbers between an interval.
@@
 
-\subsection{Quantitative Results}\label{quantitative}
+\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
 We evaluate  on held-out testing tasks,
@@
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
-  \caption{Learning curves for \system both with (\orange{in orange}) and without
+\vspace{-0.4cm}  \caption{Learning curves for \system both with (\orange{in orange}) and without
     (\teal{in teal}) the recognition model.
 Solid lines: \% holdout testing tasks solved w/ 10m timeout.
-Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}
+Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}\vspace{-0.5cm}
 \end{figure*}
 
-\begin{comment}
-\section{Related Work}
- Our work is far from the first for learning to learn programs,
- an idea that goes back to Solomonoff~\cite{solomonoff1989system}:
-
- \noindent \textbf{Deep learning:} Much recent work in the ML community has
- focused on creating neural networks that regress from
- input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. %This family of work is closest to our RF/DC baseline.
-\systemEnding's recognition model draws heavily from this line of work, particularly from~\cite{menon2013machine}.
-We see these prior works as operating in a different regime: typically, they train with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
- Our work   considers a weakly-supervised regime where ground truth programs are not provided and
- the agent must learn from at most a few hundred tasks,
- which is facilitated by our ``Helmholtz machine'' style recognition model.%(Sec.~\ref{recognitionSection}).
-
- \noindent \textbf{Inventing new subroutines for program induction:}
- Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting.
-We find this work inspiring and motivating,
- and extend it along two dimensions: (1) we propose a new algorithm for
- inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
- and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
-
-\noindent\textbf{Bayesian Program
- Learning:} Our work is an instance of
- Bayesian Program
- Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
-Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
- and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.
-  \end{comment} 
 \section{Discussion}
@@
 bootstrapping a DSL with new domain-specific primitives that the algorithm
-itself discovers, together with a neural recognition model that learns how to
-efficiently deploy the DSL on new tasks. We believe this integration of top-down
-symbolic representations and bottom-up neural nets --- both of them learned
---- can help make program induction systems more generally useful for AI. Many
-directions remain open.
+itself discovers, together with a neural recognition model that learns to
+efficiently deploy the DSL on new tasks. %% We believe this integration of top-down
+%% symbolic representations and bottom-up neural nets --- both of them learned
+%% --- can help make program induction systems more generally useful for AI. 
+%Many
+%directions remain open.
 %\paragraph{Future.}
-Two immediate goals are to integrate more sophisticated neural recognition
+Two immediate future goals are to integrate more sophisticated neural recognition
 models~\cite{devlin2017robustfill} and program
@@
 Another direction is DSL meta-learning: can we find a
-\emph{single} universal primitive set that could effectively bootstrap DSLs for
+\emph{single} universal primitive set that could bootstrap DSLs for
 new domains, including the domains considered here,  but also many others?

### 2018-06-08 short.tex
2 pages

--- 2018-06-08 nampi.tex
@@
   \end{figure}
-\begin{figure}\centering
+\begin{figure}
+\centering
 \tiny
-  \begin{tabular}{lll}
-\multicolumn{3}{c}{Tasks and synthesized programs:}\\
+\tabcolsep=2pt
+\vspace{0pt}
+\begin{tabular}{cc}
+\begin{tabular}{ccc}
+\multicolumn{3}{c}{Tasks:}\\
+\hline
 cut&F &Moss Side \\ 
@@
 cut& PCFL& Brooklands \\
+\multicolumn{3}{c}{Learned generative models:}\\
 \hline
-\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash u\textbackslash w*)*} \\
-    \bottomrule
+\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash
+u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash
+u\textbackslash w*)*} \\
+\multicolumn{3}{c}{Samples from synthesized generative models:}\\
+\hline
+ya&DQDF & Vr DR \\ 
+glrwfdcnc&F &BeF lKQ \\
+mgs&F& W\\
+piljnl& KI& kqBfZ 0 \\
+kj& F& ON\\
+zci& GL& Bttc \\
+sxpm& F& S \\
+\hline
+\\
 \end{tabular} 
-  \quad
+&
+\begin{tabular}{c}
 \begin{tabular}{l}
-\multicolumn{1}{c}{DSL:}\\
-$f_1 = $ \scode{\textbackslash u\textbackslash w*}\\
-$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}\\
-$f_3(x) =  f_2(\scode{space}) = $\scode{( \textbar\textbackslash u\textbackslash w*)*}\\
+\multicolumn{1}{c}{Learned DSL:}\\
 \hline
-$f_4(x) = $ \scode{($x$*$x$)} (equivalent to regex `plus')\\
-$f_5 = f_4(\scode{\textbackslash l}) = \scode{\textbackslash l*\textbackslash l}$\\
-    \bottomrule
+$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
+$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash
+u\textbackslash w*)*}\\
+$f_3(x) = f_2(\scode{space}) = $\scode{( \textbar\textbackslash
+u\textbackslash w*)*}\\
+\hline
+$f_4(x) = $ \scode{($x$*$x$)}\\
+\emph{ (equivalent to regex `plus')}\\
+$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash
+l*\textbackslash l}$\\
+\hline
 \end{tabular}
 \\
-\begin{tabular}{ccc}
-\multicolumn{3}{c}{samples from synthesized programs:}\\
- \scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash u\textbackslash w*)*} \\
-\hline
-ya&DQDF & Vr DR \\ 
-fcbadcvoaobvjf&F &BeF lKQ \\
-mgs&F& W\\
-piljnl& KI& kqBfZ 0 \\
-    \bottomrule
+\\
+\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
 \end{tabular}
-            \centering 
-            \includegraphics[width=.25\textwidth]{figures/nampiregexjoshcurve.png}   
-\caption{regex stuff}\label{regularExpressions}
+\end{tabular}
+\vspace{-.3cm}
+\caption{regex stuff}
 \end{figure}

### 2018-06-08 short.tex
shortened

--- 2018-06-08 nampi.tex
@@
 
-\subsection{Learning from Scratch}
-A long-standing dream within the program induction community
-is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
-and then learning to solve a wide swath of
-induction problems~\cite{solomonoff1964formal}.
-All existing systems,
-including ours,
-fall far short of this dream,
-%and it is unclear (and we believe unlikely)
-and we believe it is unlikely
-that this dream could ever be fully realized.
-%How far can we push in this direction?
-``Learning from scratch'' is subjective, but a reasonable
-starting point is the set of primitives provided in 1959
-Lisp~\cite{mccarthy1960recursive}: these include
-conditionals, recursion, arithmetic, and the 
-list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
-A  basic first goal is to start with
-these primitives,
-and then recover a DSL that
-more closely resembles modern functional languages like Haskell and OCaml.
-Recall (Sec.~\ref{sequences})
-that we initially provided our system with functional programming routines like
-\code{map} and \code{fold}.
-
-We ran the following experiment: \system was given a subset of
-the 1959 Lisp primitives, and tasked with solving 22
-programming exercises.  After running for 93 hours on 64 CPUs, our
-algorithm solves these exercises, along the way assembling a DSL
-with a modern repertoire of
-functional programming idioms and subroutines, including \code{map},
-\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
-and  arithmetic operations like 
-building lists of natural numbers between an interval.
-%% We did not use the recognition model for this experiment:
-%% a bottom-up pattern recognizer is of little use
-%% for acquiring this abstract knowledge
-%% from less than two dozen problems.
-
-We believe that program learners should \emph{not}
-start from scratch,
-but instead should start from
-a rich, domain-agnostic
-basis like those embodied in the standard libraries of modern  languages.
-What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
-and can in principle recover many of the amenities of modern programming systems,
-provided it is given enough computational power and a suitable
-spectrum of tasks.% programming exercises.
-
-\subsection{Learning Generative Models}
-
-We apply \system to learning generative models for images and text (Fig.~\ref{geomCompiled}-\ref{regularExpressions}).
-For images, we learn programs controlling a simulated ``pen,''
-and the task is to look at an image and explain it in terms of a graphics program.
-For text, we learn probabilistic regular expressions -- a simple probabilistic program for which inference is always tractable -- and the task is to
-infer a regex from a collection of strings.
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{figures/geomCompiled.eps} 
-  \caption{
-  Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: \% holdout testing tasks solved (middle); on the sides are averaged samples from the DSL
-  before any training (left) and after last iteration (right).}\label{geomCompiled}
-
-  \end{figure}
-\begin{figure}\centering
-\tiny
-  \begin{tabular}{lll}
-\multicolumn{3}{c}{Tasks and synthesized programs:}\\
-cut&F &Moss Side \\ 
-control&CL &Burnage \\
-  control&F & City Centre\\
-cut& PCFL& Brooklands \\
-    \hline
-\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash u\textbackslash w*)*} \\
-    \bottomrule
-  \end{tabular}
-  \quad
-    \begin{tabular}{l}
-\multicolumn{1}{c}{DSL:}\\
-$f_1 = $ \scode{\textbackslash u\textbackslash w*}\\
-$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}\\
-$f_3(x) =  f_2(\scode{space}) = $\scode{( \textbar\textbackslash u\textbackslash w*)*}\\
-\hline
-$f_4(x) = $ \scode{($x$*$x$)} (equivalent to regex `plus')\\
-$f_5 = f_4(\scode{\textbackslash l}) = \scode{\textbackslash l*\textbackslash l}$\\
-    \bottomrule
-  \end{tabular}
-\\
-\begin{tabular}{ccc}
-\multicolumn{3}{c}{samples from synthesized programs:}\\
- \scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash u\textbackslash w*)*} \\
-\hline
-ya&DQDF & Vr DR \\ 
-fcbadcvoaobvjf&F &BeF lKQ \\
-mgs&F& W\\
-piljnl& KI& kqBfZ 0 \\
-    \bottomrule
-  \end{tabular}
-            \centering 
-            \includegraphics[width=.25\textwidth]{figures/nampiregexjoshcurve.png}   
-\caption{regex stuff}\label{regularExpressions}
-\end{figure}
-
 \subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}

### 2018-06-08 short.tex
eviscerating the short paper

@@
 \end{table*}
-Because any learning problem can in principle be cast as
-program induction, it is important to delimit our focus.  In contrast
+%% Because any learning problem can in principle be cast as
+%% program induction, it is important to delimit our focus.  
+In contrast
 to computer assisted programming~\cite{solar2008program} or genetic
@@
 in new domains --- to learn what you need to know about a domain so you
-don't have to solve new problems starting from scratch.  We are
-focused on problems that people solve relatively
-quickly, once they acquire the relevant domain expertise.  These
-correspond to tasks solved by short programs --- if you have an
-expressive DSL.  %% Even with a good DSL, program search may be
+don't have to solve new problems starting from scratch.  %% We are
+%% focused on problems that people solve relatively
+%% quickly, once they acquire the relevant domain expertise.  These
+%% correspond to tasks solved by short programs --- if you have an
+%% expressive DSL.
+%% Even with a good DSL, program search may be
 %% intractable;
@@
 
-%\section{Overview of \system}
+\section{The \system Algorithm}
 
@@
 
-Our algorithm is called \system because it is based on a
+Our algorithm is called \system because it incorporates these insights into a
 novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
@@
 recognition model ensures that searching for programs remains
-tractable even as the DSL %% (and hence the search space for programs)
+tractable even as the DSL (and hence the search space for programs)
 expands.
@@
 %% algorithm incorporates these insights by iterating
-Concretely, our algorithm iterates through three steps. 
-The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
-several hundred, and searches for compact programs that solve these tasks,
-guided by the current DSL and neural network.
-The \textbf{Sleep-G} step grows the DSL (or \textbf{G}enerative model) 
+Concretely, our algorithm iterates through three cycles:
+\\\noindent\textbf{Wake Cycle}: Here we take a given set of \textbf{tasks}, typically
+several hundred, and search for compact programs that solve these tasks.
+Our current implementation uses a simple and generic enumerative program synthesis algorithm,
+enumerating programs written in the current DSL in order of their probability according to the neural network.
+\\\noindent\textbf{Sleep-G Cycle}: Here we grow the the DSL (or \textbf{G}enerative model),
 which allows the agent to more compactly
-write programs in the domain; it modifies the structure of the DSL by
+write programs in the domain. We modify the structure of the DSL by
 discovering regularities across programs found during waking, compressing
 them to distill out common code fragments across successful programs.
-The \textbf{Sleep-R} step improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
-write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}. %%  (CITE ,
-%% and others too on amortized inference.)
-%Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
-%% \begin{figure}[h]\centering
-%%   \begin{tikzpicture}
-%%     \begin{scope}[shift = {(1,-1)}]
-%%     \node[align = center](synthesis) at (6,4) {Search for \\programs};
-%%     \node[align = center](DSL) at (3,1) {DSL};
-%%     \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
-
-%%     \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-%%     \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-%%     \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-%%     \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-%%     \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-%%   \end{scope}
-%%     \end{tikzpicture}
-%%   \caption{Algorithm solves for programs, the DSL, and a recognition model. These steps bootstrap off each other in a Helmholtz-machine inspired wake/sleep inference strategy.}  \label{feeding}
-%% \end{figure}
-
-%% We apply \system to three domains:
-%% list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
-%%  For each of these we initially provide a generic
-%%  set of programming primitives.
-%% Our algorithm then constructs
-%% its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
-
-\textbf{Related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
-both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
-FlashFill:~\cite{gulwani2011automating}),
-neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
-A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
-which also learns a library of subroutines.
-We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
-%%   \setlength\itemsep{0}
-%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
-%%   \item We give an algorithm for learning DSLs, built on a formalism known
-%%     as Fragment Grammars~\cite{tim}; and 
-%%   \item We give a hierarchical Bayesian framing of the problem that allows a
-%%     joint inference of the DSL and recognition model.
-%% \end{itemize}
-\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
-(2) We give an algorithm for learning DSLs,
-built on a formalism known as Fragment Grammars~\cite{tim};
-and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.
-
-\section{The \system Algorithm}
-
-\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
-%% \footnote{For example, for string editing,
-%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
-%% and 0 otherwise.}
-Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL, written $\mathcal{D}$,
-which is a generative model over programs, written $\probability[p |\mathcal{D}]$.
-We frame our goal as maximum a posteriori (MAP) inference of $\mathcal{D}$ given $X$:
-$$
-\argmax_{\mathcal{D}}\probability[\mathcal{D}]\prod_{x\in X}\sum_p\probability[x|p]\probability[p|\mathcal{D}]
-$$
-Even evaluating the above objective is intractable due to the sum over all programs,
-and so we instead marginalize over finite sets of programs, called \textbf{frontiers}, one for each task $x\in X$, written $\mathcal{F}_x$, giving the following lower bound on the marginal, which we call $\lowerBound$:
-\begin{equation}\label{lowerBound}
-  \lowerBound  \triangleq\probability[\mathcal{D}]\prod_{x\in X}\sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D}]
-\end{equation}
-We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ %% (\textbf{Wake})
-and $\mathcal{D}$:  %% (\textbf{Sleep-G})
-
-\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $\mathcal{D}$ is fixed and we
-want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D}]$
-is large.
-Here we use a simple and generic enumerative program synthesis algorithm.
-%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
-%% where  is large.
-\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\mathcal{D}$.
-To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.
-
-Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
-during the \textbf{Sleep-R} phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$,
+The technical machinery behind this DSL learning
+comes from a formalism known as Fragment Grammars~\cite{tim},
+which casts DSL induction as grammar induction.
+\\\noindent\textbf{Sleep-R Cycle}: Here we improve the search procedure by training a neural network (the \textbf{R}ecognition model) to
+predict a distribution over programs written in the current DSL, in the spirit of ``amortized'' or
+``compiled'' inference~\cite{le2016inference}.
+The recognition model is trained, conditioned on a task, to assign high probability to
+programs that (1) have high prior probability according to the current DSL,
+while (2) also assigning high likelihood to the task at hand,
 thus amortizing the cost of finding programs with high posterior probability.
-
-\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
-  frontiers.}  Here we train %% a neural network, $q$, to predict a
-%% distribution over programs conditioned on a task. The objective of $q$
-%% is
-$q(p|x)$ to assign high probability to programs $p$ where
-$\probability[x|p]\probability[p|\mathcal{D}]$ is large, because including those programs
-in the frontiers will most increase $\lowerBound$.
-Concretely, the loss for $q(\cdot |x)$ is:
-$$
-\expect_X\left[ \text{KL}\left(\probability[\cdot |x,\mathcal{D}]\; ||\; q(\cdot |x) \right) \right]
-$$
-which minimizes the KL between the distribution predicted by $q$ and the true posterior over programs solving task $x$.
 
@@
 %% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
-\begin{figure}[b]\centering
-\vspace{-0.5cm}  \begin{tabular}{lll}
-    \toprule
-    Name & Input & Output \\\midrule
-    repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
-%    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-%    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-    keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
-    product & [7\, 1\, 6\, 2] & 84 \\
-    \bottomrule
-  \end{tabular}
-  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
-\end{figure}
+%% \begin{figure}[b]\centering
+%% \vspace{-0.5cm}  \begin{tabular}{lll}
+%%     \toprule
+%%     Name & Input & Output \\\midrule
+%%     repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
+%% %    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+%%     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
+%% %    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+%%     keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
+%%     product & [7\, 1\, 6\, 2] & 84 \\
+%%     \bottomrule
+%%   \end{tabular}
+%%   \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
+%% \end{figure}
 

### 2018-06-08 short.tex
only slightly longer than two pages

@@
 
-Much of everyday human thinking and learning can be understood in
-terms of program induction: constructing a procedure that maps inputs
-to desired outputs, based on observing example input-output pairs.
-People can induce programs flexibly across many different domains, often from just one or a few examples.  For instance, if
+%% Much of everyday human thinking and learning can be understood in
+%% terms of program induction: constructing a procedure that maps inputs
+%% to desired outputs, based on observing example input-output pairs.
+Humans can induce programs flexibly across many different domains, often from just one or a few examples.  If
 shown that a text-editing program should map ``Jane Morris Goodall''
 to ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey''
-to ``R. E. Leakey''; if instead the first input mapped to
+to ``R. E. Leakey''; if the first input mapped to
 ``Dr. Jane'' or ``Goodall, Jane'', we might guess
-the latter should map to ``Dr. Richard'' or ``Leakey, Richard'',
+the latter mapped to ``Dr. Richard'' or ``Leakey, Richard'',
  respectively.
@@
 programming languages and AI have had successes in many domains, such as handwriting recognition and
-generation~\cite{lake2015human}, procedural
-graphics~\cite{ellis2017learning}, question
+generation~\cite{lake2015human}, question
 answering~\cite{johnson2017clevr} and robot motion
@@
 work in different ways, but most hinge upon a carefully engineered
-\textbf{Domain Specific Language (DSL)}.  This is especially true for
-systems like FlashFill that induce a wide range of programs
-very quickly, in a few seconds or less.  DSLs constrain the search
+\textbf{Domain Specific Language (DSL)}.  %% This is especially true for
+%% systems like FlashFill that induce a wide range of programs
+%% very quickly, in a few seconds or less.  
+DSLs constrain the search
 over programs with strong prior knowledge in the form of a restricted
-set of programming primitives finely tuned to the domain: for
+library of primitives tuned to the domain: for
 text editing, these are operations like appending and
-splitting on characters.
+splitting on strings.
 
@@
 program induction tasks, and also the problem of acquiring the prior knowledge
-necessary to quickly solve these tasks in a new domain.  Representative problems
-in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
+necessary to quickly solve these tasks in a new domain; see
+ Table~\ref{initialExampleDSL}.  Our solution is an
 algorithm that grows or boostraps a DSL while jointly training a neural network
@@
 Concretely, our algorithm iterates through three cycles:
-\\\noindent\textbf{Wake Cycle}: Here we take a given set of \textbf{tasks}, typically
-several hundred, and search for compact programs that solve these tasks.
-Our current implementation uses a simple and generic enumerative program synthesis algorithm,
+\\\noindent\textbf{Wake Cycle}: Given a set of \textbf{tasks}, typically
+several hundred, we search for compact programs that solve these tasks.
+Currently we take the simple strategy of
 enumerating programs written in the current DSL in order of their probability according to the neural network.
@@
 which casts DSL induction as grammar induction.
-\\\noindent\textbf{Sleep-R Cycle}: Here we improve the search procedure by training a neural network (the \textbf{R}ecognition model) to
+\\\noindent\textbf{Sleep-R Cycle}: This trains a neural net (the \textbf{R}ecognition model) to
 predict a distribution over programs written in the current DSL, in the spirit of ``amortized'' or
 ``compiled'' inference~\cite{le2016inference}.
-The recognition model is trained, conditioned on a task, to assign high probability to
-programs that (1) have high prior probability according to the current DSL,
-while (2) also assigning high likelihood to the task at hand,
+The network is trained, conditioned on a task, to assign high probability to
+programs that have high prior probability according to the current DSL,
+while also assigning high likelihood to the task at hand,
 thus amortizing the cost of finding programs with high posterior probability.
@@
 
-\subsection{Programs that manipulate sequences}\label{sequences}
 We apply \system to list processing %% (Section~\ref{listSection})
 and text editing, %% (Section~\ref{textSection})
-using a GRU~\cite{cho2014learning} for
+using a recurrent network for
 the recognition model, and initially providing the system with generic sequence manipulation primitives:
@@
 We consider this problem within the context of learning functions that
-manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
-We created 236 human-interpretable list manipulation tasks, each with 15
-input/output examples.
+manipulate lists,
+creating 236 human-interpretable list manipulation tasks (Table~\ref{initialExampleDSL}, left).
 %% Our data set is interesting in three major ways: many of the tasks
@@
 
-We trained our system on 109 automatically-generated text editing tasks, with 4 input/output examples each.
-After three iterations, it assembles a DSL containing a dozen new functions (Fig.~\ref{initialExampleDSL}, center) solving
-all the training tasks.
-But, how well does the  learned DSL generalized to real text-editing scenarios?
-We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+We trained our system on 109 automatically-generated text editing tasks.
+After three wake/sleep cycles, it assembles a DSL containing a dozen new functions (Fig.~\ref{initialExampleDSL}, center) solving
+the training tasks.
+%But, how well does the  learned DSL generalized to real text-editing scenarios?
+We also tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
 \system solves 3.7\% of the problems with an average search time of 235 seconds.
@@
 
-\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
+\textbf{Symbolic Regression: Programs from visual input.}
 We apply \system
@@
 what kind of function best explains it.
-These programs can contain real numbers,
-which we set with gradient descent,
-and penalize continuous parameters by
-incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
+%% These programs can contain real numbers,
+%% which we set with gradient descent,
+%% and penalize continuous parameters by
+%% incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
 %% A key difficulty, however, is
@@
 \end{figure}
-We learn a DSL containing 13 new functions,
-mainly templates for polynomials of different orders or ratios of polynomials.
-The model also learns to find programs minimizing the number of continuous parameters ---
-learning to represent linear functions with 
-\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
-which has five continuous parameters.
-This phenomenon arises from our Bayesian framing.
+%% We learn a DSL containing 13 new functions,
+%% mainly templates for polynomials of different orders or ratios of polynomials.
+%% The model also learns to find programs minimizing the number of continuous parameters ---
+%% learning to represent linear functions with 
+%% \code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
+%% which has five continuous parameters.
+%% This phenomenon arises from our Bayesian framing.
 
-\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
-We evaluate  on held-out testing tasks,
-measuring how many
-tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
-Prior to any learning,
-the system cannot find solutions for most of the tasks,
-and those it does solve take a long time;
-with more wake/sleep iterations,
-we converge upon DSLs and recognition models 
-more closely matching the domain.
+%% \subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
+%% We evaluate  on held-out testing tasks,
+%% measuring how many
+%% tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
+%% Prior to any learning,
+%% the system cannot find solutions for most of the tasks,
+%% and those it does solve take a long time;
+%% with more wake/sleep iterations,
+%% we converge upon DSLs and recognition models 
+%% more closely matching the domain.
 
@@
     (\teal{in teal}) the recognition model.
+Iterations: \# wake/sleep cycles.
 Solid lines: \% holdout testing tasks solved w/ 10m timeout.
-Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}\vspace{-0.5cm}
+Dashed lines: Average solve time.}\label{learningCurves}\vspace{-0.5cm}
 \end{figure*}

### 2018-06-12 short.tex
workshop cross submission

@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage[nohyperref]{icml2018}
+\usepackage[nohyperref,accepted]{icml2018}
 
@@
 Microsoft Excel solves problems such as these and is probably the best
-known practical program-induction algorithm, but researchers in
-programming languages and AI have had successes in many domains, such as handwriting recognition and
-generation~\cite{lake2015human}, question
-answering~\cite{johnson2017clevr} and robot motion
+known practical program-induction algorithm, but researchers %%  in
+%% programming languages and AI 
+have had many successes: handwriting recognition and
+generation~\cite{lake2015human}; question
+answering~\cite{johnson2017clevr}; and robot motion
 planning~\cite{devlin2017neural}.  These systems
@@
 %% very quickly, in a few seconds or less.  
-DSLs constrain the search
-over programs with strong prior knowledge in the form of a restricted
-library of primitives tuned to the domain: for
+DSLs constrain the space
+of programs with strong prior knowledge in the form of a restricted
+set of primitives tuned to the domain: for
 text editing, these are operations like appending and
-splitting on strings.
+splitting on characters.
 
@@
 
-We take inspiration from several ways that skilled human
+We are inspired by several ways that skilled human
 programmers have learned to code: Skilled coders build libraries of
-reusable subroutines that are shared across related programming tasks,
-and can be composed to generate increasingly complex
+reusable subroutines shared across related programming tasks,
+that can be composed to generate increasingly complex
 subroutines.  In text editing, a good library should support routines
 for splitting on characters, but also specialize these routines to
-split on frequent delimiters such as spaces or commas. Skilled coders
-also learn to recognize what kinds of programming idioms and library
-routines would be useful for solving the task at hand, even if they
-cannot instantly work out the details.  In text editing, one might
-learn that if outputs are consistently shorter than inputs, removing
-characters is likely part of the solution.%% ; if every output
+split on frequent delimiters like spaces. Skilled coders
+also learn to recognize which programming idioms and library
+routines are useful for solving the task at hand, even if they
+cannot instantly work out the details.  %% In text editing, one might
+%% learn that if outputs are consistently shorter than inputs, removing
+%% characters is likely part of the solution.
+%% ; if every output
 %% contains a constant substring (e.g., ``Dr.''), inserting or appending
@@
 Our algorithm is called \system because it incorporates these insights into a
-novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
+novel kind of ``wake-sleep'' learning (c.f. \citep{hinton1995wake}), iterating
 between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve tasks; creating a DSL by discovering
+programs that solve tasks; growing a DSL by discovering
 and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL.
+that  guides search for programs in the DSL.
 The learned DSL effectively encodes a prior on programs likely to
-solve tasks in the domain, while the neural net looks at the example
-input-output pairs for a specific task and produces a ``posterior''
+solve tasks in the domain, while the neural net looks at %the example
+%input-output pairs for
+a specific task and produces a ``posterior''
 for programs likely to solve that specific task.  The neural network
@@
 recognition model ensures that searching for programs remains
-tractable even as the DSL (and hence the search space for programs)
+tractable even as the DSL 
 expands.
@@
 Concretely, our algorithm iterates through three cycles:
-\\\noindent\textbf{Wake Cycle}: Given a set of \textbf{tasks}, typically
-several hundred, we search for compact programs that solve these tasks.
-Currently we take the simple strategy of
+\\\noindent\textbf{Wake Cycle}: Given a set of \textbf{tasks}, 
+ we search for compact programs that solve these tasks,
+taking the simple strategy of
 enumerating programs written in the current DSL in order of their probability according to the neural network.
-\\\noindent\textbf{Sleep-G Cycle}: Here we grow the the DSL (or \textbf{G}enerative model),
-which allows the agent to more compactly
-write programs in the domain. We modify the structure of the DSL by
+\\\noindent\textbf{Sleep-G Cycle}: Here we grow  the DSL (\textbf{G}enerative model),
+allowing the agent to more compactly
+write programs in the domain. We modify the DSL by
 discovering regularities across programs found during waking, compressing
-them to distill out common code fragments across successful programs.
+them to distill out common code fragments. % across successful programs.
 The technical machinery behind this DSL learning
-comes from a formalism known as Fragment Grammars~\cite{tim},
-which casts DSL induction as grammar induction.
-\\\noindent\textbf{Sleep-R Cycle}: This trains a neural net (the \textbf{R}ecognition model) to
-predict a distribution over programs written in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}.
+comes from a formalism known as Fragment Grammars~\cite{tim}.
+%which casts DSL induction as grammar induction.
+\\\noindent\textbf{Sleep-R Cycle}: This trains a neural net (\textbf{R}ecognition model) to
+predict a distribution over programs in the current DSL, in the spirit of ``amortized''  inference~\cite{le2016inference}.
 The network is trained, conditioned on a task, to assign high probability to
-programs that have high prior probability according to the current DSL,
+programs that have high prior probability according to the DSL,
 while also assigning high likelihood to the task at hand,
@@
 using a recurrent network for
-the recognition model, and initially providing the system with generic sequence manipulation primitives:
+the recognition model, initially providing the system with generic sequence manipulation primitives:
 \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
@@
 %% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
-\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
-  \begin{minipage}[c]{0.3\columnwidth}%\vspace{-0.3cm}
+\begin{figure}\vspace{-0.1cm} \newcommand{\functionSize}{1.1cm}\centering
+  \begin{minipage}[c]{0.3\columnwidth}\vspace{0.5cm}
   \includegraphics[width = \functionSize]{figures/functions/6.png}
@@
   \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
-\vspace{-0.4cm}  \caption{Learning curves for \system both with (\orange{in orange}) and without
+  %\vspace{-0.4cm}
+  \caption{Learning curves for \system both with (\orange{in orange}) and without
     (\teal{in teal}) the recognition model.
@@
 %\paragraph{Future.}
-Two immediate future goals are to integrate more sophisticated neural recognition
-models~\cite{devlin2017robustfill} and program
-synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the generic methods used here.
-Another direction is DSL meta-learning: can we find a
+%% Two immediate future goals are to integrate more sophisticated neural recognition
+%% models~\cite{devlin2017robustfill} and program
+%% synthesizers~\cite{solar2008program}, which may improve performance in some
+%% domains over the generic methods used here.
+One future direction is DSL meta-learning: can we find a
 \emph{single} universal primitive set that could bootstrap DSLs for

### 2018-06-13 archive.tex
progress on archive posting

--- 2018-06-08 nampi.tex
@@
 generation~\cite{lake2015human}, procedural
-graphics~\cite{ellis2017learning}, question
+graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
 answering~\cite{johnson2017clevr} and robot motion
@@
 over programs with strong prior knowledge in the form of a restricted
-set of programming primitives finely tuned to the domain: for
+inventory of programming primitives finely tuned to the domain: for
 text editing, these are operations like appending and
@@
 to help write programs in the increasingly rich DSL.
-%
-% (JBT: MAYBE INSERT FIGURE FROM ICML PAPER?)
-%
+
 \begin{table*} %[t!]
@@
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}\vspace{-0.5cm}
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}%\vspace{-0.5cm}
 \end{table*}
+
 Because any learning problem can in principle be cast as
@@
 
-We take inspiration from several ways that skilled human
+Our algorithm takes inspiration from several ways that skilled human
 programmers have learned to code: Skilled coders build libraries of
 reusable subroutines that are shared across related programming tasks,
-and can be composed to generate increasingly complex
+and can be composed to generate increasingly complex and powerful
 subroutines.  In text editing, a good library should support routines
 for splitting on characters, but also specialize these routines to
-split on frequent delimiters such as spaces or commas. Skilled coders
+split on particular characters such as spaces or commas that are
+frequently used to delimit substrings across tasks.  Skilled coders
 also learn to recognize what kinds of programming idioms and library
@@
 learn that if outputs are consistently shorter than inputs, removing
-characters is likely part of the solution.%% ; if every output
-%% contains a constant substring (e.g., ``Dr.''), inserting or appending
-%% that constant string is likely to be a subroutine.
+characters is likely to be part of the solution; if every output
+contains a constant substring (e.g., ``Dr.''), inserting or appending
+that constant string is likely to be a subroutine.
 
@@
 %% algorithm incorporates these insights by iterating
-Concretely, our algorithm iterates through three steps. 
-The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
+Concretely, our algorithm iterates through three phases. 
+The \textbf{Wake} phase takes a given set of \textbf{tasks}, typically
 several hundred, and searches for compact programs that solve these tasks,
 guided by the current DSL and neural network.
-The \textbf{Sleep-G} step grows the DSL (or \textbf{G}enerative model) 
+The \textbf{Sleep-G} phase grows the DSL (or \textbf{G}enerative model),
 which allows the agent to more compactly
-write programs in the domain; it modifies the structure of the DSL by
-discovering regularities across programs found during waking, compressing
-them to distill out common code fragments across successful programs.
-The \textbf{Sleep-R} step improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
+write programs in the domain. We modify the structure of the DSL by
+discovering regularities across programs, compressing
+them to distill out common abstractions across  programs discovered during waking.\footnote{This is loosely biologically
+    inspired by the formation of abstractions during sleep memory
+    consolidation~\cite{DUDAI201520}}
+The \textbf{Sleep-R} phase improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
 write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}. %%  (CITE ,
-%% and others too on amortized inference.)
-%Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
-%% \begin{figure}[h]\centering
-%%   \begin{tikzpicture}
-%%     \begin{scope}[shift = {(1,-1)}]
-%%     \node[align = center](synthesis) at (6,4) {Search for \\programs};
-%%     \node[align = center](DSL) at (3,1) {DSL};
-%%     \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};
-
-%%     \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-%%     \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-%%     \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-%%     \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-%%     \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-%%   \end{scope}
-%%     \end{tikzpicture}
-%%   \caption{Algorithm solves for programs, the DSL, and a recognition model. These steps bootstrap off each other in a Helmholtz-machine inspired wake/sleep inference strategy.}  \label{feeding}
-%% \end{figure}
-
-%% We apply \system to three domains:
-%% list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
-%%  For each of these we initially provide a generic
-%%  set of programming primitives.
-%% Our algorithm then constructs
-%% its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
-
-\textbf{Related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
+``compiled'' inference~\cite{le2016inference}.
+We train the recognition model on  two data sources: samples
+from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
+programs found during waking.%%  As the DSL
+%% improves and as search finds more programs, the recognition model gets
+%% both more data to train on, and better data.
+\footnote{These two sources are also
+    loosely biologically inspired by the distinct episodic replay and
+    hallucination components of dream sleep~\cite{fosse2003dreaming}}
+
+\textbf{Brisk overview of related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
 both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
@@
 and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.
+up
 
@@
 
-\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.
-%% \footnote{For example, for string editing,
-%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
-%% and 0 otherwise.}
+ We first mathematically describe our 3-step algorithm as
+ an inference procedure for a hierarchical Bayesian model (Section~\ref{mathematicalFraming}),
+  and then describe each step algorithmically in detail (Section~\ref{explorationSection}-\ref{grammarInductionSection}).
+
+ \subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}
+
+\system takes as input a set of \textbf{tasks}, written $X$, each of which is a program synthesis problem.
+It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.\footnote{For example, for string editing,
+  the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
+  and 0 otherwise; when learning a generative model or probabilistic program, the likelihood is
+the probability of the program sampling the observation.}
 Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL, written $\mathcal{D}$,
-which is a generative model over programs, written $\probability[p |\mathcal{D}]$.
-We frame our goal as maximum a posteriori (MAP) inference of $\mathcal{D}$ given $X$:
-$$
-\argmax_{\mathcal{D}}\probability[\mathcal{D}]\prod_{x\in X}\sum_p\probability[x|p]\probability[p|\mathcal{D}]
-$$
-Even evaluating the above objective is intractable due to the sum over all programs,
-and so we instead marginalize over finite sets of programs, called \textbf{frontiers}, one for each task $x\in X$, written $\mathcal{F}_x$, giving the following lower bound on the marginal, which we call $\lowerBound$:
-\begin{equation}\label{lowerBound}
-  \lowerBound  \triangleq\probability[\mathcal{D}]\prod_{x\in X}\sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D}]
-\end{equation}
-We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ %% (\textbf{Wake})
-and $\mathcal{D}$:  %% (\textbf{Sleep-G})
-
-\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $\mathcal{D}$ is fixed and we
+and also to infer a DSL, written $\mathcal{D}$.
+We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
+$(\mathcal{D},\theta)$ define a generative model over programs.
+We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
+Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
+\begin{align}\label{intractableObjectives}
+\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta &\qquad
+  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
+\end{align}
+
+%% Inference in this model is
+%% difficult because the programs are unobserved,
+%% and so we must solve a hard search problem to recover them. To make
+%% search tractable we learn a bottom-up \emph{recognition
+%%   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
+%% The recognition model $q(\cdot )$ is a neural
+%% network that regresses from input/output pairs to a distribution over
+%% programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
+%% inference
+%% scheme~\cite{le2016inference}.
+%%  The neural recognition model and the
+%% generative model embodied in the DSL jointly train each other, as they
+%% iteratively learn to solve more programming tasks.
+
+The above equations summarize the problem from the point of view of an ideal Bayesian learner.
+However, Eq.~\ref{intractableObjectives}
+is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
+summing over the  infinite set of all programs.
+In practice we will only ever be able to sum over a finite set of programs.
+So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
+\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
+
+Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
+\begin{align}
+ J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
+\end{align}
+We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
+and $(\mathcal{D},\theta)$ (\textbf{Sleep-G}):
+
+\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D}]$
-is large.
-Here we use a simple and generic enumerative program synthesis algorithm.
+$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
+is large (i.e., programs with high posterior probability).
+%Here we use a simple and generic enumerative program synthesis algorithm.
 %% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
@@
 \\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\mathcal{D}$.
-To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.
+Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
+and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
+%To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.
 
@@
 during the \textbf{Sleep-R} phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$,
-  thus amortizing the cost of finding programs with high posterior probability.
+posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$.
+  Thus training the neural network amortizes the cost of finding programs with high posterior probability.
 
@@
 $q(p|x)$ to assign high probability to programs $p$ where
-$\probability[x|p]\probability[p|\mathcal{D}]$ is large, because including those programs
-in the frontiers will most increase $\lowerBound$.
-Concretely, the loss for $q(\cdot |x)$ is:
-$$
-\expect_X\left[ \text{KL}\left(\probability[\cdot |x,\mathcal{D}]\; ||\; q(\cdot |x) \right) \right]
-$$
-which minimizes the KL between the distribution predicted by $q$ and the true posterior over programs solving task $x$.
+$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
+into the frontiers will most increase $\lowerBound$.
+
+Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figure~\ref{feeding}).
+As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
+the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
+samples from the DSL and programs found during waking,
+the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
+\begin{figure}\centering
+  \begin{tikzpicture}
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
+    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
+
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+  \caption{\system solves for programs, the DSL, and a recognition model.
+Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
+\end{figure}
+\begin{figure}[h]
+\centering
+%  \includegraphics{figures/iterations.eps}
+\begin{tikzpicture}
+  \begin{scope}[shift={(2.5,0)}]
+    \node[obs] at (3,3) (dx){$\mathcal{D}$};
+    \node[latent] at (3.5,1.75) (zp){$p$};
+%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
+    \node[obs] at (3.5,0.7) (xp) {$x$};
+    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
+    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
+%    \draw [->,red] (tx) -- (zp);
+    \draw [->] (dx) -- (zp);
+    \node at (nn) {
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
+  \end{scope}
+
+  \begin{scope}[shift={(0.7,-4.2)}]
+    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+    \draw [->] (p2.south) -- (x2.north);
+    \draw [->] (dt.south) -- (p2.north);
+
+    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+    \draw [->] (p1.south) -- (x1.north);
+    \draw [->] (dt.south) -- (p1.north);
+
+    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+    \draw [->] (p3.south) -- (x3.north);
+    \draw [->] (dt.south) -- (p3.north);
+
+    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+    \draw [->] (zp.south) -- (xp.north);
+    \plate {}{(zp)(xp)}{$x\in X$};
+
+    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
+    rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
+    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
+  \end{scope}
+
+  \begin{scope}[shift={(7.7,-4)}]
+    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+    \node[latent] at (1.5,3) (t){$\theta$};
+    \node[obs] at (1,1.75) (z){$p$};
+%     \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+    \node[obs] at (1,0.5) (x) {$x$};
+    \edge {z}{x};
+    \edge {d,t}{z};
+    \plate {}{(z)(x)}{$x\in X$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
+  \end{scope}
+
+\end{tikzpicture}
+\end{figure}
+
+\subsection{Wake: Searching for Programs}\label{explorationSection}
+
+During waking, the agent's goal is to search for programs solving the tasks.  We use the simple approach of enumerating programs from
+the DSL  in decreasing order of their probability according to the recognition model,
+and then checking if a program $p$ assigns positive
+probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
+into the frontier $\mathcal{F}_x$.
+
+To make this concrete we need to define what programs actually are and
+what form $\probability[p |\mathcal{D},\theta]$ takes.
+We represent programs as $\lambda$-calculus expressions.
+$\lambda$-calculus is a formalism for expressing functional programs
+that closely resembles Lisp,
+including variables, function application, and the ability to create new functions.
+Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
+Our programs are all strongly typed.
+We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
+used in functional programming languages like OCaml and Haskell.
+%% Type variables are always written using lowercase Greek letters
+%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
+%% and returns something of type $\beta$.
+%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
+%% has the type $\tau$.
+%% For example, to describe the type of the identity function
+%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
+%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
+%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
+%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
+%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
+%% See Supplement for more detail on program representation.
+We now define DSLs:
+
+\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
+A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
+%\\\noindent\textbf{Definition.}
+A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
+one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
+and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
+
+Together with its weight vector,
+a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
+In the supplement, we define this distribution  
+by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$.
+Care must be taken to ensure that programs are well-typed
+ and that variable scoping rules are obeyed.
+%% With this distribution in hand, we search for programs by enumerating
+%% $\lambda$-calculus expressions in decreasing order of their probability under
+%% $(\mathcal{D},\theta)$.
+
+Why enumerate, when the program synthesis community has invented many
+sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
+We have two reasons:
+(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
+(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
+  the space of  programs.
+
+  However, a drawback of   enumerative search  is that we have no
+efficient means of solving for arbitrary constants that might occur in a
+program. In Sec.~\ref{regressionSection},
+we will show how to find programs with real-valued constants
+by automatically differentiating through the program and setting the constants using gradient descent.
+%% In Sec.~\ref{textSection}
+%% we will show that the bottom-up neural recognition model can learn
+%% which discrete constants should be included in a program.
+
+\subsection{Sleep-R: Training a Neural Recognition Model}\label{recognitionSection}
+
+The purpose of training the recognition model is to amortize the cost
+of searching for programs.  It does this by learning to predict, for
+each task, programs with high likelihood according to
+$\probability[x|p]$ while also being probable under the prior
+$(\mathcal{D},\theta)$.  Concretely, the recognition model $q$
+predicts, for each task $x\in X$, a weight vector $q(x) =
+\theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the
+DSL, this defines a distribution over programs,
+$\probability[p|\mathcal{D},\theta = q(x)]$.  We abbreviate this
+distribution as $q(p|x)$.  The crucial aspect of this framing is that
+the neural network leverages the structure of the learned DSL, so it
+is \emph{not} responsible for generating programs wholesale.  We share
+this aspect with DeepCoder~\cite{balog2016deepcoder}
+and~\cite{menon2013machine}.
+
+How should we get the data to train $q$?
+This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
+One approach is to sample programs from the DSL,
+run them to get their input/outputs,
+and then train $q$ to predict the program from the input/outputs.
+This is like how a Helmholtz machine
+trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
+The advantage of ``Helmholtz machine'' training is that
+we can draw unlimited samples from the DSL,
+training on a large amount of data.
+Another approach is
+ self-supervised learning,
+training $q$ on the (program, task)
+pairs discovered during waking.
+The advantage of self-supervised learning
+is that the training data is much higher quality,
+because we are training on the actual tasks.
+Due to these complementary advantages,
+we train on both these sources of data.
+
+Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
+ equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
+%% \begin{equation*}
+%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
+%% \end{equation*}
+ where the expectation is taken over tasks.
+Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
+ The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives:
+ \begin{align}
+   \mathcal{L}_{\text{RM}} &= \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}\\
+\nonumber\mathcal{L}_{\text{HM}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\\
+\nonumber\mathcal{L}_{\text{SS}} &= \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
+\end{align}
+
+  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
+all of our experiments train \system on only a few hundred tasks, which is too little for
+a high-capacity neural network $q$.
+Once we bootstrap a $(\mathcal{D},\theta)$,
+we can draw unlimited samples from $(\mathcal{D},\theta)$
+and train $q$ on those samples.
+
+Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
+the current DSL, running them to get their outputs,
+and then training $q$ to regress from the input/outputs to the program.
+Since these programs map inputs to outputs,
+we need to sample the inputs as well.
+Our solution is to sample the inputs
+from the empirical observed distribution of inputs in $X$.
+
+\subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
+
+The purpose of the DSL is to
+offer a set of abstractions
+that allow an agent to easily express solutions to the tasks at hand.
+%In the \system algorithm we infer the DSL from a collection of frontiers.
+Intuitively, we want the algorithm to
+look at  the frontiers and
+generalize beyond them, 
+both so the DSL can better express the current solutions,
+and  also so that the DSL might expose new abstractions
+which will later be used to
+discover more programs.
+ Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
+We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
+\begin{align}
+\nonumber      \log \probability[\mathcal{D}]& + 
+\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
+\label{AIC}
+  \end{align}
+We induce a DSL by searching locally through the space of DSLs,
+proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
+The search moves work by introducing new
+$\lambda$-expressions into the DSL.
+We propose these new expressions by extracting fragments of
+programs already in the frontiers (Tbl.~\ref{fragmentExample}).
+An important point here is that we are \emph{not} simply adding
+subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
+extracting fragments that unify with programs in the frontiers.  This
+idea of storing and reusing fragments of expressions comes from
+Fragment Grammars~\cite{tim} and Tree-Substitution
+Grammars~\cite{cohn2010inducing}, and is closely related to the idea
+of antiunification~\cite{henderson2013cumulative}.
+\begin{figure*}[!ht]\tabcolsep=2pt
+\centering  \begin{minipage}[c]{0.2\textwidth}
+  \begin{tikzpicture}[scale=0.7]
+    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
+    %%     \footnotesize Program:\\
+    %%     \code{($\lambda$ (x) (+ (- x) 1))}
+    %% \end{tabular}};
+    %\node at (,0.5) {\code{cons}};
+    \node [rotate=90] at (-2.3,-0.5) {\small program};
+
+          \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  \node(x1) at (0,-1) {\code{1}};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+  \node(t) at (-0.5,0.5) {};
+  \draw (l1.south) -- (t.south);
+  \node(c) at (-1.5,-0.2) {\code{cons}};
+  \draw (t.south) -- (c.north);
+
+%    \draw  (l1.south) -- (-0.5,0.5);
+
+  %% \node(c) at (-0.5,-1.5) {\code{-}};
+  %% \node(z) at (0.5,-1.5) {\code{x}};
+
+  %% \draw (0,-1) -- (c.north);
+  %% \draw (0,-1) -- (z.north);
+
+  \begin{scope}[shift={(0,-2.5)}]
+      \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  %\node(x1) at (0,-1) {};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+  \node(c) at (-0.5,-1.5) {\code{car}};
+  \node(z) at (0.5,-1.5) {\code{z}};
+
+  \draw (0,-1) -- (c.north);
+  \draw (0,-1) -- (z.north);
+
+  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+  \end{scope}
+
+\begin{scope}[shift={(0,-5)}]
+  \node[pop3](p1) at (-1,-1) {\code{+}};
+  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
+  \node[pop3](a) at (0,-1) {\code{ }};
+  %\node(x1) at (0,-1) {};
+  \draw[pop3] (0,0) -- (p1.north);
+  \draw[pop3] (0,0) -- (n1.north);
+  \draw[pop3] (-0.55,-0.4) -- (a.north);
+  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
+
+  \end{scope}
+
+\end{tikzpicture}
+  \end{minipage}
+\hspace{0.1cm}\begin{tabular}{ll}
+    \toprule
+    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
+    \begin{tabular}{l}
+      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
+      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
+      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
+      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
+    \end{tabular}&
+    \begin{tabular}{l}
+      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
+      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
+    \end{tabular}\\\midrule
+    \begin{tabular}{l}
+      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
+      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
+      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
+      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
+      \end{tabular}&
+    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
+      \end{tabular}
+\\\bottomrule\\
+\end{tabular}
+\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
+highlighted in {\orange{orange}},
+  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
+  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
+\end{figure*}
+
+To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+  and $\lambda$ controls how strongly we regularize the size of the DSL.
+  We place a symmetric Dirichlet prior over the weight vector $\theta$,
+  defining 
+$  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
+ $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
+%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
+
+To appropriately score each proposed $\mathcal{D}$ we must reestimate
+ the weight vector $\theta$.
+Although this  may seem 
+very similar to estimating the parameters of a probabilistic context free grammar,
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+our DSLs are context-sensitive due to the presence of variables
+in the programs and also due to the polymorphic typing system.
+In the Supplement we derive a tractable MAP estimator for $\theta$.
+%% Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+%% recognition model training, and DSL induction.
+
+\begin{comment}
+  \begin{wrapfigure}{R}{0.6\textwidth}
+  \begin{minipage}{0.6\textwidth}    
+    \begin{algorithm}[H]
+      \caption{The \system Algorithm}
+      \label{mainAlgorithm}
+      \begin{algorithmic}
+        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
+        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
+        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
+        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
+        \FOR{$i=1$ {\bfseries to} $I$}
+        %     \FOR{$x:\tau\in X$}
+        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
+        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
+        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
+        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
+        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
+        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
+        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}
+
+        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
+        %% \STATE Define $Q_x(z) \propto \begin{cases}
+        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
+        %%   0&x\not \in \mathcal{F}_x
+        %% \end{cases}
+        \ENDFOR
+        \STATE \textbf{return} $\mathcal{D},\theta,q$
+      \end{algorithmic}
+    \end{algorithm}
+  \end{minipage}
+\end{wrapfigure}
+  \end{comment}
 

### 2018-06-14 archive.tex
more writing for archive posting

@@
 
+\usepackage[section]{placeins}
 % hyperref makes hyperlinks in the resulting PDF.
@@
 \newcommand{\teal}[1]{\textcolor{teal}{#1}}
+
+\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
 
@@
 
-\begin{table*} %[t!]
+\begin{table*}%[t!]
   \makebox[\textwidth][c]{
@@
 of antiunification~\cite{henderson2013cumulative}.
-\begin{figure*}[!ht]\tabcolsep=2pt
+
+%\vfill\null
+
+%\columnbreak
+
+To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+  and $\lambda$ controls how strongly we regularize the size of the DSL.
+  We place a symmetric Dirichlet prior over the weight vector $\theta$,
+  defining 
+$  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
+ $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
+%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
+
+To appropriately score each proposed $\mathcal{D}$ we must reestimate
+ the weight vector $\theta$.
+Although this  may seem 
+very similar to estimating the parameters of a probabilistic context free grammar,
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+our DSLs are context-sensitive due to the presence of variables
+in the programs and also due to the polymorphic typing system.
+In the Supplement we derive a tractable MAP estimator for $\theta$.
+%% Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+%% recognition model training, and DSL induction.
+
+\begin{figure*}[!htb]\tabcolsep=2pt
 \centering  \begin{minipage}[c]{0.2\textwidth}
@@
 \end{figure*}
-
-To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-  and $\lambda$ controls how strongly we regularize the size of the DSL.
-  We place a symmetric Dirichlet prior over the weight vector $\theta$,
-  defining 
-$  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
- $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
-
-To appropriately score each proposed $\mathcal{D}$ we must reestimate
- the weight vector $\theta$.
-Although this  may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-our DSLs are context-sensitive due to the presence of variables
-in the programs and also due to the polymorphic typing system.
-In the Supplement we derive a tractable MAP estimator for $\theta$.
-%% Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-%% recognition model training, and DSL induction.
 
@@
     repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
-%    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-%    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
     keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
@@
 what kind of function best explains it.
-These programs can contain real numbers,
-which we set with gradient descent,
-and penalize continuous parameters by
-incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
-%% A key difficulty, however, is
-%% that these problems are best solved with programs containing real
-%% numbers.  Our solution to this difficulty is to enumerate
-%%  programs with real-valued parameters, and then fit those
-%% parameters by automatically differentiating through the programs the
-%% system writes and use gradient descent to fit the parameters.
-%% We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
-%% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
+A key difficulty, however, is
+that these problems are best solved with programs containing real
+numbers.  Our solution to this difficulty is to enumerate
+ programs with real-valued parameters, and then fit those
+parameters by automatically differentiating through the programs the
+system writes and use gradient descent to fit the parameters.
+We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
+and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
 \begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
@@
 which has five continuous parameters.
-This phenomenon arises from our Bayesian framing.
+This phenomenon arises from our Bayesian framing:
+both the generative model's bias toward shorter programs,
+and the likelihood model's BIC penalty.
 
@@
 fall far short of this dream,
-%and it is unclear (and we believe unlikely)
-and we believe it is unlikely
+and it is unclear (and we believe unlikely)
 that this dream could ever be fully realized.
-%How far can we push in this direction?
+How far can we push in this direction?
 ``Learning from scratch'' is subjective, but a reasonable
@@
 the 1959 Lisp primitives, and tasked with solving 22
-programming exercises.  After running for 93 hours on 64 CPUs, our
-algorithm solves these exercises, along the way assembling a DSL
+programming exercises. A key difference between
+this setup and our previous experiments is that,
+for this experiment,
+the system is
+given primitive recursion,
+whereas previously we had sequestered recursion within
+higher-order functions like \code{map}, \code{fold}, and \code{unfold}.
+
+After running for 93 hours on 64 CPUs, our
+algorithm solves these 22 exercises, along the way assembling a DSL
 with a modern repertoire of
@@
 and  arithmetic operations like 
-building lists of natural numbers between an interval.
-%% We did not use the recognition model for this experiment:
-%% a bottom-up pattern recognizer is of little use
-%% for acquiring this abstract knowledge
-%% from less than two dozen problems.
+building lists of natural numbers between an interval (see Figure~\ref{learningFromScratch}).
+We did not use the recognition model for this experiment:
+a bottom-up pattern recognizer is of little use
+for acquiring this abstract knowledge
+from less than two dozen problems.
+
+\begin{figure}  \newcommand{\helpSize}{0.25cm}
+  \begin{tabular}{cc}\toprule
+    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}\\\midrule
+    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+    \begin{tabular}{l}
+      \greenCode{$f_0($x$)\,=\,$(+ x 1)}\\
+      \hspace{\helpSize}($f_0$: \emph{increment})\\
+      \greenCode{$f_1($x$)\,=\,$(- x 1)}\\
+      \hspace{\helpSize}($f_1$: \emph{decrement})\\
+        \popp{\code{$f_2($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
+      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_2$ (n x))))}}\\
+        \hspace{\helpSize}($f_2$: \emph{unfold})\\
+        \popp{\code{$f_3($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
+      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{($f_3$ ($f_1$ i) (cdr l))))}}\\
+        \hspace{\helpSize}($f_3$: \emph{index})\\
+        \popp{\code{$f_4($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
+      \phantom{\code{$f_4($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_4$ (cdr l))))}}\\
+        \hspace{\helpSize}($f_4$: \emph{fold})\\
+
+      \end{tabular}
+    %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
+    %% (lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) ($1 ($3 $0)) ($4 $0) ($5 $0)))))))))
+    %% (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (#(lambda (lambda (fix1 $0 (lambda (lambda (if (empty?
+$0) empty (cons ($3 (car $0)) ($1 (cdr $0))))))))) (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
+$0 $3)))))))
+    \\\bottomrule 
+    \end{tabular}
+  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
+  \end{figure}
 

### 2018-06-15 archive.tex
more learning from scratch

@@
 \newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
+\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}
 
@@
 and then learning to solve a wide swath of
-induction problems~\cite{solomonoff1964formal}.
+induction problems~\cite{solomonoff1964formal,schmidhuber2004optimal,hutter2004universal,solomonoff1989system}.
 All existing systems,
@@
 \begin{figure}  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{cc}\toprule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}\\\midrule
+  \begin{tabular}{cl}\toprule
+    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
+    \begin{tabular}{l}
+      \code{[4\, 7\, 6\, 9\, 0]$\to$5}\\
+\code{[3\, 8\, 2\, 1\, 4]$\to$4}\\
+\code{[2\, 2\, 9\, 4\, 8]$\to$3}\\
+\code{[0\, 9\, 8\, 9\, 9\, 8]$\to$1}\\
+\blueCode{$f(\ell) = $($f_0$ (car $\ell$))}\\\\
+
+\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
+\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
+\blueCode{$f(\ell) = $($f_4$ + $\ell$ 0)}\\\\
+
+\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
+\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
+\blueCode{$f(\ell) = $($f_5$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
+
+      \code{[5\, 2\, 9]}$\to$\code{[[2\, 9]\, [9]\, []]}\\
+      \code{[3\, 8\, 1\, 3]}$\to$\code{[[8\, 1\, 3]\, [1\, 3]\, [3]\, []]}\\
+      \blueCode{$f(\ell) = $($f_2$ empty? cdr cdr $\ell$)}
+      \\
+
+      \end{tabular}
+    \\\midrule
     \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
@@
         \hspace{\helpSize}($f_4$: \emph{fold})\\
+        \greenCode{$f_5($f$,$l$)\,=\,$(if (empty? l) nil)}\\
+        \phantom{\greenCode{$f_5($f$,$l$)\,=\,$(if }}\greenCode{(cons (f (car l)) ($f_5$ (cdr l)))}\\
+        \hspace{\helpSize}($f_5$: \emph{map})
+
+%        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
+$0 $3)))))))
 

### 2018-07-03 nampi.tex
first poster commit

@@
 \icmlauthor{Lucas Morales}{MIT}
-\icmlauthor{Matthias}{MIT}
-\icmlauthor{Maxwell Nye}{MIT}
+\icmlauthor{Mathias Sabl\'e Meyer}{MIT}
+%\icmlauthor{Maxwell Nye}{MIT}
 \icmlauthor{Armando Solar-Lezama}{MIT}
@@
 
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
 %\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

### 2018-07-03 poster.tex
first poster commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Portrait Poster
% LaTeX Template
% Version 1.0 (31/08/2015)
% (Based on Version 1.0 (29/03/13) of the landscape template
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% Portrait version by:
% John Hammersley
%
% The landscape version of this template was downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{36in} % A0 width: 46.8in
\setlength{\paperheight}{54in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}

\title{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning} % Poster title

\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)

\institute{Massachusetts Institute of Technology} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Wake/Sleep DSL Induction}

Lorem ipsum dolor sit amet, consectetur, nunc tellus pulvinar tortor, commodo eleifend risus arcu sed odio:
\begin{itemize}
\item Mollis dignissim, magna augue tincidunt dolor, interdum vestibulum urna
\item Sed aliquet luctus lectus, eget aliquet leo ullamcorper consequat. Vivamus eros sem, iaculis ut euismod non, sollicitudin vel orci.
\item Nascetur ridiculus mus.  
\item Euismod non erat. Nam ultricies pellentesque nunc, ultrices volutpat nisl ultrices a.
\end{itemize}

\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a. \textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est. In ut orci purus, interdum ornare nibh. Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla. Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam. Maecenas sed diam augue.

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a.

This statement requires citation \cite{Smith:2012qr}.

\end{block}

%------------------------------------------------

\begin{figure}
\includegraphics[width=0.8\linewidth]{placeholder.jpg}
\caption{Figure caption}
\end{figure}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Materials}

The following materials were required to complete the research:

\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\item Eu facilisis est tempus quis
\end{itemize}

The materials were prepared according to the steps outlined below:

\begin{enumerate}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\item Curabitur pellentesque dignissim
\end{enumerate}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------

\begin{block}{Methods}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a. \textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est. In ut orci purus, interdum ornare nibh. Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla. Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam. Maecenas sed diam augue.

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

%----------------------------------------------------------------------------------------
%	IMPORTANT RESULT
%----------------------------------------------------------------------------------------

\begin{alertblock}{Important Result}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit.

\end{alertblock}

%----------------------------------------------------------------------------------------

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again

\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATHEMATICAL SECTION
%----------------------------------------------------------------------------------------

\begin{block}{Mathematical Section}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\begin{block}{Results}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}

\end{block}

%----------------------------------------------------------------------------------------
%	ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Additional Information}

\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\begin{block}{References}

\nocite{*} % Insert publications even if they are not cited in the poster
\small{\bibliographystyle{unsrt}
\bibliography{sample}\vspace{0.75in}}

\end{block}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color

\begin{block}{Acknowledgements}

\small{\rmfamily{Nam mollis tristique neque eu luctus. Suspendisse rutrum congue nisi sed convallis. Aenean id neque dolor. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\

\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

\begin{alertblock}{Contact Information}

\begin{itemize}
\item Web: \href{http://www.my.edu/smithlab}{http://www.my.edu/smithlab}
\item Email: \href{mailto:john@smith.com}{john@smith.com}
\item Phone: +1 (000) 111 1111
\end{itemize}

\end{alertblock}

\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
\end{tabular}
\end{center}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2018-07-03 short.tex
first poster commit

@@
 \icmlauthor{Lucas Morales}{MIT}
-\icmlauthor{Matthias}{MIT}
-\icmlauthor{Maxwell Nye}{MIT}
+\icmlauthor{Mathias Sabl\'e Meyer}{MIT}%\icmlauthor{Matthias}{MIT}
 \icmlauthor{Armando Solar-Lezama}{MIT}
@@
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}\vspace{-0.5cm}
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}%\vspace{-0.5cm}
 \end{table*}

### 2018-07-04 poster.tex
progress on poster

@@
 \usetheme{confposter} % Use the confposter theme supplied with this template 
-
+\usepackage{exscale}
 \setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
@@
 %-----------------------------------------------------------
+\usepackage{dsfont}
+\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
+\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
+
+\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
+\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em E}}
+\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
+\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}
+\newcommand{\lowerBound}{\mathscr{L}}
+\newcommand{\code}[1]{{\texttt{#1}}}
+\newcommand{\scode}[1]{{\texttt{#1}}}
+\newcommand{\codechar}[1]{{\texttt{"#1"}}}
+
+\usepackage{xcolor}
+\definecolor{pop1}{HTML}{1F78b4}
+\definecolor{pop2}{HTML}{164C13}
+\definecolor{pop3}{HTML}{d95F02}
+\definecolor{orange}{HTML}{d95F02}
+\definecolor{teal}{HTML}{1b9e77}
+\newcommand{\pop}[1]{\textcolor{pop1}{#1}}
+\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
+\newcommand{\tree}[1]{\textcolor{pop3}{#1}}
+\newcommand{\orange}[1]{\textcolor{orange}{#1}}
+\newcommand{\teal}[1]{\textcolor{teal}{#1}}
 
@@
 \usepackage{booktabs} % Top and bottom rules for tables
-
+\usepackage{tabularx}
 %----------------------------------------------------------------------------------------
@@
 
-\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
+\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Maxwell Nye, Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
 
@@
 \begin{alertblock}{Wake/Sleep DSL Induction}
+  \textbf{Domain Specific Language (DSL):}
+  A finely-tuned program representation, specialized to a domain of programming tasks.
+  Prior work in program learning  largely uses hand-engineered DSLs.
 
-Lorem ipsum dolor sit amet, consectetur, nunc tellus pulvinar tortor, commodo eleifend risus arcu sed odio:
+  \textbf{Contribution:} the \system algorithm,
+  which bootstraps a learned DSL while jointly training
+  a neural net to search for programs in the learned DSL
+
+  \textbf{Approach:} given a few hundred programming tasks, alternatingly:
   \begin{itemize}
-\item Mollis dignissim, magna augue tincidunt dolor, interdum vestibulum urna
-\item Sed aliquet luctus lectus, eget aliquet leo ullamcorper consequat. Vivamus eros sem, iaculis ut euismod non, sollicitudin vel orci.
-\item Nascetur ridiculus mus.  
-\item Euismod non erat. Nam ultricies pellentesque nunc, ultrices volutpat nisl ultrices a.
+  \item \textbf{Wake}: synthesize programs
+  \item \textbf{Sleep-R}: train neural net (\textbf{R}ecognition model)
+    \item \textbf{Sleep-G}: improve DSL (\textbf{G}enerative model)
     \end{itemize}
+
+  \textbf{Program representation:}
+  Lisp-like; conditionals, variables, local functions
 
@@
 
-\begin{block}{Introduction}
+\begin{block}{Bayesian framing}
 
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed commodo molestie porta.
-Sed ultrices scelerisque sapien ac commodo.
-Donec ut volutpat elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
-\textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est.
-In ut orci purus, interdum ornare nibh.
-Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla.
-Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam.
-Maecenas sed diam augue.
+  Observe $N$ \textbf{tasks}, written $\left\{x_n \right\}_{n = 1}^N$, each a program synthesis problem.
 
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed commodo molestie porta.
-Sed ultrices scelerisque sapien ac commodo.
-Donec ut volutpat elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
+  Solve task $x_n$ with latent program $p_n$
 
-This statement requires citation \cite{Smith:2012qr}.
+  \textbf{Likelihood model} $\probability[x_n|p_n]$
+  scores program $p_n$ on task $x_n$
+
+  Latent  \textbf{DSL $\mathcal{D}$} acts as generative model over programs: $\probability[x|\mathcal{D}]$
+
+$$
+\underbrace{p_n^* =   \argmax_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}^*]}_{\text{\textbf{Wake}}}
+$$
+\begin{equation*}
+\underbrace{\mathcal{D}^* = \argmax_{\mathcal{D}}    \probability[\mathcal{D}]\prod_n\sum_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}]}_{\textbf{\text{Sleep-G}}}
+  \end{equation*}
+
+\end{block}
+
+\begin{block}{Neural recognition model}
+
+  Neural network $Q(p|x)$
+  predicts distribution over programs conditioned on tasks.
+  Simple $Q$: just predicts probabilities of DSL productions.  Goal: learn to invert generative model
+$$\underbrace{\min_Q \text{KL}\left(\probability[p|x,\mathcal{D}] || Q(p|x)\right) }_{\text{\textbf{Sleep-R}}}$$
+
+Train on two sources of data:
+
+  \begin{itemize}
+  \item \textbf{Samples (``Dreams'') from DSL}: Unlimited data, but only high-quality if generative model $\mathcal{D}$ is good. Like Helmholtz Machine's recognition model training. Loss:
+    \begin{equation*}
+\expect_{(p,x)\sim\mathcal{D} }\left[\log Q(p|x)\right]
+      \end{equation*}
+  \item \textbf{Self-Supervised}: $(x_n,p_n)$ pairs discovered during waking. Loss:
+    \begin{equation*}
+ %\sum_{n = 1}^N\left[\sum_{(x_n,p_n)}
+  \frac{\probability\left[x_n,p_n|\mathcal{D}\right]}{\sum_{(x_n,p_n')}\probability\left[x_n,p_n|\mathcal{D} \right]}\log Q(p_n|x_n)%\right]
+      \end{equation*}
+  \end{itemize}
 
@@
 %------------------------------------------------
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{placeholder.jpg}
-\caption{Figure caption}
-\end{figure}
 
@@
 
-\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)
+\begin{column}{\threecolwid} % Begin a column which is two columns wide (column 2)
 
-\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column
+\begin{columns}[t,totalwidth=\threecolwid] % Split up the two columns wide column
 
-\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
+\begin{column}{\threecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
 
@@
 
-\begin{block}{Materials}
+\begin{block}{Model outputs for three different task domains}
 
-The following materials were required to complete the research:
+\begin{table*} %[t!]
+  \makebox[\textwidth][c]{
+%    \scriptsize
+  \tabcolsep=4pt
+  %% \renewcommand\code\texttt
+  %% \renewcommand\codechar[1]{\texttt{"#1"}}
+  \newcommand{\helpSize}{0.25cm}
+  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{1em}}>{\hspace{1em}}c<{\hspace{1em}}>{\hspace{-1em}}c<{\hspace{1em}}>{\hspace{0em}}c<{\hspace{-1em}}}
+    \toprule
+    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
+    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
+      \begin{tabular}{cc}
+        \begin{tabular}{c}
+          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
+          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
+          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
+          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
+          \\
+          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
+          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
+          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
+        \end{tabular}
+        &
+        \hspace{-0.3cm}\begin{tabular}{c}
+          \code{[7\, 3]}$\to $\code{False}                              \\
+          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
+          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
+          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
+          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
+          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
+        \end{tabular}
+      \end{tabular}
+    }
+    &
+    \hspace{-0.5cm}\begin{tabular}{c}
+      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
+      +83 973-831$\to $83.973.831\\
+      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
+      }}\\
+      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
+      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
+      ~\\
+      Temple Anna H $\to $TAH\\
+      Lara Gregori$\to $LG\\
+      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
+    \end{tabular}
+    &
+    \begin{tabular}{ccc}
+      \includegraphics[width = 3em]{figures/functions/4.png}&&
+      \includegraphics[width = 3em]{figures/functions/146}\\
+      \pop{\code{$f($x$) = $($f_1$ x)}}& &   \pop{\code{$f($x$) = $($f_6$ x)}}\\
+      ~\\
+      \includegraphics[width = 3em]{figures/functions/112.png}&&
+        \includegraphics[width = 3em]{figures/functions/92.png}
+      \\
+      \pop{\code{$f($x$) = $($f_4$ x)}}& &   \pop{\code{$f($x$) = $($f_3$ x)}}\\
 
-\begin{itemize}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\item Eu facilisis est tempus quis
-\end{itemize}
+    \end{tabular}
+    ~\\
+    \midrule
+    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+    \hspace{0cm}\begin{tabular}{l}
+      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
+      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
+      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
+      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
+      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
+      %NOTE: this is the actual invention, but I removed a redundant lambda
+      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
+      %% \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
+      %% \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
+      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
+      \hspace{4cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
+      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\\\
+      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
+      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
+      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
+      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\\\
+      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
+      \phantom{$f_1(\ell,$}
+      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
+      \hspace{\helpSize}($f_3$: \emph{Whether $\ell$ contains }\code{k})
+    \end{tabular}&
 
-The materials were prepared according to the steps outlined below:
+  \hspace{0.5cm}\begin{tabular}{l}
+    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
+    (x)}}\\
+    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
+      \hspace{\helpSize}($f_0$: \emph{substitutes characters)}\\\\
+      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
+      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
+      \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\\
+      \hspace{\helpSize}\phantom{($f_1$: }\emph{ until  }\code{c}\emph{ reached})\\\\
+      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
+      }}\\
+      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
+        \hspace{\helpSize}($f_2$: \emph{Abbreviates  words})
+      %%   \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
+      %% \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
+  \end{tabular}&
 
-\begin{enumerate}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\item Curabitur pellentesque dignissim
-\end{enumerate}
+  \begin{tabular}{l}
+    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
+    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
+    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
+    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
+    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
+    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\\\
+    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
+    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
+    \hspace{\helpSize}\emph{($f_6$: rational function)}\\
 
+  \end{tabular}
+  \\\bottomrule\\
+\end{tabular}}\vspace{-0.5cm}
+%}\label{initialExampleDSL}%\vspace{-0.5cm}
+\end{table*}
+Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
+Bottom: Several examples from learned DSL.
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)
 \end{block}
@@
 \end{column} % End of column 2.1
-
-\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)
-
-%----------------------------------------------------------------------------------------
-%	METHODS
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Methods}
-
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
-\textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est.
-In ut orci purus, interdum ornare nibh.
-Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla.
-Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam.
-Maecenas sed diam augue.
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
-
-\end{column} % End of column 2.2
 
@@
 
-\begin{alertblock}{Important Result}
-
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit.
-
-\end{alertblock}
-
 %----------------------------------------------------------------------------------------
 
-\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again
+\begin{columns}[t,totalwidth=\threecolwid] % Split up the two columns wide column again
 
-\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
+\begin{column}{\twocolwid} % The first column within column 2 (column 2.1)
 
@@
 
-\begin{block}{Mathematical Section}
+\begin{block}{Fragment Grammars: Inducing a DSL}
+Fragment grammars: first introduced in computational linguistics~\cite{tim}
+\begin{figure}[b!]\tabcolsep=2pt
+\centering  \begin{minipage}[c]{7.5cm}
+  \begin{tikzpicture}[scale=2]
+    \node [rotate=90] at (-2.5,-0.5) {prog.};
+
+          \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  \node(x1) at (0,-1) {\code{1}};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+  \node(t) at (-0.5,0.5) {};
+  \draw (l1.south) -- (t.south);
+  \node(c) at (-1.5,-0.2) {\code{cons}};
+  \draw (t.south) -- (c.north);
+
+%    \draw  (l1.south) -- (-0.5,0.5);
+
+  %% \node(c) at (-0.5,-1.5) {\code{-}};
+  %% \node(z) at (0.5,-1.5) {\code{x}};
+
+  %% \draw (0,-1) -- (c.north);
+  %% \draw (0,-1) -- (z.north);
+
+  \begin{scope}[shift={(0,-2.5)}]
+      \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  %\node(x1) at (0,-1) {};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+  \node(c) at (-0.5,-1.5) {\code{car}};
+  \node(z) at (0.5,-1.5) {\code{z}};
+
+  \draw (0,-1) -- (c.north);
+  \draw (0,-1) -- (z.north);
+
+  \node [rotate=90] at (-2.3,-0.7) {prog.};
+
+  \end{scope}
+
+\begin{scope}[shift={(0,-5)}]
+  \node[pop3](p1) at (-1,-1) {\code{+}};
+  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
+  \node[pop3](a) at (0,-1) {\code{ }};
+  %\node(x1) at (0,-1) {};
+  \draw[pop3] (0,0) -- (p1.north);
+  \draw[pop3] (0,0) -- (n1.north);
+  \draw[pop3] (-0.55,-0.4) -- (a.north);
+  \node [rotate=90] at (-2.3,-0.7) {frag.};
+
+  \end{scope}
+
+\end{tikzpicture}
+  \end{minipage}
+\hspace{0.1cm}\begin{tabular}{ll}
+    \toprule
+    Example synthesized programs &Proposed $\lambda$-expression\\\midrule
+    %% \begin{tabular}{l}
+    %%   \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
+    %%   \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
+    %%   \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
+    %%   \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
+    %% \end{tabular}&
+    %% \begin{tabular}{l}
+    %%   \code{(map ($\lambda$ (x) (index x $\ell$))}\\
+    %%   \phantom{\code{(map)}}\code{(range $\alpha$))}\\
+    %% \end{tabular}\\\midrule
+    \begin{tabular}{l}
+      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
+      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
+      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
+      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
+      \end{tabular}&
+    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
+      \end{tabular}
+\\\bottomrule\\
+\end{tabular}
+\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
+highlighted in {\orange{orange}},
+  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
+  from which we extract fragments that perform character substitutions.}\label{fragmentExample}
+\end{figure}
 
@@
 \begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
-
-%----------------------------------------------------------------------------------------
-%	RESULTS
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Results}
+\begin{block}{Ongoing work: Generative models for text}
+\begin{figure}
+\centering
+\tiny
+\tabcolsep=2pt
+\vspace{0pt}
+\begin{tabular}{cc}
+\begin{tabular}{ccc}
+\multicolumn{3}{c}{Tasks:}\\
+\hline
+cut&F &Moss Side \\ 
+control&CL &Burnage \\
+control&F & City Centre\\
+cut& PCFL& Brooklands \\
+\multicolumn{3}{c}{Learned generative models:}\\
+\hline
+\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash
+u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash
+u\textbackslash w*)*} \\
+\multicolumn{3}{c}{Samples from synthesized generative models:}\\
+\hline
+ya&DQDF & Vr DR \\ 
+glrwfdcnc&F &BeF lKQ \\
+mgs&F& W\\
+piljnl& KI& kqBfZ 0 \\
+kj& F& ON\\
+zci& GL& Bttc \\
+sxpm& F& S \\
+\hline
+\\
+\end{tabular} 
+&
+\begin{tabular}{c}
+\begin{tabular}{l}
+\multicolumn{1}{c}{Learned DSL:}\\
+\hline
+$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
+$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash
+u\textbackslash w*)*}\\
+$f_3(x) = f_2(\scode{space}) = $\scode{( \textbar\textbackslash
+u\textbackslash w*)*}\\
+\hline
+$f_4(x) = $ \scode{($x$*$x$)}\\
+\emph{ (equivalent to regex `plus')}\\
+$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash
+l*\textbackslash l}$\\
+\hline
+\end{tabular}
+\\
+\\
+\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
+\end{tabular}
+\end{tabular}
+\vspace{-.3cm}
+\caption{regex stuff}
+\end{figure}
 
@@
 
-%----------------------------------------------------------------------------------------
+\begin{block}{Learning from Scratch?}
 
-\end{column} % End of column 2.2
+  \end{block}
+\end{column}
 
@@
 
-\begin{column}{\onecolwid} % The third column
+%% \begin{column}{\onecolwid} % The third column
 
-%----------------------------------------------------------------------------------------
-%	CONCLUSION
-%----------------------------------------------------------------------------------------
+%% %----------------------------------------------------------------------------------------
+%% %	CONCLUSION
+%% %----------------------------------------------------------------------------------------
 
-\begin{block}{Conclusion}
+%% \begin{block}{Conclusion}
 
-\end{block}
+%% \end{block}
 
-%----------------------------------------------------------------------------------------
-%	ADDITIONAL INFORMATION
-%----------------------------------------------------------------------------------------
+%% %----------------------------------------------------------------------------------------
+%% %	ADDITIONAL INFORMATION
+%% %----------------------------------------------------------------------------------------
 
-\begin{block}{Additional Information}
+%% \begin{block}{Additional Information}
 
-\end{block}
+%% \end{block}
 
-%----------------------------------------------------------------------------------------
-%	REFERENCES
-%----------------------------------------------------------------------------------------
+%% %----------------------------------------------------------------------------------------
+%% %	REFERENCES
+%% %----------------------------------------------------------------------------------------
 
-\begin{block}{References}
+%% \begin{block}{References}
 
-\nocite{*} % Insert publications even if they are not cited in the poster
-\small{\bibliographystyle{unsrt}
-\bibliography{sample}\vspace{0.75in}}
+%% \nocite{*} % Insert publications even if they are not cited in the poster
+%% \small{\bibliographystyle{unsrt}
+%% \bibliography{sample}\vspace{0.75in}}
 
-\end{block}
+%% \end{block}
 
-%----------------------------------------------------------------------------------------
-%	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
+%% %----------------------------------------------------------------------------------------
+%% %	ACKNOWLEDGEMENTS
+%% %----------------------------------------------------------------------------------------
 
-\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
+%% \setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
 
-\begin{block}{Acknowledgements}
+%% \begin{block}{Acknowledgements}
 
-\small{\rmfamily{Nam mollis tristique neque eu luctus.
+%% \small{\rmfamily{Nam mollis tristique neque eu luctus.
 Suspendisse rutrum congue nisi sed convallis.
@@
 
-\end{block}
+%% \end{block}
 
-%----------------------------------------------------------------------------------------
-%	CONTACT INFORMATION
-%----------------------------------------------------------------------------------------
+%% %----------------------------------------------------------------------------------------
+%% %	CONTACT INFORMATION
+%% %----------------------------------------------------------------------------------------
 
-\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
-\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors
+%% \setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
+%% \setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors
 
-\begin{alertblock}{Contact Information}
+%% \begin{alertblock}{Contact Information}
 
-\begin{itemize}
-\item Web: \href{http://www.my.edu/smithlab}{http://www.my.edu/smithlab}
-\item Email: \href{mailto:john@smith.com}{john@smith.com}
-\item Phone: +1 (000) 111 1111
-\end{itemize}
+%% \begin{itemize}
+%% \item Web: \href{http://www.my.edu/smithlab}{http://www.my.edu/smithlab}
+%% \item Email: \href{mailto:john@smith.com}{john@smith.com}
+%% \item Phone: +1 (000) 111 1111
+%% \end{itemize}
 
-\end{alertblock}
+%% \end{alertblock}
 
-\begin{center}
-\begin{tabular}{ccc}
-\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
-\end{tabular}
-\end{center}
+%% \begin{center}
+%% \begin{tabular}{ccc}
+%% \includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
+%% \end{tabular}
+%% \end{center}
 
-%----------------------------------------------------------------------------------------
+%% %----------------------------------------------------------------------------------------
 
-\end{column} % End of the third column
+%% \end{column}
+% End of the third column
 

### 2018-07-04 poster.tex
learning from scratch
poster stuff

@@
 \usepackage{graphicx}  % Required for including images
+
+\usepackage{tikz}
+\usetikzlibrary{arrows,shapes}
 
@@
 \begin{block}{Fragment Grammars: Inducing a DSL}
-Fragment grammars: first introduced in computational linguistics~\cite{tim}
+Fragment grammars: introduced in computational linguistics~(O'Donnell~2015)
 \begin{figure}[b!]\tabcolsep=2pt
@@
 
-%----------------------------------------------------------------------------------------
+\begin{block}{Ongoing work: Generative models}
+  Learn probabilistic program (a regex) $p_n$ from $K$ strings $x_n = \left\{y_n^k \right\}_{k = 1}^k$.
 
-\end{column} % End of column 2.1
-
-\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
-\begin{block}{Ongoing work: Generative models for text}
+  Likelihood model:
+  $$\probability [x_n|p_n] = \prod_{k = 1}^K \probability[y_n^k|p_n]$$
 \begin{figure}
 \centering
-\tiny
+%\tiny
 \tabcolsep=2pt
@@
 \vspace{-.3cm}
-\caption{regex stuff}
+%\caption{regex stuff}
 \end{figure}
@@
 
-\begin{block}{Learning from Scratch?}
+%----------------------------------------------------------------------------------------
+
+\end{column} % End of column 2.1
+
+\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
+
+\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
+\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}
+
+\begin{block}{Why this works: Bootstrapping}
+\begin{figure}\centering
+  \begin{tikzpicture}[scale=3]
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
+    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
+
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+ % \caption{\system solves for programs, the DSL, and a recognition model.
+Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
+\end{figure}
 
 \end{block}
+
+\begin{block}{Learning from Scratch}
+  Start w/ McCarthy 1959 Lisp: recursion, conditionals, lists.
+  Train on 22 programming exercises.
+  After 93 hours on 64 CPUs, rediscovers 9 functional programming staples: \code{map},
+  \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length}, \code{range}, \code{incr}, \code{decr}.
+
+\end{block}
+
+%% \begin{block}{References}
+
+%% \nocite{*} % Insert publications even if they are not cited in the poster
+%% \small{\bibliographystyle{unsrt}
+%% \bibliography{sample}\vspace{0.75in}}
+
+%% \end{block}
+
+%% %----------------------------------------------------------------------------------------
+%% %	ACKNOWLEDGEMENTS
+%% %----------------------------------------------------------------------------------------
+
+%% \setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
+
+%% \begin{block}{Acknowledgements}
+
+%% \small{\rmfamily{Nam mollis tristique neque eu luctus.
+Suspendisse rutrum congue nisi sed convallis.
+Aenean id neque dolor.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\
+
+%% \end{block}
+
 \end{column}

### 2018-07-05 archive.tex
algorithm flowchart

@@
 %\usetikzlibrary{decorations.text}
-%\usetikzlibrary{decorations.pathmorphing}
+\usetikzlibrary{decorations.pathmorphing}
+\tikzset{squiggle/.style={decorate, decoration=snake}}
 \usepackage{dsfont}
@@
 \end{figure}
+
+\begin{figure}
+  \begin{tikzpicture}
+    \node at (0,0) (d){DSL};
+    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+
+    \node at ([xshift = 2cm]t) (nn){
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
+    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.5cm,yshift = -1cm]) (s){Search};
+
+    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+    \node at (7,-1) (f) {Frontier};
+    \draw [->  ] (s.east) -- (f.west);
+
+    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
+
+    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
+
+  \end{tikzpicture}
+
+  \vspace{1cm}
+
+  \begin{tikzpicture}
+    \node at (0,0) (d){DSL};
+    \node at ([xshift = 3cm]d.east) (p2){program};
+    \node at ([yshift = 1.5cm]p2) (p1){program};
+    \node at ([yshift = -1.5cm]p2) (p3){program};
+
+    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
+    \draw[squiggle,-> ] (d.east) -- (p1.west);
+    \draw[squiggle,-> ] (d.east) -- (p3.west);
+
+    \node at ([xshift = 2cm]p1.east) (t1){task};
+    \node at ([xshift = 2cm]p2.east) (t2){task};
+    \node at ([xshift = 2cm]p3.east) (t3){task};
+    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
+    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
+    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
+
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
+  \end{tikzpicture}
+
+  \begin{tikzpicture}
+    \node at (0,0) (f1){Frontier$_1$};
+    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
+
+    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
+    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
+    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
+
+    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
+    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
+    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
+
+    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
+    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
+    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
+
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
+  \end{tikzpicture}
+  \vspace{2cm}
+
+  \begin{tikzpicture}
+    \node at (0,0) (f1){Frontier$_1$};
+    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+    \node at ([yshift = -1cm]f2.south) (ff){$\vdots$};
+    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
+
+    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
+    \node[anchor=north] at (c.north) {Compression};
+
+    \draw [-> ] (f1.east) -- (c.west|-f1.east);
+    \draw [-> ] (f2.east) -- (c.west|-f2.east);
+    \draw [-> ] (f3.east) -- (c.west|-f3.east);
+
+    \node[align = right](d) at ([xshift = 1.5cm,yshift = 1cm]c.east) {DSL $\mathcal{D}$};
+    \node[align = right](t) at ([xshift = 1.5cm,yshift = -1cm]c.east) {Parameters $\theta$};
+    \draw [-> ] (c.east) -- (d.west);
+    \draw [-> ] (c.east) -- (t.west);
+
+    \node at (c.center) {
+\begin{tikzpicture}[scale=0.7]
+    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
+    %%     \footnotesize Program:\\
+    %%     \code{($\lambda$ (x) (+ (- x) 1))}
+    %% \end{tabular}};
+    %\node at (,0.5) {\code{cons}};
+%    \node [rotate=90] at (-2.3,-0.5) {\small program};
+
+          \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  \node(x1) at (0,-1) {\code{1}};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+  \node(t) at (-0.5,0.5) {};
+  \draw (l1.south) -- (t.south);
+  \node(c) at (-1.5,-0.2) {\code{cons}};
+  \draw (t.south) -- (c.north);
+
+%    \draw  (l1.south) -- (-0.5,0.5);
+
+  %% \node(c) at (-0.5,-1.5) {\code{-}};
+  %% \node(z) at (0.5,-1.5) {\code{x}};
+
+  %% \draw (0,-1) -- (c.north);
+  %% \draw (0,-1) -- (z.north);
+
+  \begin{scope}[shift={(-1,-2.5)}]
+      \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  %\node(x1) at (0,-1) {};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+  \node(c) at (-0.5,-1.5) {\code{car}};
+  \node(z) at (0.5,-1.5) {\code{z}};
+
+  \draw (0,-1) -- (c.north);
+  \draw (0,-1) -- (z.north);
+
+%  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+  \end{scope}
+
+\begin{scope}[shift={(0,-5)}]
+  \node[pop3](p1) at (-1,-1) {\code{+}};
+  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
+  \node[pop3](a) at (0,-1) {\code{ }};
+  %\node(x1) at (0,-1) {};
+  \draw[pop3] (0,0) -- (p1.north);
+  \draw[pop3] (0,0) -- (n1.north);
+  \draw[pop3] (-0.55,-0.4) -- (a.north);
+%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
+
+  \end{scope}
+
+\end{tikzpicture}
+    };
+
+    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};
+
+    \end{tikzpicture}
+  \end{figure}
+
 \begin{figure}[h]

### 2018-07-05 poster.tex
algorithm flowchart

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,shapes}
-
+\usetikzlibrary{arrows.meta}
+%\usetikzlibrary{arrows,shapes}
+\usetikzlibrary{fit,bayesnet}
 \usepackage{booktabs} % Top and bottom rules for tables
@@
 
-  \textbf{Contribution:} the \system algorithm,
+  \textbf{Approach:}  \system algorithm,
   which bootstraps a learned DSL while jointly training
-  a neural net to search for programs in the learned DSL
-
-  \textbf{Approach:} given a few hundred programming tasks, alternatingly:
+  a neural net to search for programs in the learned DSL.
+Given a few hundred programming tasks, alternatingly:
   \begin{itemize}
@@
   \textbf{Program representation:}
-  Lisp-like; conditionals, variables, local functions
+  $\approx $Lisp; conditionals, variables, $\lambda$~abstraction
 
@@
 \begin{block}{Bayesian framing}
+
+\begin{tikzpicture}[scale=4,line width=1.5mm]
+
+  \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
+  \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
+  \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
+  \plate {}{(zp)(xp)}{};
+%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
+  \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
+%  \draw [->,red] (tx.west) -- (zp.east);
+  \draw [->] (dx.east) -- (zp.west);
+  \draw [->] (zp.east) -- (xp.west);
+
+  \node at (nn) {
+    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}
+  };
+  \node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+
+  \end{tikzpicture}
 
@@
 \centering  \begin{minipage}[c]{7.5cm}
-  \begin{tikzpicture}[scale=2]
+  \begin{tikzpicture}[scale=2,line width=1mm]
     \node [rotate=90] at (-2.5,-0.5) {prog.};
@@
 \begin{block}{Why this works: Bootstrapping}
-\begin{figure}\centering
-  \begin{tikzpicture}[scale=3]
+
+\begin{figure}
+\hspace*{-1.5cm}  \begin{tikzpicture}[scale=3,line width=1.5mm]
     \begin{scope}[shift = {(1,-1)}]
@@
 
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
 
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
   \end{scope}
@@
 \end{figure}
+\begin{itemize}
+\item Search finds new programs$\implies$ DSL+recognition model get more data
+\item DSL improves$\implies$easier search, recognition model gets better data
+  \item Recognition model improves$\implies$ easier search
+  \end{itemize}
 

### 2018-07-07 archive.tex
poster acknowledgments

@@
 \usetikzlibrary{decorations.pathmorphing}
-\tikzset{squiggle/.style={decorate, decoration=snake}}
+\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
 \usepackage{dsfont}
@@
 
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.5cm,yshift = -1cm]) (s){Search};
+    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
 
@@
 
-    \node at (7,-1) (f) {Frontier};
+    \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
     \draw [->  ] (s.east) -- (f.west);
@@
   \begin{tikzpicture}
-    \node at (0,0) (d){DSL};
+    \node at (0,0) (f1){Frontier$_1$};
+    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
+
+    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
+    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
+    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
+
+    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
+    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
+    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
+
+    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
+    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
+    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
+
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
+  \end{tikzpicture}
+
+  \begin{tikzpicture}
+    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
     \node at ([xshift = 3cm]d.east) (p2){program};
@@
 
+  \vspace{2cm}
+
   \begin{tikzpicture}
@@
     \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
-
-    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
-    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
-    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
-
-    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
-    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
-    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
-
-    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
-    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
-    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
-  \end{tikzpicture}
-  \vspace{2cm}
-
-  \begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -1cm]f2.south) (ff){$\vdots$};
+    \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
+    \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
     \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
@@
 
-    \node[align = right](d) at ([xshift = 1.5cm,yshift = 1cm]c.east) {DSL $\mathcal{D}$};
-    \node[align = right](t) at ([xshift = 1.5cm,yshift = -1cm]c.east) {Parameters $\theta$};
+    \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
+    \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
     \draw [-> ] (c.east) -- (d.west);

### 2018-07-07 nips_2018.tex
first poster commit
poster acknowledgments

@@
 
-\parbox{0.58\textwidth}{
+%\parbox{0.58\textwidth}{
 
@@
 domains over the generic methods used here.
-We are in the process of applying our algorithm to generative programs and
-prototyped this with a turtle-like domain: Figure~\ref{geomCompiled} gives some
-preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
-details.  Another direction is to explore DSL meta-learning: can we find a
+%% We are in the process of applying our algorithm to generative programs and
+%% prototyped this with a turtle-like domain: Figure~\ref{geomCompiled} gives some
+%% preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
+%% details.
+Another direction is to explore DSL meta-learning: can we find a
 \emph{single} universal primitive set that could effectively bootstrap DSLs for
 new domains, including the three domains considered,  but also many others?
-}\hfill
-\begin{minipage}{0.4\textwidth}
-  \centering
-  \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
-  \captionof{figure}{
-  Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
-  before any training (left) and after last iteration (right).}
-\label{geomCompiled}
-\end{minipage}
+%% }\hfill
+%% \begin{minipage}{0.4\textwidth}
+%%   \centering
+%%   \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
+%%   \captionof{figure}{
+%%   Top left: Example training tasks. Top right: samples from the learned DSL.
+%%  Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
+%%   before any training (left) and after last iteration (right).}
+%% \label{geomCompiled}
+%% \end{minipage}
 

### 2018-07-07 poster.tex
poster acknowledgments
single character

@@
 \begin{block}{Ongoing work: Generative models}
-  Learn probabilistic program (a regex) $p_n$ from $K$ strings $x_n = \left\{y_n^k \right\}_{k = 1}^k$.
+  Learn probabilistic program (a regex) $p_n$ from $K$ strings $x_n = \left\{y_n^k \right\}_{k = 1}^K$.
 
@@
 
-%% \setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
+\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
 
-%% \begin{block}{Acknowledgements}
+\begin{block}{Acknowledgements}
 
-%% \small{\rmfamily{Nam mollis tristique neque eu luctus.
-Suspendisse rutrum congue nisi sed convallis.
-Aenean id neque dolor.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\
+  We gratefully acknowledge collaboration with Eyal Dechter, whose EC algorithm (Dechter et al, IJCAI 2013) provided the inspiration for DreamCoder, and Luke Hewitt, who graciously provided us with
+  a regex learning data set.
 
-%% \end{block}
+\end{block}
 

### 2018-07-09 poster.tex
affiliation
turtle graphics

@@
 
-\title{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning} % Poster title
+\title{\huge \systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning} % Poster title
 
@@
 
-\institute{Massachusetts Institute of Technology} % Institution(s)
+\institute{Massachusetts Institute of Technology \& MIT-IBM Watson AI lab} % Institution(s)
 
@@
 
-\begin{block}{Ongoing work: Generative models}
+\begin{block}{Turtle/LOGO Graphics}
+
+  \newcommand{\dreamSize}{5.5cm}
+  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{9 (of 132) Tasks}}\\\hline
+    \includegraphics[width = \dreamSize]{turtleTasks/square/trimmed/train_square_3_h.png}&
+    \includegraphics[width = \dreamSize]{turtleTasks/square/trimmed/train_square_4_h.png}&
+    \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/snow1_h.png}\\
+    \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/snow2_random_1.png}&
+    \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/hexa_vh.png}&
+        \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/star_5_h.png}\\
+%    \includegraphics[width = \dreamSize]{turtleTasks/circle/trimmed/train_circle_10_h.png}\\
+    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/smooth_spiral4_h.png}&
+    \includegraphics[width = \dreamSize]{turtleTasks/los/trimmed/train_los_3_1_h.png}&
+    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/iter2_h.png}\\\hline
+%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/1748505267_norm.png}&
+  \end{tabular}\hspace{4cm}%
+  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{9 samples from learned DSL}}\\\hline
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/4361002260.png}&
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/5809723215.png}&
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/5818934795_norm.png}\\
+%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/6027771181.png}\\
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/1950268419.png}&
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/6770710484.png}&
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/7045328293.png}\\
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/7623700519.png}&
+%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/8879622788.png}&
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/9150025422.png}&
+    \includegraphics[width = \dreamSize]{dreams/final/trimmed/9666602433.png}\\\hline
+%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/1748505267_norm.png}&
+    \end{tabular}
   Learn probabilistic program (a regex) $p_n$ from $K$ strings $x_n = \left\{y_n^k \right\}_{k = 1}^K$.

### 2018-07-10 poster.tex
final logo figured
moved around sections and poster

@@
 
-\begin{block}{Turtle/LOGO Graphics}
+\begin{block}{Generative models of images: Turtle/LOGO Graphics}
 
   \newcommand{\dreamSize}{5.5cm}
-  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{9 (of 132) Tasks}}\\\hline
+  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{12 (of 132) Tasks}}\\\hline
     \includegraphics[width = \dreamSize]{turtleTasks/square/trimmed/train_square_3_h.png}&
@@
     \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/star_5_h.png}\\
+        \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/star_vh.png}&
+    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/spiral6_h.png}&
+        \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/star_7_h.png}\\
+
 %    \includegraphics[width = \dreamSize]{turtleTasks/circle/trimmed/train_circle_10_h.png}\\
@@
   \end{tabular}
-  Learn probabilistic program (a regex) $p_n$ from $K$ strings $x_n = \left\{y_n^k \right\}_{k = 1}^K$.
 
-  Likelihood model:
-  $$\probability [x_n|p_n] = \prod_{k = 1}^K \probability[y_n^k|p_n]$$
+  \vspace{2cm} \emph{Left:}  Agent controls a `pen' -- tasked with drawing pictures. \\\emph{Right:} During Sleep-R,
+  agent `dreams' by sampling programs from learned DSL and rendering them.
+
+\end{block}
+
+%----------------------------------------------------------------------------------------
+
+\end{column} % End of column 2.1
+
+\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
+
+\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
+\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}
+
+\begin{block}{Why this works: Bootstrapping}
+
 \begin{figure}
+\hspace*{-1.5cm}  \begin{tikzpicture}[scale=3,line width=1.5mm]
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
+    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
+
+    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+
+    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+ % \caption{\system solves for programs, the DSL, and a recognition model.
+Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
+\end{figure}
+%% \begin{itemize}
+%% \item Search finds new programs$\implies$ DSL+recognition model get more data
+%% \item DSL improves$\implies$easier search, recognition model gets better data
+%%   \item Recognition model improves$\implies$ easier search
+%%   \end{itemize}
+
+\end{block}
+
+\begin{block}{Learning from Scratch}
+  Start w/ McCarthy 1959 Lisp: recursion, conditionals, lists.
+  Train on 22 programming exercises.
+  After 93 hours on 64 CPUs, rediscovers 9 functional programming staples: \code{map},
+  \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length}, \code{range}, \code{incr}, \code{decr}.
+
+\end{block}
+
+\begin{block}{Generative models of text}
+    Learn probabilistic program: a stochastic regex\begin{figure}
 \centering
@@
 \end{figure}
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
-
-\end{column} % End of column 2.1
-
-\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
-
-\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
-\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}
-
-\begin{block}{Why this works: Bootstrapping}
-
-\begin{figure}
-\hspace*{-1.5cm}  \begin{tikzpicture}[scale=3,line width=1.5mm]
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
-    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
-
-    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
- % \caption{\system solves for programs, the DSL, and a recognition model.
-Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
-\end{figure}
-\begin{itemize}
-\item Search finds new programs$\implies$ DSL+recognition model get more data
-\item DSL improves$\implies$easier search, recognition model gets better data
-  \item Recognition model improves$\implies$ easier search
-  \end{itemize}
-
-\end{block}
-
-\begin{block}{Learning from Scratch}
-  Start w/ McCarthy 1959 Lisp: recursion, conditionals, lists.
-  Train on 22 programming exercises.
-  After 93 hours on 64 CPUs, rediscovers 9 functional programming staples: \code{map},
-  \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length}, \code{range}, \code{incr}, \code{decr}.
 

### 2018-07-10 poster.tex
final poster

@@
 
-\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Maxwell Nye, Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
+\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Maxwell Nye, Luke Hewitt,\\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
 
@@
 \end{figure}
-
+\vspace{-1cm}
 \end{block}
@@
   \newcommand{\dreamSize}{5.5cm}
-  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{12 (of 132) Tasks}}\\\hline
+  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{12 (of 136) Tasks}}\\\hline
     \includegraphics[width = \dreamSize]{turtleTasks/square/trimmed/train_square_3_h.png}&
@@
 
-  \vspace{2cm} \emph{Left:}  Agent controls a `pen' -- tasked with drawing pictures. \\\emph{Right:} During Sleep-R,
-  agent `dreams' by sampling programs from learned DSL and rendering them.
+  \vspace{2cm} \textbf{Left:}  Agent controls a `pen' -- tasked with drawing pictures. \textbf{Right:} During Sleep-R,
+   `dream' by sampling programs from learned DSL and rendering them.
+
+\end{block}
+
+\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
+
+\begin{block}{Acknowledgements}
+
+  We gratefully acknowledge collaboration with Eyal Dechter, whose EC algorithm (Dechter et al, IJCAI 2013) provided the inspiration for DreamCoder, and Luke Hewitt, who graciously provided us with
+  a regex learning data set.
 
@@
 
-\begin{block}{Learning from Scratch}
-  Start w/ McCarthy 1959 Lisp: recursion, conditionals, lists.
-  Train on 22 programming exercises.
-  After 93 hours on 64 CPUs, rediscovers 9 functional programming staples: \code{map},
-  \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length}, \code{range}, \code{incr}, \code{decr}.
-
-\end{block}
-
 \begin{block}{Generative models of text}
-    Learn probabilistic program: a stochastic regex\begin{figure}
+  \begin{figure}
 \centering
 %\tiny
-\tabcolsep=2pt
-\vspace{0pt}
-\begin{tabular}{cc}
-\begin{tabular}{ccc}
-\multicolumn{3}{c}{Tasks:}\\
+\tabcolsep=12pt
+\begin{tabular}{ccc}\toprule 
+\multicolumn{3}{c}{Three Tasks}\\
 \hline
-cut&F &Moss Side \\ 
-control&CL &Burnage \\
-control&F & City Centre\\
-cut& PCFL& Brooklands \\
-\multicolumn{3}{c}{Learned generative models:}\\
+1.14531&F&110.9 \\ 
+?&CL&163.2 \\
+1.29857&F&207.3\\
+?&PCFL&143.3 \\\\\midrule 
+\multicolumn{3}{c}{Three learned generative models}\\
 \hline
-\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash
-u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash
-u\textbackslash w*)*} \\
-\multicolumn{3}{c}{Samples from synthesized generative models:}\\
+ \pop{\scode{?\textbar(1.\textbackslash d+)}}& \pop{\scode{((\textbackslash u\textbackslash u)*)\textbar F}}& \pop{\scode{\textbackslash d\textbackslash d\textbackslash d.\textbackslash d}} \\\\\midrule
+ \multicolumn{3}{c}{Samples from generative models}\\
 \hline
-ya&DQDF & Vr DR \\ 
-glrwfdcnc&F &BeF lKQ \\
-mgs&F& W\\
-piljnl& KI& kqBfZ 0 \\
-kj& F& ON\\
-zci& GL& Bttc \\
-sxpm& F& S \\
-\hline
+1.61&DQDF& 343.8 \\ 
+?&F&241.2\\
+?&F&647.5\\
+1.2&KI&246.8%% \\
+%% 1.5080987&F&728.6\\
+%% ?&GL&029.3\\
+%% 1.453&F&289.1\\
+%%     \hline
 \\
-\end{tabular} 
-&
-\begin{tabular}{c}
-\begin{tabular}{l}
-\multicolumn{1}{c}{Learned DSL:}\\
-\hline
-$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
-$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash
-u\textbackslash w*)*}\\
-$f_3(x) = f_2(\scode{space}) = $\scode{( \textbar\textbackslash
-u\textbackslash w*)*}\\
-\hline
-$f_4(x) = $ \scode{($x$*$x$)}\\
-\emph{ (equivalent to regex `plus')}\\
-$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash
-l*\textbackslash l}$\\
-\hline
-\end{tabular}
-\\
-\\
-\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
-\end{tabular}
+
+\midrule \multicolumn{3}{c}{Learned DSL}\\\midrule
+\multicolumn{3}{l}{\popp{\scode{$f_1 = $ \textbackslash u\textbackslash w*}}
+\hspace{3cm}\popp{\scode{$f_2(x) = $($x$|$f_1$)*}}}\\% $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}}\\
+\multicolumn{3}{l}{\popp{\scode{$f_3 = f_2(\scode{`` ''}) = $( \textbar\textbackslash u\textbackslash w*)*}}}\\
+\multicolumn{3}{l}{\hspace{1cm}\emph{($f_3$: whitespace delimited words)}}\\
+\multicolumn{3}{l}{\popp{$f_4(x) = $ \scode{($x$*$x$)}}\hspace{1cm}\emph{($f_4$: regex `plus')}}
 \end{tabular}
@@
 \end{figure}
+
+\end{block}
+\begin{block}{Learning from Scratch}
+  Start w/ McCarthy 1959 Lisp: recursion, conditionals, lists.
+  Train on 22 programming exercises.
+  After 93 hours on 64 CPUs, rediscovers: \code{map},
+  \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length}, \code{range}, \code{incr}, \code{decr}.
 
@@
 %% %----------------------------------------------------------------------------------------
-
-\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
-
-\begin{block}{Acknowledgements}
-
-  We gratefully acknowledge collaboration with Eyal Dechter, whose EC algorithm (Dechter et al, IJCAI 2013) provided the inspiration for DreamCoder, and Luke Hewitt, who graciously provided us with
-  a regex learning data set.
-
-\end{block}
 

### 2018-07-19 talk.tex
talk

--- 2018-07-07 archive.tex
@@
-%%%%%%%% ICML 2018 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
-\documentclass{article}
-
-\usepackage{dblfloatfix}    % To enable figures at the bottom of page
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
-\usepackage{graphicx}
-%\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
-
-\usepackage[section]{placeins}
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
-%\usepackage{hyperref}
-
-% Attempt to make hyperref and algorithmic work together better:
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-\newcommand{\system}{\textsc{DreamCoder}~}
-\newcommand{\systemEnding}{\textsc{DreamCoder}}
-\newcommand{\lowerBound}{\mathscr{L}}
-\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
-\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
-\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
-% Use the following line for the initial blind version submitted for review:
-\usepackage[nohyperref]{icml2018}
-
+\documentclass{beamer}
+\usepackage{listings}
+\usepackage{amsfonts}
+\usepackage{dsfont}
+\usepackage{amsmath}
+\usepackage{bbm}
+\usepackage{verbatim}
+\usepackage{listings}
+%\usepackage{tipa}
+\usepackage{tikz}
+\usetikzlibrary{fit}
+\usepackage{color}
+\usepackage{booktabs}
+\usepackage{tipa}
+\usepackage{amssymb}
+\usepackage{verbatim}
+\usepackage[absolute,overlay]{textpos}
+\usepackage{pifont}% http://ctan.org/pkg/pifont
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\usetikzlibrary{bayesnet}
+\usetikzlibrary{decorations.markings}
+\usetikzlibrary{decorations.pathmorphing}
+\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
 \usepackage{xcolor}
@@
 
-\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
-\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}
+\newcommand{\nextForm}[1]{\rotatebox[origin=c]{270}{$_{\curvearrowright}$}$_{#1}$}
 
-%\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{mathrsfs}
-\usepackage{listings}
-\usepackage{amsthm}
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-%\usepackage{epsfig} % less modern
-\usepackage{subfig} 
+\usepackage{amsfonts}
+\usepackage{tabularx}
+%\usepackage{color}
+\usepackage{graphicx}
+\usepackage{booktabs}
+\usepackage{xcolor}
+\usepackage{tikz}
+\usetikzlibrary{trees}
+\usetikzlibrary{fit}
+\usetikzlibrary{calc}
+\usetikzlibrary{bayesnet}
+\usepackage[absolute,overlay]{textpos}
+\usepackage{stmaryrd}
+\newcommand{\sem}[1]{\llbracket #1\rrbracket}
+\newcommand{\tuple}[1]{\ensuremath{\left \langle #1\right \rangle}}
+\usepackage{booktabs}
+\usepackage{tipa}
+\usepackage{amssymb}
+\usepackage{verbatim}
+\usepackage[absolute,overlay]{textpos}
+\usepackage{pifont}% http://ctan.org/pkg/pifont
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\usetikzlibrary{bayesnet}
+\usetikzlibrary{decorations.markings}
+
+\usepackage[utf8]{inputenc}
+
+\usepackage{amssymb}% http://ctan.org/pkg/amssymb
+\usepackage{pifont}% http://ctan.org/pkg/pifont
+
 \usepackage{fancyvrb}
 
-\usepackage{caption}
-%\usepackage{subcaption}
+%% Program ::=
+%%   (if Bool List
+%%     (append RecursiveList
+%%             RecursiveList
+%%             RecursiveList))
+%% RecursiveList ::= List
+%%          | (recurse List)
 
-\fvset{fontsize=\footnotesize}
+\begin{SaveVerbatim}[]{ListSketch}
+Bool ::= (<= Int) | (>= Int)
+       | (= Int)
+Int  ::= 0
+       | (+1 Int) | (-1 Int)
+       | (length List)
+       | (head List)
+List ::= nil | X
+       | (filter Bool List)
+       | (tail List)
+       | (list Int)
+\end{SaveVerbatim}
 
-\usepackage{amssymb}
-\usepackage{listings}
-\usepackage{wrapfig}
-\usepackage{tabularx}
+\begin{SaveVerbatim}[]{ListSketchFull}
+Program ::= (if (Bool Int)
+                List
+                (append RecursiveList
+                        RecursiveList
+                        RecursiveList))
+RecursiveList ::= List
+                | (recurse List)
+Bool ::= (<= Int) | (>= Int)
+       | (= Int)
+Int  ::= 0
+       | (+1 Int) | (-1 Int)
+       | (length List)
+       | (head List)
+List ::= nil | X
+       | (filter Bool List)
+       | (tail List)
+       | (list Int)
+\end{SaveVerbatim}
 
-\usepackage{verbatim}
- \usepackage{booktabs}
- % For algorithms
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{tikz}
-\usepackage{circuitikz}
-\usetikzlibrary{fit,bayesnet}
-%\usetikzlibrary{arrows.meta}
-%\usetikzlibrary{positioning}
-%\usetikzlibrary{decorations.text}
-\usetikzlibrary{decorations.pathmorphing}
-\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
-\usepackage{dsfont}
-\usepackage{amsmath}
+\begin{SaveVerbatim}[]{Reverse}
+(if (<= (+1 0) (length X))
+    X
+    (append (recurse (tail X))
+            (list (head X))))
+\end{SaveVerbatim}
+\begin{SaveVerbatim}[]{Reversep}
+(if (= (+1 0) (length X))
+    X
+    (append (recurse (filter (<= (+1 0)) (tail X)))
+            (list (head X))))
+\end{SaveVerbatim}
 
-\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
-\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
+\begin{SaveVerbatim}[]{Count}
+(length (filter (= (head X)) (tail X)))
+\end{SaveVerbatim}
 
-% Packages hyperref and algorithmic misbehave sometimes.  We can fix
-% this with the following command.
+\begin{SaveVerbatim}[]{Countp}
+(-1 (length (filter (= (head X)) X)))
+\end{SaveVerbatim}
+
+\begin{SaveVerbatim}[]{Sort}
+(if (<= (+1 0) (length X))
+    X
+    (append (recurse (filter (<= (head X)) (tail X)))
+            (list (head X))
+            (recurse (filter (>= (head X)) (tail X)))))
+\end{SaveVerbatim}
+
+\begin{SaveVerbatim}[]{Sortp}
+(if (= (+1 0) (length X))
+    X
+    (append (recurse (tail X))
+            (list (length X))))
+\end{SaveVerbatim}
+
+\begin{SaveVerbatim}[]{TextSketch}
+Program ::= Term
+          | Program + Term
+Term    ::= String
+          | substr(Pos,Pos)
+Pos     ::= Number
+          | pos(Str,Str,Num)
+Num     ::= 0 | 1 | 2 | ... 
+          | -1 | -2 | ...
+Str     ::= Character 
+          | Character + String
+Character ::= a | b | c | ...
+\end{SaveVerbatim}
+
+\newcommand{\theSystem}{\textsc{ProgramSample}}
+\usepackage{arydshln}
 
 \newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
-\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em E}}
-\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
-\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}
+\newcommand{\Probability}{\mathds{P}} %{{\rm I\kern-.3em P}}
 
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2018}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
+\DeclareMathOperator*{\argmax}{arg\,max}
+%Information to be included in the title page:
+\title{Growing libraries of concepts with wake-sleep program induction}
+\author{Kevin Ellis \& Mathias Sabl\'e Meyer\\Joint work with: Lucas Morales, Armando Solar-Lezama, Joshua B. Tenenbaum\\Heavy inspiration from: Eyal Dechter}
+\institute{MIT} 
+%\date{November 2017}
 
@@
 
-\twocolumn[
-\icmltitle{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
+\frame{\titlepage}
 
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2018
-% package.
+\begin{frame}{The Language of Thought}
+\begin{tikzpicture}
+    \node at (-3,1) {\includegraphics[width = 6cm]{Fodor.jpg}};
+  \node at (2,2) {\includegraphics[width = 8cm]{Solomon.png}};
+%
+    \end{tikzpicture}
+\end{frame}
 
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
+\begin{frame}{Engineering the language of thought}
+ \begin{tikzpicture}
+    \node () {\includegraphics[width = 15.6cm]{theory.png}};
+    \node at (-5,-4) {  Ullman et al 2012};
+    \end{tikzpicture}
+\end{frame}
 
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
+\begin{frame}{Engineering the language of thought}
+  \begin{tikzpicture}
+    \node at(0,0) {\includegraphics[width = 8cm]{characters.png}};
+    \node at(-3,2) {\includegraphics[width = 8cm]{Brendan.jpg}};
 
-\begin{icmlauthorlist}
-\icmlauthor{Kevin Ellis}{MIT}
-\icmlauthor{Lucas Morales}{MIT}
-\icmlauthor{Matthias}{MIT}
-\icmlauthor{Maxwell Nye}{MIT}
-\icmlauthor{Armando Solar-Lezama}{MIT}
-\icmlauthor{Joshua B. Tenenbaum}{MIT}
-\end{icmlauthorlist}
+    \node at (-5,-4) {  Lake et al 2015};
+    \end{tikzpicture}
+\end{frame}
 
-%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
-\icmlaffiliation{MIT}{MIT}
-%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
+\begin{frame}{Growing a domain-specific language of thought}
 
-\icmlcorrespondingauthor{Kevin Ellis}{ellisk@mit.edu}
-%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
+  %  \Large
+  Goal: acquire domain-specific knowledge needed to induce a class of programs
 
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Machine Learning, ICML}
+  \pause
+  \vspace{1cm}
 
-\vskip 0.3in
-]
+  \begin{itemize}
+  \item Library of concepts (declarative knowledge)
+    \item Search strategy (procedural knowledge)
+    \end{itemize}
+\end{frame}
 
-% this must go after the closing bracket ] following \twocolumn[ ...
+\begin{frame}{DSL: Library of concepts}
+\includegraphics[width = 11cm]{3DSL.png}
+\end{frame}
 
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
+\begin{frame}{DSL learning as Bayesian inference}
+  \centering
+  \begin{tikzpicture}
+    \node[latent,align=center] (g) {DSL: \\concept library};
+    \node[latent,below = of g,xshift = -2cm] (e1) {$\text{program}_1$};
+    \node[latent,below = of g,xshift = 0cm] (e2) {$\text{program}_2$};
+ \node[latent,below = of g,xshift = 2cm] (e3) {$\text{program}_3$};
+    \node[obs,below = of e1] (x1) {$\text{task}_1$};
+    \node[obs,below = of e2] (x2) {$\text{task}_2$};
+    \node[obs,below = of e3] (x3) {$\text{task}_3$};
+    \edge {g} {e1,e2,e3};
+    \edge {e1} {x1};
+    \edge {e2} {x2};
+    \edge {e3} {x3};
+  \end{tikzpicture}
 
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
+\textbf{[Dechter et al., 2013]}  [Liang et al, 2010]; [Lake et al, 2015]
 
-%% \begin{abstract}
-%%   Successful approaches to program induction require a hand-engineered
-%%   domain-specific language (DSL), constraining the space of allowed
-%%   programs and imparting prior knowledge of the domain.  We contribute
-%%   a program induction algorithm called \system that learns a DSL while
-%%   jointly training a neural network to efficiently search for programs
-%%   in the learned DSL.\@ We use our model to synthesize functions on lists,
-%%   edit text, and solve symbolic regression problems, showing how the
-%%   model learns a domain-specific library of program components for
-%%   expressing solutions to problems in the domain.
-%% \end{abstract}
+  \end{frame}
 
-\section{Introduction}
+\begin{frame}{Amortized program inference}
 
-Much of everyday human thinking and learning can be understood in
-terms of program induction: constructing a procedure that maps inputs
-to desired outputs, based on observing example input-output pairs.
-People can induce programs flexibly across many different domains, often from just one or a few examples.  For instance, if
-shown that a text-editing program should map ``Jane Morris Goodall''
-to ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey''
-to ``R. E. Leakey''; if instead the first input mapped to
-``Dr. Jane'' or ``Goodall, Jane'', we might guess
-the latter should map to ``Dr. Richard'' or ``Leakey, Richard'',
- respectively.
+  \includegraphics[width = 9cm]{rm.png}
 
-The FlashFill system~\cite{gulwani2011automating} embedded in
-Microsoft Excel solves problems such as these and is probably the best
-known practical program-induction algorithm, but researchers in
-programming languages and AI have had successes in many domains, such as handwriting recognition and
-generation~\cite{lake2015human}, procedural
-graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
-answering~\cite{johnson2017clevr} and robot motion
-planning~\cite{devlin2017neural}.  These systems
-work in different ways, but most hinge upon a carefully engineered
-\textbf{Domain Specific Language (DSL)}.  This is especially true for
-systems like FlashFill that induce a wide range of programs
-very quickly, in a few seconds or less.  DSLs constrain the search
-over programs with strong prior knowledge in the form of a restricted
-inventory of programming primitives finely tuned to the domain: for
-text editing, these are operations like appending and
-splitting on characters.
+  \begin{itemize}
+  \item $\mathcal{D}$: DSL
+  \item $p_n$: program
+  \item $x_n$: task
+  \item $Q$: Recognition model
+    \end{itemize}
 
-In this work, we consider the problem of building agents that learn to solve
-program induction tasks, and also the problem of acquiring the prior knowledge
-necessary to quickly solve these tasks in a new domain.  Representative problems
-in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
-algorithm that grows or boostraps a DSL while jointly training a neural network
-to help write programs in the increasingly rich DSL.
+\end{frame}
 
-\begin{table*}%[t!]
-  \makebox[\textwidth][c]{
-    \scriptsize
-  \tabcolsep=4pt
-  \renewcommand\code\texttt
-  \renewcommand\codechar[1]{\texttt{"#1"}}
-  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
-    \toprule
-    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
-      \begin{tabular}{cc}
-        \begin{tabular}{c}
-          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
-          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
-          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
-          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
-          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
-          \\
-          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
-          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
-        \end{tabular}
-        &
-        \hspace{-0.3cm}\begin{tabular}{c}
-          \code{[7\, 3]}$\to $\code{False}                              \\
-          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
-          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
-          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
-          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
-        \end{tabular}
-      \end{tabular}
-    }
-    &
-    \hspace{-0.5cm}\begin{tabular}{c}
-      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
-      +83 973-831$\to $83.973.831\\
-      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
-      }}\\
-      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
-      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
-      ~\\
-      Temple Anna H $\to $TAH\\
-      Lara Gregori$\to $LG\\
-      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
-    \end{tabular}
-    &
-    \begin{tabular}{cc}
-      \includegraphics[width = 3em]{figures/functions/4.png}&
-      \includegraphics[width = 3em]{figures/functions/146}\\
-      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
-      ~\\
-      \includegraphics[width = 3em]{figures/functions/112.png}&
-        \includegraphics[width = 3em]{figures/functions/92.png}
-      \\
-      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\
+\begin{frame}{DreamCoder}
+  \begin{itemize}
+  \item   \textbf{Wake:} Solve problems by writing programs
+  \item \textbf{Sleep:} Improve DSL and neural recognition model:
+    \begin{itemize}
+    \item \textbf{Sleep-G:} Improve DSL (\textbf{G}enerative model)
+      \item \textbf{Sleep-R:} Improve \textbf{R}ecognition model
+      \end{itemize}
+  \end{itemize}
+  Based on Exploration-Compression algorithm by Eyal Dechter
 
-    \end{tabular}
-    ~\\
-    \midrule
-    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
-    \hspace{0cm}\begin{tabular}{l}
-      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
-      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
-      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
-      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
-      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
-      %NOTE: this is the actual invention, but I removed a redundant lambda
-      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
-      %% \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
-      %% \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
-      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
-      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
-      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
-      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
-      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
-      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
-      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
-      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
-      \phantom{$f_1(\ell,$}
-      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
-      \hspace{\helpSize}($f_3$: \emph{Whether $\ell$ contains }\code{k})\\
-    \end{tabular}&
+\pause
+  \vspace{1cm}
 
-  \hspace{0.5cm}\begin{tabular}{l}
-    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
-    (x)}}\\
-    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
-      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
-      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
-      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
-        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
-      }}\\
-      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
-        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
-      %%   \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
-      %% \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
-  \end{tabular}&
-
-  \begin{tabular}{l}
-    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
-    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
-    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
-    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
-    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
-    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
-    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
-    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
-    \hspace{\helpSize}\emph{($f_6$: rational function)}\\
-
-  \end{tabular}
-  \\\bottomrule\\
-\end{tabular}}\vspace{-0.5cm}
-\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
-Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}%\vspace{-0.5cm}
-\end{table*}
-
-Because any learning problem can in principle be cast as
-program induction, it is important to delimit our focus.  In contrast
-to computer assisted programming~\cite{solar2008program} or genetic
-programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
-engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
-capturing the human ability to learn to think flexibly and efficiently
-in new domains --- to learn what you need to know about a domain so you
-don't have to solve new problems starting from scratch.  We are
-focused on problems that people solve relatively
-quickly, once they acquire the relevant domain expertise.  These
-correspond to tasks solved by short programs --- if you have an
-expressive DSL.  %% Even with a good DSL, program search may be
-%% intractable;
-%
-% , and adding new library routines to the DSL only broadens the
-% search space;
-%
-%% so we train a
-%% neural network to guide the search procedure.
-
-%\section{Overview of \system}
-
-Our algorithm takes inspiration from several ways that skilled human
-programmers have learned to code: Skilled coders build libraries of
-reusable subroutines that are shared across related programming tasks,
-and can be composed to generate increasingly complex and powerful
-subroutines.  In text editing, a good library should support routines
-for splitting on characters, but also specialize these routines to
-split on particular characters such as spaces or commas that are
-frequently used to delimit substrings across tasks.  Skilled coders
-also learn to recognize what kinds of programming idioms and library
-routines would be useful for solving the task at hand, even if they
-cannot instantly work out the details.  In text editing, one might
-learn that if outputs are consistently shorter than inputs, removing
-characters is likely to be part of the solution; if every output
-contains a constant substring (e.g., ``Dr.''), inserting or appending
-that constant string is likely to be a subroutine.
-
-Our algorithm is called \system because it is based on a
-novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
-between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve tasks; creating a DSL by discovering
-and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL.
-The learned DSL effectively encodes a prior on programs likely to
-solve tasks in the domain, while the neural net looks at the example
-input-output pairs for a specific task and produces a ``posterior''
-for programs likely to solve that specific task.  The neural network
-thus functions as a \textbf{recognition model} supporting a form of
-approximate Bayesian program induction, jointly trained with a
-\textbf{generative model} for programs encoded in the DSL, in the
-spirit of the Helmholtz machine~\cite{hinton1995wake}. The
-recognition model ensures that searching for programs remains
-tractable even as the DSL (and hence the search space for programs)
-expands.
-
-%% Our \system
-%% algorithm incorporates these insights by iterating
-Concretely, our algorithm iterates through three phases. 
-The \textbf{Wake} phase takes a given set of \textbf{tasks}, typically
-several hundred, and searches for compact programs that solve these tasks,
-guided by the current DSL and neural network.
-The \textbf{Sleep-G} phase grows the DSL (or \textbf{G}enerative model),
-which allows the agent to more compactly
-write programs in the domain. We modify the structure of the DSL by
-discovering regularities across programs, compressing
-them to distill out common abstractions across  programs discovered during waking.\footnote{This is loosely biologically
-    inspired by the formation of abstractions during sleep memory
-    consolidation~\cite{DUDAI201520}}
-The \textbf{Sleep-R} phase improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
-write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}.
-We train the recognition model on  two data sources: samples
-from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
-programs found during waking.%%  As the DSL
-%% improves and as search finds more programs, the recognition model gets
-%% both more data to train on, and better data.
-\footnote{These two sources are also
-    loosely biologically inspired by the distinct episodic replay and
-    hallucination components of dream sleep~\cite{fosse2003dreaming}}
-
-\textbf{Brisk overview of related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
-both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
-FlashFill:~\cite{gulwani2011automating}),
-neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
-A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
-which also learns a library of subroutines.
-We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
-%%   \setlength\itemsep{0}
-%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
-%%   \item We give an algorithm for learning DSLs, built on a formalism known
-%%     as Fragment Grammars~\cite{tim}; and 
-%%   \item We give a hierarchical Bayesian framing of the problem that allows a
-%%     joint inference of the DSL and recognition model.
-%% \end{itemize}
-\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
-(2) We give an algorithm for learning DSLs,
-built on a formalism known as Fragment Grammars~\cite{tim};
-and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.
-up
-
-\section{The \system Algorithm}
-
- We first mathematically describe our 3-step algorithm as
- an inference procedure for a hierarchical Bayesian model (Section~\ref{mathematicalFraming}),
-  and then describe each step algorithmically in detail (Section~\ref{explorationSection}-\ref{grammarInductionSection}).
-
- \subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}
-
-\system takes as input a set of \textbf{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.\footnote{For example, for string editing,
-  the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
-  and 0 otherwise; when learning a generative model or probabilistic program, the likelihood is
-the probability of the program sampling the observation.}
-Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL, written $\mathcal{D}$.
-We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
-$(\mathcal{D},\theta)$ define a generative model over programs.
-We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
-Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
-\begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta &\qquad
-  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
-\end{align}
-
-%% Inference in this model is
-%% difficult because the programs are unobserved,
-%% and so we must solve a hard search problem to recover them. To make
-%% search tractable we learn a bottom-up \emph{recognition
-%%   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
-%% The recognition model $q(\cdot )$ is a neural
-%% network that regresses from input/output pairs to a distribution over
-%% programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
-%% inference
-%% scheme~\cite{le2016inference}.
-%%  The neural recognition model and the
-%% generative model embodied in the DSL jointly train each other, as they
-%% iteratively learn to solve more programming tasks.
-
-The above equations summarize the problem from the point of view of an ideal Bayesian learner.
-However, Eq.~\ref{intractableObjectives}
-is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
-summing over the  infinite set of all programs.
-In practice we will only ever be able to sum over a finite set of programs.
-So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
-\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
-is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
-
-Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
-\begin{align}
- J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
-\end{align}
-We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
-and $(\mathcal{D},\theta)$ (\textbf{Sleep-G}):
-
-\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
-want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
-is large (i.e., programs with high posterior probability).
-%Here we use a simple and generic enumerative program synthesis algorithm.
-%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
-%% where  is large.
-\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
-and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
-%To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.
-
-Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
-during the \textbf{Sleep-R} phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$.
-  Thus training the neural network amortizes the cost of finding programs with high posterior probability.
-
-\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
-  frontiers.}  Here we train %% a neural network, $q$, to predict a
-%% distribution over programs conditioned on a task. The objective of $q$
-%% is
-$q(p|x)$ to assign high probability to programs $p$ where
-$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
-into the frontiers will most increase $\lowerBound$.
-
-Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figure~\ref{feeding}).
-As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
-the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
-samples from the DSL and programs found during waking,
-the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
-\begin{figure}\centering
-  \begin{tikzpicture}
+  \centering\begin{tikzpicture}
     \begin{scope}[shift = {(1,-1)}]
@@
     \end{tikzpicture}
-  \caption{\system solves for programs, the DSL, and a recognition model.
-Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
-\end{figure}
 
-\begin{figure}
-  \begin{tikzpicture}
+\end{frame}
+
+\begin{frame}{Wake phase}
+  \centering\begin{tikzpicture}[scale=1.3]
     \node at (0,0) (d){DSL};
@@
 
+\end{frame}
+
+\begin{frame}{Sleep-G Phase}
+  Fragment grammar (\textbf{O'Donnell 2015})
   \vspace{1cm}
 
-  \begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
-
-    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
-    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
-    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
-
-    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
-    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
-    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
-
-    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
-    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
-    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
-  \end{tikzpicture}
-
-  \begin{tikzpicture}
-    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
-    \node at ([xshift = 3cm]d.east) (p2){program};
-    \node at ([yshift = 1.5cm]p2) (p1){program};
-    \node at ([yshift = -1.5cm]p2) (p3){program};
-
-    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
-    \draw[squiggle,-> ] (d.east) -- (p1.west);
-    \draw[squiggle,-> ] (d.east) -- (p3.west);
-
-    \node at ([xshift = 2cm]p1.east) (t1){task};
-    \node at ([xshift = 2cm]p2.east) (t2){task};
-    \node at ([xshift = 2cm]p3.east) (t3){task};
-    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
-    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
-    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
-  \end{tikzpicture}
-
-  \vspace{2cm}
-
-  \begin{tikzpicture}
+  \centering\begin{tikzpicture}
     \node at (0,0) (f1){Frontier$_1$};
@@
     \end{tikzpicture}
-  \end{figure}
 
-\begin{figure}[h]
-\centering
-%  \includegraphics{figures/iterations.eps}
-\begin{tikzpicture}
-  \begin{scope}[shift={(2.5,0)}]
-    \node[obs] at (3,3) (dx){$\mathcal{D}$};
-    \node[latent] at (3.5,1.75) (zp){$p$};
-%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-    \node[obs] at (3.5,0.7) (xp) {$x$};
-    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
-    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
-%    \draw [->,red] (tx) -- (zp);
-    \draw [->] (dx) -- (zp);
-    \node at (nn) {
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
+\end{frame}
+
+\begin{frame}{Sleep-R Phase}
+Train recognition model on two sources of data:
+  \begin{itemize}
+  \item Samples from generative model
+    \item Programs found during waking
+    \end{itemize}
+
+\end{frame}
+
+\begin{frame}{Sleep-R: Train on samples from generative model}
+\centering  \begin{tikzpicture}[scale=1.2]
+    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
+    \node at ([xshift = 3cm]d.east) (p2){program};
+    \node at ([yshift = 1.5cm]p2) (p1){program};
+    \node at ([yshift = -1.5cm]p2) (p3){program};
+
+    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
+    \draw[squiggle,-> ] (d.east) -- (p1.west);
+    \draw[squiggle,-> ] (d.east) -- (p3.west);
+
+    \node at ([xshift = 2cm]p1.east) (t1){task};
+    \node at ([xshift = 2cm]p2.east) (t2){task};
+    \node at ([xshift = 2cm]p3.east) (t3){task};
+    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
+    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
+    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
+
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
   \end{tikzpicture}
-    };
-    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-  \end{scope}
 
-  \begin{scope}[shift={(0.7,-4.2)}]
-    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-    \draw [->] (p2.south) -- (x2.north);
-    \draw [->] (dt.south) -- (p2.north);
+\end{frame}
 
-    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-    \draw [->] (p1.south) -- (x1.north);
-    \draw [->] (dt.south) -- (p1.north);
+\begin{frame}{Sleep-R: Train on programs found during waking}
+\centering  \begin{tikzpicture}[scale=1.2]
+    \node at (0,0) (f1){Frontier$_1$};
+    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
 
-    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-    \draw [->] (p3.south) -- (x3.north);
-    \draw [->] (dt.south) -- (p3.north);
+    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
+    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
+    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
 
-    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
+    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
+    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
 
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-    \draw [->] (zp.south) -- (xp.north);
-    \plate {}{(zp)(xp)}{$x\in X$};
+    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
+    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
+    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
 
-    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
-    rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
-  \end{scope}
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
+  \end{tikzpicture}
 
-  \begin{scope}[shift={(7.7,-4)}]
-    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-    \node[latent] at (1.5,3) (t){$\theta$};
-    \node[obs] at (1,1.75) (z){$p$};
-%     \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-    \node[obs] at (1,0.5) (x) {$x$};
-    \edge {z}{x};
-    \edge {d,t}{z};
-    \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
-  \end{scope}
+\end{frame}
 
-\end{tikzpicture}
-\end{figure}
+\begin{frame}{List functions}
+  Created+investigated by Lucas Morales
 
-\subsection{Wake: Searching for Programs}\label{explorationSection}
+  \vspace{1cm}
 
-During waking, the agent's goal is to search for programs solving the tasks.  We use the simple approach of enumerating programs from
-the DSL  in decreasing order of their probability according to the recognition model,
-and then checking if a program $p$ assigns positive
-probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
-into the frontier $\mathcal{F}_x$.
-
-To make this concrete we need to define what programs actually are and
-what form $\probability[p |\mathcal{D},\theta]$ takes.
-We represent programs as $\lambda$-calculus expressions.
-$\lambda$-calculus is a formalism for expressing functional programs
-that closely resembles Lisp,
-including variables, function application, and the ability to create new functions.
-Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
-Our programs are all strongly typed.
-We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
-used in functional programming languages like OCaml and Haskell.
-%% Type variables are always written using lowercase Greek letters
-%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
-%% and returns something of type $\beta$.
-%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
-%% has the type $\tau$.
-%% For example, to describe the type of the identity function
-%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
-%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
-%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
-%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
-%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
-%% See Supplement for more detail on program representation.
-We now define DSLs:
-
-\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
-A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-%\\\noindent\textbf{Definition.}
-A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
-and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
-
-Together with its weight vector,
-a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
-In the supplement, we define this distribution  
-by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$.
-Care must be taken to ensure that programs are well-typed
- and that variable scoping rules are obeyed.
-%% With this distribution in hand, we search for programs by enumerating
-%% $\lambda$-calculus expressions in decreasing order of their probability under
-%% $(\mathcal{D},\theta)$.
-
-Why enumerate, when the program synthesis community has invented many
-sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
-We have two reasons:
-(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
-(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-  the space of  programs.
-
-  However, a drawback of   enumerative search  is that we have no
-efficient means of solving for arbitrary constants that might occur in a
-program. In Sec.~\ref{regressionSection},
-we will show how to find programs with real-valued constants
-by automatically differentiating through the program and setting the constants using gradient descent.
-%% In Sec.~\ref{textSection}
-%% we will show that the bottom-up neural recognition model can learn
-%% which discrete constants should be included in a program.
-
-\subsection{Sleep-R: Training a Neural Recognition Model}\label{recognitionSection}
-
-The purpose of training the recognition model is to amortize the cost
-of searching for programs.  It does this by learning to predict, for
-each task, programs with high likelihood according to
-$\probability[x|p]$ while also being probable under the prior
-$(\mathcal{D},\theta)$.  Concretely, the recognition model $q$
-predicts, for each task $x\in X$, a weight vector $q(x) =
-\theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the
-DSL, this defines a distribution over programs,
-$\probability[p|\mathcal{D},\theta = q(x)]$.  We abbreviate this
-distribution as $q(p|x)$.  The crucial aspect of this framing is that
-the neural network leverages the structure of the learned DSL, so it
-is \emph{not} responsible for generating programs wholesale.  We share
-this aspect with DeepCoder~\cite{balog2016deepcoder}
-and~\cite{menon2013machine}.
-
-How should we get the data to train $q$?
-This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
-One approach is to sample programs from the DSL,
-run them to get their input/outputs,
-and then train $q$ to predict the program from the input/outputs.
-This is like how a Helmholtz machine
-trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
-The advantage of ``Helmholtz machine'' training is that
-we can draw unlimited samples from the DSL,
-training on a large amount of data.
-Another approach is
- self-supervised learning,
-training $q$ on the (program, task)
-pairs discovered during waking.
-The advantage of self-supervised learning
-is that the training data is much higher quality,
-because we are training on the actual tasks.
-Due to these complementary advantages,
-we train on both these sources of data.
-
-Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
- equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
-%% \begin{equation*}
-%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
-%% \end{equation*}
- where the expectation is taken over tasks.
-Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
- The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives:
- \begin{align}
-   \mathcal{L}_{\text{RM}} &= \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}\\
-\nonumber\mathcal{L}_{\text{HM}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\\
-\nonumber\mathcal{L}_{\text{SS}} &= \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
-\end{align}
-
-  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
-all of our experiments train \system on only a few hundred tasks, which is too little for
-a high-capacity neural network $q$.
-Once we bootstrap a $(\mathcal{D},\theta)$,
-we can draw unlimited samples from $(\mathcal{D},\theta)$
-and train $q$ on those samples.
-
-Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
-the current DSL, running them to get their outputs,
-and then training $q$ to regress from the input/outputs to the program.
-Since these programs map inputs to outputs,
-we need to sample the inputs as well.
-Our solution is to sample the inputs
-from the empirical observed distribution of inputs in $X$.
-
-\subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
-
-The purpose of the DSL is to
-offer a set of abstractions
-that allow an agent to easily express solutions to the tasks at hand.
-%In the \system algorithm we infer the DSL from a collection of frontiers.
-Intuitively, we want the algorithm to
-look at  the frontiers and
-generalize beyond them, 
-both so the DSL can better express the current solutions,
-and  also so that the DSL might expose new abstractions
-which will later be used to
-discover more programs.
- Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
-We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
-\begin{align}
-\nonumber      \log \probability[\mathcal{D}]& + 
-\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
-\label{AIC}
-  \end{align}
-We induce a DSL by searching locally through the space of DSLs,
-proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
-The search moves work by introducing new
-$\lambda$-expressions into the DSL.
-We propose these new expressions by extracting fragments of
-programs already in the frontiers (Tbl.~\ref{fragmentExample}).
-An important point here is that we are \emph{not} simply adding
-subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
-extracting fragments that unify with programs in the frontiers.  This
-idea of storing and reusing fragments of expressions comes from
-Fragment Grammars~\cite{tim} and Tree-Substitution
-Grammars~\cite{cohn2010inducing}, and is closely related to the idea
-of antiunification~\cite{henderson2013cumulative}.
-
-%\vfill\null
-
-%\columnbreak
-
-To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-  and $\lambda$ controls how strongly we regularize the size of the DSL.
-  We place a symmetric Dirichlet prior over the weight vector $\theta$,
-  defining 
-$  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
- $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
-
-To appropriately score each proposed $\mathcal{D}$ we must reestimate
- the weight vector $\theta$.
-Although this  may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-our DSLs are context-sensitive due to the presence of variables
-in the programs and also due to the polymorphic typing system.
-In the Supplement we derive a tractable MAP estimator for $\theta$.
-%% Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-%% recognition model training, and DSL induction.
-
-\begin{figure*}[!htb]\tabcolsep=2pt
-\centering  \begin{minipage}[c]{0.2\textwidth}
-  \begin{tikzpicture}[scale=0.7]
-    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
-    %%     \footnotesize Program:\\
-    %%     \code{($\lambda$ (x) (+ (- x) 1))}
-    %% \end{tabular}};
-    %\node at (,0.5) {\code{cons}};
-    \node [rotate=90] at (-2.3,-0.5) {\small program};
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-%    \draw  (l1.south) -- (-0.5,0.5);
-
-  %% \node(c) at (-0.5,-1.5) {\code{-}};
-  %% \node(z) at (0.5,-1.5) {\code{x}};
-
-  %% \draw (0,-1) -- (c.north);
-  %% \draw (0,-1) -- (z.north);
-
-  \begin{scope}[shift={(0,-2.5)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-  \end{scope}
-
-\begin{scope}[shift={(0,-5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  %\node(x1) at (0,-1) {};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.55,-0.4) -- (a.north);
-  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
-
-  \end{scope}
-
-\end{tikzpicture}
-  \end{minipage}
-\hspace{0.1cm}\begin{tabular}{ll}
-    \toprule
-    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
-    \begin{tabular}{l}
-      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
-      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
-    \end{tabular}&
-    \begin{tabular}{l}
-      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
-    \end{tabular}\\\midrule
-    \begin{tabular}{l}
-      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
-      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
-      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
-      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
-      \end{tabular}&
-    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
-      \end{tabular}
-\\\bottomrule\\
-\end{tabular}
-\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
-highlighted in {\orange{orange}},
-  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
-  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
-\end{figure*}
-
-\begin{comment}
-  \begin{wrapfigure}{R}{0.6\textwidth}
-  \begin{minipage}{0.6\textwidth}    
-    \begin{algorithm}[H]
-      \caption{The \system Algorithm}
-      \label{mainAlgorithm}
-      \begin{algorithmic}
-        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-        \FOR{$i=1$ {\bfseries to} $I$}
-        %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
-        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
-        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
-        %% \STATE Define $Q_x(z) \propto \begin{cases}
-        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-        %%   0&x\not \in \mathcal{F}_x
-        %% \end{cases}
-        \ENDFOR
-        \STATE \textbf{return} $\mathcal{D},\theta,q$
-      \end{algorithmic}
-    \end{algorithm}
-  \end{minipage}
-\end{wrapfigure}
-  \end{comment}
-
-\section{Experiments}
-
-\subsection{Programs that manipulate sequences}\label{sequences}
-We apply \system to list processing %% (Section~\ref{listSection})
-and text editing, %% (Section~\ref{textSection})
-using a GRU~\cite{cho2014learning} for
-the recognition model, and initially providing the system with generic sequence manipulation primitives:
-\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
-\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
-\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
-
-%\subsubsection{List Processing}\label{listSection}
-\textbf{List Processing:} Synthesizing programs that manipulate data structures is a widely studied
-problem in the programming languages community~\cite{feser2015synthesizing}.
-%with applications to computer aided programming~\cite{solar2008program}.
-We consider this problem within the context of learning functions that
-manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
-We created 236 human-interpretable list manipulation tasks, each with 15
-input/output examples.
-%% Our data set is interesting in three major ways: many of the tasks
-%% require complex solutions; the tasks were not generated from some
-%% latent DSL, and the agent must learn to solve these complicated
-%% problems from only 236 tasks.
-%% Our data set assumes arithmetic operations as well as sequence operations,
-%% so we additionally provide our system with the following arithmetic
-%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
-%% \code{is-prime}.
-%% evaluating  on a 50/50 test/train split.
-In solving these tasks, the system
-composed 38 new subroutines, and rediscovered the higher-order
-function \code{filter} ($f_1$ in Table~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
-%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
   \begin{figure}[b]\centering
@@
   \end{tabular}
-  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
+  %\captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
 \end{figure}
 
-%\subsubsection{Text Editing}\label{textSection}
-\textbf{Text Editing:} Synthesizing programs that edit text is a classic problem in the
-programming languages and AI literatures~\cite{gulwani2011automating,lau2001programming}.
-This prior work uses hand-engineered DSLs.
-Here, we instead start out with generic sequence manipulation
-primitives and recover many of the higher-level building blocks that 
-have made these other systems successful.
-%% Because our enumerative search procedure cannot generate string %
-%% constants, we instead enumerate programs with string-valued
-%% parameters.  For example, to learn a program that prepends ``Dr.'', we
-%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
-%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
-%% define $\probability[x|p]$ by approximately marginalizing out the
-%% string parameters via a simple dynamic program.
-%% In Sec.~\ref{regressionSection}, we will use a similar trick to
-%% synthesize programs containing real numbers, but using gradient
-%% descent instead of dynamic programming.
+  Discovers 38 concepts, including `filter'
+\end{frame}
 
-We trained our system on 109 automatically-generated text editing tasks, with 4 input/output examples each.
-After three iterations, it assembles a DSL containing a dozen new functions (Fig.~\ref{initialExampleDSL}, center) solving
-all the training tasks.
-But, how well does the  learned DSL generalized to real text-editing scenarios?
-We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
-\system solves 3.7\% of the problems with an average search time of 235 seconds.
-After learning,
-it solves 74.1\%, and does so much faster,
-solving them in an average of 29 seconds.
-As of the 2017 SyGuS competition,
-the best-performing algorithm solves 79.6\% of the problems.
-But, SyGuS comes with a
-different hand-engineered DSL \emph{for each text editing problem}.
-%\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
-For these experiments, our system did not use this assistance.}
-Here  we learned a single DSL
-that applied generically to
-all of the tasks,
-and perform comparably to the best
-prior work.
+\begin{frame}{Text editing}
+  In the style of FlashFill (Gulwani 2012)
 
-\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
-We apply \system
-to symbolic regression problems.  Here, the
-agent observes points along the curve of a function, and must write a
-program that fits those points.  We initially equip our learner with
-addition, multiplication, and division, and task it with solving
-100 %% symbolic regression
-problems, each either a polynomial or rational function.  The recognition model is a
-convnet that observes an image of the target function's
-graph (Fig.~\ref{functions}) --- visually, different kinds of
-polynomials and rational functions produce different kinds of graphs,
-and so the convnet can  look at a graph and predict
-what kind of function best explains it.
-A key difficulty, however, is
-that these problems are best solved with programs containing real
-numbers.  Our solution to this difficulty is to enumerate
- programs with real-valued parameters, and then fit those
-parameters by automatically differentiating through the programs the
-system writes and use gradient descent to fit the parameters.
-We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
-and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
-\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
-  \begin{minipage}[c]{0.3\columnwidth}%\vspace{-0.3cm}
-  \includegraphics[width = \functionSize]{figures/functions/6.png}
-  \includegraphics[width = \functionSize]{figures/functions/48.png}\\
-  \includegraphics[width = \functionSize]{figures/functions/102.png}
-  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
-  \vspace{1pt}%\includegraphics[width = \functionSize]{figures/functions/181.png}
-  %\includegraphics[width = \functionSize]{figures/functions/160.png}
-  \includegraphics[width = \functionSize]{figures/functions/160.png}
-    \includegraphics[width = \functionSize]{figures/functions/149.png}
-  \end{minipage}
-  \begin{minipage}[c]{0.69\columnwidth}    
-    \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}\vspace{-1cm}
-        \end{minipage}
-\end{figure}
-We learn a DSL containing 13 new functions,
-mainly templates for polynomials of different orders or ratios of polynomials.
-The model also learns to find programs minimizing the number of continuous parameters ---
-learning to represent linear functions with 
-\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
-which has five continuous parameters.
-This phenomenon arises from our Bayesian framing:
-both the generative model's bias toward shorter programs,
-and the likelihood model's BIC penalty.
+  \centering  \includegraphics[width = 5cm]{textColumn.png}
 
-\subsection{Learning from Scratch}
-A long-standing dream within the program induction community
-is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
-and then learning to solve a wide swath of
-induction problems~\cite{solomonoff1964formal,schmidhuber2004optimal,hutter2004universal,solomonoff1989system}.
-All existing systems,
-including ours,
-fall far short of this dream,
-and it is unclear (and we believe unlikely)
-that this dream could ever be fully realized.
-How far can we push in this direction?
-``Learning from scratch'' is subjective, but a reasonable
-starting point is the set of primitives provided in 1959
-Lisp~\cite{mccarthy1960recursive}: these include
-conditionals, recursion, arithmetic, and the 
-list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
-A  basic first goal is to start with
-these primitives,
-and then recover a DSL that
-more closely resembles modern functional languages like Haskell and OCaml.
-Recall (Sec.~\ref{sequences})
-that we initially provided our system with functional programming routines like
-\code{map} and \code{fold}.
+  SyGuS problems: solves 3\% before learning, vs 74\% after learning. Best prior work: 80\%
 
-We ran the following experiment: \system was given a subset of
-the 1959 Lisp primitives, and tasked with solving 22
-programming exercises. A key difference between
-this setup and our previous experiments is that,
-for this experiment,
-the system is
-given primitive recursion,
-whereas previously we had sequestered recursion within
-higher-order functions like \code{map}, \code{fold}, and \code{unfold}.
+  \end{frame}
 
-After running for 93 hours on 64 CPUs, our
-algorithm solves these 22 exercises, along the way assembling a DSL
-with a modern repertoire of
-functional programming idioms and subroutines, including \code{map},
-\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
-and  arithmetic operations like 
-building lists of natural numbers between an interval (see Figure~\ref{learningFromScratch}).
-We did not use the recognition model for this experiment:
-a bottom-up pattern recognizer is of little use
-for acquiring this abstract knowledge
-from less than two dozen problems.
+\begin{frame}{Symbolic regression from visual input}
+\centering\includegraphics[width = 5cm]{symbolicRegression.png}
+\end{frame}
 
-\begin{figure}  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{cl}\toprule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
-    \begin{tabular}{l}
-      \code{[4\, 7\, 6\, 9\, 0]$\to$5}\\
-\code{[3\, 8\, 2\, 1\, 4]$\to$4}\\
-\code{[2\, 2\, 9\, 4\, 8]$\to$3}\\
-\code{[0\, 9\, 8\, 9\, 9\, 8]$\to$1}\\
-\blueCode{$f(\ell) = $($f_0$ (car $\ell$))}\\\\
+\begin{frame}{Turtle graphics}
+  Created+Investigated by Mathias Sabl\'e Meyer
 
-\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
-\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
-\blueCode{$f(\ell) = $($f_4$ + $\ell$ 0)}\\\\
+  \vspace{0.5cm}
 
-\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
-\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
-\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
-\blueCode{$f(\ell) = $($f_5$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
+  \begin{itemize}
+  \item pen up/pen down
+  \item move pen (forward+rotate)
+  \item get/set pen state
+  \item `for' loops
+  \end{itemize}
 
-      \code{[5\, 2\, 9]}$\to$\code{[[2\, 9]\, [9]\, []]}\\
-      \code{[3\, 8\, 1\, 3]}$\to$\code{[[8\, 1\, 3]\, [1\, 3]\, [3]\, []]}\\
-      \blueCode{$f(\ell) = $($f_2$ empty? cdr cdr $\ell$)}
-      \\
+  \centering
+  \includegraphics[width = 8cm]{exampleGraphicsPrograms.png}
 
-      \end{tabular}
-    \\\midrule
-    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
-    \begin{tabular}{l}
-      \greenCode{$f_0($x$)\,=\,$(+ x 1)}\\
-      \hspace{\helpSize}($f_0$: \emph{increment})\\
-      \greenCode{$f_1($x$)\,=\,$(- x 1)}\\
-      \hspace{\helpSize}($f_1$: \emph{decrement})\\
-        \popp{\code{$f_2($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_2$ (n x))))}}\\
-        \hspace{\helpSize}($f_2$: \emph{unfold})\\
-        \popp{\code{$f_3($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{($f_3$ ($f_1$ i) (cdr l))))}}\\
-        \hspace{\helpSize}($f_3$: \emph{index})\\
-        \popp{\code{$f_4($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
-      \phantom{\code{$f_4($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_4$ (cdr l))))}}\\
-        \hspace{\helpSize}($f_4$: \emph{fold})\\
-        \greenCode{$f_5($f$,$l$)\,=\,$(if (empty? l) nil)}\\
-        \phantom{\greenCode{$f_5($f$,$l$)\,=\,$(if }}\greenCode{(cons (f (car l)) ($f_5$ (cdr l)))}\\
-        \hspace{\helpSize}($f_5$: \emph{map})
+\end{frame}
+\begin{frame}{Turtle graphics}
+  Created+Investigated by Mathias Sabl\'e Meyer
 
-%        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
-$0 $3)))))))
+  \vspace{0.5cm}
 
-      \end{tabular}
-    %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
-    %% (lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) ($1 ($3 $0)) ($4 $0) ($5 $0)))))))))
-    %% (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (#(lambda (lambda (fix1 $0 (lambda (lambda (if (empty?
-$0) empty (cons ($3 (car $0)) ($1 (cdr $0))))))))) (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
-$0 $3)))))))
-    \\\bottomrule 
-    \end{tabular}
-  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
-  \end{figure}
+  \includegraphics[width = 10cm]{turtleFigure.png}
 
-We believe that program learners should \emph{not}
-start from scratch,
-but instead should start from
-a rich, domain-agnostic
-basis like those embodied in the standard libraries of modern  languages.
-What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
-and can in principle recover many of the amenities of modern programming systems,
-provided it is given enough computational power and a suitable
-spectrum of tasks.% programming exercises.
+  \end{frame}
 
-\subsection{Learning Generative Models}
+\begin{frame}{Conclusion}
 
-We apply \system to learning generative models for images and text (Fig.~\ref{geomCompiled}-\ref{regularExpressions}).
-For images, we learn programs controlling a simulated ``pen,''
-and the task is to look at an image and explain it in terms of a graphics program.
-For text, we learn probabilistic regular expressions -- a simple probabilistic program for which inference is always tractable -- and the task is to
-infer a regex from a collection of strings.
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{figures/geomCompiled.eps} 
-  \caption{
-  Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: \% holdout testing tasks solved (middle); on the sides are averaged samples from the DSL
-  before any training (left) and after last iteration (right).}\label{geomCompiled}
+  \textbf{Takeaways:}
+  \begin{itemize}
+  \item Language of thought models can be engineered with fewer prior assumptions and less hand engineering
+    \item Humans can flexibly adapt to new problem domains --- DreamCoder takes a step in this direction
+  \end{itemize}
+  \pause
 
-  \end{figure}
-\begin{figure}
-\centering
-\tiny
-\tabcolsep=2pt
-\vspace{0pt}
-\begin{tabular}{cc}
-\begin{tabular}{ccc}
-\multicolumn{3}{c}{Tasks:}\\
-\hline
-cut&F &Moss Side \\ 
-control&CL &Burnage \\
-control&F & City Centre\\
-cut& PCFL& Brooklands \\
-\multicolumn{3}{c}{Learned generative models:}\\
-\hline
-\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash
-u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash
-u\textbackslash w*)*} \\
-\multicolumn{3}{c}{Samples from synthesized generative models:}\\
-\hline
-ya&DQDF & Vr DR \\ 
-glrwfdcnc&F &BeF lKQ \\
-mgs&F& W\\
-piljnl& KI& kqBfZ 0 \\
-kj& F& ON\\
-zci& GL& Bttc \\
-sxpm& F& S \\
-\hline
-\\
-\end{tabular} 
-&
-\begin{tabular}{c}
-\begin{tabular}{l}
-\multicolumn{1}{c}{Learned DSL:}\\
-\hline
-$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
-$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash
-u\textbackslash w*)*}\\
-$f_3(x) = f_2(\scode{space}) = $\scode{( \textbar\textbackslash
-u\textbackslash w*)*}\\
-\hline
-$f_4(x) = $ \scode{($x$*$x$)}\\
-\emph{ (equivalent to regex `plus')}\\
-$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash
-l*\textbackslash l}$\\
-\hline
-\end{tabular}
-\\
-\\
-\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
-\end{tabular}
-\end{tabular}
-\vspace{-.3cm}
-\caption{regex stuff}
-\end{figure}
+  \textbf{Future work}
 
-\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
-We evaluate  on held-out testing tasks,
-measuring how many
-tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
-Prior to any learning,
-the system cannot find solutions for most of the tasks,
-and those it does solve take a long time;
-with more wake/sleep iterations,
-we converge upon DSLs and recognition models 
-more closely matching the domain.
+  \begin{itemize}
+  \item A real solution to the problem of searching in the language of thought: recognition model can only go so far
+    \pause
+    \item Generative modeling via probabilistic programs: input/output mappings do not capture many of our concepts
+    \end{itemize}
 
-%% We compare with ablations of our model on held out tasks.
-%% The purpose of this ablation study is 
-%% both to examine the role of each component of \systemEnding,
-%% as well as to compare with
-%% prior approaches in the literature:
-%% a head-to-head
-%% comparison of program synthesizers is complicated by the fact that
-%% each system, including ours, makes idiosyncratic 
-%% assumptions about the space of programs and the statement of tasks.
+  \end{frame}
 
-%% Nevertheless, much prior work can be modeled within our setup. 
-%% We compare with the following ablations (Tbl~\ref{baselineComparisons};
-%% Fig~\ref{learningCurves}):
-%% \\\noindent \textbf{No NN:} lesions the recognition model.
-%% \\\noindent \textbf{NPS}, which does not learn the DSL,
-%% instead learning the recognition model
-%% from samples drawn from the fixed DSL.
-%% We call this NPS (Neural Program Synthesis)
-%% because this is closest to how
-%% RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-%% \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-%% only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
-This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
-%% \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-%% but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-%% \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
-%For each domain,
-%% We are interested both in how many tasks the
-%% agent can solve and how quickly it can find those solutions.
-%% Tbl.~\ref{baselineComparisons}
-%% compares our model against these alternatives.
-%% We consistently
-%% improve on the baselines,
-%% and find that lesioning the recognition model
-%% and lesioning it also slows down the convergence of the algorithm,
-%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
-%% This supports a view of the recognition model as a way of amortizing the cost of search.
-
-\begin{figure*}\centering
-  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
-  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
-  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
-\vspace{-0.4cm}  \caption{Learning curves for \system both with (\orange{in orange}) and without
-    (\teal{in teal}) the recognition model.
-Solid lines: \% holdout testing tasks solved w/ 10m timeout.
-Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}\vspace{-0.5cm}
-\end{figure*}
-
-\section{Discussion}
-
-%\paragraph{Outlook.}
-We contribute an algorithm, \systemEnding, that learns to program by
-bootstrapping a DSL with new domain-specific primitives that the algorithm
-itself discovers, together with a neural recognition model that learns to
-efficiently deploy the DSL on new tasks. %% We believe this integration of top-down
-%% symbolic representations and bottom-up neural nets --- both of them learned
-%% --- can help make program induction systems more generally useful for AI. 
-%Many
-%directions remain open.
-%\paragraph{Future.}
-Two immediate future goals are to integrate more sophisticated neural recognition
-models~\cite{devlin2017robustfill} and program
-synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the generic methods used here.
-Another direction is DSL meta-learning: can we find a
-\emph{single} universal primitive set that could bootstrap DSLs for
-new domains, including the domains considered here,  but also many others?
-
-\bibliography{main}
-\bibliographystyle{icml2018}
 \end{document}

### 2018-07-19 talk.tex
metropolis
more talk stuff

@@
 \documentclass{beamer}
+\usetheme{metropolis}
 \usepackage{listings}
@@
 \title{Growing libraries of concepts with wake-sleep program induction}
-\author{Kevin Ellis \& Mathias Sabl\'e Meyer\\Joint work with: Lucas Morales, Armando Solar-Lezama, Joshua B. Tenenbaum\\Heavy inspiration from: Eyal Dechter}
+\author{Kevin Ellis \& Mathias Sabl\'e Meyer\\Joint with: Lucas Morales, Armando Solar-Lezama, Joshua B. Tenenbaum\\Heavy inspiration from: Eyal Dechter}
 \institute{MIT} 
@@
   \centering
-  \begin{tikzpicture}
-    \node[latent,align=center] (g) {DSL: \\concept library};
-    \node[latent,below = of g,xshift = -2cm] (e1) {$\text{program}_1$};
-    \node[latent,below = of g,xshift = 0cm] (e2) {$\text{program}_2$};
- \node[latent,below = of g,xshift = 2cm] (e3) {$\text{program}_3$};
-    \node[obs,below = of e1] (x1) {$\text{task}_1$};
-    \node[obs,below = of e2] (x2) {$\text{task}_2$};
-    \node[obs,below = of e3] (x3) {$\text{task}_3$};
-    \edge {g} {e1,e2,e3};
-    \edge {e1} {x1};
-    \edge {e2} {x2};
-    \edge {e3} {x3};
+
+\begin{tikzpicture}[scale=2,line width=1.5mm]
+
+  \node[latent,scale=2,align=center] at (3.5,3) (dx){DSL};
+  \node[ellipse,fill=white,draw=black,inner sep=1pt,
+minimum size=20pt, font=\fontsize{10}{10}\selectfont, node distance=1,scale=2] at ([xshift=2cm]dx) (zp){program};
+  \node[obs,scale=2] at ([xshift=2cm]zp) (xp) {task};
+  \plate {}{(zp)(xp)}{};
+%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
+  %% \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
+%  \draw [->,red] (tx.west) -- (zp.east);
+  \draw [->] (dx.east) -- (zp.west);
+  \draw [->] (zp.east) -- (xp.west);
+
   \end{tikzpicture}
+ %%  \begin{tikzpicture}
+ %%    \node[latent,align=center] (g) {DSL: \\concept library};
+ %%    \node[latent,below = of g,xshift = -2cm] (e1) {$\text{program}_1$};
+ %%    \node[latent,below = of g,xshift = 0cm] (e2) {$\text{program}_2$};
+ %% \node[latent,below = of g,xshift = 2cm] (e3) {$\text{program}_3$};
+ %%    \node[obs,below = of e1] (x1) {$\text{task}_1$};
+ %%    \node[obs,below = of e2] (x2) {$\text{task}_2$};
+ %%    \node[obs,below = of e3] (x3) {$\text{task}_3$};
+ %%    \edge {g} {e1,e2,e3};
+ %%    \edge {e1} {x1};
+ %%    \edge {e2} {x2};
+ %%    \edge {e3} {x3};
+ %%  \end{tikzpicture}
 
@@
   \end{itemize}
-  Based on Exploration-Compression algorithm by Eyal Dechter
-
-\pause
-  \vspace{1cm}
-
-  \centering\begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
-    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
+  Combines ideas from Wake-Sleep \& Exploration-Compression algorithm by Eyal Dechter
 

### 2018-07-19 talk.tex
beautiful table
Added figures

@@
     \end{tikzpicture}
+\end{frame}
+
+\begin{frame}[t]{EC}
+  \includegraphics[width = 11cm]{figures/animation/EC.eps}
+\end{frame}
+\begin{frame}[t]{DC}
+  \includegraphics[width = 11cm]{figures/animation/DC.eps}
+\end{frame}
+\begin{frame}[t]{Wake}
+  \includegraphics[width = 11cm]{figures/animation/Wake.eps}
+\end{frame}
+\begin{frame}[t]{Sleep-G}
+  \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
+\end{frame}
+\begin{frame}[t]{Sleep-R-1}
+  \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
+\end{frame}
+\begin{frame}[t]{Sleep-R-2}
+  \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
+
 \end{frame}

### 2018-07-20 talk.tex
labels for sections

@@
 \usetikzlibrary{decorations.markings}
+
+\newcommand\Wider[2][3em]{%
+\makebox[\linewidth][c]{%
+  \begin{minipage}{\dimexpr\textwidth+#1\relax}
+  \raggedright#2
+  \end{minipage}%
+  }%
+}
 
@@
 
-\begin{frame}[t]{EC}
-  \includegraphics[width = 11cm]{figures/animation/EC.eps}
-\end{frame}
-\begin{frame}[t]{DC}
-  \includegraphics[width = 11cm]{figures/animation/DC.eps}
-\end{frame}
-\begin{frame}[t]{Wake}
-  \includegraphics[width = 11cm]{figures/animation/Wake.eps}
-\end{frame}
-\begin{frame}[t]{Sleep-G}
-  \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
-\end{frame}
-\begin{frame}[t]{Sleep-R-1}
-  \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
-\end{frame}
-\begin{frame}[t]{Sleep-R-2}
-  \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
-
-\end{frame}
-
 \begin{frame}{Growing a domain-specific language of thought}
@@
 \begin{frame}{DSL: Library of concepts}
-\includegraphics[width = 11cm]{3DSL.png}
-\end{frame}
-
-\begin{frame}{DSL learning as Bayesian inference}
-  \centering
-
-\begin{tikzpicture}[scale=2,line width=1.5mm]
-
-  \node[latent,scale=2,align=center] at (3.5,3) (dx){DSL};
-  \node[ellipse,fill=white,draw=black,inner sep=1pt,
-minimum size=20pt, font=\fontsize{10}{10}\selectfont, node distance=1,scale=2] at ([xshift=2cm]dx) (zp){program};
-  \node[obs,scale=2] at ([xshift=2cm]zp) (xp) {task};
-  \plate {}{(zp)(xp)}{};
-%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
-  %% \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
-%  \draw [->,red] (tx.west) -- (zp.east);
-  \draw [->] (dx.east) -- (zp.west);
-  \draw [->] (zp.east) -- (xp.west);
-
-  \end{tikzpicture}
- %%  \begin{tikzpicture}
- %%    \node[latent,align=center] (g) {DSL: \\concept library};
- %%    \node[latent,below = of g,xshift = -2cm] (e1) {$\text{program}_1$};
- %%    \node[latent,below = of g,xshift = 0cm] (e2) {$\text{program}_2$};
- %% \node[latent,below = of g,xshift = 2cm] (e3) {$\text{program}_3$};
- %%    \node[obs,below = of e1] (x1) {$\text{task}_1$};
- %%    \node[obs,below = of e2] (x2) {$\text{task}_2$};
- %%    \node[obs,below = of e3] (x3) {$\text{task}_3$};
- %%    \edge {g} {e1,e2,e3};
- %%    \edge {e1} {x1};
- %%    \edge {e2} {x2};
- %%    \edge {e3} {x3};
- %%  \end{tikzpicture}
-
-\textbf{[Dechter et al., 2013]}  [Liang et al, 2010]; [Lake et al, 2015]
-
-  \end{frame}
-
-\begin{frame}{Amortized program inference}
-
-  \includegraphics[width = 9cm]{rm.png}
-
-  \begin{itemize}
-  \item $\mathcal{D}$: DSL
-  \item $p_n$: program
-  \item $x_n$: task
-  \item $Q$: Recognition model
-    \end{itemize}
-
+\Wider[5.4em]{
+  \renewcommand\code\texttt
+  \renewcommand\codechar[1]{\texttt{"#1"}}
+  \newcommand{\helpSize}{0.25cm}
+  \footnotesize\begin{tabular}{cc}
+    \toprule
+Tasks and     \pop{Programs} & DSL\\
+    \midrule
+      \begin{tabular}{cc}
+        \begin{tabular}{c}
+          \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+          \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
+          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
+          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
+          \\
+          \\
+          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
+          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
+          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
+        \end{tabular}
+        &
+        \hspace{-0.3cm}\begin{tabular}{c}
+          \code{[7\, 3]}$\to $\code{False}                              \\
+          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
+          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
+          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
+          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
+          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
+        \end{tabular}
+      \end{tabular}
+    &
+    \begin{tabular}{l}
+      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
+      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
+      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
+      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
+      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
+      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
+      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
+      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
+      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
+      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
+      \phantom{$f_1(\ell,$}
+      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
+      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
+    \end{tabular}
+  \\\bottomrule
+  \end{tabular}
+}
 \end{frame}
@@
   Combines ideas from Wake-Sleep \& Exploration-Compression algorithm by Eyal Dechter
+
+\end{frame}
+
+\begin{frame}[t]{DSL learning as Bayesian inference}
+  \includegraphics[width = 11cm]{figures/animation/EC.eps}
+
+\vspace{1cm}
+
+  \textbf{[Dechter et al., 2013]}  [Liang et al, 2010]; [Lake et al, 2015]
+
+  \vfill
+  Gray: Observed.\\
+  White: Latent.\\
+  Boxed (plate): Repeated.\\
+
+\end{frame}
+\begin{frame}[t]{DSL learning as \alert{amortized} Bayesian inference}
+  \includegraphics[width = 11cm]{figures/animation/DC.eps}
+
+  \vfill
+  Recognition model: Neural net.\\
+  Red: Bottom-up inference.\\
+  Gray: Observed.\\
+  White: Latent.\\
+  Boxed (plate): Repeated.\\
+
+\end{frame}
+\begin{frame}[t]{Wake: Problem solving}
+  \includegraphics[width = 11cm]{figures/animation/Wake.eps}
+
+    \vfill
+  Recognition model: Neural net.\\
+  Red: Bottom-up inference.\\
+  Gray: Observed.\\
+  White: Latent.\\
+  Boxed (plate): Repeated.\\
+
+\end{frame}
+\begin{frame}[t]{Sleep-G: Memory consolidation}
+  \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
+  \vfill
+  \textbf{Fragment Grammars: O'Donnell 2015.}\\
+  Orange: Code fragments.\\
+  %% Gray: Observed.\\
+  %% White: Latent.\\
+  %% Boxed (plate): Repeated.\\
+
+\end{frame}
+\begin{frame}[t]{Sleep-R: Experience replay}
+  \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
+\end{frame}
+\begin{frame}[t]{Sleep-R: Dreaming}
+  \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
 

### 2018-07-20 talk.tex
PDF

@@
 
-\begin{frame}{Wake phase}
-  \centering\begin{tikzpicture}[scale=1.3]
-    \node at (0,0) (d){DSL};
-    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
-
-    \node at ([xshift = 2cm]t) (nn){
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
-    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
-
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
-    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
-
-    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
-
-    \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
-    \draw [->  ] (s.east) -- (f.west);
-
-    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
-
-    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
-
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}{Sleep-G Phase}
-  Fragment grammar (\textbf{O'Donnell 2015})
-  \vspace{1cm}
-
-  \centering\begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
-    \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
-    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
-
-    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
-    \node[anchor=north] at (c.north) {Compression};
-
-    \draw [-> ] (f1.east) -- (c.west|-f1.east);
-    \draw [-> ] (f2.east) -- (c.west|-f2.east);
-    \draw [-> ] (f3.east) -- (c.west|-f3.east);
-
-    \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
-    \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
-    \draw [-> ] (c.east) -- (d.west);
-    \draw [-> ] (c.east) -- (t.west);
-
-    \node at (c.center) {
-\begin{tikzpicture}[scale=0.7]
-    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
-    %%     \footnotesize Program:\\
-    %%     \code{($\lambda$ (x) (+ (- x) 1))}
-    %% \end{tabular}};
-    %\node at (,0.5) {\code{cons}};
-%    \node [rotate=90] at (-2.3,-0.5) {\small program};
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-%    \draw  (l1.south) -- (-0.5,0.5);
-
-  %% \node(c) at (-0.5,-1.5) {\code{-}};
-  %% \node(z) at (0.5,-1.5) {\code{x}};
-
-  %% \draw (0,-1) -- (c.north);
-  %% \draw (0,-1) -- (z.north);
-
-  \begin{scope}[shift={(-1,-2.5)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-%  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-  \end{scope}
-
-\begin{scope}[shift={(0,-5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  %\node(x1) at (0,-1) {};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.55,-0.4) -- (a.north);
-%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
-
-  \end{scope}
-
-\end{tikzpicture}
-    };
-
-    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};
-
-    \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}{Sleep-R Phase}
-Train recognition model on two sources of data:
-  \begin{itemize}
-  \item Samples from generative model
-    \item Programs found during waking
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Sleep-R: Train on samples from generative model}
-\centering  \begin{tikzpicture}[scale=1.2]
-    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
-    \node at ([xshift = 3cm]d.east) (p2){program};
-    \node at ([yshift = 1.5cm]p2) (p1){program};
-    \node at ([yshift = -1.5cm]p2) (p3){program};
-
-    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
-    \draw[squiggle,-> ] (d.east) -- (p1.west);
-    \draw[squiggle,-> ] (d.east) -- (p3.west);
-
-    \node at ([xshift = 2cm]p1.east) (t1){task};
-    \node at ([xshift = 2cm]p2.east) (t2){task};
-    \node at ([xshift = 2cm]p3.east) (t3){task};
-    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
-    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
-    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}{Sleep-R: Train on programs found during waking}
-\centering  \begin{tikzpicture}[scale=1.2]
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
-
-    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
-    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
-    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
-
-    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
-    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
-    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
-
-    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
-    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
-    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
-  \end{tikzpicture}
-
-\end{frame}
-
 \begin{frame}{List functions}
@@
 
-\begin{frame}{Text editing}
-  In the style of FlashFill (Gulwani 2012)
+\begin{frame}{List functions: Learning curves on hold out tasks}
 
-  \centering  \includegraphics[width = 5cm]{textColumn.png}
+  \begin{center}
+      \includegraphics[width = 6.5cm]{figures/listLearningCurve.eps} 
+    \end{center}
 
-  SyGuS problems: solves 3\% before learning, vs 74\% after learning. Best prior work: 80\%
+Learning curves for DreamCoder both with (\orange{in orange}) and without
+    (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved w/ 10m timeout. Dashed lines: Average solve time, averaged only over tasks that are solved.
 
@@
 
-\begin{frame}{Symbolic regression from visual input}
-\centering\includegraphics[width = 5cm]{symbolicRegression.png}
-\end{frame}
+%% \begin{frame}{Text editing}
+%%   In the style of FlashFill (Gulwani 2012)
+
+%%   \centering  \includegraphics[width = 5cm]{textColumn.png}
+
+%%   SyGuS problems: solves 3\% before learning, vs 74\% after learning. Best prior work: 80\%
+
+%%   \end{frame}
+
+%% \begin{frame}{Symbolic regression from visual input}
+%% \centering\includegraphics[width = 5cm]{symbolicRegression.png}
+%% \end{frame}
 
@@
 
-\begin{frame}{Conclusion}
+\begin{frame}{Contributions}
 
-  \textbf{Takeaways:}
+  \textbf{Takeaway:}
   \begin{itemize}
-  \item Language of thought models can be engineered with fewer prior assumptions and less hand engineering
-    \item Humans can flexibly adapt to new problem domains --- DreamCoder takes a step in this direction
+    \item Humans flexibly adapt to diverse sets of new problem domains --- DreamCoder takes a step in this direction
   \end{itemize}
@@
   \begin{itemize}
-  \item A real solution to the problem of searching in the language of thought: recognition model can only go so far
+  \item More human-like learning: intelligently composing new tasks, 
     \pause
-    \item Generative modeling via probabilistic programs: input/output mappings do not capture many of our concepts
+  \item Theory learning
+    \item Planning
     \end{itemize}

### 2018-07-20 talk.tex
Turtle part of the slides

--- 2018-07-19 talk.tex
@@
 \usepackage{verbatim}
-\usepackage{listings}
 %\usepackage{tipa}
@@
 
+\newcommand\Wider[2][3em]{%
+\makebox[\linewidth][c]{%
+  \begin{minipage}{\dimexpr\textwidth+#1\relax}
+  \raggedright#2
+  \end{minipage}%
+  }%
+}
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage{fancyvrb}
+
+\usepackage[most]{tcolorbox}
+\definecolor{block-gray}{gray}{0.10}
+\newtcolorbox{mycode}{colback=block-gray,grow to right by=0mm,grow to left
+by=0mm, boxrule=0pt,boxsep=0pt,breakable,fontupper=\color{white}}
 
@@
 \begin{frame}{DSL: Library of concepts}
-\includegraphics[width = 11cm]{3DSL.png}
+\Wider[5.4em]{
+  \renewcommand\code\texttt
+  \renewcommand\codechar[1]{\texttt{"#1"}}
+  \newcommand{\helpSize}{0.25cm}
+  \footnotesize\begin{tabular}{cc}
+    \toprule
+Tasks and     \pop{Programs} & DSL\\
+    \midrule
+      \begin{tabular}{cc}
+        \begin{tabular}{c}
+          \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+          \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
+          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
+          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
+          \\
+          \\
+          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
+          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
+          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
+        \end{tabular}
+        &
+        \hspace{-0.3cm}\begin{tabular}{c}
+          \code{[7\, 3]}$\to $\code{False}                              \\
+          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
+          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
+          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
+          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
+          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
+        \end{tabular}
+      \end{tabular}
+    &
+    \begin{tabular}{l}
+      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
+      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
+      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
+      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
+      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
+      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
+      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
+      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
+      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
+      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
+      \phantom{$f_1(\ell,$}
+      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
+      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
+    \end{tabular}
+  \\\bottomrule
+  \end{tabular}
+}
 \end{frame}
@@
 \begin{frame}{Turtle graphics}
-  Created+Investigated by Mathias Sabl\'e Meyer
+
+  \begin{block}{DSL}
+    \texttt{FW x | RT x | UP | DOWN | SET state}
+  \end{block}
+
+  \begin{block}{Tasks}
+    \texttt{unit -> image}
+  \end{block}
 
@@
 
-  \begin{itemize}
-  \item pen up/pen down
-  \item move pen (forward+rotate)
-  \item get/set pen state
-  \item `for' loops
-  \end{itemize}
+  \only<1>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{FW 1}\\
+          \vspace*{5\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/line.eps}
+    \end{columns}
+  }
+  \only<2>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{FW 1}\\
+          \texttt{RT} $\frac\pi2$\\
+          \texttt{FW 1}\\
+          \vspace*{3\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/angle.eps}
+    \end{columns}
+  }
+  \only<3>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{FW 1}\\
+          \texttt{RT} $\frac\pi2$\\
+          \texttt{FW 1}\\
+          \texttt{RT} $\frac\pi2$\\
+          \texttt{FW 1}\\
+          \vspace*{1\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/angle2.eps}
+    \end{columns}
+  }
+  \only<4>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{for i in range(4)}\\
+          \texttt{> FW 1}\\
+          \texttt{> RT} $\frac\pi2$\\
+          \vspace*{3\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/square.eps}
+    \end{columns}
+  }
+  \only<5>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{for i in range(8)}\\
+          \texttt{> FW 1} \\
+          \texttt{> SET origin} \\
+          \texttt{> RT} $\frac{2\pi}{8}$\\
+          \vspace*{2\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/star.eps}
+    \end{columns}
+  }
+  \only<6>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{for i in range(8)}\\
+          \texttt{> PU} \\
+          \texttt{> FW} $\frac{\text{\texttt{i}}}{2}$\\
+          \texttt{> PD} \\
+          \texttt{> FW} $\frac{\text{\texttt{i}}}{2}$\\
+          \texttt{> RT} $\frac\pi2$\\
+          \vspace*{0\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/spiral.eps}
+    \end{columns}
+  }
+\end{frame}
 
+\begin{frame}{Turtle graphics}
   \centering
-  \includegraphics[width = 8cm]{exampleGraphicsPrograms.png}
-
-\end{frame}
-\begin{frame}{Turtle graphics}
-  Created+Investigated by Mathias Sabl\'e Meyer
-
-  \vspace{0.5cm}
-
-  \includegraphics[width = 10cm]{turtleFigure.png}
-
+  \includegraphics[width=11cm]{figures/tasksMinusBehaviour.png}
 \end{frame}

### 2018-07-21 talk.tex
small edits to slides

--- 2018-07-20 talk.tex
@@
 \end{frame}
+\begin{frame}[t]{Sleep-R: Objective}
+  \includegraphics[width = 11cm]{figures/animation/Wake.eps}
+
+    \vfill
+    Recognition model predicts distribution over program, conditional on task.\\
+    Training: (program, task) pairs\\
+    Objective: Predict program w/ (1) high prior under DSL \& \\(2) high likelihood for task
+
+\end{frame}
 \begin{frame}[t]{Sleep-R: Experience replay}
   \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
+
+    \textbf{Train on (program, task) pairs found during waking}
 \end{frame}
@@
   \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
+
+    \textbf{Train on (program, task) pairs sampled from DSL}
 

### 2018-07-21 talk.tex
Minor tweaking

--- 2018-07-20 talk.tex
@@
 
-\begin{frame}{List functions}
-  Created+investigated by Lucas Morales
+\begin{frame}{List functions --- \small{Created \& investigated by Lucas
+  Morales}}
 
@@
 
-\begin{frame}{Turtle graphics}
+\begin{frame}{Turtle graphics --- \small{Created \& investigated by Mathias
+  Sablé--Meyer}}
 
   \begin{block}{DSL}
-    \texttt{FW x | RT x | UP | DOWN | SET state}
+    \texttt{OP ::= FW x | RT x | UP | DOWN | SET state}
   \end{block}
@@
   \begin{block}{Tasks}
-    \texttt{unit -> image}
+    \texttt{task : unit -> image}
   \end{block}
 
-  \vspace{0.5cm}
+  \vspace{0.2cm}
 
@@
   }
+  \only<7>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{for i in range($\infty$)}\\
+          \texttt{> FW $\varepsilon$}\\
+          \texttt{> RT} $\varepsilon$\\
+          \vspace*{3\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/circle.eps}
+    \end{columns}
+  }
+  \only<8-9>{
+    \begin{columns}[T]
+      \column{0.5 \textwidth}
+        \begin{mycode}
+          \texttt{for i in range($5\times\infty$)}\\
+          \texttt{> FW $\text{\texttt{i}}\times\varepsilon$}\\
+          \texttt{> RT} $\varepsilon$\\
+          \vspace*{3\baselineskip}
+        \end{mycode}
+      \column{0.5 \textwidth}
+        \includegraphics[width = 4cm]{figures/teachLogo/sspiral.eps}
+    \end{columns}
+  }
+  \visible<9>{
+    \vspace{0.2cm}
+    \texttt{\alert{NUM} ::= 1 | $\pi$ | $\infty$ | $\varepsilon$ | + | - | * | / }
+  }
 \end{frame}
 
-\begin{frame}{Turtle graphics}
+\begin{frame}{Turtle graphics --- Training tasks}
   \centering

### 2018-07-21 misc2.tex
misc2.c

--- 2018-07-10 poster.tex
@@
 \begin{alertblock}{Wake/Sleep DSL Induction}
+  \flushleft
   \textbf{Domain Specific Language (DSL):}
@@
 $$
-\begin{equation*}
+\vspace{1cm}
+$$
 \underbrace{\mathcal{D}^* = \argmax_{\mathcal{D}}    \probability[\mathcal{D}]\prod_n\sum_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}]}_{\textbf{\text{Sleep-G}}}
-  \end{equation*}
+$$
 
@@
   Simple $Q$: just predicts probabilities of DSL productions.  Goal: learn to invert generative model
+
 $$\underbrace{\min_Q \text{KL}\left(\probability[p|x,\mathcal{D}] || Q(p|x)\right) }_{\text{\textbf{Sleep-R}}}$$
@@
 
-\begin{column}{\twocolwid} % The first column within column 2 (column 2.1)
+\begin{column}{\threecolwid} % The first column within column 2 (column 2.1)
 
@@
 
-\begin{block}{Why this works: Bootstrapping}
-
-\begin{figure}
-\hspace*{-1.5cm}  \begin{tikzpicture}[scale=3,line width=1.5mm]
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
-    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
-
-    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
-  \end{scope}
-    \end{tikzpicture}
- % \caption{\system solves for programs, the DSL, and a recognition model.
-Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
-\end{figure}
-%% \begin{itemize}
-%% \item Search finds new programs$\implies$ DSL+recognition model get more data
-%% \item DSL improves$\implies$easier search, recognition model gets better data
-%%   \item Recognition model improves$\implies$ easier search
-%%   \end{itemize}
-
-\end{block}
-
-\begin{block}{Generative models of text}
-  \begin{figure}
-\centering
-%\tiny
-\tabcolsep=12pt
-\begin{tabular}{ccc}\toprule 
-\multicolumn{3}{c}{Three Tasks}\\
-\hline
-1.14531&F&110.9 \\ 
-?&CL&163.2 \\
-1.29857&F&207.3\\
-?&PCFL&143.3 \\\\\midrule 
-\multicolumn{3}{c}{Three learned generative models}\\
-\hline
- \pop{\scode{?\textbar(1.\textbackslash d+)}}& \pop{\scode{((\textbackslash u\textbackslash u)*)\textbar F}}& \pop{\scode{\textbackslash d\textbackslash d\textbackslash d.\textbackslash d}} \\\\\midrule
- \multicolumn{3}{c}{Samples from generative models}\\
-\hline
-1.61&DQDF& 343.8 \\ 
-?&F&241.2\\
-?&F&647.5\\
-1.2&KI&246.8%% \\
-%% 1.5080987&F&728.6\\
-%% ?&GL&029.3\\
-%% 1.453&F&289.1\\
-%%     \hline
-\\
-
-\midrule \multicolumn{3}{c}{Learned DSL}\\\midrule
-\multicolumn{3}{l}{\popp{\scode{$f_1 = $ \textbackslash u\textbackslash w*}}
-\hspace{3cm}\popp{\scode{$f_2(x) = $($x$|$f_1$)*}}}\\% $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}}\\
-\multicolumn{3}{l}{\popp{\scode{$f_3 = f_2(\scode{`` ''}) = $( \textbar\textbackslash u\textbackslash w*)*}}}\\
-\multicolumn{3}{l}{\hspace{1cm}\emph{($f_3$: whitespace delimited words)}}\\
-\multicolumn{3}{l}{\popp{$f_4(x) = $ \scode{($x$*$x$)}}\hspace{1cm}\emph{($f_4$: regex `plus')}}
-\end{tabular}
-\vspace{-.3cm}
-%\caption{regex stuff}
-\end{figure}
-
-\end{block}
-\begin{block}{Learning from Scratch}
-  Start w/ McCarthy 1959 Lisp: recursion, conditionals, lists.
-  Train on 22 programming exercises.
-  After 93 hours on 64 CPUs, rediscovers: \code{map},
-  \code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length}, \code{range}, \code{incr}, \code{decr}.
-
-\end{block}
-
 %% \begin{block}{References}

### 2018-07-22 misc2.tex
improvements to poster

@@
 
-\title{\huge \systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning} % Poster title
+\title{\huge \systemEnding: Growing Libraries of Concepts with Wake-Sleep Program Induction} % Poster title
 
-\author{Kevin Ellis, Lucas Morales, Mathias Sabl\'e Meyer, Maxwell Nye, Luke Hewitt,\\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
+\author{Kevin Ellis, Mathias Sabl\'e Meyer, Lucas Morales, Maxwell Nye, Luke Hewitt,\\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
 
@@
   \flushleft
-  \textbf{Domain Specific Language (DSL):}
-  A finely-tuned program representation, specialized to a domain of programming tasks.
-  Prior work in program learning  largely uses hand-engineered DSLs.
+  Goal: acquire domain-specific knowledge needed to induce a class of programs
+  \begin{itemize}
+  \item Declarative knowledge: library of concepts, `DSL' (domain specific language)
+    \item Procedural knowledge: knowing how and when to use those concepts; search strategy
+    \end{itemize}
 
-  \textbf{Approach:}  \system algorithm,
-  which bootstraps a learned DSL while jointly training
-  a neural net to search for programs in the learned DSL.
-Given a few hundred programming tasks, alternatingly:
+%%   \system algorithm,
+%%   which bootstraps a learned DSL while jointly training
+%%   a neural net to search for programs in the learned DSL.
+%% Given a few hundred programming tasks, alternatingly:
   \begin{itemize}
@@
 
-\begin{tikzpicture}[scale=4,line width=1.5mm]
+    \includegraphics[width = 20cm]{figures/animation/DC.eps}
 
-  \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
-  \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
-  \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
-  \plate {}{(zp)(xp)}{};
-%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
-  \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
-%  \draw [->,red] (tx.west) -- (zp.east);
-  \draw [->] (dx.east) -- (zp.west);
-  \draw [->] (zp.east) -- (xp.west);
+%% \begin{tikzpicture}[scale=4,line width=1.5mm]
 
-  \node at (nn) {
-    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
+%%   \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
+%%   \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
+%%   \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
+%%   \plate {}{(zp)(xp)}{};
+%% %  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
+%%   \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
+%% %  \draw [->,red] (tx.west) -- (zp.east);
+%%   \draw [->] (dx.east) -- (zp.west);
+%%   \draw [->] (zp.east) -- (xp.west);
+
+%%   \node at (nn) {
+%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
+%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%       \node[rectangle] at (1,1) {};
+%%       \foreach \name / \y in {1,...,4}
+%%           \node[neuron] (I-\name) at (0,-\y) {};
+%%       \foreach \name / \y in {1,...,3}
+%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%       \foreach \name / \y in {1,...,4}
+%%           \node[neuron] (O-\name) at (2,-\y) {};
+%%       \foreach \source in {1,...,4}
+%%           \foreach \dest in {1,...,3}
+%%               \draw [-latex] (I-\source) -- (H-\dest);
+%%       \foreach \source in {1,...,3}
+%%           \foreach \dest in {1,...,4}
+%%               \draw [-latex] (H-\source) -- (O-\dest);
+%%     \end{tikzpicture}
+%%   };
+%%   \node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+
+%%   \end{tikzpicture}
+
+  Observe \textbf{tasks} (program induction problems). 
+  Solve tasks w/ latent programs.
+  %% \textbf{Likelihood model} $\probability[x_n|p_n]$
+  %% scores program $p_n$ on task $x_n$
+  Latent  \textbf{DSL} acts as generative model over programs: `library of concepts'.
+  \textbf{Recognition model}: a neural net that predicts distribution over programs conditioned on task: `procedural knowledge'
+
+\end{block}
+
+\begin{block}{Wake: Problem solving}
+  \begin{tikzpicture}[scale=2.5]
+    \node at (0,0) (d){DSL};
+    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+
+    \node at ([xshift = 2cm]t) (nn){
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.4,shift={+(-1,2.5)}}]
         \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
@@
     };
-  \node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+    \node[align = center, text width = 1cm] at ([yshift = 0.6cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
+    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+%    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
+
+    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+    \node[align=left] at (7,-1) (f) {Programs};
+    \draw [->  ] (s.east) -- (f.west);
+
+    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+%    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
+
+ %   \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
 
@@
 
-  Observe $N$ \textbf{tasks}, written $\left\{x_n \right\}_{n = 1}^N$, each a program synthesis problem.
-
-  Solve task $x_n$ with latent program $p_n$
-
-  \textbf{Likelihood model} $\probability[x_n|p_n]$
-  scores program $p_n$ on task $x_n$
-
-  Latent  \textbf{DSL $\mathcal{D}$} acts as generative model over programs: $\probability[x|\mathcal{D}]$
-
-$$
-\underbrace{p_n^* =   \argmax_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}^*]}_{\text{\textbf{Wake}}}
-$$
-\vspace{1cm}
-$$
-\underbrace{\mathcal{D}^* = \argmax_{\mathcal{D}}    \probability[\mathcal{D}]\prod_n\sum_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}]}_{\textbf{\text{Sleep-G}}}
-$$
-
 \end{block}
 
-\begin{block}{Neural recognition model}
+\begin{block}{Sleep-R: Procedural knowledge}
 
-  Neural network $Q(p|x)$
+  Neural network
   predicts distribution over programs conditioned on tasks.
-  Simple $Q$: just predicts probabilities of DSL productions.  Goal: learn to invert generative model
-
-$$\underbrace{\min_Q \text{KL}\left(\probability[p|x,\mathcal{D}] || Q(p|x)\right) }_{\text{\textbf{Sleep-R}}}$$
+  Simple model: just predicts probabilities of DSL productions.  Goal: learn to invert generative model from (program, task) pairs
 
@@
   \begin{itemize}
-  \item \textbf{Samples (``Dreams'') from DSL}: Unlimited data, but only high-quality if generative model $\mathcal{D}$ is good. Like Helmholtz Machine's recognition model training. Loss:
-    \begin{equation*}
-\expect_{(p,x)\sim\mathcal{D} }\left[\log Q(p|x)\right]
-      \end{equation*}
-  \item \textbf{Self-Supervised}: $(x_n,p_n)$ pairs discovered during waking. Loss:
-    \begin{equation*}
- %\sum_{n = 1}^N\left[\sum_{(x_n,p_n)}
-  \frac{\probability\left[x_n,p_n|\mathcal{D}\right]}{\sum_{(x_n,p_n')}\probability\left[x_n,p_n|\mathcal{D} \right]}\log Q(p_n|x_n)%\right]
-      \end{equation*}
+  \item \textbf{``Dreaming'' from DSL}:
+
+    \includegraphics[width = 20cm]{figures/animation/Sleep-R-2.eps}
+  \item \textbf{Experience replay}: (program, task) pairs discovered during waking.
+  \includegraphics[width = 20cm]{figures/animation/Sleep-R-1.eps}    
   \end{itemize}
@@
 
-\begin{block}{Fragment Grammars: Inducing a DSL}
+\begin{block}{Sleep-G: Growing a DSL}
 Fragment grammars: introduced in computational linguistics~(O'Donnell~2015)

### 2018-07-22 misc2.tex
Improvements

--- 2018-07-21 misc2.tex
@@
   \flushleft
-  \textbf{Domain Specific Language (DSL):}
-  A finely-tuned program representation, specialized to a domain of programming tasks.
+  \textbf{DSL:}
+  Domain Specific Language, a finely-tuned program representation, specialized to a domain of programming tasks.
   Prior work in program learning  largely uses hand-engineered DSLs.
@@
   \textbf{Program representation:}
-  $\approx $Lisp; conditionals, variables, $\lambda$~abstraction
+  $\approx $~Lisp; conditionals, variables, $\lambda$~abstraction
 
@@
 
-\begin{tikzpicture}[scale=4,line width=1.5mm]
-
-  \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
-  \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
-  \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
-  \plate {}{(zp)(xp)}{};
-%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
-  \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
-%  \draw [->,red] (tx.west) -- (zp.east);
-  \draw [->] (dx.east) -- (zp.west);
-  \draw [->] (zp.east) -- (xp.west);
-
-  \node at (nn) {
-    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
-      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-      \node[rectangle] at (1,1) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (I-\name) at (0,-\y) {};
-      \foreach \name / \y in {1,...,3}
-          \node[neuron] (H-\name) at (1,-\y-0.5) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (O-\name) at (2,-\y) {};
-      \foreach \source in {1,...,4}
-          \foreach \dest in {1,...,3}
-              \draw [-latex] (I-\source) -- (H-\dest);
-      \foreach \source in {1,...,3}
-          \foreach \dest in {1,...,4}
-              \draw [-latex] (H-\source) -- (O-\dest);
-    \end{tikzpicture}
-  };
-  \node[shift={+(0,-1.7)}] at (nn) { $Q$  };
-
-  \end{tikzpicture}
+  \includegraphics{figures/animation/DC.eps}
 
@@
   \textbf{Likelihood model} $\probability[x_n|p_n]$
-  scores program $p_n$ on task $x_n$
+  scores program $p_n$ on task $x_n.
 
@@
   \begin{itemize}
-  \item \textbf{Samples (``Dreams'') from DSL}: Unlimited data, but only high-quality if generative model $\mathcal{D}$ is good. Like Helmholtz Machine's recognition model training. Loss:
+  \item \textbf{``Dreams'' from DSL}: Can sample unlimited data, quality
+    positively correlates with fitness of generative model $\mathcal{D}$. Like Helmholtz Machine's recognition model training. Loss:
     \begin{equation*}
@@
 
+\begin{columns}[t,totalwidth=\threecolwid] % Split up the two columns wide column again
+
+\begin{column}{\twocolwid} % The first column within column 2 (column 2.1)
+
 %----------------------------------------------------------------------------------------
@@
 
-\begin{block}{Generative models of images: Turtle/LOGO Graphics}
+\end{column}
+\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
 
-  \newcommand{\dreamSize}{5.5cm}
-  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{12 (of 136) Tasks}}\\\hline
-    \includegraphics[width = \dreamSize]{turtleTasks/square/trimmed/train_square_3_h.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/square/trimmed/train_square_4_h.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/snow1_h.png}\\
-    \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/snow2_random_1.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/hexa_vh.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/star_5_h.png}\\
-        \includegraphics[width = \dreamSize]{turtleTasks/six/trimmed/star_vh.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/spiral6_h.png}&
-        \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/star_7_h.png}\\
+\begin{block}{Bootstrapping}
 
-%    \includegraphics[width = \dreamSize]{turtleTasks/circle/trimmed/train_circle_10_h.png}\\
-    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/smooth_spiral4_h.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/los/trimmed/train_los_3_1_h.png}&
-    \includegraphics[width = \dreamSize]{turtleTasks/logoTasks/trimmed/iter2_h.png}\\\hline
-%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/1748505267_norm.png}&
-  \end{tabular}\hspace{4cm}%
-  \begin{tabular}{|lll|}\multicolumn{3}{c}{\textbf{9 samples from learned DSL}}\\\hline
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/4361002260.png}&
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/5809723215.png}&
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/5818934795_norm.png}\\
-%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/6027771181.png}\\
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/1950268419.png}&
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/6770710484.png}&
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/7045328293.png}\\
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/7623700519.png}&
-%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/8879622788.png}&
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/9150025422.png}&
-    \includegraphics[width = \dreamSize]{dreams/final/trimmed/9666602433.png}\\\hline
-%    \includegraphics[width = \dreamSize]{dreams/final/trimmed/1748505267_norm.png}&
-  \end{tabular}
+\begin{figure}
+\hspace*{-1.5cm}  \begin{tikzpicture}[scale=3,line width=1.5mm]
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
+    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
 
-  \vspace{2cm} \textbf{Left:}  Agent controls a `pen' -- tasked with drawing pictures. \textbf{Right:} During Sleep-R,
-   `dream' by sampling programs from learned DSL and rendering them.
+    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
+    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
+    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
+
+    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
+    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+  \end{scope}
+    \end{tikzpicture}
+ % \caption{\system solves for programs, the DSL, and a recognition model.
+Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
+\end{figure}
+%% \begin{itemize}
+%% \item Search finds new programs$\implies$ DSL+recognition model get more data
+%% \item DSL improves$\implies$easier search, recognition model gets better data
+%%   \item Recognition model improves$\implies$ easier search
+%%   \end{itemize}
+
+\end{block}
+
+\end{column}
+\end{columns}
+
+\begin{block}{Generative models of images: \texttt{Turtle}s all the way down}
+
+Agent controls a `pen' tasked with drawing pictures.
+
+  \includegraphics[width=20cm]{figures/poster2/sampleTasks.png}
 

### 2018-07-22 misc2.tex
improvements to talk and poster
Improving the LOGO part

@@
 
-Agent controls a `pen' tasked with drawing pictures.
+\begin{columns}[t,totalwidth=\threecolwid] % Split up the two columns wide column again
 
-  \includegraphics[width=20cm]{figures/poster2/sampleTasks.png}
+\begin{column}{\twocolwid} % The first column within column 2 (column 2.1)
+
+  \includegraphics[width=\twocolwid]{figures/dreams/montageLearningAndShapes.eps}
+
+  \textbf{Left column:} 16 out of 92 example tasks to hit in an all-or-none
+  fashion --- training set for the \system algorithm.
+  \textbf{Right column:} Learning curve for a run with a timeout of search time
+  per task of 5 minutes (blue, bottom) and of 120 minutes (red, top). On the
+  bottom are dreams, from left to right: before any training (original DSL, flat
+  priors), at the end of the 5-minutes-timeout run, and at the end of the
+  12-minutes-timeout run.
+
+\end{column}
+\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
+  The agent controls a `pen' on a 2D plant. Primitives are those of the LOGO
+  language, modified to have a functional programing structure.
+\end{column}
+\end{columns}
 

### 2018-07-22 talk.tex
vision slide

@@
   \vfill
+  \only<1>{
   Recognition model: Neural net.\\
@@
   Boxed (plate): Repeated.\\
+  }
+  \only<2>{
+  \begin{tikzpicture}[xscale=1.25,line width=1mm]
+    \node at (0,0) (d){DSL};
+    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+
+    \node at ([xshift = 2cm]t) (nn){
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.3,shift={+(-1,2.5)}}]
+       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[align = center, text width = 1cm] at ([yshift = 0.7cm]nn.north) {\baselineskip=0pt \Small Recog. model\par};
+    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+%    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
+
+    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+    \node[align=left] at (7,-1) (f) {Programs};
+    \draw [->  ] (s.east) -- (f.west);
+
+    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+  \end{tikzpicture}}
+  %% Recognition model: Neural net.\\
+  %% Red: Bottom-up inference.\\
+  %% Gray: Observed.\\
+  %% White: Latent.\\
+  %% Boxed (plate): Repeated.\\
 
@@
 
-\begin{frame}{Contributions}
+\begin{frame}{Vision}
 
@@
 
-  \textbf{Future work}
+  %% \textbf{Future work}
 
-  \begin{itemize}
-  \item More human-like learning: intelligently composing new tasks, 
-    \pause
-  \item Theory learning
-    \item Planning
-    \end{itemize}
+  %% \begin{itemize}
+  %%   \pause
+  %% \item Theory learning
+  %%   \item Planning
+  %%   \end{itemize}
 

### 2018-07-23 misc2.tex
changes to poster

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+s%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Jacobs Portrait Poster
@@
 
-    \includegraphics[width = 20cm]{figures/animation/DC.eps}
+\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
 
-%% \begin{tikzpicture}[scale=4,line width=1.5mm]
+\begin{tikzpicture}[scale=4,line width=1.5mm]
 
-%%   \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
-%%   \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
-%%   \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
-%%   \plate {}{(zp)(xp)}{};
-%% %  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
-%%   \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
-%% %  \draw [->,red] (tx.west) -- (zp.east);
-%%   \draw [->] (dx.east) -- (zp.west);
-%%   \draw [->] (zp.east) -- (xp.west);
+  \node[latent,scale=3] at (3.5,3) (dx){DSL};
+  \node[latent,scale=3] at ([yshift=-1.5cm]dx) (zp){prog};
+  \node[obs,scale=3] at ([yshift=-2cm]zp) (xp) {data};
+  \node[latent,scale=3] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=3] at ([xshift=2cm]xp) (xp1) {data};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+%  \node at (nn) {\NeuralNetwork{0.5}};
 
-%%   \node at (nn) {
-%%     \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
-%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-%%       \node[rectangle] at (1,1) {};
-%%       \foreach \name / \y in {1,...,4}
-%%           \node[neuron] (I-\name) at (0,-\y) {};
-%%       \foreach \name / \y in {1,...,3}
-%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
-%%       \foreach \name / \y in {1,...,4}
-%%           \node[neuron] (O-\name) at (2,-\y) {};
-%%       \foreach \source in {1,...,4}
-%%           \foreach \dest in {1,...,3}
-%%               \draw [-latex] (I-\source) -- (H-\dest);
-%%       \foreach \source in {1,...,3}
-%%           \foreach \dest in {1,...,4}
-%%               \draw [-latex] (H-\source) -- (O-\dest);
-%%     \end{tikzpicture}
-%%   };
-%%   \node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+  \node[latent,scale=3] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=3] at ([xshift=-2cm]xp) (xp1) {data};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+%  \node at (nn) {\NeuralNetwork{0.5}};
 
-%%   \end{tikzpicture}
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \node at (nn) {\NeuralNetwork{0.5}};
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
 
-  Observe \textbf{tasks} (program induction problems). 
-  Solve tasks w/ latent programs.
+  %\node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+
+  \end{tikzpicture}
+
+  Observe program induction problems, 
+  Solve  w/ latent programs.
   %% \textbf{Likelihood model} $\probability[x_n|p_n]$
@@
   Latent  \textbf{DSL} acts as generative model over programs: `library of concepts'.
-  \textbf{Recognition model}: a neural net that predicts distribution over programs conditioned on task: `procedural knowledge'
+{\color{red}  \textbf{Recognition model}}: a neural net that predicts distribution over programs conditioned on task: `procedural knowledge'
 
@@
 \begin{block}{Wake: Problem solving}
-  \begin{tikzpicture}[scale=2.5]
+  \begin{tikzpicture}[scale=2.5,line width=1.3mm]
     \node at (0,0) (d){DSL};
@@
     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-%    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
-
- %   \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
-
   \end{tikzpicture}
@@
 
-  Neural network
-  predicts distribution over programs conditioned on tasks.
-  Simple model: just predicts probabilities of DSL productions.  Goal: learn to invert generative model from (program, task) pairs
+  %% Neural network
+  %% predicts distribution over programs conditioned on tasks.
+  %% Simple model: just predicts probabilities of DSL productions.  
+  Goal: learn to invert generative model from (program, task) pairs
 

### 2018-07-23 misc2.tex
More changes to the poster

@@
   priors), at the end of the 5-minutes-timeout run, and at the end of the
-  12-minutes-timeout run.
+  12-minutes-timeout run. \textbf{The color gradient} indicates the temporal
+  aspect of the drawing part, or the pen, from blue (origin) to purple (end).
 
@@
 \begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
-  The agent controls a `pen' on a 2D plant. Primitives are those of the LOGO
-  language, modified to have a functional programing structure.
+  The agent controls a `pen' on a 2 dimentional plane. Primitives are those of
+the LOGO language, modified to have a functional programing structure. Task
+  don't take arguments (they are of type \texttt{unit -> image}) and the
+  likelyhood function is all-or-none.
+
+  Invented abstractions include:
+  \begin{itemize}
+    \item $\frac\pi2$ since it has $\pi$, $2$ and $/$ and needs right angles
+    \item A line of any length followed with a right angle
+    \item Loops of length $n$ that uses the number $n$ inside.
+    \item A single line of unit length, followed by a set at the origin of the
+      line
+  \end{itemize}
+
+  The deepest observed level of embedding of invented primitives on the 2 hours
+  run was $7$, meaning that some invented primitives used invented and so on,
+  seven times.
+
 \end{column}

### 2018-07-23 talk.tex
More figures, slides for the talk
figure changes

@@
 \begin{frame}[t]{DSL learning as Bayesian inference}
-  \includegraphics[width = 11cm]{figures/animation/EC.eps}
+  %  \includegraphics[width = 11cm]{figures/animation/EC.eps}
+\centering  \begin{tikzpicture}[scale=1.3,line width=0.5mm]
+
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-2cm]zp) (xp) {data};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {data};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  %\draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+%  \node at (nn) {\NeuralNetwork{0.5}};
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {data};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+%  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+%  \node at (nn) {\NeuralNetwork{0.5}};
+
+%  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+ % \node at (nn) {\NeuralNetwork{0.5}};
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  %\node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+
+  \end{tikzpicture}
 
@@
 
-  \vfill
-  Gray: Observed.\\
-  White: Latent.\\
-  Boxed (plate): Repeated.\\
+  %% \vfill
+  %% Gray: Observed.\\
+  %% White: Latent.\\
+  %% Boxed (plate): Repeated.\\
 
 \end{frame}
+\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
 \begin{frame}[t]{DSL learning as \alert{amortized} Bayesian inference}
-  \includegraphics[width = 11cm]{figures/animation/DC.eps}
+\centering  \begin{tikzpicture}[scale=1.3,line width=0.5mm]
 
-  \vfill
-  Recognition model: Neural net.\\
-  Red: Bottom-up inference.\\
-  Gray: Observed.\\
-  White: Latent.\\
-  Boxed (plate): Repeated.\\
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-2cm]zp) (xp) {data};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {data};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+%  \node at (nn) {\NeuralNetwork{0.5}};
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {data};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+%  \node at (nn) {\NeuralNetwork{0.5}};
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \node at (nn) {\NeuralNetwork{0.25}};
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  %\node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+
+  \end{tikzpicture}
 
@@
 
+\begin{frame}{Turtle graphics --- Some results}
+  \centering
+  \includegraphics[width=9cm]{figures/dreams/montageTalk.eps}
+\end{frame}
+
 \begin{frame}{Vision}

### 2018-07-23 misc2.tex
More improvements to the poster.
poster stuff

--- 2018-07-23 misc2.tex
@@
-s%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Jacobs Portrait Poster
@@
 
-\author{Kevin Ellis, Mathias Sabl\'e Meyer, Lucas Morales, Maxwell Nye, Luke Hewitt,\\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
+\author{Kevin Ellis, Mathias Sabl\'e Meyer, Lucas Morales, %% Maxwell Nye, Luke Hewitt,
+  \\ Armando Solar-Lezama, Joshua B. Tenenbaum} % Author(s)
 
@@
 \begin{block}{Wake: Problem solving}
-  \begin{tikzpicture}[scale=2.5,line width=1.3mm]
+  \begin{tikzpicture}[scale=2.5,line width=1.3mm,arrowhead=0.1mm]
     \node at (0,0) (d){DSL};
@@
     \node[align = center, text width = 1cm] at ([yshift = 0.6cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
-    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+    \draw [double,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
 
@@
 
-    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+    \draw [double,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+    \draw [->,rounded corners,double,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
 
     \node[align=left] at (7,-1) (f) {Programs};
-    \draw [->  ] (s.east) -- (f.west);
+    \draw [double,->  ] (s.east) -- (f.west);
 
-    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+    \draw [->  ,rounded corners,double,] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
   \end{tikzpicture}
@@
   %% Simple model: just predicts probabilities of DSL productions.  
-  Goal: learn to invert generative model from (program, task) pairs
+  Goal: learn to invert generative model, trained on (program, task) pairs
 
@@
 
-  \textbf{Left column:} 16 out of 92 example tasks to hit in an all-or-none
-  fashion --- training set for the \system algorithm.
-  \textbf{Right column:} Learning curve for a run with a timeout of search time
-  per task of 5 minutes (blue, bottom) and of 120 minutes (red, top). On the
-  bottom are dreams, from left to right: before any training (original DSL, flat
-  priors), at the end of the 5-minutes-timeout run, and at the end of the
-  12-minutes-timeout run.
+  \textbf{Left column:} 16 out of 92 example tasks %% to hit in an all-or-none
+  %% fashion
+ (training set). % for the \system algorithm.
+  \textbf{Top right:} Learning curves for  runs w/ search timeout
+    5 minutes (blue, bottom) and  120 minutes (red, top). \textbf{Bottom right:}
+   dreams, from left to right: before learning, at the end of the 5-minute-timeout run, and at the end of the
+  120-minute-timeout run. \textbf{Color gradient:} pen trajectory over time,
+  %% aspect of the drawing part, or the pen,
+  from blue (origin) to purple (end).
 
@@
 \begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
-  The agent controls a `pen' on a 2D plant. Primitives are those of the LOGO
-  language, modified to have a functional programing structure.
+  Agent controls a `pen' on a 2 dimentional plane. LOGO/Turtle primitives.%% Primitives are those of
+%% the LOGO language, modified to have a functional programing structure. Task
+%%   don't take arguments (they are of type \texttt{unit -> image}) and the
+%%   likelyhood function is all-or-none.
+
+  Invented abstractions include:
+  \begin{itemize}
+    \item $\frac\pi2$ since it has $\pi$, $2$ and $/$ and needs right angles
+    \item A line of any length followed with a right angle
+    \item Loops of length $n$ that uses the number $n$ inside.
+    \item A single line of unit length, followed by a set at the origin of the
+      line
+  \end{itemize}
+
+  Deep embedding of invented abstractions: up to $7$, %observed level of embedding of invented primitives on the 2 hours
+  %% run was $7$
+  meaning that some invented primitives used invented and so on,
+  seven times.
+
 \end{column}
@@
 
-  We gratefully acknowledge collaboration with Eyal Dechter, whose EC algorithm (Dechter et al, IJCAI 2013) provided the inspiration for DreamCoder, and Luke Hewitt, who graciously provided us with
-  a regex learning data set.
+  We gratefully acknowledge collaboration with Eyal Dechter, whose EC algorithm (Dechter et al, IJCAI 2013) provided the inspiration for DreamCoder.%% , and Luke Hewitt, who graciously provided us with
+  %% a regex learning data set.
 

### 2018-07-23 misc2.tex
Figure ?improvements?

@@
   \node[latent,scale=3] at (3.5,3) (dx){DSL};
-  \node[latent,scale=3] at ([yshift=-1.5cm]dx) (zp){prog};
+  \node[latent,scale=3] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
   \node[obs,scale=3] at ([yshift=-2cm]zp) (xp) {data};
@@
   \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-%  \node at (nn) {\NeuralNetwork{0.5}};
 
@@
   \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-%  \node at (nn) {\NeuralNetwork{0.5}};
 
   \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
-  \node at (nn) {\NeuralNetwork{0.5}};
   \draw [->] (dx.south) -- (zp.north);
@@
 
-  %\node[shift={+(0,-1.7)}] at (nn) { $Q$  };
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.5}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
 
@@
     \end{tikzpicture}
- % \caption{\system solves for programs, the DSL, and a recognition model.
+  \caption{~\system solves for programs, the DSL, and a recognition model.
 Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}

### 2018-07-23 talk.tex
Animated unfolding of DreamCoder with figures

@@
 \frame{\titlepage}
+
+\begin{frame}{EC --- Wake}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-0.pdf}
+  }
+  \only<2>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-1.pdf}
+  }
+  \only<3>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-2.pdf}
+  }
+  \only<4>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-3.pdf}
+  }
+  \only<5>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-4.pdf}
+  }
+
+\end{frame}
+
+\begin{frame}{DreamCoder --- Wake}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-0.pdf}
+  }
+  \only<2>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-1.pdf}
+  }
+  \only<3>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-2.pdf}
+  }
+  \only<4>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-3.pdf}
+  }
+  \only<5>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-4.pdf}
+  }
+
+\end{frame}
+
+\begin{frame}{DreamCoder --- Sleep-G}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-g/slide-4.pdf}
+  }
+\end{frame}
+\begin{frame}{DreamCoder --- Sleep-R-1}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-1/slide-4.pdf}
+  }
+\end{frame}
+\begin{frame}{DreamCoder --- Sleep-R-2}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-2/slide-4.pdf}
+  }
+\end{frame}
 

### 2018-07-23 talk.tex
Sequences unfolding and XKCD (?)

@@
 \usepackage{pifont}% http://ctan.org/pkg/pifont
+\usepackage{caption}
+\usepackage{subcaption}
 \newcommand{\cmark}{\ding{51}}%
@@
 
-\begin{frame}{Turtle graphics --- Some results}
+\begin{frame}{Turtle graphics --- Learning curves and dreams}
   \centering
   \includegraphics[width=9cm]{figures/dreams/montageTalk.eps}
+\end{frame}
+
+\begin{frame}{Turtle graphics --- Improved DSL}
+  \begin{itemize}
+    \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
+    \item A line of length \texttt{n} followed with a right angle
+    \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
+    \item Unit line then teleport back to origin
+    \item ...
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Turtle graphics --- All the way down}
+  \centering
+  \begin{center}
+  \only<1>{
+    \begin{figure}
+      \includegraphics[width=5cm]{figures/teachLogo/star5_black.eps}
+    \end{figure}
+  }
+  \only<2>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_colour.eps}
+    \end{figure}
+  }
+  \only<3>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq1.eps}
+    \end{figure}
+  }
+  \only<4>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq2.eps}
+    \end{figure}
+  }
+  \only<5>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq3.eps}
+    \end{figure}
+  }
+  \only<6>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq4.eps}
+    \end{figure}
+  }
+  \only<7>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq5.eps}
+    \end{figure}
+  }
+  \only<8>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq1.eps}
+    \end{figure}
+  }
+  \only<9>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq2.eps}
+    \end{figure}
+  }
+  \only<10>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq3.eps}
+    \end{figure}
+  }
+  \only<11>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq4.eps}
+    \end{figure}
+  }
+  \only<12>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq5.eps}
+    \end{figure}
+  }
+  \only<13>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/star5_seq1.eps}
+    \end{figure}
+  }
+  \only<14>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp000.png}
+    \end{figure}
+  }
+  \only<15>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp002.png}
+    \end{figure}
+  }
+  \only<16>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp004.png}
+    \end{figure}
+  }
+  \only<17>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp006.png}
+    \end{figure}
+  }
+  \only<18>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp008.png}
+    \end{figure}
+  }
+  \only<19>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp010.png}
+    \end{figure}
+  }
+  \only<20>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp012.png}
+    \end{figure}
+  }
+  \only<21>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp014.png}
+    \end{figure}
+  }
+  \only<22>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp016.png}
+    \end{figure}
+  }
+  \only<23>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp018.png}
+    \end{figure}
+  }
+  \only<24>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp020.png}
+    \end{figure}
+  }
+  \only<25>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp022.png}
+    \end{figure}
+  }
+  \only<26>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp024.png}
+    \end{figure}
+  }
+  \only<27>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp026.png}
+    \end{figure}
+  }
+  \only<28>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp028.png}
+    \end{figure}
+  }
+  \only<29>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp030.png}
+    \end{figure}
+  }
+  \only<30>{
+    \begin{figure}
+    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp032.png}
+    \end{figure}
+  }
+\end{center}
 \end{frame}
@@
 
+    \pause{}
+
+    \begin{figure}
+        \includegraphics[scale=0.44]{figures/xkcd_star.png}
+        \caption*{Thanks to \textsc{xkcd} we can rule out $36^\circ$ as a primitive.}
+    \end{figure}
+
   \end{frame}

### 2018-07-24 talk.tex
talk improvements

@@
 
-\begin{frame}{EC --- Wake}
-  \only<1>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-0.pdf}
-  }
-  \only<2>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-1.pdf}
-  }
-  \only<3>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-2.pdf}
-  }
-  \only<4>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-3.pdf}
-  }
-  \only<5>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-4.pdf}
-  }
-
-\end{frame}
-
-\begin{frame}{DreamCoder --- Wake}
-  \only<1>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-0.pdf}
-  }
-  \only<2>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-1.pdf}
-  }
-  \only<3>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-2.pdf}
-  }
-  \only<4>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-3.pdf}
-  }
-  \only<5>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-4.pdf}
-  }
-
-\end{frame}
-
-\begin{frame}{DreamCoder --- Sleep-G}
-  \only<1>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-g/slide-4.pdf}
-  }
-\end{frame}
-\begin{frame}{DreamCoder --- Sleep-R-1}
-  \only<1>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-1/slide-4.pdf}
-  }
-\end{frame}
-\begin{frame}{DreamCoder --- Sleep-R-2}
-  \only<1>{
-    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-2/slide-4.pdf}
-  }
-\end{frame}
-
 \begin{frame}{The Language of Thought}
@@
 \end{frame}
-
 \begin{frame}[t]{DSL learning as Bayesian inference}
@@
 
+\pause
+
+\textbf{New:} amortized inference +\\ better program representation (Lisp) + \\better DSL inference
+
 \end{frame}
-\begin{frame}[t]{Wake: Problem solving}
-  \includegraphics[width = 11cm]{figures/animation/Wake.eps}
-
-  \vfill
+\begin{frame}{Wake --- as in Exploration/Compression Algorithm}
   \only<1>{
-  Recognition model: Neural net.\\
-  Red: Bottom-up inference.\\
-  Gray: Observed.\\
-  White: Latent.\\
-  Boxed (plate): Repeated.\\
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-0.pdf}
   }
   \only<2>{
-  \begin{tikzpicture}[xscale=1.25,line width=1mm]
-    \node at (0,0) (d){DSL};
-    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-1.pdf}
+  }
+  \only<3>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-2.pdf}
+  }
+  \only<4>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-3.pdf}
+  }
+  \only<5>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-ec/slide-4.pdf}
+  }
 
-    \node at ([xshift = 2cm]t) (nn){
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.3,shift={+(-1,2.5)}}]
-       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[align = center, text width = 1cm] at ([yshift = 0.7cm]nn.north) {\baselineskip=0pt \Small Recog. model\par};
-    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+\end{frame}
 
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
-%    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
+\begin{frame}{DreamCoder --- Wake}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-0.pdf}
+  }
+  \only<2>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-1.pdf}
+  }
+  \only<3>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-2.pdf}
+  }
+  \only<4>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-3.pdf}
+  }
+  \only<5>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-dc/slide-4.pdf}
+  }
 
-    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+\end{frame}
 
-    \node[align=left] at (7,-1) (f) {Programs};
-    \draw [->  ] (s.east) -- (f.west);
+\begin{frame}{DreamCoder --- Sleep-G}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-g/slide-4.pdf}
+  }
+\end{frame}
+\begin{frame}{DreamCoder --- Sleep-R (\textbf{Experience Replay})}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-1/slide-4.pdf}
+  }
+\end{frame}
+\begin{frame}{DreamCoder --- Sleep-R (\textbf{Dreaming})}
+  \only<1>{
+    \includegraphics[width=11cm]{figures/teachDC/.inkslides-sleep-r-2/slide-4.pdf}
+  }
+\end{frame}
 
-    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-  \end{tikzpicture}}
+%% \begin{frame}[t]{Wake: Problem solving}
+%%   \includegraphics[width = 11cm]{figures/animation/Wake.eps}
+
+%%   \vfill
+%%   \only<1>{
 %%   Recognition model: Neural net.\\
@@
 %%   Boxed (plate): Repeated.\\
+%%   }
+%%   \only<2>{
+%%   \begin{tikzpicture}[xscale=1.25,line width=1mm]
+%%     \node at (0,0) (d){DSL};
+%%     \node at ([yshift = -2cm]d) (t){$\text{Task}$};
 
-\end{frame}
-\begin{frame}[t]{Sleep-G: Memory consolidation}
-  \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
-  \vfill
-  \textbf{Fragment Grammars: O'Donnell 2015.}\\
-  Orange: Code fragments.\\
-  %% Gray: Observed.\\
-  %% White: Latent.\\
-  %% Boxed (plate): Repeated.\\
+%%     \node at ([xshift = 2cm]t) (nn){
+%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.3,shift={+(-1,2.5)}}]
+%%        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%         \node[rectangle] at (1,1) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (I-\name) at (0,-\y) {};
+%%         \foreach \name / \y in {1,...,3}
+%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (O-\name) at (2,-\y) {};
+%%         \foreach \source in {1,...,4}
+%%             \foreach \dest in {1,...,3}
+%%                 \draw [-latex] (I-\source) -- (H-\dest);
+%%         \foreach \source in {1,...,3}
+%%             \foreach \dest in {1,...,4}
+%%                 \draw [-latex] (H-\source) -- (O-\dest);
+%%       \end{tikzpicture}
+%%     };
+%%     \node[align = center, text width = 1cm] at ([yshift = 0.7cm]nn.north) {\baselineskip=0pt \Small Recog. model\par};
+%%     \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
 
-\end{frame}
-\begin{frame}[t]{Sleep-R: Objective}
-  \includegraphics[width = 11cm]{figures/animation/Wake.eps}
+%%     \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+%% %    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
 
-    \vfill
-    Recognition model predicts distribution over program, conditional on task.\\
-    Training: (program, task) pairs\\
-    Objective: Predict program w/ (1) high prior under DSL \& \\(2) high likelihood for task
+%%     \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+%%     \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
 
-\end{frame}
-\begin{frame}[t]{Sleep-R: Experience replay}
-  \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
+%%     \node[align=left] at (7,-1) (f) {Programs};
+%%     \draw [->  ] (s.east) -- (f.west);
 
-    \textbf{Train on (program, task) pairs found during waking}
-\end{frame}
-\begin{frame}[t]{Sleep-R: Dreaming}
-  \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
+%%     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+%%   \end{tikzpicture}}
+%%   %% Recognition model: Neural net.\\
+%%   %% Red: Bottom-up inference.\\
+%%   %% Gray: Observed.\\
+%%   %% White: Latent.\\
+%%   %% Boxed (plate): Repeated.\\
 
-    \textbf{Train on (program, task) pairs sampled from DSL}
+%% \end{frame}
+%% \begin{frame}[t]{Sleep-G: Memory consolidation}
+%%   \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
+%%   \vfill
+%%   \textbf{Fragment Grammars: O'Donnell 2015.}\\
+%%   Orange: Code fragments.\\
+%%   %% Gray: Observed.\\
+%%   %% White: Latent.\\
+%%   %% Boxed (plate): Repeated.\\
 
-\end{frame}
+%% \end{frame}
+%% \begin{frame}[t]{Sleep-R: Objective}
+%%   \includegraphics[width = 11cm]{figures/animation/Wake.eps}
+
+%%     \vfill
+%%     Recognition model predicts distribution over program, conditional on task.\\
+%%     Training: (program, task) pairs\\
+%%     Objective: Predict program w/ (1) high prior under DSL \& \\(2) high likelihood for task
+
+%% \end{frame}
+%% \begin{frame}[t]{Sleep-R: Experience replay}
+%%   \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
+
+%%     \textbf{Train on (program, task) pairs found during waking}
+%% \end{frame}
+%% \begin{frame}[t]{Sleep-R: Dreaming}
+%%   \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
+
+%%     \textbf{Train on (program, task) pairs sampled from DSL}
+
+%% \end{frame}
 
@@
 
-  Discovers 38 concepts, including `filter'
+  Discovers 38 concepts, including `filter'. With suitable curriculum can also learn `map', `fold', etc.
 \end{frame}
 
-\begin{frame}{List functions: Learning curves on hold out tasks}
+\begin{frame}{List functions \& Text editing: Learning curves on hold out tasks}
 
   \begin{center}
-      \includegraphics[width = 6.5cm]{figures/listLearningCurve.eps} 
+    \includegraphics[width = 5cm]{figures/listLearningCurve.eps}
+\hfill    \includegraphics[width = 5cm]{figures/textLearningCurve.eps} 
     \end{center}
@@
   \begin{block}{Tasks}
-    \texttt{task : unit -> image}
+    \texttt{task : image}
   \end{block}
@@
 
-  \textbf{Takeaway:}
-  \begin{itemize}
-    \item Humans flexibly adapt to diverse sets of new problem domains --- DreamCoder takes a step in this direction
-  \end{itemize}
+   \underline{More human-like machine intelligence:}\\Flexibly adapting to new problem domains by
+   acquiring a domain-specific language-of-thought (DSL), while jointly learning how to use the language of thought (recognition model)
+
   \pause
 
-  %% \textbf{Future work}
-
-  %% \begin{itemize}
-  %%   \pause
-  %% \item Theory learning
-  %%   \item Planning
-  %%   \end{itemize}
-
-    \pause{}
-
-    \begin{figure}
-        \includegraphics[scale=0.44]{figures/xkcd_star.png}
-        \caption*{Thanks to \textsc{xkcd} we can rule out $36^\circ$ as a primitive.}
-    \end{figure}
-
+\hspace{-1cm}\includegraphics[width = 12cm]{figures/finale.png} 
   \end{frame}

### 2018-07-24 talk.tex
Figure tweaking and more dreams

@@
 %  \node at (nn) {\NeuralNetwork{0.5}};
+  % HACK : "invisible" arrow to phantom and push to the left and align with next
+  % slide
+  \draw [->,red,opacity=0.0001] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
 
@@
 
-\begin{frame}{Turtle graphics --- Learning curves and dreams}
+\begin{frame}{Turtle graphics --- Learning curves}
   \centering
-  \includegraphics[width=9cm]{figures/dreams/montageTalk.eps}
-\end{frame}
-
-\begin{frame}{Turtle graphics --- Improved DSL}
+  \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
+  \pause{}
   \begin{itemize}
@@
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Turtle graphics --- Dreams}
+  \centering
+  \includegraphics[width=7cm]{figures/dreams/montageTalkDreams.eps}
+\end{frame}
+
+\begin{frame}{Turtle graphics --- More dreams, 5 minutes, 1st iteration}
+  \centering
+  \includegraphics[width=8cm]{figures/dreams/5m_firstIteration.eps}
+\end{frame}
+
+\begin{frame}{Turtle graphics --- More dreams, 5 minutes, last iteration}
+  \centering
+  \includegraphics[width=8cm]{figures/dreams/5m_final.eps}
 \end{frame}

### 2018-07-24 talk.tex
final slide

@@
 
-   \underline{More human-like machine intelligence:}\\Flexibly adapting to new problem domains by
-   acquiring a domain-specific language-of-thought (DSL), while jointly learning how to use the language of thought (recognition model)
+   \underline{More human-like machine intelligence}\\Flexibly adapting to new problem domains:
+   acquiring a domain-specific language-of-thought (DSL); learning how to use the language of thought (recognition model)
 

### 2018-07-24 talk.tex
improvements to final slide
even more tweaks to the final slide

@@
   \item Library of concepts (declarative knowledge)
-    \item Search strategy (procedural knowledge)
+    \item Inference strategy (procedural knowledge)
     \end{itemize}
@@
 
-   \underline{More human-like machine intelligence}\\Flexibly adapting to new problem domains:
-   acquiring a domain-specific language-of-thought (DSL); learning how to use the language of thought (recognition model)
+   \underline{More human-like machine intelligence}\\%Flexibly adapting to new problem domains:
+   \begin{itemize}
+   \item    Acquiring a domain-specific language-of-thought (DSL)
+     \item Learning  to use the language of thought (recognition model)
+   \end{itemize}
+   DreamCoder: an algorithm for jointly realizing these goals
+
+   \hspace{-1cm}\includegraphics[width = 12cm]{figures/finale.png}
 
@@
 
-\hspace{-1cm}\includegraphics[width = 12cm]{figures/finale.png} 
+   \Huge \centering \textbf{The End.}
   \end{frame}

### 2018-07-30 nips_2018_author_response.tex
Tex update

\documentclass{article}

\usepackage{nips_2018_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\begin{document}

Something something thank you guys for your rich on point comments that are so
very useful for making science what it is.

\subsubsection*{Choice of original \texttt{DSL}}

As several reviewers pointed out the original \texttt{DSL} influences the efficiency and
accuracy of the algorithm. Reviewer 1 points out ``What happens, for example, if
length is removed [from the original list \texttt{DSL}]?''. To explore that question we
took the original \texttt{Lisp} specification from [citation needed] as well as
tasks from \texttt{Lisp} textbooks to see whether higher level functions would
be discovered. While this example was computationally expensive, it successfully
rediscovered common abstractions from the functional programming world,
including but not limited to \texttt{fold}, \texttt{unfold}, \texttt{map},
\texttt{length}, \texttt{zip}, \texttt{range}, etc.

That being said, the intended goal of the algorithm is not to rediscover
well-known useful abstractions but rather to adapt such abstractions to the
tasks at hands, through more abstractions if need be but not with that as a goal
\textit{per se}.

\paragraph*{Example of evolution for a task through \texttt{DSL}s}

Find some logs and give an example I guess? I have a few available for geometry
but hey :/

\subsubsection*{Relationship to Schmid and Kitzelmann}

We thank reviewer three for suggesting additions to the related work. Our
revision will prominently highlight the work of Kitzelmann, Schmid and Katayama.

The \textsc{Igor} systems introduced by Kitzelmann and Schmid is a sophisticated
algorithm for analytical inductive programming that handles a certain classe of
programs, much like {Sketch} and {FlashFill} are sophisticated
program synthesiser for certain classes of programs.

We see these line of work as orthogonal and complementary to that of SCC\@. Our
goal is to take a black-box synthesiser and learn how to solve a class of
problems. One could in principle replace our enumerative search based solver
with \textsc{Igor}, Sketch or \textsc{MagicHaskeller}. We chose however to
use a more generic and simple enumerative approach for it allows us to apply SCC
to a diverse spread of problems: functional programming exercises, text editing,
graphics programming and symbolic regressions from images.

\subsubsection*{Reproducibility}

Our project --- including source code, data and automation scripts --- is
already available on \url{GitHub} and future revision will include a link to it
once the anonymity of the authors can be unveiled. No technical knowledge will
be required to reproduced the results.

\end{document}

### 2018-07-30 nips_2018_author_response.tex
Figure, include R2 in the reponse

@@
 \usepackage{microtype}      % microtypography
+
+\usepackage[rightcaption]{sidecap}
+\usepackage{graphicx}
 
@@
 
-\paragraph*{Example of evolution for a task through \texttt{DSL}s}
-
-Find some logs and give an example I guess? I have a few available for geometry
-but hey :/
+\begin{SCfigure}[2][h]
+  \centering
+  \includegraphics[width=7cm]{figures/rebutal/spirals.eps}\label{geomCompiled}
+  \caption{Practical example of DSL evolution: originally, the preferred program
+  for the leftmost spiral does not use loops while that of the two other does.
+  After an iteration the structure of the two rightmost spiral is abstracted.
+  For the next iteration the preferred program for the leftmost spiral leverages
+  on that and changes in a meaningful way.}
+\end{SCfigure}
 
@@
 algorithm for analytical inductive programming that handles a certain classe of
-programs, much like {Sketch} and {FlashFill} are sophisticated
-program synthesiser for certain classes of programs.
+programs, much like {Sketch} and {FlashFill} are sophisticated program
+synthesiser for certain classes of programs. This also relates to an insightful
+and suggestion for comparison by reviewer 2: FlashFill's DSL is fixed and
+manually curated and enhanced for a specific class of problems, over which is
+performs well at the loss of agility.
 

### 2018-07-30 nips_2018_author_response.tex
Added some answers to reviewer 2

@@
 
+\subsubsection*{More specific questions}
+
+\paragraph*{Reviewer 2: Origin of list task} All the list tasks were manually
+created in order to be interesting, challeging, but also human-interpretable for
+a prior behaviour experiment. They pre-existed our algorithm.
+
+\paragraph*{Reviewer 2: assesment of the recognition model} The recognition
+model's contribution was measured on lesions blabla
+
 \subsubsection*{Reproducibility}

### 2018-07-30 nips_2018_author_response.tex
re- factored response

@@
 We thank reviewer three for suggesting additions to the related work. Our
-revision will prominently highlight the work of Kitzelmann, Schmid and Katayama.
+revision will prominently highlight the work of Kitzelmann, Schmid and Katayama,
+particularly Kitzelmann and Schmid, whose cognitive inspiration
+is closely allied with our motivations.
 
-The \textsc{Igor} systems introduced by Kitzelmann and Schmid is a sophisticated
-algorithm for analytical inductive programming that handles a certain classe of
-programs, much like {Sketch} and {FlashFill} are sophisticated program
-synthesiser for certain classes of programs. This also relates to an insightful
+Kitzelmann and Schmid's \textsc{Igor} systems are sophisticated
+analytical inductive programming algorithms that handle certain classes of
+problems, much like {Sketch} and {FlashFill} are sophisticated program
+synthesizers for certain classes of programs.
+We see these lines of work as orthogonal and complementary to that of SCC\@. Our
+goal is to take a black-box synthesizer and then learn how to use it to solve a class of
+problems. One could replace our enumerative synthesizer
+with \textsc{Igor}, Sketch or \textsc{MagicHaskeller}. We chose however to
+use a generic and simple enumerative approach because it allowed us to apply SCC
+to a diverse spread of problems: functional programming exercises, text editing,
+turtle graphics programming, and symbolic regression from images.
+
+This also relates to an insightful
 and suggestion for comparison by reviewer 2: FlashFill's DSL is fixed and
@@
 performs well at the loss of agility.
-
-We see these line of work as orthogonal and complementary to that of SCC\@. Our
-goal is to take a black-box synthesiser and learn how to solve a class of
-problems. One could in principle replace our enumerative search based solver
-with \textsc{Igor}, Sketch or \textsc{MagicHaskeller}. We chose however to
-use a more generic and simple enumerative approach for it allows us to apply SCC
-to a diverse spread of problems: functional programming exercises, text editing,
-graphics programming and symbolic regressions from images.
 
@@
 \subsubsection*{Reproducibility}
+Our source code is publicly available on GitHub,
+and a link will be added upon publication (in NIPS or elsewhere),
+along with the specific command lines  needed to run
+the experiments in the paper.
 
-Our project --- including source code, data and automation scripts --- is
-already available on \url{GitHub} and future revision will include a link to it
-once the anonymity of the authors can be unveiled. No technical knowledge will
-be required to reproduced the results.
+%% Our project --- including source code, data and automation scripts --- is
+%% already available on \url{GitHub} and future revision will include a link to it
+%% once the anonymity of the authors can be unveiled. No technical knowledge will
+%% be required to reproduced the results.
 

### 2018-07-30 nips_2018_author_response.tex
initial programming primitives section

@@
 
-\subsubsection*{Choice of original \texttt{DSL}}
+\subsubsection*{Choice of initial programming primitives}
 
-As several reviewers pointed out the original \texttt{DSL} influences the efficiency and
-accuracy of the algorithm. Reviewer 1 points out ``What happens, for example, if
-length is removed [from the original list \texttt{DSL}]?''. To explore that question we
-took the original \texttt{Lisp} specification from [citation needed] as well as
-tasks from \texttt{Lisp} textbooks to see whether higher level functions would
-be discovered. While this example was computationally expensive, it successfully
-rediscovered common abstractions from the functional programming world,
-including but not limited to \texttt{fold}, \texttt{unfold}, \texttt{map},
-\texttt{length}, \texttt{zip}, \texttt{range}, etc.
+Reviewer 1 astutely observes that our initial set of programming primitives is
+not as minimal as it could have been,
+raising intriguing questions
+about the importance and role of the initial basis.
+We believe that general program-learners should not start with a \emph{minimal} basis,
+but rather a \emph{domain-agnostic}
+basis,
+like those embodied in the standard libraries of modern programming languages.
+An open question however is the extent to which one could ``learn from scratch.''
+As a step toward answering this question,
+we ran the following experiment:
+SCC was provided with a subset of primitives in 1959 Lisp, crucially including primitive recursion,
+and tasked with solving a small set of functional programming exercises, like appending lists or summing the elements of a list.
+After 93 hours on 64 CPUs,
+SCC rediscovered \texttt{fold}, \texttt{unfold}, \texttt{map},
+\texttt{length}, \texttt{zip}, \texttt{range}, \texttt{index},
+and a handful of arithmetic operations.
+Although this result is suggestive,
+we believe that in practice learners should start from a rich yet domain-agnostic
+inventory of primitives, like human programmers do.
 
-That being said, the intended goal of the algorithm is not to rediscover
-well-known useful abstractions but rather to adapt such abstractions to the
-tasks at hands, through more abstractions if need be but not with that as a goal
-\textit{per se}.
+%% As several reviewers pointed out the original \texttt{DSL} influences the efficiency and
+%% accuracy of the algorithm. Reviewer 1 points out ``What happens, for example, if
+%% length is removed [from the original list \texttt{DSL}]?''. To explore that question we
+%% took the original \texttt{Lisp} specification from [citation needed] as well as
+%% tasks from \texttt{Lisp} textbooks to see whether higher level functions would
+%% be discovered. While this example was computationally expensive, it successfully
+%% rediscovered common abstractions from the functional programming world,
+%% including but not limited to \texttt{fold}, \texttt{unfold}, \texttt{map},
+%% \texttt{length}, \texttt{zip}, \texttt{range}, etc.
+
+%% That being said, the intended goal of the algorithm is not to rediscover
+%% well-known useful abstractions but rather to adapt such abstractions to the
+%% tasks at hands, through more abstractions if need be but not with that as a goal
+%% \textit{per se}.
 
@@
 
-\subsubsection*{Relationship to Schmid and Kitzelmann}
+Reviewer 2 suggests comparing with FlashFill, the state-of-the-art program synthesizer for text editing.
+We very much would like to do this comparison, but this comparison
+is encumbered by the fact that FlashFill is not freely available nor accessible outside Microsoft Excel.
+We strongly believe that FlashFill would easily solve essentially all of our text editing problems.
+Instead, we compared with the winner of the 2017 SyGuS program synthesis competition (Alur et al. 2017), 
+showing that, without hand-engineering of the DSL,
+we achieve results comparable to the 2017 winning solver (CVC4),
+when CVC4 was provided with hand-engineered DSLs (Section 3.2 of paper).
+\subsubsection*{Related Work}
 
-We thank reviewer three for suggesting additions to the related work. Our
-revision will prominently highlight the work of Kitzelmann, Schmid and Katayama,
-particularly Kitzelmann and Schmid, whose cognitive inspiration
-is closely allied with our motivations.
-
-Kitzelmann and Schmid's \textsc{Igor} systems are sophisticated
-analytical inductive programming algorithms that handle certain classes of
-problems, much like {Sketch} and {FlashFill} are sophisticated program
-synthesizers for certain classes of programs.
-We see these lines of work as orthogonal and complementary to that of SCC\@. Our
-goal is to take a black-box synthesizer and then learn how to use it to solve a class of
-problems. One could replace our enumerative synthesizer
-with \textsc{Igor}, Sketch or \textsc{MagicHaskeller}. We chose however to
+We thank reviewer three for suggesting points of contact between our algorithm and the broader program induction literature.
+We are revising our paper to prominently cite the work of Schmid and Kitzelmann,
+whose \textsc{Igor} systems, like ours,
+are motivated by the flexibility and generality of human learning.
+Katayama's \textsc{MagicHaskeller} synthesizer for functional programs will also be cited.
+More broadly,
+we see these pieces of related work as
+orthogonal and complementary to
+our system:
+we take a black-box synthesizer and
+learn to use it by growing the DSL and
+training a recognition model,
+and one could replace our enumerative synthesizer with
+\textsc{Igor}, Sketch (Solar-Lezama 2008), $\lambda^2$ (Feser et al. 2015), \textsc{MagicHaskeller},
+or many other synthesis algorithms.
+We chose however to
 use a generic and simple enumerative approach because it allowed us to apply SCC
@@
 turtle graphics programming, and symbolic regression from images.
+
+%% revision will prominently highlight the work of Kitzelmann, Schmid and Katayama,
+%% particularly Kitzelmann and Schmid, whose cognitive inspiration
+%% is closely allied with our motivations.
+
+%% Kitzelmann and Schmid's \textsc{Igor} systems are sophisticated
+%% analytical inductive programming algorithms that handle certain classes of
+%% problems, much like {Sketch} and {FlashFill} are sophisticated program
+%% synthesizers for certain classes of programs.
+%% We see these lines of work as orthogonal and complementary to that of SCC\@. Our
+%% goal is to take a black-box synthesizer and then learn how to use it to solve a class of
+%% problems. One could replace our enumerative synthesizer
+%% with \textsc{Igor}, Sketch or \textsc{MagicHaskeller}.
 
@@
 
+\begin{comment}
+
+Reviewer #1
+Questions
+
+    1. Please provide an "overall score" for this submission.
+        8: Top 50% of accepted NIPS papers. A very good submission; a clear accept. I vote and argue for accepting this submission.
+    2. Please provide a "confidence score" for your assessment of this submission.
+        4: You are confident in your assessment, but not absolutely certain.
+It is unlikely, but not impossible, that you did not understand some parts of the submission or that you are unfamiliar with some pieces of related work.
+    3.
+Please provide detailed comments that explain your "overall score" and "confidence score" for this submission.
+You should summarize the main ideas of the submission and relate these ideas to previous work at NIPS and in other archival conferences and journals.
+You should then summarize the strengths and weaknesses of the submission, focusing on each of the following four criteria: quality, clarity, originality, and significance.
+        = Summary
+
+        A method for learning a DSL for program synthesis together with a search
+        algorithm in that DSL is presented. The method proceeds iteratively,
+        trying to solve tasks with the current DSL, and then extracting new DSL
+        components from the solutions. Experiments show that bootstrapping the
+        method with a DSL made up of trivial primitives is sufficient to
+        discover common high-level constructs present in manually constructed
+        DSLs.
+
+        The paper tackles an important problem (DSL design) in an elegant and
+        novel way. The clarity of the paper is not perfect, as the details of the
+        idea require more space than the 8 pages available, but it clearly is
+        stepping stone towards a new generation of program synthesis approaches.
+
+        = Quality
+
+        The paper precisely defines its main algorithm and the core concepts of
+        the approach. There are a few omissions (e.g., Sect. 2.4 does not
+        explain the exact mechanism to extract new constructs; the supplement
+        gives an algorithm, but omits a clear definition of how "fragments" are
+        found), but overall the paper is precise and explains core concepts well.
+        The experiments are very thorough, comparing the method to interesting
+        ablations and baselines.
+        What I am missing is an analysis of the influence of the initial set of
+        primitives. For example, the list processing set starts with a
+        non-minimal set of primitives (as length is "foldr (\lambda (n) -> 1 + n) 0"
+        and index can be implemented using the others as well, though
+        painfully). What happens, for example, if length is removed?
+
+        As an additional data point, I would have appreciated a look at the
+        "evolution" of solutions to a fixed task through the iterations of the
+        main algorithm, i.e., show an initial, early and late solution in the
+        style of Table 1.
+
+        = Clarity
+
+        Many part of the paper are better understood with the supplement in hand
+        (Sect. 2.3 is not discussing details of how q is modeled, but Supplement
+        Sect. 5 makes clear that is is essentially like DeepCoder; Sect. 2.4 is
+        missing details as discussed above).
+        Overall, the paper is easily understood, but not entirely self-contained
+        and not precise enough to allow re-implementation by third parties.
+
+        The supplement is very unpolished (there are many grammatical errors and
+        half-finished sentences, e.g. in line 6, line 33, lines 45/46, ...)
+
+        = Originality
+
+        While drawing on the program synthesis literature of the last few years
+        (e.g., the guided enumerative search), the core contributions of the
+        paper are completely new and are a substantial step forward for the
+        field.
+
+        = Significance
+
+        Designing a suitable DSL is often one of the hardest tasks in applying
+        program synthesis to a new domain; the promise of automatically learning
+        a DSL (together with a search algorithm) is obvious. I am not entirely
+        convinced by the authors idea of generalizing to other generative
+        programs, but I am looking forward to being proven wrong.
+    4. How confident are you that this submission could be reproduced by others, assuming equal access to data and resources?
+        2: Somewhat confident
+
+Reviewer #2
+Questions
+
+    1. Please provide an "overall score" for this submission.
+        8: Top 50% of accepted NIPS papers. A very good submission; a clear accept. I vote and argue for accepting this submission.
+    2. Please provide a "confidence score" for your assessment of this submission.
+        3: You are fairly confident in your assessment.
+It is possible that you did not understand some parts of the submission or that you are unfamiliar with some pieces of related work.
+Math/other details were not carefully checked.
+    3.
+Please provide detailed comments that explain your "overall score" and "confidence score" for this submission.
+You should summarize the main ideas of the submission and relate these ideas to previous work at NIPS and in other archival conferences and journals.
+You should then summarize the strengths and weaknesses of the submission, focusing on each of the following four criteria: quality, clarity, originality, and significance.
+        This paper proposes an algorithm for program synthesis in an incremental way.
+There are three steps to it: 1) searching for program solutions, 2) expandind the dsl by discovering useful reusable components, 3) training a neural network that primes the program solution search in 1).
+The interesting part of the contribution is the focus on roughly mimicking the way humans refactor, recompose and reuse when coding.
+
+        Strengths
+        - the problem presented is interesting, and the scc algorithm is novel. the theoretical framework is sound, and approximations made are reasonable and useful
+        - the paper presents a good evaluation framework on three different tasks and through the ablation analysis that clearly shows the contributions of each part of the model to the final results
+        - the results are impressive - excellent performance on all three tasks, both accuracy-wise and time-wise
+
+        Weaknesses
+        - this is quite a dense paper and it packs a lot of information
+        - the paper could benefit from a comparison to a well known systems such as flashfill, where the dsl is fixed
+
+        Questions:
+        - 125-126 the probability of a variable occurring in a program. could you provide an example here?
+        - equation 3 - is this the counting norm? if so, can you provide a quick rationale why it is used
+        - 167 - can you explicitly say what AIC is and/or cite it?
+        - 181-183 - how is this unification done?
+        - 200 - was this manually created?
+        - what was the recognition model validated on?
+
+        Other:
+        - figure 4 position is weird
+        - figure 4 bottom left and right are not clear
+    4. How confident are you that this submission could be reproduced by others, assuming equal access to data and resources?
+        1: Not confident
+
+Reviewer #3
+Questions
+
+    1. Please provide an "overall score" for this submission.
+        7: A good submission; an accept. I vote for accepting this submission, although I would not be upset if it were rejected.
+    2. Please provide a "confidence score" for your assessment of this submission.
+        3: You are fairly confident in your assessment.
+It is possible that you did not understand some parts of the submission or that you are unfamiliar with some pieces of related work.
+Math/other details were not carefully checked.
+    3.
+Please provide detailed comments that explain your "overall score" and "confidence score" for this submission.
+You should summarize the main ideas of the submission and relate these ideas to previous work at NIPS and in other archival conferences and journals.
+You should then summarize the strengths and weaknesses of the submission, focusing on each of the following four criteria: quality, clarity, originality, and significance.
+        The authors present an inductive programming approach where a DSL is learned by a search, compile, and compress strategy.
+They evaluate their approach in three different domains -- list processing, text editing, and symbolic regression.
+They motivate their approach by pointing out the effectivenes and flexibility of human learning.
+Similar arguments are presented by Schmid & Kitzelmann, Cognitive Systems Research 2011 and the approach has some similarities with the MagicHaskeler of Katayama.Although some relation to pevious research is missing (mainly the two groups named above), the presented approach is original.
+I am most impressed by the empirical demonstration of applicability in three rather distinct domains (but: see also Schmid/Kitzelmann, 2011).
+The presentation is technically sound and the empirical evaluation is very convincing.
+    4. How confident are you that this submission could be reproduced by others, assuming equal access to data and resources?
+        3: Very confident
+
+  \end{comment}
+
 \end{document}

### 2018-07-31 nips_2018_author_response.tex
a bunch of editing

@@
 
+\subsubsection*{The evolution of task solutions over time}
+Reviewer 1 asks about the ``evolution of solutions to a fixed task through the iterations of the main algorithm.''
+In most cases the solution to a fixed task
+stays the same,
+modulo rewriting the solution in terms of new subroutines.
+There are interesting cases however where the
+semantics of the program changes as the algorithms constructs more subroutines. Figure~\cite{geomCompiled}
+illustrates one case for turtle graphics.
 \begin{SCfigure}[2][h]
@@
   \includegraphics[width=7cm]{figures/rebutal/spirals.eps}\label{geomCompiled}
-  \caption{Practical example of DSL evolution: originally, the preferred program
-  for the leftmost spiral does not use loops while that of the two other does.
-  After an iteration the structure of the two rightmost spiral is abstracted.
-  For the next iteration the preferred program for the leftmost spiral leverages
-  on that and changes in a meaningful way.}
+  \caption{Three turtle graphics tasks: the task is to write a program that draws the image.
+Initially SSC solves the leftmost task without using a loop, but after solving the middle and rightmost tasks,
+  it constructs a parametric `spiral' subroutine, which uses a loop -- on subsequent iterations it solves the leftmost task using the `spiral' routine.}
 \end{SCfigure}

### 2018-07-31 nips_2018_author_response.tex
completed a draft of response

@@
 \usepackage{graphicx}
-
+\usepackage{verbatim}
 \begin{document}
 
-Something something thank you guys for your rich on point comments that are so
-very useful for making science what it is.
+We thank all of our reviewers for their thoughtful and detailed comments, and we are revising our our manuscript in light of this helpful feedback.
+Here, we discuss a handful of specific points:
+%\subsubsection*{Choice of initial programming primitives}
 
-\subsubsection*{Choice of initial programming primitives}
-
-Reviewer 1 astutely observes that our initial set of programming primitives is
+\textbf{Choice of initial programming primitives:} Reviewer 1 observes that our initial set of programming primitives is
 not as minimal as it could have been,
@@
 we ran the following experiment:
-SCC was provided with a subset of primitives in 1959 Lisp, crucially including primitive recursion,
+SCC was provided with a minimal subset of primitives in 1959 Lisp, crucially including primitive recursion,
 and tasked with solving a small set of functional programming exercises, like appending lists or summing the elements of a list.
@@
 
-\subsubsection*{The evolution of task solutions over time}
+\textbf{The evolution of task solutions over time:} %\subsubsection*{The evolution of task solutions over time}
 Reviewer 1 asks about the ``evolution of solutions to a fixed task through the iterations of the main algorithm.''
@@
 There are interesting cases however where the
-semantics of the program changes as the algorithms constructs more subroutines. Figure~\cite{geomCompiled}
+semantics of the program changes as the algorithms constructs more subroutines. Figure~1
 illustrates one case for turtle graphics.
@@
   \centering
-  \includegraphics[width=7cm]{figures/rebutal/spirals.eps}\label{geomCompiled}
+  \includegraphics[width=7cm]{figures/rebutal/spirals.eps}
   \caption{Three turtle graphics tasks: the task is to write a program that draws the image.
@@
 
-Reviewer 2 suggests comparing with FlashFill, the state-of-the-art program synthesizer for text editing.
+\textbf{Fragment Grammars:} Reviewers 1 \& 2 both request clarification on the mechanics of the DSL induction,
+which employs a representation from computational linguistics called Fragment Grammars (O'Donnell 2015).
+We are revising the Supplement to include a brief overview of this formalism.
+
+\textbf{Reproducibility:} %\subsubsection*{Reproducibility}
+Our source code is publicly available on GitHub,
+and a link will be added upon publication (in NIPS or elsewhere),
+along with the specific command lines  needed to run
+the experiments in the paper.
+
+\textbf{Specific questions from Reviewer 2:}
+The 236 list tasks were created by one of the authors of this paper
+several months before this project began
+and without this system (SCC) in mind,
+with the goal of
+being human-interpretable
+yet reasonably challenging for
+program synthesizers.
+The AIC (Akaike Information Criterion) is a model comparison criteria 
+similar to the BIC (Bishop 2006: ``Pattern Recognition and Machine Learning'' has an overview),
+which introduces the counting norm term in Eq.~3.
+The $\theta_{var}$ parameter
+is the probability that
+a program sampled from the prior will
+have a variable at a particular node it syntax tree -- this is specified in Algorithm 1 of the supplement,
+but we admit that this is quite buried,
+so we will include a pointer to this part of the supplement in the revised manuscript.
+We compare our full system against an ablation 
+without the recognition model,
+finding that
+the full system solves more programming tasks and takes less time to
+solve those tasks.
+
+\textbf{FlashFill:} Reviewer 2 suggests comparing with FlashFill, the state-of-the-art program synthesizer for text editing.
 We very much would like to do this comparison, but this comparison
@@
 we achieve results comparable to the 2017 winning solver (CVC4),
-when CVC4 was provided with hand-engineered DSLs (Section 3.2 of paper).
-\subsubsection*{Related Work}
+whereas CVC4 was provided with hand-engineered DSLs (Section 3.2 of paper).
+%\subsubsection*{Related Work}
 
-We thank reviewer three for suggesting points of contact between our algorithm and the broader program induction literature.
+\textbf{Related Work:} We thank reviewer three for suggesting points of contact between our algorithm and other work in program induction.
 We are revising our paper to prominently cite the work of Schmid and Kitzelmann,
@@
 More broadly,
-we see these pieces of related work as
-orthogonal and complementary to
+we see these  related works as
+ complementary to
 our system:
@@
 \textsc{Igor}, Sketch (Solar-Lezama 2008), $\lambda^2$ (Feser et al. 2015), \textsc{MagicHaskeller},
-or many other synthesis algorithms.
+genetic programming, or other synthesis algorithms.
 We chose however to
@@
 
-This also relates to an insightful
-and suggestion for comparison by reviewer 2: FlashFill's DSL is fixed and
-manually curated and enhanced for a specific class of problems, over which is
-performs well at the loss of agility.
+%% This also relates to an insightful
+%% and suggestion for comparison by reviewer 2: FlashFill's DSL is fixed and
+%% manually curated and enhanced for a specific class of problems, over which is
+%% performs well at the loss of agility.
 
-\subsubsection*{More specific questions}
+%j\subsubsection*{More specific questions}
 
-\paragraph*{Reviewer 2: Origin of list task} All the list tasks were manually
-created in order to be interesting, challeging, but also human-interpretable for
-a prior behaviour experiment. They pre-existed our algorithm.
-
-\paragraph*{Reviewer 2: assesment of the recognition model} The recognition
-model's contribution was measured on lesions blabla
-
-\subsubsection*{Reproducibility}
-Our source code is publicly available on GitHub,
-and a link will be added upon publication (in NIPS or elsewhere),
-along with the specific command lines  needed to run
-the experiments in the paper.
+%% \paragraph*{Reviewer 2: Origin of list task} All the list tasks were manually
+%% created in order to be interesting, challeging, but also human-interpretable for
+%% a prior behaviour experiment. They pre-existed our algorithm.
 

### 2018-07-31 nips_2018_author_response.tex
small edits to response

@@
 SCC was provided with a minimal subset of primitives in 1959 Lisp, crucially including primitive recursion,
-and tasked with solving a small set of functional programming exercises, like appending lists or summing the elements of a list.
+and tasked with solving a moderate set of functional programming exercises, like appending lists or summing the elements of a list.
 After 93 hours on 64 CPUs,
@@
 and a handful of arithmetic operations.
+(We also tried the exact experiment Reviewer 1 suggests---removing only \texttt{length}---and confirmed that SCC rediscovers \texttt{length} when trained on our list processing domain.)
 Although this result is suggestive,
@@
 and one could replace our enumerative synthesizer with
-\textsc{Igor}, Sketch (Solar-Lezama 2008), $\lambda^2$ (Feser et al. 2015), \textsc{MagicHaskeller},
+\textsc{Igor}, Sketch (Solar-Lezama 2008), \textsc{MagicHaskeller},
 genetic programming, or other synthesis algorithms.

### 2018-07-31 nips_2018_author_response.tex
SCC typo

@@
   \caption{Three turtle graphics tasks: the task is to write a program that draws the image.
-Initially SSC solves the leftmost task without using a loop, but after solving the middle and rightmost tasks,
+Initially SCC solves the leftmost task without using a loop, but after solving the middle and rightmost tasks,
   it constructs a parametric `spiral' subroutine, which uses a loop -- on subsequent iterations it solves the leftmost task using the `spiral' routine.}

### 2018-09-03 archive.tex
final poster
compression notes
edits to introduction

@@
 %\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
-
+\usepackage{booktabs} \usepackage{stmaryrd}% for professional tables
+\usepackage{stmaryrd}(
 \usepackage[section]{placeins}
@@
 \newcommand{\lowerBound}{\mathscr{L}}
+\newcommand{\denotation}[1]{{\llbracket #1 \rrbracket}}
 \newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
@@
 
-\section{Introduction}
+\section{Introduction: Expertise}
+
+Human learners acquire expertise in a wide range of domains: some of
+us become experts in calculus, or cooking, or biology, music, tennis,
+or software engineering, to name just a few, and every child develops
+expertise in natural language, intuitive physics~\cite{Pete?},
+intuitive psychology (theory-of-mind), and motor control.  This
+picture contrasts sharply with the current state of machine
+intelligence, where a machine is built to be an expert in a single
+domain, like boardgames~\cite{alphaGo}, medical
+diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
+proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
+Thus an outstanding challenge in the long-term program of building
+more humanlike machines is to develop an algorithm that, like people,
+autonomously acquires expertise across many different kinds of
+domains.
+
+We contribute a model of the development of expertise that combines
+two key ingredients.  First, an expert needs a sufficiently expressive
+knowledge representation.  Following a long tradition in cognitive
+science and AI, we represent knowledge as programs, which prior work
+has used to represent expertise in recognition and generation of
+handwriting and speech~\cite{lake2015human}, intuitive theories (of
+kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
+grammar and
+semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
+Second, atop this program representation, experts possess two kinds of
+domain expertise.  They have at their disposal a powerful, yet
+specialized, repertoire of concepts and abstractions: e.g., in
+architecture, these are concepts like `arch' or `foundation'; in
+software engineering these are libraries of code and domain-specific
+languages (DSLs).  Here our model represents knowledge as programs,
+and so we identify this kind of expertise with a DSL.  Experts also
+have knowledge of when and how to deploy these domain-specific
+concepts efficiently when solving new problems: e.g., mathematicians
+efficiently search the space of proofs, intuiting which lemmas are
+appropriate when; expert chefs intuit which compositions of
+ingredients are likely tasty, before they actually start cooking.  For
+our model, this aspect of expertise corresponds to the ability to
+quickly assemble new, useful programs out of its DSL.
+
+We integrate these ideas into a model called DreamCoder which acquires
+expertise through a novel kind of wake/sleep or `dream' learning. The
+model iterates through wake cycles -- where it solves problems by
+writing programs -- and a pair of sleep cycles: a sleep cycle that
+grows its DSL by replaying experiences from waking and consolidating
+them into new abstractions, and a sleep cycle that improves its
+knowledge of how to write programs by training a neural network on
+replayed experiences as well as `dreams', or samples, from its DSL.
+
+DreamCoder builds on multiple generations of AI research, going back
+to the 1960's~\cite{solomonoff1964formal} when program-learning was
+proposed as a paradigm for general AI.  Broadly speaking recent work
+has either developed neural approaches for learning to efficiently
+deploy a fixed
+DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
+developed symbolic approaches for representing and searching through
+spaces of
+programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
+We were motivated by approaches that learn or grow the
+DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
+Our goal with DreamCoder is to show that the combination of
+neurally-guided search and DSL learning is a uniquely powerful way of
+building systems that, like human learners, autonomously acquire the
+expertise needed to navigate a new domain of problems.
+
+\section{Introduction: AI/PL/dreaming}
+
+An age-old dream within AI is a machine that learns and reasons by
+writing its own programs.  This vision stretches back to the
+1960's~\cite{solomonoff1964formal} and cuts to the core of much of
+what it would take to build machines that learn and think like humans.
+Computational models of cognition often explain the flexibility and
+richness of human thinking in terms of program learning: from everyday
+thinking and problem solving (motor program induction as an account of
+recognition and generation of handwriting and
+speech~\cite{lake2015human}; functional programs as a model of natural
+language semantics~\cite{SOMETHING}) to learning problems that unfold
+over longer developmental time scales: the child's acquisition of
+intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
+natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
+just a few.  An outstanding challenge, however, is to engineer
+program-learners that display the same level of domain-generality as
+the humans they are meant to model.
+
+Recent program-learning systems developed within the AI and machine
+learning community are impressive along many dimensions, authoring
+programs for problem domains like drawing
+pictures~\cite{spiral,ellis2017learning}, transforming
+text~\cite{gulwani2011automating} and numerical
+sequences~\cite{balog2016deepcoder}, and reasoning over common sense
+knowledge bases~\cite{muggleton2015meta}.  These systems work in
+different ways, but typically hinge upon a carefully hand-engineered
+Domain Specific Language (DSL).  The DSL restricts the space of
+programs to contain the kinds of concepts needed for one specific
+domain.  For example, a picture-drawing DSL could include concepts
+like circles and spirals, and a DSL for numerical sequences could
+include sorting and reversing lists of numbers.  Modern systems also
+learn how to efficiently deploy the DSL on new
+problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
+unlike human learners -- do not discover the underlying system of
+concepts needed to navigate the domain.
+
+We contribute a program-induction system that learns the
+domain-specific concepts (DSL) while jointly learning how to use those
+concepts.  This joint learning problem models two complementary
+notions of domain expertise: (1) domain experts have at their disposal
+a powerful, yet specialized repertoire of concepts and abstractions
+(analogous to the DSL) while also (2) having accurate intuitions about
+when and how to use those concepts to solve new problems.
+
+Our system is called DreamCoder because it acquires these two kinds of
+expertise through a novel kind of wake/sleep or `dream'
+learning~\cite{hinton1995wake}, iterating through a wake cycle --
+where it solves problems by writing programs -- and a pair of sleep
+cycles: a sleep cycle that grows the DSL by replaying experiences from
+waking and consolidating them into new abstractions (modeled as
+fragments of code), and a sleep cycle that improves its knowledge of
+how to write programs by training a neural network based on replayed
+experiences as well as `dreams', or samples, from the DSL.
+
+DreamCoder builds on two lines of prior work.  One line of work
+considers the problem of learning new concepts, abstractions, or
+`options' from
+experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
+while the other line of work considers the problem of learning how to
+deploy those concepts
+efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
+goal with DreamCoder is to show that the combination of these ideas is
+uniquely powerful, and pushes us toward program-writing systems that,
+like human learners, autonomously acquire the expertise needed to
+solve a new class of problems.
+
+\section{Introduction: NIPS}
 
@@
 to desired outputs, based on observing example input-output pairs.
-People can induce programs flexibly across many different domains, often from just one or a few examples.  For instance, if
-shown that a text-editing program should map ``Jane Morris Goodall''
-to ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey''
-to ``R. E. Leakey''; if instead the first input mapped to
-``Dr. Jane'' or ``Goodall, Jane'', we might guess
-the latter should map to ``Dr. Richard'' or ``Leakey, Richard'',
- respectively.
+People can induce programs flexibly across many different domains,
+often from just one or a few examples.  For instance, if shown that a
+text-editing program should map ``Jane Morris Goodall'' to
+``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey'' to
+``R. E. Leakey''; if instead the first input mapped to ``Dr. Jane'' or
+``Goodall, Jane'', we might guess the latter should map to
+``Dr. Richard'' or ``Leakey, Richard'', respectively.
 
@@
   \end{figure}
+
 \begin{figure}
@@
 \hline
-cut&F &Moss Side \\ 
-control&CL &Burnage \\
-control&F & City Centre\\
-cut& PCFL& Brooklands \\
+1.14531&F&110.9 \\ 
+?&CL&163.2 \\
+1.29857&F&207.3\\
+?&PCFL&143.3 \\
 \multicolumn{3}{c}{Learned generative models:}\\
 \hline
-\scode{\textbackslash l*\textbackslash l} & \scode{((\textbackslash
-u\textbackslash u)*)\textbar F}& \scode{( \textbar\textbackslash
-u\textbackslash w*)*} \\
+ \scode{\textbackslash?\textbar(1\textbackslash .\textbackslash d+)}& \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{\textbackslash d\textbackslash d\textbackslash d\textbackslash.\textbackslash d} \\
  \multicolumn{3}{c}{Samples from synthesized generative models:}\\
 \hline
-ya&DQDF & Vr DR \\ 
-glrwfdcnc&F &BeF lKQ \\
-mgs&F& W\\
-piljnl& KI& kqBfZ 0 \\
-kj& F& ON\\
-zci& GL& Bttc \\
-sxpm& F& S \\
+1.61&DQDF& 343.8 \\ 
+?&F&241.2\\
+?&F&647.5\\
+1.2&KI&246.8\\
+1.5080987&F&728.6\\
+?&GL&029.3\\
+1.453&F&289.1\\
     \hline
@@
 $f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
-$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash
-u\textbackslash w*)*}\\
-$f_3(x) = f_2(\scode{space}) = $\scode{( \textbar\textbackslash
-u\textbackslash w*)*}\\
+$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}\\
+$f_3(x) =  f_2(\scode{space}) = $\scode{( \textbar\textbackslash u\textbackslash w*)*}\\
 \hline
@@
 \emph{ (equivalent to regex `plus')}\\
-$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash
-l*\textbackslash l}$\\
+$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash l*\textbackslash l}$\\
     \hline

### 2018-09-20 nips_2018.tex
got rid of old graphics content; deanonymized

@@
 % ready for submission
-\usepackage{nips_2018}
+\usepackage[final]{nips_2018}
 
@@
 
-\title{Search, Compress, Compile: Library Learning in Neurally-Guided Bayesian Program Learning}
+\title{Library Learning in Neurally-Guided Bayesian Program Learning}
 
@@
 \author{
-  David S.~Hippocampus\thanks{Use footnote for providing further
-    information about author (webpage, alternative
-    address)---\emph{not} for acknowledging funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  %% examples of more authors
+  Kevin Ellis\\MIT\And Lucas Morales\\MIT\And Mathias Sabl\'e Meyer\\France
+  \AND
+  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
   %% \And
@@
           \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
-          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
+          \pop{\code{$f(\ell) = $}\code{($f_0$ $\ell$ ($\lambda$ (x)}}\\
           \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
@@
           \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
+          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$)}}
         \end{tabular}
@@
           \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
+          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$ 0)}}
         \end{tabular}
@@
       % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
-      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
-      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
-      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
+      %% \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
+      %% \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
+      \popp{$f_0(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
       \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
-      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
+      \hspace{\helpSize}($f_0$: \emph{Higher-order filter function})\\
       %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
-      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
-      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
-      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
-      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
+      \popp{$f_1(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
+      \popp{\phantom{$f_1(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
+      \hspace{\helpSize}($f_1$: \emph{Maximum element in list $\ell$})\\
+      \popp{$f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
       \phantom{$f_1(\ell,$}
@@
 Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ in List Functions)}\label{initialExampleDSL}
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_0$ in List Functions)}\label{initialExampleDSL}
 \end{table}
@@
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
-\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments
-without training the recognition model on ``Helmholtz machine'' style data. We
-are rerunning this experiment and anticipate improved quantitative results for
-list processing --- notice the recognition model helps much more for text editing and symbolic regression.} &
+&
 %For each domain,

### 2018-09-20 supplement.tex
vfill + compile
Updated colors and compiling
More last minute teaking
got rid of old graphics content; deanonymized

--- 2018-05-18 supplement.tex
@@
 \maketitle
-
-\section{Learning abstract geometry}
-The brifly presented geometry project is a recent addition to the set of
-problems adressed with our method.
-
-The original DSL, while inspired by the Logo language~(\cite{abelson1974logo}) in the sense
-that ones controls a moving on-off point, was designed to be congitively
-plausible in the sense that some abstractions are baked in for curves and a few
-structures --- it was first described in~\cite{sablemeyer2017geom}.
-
-It is to be thought as a turtle with as internal state a position, a facing
-direction, a speed, an acceleration, an angular speed and an angular
-acceleration.
-
-The DSL is the following:
-
-\begin{center}
-\tt
-\begin{tabular}{rl}
-  \toprule
-  \textrm{\emph{name}}    & \textrm{\emph{type}} \\
-  \midrule
-  Concat    & \text{prog}      \to \text{prog}        \to \text{prog} \\
-  Repeat    & var option        \to \text{prog}       \to \text{prog} \\
-  Embed     & \text{prog}      \to \text{prog}                        \\
-  Define    & \text{var}        \to \text{prog}                       \\
-  Turn      & \text{var option} \to \text{prog}                       \\
-  Integrate & \text{var option} \to \text{bool}       \to             \\
-            & \text{var option} \to \text{var option} \to             \\
-            & \text{prog}                                             \\
-  \midrule
-  True      & bool                                                    \\
-  False     & bool                                                    \\
-  \midrule
-  Nothing   & \text{var option}                                       \\
-  Just      & \text{var}        \to \text{var option}                 \\
-  \midrule
-  Unit      & var                                                     \\
-  Name      & var                                                     \\
-  Next      & \text{var}        \to \text{var}                        \\
-  Prev      & \text{var}        \to \text{var}                        \\
-  Double    & \text{var}        \to \text{var}                        \\
-  Half      & \text{var}        \to \text{var}                        \\
-  Opposite  & \text{var}        \to \text{var}                        \\
-  \bottomrule
-\end{tabular}
-\end{center}
-
-In this context programs to be learned correspond to a single shape since they
-don't take an input. All tasks will ignore input and produces a single output.
-This differs from the previous examples where each target could be evaluated on
-various input.
-
-Some elements of the semantics are common, the others are as follow.
-\texttt{Repeat} takes a \texttt{variable} and a \texttt{prog} and repeats
-said \texttt{prog} $n$ times where $n$ is the evaluation of the
-\texttt{variable} --- if not set it is defaulted to two. \texttt{Embed} of a
-\texttt{prog} means that said \texttt{prog} will be executed and then returns to
-the current state --- leaving what has been drawn in the meantime on the canvas.
-
-\begin{wrapfigure}{r}{0.25\textwidth}
-  \begin{center}
-    \includegraphics[width = 0.25\textwidth]{figures/DSLGeom.eps}
-  \end{center}\label{dslgeom}\caption{Top: Some defaults for \texttt{Integrate}.
-  Bottom, the result of the \texttt{Cross} example bellow.}
-\end{wrapfigure}
-
-\texttt{Integrate} is the main instruction and the only one that draws anything:
-\texttt{Integrate(t,p,a,c)} takes a time \texttt{var}, a pen \texttt{bool}, an
-acceleration \texttt{var} and an angular speed \texttt{var}, and it moves the
-turtle according to these parameters --- with or without actually drawing
-depending on the pen variable \texttt{p}.
-
-The default values of these parameters are set such that \texttt{Integrate(nothing,
-true, nothing, nothing)} draws a unit segment, \texttt{Integrate(nothing, true,
-nothing, Just(Unit))} draws a circle of unit length, and
-\texttt{Integrate(nothing, true, Just(Unit), Just(unit))} draws the first spire
-of a spiral. Playing with the first arguments decides on the duration during
-which the arguments are integrated.
-
-In the \texttt{var} type, most are self explicit, while \texttt{Opposite(v)}
-takes the opposite of \texttt{v}. The behaviour of \texttt{Name} was originally
-designed to handle arbitrary variables in a call by name fashion but was latter
-reduced to a single storage location, which proves to be enough for the
-targeted shapes. One can therefore store elements in that placeholder using
-\texttt{Define} and retrieve it through \texttt{Name}.
-
-Because of these defaults, the following program draws a cross:
-
-\[
-  \text{\texttt{Cross = Repeat(Double(Double(Unit)), Concat(Embed(Integrate), Turn(None)))}}
-\]
-
-Describing highly regular complex shapes in this language is easy to do as a
-human but quickly escapes the reach of naïve enumeration search. By using a
-curriculum of shapes our approach compresses the search in the corresponding
-directions --- another way to say this is that upon being given a dataset of
-shapes, it picks up the simple ones and abstract them as building blocks for
-latter staged of search.
-
-For example, the simplest way to draw a square in this language is already of
-length 12. Placing several around, for example to draw a grid is out of reach of
-the initial search. However by first abstracting as a primitive the segment ---
-thus reducing the length by four ---, then assuming that after a segment it
-often needs to draw something else, then abstracting the first arguments of
-\texttt{Repeat} --- further reducing the length by two --- as well as the one of
-\texttt{Turn} and finally making the square a primitive on its own once it
-starts using it often enough, the length of this particular shape drops.
-
-\begin{table}[htbp]
-  \begin{center}
-    \begin{tabular}{rcl}
-      \toprule
-      New Primitive & Type & Definition \\
-      \midrule
-      Segment = $f_0$ & \texttt{prog} & \texttt{Concat(Nothing, True, Nothing, Nothing)} \\
-      Right-angle = $f_1$ & \texttt{prog} & \texttt{Turn(Nothing)} \\
-      RepeatTwice = $f_2$ & \texttt{prog} $\to$ \texttt{prog} & $\lambda p.$ \texttt{Repeat(nothing,} $p$) \\
-      AddToUnitSegment = $f_3$ & \texttt{prog} $\to$ \texttt{prog} & $\lambda p.$ \texttt{Concat(}$f_0, p$) \\
-      Square = $f_4$ & \texttt{prog} & $f_3(f_3(f_2(f_1)))$ \\
-      \bottomrule
-    \end{tabular}
-  \end{center}
-  \caption{How our method compresses the square step by step. On the example
-  given in the main article this was produced by the compressor after the second
-  search phase and leads to the second jump in success, the first one being the
-  abstraction of the Segment. Names are \emph{not} produced by the compressor
-  and are here as indication to help the reader.}
-\end{table}
-
-The underlying hypothesis is that the new primitive distort the space of search
-toward something that looks more like what human actually produce and moves away
-from semantically valid but meaningless programs --- in a sense, learns to care
-about what \emph{matters} rather than what is \emph{true} in a very
-pragmatic-like way.
-
-In Figure \ref{alltasksgeom} are listed all the shapes used for this project
-without particular order.
-
-\begin{figure}[htp]
-\centering
-\includegraphics[width = \textwidth]{figures/geomAllTasks.png}
-\caption{The set of tasks for the geometry domain}
-\label{alltasksgeom}
-\end{figure}
-
-\begin{figure}[htp]
-  \begin{center}
-  \includegraphics[width = 0.75\textwidth]{figures/moreDreams.png}
-  \end{center}
-\caption{Another example of compiled figures generated by our method.}\label{moreDreams}
-\end{figure}
-
-In Figure~\ref{moreDreams} is another montage with several compiled shapes generated by our
-method after some training. While not all are that regular, often offer high
-level structure in the latter run while no structure is to be found at the
-begining.
-
-Since this is a first step in broader project of program induction for abstract
-geometry the likelyhood is currently all-or-none --- ongoing work moves this to
-a neural net based distance function to abstract away from noise in the shapes.
-
-The result presented in the main article desribe a sample of tasks and compiled
-new shapes on the top row. On the bottom row is a condesed description of a run
-where are displayed the mean of typical compiled programs on both sides, once
-before any training and once after the last iteration --- note how the
-probability weighting shaped the output space to add structure. In the middle is
-the learning rate measured by holding some tasks out, training on all the other,
-and with each iteration measuring the success rate on the unseen tasks. In the
-given example the split was $25\%$ test and $73\%$ train.
-
-\newpage
-
-\section{An Illustration of the 3 iterations of our algorithm}
-Below we diagram the iterations employed by our algorithm. At each stage of the algorithm,
-we have shaded the observed variables in gray and left the unobserved variables white.
-Black lines correspond to a connection from the top-down
-generative model, while red lines correspond to connections from the bottom-up recognition model.
-
-\begin{figure}[h]
-\centering
-  \includegraphics{figures/iterations.eps}
-%\begin{tikzpicture}
-  %\begin{scope}[shift={(2.5,0)}]
-    %\node[obs] at (3,3) (dx){$\mathcal{D}$};
-    %\node[latent] at (3.5,1.75) (zp){$p$};
-    %\node[obs] at (4,3) (tx){$\theta^{(x)}$};
-    %\node[obs] at (3.5,0.7) (xp) {$x$};
-    %\node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Search: Infer $p$};
-    %\draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
-    %\draw [->,red] (tx) -- (zp);
-    %\draw [->] (dx) -- (zp);
-    %\node at (nn) {
-      %\begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        %\tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        %\fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        %\node[rectangle] at (1,1) {};
-        %\foreach \name / \y in {1,...,4}
-            %\node[neuron] (I-\name) at (0,-\y) {};
-        %\foreach \name / \y in {1,...,3}
-            %\node[neuron] (H-\name) at (1,-\y-0.5) {};
-        %\foreach \name / \y in {1,...,4}
-            %\node[neuron] (O-\name) at (2,-\y) {};
-        %\foreach \source in {1,...,4}
-            %\foreach \dest in {1,...,3}
-                %\draw [-latex] (I-\source) -- (H-\dest);
-        %\foreach \source in {1,...,3}
-            %\foreach \dest in {1,...,4}
-                %\draw [-latex] (H-\source) -- (O-\dest);
-      %\end{tikzpicture}
-    %};
-    %\node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-  %\end{scope}
-
-  %\begin{scope}[shift={(0.7,-4.2)}]
-    %\node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-    %\draw [->] (p2.south) -- (x2.north);
-    %\draw [->] (dt.south) -- (p2.north);
-
-    %\node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-    %\draw [->] (p1.south) -- (x1.north);
-    %\draw [->] (dt.south) -- (p1.north);
-
-    %\node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-    %\node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-    %\draw [->] (p3.south) -- (x3.north);
-    %\draw [->] (dt.south) -- (p3.north);
-
-    %\node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-    %\node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
-
-    %\node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-    %\node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-    %\draw [->] (zp.south) -- (xp.north);
-    %\plate {}{(zp)(xp)}{$x\in X$};
-
-    %\draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
-    %rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
-    %\node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
-  %\end{scope}
-
-  %\begin{scope}[shift={(7.7,-4)}]
-    %\node[latent] at (0.5,3) (d){$\mathcal{D}$};
-    %\node[latent] at (1.5,3) (t){$\theta$};
-    %\node[obs] at (1,1.75) (z){$p$};
-     %\node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-    %\node[obs] at (1,0.5) (x) {$x$};
-    %\edge {z}{x};
-    %\edge {d,t}{z};
-    %\plate {}{(z)(x)}{$x\in X$};
-    %\node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
-  %\end{scope}
-
-%\end{tikzpicture}
-\end{figure}
 
@@
 
-\vfill
-
 \section{List Processing Data Set}
@@
 
-\subsection{Geometry}
-\begin{lstlisting}
-#(var_half var_name)
-#(var_double var_name)
-#(var_next #(var_half var_name))
-#(concat (turn (just #(var_half var_name))))
-#(@$\lambda$@ (integrate nothing $0 nothing nothing))
-#(@$\lambda$@ (integrate $0 true nothing (just var_name)))
-#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1)))))
-#(integrate (just #(var_half var_name)) true nothing nothing)
-#(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)
-#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))
-#(@$\lambda$@ (@$\lambda$@ (integrate (just #(var_double var_name)) true $0 (just $1))))
-#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))
-#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)))
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2)))))
-#(#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) nothing)
-#(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name))))
-#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0)))
-#(concat #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name)))))
-#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (embed $0)))
-#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false)))))
-#(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn $0) nothing nothing))
-#(@$\lambda$@ (#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (integrate $0 false nothing nothing)))
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit))))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (embed (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))))
-#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit)))))) nothing)
-#(@$\lambda$@ (concat (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true) (turn nothing)) (integrate (just $0) true nothing nothing)))
-#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0)))
-#(#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) nothing)
-#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (integrate nothing true nothing (just #(var_half var_name))) nothing)
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (integrate nothing $0 nothing $1) $2 nothing))))
-#(@$\lambda$@ (@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) $1)))))
-#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false) (just var_name) nothing)
-#(#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) (just #(var_half var_name)))
-#(#(@$\lambda$@ (concat $0 (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name)))))
-#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) (just (var_double $0))))
-#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (define #(var_half var_name)) (just #(var_half var_name)) $0)))
-#(#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing))
-#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (concat (define $0) #(integrate (just #(var_half var_name)) true nothing nothing))))
-#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit)))))) nothing (just (var_half $0)) false)))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 #(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name))))))
-#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (embed $0))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just $0))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) $0)))
-#(embed (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 #(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name)))))) nothing))
-#(@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (embed (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))) $0)))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (integrate nothing $0 nothing $1) $2 nothing)))) (just #(var_half var_name)) (just $0) true))
-#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) $0)))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn nothing) nothing $0))) nothing)))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (turn $0) nothing nothing)) (just #(var_half var_name)))))
-#(#(@$\lambda$@ (@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) $1))))) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false) true)
-#(repeat nothing (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) (integrate $1 true $2 (just var_unit)))))) nothing (just (var_half $0)) false))) #(var_half var_name)))
-#(#(@$\lambda$@ (@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (concat (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) $0) $1))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing) false)
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 (embed (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true)))) nothing) (just $0) nothing))
-#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing (#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) $0))) (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) $0)))
-#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (repeat nothing (concat (integrate $0 true nothing $1) $2))))) #(#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) nothing) nothing $0)))
-#(repeat nothing (concat #(#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)) #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name))))))
-#(@$\lambda$@ (repeat (just $0) #(repeat nothing (concat #(#(@$\lambda$@ (repeat nothing (repeat nothing (concat (embed $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) false))))) (#(@$\lambda$@ (integrate $0 true nothing (just var_name))) nothing)) #(#(@$\lambda$@ (integrate $0 true nothing (just var_name))) (just (var_half #(var_half var_name))))))))
-#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (repeat nothing (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) nothing $0))) (embed $0))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (repeat nothing (concat $0 (turn $1))))) $0 #(embed (#(@$\lambda$@ (concat (turn $0) (#(@$\lambda$@ (integrate nothing $0 nothing nothing)) true))) (just #(var_half var_name)))))) (just $0))))
-\end{lstlisting}
-
 \bibliography{main}

### 2018-09-22 nips_2018.tex
ec2

@@
 
-\newcommand{\system}{\textsc{SCC} }
-\newcommand{\systemEnding}{\textsc{SCC}}
+\newcommand{\system}{\textsc{EC$^2$} }
+\newcommand{\systemEnding}{\textsc{EC$^2$}}
 
@@
 \author{
-  Kevin Ellis\\MIT\And Lucas Morales\\MIT\And Mathias Sabl\'e Meyer\\France
+  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay
   \AND
@@
 
-Our \system (\textbf{S}earch/\textbf{C}ompress/\textbf{C}ompile)
+Our \system (ECC, for \textbf{E}xplore/\textbf{C}ompress/\textbf{C}ompile)
 algorithm incorporates these insights by iterating through three steps. 
-The \textbf{Search} step takes a given set of \textbf{tasks}, typically
-several hundred, and searches for compact programs that solve these tasks
+The \textbf{Explore} step takes a given set of \textbf{tasks}, typically
+several hundred, and explores the space of programs, searching for compact programs that solve these tasks,
 guided by the current DSL and neural network.
@@
 write programs in the domain; it modifies the structure of the DSL by
-discovering regularities across programs found during search, compressing
+discovering regularities across programs found during exploration, compressing
 them to distill out common code fragments across successful programs.
@@
 \system does approximate MAP inference  by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t.\ the frontiers (Search) and the DSL (Compression):
-\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
+alternating maximization w.r.t.\ the frontiers (Explore) and the DSL (Compression):
+\\\noindent \textbf{Explore: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
 want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
@@
 
-\subsection{Searching for Programs}\label{explorationSection}
+\subsection{Exploration: Searching for Programs}\label{explorationSection}
 
@@
   \% solved&\textbf{79\%}\footnotemark & 
-  76\% &71\%&35\%&62\%&37\%\\
-  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
+  94\% &71\%&35\%&62\%&37\%\\
+  Solve time&  88s&6s&11s&35s&44s&20s\\
   \midrule
@@
 but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.
 &

### 2018-09-22 supplement.tex
really terrible overview of fragment grammars

@@
 
-\subsection{Structure Learning}
+\subsection{Fragment grammars}
 
-We use Alg.~\ref{grammarInductionAlgorithm} to search for the structure of the DSL that best explains the frontiers.
+`Fragment grammars'~\cite{tim} is a formalism from computational linguistics
+developed for the purpose of modeling
+the reuse of structure
+in natural language.
+Here,
+we will adapt the fragment grammar formalism
+to model the reuse of structure in
+a formal language ($\lambda$ calculus),
+interpreting reuse of structure as subroutines.
+
+In the original fragment grammar formalism,
+a fragment grammar
+contains a probabilistic CFG called its \emph{base grammar};
+a \emph{fragment} is a tree drawn from the base grammar,
+but which can contain nonterminal symbols.
+As a motivating example,
+consider the following base grammar:
+\begin{align*}
+  S\to&(+ \;S\; S)\\
+  S\to&(\times \;S  \;S)\\
+  S\to&0\\
+      S\to&1  
+\end{align*}
+An example expression drawn from this grammar is $(+ \;1\; (\times \;0 \;0))$.
+An example \emph{fragment} drawn from this grammar $(+ \;1\; (\times \;S\; S))$.
+
+When extending fragment grammars to $\lambda$-calculus we write nonterminal symbols as free variables,
+and use the (current) DSL as our base grammar.
+For example, if our DSL includes addition, multiplication, and the numbers zero \& one (as in our example base grammar),
+then a possible  $\lambda$-fragment would be $(+ \;1\; (\times \;x\; y))$.
+
+In order to use fragments
+for DSL induction,
+we need several
+pieces: (1) a procedure for proposing fragments from programs found in the frontiers (described in figure~\ref{extractFragments});
+(2) a procedure for evaluating
+the likelihood of a program given that a fragment
+has been added to the DSL (defined in figure~\ref{match});
+and (3) a procedure for converting a fragment to a subroutine (i.e., a closed, typed $\lambda$-calculus expression; described in Figure~\ref{close}).
+Putting these procedures together, Algorithm~\ref{grammarInductionAlgorithm}
+is used to induce a new DSL.
+
+%\vfill\pagebreak
+\begin{figure}  
+  \begin{minipage}[c]{\textwidth}\toprule 
+    \begin{align*}
+  \text{fragments}(\lambda z.e)& = \text{fragments}'(\lambda z.e)\cup \text{fragments}(e)\\
+  \text{fragments}(f \; x)& = \text{fragments}'(f \; x)\cup \text{fragments}(f)\cup\text{fragments}(x)\\
+  \text{fragments}(e)& = \varnothing\text{, if $e$ is a variable or primitive.}\\
+  \\
+  \text{fragments}'(\lambda z.e)& = \left\{\lambda z.e' | e'\in \text{fragments}'(e) \right\}\cup \left\{v \right\}\text{, $v$ a free variable}\\
+  \text{fragments}'(f\; x)& = \left\{f'\; x' | f'\in \text{fragments}'(f),\;x'\in \text{fragments}'(x) \right\}\cup \left\{v \right\}\text{, $v$ a free variable}\\
+    \text{fragments}'(e)& = \left\{e\right\}\cup \left\{v \right\}\text{, $v$ a free variable, $e$ a variable or primitive}
+\end{align*}\bottomrule 
+    \end{minipage}
+\caption{Procedure for extracting possible fragments from a program.
+To prevent an explosion in the number of fragments, we also cap the number of free variables in a fragment to 3.}\label{extractFragments}
+\end{figure}
+\begin{figure}
+  \toprule \begin{minipage}[c]{\textwidth}
+    \begin{align*}
+      \text{closeFragment}(e)& = \text{closeFragment}(\lambda z.e)\text{, if $z$ is a free variable in $e$}\\
+      \text{closeFragment}(e)& = e\text{, if $e$ has no free variables.}
+    \end{align*}
+    \bottomrule 
+  \end{minipage}
+  \caption{Procedure for converting a fragment to a subroutine.}\label{close}
+\end{figure}
+
+\begin{figure}
+  \toprule \begin{minipage}[c]{\textwidth}
+    \begin{align*}
+      \probability[p|\mathcal{D},\theta]& = P_{\mathcal{D},\theta} (p|\tau,\varnothing )\text{, where $p:\tau$}\\
+      P_{\mathcal{D},\theta} (\lambda z.p|\alpha\to\beta,\mathcal{E})& =
+      P_{\mathcal{D},\theta}(p|\beta,[\tau\mapsto\alpha]\mathcal{E})\text{, where $\tau$ a fresh type variable}\\
+      P_{\mathcal{D},\theta}(e|\tau,\mathcal{E})& = \underbrace{\sum_{f,xs\in \text{parse}(e)} P'_{\mathcal{D},\theta}(f,xs|\tau,\mathcal{E})}_{\text{Marginalizing out how $\mathcal{D}$ produced $e$}}\text{, when $e$ not a $\lambda$-abstraction}\\
+      \\
+      P'_{\mathcal{D},\theta}(f,xs|\tau,\mathcal{E})& = \underbrace{\sum_{f':\tau_1\to\tau_2\to\cdots\to\tau\in \mathcal{D}}\frac{\theta_f}{Z_{\tau,\mathcal{D},\theta,\mathcal{E}}}
+        \indicator[\text{match}(f',f)\not=\bot]}_{\text{Marginalize over which fragment produced $f,xs$}}\times\\&\prod_{n=1}^{|xs|}P_{\mathcal{D},\theta}(x_n|\tau_n,\mathcal{E})\prod_{e,\tau'\in \text{match}(f',f)}P_{\mathcal{D},\theta}(e|\tau',\mathcal{E})\\
+      \\
+      \text{match}(\lambda z.b,\lambda z.b')& = \text{match}(b,b')\\
+      \text{match}(f\;x,f'\;x')& = \text{match}(f,f')\cup\text{match}(x,x')\\
+      \text{match}(p,p)& = \varnothing \text{, $p$ a primitive or variable bound in fragment}\\
+      \text{match}(v,p)& = \left\{v\mapsto p   \right\} \text{, $v$ a variable free in fragment}\\
+      \text{match}(e,e')& = \bot\text{, otherwise.}
+      \\
+      \\
+      Z_{\tau,\mathcal{D},\theta,\mathcal{E}}& = \underbrace{\sum_{f:\tau_1\to\tau_2\to\cdots\to\tau\in \mathcal{D}\cup\mathcal{E}}\theta_f}_{\text{Normalizing constant for type $\tau$}}\\\\
+      \text{parse}(f\;x)& = \left\{f',xs+[x] \;|\; f',xs\in \text{parse}(f) \right\}\cup\left\{f\;x \right\}\\
+      \text{parse}(e)& = \left\{e,[] \right\} \text{, $e$ not an application}
+    \end{align*}
+  \end{minipage}
+  \bottomrule
+  \caption{Procedure for calculating likelihood of a program after a new fragment has been added to the DSL.
+For simplicity we omit typing context and reuse type variables to indicate unification, as is commonly done in e.g.
+Prolog.}\label{match}
+  \end{figure}
 
@@
   \centering
-  \begin{minipage}{0.675\textwidth}    
+  \begin{minipage}{\textwidth}    
     \begin{algorithm}[H]
@@
          \WHILE {true}
-         \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\text{ a fragment  of }p\}$
+         \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\in \text{fragments}(p)\}$
          \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
-         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
+         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\left\{\text{closeFragment}(s)|s\in \mathcal{D} \right\},\theta^*(\mathcal{D})$
          \STATE $\mathcal{D}\gets\mathcal{D}'$

### 2018-09-22 supplement.tex
minor change to last commit
author list on supplement

@@
 \author{
-  David S.~Hippocampus\thanks{Use footnote for providing further
-    information about author (webpage, alternative
-    address)---\emph{not} for acknowledging funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  %% examples of more authors
+  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay
+  \AND
+  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
   %% \And
@@
       \text{match}(p,p)& = \varnothing \text{, $p$ a primitive or variable bound in fragment}\\
-      \text{match}(v,p)& = \left\{v\mapsto p   \right\} \text{, $v$ a variable free in fragment}\\
+      \text{match}(v,p)& = \left\{v\mapsto p   \right\} \text{, $v$ a variable free in fragment, $p$ not containing free variables bound in fragment}\\
       \text{match}(e,e')& = \bot\text{, otherwise.}

### 2018-09-26 nips_2018.tex
acknowledgments and extra citation

@@
 algorithm, but researchers in programming languages and AI have built
-successful program induction algorithms for many applications, 
-such as handwriting recognition and generation~\cite{lake2015human},
-procedural graphics~\cite{ellis2017learning}, question
+successful program induction algorithms for many applications, such as
+handwriting recognition and generation~\cite{lake2015human},
+procedural graphics~\cite{ellis2017learning}, cognitive
+modeling~\cite{DBLP:journals/cogsr/SchmidK11}, question
 answering~\cite{johnson2017clevr} and robot motion
-planning~\cite{devlin2017neural}, to name just a few.
-These systems work in different ways, but most hinge upon having a
-carefully engineered \textbf{Domain Specific Language (DSL)}.  This is
+planning~\cite{devlin2017neural}, to name just a few.  These systems
+work in different ways, but most hinge upon having a carefully
+engineered \textbf{Domain Specific Language (DSL)}.  This is
 especially true for systems such as FlashFill that aim to induce a
@@
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
 
@@
  Bayesian Program
- Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}).
+ Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10,ellis2015unsupervised}).
 Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
@@
 
+\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work.
+Funding from NSF GRFP and the MUSE program (Darpa grant FA8750-14-2-0242).
+
 \bibliographystyle{unsrt}

### 2018-10-02 pp.tex
probabilistic programming talk

--- 2018-07-24 talk.tex
@@
 \title{Growing libraries of concepts with wake-sleep program induction}
-\author{Kevin Ellis \& Mathias Sabl\'e Meyer\\Joint with: Lucas Morales, Armando Solar-Lezama, Joshua B. Tenenbaum\\Heavy inspiration from: Eyal Dechter}
+\author{Kevin Ellis\\Joint with: Lucas Morales, Mathias Sabl\'e Meyer, Armando Solar-Lezama, Joshua B. Tenenbaum\\Heavy inspiration from: Eyal Dechter}
 \institute{MIT} 
-%\date{November 2017}
+\date{October 2018}
 
@@
 
-\begin{frame}{Engineering the language of thought}
- \begin{tikzpicture}
-    \node () {\includegraphics[width = 15.6cm]{theory.png}};
-    \node at (-5,-4) {  Ullman et al 2012};
-    \end{tikzpicture}
-\end{frame}
+%% \begin{frame}{Engineering the language of thought}
+%%  \begin{tikzpicture}
+%%     \node () {\includegraphics[width = 15.6cm]{theory.png}};
+%%     \node at (-5,-4) {  Ullman et al 2012};
+%%     \end{tikzpicture}
+%% \end{frame}
 
@@
 \end{frame}
+
+\begin{frame}{DreamCoder --- Sleep-G (\textbf{Refactoring})}
+  \textbf{Learning higher-order \texttt{map} function}
+
+  \begin{tabular}{cc}
+    \toprule Task&Program\\\midrule
+    \begin{tabular}{l}
+      \texttt{(1 2 3)}$\to$\texttt{(2 4 6)}\\
+      \texttt{(1 9 2)}$\to$\texttt{(2 18 4)}\\
+    \end{tabular}&
+
+    \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (+ (car l) (car l))}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}\\\midrule
+    \begin{tabular}{l}
+      \texttt{(1 2 3)}$\to$\texttt{(2 3 4)}\\
+      \texttt{(1 9 2)}$\to$\texttt{(2 10 3)}\\
+    \end{tabular}&
+    \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (+ (car l) 1)}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+
+    \\\bottomrule
+  \end{tabular}
+
+     \visible<2->{
+      \centering
+      \begin{tabular}{l}
+          \texttt{\emph{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}
+      \end{tabular}
+}
+\end{frame}
+
+\begin{frame}{DreamCoder --- Sleep-G (\textbf{Refactoring})}
+  \textbf{Learning higher-order \texttt{map} function}
+
+  \begin{tabular}{cc}
+    \toprule Task&Program\\\midrule
+    \begin{tabular}{l}
+      \texttt{(1 2)}$\to$\texttt{(2 4)}\\
+      \texttt{(1 9)}$\to$\texttt{(2 18)}\\
+    \end{tabular}&
+\begin{tabular}{l}
+\texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(r (cdr l)))))))}}\\
+\texttt{ ($\lambda$ (z) (+ z z)))}
+      \end{tabular}
+
+    \begin{tabular}{l}
+    \end{tabular}\\\midrule
+    \begin{tabular}{l}
+      \texttt{(1 2)}$\to$\texttt{(2 3)}\\
+      \texttt{(1 9)}$\to$\texttt{(2 10)}\\
+    \end{tabular}&
+    \begin{tabular}{l}
+\begin{tabular}{l}
+\texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(r (cdr l)))))))}}\\
+\texttt{ ($\lambda$ (z) (+ z 1)))}
+      \end{tabular}
+    \end{tabular}
+
+    \\\bottomrule
+  \end{tabular}
+
+      \centering
+\begin{tabular}{l}
+          \texttt{\emph{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}
+      \end{tabular}
+
+\end{frame}
+
+\begin{frame}[plain]{}\footnotesize
+  \Wider[5.4em]{
+    \begin{tikzpicture}[every node/.style={inner sep=0,outer sep=0}]
+    \node(p1)[draw] at (-1,0) {
+      \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (+ (car l) (car l))}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+    };
+    \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-3cm]p1.south) {
+      \begin{tabular}{l}
+  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
+\texttt{ ($\lambda$ (z) (+ z z)))}
+      \end{tabular}
+    };
+
+        \node(p2)[draw] at ([xshift=3.5cm]p1.east) {
+    \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (+ (car l) 1)}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+
+    };
+    \node(r2)[draw] at ([yshift=-3cm]p2.south) {
+\begin{tabular}{l}
+  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
+\texttt{ ($\lambda$ (z) (+ z 1)))}
+      \end{tabular}
+
+    };
+
+    \draw [->] (p1.south)  --(r1.north) node[fill=white,midway] {refactor};
+    \draw [->] (p2.south)  --(r2.north) node[fill=white,midway] {refactor};
+
+    \node[draw](m) at (2,-7) {
+\begin{tabular}{l}
+          \texttt{\emph{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}
+      \end{tabular}      
+    };
+    \draw [->](r1.south)--(m.north);
+    \draw [->](r2.south)--(m.north);
+    \node[fill=white] at ([yshift=0.5cm]m.north) {\textbf{Compress (MDL/Bayes objective)}};
+    \visible<2-3>{
+      \node[fill=orange!50!white,align=center] at (r1) {\Huge $10^{14}$\\\Huge  refactorings };
+      \node[fill=orange!50!white,align=center] at (r2) {\Huge $10^{13}$\\\Huge  refactorings };
+    }
+    \visible<3>{
+      \node[fill=orange!50!white,align=center] at ([yshift=7cm]m.north) {\Huge version space: set of programs\\
+      \huge Lau 2003; Gulwani 2012};
+      }
+    \visible<4>{
+      \node[fill=orange!50!white,align=center] at (r1) {\Huge $\leq 10^6$\\\Huge version spaces };
+      \node[fill=orange!50!white,align=center] at (r2) {\Huge $\leq 10^6$\\\Huge version spaces };
+    }
+    \end{tikzpicture}
+    }
+  \end{frame}
+
 \begin{frame}{DreamCoder --- Sleep-R (\textbf{Experience Replay})}
@@
 \end{frame}
-
-%% \begin{frame}[t]{Wake: Problem solving}
-%%   \includegraphics[width = 11cm]{figures/animation/Wake.eps}
-
-%%   \vfill
-%%   \only<1>{
-%%   Recognition model: Neural net.\\
-%%   Red: Bottom-up inference.\\
-%%   Gray: Observed.\\
-%%   White: Latent.\\
-%%   Boxed (plate): Repeated.\\
-%%   }
-%%   \only<2>{
-%%   \begin{tikzpicture}[xscale=1.25,line width=1mm]
-%%     \node at (0,0) (d){DSL};
-%%     \node at ([yshift = -2cm]d) (t){$\text{Task}$};
-
-%%     \node at ([xshift = 2cm]t) (nn){
-%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.3,shift={+(-1,2.5)}}]
-%%        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-%%         \node[rectangle] at (1,1) {};
-%%         \foreach \name / \y in {1,...,4}
-%%             \node[neuron] (I-\name) at (0,-\y) {};
-%%         \foreach \name / \y in {1,...,3}
-%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
-%%         \foreach \name / \y in {1,...,4}
-%%             \node[neuron] (O-\name) at (2,-\y) {};
-%%         \foreach \source in {1,...,4}
-%%             \foreach \dest in {1,...,3}
-%%                 \draw [-latex] (I-\source) -- (H-\dest);
-%%         \foreach \source in {1,...,3}
-%%             \foreach \dest in {1,...,4}
-%%                 \draw [-latex] (H-\source) -- (O-\dest);
-%%       \end{tikzpicture}
-%%     };
-%%     \node[align = center, text width = 1cm] at ([yshift = 0.7cm]nn.north) {\baselineskip=0pt \Small Recog. model\par};
-%%     \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
-
-%%     \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
-%% %    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
-
-%%     \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-%%     \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
-
-%%     \node[align=left] at (7,-1) (f) {Programs};
-%%     \draw [->  ] (s.east) -- (f.west);
-
-%%     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-%%   \end{tikzpicture}}
-%%   %% Recognition model: Neural net.\\
-%%   %% Red: Bottom-up inference.\\
-%%   %% Gray: Observed.\\
-%%   %% White: Latent.\\
-%%   %% Boxed (plate): Repeated.\\
-
-%% \end{frame}
-%% \begin{frame}[t]{Sleep-G: Memory consolidation}
-%%   \includegraphics[width = 11cm]{figures/animation/Sleep-G.eps}
-%%   \vfill
-%%   \textbf{Fragment Grammars: O'Donnell 2015.}\\
-%%   Orange: Code fragments.\\
-%%   %% Gray: Observed.\\
-%%   %% White: Latent.\\
-%%   %% Boxed (plate): Repeated.\\
-
-%% \end{frame}
-%% \begin{frame}[t]{Sleep-R: Objective}
-%%   \includegraphics[width = 11cm]{figures/animation/Wake.eps}
-
-%%     \vfill
-%%     Recognition model predicts distribution over program, conditional on task.\\
-%%     Training: (program, task) pairs\\
-%%     Objective: Predict program w/ (1) high prior under DSL \& \\(2) high likelihood for task
-
-%% \end{frame}
-%% \begin{frame}[t]{Sleep-R: Experience replay}
-%%   \includegraphics[width = 11cm]{figures/animation/Sleep-R-1.eps}
-
-%%     \textbf{Train on (program, task) pairs found during waking}
-%% \end{frame}
-%% \begin{frame}[t]{Sleep-R: Dreaming}
-%%   \includegraphics[width = 11cm]{figures/animation/Sleep-R-2.eps}
-
-%%     \textbf{Train on (program, task) pairs sampled from DSL}
-
-%% \end{frame}
 
@@
 
-  Discovers 38 concepts, including `filter'. With suitable curriculum can also learn `map', `fold', etc.
+  Discovers 38 concepts, including `filter'. With different tasks will also learn `map', `fold', `unfold', etc. starting with 1950's Lisp
 \end{frame}
 
+\begin{frame}{Text editing}
+  In the style of FlashFill (Gulwani 2012)
+
+  \centering  \includegraphics[width = 5cm]{textColumn.png}
+
+\vspace{-0.5cm}  SyGuS problems: solves 3\% before learning, vs 75\% after learning. Best prior work: 80\%
+
+\end{frame}
 \begin{frame}{List functions \& Text editing: Learning curves on hold out tasks}
@@
 
-%% \begin{frame}{Text editing}
-%%   In the style of FlashFill (Gulwani 2012)
-
-%%   \centering  \includegraphics[width = 5cm]{textColumn.png}
-
-%%   SyGuS problems: solves 3\% before learning, vs 74\% after learning. Best prior work: 80\%
-
-%%   \end{frame}
-
-%% \begin{frame}{Symbolic regression from visual input}
-%% \centering\includegraphics[width = 5cm]{symbolicRegression.png}
-%% \end{frame}
+\begin{frame}{Symbolic regression from visual input}
+\centering\includegraphics[width = 5cm]{symbolicRegression.png}
+\end{frame}
 
@@
   }
-  \only<8-9>{
-    \begin{columns}[T]
-      \column{0.5 \textwidth}
-        \begin{mycode}
-          \texttt{for i in range($5\times\infty$)}\\
-          \texttt{> FW $\text{\texttt{i}}\times\varepsilon$}\\
-          \texttt{> RT} $\varepsilon$\\
-          \vspace*{3\baselineskip}
-        \end{mycode}
-      \column{0.5 \textwidth}
-        \includegraphics[width = 4cm]{figures/teachLogo/sspiral.eps}
-    \end{columns}
-  }
-  \visible<9>{
+  \visible<7>{
     \vspace{0.2cm}
@@
   \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
-  \pause{}
   \begin{itemize}
@@
 
-\begin{frame}{Turtle graphics --- All the way down}
+\begin{frame}{Turtle graphics --- Dreaming from learned generative model}
   \centering
-  \begin{center}
   \only<1>{
     \begin{figure}
-      \includegraphics[width=5cm]{figures/teachLogo/star5_black.eps}
+      \includegraphics[width=6cm]{figures/dreams/1.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_colour.eps}
+      \includegraphics[width=6cm]{figures/dreams/2.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq1.eps}
+      \includegraphics[width=6cm]{figures/dreams/3.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq2.eps}
+      \includegraphics[width=6cm]{figures/dreams/4.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq3.eps}
+      \includegraphics[width=6cm]{figures/dreams/6.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq4.eps}
+      \includegraphics[width=6cm]{figures/dreams/7.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq5.eps}
+      \includegraphics[width=6cm]{figures/dreams/8.png}
     \end{figure}
@@
     \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq1.eps}
+      \includegraphics[width=6cm]{figures/dreams/9.png}
     \end{figure}
   }
-  \only<9>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq2.eps}
-    \end{figure}
-  }
-  \only<10>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq3.eps}
-    \end{figure}
-  }
-  \only<11>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq4.eps}
-    \end{figure}
-  }
-  \only<12>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq5.eps}
-    \end{figure}
-  }
-  \only<13>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/star5_seq1.eps}
-    \end{figure}
-  }
-  \only<14>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp000.png}
-    \end{figure}
-  }
-  \only<15>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp002.png}
-    \end{figure}
-  }
-  \only<16>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp004.png}
-    \end{figure}
-  }
-  \only<17>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp006.png}
-    \end{figure}
-  }
-  \only<18>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp008.png}
-    \end{figure}
-  }
-  \only<19>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp010.png}
-    \end{figure}
-  }
-  \only<20>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp012.png}
-    \end{figure}
-  }
-  \only<21>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp014.png}
-    \end{figure}
-  }
-  \only<22>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp016.png}
-    \end{figure}
-  }
-  \only<23>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp018.png}
-    \end{figure}
-  }
-  \only<24>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp020.png}
-    \end{figure}
-  }
-  \only<25>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp022.png}
-    \end{figure}
-  }
-  \only<26>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp024.png}
-    \end{figure}
-  }
-  \only<27>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp026.png}
-    \end{figure}
-  }
-  \only<28>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp028.png}
-    \end{figure}
-  }
-  \only<29>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp030.png}
-    \end{figure}
-  }
-  \only<30>{
-    \begin{figure}
-    \includegraphics[width=5cm]{figures/teachLogo/spiral/seqSp032.png}
-    \end{figure}
-  }
-\end{center}
+\end{frame}
+\begin{frame}{Tower building in blocks world}
+  \large
+\centering  
+  Control a hand that puts down blocks \\
+  \hspace{1cm}(turtle: control a pen that puts down ink)
+\only<1>{\includegraphics[width = \textwidth]{figures/3t.png}}
+\only<2>{\includegraphics[width = 0.6\textwidth]{figures/every_tower.png}}
+
+  \end{frame}
+
+\begin{frame}{Tower building in blocks world: Learned concepts}
+  Parametric planning primitives. Example pyramid concept:
+
+  \includegraphics[width = \textwidth]{figures/pyramid.png}
+
+\end{frame}
+
+\begin{frame}{Tower building in blocks world: Learned concepts}
+  Parametric planning primitives. Example brickwall concept:
+
+  \includegraphics[width = \textwidth]{figures/brickwall.png}
+
   \end{frame}
@@
    \begin{itemize}
-   \item    Acquiring a domain-specific language-of-thought (DSL)
-     \item Learning  to use the language of thought (recognition model)
+   \item    Acquiring a domain-specific representation (DSL)
+     \item Learning  to write programs (recognition model)
    \end{itemize}

### 2018-10-02 pp.tex
progress on presentation

@@
 %Information to be included in the title page:
-\title{Growing libraries of concepts with wake-sleep program induction}
+\title{\texttt{DreamCoder}: Growing libraries of concepts with wake-sleep program induction}
 \author{Kevin Ellis\\Joint with: Lucas Morales, Mathias Sabl\'e Meyer, Armando Solar-Lezama, Joshua B. Tenenbaum\\Heavy inspiration from: Eyal Dechter}
@@
 
-\begin{frame}{The Language of Thought}
+\begin{frame}{Human program induction everywhere}
 \begin{tikzpicture}
-    \node at (-3,1) {\includegraphics[width = 6cm]{Fodor.jpg}};
-  \node at (2,2) {\includegraphics[width = 8cm]{Solomon.png}};
-%
+  \node(Lisp) at (0,0) {\includegraphics[width = 5cm]{figures/1975.png}};
+  \visible<2->{
+    \node(colorless) at (6,0) {\includegraphics[width = 5cm]{figures/colorless.png}};
+    \node at ([yshift=-2cm]colorless.south) {\includegraphics[width = 4cm]{figures/child.jpg}};
+  }
+  \visible<3->{
+    \node(aqueduct) at ([yshift=-1cm,xshift=0cm]Lisp.south) {\includegraphics[width = 5cm]{figures/aqueduct.jpg}};
+    \node(arch) at ([xshift=1cm,yshift=0cm]aqueduct.south) {\includegraphics[width = 1.5cm]{figures/arch.jpg}};
+  }
+  \visible<4->{
+    \node at ([yshift=-1cm,xshift=-2cm]aqueduct.south) {\includegraphics[width = 2cm]{figures/tapestry.jpg}};
+    \node(wiringDiagram) at ([yshift=-1cm,xshift=1.6cm]aqueduct.south) {\includegraphics[width = 5cm]{figures/wiringDiagram.jpg}};
+  }
+  \visible<5->{
+    \node at ([xshift=2cm,yshift=1cm]wiringDiagram.east) {\includegraphics[width = 5cm]{tree.jpg}};
+    }
   \end{tikzpicture}
+
 \end{frame}
@@
 
-\begin{frame}{Engineering the language of thought}
+\begin{frame}{Human program induction everywhere}
   \begin{tikzpicture}
@@
   Combines ideas from Wake-Sleep \& Exploration-Compression algorithm by Eyal Dechter
+\Wider[5.4em]{
+  \begin{tikzpicture}
+    \node at (6,0) {  \includegraphics[width = 4cm]{figures/sleepingChild.jpg}};
+    \visible<2->{
+      \node(Lisp) at (-3,0) {\includegraphics[width = 3cm]{figures/Lisp.png}};
+      \node(tower) at ([xshift=1.5cm]Lisp.east) {\includegraphics[width = 3cm]{figures/pyramid1.png}};
+      \node(logo) at ([xshift=1.5cm]tower.east) {\includegraphics[width =3cm]{figures/9logo.png}}; 
+    }
+    \end{tikzpicture}}
 
@@
 
-\begin{frame}{Symbolic regression from visual input}
-\centering\includegraphics[width = 5cm]{symbolicRegression.png}
-\end{frame}
+%% \begin{frame}{Symbolic regression from visual input}
+%% \centering\includegraphics[width = 5cm]{symbolicRegression.png}
+%% \end{frame}
 

### 2018-10-04 pp.tex
possibly final talk slides

@@
 \newcommand{\teal}[1]{\textcolor{teal}{#1}}
+\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
+\newcommand{\greenCode}[1]{{\footnotesize\popp{\code{#1}}}}
+\newcommand{\blueCode}[1]{{\footnotesize\pop{\code{#1}}}}
+
+\usepackage{pgf}
 
@@
 
-\begin{frame}{Growing a domain-specific language of thought}
+\begin{frame}{Growing domain-specific knowledge}
 
@@
 
-  \pause
   \vspace{1cm}
@@
   \begin{itemize}
-  \item Library of concepts (declarative knowledge)
+  \item Library of concepts (declarative knowledge; generative model over programs)
     \item Inference strategy (procedural knowledge)
   \end{itemize}
+  \only<2>{
+  \begin{tikzpicture}
+    \node(problem) at (0,0) {\includegraphics[width = 2cm]{figures/functions/48.png}};
+    \node(synthesizer)[draw,align=center] at ([xshift=3cm]problem.east) {Learned \\program inducer};
+    \draw[->] (problem.east) -- (synthesizer.west);
+    \node(program)[draw, align=center] at ([xshift=3cm]synthesizer.east) {program:\\$f(x) = 0.7x^2+1.2$};
+    \draw[->] (synthesizer.east) -- (program.west);
+  \end{tikzpicture}}
+    \renewcommand\code\texttt
+    \only<3>{
+\Wider[4em]{  \begin{tikzpicture}
+    \node[draw,align=center](problem) at (0,0) {
+            \begin{tabular}{l}
+          \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+          \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
+          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
+          \end{tabular}
+};
+    \node(synthesizer)[draw,align=center] at ([xshift=2.5cm]problem.east) {Learned \\program inducer};
+    \draw[->] (problem.east) -- (synthesizer.west);
+    \node(program)[draw, align=center] at ([xshift=2.5cm]synthesizer.east) {program:\\\code{(filter ($\lambda$ (x)}\\\code{ (> x 2)))}};
+    \draw[->] (synthesizer.east) -- (program.west);
+  \end{tikzpicture}
+
+  \vspace{0.5cm}Concepts: \code{filter}, etc\\Inference strategy: neurosymbolic search for programs
+    }}
+
 \end{frame}
@@
       \node(Lisp) at (-3,0) {\includegraphics[width = 3cm]{figures/Lisp.png}};
-      \node(tower) at ([xshift=1.5cm]Lisp.east) {\includegraphics[width = 3cm]{figures/pyramid1.png}};
+      \node[align=center](tower) at ([xshift=1cm]Lisp.east) {\includegraphics[width = 1cm]{figures/functions/627.png}\hspace{0.05cm}\includegraphics[width = 1cm]{figures/functions/615.png}\\
+        \includegraphics[width = 1cm]{figures/functions/382.png}\hspace{0.05cm}\includegraphics[width = 1cm]{figures/functions/0.png}%
+      };
       \node(logo) at ([xshift=1.5cm]tower.east) {\includegraphics[width =3cm]{figures/9logo.png}}; 
@@
 
+\begin{comment}
+
 \begin{frame}{DreamCoder --- Sleep-G (\textbf{Refactoring})}
   \textbf{Learning higher-order \texttt{map} function}
-
+  \Wider[5.4em]{
   \begin{tabular}{cc}
@@
     \\\bottomrule
-  \end{tabular}
+  \end{tabular}}
 
@@
   \textbf{Learning higher-order \texttt{map} function}
-
+    \Wider[5.4em]{
   \begin{tabular}{cc}
@@
     \begin{tabular}{l}
-      \texttt{(1 2)}$\to$\texttt{(2 4)}\\
-      \texttt{(1 9)}$\to$\texttt{(2 18)}\\
+      \texttt{(1 2 3)}$\to$\texttt{(2 4 6)}\\
+      \texttt{(1 9 2)}$\to$\texttt{(2 18 4)}\\
     \end{tabular}&
@@
     \begin{tabular}{l}
-      \texttt{(1 2)}$\to$\texttt{(2 3)}\\
-      \texttt{(1 9)}$\to$\texttt{(2 10)}\\
-    \end{tabular}&
+      \texttt{(1 2 3)}$\to$\texttt{(2 3 4)}\\
+      \texttt{(1 9 2)}$\to$\texttt{(2 10 3)}\\
+    \end{tabular}
+    &
     \begin{tabular}{l}
@@
 \end{frame}
+ \end{comment}
 
@@
 
-  Discovers 38 concepts, including `filter'. With different tasks will also learn `map', `fold', `unfold', etc. starting with 1950's Lisp
+  Discovers 38 concepts, including `filter'. %With different tasks will also learn `map', `fold', `unfold', etc. starting with 1950's Lispthe
+
+\begin{picture}(50,70) \put(250,0){\hbox{\includegraphics[width = 3cm]{figures/Lucas}}} \end{picture} 
+\end{frame}
+\begin{frame}{Learning the basics of programming}
+  \includegraphics[width = \textwidth]{figures/McCarthy.png}
+
+\centering  McCarthy 1959 Lisp  $\longrightarrow$ Modern functional programming
+
+  22 tasks. 64 CPUs. 93 hours.
+
   \end{frame}
@@
 
+\begin{frame}{Learned text processing DSL}
+  \only<1>{  \includegraphics[width = \textwidth]{figures/textPrimitives.pdf}}
+  \only<2>{  \includegraphics[width = \textwidth]{figures/textPrimitives.png}}
+
+  \end{frame}
+
 %% \begin{frame}{Symbolic regression from visual input}
@@
       Sablé--Meyer}}
+    \begin{picture}(50,50) \put(280,-20){\hbox{\includegraphics[width = 2cm]{figures/French}}} \end{picture} \vspace{-2cm}
 
@@
   }
+
 \end{frame}
@@
   \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
-  \begin{itemize}
-    \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
-    \item A line of length \texttt{n} followed with a right angle
-    \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
-    \item Unit line then teleport back to origin
-    \item ...
-  \end{itemize}
+  %% \begin{itemize}
+  %%   \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
+  %%   \item A line of length \texttt{n} followed with a right angle
+  %%   \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
+  %%   \item Unit line then teleport back to origin
+  %%   \item ...
+  %% \end{itemize}
+\end{frame}
+
+\begin{frame}{Turtle graphics --- Illustration of learned DSL}
+
+  \only<1>{\centering\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_4.png}}
+  \only<2>{\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_1.png}}
+  \only<3>{\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_7.png}}
+  \only<4>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_12.png}}
+  \only<5>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_14.png}}
+  \only<6>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_5.png}}
+  \only<7>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_13.png}}
+
   \end{frame}
@@
 
-\begin{frame}{Turtle graphics --- More dreams, 5 minutes, 1st iteration}
-  \centering
-  \includegraphics[width=8cm]{figures/dreams/5m_firstIteration.eps}
-\end{frame}
+%% \begin{frame}{Turtle graphics --- More dreams, 5 minutes, 1st iteration}
+%%   \centering
+%%   \includegraphics[width=8cm]{figures/dreams/5m_firstIteration.eps}
+%% \end{frame}
 
-\begin{frame}{Turtle graphics --- More dreams, 5 minutes, last iteration}
-  \centering
-  \includegraphics[width=8cm]{figures/dreams/5m_final.eps}
-\end{frame}
+%% \begin{frame}{Turtle graphics --- More dreams, 5 minutes, last iteration}
+%%   \centering
+%%   \includegraphics[width=8cm]{figures/dreams/5m_final.eps}
+%% \end{frame}
 
@@
 \end{frame}
+
+\begin{comment}
 \begin{frame}{Tower building in blocks world}
@@
 \end{frame}
+  \end{comment}
 

### 2018-10-06 pp.tex
last-minute tweaks

@@
 
-\begin{SaveVerbatim}[]{ListSketch}
-Bool ::= (<= Int) | (>= Int)
-       | (= Int)
-Int  ::= 0
-       | (+1 Int) | (-1 Int)
-       | (length List)
-       | (head List)
-List ::= nil | X
-       | (filter Bool List)
-       | (tail List)
-       | (list Int)
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{ListSketchFull}
-Program ::= (if (Bool Int)
-                List
-                (append RecursiveList
-                        RecursiveList
-                        RecursiveList))
-RecursiveList ::= List
-                | (recurse List)
-Bool ::= (<= Int) | (>= Int)
-       | (= Int)
-Int  ::= 0
-       | (+1 Int) | (-1 Int)
-       | (length List)
-       | (head List)
-List ::= nil | X
-       | (filter Bool List)
-       | (tail List)
-       | (list Int)
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{Reverse}
-(if (<= (+1 0) (length X))
-    X
-    (append (recurse (tail X))
-            (list (head X))))
-\end{SaveVerbatim}
-\begin{SaveVerbatim}[]{Reversep}
-(if (= (+1 0) (length X))
-    X
-    (append (recurse (filter (<= (+1 0)) (tail X)))
-            (list (head X))))
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{Count}
-(length (filter (= (head X)) (tail X)))
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{Countp}
-(-1 (length (filter (= (head X)) X)))
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{Sort}
-(if (<= (+1 0) (length X))
-    X
-    (append (recurse (filter (<= (head X)) (tail X)))
-            (list (head X))
-            (recurse (filter (>= (head X)) (tail X)))))
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{Sortp}
-(if (= (+1 0) (length X))
-    X
-    (append (recurse (tail X))
-            (list (length X))))
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}[]{TextSketch}
-Program ::= Term
-          | Program + Term
-Term    ::= String
-          | substr(Pos,Pos)
-Pos     ::= Number
-          | pos(Str,Str,Num)
-Num     ::= 0 | 1 | 2 | ... 
-          | -1 | -2 | ...
-Str     ::= Character 
-          | Character + String
-Character ::= a | b | c | ...
-\end{SaveVerbatim}
-
-\newcommand{\theSystem}{\textsc{ProgramSample}}
 \usepackage{arydshln}
@@
   }
-  \visible<5->{
-    \node at ([xshift=2cm,yshift=1cm]wiringDiagram.east) {\includegraphics[width = 5cm]{tree.jpg}};
-    }
+  %% \visible<5->{
+  %%   \node at ([xshift=2cm,yshift=1cm]wiringDiagram.east) {\includegraphics[width = 5cm]{tree.jpg}};
+  %%   }
   \end{tikzpicture}
@@
     \draw[->] (synthesizer.east) -- (program.west);
-  \end{tikzpicture}}
+  \end{tikzpicture}
+
+    \vspace{0.2cm}Concepts: $x^2$, etc\\Inference strategy: neurosymbolic search for programs}
   \renewcommand\code\texttt
     \only<3>{
-\Wider[4em]{  \begin{tikzpicture}
-    \node[draw,align=center](problem) at (0,0) {
-            \begin{tabular}{l}
-          \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
-          \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
-          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
-          \end{tabular}
-};
-    \node(synthesizer)[draw,align=center] at ([xshift=2.5cm]problem.east) {Learned \\program inducer};
+  \begin{tikzpicture}
+    \node(problem) at (0,0) {\includegraphics[width = 2cm]{figures/logoTasks/logo43.png}};
+    \node(synthesizer)[draw,align=center] at ([xshift=3cm]problem.east) {Learned \\program inducer};
     \draw[->] (problem.east) -- (synthesizer.west);
-    \node(program)[draw, align=center] at ([xshift=2.5cm]synthesizer.east) {program:\\\code{(filter ($\lambda$ (x)}\\\code{ (> x 2)))}};
+    \node(program)[draw, align=center] at ([xshift=3cm]synthesizer.east) {program:\\\code{(for (i 4)}\\ \code{ (circle i))}};
     \draw[->] (synthesizer.east) -- (program.west);
@@
 
-  \vspace{0.5cm}Concepts: \code{filter}, etc\\Inference strategy: neurosymbolic search for programs
-    }}
+  \vspace{0.2cm}Concepts: \code{circle}, etc\\Inference strategy: neurosymbolic search for programs
+    }
+%%     \only<4>{
+%% \Wider[4em]{  \begin{tikzpicture}
+%%     \node[draw,align=center](problem) at (0,0) {
+%%             \begin{tabular}{l}
+%%           \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+%%           \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
+%%           \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
+%%           \end{tabular}
+%% };
+%%     \node(synthesizer)[draw,align=center] at ([xshift=2.5cm]problem.east) {Learned \\program inducer};
+%%     \draw[->] (problem.east) -- (synthesizer.west);
+%%     \node(program)[draw, align=center] at ([xshift=2.5cm]synthesizer.east) {program:\\\code{(filter ($\lambda$ (x)}\\\code{ (> x 2)))}};
+%%     \draw[->] (synthesizer.east) -- (program.west);
+%%   \end{tikzpicture}
+
+%%   \vspace{0.5cm}Concepts: \code{filter}, etc\\Inference strategy: neurosymbolic search for programs
+%%     }}
 
@@
 
-\vspace{1cm}
+\vspace{0.5cm}
 
 \textbf{[Dechter et al., 2013]}  [Liang et al, 2010]; [Lake et al, 2015]
+
+\textbf{Dechter et al.}: Exploration-Compression. Inspiration for DreamCoder.
 
@@
 \end{tikzpicture}
-
-\pause
 
@@
 \end{frame}
-\begin{frame}{Learning the basics of programming}
-  \includegraphics[width = \textwidth]{figures/McCarthy.png}
-
-\centering  McCarthy 1959 Lisp  $\longrightarrow$ Modern functional programming
-
-  22 tasks. 64 CPUs. 93 hours.
-
-  \end{frame}
 
@@
 \end{frame}
+
 \begin{frame}{List functions \& Text editing: Learning curves on hold out tasks}
@@
 
-%% \begin{frame}{Symbolic regression from visual input}
-%% \centering\includegraphics[width = 5cm]{symbolicRegression.png}
-%% \end{frame}
+\begin{frame}{Learning the fundamentals of programming}
+  \includegraphics[width = \textwidth]{figures/McCarthy.png}
+
+\centering  McCarthy 1959 Lisp  $\longrightarrow$ Modern functional programming
+
+  22 tasks. 64 CPUs. 93 hours.
+
+  \end{frame}
+
+\begin{frame}{Symbolic regression from visual input}
+\centering\includegraphics[width = 5cm]{symbolicRegression.png}
+\end{frame}
 
@@
 
-\begin{frame}{Turtle graphics --- Learning curves}
-  \centering
-  \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
-  %% \begin{itemize}
-  %%   \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
-  %%   \item A line of length \texttt{n} followed with a right angle
-  %%   \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
-  %%   \item Unit line then teleport back to origin
-  %%   \item ...
-  %% \end{itemize}
-\end{frame}
+%% \begin{frame}{Turtle graphics --- Learning curves}
+%%   \centering
+%%   \includegraphics[width=9cm]{figures/dreams/montageTalkLC.eps}
+%%   %% \begin{itemize}
+%%   %%   \item $\frac\pi2$ and $\frac\pi4$ from $\pi$, $2$, $+$ and $/$
+%%   %%   \item A line of length \texttt{n} followed with a right angle
+%%   %%   \item Loops of length \texttt{n} that uses the number \texttt{n} inside.
+%%   %%   \item Unit line then teleport back to origin
+%%   %%   \item ...
+%%   %% \end{itemize}
+%% \end{frame}
 
@@
   \only<6>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_5.png}}
-  \only<7>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_13.png}}
+%  \only<7>{\centering\includegraphics[width = \textwidth]{figures/logo_primitives/logo_primitive_13.png}}
 
@@
    \item    Acquiring a domain-specific representation (DSL)
-     \item Learning  to write programs (recognition model)
+     \item Learning  to use that representation (recognition model)
    \end{itemize}

### 2018-10-16 nips_2018.tex
camera ready?

@@
 \author{
-  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay
+  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucas@lucasem.com}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
   \AND
-  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
+  Armando Solar-Lezama\\MIT\\\texttt{asolar@csail.mit.edu}\And Joshua B. Tenenbaum\\MIT\\\texttt{jbt@mit.edu}
   %% \And
@@
   \end{equation}
-
-\begin{wrapfigure}{R}{0.6\textwidth}
-  \begin{minipage}{0.6\textwidth}    
-    \begin{algorithm}[H]
-      \caption{The \system Algorithm}
-      \label{mainAlgorithm}
-      \begin{algorithmic}
-        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-        \FOR{$i=1$ {\bfseries to} $I$}
-        %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
-        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
-        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
-        %% \STATE Define $Q_x(z) \propto \begin{cases}
-        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-        %%   0&x\not \in \mathcal{F}_x
-        %% \end{cases}
-        \ENDFOR
-        \STATE \textbf{return} $\mathcal{D},\theta,q$
-      \end{algorithmic}
-    \end{algorithm}
-  \end{minipage}
-\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
-
 We induce a DSL by searching locally through the space of DSLs,
@@
 Grammars~\cite{cohn2010inducing}, and is closely related to the idea
-of antiunification~\cite{henderson2013cumulative}.
-
+of antiunification~\cite{henderson2013cumulative,hwang2011inducing}.
+Care must be taken
+to ensure that
+this `fragmenting' obeys variable scoping rules;
+Section 4 of the supplement gives an overview of Fragment Grammars
+and how we adapt them to the lexical scoping rules of $\lambda$-calculus.
 \begin{figure}[b!]\tabcolsep=2pt
@@
 \end{figure}
-
+\begin{wrapfigure}{R}{0.6\textwidth}\vspace{-1cm}
+  \begin{minipage}{0.6\textwidth}    
+    \begin{algorithm}[H]
+      \caption{The \system Algorithm}
+      \label{mainAlgorithm}
+      \begin{algorithmic}
+        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
+        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
+        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
+        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
+        \FOR{$i=1$ {\bfseries to} $I$}
+        %     \FOR{$x:\tau\in X$}
+        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
+        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
+        \footnotesize{\hspace{0.0cm}(\textbf{Explore})}
+        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
+        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
+        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
+        \footnotesize{\hspace{0.0cm}\textbf{(Explore)}}
+
+        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
+        %% \STATE Define $Q_x(z) \propto \begin{cases}
+        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
+        %%   0&x\not \in \mathcal{F}_x
+        %% \end{cases}
+        \ENDFOR
+        \STATE \textbf{return} $\mathcal{D},\theta,q$
+      \end{algorithmic}
+    \end{algorithm}
+  \end{minipage}\vspace{-1cm}
+\end{wrapfigure}
 To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
@@
 
+\pagebreak
 \section{Programs that manipulate sequences}\label{sequences}
@@
 \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
+
+\vspace{0.5cm}
 
@@
   \midrule
-  \% solved&\textbf{79\%}\footnotemark & 
-  94\% &71\%&35\%&62\%&37\%\\
-  Solve time&  88s&6s&11s&35s&44s&20s\\
+  \% solved&\textbf{94\%} & 
+  79\% &71\%&35\%&62\%&37\%\\
+  Solve time&  88s&39s&11s&35s&44s&20s\\
   \midrule
@@
 \begin{figure}\centering
-  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps}
+  \includegraphics[width = 4.35cm]{figures/listLearningCurve.eps}
   \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}
@@
  and (2) we show how to combine these techniques with bottom-up neural recognition models.
- Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
+ Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, Bayesian program merging~\cite{hwang2011inducing}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
 

### 2018-10-18 nips_2018.tex
fixed color in learning curve figure
included some more discussion

@@
 \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.
-&
+
 %For each domain,
-%% We are interested both in how many tasks the
-%% agent can solve and how quickly it can find those solutions.
-%% Tbl.~\ref{baselineComparisons}
-%% compares our model against these alternatives.
-%% We consistently
-%% improve on the baselines,
-%% and find that lesioning the recognition model
-%% and lesioning it also slows down the convergence of the algorithm,
-%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
-%% This supports a view of the recognition model as a way of amortizing the cost of search.
+We are interested both in how many tasks the
+agent can solve and how quickly it can find those solutions.
+Tbl.~\ref{baselineComparisons}
+compares our model against these alternatives.
+We consistently
+improve on the baselines,
+and also find that lesioning the recognition model
+impairs the convergence of the algorithm,
+causing it to hit a `plateau' after which it stops solving new tasks after an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
+This `plateau' supports a view of the recognition model as a way of amortizing the cost of search.
 
 \begin{figure}\centering
-  \includegraphics[width = 4.35cm]{figures/listLearningCurve.eps}
+  \includegraphics[width = 4.35cm]{figures/listLearningCurve_color.eps}
   \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}

### 2018-10-19 nips_2018.tex
tweaks to spacing

@@
 \author{
-  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucas@lucasem.com}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
+  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucasem@mit.edu}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
   \AND
@@
 Our algorithm takes inspiration from several ways that skilled human
-programmers have learned to code: Skilled coders build libraries of
+programmers have learned to code: skilled coders build libraries of
 reusable subroutines that are shared across related programming tasks,
@@
 How should we get the data to train $q$?
-This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
+This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
 One approach is to sample programs from the DSL,
@@
 Our data set is interesting in three major ways: many of the tasks require
-complex solutions; the tasks were not generated from some latent DSL, and
+complex solutions; the tasks were not generated from some latent DSL; and
 the agent must learn to solve these complicated problems from only 236
@@
 \begin{wraptable}{r}{0.55\textwidth}
-\vspace{-5pt}
+\vspace{8pt}
 %% This radically shrinks the table
@@
   \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
-  Solve time& 29s&65s&38s&80s&--&235s\\
+  Solve time& 29s&49s % FIXME: 65 in the table but 49 on the graph? 65s
+  &38s&80s&--&235s\\
   \midrule
@@
   %% proxy for \# of continuous parameters.
-}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
+}\vspace{-0cm}\label{baselineComparisons} \end{wraptable}
 Nevertheless, much prior work can be modeled within our setup. 
@@
 impairs the convergence of the algorithm,
-causing it to hit a `plateau' after which it stops solving new tasks after an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
-This `plateau' supports a view of the recognition model as a way of amortizing the cost of search.
+causing it to hit a lower `plateau' after which it stops solving new tasks, following an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
+This lowered `plateau' supports a view of the recognition model as a way of amortizing the cost of search.
 

### 2018-10-22 nips_2018.tex
changed my email address to MIT, some tiny edits

@@
 \begin{wraptable}{r}{0.55\textwidth}
-\vspace{8pt}
+\vspace{-5pt}
 %% This radically shrinks the table
@@
   \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
-  Solve time& 29s&49s % FIXME: 65 in the table but 49 on the graph? 65s
-  &38s&80s&--&235s\\
+  Solve time& 29s&65s&38s&80s&--&235s\\
   \midrule
@@
   %% proxy for \# of continuous parameters.
-}\vspace{-0cm}\label{baselineComparisons} \end{wraptable}
+}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
 Nevertheless, much prior work can be modeled within our setup. 
@@
 impairs the convergence of the algorithm,
-causing it to hit a lower `plateau' after which it stops solving new tasks, following an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
-This lowered `plateau' supports a view of the recognition model as a way of amortizing the cost of search.
+causing it to hit a `plateau' after which it stops solving new tasks after an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
+This `plateau' supports a view of the recognition model as a way of amortizing the cost of search.
 

### 2018-10-26 nips_2018.tex
more writing

--- 2018-10-19 nips_2018.tex
@@
 
-\title{Library Learning in Neurally-Guided Bayesian Program Learning}
+\title{Library Learning in Neurally-Guided Bayesian Program Induction}
 
@@
         &
-        \hspace{-0.3cm}\begin{tabular}{c}
+        \hspace{0.1cm}\begin{tabular}{c}
           \code{[7\, 3]}$\to $\code{False}                              \\
@@
 write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
+``compiled'' inference~\cite{le2016inference,stuhlmuller2013learning}.%%  (CITE Goodman and Stuhlmuller,
 %% and others too on amortized inference.)
@@
 %% \end{itemize}
-\noindent\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
+%% \noindent\\\textbf{Contributions.} 
+(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
 (2) We give an algorithm for learning DSLs,
@@
 Why enumerate, when the program synthesis community has invented many
-sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
+sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta,polikarpova2016program}.
 We have two reasons:
@@
 \end{figure}
-\begin{wrapfigure}{R}{0.6\textwidth}\vspace{-1cm}
-  \begin{minipage}{0.6\textwidth}    
+\begin{figure}%{R}{0.6\textwidth}\vspace{-1cm}
+\centering  \begin{minipage}{0.9\textwidth}    
     \begin{algorithm}[H]
@@
         %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
+        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
         \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.0cm}(\textbf{Explore})}
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
+        \footnotesize{\hfill(\textbf{Explore})}
+        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\footnotesize{\textbf{Compile}})
+        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
         \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.0cm}\textbf{(Explore)}}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
+        \footnotesize{\hfill\textbf{(Explore)}}
+
+        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\footnotesize{\hfill\textbf{(Compress)}}
         %% \STATE Define $Q_x(z) \propto \begin{cases}
@@
     \end{algorithm}
-  \end{minipage}\vspace{-1cm}
-\end{wrapfigure}
+  \end{minipage}
+\end{figure}
 To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
@@
 recognition model training, and DSL induction.
+For added robustness,
+we interleave an extra program search step (Exploration) before training the recognition model,
+and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Exploration step.
 
@@
 assumptions about the space of programs and the statement of tasks.
+Nevertheless, much prior work can be modeled within our setup. 
+We compare with the following ablations (Tbl~\ref{baselineComparisons};
+Fig~\ref{learningCurves}):
+\\\noindent \textbf{No NN:} lesions the recognition model.
+\\\noindent \textbf{NPS}, which does not learn the DSL,
+instead learning the recognition model
+from samples drawn from the fixed DSL.
+We call this NPS (Neural Program Synthesis)
+because this is closest to how
+RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
+\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
+only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
+This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
+\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
+but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.
+
 \begin{wraptable}{r}{0.55\textwidth}
-\vspace{8pt}
+\vspace{-10pt}
 %% This radically shrinks the table
@@
   %% proxy for \# of continuous parameters.
-}\vspace{-0cm}\label{baselineComparisons} \end{wraptable}
-Nevertheless, much prior work can be modeled within our setup. 
-We compare with the following ablations (Tbl~\ref{baselineComparisons};
-Fig~\ref{learningCurves}):
-\\\noindent \textbf{No NN:} lesions the recognition model.
-\\\noindent \textbf{NPS}, which does not learn the DSL,
-instead learning the recognition model
-from samples drawn from the fixed DSL.
-We call this NPS (Neural Program Synthesis)
-because this is closest to how
-RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
-This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
-\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.
-
+}\vspace{-1cm}\label{baselineComparisons} \end{wraptable}
 %For each domain,
@@
  and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.
+
+ \section{Limitations of EC$^2$}
+
+ We have presented an algorithm that, starting from relatively little,
+ learned to solve nontrivial program synthesis problems in three
+ different domains.  However, our system was also given (and also,
+ critically \emph{needs}) a corpus of training tasks to solve for each
+ domain.  Is constructing (or curating) corpra of tasks any easier or
+ better than hand-engineering DSLs?  In the immediate future, we
+ expect hand-engineering of DSLs to largely continue, especially in
+ domains where humans have strong intuitions about the underlying
+ system of domain specific concepts, like text editing. However, if
+ program induction is to become a standard part of the AI toolkit,
+ then, in the long-term, we need to build agents that autonomously
+ acquire the domain specific knowledge needed to navigate a new
+ domain.  So, through the lens of program synthesis, EC$^2$ carries
+ the limitation that it requires a high-quality corpus of training
+ tasks; but, for the program-induction approach to AI, this limitation
+ is natural.
+
+ The particular implementation of EC$^2$ developed here
+ inherits a key limitation from a specific arbitrary engineering decision:
+ our enumerative program search strategy cannot possibly
+ scale to large bodies of code, even with a good recognition model --- but one could in principle replace
+ our enumerative approach with stochastic search~\cite{schkufza2013stochastic,DBLP:books/daglib/0070933},
+ constraint solving~\cite{solar2008program}, type-directed synthesis~\cite{polikarpova2016program},
+ or other techniques from the program synthesis community.
 
@@
 
-\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work.
-Funding from NSF GRFP and the MUSE program (Darpa grant FA8750-14-2-0242).
+\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work, and for funding from the NSF GRFP, the MUSE program (Darpa grant FA8750-14-2-0242),
+and an AWS ML Research Award.
 

### 2018-10-26 supplement.tex
more writing

@@
 
-\title{Supplement to: Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning}
+\title{Supplement to: Library Learning in Neurally-Guided Bayesian Program Induction}
 
@@
 \author{
-  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay
+  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucasem@mit.edu}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
   \AND
-  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
+  Armando Solar-Lezama\\MIT\\\texttt{asolar@csail.mit.edu}\And Joshua B. Tenenbaum\\MIT\\\texttt{jbt@mit.edu}
   %% \And

### 2018-10-26 nips_2018.tex
tweaks to discussion

@@
  better than hand-engineering DSLs?  In the immediate future, we
- expect hand-engineering of DSLs to largely continue, especially in
- domains where humans have strong intuitions about the underlying
+ expect some degree of hand-engineering of DSLs to continue, especially
+ in domains where humans have strong intuitions about the underlying
  system of domain specific concepts, like text editing. However, if

### 2018-10-26 nips_2018.tex
more related work

@@
  Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, Bayesian program merging~\cite{hwang2011inducing}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
+ Concurrent with this work, closely related ideas were applied to mining `code idioms' from imperative programs~\cite{8355713}
+ and for synthesizing functional programs from natural language~\cite{Alex}.
 
@@
  We have presented an algorithm that, starting from relatively little,
- learned to solve nontrivial program synthesis problems in three
+ learned to solve program synthesis problems in three
  different domains.  However, our system was also given (and also,

### 2018-10-26 nips_2018.tex
almost, ready

@@
 write programs in the domain; it modifies the structure of the DSL by
-discovering regularities across programs found during exploration, compressing
+discovering regularities across programs found during the Explore step, compressing
 them to distill out common code fragments across successful programs.
@@
 So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
-\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+
+\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
 is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
@@
 \system does approximate MAP inference  by maximizing this lower bound on the joint probability,
-alternating maximization w.r.t.\ the frontiers (Explore) and the DSL (Compression):
+alternating maximization w.r.t.\ the frontiers (Explore) and the DSL (Compress):
 \\\noindent \textbf{Explore: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
@@
 %% where  is large.
-\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
+\\\noindent \textbf{Compress: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
@@
 of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
-during the compilation phase: $q$ is trained to approximate the
+during the Compile step: $q$ is trained to approximate the
 posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D},\theta]\propto\probability[x|p]\probability[p|\mathcal{D},\theta}]$,
@@
 
-\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
+\textbf{Compile: learning to tractably maxing $\lowerBound$ w.r.t. the
   frontiers.}  Here we train %% a neural network, $q$, to predict a
@@
 $\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
-in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
+in the frontiers will most increase $\lowerBound$.
+We train $q$ both on programs found during the Explore step and
+on samples from the current DSL.
+
+\begin{wrapfigure}{r}{0.6\textwidth}\vspace{-0.55cm}\centering
+  \begin{tikzpicture}[scale=0.9]
+    {\footnotesize
+    \begin{scope}[shift = {(1,-1)}]
+    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
+    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
+
+    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compress)}} (DSL.30);
+    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Compile)}} (recognitionModel.150);
+    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Compile)}} (recognitionModel.west);
+
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Explore)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Explore)}} (synthesis.east);
+  \end{scope}}
+    \end{tikzpicture}
+  \caption{How these steps bootstrap each other.}  \label{feeding}\vspace{-1cm}
+\end{wrapfigure}%
+Crucially, each of these three steps bootstraps the others (Fig.~\ref{feeding}): improving either the DSL or the recognition model
+makes search easier, so we find more programs solving tasks;
+both improving the DSL and solving more tasks expands the training data for the recognition model;
+and finding more programs that solve tasks
+gives more data from which to learn a DSL.
+
+%% With $q$ in hand we can find programs for
 %% the frontier $\mathcal{F}_x$ by searching for programs that maximize
@@
 %% bottom-up \emph{recognition model}
-
-\subsection{Exploration: Searching for Programs}\label{explorationSection}
+\vspace{0.5cm}Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+recognition model training, and DSL induction.
+For added robustness,
+we interleave an extra program search step (Explore) before training the recognition model,
+and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Explore step.
+
+\begin{figure}[h]
+\centering  \begin{minipage}{0.9\textwidth}    
+    \begin{algorithm}[H]
+      \caption{The \system Algorithm}
+      \label{mainAlgorithm}
+      \begin{algorithmic}
+        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
+        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
+        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
+        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
+        \FOR{$i=1$ {\bfseries to} $I$}
+        %     \FOR{$x:\tau\in X$}
+        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
+        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
+        \hfill(\textbf{Explore})
+        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\textbf{Compile})
+        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
+        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
+        \hfill\textbf{(Explore)}
+
+        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\hfill\textbf{(Compress)}
+        %% \STATE Define $Q_x(z) \propto \begin{cases}
+        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
+        %%   0&x\not \in \mathcal{F}_x
+        %% \end{cases}
+        \ENDFOR
+        \STATE \textbf{return} $\mathcal{D},\theta,q$
+      \end{algorithmic}
+    \end{algorithm}
+  \end{minipage}
+\end{figure}
+
+\subsection{Explore: Searching for Programs}\label{explorationSection}
 
@@
 
-\subsection{Compilation: Learning a Neural Recognition Model}\label{recognitionSection}
+\subsection{Compile: Learning a Neural Recognition Model}\label{recognitionSection}
 
@@
 and then train $q$ to predict the program from the input/outputs.
-This is like how a Helmholtz machine
-trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
-The advantage of ``Helmholtz machine'' training is that
+This is like how the wake-sleep algorithm for the Helmholtz machine
+trains its recognition model during its sleep phase~\cite{dayan1995helmholtz}.
+The advantage of training on samples, or ``fantasies,'' is that
 we can draw unlimited samples from the DSL,
@@
 Another approach is
- self-supervised learning,
-training $q$ on the (program, task)
-pairs discovered by search.
-The advantage of self-supervised learning
+to train $q$ on the (program, task)
+pairs discovered by the Explore step.
+The advantage here
 is that the training data is much higher quality,
-because we are training on the actual tasks.
+because we are training on real tasks.
 Due to these complementary advantages,
@@
  where the expectation is taken over tasks.
-Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
- The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
+Taking this expectation over the empirical distribution of tasks trains $q$ on the real data; taking it over samples from the generative model trains $q$ on ``fantasies.''
+ The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the fantasy ($\mathcal{L}_{\text{f}}$) and real-data ($\mathcal{L}_{\text{r}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{r} + \mathcal{L}_\text{f}$:
 \begin{align}\nonumber
-\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
-\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+\mathcal{L}_{\text{f}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
+\mathcal{L}_{\text{r}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
   \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
@@
 
-\subsection{Compression: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
+\subsection{Compress: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
 
@@
 and how we adapt them to the lexical scoping rules of $\lambda$-calculus.
-\begin{figure}[b!]\tabcolsep=2pt
+\begin{figure}[h!]\tabcolsep=2pt
 \centering  \begin{minipage}[c]{0.2\textwidth}
@@
 \end{figure}
-\begin{figure}%{R}{0.6\textwidth}\vspace{-1cm}
-\centering  \begin{minipage}{0.9\textwidth}    
-    \begin{algorithm}[H]
-      \caption{The \system Algorithm}
-      \label{mainAlgorithm}
-      \begin{algorithmic}
-        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-        \FOR{$i=1$ {\bfseries to} $I$}
-        %     \FOR{$x:\tau\in X$}
-        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
-        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hfill(\textbf{Explore})}
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\footnotesize{\textbf{Compile}})
-        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
-        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hfill\textbf{(Explore)}}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\footnotesize{\hfill\textbf{(Compress)}}
-        %% \STATE Define $Q_x(z) \propto \begin{cases}
-        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-        %%   0&x\not \in \mathcal{F}_x
-        %% \end{cases}
-        \ENDFOR
-        \STATE \textbf{return} $\mathcal{D},\theta,q$
-      \end{algorithmic}
-    \end{algorithm}
-  \end{minipage}
-\end{figure}
 To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
@@
 %% In the Supplement we derive a tractable MAP estimator for $\theta$.
-Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
-For added robustness,
-we interleave an extra program search step (Exploration) before training the recognition model,
-and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Exploration step.
-
-\begin{comment}
-\subsection{Implementing \system}
-
-Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
-We note the following implementation details:
-(1) We perform an search cycle before each compression and compilation  cycle.
-(2) On the first iteration, we do \emph{not} train the recognition model on samples from
-the generative model because a generative model has not yet been learned --
-we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
-%% (3) Because the frontiers can grow very large,
-%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
-%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
-(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
-by only summing over the top $K$
-programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
-(4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.
-\end{comment}
 
@@
 
-\pagebreak
+%\pagebreak
 \section{Programs that manipulate sequences}\label{sequences}
@@
 but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.
+\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first Explore step.
 
@@
 
- \section{Limitations of EC$^2$}
-
- We have presented an algorithm that, starting from relatively little,
- learned to solve program synthesis problems in three
- different domains.  However, our system was also given (and also,
- critically \emph{needs}) a corpus of training tasks to solve for each
- domain.  Is constructing (or curating) corpra of tasks any easier or
- better than hand-engineering DSLs?  In the immediate future, we
- expect some degree of hand-engineering of DSLs to continue, especially
- in domains where humans have strong intuitions about the underlying
- system of domain specific concepts, like text editing. However, if
- program induction is to become a standard part of the AI toolkit,
- then, in the long-term, we need to build agents that autonomously
- acquire the domain specific knowledge needed to navigate a new
- domain.  So, through the lens of program synthesis, EC$^2$ carries
- the limitation that it requires a high-quality corpus of training
- tasks; but, for the program-induction approach to AI, this limitation
- is natural.
-
- The particular implementation of EC$^2$ developed here
- inherits a key limitation from a specific arbitrary engineering decision:
- our enumerative program search strategy cannot possibly
- scale to large bodies of code, even with a good recognition model --- but one could in principle replace
- our enumerative approach with stochastic search~\cite{schkufza2013stochastic,DBLP:books/daglib/0070933},
- constraint solving~\cite{solar2008program}, type-directed synthesis~\cite{polikarpova2016program},
- or other techniques from the program synthesis community.
+ %% \section{Limitations of EC$^2$}
+
+ %% We have presented an algorithm that, starting from relatively little,
+ %% learned to solve program synthesis problems in three
+ %% different domains.  
+ %% The particular implementation of EC$^2$ developed here
+ %% inherits a key limitation from a specific arbitrary engineering decision:
+ %% our enumerative program search strategy cannot possibly
+ %% scale to large bodies of code, even with a good recognition model --- but one could in principle replace
+ %% our enumerative approach with stochastic search~\cite{schkufza2013stochastic,DBLP:books/daglib/0070933},
+ %% constraint solving~\cite{solar2008program}, type-directed synthesis~\cite{polikarpova2016program},
+ %% or other techniques from the program synthesis community.
 
@@
 
-%\parbox{0.58\textwidth}{
-
-%\paragraph{Outlook.}
 We contribute an algorithm, \systemEnding, that learns to program by
@@
 symbolic representations and bottom-up neural networks --- both of them learned
---- helps make program induction systems more generally useful for AI. Many
+--- helps make program induction systems more generally useful for AI.
+
+A feature of our system
+is that it learns from (and also, critically \emph{needs})
+a corpus of training tasks.  Is constructing (or curating) corpra of tasks any easier or
+ better than hand-engineering DSLs?  In the immediate future, we
+ expect some degree of hand-engineering of DSLs to continue, especially
+ in domains where humans have strong intuitions about the underlying
+ system of domain-specific concepts, like text editing. However, if
+ program induction is to become a standard part of the AI toolkit,
+ then, in the long-term, we need to build agents that autonomously
+ acquire the  knowledge needed to navigate a new
+ domain.  So, through the lens of program synthesis, EC$^2$ carries
+ the restriction that it requires a high-quality corpus of training
+ tasks; but, for the program-induction approach to AI, this restriction
+ is a feature, not a bug.
+
+Many
 directions remain open.
-%\paragraph{Future.}
 Two immediate goals are to integrate more sophisticated neural recognition
@@
 synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the generic methods used here.
-%% We are in the process of applying our algorithm to generative programs and
-%% prototyped this with a turtle-like domain: Figure~\ref{geomCompiled} gives some
-%% preliminary results for turtle graphics  --- see Section 1 of the supplementary material for more
-%% details.
+domains over the generic methods used here:
+while our focus in this work was
+learning to quickly write small programs,
+we believe more sophisticated neural models,
+coupled with more powerful program search algorithms,
+could extend our approach to
+synthesize larger bodies of code.
 Another direction is to explore DSL meta-learning: can we find a
@@
 new domains, including the three domains considered,  but also many others?
-%% }\hfill
-%% \begin{minipage}{0.4\textwidth}
-%%   \centering
-%%   \includegraphics[width=\textwidth]{figures/geomCompiled.eps} 
-%%   \captionof{figure}{
-%%   Top left: Example training tasks. Top right: samples from the learned DSL.
-%%  Bottom: \% holdout testing tasks solved (middle), on sides are the averaged samples from the DSL
-%%   before any training (left) and after last iteration (right).}
-%% \label{geomCompiled}
-%% \end{minipage}
 

### 2018-10-26 supplement.tex
almost, ready

@@
 \maketitle
+
+\section{An Illustration of the three steps of our algorithm}
+Below we diagram the steps employed by our algorithm. At each step of the algorithm,
+we have shaded the observed variables in gray and left the unobserved variables white.
+Black lines correspond to a connection from the top-down
+generative model, while red lines correspond to connections from the bottom-up recognition model.
+\begin{figure}[h]
+\centering
+%  \includegraphics{figures/iterations.eps}
+\begin{tikzpicture}
+  \begin{scope}[shift={(2.5,0)}]
+    \node[obs] at (3,3) (dx){$\mathcal{D}$};
+    \node[latent] at (3.5,1.75) (zp){$p$};
+    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
+    \node[obs] at (3.5,0.7) (xp) {$x$};
+    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Explore: Infer $p$};
+    \draw [->,red] (xp.east) to[out = 0,in = 0] node(nn){} (tx.east);
+    \draw [->,red] (tx) -- (zp);
+    \draw [->] (dx) -- (zp);
+    \node at (nn) {
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
+  \end{scope}
+  \begin{scope}[shift={(0.7,-4.2)}]
+    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+    \draw [->] (p2.south) -- (x2.north);
+    \draw [->] (dt.south) -- (p2.north);
+    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+    \draw [->] (p1.south) -- (x1.north);
+    \draw [->] (dt.south) -- (p1.north);
+    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+    \draw [->] (p3.south) -- (x3.north);
+    \draw [->] (dt.south) -- (p3.north);
+    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+    \draw [->] (zp.south) -- (xp.north);
+    \plate {}{(zp)(xp)}{$x\in X$};
+    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
+    rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
+    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Compile: Train $q$\\training data: cyan $(x,p)$};
+  \end{scope}
+
+  \begin{scope}[shift={(7.7,-4)}]
+    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+    \node[latent] at (1.5,3) (t){$\theta$};
+    \node[obs] at (1,1.75) (z){$p$};
+%     \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+    \node[obs] at (1,0.5) (x) {$x$};
+    \edge {z}{x};
+    \edge {d,t}{z};
+    \plate {}{(z)(x)}{$x\in X$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Compress: Induce $(\mathcal{D},\theta)$};
+  \end{scope}
+\end{tikzpicture}
+\end{figure}
 

### 2018-10-31 archive.tex
got rid of old graphics content; deanonymized
possibly final talk slides
version space archive progress

@@
  % For algorithms
-\usepackage{algorithm}
-\usepackage{algorithmic}
+%\usepackage{algorithm}
+%\usepackage{algorithmic}
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
 \usepackage{tikz}
@@
 \newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}
+\newcommand{\shift}[1]{\uparrow^{#1}}
+\newcommand{\substitute}[2]{[\$ #1 \mapsto #2]}
+\newcommand{\reduce}{\longrightarrow}
+\newcommand{\manyReduce}{\longrightarrow^*}
+
+\newtheorem{definition}{Definition}
 
@@
 \icmlauthor{Lucas Morales}{MIT}
-\icmlauthor{Matthias}{MIT}
+\icmlauthor{Mathias}{MIT}
 \icmlauthor{Maxwell Nye}{MIT}
@@
 that allow an agent to easily express solutions to the tasks at hand.
-%In the \system algorithm we infer the DSL from a collection of frontiers.
-Intuitively, we want the algorithm to
-look at  the frontiers and
-generalize beyond them, 
-both so the DSL can better express the current solutions,
-and  also so that the DSL might expose new abstractions
-which will later be used to
-discover more programs.
+Intuitively, we want the algorithm to look at the programs found
+during waking and generalize beyond them, both so the DSL can better
+express the current solutions, and also so that the DSL might expose
+new abstractions which will later be used to discover more programs.
+
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
@@
 \end{align}
-We induce a DSL by searching locally through the space of DSLs,
-proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
-The search moves work by introducing new
-$\lambda$-expressions into the DSL.
-We propose these new expressions by extracting fragments of
-programs already in the frontiers (Tbl.~\ref{fragmentExample}).
-An important point here is that we are \emph{not} simply adding
-subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
+At a high level, our approach is to search locally through the space
+of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
+fails to increase.  These small changes consist of introducing new
+candidate $\lambda$-expressions into the DSL.
+
+However, there is a snag with this simple approach:
+whenever we add a new expression $e$ to the DSL,
+the programs found during waking ($\left\{\mathcal{F}_x \right\}$)
+are not written in terms of $e$.
+Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
+after having found the programs
+\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
+As human programmers,
+we can look at these pieces of code and recognize that,
+if we define a new procedure called \code{double},
+defined as \code{($\lambda$ (x) (+ x x))},
+then we can rewrite the original programs as
+\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
+This process is a kind of refactoring
+where a new subroutine is defined (\code{double})
+and the old programs rewritten
+in terms of the new subroutine.
+Figure~\ref{mapFactor}
+diagrams this refactoring
+process for a more complicated
+setting,
+where the agent must rediscover the higher-order function \code{map}
+starting from the basics of Lisp and the Y-combinator.
+
+Formally, we want a DSL
+which puts high probability on
+any refactoring of the programs found during waking, maximizing:
+\begin{align}
+\nonumber      \log \probability[\mathcal{D}]& + 
+\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
+\label{factorObjective}
+\end{align}
+where $p'\manyReduce p$
+is the standard notation for ``expression $p'$ evaluates to $p$ by the standard rules of $\lambda$-calculus''~\cite{pierce}.
+
+Equation~\ref{factorObjective} captures the idea that we want
+to add new components to the DSL while jointly refactoring our old
+programs in terms of these new components.  But
+this joint inference problem is intractable, because there are
+infinitely many ways of refactoring a program.
+We combine two ideas to make this refactoring process tractable:
+
+\textbf{Idea 1:} Limit the degree to which
+a piece of code can be refactored.
+Instead of considering every refactoring,
+bound the number of $\lambda$-calculus evaluation steps
+separating a refactoring from its original program.
+Formally,
+we define the set of $n$-step refactorings as:
+\begin{equation}
+  R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
+\end{equation}
+where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
+in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
+For example,
+\begin{align*}
+  \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
+  \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
+  \code{(lambda (y) y)}&
+\end{align*}
+Returning to Equation~\ref{factorObjective},
+this approximation gives the following objective:
+\begin{align}
+\nonumber      \log \probability[\mathcal{D}]& + 
+\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\+&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
+\label{limitedObjective}
+\end{align}
+In practice, setting the number of refactoring steps $n$ to 3 suffices
+to give a competent DSL learning algorithm.  Although the number of
+refactorings is now finite, it is still prohibitively large, and in
+fact grows exponentially quickly both as a function of $n$ and as a
+function of the size of the program being refactored.  Next, we show
+how to tame this exponential explosion.
+
+\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
+we developed a technique for compactly representing and efficiently
+calculating the set of all refactorings.  The key idea is to represent
+the set of refactorings using a \emph{version
+  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
+A version space is a tree-shaped data structure
+that compactly represents a large set of programs.
+We first define version spaces over $\lambda$-calculus expressions,
+and then give a dynamic program for constructing a version space
+that represents the set of possible refactorings (i.e., $R_n(p)$).
+
+A version space is either:
+\begin{itemize}
+\item A deBuijn index: written $\$i$, where $i$ is a natural number
+\item An abstraction: written $\lambda v$, where $v$ is a version space
+\item A  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
+\item A  union: $\uplus V$, where $V$ is a set of version spaces
+\item The empty set, $\varnothing$
+\item The set of all $\lambda$-calculus expressions, $\Lambda$
+\end{itemize}
+The purpose of a version space to compactly represent a set of programs.
+We refer to this set as the \textbf{extension} of the version space:
+\begin{definition}
+  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
+  and is defined recursively as:
+  \begin{align*}
+    \denotation{\$i}& = \left\{\$i \right\}\\
+    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}\\
+    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
+    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}\\
+    \denotation{\varnothing}& = \varnothing\\
+    \denotation{\Lambda}& = \Lambda
+    \end{align*}
+\end{definition}
+Version spaces also support efficient membership checking, which we
+write as $e\in v$, and which is equivalent to $e\in \denotation{v}$.
+Important for our purposes, it is also efficient to refactor the
+members of a version space's extension in terms of a new DSL.
+We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
+\begin{align*}
+  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
+    e\text{, if $e\in \mathcal{D}$ and $e\in v$}\\
+    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
+  \end{cases}\\
+  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\
+  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\
+  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\
+  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\ell(e|\mathcal{D})
+  \end{align*}
+where $\ell(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
+is the size of the syntax tree of $e$,
+when members of $\mathcal{D}$ are counted as having size 1.
+Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\ell(p|\mathcal{D})$.
+
+In Appendix~\ref{appendixVersion}, we describe a dynamic program for
+efficiently constructing a version space containing every $n$-step
+refactoring.
+This dynamic program, which we call $I\beta^n(p)$,
+satisfies $\denotation{I\beta^n(p)} =  R_n(p)$.
+In other words,
+this dynamic program
+builds a data structure
+that represents the entire set of
+refactorings -- but without having to explicitly enumerate all of the refactorings.
+
+With this machinery in hand, we now have all the pieces needed to
+learn a DSL (Algorithm~\ref{grammarInductionAlgorithm})
+
+    \begin{algorithm*}[H]
+    %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
+       \caption{DSL Induction Algorithm}
+       \label{grammarInductionAlgorithm}
+       \begin{algorithmic}[1]
+         \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
+%         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
+         \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+         \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
+         \WHILE {true}         
+         \STATE $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
+         \STATE Define $\text{refactor}(p|\mathcal{D}') = \argmin_{e\in \denotation{v_p}} \ell (e|\mathcal{D}')$ \Comment{Rewrites $p$ assuming $\mathcal{D}'$ is the DSL}
+         \STATE Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
+         \STATE Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
+         \STATE Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
+         \STATE components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
+         \STATE proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
+         \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
+         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
+         \STATE $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
+         \STATE $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
+         \ENDWHILE
+       \end{algorithmic}
+    \end{algorithm}
+
+\begin{figure*}
+  \centering\begin{tikzpicture}[every node/.style={inner sep=1,outer sep=0}]
+    \node(p1)[draw] at (-1,0) {
+      \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (+ (car l) (car l))}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+    };
+    \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-3cm]p1.south) {
+      \begin{tabular}{l}
+  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
+\texttt{ ($\lambda$ (z) (+ z z)))}
+      \end{tabular}
+    };
+
+        \node(p2)[draw] at ([xshift=4.5cm]p1.east) {
+    \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (- (car l) 1)}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+
+    };
+    \node(r2)[draw] at ([yshift=-3cm]p2.south) {
+\begin{tabular}{l}
+  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
+\texttt{ ($\lambda$ (z) (- z 1)))}
+      \end{tabular}
+
+    };
+
+    \draw [->] (p1.south)  --(r1.north) node[fill=white,midway] {refactor};
+    \draw [->] (p2.south)  --(r2.north) node[fill=white,midway] {refactor};
+
+    \node[draw](m) at (3,-7) {
+\begin{tabular}{l}
+          \fbox{\textsc{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
+  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}\\
+  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z z)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
+  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z 1)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
+      \end{tabular}      
+    };
+    \draw [->](r1.south)--(m.north);
+    \draw [->](r2.south)--(m.north);
+    \node[fill=white] at ([yshift=0.5cm]m.north) {\textbf{Compress (MDL/Bayes objective)}};
+    %% \visible<2-3>{
+    %%   \node[fill=orange!50!white,align=center] at (r1) {\Huge $10^{14}$\\\Huge  refactorings };
+    %%   \node[fill=orange!50!white,align=center] at (r2) {\Huge $10^{13}$\\\Huge  refactorings };
+    %% }
+    %% \visible<3>{
+    %%   \node[fill=orange!50!white,align=center] at ([yshift=7cm]m.north) {\Huge version space: set of programs\\
+    %%   \huge Lau 2003; Gulwani 2012};
+    %%   }
+    %% \visible<4>{
+    %%   \node[fill=orange!50!white,align=center] at (r1) {\Huge $\leq 10^6$\\\Huge version spaces };
+    %%   \node[fill=orange!50!white,align=center] at (r2) {\Huge $\leq 10^6$\\\Huge version spaces };
+    %% }
+
+       \node(t1)[draw] at ([yshift=1.5cm]p1.north) {\begin{tabular}{ll}
+           \textbf{Task}:&\texttt{(1 2 3)$\to$(2 4 6)}\\
+           &\texttt{(4 3 4)$\to$(8 6 8)}
+       \end{tabular}};
+       \draw [->] (t1.south)  --(p1.north) node[fill=white,midway] {program search};
+       \node(t2)[draw] at ([yshift=1.5cm]p2.north) {\begin{tabular}{ll}
+           \textbf{Task}:&\texttt{(1 2 3)$\to$(0 1 2)}\\
+           &\texttt{(4 3 4)$\to$(3 2 3)}
+       \end{tabular}};
+       \draw [->] (t2.south)  --(p2.north) node[fill=white,midway] {program search};
+
+  \end{tikzpicture}
+  \caption{DSL learning as code refactoring.
+For each task we discover programs during waking, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}).
+Common subprograms are incorporated into the DSL when they increase a Bayesian objective.
+Intuitively, these new DSL components best compress the programs found during waking.}\label{mapFactor}
+  \end{figure*}
+
+\begin{comment}
+
+We propose these new expressions
+by extracting fragments of programs already in the frontiers
+(Tbl.~\ref{fragmentExample}).  An important point here is that we are
+\emph{not} simply adding subexpressions of programs to $\mathcal{D}$,
+as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}
+and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
 extracting fragments that unify with programs in the frontiers.  This
@@
 
+  \end{comment}
 \begin{comment}
@@
 
-\begin{figure}  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{cl}\toprule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
+\begin{figure*}  \newcommand{\helpSize}{0.25cm}
+  \begin{tabular}{cc}\toprule
+    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule
+
+    %\rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
     \begin{tabular}{l}
-      \code{[4\, 7\, 6\, 9\, 0]$\to$5}\\
-\code{[3\, 8\, 2\, 1\, 4]$\to$4}\\
-\code{[2\, 2\, 9\, 4\, 8]$\to$3}\\
-\code{[0\, 9\, 8\, 9\, 9\, 8]$\to$1}\\
-\blueCode{$f(\ell) = $($f_0$ (car $\ell$))}\\\\
+\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
+%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
+\code{[9\, 8]$\to$[9\, 8\, 0]}\\
+\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\
 
@@
 \code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
-\blueCode{$f(\ell) = $($f_4$ + $\ell$ 0)}\\\\
+\blueCode{$f(\ell) = $($f_2$ + $\ell$ 0)}\\\\
 
 \code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
-\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
 \code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
-\blueCode{$f(\ell) = $($f_5$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
-
-      \code{[5\, 2\, 9]}$\to$\code{[[2\, 9]\, [9]\, []]}\\
-      \code{[3\, 8\, 1\, 3]}$\to$\code{[[8\, 1\, 3]\, [1\, 3]\, [3]\, []]}\\
-      \blueCode{$f(\ell) = $($f_2$ empty? cdr cdr $\ell$)}
+\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
+
+      \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
+      \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
+      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
+      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
       \\
 
-      \end{tabular}
-    \\\midrule
-    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+      \end{tabular}&
+
+%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
     \begin{tabular}{l}
-      \greenCode{$f_0($x$)\,=\,$(+ x 1)}\\
-      \hspace{\helpSize}($f_0$: \emph{increment})\\
-      \greenCode{$f_1($x$)\,=\,$(- x 1)}\\
-      \hspace{\helpSize}($f_1$: \emph{decrement})\\
-        \popp{\code{$f_2($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_2$ (n x))))}}\\
-        \hspace{\helpSize}($f_2$: \emph{unfold})\\
-        \popp{\code{$f_3($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{($f_3$ ($f_1$ i) (cdr l))))}}\\
-        \hspace{\helpSize}($f_3$: \emph{index})\\
-        \popp{\code{$f_4($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
-      \phantom{\code{$f_4($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_4$ (cdr l))))}}\\
-        \hspace{\helpSize}($f_4$: \emph{fold})\\
-        \greenCode{$f_5($f$,$l$)\,=\,$(if (empty? l) nil)}\\
-        \phantom{\greenCode{$f_5($f$,$l$)\,=\,$(if }}\greenCode{(cons (f (car l)) ($f_5$ (cdr l)))}\\
-        \hspace{\helpSize}($f_5$: \emph{map})
+        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
+      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
+        \hspace{\helpSize}($f_0$: \emph{unfold})\\
+        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
+      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
+        \hspace{\helpSize}($f_1$: \emph{index})\\
+        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
+      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
+        \hspace{\helpSize}($f_2$: \emph{fold})\\
+        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
+        \hspace{\helpSize}($f_3$: \emph{map})\\
+        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
+        \hspace{\helpSize}($f_4$: \emph{length})\\
+        \greenCode{$f_5(n)$\,=\,($f_0$ (= $n$) ($\lambda$ (x) x) (+ 1) 0)}\\
+        \hspace{\helpSize}($f_5$: \emph{range})
 
@@
   \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
-  \end{figure}
+  \end{figure*}
 
@@
 \bibliographystyle{icml2018}
+
+\appendix
+
+\section{Appendix}
+
+\subsection{Refactoring code with version spaces}\label{appendixVersion}
+Observe that every $\lambda$-calculus expression is also a version space,
+and that, if $e$ is an expression, then $\denotation{e} = \left\{e \right\}$.
+
+\begin{align}
+  S_n(v)& = \left\{\shift{n}v\mapsto \$n \right\}\cup
+  \begin{cases}
+    \text{if $v$ is primitive:}&\left\{\Lambda\mapsto v \right\}\\
+    \text{if $v = \$i$ and $i < n$:}&\left\{\Lambda\mapsto \$i \right\}\\
+    \text{if $v = \$i$ and $i\geq n$:}&\left\{\Lambda\mapsto \$(i + 1) \right\}\\
+    \text{if $v = \lambda b$:}&\left\{v'\mapsto \lambda b' \;:\; v'\mapsto b'\in S_{n + 1}(b)  \right\}\\
+    \text{if $v = (f\;x)$:}&\left\{v_1\cap v_2\mapsto (f'\;x') \;:\; v_1\mapsto f'\in S_n(f),\; v_2\mapsto x'\in S_n(x) \right\}\\
+    \text{if $v = \uplus V$:}&\bigcup_{v'\in V}S_n(v')\\
+    \text{if $v$ is $\varnothing$:}&\varnothing\\
+    \text{if $v$ is $\Lambda$:}&\left\{\Lambda\mapsto\Lambda \right\}\\
+  \end{cases}
+  \\
+
+  \end{align}
 \end{document}

### 2018-10-31 nips_2018.tex
version space archive progress

@@
 
-\title{Library Learning in Neurally-Guided Bayesian Program Induction}
+\title{Library Learning for Neurally-Guided Bayesian Program Induction}
 
@@
 
-\textbf{Compile: learning to tractably maxing $\lowerBound$ w.r.t. the
+\textbf{Compile: learning to tractably maximize $\lowerBound$ w.r.t. the
   frontiers.}  Here we train %% a neural network, $q$, to predict a
@@
 
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Explore)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Explore)}} (synthesis.east);
+    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,inner sep=0pt,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Explore)}} (synthesis.west);
+    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,inner sep=0pt,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Explore)}} (synthesis.east);
   \end{scope}}
@@
 %% bottom-up \emph{recognition model}
-\vspace{0.5cm}Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-recognition model training, and DSL induction.
-For added robustness,
-we interleave an extra program search step (Explore) before training the recognition model,
-and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Explore step.
-
-\begin{figure}[h]
-\centering  \begin{minipage}{0.9\textwidth}    
-    \begin{algorithm}[H]
-      \caption{The \system Algorithm}
-      \label{mainAlgorithm}
-      \begin{algorithmic}
-        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-        \FOR{$i=1$ {\bfseries to} $I$}
-        %     \FOR{$x:\tau\in X$}
-        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
-        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \hfill(\textbf{Explore})
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\textbf{Compile})
-        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
-        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \hfill\textbf{(Explore)}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\hfill\textbf{(Compress)}
-        %% \STATE Define $Q_x(z) \propto \begin{cases}
-        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-        %%   0&x\not \in \mathcal{F}_x
-        %% \end{cases}
-        \ENDFOR
-        \STATE \textbf{return} $\mathcal{D},\theta,q$
-      \end{algorithmic}
-    \end{algorithm}
-  \end{minipage}
-\end{figure}
 
@@
 
+%\vspace{0.5cm}
+Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
+recognition model training, and DSL induction.
+For added robustness,
+we interleave an extra program search step (Explore) before training the recognition model,
+and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Explore step.
+
+\begin{figure}[h]
+\centering  \begin{minipage}{0.9\textwidth}    
+    \begin{algorithm}[H]
+      \caption{The \system Algorithm}
+      \label{mainAlgorithm}
+      \begin{algorithmic}
+        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
+        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
+        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
+        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
+        \FOR{$i=1$ {\bfseries to} $I$}
+        %     \FOR{$x:\tau\in X$}
+        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
+        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
+        \hfill(\textbf{Explore})
+        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\textbf{Compile})
+        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
+        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
+        \hfill\textbf{(Explore)}
+
+        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\hfill\textbf{(Compress)}
+        %% \STATE Define $Q_x(z) \propto \begin{cases}
+        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
+        %%   0&x\not \in \mathcal{F}_x
+        %% \end{cases}
+        \ENDFOR
+        \STATE \textbf{return} $\mathcal{D},\theta,q$
+      \end{algorithmic}
+    \end{algorithm}
+  \end{minipage}
+\end{figure}
+
 %\section{Experiments}
@@
 
-\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
+\begin{wrapfigure}{r}{0.55\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
   \includegraphics[width = \functionSize]{figures/functions/6.png}
@@
   \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
+DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-0.5cm}
 \end{wrapfigure}
@@
 
-\begin{figure}\centering
+\begin{figure}[h]\centering
   \includegraphics[width = 4.35cm]{figures/listLearningCurve_color.eps}
@@
  Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, Bayesian program merging~\cite{hwang2011inducing}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
- Concurrent with this work, closely related ideas were applied to mining `code idioms' from imperative programs~\cite{8355713}
- and for synthesizing functional programs from natural language~\cite{Alex}.
+ Closely allied ideas have been applied to mining `code idioms' from  programs~\cite{Allamanis:2014:MIS:2635868.2635901}, and, concurrent with this work, using those idioms to better synthesize functional programs from natural language~\cite{Alex}.
 
@@
 
-\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work, and for funding from the NSF GRFP, the MUSE program (Darpa grant FA8750-14-2-0242),
+\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work, and for funding from the NSF GRFP, AFOSR award FA9550-16-1-0012, the MIT-IBM Watson AI Lab, the MUSE program (Darpa grant FA8750-14-2-0242),
 and an AWS ML Research Award.
+This material is based upon work supported by the Center for Brains, Minds and Machines (CBMM), funded by NSF STC award CCF-1231216.
 

### 2018-11-05 archive.tex
compression appendix

@@
 \documentclass{article}
-
+\usepackage[margin=1in]{geometry}
 \usepackage{dblfloatfix}    % To enable figures at the bottom of page
@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage[nohyperref]{icml2018}
+%\usepackage[nohyperref]{icml2018}
 
@@
  % For algorithms
-%\usepackage{algorithm}
+\usepackage{algorithm}
 %\usepackage{algorithmic}
@@
 
-% If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2018}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
-
 \begin{document}
-
-\twocolumn[
-\icmltitle{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2018
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Kevin Ellis}{MIT}
-\icmlauthor{Lucas Morales}{MIT}
-\icmlauthor{Mathias}{MIT}
-\icmlauthor{Maxwell Nye}{MIT}
-\icmlauthor{Armando Solar-Lezama}{MIT}
-\icmlauthor{Joshua B. Tenenbaum}{MIT}
-\end{icmlauthorlist}
-
-%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
-\icmlaffiliation{MIT}{MIT}
-%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}
-
-\icmlcorrespondingauthor{Kevin Ellis}{ellisk@mit.edu}
-%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Machine Learning, ICML}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
-%% \begin{abstract}
-%%   Successful approaches to program induction require a hand-engineered
-%%   domain-specific language (DSL), constraining the space of allowed
-%%   programs and imparting prior knowledge of the domain.  We contribute
-%%   a program induction algorithm called \system that learns a DSL while
-%%   jointly training a neural network to efficiently search for programs
-%%   in the learned DSL.\@ We use our model to synthesize functions on lists,
-%%   edit text, and solve symbolic regression problems, showing how the
-%%   model learns a domain-specific library of program components for
-%%   expressing solutions to problems in the domain.
-%% \end{abstract}
 
@@
 $\probability[x|p]$ while also being probable under the prior
-$(\mathcal{D},\theta)$.  Concretely, the recognition model $q$
-predicts, for each task $x\in X$, a weight vector $q(x) =
-\theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the
-DSL, this defines a distribution over programs,
-$\probability[p|\mathcal{D},\theta = q(x)]$.  We abbreviate this
-distribution as $q(p|x)$.  The crucial aspect of this framing is that
-the neural network leverages the structure of the learned DSL, so it
-is \emph{not} responsible for generating programs wholesale.  We share
-this aspect with DeepCoder~\cite{balog2016deepcoder}
-and~\cite{menon2013machine}.
-
+$(\mathcal{D},\theta)$.  Concretely, the recognition model $Q$
+predicts, for each task $x\in X$, a distribution over programs,
+$Q(\cdot |x)$.
+We want $Q(\cdot |x)$ to approximate the posterior $\probability[\cdot |x,\mathcal{D},\theta]$.
+
+\subsubsection{Training $Q$}
 How should we get the data to train $q$?
@@
 
+\subsubsection{Parameterizing $Q$}
+
+Broadly the literature contains two different approaches to
+parameterizing conditional distributions over programs.  The first
+approach, seen in systems like RobustFill~\cite{devlin2017robustfill}
+and~\cite{zavershynskyi2018naps}, is to use a recurrent neural network
+to predict the program token-by-token ala Seq2Seq or Seq2Tree.  The
+advantage of this approach is that the neural net can predict the
+entire program, so if the network is sufficiently powerful, it can
+completely solve the synthesis problem.  There are two disadvantages
+to this approach.  First, these models perform poorly at out-of-sample
+generalization~\cite{}, which is critical for our setting, as the
+agent may need to solve new tasks that are qualitatively different
+from the tasks it has solved so far.  Second, a powerful deep
+recurrent network may be costly to sample or enumerate from ---
+so if the network cannot easily solve a task, we cannot compensate
+with rapid sampling or enumeration.  In contrast, state-of-the-art
+enumerative program synthesizers evaluate millions of programs per
+second~\cite{feser2015synthesizing}.
+
+The second approach is to have $Q$ predict a fixed-dimensional weight
+vector, which then biases a fast enumerator~\cite{balog2016deepcoder}
+or sampler~\cite{menon2013machine}.  This approach can enjoy strong
+out-of-sample generalization, because it can fall back on enumeration
+or sampling when the target program is unlike the training programs.
+A main drawback is that the neural net is deliberately handicapped,
+and can only send so much information about the target program.
+
+We adopt a middle ground between these two extremes.  Our recognition
+model predicts a distribution over primitives in the DSL,
+conditioned on the local context in the syntax tree of the
+program. When predicting the next node to add to the syntax tree of a program,
+the recognition model
+conditions on the parent node, as well as
+which argument is being generated.
+This is a kind of `bigram' model over trees,
+where the bigrams
+take the form of (parent, child, argument index).
+See Figure~\ref{bg}.
+This parameterization
+has three main advantages:
+(1) it supports fast enumeration
+and sampling of programs,
+because the
+recognition model
+only needs to run once for each task;
+(2) it allows the recognition model to provide fine-grained
+information about the structure of the target program;
+and (3)
+training this recognition model
+causes it to learn to break symmetries in the space of programs,
+described next.
+\begin{figure}
+\centering  \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
+    \node(l)[draw] at (0,0) {\code{$\lambda$ (x)}};
+    \node(k)[draw] at ([yshift=-50]l.south) {\code{+}};
+    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\code{1}};
+    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\code{*}};
+    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\code{x}};
+    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\code{x}};
+
+    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{+}|\text{start})$}(k.north);
+    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{1}|\code{+}_1)$}(o.north);
+    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{*}|\code{+}_2)$}(m.north);
+    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{x}|\code{*}_2)$}(x1.north);
+    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{x}|\code{*}_1)$}(x2.north);
+\end{tikzpicture}
+\caption{Parameterization of distribution over programs predicted by recognition model.
+  Here the program (syntax tree shown above) is \code{($\lambda$ (x) (+ 1 (* x x)))}.
+Each conditional distribution predicted by the recognition model is written $q(\text{child}|\text{parent}_{\text{argument index}})$.}
+  \end{figure}
+
 \subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
@@
 new abstractions which will later be used to discover more programs.
-
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
@@
 where $p'\manyReduce p$
-is the standard notation for ``expression $p'$ evaluates to $p$ by the standard rules of $\lambda$-calculus''~\cite{pierce}.
+is the standard notation for ``expression $p'$ evaluates to $p$ by the rules of $\lambda$-calculus''~\cite{pierce}.
 
@@
 fact grows exponentially quickly both as a function of $n$ and as a
-function of the size of the program being refactored.  Next, we show
-how to tame this exponential explosion.
+function of the size of the program being refactored.  For example,
+for the programs in Figure~\ref{mapFactor}, there are approximately
+$10^{14}$ possible refactorings.  Next, we show how to tame this
+exponential explosion.
 
@@
 that represents the set of possible refactorings (i.e., $R_n(p)$).
-
-A version space is either:
+This technique is astronomically more efficient than
+explicitly representing
+the space of possible refactorings:
+for the example in Figure~\ref{mapFactor},
+we represent
+the space of refactorings using a version space with
+$10^6$ nodes,
+which encodes $10^{14}$ refactorings.
+Formally, a version space is either:
 \begin{itemize}
-\item A deBuijn index: written $\$i$, where $i$ is a natural number
+\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus.
+When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to.
+For example, $\lambda x.\lambda y.
+(x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices.
+See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
 \item An abstraction: written $\lambda v$, where $v$ is a version space
-\item A  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
+\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
 \item A  union: $\uplus V$, where $V$ is a set of version spaces
@@
 We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
-\begin{align*}
+\begin{align}
   \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
-    e\text{, if $e\in \mathcal{D}$ and $e\in v$}\\
+    e\text{, if $e\in \mathcal{D}$ and $e\in v$. Exploits the fact that $e\in v$ can be efficiently computed.}\\
     \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
-  \end{cases}\\
-  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\
-  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\
-  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\
-  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\ell(e|\mathcal{D})
+  \end{cases}\label{FACTORING}\\\nonumber
+  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\\nonumber
+  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
+  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\\nonumber
+  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
   \end{align*}
-where $\ell(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
+where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
 is the size of the syntax tree of $e$,
 when members of $\mathcal{D}$ are counted as having size 1.
-Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\ell(p|\mathcal{D})$.
+Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.
 
@@
 With this machinery in hand, we now have all the pieces needed to
-learn a DSL (Algorithm~\ref{grammarInductionAlgorithm})
-
-    \begin{algorithm*}[H]
+learn a DSL (Algorithm~\ref{grammarInductionAlgorithm}).
+To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+  and $\lambda$ controls how strongly we regularize the size of the DSL.
+  We place a symmetric Dirichlet prior over the weight vector $\theta$.
+
+To appropriately score each proposed $\mathcal{D}$ we must reestimate
+ the weight vector $\theta$ (Algorithm~\ref{grammarInductionAlgorithm}, line 7).
+Although this  may seem 
+very similar to estimating the parameters of a probabilistic context free grammar,
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+our DSLs are context-sensitive due to the presence of variables
+in the programs and also due to the polymorphic typing system.
+Appendix~\ref{mapAppendix} derives a tractable MAP estimator for $\theta$.
+
+\begin{algorithm}%[H]
   %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
@@
   \begin{algorithmic}[1]
-         \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
+    \State {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-         \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-         \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-         \WHILE {true}         
-         \STATE $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
-         \STATE Define $\text{refactor}(p|\mathcal{D}') = \argmin_{e\in \denotation{v_p}} \ell (e|\mathcal{D}')$ \Comment{Rewrites $p$ assuming $\mathcal{D}'$ is the DSL}
-         \STATE Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
-         \STATE Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
-         \STATE Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
-         \STATE components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
-         \STATE proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
-         \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
-         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
-         \STATE $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
-         \STATE $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
-         \ENDWHILE
+    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
+    \While{true}         
+    \State $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
+    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
+    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
+    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
+    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
+    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
+    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
+    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
+    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
+    \State $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
+    \EndWhile
   \end{algorithmic}
@@
   \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z z)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
-  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z 1)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
+  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (- z 1)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
       \end{tabular}      
@@
 \bibliography{main}
-\bibliographystyle{icml2018}
+\bibliographystyle{plain}
 
@@
 \subsection{Refactoring code with version spaces}\label{appendixVersion}
-Observe that every $\lambda$-calculus expression is also a version space,
-and that, if $e$ is an expression, then $\denotation{e} = \left\{e \right\}$.
-
-\begin{align}
-  S_n(v)& = \left\{\shift{n}v\mapsto \$n \right\}\cup
+Recall that our goal is to define an operator over version spaces, $I\beta_n$,
+which calculates the set of $n$-step refactorings.
+We define this operator recursively as follows:
+\begin{align*}
+  I\beta_n(v)& = \uplus \left\{ \underbrace{I\beta'(I\beta'(I\beta'(\cdots}_{i \text{ times}} v))) \;:\; 0\leq i \leq n \right\}\\
+  I\beta'(v)& = \uplus \left\{(\lambda b)v\;:\;v\mapsto b\in S_0(v)\text{, when }v\not=\Lambda \right\}\cup
+  \begin{cases}
+    \text{if $v$ is a primitive or index:}&\varnothing\\
+    \text{if $v = \lambda b$:}&\lambda I\beta'(b)\\
+    \text{if $v = (f\;x)$:}&(I\beta'(f)\;I\beta'(x))\\
+  \end{cases}
+  \\
+  S_k(v)& = \left\{\shift{k}v\mapsto \$k \right\}\cup
   \begin{cases}
     \text{if $v$ is primitive:}&\left\{\Lambda\mapsto v \right\}\\    
-    \text{if $v = \$i$ and $i < n$:}&\left\{\Lambda\mapsto \$i \right\}\\
-    \text{if $v = \$i$ and $i\geq n$:}&\left\{\Lambda\mapsto \$(i + 1) \right\}\\
-    \text{if $v = \lambda b$:}&\left\{v'\mapsto \lambda b' \;:\; v'\mapsto b'\in S_{n + 1}(b)  \right\}\\
-    \text{if $v = (f\;x)$:}&\left\{v_1\cap v_2\mapsto (f'\;x') \;:\; v_1\mapsto f'\in S_n(f),\; v_2\mapsto x'\in S_n(x) \right\}\\
+    \text{if $v = \$i$ and $i < k$:}&\left\{\Lambda\mapsto \$i \right\}\\
+    \text{if $v = \$i$ and $i\geq k$:}&\left\{\Lambda\mapsto \$(i + 1) \right\}\\
+    \text{if $v = \lambda b$:}&\left\{v'\mapsto \lambda b' \;:\; v'\mapsto b'\in S_{k + 1}(b)  \right\}\\
+    \text{if $v = (f\;x)$:}&\left\{v_1\cap v_2\mapsto (f'\;x') \;:\; v_1\mapsto f'\in S_k(f),\; v_2\mapsto x'\in S_k(x) \right\}\\
     \text{if $v = \uplus V$:}&\bigcup_{v'\in V}S_n(v')\\
     \text{if $v$ is $\varnothing$:}&\varnothing\\
-    \text{if $v$ is $\Lambda$:}&\left\{\Lambda\mapsto\Lambda \right\}\\
+    \text{if $v$ is $\Lambda$:}&\left\{\Lambda\mapsto\Lambda \right\}
   \end{cases}
-  \\
-
+\end{align*}
+where $\shift^k$ is the shifting operator~\cite{pierce},
+which adds $k$ to all of the free variables in a $\lambda$-expression or version space.
+We have written the definition of $I\beta_n$ recursively,
+but implement it using a dynamic program:
+we hash cons each version space,
+and only calculate the operators $I\beta_n$,
+$I\beta'$, and $S_k$ once per each version space.
+Refactoring is similarly done more quickly with dynamic programming (see Equation~\ref{FACTORING} for the recursive definition of \textsc{refactor}).
+
+\subsubsection{Computational complexity of DSL learning}
+
+How long does each update to the DSL in
+Algorithm~\ref{grammarInductionAlgorithm} take?  Constructing the
+version spaces takes time linear in the number of programs (written
+$P$) in the frontiers (Algorithm~\ref{grammarInductionAlgorithm}, line
+5), and, in the worst case, exponential time as a function of the
+number of refactoring steps $n$ --- but we bound the number of steps
+to be a small number (typically $n = 3$).  Writing $V$ for the number
+of version spaces, this means that $V$ is $O(P2^n)$.  The number of
+proposals (line 10) is linear in the number of distinct version
+spaces, so is $O(V)$.  For each proposal we have to refactor every
+program (line 6), so this means we spend $O(V^2) = O(P^22^n)$ per DSL
+update.  In practice this quadratic dependence on $P$ (the number of
+programs) is prohibitively slow.
+We now describe a linear time approximation to the
+refactor step in Algorithm~\ref{grammarInductionAlgorithm} based on beam search.
+
+For each version space $v$ we calculate a \emph{beam}, which is a
+function from a DSL $\mathcal{D}$ to a shortest program in
+$\denotation{v}$ using primitives in $\mathcal{D}$.  Our strategy will
+be to only maintain the top $B$ shortest programs in the beam;
+throughout all of the experiments in this paper, we set $B = 10^6$,
+and in the limit $B\to\infty$ we recover the exact behavior of \textsc{refactor}.
+The following recursive equations
+define how we calculate these beams;
+the set `proposals' is defined in line 10 of Algorithm~\ref{grammarInductionAlgorithm},
+and $\mathcal{D}$ is the current DSL:
+\begin{align*}
+  \text{beam}_v(\mathcal{D}')& = \begin{cases}
+    \text{if }\mathcal{D}'\in \text{dom}(b_v)\text{: }&b_v(\mathcal{D}')\\
+    \text{if }\mathcal{D}'\not\in \text{dom}(b_v)\text{: }&\textsc{refactor}(v,\mathcal{D})
+  \end{cases}\\
+  b_v& = \text{the $B$ pairs $(\mathcal{D}'\mapsto p)$ in $b_v'$ where the syntax tree of $p$ is smallest}\\
+  b_v'(\mathcal{D}')& = \begin{cases}
+    \text{if $\mathcal{D}'\in \text{proposals}$ and $e\in \mathcal{D}'$ and  $e\in v$: }e\\
+    \text{otherwise if $v$ is a primitive or index:} $v$
+    \text{otherwise if $v = \lambda b$: } $\lambda $\text{beam}_b(\mathcal{D}')\\
+    \text{otherwise if $v = (f\;x)$: } $(\text{beam}_f(\mathcal{D}')\;\text{beam}_x\mathcal{D}'))\\
+    \text{otherwise if $v = \uplus V$: } \argmin_{e\in \left\{b'_{v'}(\mathcal{D}')\;:\;v'\in V \right\}}\text{size}(e|\mathcal{D}')
+    \end{cases}
+  \end{align*}
+We calculate $\text{beam}_v(\cdot )$ for each version space using
+dynamic programming.  Using a minheap to represent
+$\text{beam}_v(\cdot )$, this takes time $O(VB\log B)$, replacing the
+quadratic dependence on $V$ (and therefore the number of programs,
+$P$) with a $B\log B$ term, where the parameter $B$ can be chosen
+freely, but at the cost of a less accurate beam search.
+
+After performing this beam search,
+we take only the top $I$ proposals as measured by $-\sum_x\min_{p\in \mathcal{F}_x}\text{beam}_{v_p}(\mathcal{D}')$.
+We set $I = 300$ in all of our experiments,
+so $I \ll B$.
+The reason why we
+only take the top $I$ proposals (rather than take the top $B$) is because
+parameter estimation (estimating $\theta$ for each proposal) is much more expensive than
+performing the beam search ---
+so we perform a very wide beam search and then at the very end
+tim the beam down to
+only $I = 300$ proposals.
+Next,
+we describe our MAP estimator for the continuous parameters ($\theta$) of the DSL.
+
+\subsection{Estimating the continuous parameters $\theta$ of a DSL}\label{mapAppendix}
+We use an EM algorithm to estimate the continuous parameters of the DSL, i.e. $\theta$.
+Suppressing dependencies on $\mathcal{D}$, the EM updates are
+\begin{align}
+\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
+  Q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{F}_x \right]
 \end{align}
+In the M step of EM we will update $\theta$ by instead maximizing a lower bound on $\log \probability[p|\theta]$,
+making our approach an instance of Generalized EM.
+
+We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$; $c(p)= \sum_{e\in \mathcal{D}}c(e,p)$ to mean the total number of primitives used in program $p$; $c(\tau,p)$ to mean the number of times that type $\tau$ was the input to sample in Algorithm~\ref{sampleProgram} while sampling program $p$.
+Jensen's inequality gives a lower bound on the likelihood:
+\begin{align*}
+  &\sum_x\expect_{Q_x}\left[  \log \probability[p|\theta] \right] =\\
+  &\sum_{e\in \mathcal{D}} \log \theta_e \sum_x\expect_{Q_x}\left[c(e,p_x) \right] -
+  \sum_\tau\expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+ =   &\sum_e C(e)\log \theta_e  - \beta\sum_\tau\frac{\expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]}{\beta}\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+ \geq    &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{\expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+     =     &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{R(\tau)}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  
+  %% &\geq\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \frac{1}{c(p)}\sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
+  %% & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log\frac{1}{c(p)} \sum_{e\in \mathcal{D}} r(e,p)\theta_e
+\end{align*}
+where we have defined
+\begin{align*}
+  C(e)&\triangleq  \sum_x\expect_{Q_x}\left[c(e,p_x) \right]\\
+  R(\tau)&\triangleq \expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]\\
+  \beta&\triangleq\sum_\tau \expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]
+\end{align*}
+Crucially it was defining $\beta$ that let us use Jensen's inequality. 
+Recalling from the main paper that $P(\theta)\triangleq\text{Dir}(\alpha)$,
+we have the following lower bound on M-step objective:
+\begin{align}
+\sum_e (C(e) + \alpha)\log \theta_e  - \beta\log \sum_\tau\frac{R(\tau)}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e    
+\end{align}
+Differentiate with respect to $\theta_e$, where $e:\tau$, and set to zero to obtain:
+\begin{align}
+  &  \frac{C(e) + \alpha}{\theta_e}\propto\sum_{\tau'}\indicator\left[\text{unify}(\tau,\tau') \right] R(\tau')\\
+&  \theta_e\propto\frac{C(e) + \alpha}{\sum_{\tau'}\indicator\left[\text{unify}(\tau,\tau') \right] R(\tau')}
+\end{align}
+The above is our estimator for $\theta_e$.
+The above estimator has an intuitive interpretation.
+The quantity $C(e)$ is the expected number of times that we used $e$.
+The quantity $\sum_{\tau'}\indicator\left[\text{unify}(\tau,\tau') \right] R(\tau')$
+is the expected number of times that we \emph{could have} used $e$.
+The hyperparameter $\alpha$ acts as pseudocounts that are
+added to the number of times that we used each primitive,
+and are not added to the number of times that we could have used each primitive.
+
+We are only maximizing a lower bound on the log posterior; when is this lower bound tight?
+This lower bound is tight whenever all
+of the types of the expressions in the DSL are not polymorphic, in which case our DSL is equivalent to a PCFG
+and this estimator is equivalent to the inside/outside algorithm.
+Polymorphism introduces context-sensitivity to the DSL,
+and exactly maximizing the likelihood with respect to $\theta$
+becomes intractable,
+so for domains with polymorphic types we use this estimator.
+
 \end{document}

### 2018-11-11 archive.tex
working on archive

@@
 \usepackage{booktabs} \usepackage{stmaryrd}% for professional tables
-\usepackage{stmaryrd}(
+
 \usepackage[section]{placeins}
@@
 \usepackage{circuitikz}
-\usetikzlibrary{fit,bayesnet}
-%\usetikzlibrary{arrows.meta}
-%\usetikzlibrary{positioning}
-%\usetikzlibrary{decorations.text}
+\usetikzlibrary{fit,bayesnet,calc}
+\usetikzlibrary{arrows.meta}
+\usetikzlibrary{positioning}
+\usetikzlibrary{decorations.text}
 \usetikzlibrary{decorations.pathmorphing}
@@
 So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
-\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+\begin{definition}
+   A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
 is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
-
+  \end{definition}
 Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
@@
 and $(\mathcal{D},\theta)$ (\textbf{Sleep-G}):
-
 \\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
@@
 Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $q(\cdot |\cdot )$,
-during the \textbf{Sleep-R} phase: $q$ is trained to approximate the
-posterior over programs, $q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}}]$.
+of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $Q(\cdot |\cdot )$,
+during the \textbf{Sleep-R} phase: $Q$ is trained to approximate the
+posterior over programs, $Q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
   Thus training the neural network amortizes the cost of finding programs with high posterior probability.
 
-\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
+\noindent\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
   frontiers.}  Here we train %% a neural network, $q$, to predict a
@@
 %% is
-$q(p|x)$ to assign high probability to programs $p$ where
+$Q(p|x)$ to assign high probability to programs $p$ where
 $\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
@@
     \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $q$};
+    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
 
@@
 Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
-\end{figure}
-
-\begin{figure}
-  \begin{tikzpicture}
-    \node at (0,0) (d){DSL};
-    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
-
-    \node at ([xshift = 2cm]t) (nn){
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
-    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
-
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
-    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
-
-    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
-
-    \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
-    \draw [->  ] (s.east) -- (f.west);
-
-    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
-
-    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
-
-  \end{tikzpicture}
-
-  \vspace{1cm}
-
-  \begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
-
-    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
-    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
-    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
-
-    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
-    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
-    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
-
-    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
-    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
-    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
-  \end{tikzpicture}
-
-  \begin{tikzpicture}
-    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
-    \node at ([xshift = 3cm]d.east) (p2){program};
-    \node at ([yshift = 1.5cm]p2) (p1){program};
-    \node at ([yshift = -1.5cm]p2) (p3){program};
-
-    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
-    \draw[squiggle,-> ] (d.east) -- (p1.west);
-    \draw[squiggle,-> ] (d.east) -- (p3.west);
-
-    \node at ([xshift = 2cm]p1.east) (t1){task};
-    \node at ([xshift = 2cm]p2.east) (t2){task};
-    \node at ([xshift = 2cm]p3.east) (t3){task};
-    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
-    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
-    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
-  \end{tikzpicture}
-
-  \vspace{2cm}
-
-  \begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
-    \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
-    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
-
-    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
-    \node[anchor=north] at (c.north) {Compression};
-
-    \draw [-> ] (f1.east) -- (c.west|-f1.east);
-    \draw [-> ] (f2.east) -- (c.west|-f2.east);
-    \draw [-> ] (f3.east) -- (c.west|-f3.east);
-
-    \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
-    \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
-    \draw [-> ] (c.east) -- (d.west);
-    \draw [-> ] (c.east) -- (t.west);
-
-    \node at (c.center) {
-\begin{tikzpicture}[scale=0.7]
-    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
-    %%     \footnotesize Program:\\
-    %%     \code{($\lambda$ (x) (+ (- x) 1))}
-    %% \end{tabular}};
-    %\node at (,0.5) {\code{cons}};
-%    \node [rotate=90] at (-2.3,-0.5) {\small program};
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-%    \draw  (l1.south) -- (-0.5,0.5);
-
-  %% \node(c) at (-0.5,-1.5) {\code{-}};
-  %% \node(z) at (0.5,-1.5) {\code{x}};
-
-  %% \draw (0,-1) -- (c.north);
-  %% \draw (0,-1) -- (z.north);
-
-  \begin{scope}[shift={(-1,-2.5)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-%  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-  \end{scope}
-
-\begin{scope}[shift={(0,-5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  %\node(x1) at (0,-1) {};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.55,-0.4) -- (a.north);
-%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
-
-  \end{scope}
-
-\end{tikzpicture}
-    };
-
-    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};
-
-    \end{tikzpicture}
-  \end{figure}
-
-\begin{figure}[h]
-\centering
-%  \includegraphics{figures/iterations.eps}
-\begin{tikzpicture}
-  \begin{scope}[shift={(2.5,0)}]
-    \node[obs] at (3,3) (dx){$\mathcal{D}$};
-    \node[latent] at (3.5,1.75) (zp){$p$};
-%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-    \node[obs] at (3.5,0.7) (xp) {$x$};
-    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
-    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
-%    \draw [->,red] (tx) -- (zp);
-    \draw [->] (dx) -- (zp);
-    \node at (nn) {
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-  \end{scope}
-
-  \begin{scope}[shift={(0.7,-4.2)}]
-    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-    \draw [->] (p2.south) -- (x2.north);
-    \draw [->] (dt.south) -- (p2.north);
-
-    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-    \draw [->] (p1.south) -- (x1.north);
-    \draw [->] (dt.south) -- (p1.north);
-
-    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-    \draw [->] (p3.south) -- (x3.north);
-    \draw [->] (dt.south) -- (p3.north);
-
-    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
-
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-    \draw [->] (zp.south) -- (xp.north);
-    \plate {}{(zp)(xp)}{$x\in X$};
-
-    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
-    rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
-  \end{scope}
-
-  \begin{scope}[shift={(7.7,-4)}]
-    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-    \node[latent] at (1.5,3) (t){$\theta$};
-    \node[obs] at (1,1.75) (z){$p$};
-%     \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-    \node[obs] at (1,0.5) (x) {$x$};
-    \edge {z}{x};
-    \edge {d,t}{z};
-    \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
-  \end{scope}
-
-\end{tikzpicture}
 \end{figure}
@@
 
-\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
+\begin{definition}
 A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-%\\\noindent\textbf{Definition.}
 A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
@@
 and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
-
-Together with its weight vector,
-a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
-In the supplement, we define this distribution  
-by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$.
-Care must be taken to ensure that programs are well-typed
- and that variable scoping rules are obeyed.
-%% With this distribution in hand, we search for programs by enumerating
-%% $\lambda$-calculus expressions in decreasing order of their probability under
-%% $(\mathcal{D},\theta)$.
+\end{definition}
+
+Together with its weight vector, a DSL defines a distribution over
+programs, $\probability[p|\mathcal{D},\theta]$.  We define this
+distribution by specifying a procedure for drawing samples from
+$\probability[p|\mathcal{D},\theta]$ (Algorithm~\ref{sampleProgram}).
+Care must be taken to ensure that variable scoping rules are obeyed
+and that programs are well-typed.  We ensure well-typed programs by
+performing Hindley-Milner type inference~\cite{pierce} during
+sampling, and assume that each task is annotated with the type of the
+program that will solve it.
+ \begin{algorithm}
+   \caption{Generative model over programs}
+   \label{sampleProgram}
+   \begin{algorithmic}[1]
+     \State \textbf{function} sample$(\mathcal{D}, \theta, \tau)$:
+     \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, type $\tau$
+     \State \textbf{Output:} a program whose type unifies with $\tau$
+     \State  \textbf{return }sample$'(\mathcal{D}, \theta, \varnothing, \tau)$
+
+     \Statex
+     \State \textbf{function} sample$'(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
+     \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$\Comment{Environment $\mathcal{E}$ starts out as $\varnothing$}
+     \State \textbf{Output:} a program whose type unifies with $\tau$
+     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
+     \State var $\gets$ an unused variable name
+     \State body $\sim$ sample$'(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$\Comment{Recursively sample function body}
+     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
+     %\Endif
+     \Else\Comment{Build an application to give something w/ type $\tau$}
+     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
+     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
+     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
+       \theta_e&\text{ if }e\in \mathcal{D}\\
+\theta_{var}/|\text{variables}|&\text{ if }e\in \mathcal{E}
+       \end{cases}$
+     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
+     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
+     %   \State unify$(\tau,\beta)$
+     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
+     \State $a_k\sim\text{sample}'(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
+     \EndFor
+     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
+     \EndIf
+     \Statex
+     \Statex\textbf{where:}
+     \State yield$(\tau) = \begin{cases}
+       \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+       \tau   &\text{ otherwise.}
+     \end{cases}$ \Comment{Final return type of $\tau$}
+     \State  args$(\tau) = \begin{cases}
+       [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+       []   &\text{ otherwise.}
+     \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
+   \end{algorithmic}
+ \end{algorithm}
 
@@
 $\probability[x|p]$ while also being probable under the prior
-$(\mathcal{D},\theta)$.  Concretely, the recognition model $Q$
-predicts, for each task $x\in X$, a distribution over programs,
-$Q(\cdot |x)$.
-We want $Q(\cdot |x)$ to approximate the posterior $\probability[\cdot |x,\mathcal{D},\theta]$.
-
-\subsubsection{Training $Q$}
-How should we get the data to train $q$?
-This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
-One approach is to sample programs from the DSL,
+$(\mathcal{D},\theta)$,
+thereby approximating the posterior $\probability[\cdot |x,\mathcal{D},\theta]$.
+
+\subsubsection{Training $Q$}\label{recognitionTraining}
+How should we get the data to train $Q$?
+This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
+One approach is to draw programs from the DSL,
 run them to get their input/outputs,
-and then train $q$ to predict the program from the input/outputs.
-This is like how a Helmholtz machine
-trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
-The advantage of ``Helmholtz machine'' training is that
+and then train $Q$ to predict the program from the input/outputs.
+This is like how the wake-sleep algorithm for the Helmholtz machine
+trains its recognition model during its sleep phase~\cite{dayan1995helmholtz}.
+The advantage of training on samples, or ``dreams,'' is that
 we can draw unlimited samples from the DSL,
@@
 Another approach is
- self-supervised learning,
-training $q$ on the (program, task)
-pairs discovered during waking.
-The advantage of self-supervised learning
+to train $Q$ on the (program, task)
+pairs discovered by the Wake phase.
+The advantage here
 is that the training data is much higher quality,
-because we are training on the actual tasks.
+because we are training on real tasks.
 Due to these complementary advantages,
@@
 
-Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
- equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
-%% \begin{equation*}
-%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
-%% \end{equation*}
- where the expectation is taken over tasks.
-Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
- The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives:
- \begin{align}
-   \mathcal{L}_{\text{RM}} &= \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}\\
-\nonumber\mathcal{L}_{\text{HM}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\\
-\nonumber\mathcal{L}_{\text{SS}} &= \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
-\end{align}
-
-  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
+Formally, $Q$ should approximate the true posteriors over programs.
+We can either train $Q$ to perform full posterior inference by minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|Q(p|x) \right) \right]$,
+or we can train $Q$ to perform MAP inference
+by maximizing $\expect\left[\max_{p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]} \log Q(p|x) \right]$,
+where in both cases the expectation is taken over tasks.
+Taking this expectation over the empirical distribution of tasks trains $Q$ on the real data; taking it over samples from the generative model trains $Q$ on ``dreams.''
+We define a pair of alternative objectives for the recognition model,
+$\mathcal{L}_{\text{RM}}^{\text{posterior}}$ and $\mathcal{L}_{\text{RM}}^{\text{MAP}}$,
+which either train $Q$ to perform full posterior inference or MAP inference, respectively.
+These objectives combine real-data and dream data:
+\begin{align*}
+  \mathcal{L}_{\text{RM}}^{\text{posterior}} &= \mathcal{L}_{\text{Real}}^{\text{posterior}} + \mathcal{L}_{\text{Dream}}^{\text{posterior}}&
+  \mathcal{L}_{\text{RM}}^{\text{MAP}} &= \mathcal{L}_{\text{Real}}^{\text{MAP}} + \mathcal{L}_{\text{Dream}}^{\text{MAP}}\\
+  \mathcal{L}_{\text{Real}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
+\mathcal{L}_{\text{Real}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
+  \mathcal{L}_{\text{Dream}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
+    \mathcal{L}_{\text{Dream}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
+\end{align*}
+  The `dream' objectives are essential for data efficiency:
 all of our experiments train \system on only a few hundred tasks, which is too little for
-a high-capacity neural network $q$.
+a high-capacity neural network.
 Once we bootstrap a $(\mathcal{D},\theta)$,
@@
 and train $q$ on those samples.
-
-Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
+But, evaluating $\mathcal{L}_{\text{Dream}}$ involves drawing programs from
 the current DSL, running them to get their outputs,
-and then training $q$ to regress from the input/outputs to the program.
+and then training $Q$ to regress from the input/outputs to the program.
 Since these programs map inputs to outputs,
@@
 
-\subsubsection{Parameterizing $Q$}
+The $\mathcal{L}_{\text{Dream}}^{\text{MAP}}$ objective involves
+finding the MAP program solving a task drawn from the DSL.
+To make this tractable,
+rather than \emph{sample} programs as training data for $\mathcal{L}_{\text{Dream}}^{\text{MAP}}$,
+we \emph{enumerate} programs
+in decreasing order of their prior probability,
+tracking,
+for each dreamed task $x$,
+the set of enumerated programs maximizing
+$\probability[x,p|\mathcal{D},\theta]$.
+
+\subsubsection{Parameterizing $Q$}\label{recognitionParameterization}
 
@@
 and~\cite{zavershynskyi2018naps}, is to use a recurrent neural network
-to predict the program token-by-token ala Seq2Seq or Seq2Tree.  The
+to predict the program token-by-token as in Seq2Seq or Seq2Tree.  The
 advantage of this approach is that the neural net can predict the
@@
 completely solve the synthesis problem.  There are two disadvantages
-to this approach.  First, these models perform poorly at out-of-sample
+to this approach.  First, these models can perform poorly at out-of-sample
 generalization~\cite{}, which is critical for our setting, as the
@@
 take the form of (parent, child, argument index).
-See Figure~\ref{bg}.
+
+Formally,
+$Q$ predicts a $(|\mathcal{D}|+2)\times (|\mathcal{D}|+1)\times A$-dimensional tensor,
+where $A$ is the maximum arity\footnote{The arity of a function is the number of arguments that it takes as input.} of any primitive in the DSL.
+Slightly abusing notation, we write this tensor as $Q_{ijk}(x)$,
+where $x$ is a task,
+$i\in \mathcal{D}\cup\left\{\text{start},\text{var}\right\}$,
+$j\in \mathcal{D}\cup\left\{\text{var} \right\}$,
+and $k\in \left\{1,2,\cdots,A \right\}$.
+The output $Q_{ijk}(x)$
+controls the probability of
+sampling primitive $j$ given that
+$i$ is the parent node in the syntax tree
+and we are sampling the $k^{\text{th}}$ argument.
+Figure~\ref{bg} diagrams this generative process, Algorithm~\ref{recognitionSample}
+specifies a sampling procedure for $Q(\cdot |x)$,
+and Figure~\ref{neuralPipeline} diagrams the neurally guided program inference procedure.
 This parameterization
@@
 \centering  \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
-    \node(l)[draw] at (0,0) {\code{$\lambda$ (x)}};
-    \node(k)[draw] at ([yshift=-50]l.south) {\code{+}};
-    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\code{1}};
-    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\code{*}};
-    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\code{x}};
-    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\code{x}};
-
-    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{+}|\text{start})$}(k.north);
-    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{1}|\code{+}_1)$}(o.north);
-    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{*}|\code{+}_2)$}(m.north);
-    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{x}|\code{*}_2)$}(x1.north);
-    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{x}|\code{*}_1)$}(x2.north);
+    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
+    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
+    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
+    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
+    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
+    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};
+
+    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
+    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
+    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
+    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
+    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
 \end{tikzpicture}
 \caption{Parameterization of distribution over programs predicted by recognition model.
-  Here the program (syntax tree shown above) is \code{($\lambda$ (x) (+ 1 (* x x)))}.
-Each conditional distribution predicted by the recognition model is written $q(\text{child}|\text{parent}_{\text{argument index}})$.}
+  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
+Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task.}\label{bg}
 \end{figure}
+ \begin{algorithm}
+   \caption{Drawing from distribution over programs predicted by recognition model. Compare w/ Algorithm~\ref{sampleProgram}}
+   \label{recognitionSample}
+   \begin{algorithmic}[1]
+     \State \textbf{function} recognitionSample$(Q, x, \mathcal{D}, \tau)$:
+     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, type $\tau$
+     \State \textbf{Output:} a program whose type unifies with $\tau$
+     \State \textbf{return } $\text{recognitionSample}'(Q,x,\text{start},1,\mathcal{D},\varnothing,\tau)$
+     \Statex
+
+     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
+     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
+     \State \textbf{Output:} a program whose type unifies with $\tau$
+     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
+     \State var $\gets$ an unused variable name
+     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
+     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
+     %\Endif
+     \Else\Comment{Build an application to give something w/ type $\tau$}
+     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
+     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
+     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
+       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
+Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
+       \end{cases}$
+     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
+     \State newParent$\gets\begin{cases}
+     e &\text{ if }e\in \mathcal{D}\\
+     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
+     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
+     %   \State unify$(\tau,\beta)$
+     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
+     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
+     \EndFor
+     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
+     \EndIf
+     %% \Statex
+     %% \Statex\textbf{where:}
+     %% \State yield$(\tau) = \begin{cases}
+     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+     %%   \tau   &\text{ otherwise.}
+     %% \end{cases}$ \Comment{Final return type of $\tau$}
+     %% \State  args$(\tau) = \begin{cases}
+     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+     %%   []   &\text{ otherwise.}
+     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
+   \end{algorithmic}
+ \end{algorithm}
+ \begin{figure}
+   \begin{tikzpicture}
+     \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
+     \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
+       e.g., RNN or CNN};
+     \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
+     \node[draw,align = center](s) at ([xshift=3cm]m.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
+     \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};
+
+     \draw[->] (x.east) -- (e.west);
+     \draw[->] (e.east) -- (m.west);
+     \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
+     \draw[->] (s.east) -- (f.west);
+     \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
+     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
+     ([xshift = -5,yshift = -25]e.west) -- node[below = 3pt] {{Recognition model $Q(p|x)$}} ([xshift = 155,yshift = -25]e.west);
+     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
+     ([xshift = -5,yshift = -25]s.west) -- node[below = 3pt] {{ Search}} ([xshift = 5,yshift = -25]f.east);
+     %% \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
+     %% ([xshift = -5,yshift = -25]p.west) -- node[below = 3pt] {{\small Params}} ([xshift = 5,yshift = -25]p.east);
+   \end{tikzpicture}
+   \caption{Neurally-guided program inference pipeline.
+Recognition model outputs distribution over program $Q(p|x)$ parameterized by tensor $Q_{ijk}$.
+Program output by enumerative search incorporated into frontier if likelihood $\probability[x|p] > 0$}
+   \label{neuralPipeline}
+   \end{figure}
+
+\subsubsection{Learning to break symmetries in program space}
+
+A good DSL not only exposes high-level building blocks,
+but also carefully restricts the ways in which those building blocks are allowed to compose.
+For example,
+a DSL for list manipulation should contain both the empty list and
+a routine for appending lists,
+but should not allow
+appending the empty list.
+Similarly a DSL for arithmetic should contain both addition and
+the number zero but
+disallow adding zero.
+These restrictions break symmetries in the space of programs.
+Here we show that the
+combination of the $\mathcal{L}^{\text{MAP}}$ training objective (Section~\ref{recognitionTraining}) and bigram parameterization (Section~\ref{recognitionParameterization})
+interact in a way that spontaneously breaks symmetries in the program space,
+allowing the agent to more efficiently
+explore the space of programs.
+
+Consider an agent tasked with writing programs built from addition, multiplication, and the constants zero and one.
+A bigram parameterization of $Q$ allows
+it to represent the fact that it should never add zero ($Q_{\code{+},\code{0},0} = Q_{\code{+},\code{0},1} = 0$)
+or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
+that addition should always associate to the right
+($Q_{\code{+},\code{+},0} = 0$).
+The $\mathcal{L}^{\text{MAP}}$ training objective encourages
+learning these canonical forms.
+Consider two recognition models, $Q_1$ and $Q_2$,
+and two programs in frontier $\mathcal{F}_x$,
+$p_1=\code{(+ (+ 1 1) 1)}$ and $p_2=\code{(+ 1 (+ 1 1))}$,
+where
+\begin{align*}
+  Q_1(p_1|x) = \frac{\epsilon}{2}&\qquad Q_1(p_2|x) = \frac{\epsilon}{2}\\
+  Q_2(p_1|x) = 0&\qquad  Q_2(p_2|x) = \epsilon
+\end{align*}
+i.e., $Q_2$ breaks a symmetry by forcing right associative addition,
+but $Q_1$ does not, instead splitting its probability mass equally between $p_1$ and $p_2$.
+Now because $\probability[p_1|\mathcal{D},\theta] = \probability[p_2|\mathcal{D},\theta]$
+(Algorithm~\ref{sampleProgram}),
+we have
+\begin{align*}
+  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_1)& = \frac{\probability[p_1|\mathcal{D},\theta]\log \frac{\epsilon}{2} + \probability[p_2|\mathcal{D},\theta]\log \frac{\epsilon}{2}}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = \log \frac{\epsilon}{2}\\
+  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_2)& = \frac{\probability[p_1|\mathcal{D},\theta]\log 0 + \probability[p_2|\mathcal{D},\theta]\log \epsilon}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = +\infty\\
+  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_1)& = \log Q_1(p_1)           = \log Q_1(p_2)   = \log \frac{\epsilon}{2}\\
+  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_2)& = \log Q_2(p_2) \phantom{ = \log Q_1(p_2) } = \log \epsilon\\
+\end{align*}
+So the $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking recognition model),
+while $\mathcal{L}^{\text{posterior}}$ reverses this preference.
+We can confirm this experimentally by training
+four recognition models:
+a
 
@@
 We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
-\begin{align}
-\nonumber      \log \probability[\mathcal{D}]& + 
-\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
+\begin{equation}
+        \log \probability[\mathcal{D}] + 
+\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
 \label{AIC}
-\end{align}
+  \end{equation}
 At a high level, our approach is to search locally through the space 
@@
 any refactoring of the programs found during waking, maximizing:
-\begin{align}
-\nonumber      \log \probability[\mathcal{D}]& + 
-\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
+\begin{equation}
+        \log \probability[\mathcal{D}] + 
+\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
 \label{factorObjective}
-\end{align}
+  \end{equation}
 where $p'\manyReduce p$
@@
 this approximation gives the following objective:
-\begin{align}
-\nonumber      \log \probability[\mathcal{D}]& + 
-\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\+&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
+\begin{equation}
+        \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
 \label{limitedObjective}
-\end{align}
+  \end{equation}
 In practice, setting the number of refactoring steps $n$ to 3 suffices
@@
 refactoring.
-This dynamic program, which we call $I\beta^n(p)$,
-satisfies $\denotation{I\beta^n(p)} =  R_n(p)$.
+This dynamic program, which we call $I\beta_n(p)$,
+satisfies $\denotation{I\beta_n(p)} =  R_n(p)$.
 In other words,
@@
 prior work.
+
+\subsection{Programs that make plans and take actions}
+
+We apply \system to two domains where the agent plans a series of actions
+in order to draw a picture (Sec.~\ref{logoSection}) or build a tower out of blocks (Sec.~\ref{towerSection}).
+For drawing pictures,
+the agent controls a `pen' that it can pick up or place down while moving across a canvas.
+For building towers,
+the agent controls a `hand' that it moves through a simulated world while
+placing down differently sized blocks.
+For each of these domains,
+the agent observes a target picture or block tower,
+and must either draw the picture or build the tower.
+The recognition model is a CNN that
+observes an image of the target picture or tower.
+The system is initially provided with two control flow operators:
+\code{loop} (a `for' loop) and \code{get/set},
+which saves and then restores the current state of the pen or hand.
+
+\subsubsection{Programs that draw pictures}\label{logoSection}
+
+We take inspiration from LOGO Turtle graphics~\cite{turtle},
+and task our agent with drawing 90 different pictures (Figure~\ref{everyLogo}).
+The agent has primitives for moving the pen forward and rotating it (\code{move}),
+picking the pen up and then putting it down (\code{pen-up}),
+arithmetic operations on angles and distances,
+and constants ($2\pi$, 0 through 9, and $\infty$\footnote{To ensure that programs terminate, we set $\infty = 20$}).
+
+Over the course of 5 wake/sleep iterations,
+the system authors a DSL
+containing parametric and, sometimes, higher-order drawing routines,
+allowing it to write programs for objects like
+polygons, circles, spirals, and snowflakes.
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/logo90.png}
+  \caption{LOGO graphics tasks. The agent must learn to write programs that draw these 90 pictures.}\label{everyLogo}
+\end{figure}
+\begin{figure}
+\centering  \begin{tabular}{c|c}
+      \includegraphics[width = 0.5\textwidth]{figures/initialDreams/montage.png}&
+  \includegraphics[width = 0.5\textwidth]{figures/finalDreams/montage.png}
+    \end{tabular}
+  \caption{Dreams, or samples, from the DSL before (left) and after (right) training on tasks in Figure~\ref{everyLogo}.
+Blue: where the agent started drawing.
+Pink: where the agent ended drawing.}\label{logoDreams}
+\end{figure}
+
+\subsubsection{Building towers out of `Lego' blocks}\label{towerSection}
 
@@
 \end{align*}
-where $\shift^k$ is the shifting operator~\cite{pierce},
+where $\shift{k}$ is the shifting operator~\cite{pierce},
 which adds $k$ to all of the free variables in a $\lambda$-expression or version space.
@@
     \text{if $\mathcal{D}'\in \text{proposals}$ and $e\in \mathcal{D}'$ and  $e\in v$: }e\\
-    \text{otherwise if $v$ is a primitive or index:} $v$
-    \text{otherwise if $v = \lambda b$: } $\lambda $\text{beam}_b(\mathcal{D}')\\
-    \text{otherwise if $v = (f\;x)$: } $(\text{beam}_f(\mathcal{D}')\;\text{beam}_x\mathcal{D}'))\\
+    \text{otherwise if $v$ is a primitive or index:} v
+    \text{otherwise if $v = \lambda b$: } \lambda \text{beam}_b(\mathcal{D}')\\
+    \text{otherwise if $v = (f\;x)$: } (\text{beam}_f(\mathcal{D}')\;\text{beam}_x\mathcal{D}'))\\
     \text{otherwise if $v = \uplus V$: } \argmin_{e\in \left\{b'_{v'}(\mathcal{D}')\;:\;v'\in V \right\}}\text{size}(e|\mathcal{D}')
@@
 \begin{align}
-\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{Q_x}\left[\log \probability\left[p|\theta \right] \right]\\
-  Q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{F}_x \right]
+\label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{q_x}\left[\log \probability\left[p|\theta \right] \right]\\
+  q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{F}_x \right]
 \end{align}
@@
 \begin{align*}
-  &\sum_x\expect_{Q_x}\left[  \log \probability[p|\theta] \right] =\\
-  &\sum_{e\in \mathcal{D}} \log \theta_e \sum_x\expect_{Q_x}\left[c(e,p_x) \right] -
-  \sum_\tau\expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
- =   &\sum_e C(e)\log \theta_e  - \beta\sum_\tau\frac{\expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]}{\beta}\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
- \geq    &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{\expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+  &\sum_x\expect_{q_x}\left[  \log \probability[p|\theta] \right] =\\
+  &\sum_{e\in \mathcal{D}} \log \theta_e \sum_x\expect_{q_x}\left[c(e,p_x) \right] -
+  \sum_\tau\expect_{q_x}\left[\sum_x c(\tau,p_x)  \right]\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+ =   &\sum_e C(e)\log \theta_e  - \beta\sum_\tau\frac{\expect_{q_x}\left[\sum_x c(\tau,p_x)  \right]}{\beta}\log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
+ \geq    &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{\expect_{q_x}\left[\sum_x c(\tau,p_x)  \right]}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  \\
      =     &\sum_e C(e)\log \theta_e  - \beta\log \sum_\tau\frac{R(\tau)}{\beta}\sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e  
@@
 \begin{align*}
-  C(e)&\triangleq  \sum_x\expect_{Q_x}\left[c(e,p_x) \right]\\
-  R(\tau)&\triangleq \expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]\\
-  \beta&\triangleq\sum_\tau \expect_{Q_x}\left[\sum_x c(\tau,p_x)  \right]
+  C(e)&\triangleq  \sum_x\expect_{q_x}\left[c(e,p_x) \right]\\
+  R(\tau)&\triangleq \expect_{q_x}\left[\sum_x c(\tau,p_x)  \right]\\
+  \beta&\triangleq\sum_\tau \expect_{q_x}\left[\sum_x c(\tau,p_x)  \right]
 \end{align*}
@@
 
+\begin{figure}
+  \begin{tikzpicture}
+    \node at (0,0) (d){DSL};
+    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+
+    \node at ([xshift = 2cm]t) (nn){
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
+    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
+
+    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+    \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
+    \draw [->  ] (s.east) -- (f.west);
+
+    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
+
+    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
+
+  \end{tikzpicture}
+
+  \vspace{1cm}
+
+  \begin{tikzpicture}
+    \node at (0,0) (f1){Frontier$_1$};
+    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
+
+    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
+    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
+    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
+
+    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
+    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
+    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
+
+    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
+    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
+    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
+
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
+  \end{tikzpicture}
+
+  \begin{tikzpicture}
+    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
+    \node at ([xshift = 3cm]d.east) (p2){program};
+    \node at ([yshift = 1.5cm]p2) (p1){program};
+    \node at ([yshift = -1.5cm]p2) (p3){program};
+
+    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
+    \draw[squiggle,-> ] (d.east) -- (p1.west);
+    \draw[squiggle,-> ] (d.east) -- (p3.west);
+
+    \node at ([xshift = 2cm]p1.east) (t1){task};
+    \node at ([xshift = 2cm]p2.east) (t2){task};
+    \node at ([xshift = 2cm]p3.east) (t3){task};
+    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
+    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
+    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
+
+    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
+  \end{tikzpicture}
+
+  \vspace{2cm}
+
+  \begin{tikzpicture}
+    \node at (0,0) (f1){Frontier$_1$};
+    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+    \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
+    \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
+    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
+
+    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
+    \node[anchor=north] at (c.north) {Compression};
+
+    \draw [-> ] (f1.east) -- (c.west|-f1.east);
+    \draw [-> ] (f2.east) -- (c.west|-f2.east);
+    \draw [-> ] (f3.east) -- (c.west|-f3.east);
+
+    \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
+    \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
+    \draw [-> ] (c.east) -- (d.west);
+    \draw [-> ] (c.east) -- (t.west);
+
+    \node at (c.center) {
+\begin{tikzpicture}[scale=0.7]
+    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
+    %%     \footnotesize Program:\\
+    %%     \code{($\lambda$ (x) (+ (- x) 1))}
+    %% \end{tabular}};
+    %\node at (,0.5) {\code{cons}};
+%    \node [rotate=90] at (-2.3,-0.5) {\small program};
+
+          \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  \node(x1) at (0,-1) {\code{1}};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+  \node(t) at (-0.5,0.5) {};
+  \draw (l1.south) -- (t.south);
+  \node(c) at (-1.5,-0.2) {\code{cons}};
+  \draw (t.south) -- (c.north);
+
+%    \draw  (l1.south) -- (-0.5,0.5);
+
+  %% \node(c) at (-0.5,-1.5) {\code{-}};
+  %% \node(z) at (0.5,-1.5) {\code{x}};
+
+  %% \draw (0,-1) -- (c.north);
+  %% \draw (0,-1) -- (z.north);
+
+  \begin{scope}[shift={(-1,-2.5)}]
+      \node(l1) at (0,0) {};
+  \node[color=pop3](p1) at (-1,-1) {\code{+}};
+  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+  %\node(x1) at (0,-1) {};
+  \draw[color=pop3] (l1.south) -- (p1.north);
+  \draw[color=pop3] (l1.south) -- (n1.north);
+  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+  \node(c) at (-0.5,-1.5) {\code{car}};
+  \node(z) at (0.5,-1.5) {\code{z}};
+
+  \draw (0,-1) -- (c.north);
+  \draw (0,-1) -- (z.north);
+
+%  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+  \end{scope}
+
+\begin{scope}[shift={(0,-5)}]
+  \node[pop3](p1) at (-1,-1) {\code{+}};
+  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
+  \node[pop3](a) at (0,-1) {\code{ }};
+  %\node(x1) at (0,-1) {};
+  \draw[pop3] (0,0) -- (p1.north);
+  \draw[pop3] (0,0) -- (n1.north);
+  \draw[pop3] (-0.55,-0.4) -- (a.north);
+%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
+
+  \end{scope}
+
+\end{tikzpicture}
+    };
+
+    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};
+
+    \end{tikzpicture}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+%  \includegraphics{figures/iterations.eps}
+\begin{tikzpicture}
+  \begin{scope}[shift={(2.5,0)}]
+    \node[obs] at (3,3) (dx){$\mathcal{D}$};
+    \node[latent] at (3.5,1.75) (zp){$p$};
+%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
+    \node[obs] at (3.5,0.7) (xp) {$x$};
+    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
+    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
+%    \draw [->,red] (tx) -- (zp);
+    \draw [->] (dx) -- (zp);
+    \node at (nn) {
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
+      \end{tikzpicture}
+    };
+    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
+  \end{scope}
+
+  \begin{scope}[shift={(0.7,-4.2)}]
+    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
+    \draw [->] (p2.south) -- (x2.north);
+    \draw [->] (dt.south) -- (p2.north);
+
+    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
+    \draw [->] (p1.south) -- (x1.north);
+    \draw [->] (dt.south) -- (p1.north);
+
+    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
+    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
+    \draw [->] (p3.south) -- (x3.north);
+    \draw [->] (dt.south) -- (p3.north);
+
+    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
+    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
+
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
+    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
+    \draw [->] (zp.south) -- (xp.north);
+    \plate {}{(zp)(xp)}{$x\in X$};
+
+    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
+    rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
+    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
+  \end{scope}
+
+  \begin{scope}[shift={(7.7,-4)}]
+    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
+    \node[latent] at (1.5,3) (t){$\theta$};
+    \node[obs] at (1,1.75) (z){$p$};
+%     \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
+    \node[obs] at (1,0.5) (x) {$x$};
+    \edge {z}{x};
+    \edge {d,t}{z};
+    \plate {}{(z)(x)}{$x\in X$};
+    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
+  \end{scope}
+
+\end{tikzpicture}
+\end{figure}
 \end{document}

### 2018-11-12 archive.tex
symmetry breaking section

@@
 program that will solve it.
+Appendix~\ref{enumerationAppendix}
+explains how we enumerate,
+rather than sample,
+programs generated by Algorithm~\ref{sampleProgram}.
  \begin{algorithm}
@@
 
-Consider an agent tasked with writing programs built from addition, multiplication, and the constants zero and one.
+Consider an agent tasked with writing programs built from addition and the constants zero and one.
 A bigram parameterization of $Q$ allows
 it to represent the fact that it should never add zero ($Q_{\code{+},\code{0},0} = Q_{\code{+},\code{0},1} = 0$)
-or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
-that addition should always associate to the right
+%or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
+or that addition should always associate to the right
 ($Q_{\code{+},\code{+},0} = 0$).
@@
 \end{align*}
-So the $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking recognition model),
-while $\mathcal{L}^{\text{posterior}}$ reverses this preference.
-We can confirm this experimentally by training
-four recognition models:
-a
+So $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking
+recognition model), while $\mathcal{L}^{\text{posterior}}$ reverses
+this preference.  We experimentally confirm this symmetry-breaking
+behavior by training recognition models that minimize either
+$\mathcal{L}^{\text{MAP}}$/$\mathcal{L}^\text{posterior}$ and which
+use either a bigram parameterization/unigram\footnote{In the unigram variant $Q$ predicts a $|\mathcal{D}| + 1$-dimensional vector: $Q(p|x) = \probability[p|\mathcal{D},\theta_i = Q_i(x)]$,
+  and was used in our prior work~\cite{ecc}} parameterization.
+Figure~\ref{symmetry} shows the result of training $Q$ in these four regimes for a DSL containing \code{+}, \code{0}, and \code{1}
+and then sampling programs.
+On this particular run,
+the combination of
+bigrams and $\mathcal{L}^{\text{MAP}}$ learns to
+avoid adding zero and associate addition to the right ---
+different random initializations
+lead to either right or left association.
+
+To be clear,
+our recognition model does not learn to break \emph{every} possible symmetry in every possible DSL.
+But in practice we found that a bigrams combined with $\mathcal{L}^{\text{MAP}}$
+works well,
+and we use with this combination throughout the rest of the paper.
+\begin{figure}
+\centering  \begin{tabular}{ccc}
+     \toprule
+&     Unigram&Bigram \\\midrule
+     $\mathcal{L}^{\text{posterior}}$&
+     \begin{tabular}{l}
+       \emph{Three samples:}\\
+       \code{(+ 1 0)}\\
+\code{(+ (+ 0 0) (+ 1 0))}\\
+\code{(+ 1 1)}
+\\
+63.0\% right-associative; 37.4\% \code{+0}'s
+       \end{tabular}
+     &
+     \begin{tabular}{l}
+       \emph{Three samples:}\\
+       \code{0}\\
+\code{(+ (+ (+ 0 0) (+ 0 1)) 1)}\\
+\code{1}\\
+55.8\% right-associative; 31.9\% \code{+0}'s
+       \end{tabular}
+     \\\\
+     $\mathcal{L}^{\text{MAP}}$&
+     \begin{tabular}{l}
+       \emph{Three samples:}\\
+       \code{1}\\
+       \code{(+ 1 (+ 1 (+ (+ 1 (+ 1 1)) 1)))}\\
+       \code{(+ (+ 1 1) 1)}
+\\48.6\% right-associative;
+       0.5\% \code{+0}'s       
+       \end{tabular}
+     &
+     \begin{tabular}{l}
+       \emph{Three Samples:}\\
+\code{(+ 1 (+ 1 (+ 1 (+ 1 (+ 1 1)))))}\\
+\code{0}\\
+\code{(+ 1 (+ 1 (+ 1 1)))}
+       \\
+       \textbf{97.9\% right-associative};
+       2.5\% \code{+0}'s       
+       \end{tabular}
+\\    \bottomrule 
+  \end{tabular}
+  \caption{Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero.
+\% right-associative calculated by drawing 500 samples from $Q$.
+$\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition.
+Tasked with building programs from \code{+}, \code{1}, and \code{0}.
+}\label{symmetry}
+  \end{figure}
 
@@
     };
-    \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-3cm]p1.south) {
+    \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-2.5cm]p1.south) {
       \begin{tabular}{l}
@@
     };
-    \node(r2)[draw] at ([yshift=-3cm]p2.south) {
+    \node(r2)[draw] at ([yshift=-2.5cm]p2.south) {
 \begin{tabular}{l}
@@
 
-    \node[draw](m) at (3,-7) {
+    \node[draw](m) at (3,-6.5) {
 \begin{tabular}{l}
@@
 
+\subsection{Enumerative program search}\label{enumerationAppendix}
+
+Our current implementation of \system takes the simple and generic strategy of enumerating programs in
+descending order of their probability under either $(\mathcal{D},\theta)$ or $Q(p|x)$.
+Algorithm~\ref{sampleProgram} and~\ref{recognitionSample}
+specify procedures for sampling
+from these distributions,
+but not for enumerating from them.
+We combine two different enumeration strategies,
+which allowed us to build a massively parallel program enumerator:
+\begin{itemize}
+\item \textbf{Best-first search:} Best-first search maintains a heap of
+  partial programs ordered by their
+  probability --- here a partial program means a program whose syntax tree
+  may contain unspecified `holes'. Best-first search
+  is guaranteed to enumerate programs in decreasing order of their probability,
+  and has memory requirements that in general grow exponentially as a function of the description length of programs in the heap (thus linearly as a function of run time).
+\item \textbf{Depth-first search:} Depth first search
+  recursively explores the space of
+  execution traces through Algorithm~\ref{sampleProgram} and~\ref{recognitionSample},
+  equivalently maintaining a stack of partial programs.  
+  In general it does not enumerate programs
+  in decreasing order of probability,
+  but has memory requirements that grow linearly as a function of the description length of the programs in the stack (thus logarithmically as a function of run time).
+\end{itemize}
+
+Our parallel enumeration algorithm (Algorithm~\ref{enumerationAlgorithm})
+first performs a best-first search
+until the best-first heap
+is much larger than the number of CPUs.
+At this point,
+it switches to performing many depth-first searches in parallel,
+initializing a depth first search
+with one of the
+entries in the best-first heap.
+Because depth-first search does not
+produce programs in decreasing order of their probability,
+we wrap this entire procedure up into an outer loop
+that first enumerates programs whose description length is between
+$0$ to $\Delta$,
+then programs with description length between $\Delta$ and $2\Delta$,
+then $2\Delta$ to $3\Delta$, etc.,
+until a timeout is reached.
+This is similar in spirit to iterative deepening depth first search~\cite{Russell:2003:AIM:773294}.
+ \begin{algorithm}
+   \caption{Parallel enumerative program search algorithm }
+   \label{enumerationAlgorithm}
+   \begin{algorithmic}[1]
+     \State \textbf{function} enumerate$(\mu, T, \text{CPUs})$:
+     \State {\bfseries Input:} Distribution over programs $\mu$, timeout $T$, CPU count
+     \State \textbf{Output:} stream of programs in approximately descending order of probability under $\mu$
+     \State \textbf{Hyperparameter:} nat increase rate $\Delta$\Comment{We set $\Delta = 1.5$}
+     \State lowerBound$\gets 0$
+     \While{total elapsed time $ < T$}
+     \State heap$\gets$newMaxHeap()\Comment{Heap for best-first search}
+     \State heap.insert$(\text{priority} = 0,\text{value} = \text{empty syntax tree})$\Comment{Initialize heap with start state of search space}
+     \While{$0 < |\text{heap}|\leq 10\times \text{CPUs}$}\Comment{Each CPU will get approximately 10 jobs (a partial program)}
+     \State priority, partialProgram $\gets$ heap.popMaximum()
+     \If{partialProgram is finished}\Comment{Nothing more to fill in in the syntax tree}
+     \If{$\text{lowerBound}\leq -\text{priority} < \text{lowerBound} + \Delta$}
+     \State \textbf{yield }partialProgram
+     \EndIf
+     \Else
+     \For{child$\in $children(partialProgram)}\Comment{children$(\cdot )$ fills in next random choice in syntax tree.}
+     \If{$-\log \mu(\text{child}) < \text{lowerBound} + \Delta$}\Comment{Child's description length small enough}
+     \State heap.insert$(\text{priority} = \log \mu(\text{child}),\text{value} = \text{child})$
+     \EndIf
+     \EndFor
+     \EndIf
+     \EndWhile
+     \State \textbf{yield from }ParallelMap$_\text{CPUs}(\text{depthFirst}(\mu,T - \text{elapsed time}, \text{lowerBound}, \cdot ), \text{heap.values()})$%\Comment{Launch parallel workers}
+     \State $\text{lowerBound}\gets\text{lowerBound} + \Delta$\Comment{Push up lower bound on MDL by $\Delta$}
+     \EndWhile
+     \Statex
+     \State \textbf{function} depthFirst$(\mu,T,\text{lowerBound},\text{partialProgram})$: \Comment{Each worker does a depth first search.
+Enumerates completions of partialProgram whose MDL is between lowerBound and $\text{lowerBound} + \Delta$}
+     \State stack$\gets$[partialProgram]
+     \While{$\text{total elapsed time} < T$ and $\text{stack}$ is not empty}
+     \State partialProgram$\gets$stack.pop()
+     \If{partialProgram is finished}
+     \If{$\text{lowerBound}\leq-\log \mu(\text{partialProgram}) < \text{lowerBound} + \Delta$}
+     \State \textbf{yield }partialProgram
+     \EndIf
+     \Else
+     \For{$\text{child}\in \text{children}(\text{partialProgram})$}
+     \If{$-\log \mu(\text{child}) < \text{lowerBound} + \Delta$}\Comment{Child's description length small enough}
+     \State stack.push$(\text{child})$
+     \EndIf
+     \EndFor
+     \EndIf
+
+     \EndWhile
+   \end{algorithmic}
+ \end{algorithm}
+
 \subsection{Refactoring code with version spaces}\label{appendixVersion}

### 2018-11-28 archive.tex
soundness completeness theorems

@@
 \newcommand{\shift}[1]{\uparrow^{#1}}
+\newcommand{\downshift}[1]{\downarrow^{#1}}
 \newcommand{\substitute}[2]{[\$ #1 \mapsto #2]}
@@
 \newtheorem{definition}{Definition}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 polygons, circles, spirals, and snowflakes.
+Figure~\ref{logoPrimitives}
+illustrates several learned DSL subroutines.
+Once it has synthesized these new subroutines,
+the agent generates a rich and highly varied
+data set of dreams on which to train the recognition model.
+Figure~\ref{logoDreams}
+illustrates  dreams before and after learning ---
+one sees that
+the system has learned to recombine
+the building blocks and motifs in the training data.
 \begin{figure}
@@
 \begin{figure}
+\centering  \begin{tabular}{cc}
+    \toprule
+    Parametric drawing routines&Higher-order drawing routine\\\midrule
+    \begin{tabular}{rl}
+      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
+      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
+      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
+      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
+        Polygons \& Stars:
+&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
+    \end{tabular}
+    &
+    \begin{tabular}{c}
+      Repeat and rotate subprogram:\\
+          \includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_18.png}
+      \end{tabular}
+
+    \\\bottomrule 
+  \end{tabular}
+  \caption{Example primitives learned by \system when trained on tasks in Figure~\ref{everyLogo}.
+Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right).
+Each row of images on the left is the same code executed with different parameters.
+Each image on the right is the same code executed with different parameters and with a different subprogram provided as input.}\label{logoPrimitives}
+  \end{figure}
+\begin{figure}
 \centering  \begin{tabular}{c|c}
@@
 \subsubsection{Building towers out of `Lego' blocks}\label{towerSection}
+
+Inspired by the classic AI `copy task' --- where an agent must look at
+an image of a tower made of toy blocks and re-create the
+tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
+(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
+tower and the locations of each of its blocks, and must write a
+program that plans how a simulated hand would build the tower.
+These towers are built from Lego-style blocks
+that snap together on a discrete grid (e.g., we do not model gravity
+and discretize the coordinates of the blocks).
+
+The system starts out with four domain-specific primitives:
+\code{left} and \code{right}, which take an integer $n$ as input and
+move the hand left/right $n$ Lego grid spaces; along with
+\code{horizontal} and \code{vertical}, which put down
+horizontal/vertical Lego blocks.  As with LOGO graphics we also
+include our two standard planning primitives (\code{for} and
+\code{get/set}).
+Starting with this basis
+the system can only solve
+12.5\% of the copy tasks.
+With successive wake/sleep cycles
+we assemble a DSL containing
+parametric option-like subroutines (Figure~\ref{towerPrimitives}),
+allowing the agent to solve 90\% of the tasks.
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/every_tower.png}
+  \caption{}\label{towerTasks}
+\end{figure}
+\begin{figure}
+  \caption{Learned subroutines for building towers out of Lego-style blocks when trained on tasks in Figure~\ref{towerTasks}. These subroutines act like parametric options~\cite{options},
+  giving higher-level building blocks that the agent can use to plan.}
+  \end{figure}
 
@@
 infer a regex from a collection of strings.
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{figures/geomCompiled.eps} 
-  \caption{
-  Top left: Example training tasks. Top right: samples from the learned DSL.
- Bottom: \% holdout testing tasks solved (middle); on the sides are averaged samples from the DSL
-  before any training (left) and after last iteration (right).}\label{geomCompiled}
-
-  \end{figure}
-
-\begin{figure}
-\centering
-\tiny
-\tabcolsep=2pt
-\vspace{0pt}
-\begin{tabular}{cc}
-\begin{tabular}{ccc}
-\multicolumn{3}{c}{Tasks:}\\
-\hline
-1.14531&F&110.9 \\ 
-?&CL&163.2 \\
-1.29857&F&207.3\\
-?&PCFL&143.3 \\
-\multicolumn{3}{c}{Learned generative models:}\\
-\hline
- \scode{\textbackslash?\textbar(1\textbackslash .\textbackslash d+)}& \scode{((\textbackslash u\textbackslash u)*)\textbar F}& \scode{\textbackslash d\textbackslash d\textbackslash d\textbackslash.\textbackslash d} \\
- \multicolumn{3}{c}{Samples from synthesized generative models:}\\
-\hline
-1.61&DQDF& 343.8 \\ 
-?&F&241.2\\
-?&F&647.5\\
-1.2&KI&246.8\\
-1.5080987&F&728.6\\
-?&GL&029.3\\
-1.453&F&289.1\\
-    \hline
-        \\
-\end{tabular} 
-&
-\begin{tabular}{c}
-\begin{tabular}{l}
-\multicolumn{1}{c}{Learned DSL:}\\
-\hline
-$f_1() = $ \scode{\textbackslash u\textbackslash w*}\\
-$f_2(x) = $ \scode{($x$|$f_1$)*} $=$ \scode{($x$|\textbackslash u\textbackslash w*)*}\\
-$f_3(x) =  f_2(\scode{space}) = $\scode{( \textbar\textbackslash u\textbackslash w*)*}\\
-\hline
-$f_4(x) = $ \scode{($x$*$x$)}\\
-\emph{ (equivalent to regex `plus')}\\
-$f_5() = f_4(\scode{\textbackslash l}) = \scode{\textbackslash l*\textbackslash l}$\\
-    \hline
-\end{tabular}
-\\
-\\
-\includegraphics[width=.4\linewidth]{figures/nampiregexjoshcurve3.eps} 
-\end{tabular}
-\end{tabular}
-\vspace{-.3cm}
-\caption{regex stuff}
-\end{figure}
 
@@
   \\
-  S_k(v)& = \left\{\shift{k}v\mapsto \$k \right\}\cup
+  S_k(v)& = \left\{\downshift{k}_0v\mapsto \$k \right\}\cup
   \begin{cases}
@@
     \text{if $v$ is $\Lambda$:}&\left\{\Lambda\mapsto\Lambda \right\}
-  \end{cases}
+  \end{cases}\\
+  \downshift{k}_c \$i& = \$i\text{, when $i < c$}\\
+  \downshift{k}_c \$i& = \$(i - k)\text{, when $i\geq c + k$}\\
+  \downshift{k}_c \$i& = \varnothing \text{, when $c\leq i <  c + k$}\\
+  \downshift{k}_c \lambda b &= \lambda\downshift{k}_{c + 1}b\\
+  \downshift{k}_c (f\;x)& = (\downshift{k}_cf\;\downshift{k}_cx)\\
+  \downshift{k}_c \uplus V& = \uplus \left\{\downshift{k}_c v \;|\;v\in V \right\}\\
+  \downshift{k}_c v& = v\text{, when }v\text{ is a primitive or }\varnothing \text{ or }\Lambda
 \end{align*}
@@
 Refactoring is similarly done more quickly with dynamic programming (see Equation~\ref{FACTORING} for the recursive definition of \textsc{refactor}).
+
+We now formally proved that $I\beta$ exhaustively enumerates the space of possible refactorings.
+Our approach is to first prove that $S_k$ exhaustively enumerates
+the space of possible substitutions that
+could give rise to a program.
+The following pair of technical lemmas are useful; both are easily proven by structural induction.
+\begin{lemma}
+  Let $e$ be a program or version space and $n$, $c$ be natural numbers. \\Then $\shift{-1}_{n + c}\shift{n + 1}_ce = \shift{n}_c e$,
+  and in particular $\shift{-1}_{n}\shift{n + 1}e = \shift{n} e$.
+\label{neutralizeShift}\end{lemma}
+\begin{lemma}
+  Let $e$ be a program or version space and $n$, $c$ be natural numbers. \\Then $\downshift{n}_{c}\shift{n}_ce =  e$,
+  and in particular $\downshift{n}\shift{n}e =  e$.
+  \label{neutralizeDown}
+\end{lemma}
+
+\begin{theorem}
+  \textbf{Consistency of $S_n$}. \\If $(v\mapsto b)\in S_n(u)$ then
+  for every $v'\in v$ and $b'\in b$ we have $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b'\in u$.
+\end{theorem}
+\begin{proof}
+  Suppose $b = \$n$ and therefore, by the definition of $S_n$, also $v = \downshift{n}_0 u$.
+  Invoking Lemmas~\ref{neutralizeShift} and~\ref{neutralizeDown}
+  we know that $u = \shift{-1}_n\shift{n + 1}v$
+  and so for every $v'\in v$ we have $\shift{-1}_n\shift{n + 1}v'\in u$.
+  Because $b = \$n = b'$ we can rewrite this to $\shift{-1}_n[\$n\mapsto \shift{n + 1}v']b'\in u$.
+
+  Otherwise assume $b\not= \$n$ and proceed by structural induction on $u$:
+
+  \begin{itemize}
+  \item If $u = \$i < n$ then we have to consider the case that $v = \Lambda$ and $b = u = \$i = b'$.
+    Pick $v'\in \Lambda$ arbitrarily. Then $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b' = \shift{-1}_n\$i = \$i\in u$.
+  \item If $u = \$i\geq n$ then we have consider the case that $v = \Lambda$ and $b = \$(i + 1) = b'$.
+    Pick $v'\in \Lambda$ arbitrarily. Then $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b' = \shift{-1}_n\$(i + 1) = \$i\in u$.
+  \item If $u$ is primitive then we have to consider the case that $v = \Lambda$ and $b = u = b'$.
+    Pick $v'\in \Lambda$ arbitrarily. Then $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b' = \shift{-1}_nu = u\in u$.
+  \item If $u$ is of the form $\lambda a$,
+    then $S_n(u)\subset\left\{v\mapsto \lambda b\;|\; (v\mapsto b)\in S_{n + 1}(a) \right\}$.
+    Let $v\mapsto \lambda b\in S_n(u)$.
+    By induction for every $v'\in v$ and $b'\in b$ we have $\shift{-1}_{n + 1}[\$n\mapsto \shift{2 + n}v']b'\in a$,
+    which we can rewrite to $\shift{-1}_{n}[\$n\mapsto \shift{1 + n}v']\lambda b'\in \lambda a = u$.
+  \item If $u$ is of the form $(f\;x)$ then
+    then $S_n(u)\subset\left\{v_f\cap v_x\mapsto (b_f\;b_x)\;|\; (v_f\mapsto b_f)\in S_{n}(f),\;(v_x\mapsto b_x)\in S_{n}(x) \right\}$.
+    Pick $v'\in v_f\cap v_x$ arbitrarily.
+    By induction for every $v_f'\in v_f$, $v_x'\in v_x$, $b_f'\in b_f$, $b_x'\in b_x$
+    we have $\shift{-1}_{n}[\$n\mapsto \shift{1 + n}v_f'] b_f' \in f$ and $\shift{-1}_{n}[\$n\mapsto \shift{1 + n}v_x'] b_x' \in x$.
+    Combining these facts gives
+    $\shift{-1}_{n}[\$n\mapsto \shift{1 + n}v'] (b_f'\;b_x') \in (f\;x) = u$.
+  \item If $u$ is of the form $\uplus U$ then
+    pick $(v\mapsto b)\in S_n(u)$ arbitrarily.
+    By the definition of $S_n$ there is a $z$ such that $(v\mapsto b)\in S_n(z)$,
+    and the theorem holds immediately by induction.
+  \item If $u$  is $\varnothing $ or $\Lambda$ then the theorem holds vacuously.
+    \end{itemize}
+\end{proof}
+\begin{theorem}
+  \textbf{Completeness of $S_n$}.\\
+  If there exists programs $v'$  and $b'$, and a version space $u$, such that $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b'\in u$,
+  then there also exists $(v\mapsto b)\in S_n(u)$ such that $v'\in v$ and $b'\in b$.
+\end{theorem}
+\begin{proof}
+  As before we first consider the case that $b' = \$n$.
+  If so then
+  $\shift{-1}_n\shift{1 + n}v'\in u$ or (invoking Lemma~\ref{neutralizeShift}) that $\shift{n}v'\in u$
+  and (invoking Lemma~\ref{neutralizeDown}) that $v'\in \downshift{n}u$.
+  From the definition of $S_n$ we know that $(\downshift{n}u\mapsto \$n)\in S_n(u)$ which is what was to be shown.
+
+  Otherwise assume that $b'\not=\$n$. Proceeding by structural induction on $u$:
+  \begin{itemize}
+  \item If $u = \$i$ then,
+    because $b'$ is not $\$n$,
+    we have $\shift{-1}_nb' = \$i$.
+    Let $b' = \$j$,
+    and so
+    $$
+    i =
+    \begin{cases}
+      j&\text{ if }j < n\\
+      j - 1&\text{ if }j > n
+    \end{cases}
+    $$
+    where $j = n$ is impossible because by assumption $b'\not=\$n$.
+
+    If $j < n$ then
+    $i = j$ and so $u = b'$.
+    By the definition of $S_n$ we have
+    $(\Lambda\mapsto \$i)\in S_n(u)$,
+    completing this inductive step because $v'\in \Lambda$ and $b'\in \$i$.
+    Otherwise assume $j > n$
+    and so $\$i = \$(j - 1) = u$.
+    By the definition of $S_n$ we have
+    $(\Lambda\mapsto \$(i + 1))\in S_n(u)$,
+    completing this inductive step because $v'\in \Lambda$
+    and $b' = \$j = \$(i + 1)$.
+  \item If $u$ is a primitive then,
+    because $b'$ is not $\$n$,
+    we have $\shift{-1}_nb' = u$,
+    and so $b' = u$.
+    By the definition of $S_n$ we have $(\Lambda\mapsto u)\in S_n(u)$
+    completing this inductive step because $v'\in \Lambda$ and $b' = u$.
+  \item If $u$ is of the form $\lambda a$ then,
+    because of the assumption that $b'\not=\$n$,
+    we know that $b'$ is of the form $\lambda c'$ 
+    and that $\lambda \shift{-1}_{n + 1}[\$(n + 1)\mapsto \shift{2 + n}v']c'\in \lambda a$.
+    By induction this means that there is a $(v\mapsto c)\in S_{n + 1}(a)$
+    satisfying $v'\in v$ and $c'\in c$.
+    By the definition of $S_n$ we also know that
+    $(v\mapsto  \lambda c)\in S_n(u)$,
+    completing this inductive step because $b' = \lambda c'\in \lambda c$.
+  \item If $u$ is of the form $(f\;x)$
+    then,
+    because of the assumption that $b'\not=\$n$,
+    we know that $b'$ is of the form $(b_f'\;b_x')$
+    and that both
+    $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b_f'\in f$ and $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b_x'\in x$.
+    Invoking the inductive hypothesis twice
+    gives a
+    $(v_f\mapsto b_f)\in S_n(f)$ satisfying $v'\in v_f$, $b_f'\in b_f$
+    and a
+    $(v_x\mapsto b_x)\in S_n(x)$ satisfying $v'\in v_x$, $b_x'\in b_x$.
+    By the definition of $S_n$
+    we know that $(v_f\cap v_x\mapsto b_f\;b_x)\in S_n(u)$
+    completing the inductive step because $v'$ is guaranteed to be
+    in both $v_f$ and $v_x$ and we know that
+    $b' = (b_f'\;b_x')\in (b_f\;b_x)$.
+
+  \item If $u$ is of the form $\uplus U$
+    then there must be a $z\in U$ such that $\shift{-1}_n[\$n\mapsto \shift{1 + n}v']b'\in z$.
+    By induction there is a $(v\mapsto b)\in S_n(z)$ such that $v'\in v$ and $b'\in v$.
+    By the definition of $S_n$ we know that $(v\mapsto b)$ is also in $S_n(u)$ completing the inductive step.
+  \item If $u$  is $\varnothing $ or $\Lambda$ then the theorem holds vacuously.    
+    \end{itemize}
+\end{proof}
 
@@
 
-\begin{figure}[h]
-\centering
-%  \includegraphics{figures/iterations.eps}
-\begin{tikzpicture}
-  \begin{scope}[shift={(2.5,0)}]
-    \node[obs] at (3,3) (dx){$\mathcal{D}$};
-    \node[latent] at (3.5,1.75) (zp){$p$};
-%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
-    \node[obs] at (3.5,0.7) (xp) {$x$};
-    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
-    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
-%    \draw [->,red] (tx) -- (zp);
-    \draw [->] (dx) -- (zp);
-    \node at (nn) {
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
-  \end{scope}
-
-  \begin{scope}[shift={(0.7,-4.2)}]
-    \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
-    \draw [->] (p2.south) -- (x2.north);
-    \draw [->] (dt.south) -- (p2.north);
-
-    \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
-    \draw [->] (p1.south) -- (x1.north);
-    \draw [->] (dt.south) -- (p1.north);
-
-    \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
-    \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
-    \draw [->] (p3.south) -- (x3.north);
-    \draw [->] (dt.south) -- (p3.north);
-
-    \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
-    \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};
-
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
-    \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
-    \draw [->] (zp.south) -- (xp.north);
-    \plate {}{(zp)(xp)}{$x\in X$};
-
-    \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -2]p1.west)
-    rectangle  ([yshift = -0.45cm,xshift = +2]xp.east);
-    \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {Sleep-R: Train $q$\\training data: cyan $(x,p)$};
-  \end{scope}
-
-  \begin{scope}[shift={(7.7,-4)}]
-    \node[latent] at (0.5,3) (d){$\mathcal{D}$};
-    \node[latent] at (1.5,3) (t){$\theta$};
-    \node[obs] at (1,1.75) (z){$p$};
-%     \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
-    \node[obs] at (1,0.5) (x) {$x$};
-    \edge {z}{x};
-    \edge {d,t}{z};
-    \plate {}{(z)(x)}{$x\in X$};
-    \node[align = center] at ([yshift = -1cm]x.south) {Sleep-G: Induce $(\mathcal{D},\theta)$};
-  \end{scope}
-
-\end{tikzpicture}
-\end{figure}
+\begin{table*}%[t!]
+  \makebox[\textwidth][c]{
+    \scriptsize
+  \tabcolsep=4pt
+  \renewcommand\code\texttt
+  \renewcommand\codechar[1]{\texttt{"#1"}}
+  \newcommand{\helpSize}{0.25cm}
+  \begin{tabular}{cccc}
+    \toprule
+    &{\normalsize Symbolic regression}&{\normalsize Laws of motion}&\\\midrule
+    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
+    \begin{tabular}{cc}
+      \includegraphics[width = 3em]{figures/functions/4.png}&
+      \includegraphics[width = 3em]{figures/functions/146}\\
+      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
+      ~\\
+      \includegraphics[width = 3em]{figures/functions/112.png}&
+        \includegraphics[width = 3em]{figures/functions/92.png}
+      \\
+      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\
+
+    \end{tabular}
+    }
+    &
+    \begin{tabular}{cc}
+      \includegraphics[width = 15em]{figures/massOnSpring.png} &
+            \includegraphics[width = 15em]{figures/planets.png}
+      \\
+      \begin{tabular}{l}
+              \blueCode{$f($o,$\Delta) = $($f_3$ o $\Delta$}\\
+      \hspace{1cm}\blueCode{(- (* $k$ (pos o))}\\
+      \hspace{1cm}\phantom{\code{(- }}\blueCode{(* $-9.8$ $\hat{x}$)))}\\
+      \end{tabular}&
+      \begin{tabular}{l}
+              \blueCode{$f($a,b,$\Delta) = $($f_3$ a $\Delta$}\\
+      \hspace{1cm}\blueCode{(/ (* $G$ (mass a) (mass b))}\\
+      \hspace{1cm}\phantom{\code{(- }}\blueCode{(square (- (pos a) (pos b)))))}\\
+        \end{tabular}
+      \end{tabular}
+    &
+
+    ~\\
+    \midrule
+    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+      \begin{tabular}{l}
+    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
+    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
+    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
+    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
+    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
+    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
+    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
+    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
+    \hspace{\helpSize}\emph{($f_6$: rational function)}\\
+
+  \end{tabular}
+      &
+      \begin{tabular}{l}
+        \greenCode{$f_0($\code{o},$\Delta)\, = \,$(set-pos o (+ (pos o) (* $\Delta$ (vel o))))}\\
+        \hspace{\helpSize}\emph{($f_0$: integrates position)}\\
+        \greenCode{$f_1($\code{o},\code{a},$\Delta)\, = \,$(set-vel o (+ (vel o) (* $\Delta$ a)))}\\
+        \hspace{\helpSize}\emph{($f_1$: integrates velocity)}\\
+        \greenCode{$f_2($\code{o},\code{F}$)\, = \,$(/ F (mass o))}\\
+        \hspace{\helpSize}\emph{($f_2$: Newton's second law)}\\
+        \greenCode{$f_3($\code{o},$\Delta$,\code{F}$)\, = \,$($f_0$ o $\Delta$ ($f_1$ o ($f_2$ o F) $\Delta$))}\\
+        \hspace{\helpSize}\emph{($f_3$: applies Newton's second law and integrates)}
+
+        \end{tabular}
+
+&
+
+  \\\bottomrule\\
+\end{tabular}}\vspace{-0.5cm} 
+\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
+Bottom: Several examples from learned DSL.
+Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}%\vspace{-0.5cm}
+\end{table*}
 \end{document}

### 2018-11-29 spotlight.tex
spotlight

\documentclass[aspectratio=169,final]{beamer}
\setbeamersize{text margin left=5pt,text margin right=5pt}
\setbeamertemplate{navigation symbols}{}

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
%\usetikzlibrary{arrows,shapes}
\usetikzlibrary{fit,bayesnet}

\usepackage{amsmath}
\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}
\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em P}}
\DeclareMathOperator*{\argmax}{arg\,max} % thin space, limits underneath in displays
\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in displays
\newcommand{\code}[1]{{\texttt{#1}}}
\newcommand{\scode}[1]{{\texttt{#1}}}
\newcommand{\codechar}[1]{{\texttt{"#1"}}}

\usepackage{xcolor}
\definecolor{pop1}{HTML}{1F78b4}
\definecolor{pop2}{HTML}{164C13}
\definecolor{pop3}{HTML}{d95F02}
\definecolor{orange}{HTML}{d95F02}
\definecolor{teal}{HTML}{1b9e77}
\newcommand{\pop}[1]{\textcolor{pop1}{#1}}
\newcommand{\popp}[1]{\textcolor{pop2}{#1}}
\newcommand{\tree}[1]{\textcolor{pop3}{#1}}
\newcommand{\orange}[1]{\textcolor{orange}{#1}}
\newcommand{\teal}[1]{\textcolor{teal}{#1}}
\newcommand\Wider[2][3em]{%
\makebox[\linewidth][c]{%
  \begin{minipage}{\dimexpr\textwidth+#1\relax}
  \raggedright#2
  \end{minipage}%
  }%
}

\begin{document}

\begin{frame}{\centering  Library Learning for Neurally-Guided Bayesian Program Induction\\
    \centering\small  Kevin Ellis$^1$, Lucas Morales$^1$, Mathias Sabl\'e-Meyer$^2$,\\ Armando Solar-Lezama$^1$, Joshua B. Tenenbaum$^1$\\
  \small   $^1$: MIT. $^2$: ENS Paris-Saclay.}

  \begin{minipage}[c]{0.3\textwidth}
\centering        \underline{List processing:}\vspace{-0.3cm}\begin{align*}
    \code{[5 2 9]}&\to\code{[9 2 5]}\\
    \code{[1 1 2 2]}&\to\code{[2 2 1 1]}\\
    \code{[1 2 3 2]}&\to\code{[2 3 2 1]}\\
    %% &\\
    %% \code{[9 8 2]}&\to\code{[8 2]}\\
    %% \code{[0 3 0 0]}&\to\code{[3 0 0]}\\
    %% \code{[2 1]}&\to\code{[1]}
    \end{align*}
  \end{minipage} $\quad$
  \begin{minipage}[c]{0.3\textwidth}
\centering            \underline{Text editing:}\vspace{-0.3cm}\begin{align*}
      \text{P Kohli}&\to\text{Dr. Kohli}\\
      \text{Sumit Gulwani}&\to\text{Dr. Gulwani}\\
      \text{Danny Tarlow}&\to\text{Dr. Tarlow}\\
    %%   &\\
    %% \text{+106 769-438}&\to\text{106.769.438}\\
    %% \text{+83 973-831}&\to\text{83.973.831}\\
    %% \text{+882 101-723}&\to\text{882.101.723}    
      \end{align*}    
    \end{minipage} $\quad$
  \begin{minipage}[c]{0.3\textwidth}
\centering            \underline{Symbolic regression:}\\\begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/92.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \begin{tabular}{c}
                  $ax^3 + bx^2 +$\\$ cx + d$
        \end{tabular}
      &    $a/(x - b)$\\
      ~\\
      %% \includegraphics[width = 3em]{figures/functions/112.png}&
      %%   \includegraphics[width = 3em]{figures/functions/92.png}
%%         \\
%%         \begin{tabular}{c}
%% $          ax^4 + bx^3 + $\\$cx^2 + dx + e$
%%           \end{tabular}

%%         &\begin{tabular}{c}
%%            $ax^3 + bx^2 +$\\$ cx + d$
%%            \end{tabular}

    \end{tabular}

  \end{minipage}

  \vspace{1cm}
  Explore/Compress/Compile (EC$^2$) learns to solve programming tasks like these by growing a library of code and training a neural net to search for programs written using the library

  \end{frame}

\begin{frame}{Library Learning}
\centering
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
\begin{tabular}{cc}
    \toprule
Tasks and     \pop{Programs} & DSL\\
    \midrule
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    &
    \begin{tabular}{l}
      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
      \phantom{$f_1(\ell,$}
      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
    \end{tabular}
  \\\bottomrule
  \end{tabular}
  \begin{itemize}
  \item Learned DSL primitives can call each other
    \item Rediscovers higher-order functions like \code{filter}
    \end{itemize}
\end{frame}

\begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}

\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}}

\begin{minipage}[c]{0.4\textwidth}
  \begin{tikzpicture}[scale=1,line width=0.5mm]

  \node[latent,scale=1] at (3.5,3) (dx){DSL};
  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
  \draw [->] (zp1.south) -- (xp1.north);
  \draw [->] (dx.south) -- (zp1.north);
  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);

  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
  \draw [->] (zp1.south) -- (xp1.north);
  \draw [->] (dx.south) -- (zp1.north);
  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);

  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
  \draw [->] (dx.south) -- (zp.north);
  \draw [->] (zp.south) -- (xp.north);

  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
  \end{tikzpicture}

  \vspace{0.5cm}

\visible<4->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
    \node at (0,0) {\underline{\textbf{Compile}: Train {\color{red} recognition model}}};
          \node[align=center] at (-3,-1) (d){DSL};
      %    \node at ([xshift = 3cm]d.east) (p2){program};
      \node at ([xshift=2.5cm,yshift = 0cm]d.east) (p1){program};

      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\footnotesize  sample} (p1.west);

      \node at ([xshift = 2.5cm]p1.east) (t1){ task};

      \draw [-> ] (p1.east) -- node[above]{\footnotesize  run} (t1.west);

      %% \node(f1) at ([yshift=-0.9cm,xshift=1cm]d.south) {\footnotesize progs. for task$_i$};
      %% \node(t) at ([xshift = 2cm]f1.east) (p11){\footnotesize program$_i$, task$_i$};
      %% \draw [->,squiggle ] (f1.east) -- node[above]{\footnotesize  sample} (p11.west);

%      \draw[dashed,cyan,very thick] ([yshift=0.5cm]p1.west) rectangle  ([yshift=-0.5cm,xshift=0.5cm]p11.east);

      \node(n) at ([yshift=-0.8cm,xshift=1.6cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [->,red] ([xshift=-0.4cm]n.west) to[out = -180,in = -70]  ([xshift=-0.25cm,yshift=0.1cm]p1.south);

    \end{tikzpicture}}
\end{minipage}
\begin{minipage}[c]{0.55\textwidth}
  \visible<2->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
    \node at (2,1) {\underline{\textbf{Explore}: Infer programs, fixing DSL and neural net}};
    \node at (0,0) (d){DSL};
    \node at ([yshift = -2cm]d) (t){$\text{Task}$};

    \node at ([xshift = 2cm]t) (nn){
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[align = center, text width = 1cm] at ([yshift = 0.9cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
    \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);

    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};

    \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
    \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

    \node[align=left] at ([xshift=2cm]s.east) (f) {Programs};
    \draw [->  ] (s.east) -- (f.west);

    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
  \end{tikzpicture}}

  \vspace{1cm}

\visible<3->{  \begin{tikzpicture}[scale=0.4,line width=0.5mm]

    \node at (0,0) (f1){progs. for task$_1$};
    \node at ([yshift = -1.4cm]f1.south) (f2){progs. for task$_2$};
    \node at ([yshift = -1.4cm,xshift=1.5cm]f2.south) (f3){DSL};

    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 2.5cm, anchor = north west] at (5.2,1) {};
%    \node[anchor=north] at (c.north) {Compression};
    \node at ([yshift=1cm,xshift=-5cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
    \draw [-> ] (f1.east) -- (c.west|-f1.east);
    \draw [-> ] (f2.east) -- (c.west|-f2.east);
    \draw [-> ] (f3.east) -- (c.west|-f3.east);

    \node[right](d) at ([xshift = 1.6cm,yshift = -0.0cm]c.east) {
      \begin{tikzpicture}[scale=0.6]
        \node[align=center] at (0,0) {DSL+};
\begin{scope}[shift={(0.6,-0.5)}]
  \node[pop3](p1) at (-1,-1) {\code{+}};
  \node[pop3](n1) at (0.6,-0.9) {\code{1}};
  \node[pop3](a) at (0,-1) {\code{ }};
  \draw[pop3] (0,0) -- (p1.north);
  \draw[pop3] (0,0) -- (n1.north);
  \draw[pop3] (-0.3,-0.3) -- (a.north);
  \end{scope}
        \end{tikzpicture}
      };
    \draw [-> ] (c.east) -- (d.west);

    \node at (c.center) {
\begin{tikzpicture}[scale=0.6]

          \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  \node(x1) at (0,-1) {\code{1}};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);

  \node(t) at (-0.5,0.5) {};
  \draw (l1.south) -- (t.south);
  \node(c) at (-1.5,-0.2) {\code{cons}};
  \draw (t.south) -- (c.north);

  \begin{scope}[shift={(0.9,-1.3)}]
      \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  %\node(x1) at (0,-1) {};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);

  \node(c) at (-0.5,-1.5) {\code{car}};
  \node(z) at (0.5,-1.5) {\code{z}};

  \draw (0,-1) -- (c.north);
  \draw (0,-1) -- (z.north);

%  \node [rotate=90] at (-2.3,-0.7) {\small program};

  \end{scope}

\end{tikzpicture}
    };

    \end{tikzpicture}}

  \end{minipage}

\end{frame}

\begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}

\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}}

\begin{minipage}[c]{0.4\textwidth}
  \begin{tikzpicture}[scale=1,line width=0.5mm]

  \node[latent,scale=1] at (3.5,3) (dx){DSL};
  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
  \draw [->] (zp1.south) -- (xp1.north);
  \draw [->] (dx.south) -- (zp1.north);
  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);

  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
  \draw [->] (zp1.south) -- (xp1.north);
  \draw [->] (dx.south) -- (zp1.north);
  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);

  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
  \draw [->] (dx.south) -- (zp.north);
  \draw [->] (zp.south) -- (xp.north);

  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
  \end{tikzpicture}

  \vspace{0.5cm}

\visible<4->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
    \node at (0,0) {\underline{\textbf{Compile}: Train {\color{red} recognition model}}};
          \node[align=center] at (-3,-1) (d){DSL};
      %    \node at ([xshift = 3cm]d.east) (p2){program};
      \node at ([xshift=2.5cm,yshift = 0cm]d.east) (p1){program};

      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\footnotesize  sample} (p1.west);

      \node at ([xshift = 2.5cm]p1.east) (t1){ task};

      \draw [-> ] (p1.east) -- node[above]{\footnotesize  run} (t1.west);

      %% \node(f1) at ([yshift=-0.9cm,xshift=1cm]d.south) {\footnotesize progs. for task$_i$};
      %% \node(t) at ([xshift = 2cm]f1.east) (p11){\footnotesize program$_i$, task$_i$};
      %% \draw [->,squiggle ] (f1.east) -- node[above]{\footnotesize  sample} (p11.west);

%      \draw[dashed,cyan,very thick] ([yshift=0.5cm]p1.west) rectangle  ([yshift=-0.5cm,xshift=0.5cm]p11.east);

      \node(n) at ([yshift=-0.8cm,xshift=1.6cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [->,red] ([xshift=-0.4cm]n.west) to[out = -180,in = -70]  ([xshift=-0.25cm,yshift=0.1cm]p1.south);

    \end{tikzpicture}}
\end{minipage}
\begin{minipage}[c]{0.55\textwidth}
  \visible<2->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
    \node at (2,1) {\underline{\textbf{Explore}: Infer programs, fixing DSL and neural net}};
    \node at (0,0) (d){DSL};
    \node at ([yshift = -2cm]d) (t){$\text{Task}$};

    \node at ([xshift = 2cm]t) (nn){
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[align = center, text width = 1cm] at ([yshift = 0.9cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
    \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);

    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};

    \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
    \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

    \node[align=left] at ([xshift=2cm]s.east) (f) {Programs};
    \draw [->  ] (s.east) -- (f.west);

    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
  \end{tikzpicture}}

  \vspace{1cm}

\visible<3->{  \begin{tikzpicture}[scale=0.4,line width=0.5mm]

    \node at (0,0) (f1){progs. for task$_1$};
    \node at ([yshift = -1.4cm]f1.south) (f2){progs. for task$_2$};
    \node at ([yshift = -1.4cm,xshift=1.5cm]f2.south) (f3){DSL};

    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 2.5cm, anchor = north west] at (5.2,1) {};
%    \node[anchor=north] at (c.north) {Compression};
    \node at ([yshift=1cm,xshift=-5cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
    \draw [-> ] (f1.east) -- (c.west|-f1.east);
    \draw [-> ] (f2.east) -- (c.west|-f2.east);
    \draw [-> ] (f3.east) -- (c.west|-f3.east);

    \node[right](d) at ([xshift = 1.6cm,yshift = -0.0cm]c.east) {
      \begin{tikzpicture}[scale=0.6]
        \node[align=center] at (0,0) {DSL+};
\begin{scope}[shift={(0.6,-0.5)}]
  \node[pop3](p1) at (-1,-1) {\code{+}};
  \node[pop3](n1) at (0.6,-0.9) {\code{1}};
  \node[pop3](a) at (0,-1) {\code{ }};
  \draw[pop3] (0,0) -- (p1.north);
  \draw[pop3] (0,0) -- (n1.north);
  \draw[pop3] (-0.3,-0.3) -- (a.north);
  \end{scope}
        \end{tikzpicture}
      };
    \draw [-> ] (c.east) -- (d.west);

    \node at (c.center) {
\begin{tikzpicture}[scale=0.6]

          \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  \node(x1) at (0,-1) {\code{1}};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);

  \node(t) at (-0.5,0.5) {};
  \draw (l1.south) -- (t.south);
  \node(c) at (-1.5,-0.2) {\code{cons}};
  \draw (t.south) -- (c.north);

  \begin{scope}[shift={(0.9,-1.3)}]
      \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  %\node(x1) at (0,-1) {};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);

  \node(c) at (-0.5,-1.5) {\code{car}};
  \node(z) at (0.5,-1.5) {\code{z}};

  \draw (0,-1) -- (c.north);
  \draw (0,-1) -- (z.north);

%  \node [rotate=90] at (-2.3,-0.7) {\small program};

  \end{scope}

\end{tikzpicture}
    };

    \end{tikzpicture}}

  \end{minipage}

  \end{frame}

%% \begin{frame}{Compress: Growing the DSL}
%%   \begin{figure}[h!]\tabcolsep=2pt
%% \centering  \begin{minipage}[c]{0.2\textwidth}
%%   \begin{tikzpicture}[scale=0.7]
%%     %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
%%     %%     \footnotesize Program:\\
%%     %%     \code{($\lambda$ (x) (+ (- x) 1))}
%%     %% \end{tabular}};
%%     %\node at (,0.5) {\code{cons}};
%%     \node [rotate=90] at (-2.3,-0.5) {\small program};

%%           \node(l1) at (0,0) {};
%%   \node[color=pop3](p1) at (-1,-1) {\code{+}};
%%   \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
%%   \node(x1) at (0,-1) {\code{1}};
%%   \draw[color=pop3] (l1.south) -- (p1.north);
%%   \draw[color=pop3] (l1.south) -- (n1.north);
%%   \draw[color=pop3] (-0.5,-0.45) -- (x1.north);

%%   \node(t) at (-0.5,0.5) {};
%%   \draw (l1.south) -- (t.south);
%%   \node(c) at (-1.5,-0.2) {\code{cons}};
%%   \draw (t.south) -- (c.north);

%% %    \draw  (l1.south) -- (-0.5,0.5);

%%   %% \node(c) at (-0.5,-1.5) {\code{-}};
%%   %% \node(z) at (0.5,-1.5) {\code{x}};

%%   %% \draw (0,-1) -- (c.north);
%%   %% \draw (0,-1) -- (z.north);

%%   \begin{scope}[shift={(0,-2.5)}]
%%       \node(l1) at (0,0) {};
%%   \node[color=pop3](p1) at (-1,-1) {\code{+}};
%%   \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
%%   %\node(x1) at (0,-1) {};
%%   \draw[color=pop3] (l1.south) -- (p1.north);
%%   \draw[color=pop3] (l1.south) -- (n1.north);
%%   \draw[color=pop3] (-0.5,-0.45) -- (0,-1);

%%   \node(c) at (-0.5,-1.5) {\code{car}};
%%   \node(z) at (0.5,-1.5) {\code{z}};

%%   \draw (0,-1) -- (c.north);
%%   \draw (0,-1) -- (z.north);

%%   \node [rotate=90] at (-2.3,-0.7) {\small program};

%%   \end{scope}

%% \begin{scope}[shift={(0,-5)}]
%%   \node[pop3](p1) at (-1,-1) {\code{+}};
%%   \node[pop3](n1) at (0.8,-0.7) {\code{1}};
%%   \node[pop3](a) at (0,-1) {\code{ }};
%%   %\node(x1) at (0,-1) {};
%%   \draw[pop3] (0,0) -- (p1.north);
%%   \draw[pop3] (0,0) -- (n1.north);
%%   \draw[pop3] (-0.55,-0.4) -- (a.north);
%%   \node [rotate=90] at (-2.3,-0.7) {\small fragment};

%%   \end{scope}

%% \end{tikzpicture}
%%   \end{minipage}
%% \hspace{0.1cm}\begin{tabular}{ll}
%%     \toprule
%%     Programs discovered during Explore&Proposed $\lambda$-expression\\\midrule
%%     %% \begin{tabular}{l}
%%     %%   \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
%%     %%   \phantom{\code{($\lambda$ ($\ell$) (}}\code{(range (- (length $\ell$) 1))))}\\
%%     %%   \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
%%     %%   \phantom{\code{($\lambda$ ($\ell$) (}}\code{(range (+ 1 1))))}\\
%%     %% \end{tabular}&
%%     %% \begin{tabular}{l}
%%     %%   \code{(map ($\lambda$ (x) (index x $\ell$))}\\
%%     %%   \phantom{\code{(map)}}\code{(range $\alpha$))}\\
%%     %% \end{tabular}\\\midrule
%%     \begin{tabular}{l}
%%       \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
%%       \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
%%       \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
%%       \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
%%       \end{tabular}&
%%     \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
%%       \end{tabular}
%% \\\bottomrule\\
%% \end{tabular}
%% \caption{\textbf{Left:} syntax trees of two programs sharing common structure,
%% highlighted in {\orange{orange}},
%%   from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
%%   from which we extract fragments that perform character substitutions.}\label{fragmentExample}
%% \end{figure}

%% \end{frame}

%% \begin{frame}{Compile: Learning to search for programs}
%%   \begin{tikzpicture}[scale=0.85,line width=0.5mm]
%%     \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
%%     \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Encoder\\
%%       e.g., RNN or CNN};
%%     \node(m)[draw,align = center] at ([xshift=1.8cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
%%     \node[draw,align = center](s) at ([xshift=3cm]m.east) {Enumerate\\(Explore)};
%%     \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Programs\\solving $x$};

%%     \draw[->] (x.east) -- (e.west);
%%     \draw[->] (e.east) -- (m.west);
%%     \draw[->] (m.east) -- (s.west);
%%     \draw[->] (s.east) -- (f.west);
%%     \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway](likelihood){\small likelihood $\probability[\text{task}|\text{prog}]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
%%     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
%%     ([xshift = -5,yshift = -25]e.west) -- node[below = 3pt] {{Recognition model $Q(\text{prog}|\text{task})$}} ([xshift = 180,yshift = -25]e.west);
%%     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
%%     ([xshift = -5,yshift = -25]s.west) -- node[below = 3pt] {{ Search}} ([xshift = 5,yshift = -25]f.east);
%%     %% \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
%%     %% ([xshift = -5,yshift = -25]p.west) -- node[below = 3pt] {{\small Params}} ([xshift = 5,yshift = -25]p.east);
%%     \node at ([yshift=0.5cm]likelihood.north) {\textbf{\underline{Inference at test time (\underline{Explore Step})}}};
%%   \end{tikzpicture}

%%   \vspace{0.5cm}

%%   \centering\emph{\textbf{Minimize Loss}}: $\text{KL}(\text{P}\left[\text{prog}|\text{DSL},\text{task} \right] || Q(\text{prog}|\text{task}))$

%%   \vspace{0.5cm}
%%   \begin{tikzpicture}[scale=1,line width=0.5mm]
%% \visible<2->{    \node at (0,0) (f1){progs. for task$_1$};
%%     \node at ([yshift = -0.6cm]f1.south) (f2){progs. for task$_2$};
%%     %    \node at ([yshift = -1cm]f2.south) (f3){progs. for task$_3$};

%%     \node at ([xshift = 2cm]f1.east) (p1){program$_1$, task$_1$};
%%     \node at ([xshift = 2cm]f2.east) (p2){program$_2$, task$_2$};
%%     %    \node at ([xshift = 2cm]f3.east) (p3){program$_3$, task$_3$};

%%     \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
%%     \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
%%     %    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);

%%     \node at ([yshift = 0.5cm,xshift=-1.3cm]p1.north) {\textsc{\textbf{\underline{Training on real data}}}};
%%     }
%%     %% \end{tikzpicture}%
%%     %%   \begin{tikzpicture}[scale=3,line width=1.5mm]
%%     \begin{scope}[shift = {(6cm,-0.5cm)}]

%% \visible<3>{      \node[align=center] at (0,0) (d){DSL};
%%       %    \node at ([xshift = 3cm]d.east) (p2){program};
%%       \node at ([xshift=2cm,yshift = 0.5cm]d.east) (p1){program};
%%       \node at ([xshift=2cm,yshift = -0.5cm]d.east) (p3){program};

%%       \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
%%       \draw[squiggle,-> ] (d.east) -- node[sloped,below]{\small sample} (p3.west);

%%       \node at ([xshift = 2cm]p1.east) (t1){task};
%%       \node at ([xshift = 2cm]p3.east) (t3){task};
%%       \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
%%       \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);

%%       \node at ([yshift = 0.5cm]p1.north) {\textsc{\textbf{\underline{Training on DSL samples}}}};
%% }    \end{scope}
%%   \end{tikzpicture}
%% \end{frame}

\begin{frame}{Domain: List processing}
      Starts with: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
\code{car}, \code{cdr}, \code{nil}, \code{is-nil},  \code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
\\236 human-interpretable list processing tasks.
\\Discovers 38 new DSL primitives, including \code{filter}

\vspace{0.25cm}

\centering \begin{minipage}[c]{0.6\textwidth}
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
\end{minipage}%
%% \begin{minipage}[c]{0.3\textwidth}
%%   \includegraphics[width = \textwidth]{figures/listLearningCurve_color.eps}
%% {   \color{orange}orange: } full system. {\color{teal}teal: } no neural recognition model
%% \end{minipage}

\vspace{0.25cm}

\centering \textbf{With functional programming ``problem set'' + 93 hours on 64 CPUs, rediscovers:}\\
\textbf{\code{map}, \code{foldr}, \code{unfold}, \code{range}, \code{length}, \code{index}, \code{zip}}, +some arithmetic routines

  \end{frame}

\begin{frame}{Domain: Text Editing}
  In the style of FlashFill (Gulwani 2012). Starts with: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
\code{car}, \code{cdr}, \code{nil}, \code{is-nil}
, plus string \& character constants.\\

\begin{minipage}[c]{0.4\textwidth}
  \begin{tabular}{ll}
    \toprule
    Input&Output\\\midrule
    +106 769-438&106.769.438\\
    +83 973-831&83.973.831
    \\\midrule
          Temple Anna H&TAH\\
      Lara Gregori&LG
    \\\bottomrule 
    \end{tabular}
  %% \begin{figure}
  %% \includegraphics[width = 0.8\textwidth]{figures/textLearningCurve.eps}
  %%   \end{figure}
  \end{minipage}
\begin{minipage}[c]{0.5\textwidth}
      \begin{figure}        
        \includegraphics[width = \textwidth]{figures/textPrimitives.png}\vspace{-0.8cm}
      \caption{11 learned DSL primitives over 3 successive iterations (3 columns). Learned primitives call each other (arrows).}
      \end{figure}  
  \end{minipage}
\end{frame}

\begin{frame}{Domain: Symbolic regression from visual input}
  Starts with: \code{plus}, \code{times}, \code{divide}, \code{real-number}. Autograds through program
  \\  Bayesian likelihood $\probability[\text{data}|\text{prog}]$ favors fewer continuous parameters

  \centering
  %% \begin{minipage}[c]{0.4\textwidth}
  %% \begin{figure}
  %%   \newcommand{\functionSize}{1.5cm}\centering
  %% \includegraphics[width = \functionSize]{figures/functions/6.png}
  %% \includegraphics[width = \functionSize]{figures/functions/48.png}
  %% \includegraphics[width = \functionSize]{figures/functions/102.png}
  %% \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  %% \vspace{0.0pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  %% \includegraphics[width = \functionSize]{figures/functions/160.png}
  %% \includegraphics[width = \functionSize]{figures/functions/149.png}
  %%   \includegraphics[width = \functionSize]{figures/functions/187.png}
  %% \caption{Recognition model input. DSL learns subroutines for polynomials/rational functions (top/bottom) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-0.5cm}
  %%   \end{figure}    
  %% \end{minipage}
  %
\centering

\vspace{1cm}\begin{tabular}{cc}
    \toprule
\normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule 
\small    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_5$ x)}}\\
%      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

    \end{tabular}
&    \small
    \begin{tabular}{l}
    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real ($f_0$ x))}}\\
    \hspace{\helpSize}\emph{($f_5$: rational function)}\\

  \end{tabular}
  \\\bottomrule\\
\end{tabular}
  \end{frame}

\begin{frame}{New domain: Generative  graphics  programs (Turtle/LOGO)}
  \begin{tabular}{cc|c}
    Training tasks& DSL samples before learning&DSL samples after learning\\
              \raisebox{-.5\height}{\includegraphics[width = 0.2\textwidth]{figures/logo90.png}$\quad\quad$ }&
      \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/initialDreams/montage.png}}&
  \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/finalDreams/montage.png}}
  \end{tabular}

  \vspace{0.3cm}

  \pause

  Learned DSL contains parametric drawing routines:
      \begin{tabular}{rlrl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}&
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}&
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
        &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}&
                Circles:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}
    \end{tabular}

  \end{frame}

\end{document}

### 2018-12-05 archive.tex
spotlight before Armando revisions

@@
 
+\begin{tabular}{}
+  \end{tabular}
+
 \begin{table*}%[t!]
@@
     \toprule
-    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
+    &{\normalsize List Processing}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
     \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
@@
 into the frontiers will most increase $\lowerBound$.
-
-Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figure~\ref{feeding}).
+This neural net is trained on two sources of data: samples, or `dreams',
+drawn from the DSL,
+as well as
+replays of (program, task) pairs discovered during waking.
+
+Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figures~\ref{feeding}~and~\ref{bayesianSummary}).
 As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
@@
 we can draw unlimited samples from $(\mathcal{D},\theta)$
-and train $q$ on those samples.
+and train $Q$ on those samples.
 But, evaluating $\mathcal{L}_{\text{Dream}}$ involves drawing programs from
@@
 
-    %\rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
     \begin{tabular}{l}
+      \code{[1\, 9]}$\to $\code{2}\\
+      \code{[5\, 3\, 8]$\to$3}\\
+    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
+      \code{[true\, false]}$\to $\code{2}\\
+      \code{[false\, false\, false]$\to$3}\\
+    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
+
 \code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
@@
 
+\code{[2\, 1\, 4]$\to$[2\, 1]}\\
+%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
+\code{[9\, 8]$\to$[9]}\\
+\blueCode{$f(\ell) = $($f_0$ ($\lambda$ (z) (empty? (cdr z))) car cdr $\ell$)}\\\\
+
 \code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
@@
 \blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
+\code{[4\, 2\, 6\, 4]$\to$[-4\, -2\, -6\, -4]}\\
+%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+\code{[2\, 3\, 0\, 7]$\to$[-2\, -3\, -0\, -7]}\\
+\blueCode{$f(\ell) = $($f_3$ (- 0) $\ell$)}\\\\
+\code{[4\, 2\, 6\, 4]$\to$[5\, 3\, 7\, 5]}\\
+%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+\code{[2\, 3\, 0\, 7]$\to$[3\, 4\, 1\, 8]}\\
+\blueCode{$f(\ell) = $($f_3$ (+ 1) $\ell$)}\\\\
 
@@
       \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
-      \\
+      \\\\
+
+      \code{3$\to $[0\, 1\, 2]}\\
+      \code{2$\to $[0\, 1]}\\
+      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 $n$)}\\\\
+
+      \code{3$\to $[0\, 1\, 2\, 3]}\\
+      \code{2$\to $[0\, 1\, 2]}\\
+      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 (+ 1 $n$))}\\\\
 
@@
         \hspace{\helpSize}($f_4$: \emph{length})\\
-        \greenCode{$f_5(n)$\,=\,($f_0$ (= $n$) ($\lambda$ (x) x) (+ 1) 0)}\\
-        \hspace{\helpSize}($f_5$: \emph{range})
+        \greenCode{$f_5(\code{f},\code{m},\code{n})$\,=\,($f_0$ (= m) f (+ 1) n)}\\
+        \hspace{\helpSize}($f_5$: \emph{generalization of range})
 
@@
 which calculates the set of $n$-step refactorings.
-We define this operator recursively as follows:
+We define this operator recursively in terms of another operator, $I\beta'$, which performs a single step of refactoring:
 \begin{align*}
-  I\beta_n(v)& = \uplus \left\{ \underbrace{I\beta'(I\beta'(I\beta'(\cdots}_{i \text{ times}} v))) \;:\; 0\leq i \leq n \right\}\\
-  I\beta'(v)& = \uplus \left\{(\lambda b)v\;:\;v\mapsto b\in S_0(v)\text{, when }v\not=\Lambda \right\}\cup
+  I\beta_n(v)& = \uplus \left\{ \underbrace{I\beta'(I\beta'(I\beta'(\cdots}_{i \text{ times}} v))) \;:\; 0\leq i \leq n \right\}
+\end{align*}
+where
+  \begin{align*}
+    I\beta'(u)& = \uplus \left\{(\lambda b)v\;:\;v\mapsto b\in S_0(u)%\text{, when }v\not=\Lambda
+    \right\}\cup
   \begin{cases}
-    \text{if $v$ is a primitive or index:}&\varnothing\\
-    \text{if $v = \lambda b$:}&\lambda I\beta'(b)\\
-    \text{if $v = (f\;x)$:}&(I\beta'(f)\;I\beta'(x))\\
+    \text{if $u$ is a primitive or index or $\varnothing $ or $\Lambda$:}&\varnothing\\
+    \text{if $u = \lambda b$:}&\lambda I\beta'(b)\\
+    \text{if $u = (f\;x)$:}&(I\beta'(f)\;I\beta'(x))\\
+    \text{if $u = \uplus V$:}&\uplus\left\{I\beta'(u')\;|\;u'\in V \right\}
   \end{cases}  
-  \\
+  \end{align*}
+  where we have defined $I\beta'$
+  in terms of another operator, $S_k:\text{VS}\to 2^{\text{VS}\times\text{VS}}$,
+  whose purpose is to construct the set of
+  substitutions that are refactorings of a program in a version space.
+We define $S$ as:
+  \begin{align*}
   S_k(v)& = \left\{\downshift{k}_0v\mapsto \$k \right\}\cup
@@
 where $\shift{k}$ is the shifting operator~\cite{pierce},
-which adds $k$ to all of the free variables in a $\lambda$-expression or version space.
-We have written the definition of $I\beta_n$ recursively,
-but implement it using a dynamic program:
+which adds $k$ to all of the free variables in a $\lambda$-expression or version space,
+and we have defined a new operator, $\downshift$, whose purpose is to
+undo the action of $\shift$.
+We have written definitions recursively,
+but implement them using a dynamic program:
 we hash cons each version space,
@@
 
+From these results the consistency and completeness of $I\beta$ follows immediately by its definition.
+
+%% First observe that
+%% we do not \emph{enumerate} every refactoring,
+%% but only those for which every expression of the form $(\lambda x.b)v$ has
+%% $x$ free in $b$.
+%% For example, we do not refactor $\code{+}$ into $(\lambda x. \code{+})v$ where $v$ is arbitrary,
+%% because there are infinitely many such refactorings.
+%% Define the following relation,
+%% which captures this restricted kind of
+%% refactoring:
+%% \begin{gather*}
+%%   \frac{f\reduce^\sim f'}{f\;x \reduce^\sim f'\;x}\qquad
+%%   \frac{x\reduce^\sim x'}{f\;x \reduce^\sim f\;x'}\qquad
+%%   \frac{b\reduce^\sim b'}{\lambda b \reduce^\sim \lambda b'}\qquad
+%%   \frac{\$0 \text{ occurs free in }b}{(\lambda b)v\reduce^\sim \shift{-1}[\$0\mapsto \shift{1}v]b}
+%%   \end{gather*}
+%% \begin{theorem}
+%%   \textbf{Completeness of $I\beta'$.}
+%%   Let $v$ be a version space and let $p\in v$.
+%%   If $p'\reduce^\sim p$
+%%   then $p'\in I\beta'(v)$.
+%% \end{theorem}
+%% \begin{proof}
+%%   We know (from the definition of $\reduce^\sim$) that $p'$ and $p$ differ only
+%%   in that there exists a single subexpression in
+%%   $p'$ of the form $(\lambda b)v$
+%%   where that subexpression occurs in
+%%   $p$ as $\shift{-1}[\$0\mapsto \shift{1}v]b$.
+%%   From Theorem~\ref{}
+%% \end{proof}
+
 \subsubsection{Computational complexity of DSL learning}

### 2018-12-05 spotlight.tex
spotlight before Armando revisions

@@
 
+\usepackage{amssymb,amsmath}
 \usepackage{booktabs}
@@
 
-\begin{minipage}[c]{0.4\textwidth}
-  \begin{tikzpicture}[scale=1,line width=0.5mm]
-
-  \node[latent,scale=1] at (3.5,3) (dx){DSL};
-  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
-  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
-  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
-  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
-  \draw [->] (zp1.south) -- (xp1.north);
-  \draw [->] (dx.south) -- (zp1.north);
-  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-
-  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
-  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
-  \draw [->] (zp1.south) -- (xp1.north);
-  \draw [->] (dx.south) -- (zp1.north);
-  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-
-  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
-  \draw [->] (dx.south) -- (zp.north);
-  \draw [->] (zp.south) -- (xp.north);
-
-  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
-  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
-  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
-  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
-  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
-  \end{tikzpicture}
-
-  \vspace{0.5cm}
-
-\visible<4->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
-    \node at (0,0) {\underline{\textbf{Compile}: Train {\color{red} recognition model}}};
-          \node[align=center] at (-3,-1) (d){DSL};
-      %    \node at ([xshift = 3cm]d.east) (p2){program};
-      \node at ([xshift=2.5cm,yshift = 0cm]d.east) (p1){program};
-
-      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\footnotesize  sample} (p1.west);
-
-      \node at ([xshift = 2.5cm]p1.east) (t1){ task};
-
-      \draw [-> ] (p1.east) -- node[above]{\footnotesize  run} (t1.west);
-
-      %% \node(f1) at ([yshift=-0.9cm,xshift=1cm]d.south) {\footnotesize progs. for task$_i$};
-      %% \node(t) at ([xshift = 2cm]f1.east) (p11){\footnotesize program$_i$, task$_i$};
-      %% \draw [->,squiggle ] (f1.east) -- node[above]{\footnotesize  sample} (p11.west);
-
-%      \draw[dashed,cyan,very thick] ([yshift=0.5cm]p1.west) rectangle  ([yshift=-0.5cm,xshift=0.5cm]p11.east);
-
-      \node(n) at ([yshift=-0.8cm,xshift=1.6cm]p1.south) {
-        \NeuralNetwork{0.17}};
-      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
-      \draw [->,red] ([xshift=-0.4cm]n.west) to[out = -180,in = -70]  ([xshift=-0.25cm,yshift=0.1cm]p1.south);
-
-    \end{tikzpicture}}
-\end{minipage}
-\begin{minipage}[c]{0.55\textwidth}
+\begin{minipage}[c]{0.6\textwidth}
   \visible<2->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
       \node at (2,1) {\underline{\textbf{Explore}: Infer programs, fixing DSL and neural net}};
+      \begin{scope}[shift = {(-3,0)}]
     \node at (0,0) (d){DSL};
@@
     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+      \end{scope}
+      \visible<5->{
+        \node(repeat)[scale=8] at (6,-3.5) {$\circlearrowleft$};
+        \node at (repeat) {\emph{Repeat!}};
+      }
   \end{tikzpicture}}
 
-  \vspace{1cm}
+  \vspace{-1.5cm}
 
 \visible<3->{  \begin{tikzpicture}[scale=0.4,line width=0.5mm]
-
     \node at (0,0) (f1){progs. for task$_1$};
@@
 %    \node[anchor=north] at (c.north) {Compression};
-    \node at ([yshift=1cm,xshift=-5cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
+    \node at ([yshift=1cm,xshift=-3cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
     \draw [-> ] (f1.east) -- (c.west|-f1.east);
@@
 
-  \end{minipage}
-
-\end{frame}
-
-\begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}
-
-\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
-      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-      \node[rectangle] at (1,1) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (I-\name) at (0,-\y) {};
-      \foreach \name / \y in {1,...,3}
-          \node[neuron] (H-\name) at (1,-\y-0.5) {};
-      \foreach \name / \y in {1,...,4}
-          \node[neuron] (O-\name) at (2,-\y) {};
-      \foreach \source in {1,...,4}
-          \foreach \dest in {1,...,3}
-              \draw [-latex] (I-\source) -- (H-\dest);
-      \foreach \source in {1,...,3}
-          \foreach \dest in {1,...,4}
-              \draw [-latex] (H-\source) -- (O-\dest);
-    \end{tikzpicture}}
-
-\begin{minipage}[c]{0.4\textwidth}
+\end{minipage}%
+\begin{minipage}[c]{0.28\textwidth}
+  \vspace{-1cm}
   \begin{tikzpicture}[scale=1,line width=0.5mm]
@@
 
-      %% \node(f1) at ([yshift=-0.9cm,xshift=1cm]d.south) {\footnotesize progs. for task$_i$};
-      %% \node(t) at ([xshift = 2cm]f1.east) (p11){\footnotesize program$_i$, task$_i$};
-      %% \draw [->,squiggle ] (f1.east) -- node[above]{\footnotesize  sample} (p11.west);
-
-%      \draw[dashed,cyan,very thick] ([yshift=0.5cm]p1.west) rectangle  ([yshift=-0.5cm,xshift=0.5cm]p11.east);
-
-      \node(n) at ([yshift=-0.8cm,xshift=1.6cm]p1.south) {
+      \node(n) at ([yshift=-1.2cm,xshift=1.6cm]p1.south) {
         \NeuralNetwork{0.17}};
@@
       \draw [->,red] ([xshift=-0.4cm]n.west) to[out = -180,in = -70]  ([xshift=-0.25cm,yshift=0.1cm]p1.south);
+%      \node at ([xshift=-0.5cm]$(p1.south)!0.5!(n.west)$) {\textbf{Loss4%}};
 
@@
 \end{minipage}
-\begin{minipage}[c]{0.55\textwidth}
-  \visible<2->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
-    \node at (2,1) {\underline{\textbf{Explore}: Infer programs, fixing DSL and neural net}};
-    \node at (0,0) (d){DSL};
-    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
-
-    \node at ([xshift = 2cm]t) (nn){
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[align = center, text width = 1cm] at ([yshift = 0.9cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
-    \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
-
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
-
-    \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-    \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
-
-    \node[align=left] at ([xshift=2cm]s.east) (f) {Programs};
-    \draw [->  ] (s.east) -- (f.west);
-
-    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-  \end{tikzpicture}}
-
-  \vspace{1cm}
-
-\visible<3->{  \begin{tikzpicture}[scale=0.4,line width=0.5mm]
-
-    \node at (0,0) (f1){progs. for task$_1$};
-    \node at ([yshift = -1.4cm]f1.south) (f2){progs. for task$_2$};
-    \node at ([yshift = -1.4cm,xshift=1.5cm]f2.south) (f3){DSL};
-
-    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 2.5cm, anchor = north west] at (5.2,1) {};
-%    \node[anchor=north] at (c.north) {Compression};
-    \node at ([yshift=1cm,xshift=-5cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
-    \draw [-> ] (f1.east) -- (c.west|-f1.east);
-    \draw [-> ] (f2.east) -- (c.west|-f2.east);
-    \draw [-> ] (f3.east) -- (c.west|-f3.east);
-
-    \node[right](d) at ([xshift = 1.6cm,yshift = -0.0cm]c.east) {
-      \begin{tikzpicture}[scale=0.6]
-        \node[align=center] at (0,0) {DSL+};
-\begin{scope}[shift={(0.6,-0.5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.6,-0.9) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.3,-0.3) -- (a.north);
-  \end{scope}
-        \end{tikzpicture}
-      };
-    \draw [-> ] (c.east) -- (d.west);
-
-    \node at (c.center) {
-\begin{tikzpicture}[scale=0.6]
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-  \begin{scope}[shift={(0.9,-1.3)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-%  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-  \end{scope}
-
-\end{tikzpicture}
-    };
-
-    \end{tikzpicture}}
-
-  \end{minipage}
 
 \end{frame}
+
+%% \begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}
+
+%% \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+%%       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%       \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%       \node[rectangle] at (1,1) {};
+%%       \foreach \name / \y in {1,...,4}
+%%           \node[neuron] (I-\name) at (0,-\y) {};
+%%       \foreach \name / \y in {1,...,3}
+%%           \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%       \foreach \name / \y in {1,...,4}
+%%           \node[neuron] (O-\name) at (2,-\y) {};
+%%       \foreach \source in {1,...,4}
+%%           \foreach \dest in {1,...,3}
+%%               \draw [-latex] (I-\source) -- (H-\dest);
+%%       \foreach \source in {1,...,3}
+%%           \foreach \dest in {1,...,4}
+%%               \draw [-latex] (H-\source) -- (O-\dest);
+%%     \end{tikzpicture}}
+
+%% \begin{minipage}[c]{0.4\textwidth}
+%%   \begin{tikzpicture}[scale=1,line width=0.5mm]
+
+%%   \node[latent,scale=1] at (3.5,3) (dx){DSL};
+%%   \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+%%   \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+%%   \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+%%   \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+%%   \draw [->] (zp1.south) -- (xp1.north);
+%%   \draw [->] (dx.south) -- (zp1.north);
+%%   \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+%%   \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+%%   \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+%%   \draw [->] (zp1.south) -- (xp1.north);
+%%   \draw [->] (dx.south) -- (zp1.north);
+%%   \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+%%   \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+%%   \draw [->] (dx.south) -- (zp.north);
+%%   \draw [->] (zp.south) -- (xp.north);
+
+%%   \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+%%   \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+%%   \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
+%%   \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+%%   ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+%%   \end{tikzpicture}
+
+%%   \vspace{0.5cm}
+
+%% \visible<4->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
+%%     \node at (0,0) {\underline{\textbf{Compile}: Train {\color{red} recognition model}}};
+%%           \node[align=center] at (-3,-1) (d){DSL};
+%%       %    \node at ([xshift = 3cm]d.east) (p2){program};
+%%       \node at ([xshift=2.5cm,yshift = 0cm]d.east) (p1){program};
+
+%%       \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\footnotesize  sample} (p1.west);
+
+%%       \node at ([xshift = 2.5cm]p1.east) (t1){ task};
+
+%%       \draw [-> ] (p1.east) -- node[above]{\footnotesize  run} (t1.west);
+
+%%       %% \node(f1) at ([yshift=-0.9cm,xshift=1cm]d.south) {\footnotesize progs. for task$_i$};
+%%       %% \node(t) at ([xshift = 2cm]f1.east) (p11){\footnotesize program$_i$, task$_i$};
+%%       %% \draw [->,squiggle ] (f1.east) -- node[above]{\footnotesize  sample} (p11.west);
+
+%% %      \draw[dashed,cyan,very thick] ([yshift=0.5cm]p1.west) rectangle  ([yshift=-0.5cm,xshift=0.5cm]p11.east);
+
+%%       \node(n) at ([yshift=-0.8cm,xshift=1.6cm]p1.south) {
+%%         \NeuralNetwork{0.17}};
+%%       \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
+%%       \draw [->,red] ([xshift=-0.4cm]n.west) to[out = -180,in = -70]  ([xshift=-0.25cm,yshift=0.1cm]p1.south);
+
+%%     \end{tikzpicture}}
+%% \end{minipage}
+%% \begin{minipage}[c]{0.55\textwidth}
+%%   \visible<2->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
+%%     \node at (2,1) {\underline{\textbf{Explore}: Infer programs, fixing DSL and neural net}};
+%%     \node at (0,0) (d){DSL};
+%%     \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+
+%%     \node at ([xshift = 2cm]t) (nn){
+%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+%%         \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%         \node[rectangle] at (1,1) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (I-\name) at (0,-\y) {};
+%%         \foreach \name / \y in {1,...,3}
+%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (O-\name) at (2,-\y) {};
+%%         \foreach \source in {1,...,4}
+%%             \foreach \dest in {1,...,3}
+%%                 \draw [-latex] (I-\source) -- (H-\dest);
+%%         \foreach \source in {1,...,3}
+%%             \foreach \dest in {1,...,4}
+%%                 \draw [-latex] (H-\source) -- (O-\dest);
+%%       \end{tikzpicture}
+%%     };
+%%     \node[align = center, text width = 1cm] at ([yshift = 0.9cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
+%%     \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+%%     \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+
+%%     \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+%%     \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+%%     \node[align=left] at ([xshift=2cm]s.east) (f) {Programs};
+%%     \draw [->  ] (s.east) -- (f.west);
+
+%%     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+%%   \end{tikzpicture}}
+
+%%   \vspace{1cm}
+
+%% \visible<3->{  \begin{tikzpicture}[scale=0.4,line width=0.5mm]
+
+%%     \node at (0,0) (f1){progs. for task$_1$};
+%%     \node at ([yshift = -1.4cm]f1.south) (f2){progs. for task$_2$};
+%%     \node at ([yshift = -1.4cm,xshift=1.5cm]f2.south) (f3){DSL};
+
+%%     \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 2.5cm, anchor = north west] at (5.2,1) {};
+%% %    \node[anchor=north] at (c.north) {Compression};
+%%     \node at ([yshift=1cm,xshift=-5cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
+%%     \draw [-> ] (f1.east) -- (c.west|-f1.east);
+%%     \draw [-> ] (f2.east) -- (c.west|-f2.east);
+%%     \draw [-> ] (f3.east) -- (c.west|-f3.east);
+
+%%     \node[right](d) at ([xshift = 1.6cm,yshift = -0.0cm]c.east) {
+%%       \begin{tikzpicture}[scale=0.6]
+%%         \node[align=center] at (0,0) {DSL+};
+%% \begin{scope}[shift={(0.6,-0.5)}]
+%%   \node[pop3](p1) at (-1,-1) {\code{+}};
+%%   \node[pop3](n1) at (0.6,-0.9) {\code{1}};
+%%   \node[pop3](a) at (0,-1) {\code{ }};
+%%   \draw[pop3] (0,0) -- (p1.north);
+%%   \draw[pop3] (0,0) -- (n1.north);
+%%   \draw[pop3] (-0.3,-0.3) -- (a.north);
+%%   \end{scope}
+%%         \end{tikzpicture}
+%%       };
+%%     \draw [-> ] (c.east) -- (d.west);
+
+%%     \node at (c.center) {
+%% \begin{tikzpicture}[scale=0.6]
+
+%%           \node(l1) at (0,0) {};
+%%   \node[color=pop3](p1) at (-1,-1) {\code{+}};
+%%   \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+%%   \node(x1) at (0,-1) {\code{1}};
+%%   \draw[color=pop3] (l1.south) -- (p1.north);
+%%   \draw[color=pop3] (l1.south) -- (n1.north);
+%%   \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+%%   \node(t) at (-0.5,0.5) {};
+%%   \draw (l1.south) -- (t.south);
+%%   \node(c) at (-1.5,-0.2) {\code{cons}};
+%%   \draw (t.south) -- (c.north);
+
+%%   \begin{scope}[shift={(0.9,-1.3)}]
+%%       \node(l1) at (0,0) {};
+%%   \node[color=pop3](p1) at (-1,-1) {\code{+}};
+%%   \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+%%   %\node(x1) at (0,-1) {};
+%%   \draw[color=pop3] (l1.south) -- (p1.north);
+%%   \draw[color=pop3] (l1.south) -- (n1.north);
+%%   \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+%%   \node(c) at (-0.5,-1.5) {\code{car}};
+%%   \node(z) at (0.5,-1.5) {\code{z}};
+
+%%   \draw (0,-1) -- (c.north);
+%%   \draw (0,-1) -- (z.north);
+
+%% %  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+%%   \end{scope}
+
+%% \end{tikzpicture}
+%%     };
+
+%%     \end{tikzpicture}}
+
+%%   \end{minipage}
+
+%%   \end{frame}
 
@@
 
+\begin{frame}{\Huge Learning to program: Poster AB \#24}
+
+\centering\includegraphics[width = \textwidth]{figures/finale.png}
+
+  \end{frame}
 \end{document}

### 2018-12-16 archive.tex
3-panel figure

@@
 
+\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
+\newcommand{\spiral}[2]{
+  \draw[ultra thick,->] ([shift={#1}]-30:#2) arc [radius = #2, start angle = -30, end angle = 90];
+  \draw[ultra thick,->] ([shift={#1}]-30:#2) arc [radius = #2, start angle = -30, end angle = 95];
+
+      \draw[ultra thick,->] ([shift={#1}]90:#2) arc [radius = #2, start angle = 90, end angle = 210];
+      \draw[ultra thick,->] ([shift={#1}]90:#2) arc [radius = #2, start angle = 90, end angle = 205];
+
+      \draw[ultra thick,->] ([shift={#1}]210:#2) arc [radius = #2, start angle = 210, end angle = 340];
+      \draw[ultra thick,->] ([shift={#1}]210:#2) arc [radius = #2, start angle = 210, end angle = 335];
+}
+\newcommand{\legend}{
+  \begin{tikzpicture}
+    \node at (0,0) (uses){is};
+    \draw[->,red] ([xshift=-0.6cm]uses.west)  -- (uses.west);
+    \node at ([xshift=0.4cm]uses.east) {\NeuralNetwork{0.15}};
+    \draw[thin] (-1,-0.4) rectangle (1.2,0.4);
+  \end{tikzpicture}
+  }
+
 \begin{document}
 
-\section{Introduction: Expertise}
-
-Human learners acquire expertise in a wide range of domains: some of
-us become experts in calculus, or cooking, or biology, music, tennis,
-or software engineering, to name just a few, and every child develops
-expertise in natural language, intuitive physics~\cite{Pete?},
-intuitive psychology (theory-of-mind), and motor control.  This
-picture contrasts sharply with the current state of machine
-intelligence, where a machine is built to be an expert in a single
-domain, like boardgames~\cite{alphaGo}, medical
-diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
-proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
-Thus an outstanding challenge in the long-term program of building
-more humanlike machines is to develop an algorithm that, like people,
-autonomously acquires expertise across many different kinds of
-domains.
-
-We contribute a model of the development of expertise that combines
-two key ingredients.  First, an expert needs a sufficiently expressive
-knowledge representation.  Following a long tradition in cognitive
-science and AI, we represent knowledge as programs, which prior work
-has used to represent expertise in recognition and generation of
-handwriting and speech~\cite{lake2015human}, intuitive theories (of
-kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
-grammar and
-semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
-Second, atop this program representation, experts possess two kinds of
-domain expertise.  They have at their disposal a powerful, yet
-specialized, repertoire of concepts and abstractions: e.g., in
-architecture, these are concepts like `arch' or `foundation'; in
-software engineering these are libraries of code and domain-specific
-languages (DSLs).  Here our model represents knowledge as programs,
-and so we identify this kind of expertise with a DSL.  Experts also
-have knowledge of when and how to deploy these domain-specific
-concepts efficiently when solving new problems: e.g., mathematicians
-efficiently search the space of proofs, intuiting which lemmas are
-appropriate when; expert chefs intuit which compositions of
-ingredients are likely tasty, before they actually start cooking.  For
-our model, this aspect of expertise corresponds to the ability to
-quickly assemble new, useful programs out of its DSL.
-
-We integrate these ideas into a model called DreamCoder which acquires
-expertise through a novel kind of wake/sleep or `dream' learning. The
-model iterates through wake cycles -- where it solves problems by
-writing programs -- and a pair of sleep cycles: a sleep cycle that
-grows its DSL by replaying experiences from waking and consolidating
-them into new abstractions, and a sleep cycle that improves its
-knowledge of how to write programs by training a neural network on
-replayed experiences as well as `dreams', or samples, from its DSL.
-
-DreamCoder builds on multiple generations of AI research, going back
-to the 1960's~\cite{solomonoff1964formal} when program-learning was
-proposed as a paradigm for general AI.  Broadly speaking recent work
-has either developed neural approaches for learning to efficiently
-deploy a fixed
-DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
-developed symbolic approaches for representing and searching through
-spaces of
-programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
-We were motivated by approaches that learn or grow the
-DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
-Our goal with DreamCoder is to show that the combination of
-neurally-guided search and DSL learning is a uniquely powerful way of
-building systems that, like human learners, autonomously acquire the
-expertise needed to navigate a new domain of problems.
-
-\section{Introduction: AI/PL/dreaming}
+\begin{figure}
+  \begin{tikzpicture}[line width=0.4mm]
+    \draw (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
+    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
+
+    \begin{scope}[shift={(0.5,0.5)}]
+      \node[align=right] at (0,0) (d){\textbf{DSL}\\
+        $f_1$, $f_2$, ...};
+      \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
+        \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+        \footnotesize    \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} };
+
+      \node at ([xshift = 1.25cm]t.east) (nn){\NeuralNetwork{0.25}};
+      \node[align = center, text width = 1cm] at ([yshift = 0.7cm,xshift=0cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
+      \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+      \node[draw,rounded corners, align=center, inner sep = 10] at ([xshift = 4.2cm,yshift = 1cm]t.east) (s){Neurally-Guided\\ Enumerative Search};
+
+      \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+      \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+      \node[align=left] at ([xshift=3cm]s.east) (f) {\textbf{Programs:}\\
+        \small    \code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}\\
+        \small $\cdots\cdots\cdots$};
+      \draw [->  ] (s.east) -- (f.west);
+
+      \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+    \end{scope}
+    \begin{scope}[shift={(9.4,-3.5)},scale=0.6,line width=0.05mm]
+      \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
+      \node[latent,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+      \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
+      \node[latent,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
+      \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+      \node[latent,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
+      \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->] (dx.south) -- (zp.north);
+      \draw [->] (zp.south) -- (xp.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+    \end{scope}
+
+    \node at (0.5,-4.75) {\textbf{\textsc{Memory Consolidation}}};
+    \draw (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
+    \node at (11.5,-4.75) {\textbf{\textsc{Dreaming}}};
+    \draw (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
+
+    \begin{scope}[shift={(9,-4.5)}]
+      \node(dreaming) at (2,-1) {\underline{Fantasies}};
+      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
+      \node at ([xshift=1.5cm]d.east) (p1){prog.};
+      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
+      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
+      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
+      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
+        \NeuralNetwork{0.17}};
+      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
+      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
+
+      \node(replay) at (2,-5) {\underline{Replays}};
+      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
+      \node at ([xshift=1.5cm]d.east) (p1){prog.};
+      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
+      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
+      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
+      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
+        \NeuralNetwork{0.17}};
+      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
+      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
+
+      \begin{scope}[shift={(-3.3,-8)},scale=0.6,line width=0.05mm]
+        \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
+        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \draw [->] (dx.south) -- (zp.north);
+        \draw [->] (zp.south) -- (xp.north);
+        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+        \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+      \end{scope}
+
+      \end{scope}
+
+    %%}b memory consolidation
+    \begin{scope}[shift={(-2,-4.5)}]
+      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\scode{(+ (car z) 1)}};
+      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\scode{(cons (+ 1 1))}};
+      \node[align=center] at ([yshift = 0cm,xshift=1cm]f2.east) (f3){\textbf{DSL}\\\scode{cons}, \scode{+}, \\...};
+      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm]f2.south) {};
+      \draw [-> ] (f1.south) -- (f1.south|-c.north);
+      \draw [-> ] (f2.south) -- (f2.south|-c.north);
+      \draw [-> ] (f3.south) -- (f3.south|-c.north);
+
+      \node(d) at ([yshift = -1.8cm]c.south) {
+        \begin{tikzpicture}[scale=0.9,line width=0.5mm]
+          \node[align=center] at (0,0) {DSL w/ \code{(+ x 1)}:};
+          \begin{scope}[shift={(0.6,-0.5)}]
+            \node[pop3](p1) at (-1,-1) {\code{+}};
+            \node[pop3](n1) at (0.6,-0.9) {\code{1}};
+            \node[pop3](a) at (0,-1) {\code{ }};
+            \draw[pop3] (0,0) -- (p1.north);
+            \draw[pop3] (0,0) -- (n1.north);
+            \draw[pop3] (-0.3,-0.3) -- (a.north);
+          \end{scope}
+      \end{tikzpicture}};
+      \draw [-> ] (c.south) -- (d.north);
+
+      \node at (c.center) {
+        \begin{tikzpicture}[scale=0.9,line width=0.35mm]
+
+          \node(l1) at (0,0) {};
+          \node[color=pop3](p1) at (-1,-1) {\code{+}};
+          \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+          \node(x1) at (0,-1) {\code{1}};
+          \draw[color=pop3] (l1.south) -- (p1.north);
+          \draw[color=pop3] (l1.south) -- (n1.north);
+          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+          \node(t) at (-0.5,0.5) {};
+          \draw (l1.south) -- (t.south);
+          \node(c) at (-1.5,-0.2) {\code{cons}};
+          \draw (t.south) -- (c.north);
+
+          \begin{scope}[shift={(1,2)}]
+            \node(l1) at (0,0) {};
+            \node[color=pop3](p1) at (-1,-1) {\code{+}};
+            \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+            %\node(x1) at (0,-1) {};
+            \draw[color=pop3] (l1.south) -- (p1.north);
+            \draw[color=pop3] (l1.south) -- (n1.north);
+            \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+            \node(c) at (-0.5,-1.5) {\code{car}};
+            \node(z) at (0.5,-1.5) {\code{z}};
+
+            \draw (0,-1) -- (c.north);
+            \draw (0,-1) -- (z.north);
+
+            %  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+          \end{scope}
+
+        \end{tikzpicture}
+      };
+
+      \begin{scope}[shift={(4,-8)},scale=0.6,line width=0.05mm]
+        \node[latent,scale=0.7] at (3.5,3) (dx){DSL};
+        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \draw [->] (dx.south) -- (zp.north);
+        \draw [->] (zp.south) -- (xp.north);
+      \end{scope}
+
+      \end{scope}
+
+    %% center spiral
+    \begin{scope}[shift={(3.25,-7.9)},scale=0.8]    
+      \spiral{(3.5,1)}{3.5}
+      \node[latent,scale=1] at (3.5,3) (dx){DSL};
+      \node[latent,scale=1] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+      \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+      \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+      \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+      \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+      \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+      \draw [->] (dx.south) -- (zp.north);
+      \draw [->] (zp.south) -- (xp.north);
+
+      \node at ([yshift=-0.6cm]xp.south) {\legend};
+
+    \end{scope}
+
+    \end{tikzpicture}
+  \end{figure}
+
+%% \section{Introduction: Expertise}
+
+%% Human learners acquire expertise in a wide range of domains: some of
+%% us become experts in calculus, or cooking, or biology, music, tennis,
+%% or software engineering, to name just a few, and every child develops
+%% expertise in natural language, intuitive physics~\cite{Pete?},
+%% intuitive psychology (theory-of-mind), and motor control.  This
+%% picture contrasts sharply with the current state of machine
+%% intelligence, where a machine is built to be an expert in a single
+%% domain, like boardgames~\cite{alphaGo}, medical
+%% diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
+%% proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
+%% Thus an outstanding challenge in the long-term program of building
+%% more humanlike machines is to develop an algorithm that, like people,
+%% autonomously acquires expertise across many different kinds of
+%% domains.
+
+%% We contribute a model of the development of expertise that combines
+%% two key ingredients.  First, an expert needs a sufficiently expressive
+%% knowledge representation.  Following a long tradition in cognitive
+%% science and AI, we represent knowledge as programs, which prior work
+%% has used to represent expertise in recognition and generation of
+%% handwriting and speech~\cite{lake2015human}, intuitive theories (of
+%% kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
+%% grammar and
+%% semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
+%% Second, atop this program representation, experts possess two kinds of
+%% domain expertise.  They have at their disposal a powerful, yet
+%% specialized, repertoire of concepts and abstractions: e.g., in
+%% architecture, these are concepts like `arch' or `foundation'; in
+%% software engineering these are libraries of code and domain-specific
+%% languages (DSLs).  Here our model represents knowledge as programs,
+%% and so we identify this kind of expertise with a DSL.  Experts also
+%% have knowledge of when and how to deploy these domain-specific
+%% concepts efficiently when solving new problems: e.g., mathematicians
+%% efficiently search the space of proofs, intuiting which lemmas are
+%% appropriate when; expert chefs intuit which compositions of
+%% ingredients are likely tasty, before they actually start cooking.  For
+%% our model, this aspect of expertise corresponds to the ability to
+%% quickly assemble new, useful programs out of its DSL.
+
+%% We integrate these ideas into a model called DreamCoder which acquires
+%% expertise through a novel kind of wake/sleep or `dream' learning. The
+%% model iterates through wake cycles -- where it solves problems by
+%% writing programs -- and a pair of sleep cycles: a sleep cycle that
+%% grows its DSL by replaying experiences from waking and consolidating
+%% them into new abstractions, and a sleep cycle that improves its
+%% knowledge of how to write programs by training a neural network on
+%% replayed experiences as well as `dreams', or samples, from its DSL.
+
+%% DreamCoder builds on multiple generations of AI research, going back
+%% to the 1960's~\cite{solomonoff1964formal} when program-learning was
+%% proposed as a paradigm for general AI.  Broadly speaking recent work
+%% has either developed neural approaches for learning to efficiently
+%% deploy a fixed
+%% DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
+%% developed symbolic approaches for representing and searching through
+%% spaces of
+%% programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
+%% We were motivated by approaches that learn or grow the
+%% DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
+%% Our goal with DreamCoder is to show that the combination of
+%% neurally-guided search and DSL learning is a uniquely powerful way of
+%% building systems that, like human learners, autonomously acquire the
+%% expertise needed to navigate a new domain of problems.
+
+\section{Introduction}
 
@@
 
-Our system is called DreamCoder because it acquires these two kinds of
+We call our system `DreamCoder' because it acquires these two kinds of
 expertise through a novel kind of wake/sleep or `dream'
@@
 where it solves problems by writing programs -- and a pair of sleep
-cycles: a sleep cycle that grows the DSL by replaying experiences from
-waking and consolidating them into new abstractions (modeled as
-fragments of code), and a sleep cycle that improves its knowledge of
-how to write programs by training a neural network based on replayed
-experiences as well as `dreams', or samples, from the DSL.
-
-DreamCoder builds on two lines of prior work.  One line of work
+cycles, both of which are loosely biologically inspired by actual
+human sleep.  The first sleep cycle, which we refer to as
+\textbf{memory consolidation}, grows the DSL by replying experiences
+from waking and consolidating them into new code abstractions.  This
+cycle is inspired by the formation of abstractions during sleep memory
+consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
+refer to as \textbf{dreaming}, improves the agents knowledge of how to
+write programs by training a neural network on replayed experiences as
+well as `fantasies', or samples, from the DSL.  These two kinds of
+dreams are inspired by the distinct episodic replay and hallucination
+components of dream sleep~\cite{fosse2003dreaming}.
+
+This dream-learning architecture unifies two lines of prior work.
+One line of work
 considers the problem of learning new concepts, abstractions, or
@@
 like human learners, autonomously acquire the expertise needed to
-solve a new class of problems.
-
-\section{Introduction: NIPS}
-
-Much of everyday human thinking and learning can be understood in
-terms of program induction: constructing a procedure that maps inputs
-to desired outputs, based on observing example input-output pairs.
-People can induce programs flexibly across many different domains,
-often from just one or a few examples.  For instance, if shown that a
-text-editing program should map ``Jane Morris Goodall'' to
-``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey'' to
-``R. E. Leakey''; if instead the first input mapped to ``Dr. Jane'' or
-``Goodall, Jane'', we might guess the latter should map to
-``Dr. Richard'' or ``Leakey, Richard'', respectively.
-
-The FlashFill system~\cite{gulwani2011automating} embedded in
-Microsoft Excel solves problems such as these and is probably the best
-known practical program-induction algorithm, but researchers in
-programming languages and AI have had successes in many domains, such as handwriting recognition and
-generation~\cite{lake2015human}, procedural
-graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
-answering~\cite{johnson2017clevr} and robot motion
-planning~\cite{devlin2017neural}.  These systems
-work in different ways, but most hinge upon a carefully engineered
-\textbf{Domain Specific Language (DSL)}.  This is especially true for
-systems like FlashFill that induce a wide range of programs
-very quickly, in a few seconds or less.  DSLs constrain the search
-over programs with strong prior knowledge in the form of a restricted
-inventory of programming primitives finely tuned to the domain: for
-text editing, these are operations like appending and
-splitting on characters.
-
-In this work, we consider the problem of building agents that learn to solve
-program induction tasks, and also the problem of acquiring the prior knowledge
-necessary to quickly solve these tasks in a new domain.  Representative problems
-in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
-algorithm that grows or boostraps a DSL while jointly training a neural network
-to help write programs in the increasingly rich DSL.
-
-\begin{tabular}{}
-  \end{tabular}
+navigate a new domain of problems.
+
+\section{Deep Wake/Sleep Program Induction}
+
+\system~takes as its goal to
+acquire domain-specific expertise.
+It gets this expertise from
+solving a collection of
+programming \textbf{tasks} -- typically about a hundred --
+while jointly
+authoring a library of code (the DSL)
+and training a neural network
+
+%% \section{Introduction: NIPS}
+
+%% Much of everyday human thinking and learning can be understood in
+%% terms of program induction: constructing a procedure that maps inputs
+%% to desired outputs, based on observing example input-output pairs.
+%% People can induce programs flexibly across many different domains,
+%% often from just one or a few examples.  For instance, if shown that a
+%% text-editing program should map ``Jane Morris Goodall'' to
+%% ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey'' to
+%% ``R. E. Leakey''; if instead the first input mapped to ``Dr. Jane'' or
+%% ``Goodall, Jane'', we might guess the latter should map to
+%% ``Dr. Richard'' or ``Leakey, Richard'', respectively.
+
+%% The FlashFill system~\cite{gulwani2011automating} embedded in
+%% Microsoft Excel solves problems such as these and is probably the best
+%% known practical program-induction algorithm, but researchers in
+%% programming languages and AI have had successes in many domains, such as handwriting recognition and
+%% generation~\cite{lake2015human}, procedural
+%% graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
+%% answering~\cite{johnson2017clevr} and robot motion
+%% planning~\cite{devlin2017neural}.  These systems
+%% work in different ways, but most hinge upon a carefully engineered
+%% \textbf{Domain Specific Language (DSL)}.  This is especially true for
+%% systems like FlashFill that induce a wide range of programs
+%% very quickly, in a few seconds or less.  DSLs constrain the search
+%% over programs with strong prior knowledge in the form of a restricted
+%% inventory of programming primitives finely tuned to the domain: for
+%% text editing, these are operations like appending and
+%% splitting on characters.
+
+%% In this work, we consider the problem of building agents that learn to solve
+%% program induction tasks, and also the problem of acquiring the prior knowledge
+%% necessary to quickly solve these tasks in a new domain.  Representative problems
+%% in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
+%% algorithm that grows or boostraps a DSL while jointly training a neural network
+%% to help write programs in the increasingly rich DSL.
 
@@
  \begin{figure}
-   \begin{tikzpicture}
+   \begin{tikzpicture}[every node/.style={rounded corners,thick}]
      \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
@@
 \begin{figure*}
-  \centering\begin{tikzpicture}[every node/.style={inner sep=1,outer sep=0}]
-    \node(p1)[draw] at (-1,0) {
+  \centering\begin{tikzpicture}[every node/.style={inner sep=1,outer sep=0,rounded corners,thick}]
+    \node(p1)[draw,rounded corners,thick] at (-1,0) {
       \begin{tabular}{l}
@@
       Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
+      Circles:
+      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
       Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\

### 2018-12-16 spotlight.tex
3-panel figure

@@
   \end{frame}
-
-\begin{frame}{Library Learning}
+\begin{frame}{Library learning}
+
   \centering
-  \renewcommand\code\texttt
-  \renewcommand\codechar[1]{\texttt{"#1"}}
-  \newcommand{\helpSize}{0.25cm}
-\begin{tabular}{cc}
-    \toprule
-Tasks and     \pop{Programs} & DSL\\
-    \midrule
-      \begin{tabular}{cc}
+\Large
   \begin{tabular}{c}
@@
           \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
-          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\
-          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
-          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
-          \\
-          \\
-          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
-          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
+          \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} \\\\
+          \visible<3->{
+          \code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}}
   \end{tabular}
-        &
-        \hspace{-0.3cm}\begin{tabular}{c}
-          \code{[7\, 3]}$\to $\code{False}                              \\
-          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
-          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
-          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
-          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
+  \vspace{2cm}
+
+  \visible<2->{
+    \textbf{Library:}\\
+          $f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a) (if (p x) (cons x a) a)))}\\
+    \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
+    \hspace{\helpSize}(\emph{Get elements} $x$ \emph{ from $\ell$ where \code{(p $x$)} returns true})
+    }
+
+\end{frame}
+
+\begin{frame}{Subset of 38 learned library routines for list processing}
+  \Large\centering
+  \begin{tabular}{ll}
+    \begin{tabular}{l}
+            $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}
+      \end{tabular}&$f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$}\\\\
+    \begin{tabular}{l}
+      $f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}\\
+  \hspace{0.5cm}\code{(if (p x) (cons x a) a)))}
+      \end{tabular}&    
+$f_1$: \emph{Higher-order filter function}\\\\
+    \begin{tabular}{l}
+      $f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\
+\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}
       \end{tabular}
+&$f_2$: \emph{Maximum element in list $\ell$}\\\\
+    \begin{tabular}{l}
+      $f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
+      \phantom{$f_1(\ell,$}
+\code{($\lambda$ (x a) (if a a (= k x))))}
       \end{tabular}
-    &
-    \begin{tabular}{l}
-      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
-      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
-      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
-      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
-      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
-      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
-      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
-      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
-      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
-      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
-      \phantom{$f_1(\ell,$}
-      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
-      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
+&$f_2$: \emph{Whether $\ell$ contains }\code{k}
     \end{tabular}
-  \\\bottomrule
-  \end{tabular}
-  \begin{itemize}
-  \item Learned DSL primitives can call each other
-    \item Rediscovers higher-order functions like \code{filter}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}
+  \end{frame}
+
+\begin{frame}{Explore/Compress/Compile as Bayesian Inference}
+
+\centering
+
+  \begin{tikzpicture}[scale=1,line width=0.5mm]
+
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \end{tikzpicture}
+\end{frame}
+\begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}\centering
 
@@
     \end{tikzpicture}}
-
-\begin{minipage}[c]{0.6\textwidth}
-  \visible<2->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
-      \node at (2,1) {\underline{\textbf{Explore}: Infer programs, fixing DSL and neural net}};
-      \begin{scope}[shift = {(-3,0)}]
-    \node at (0,0) (d){DSL};
-    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
-
-    \node at ([xshift = 2cm]t) (nn){
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+\begin{tikzpicture}[scale=1,line width=0.5mm]
+
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}{\underline{\textbf{Explore}}/Compress/Compile as {\color{red}Amortized} Bayesian Inference}\centering
+  \underline{\textbf{Explore}: Infer programs, fixing DSL and neural recognition model}\vspace{1cm}
+
+\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
       \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
@@
               \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
+\begin{tikzpicture}[scale=1,line width=0.5mm]
+
+  \node[obs,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+  \end{frame}
+
+\begin{frame}{\underline{\textbf{Explore}}/Compress/Compile as {\color{red}Amortized} Bayesian Inference}\centering
+  \underline{\textbf{Explore}: Infer programs, fixing DSL and neural recognition model}%
+
+%  \vspace{1cm}
+
+    \begin{tikzpicture}[scale=1.2,line width=0.5mm]
+
+      \begin{scope}[shift = {(-3,0)}]
+        \node[align=right] at (0,0) (d){\textbf{DSL}\\
+        $f_1$, $f_2$, ...};
+    \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
+      \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+      \footnotesize    \code{[1\, 2\, 3\, 4]}$\to$\code{[3\, 4]} \\
+      \footnotesize    \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} };
+
+    \node at ([xshift = 1.3cm]t.east) (nn){
+      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+        \node[rectangle] at (1,1) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (I-\name) at (0,-\y) {};
+        \foreach \name / \y in {1,...,3}
+            \node[neuron] (H-\name) at (1,-\y-0.5) {};
+        \foreach \name / \y in {1,...,4}
+            \node[neuron] (O-\name) at (2,-\y) {};
+        \foreach \source in {1,...,4}
+            \foreach \dest in {1,...,3}
+                \draw [-latex] (I-\source) -- (H-\dest);
+        \foreach \source in {1,...,3}
+            \foreach \dest in {1,...,4}
+                \draw [-latex] (H-\source) -- (O-\dest);
       \end{tikzpicture}
     };
-    \node[align = center, text width = 1cm] at ([yshift = 0.9cm,xshift=-0.5cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
+    \node[align = center, text width = 1cm] at ([yshift = 0.5cm,xshift=0cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
     \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
 
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+    \node[draw,rounded corners, align=center, inner sep = 10] at ([xshift = 4.2cm,yshift = 1cm]t.east) (s){Neurally-Guided\\ Enumerative Search};
 
@@
 
-    \node[align=left] at ([xshift=2cm]s.east) (f) {Programs};
+    \node[align=left] at ([xshift=3cm]s.east) (f) {\textbf{Programs:}\\
+\small    \code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}};
     \draw [->  ] (s.east) -- (f.west);
@@
       \end{scope}
-      \visible<5->{
-        \node(repeat)[scale=8] at (6,-3.5) {$\circlearrowleft$};
-        \node at (repeat) {\emph{Repeat!}};
-      }
+
+      \node at ([yshift=-2.8cm,xshift=-0.5cm]f.south) {
+  \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
     \end{tikzpicture}}
-
-  \vspace{-1.5cm}
-
-\visible<3->{  \begin{tikzpicture}[scale=0.4,line width=0.5mm]
+\begin{tikzpicture}[scale=0.8,line width=0.25mm]
+
+  \node[obs,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.12}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+      };
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}{Explore/\underline{\textbf{Compress}}/Compile as {\color{red}Amortized} Bayesian Inference}\centering
+  \underline{\textbf{Compress}: Update DSL, fixing programs}
+\vspace{1cm}
+
+\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
+\begin{tikzpicture}[scale=1,line width=0.5mm]
+
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[obs,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[obs,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[obs,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.17}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+  \end{frame}
+
+\begin{frame}{Explore/\underline{\textbf{Compress}}/Compile as {\color{red}Amortized} Bayesian Inference}\centering
+  \underline{\textbf{Compress}: Update DSL, fixing programs}
+
+  \begin{tikzpicture}[scale=0.4,line width=0.5mm]
     \node at (0,0) (f1){progs. for task$_1$};
@@
 
-    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 2.5cm, anchor = north west] at (5.2,1) {};
-%    \node[anchor=north] at (c.north) {Compression};
-    \node at ([yshift=1cm,xshift=-3cm]c.north) {\underline{\textbf{Compress}: Update DSL, fixing programs}};
+    \node(c)[rectangle, rounded corners, draw, minimum width = 8cm, minimum height = 3cm, anchor = north west] at (5.2,1) {};
     \draw [-> ] (f1.east) -- (c.west|-f1.east);
@@
     \node[right](d) at ([xshift = 1.6cm,yshift = -0.0cm]c.east) {
-      \begin{tikzpicture}[scale=0.6]
+      \begin{tikzpicture}[scale=0.9,line width=0.5mm]
         \node[align=center] at (0,0) {DSL+};
@@
         \end{scope}
-        \end{tikzpicture}
-      };
+    \end{tikzpicture}};
     \draw [-> ] (c.east) -- (d.west);
@@
     \node at (c.center) {
-\begin{tikzpicture}[scale=0.6]
+\begin{tikzpicture}[scale=1.2,line width=0.5mm]
 
@@
 
-  \begin{scope}[shift={(0.9,-1.3)}]
+  \begin{scope}[shift={(3,0.5)}]
       \node(l1) at (0,0) {};
@@
 
+    \node at ([yshift=-5.8cm,xshift=-0.5cm]d.south) {
+  \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
     \end{tikzpicture}}
-
-\end{minipage}%
-\begin{minipage}[c]{0.28\textwidth}
-  \vspace{-1cm}
+\begin{tikzpicture}[scale=0.8,line width=0.25mm]
+
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[obs,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[obs,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[obs,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.12}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+      };
+
+  \end{tikzpicture}
+
+\end{frame}
+\begin{frame}{Explore/Compress/\underline{\textbf{Compile}} as {\color{red}Amortized} Bayesian Inference}\centering
+\underline{\textbf{Compile}: Train {\color{red} recognition model}}
+\vspace{1cm}
+
+\newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
 \begin{tikzpicture}[scale=1,line width=0.5mm]
 
-  \node[latent,scale=1] at (3.5,3) (dx){DSL};
-  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at (3.5,3) (dx){DSL};
+  \node[obs,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
   \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
-  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]zp) (zp1){prog};
   \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
@@
 
-  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
   \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
@@
   \end{tikzpicture}
-
-  \vspace{0.5cm}
-
-\visible<4->{  \begin{tikzpicture}[scale=0.8,line width=0.5mm]
-    \node at (0,0) {\underline{\textbf{Compile}: Train {\color{red} recognition model}}};
-          \node[align=center] at (-3,-1) (d){DSL};
+  \end{frame}
+\begin{frame}{Explore/Compress/\underline{\textbf{Compile}} as {\color{red}Amortized} Bayesian Inference}\centering
+\underline{\textbf{Compile}: Train {\color{red} recognition model}}
+\vspace{1cm}
+
+  \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
+\hspace{1.5cm}\begin{tikzpicture}[scale=1.2,line width=0.5mm]
+            \node[align=center] at (-3,-1) (d){\Large DSL};
       %    \node at ([xshift = 3cm]d.east) (p2){program};
-      \node at ([xshift=2.5cm,yshift = 0cm]d.east) (p1){program};
-
-      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\footnotesize  sample} (p1.west);
-
-      \node at ([xshift = 2.5cm]p1.east) (t1){ task};
-
-      \draw [-> ] (p1.east) -- node[above]{\footnotesize  run} (t1.west);
+      \node at ([xshift=2.5cm,yshift = 0cm]d.east) (p1){\Large  program};
+
+      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{sample} (p1.west);
+
+      \node at ([xshift = 2.5cm]p1.east) (t1){ \Large  task};
+
+      \draw [-> ] (p1.east) -- node[above]{ run} (t1.west);
 
@@
       \draw [->,red] ([xshift=-0.4cm]n.west) to[out = -180,in = -70]  ([xshift=-0.25cm,yshift=0.1cm]p1.south);
-%      \node at ([xshift=-0.5cm]$(p1.south)!0.5!(n.west)$) {\textbf{Loss4%}};
-
+
+    \node at ([yshift=-3cm,xshift=2cm]t1.south) {
+  \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
     \end{tikzpicture}}
-\end{minipage}
-
+\begin{tikzpicture}[scale=0.8,line width=0.25mm]
+
+  \node[obs,scale=1] at (3.5,3) (dx){DSL};
+  \node[obs,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[obs,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[obs,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.12}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+      };
+
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}{Explore/Compress/Compile as {\color{red}Amortized} Bayesian Inference}
+
+  \begin{tikzpicture}[line width=0.5mm]
+    \node at (0,0) (e){
+  \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
+\begin{tikzpicture}[scale=0.7,line width=0.25mm]
+
+  \node[obs,scale=1] at (3.5,3) (dx){DSL};
+  \node[latent,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \draw [->,red] ([xshift=2cm]dx.east) -- ([xshift=3cm]dx.east);
+  \node at ([xshift=3.2cm]dx.east]) (uses){is};k
+  \node at ([xshift=0.5cm]uses.east) (nnn){\NeuralNetwork{0.12}};
+  \draw[thin] ([xshift=1.9cm,yshift=0.4cm]dx.east) rectangle
+  ([xshift=0.5cm,yshift=-0.3cm]nnn.south);
+  \end{tikzpicture}
+      };
+
+    \node at ([xshift=5cm,yshift=-2cm]e.south) (cc){
+  \newcommand{\NeuralNetwork}[1]{    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=#1,shift={+(-1,2.5)}}]
+      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+      \node[rectangle] at (1,1) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (I-\name) at (0,-\y) {};
+      \foreach \name / \y in {1,...,3}
+          \node[neuron] (H-\name) at (1,-\y-0.5) {};
+      \foreach \name / \y in {1,...,4}
+          \node[neuron] (O-\name) at (2,-\y) {};
+      \foreach \source in {1,...,4}
+          \foreach \dest in {1,...,3}
+              \draw [-latex] (I-\source) -- (H-\dest);
+      \foreach \source in {1,...,3}
+          \foreach \dest in {1,...,4}
+              \draw [-latex] (H-\source) -- (O-\dest);
+    \end{tikzpicture}}
+\begin{tikzpicture}[scale=0.8,line width=0.25mm]
+
+  \node[obs,scale=1] at (3.5,3) (dx){DSL};
+  \node[obs,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[obs,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[obs,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \end{tikzpicture}
+    };
+      \node at ([xshift=-5cm,yshift=-2cm]e.south) (c){
+\begin{tikzpicture}[scale=0.8,line width=0.25mm]
+
+  \node[latent,scale=1] at (3.5,3) (dx){DSL};
+  \node[obs,scale=1] at ([yshift=-1.5cm,xshift=2cm]dx) (zp){prog};
+  \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+  \node[obs,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \node[obs,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+  \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+  \draw [->] (zp1.south) -- (xp1.north);
+  \draw [->] (dx.south) -- (zp1.north);
+  \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+  \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+  \draw [->] (dx.south) -- (zp.north);
+  \draw [->] (zp.south) -- (xp.north);
+
+  \end{tikzpicture}
+      };
+      \node at (e.north) {\textbf{\underline{Explore}}};
+      \node at ([yshift=-0.25cm]cc.south) {\textbf{\underline{Compile}}};
+      \node at ([yshift=-0.25cm]c.south) {\textbf{\underline{Compress}}};
+      \draw[->] (c.east) -- (cc.west);
+      \draw[->] (e.west) -- (c.north);
+      \draw[->] (cc.north) -- (e.east);
+    \end{tikzpicture}
   \end{frame}
@@
 
-\begin{frame}{Domain: List processing}
-      Starts with: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
-\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
-\code{car}, \code{cdr}, \code{nil}, \code{is-nil},  \code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
-\\236 human-interpretable list processing tasks.
-\\Discovers 38 new DSL primitives, including \code{filter}
-
-\vspace{0.25cm}
-
-\centering \begin{minipage}[c]{0.6\textwidth}
-  \begin{tabular}{lll}
-    \toprule
-    Name & Input & Output \\\midrule
-    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
-    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
-    product & [7\, 1\, 6\, 2] & 84 \\
-    \bottomrule
-  \end{tabular}
-\end{minipage}%
-%% \begin{minipage}[c]{0.3\textwidth}
-%%   \includegraphics[width = \textwidth]{figures/listLearningCurve_color.eps}
-%% {   \color{orange}orange: } full system. {\color{teal}teal: } no neural recognition model
-%% \end{minipage}
-
-\vspace{0.25cm}
-
-\centering \textbf{With functional programming ``problem set'' + 93 hours on 64 CPUs, rediscovers:}\\
-\textbf{\code{map}, \code{foldr}, \code{unfold}, \code{range}, \code{length}, \code{index}, \code{zip}}, +some arithmetic routines
-
-  \end{frame}
+%% \begin{frame}{Domain: List processing}
+%%       Starts with: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
+%% \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
+%% \code{car}, \code{cdr}, \code{nil}, \code{is-nil},  \code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.
+%% \\236 human-interpretable list processing tasks.
+%% \\Discovers 38 new DSL primitives, including \code{filter}
+
+%% \vspace{0.25cm}
+
+%% \centering \begin{minipage}[c]{0.6\textwidth}
+%%   \begin{tabular}{lll}
+%%     \toprule
+%%     Name & Input & Output \\\midrule
+%%     repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
+%%     drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+%%     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
+%%     count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+%%     keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
+%%     product & [7\, 1\, 6\, 2] & 84 \\
+%%     \bottomrule
+%%   \end{tabular}
+%% \end{minipage}%
+%% %% \begin{minipage}[c]{0.3\textwidth}
+%% %%   \includegraphics[width = \textwidth]{figures/listLearningCurve_color.eps}
+%% %% {   \color{orange}orange: } full system. {\color{teal}teal: } no neural recognition model
+%% %% \end{minipage}
+
+%% \vspace{0.25cm}
+
+%% \centering \textbf{With functional programming ``problem set'' + 93 hours on 64 CPUs, rediscovers:}\\
+%% \textbf{\code{map}, \code{foldr}, \code{unfold}, \code{range}, \code{length}, \code{index}, \code{zip}}, +some arithmetic routines
+
+%%   \end{frame}
 
 \begin{frame}{Domain: Text Editing}
-  In the style of FlashFill (Gulwani 2012). Starts with: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
-\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
-\code{car}, \code{cdr}, \code{nil}, \code{is-nil}
-, plus string \& character constants.\\
-
-\begin{minipage}[c]{0.4\textwidth}
+  In the style of FlashFill (Gulwani 2012). Starts with \code{map}, \code{fold}, etc.
+
+  \centering
+\vspace{1cm}
   \begin{tabular}{ll}
+
     \toprule
@@
   \end{tabular}
-  %% \begin{figure}
-  %% \includegraphics[width = 0.8\textwidth]{figures/textLearningCurve.eps}
-  %%   \end{figure}
-  \end{minipage}
-\begin{minipage}[c]{0.5\textwidth}
-      \begin{figure}        
-        \includegraphics[width = \textwidth]{figures/textPrimitives.png}\vspace{-0.8cm}
-      \caption{11 learned DSL primitives over 3 successive iterations (3 columns). Learned primitives call each other (arrows).}
-      \end{figure}  
-  \end{minipage}
-\end{frame}
-
-\begin{frame}{Domain: Symbolic regression from visual input}
-  Starts with: \code{plus}, \code{times}, \code{divide}, \code{real-number}. Autograds through program
-  \\  Bayesian likelihood $\probability[\text{data}|\text{prog}]$ favors fewer continuous parameters
+\end{frame}
+\begin{frame}{Text editing: Library learning builds on itself}
+\centering        \includegraphics[width=0.6\textwidth]{figures/textPrimitives.png}
+\\      Learned DSL primitives over 3 iterations (3 columns). Learned primitives call each other (arrows).
+
+  \end{frame}
+\begin{frame}{Programs with numerical parameters: Symbolic regression from visual input}
+Fits parameters by autograd-ing thru program
+\\
+\\
+\textbf{Recognition model looks at picture of function's graph}
 
@@
 \begin{frame}{New domain: Generative  graphics  programs (Turtle/LOGO)}
-  \begin{tabular}{cc|c}
-    Training tasks& DSL samples before learning&DSL samples after learning\\
-              \raisebox{-.5\height}{\includegraphics[width = 0.2\textwidth]{figures/logo90.png}$\quad\quad$ }&
-      \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/initialDreams/montage.png}}&
-  \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/finalDreams/montage.png}}
+\centering  Training tasks:
+
+\includegraphics[width = 0.4\textwidth]{figures/logo90.png}
+\end{frame}
+\begin{frame}{Generative graphics programs: Samples from DSL}\centering
+  \begin{tabular}{c|c}
+     DSL samples before learning&DSL samples after learning\\
+          \raisebox{-.5\height}{\includegraphics[width = 0.45\textwidth]{figures/initialDreams/montage.png}}&
+  \raisebox{-.5\height}{\includegraphics[width = 0.45\textwidth]{figures/finalDreams/montage.png}}
   \end{tabular}
-
-  \vspace{0.3cm}
-
-  \pause
-
-  Learned DSL contains parametric drawing routines:
-      \begin{tabular}{rlrl}
+  \end{frame}
+
+\begin{frame}{Generative graphics programs: Learned library contains parametric drawing routines}
+\centering      \begin{tabular}{rlrl}
       Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}&

### 2018-12-16 archive.tex
completeness and consistency theorems are complete and consistent

@@
   \end{tikzpicture}
+  \caption{}\label{threeCycles}
   \end{figure}
@@
 \system~takes as its goal to
-acquire domain-specific expertise.
-It gets this expertise from
+acquire domain-specific expertise,
+which it learns by
 solving a collection of
-programming \textbf{tasks} -- typically about a hundred --
-while jointly
-authoring a library of code (the DSL)
-and training a neural network
+programming \textbf{tasks} -- typically about a hundred.
+During its Wake phase,
+it searches for programs solving tasks,
+guided by its current DSL and recognition model.
+It cycles through Wake
 
@@
   \begin{cases}
-    \text{if $u$ is a primitive or index or $\varnothing $ or $\Lambda$:}&\varnothing\\
-    \text{if $u = \lambda b$:}&\lambda I\beta'(b)\\
-    \text{if $u = (f\;x)$:}&(I\beta'(f)\;I\beta'(x))\\
-    \text{if $u = \uplus V$:}&\uplus\left\{I\beta'(u')\;|\;u'\in V \right\}
+    \text{if $u$ is a primitive or index or $\varnothing $:}&\varnothing\\
+    \text{if $u$ is $\Lambda$:}&\left\{\Lambda \right\}\\
+    \text{if $u = \lambda b$:}&\left\{\lambda I\beta'(b) \right\}\\
+    \text{if $u = (f\;x)$:}&\left\{(I\beta'(f)\;x),\;(f\;I\beta'(x)) \right\}\\
+    \text{if $u = \uplus V$:}&\left\{I\beta'(u')\;|\;u'\in V \right\}
   \end{cases}  
@@
 
-From these results the consistency and completeness of $I\beta$ follows immediately by its definition.
+From these results the consistency and completeness of $I\beta$ follows:
+\begin{theorem}
+  \textbf{Consistency of $I\beta'$.}\\
+  If $p\in \denotation{I\beta'(u)}$ then there exists $p'\in \denotation{u}$ such that $p\reduce p'$.
+\end{theorem}
+\begin{proof}
+  Proceed by structural induction on $u$. If $p\in \denotation{I\beta'(u)}$ then, from the definition of $I\beta'$ and $\denotation{\cdot }$, at least one of the following holds:
+  \begin{itemize}
+  \item Case $p = (\lambda b')v'$ where $v'\in v$, $b'\in b$, and $v\mapsto b\in S_0(u)$:
+    From the definition of $\beta$-reduction we know that $p\reduce \shift{-1}[\$0\mapsto \shift{1}v']b'$.
+    From the consistency of $S_n$ we know that $\shift{-1}[\$0\mapsto \shift{1}v']b'\in u$.
+    Identify $p' = \shift{-1}[\$0\mapsto \shift{1}v']b'$.
+  \item Case $u = \lambda b$ and $p = \lambda b'$
+    where $b'\in \denotation{I\beta'(b)}$:
+    By induction there exists $b''\in \denotation{b}$ such that $b'\reduce b''$.
+    So $p \reduce \lambda b''$.
+    But $\lambda b''\in \denotation{\lambda b} = \denotation{u}$,
+    so identify $p' = \lambda b''$.
+  \item Case $u = (f\; x)$ and $p = (f'\; x')$ where $f'\in \denotation{I\beta'(f)}$ and $x'\in \denotation{x}$:
+    By induction there exists $f''\in \denotation{f}$
+    such that $f'\reduce f''$.
+    So $(f'\;x')\reduce (f''\;x')$.
+    But $(f''\;x')\in \denotation{(f\;x)} = \denotation{u}$,
+    so identify $p' = (f''\;x')$.
+  \item  Case $u = (f\; x)$ and $p = (f'\; x')$ where $x'\in \denotation{I\beta'(x)}$ and $f'\in \denotation{f}$: Symmetric to the previous case.
+  \item Case $u = \uplus U$ and $p \in \denotation{I\beta'(u')}$ where $u'\in U$:
+    By induction there is a $p'\in \denotation{u'}$ satisfying $p'\reduce p$.
+    But $\denotation{u'}\subseteq \denotation{u}$, so also $p'\in \denotation{u}$.
+  \item Case $u$ is an index, primitive, $\varnothing$, or $\Lambda$: The theorem holds vacuously.    
+    \end{itemize}  
+\end{proof}
+\begin{theorem}
+  \textbf{Completeness of $I\beta'$.}\\
+  Let $p\reduce p'$ and $p'\in \denotation{u}$.
+  Then $p\in \denotation{I\beta'(u)}$.
+\end{theorem}
+\begin{proof}
+  Structural induction on $u$.
+  If $u = \uplus V$
+  then there is a $v\in V$
+      such that $p' \in \denotation{v}$;
+      by induction on $v$ combined with the definition of $I\beta'$ we have $p\in \denotation{I\beta'(v)}\subseteq \denotation{I\beta'(u)}$,
+      which is what we were to show.
+      Otherwise assume that $u\not=\uplus V$.
+
+  From the definition of $p\reduce p'$
+  at least one of these cases must hold:
+  \begin{itemize}
+  \item Case $p = (\lambda b')v'$ and $p' = \shift{-1}[\$0\mapsto \shift{1}v']b'$:
+    Using the fact that $\shift{-1}[\$0\mapsto \shift{1}v']b'\in \denotation{u}$, we can invoke the completeness of $S_n$ to construct a $(v\mapsto  b)\in S_0(u)$
+    such that $v'\in \denotation{v}$
+    and $b'\in \denotation{b}$.
+    Combine these facts with the definition of $I\beta'$
+    to get
+    $p = (\lambda b')v'\in \denotation{(\lambda b)v}\subseteq I\beta'(u)$.
+  \item Case $p = \lambda b$ and $p' = \lambda b'$ where
+    $b\reduce b'$:
+    Because $p' = \lambda b'\in \denotation{u}$
+    and by assumption $u\not=\uplus V$,
+    we know that $u = \lambda v$ and $b'\in \denotation{v}$. By
+    induction $b\in \denotation{I\beta'(v)}$.
+    Combine with the definition of $I\beta'$ to get $p = \lambda b \in \denotation{\lambda I\beta'(v)}\subseteq \denotation{I\beta'(u)}$.
+  \item Case $p = (f\;x)$ and $p' = (f'\;x)$ where $f\reduce f'$:
+    Because $p' = (f'\;x)\in \denotation{u}$
+    and by assumption $u\not=\uplus V$
+    we know that $u = (a\;b)$
+    where $f'\in \denotation{a}$
+    and $x\in \denotation{b}$.
+    By induction on $a$
+    we know
+    $f\in \denotation{I\beta'(a)}$.
+    Therefore
+    $p = (f\;x)\in \denotation{(I\beta'(a)\; b)}\subseteq\denotation{I\beta'((a\;b))}\subseteq\denotation{I\beta'(u)}$.
+  \item Case $p = (f\;x)$ and $p' = (f\;x')$ where $x\reduce x'$: Symmetric to the previous case.    
+  \end{itemize}
+\end{proof}
+Finally we have our main result:
+\begin{theorem}
+  \textbf{Consistency and completeness of $I\beta_n$.}
+  Let $p$ and $p'$
+  be programs.
+  Then $p\underbrace{\reduce q\reduce\cdots\reduce}_{\text{$\leq n$ times}} p'$  if and only
+  if $p\in \denotation{I\beta_n(p')}$.
+\end{theorem}
+\begin{proof}
+  Induction on $n$.
+
+  If $n = 0$
+  then $\denotation{I\beta_n(p')} = \left\{p \right\}$
+  and $p = p'$;
+  the theorem holds immediately. Assume $n > 0$.
+
+  If $p\underbrace{\reduce q\reduce\cdots\reduce}_{\text{$\leq n$ times}} p'$
+  then $q\underbrace{\reduce\cdots\reduce}_{\text{$\leq n - 1$ times}} p'$;
+  induction on $n$
+  gives $q\in \denotation{I\beta_{n - 1}(p')}$.
+  Combined with $p\reduce q$ we can invoke the completeness of $I\beta'$
+  to get $p\in \denotation{I\beta'(I\beta_{n - 1}(p'))}\subset\denotation{I\beta_n(p')}$.
+
+  If $p\in \denotation{I\beta_n(p')}$
+  then
+  there exists a $i\leq n$
+  such that $p\in \denotation{\underbrace{I\beta'(I\beta'(I\beta'(\cdots}_{i \text{ times}} p')))}$.
+  If $i = 0$
+  then $p = p'$ and $p$ reduces to $p'$ in $0\leq n$ steps.
+  Otherwise $i > 0$
+  and $p\in \denotation{I\beta'(\underbrace{I\beta'(I\beta'(\cdots}_{i - 1 \text{ times}} p'))}$.
+  Invoking the
+  consistency of $I\beta'$
+  we know that $p\reduce q$
+  for a program $q\in \denotation{\underbrace{I\beta'(I\beta'(\cdots}_{i - 1 \text{ times}} p'))}\subseteq\denotation{I\beta_{i - 1}(p')}$.
+  By induction $q\underbrace{\reduce\cdots\reduce}_{\text{$\leq i - 1$ times}} p'$,
+  which combined with $p\reduce q$
+  gives  $p\underbrace{\reduce q\reduce\cdots\reduce}_{\text{$\leq i\leq n$ times}} p'$.  
+\end{proof}
 
@@
 update.  In practice this quadratic dependence on $P$ (the number of
-programs) is prohibitively slow.
-We now describe a linear time approximation to the
-refactor step in Algorithm~\ref{grammarInductionAlgorithm} based on beam search.
+programs) is prohibitively slow.  We now describe a linear time
+approximation to the refactor step in
+Algorithm~\ref{grammarInductionAlgorithm} based on beam search.
 

### 2018-12-17 archive.tex
progress on DSL figure

@@
 \newcommand{\scode}[1]{{\tiny\texttt{#1}}}
+\newcommand{\mcode}[1]{{\scriptsize\texttt{#1}}}
 \newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
@@
 \usepackage{circuitikz}
-\usetikzlibrary{fit,bayesnet,calc}
+\usetikzlibrary{fit,bayesnet,calc,tikzmark}
 \usetikzlibrary{arrows.meta}
@@
   \begin{tikzpicture}[line width=0.4mm]
-    \draw (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
+    \draw[fill=white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
     \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
@@
 
-    \node at (0.5,-4.75) {\textbf{\textsc{Memory Consolidation}}};
+    \node at (0.5,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
     \draw (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
-    \node at (11.5,-4.75) {\textbf{\textsc{Dreaming}}};
+    \node at (11.5,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
     \draw (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
@@
       \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
-      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
+      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
 
@@
     \begin{scope}[shift={(-2,-4.5)}]
-      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\scode{(+ (car z) 1)}};
-      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\scode{(cons (+ 1 1))}};
-      \node[align=center] at ([yshift = 0cm,xshift=1cm]f2.east) (f3){\textbf{DSL}\\\scode{cons}, \scode{+}, \\...};
-      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm]f2.south) {};
+      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\mcode{(+ (car z) 1)}};
+      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\mcode{(cons (+ 1 1))}};
+      \node[align=center] at ([yshift = 0cm,xshift=0.4cm]f2.east) (f3){\textbf{DSL}\\\mcode{cons}, \mcode{+}, \\...};
+      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm,xshift=-0.3cm]f2.south) {};
       \draw [-> ] (f1.south) -- (f1.south|-c.north);
@@
   \caption{}\label{threeCycles}
+  \end{figure}
+
+\begin{figure}
+  \begin{tikzpicture}
+%    \draw (0,0) rectangle (10,5);
+    \node(t1)[align=left] at (1,-1) {
+      \small +106 769-43$\to $106.769.43\\
+      \small +83 973-831$\to $83.973.831\\
+      \small $f(\text{\code{s}}) = $\code{(}\subnode{f01}{$f_0$}\code{  \codechar{.} \codechar{-}}\\
+      \small \hspace{1.25cm}\code{($f_0$ \codechar{.} \codechar{ }}\\
+      \small \hspace{1.4cm}\code{(cdr s)))}
+    };
+    \node(t2)[align=left] at ([xshift=2cm]t1.east) {
+      \small Dream Coder$\to$Dr. Dream\\
+      \small Andy Cencici$\to$Dr. Andy\\
+      \small Jan Kotas$\to$Dr. Jan\\
+      \small Nancy Jones$\to$Dr. Nancy\\
+      $f(\text{\code{s}}) = $\code{($f_3$ ($f_4$ \codechar{ } s))}
+    };
+    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
+      \small Dream (Coder)$\to$Coder\\
+      \small 1 (2) 3 4$\to$2\\
+      \small the (dog) ran$\to$ran\\
+      \small +817(33)99$\to$33\\
+      $f(\text{\code{s}}) = $\code{($f_6$ \codechar{(} \codechar{)} s)}
+    };
+
+    \node(substitution)[draw,rounded corners,align=center] at (1,-4) {
+      \small $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\\
+       \mcode{(map ($\lambda$ (x) (if}\\
+      \mcode{(= x a) b x)) s)}
+    };
+    \node[anchor = north] at (substitution.south) {substitute characters};
+    \node(cat)[draw,rounded corners,align=center] at (1,-6) {
+      \small $f_1($\code{a}$,$\code{b}$) \,=\, $\\
+       \mcode{(fold ($\lambda$ (x y)}\\\mcode{ (cons x y)) a b))}
+    };
+    \node[anchor = north] at (cat.south) {concat strings};
+    \node(looping)[draw,rounded corners,align=center] at (1,-8) {
+      \small $f_2($\code{c}$,$\code{p}$,$\code{s}$) \,=\, $\\
+       \mcode{(unfold l ($\lambda$ (x)}\\\mcode{(= c (car x)))}\\\mcode{ p ($\lambda$ (z) (cdr z)))}
+    };
+    \node[anchor = north] at (looping.south) {loop until char};
+
+    \node(constant)[draw,rounded corners,align=center] at (4.5,-4) {
+      \small $f_3($\code{s}$) \,=\, $\\
+       \mcode{($f_0$ s string)}
+    };
+    \node[anchor = north] at (constant.south) {prepend constant};
+    \node(drop)[draw,rounded corners,align=center] at (4.5,-6) {
+      \small $f_4($\code{c}$) \,=\, $\\
+      \mcode{($f_2$ c ($\lambda$ }\\\mcode{(z) (car z)))}
+    };
+    \node[anchor = north] at (drop.south) {prefix};
+    \node(suffix)[draw,rounded corners,align=center] at ([yshift=-2cm]drop.center) {
+      \small $f_5($\code{s}$,$\code{c}$) \,=\, $\\
+      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
+    };
+    \node[anchor = north] at (suffix.south) {suffix};
+    \node(delimited)[draw, rounded corners, align=center,text width=2.5cm] at ([xshift=3.5cm]constant.center) {
+      \code{$f_6($x$,$y$,$s$)\,=\,$}\\
+      \mcode{($f_4$ x ($f_5$ y c))}};
+    \node[anchor = north] at (delimited.south) {delimited substring};
+    \draw[->] (delimited.west) -- (drop.east);
+    \draw[->] (delimited.west) -- (suffix.east);
+
+    \draw[->] (drop.west) -- (looping.east);
+    \draw[->] (constant.west) -- (cat.east);
+
+    \draw[->] ([xshift=-0.5cm]t2.south) -- (constant.north);
+    \draw[->] ([xshift=0.3cm]t2.south) to[in=30,out=-30] ([xshift=1cm]drop.north);
+
+    \draw[->] ([yshift=0.5cm,xshift=-0.4cm]t1.south) to[in=100,out=200] (substitution.north);
+    \draw[->] ([yshift=0.8cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
+    \draw[->] ([xshift=-0.5cm]t3.south) -- (delimited.north);
+
+    \end{tikzpicture}
   \end{figure}
@@
 
-This dream-learning architecture unifies two lines of prior work.
+This dream-learning architecture brings together two lines of prior work.
 One line of work
@@
 
-\system~takes as its goal to
-acquire domain-specific expertise,
-which it learns by
-solving a collection of
-programming \textbf{tasks} -- typically about a hundred.
-During its Wake phase,
-it searches for programs solving tasks,
-guided by its current DSL and recognition model.
-It cycles through Wake
+\system~takes as its goal to acquire domain-specific expertise, which
+it learns by solving a collection of programming \textbf{tasks} --
+typically about a hundred.  It alternatingly finds programs that solve
+tasks (Wake -- Figure~\ref{threeCycles} top); improving its DSL by
+discovering and reusing domain-specific subroutines (Memory
+Consolidation -- Figure~\ref{threeCycles} left); and training a neural
+network that efficiently guides search for programs in the DSL
+(Dreaming -- Figure~\ref{threeCycles} right).  The learned DSL
+effectively encodes a prior on programs likely to solve tasks in the
+domain, while the neural net looks at the example input-output pairs
+for a specific task and produces a ``posterior'' for programs likely
+to solve that specific task (Figure~\ref{threeCycles} middle).  The
+neural network thus functions as a \textbf{recognition model}
+supporting a form of approximate Bayesian program induction, jointly
+trained with a \textbf{generative model} for programs encoded in the
+DSL, in the spirit of the Helmholtz machine~\cite{hinton1995wake}. The
+recognition model ensures that searching for programs remains
+tractable even as the DSL (and hence the search space for programs)
+expands.
 
@@
 \begin{table*}%[t!]
-  \makebox[\textwidth][c]{
-    \scriptsize
-  \tabcolsep=4pt
-  \renewcommand\code\texttt
-  \renewcommand\codechar[1]{\texttt{"#1"}}
+  \makebox[\textwidth][c]{ \scriptsize \tabcolsep=4pt
+    \renewcommand\code\texttt \renewcommand\codechar[1]{\texttt{"#1"}}
     \newcommand{\helpSize}{0.25cm}
@@
   \begin{align*}
-    \denotation{\$i}& = \left\{\$i \right\}\\
-    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}\\
+    \denotation{\$i}& = \left\{\$i \right\}&
+    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}&
     \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
-    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}\\
-    \denotation{\varnothing}& = \varnothing\\
+    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}&
+    \denotation{\varnothing}& = \varnothing&
     \denotation{\Lambda}& = \Lambda
@@
 Version spaces also support efficient membership checking, which we
-write as $e\in v$, and which is equivalent to $e\in \denotation{v}$.
+write as  $e\in \denotation{v}$.
 Important for our purposes, it is also efficient to refactor the
@@
   \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
-    e\text{, if $e\in \mathcal{D}$ and $e\in v$. Exploits the fact that $e\in v$ can be efficiently computed.}\\
+    e\text{, if $e\in \mathcal{D}$ and $e\in \denotation{v}$. Exploits the fact that $\denotation{e}\in v$ can be efficiently computed.}\\
     \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
   \end{cases}\label{FACTORING}\\\nonumber    
-  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\\nonumber
+  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber
   \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
-  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\\nonumber
+  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})&%\\\nonumber
   \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber

### 2018-12-18 archive.tex
shorten some equations (vs section)

@@
   \begin{tikzpicture}[line width=0.4mm]
-    \draw[fill=white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
+    \draw[fill=teal!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
     \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
@@
     \node at (0.5,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
-    \draw (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
+    \draw[fill=teal!5!white] (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
     \node at (11.5,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
-    \draw (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
+    \draw[fill=teal!5!white] (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
 
@@
 
+    \draw (-2,-2.5) -- (15,-2.5);
+    \node[align=left] at (-1.5,-2.3) {Tasks and\\\texttt{Programs}\\\\DSL};
+
     \node(substitution)[draw,rounded corners,align=center] at (1,-4) {
@@
     \draw[->] (delimited.west) -- (suffix.east);
+
+    \node(appendSingle)[draw, rounded corners, align=center,text width=2.5cm] at ([yshift=-2cm]delimited.center) {
+      \code{$f_7($c$,$s$)\,=\,$}\\
+      \mcode{($f_1$ (cons c empty) s)}};
+    \node[anchor = north] at (appendSingle.south) {append character};
+
+    \node(surround)[draw, rounded corners, align=center,text width=2.5cm] at ([yshift=-2cm]appendSingle.center) {
+      \code{$f_8($s$)\,=\,$}\\
+      \mcode{(cons \codechar{(} ($f_7$ \codechar{)} s))}};
+    \node[anchor = north] at (surround.south) {surround in parentheses};
+    \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
+    \draw[->] (appendSingle.west) to[in=40,out=120] ([yshift=0.25cm]cat.east);
 
@@
 We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
-\begin{align}
+\begin{align*}
   \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
@@
     \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
-  \end{cases}\label{FACTORING}\\\nonumber    
+  \end{cases}
+\end{align*}\vspace{-\baselineskip}%
+\begin{align*}
   \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber

### 2018-12-19 archive.tex
progress on deep program learning figure

@@
   \begin{tikzpicture}
+    \renewcommand{\baselinestretch}{0.5}
+
 %    \draw (0,0) rectangle (10,5);
     \node(t1)[align=left] at (1,-1) {
-      \small +106 769-43$\to $106.769.43\\
-      \small +83 973-831$\to $83.973.831\\
-      \small $f(\text{\code{s}}) = $\code{(}\subnode{f01}{$f_0$}\code{  \codechar{.} \codechar{-}}\\
-      \small \hspace{1.25cm}\code{($f_0$ \codechar{.} \codechar{ }}\\
-      \small \hspace{1.4cm}\code{(cdr s)))}
+      \footnotesize +106 769-43$\to $106.769.43\\
+      \footnotesize +83 973-831$\to $83.973.831\\
+      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{orange}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
+      \footnotesize \hspace{1.25cm}\mcode{(${\color{orange}f_0}$ \codechar{.} \codechar{ }}\\
+      \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
     };
     \node(t2)[align=left] at ([xshift=2cm]t1.east) {
-      \small Dream Coder$\to$Dr. Dream\\
-      \small Andy Cencici$\to$Dr. Andy\\
-      \small Jan Kotas$\to$Dr. Jan\\
-      \small Nancy Jones$\to$Dr. Nancy\\
-      $f(\text{\code{s}}) = $\code{($f_3$ ($f_4$ \codechar{ } s))}
+      \footnotesize Dream Coder$\to$Dr. Dream\\
+      \footnotesize Andy Cencici$\to$Dr. Andy\\
+      \footnotesize Jan Kotas$\to$Dr. Jan\\
+      \footnotesize Nancy Jones$\to$Dr. Nancy\\
+      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{orange}f_3}$ (${\color{orange}f_4}$ \codechar{ } s))}
     };
     \node(t3)[align=left] at ([xshift=2cm]t2.east) {
-      \small Dream (Coder)$\to$Coder\\
-      \small 1 (2) 3 4$\to$2\\
-      \small the (dog) ran$\to$ran\\
-      \small +817(33)99$\to$33\\
-      $f(\text{\code{s}}) = $\code{($f_6$ \codechar{(} \codechar{)} s)}
+      \footnotesize Dream (Coder)$\to$Coder\\
+      \footnotesize 1 (2) 3 4$\to$2\\
+      \footnotesize the (dog) ran$\to$ran\\
+      \footnotesize +817(33)99$\to$33\\
+      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{orange}f_6}$ \codechar{(} \codechar{)} s)}
     };
@@
     \draw (-2,-2.5) -- (15,-2.5);
-    \node[align=left] at (-1.5,-2.3) {Tasks and\\\texttt{Programs}\\\\DSL};
-
-    \node(substitution)[draw,rounded corners,align=center] at (1,-4) {
-      \small $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\\
-       \mcode{(map ($\lambda$ (x) (if}\\
-      \mcode{(= x a) b x)) s)}
+    \node[align=left,rotate=90] at (-1.5,-1.3) {Tasks and\\\\\texttt{Programs}};
+
+    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)};
+
+    \renewcommand{\horizontalSpacing}{3.5cm}
+    \renewcommand{\verticalSpacing}{2cm}
+
+    \newcommand{\invention}[4]{
+      \node(#1)[inner sep=1,anchor=center,text width=2cm,draw,rounded corners,align=center] at (#2) {
+        \footnotesize #3
       };
-    \node[anchor = north] at (substitution.south) {substitute characters};
-    \node(cat)[draw,rounded corners,align=center] at (1,-6) {
-      \small $f_1($\code{a}$,$\code{b}$) \,=\, $\\
-       \mcode{(fold ($\lambda$ (x y)}\\\mcode{ (cons x y)) a b))}
-    };
-    \node[anchor = north] at (cat.south) {concat strings};
-    \node(looping)[draw,rounded corners,align=center] at (1,-8) {
-      \small $f_2($\code{c}$,$\code{p}$,$\code{s}$) \,=\, $\\
-       \mcode{(unfold l ($\lambda$ (x)}\\\mcode{(= c (car x)))}\\\mcode{ p ($\lambda$ (z) (cdr z)))}
-    };
-    \node[anchor = north] at (looping.south) {loop until char};
-
-    \node(constant)[draw,rounded corners,align=center] at (4.5,-4) {
-      \small $f_3($\code{s}$) \,=\, $\\
-       \mcode{($f_0$ s string)}
-    };
-    \node[anchor = north] at (constant.south) {prepend constant};
-    \node(drop)[draw,rounded corners,align=center] at (4.5,-6) {
-      \small $f_4($\code{c}$) \,=\, $\\
-      \mcode{($f_2$ c ($\lambda$ }\\\mcode{(z) (car z)))}
-    };
-    \node[anchor = north] at (drop.south) {prefix};
-    \node(suffix)[draw,rounded corners,align=center] at ([yshift=-2cm]drop.center) {
-      \small $f_5($\code{s}$,$\code{c}$) \,=\, $\\
+      \node[anchor=north] at ([yshift=3.5pt]#1.south) {#4};
+    }
+    \newcommand{\vinvention}[4]{
+      \invention{#1}{[yshift=-\verticalSpacing]#2.center}{#3}{#4}
+    }
+    \newcommand{\hinvention}[4]{
+      \invention{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
+    }
+
+    \invention{substitution}{1,-4}{
+      ${\color{orange}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
+       \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
+    }{substitute chars}
+    \vinvention{cat}{substitution}{
+      ${\color{orange}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
+      \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
+    }{concat strings}
+    \vinvention{looping}{cat}{
+       ${\color{orange}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
+      \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
+    }{loop until char}
+
+    \hinvention{constant}{substitution}{
+      ${\color{orange}f_3}($\mcode{s}$) \,=\, $\\
+      \mcode{(${\color{orange}f_1}$ s string)}
+    }{prepend constant}
+    \vinvention{drop}{constant}{
+      ${\color{orange}f_4}($\mcode{c}$) \,=\, $\\
+      \mcode{(${\color{orange}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
+    }{prefix}
+    \vinvention{suffix}{drop}{
+      ${\color{orange}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
       \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
-    };
-    \node[anchor = north] at (suffix.south) {suffix};
-    \node(delimited)[draw, rounded corners, align=center,text width=2.5cm] at ([xshift=3.5cm]constant.center) {
-      \code{$f_6($x$,$y$,$s$)\,=\,$}\\
-      \mcode{($f_4$ x ($f_5$ y c))}};
-    \node[anchor = north] at (delimited.south) {delimited substring};
+    }{suffix}
+
+    \hinvention{delimited}{constant}{
+      \mcode{${\color{orange}f_6}($x$,$y$,$s$)\,=\,$}\\
+      \mcode{(${\color{orange}f_4}$ x (${\color{orange}f_5}$ y c))}
+    }{delimited substring}
     \draw[->] (delimited.west) -- (drop.east);
@@
 
-    \node(appendSingle)[draw, rounded corners, align=center,text width=2.5cm] at ([yshift=-2cm]delimited.center) {
-      \code{$f_7($c$,$s$)\,=\,$}\\
-      \mcode{($f_1$ (cons c empty) s)}};
-    \node[anchor = north] at (appendSingle.south) {append character};
-
-    \node(surround)[draw, rounded corners, align=center,text width=2.5cm] at ([yshift=-2cm]appendSingle.center) {
-      \code{$f_8($s$)\,=\,$}\\
-      \mcode{(cons \codechar{(} ($f_7$ \codechar{)} s))}};
-    \node[anchor = north] at (surround.south) {surround in parentheses};
+    \vinvention{appendSingle}{delimited}{
+      \mcode{${\color{orange}f_7}($c$,$s$)\,=\,$}\\
+      \mcode{(${\color{orange}f_1}$ (cons c nil) s)}}{append character}
+
+\vinvention{surround}{appendSingle}{
+      \mcode{${\color{orange}f_8}($s$)\,=\,$}\\
+      \mcode{(cons \codechar{(} (${\color{orange}f_7}$ \codechar{)} s))}}{surround in parens}
     \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
@@
 
-    \draw[->] ([yshift=0.5cm,xshift=-0.4cm]t1.south) to[in=100,out=200] (substitution.north);
-    \draw[->] ([yshift=0.8cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
+    \draw[->] ([yshift=0.2cm,xshift=-0.2cm]t1.south) to[in=100,out=200] (substitution.north);
+    \draw[->] ([yshift=0.5cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
     \draw[->] ([xshift=-0.5cm]t3.south) -- (delimited.north);
+
+    \begin{scope}[shift={(0,-15)}]
+      \invention{filter}{0,0}{
+        ${\color{orange}f_0} = $\mcode{($\lambda$ (p l)\\ (fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil l))} 
+        }{higher-order `filter'}
+
+      \invention{three}{[yshift=-\verticalSpacing]filter.center}{\footnotesize ${\color{orange}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
+      \invention{prefix}{[yshift=-\verticalSpacing]three.center}{\footnotesize ${\color{orange}f_2} = $\mcode{($\lambda$ (n l) (map ($\lambda$ (i) (index i l)) (range (+ 1 n))))}}{First $n$ elements}
+      \invention{single}{[xshift=\horizontalSpacing]filter.center}{\footnotesize ${\color{orange}f_3} = $\mcode{($\lambda$ (z l) (fold ($\lambda$ (x y) (cons x y)) z l))}}{Append element}
+      \invention{for}{[yshift=-\verticalSpacing]single.center}{\footnotesize ${\color{orange}f_4} = $\mcode{(+ 1 ${\color{orange}f_1}$)}}{The number `4'}
+      \invention{Max}{[yshift=-\verticalSpacing]for.center}{\footnotesize ${\color{orange}f_5} = $\mcode{($\lambda$ (l) %
+          (car (${\color{orange}f_0}$ ($\lambda$ (x) (nil? (${\color{orange}f_0}$ l ($\lambda$ (z) (> z x))))) l)))}}{Max of list}
+      \invention{largest}{[xshift=\horizontalSpacing]single.center}{\footnotesize ${\color{orange}f_6} = $\mcode{($\lambda$ (l n) %
+          (${\color{orange}f_5}$ (${\color{orange}f_0}$ ($\lambda$ (x) (> n (length (${\color{orange}f_0}$ ($\lambda$ (y) (> x y)) l)))) l)))}}{$n^{\text{th}}$ largest element}
+      \vinvention{count}{largest}{\footnotesize ${\color{orange}f_7} = $\mcode{($\lambda$ (x l) %
+          (length (${\color{orange}f_0}$ ($\lambda$ (y) (= x y)) l)))}}{count}
+      \vinvention{weird}{count}{\footnotesize ${\color{orange}f_8} = $\mcode{($\lambda$ (p l) %
+          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
+                 {Higher-order incrementer}
+
+      \draw[->] (for.west) -- (three.east);
+      \draw[->] ([yshift=-0.5cm]largest.west) -- (Max.east);
+      \draw[->] ([yshift=-0.5cm]largest.west) to[out=240,in=340] ([yshift=-0.8cm]filter.east);
+      \draw[->] (Max.west) -- (filter.east);
+      \draw[->] ([yshift=0.1cm]count.west) to[out=90,in=300] ([yshift=-0.3cm]filter.east);
+
+      \node(t1)[align=left] at (0,3) {
+        \mcode{[1 4 5 8]$\to$[5 8]}\\
+        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
+        \mcode{[3 4 5 6]$\to$[5 6]}\\
+        \mcode{[6 5 4 3]$\to$[6 5]}\\
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{orange}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{orange}f_4}$)) z)}
+      };
+      \node(t2)[align=left] at ([xshift=3.75cm]t1.center) {
+        \mcode{[9 2 4]$\to$[4 2 9]}\\
+        \mcode{[4 3]$\to$[3 4]}\\
+        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
+        \mcode{[]$\to$[]}\\
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{orange}f_3}$ x a)) nil z)}
+      };
+      \node(t3)[align=left] at ([xshift=3.75cm]t2.center) {
+        \mcode{[9 2 3]$\to$[2 3 9]}\\
+        \mcode{[2 1]$\to$[1 2]}\\
+        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{ (${\color{orange}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range (length z)))}
+      };
+
+      \draw[->] ([yshift=0.5cm,xshift=-0.2cm]t1.south) to[out=250] (filter.north);
+      \draw[->] ([yshift=0.0cm,xshift=0.3cm]t1.south) to[out=350,in=100] (for.west);
+      \draw[->] ([xshift=-0.7cm]t2.south) -- (single.north);
+      \draw[->] ([xshift=-1.25cm,yshift=0.5cm]t3.south) to[out=240] (largest.north);
+
+    \end{scope}
 

### 2018-12-19 archive.tex
color-coded and cleaned up arrows

@@
     \renewcommand{\baselinestretch}{0.5}
+    \renewcommand{\inventionColor}{orange}
 
@@
       \footnotesize +83 973-831$\to $83.973.831\\
-      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{orange}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
-      \footnotesize \hspace{1.25cm}\mcode{(${\color{orange}f_0}$ \codechar{.} \codechar{ }}\\
+      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{\inventionColor}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
+      \footnotesize \hspace{1.25cm}\mcode{(${\color{\inventionColor}f_0}$ \codechar{.} \codechar{ }}\\
       \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
@@
       \footnotesize Nancy Jones$\to$Dr. Nancy\\
-      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{orange}f_3}$ (${\color{orange}f_4}$ \codechar{ } s))}
+      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_3}$ (${\color{\inventionColor}f_4}$ \codechar{ } s))}
     };
@@
       \footnotesize +817(33)99$\to$33\\
-      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{orange}f_6}$ \codechar{(} \codechar{)} s)}
+      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_6}$ \codechar{(} \codechar{)} s)}
     };
@@
 
-    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)};
+    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)\\\\System discovers subroutines in {\color{\inventionColor}orange}};
 
@@
     \invention{substitution}{1,-4}{
-      ${\color{orange}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
+      ${\color{\inventionColor}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
        \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
@@
     \vinvention{cat}{substitution}{
-      ${\color{orange}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
+      ${\color{\inventionColor}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
       \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
@@
     \vinvention{looping}{cat}{
-       ${\color{orange}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
+       ${\color{\inventionColor}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
       \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
@@
     \hinvention{constant}{substitution}{
-      ${\color{orange}f_3}($\mcode{s}$) \,=\, $\\
-      \mcode{(${\color{orange}f_1}$ s string)}
+      ${\color{\inventionColor}f_3}($\mcode{s}$) \,=\, $\\
+      \mcode{(${\color{\inventionColor}f_1}$ s string)}
     }{prepend constant}
     \vinvention{drop}{constant}{
-      ${\color{orange}f_4}($\mcode{c}$) \,=\, $\\
-      \mcode{(${\color{orange}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
+      ${\color{\inventionColor}f_4}($\mcode{c}$) \,=\, $\\
+      \mcode{(${\color{\inventionColor}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
     }{prefix}
     \vinvention{suffix}{drop}{
-      ${\color{orange}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
+      ${\color{\inventionColor}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
       \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
@@
     \hinvention{delimited}{constant}{
-      \mcode{${\color{orange}f_6}($x$,$y$,$s$)\,=\,$}\\
-      \mcode{(${\color{orange}f_4}$ x (${\color{orange}f_5}$ y c))}
+      \mcode{${\color{\inventionColor}f_6}($x$,$y$,$s$)\,=\,$}\\
+      \mcode{(${\color{\inventionColor}f_4}$ x (${\color{\inventionColor}f_5}$ y c))}
     }{delimited substring}
@@
     \vinvention{appendSingle}{delimited}{
-      \mcode{${\color{orange}f_7}($c$,$s$)\,=\,$}\\
-      \mcode{(${\color{orange}f_1}$ (cons c nil) s)}}{append character}
+      \mcode{${\color{\inventionColor}f_7}($c$,$s$)\,=\,$}\\
+      \mcode{(${\color{\inventionColor}f_1}$ (cons c nil) s)}}{append character}
 
 \vinvention{surround}{appendSingle}{
-      \mcode{${\color{orange}f_8}($s$)\,=\,$}\\
-      \mcode{(cons \codechar{(} (${\color{orange}f_7}$ \codechar{)} s))}}{surround in parens}
+      \mcode{${\color{\inventionColor}f_8}($s$)\,=\,$}\\
+      \mcode{(cons \codechar{(} (${\color{\inventionColor}f_7}$ \codechar{)} s))}}{surround in parens}
     \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
@@
       \invention{filter}{0,0}{
-        ${\color{orange}f_0} = $\mcode{($\lambda$ (p l)\\ (fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil l))} 
+        ${\color{\inventionColor}f_0} = $\mcode{($\lambda$ (p l)\\ (fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil l))} 
         }{higher-order `filter'}
 
-      \invention{three}{[yshift=-\verticalSpacing]filter.center}{\footnotesize ${\color{orange}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
-      \invention{prefix}{[yshift=-\verticalSpacing]three.center}{\footnotesize ${\color{orange}f_2} = $\mcode{($\lambda$ (n l) (map ($\lambda$ (i) (index i l)) (range (+ 1 n))))}}{First $n$ elements}
-      \invention{single}{[xshift=\horizontalSpacing]filter.center}{\footnotesize ${\color{orange}f_3} = $\mcode{($\lambda$ (z l) (fold ($\lambda$ (x y) (cons x y)) z l))}}{Append element}
-      \invention{for}{[yshift=-\verticalSpacing]single.center}{\footnotesize ${\color{orange}f_4} = $\mcode{(+ 1 ${\color{orange}f_1}$)}}{The number `4'}
-      \invention{Max}{[yshift=-\verticalSpacing]for.center}{\footnotesize ${\color{orange}f_5} = $\mcode{($\lambda$ (l) %
-          (car (${\color{orange}f_0}$ ($\lambda$ (x) (nil? (${\color{orange}f_0}$ l ($\lambda$ (z) (> z x))))) l)))}}{Max of list}
-      \invention{largest}{[xshift=\horizontalSpacing]single.center}{\footnotesize ${\color{orange}f_6} = $\mcode{($\lambda$ (l n) %
-          (${\color{orange}f_5}$ (${\color{orange}f_0}$ ($\lambda$ (x) (> n (length (${\color{orange}f_0}$ ($\lambda$ (y) (> x y)) l)))) l)))}}{$n^{\text{th}}$ largest element}
-      \vinvention{count}{largest}{\footnotesize ${\color{orange}f_7} = $\mcode{($\lambda$ (x l) %
-          (length (${\color{orange}f_0}$ ($\lambda$ (y) (= x y)) l)))}}{count}
-      \vinvention{weird}{count}{\footnotesize ${\color{orange}f_8} = $\mcode{($\lambda$ (p l) %
+      \invention{three}{[yshift=-\verticalSpacing]filter.center}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
+      \invention{prefix}{[yshift=-\verticalSpacing]three.center}{\footnotesize ${\color{\inventionColor}f_2} = $\mcode{($\lambda$ (n l) (map ($\lambda$ (i) (index i l)) (range (+ 1 n))))}}{First $n$ elements}
+      \invention{single}{[xshift=\horizontalSpacing]filter.center}{\footnotesize ${\color{\inventionColor}f_3} = $\mcode{($\lambda$ (z l) (fold ($\lambda$ (x y) (cons x y)) z l))}}{Append element}
+      \invention{for}{[yshift=-\verticalSpacing]single.center}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
+      \invention{Max}{[yshift=-\verticalSpacing]for.center}{\footnotesize ${\color{\inventionColor}f_5} = $\mcode{($\lambda$ (l) %
+          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l)))}}{Max of list}
+      \invention{largest}{[xshift=\horizontalSpacing]single.center}{\footnotesize ${\color{\inventionColor}f_6} = $\mcode{($\lambda$ (l n) %
+          (${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l)))}}{$n^{\text{th}}$ largest element}
+      \vinvention{count}{largest}{\footnotesize ${\color{\inventionColor}f_7} = $\mcode{($\lambda$ (x l) %
+          (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l)))}}{count}
+      \vinvention{weird}{count}{\footnotesize ${\color{\inventionColor}f_8} = $\mcode{($\lambda$ (p l) %
           (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
@@
       \draw[->] (for.west) -- (three.east);
-      \draw[->] ([yshift=-0.5cm]largest.west) -- (Max.east);
-      \draw[->] ([yshift=-0.5cm]largest.west) to[out=240,in=340] ([yshift=-0.8cm]filter.east);
+      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
+      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
       \draw[->] (Max.west) -- (filter.east);
-      \draw[->] ([yshift=0.1cm]count.west) to[out=90,in=300] ([yshift=-0.3cm]filter.east);
+      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);
 
@@
         \mcode{[6 5 4 3]$\to$[6 5]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{orange}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{orange}f_4}$)) z)}
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}
       };
@@
         \mcode{[]$\to$[]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{orange}f_3}$ x a)) nil z)}
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}
       };
@@
         \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{ (${\color{orange}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range (length z)))}
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{ (${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range (length z)))}
       };

### 2018-12-20 archive.tex
improvements to figures

@@
   \begin{tikzpicture}[line width=0.4mm]
-    \draw[fill=teal!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
+    \draw[fill=white!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
     \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
@@
     \node at (0.5,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
-    \draw[fill=teal!5!white] (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
+    \draw[fill=white!5!white] (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
     \node at (11.5,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
-    \draw[fill=teal!5!white] (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
+    \draw[fill=white!5!white] (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
 
@@
     %%}b memory consolidation
-    \begin{scope}[shift={(-2,-4.5)}]
-      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\mcode{(+ (car z) 1)}};
-      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\mcode{(cons (+ 1 1))}};
-      \node[align=center] at ([yshift = 0cm,xshift=0.4cm]f2.east) (f3){\textbf{DSL}\\\mcode{cons}, \mcode{+}, \\...};
-      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm,xshift=-0.3cm]f2.south) {};
+    \begin{scope}[shift={(-1.9,-4.5)}]
+      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\%
+        \mcode{(+ (car z)}\vspace{-5pt}\\\mcode{1)  }};
+      \node[align=center] at ([xshift = 0.7cm]f1.east) (f2){progs. for\\task$_2$:\\\mcode{(cons }\vspace{-5pt}\\\mcode{(+ 1 1))}};
+      \node[align=center] at ([xshift=0.7cm]f2.east) (f3){\textbf{DSL}\\\mcode{cons}, \mcode{+}, \\...};
+      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm,xshift=-0.0cm]f2.south) {};
       \draw [-> ] (f1.south) -- (f1.south|-c.north);
@@
         \begin{tikzpicture}[scale=0.9,line width=0.5mm]
-          \node[align=center] at (0,0) {DSL w/ \code{(+ x 1)}:};
+          \node[align=center] at (0,0) {\textbf{new DSL} w/ \texttt{(+ x 1)}:};
           \begin{scope}[shift={(0.6,-0.5)}]
-            \node[pop3](p1) at (-1,-1) {\code{+}};
-            \node[pop3](n1) at (0.6,-0.9) {\code{1}};
-            \node[pop3](a) at (0,-1) {\code{ }};
+            \node[pop3](p1) at (-1,-1) {\texttt{+}};
+            \node[pop3](n1) at (0.6,-0.9) {\texttt{1}};
+            \node[pop3](a) at (0,-1) {\texttt{ }};
             \draw[pop3] (0,0) -- (p1.north);
@@
           \node(l1) at (0,0) {};
-          \node[color=pop3](p1) at (-1,-1) {\code{+}};
-          \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-          \node(x1) at (0,-1) {\code{1}};
+          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
+          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
+          \node(x1) at (0,-1) {\texttt{1}};
           \draw[color=pop3] (l1.south) -- (p1.north);
@@
           \draw (l1.south) -- (t.south);
-          \node(c) at (-1.5,-0.2) {\code{cons}};
+          \node(c) at (-1.5,-0.2) {\texttt{cons}};
           \draw (t.south) -- (c.north);
@@
             \node(l1) at (0,0) {};
-            \node[color=pop3](p1) at (-1,-1) {\code{+}};
-            \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+            \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
+            \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
             %\node(x1) at (0,-1) {};
@@
 
-            \node(c) at (-0.5,-1.5) {\code{car}};
-            \node(z) at (0.5,-1.5) {\code{z}};
+            \node(c) at (-0.5,-1.5) {\texttt{car}};
+            \node(z) at (0.5,-1.5) {\texttt{z}};
 
@@
     \renewcommand{\inventionColor}{orange}
+    \newcommand{\name}[2]{
+      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
+    }
+    \newcommand{\noArguments}[1]{
+      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
+    }
+
+    \renewcommand{\horizontalSpacing}{3.8cm}
+    \renewcommand{\verticalSpacing}{2cm}
 
@@
     };
-    \node(t2)[align=left] at ([xshift=2cm]t1.east) {
+    \node(t2)[align=left,anchor=center] at ([xshift=\horizontalSpacing]t1.center) {
       \footnotesize Dream Coder$\to$Dr. Dream\\
@@
 
-    \renewcommand{\horizontalSpacing}{3.5cm}
-    \renewcommand{\verticalSpacing}{2cm}
-
     \newcommand{\invention}[4]{
@@
       \invention{filter}{0,0}{
-        ${\color{\inventionColor}f_0} = $\mcode{($\lambda$ (p l)\\ (fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil l))} 
+        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
         }{higher-order `filter'}
 
-      \invention{three}{[yshift=-\verticalSpacing]filter.center}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
-      \invention{prefix}{[yshift=-\verticalSpacing]three.center}{\footnotesize ${\color{\inventionColor}f_2} = $\mcode{($\lambda$ (n l) (map ($\lambda$ (i) (index i l)) (range (+ 1 n))))}}{First $n$ elements}
-      \invention{single}{[xshift=\horizontalSpacing]filter.center}{\footnotesize ${\color{\inventionColor}f_3} = $\mcode{($\lambda$ (z l) (fold ($\lambda$ (x y) (cons x y)) z l))}}{Append element}
-      \invention{for}{[yshift=-\verticalSpacing]single.center}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
-      \invention{Max}{[yshift=-\verticalSpacing]for.center}{\footnotesize ${\color{\inventionColor}f_5} = $\mcode{($\lambda$ (l) %
-          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l)))}}{Max of list}
-      \invention{largest}{[xshift=\horizontalSpacing]single.center}{\footnotesize ${\color{\inventionColor}f_6} = $\mcode{($\lambda$ (l n) %
-          (${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l)))}}{$n^{\text{th}}$ largest element}
-      \vinvention{count}{largest}{\footnotesize ${\color{\inventionColor}f_7} = $\mcode{($\lambda$ (x l) %
-          (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l)))}}{count}
-      \vinvention{weird}{count}{\footnotesize ${\color{\inventionColor}f_8} = $\mcode{($\lambda$ (p l) %
+      \vinvention{three}{filter}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
+      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
+      \hinvention{single}{filter}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
+      \vinvention{for}{single}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
+      \vinvention{Max}{for}{\name{5}{l}\\\mcode{%
+          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
+      \hinvention{largest}{single}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
+      \vinvention{count}{largest}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
+      \vinvention{weird}{count}{\name{8}{p$,$l}\mcode{
           (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
@@
 
-      \node(t1)[align=left] at (0,3) {
+      \node(t1)[align=left,anchor=center] at (0,3) {
         \mcode{[1 4 5 8]$\to$[5 8]}\\
@@
       };
-      \node(t2)[align=left] at ([xshift=3.75cm]t1.center) {
+      \node(t2)[align=left] at ([xshift=\horizontalSpacing]t1.center) {
         \mcode{[9 2 4]$\to$[4 2 9]}\\
@@
       };
-      \node(t3)[align=left] at ([xshift=3.75cm]t2.center) {
+      \node(t3)[align=left] at ([xshift=\horizontalSpacing]t2.center) {
         \mcode{[9 2 3]$\to$[2 3 9]}\\
@@
         \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{ (${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range (length z)))}
+        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}
       };
 
-      \draw[->] ([yshift=0.5cm,xshift=-0.2cm]t1.south) to[out=250] (filter.north);
+      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]t1.south) to[out=250] (filter.north);
       \draw[->] ([yshift=0.0cm,xshift=0.3cm]t1.south) to[out=350,in=100] (for.west);
       \draw[->] ([xshift=-0.7cm]t2.south) -- (single.north);
-      \draw[->] ([xshift=-1.25cm,yshift=0.5cm]t3.south) to[out=240] (largest.north);
+      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]t3.south) to[out=240] (largest.north);
 
@@
     \end{tikzpicture}
+\end{figure}
+
+\begin{figure}
+  \begin{tikzpicture}
+    \renewcommand{\baselinestretch}{0.5}
+    \renewcommand{\inventionColor}{orange}
+
+    \newcommand{\name}[2]{
+      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
+    }
+    \newcommand{\noArguments}[1]{
+      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
+    }
+    \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
+    \renewcommand{\horizontalSpacing}{3.9cm}
+    \renewcommand{\verticalSpacing}{2cm}
+
+    \newcommand{\invention}[4]{
+      \node(#1)[inner sep=1,anchor=south,text width=2cm,draw,rounded corners,align=center] at (#2) {
+        \footnotesize #3
+      };
+      \node[anchor=north,align=center] at ([yshift=3.5pt]#1.south) {#4};
+    }
+    \newcommand{\vinvention}[4]{
+      \invention{#1}{[yshift=-\verticalSpacing]#2.south}{#3}{#4}
+    }
+    \newcommand{\hinvention}[4]{
+      \invention{#1}{[xshift=\horizontalSpacing]#2.south}{#3}{#4}
+    }
+    \newcommand{\exampleFunction}[2]{
+      \includegraphics[width = 0.5cm]{#1}, \includegraphics[width = 0.5cm]{#2}
+    }
+    \node[anchor=center,align=center](t1) at (2,0) {
+      \includegraphics[width = 2cm]{figures/logoTasks/logo144.png}\\
+    \mcode{(for (i 4) (\use{3} i))}};
+    \node[anchor=center,align=center](t2) at ([xshift=\horizontalSpacing]t1.center) {
+      \includegraphics[width = 2cm]{figures/logoTasks/logo121.png}\\
+    \mcode{(\use{7} 6 (\use{1} \use{0}))}};
+    \node[anchor=center,align=center](t3) at ([xshift=\horizontalSpacing]t2.center) {
+      \includegraphics[width = 2cm]{figures/logoTasks/logo133.png}\\
+      \mcode{(\use{7} 7 (\use{8} \use{6}))}};
+    \invention{circle}{2,-3}{
+      \name{0}{r}\mcode{(for (i $\infty$) (move r $\epsilon$) (move $\epsilon$ $\epsilon$))}
+    }{circle\\\exampleFunction{figures/logoTasks/logo51}{figures/logoTasks/logo63_dilated}}
+    \vinvention{pu}{circle}{
+      \noArguments{1}\mcode{(pen-up (move 1 0))}
+    }{pick up pen\\and move}
+    \vinvention{repeated}{pu}{
+      \name{2}{n$,$d$,$a}\mcode{(for (i n) (move d a))}
+    }{repeated segment\\\exampleFunction{figures/logoTasks/logoDemo1_dilated}{figures/logoTasks/logoDemo0_dilated.png}}
+    \hinvention{square}{circle}{
+      \name{3}{d}\mcode{(\use{2} 4 d (/ $2\pi$ 4))}
+    }{square\\\exampleFunction{figures/logoTasks/logo139_dilated}{figures/logoTasks/logo142_dilated}}
+    \vinvention{smooth}{square}{
+      \name{4}{d$,$a}\mcode{(\use{2} $\infty$ d a)}
+    }{Smooth curve\\\exampleFunction{figures/logoTasks/logoDemo2_dilated}{figures/logoTasks/logoDemo3_dilated}}
+    \vinvention{angle}{smooth}{
+      \name{5}{n}\mcode{(/ $2\pi$ n)}
+    }{$\frac{2\pi}{n}$}
+    \hinvention{semi}{square}{
+      \noArguments{6}\mcode{(\use{4} $\epsilon$ $\epsilon$)}
+    }{semicircle\\\exampleFunction{figures/logoTasks/logo55_dilated}{figures/logoTasks/logo62_dilated}}
+    \vinvention{snow}{semi}{
+      \name{7}{n$,$f}\mcode{(for (i n) (move 0 (\use{5} n)) (get/set f))}
+    }{rotational symmetry\\\exampleFunction{figures/logoTasks/logo119_dilated}{figures/logoTasks/logo68_dilated}}
+    \vinvention{line}{snow}{
+      \noArguments{8}\mcode{(move 1 0)}
+    }{line segment}
+
+    \draw[->] (t1.south) -- (square.north);
+    \draw[->] (t2.south) -- (circle.north);
+    \draw[->] (t3.south) -- (semi.north);
+    \draw[->] (t2.south) -- (snow.west);
+    \draw[->] (t2.south) -- (pu.east);
+    \draw[->] (t3.south) to[out=320,in=45] (line.east);
+    \draw[->] (t3.south) to[out=200,in=120] (snow.west);
+
+    \draw[->] (snow.west) -- (angle.east);
+    \draw[->] (smooth.west) -- (repeated.east);
+    \draw[->] (semi.west) -- (smooth.east);
+    \draw[->] (square.west) -- (repeated.east);
+  \end{tikzpicture}
+
   \end{figure}

### 2018-12-23 archive.tex
improve consolidation figure

@@
     \begin{scope}[shift={(0.5,0.5)}]
-      \node[align=right] at (0,0) (d){\textbf{DSL}\\
+      \node[align=right] at (0,-0.5) (d){\textbf{DSL}\\
         $f_1$, $f_2$, ...};
@@
 
-    \node at (0.5,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
-    \draw[fill=white!5!white] (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
-    \node at (11.5,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
-    \draw[fill=white!5!white] (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);
-
-    \begin{scope}[shift={(9,-4.5)}]
+    \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
+    \draw[fill=white!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
+    \node at (12,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
+    \draw[fill=white!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);
+
+    \begin{scope}[shift={(9.5,-4.5)}]
       \node(dreaming) at (2,-1) {\underline{Fantasies}};
@@
 
-      \begin{scope}[shift={(-3.3,-8)},scale=0.6,line width=0.05mm]
+      \begin{scope}[shift={(-3.8,-8)},scale=0.6,line width=0.05mm]
         \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
@@
 
-    %%}b memory consolidation
-    \begin{scope}[shift={(-1.9,-4.5)}]
-      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\%
-        \mcode{(+ (car z)}\vspace{-5pt}\\\mcode{1)  }};
-      \node[align=center] at ([xshift = 0.7cm]f1.east) (f2){progs. for\\task$_2$:\\\mcode{(cons }\vspace{-5pt}\\\mcode{(+ 1 1))}};
-      \node[align=center] at ([xshift=0.7cm]f2.east) (f3){\textbf{DSL}\\\mcode{cons}, \mcode{+}, \\...};
-      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm,xshift=-0.0cm]f2.south) {};
-      \draw [-> ] (f1.south) -- (f1.south|-c.north);
-      \draw [-> ] (f2.south) -- (f2.south|-c.north);
-      \draw [-> ] (f3.south) -- (f3.south|-c.north);
+    % memory consolidation
+    \begin{scope}[shift={(-2,-4.5)}]
+
+      %% defined routines for creating fragmented syntax trees
+      \renewcommand{\syntaxOne}[1]{
+        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
+          \node(l1) at (0,0) {};
+          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
+          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
+          \node(x1) at (0,-1) {\texttt{1}};
+          \draw[color=pop3] (l1.south) -- (p1.north);
+          \draw[color=pop3] (l1.south) -- (n1.north);
+          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+          \node(t) at (-0.5,0.5) {};
+          \draw (l1.south) -- (t.south);
+          \node(c) at (-1.5,-0.2) {\texttt{cons}};
+          \draw (t.south) -- (c.north);
+        \end{tikzpicture}
+      }
+      \renewcommand{\syntaxTo}[1]{
+        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
+            \node(l1) at (0,0) {};
+            \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
+            \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
+            \draw[color=pop3] (l1.south) -- (p1.north);
+            \draw[color=pop3] (l1.south) -- (n1.north);
+            \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+            \node(c) at (-0.5,-1.5) {\texttt{car}};
+            \node(z) at (0.5,-1.5) {\texttt{z}};
+            \draw (0,-1) -- (c.north);
+            \draw (0,-1) -- (z.north);
+        \end{tikzpicture}
+      }
+
+      \node[align=center,anchor=center] at (0.4,-1.2) (f1){\textbf{Progs. for Task 1}:\\\code{(+ (car z) 1)}};
+      \node[align=center] at ([xshift = 1.75cm]f1.east) (f2){\textbf{Progs. for Task 2}:\\\code{(cons (+ 1 1))}};
+      \node(s1) at ([yshift=-0.5cm]f1.south) {\syntaxOne{0.8}};
+      \node(s2) at ([yshift=-0.5cm]f2.south) {\syntaxTo{0.8}};
+      \node(c)[align=center,rectangle, rounded corners, draw, minimum width = 4cm, minimum height = 1.5cm, anchor = north] at ($(s1.south)!0.5!(s2.south) + (0,-1)$) {Refactoring Algorithm};
+      \draw [-> ] (s1.south) -- (s1.south|-c.north);
+      \draw [-> ] (s2.south) -- (s2.south|-c.north);
 
@@
       \draw [-> ] (c.south) -- (d.north);
-
-      \node at (c.center) {
-        \begin{tikzpicture}[scale=0.9,line width=0.35mm]
-
-          \node(l1) at (0,0) {};
-          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
-          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
-          \node(x1) at (0,-1) {\texttt{1}};
-          \draw[color=pop3] (l1.south) -- (p1.north);
-          \draw[color=pop3] (l1.south) -- (n1.north);
-          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-          \node(t) at (-0.5,0.5) {};
-          \draw (l1.south) -- (t.south);
-          \node(c) at (-1.5,-0.2) {\texttt{cons}};
-          \draw (t.south) -- (c.north);
-
-          \begin{scope}[shift={(1,2)}]
-            \node(l1) at (0,0) {};
-            \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
-            \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
-            %\node(x1) at (0,-1) {};
-            \draw[color=pop3] (l1.south) -- (p1.north);
-            \draw[color=pop3] (l1.south) -- (n1.north);
-            \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-            \node(c) at (-0.5,-1.5) {\texttt{car}};
-            \node(z) at (0.5,-1.5) {\texttt{z}};
-
-            \draw (0,-1) -- (c.north);
-            \draw (0,-1) -- (z.north);
-
-            %  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-          \end{scope}
-
-        \end{tikzpicture}
-      };
 
@@
 
-\begin{figure}
+\begin{figure}\centering
   \begin{tikzpicture}
@@
     \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
-    \renewcommand{\horizontalSpacing}{3.9cm}
+    \renewcommand{\horizontalSpacing}{4.2cm}
     \renewcommand{\verticalSpacing}{2cm}
@@
     \newcommand{\invention}[4]{
-      \node(#1)[inner sep=1,anchor=south,text width=2cm,draw,rounded corners,align=center] at (#2) {
+      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,draw,rounded corners,align=center] at (#2) {
         \footnotesize #3
@@
     }
-    \node[anchor=center,align=center](t1) at (2,0) {
-      \includegraphics[width = 2cm]{figures/logoTasks/logo144.png}\\
-    \mcode{(for (i 4) (\use{3} i))}};
-    \node[anchor=center,align=center](t2) at ([xshift=\horizontalSpacing]t1.center) {
-      \includegraphics[width = 2cm]{figures/logoTasks/logo121.png}\\
-    \mcode{(\use{7} 6 (\use{1} \use{0}))}};
-    \node[anchor=center,align=center](t3) at ([xshift=\horizontalSpacing]t2.center) {
-      \includegraphics[width = 2cm]{figures/logoTasks/logo133.png}\\
-      \mcode{(\use{7} 7 (\use{8} \use{6}))}};
-    \invention{circle}{2,-3}{
+    \renewcommand{\task}[4]{
+      \node[align=left,anchor=center](t#1) at (#2) {#3};
+      \node[align=left](p#1) at ([yshift=-0.75cm]t#1.south) {#4};
+    }
+    \renewcommand{\nextTask}[4]{
+      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
+    }
+
+    \begin{scope}[shift={(0,8)}]
+      \task{1}{2,3}{
+        \textsc{Take if $\geq 5$}\\
+        \mcode{[1 4 5 8]$\to$[5 8]}\\
+        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
+        \mcode{[3 4 5 6]$\to$[5 6]}\\
+        \mcode{[6 5 4 3]$\to$[6 5]}
+      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}}
+      \nextTask{2}{t1}{
+        \textsc{Reverse}\\
+        \mcode{[9 2 4]$\to$[4 2 9]}\\
+        \mcode{[4 3]$\to$[3 4]}\\
+        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
+        \mcode{[]$\to$[]}
+      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}}
+
+      \nextTask{3}{t2}{
+        \textsc{Sort}\\
+        \mcode{[9 2 3]$\to$[2 3 9]}\\
+        \mcode{[2 1]$\to$[1 2]}\\
+        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
+        \mcode{[9 8 9]$\to$[8 9 9]}
+      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}}
+
+      \vinvention{filter}{p1}{
+        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
+      }{higher-order `filter'}
+
+      \vinvention{three}{filter}{\footnotesize \noArguments{1}\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
+      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
+      \hinvention{single}{filter}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
+      \vinvention{for}{single}{\noArguments{4}\\\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
+      \vinvention{Max}{for}{\name{5}{l}\\\mcode{%
+          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
+      \hinvention{largest}{single}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
+      \vinvention{count}{largest}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
+      \vinvention{weird}{count}{\name{8}{p$,$l}\mcode{
+          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0)))}}{Higher-order incrementer}
+
+      \node[rotate=90,align=center] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
+      \node[rotate=90,align=center] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
+      \node[rotate=90,align=center] at ([xshift=-2cm]three.center) {\textbf{\textsc{Domain Specific Language}}\\
+        \small Learned routines in {\color{orange}orange}};
+
+      \draw[->] (for.west) -- (three.east);
+      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
+      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
+      \draw[->] (Max.west) -- (filter.east);
+      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);
+
+      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]p1.south) to[out=250] (filter.north);
+      \draw[->] ([yshift=0.0cm,xshift=0.3cm]p1.south) to[out=350,in=100] (for.west);
+      \draw[->] ([xshift=-0.7cm]p2.south) -- (single.north);
+      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]p3.south) to[out=240] (largest.north);
+
+    \end{scope}
+
+    \renewcommand{\task}[4]{
+      \node(t#1) at (#2) {\includegraphics[width = 2cm]{#3}};
+      \node(p#1) at ([yshift=-0.75cm]t#1.south) {#4};
+    }
+    \newcommand{\nextTask}[4]{
+      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
+    }
+
+    \task{1}{2,0}{figures/logoTasks/logo144.png}{\mcode{(for (i 4) (\use{3} i))}}
+    \nextTask{2}{t1}{figures/logoTasks/logo121.png}{\mcode{(\use{7} 6 (\use{1}; (\use{0} 1)))}}
+    \nextTask{3}{t2}{figures/logoTasks/logo133.png}{\mcode{(\use{7} 7 (\use{8}; \use{6}))}}
+
+    \vinvention{circle}{p1}{
       \name{0}{r}\mcode{(for (i $\infty$) (move r $\epsilon$) (move $\epsilon$ $\epsilon$))}
@@
     \vinvention{snow}{semi}{
-      \name{7}{n$,$f}\mcode{(for (i n) (move 0 (\use{5} n)) (get/set f))}
+      \name{7}{n$,$f}\mcode{(for (i n) \\(move 0 (\use{5} n))\\ (get/set f))}
     }{rotational symmetry\\\exampleFunction{figures/logoTasks/logo119_dilated}{figures/logoTasks/logo68_dilated}}
@@
 
-    \draw[->] (t1.south) -- (square.north);
-    \draw[->] (t2.south) -- (circle.north);
-    \draw[->] (t3.south) -- (semi.north);
-    \draw[->] (t2.south) -- (snow.west);
-    \draw[->] (t2.south) -- (pu.east);
-    \draw[->] (t3.south) to[out=320,in=45] (line.east);
-    \draw[->] (t3.south) to[out=200,in=120] (snow.west);
-
-    \draw[->] (snow.west) -- (angle.east);
-    \draw[->] (smooth.west) -- (repeated.east);
-    \draw[->] (semi.west) -- (smooth.east);
-    \draw[->] (square.west) -- (repeated.east);
+    \draw[->] (p1.south) -- (square.north);
+    \draw[->] (p2.south) -- (circle.north);
+    \draw[->] (p3.south) -- (semi.north);
+    \draw[->] (p2.south) -- (snow.west);
+    \draw[->] (p2.south) -- (pu.east);
+    \draw[->] (p3.south) to[out=320,in=45] (line.east);
+    \draw[->] (p3.south) to[out=200,in=120] (snow.west);
+
+    \draw[->,orange] (snow.west) -- (angle.east);
+    \draw[->,orange] (smooth.west) -- (repeated.east);
+    \draw[->,orange] (semi.west) -- (smooth.east);
+    \draw[->,orange] (square.west) -- (repeated.east);
+
+    \node[align=center,rotate=90] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
+    \node(programLabel)[align=center,rotate=90] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
+    \node[rotate=90,align=center] at ([xshift=-2cm]pu.center) {\textbf{\textsc{Domain Specific Language}}\\
+      \small Learned routines in {\color{orange}orange}};
+
+    \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
+    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};
   \end{tikzpicture}
-
+  \caption{Two of the six domains we apply \system to.
+Agent observes tasks (top rows) which it solves by writing programs (middle rows) while jointly growing a library (DSL; bottom rows).
+Learned DSLs rediscover multiple higher-order functions (\code{filter} for list functions and rotational symmetry for generative graphics).
+Learned DSL components call each other (arrows).}
 \end{figure}

### 2018-12-29 archive.tex
lots of improvements

@@
 
+\tikzstyle{vecArrow} = [thick, decoration={markings,mark=at position                                
+   1 with {\arrow[semithick]{open triangle 60}}},                                                   
+   double distance=1.4pt, shorten >= 5.5pt,                                                         
+   preaction = {decorate},                                                                          
+   postaction = {draw,line width=1.4pt, white,shorten >= 4.5pt}]                                    
+\tikzstyle{innerWhite} = [semithick, white,line width=1.4pt, shorten >= 4.5pt]
+
 \begin{document}
@@
 \begin{figure}
-  \begin{tikzpicture}[line width=0.4mm]
-    \draw[fill=white!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
-    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
-
-    \begin{scope}[shift={(0.5,0.5)}]
-      \node[align=right] at (0,-0.5) (d){\textbf{DSL}\\
-        $f_1$, $f_2$, ...};
-      \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
-        \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
-        \footnotesize    \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} };
-
-      \node at ([xshift = 1.25cm]t.east) (nn){\NeuralNetwork{0.25}};
-      \node[align = center, text width = 1cm] at ([yshift = 0.7cm,xshift=0cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
-      \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
-
-      \node[draw,rounded corners, align=center, inner sep = 10] at ([xshift = 4.2cm,yshift = 1cm]t.east) (s){Neurally-Guided\\ Enumerative Search};
-
-      \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-      \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
-
-      \node[align=left] at ([xshift=3cm]s.east) (f) {\textbf{Programs:}\\
-        \small    \code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}\\
-        \small $\cdots\cdots\cdots$};
-      \draw [->  ] (s.east) -- (f.west);
-
-      \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-    \end{scope}
-    \begin{scope}[shift={(9.4,-3.5)},scale=0.6,line width=0.05mm]
-      \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
-      \node[latent,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
-      \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
-      \node[latent,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
-      \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
-      \draw [->] (zp1.south) -- (xp1.north);
-      \draw [->] (dx.south) -- (zp1.north);
-      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-      \node[latent,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
-      \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
-      \draw [->] (zp1.south) -- (xp1.north);
-      \draw [->] (dx.south) -- (zp1.north);
-      \draw [->] (dx.south) -- (zp.north);
-      \draw [->] (zp.south) -- (xp.north);
-      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
-    \end{scope}
-
-    \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
-    \draw[fill=white!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
-    \node at (12,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
-    \draw[fill=white!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);
-
-    \begin{scope}[shift={(9.5,-4.5)}]
-      \node(dreaming) at (2,-1) {\underline{Fantasies}};
-      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
-      \node at ([xshift=1.5cm]d.east) (p1){prog.};
-      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
-      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
-      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
-      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
-        \NeuralNetwork{0.17}};
-      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
-      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
-
-      \node(replay) at (2,-5) {\underline{Replays}};
-      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
-      \node at ([xshift=1.5cm]d.east) (p1){prog.};
-      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
-      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
-      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
-      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
-        \NeuralNetwork{0.17}};
-      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
-      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
-
-      \begin{scope}[shift={(-3.8,-8)},scale=0.6,line width=0.05mm]
-        \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
-        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
-        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
-        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
-        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
-        \draw [->] (zp1.south) -- (xp1.north);
-        \draw [->] (dx.south) -- (zp1.north);
-        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
-        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
-        \draw [->] (zp1.south) -- (xp1.north);
-        \draw [->] (dx.south) -- (zp1.north);
-        \draw [->] (dx.south) -- (zp.north);
-        \draw [->] (zp.south) -- (xp.north);
-        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-        \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
-      \end{scope}
-
-      \end{scope}
-
-    % memory consolidation
-    \begin{scope}[shift={(-2,-4.5)}]
-
-      %% defined routines for creating fragmented syntax trees
-      \renewcommand{\syntaxOne}[1]{
-        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
-          \node(l1) at (0,0) {};
-          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
-          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
-          \node(x1) at (0,-1) {\texttt{1}};
-          \draw[color=pop3] (l1.south) -- (p1.north);
-          \draw[color=pop3] (l1.south) -- (n1.north);
-          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-          \node(t) at (-0.5,0.5) {};
-          \draw (l1.south) -- (t.south);
-          \node(c) at (-1.5,-0.2) {\texttt{cons}};
-          \draw (t.south) -- (c.north);
-        \end{tikzpicture}
-      }
-      \renewcommand{\syntaxTo}[1]{
-        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
-            \node(l1) at (0,0) {};
-            \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
-            \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
-            \draw[color=pop3] (l1.south) -- (p1.north);
-            \draw[color=pop3] (l1.south) -- (n1.north);
-            \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-            \node(c) at (-0.5,-1.5) {\texttt{car}};
-            \node(z) at (0.5,-1.5) {\texttt{z}};
-            \draw (0,-1) -- (c.north);
-            \draw (0,-1) -- (z.north);
-        \end{tikzpicture}
-      }
-
-      \node[align=center,anchor=center] at (0.4,-1.2) (f1){\textbf{Progs. for Task 1}:\\\code{(+ (car z) 1)}};
-      \node[align=center] at ([xshift = 1.75cm]f1.east) (f2){\textbf{Progs. for Task 2}:\\\code{(cons (+ 1 1))}};
-      \node(s1) at ([yshift=-0.5cm]f1.south) {\syntaxOne{0.8}};
-      \node(s2) at ([yshift=-0.5cm]f2.south) {\syntaxTo{0.8}};
-      \node(c)[align=center,rectangle, rounded corners, draw, minimum width = 4cm, minimum height = 1.5cm, anchor = north] at ($(s1.south)!0.5!(s2.south) + (0,-1)$) {Refactoring Algorithm};
-      \draw [-> ] (s1.south) -- (s1.south|-c.north);
-      \draw [-> ] (s2.south) -- (s2.south|-c.north);
-
-      \node(d) at ([yshift = -1.8cm]c.south) {
-        \begin{tikzpicture}[scale=0.9,line width=0.5mm]
-          \node[align=center] at (0,0) {\textbf{new DSL} w/ \texttt{(+ x 1)}:};
-          \begin{scope}[shift={(0.6,-0.5)}]
-            \node[pop3](p1) at (-1,-1) {\texttt{+}};
-            \node[pop3](n1) at (0.6,-0.9) {\texttt{1}};
-            \node[pop3](a) at (0,-1) {\texttt{ }};
-            \draw[pop3] (0,0) -- (p1.north);
-            \draw[pop3] (0,0) -- (n1.north);
-            \draw[pop3] (-0.3,-0.3) -- (a.north);
-          \end{scope}
-      \end{tikzpicture}};
-      \draw [-> ] (c.south) -- (d.north);
-
-      \begin{scope}[shift={(4,-8)},scale=0.6,line width=0.05mm]
-        \node[latent,scale=0.7] at (3.5,3) (dx){DSL};
-        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
-        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
-        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
-        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
-        \draw [->] (zp1.south) -- (xp1.north);
-        \draw [->] (dx.south) -- (zp1.north);
-        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
-        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
-        \draw [->] (zp1.south) -- (xp1.north);
-        \draw [->] (dx.south) -- (zp1.north);
-        \draw [->] (dx.south) -- (zp.north);
-        \draw [->] (zp.south) -- (xp.north);
-      \end{scope}
-
-      \end{scope}
-
-    %% center spiral
-    \begin{scope}[shift={(3.25,-7.9)},scale=0.8]    
-      \spiral{(3.5,1)}{3.5}
-      \node[latent,scale=1] at (3.5,3) (dx){DSL};
-      \node[latent,scale=1] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
-      \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
-      \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
-      \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
-      \draw [->] (zp1.south) -- (xp1.north);
-      \draw [->] (dx.south) -- (zp1.north);
-      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-
-      \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
-      \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
-      \draw [->] (zp1.south) -- (xp1.north);
-      \draw [->] (dx.south) -- (zp1.north);
-      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
-
-      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
-      \draw [->] (dx.south) -- (zp.north);
-      \draw [->] (zp.south) -- (xp.north);
-
-      \node at ([yshift=-0.6cm]xp.south) {\legend};
-
-    \end{scope}
-
-  \end{tikzpicture}
-  \caption{}\label{threeCycles}
+  \includegraphics[width = \textwidth]{figures/DeepLogo.pdf} 
   \end{figure}
@@
 
+%% \section{Introduction: Expertise}
+
+%% Human learners acquire expertise in a wide range of domains: some of
+%% us become experts in calculus, or cooking, or biology, music, tennis,
+%% or software engineering, to name just a few, and every child develops
+%% expertise in natural language, intuitive physics~\cite{Pete?},
+%% intuitive psychology (theory-of-mind), and motor control.  This
+%% picture contrasts sharply with the current state of machine
+%% intelligence, where a machine is built to be an expert in a single
+%% domain, like boardgames~\cite{alphaGo}, medical
+%% diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
+%% proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
+%% Thus an outstanding challenge in the long-term program of building
+%% more humanlike machines is to develop an algorithm that, like people,
+%% autonomously acquires expertise across many different kinds of
+%% domains.
+
+%% We contribute a model of the development of expertise that combines
+%% two key ingredients.  First, an expert needs a sufficiently expressive
+%% knowledge representation.  Following a long tradition in cognitive
+%% science and AI, we represent knowledge as programs, which prior work
+%% has used to represent expertise in recognition and generation of
+%% handwriting and speech~\cite{lake2015human}, intuitive theories (of
+%% kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
+%% grammar and
+%% semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
+%% Second, atop this program representation, experts possess two kinds of
+%% domain expertise.  They have at their disposal a powerful, yet
+%% specialized, repertoire of concepts and abstractions: e.g., in
+%% architecture, these are concepts like `arch' or `foundation'; in
+%% software engineering these are libraries of code and domain-specific
+%% languages (DSLs).  Here our model represents knowledge as programs,
+%% and so we identify this kind of expertise with a DSL.  Experts also
+%% have knowledge of when and how to deploy these domain-specific
+%% concepts efficiently when solving new problems: e.g., mathematicians
+%% efficiently search the space of proofs, intuiting which lemmas are
+%% appropriate when; expert chefs intuit which compositions of
+%% ingredients are likely tasty, before they actually start cooking.  For
+%% our model, this aspect of expertise corresponds to the ability to
+%% quickly assemble new, useful programs out of its DSL.
+
+%% We integrate these ideas into a model called DreamCoder which acquires
+%% expertise through a novel kind of wake/sleep or `dream' learning. The
+%% model iterates through wake cycles -- where it solves problems by
+%% writing programs -- and a pair of sleep cycles: a sleep cycle that
+%% grows its DSL by replaying experiences from waking and consolidating
+%% them into new abstractions, and a sleep cycle that improves its
+%% knowledge of how to write programs by training a neural network on
+%% replayed experiences as well as `dreams', or samples, from its DSL.
+
+%% DreamCoder builds on multiple generations of AI research, going back
+%% to the 1960's~\cite{solomonoff1964formal} when program-learning was
+%% proposed as a paradigm for general AI.  Broadly speaking recent work
+%% has either developed neural approaches for learning to efficiently
+%% deploy a fixed
+%% DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
+%% developed symbolic approaches for representing and searching through
+%% spaces of
+%% programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
+%% We were motivated by approaches that learn or grow the
+%% DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
+%% Our goal with DreamCoder is to show that the combination of
+%% neurally-guided search and DSL learning is a uniquely powerful way of
+%% building systems that, like human learners, autonomously acquire the
+%% expertise needed to navigate a new domain of problems.
+
+\section{Introduction}
+
+An age-old dream within AI is a machine that learns and reasons by
+writing its own programs.  This vision stretches back to the
+1960's~\cite{solomonoff1964formal} and and is central to much of
+what it would take to build machines that learn and think like humans.
+Computational models of cognition often explain the flexibility and
+richness of human thinking in terms of program learning: from everyday
+thinking and problem solving (motor program induction as an account of
+recognition and generation of handwriting and
+speech~\cite{lake2015human}; functional programs as a model of natural
+language semantics~\cite{SOMETHING}) to learning problems that unfold
+over longer developmental time scales: the child's acquisition of
+intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
+natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
+just a few.  An outstanding challenge, however, is to engineer
+program-learners that display the same level of domain-generality as
+the humans they are meant to model.
+
+Recent program-learning systems developed within the AI and machine
+learning community are impressive along many dimensions, authoring
+programs for problem domains like drawing
+pictures~\cite{spiral,ellis2017learning}, transforming
+text~\cite{gulwani2011automating} and numerical
+sequences~\cite{balog2016deepcoder}, and reasoning over common sense
+knowledge bases~\cite{muggleton2015meta}.  These systems work in
+different ways, but typically hinge upon a carefully hand-engineered
+Domain Specific Language (DSL).  The DSL restricts the space of
+programs to contain the kinds of concepts needed for one specific
+domain.  For example, a picture-drawing DSL could include concepts
+like circles and spirals, and a DSL for numerical sequences could
+include sorting and reversing lists of numbers.  Modern systems also
+learn how to efficiently deploy the DSL on new
+problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
+unlike human learners -- do not discover the underlying system of
+concepts needed to navigate the domain.
+
+We contribute a program-induction system that learns the
+domain-specific concepts (DSL) while jointly learning how to use those
+concepts.  This joint learning problem models two complementary
+notions of domain expertise: (1) domain experts have at their disposal
+a powerful, yet specialized repertoire of concepts and abstractions
+(analogous to the DSL) while also (2) having accurate intuitions about
+when and how to use those concepts to solve new problems.
+Representative domains,
+along with DSLs we learn for them,
+are shown in Figure~\ref{exampleDSL}.
+
+We call our system `DreamCoder' because it acquires these two kinds of
+expertise through a novel kind of wake/sleep or `dream'
+learning~\cite{hinton1995wake}, iterating through a wake cycle --
+where it solves problems by writing programs -- and a pair of sleep
+cycles, both of which are loosely biologically inspired by actual
+human sleep.  The first sleep cycle, which we refer to as
+\textbf{consolidation}, grows the DSL by replying experiences from
+waking and consolidating them into new code abstractions.  This cycle
+is inspired by the formation of abstractions during sleep memory
+consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
+refer to as \textbf{dreaming}, improves the agents knowledge of how to
+write programs by training a neural network to help search for
+programs. The neural net is trained on replayed experiences as well as
+`fantasies', or samples, from the DSL.  These two kinds of dreams are
+inspired by the distinct episodic replay and hallucination components
+of dream sleep~\cite{fosse2003dreaming}.
+
+This dream-learning architecture brings together two lines of prior work,
+both of which have been separately influential within
+artificial intelligence.
+One line of work
+considers the problem of learning new concepts, abstractions, or
+`options' from
+experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
+while the other line of work considers the problem of learning how to
+deploy those concepts
+efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
+goal with DreamCoder is to show that the combination of these ideas is
+uniquely powerful, and pushes us toward program-writing systems that,
+like human learners, autonomously acquire the expertise needed to
+navigate a new domain of problems.
+
+Because any learning problem can in principle be cast as program
+induction, it is important to delimit our focus.  In contrast to
+computer assisted programming~\cite{solar2008program} or genetic
+programming~\cite{DBLP:books/daglib/0070933}, our goal is not to
+automate software engineering, or to synthesize large bodies of code
+starting from scratch.  Ours is a basic AI goal: capturing the human
+ability to learn to think flexibly and efficiently in new domains ---
+to learn what you need to know about a domain so you don't have to
+solve new problems starting from scratch.  We are focused on problems
+that people solve relatively quickly, once they acquire the relevant
+domain expertise.  These correspond to tasks solved by short programs
+--- if you have an expressive DSL.
+
 \begin{figure}\centering
@@
     \newcommand{\noArguments}[1]{
-      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
+      \mcode{{\color{\inventionColor}$f_{#1}$}$\,=\,$}
     }
@@
       \task{1}{2,3}{
+        \textsc{Reverse}\\
+        \mcode{[9 2 4]$\to$[4 2 9]}\\
+        \mcode{[4 3]$\to$[3 4]}\\
+        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
+        \mcode{[]$\to$[]}
+      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}}
+      \nextTask{2}{t1}{
         \textsc{Take if $\geq 5$}\\
@@
       }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}}
-      \nextTask{2}{t1}{
-        \textsc{Reverse}\\
-        \mcode{[9 2 4]$\to$[4 2 9]}\\
-        \mcode{[4 3]$\to$[3 4]}\\
-        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
-        \mcode{[]$\to$[]}
-      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}}
 
@@
 
-      \vinvention{filter}{p1}{
+      %% \vinvention{filter}{p1}{
+      %%   \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
+      %% }{higher-order `filter'}
+      \vinvention{single}{p1}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
+
+      \vinvention{three}{single}{\footnotesize \noArguments{1}\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
+      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
+      \hinvention{filter}{single}{
         \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
       }{higher-order `filter'}
-
-      \vinvention{three}{filter}{\footnotesize \noArguments{1}\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
-      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
-      \hinvention{single}{filter}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
-      \vinvention{for}{single}{\noArguments{4}\\\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
-      \vinvention{Max}{for}{\name{5}{l}\\\mcode{%
+      \vinvention{for}{filter}{\noArguments{4}\\\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
+      \vinvention{weird}{for}{\name{8}{p$,$l}\mcode{
+          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0)))}}{Higher-order incrementer}                 
+      \hinvention{largest}{filter}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
+      \vinvention{Max}{largest}{\name{5}{l}\\\mcode{%
           (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
-      \hinvention{largest}{single}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
-      \vinvention{count}{largest}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
-      \vinvention{weird}{count}{\name{8}{p$,$l}\mcode{
-          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0)))}}{Higher-order incrementer}
+      \vinvention{count}{Max}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
 
       \node[rotate=90,align=center] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
-      \node[rotate=90,align=center] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
+      \node(programLabel)[rotate=90,align=center] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
       \node[rotate=90,align=center] at ([xshift=-2cm]three.center) {\textbf{\textsc{Domain Specific Language}}\\
         \small Learned routines in {\color{orange}orange}};
+      \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: List Processing}}};
+
+      \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
+    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};
 
       \draw[->] (for.west) -- (three.east);
-      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
-      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
+      \draw[->] (largest.east) to[out=0,in=0] (Max.east);
+      \draw[->] ([yshift=-0.cm]largest.west) -- ([yshift=0.2cm]filter.east);
       \draw[->] (Max.west) -- (filter.east);
@@
 
-      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]p1.south) to[out=250] (filter.north);
-      \draw[->] ([yshift=0.0cm,xshift=0.3cm]p1.south) to[out=350,in=100] (for.west);
-      \draw[->] ([xshift=-0.7cm]p2.south) -- (single.north);
+      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]p2.south) to[out=250] (filter.north);
+      \draw[->] ([yshift=0.0cm,xshift=0.3cm]p2.south) to[out=350,in=0] (for.east);
+      \draw[->] ([xshift=-0.7cm]p1.south) -- (single.north);
       \draw[->] ([xshift=-1.3cm,yshift=0.6cm]p3.south) to[out=240] (largest.north);
@@
       \small Learned routines in {\color{orange}orange}};
+          \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: Graphics Programming}}};
 
@@
   \end{tikzpicture}
-  \caption{Two of the six domains we apply \system to.
+  \caption{Two of the six domains we apply our system to.
 Agent observes tasks (top rows) which it solves by writing programs (middle rows) while jointly growing a library (DSL; bottom rows).
 Learned DSLs rediscover multiple higher-order functions (\code{filter} for list functions and rotational symmetry for generative graphics).
-Learned DSL components call each other (arrows).}
+Learned DSL components call each other (arrows).}  \label{exampleDSL}
 \end{figure}
 
-%% \section{Introduction: Expertise}
-
-%% Human learners acquire expertise in a wide range of domains: some of
-%% us become experts in calculus, or cooking, or biology, music, tennis,
-%% or software engineering, to name just a few, and every child develops
-%% expertise in natural language, intuitive physics~\cite{Pete?},
-%% intuitive psychology (theory-of-mind), and motor control.  This
-%% picture contrasts sharply with the current state of machine
-%% intelligence, where a machine is built to be an expert in a single
-%% domain, like boardgames~\cite{alphaGo}, medical
-%% diagnosis~\cite{mycin,CNNsForDiseaseDetection}, theorem
-%% proving~\cite{eurisko}, or visual object recognition~\cite{alexNet}.
-%% Thus an outstanding challenge in the long-term program of building
-%% more humanlike machines is to develop an algorithm that, like people,
-%% autonomously acquires expertise across many different kinds of
-%% domains.
-
-%% We contribute a model of the development of expertise that combines
-%% two key ingredients.  First, an expert needs a sufficiently expressive
-%% knowledge representation.  Following a long tradition in cognitive
-%% science and AI, we represent knowledge as programs, which prior work
-%% has used to represent expertise in recognition and generation of
-%% handwriting and speech~\cite{lake2015human}, intuitive theories (of
-%% kinship, taxonomy, etc.)~\cite{Ullman2012}, and natural language
-%% grammar and
-%% semantics~\cite{DBLP:journals/cogsr/SchmidK11,piantadosi2011learning}.
-%% Second, atop this program representation, experts possess two kinds of
-%% domain expertise.  They have at their disposal a powerful, yet
-%% specialized, repertoire of concepts and abstractions: e.g., in
-%% architecture, these are concepts like `arch' or `foundation'; in
-%% software engineering these are libraries of code and domain-specific
-%% languages (DSLs).  Here our model represents knowledge as programs,
-%% and so we identify this kind of expertise with a DSL.  Experts also
-%% have knowledge of when and how to deploy these domain-specific
-%% concepts efficiently when solving new problems: e.g., mathematicians
-%% efficiently search the space of proofs, intuiting which lemmas are
-%% appropriate when; expert chefs intuit which compositions of
-%% ingredients are likely tasty, before they actually start cooking.  For
-%% our model, this aspect of expertise corresponds to the ability to
-%% quickly assemble new, useful programs out of its DSL.
-
-%% We integrate these ideas into a model called DreamCoder which acquires
-%% expertise through a novel kind of wake/sleep or `dream' learning. The
-%% model iterates through wake cycles -- where it solves problems by
-%% writing programs -- and a pair of sleep cycles: a sleep cycle that
-%% grows its DSL by replaying experiences from waking and consolidating
-%% them into new abstractions, and a sleep cycle that improves its
-%% knowledge of how to write programs by training a neural network on
-%% replayed experiences as well as `dreams', or samples, from its DSL.
-
-%% DreamCoder builds on multiple generations of AI research, going back
-%% to the 1960's~\cite{solomonoff1964formal} when program-learning was
-%% proposed as a paradigm for general AI.  Broadly speaking recent work
-%% has either developed neural approaches for learning to efficiently
-%% deploy a fixed
-%% DSL~\cite{devlin2017robustfill,balog2016deepcoder,NGDS,spiral}, or
-%% developed symbolic approaches for representing and searching through
-%% spaces of
-%% programs~\cite{gulwani2011automating,solar2008program,DBLP:books/daglib/0070933}.
-%% We were motivated by approaches that learn or grow the
-%% DSL~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning}.
-%% Our goal with DreamCoder is to show that the combination of
-%% neurally-guided search and DSL learning is a uniquely powerful way of
-%% building systems that, like human learners, autonomously acquire the
-%% expertise needed to navigate a new domain of problems.
-
-\section{Introduction}
-
-An age-old dream within AI is a machine that learns and reasons by
-writing its own programs.  This vision stretches back to the
-1960's~\cite{solomonoff1964formal} and cuts to the core of much of
-what it would take to build machines that learn and think like humans.
-Computational models of cognition often explain the flexibility and
-richness of human thinking in terms of program learning: from everyday
-thinking and problem solving (motor program induction as an account of
-recognition and generation of handwriting and
-speech~\cite{lake2015human}; functional programs as a model of natural
-language semantics~\cite{SOMETHING}) to learning problems that unfold
-over longer developmental time scales: the child's acquisition of
-intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
-natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
-just a few.  An outstanding challenge, however, is to engineer
-program-learners that display the same level of domain-generality as
-the humans they are meant to model.
-
-Recent program-learning systems developed within the AI and machine
-learning community are impressive along many dimensions, authoring
-programs for problem domains like drawing
-pictures~\cite{spiral,ellis2017learning}, transforming
-text~\cite{gulwani2011automating} and numerical
-sequences~\cite{balog2016deepcoder}, and reasoning over common sense
-knowledge bases~\cite{muggleton2015meta}.  These systems work in
-different ways, but typically hinge upon a carefully hand-engineered
-Domain Specific Language (DSL).  The DSL restricts the space of
-programs to contain the kinds of concepts needed for one specific
-domain.  For example, a picture-drawing DSL could include concepts
-like circles and spirals, and a DSL for numerical sequences could
-include sorting and reversing lists of numbers.  Modern systems also
-learn how to efficiently deploy the DSL on new
-problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
-unlike human learners -- do not discover the underlying system of
-concepts needed to navigate the domain.
-
-We contribute a program-induction system that learns the
-domain-specific concepts (DSL) while jointly learning how to use those
-concepts.  This joint learning problem models two complementary
-notions of domain expertise: (1) domain experts have at their disposal
-a powerful, yet specialized repertoire of concepts and abstractions
-(analogous to the DSL) while also (2) having accurate intuitions about
-when and how to use those concepts to solve new problems.
-
-We call our system `DreamCoder' because it acquires these two kinds of
-expertise through a novel kind of wake/sleep or `dream'
-learning~\cite{hinton1995wake}, iterating through a wake cycle --
-where it solves problems by writing programs -- and a pair of sleep
-cycles, both of which are loosely biologically inspired by actual
-human sleep.  The first sleep cycle, which we refer to as
-\textbf{memory consolidation}, grows the DSL by replying experiences
-from waking and consolidating them into new code abstractions.  This
-cycle is inspired by the formation of abstractions during sleep memory
-consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
-refer to as \textbf{dreaming}, improves the agents knowledge of how to
-write programs by training a neural network on replayed experiences as
-well as `fantasies', or samples, from the DSL.  These two kinds of
-dreams are inspired by the distinct episodic replay and hallucination
-components of dream sleep~\cite{fosse2003dreaming}.
-
-This dream-learning architecture brings together two lines of prior work.
-One line of work
-considers the problem of learning new concepts, abstractions, or
-`options' from
-experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
-while the other line of work considers the problem of learning how to
-deploy those concepts
-efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
-goal with DreamCoder is to show that the combination of these ideas is
-uniquely powerful, and pushes us toward program-writing systems that,
-like human learners, autonomously acquire the expertise needed to
-navigate a new domain of problems.
-
-\section{Deep Wake/Sleep Program Induction}
+\section{Wake/Sleep Program Induction}\label{overviewSection}
 
 \system~takes as its goal to acquire domain-specific expertise, which
-it learns by solving a collection of programming \textbf{tasks} --
-typically about a hundred.  It alternatingly finds programs that solve
-tasks (Wake -- Figure~\ref{threeCycles} top); improving its DSL by
-discovering and reusing domain-specific subroutines (Memory
-Consolidation -- Figure~\ref{threeCycles} left); and training a neural
-network that efficiently guides search for programs in the DSL
-(Dreaming -- Figure~\ref{threeCycles} right).  The learned DSL
-effectively encodes a prior on programs likely to solve tasks in the
-domain, while the neural net looks at the example input-output pairs
-for a specific task and produces a ``posterior'' for programs likely
-to solve that specific task (Figure~\ref{threeCycles} middle).  The
-neural network thus functions as a \textbf{recognition model}
+it learns by solving a collection of programming \textbf{tasks}.  It
+alternatingly finds programs that solve tasks (Wake --
+Figure~\ref{threeCycles} top); improves its DSL by discovering and
+reusing domain-specific subroutines (Consolidation --
+Figure~\ref{threeCycles} left); and trains a neural network that
+efficiently guides search for programs in the DSL (Dreaming --
+Figure~\ref{threeCycles} right).  The learned DSL acts as a a prior on
+programs likely to solve tasks in the domain, while the neural net
+looks at a specific task and produces a ``posterior'' for programs
+likely to solve that specific task (Figure~\ref{threeCycles} middle).
+The neural network thus functions as a \textbf{recognition model}
 supporting a form of approximate Bayesian program induction, jointly
@@
 tractable even as the DSL (and hence the search space for programs)
-expands.
-
-%% \section{Introduction: NIPS}
-
-%% Much of everyday human thinking and learning can be understood in
-%% terms of program induction: constructing a procedure that maps inputs
-%% to desired outputs, based on observing example input-output pairs.
-%% People can induce programs flexibly across many different domains,
-%% often from just one or a few examples.  For instance, if shown that a
-%% text-editing program should map ``Jane Morris Goodall'' to
-%% ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey'' to
-%% ``R. E. Leakey''; if instead the first input mapped to ``Dr. Jane'' or
-%% ``Goodall, Jane'', we might guess the latter should map to
-%% ``Dr. Richard'' or ``Leakey, Richard'', respectively.
-
-%% The FlashFill system~\cite{gulwani2011automating} embedded in
-%% Microsoft Excel solves problems such as these and is probably the best
-%% known practical program-induction algorithm, but researchers in
-%% programming languages and AI have had successes in many domains, such as handwriting recognition and
-%% generation~\cite{lake2015human}, procedural
-%% graphics~\cite{ellis2017learning,DBLP:journals/corr/abs-1804-01118}, question
-%% answering~\cite{johnson2017clevr} and robot motion
-%% planning~\cite{devlin2017neural}.  These systems
-%% work in different ways, but most hinge upon a carefully engineered
-%% \textbf{Domain Specific Language (DSL)}.  This is especially true for
-%% systems like FlashFill that induce a wide range of programs
-%% very quickly, in a few seconds or less.  DSLs constrain the search
-%% over programs with strong prior knowledge in the form of a restricted
-%% inventory of programming primitives finely tuned to the domain: for
-%% text editing, these are operations like appending and
-%% splitting on characters.
-
-%% In this work, we consider the problem of building agents that learn to solve
-%% program induction tasks, and also the problem of acquiring the prior knowledge
-%% necessary to quickly solve these tasks in a new domain.  Representative problems
-%% in three domains are shown in Table~\ref{initialExampleDSL}.  Our solution is an
-%% algorithm that grows or boostraps a DSL while jointly training a neural network
-%% to help write programs in the increasingly rich DSL.
-
-\begin{table*}%[t!]
-  \makebox[\textwidth][c]{ \scriptsize \tabcolsep=4pt
-    \renewcommand\code\texttt \renewcommand\codechar[1]{\texttt{"#1"}}
-    \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
-    \toprule
-    &{\normalsize List Processing}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
-      \begin{tabular}{cc}
-        \begin{tabular}{c}
-          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
-          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
-          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
-          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
-          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
-          \\
-          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
-          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
-        \end{tabular}
-        &
-        \hspace{-0.3cm}\begin{tabular}{c}
-          \code{[7\, 3]}$\to $\code{False}                              \\
-          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
-          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
-          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
-          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
-          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
-        \end{tabular}
-      \end{tabular}
-    }
-    &
-    \hspace{-0.5cm}\begin{tabular}{c}
-      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
-      +83 973-831$\to $83.973.831\\
-      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
-      }}\\
-      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
-      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
-      ~\\
-      Temple Anna H $\to $TAH\\
-      Lara Gregori$\to $LG\\
-      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
-    \end{tabular}
-    &
-    \begin{tabular}{cc}
-      \includegraphics[width = 3em]{figures/functions/4.png}&
-      \includegraphics[width = 3em]{figures/functions/146}\\
-      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
-      ~\\
-      \includegraphics[width = 3em]{figures/functions/112.png}&
-        \includegraphics[width = 3em]{figures/functions/92.png}
-      \\
-      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\
-
-    \end{tabular}
-    ~\\
-    \midrule
-    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
-    \hspace{0cm}\begin{tabular}{l}
-      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
-      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
-      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
-      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
-      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
-      %NOTE: this is the actual invention, but I removed a redundant lambda
-      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
-      %% \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
-      %% \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
-      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
-      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
-      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
-      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
-      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
-      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
-      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
-      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
-      \phantom{$f_1(\ell,$}
-      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
-      \hspace{\helpSize}($f_3$: \emph{Whether $\ell$ contains }\code{k})\\
-    \end{tabular}&
-
-  \hspace{0.5cm}\begin{tabular}{l}
-    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
-    (x)}}\\
-    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
-      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
-      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
-      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
-        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
-      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
-      }}\\
-      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
-        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
-      %%   \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
-      %% \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
-  \end{tabular}&
-
-  \begin{tabular}{l}
-    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
-    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
-    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
-    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
-    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
-    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
-    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
-    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
-    \hspace{\helpSize}\emph{($f_6$: rational function)}\\
-
-  \end{tabular}
-  \\\bottomrule\\
-\end{tabular}}\vspace{-0.5cm} 
-\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
-Bottom: Several examples from learned DSL.
-Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}%\vspace{-0.5cm}
-\end{table*}
-
-Because any learning problem can in principle be cast as
-program induction, it is important to delimit our focus.  In contrast
-to computer assisted programming~\cite{solar2008program} or genetic
-programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
-engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
-capturing the human ability to learn to think flexibly and efficiently
-in new domains --- to learn what you need to know about a domain so you
-don't have to solve new problems starting from scratch.  We are
-focused on problems that people solve relatively
-quickly, once they acquire the relevant domain expertise.  These
-correspond to tasks solved by short programs --- if you have an
-expressive DSL.  %% Even with a good DSL, program search may be
-%% intractable;
-%
-% , and adding new library routines to the DSL only broadens the
-% search space;
-%
-%% so we train a
-%% neural network to guide the search procedure.
-
-%\section{Overview of \system}
-
-Our algorithm takes inspiration from several ways that skilled human
-programmers have learned to code: Skilled coders build libraries of
-reusable subroutines that are shared across related programming tasks,
-and can be composed to generate increasingly complex and powerful
-subroutines.  In text editing, a good library should support routines
-for splitting on characters, but also specialize these routines to
-split on particular characters such as spaces or commas that are
-frequently used to delimit substrings across tasks.  Skilled coders
-also learn to recognize what kinds of programming idioms and library
-routines would be useful for solving the task at hand, even if they
-cannot instantly work out the details.  In text editing, one might
-learn that if outputs are consistently shorter than inputs, removing
-characters is likely to be part of the solution; if every output
-contains a constant substring (e.g., ``Dr.''), inserting or appending
-that constant string is likely to be a subroutine.
-
-Our algorithm is called \system because it is based on a
-novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
-between ``wake'' and ``sleep'' phases to achieve three goals: finding
-programs that solve tasks; creating a DSL by discovering
-and reusing domain-specific subroutines; and training a neural network
-that efficiently guides search for programs in the DSL.
-The learned DSL effectively encodes a prior on programs likely to
-solve tasks in the domain, while the neural net looks at the example
-input-output pairs for a specific task and produces a ``posterior''
-for programs likely to solve that specific task.  The neural network
-thus functions as a \textbf{recognition model} supporting a form of
-approximate Bayesian program induction, jointly trained with a
-\textbf{generative model} for programs encoded in the DSL, in the
-spirit of the Helmholtz machine~\cite{hinton1995wake}. The
-recognition model ensures that searching for programs remains
-tractable even as the DSL (and hence the search space for programs)
-expands.
-
-%% Our \system
-%% algorithm incorporates these insights by iterating
-Concretely, our algorithm iterates through three phases. 
-The \textbf{Wake} phase takes a given set of \textbf{tasks}, typically
-several hundred, and searches for compact programs that solve these tasks,
-guided by the current DSL and neural network.
-The \textbf{Sleep-G} phase grows the DSL (or \textbf{G}enerative model),
-which allows the agent to more compactly
-write programs in the domain. We modify the structure of the DSL by
-discovering regularities across programs, compressing
-them to distill out common abstractions across  programs discovered during waking.\footnote{This is loosely biologically
-    inspired by the formation of abstractions during sleep memory
-    consolidation~\cite{DUDAI201520}}
-The \textbf{Sleep-R} phase improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
-write programs in the current DSL, in the spirit of ``amortized'' or
-``compiled'' inference~\cite{le2016inference}.
-We train the recognition model on  two data sources: samples
-from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
-programs found during waking.%%  As the DSL
-%% improves and as search finds more programs, the recognition model gets
-%% both more data to train on, and better data.
-\footnote{These two sources are also
-    loosely biologically inspired by the distinct episodic replay and
-    hallucination components of dream sleep~\cite{fosse2003dreaming}}
-
-\textbf{Brisk overview of related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
-both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
-FlashFill:~\cite{gulwani2011automating}),
-neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
-symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
-A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
-which also learns a library of subroutines.
-We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
-%%   \setlength\itemsep{0}
-%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
-%%   \item We give an algorithm for learning DSLs, built on a formalism known
-%%     as Fragment Grammars~\cite{tim}; and 
-%%   \item We give a hierarchical Bayesian framing of the problem that allows a
-%%     joint inference of the DSL and recognition model.
-%% \end{itemize}
-\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
-(2) We give an algorithm for learning DSLs,
-built on a formalism known as Fragment Grammars~\cite{tim};
-and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.
-up
-
-\section{The \system Algorithm}
-
- We first mathematically describe our 3-step algorithm as
- an inference procedure for a hierarchical Bayesian model (Section~\ref{mathematicalFraming}),
-  and then describe each step algorithmically in detail (Section~\ref{explorationSection}-\ref{grammarInductionSection}).
-
- \subsection{Hierarchical Bayesian Framing}\label{mathematicalFraming}
-
-\system takes as input a set of \textbf{tasks}, written $X$, each of which is a program synthesis problem.
-It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.\footnote{For example, for string editing,
-  the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
-  and 0 otherwise; when learning a generative model or probabilistic program, the likelihood is
-the probability of the program sampling the observation.}
-Its goal is to solve each of the tasks by writing a program,
-and also to infer a DSL, written $\mathcal{D}$.
-We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
-$(\mathcal{D},\theta)$ define a generative model over programs.
-We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
-Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
+expands.  The generative model, or DSL, distills out common
+abstractions across programs discovered during waking, growing a
+network of increasingly deep and specialized domain-specific concepts
+(Figure~\ref{exampleDSL}, bottom rows).
+
+Viewed as an inference problem, our goal is to start with a collection
+of tasks, written $X$, and infer both a program for each task, as well
+as a distribution over programs encoded by a DSL, written
+$\mathcal{D}$.  We equip $\mathcal{D}$ with a learned weight vector
+$\theta$, and together $(\mathcal{D},\theta)$ define a generative
+model over programs (see Appendix~\ref{sampleProgram}).  We perform
+maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given
+$X$.  Writing $J$ for the joint probability of $(\mathcal{D},\theta)$
+and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
@@
 \end{align}
-
-%% Inference in this model is
-%% difficult because the programs are unobserved,
-%% and so we must solve a hard search problem to recover them. To make
-%% search tractable we learn a bottom-up \emph{recognition
-%%   model} (written $q(\cdot )$, Fig.~\ref{graphicalModel}(b)).
-%% The recognition model $q(\cdot )$ is a neural
-%% network that regresses from input/output pairs to a distribution over
-%% programs likely to explain the input/outputs. We can also view $q(\cdot )$ as implementing an amortized
-%% inference
-%% scheme~\cite{le2016inference}.
-%%  The neural recognition model and the
-%% generative model embodied in the DSL jointly train each other, as they
-%% iteratively learn to solve more programming tasks.
-
+where $\probability[x|p]$ scores the likelihood of a task
+$x\in X$ given a program $p$.\footnote{For example, for list
+  processing, the likelihood is 1 if the program predicts the observed
+  outputs on the observed inputs, and 0 otherwise; when learning a
+  generative model or probabilistic program, the likelihood is the
+  probability of the program sampling the observation.}
 The above equations summarize the problem from the point of view of an ideal Bayesian learner.
@@
 So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
+
 \begin{definition}
-   A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
+A \textbf{frontier of task $x$}, written $\mathcal{F}_x$,
 is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
 \end{definition}
-Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
+\noindent Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
 \begin{align}
@@
 \end{align}
-We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
-and $(\mathcal{D},\theta)$ (\textbf{Sleep-G}):
+Our wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
+and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
 \\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
@@
 %% where  is large.
-\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
+\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
 and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
-%To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.
 
@@
 of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $Q(\cdot |\cdot )$,
-during the \textbf{Sleep-R} phase: $Q$ is trained to approximate the
+during the \textbf{Dreaming} phase: $Q$ is trained to approximate the
 posterior over programs, $Q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
@@
 
-\noindent\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
+\noindent\textbf{Sleep (Dreaming): tractably maxing $\lowerBound$ w.r.t. the
   frontiers.}  Here we train %% a neural network, $q$, to predict a
@@
 into the frontiers will most increase $\lowerBound$.
-This neural net is trained on two sources of data: samples, or `dreams',
-drawn from the DSL,
-as well as
-replays of (program, task) pairs discovered during waking.
-
-Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figures~\ref{feeding}~and~\ref{bayesianSummary}).
+
+Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others.
 As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
-the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
+the Consolidation cycle has more data from which to learn the DSL; and, because the recognition model is trained on both
 samples from the DSL and programs found during waking,
 the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
-\begin{figure}\centering
-  \begin{tikzpicture}
-    \begin{scope}[shift = {(1,-1)}]
-    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
-    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
-    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};
-
-    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
-    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
-    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);
-
-    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
-    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
+\begin{figure}
+  \begin{tikzpicture}[line width=0.4mm]
+    \draw[fill=white!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
+    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
+
+    \begin{scope}[shift={(0.5,0.5)}]
+      \node[align=right] at (0,-0.5) (d){\textbf{DSL}\\
+        $f_1$, $f_2$, ...};
+      \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
+        \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
+        \footnotesize    \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} };
+
+      \node at ([xshift = 1.25cm]t.east) (nn){\NeuralNetwork{0.25}};
+      \node[align = center, text width = 1cm] at ([yshift = 0.7cm,xshift=0cm]nn.north) {\baselineskip=0pt \small Recog. model\par};
+      \draw [red,-{>[scale=0.2]}] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+      \node[draw,rounded corners, align=center, inner sep = 10] at ([xshift = 4.2cm,yshift = 1cm]t.east) (s){Neurally-Guided\\ Enumerative Search};
+
+      \draw [red,->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+      \draw [->,rounded corners,] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+      \node[align=left] at ([xshift=3cm]s.east) (f) {\textbf{Programs:}\\
+        \small    \code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}\\
+        \small $\cdots\cdots\cdots$};
+      \draw [->  ] (s.east) -- (f.west);
+
+      \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
     \end{scope}
+    \begin{scope}[shift={(9.4,-3.5)},scale=0.6,line width=0.05mm]
+      \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
+      \node[latent,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+      \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
+      \node[latent,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
+      \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+      \node[latent,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
+      \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->] (dx.south) -- (zp.north);
+      \draw [->] (zp.south) -- (xp.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+    \end{scope}
+
+    \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
+    \draw[fill=white!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
+    \node at (12,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
+    \draw[fill=white!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);
+
+    \begin{scope}[shift={(9.5,-4.5)}]
+      \node(dreaming) at (2,-1) {\underline{Fantasies}};
+      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
+      \node at ([xshift=1.5cm]d.east) (p1){prog.};
+      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
+      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
+      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
+      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
+        \NeuralNetwork{0.17}};
+      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
+      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
+
+      \node(replay) at (2,-5) {\underline{Replays}};
+      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
+      \node at ([xshift=1.5cm]d.east) (p1){prog.};
+      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
+      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
+      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
+      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
+        \NeuralNetwork{0.17}};
+      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
+      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
+
+      \begin{scope}[shift={(-3.8,-8)},scale=0.6,line width=0.05mm]
+        \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
+        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \draw [->] (dx.south) -- (zp.north);
+        \draw [->] (zp.south) -- (xp.north);
+        \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+        \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+      \end{scope}
+
+      \end{scope}
+
+    % memory consolidation
+    \begin{scope}[shift={(-2,-4.5)}]
+
+      %% defined routines for creating fragmented syntax trees
+      \renewcommand{\syntaxOne}[1]{
+        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
+          \node(l1) at (0,0) {};
+          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
+          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
+          \node(x1) at (0,-1) {\texttt{1}};
+          \draw[color=pop3] (l1.south) -- (p1.north);
+          \draw[color=pop3] (l1.south) -- (n1.north);
+          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+          \node(t) at (-0.5,0.5) {};
+          \draw (l1.south) -- (t.south);
+          \node(c) at (-1.5,-0.2) {\texttt{cons}};
+          \draw (t.south) -- (c.north);
         \end{tikzpicture}
-  \caption{\system solves for programs, the DSL, and a recognition model.
-Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
+      }
+      \renewcommand{\syntaxTo}[1]{
+        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
+            \node(l1) at (0,0) {};
+            \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
+            \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
+            \draw[color=pop3] (l1.south) -- (p1.north);
+            \draw[color=pop3] (l1.south) -- (n1.north);
+            \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+            \node(c) at (-0.5,-1.5) {\texttt{car}};
+            \node(z) at (0.5,-1.5) {\texttt{z}};
+            \draw (0,-1) -- (c.north);
+            \draw (0,-1) -- (z.north);
+        \end{tikzpicture}
+      }
+
+      \node[align=center,anchor=center] at (0.4,-1.2) (f1){\textbf{Progs. for Task 1}:\\\code{(+ (car z) 1)}};
+      \node[align=center] at ([xshift = 1.75cm]f1.east) (f2){\textbf{Progs. for Task 2}:\\\code{(cons (+ 1 1))}};
+      \node(s1) at ([yshift=-0.5cm]f1.south) {\syntaxOne{0.8}};
+      \node(s2) at ([yshift=-0.5cm]f2.south) {\syntaxTo{0.8}};
+      \node(c)[align=center,rectangle, rounded corners, draw, minimum width = 4cm, minimum height = 1.5cm, anchor = north] at ($(s1.south)!0.5!(s2.south) + (0,-1)$) {Refactoring Algorithm};
+      \draw [-> ] (s1.south) -- (s1.south|-c.north);
+      \draw [-> ] (s2.south) -- (s2.south|-c.north);
+
+      \node(d) at ([yshift = -1.8cm]c.south) {
+        \begin{tikzpicture}[scale=0.9,line width=0.5mm]
+          \node[align=center] at (0,0) {\textbf{new DSL} w/ \texttt{(+ x 1)}:};
+          \begin{scope}[shift={(0.6,-0.5)}]
+            \node[pop3](p1) at (-1,-1) {\texttt{+}};
+            \node[pop3](n1) at (0.6,-0.9) {\texttt{1}};
+            \node[pop3](a) at (0,-1) {\texttt{ }};
+            \draw[pop3] (0,0) -- (p1.north);
+            \draw[pop3] (0,0) -- (n1.north);
+            \draw[pop3] (-0.3,-0.3) -- (a.north);
+          \end{scope}
+      \end{tikzpicture}};
+      \draw [-> ] (c.south) -- (d.north);
+
+      \begin{scope}[shift={(4,-8)},scale=0.6,line width=0.05mm]
+        \node[latent,scale=0.7] at (3.5,3) (dx){DSL};
+        \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+        \node[obs,scale=0.7] at ([yshift=-1.45cm]zp) (xp) {task};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]zp) (zp1){prog};
+        \node[obs,scale=0.7] at ([xshift=-1.5cm]xp) (xp1) {task};
+        \draw [->] (zp1.south) -- (xp1.north);
+        \draw [->] (dx.south) -- (zp1.north);
+        \draw [->] (dx.south) -- (zp.north);
+        \draw [->] (zp.south) -- (xp.north);
+      \end{scope}
+
+      \end{scope}
+
+    %% center spiral
+    \begin{scope}[shift={(3.25,-7.9)},scale=0.8]    
+      \spiral{(3.5,1)}{3.5}
+      \node[latent,scale=1] at (3.5,3) (dx){DSL};
+      \node[latent,scale=1] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+      \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
+      \node[latent,scale=1] at ([xshift=2cm]zp) (zp1){prog};
+      \node[obs,scale=1] at ([xshift=2cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+      \node[latent,scale=1] at ([xshift=-2cm]zp) (zp1){prog};
+      \node[obs,scale=1] at ([xshift=-2cm]xp) (xp1) {task};
+      \draw [->] (zp1.south) -- (xp1.north);
+      \draw [->] (dx.south) -- (zp1.north);
+      \draw [->,red] (xp1.east) to[out = 30,in = -30] node(nn){} (zp1.east);
+
+      \draw [->,red] (xp.east) to[out = 30,in = -30] node(nn){} (zp.east);
+      \draw [->] (dx.south) -- (zp.north);
+      \draw [->] (zp.south) -- (xp.north);
+
+      \node at ([yshift=-0.6cm]xp.south) {\legend};
+
+    \end{scope}
+
+  \end{tikzpicture}
+  \caption{\textbf{Middle:} \system as a graphical model.
+Agent observes programming tasks (e.g., input/outputs for list processing or images for graphics programs), which it explains with latent programs, while jointly inferring a latent Domain Specific Language (DSL) capturing cross-program regularities.
+A neural network, called the \emph{recognition model} (red arrows) is trained to quickly infer programs with high posterior probability.
+\textbf{Top}: Wake phase infers programs while holding the DSL and recognition model fixed.
+\textbf{Left}: Sleep (Consolidation) phase infers DSL while holding the programs fixed by refactoring programs found during waking and extracting common components.
+\textbf{Right}: Sleep (Dreaming) phase trains recognition model to predict approximate posterior over programs conditioned on task.
+Trained on `Fantasies' (programs sampled from DSL) \& `Replays' (programs found during waking).}\label{threeCycles}
 \end{figure}
 
-\subsection{Wake: Searching for Programs}\label{explorationSection}
-
-During waking, the agent's goal is to search for programs solving the tasks.  We use the simple approach of enumerating programs from
-the DSL  in decreasing order of their probability according to the recognition model,
-and then checking if a program $p$ assigns positive
-probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
-into the frontier $\mathcal{F}_x$.
-
-To make this concrete we need to define what programs actually are and
-what form $\probability[p |\mathcal{D},\theta]$ takes.
-We represent programs as $\lambda$-calculus expressions.
-$\lambda$-calculus is a formalism for expressing functional programs
-that closely resembles Lisp,
-including variables, function application, and the ability to create new functions.
-Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
-Our programs are all strongly typed.
-We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
-used in functional programming languages like OCaml and Haskell.
-%% Type variables are always written using lowercase Greek letters
-%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
-%% and returns something of type $\beta$.
-%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
-%% has the type $\tau$.
-%% For example, to describe the type of the identity function
-%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
-%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
-%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
-%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
-%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
-%% See Supplement for more detail on program representation.
-We now define DSLs:
-
-\begin{definition}
-A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
-A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
-one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
-and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
-\end{definition}
-
-Together with its weight vector, a DSL defines a distribution over
-programs, $\probability[p|\mathcal{D},\theta]$.  We define this
-distribution by specifying a procedure for drawing samples from
-$\probability[p|\mathcal{D},\theta]$ (Algorithm~\ref{sampleProgram}).
-Care must be taken to ensure that variable scoping rules are obeyed
-and that programs are well-typed.  We ensure well-typed programs by
-performing Hindley-Milner type inference~\cite{pierce} during
-sampling, and assume that each task is annotated with the type of the
-program that will solve it.
-Appendix~\ref{enumerationAppendix}
-explains how we enumerate,
-rather than sample,
-programs generated by Algorithm~\ref{sampleProgram}.
- \begin{algorithm}
-   \caption{Generative model over programs}
-   \label{sampleProgram}
-   \begin{algorithmic}[1]
-     \State \textbf{function} sample$(\mathcal{D}, \theta, \tau)$:
-     \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, type $\tau$
-     \State \textbf{Output:} a program whose type unifies with $\tau$
-     \State  \textbf{return }sample$'(\mathcal{D}, \theta, \varnothing, \tau)$
-
-     \Statex
-     \State \textbf{function} sample$'(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
-     \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$\Comment{Environment $\mathcal{E}$ starts out as $\varnothing$}
-     \State \textbf{Output:} a program whose type unifies with $\tau$
-     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
-     \State var $\gets$ an unused variable name
-     \State body $\sim$ sample$'(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$\Comment{Recursively sample function body}
-     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
-     %\Endif
-     \Else\Comment{Build an application to give something w/ type $\tau$}
-     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
-     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
-     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
-     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
-       \theta_e&\text{ if }e\in \mathcal{D}\\
-\theta_{var}/|\text{variables}|&\text{ if }e\in \mathcal{E}
-       \end{cases}$
-     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
-     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
-     %   \State unify$(\tau,\beta)$
-     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
-     \State $a_k\sim\text{sample}'(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
-     \EndFor
-     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
-     \EndIf
-     \Statex
-     \Statex\textbf{where:}
-     \State yield$(\tau) = \begin{cases}
-       \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-       \tau   &\text{ otherwise.}
-     \end{cases}$ \Comment{Final return type of $\tau$}
-     \State  args$(\tau) = \begin{cases}
-       [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-       []   &\text{ otherwise.}
-     \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
-   \end{algorithmic}
- \end{algorithm}
+\subsection{Wake: Solving tasks}\label{explorationSection}
+
+During waking we enumerate programs from the DSL in decreasing order
+of their probability according to the recognition model, and then
+check if a program $p$ assigns positive probability to a task
+($\probability[x|p] > 0$); if so, we incorporate $p$ into the frontier
+$\mathcal{F}_x$.  We represent programs as polymorphicly-typed
+$\lambda$-calculus expressions, a representation closely resembling
+Lisp and functional languages like Haskell and OCaml, including
+variables, conditionals, higher-order recursive functions, and the
+ability to create new functions.
+
+%% Our programs are all
+%% strongly typed.  We use the Hindley-Milner polymorphic typing
+%% system~\cite{pierce} which is used in functional programming languages
+%% like OCaml and Haskell.  We now define DSLs:
+%% \begin{definition}
+%% A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
+%% A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
+%% one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
+%% and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.
+%% \end{definition}
+%% \noindent Together with its weight vector, a DSL defines a distribution over
+%% programs, $\probability[p|\mathcal{D},\theta]$.  We define this
+%% distribution by specifying a procedure for drawing samples from
+%% $\probability[p|\mathcal{D},\theta]$ (Appendix~\ref{generativeAppendix}).
+%% Care must be taken to ensure that variable scoping rules are obeyed
+%% and that programs are well-typed.  We ensure well-typed programs by
+%% performing Hindley-Milner type inference~\cite{pierce} during
+%% sampling, and assume that each task is annotated with the type of the
+%% program that will solve it.
+%% Appendix~\ref{enumerationAppendix}
+%% explains how we enumerate,
+%% rather than sample,
+%% programs generated by Algorithm~\ref{sampleProgram}.
 
@@
 
-  However, a drawback of   enumerative search  is that we have no
-efficient means of solving for arbitrary constants that might occur in a
-program. In Sec.~\ref{regressionSection},
-we will show how to find programs with real-valued constants
-by automatically differentiating through the program and setting the constants using gradient descent.
+%%   However, a drawback of   enumerative search  is that we have no
+%% efficient means of solving for arbitrary constants that might occur in a
+%% program. In Sec.~\ref{regressionSection},
+%% we will show how to find programs with real-valued constants
+%% by automatically differentiating through the program and setting the constants using gradient descent.
 %% In Sec.~\ref{textSection}
@@
 
-\subsection{Sleep-R: Training a Neural Recognition Model}\label{recognitionSection}
+\subsection{Consolidation-Sleep: Growing a Domain Specific Language}
+
+The DSL offers a set of abstractions that allow an agent to concisely
+express solutions to the tasks at hand. We automatically discover
+these new abstractions by combining two ideas. First, we build on
+techniques from the programming languages community to develop a new
+algorithm for automatically refractoring programs, where this
+refactoring exposes common reused subexpressions found across the
+programs found during waking.  Second, we use this automatic refactoring
+process to search for DSLs that maximally compress these programs by
+incorporating reused subexpressions into the DSL.
+
+Mathematically this compression takes the form of
+finding the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
+We replace this marginal with an AIC approximation
+and minimize the following expression,
+which can be interpreted as a kind of compression:
+\begin{equation}
+\underbrace{-\log \probability[\mathcal{D}] + \min_{\theta}\Bigg(-\log
+\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
+  X}\underbrace{-\log \sum_{p\in
+  \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
+\label{AIC}
+  \end{equation}
+At a high level, our approach is to search locally through the space 
+of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
+fails to decrease.  These small changes consist of introducing new
+candidate $\lambda$-expressions into the DSL.
+
+However, there is a snag with this simple approach:
+whenever we add a new expression $e$ to the DSL,
+the programs found during waking ($\left\{\mathcal{F}_x \right\}$)
+are not written in terms of $e$.
+Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
+after having found the programs
+\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
+As human programmers,
+we can look at these pieces of code and recognize that,
+if we define a new procedure called \code{double},
+defined as \code{($\lambda$ (x) (+ x x))},
+then we can rewrite the original programs as
+\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
+This process is a kind of refactoring
+where a new subroutine is defined (\code{double})
+and the old programs rewritten
+in terms of the new subroutine.
+Figure~\ref{mapFactor}A
+diagrams this refactoring
+process for a more complicated
+setting,
+where the agent must rediscover the higher-order function \code{map}
+starting from the basics of Lisp and the Y-combinator.
+
+We refine our objective to compress
+\emph{refactorings} of the programs found during waking, minimizing
+\begin{equation}
+        -\log \probability[\mathcal{D}] + 
+\min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{F}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
+\label{factorObjective}
+\end{equation}
+where $p\manyReduce p'$ is the standard notation for ``expression $p$
+evaluates to $p'$ by the rules of $\lambda$-calculus''~\cite{pierce}.
+Equation~\ref{factorObjective} captures the idea that we want to add
+new components to the DSL while jointly refactoring our old programs
+in terms of these new components.  But this joint optimization is
+intractable, because there are infinitely many ways of refactoring a
+program.  To make refactoring tractable we first limit the degree to
+which a piece of code can be refactored: rather than consider every
+refactoring, we bound the number of $\lambda$-calculus evaluation
+steps separating a refactoring from its original program.  Now the
+number of refactoring this is finite but astronomically large: for the
+example in Figure~\ref{mapFactor} there are approximately $10^{14}$
+possible refactorings -- a quantity that grows exponentially both as a
+function of program size and a function of the bound on evaluation
+steps. How can we tame this combinatorial explosion?
+
+A structure called a \textbf{version space
+  algebra}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
+neatly resolves this exponential growth.  A version space is a
+tree-shaped data structure that compactly represents a large set of
+programs and supports efficient set operations like union,
+intersection, and membership checking.  In
+Appendix~\ref{appendixVersion}, we give a dynamic program that takes
+as input a program $p$ and then outputs a version space containing the
+refactorings of $p$.  Our technique is astronomically more efficient
+than explicitly representing the space of possible refactorings: for
+the example in Figure~\ref{mapFactor}A, we represent the space of
+refactorings using a version space with $10^6$ nodes, which encodes
+$10^{14}$ refactorings.
+
+%% \textbf{Idea 1:} Limit the degree to which
+%% a piece of code can be refactored.
+%% Instead of considering every refactoring,
+%% bound the number of $\lambda$-calculus evaluation steps
+%% separating a refactoring from its original program.
+%% Formally,
+%% we define the set of $n$-step refactorings as:
+%% \begin{equation}
+%%   R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
+%% \end{equation}
+%% where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
+%% in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
+%% For example,
+%% \begin{align*}
+%%   \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
+%%   \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
+%%   \code{(lambda (y) y)}&
+%% \end{align*}
+%% Returning to Equation~\ref{factorObjective},
+%% this approximation gives the following objective:
+%% \begin{equation}
+%%         \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
+%% \label{limitedObjective}
+%%   \end{equation}
+%% In practice, setting the number of refactoring steps $n$ to 3 suffices
+%% to give a competent DSL learning algorithm.  Although the number of
+%% refactorings is now finite, it is still prohibitively large, and in
+%% fact grows exponentially quickly both as a function of $n$ and as a
+%% function of the size of the program being refactored.  For example,
+%% for the programs in Figure~\ref{mapFactor}, there are approximately
+%% $10^{14}$ possible refactorings.  Next, we show how to tame this
+%% exponential explosion.
+
+%% In Appendix~\ref{appendixVersion}, we describe a dynamic program for
+%% efficiently constructing a version space containing every $n$-step
+%% refactoring.
+%% This dynamic program, which we call $I\beta_n(p)$,
+%% satisfies $\denotation{I\beta_n(p)} =  R_n(p)$.
+%% In other words,
+%% this dynamic program
+%% builds a data structure
+%% that represents the entire set of
+%% refactorings -- but without having to explicitly enumerate all of the refactorings.
+
+With this machinery in hand, we now have all the pieces needed to
+learn a DSL (Algorithm~\ref{grammarInductionAlgorithm}).
+The functions $I\beta(\cdot )$ and \textsc{refactor} construct
+a version space from a program and extract the shortest program from a version space, respectively (Algorithm~\ref{grammarInductionAlgorithm}, lines 5-6, 14; see Appendix~\ref{appendixVersion}).
+To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+  and $\lambda$ controls how strongly we regularize the size of the DSL.
+  We place a symmetric Dirichlet prior over the weight vector $\theta$.
+
+To appropriately score each proposed $\mathcal{D}$ we must reestimate
+ the weight vector $\theta$ (Algorithm~\ref{grammarInductionAlgorithm}, line 7).
+Although this  may seem 
+very similar to estimating the parameters of a probabilistic context free grammar,
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+our DSLs are context-sensitive due to the presence of variables
+in the programs and also due to the polymorphic typing system.
+Appendix~\ref{mapAppendix} derives a tractable MAP estimator for $\theta$.
+
+\begin{algorithm}%[H]
+  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
+  \caption{DSL Induction Algorithm}
+  \label{grammarInductionAlgorithm}
+  \begin{algorithmic}[1]
+    \State {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
+    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
+    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
+    \While{true}         
+    \State $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
+    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
+    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
+    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
+    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
+    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
+    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
+    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
+    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
+    \State $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
+    \EndWhile
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{figure*}
+  \centering\begin{tikzpicture}[every node/.style={inner sep=1,outer sep=0,rounded corners,thick}]
+  \node(p1)[draw,rounded corners,thick] at (-1,0) {
+    \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (+ (car l) (car l))}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+  };
+  \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-2.5cm]p1.south) {
+    \begin{tabular}{l}
+      \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
+      \phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
+      \phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+      \phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
+      \texttt{ ($\lambda$ (z) (+ z z)))}
+    \end{tabular}
+  };
+
+  \node(p2)[draw] at ([xshift=4.5cm]p1.east) {
+    \begin{tabular}{l}
+      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
+      \texttt{ (cons (- (car l) 1)}\\
+      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
+    \end{tabular}
+
+  };
+  \node(r2)[draw] at ([yshift=-2.5cm]p2.south) {
+    \begin{tabular}{l}
+      \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
+      \phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
+      \phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
+      \phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
+      \texttt{ ($\lambda$ (z) (- z 1)))}
+    \end{tabular}
+
+  };
+
+  \draw [->] (p1.south)  --(r1.north) node[fill=white,midway] {refactor};
+  \draw [->] (p2.south)  --(r2.north) node[fill=white,midway] {refactor};
+
+  \node[draw](m) at (3,-6.5) {
+    \begin{tabular}{l}
+      \fbox{\textsc{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
+      \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
+      \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}\\
+      \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z z)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
+      \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (- z 1)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
+    \end{tabular}      
+  };
+  \draw [->](r1.south)--(m.north);
+  \draw [->](r2.south)--(m.north);
+  \node[fill=white] at ([yshift=0.5cm]m.north) {\textbf{Compress (MDL/Bayes objective)}};
+
+  \node(t1)[draw] at ([yshift=1.5cm]p1.north) {\begin{tabular}{ll}
+      \textbf{Task}:&\texttt{(1 2 3)$\to$(2 4 6)}\\
+      &\texttt{(4 3 4)$\to$(8 6 8)}
+  \end{tabular}};
+  \draw [->] (t1.south)  --(p1.north) node[fill=white,midway] {Wake: program search};
+  \node(t2)[draw] at ([yshift=1.5cm]p2.north) {\begin{tabular}{ll}
+      \textbf{Task}:&\texttt{(1 2 3)$\to$(0 1 2)}\\
+      &\texttt{(4 3 4)$\to$(3 2 3)}
+  \end{tabular}};
+  \draw [->] (t2.south)  --(p2.north) node[fill=white,midway] {Wake: program search};
+
+  \node(panelA)[ultra thick, rounded corners=0, inner sep=10,outer sep=10, draw, fit=(t1) (t2) (m) (r1) (r2)] {}; \node at ($(0.5,0) + (panelA.west |- panelA.north)$) {\textbf{A}};
+
+  \footnotesize
+  \begin{scope}[shift={(2,-11)}]  
+    \node[draw, rounded corners](u1) at (0,0) {union};
+    \node[draw, rounded corners](u2) at ($(0,-1) + (u1.south)$) {\code{(}union\code{ 1)}}; \draw (u1.south) -- (u2.north);
+    \node[draw, rounded corners](u21) at ($(-1.3,-1) + (u2.south)$) {\code{(($\lambda$ (x) (x 1)) +)}}; \draw ([xshift=-0.2cm]u2.south) -- (u21.north);
+    \node[draw, rounded corners](u22) at ($(2.3,-1) + (u2.south)$) {\code{(($\lambda$ (x) (+ x)) 1)}};  \draw ([xshift=-0.2cm]u2.south) -- (u22.north);
+
+    \node[draw, rounded corners](u11) at ($(-1.75,-1) + (u1.south)$) {\code{(+ 1 1)}}; \draw (u1.south) -- (u11.north);
+    \node[draw, rounded corners](u12) at ($(2.75,-1) + (u1.south)$) {\code{(($\lambda$ (x) (x 1 1)) +)}}; \draw (u1.south) -- (u12.north);
+
+    \node(vs) at ($(u21.west)!0.5!(u12.east) + (0,-1.25)$) {$\underbrace{\hspace{8cm}}_{\text{\normalsize Subset of version space}}$};
+
+    \node[anchor=north](p) at ($(-6,0) + (u2)$) {\normalsize $\underbrace{\texttt{(+ 1 1)}}_{\text{\normalsize  Program}}$};
+    \draw[ultra thick,->] ($(0.5,0) + (p.east)$) -- ($(1.5,0) + (p.east)$);
+    \node at ($(1,0.25) + (p.east)$) {Refactors};
+    \node(g)[anchor=left] at ($(7,0) + (u12.east)$) {\includegraphics[width = 6cm]{figures/vs.eps}};
+    \node(panelB)[ultra thick, rounded corners=0, inner sep=10,outer sep=10, draw, fit=(p) (u1) (vs) (g)] {}; \node at ($(0.5,0) + (panelB.west |- panelB.north)$) {\normalsize\textbf{B}};
+
+    %% \node(panelC)[ultra thick, rounded corners=0, inner sep=8,outer sep=8, draw, fit=(g)] {}; \node at ($(0.5,0) + (panelC.west |- panelC.north)$) {\normalsize\textbf{C}}; 
+  \end{scope}
+
+  \end{tikzpicture}
+  \caption{DSL learning as code refactoring.
+\textbf{Panel A:} For each task we discover programs during waking, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}).
+Common subprograms are incorporated into the DSL when they increase a Bayesian objective.
+Intuitively, these new DSL components best compress the programs found during waking.
+\textbf{Panel B:} \# of possible refactorings grows exponentially with program size.
+We represent refactorings using version spaces, which augment syntax trees with a \emph{union} operator whose children are themselves version spaces.
+Right graph: version spaces are exponentially more efficient than explicitly constructing set of refactorings.}\label{mapFactor}
+\end{figure*}
+
+\subsection{Dream Sleep: Training a Neural Recognition Model}\label{recognitionSection}
+
+During ``dreaming'' the system learns a recognition model that guides
+program search.  It learns from (program, task) pairs drawn from two
+sources of self-supervised data: \emph{replays} of programs discovered
+during waking, and \emph{fantasies}, or programs drawn from the DSL.
+Replays ensure that the recognition model is trained on the tasks
+solved so far, and does not forget how to solve them.  Fantasies
+ensure that the recognition model has a large and highly varied corpus
+of (program, task) pairs to learn from.
+
+Formally, $Q(p|x)$ should approximate the posterior
+$\probability[p|\mathcal{D},\theta,x]$
+
+training a neural recognition model, $Q$, to regress from
+a task to a distribution over programs
+that are likely to both
+solve the task and
+be likely under the prior.
+To train $Q$ we need (program, task)
+pairs.
+
+To get the data to train $Q$,
+we take inspiration from
+the
 
@@
 
-\subsubsection{Training $Q$}\label{recognitionTraining}
-How should we get the data to train $Q$?
-This is non-obvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from task/program pairs).
-One approach is to draw programs from the DSL,
-run them to get their input/outputs,
-and then train $Q$ to predict the program from the input/outputs.
-This is like how the wake-sleep algorithm for the Helmholtz machine
-trains its recognition model during its sleep phase~\cite{dayan1995helmholtz}.
-The advantage of training on samples, or ``dreams,'' is that
-we can draw unlimited samples from the DSL,
-training on a large amount of data.
-Another approach is
-to train $Q$ on the (program, task)
-pairs discovered by the Wake phase.
-The advantage here
-is that the training data is much higher quality,
-because we are training on real tasks.
-Due to these complementary advantages,
-we train on both these sources of data.
-
-Formally, $Q$ should approximate the true posteriors over programs.
 We can either train $Q$ to perform full posterior inference by minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|Q(p|x) \right) \right]$,
@@
 
-\subsection{Sleep-G: Learning a Generative Model (a DSL)}\label{grammarInductionSection}
-
-The purpose of the DSL is to
-offer a set of abstractions
-that allow an agent to easily express solutions to the tasks at hand.
-Intuitively, we want the algorithm to look at the programs found
-during waking and generalize beyond them, both so the DSL can better
-express the current solutions, and also so that the DSL might expose
-new abstractions which will later be used to discover more programs.
-Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
-We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
-\begin{equation}
-        \log \probability[\mathcal{D}] + 
-\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
-\label{AIC}
-  \end{equation}
-At a high level, our approach is to search locally through the space 
-of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
-fails to increase.  These small changes consist of introducing new
-candidate $\lambda$-expressions into the DSL.
-
-However, there is a snag with this simple approach:
-whenever we add a new expression $e$ to the DSL,
-the programs found during waking ($\left\{\mathcal{F}_x \right\}$)
-are not written in terms of $e$.
-Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
-after having found the programs
-\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
-As human programmers,
-we can look at these pieces of code and recognize that,
-if we define a new procedure called \code{double},
-defined as \code{($\lambda$ (x) (+ x x))},
-then we can rewrite the original programs as
-\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
-This process is a kind of refactoring
-where a new subroutine is defined (\code{double})
-and the old programs rewritten
-in terms of the new subroutine.
-Figure~\ref{mapFactor}
-diagrams this refactoring
-process for a more complicated
-setting,
-where the agent must rediscover the higher-order function \code{map}
-starting from the basics of Lisp and the Y-combinator.
-
-Formally, we want a DSL
-which puts high probability on
-any refactoring of the programs found during waking, maximizing:
-\begin{equation}
-        \log \probability[\mathcal{D}] + 
-\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
-\label{factorObjective}
-  \end{equation}
-where $p'\manyReduce p$
-is the standard notation for ``expression $p'$ evaluates to $p$ by the rules of $\lambda$-calculus''~\cite{pierce}.
-
-Equation~\ref{factorObjective} captures the idea that we want
-to add new components to the DSL while jointly refactoring our old
-programs in terms of these new components.  But
-this joint inference problem is intractable, because there are
-infinitely many ways of refactoring a program.
-We combine two ideas to make this refactoring process tractable:
-
-\textbf{Idea 1:} Limit the degree to which
-a piece of code can be refactored.
-Instead of considering every refactoring,
-bound the number of $\lambda$-calculus evaluation steps
-separating a refactoring from its original program.
-Formally,
-we define the set of $n$-step refactorings as:
-\begin{equation}
-  R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
-\end{equation}
-where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
-in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
-For example,
-\begin{align*}
-  \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
-  \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
-  \code{(lambda (y) y)}&
-\end{align*}
-Returning to Equation~\ref{factorObjective},
-this approximation gives the following objective:
-\begin{equation}
-        \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
-\label{limitedObjective}
-  \end{equation}
-In practice, setting the number of refactoring steps $n$ to 3 suffices
-to give a competent DSL learning algorithm.  Although the number of
-refactorings is now finite, it is still prohibitively large, and in
-fact grows exponentially quickly both as a function of $n$ and as a
-function of the size of the program being refactored.  For example,
-for the programs in Figure~\ref{mapFactor}, there are approximately
-$10^{14}$ possible refactorings.  Next, we show how to tame this
-exponential explosion.
-
-\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
-we developed a technique for compactly representing and efficiently
-calculating the set of all refactorings.  The key idea is to represent
-the set of refactorings using a \emph{version
-  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
-A version space is a tree-shaped data structure
-that compactly represents a large set of programs.
-We first define version spaces over $\lambda$-calculus expressions,
-and then give a dynamic program for constructing a version space
-that represents the set of possible refactorings (i.e., $R_n(p)$).
-This technique is astronomically more efficient than
-explicitly representing
-the space of possible refactorings:
-for the example in Figure~\ref{mapFactor},
-we represent
-the space of refactorings using a version space with
-$10^6$ nodes,
-which encodes $10^{14}$ refactorings.
-Formally, a version space is either:
-\begin{itemize}
-\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus.
-When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to.
-For example, $\lambda x.\lambda y.
-(x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices.
-See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
-\item An abstraction: written $\lambda v$, where $v$ is a version space
-\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
-\item A  union: $\uplus V$, where $V$ is a set of version spaces
-\item The empty set, $\varnothing$
-\item The set of all $\lambda$-calculus expressions, $\Lambda$
-\end{itemize}
-The purpose of a version space to compactly represent a set of programs.
-We refer to this set as the \textbf{extension} of the version space:
-\begin{definition}
-  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
-  and is defined recursively as:
-  \begin{align*}
-    \denotation{\$i}& = \left\{\$i \right\}&
-    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}&
-    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
-    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}&
-    \denotation{\varnothing}& = \varnothing&
-    \denotation{\Lambda}& = \Lambda
-    \end{align*}
-\end{definition}
-Version spaces also support efficient membership checking, which we
-write as  $e\in \denotation{v}$.
-Important for our purposes, it is also efficient to refactor the
-members of a version space's extension in terms of a new DSL.
-We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
-\begin{align*}
-  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
-    e\text{, if $e\in \mathcal{D}$ and $e\in \denotation{v}$. Exploits the fact that $\denotation{e}\in v$ can be efficiently computed.}\\
-    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
-  \end{cases}
-\end{align*}\vspace{-\baselineskip}%
-\begin{align*}
-  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber
-  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
-  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})&%\\\nonumber
-  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
-  \end{align*}
-where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
-is the size of the syntax tree of $e$,
-when members of $\mathcal{D}$ are counted as having size 1.
-Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.
-
-In Appendix~\ref{appendixVersion}, we describe a dynamic program for
-efficiently constructing a version space containing every $n$-step
-refactoring.
-This dynamic program, which we call $I\beta_n(p)$,
-satisfies $\denotation{I\beta_n(p)} =  R_n(p)$.
-In other words,
-this dynamic program
-builds a data structure
-that represents the entire set of
-refactorings -- but without having to explicitly enumerate all of the refactorings.
-
-With this machinery in hand, we now have all the pieces needed to
-learn a DSL (Algorithm~\ref{grammarInductionAlgorithm}).
-To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-  and $\lambda$ controls how strongly we regularize the size of the DSL.
-  We place a symmetric Dirichlet prior over the weight vector $\theta$.
-
-To appropriately score each proposed $\mathcal{D}$ we must reestimate
- the weight vector $\theta$ (Algorithm~\ref{grammarInductionAlgorithm}, line 7).
-Although this  may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-our DSLs are context-sensitive due to the presence of variables
-in the programs and also due to the polymorphic typing system.
-Appendix~\ref{mapAppendix} derives a tractable MAP estimator for $\theta$.
-
-\begin{algorithm}%[H]
-  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
-  \caption{DSL Induction Algorithm}
-  \label{grammarInductionAlgorithm}
-  \begin{algorithmic}[1]
-    \State {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-    \While{true}         
-    \State $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
-    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
-    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
-    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
-    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
-    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
-    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
-    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
-    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
-    \State $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
-    \EndWhile
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{figure*}
-  \centering\begin{tikzpicture}[every node/.style={inner sep=1,outer sep=0,rounded corners,thick}]
-    \node(p1)[draw,rounded corners,thick] at (-1,0) {
-      \begin{tabular}{l}
-      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
-      \texttt{ (cons (+ (car l) (car l))}\\
-      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
-    \end{tabular}
-    };
-    \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-2.5cm]p1.south) {
-      \begin{tabular}{l}
-  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
-\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
-\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
-\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
-\texttt{ ($\lambda$ (z) (+ z z)))}
-      \end{tabular}
-    };
-
-        \node(p2)[draw] at ([xshift=4.5cm]p1.east) {
-    \begin{tabular}{l}
-      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
-      \texttt{ (cons (- (car l) 1)}\\
-      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
-    \end{tabular}
-
-    };
-    \node(r2)[draw] at ([yshift=-2.5cm]p2.south) {
-\begin{tabular}{l}
-  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
-\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
-\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
-\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
-\texttt{ ($\lambda$ (z) (- z 1)))}
-      \end{tabular}
-
-    };
-
-    \draw [->] (p1.south)  --(r1.north) node[fill=white,midway] {refactor};
-    \draw [->] (p2.south)  --(r2.north) node[fill=white,midway] {refactor};
-
-    \node[draw](m) at (3,-6.5) {
-\begin{tabular}{l}
-          \fbox{\textsc{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
-  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
-  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}\\
-  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z z)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
-  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (- z 1)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
-      \end{tabular}      
-    };
-    \draw [->](r1.south)--(m.north);
-    \draw [->](r2.south)--(m.north);
-    \node[fill=white] at ([yshift=0.5cm]m.north) {\textbf{Compress (MDL/Bayes objective)}};
-    %% \visible<2-3>{
-    %%   \node[fill=orange!50!white,align=center] at (r1) {\Huge $10^{14}$\\\Huge  refactorings };
-    %%   \node[fill=orange!50!white,align=center] at (r2) {\Huge $10^{13}$\\\Huge  refactorings };
-    %% }
-    %% \visible<3>{
-    %%   \node[fill=orange!50!white,align=center] at ([yshift=7cm]m.north) {\Huge version space: set of programs\\
-    %%   \huge Lau 2003; Gulwani 2012};
-    %%   }
-    %% \visible<4>{
-    %%   \node[fill=orange!50!white,align=center] at (r1) {\Huge $\leq 10^6$\\\Huge version spaces };
-    %%   \node[fill=orange!50!white,align=center] at (r2) {\Huge $\leq 10^6$\\\Huge version spaces };
-    %% }
-
-       \node(t1)[draw] at ([yshift=1.5cm]p1.north) {\begin{tabular}{ll}
-           \textbf{Task}:&\texttt{(1 2 3)$\to$(2 4 6)}\\
-           &\texttt{(4 3 4)$\to$(8 6 8)}
-       \end{tabular}};
-       \draw [->] (t1.south)  --(p1.north) node[fill=white,midway] {program search};
-       \node(t2)[draw] at ([yshift=1.5cm]p2.north) {\begin{tabular}{ll}
-           \textbf{Task}:&\texttt{(1 2 3)$\to$(0 1 2)}\\
-           &\texttt{(4 3 4)$\to$(3 2 3)}
-       \end{tabular}};
-       \draw [->] (t2.south)  --(p2.north) node[fill=white,midway] {program search};
-
-  \end{tikzpicture}
-  \caption{DSL learning as code refactoring.
-For each task we discover programs during waking, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}).
-Common subprograms are incorporated into the DSL when they increase a Bayesian objective.
-Intuitively, these new DSL components best compress the programs found during waking.}\label{mapFactor}
-  \end{figure*}
-
-\begin{comment}
-
-We propose these new expressions
-by extracting fragments of programs already in the frontiers
-(Tbl.~\ref{fragmentExample}).  An important point here is that we are
-\emph{not} simply adding subexpressions of programs to $\mathcal{D}$,
-as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}
-and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
-extracting fragments that unify with programs in the frontiers.  This
-idea of storing and reusing fragments of expressions comes from
-Fragment Grammars~\cite{tim} and Tree-Substitution
-Grammars~\cite{cohn2010inducing}, and is closely related to the idea
-of antiunification~\cite{henderson2013cumulative}.
-
-%\vfill\null
-
-%\columnbreak
-
-To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-  and $\lambda$ controls how strongly we regularize the size of the DSL.
-  We place a symmetric Dirichlet prior over the weight vector $\theta$,
-  defining 
-$  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
- $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
-%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.
-
-To appropriately score each proposed $\mathcal{D}$ we must reestimate
- the weight vector $\theta$.
-Although this  may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-our DSLs are context-sensitive due to the presence of variables
-in the programs and also due to the polymorphic typing system.
-In the Supplement we derive a tractable MAP estimator for $\theta$.
-%% Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
-%% recognition model training, and DSL induction.
-
-\begin{figure*}[!htb]\tabcolsep=2pt
-\centering  \begin{minipage}[c]{0.2\textwidth}
-  \begin{tikzpicture}[scale=0.7]
-    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
-    %%     \footnotesize Program:\\
-    %%     \code{($\lambda$ (x) (+ (- x) 1))}
-    %% \end{tabular}};
-    %\node at (,0.5) {\code{cons}};
-    \node [rotate=90] at (-2.3,-0.5) {\small program};
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-%    \draw  (l1.south) -- (-0.5,0.5);
-
-  %% \node(c) at (-0.5,-1.5) {\code{-}};
-  %% \node(z) at (0.5,-1.5) {\code{x}};
-
-  %% \draw (0,-1) -- (c.north);
-  %% \draw (0,-1) -- (z.north);
-
-  \begin{scope}[shift={(0,-2.5)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-  \end{scope}
-
-\begin{scope}[shift={(0,-5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  %\node(x1) at (0,-1) {};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.55,-0.4) -- (a.north);
-  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
-
-  \end{scope}
-
-\end{tikzpicture}
-  \end{minipage}
-\hspace{0.1cm}\begin{tabular}{ll}
-    \toprule
-    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
-    \begin{tabular}{l}
-      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
-      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
-    \end{tabular}&
-    \begin{tabular}{l}
-      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
-      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
-    \end{tabular}\\\midrule
-    \begin{tabular}{l}
-      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
-      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
-      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
-      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
-      \end{tabular}&
-    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
-      \end{tabular}
-\\\bottomrule\\
-\end{tabular}
-\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
-highlighted in {\orange{orange}},
-  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
-  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
-\end{figure*}
-
-  \end{comment}
-\begin{comment}
-  \begin{wrapfigure}{R}{0.6\textwidth}
-  \begin{minipage}{0.6\textwidth}    
-    \begin{algorithm}[H]
-      \caption{The \system Algorithm}
-      \label{mainAlgorithm}
-      \begin{algorithmic}
-        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
-        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
-        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
-        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
-        \FOR{$i=1$ {\bfseries to} $I$}
-        %     \FOR{$x:\tau\in X$}
-        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
-        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
-        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
-        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
-        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
-        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}
-
-        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
-        %% \STATE Define $Q_x(z) \propto \begin{cases}
-        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
-        %%   0&x\not \in \mathcal{F}_x
-        %% \end{cases}
-        \ENDFOR
-        \STATE \textbf{return} $\mathcal{D},\theta,q$
-      \end{algorithmic}
-    \end{algorithm}
-  \end{minipage}
-\end{wrapfigure}
-  \end{comment}
-
 \section{Experiments}
@@
 \subsection{Programs that manipulate sequences}\label{sequences}
-We apply \system to list processing %% (Section~\ref{listSection})
-and text editing, %% (Section~\ref{textSection})
+We first apply \system to two classic benchmark domains: list
+processing and text editing. We find that we can start with generic
+Lisp-style primitives and recover a domain-specific vocabulary for
+solving each of these classes of problems.
+
 using a GRU~\cite{cho2014learning} for
@@
 \textbf{List Processing:} Synthesizing programs that manipulate data structures is a widely studied
-problem in the programming languages community~\cite{feser2015synthesizing}.
-%with applications to computer aided programming~\cite{solar2008program}.
+problem in the programming languages community~\cite{feser2015synthesizing},
+and progress here can translate to better computer-aided programming~\cite{solar2008program}.
 We consider this problem within the context of learning functions that
@@
 \begin{figure}
-  \includegraphics[width = \textwidth]{figures/logo90.png}
+  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
+  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
   \caption{LOGO graphics tasks. The agent must learn to write programs that draw these 90 pictures.}\label{everyLogo}
@@
 allowing the agent to solve 90\% of the tasks.
-\begin{figure}
-  \includegraphics[width = \textwidth]{figures/every_tower.png}
-  \caption{}\label{towerTasks}
-\end{figure}
 \begin{figure}
@@
 
+\begin{figure}
+  \includegraphics[width = 5cm]{figures/logoLearningCurve.png}
+  \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}
+  \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}
+  \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}
+  \end{figure}
+
 \begin{figure*}\centering
@@
 \section{Appendix}
+
+\subsection{Generative model over programs}\label{generativeAppendix}
+Algorithm~\ref{sampleProgram} gives a stochastic procedure for drawing samples from $\probability[\cdot |\mathcal{D},\theta]$.
+It takes as input the desired type of the unknown program,
+and performs type inference during
+sampling to ensure that
+the program has the desired type.
+It also maintains a \emph{environment}
+mapping variables to types,
+which ensures
+that lexical scoping rules are obeyed.
+\begin{algorithm}
+  \caption{Generative model over programs}
+  \label{sampleProgram}
+  \begin{algorithmic}[1]
+    \State \textbf{function} sample$(\mathcal{D}, \theta, \tau)$:
+    \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, type $\tau$
+    \State \textbf{Output:} a program whose type unifies with $\tau$
+    \State  \textbf{return }sample$'(\mathcal{D}, \theta, \varnothing, \tau)$
+
+    \Statex
+    \State \textbf{function} sample$'(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
+    \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$\Comment{Environment $\mathcal{E}$ starts out as $\varnothing$}
+    \State \textbf{Output:} a program whose type unifies with $\tau$
+    \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
+    \State var $\gets$ an unused variable name
+    \State body $\sim$ sample$'(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$\Comment{Recursively sample function body}
+    \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
+    %\Endif
+    \Else\Comment{Build an application to give something w/ type $\tau$}
+    \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+    $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
+    \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
+    \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
+    \theta_e&\text{ if }e\in \mathcal{D}\\
+    \theta_{var}/|\text{variables}|&\text{ if }e\in \mathcal{E}
+    \end{cases}$
+    \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
+    \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
+    %   \State unify$(\tau,\beta)$
+    \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
+    \State $a_k\sim\text{sample}'(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
+    \EndFor
+    \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
+    \EndIf
+    \Statex
+    \Statex\textbf{where:}
+    \State yield$(\tau) = \begin{cases}
+      \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+      \tau   &\text{ otherwise.}
+    \end{cases}$ \Comment{Final return type of $\tau$}
+    \State  args$(\tau) = \begin{cases}
+      [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+      []   &\text{ otherwise.}
+    \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
+  \end{algorithmic}
+\end{algorithm}
 
@@
  \subsection{Refactoring code with version spaces}\label{appendixVersion}
+ Formally, a version space is either:
+\begin{itemize}
+\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus.
+When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to.
+For example, $\lambda x.\lambda y.
+(x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices.
+See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
+\item An abstraction: written $\lambda v$, where $v$ is a version space
+\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
+\item A  union: $\uplus V$, where $V$ is a set of version spaces
+\item The empty set, $\varnothing$
+\item The set of all $\lambda$-calculus expressions, $\Lambda$
+\end{itemize}
+The purpose of a version space to compactly represent a set of programs.
+We refer to this set as the \textbf{extension} of the version space:
+\begin{definition}
+  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
+  and is defined recursively as:
+  \begin{align*}
+    \denotation{\$i}& = \left\{\$i \right\}&
+    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}&
+    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
+    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}&
+    \denotation{\varnothing}& = \varnothing&
+    \denotation{\Lambda}& = \Lambda
+    \end{align*}
+\end{definition}
+Version spaces also support efficient membership checking, which we
+write as  $e\in \denotation{v}$.
+Important for our purposes, it is also efficient to refactor the
+members of a version space's extension in terms of a new DSL.
+We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
+\begin{align*}
+  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
+    e\text{, if $e\in \mathcal{D}$ and $e\in \denotation{v}$. Exploits the fact that $\denotation{e}\in v$ can be efficiently computed.}\\
+    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
+  \end{cases}
+\end{align*}\vspace{-\baselineskip}%
+\begin{align*}
+  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber
+  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
+  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})&%\\\nonumber
+  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
+  \end{align*}
+where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
+is the size of the syntax tree of $e$,
+when members of $\mathcal{D}$ are counted as having size 1.
+Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.
+
 Recall that our goal is to define an operator over version spaces, $I\beta_n$,
 which calculates the set of $n$-step refactorings.
-We define this operator recursively in terms of another operator, $I\beta'$, which performs a single step of refactoring:
+We define this operator in terms of another operator, $I\beta'$, which performs a single step of refactoring:
 \begin{align*}
@@
 $I\beta'$, and $S_k$ once per each version space.
-Refactoring is similarly done more quickly with dynamic programming (see Equation~\ref{FACTORING} for the recursive definition of \textsc{refactor}).
-
-We now formally proved that $I\beta$ exhaustively enumerates the space of possible refactorings.
+%Refactoring is similarly done more quickly with dynamic programming (see Equation~\ref{FACTORING} for the recursive definition of \textsc{refactor}).
+
+We now formally prove that $I\beta$ exhaustively enumerates the space of possible refactorings.
 Our approach is to first prove that $S_k$ exhaustively enumerates
@@
 
+\subsection{Full set of LOGO  tasks}
+
 \begin{figure}
-  \begin{tikzpicture}
-    \node at (0,0) (d){DSL};
-    \node at ([yshift = -2cm]d) (t){$\text{Task}$};
-
-    \node at ([xshift = 2cm]t) (nn){
-      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
-        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
-        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
-        \node[rectangle] at (1,1) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (I-\name) at (0,-\y) {};
-        \foreach \name / \y in {1,...,3}
-            \node[neuron] (H-\name) at (1,-\y-0.5) {};
-        \foreach \name / \y in {1,...,4}
-            \node[neuron] (O-\name) at (2,-\y) {};
-        \foreach \source in {1,...,4}
-            \foreach \dest in {1,...,3}
-                \draw [-latex] (I-\source) -- (H-\dest);
-        \foreach \source in {1,...,3}
-            \foreach \dest in {1,...,4}
-                \draw [-latex] (H-\source) -- (O-\dest);
-      \end{tikzpicture}
-    };
-    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
-    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
-
-    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
-    \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
-
-    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
-    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
-
-    \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
-    \draw [->  ] (s.east) -- (f.west);
-
-    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
-    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
-
-    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
-
-  \end{tikzpicture}
-
-  \vspace{1cm}
-
-  \begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
-
-    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
-    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
-    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
-
-    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
-    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
-    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
-
-    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
-    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
-    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
-  \end{tikzpicture}
-
-  \begin{tikzpicture}
-    \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
-    \node at ([xshift = 3cm]d.east) (p2){program};
-    \node at ([yshift = 1.5cm]p2) (p1){program};
-    \node at ([yshift = -1.5cm]p2) (p3){program};
-
-    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
-    \draw[squiggle,-> ] (d.east) -- (p1.west);
-    \draw[squiggle,-> ] (d.east) -- (p3.west);
-
-    \node at ([xshift = 2cm]p1.east) (t1){task};
-    \node at ([xshift = 2cm]p2.east) (t2){task};
-    \node at ([xshift = 2cm]p3.east) (t3){task};
-    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
-    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
-    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
-
-    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
-  \end{tikzpicture}
-
-  \vspace{2cm}
-
-  \begin{tikzpicture}
-    \node at (0,0) (f1){Frontier$_1$};
-    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
-    \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
-    \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
-    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
-
-    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
-    \node[anchor=north] at (c.north) {Compression};
-
-    \draw [-> ] (f1.east) -- (c.west|-f1.east);
-    \draw [-> ] (f2.east) -- (c.west|-f2.east);
-    \draw [-> ] (f3.east) -- (c.west|-f3.east);
-
-    \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
-    \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
-    \draw [-> ] (c.east) -- (d.west);
-    \draw [-> ] (c.east) -- (t.west);
-
-    \node at (c.center) {
-\begin{tikzpicture}[scale=0.7]
-    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
-    %%     \footnotesize Program:\\
-    %%     \code{($\lambda$ (x) (+ (- x) 1))}
-    %% \end{tabular}};
-    %\node at (,0.5) {\code{cons}};
-%    \node [rotate=90] at (-2.3,-0.5) {\small program};
-
-          \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  \node(x1) at (0,-1) {\code{1}};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
-
-  \node(t) at (-0.5,0.5) {};
-  \draw (l1.south) -- (t.south);
-  \node(c) at (-1.5,-0.2) {\code{cons}};
-  \draw (t.south) -- (c.north);
-
-%    \draw  (l1.south) -- (-0.5,0.5);
-
-  %% \node(c) at (-0.5,-1.5) {\code{-}};
-  %% \node(z) at (0.5,-1.5) {\code{x}};
+  \includegraphics[width = \textwidth]{figures/fullLogo.png}
+  \caption{Full set of LOGO graphics tasks that we apply our system to}
+\end{figure}
+
+\subsection{Full set of tower tasks}
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/fullTower.png}
+  \caption{Full set of tower building tasks that we apply our system to}
+\end{figure}
+
+%% \begin{figure}
+%%   \begin{tikzpicture}
+%%     \node at (0,0) (d){DSL};
+%%     \node at ([yshift = -2cm]d) (t){$\text{Task}$};
+
+%%     \node at ([xshift = 2cm]t) (nn){
+%%       \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
+%%         \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
+%%         \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
+%%         \node[rectangle] at (1,1) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (I-\name) at (0,-\y) {};
+%%         \foreach \name / \y in {1,...,3}
+%%             \node[neuron] (H-\name) at (1,-\y-0.5) {};
+%%         \foreach \name / \y in {1,...,4}
+%%             \node[neuron] (O-\name) at (2,-\y) {};
+%%         \foreach \source in {1,...,4}
+%%             \foreach \dest in {1,...,3}
+%%                 \draw [-latex] (I-\source) -- (H-\dest);
+%%         \foreach \source in {1,...,3}
+%%             \foreach \dest in {1,...,4}
+%%                 \draw [-latex] (H-\source) -- (O-\dest);
+%%       \end{tikzpicture}
+%%     };
+%%     \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
+%%     \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);
+
+%%     \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.2cm,yshift = -1cm]) (s){Search};
+%%     \node at ([xshift=-7pt,yshift=5pt]s.north west) {$\mathcal{D}$};
+
+%%     \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
+%%     \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);
+
+%%     \node[align=left] at (7,-1) (f) {Frontier\\{\small (set of programs)}};
+%%     \draw [->  ] (s.east) -- (f.west);
+
+%%     \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
+%%     \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};
+
+%%     \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};
+
+%%   \end{tikzpicture}
+
+%%   \vspace{1cm}
+
+%%   \begin{tikzpicture}
+%%     \node at (0,0) (f1){Frontier$_1$};
+%%     \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+%%     \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};
+
+%%     \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
+%%     \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
+%%     \node at ([xshift = 2cm]f3.east) (p3){program$_3$};
+
+%%     \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
+%%     \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
+%%     \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);
+
+%%     \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
+%%     \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
+%%     \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};
+
+%%     \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
+%%   \end{tikzpicture}
+
+%%   \begin{tikzpicture}
+%%     \node[align=center] at (0,0) (d){DSL\\$(\mathcal{D},\theta)$};
+%%     \node at ([xshift = 3cm]d.east) (p2){program};
+%%     \node at ([yshift = 1.5cm]p2) (p1){program};
+%%     \node at ([yshift = -1.5cm]p2) (p3){program};
+
+%%     \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
+%%     \draw[squiggle,-> ] (d.east) -- (p1.west);
+%%     \draw[squiggle,-> ] (d.east) -- (p3.west);
+
+%%     \node at ([xshift = 2cm]p1.east) (t1){task};
+%%     \node at ([xshift = 2cm]p2.east) (t2){task};
+%%     \node at ([xshift = 2cm]p3.east) (t3){task};
+%%     \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
+%%     \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
+%%     \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);
+
+%%     \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
+%%   \end{tikzpicture}
+
+%%   \vspace{2cm}
+
+%%   \begin{tikzpicture}
+%%     \node at (0,0) (f1){Frontier$_1$};
+%%     \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
+%%     \node at ([yshift = -0.7cm]f2.south) (ff){\textbf{$\vdots$}};
+%%     \node at ([yshift = -1.2cm]f2.south) (ff){\textbf{$\vdots$}};
+%%     \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};
+
+%%     \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
+%%     \node[anchor=north] at (c.north) {Compression};
+
+%%     \draw [-> ] (f1.east) -- (c.west|-f1.east);
+%%     \draw [-> ] (f2.east) -- (c.west|-f2.east);
+%%     \draw [-> ] (f3.east) -- (c.west|-f3.east);
+
+%%     \node[right](d) at ([xshift = 1.2cm,yshift = 0.7cm]c.east) {DSL $\mathcal{D}$};
+%%     \node[right](t) at ([xshift = 1.2cm,yshift = -0.7cm]c.east) {Weights $\theta$};
+%%     \draw [-> ] (c.east) -- (d.west);
+%%     \draw [-> ] (c.east) -- (t.west);
+
+%%     \node at (c.center) {
+%% \begin{tikzpicture}[scale=0.7]
+%%     %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
+%%     %%     \footnotesize Program:\\
+%%     %%     \code{($\lambda$ (x) (+ (- x) 1))}
+%%     %% \end{tabular}};
+%%     %\node at (,0.5) {\code{cons}};
+%% %    \node [rotate=90] at (-2.3,-0.5) {\small program};
+
+%%           \node(l1) at (0,0) {};
+%%   \node[color=pop3](p1) at (-1,-1) {\code{+}};
+%%   \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+%%   \node(x1) at (0,-1) {\code{1}};
+%%   \draw[color=pop3] (l1.south) -- (p1.north);
+%%   \draw[color=pop3] (l1.south) -- (n1.north);
+%%   \draw[color=pop3] (-0.5,-0.45) -- (x1.north);
+
+%%   \node(t) at (-0.5,0.5) {};
+%%   \draw (l1.south) -- (t.south);
+%%   \node(c) at (-1.5,-0.2) {\code{cons}};
+%%   \draw (t.south) -- (c.north);
+
+%% %    \draw  (l1.south) -- (-0.5,0.5);
+
+%%   %% \node(c) at (-0.5,-1.5) {\code{-}};
+%%   %% \node(z) at (0.5,-1.5) {\code{x}};
+
+%%   %% \draw (0,-1) -- (c.north);
+%%   %% \draw (0,-1) -- (z.north);
+
+%%   \begin{scope}[shift={(-1,-2.5)}]
+%%       \node(l1) at (0,0) {};
+%%   \node[color=pop3](p1) at (-1,-1) {\code{+}};
+%%   \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
+%%   %\node(x1) at (0,-1) {};
+%%   \draw[color=pop3] (l1.south) -- (p1.north);
+%%   \draw[color=pop3] (l1.south) -- (n1.north);
+%%   \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
+
+%%   \node(c) at (-0.5,-1.5) {\code{car}};
+%%   \node(z) at (0.5,-1.5) {\code{z}};
 
@@
 
-  \begin{scope}[shift={(-1,-2.5)}]
-      \node(l1) at (0,0) {};
-  \node[color=pop3](p1) at (-1,-1) {\code{+}};
-  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
-  %\node(x1) at (0,-1) {};
-  \draw[color=pop3] (l1.south) -- (p1.north);
-  \draw[color=pop3] (l1.south) -- (n1.north);
-  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);
-
-  \node(c) at (-0.5,-1.5) {\code{car}};
-  \node(z) at (0.5,-1.5) {\code{z}};
-
-  \draw (0,-1) -- (c.north);
-  \draw (0,-1) -- (z.north);
-
-%  \node [rotate=90] at (-2.3,-0.7) {\small program};
-
-  \end{scope}
-
-\begin{scope}[shift={(0,-5)}]
-  \node[pop3](p1) at (-1,-1) {\code{+}};
-  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
-  \node[pop3](a) at (0,-1) {\code{ }};
-  %\node(x1) at (0,-1) {};
-  \draw[pop3] (0,0) -- (p1.north);
-  \draw[pop3] (0,0) -- (n1.north);
-  \draw[pop3] (-0.55,-0.4) -- (a.north);
-%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
-
-  \end{scope}
-
-\end{tikzpicture}
-    };
-
-    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};
-
-    \end{tikzpicture}
-\end{figure}
-
-\begin{table*}%[t!]
-  \makebox[\textwidth][c]{
-    \scriptsize
-  \tabcolsep=4pt
-  \renewcommand\code\texttt
-  \renewcommand\codechar[1]{\texttt{"#1"}}
-  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{cccc}
-    \toprule
-    &{\normalsize Symbolic regression}&{\normalsize Laws of motion}&\\\midrule
-    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
-    \begin{tabular}{cc}
-      \includegraphics[width = 3em]{figures/functions/4.png}&
-      \includegraphics[width = 3em]{figures/functions/146}\\
-      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
-      ~\\
-      \includegraphics[width = 3em]{figures/functions/112.png}&
-        \includegraphics[width = 3em]{figures/functions/92.png}
-      \\
-      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\
-
-    \end{tabular}
-    }
-    &
-    \begin{tabular}{cc}
-      \includegraphics[width = 15em]{figures/massOnSpring.png} &
-            \includegraphics[width = 15em]{figures/planets.png}
-      \\
-      \begin{tabular}{l}
-              \blueCode{$f($o,$\Delta) = $($f_3$ o $\Delta$}\\
-      \hspace{1cm}\blueCode{(- (* $k$ (pos o))}\\
-      \hspace{1cm}\phantom{\code{(- }}\blueCode{(* $-9.8$ $\hat{x}$)))}\\
-      \end{tabular}&
-      \begin{tabular}{l}
-              \blueCode{$f($a,b,$\Delta) = $($f_3$ a $\Delta$}\\
-      \hspace{1cm}\blueCode{(/ (* $G$ (mass a) (mass b))}\\
-      \hspace{1cm}\phantom{\code{(- }}\blueCode{(square (- (pos a) (pos b)))))}\\
-        \end{tabular}
-      \end{tabular}
-    &
-
-    ~\\
-    \midrule
-    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
-      \begin{tabular}{l}
-    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
-    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
-    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
-    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
-    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
-    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
-    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
-    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
-    \hspace{\helpSize}\emph{($f_6$: rational function)}\\
-
-  \end{tabular}
-      &
-      \begin{tabular}{l}
-        \greenCode{$f_0($\code{o},$\Delta)\, = \,$(set-pos o (+ (pos o) (* $\Delta$ (vel o))))}\\
-        \hspace{\helpSize}\emph{($f_0$: integrates position)}\\
-        \greenCode{$f_1($\code{o},\code{a},$\Delta)\, = \,$(set-vel o (+ (vel o) (* $\Delta$ a)))}\\
-        \hspace{\helpSize}\emph{($f_1$: integrates velocity)}\\
-        \greenCode{$f_2($\code{o},\code{F}$)\, = \,$(/ F (mass o))}\\
-        \hspace{\helpSize}\emph{($f_2$: Newton's second law)}\\
-        \greenCode{$f_3($\code{o},$\Delta$,\code{F}$)\, = \,$($f_0$ o $\Delta$ ($f_1$ o ($f_2$ o F) $\Delta$))}\\
-        \hspace{\helpSize}\emph{($f_3$: applies Newton's second law and integrates)}
-
-        \end{tabular}
-
-&
-
-  \\\bottomrule\\
-\end{tabular}}\vspace{-0.5cm} 
-\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
+%% %  \node [rotate=90] at (-2.3,-0.7) {\small program};
+
+%%   \end{scope}
+
+%% \begin{scope}[shift={(0,-5)}]
+%%   \node[pop3](p1) at (-1,-1) {\code{+}};
+%%   \node[pop3](n1) at (0.8,-0.7) {\code{1}};
+%%   \node[pop3](a) at (0,-1) {\code{ }};
+%%   %\node(x1) at (0,-1) {};
+%%   \draw[pop3] (0,0) -- (p1.north);
+%%   \draw[pop3] (0,0) -- (n1.north);
+%%   \draw[pop3] (-0.55,-0.4) -- (a.north);
+%% %  \node [rotate=90] at (-2.3,-0.7) {\small fragment};
+
+%%   \end{scope}
+
+%% \end{tikzpicture}
+%%     };
+
+%%     \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};
+
+%%     \end{tikzpicture}
+%% \end{figure}
+
+%% \begin{table*}%[t!]
+%%   \makebox[\textwidth][c]{
+%%     \scriptsize
+%%   \tabcolsep=4pt
+%%   \renewcommand\code\texttt
+%%   \renewcommand\codechar[1]{\texttt{"#1"}}
+%%   \newcommand{\helpSize}{0.25cm}
+%%   \begin{tabular}{cccc}
+%%     \toprule
+%%     &{\normalsize Symbolic regression}&{\normalsize Laws of motion}&\\\midrule
+%%     \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
+%%     \begin{tabular}{cc}
+%%       \includegraphics[width = 3em]{figures/functions/4.png}&
+%%       \includegraphics[width = 3em]{figures/functions/146}\\
+%%       \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
+%%       ~\\
+%%       \includegraphics[width = 3em]{figures/functions/112.png}&
+%%         \includegraphics[width = 3em]{figures/functions/92.png}
+%%       \\
+%%       \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\
+
+%%     \end{tabular}
+%%     }
+%%     &
+%%     \begin{tabular}{cc}
+%%       \includegraphics[width = 15em]{figures/massOnSpring.png} &
+%%             \includegraphics[width = 15em]{figures/planets.png}
+%%       \\
+%%       \begin{tabular}{l}
+%%               \blueCode{$f($o,$\Delta) = $($f_3$ o $\Delta$}\\
+%%       \hspace{1cm}\blueCode{(- (* $k$ (pos o))}\\
+%%       \hspace{1cm}\phantom{\code{(- }}\blueCode{(* $-9.8$ $\hat{x}$)))}\\
+%%       \end{tabular}&
+%%       \begin{tabular}{l}
+%%               \blueCode{$f($a,b,$\Delta) = $($f_3$ a $\Delta$}\\
+%%       \hspace{1cm}\blueCode{(/ (* $G$ (mass a) (mass b))}\\
+%%       \hspace{1cm}\phantom{\code{(- }}\blueCode{(square (- (pos a) (pos b)))))}\\
+%%         \end{tabular}
+%%       \end{tabular}
+%%     &
+
+%%     ~\\
+%%     \midrule
+%%     \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+%%       \begin{tabular}{l}
+%%     \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
+%%     \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
+%%     \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
+%%     \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
+%%     \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
+%%     \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
+%%     \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
+%%     \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
+%%     \hspace{\helpSize}\emph{($f_6$: rational function)}\\
+
+%%   \end{tabular}
+%%       &
+%%       \begin{tabular}{l}
+%%         \greenCode{$f_0($\code{o},$\Delta)\, = \,$(set-pos o (+ (pos o) (* $\Delta$ (vel o))))}\\
+%%         \hspace{\helpSize}\emph{($f_0$: integrates position)}\\
+%%         \greenCode{$f_1($\code{o},\code{a},$\Delta)\, = \,$(set-vel o (+ (vel o) (* $\Delta$ a)))}\\
+%%         \hspace{\helpSize}\emph{($f_1$: integrates velocity)}\\
+%%         \greenCode{$f_2($\code{o},\code{F}$)\, = \,$(/ F (mass o))}\\
+%%         \hspace{\helpSize}\emph{($f_2$: Newton's second law)}\\
+%%         \greenCode{$f_3($\code{o},$\Delta$,\code{F}$)\, = \,$($f_0$ o $\Delta$ ($f_1$ o ($f_2$ o F) $\Delta$))}\\
+%%         \hspace{\helpSize}\emph{($f_3$: applies Newton's second law and integrates)}
+
+%%         \end{tabular}
+
+%% &
+
+%%   \\\bottomrule\\
+%% \end{tabular}}\vspace{-0.5cm} 
+%% \caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them.
 Bottom: Several examples from learned DSL.
 Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}%\vspace{-0.5cm}
-\end{table*}
+%% \end{table*}
+
 \end{document}

### 2018-12-30 archive.tex
symmetry breaking proof rough draft

@@
 ($\probability[x|p] > 0$); if so, we incorporate $p$ into the frontier
-$\mathcal{F}_x$.  We represent programs as polymorphicly-typed
-$\lambda$-calculus expressions, a representation closely resembling
-Lisp and functional languages like Haskell and OCaml, including
-variables, conditionals, higher-order recursive functions, and the
-ability to create new functions.
+$\mathcal{F}_x$ (Figure~\ref{neuralPipeline}).  We represent programs
+as polymorphicly-typed $\lambda$-calculus expressions, a
+representation closely resembling Lisp and functional languages like
+Haskell and OCaml, including variables, conditionals, higher-order
+recursive functions, and the ability to create new functions.
 
@@
 the space of  programs.
+\begin{figure}
+  \begin{tikzpicture}[every node/.style={rounded corners,thick}]
+    \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
+    \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
+      e.g., RNN or CNN};
+%    \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
+    \node[draw,align = center](s) at ([xshift=2cm]e.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
+    \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};
+
+    \draw[->] (x.east) -- (e.west);
+    \draw[->] (e.east) -- (s.west);
+%    \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
+    \draw[->] (s.east) -- (f.west);
+    \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
+    \node[align=center] at ($(0,-1) + (e.center)$) {$\underbrace{\hspace{4cm}}_{\normalsize\text{\normalsize Recognition model $Q(p|x)$}}$};
+    \node[align=center] at ($(s.west)!0.5!(f.east) + (0,-1)$) {$\underbrace{\hspace{5cm}}_{\text{\normalsize Search}}$};
+  \end{tikzpicture}
+  \caption{Neurally-guided program inference pipeline.
+Recognition model outputs distribution over program $Q(p|x)$.
+Program output by enumerative search incorporated into frontier if likelihood $\probability[x|p] > 0$}
+  \label{neuralPipeline}
+\end{figure}
 
@@
 as input a program $p$ and then outputs a version space containing the
-refactorings of $p$.  Our technique is astronomically more efficient
-than explicitly representing the space of possible refactorings: for
-the example in Figure~\ref{mapFactor}A, we represent the space of
-refactorings using a version space with $10^6$ nodes, which encodes
-$10^{14}$ refactorings.
+refactorings of $p$. Figure~\ref{mapFactor}B diagrams a subtree of a
+version space containing refactorings of a small program. Our
+technique is substantially more efficient than explicitly representing
+the space of possible refactorings: for the example in
+Figure~\ref{mapFactor}A, we represent the space of refactorings using
+a version space with $10^6$ nodes, which encodes $10^{14}$
+refactorings.
 
@@
 
-  \draw [->] (p1.south)  --(r1.north) node[fill=white,midway] {refactor};
-  \draw [->] (p2.south)  --(r2.north) node[fill=white,midway] {refactor};
+  \draw [->] (p1.south)  --(r1.north) node[fill=white,midway,align=center] {refactor\\($\approx 10^{14}$ refactorings)};
+  \draw [->] (p2.south)  --(r2.north) node[fill=white,midway,align=center] {refactor\\($\approx 10^{14}$ refactorings)};
 
@@
 Intuitively, these new DSL components best compress the programs found during waking.
-\textbf{Panel B:} \# of possible refactorings grows exponentially with program size.
-We represent refactorings using version spaces, which augment syntax trees with a \emph{union} operator whose children are themselves version spaces.
-Right graph: version spaces are exponentially more efficient than explicitly constructing set of refactorings.}\label{mapFactor}
+\textbf{Panel B:} \# of possible refactorings grows exponentially with program size, so we represent refactorings using version spaces, which augment syntax trees with a \emph{union} operator whose children are themselves version spaces.
+Right graph: version spaces are exponentially more efficient than explicitly constructing set of refactorings.
+In this graph, refactored programs are of the form $1+1+\cdots  + 1$.}\label{mapFactor}
 \end{figure*}
@@
 during waking, and \emph{fantasies}, or programs drawn from the DSL.
-Replays ensure that the recognition model is trained on the tasks
-solved so far, and does not forget how to solve them.  Fantasies
-ensure that the recognition model has a large and highly varied corpus
-of (program, task) pairs to learn from.
-
-Formally, $Q(p|x)$ should approximate the posterior
-$\probability[p|\mathcal{D},\theta,x]$
-
-training a neural recognition model, $Q$, to regress from
-a task to a distribution over programs
-that are likely to both
-solve the task and
-be likely under the prior.
-To train $Q$ we need (program, task)
-pairs.
-
-To get the data to train $Q$,
-we take inspiration from
-the
-
-The purpose of training the recognition model is to amortize the cost
-of searching for programs.  It does this by learning to predict, for
-each task, programs with high likelihood according to
-$\probability[x|p]$ while also being probable under the prior
-$(\mathcal{D},\theta)$,
-thereby approximating the posterior $\probability[\cdot |x,\mathcal{D},\theta]$.
-
+Replays ensure that the recognition model is trained on the actual
+tasks it needs to solve, and does not forget how to solve them.
+Fantasies ensure that the recognition model has a large and highly
+varied corpus of (program, task) pairs to learn from.
+
+Formally, the recognition model $Q(p|x)$ should approximate the posterior
+$\probability[p|\mathcal{D},\theta,x]$.
 We can either train $Q$ to perform full posterior inference by minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|Q(p|x) \right) \right]$,
@@
 where in both cases the expectation is taken over tasks.
-Taking this expectation over the empirical distribution of tasks trains $Q$ on the real data; taking it over samples from the generative model trains $Q$ on ``dreams.''
+Taking this expectation over the empirical distribution of tasks trains $Q$ on replays; taking it over samples from the generative model trains $Q$ on fantasies.
 We define a pair of alternative objectives for the recognition model,
-$\mathcal{L}_{\text{RM}}^{\text{posterior}}$ and $\mathcal{L}_{\text{RM}}^{\text{MAP}}$,
+$\mathcal{L}^{\text{posterior}}$ and $\mathcal{L}^{\text{MAP}}$,
 which either train $Q$ to perform full posterior inference or MAP inference, respectively.
-These objectives combine real-data and dream data:
+These objectives combine replays and fantasies:
 \begin{align*}
-  \mathcal{L}_{\text{RM}}^{\text{posterior}} &= \mathcal{L}_{\text{Real}}^{\text{posterior}} + \mathcal{L}_{\text{Dream}}^{\text{posterior}}&
-  \mathcal{L}_{\text{RM}}^{\text{MAP}} &= \mathcal{L}_{\text{Real}}^{\text{MAP}} + \mathcal{L}_{\text{Dream}}^{\text{MAP}}\\
-  \mathcal{L}_{\text{Real}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+  \mathcal{L}^{\text{posterior}} &= \mathcal{L}_{\text{Replay}}^{\text{posterior}} + \mathcal{L}_{\text{Fantasy}}^{\text{posterior}}&
+  \mathcal{L}^{\text{MAP}} &= \mathcal{L}_{\text{Replay}}^{\text{MAP}} + \mathcal{L}_{\text{Fantasy}}^{\text{MAP}}\\
+  \mathcal{L}_{\text{Replay}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
     \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
-\mathcal{L}_{\text{Real}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
-  \mathcal{L}_{\text{Dream}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
-    \mathcal{L}_{\text{Dream}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
+  \mathcal{L}_{\text{Replay}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
+  \mathcal{L}_{\text{Fantasy}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
+  \mathcal{L}_{\text{Fantasy}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
 \end{align*}
-  The `dream' objectives are essential for data efficiency:
-all of our experiments train \system on only a few hundred tasks, which is too little for
-a high-capacity neural network.
-Once we bootstrap a $(\mathcal{D},\theta)$,
-we can draw unlimited samples from $(\mathcal{D},\theta)$
-and train $Q$ on those samples.
-But, evaluating $\mathcal{L}_{\text{Dream}}$ involves drawing programs from
-the current DSL, running them to get their outputs,
-and then training $Q$ to regress from the input/outputs to the program.
-Since these programs map inputs to outputs,
-we need to sample the inputs as well.
-Our solution is to sample the inputs
-from the empirical observed distribution of inputs in $X$.
-
-The $\mathcal{L}_{\text{Dream}}^{\text{MAP}}$ objective involves
-finding the MAP program solving a task drawn from the DSL.
-To make this tractable,
-rather than \emph{sample} programs as training data for $\mathcal{L}_{\text{Dream}}^{\text{MAP}}$,
-we \emph{enumerate} programs
-in decreasing order of their prior probability,
-tracking,
-for each dreamed task $x$,
-the set of enumerated programs maximizing
-$\probability[x,p|\mathcal{D},\theta]$.
+We maximize $\mathcal{L}^{\text{MAP}}$ rather than
+$\mathcal{L}^{\text{posterior}}$ for two reasons:
+$\mathcal{L}^{\text{MAP}}$ prioritizes the shortest program solving a
+task, thus more strongly accelerating enumerative search during waking;
+and, combined with our parameterization of $Q$, described next, we
+will show that $\mathcal{L}^{\text{MAP}}$ forces the recognition model
+to break symmetries in the space of programs
+(Sec.~\ref{symmetricSection}).
 
@@
 parameterizing conditional distributions over programs.  The first
-approach, seen in systems like RobustFill~\cite{devlin2017robustfill}
-and~\cite{zavershynskyi2018naps}, is to use a recurrent neural network
-to predict the program token-by-token as in Seq2Seq or Seq2Tree.  The
-advantage of this approach is that the neural net can predict the
-entire program, so if the network is sufficiently powerful, it can
-completely solve the synthesis problem.  There are two disadvantages
-to this approach.  First, these models can perform poorly at out-of-sample
+approach~\cite{devlin2017robustfill,zavershynskyi2018naps} is to use a
+recurrent network to predict the entire program token-by-token, which
+has the advantage that, if the network is sufficiently powerful, it
+can completely solve the synthesis problem.  The disadvantage is that
+these models can perform poorly at out-of-sample
 generalization~\cite{}, which is critical for our setting, as the
 agent may need to solve new tasks that are qualitatively different
-from the tasks it has solved so far.  Second, a powerful deep
-recurrent network may be costly to sample or enumerate from ---
-so if the network cannot easily solve a task, we cannot compensate
-with rapid sampling or enumeration.  In contrast, state-of-the-art
-enumerative program synthesizers evaluate millions of programs per
-second~\cite{feser2015synthesizing}.
+from the tasks it has solved so far.
+
+%% Second, a powerful deep
+%% recurrent network may be costly to sample or enumerate from --- so if
+%% the network cannot easily solve a task, we cannot compensate with
+%% rapid sampling or enumeration.  In contrast, state-of-the-art
+%% enumerative program synthesizers evaluate millions of programs per
+%% second~\cite{feser2015synthesizing}.
 
 The second approach is to have $Q$ predict a fixed-dimensional weight
-vector, which then biases a fast enumerator~\cite{balog2016deepcoder}
+vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ec2}
 or sampler~\cite{menon2013machine}.  This approach can enjoy strong
@@
 take the form of (parent, child, argument index).
-
-Formally,
-$Q$ predicts a $(|\mathcal{D}|+2)\times (|\mathcal{D}|+1)\times A$-dimensional tensor,
-where $A$ is the maximum arity\footnote{The arity of a function is the number of arguments that it takes as input.} of any primitive in the DSL.
-Slightly abusing notation, we write this tensor as $Q_{ijk}(x)$,
-where $x$ is a task,
-$i\in \mathcal{D}\cup\left\{\text{start},\text{var}\right\}$,
-$j\in \mathcal{D}\cup\left\{\text{var} \right\}$,
-and $k\in \left\{1,2,\cdots,A \right\}$.
-The output $Q_{ijk}(x)$
-controls the probability of
-sampling primitive $j$ given that
-$i$ is the parent node in the syntax tree
-and we are sampling the $k^{\text{th}}$ argument.
-Figure~\ref{bg} diagrams this generative process, Algorithm~\ref{recognitionSample}
-specifies a sampling procedure for $Q(\cdot |x)$,
-and Figure~\ref{neuralPipeline} diagrams the neurally guided program inference procedure.
+Figure~\ref{bg} diagrams this generative process and Algorithm~\ref{recognitionSample}
+specifies a sampling procedure for $Q(\cdot |x)$.
 This parameterization
@@
 \end{figure}
- \begin{algorithm}
-   \caption{Drawing from distribution over programs predicted by recognition model. Compare w/ Algorithm~\ref{sampleProgram}}
-   \label{recognitionSample}
-   \begin{algorithmic}[1]
-     \State \textbf{function} recognitionSample$(Q, x, \mathcal{D}, \tau)$:
-     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, type $\tau$
-     \State \textbf{Output:} a program whose type unifies with $\tau$
-     \State \textbf{return } $\text{recognitionSample}'(Q,x,\text{start},1,\mathcal{D},\varnothing,\tau)$
-     \Statex
-
-     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
-     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
-     \State \textbf{Output:} a program whose type unifies with $\tau$
-     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
-     \State var $\gets$ an unused variable name
-     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
-     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
-     %\Endif
-     \Else\Comment{Build an application to give something w/ type $\tau$}
-     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
-     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
-     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
-     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
-       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
-Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
-       \end{cases}$
-     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
-     \State newParent$\gets\begin{cases}
-     e &\text{ if }e\in \mathcal{D}\\
-     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
-     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
-     %   \State unify$(\tau,\beta)$
-     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
-     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
-     \EndFor
-     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
-     \EndIf
-     %% \Statex
-     %% \Statex\textbf{where:}
-     %% \State yield$(\tau) = \begin{cases}
-     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-     %%   \tau   &\text{ otherwise.}
-     %% \end{cases}$ \Comment{Final return type of $\tau$}
-     %% \State  args$(\tau) = \begin{cases}
-     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-     %%   []   &\text{ otherwise.}
-     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
-   \end{algorithmic}
- \end{algorithm}
- \begin{figure}
-   \begin{tikzpicture}[every node/.style={rounded corners,thick}]
-     \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
-     \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
-       e.g., RNN or CNN};
-     \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
-     \node[draw,align = center](s) at ([xshift=3cm]m.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
-     \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};
-
-     \draw[->] (x.east) -- (e.west);
-     \draw[->] (e.east) -- (m.west);
-     \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
-     \draw[->] (s.east) -- (f.west);
-     \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
-     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
-     ([xshift = -5,yshift = -25]e.west) -- node[below = 3pt] {{Recognition model $Q(p|x)$}} ([xshift = 155,yshift = -25]e.west);
-     \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
-     ([xshift = -5,yshift = -25]s.west) -- node[below = 3pt] {{ Search}} ([xshift = 5,yshift = -25]f.east);
-     %% \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
-     %% ([xshift = -5,yshift = -25]p.west) -- node[below = 3pt] {{\small Params}} ([xshift = 5,yshift = -25]p.east);
-   \end{tikzpicture}
-   \caption{Neurally-guided program inference pipeline.
-Recognition model outputs distribution over program $Q(p|x)$ parameterized by tensor $Q_{ijk}$.
-Program output by enumerative search incorporated into frontier if likelihood $\probability[x|p] > 0$}
-   \label{neuralPipeline}
-   \end{figure}
-
-\subsubsection{Learning to break symmetries in program space}
-
-A good DSL not only exposes high-level building blocks,
-but also carefully restricts the ways in which those building blocks are allowed to compose.
-For example,
-a DSL for list manipulation should contain both the empty list and
-a routine for appending lists,
-but should not allow
-appending the empty list.
-Similarly a DSL for arithmetic should contain both addition and
-the number zero but
-disallow adding zero.
-These restrictions break symmetries in the space of programs.
-Here we show that the
-combination of the $\mathcal{L}^{\text{MAP}}$ training objective (Section~\ref{recognitionTraining}) and bigram parameterization (Section~\ref{recognitionParameterization})
-interact in a way that spontaneously breaks symmetries in the program space,
-allowing the agent to more efficiently
-explore the space of programs.
-
-Consider an agent tasked with writing programs built from addition and the constants zero and one.
-A bigram parameterization of $Q$ allows
-it to represent the fact that it should never add zero ($Q_{\code{+},\code{0},0} = Q_{\code{+},\code{0},1} = 0$)
-%or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
-or that addition should always associate to the right
-($Q_{\code{+},\code{+},0} = 0$).
-The $\mathcal{L}^{\text{MAP}}$ training objective encourages
-learning these canonical forms.
-Consider two recognition models, $Q_1$ and $Q_2$,
-and two programs in frontier $\mathcal{F}_x$,
-$p_1=\code{(+ (+ 1 1) 1)}$ and $p_2=\code{(+ 1 (+ 1 1))}$,
-where
-\begin{align*}
-  Q_1(p_1|x) = \frac{\epsilon}{2}&\qquad Q_1(p_2|x) = \frac{\epsilon}{2}\\
-  Q_2(p_1|x) = 0&\qquad  Q_2(p_2|x) = \epsilon
-\end{align*}
-i.e., $Q_2$ breaks a symmetry by forcing right associative addition,
-but $Q_1$ does not, instead splitting its probability mass equally between $p_1$ and $p_2$.
-Now because $\probability[p_1|\mathcal{D},\theta] = \probability[p_2|\mathcal{D},\theta]$
-(Algorithm~\ref{sampleProgram}),
-we have
-\begin{align*}
-  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_1)& = \frac{\probability[p_1|\mathcal{D},\theta]\log \frac{\epsilon}{2} + \probability[p_2|\mathcal{D},\theta]\log \frac{\epsilon}{2}}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = \log \frac{\epsilon}{2}\\
-  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_2)& = \frac{\probability[p_1|\mathcal{D},\theta]\log 0 + \probability[p_2|\mathcal{D},\theta]\log \epsilon}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = +\infty\\
-  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_1)& = \log Q_1(p_1)           = \log Q_1(p_2)   = \log \frac{\epsilon}{2}\\
-  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_2)& = \log Q_2(p_2) \phantom{ = \log Q_1(p_2) } = \log \epsilon\\
-\end{align*}
-So $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking
-recognition model), while $\mathcal{L}^{\text{posterior}}$ reverses
-this preference.  We experimentally confirm this symmetry-breaking
+
+\subsubsection{Learning to break symmetries in program space}\label{symmetricSection}
+
+A good DSL not only exposes high-level building blocks, but also
+carefully restricts the ways in which those building blocks are
+allowed to compose.  For example,
+%% a DSL for list manipulation should
+%% contain both the empty list and a routine for appending lists, but
+%% should not allow appending the empty list.  Similarly 
+a DSL for arithmetic should contain both addition and the number zero
+but disallow adding zero.  These restrictions break symmetries in the
+space of programs.  A bigram parameterization of the recognition
+model, combined with the $\mathcal{L}^{\text{MAP}}$ training
+objective, interact in a way that breaks symmetries in the program
+space, allowing the agent to more efficiently explore the space of
+programs.  This interaction occurs because the bigram parameterization
+can disallow DSL primitives depending on their local syntactic
+context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
+probability mass onto a single member of a set of syntactically
+distinct but semantically equivalent expressions
+(Appendix~\ref{recognitionAppendix}).
+
+We experimentally confirm this symmetry-breaking
 behavior by training recognition models that minimize either
@@
 
-To be clear,
-our recognition model does not learn to break \emph{every} possible symmetry in every possible DSL.
-But in practice we found that a bigrams combined with $\mathcal{L}^{\text{MAP}}$
-works well,
-and we use with this combination throughout the rest of the paper.
 \begin{figure}
@@
 We first apply \system to two classic benchmark domains: list
-processing and text editing. We find that we can start with generic
-Lisp-style primitives and recover a domain-specific vocabulary for
-solving each of these classes of problems.
-
-using a GRU~\cite{cho2014learning} for
-the recognition model, and initially providing the system with generic sequence manipulation primitives:
-\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
-\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
-\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.
-
-%\subsubsection{List Processing}\label{listSection}
-\textbf{List Processing:} Synthesizing programs that manipulate data structures is a widely studied
-problem in the programming languages community~\cite{feser2015synthesizing},
-and progress here can translate to better computer-aided programming~\cite{solar2008program}.
-We consider this problem within the context of learning functions that
-manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
-We created 236 human-interpretable list manipulation tasks, each with 15
-input/output examples.
-%% Our data set is interesting in three major ways: many of the tasks
-%% require complex solutions; the tasks were not generated from some
-%% latent DSL, and the agent must learn to solve these complicated
-%% problems from only 236 tasks.
-%% Our data set assumes arithmetic operations as well as sequence operations,
-%% so we additionally provide our system with the following arithmetic
-%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
-%% \code{is-prime}.
-%% evaluating  on a 50/50 test/train split.
-In solving these tasks, the system
-composed 38 new subroutines, and rediscovered the higher-order
-function \code{filter} ($f_1$ in Table~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
-%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
-\begin{figure}[b]\centering
-\vspace{-0.5cm}  \begin{tabular}{lll}
-    \toprule
-    Name & Input & Output \\\midrule
-    repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
-    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-    keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
-    product & [7\, 1\, 6\, 2] & 84 \\
-    \bottomrule
-  \end{tabular}
-  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
-\end{figure}
-
-%\subsubsection{Text Editing}\label{textSection}
-\textbf{Text Editing:} Synthesizing programs that edit text is a classic problem in the
-programming languages and AI literatures~\cite{gulwani2011automating,lau2001programming}.
-This prior work uses hand-engineered DSLs.
-Here, we instead start out with generic sequence manipulation
-primitives and recover many of the higher-level building blocks that 
-have made these other systems successful.
+processing and text editing. In both cases we solve tasks specified by
+a few input/output examples, using a recurrent neural network for the
+recognition model and starting with a generic functional programming
+basis: \code{foldr}, \code{unfold}, \code{if}, \code{map},
+\code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
+\code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
+\code{is-nil}.
+
+\subsubsection{List Processing}\label{listSection}
+We took 218 list manipulation tasks from our previous work~\cite{ec2},
+each with 15 input/output examples.  In solving these tasks, the
+system composed 16 new DSL components, and discovered multiple
+higher-order functions, such as \code{filter} ($f_0$ in
+Figure~\ref{initialExampleDSL}).  Each round of memory consolidation
+built on components discovered in earlier sleep cycles --- for example
+the agent first learns \code{filter}, uses filter to learn to take the
+maximum element of a list, then uses that routine to learn a new
+component for extracting the $n^{\text{th}}$ largest element of a
+list, which it finally uses to learn how to sort a list of numbers
+(Figure~\ref{initialExampleDSL}).  This incremental, modular learning
+of deep hierarchies of DSL components is characteristic of \system
+because of the alternation between code writing and code refactoring.
+
+%% \begin{figure}[b]\centering
+%% \vspace{-0.5cm}  \begin{tabular}{lll}
+%%     \toprule
+%%     Name & Input & Output \\\midrule
+%%     repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
+%%     drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
+%%     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
+%%     count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
+%%     keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
+%%     product & [7\, 1\, 6\, 2] & 84 \\
+%%     \bottomrule
+%%   \end{tabular}
+%%   \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
+%% \end{figure}
+
+\subsubsection{Text Editing}\label{textSection}
+Synthesizing programs that edit text is a classic problem in the
+programming languages and AI literatures~\cite{lau2001programming},
+and algorithms that synthesize text editing programs ship in Microsoft
+Excel~\cite{gulwani2011automatin}.  This prior work uses
+hand-engineered DSLs and hand-engineered search strategies.  Here, we
+will show that we can jointly learn both these ingredients and surpass
+the state-of-the-art domain-general program synthesizers on a standard
+text editing benchmark.
+
 %% Because our enumerative search procedure cannot generate string %
@@
 
-We trained our system on 109 automatically-generated text editing tasks, with 4 input/output examples each.
-After three iterations, it assembles a DSL containing a dozen new functions (Fig.~\ref{initialExampleDSL}, center) solving
-all the training tasks.
-But, how well does the  learned DSL generalized to real text-editing scenarios?
+We trained our system on 128 automatically-generated text editing tasks, with 4 input/output examples each.
 We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
-\system solves 3.7\% of the problems with an average search time of 235 seconds.
+\system solves 3.7\% of the problems within 10 minutes with an average search time of 235 seconds.
 After learning,
-it solves 74.1\%, and does so much faster,
-solving them in an average of 29 seconds.
+it solves 79.6\%, and does so much faster,
+solving them in an average of 40 seconds.
 As of the 2017 SyGuS competition,
-the best-performing algorithm solves 79.6\% of the problems.
-But, SyGuS comes with a
+the best-performing synthesizer (CVC4) solves 82.4\% of the problems ---
+but here, the competition conditions are 1 hour \& 8 CPUs per problem,
+and with this more generous compute budget we
+surpass
+this previous
+result and solve
+84.3\% of the problems.
+SyGuS additionally comes with a
 different hand-engineered DSL \emph{for each text editing problem}.
@@
 
-We take inspiration from LOGO Turtle graphics~\cite{turtle},
-and task our agent with drawing 90 different pictures (Figure~\ref{everyLogo}).
-The agent has primitives for moving the pen forward and rotating it (\code{move}),
-picking the pen up and then putting it down (\code{pen-up}),
-arithmetic operations on angles and distances,
-and constants ($2\pi$, 0 through 9, and $\infty$\footnote{To ensure that programs terminate, we set $\infty = 20$}).
-
-Over the course of 5 wake/sleep iterations,
-the system authors a DSL
-containing parametric and, sometimes, higher-order drawing routines,
-allowing it to write programs for objects like
-polygons, circles, spirals, and snowflakes.
-Figure~\ref{logoPrimitives}
-illustrates several learned DSL subroutines.
-Once it has synthesized these new subroutines,
-the agent generates a rich and highly varied
-data set of dreams on which to train the recognition model.
-Figure~\ref{logoDreams}
-illustrates  dreams before and after learning ---
-one sees that
-the system has learned to recombine
-the building blocks and motifs in the training data.
+Procedural visual concepts are studied across AI and cognitive science
+--- Bongard problems~\cite{Moscow}, Raven's progressive
+matrices~\cite{raven2003raven}, and Lake et al.'s BPL model of
+omniglot~\cite{lake2015human} are prominent examples.  Here we take
+inspiration from LOGO Turtle graphics~\cite{turtle}, and give our
+agent the ability to control a pen, along with arithmetic operations
+on angles and distances and simple control flow primitives.  We task
+it with drawing a small corpus of images starting from this basis.
+
+Inside its learned DSL we find interpretable parametric drawing
+routines corresponding to the families of visual objects in its
+training data, like polygons, circles, and spirals
+(Figure~\ref{logoPrimitives}, left). It additionally learns more
+abstract visual relationships, like rotational symmetry, which it
+models by incorporating a higher-order function into its DSL
+(Figure~\ref{logoPrimitives}, right).  This abstraction comes from
+jointly compressing programs for all of the training images into its
+DSL.
+
+What does \system dream of?  Prior to learning samples from the DSL
+are simple and largely unstructured (Figure~\ref{logoDreams}, left).
+After training the samples become richly structured
+(Figure~\ref{logoDreams}, right), compositionally recombining latent
+building blocks and motifs acquired from the training data. This
+offers a window into how the generative model bootstraps recognition
+model training: as the DSL grows more finely tuned to the domain, the
+neural net gets richer and more highly varied training data.
+
+%% has primitives for moving the pen forward and rotating it (\code{move}),
+%% picking the pen up and then putting it down (\code{pen-up}),
+%% arithmetic operations on angles and distances,
+%% and constants ($2\pi$, 0 through 9, and $\infty$\footnote{To ensure that programs terminate, we set $\infty = 20$}).
+
 \begin{figure}
@@
 
+\subsection{Recognition model training}\label{recognitionAppendix}
+
+Recall that our goal is to maximize either $\mathcal{L}^{posterior}$ or $\mathcal{L}^{MAP}$, defined as:
+\begin{align*}
+  \mathcal{L}^{\text{posterior}} &= \mathcal{L}_{\text{Replay}}^{\text{posterior}} + \mathcal{L}_{\text{Fantasy}}^{\text{posterior}}&
+  \mathcal{L}^{\text{MAP}} &= \mathcal{L}_{\text{Replay}}^{\text{MAP}} + \mathcal{L}_{\text{Fantasy}}^{\text{MAP}}\\
+  \mathcal{L}_{\text{Replay}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
+    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
+\mathcal{L}_{\text{Replay}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
+  \mathcal{L}_{\text{Fantasy}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
+    \mathcal{L}_{\text{Fantasy}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
+\end{align*}
+The fantasy objectives are essential for data efficiency:
+all of our experiments train \system on only a few hundred tasks, which is too little for
+a high-capacity neural network.
+Once we bootstrap a $(\mathcal{D},\theta)$,
+we can draw unlimited samples from $(\mathcal{D},\theta)$
+and train $Q$ on those samples.
+But, evaluating $\mathcal{L}_{\text{Fantasy}}$ involves drawing programs from
+the current DSL, running them to get their outputs,
+and then training $Q$ to regress from the input/outputs to the program.
+Since these programs map inputs to outputs,
+we need to sample the inputs as well.
+Our solution is to sample the inputs
+from the empirical observed distribution of inputs in $X$.
+
+The $\mathcal{L}_{\text{Fantasy}}^{\text{MAP}}$ objective involves
+finding the MAP program solving a task drawn from the DSL.  To make
+this tractable, rather than \emph{sample} programs as training data
+for $\mathcal{L}_{\text{Fantasy}}^{\text{MAP}}$, we \emph{enumerate}
+programs in decreasing order of their prior probability, tracking, for
+each dreamed task $x$, the set of enumerated programs maximizing
+$\probability[x,p|\mathcal{D},\theta]$.
+
+We parameterize $Q$ using a bigram model over syntax trees.
+Formally,
+$Q$ predicts a $(|\mathcal{D}|+2)\times (|\mathcal{D}|+1)\times A$-dimensional tensor,
+where $A$ is the maximum arity\footnote{The arity of a function is the number of arguments that it takes as input.} of any primitive in the DSL.
+Slightly abusing notation, we write this tensor as $Q_{ijk}(x)$,
+where $x$ is a task,
+$i\in \mathcal{D}\cup\left\{\text{start},\text{var}\right\}$,
+$j\in \mathcal{D}\cup\left\{\text{var} \right\}$,
+and $k\in \left\{1,2,\cdots,A \right\}$.
+The output $Q_{ijk}(x)$
+controls the probability of
+sampling primitive $j$ given that
+$i$ is the parent node in the syntax tree
+and we are sampling the $k^{\text{th}}$ argument.
+Algorithm~\ref{recognitionSample}
+specifies a procedure for drawing samples from $Q(\cdot |X)$.
+ \begin{algorithm}
+   \caption{Drawing from distribution over programs predicted by recognition model. Compare w/ Algorithm~\ref{sampleProgram}}
+   \label{recognitionSample}
+   \begin{algorithmic}[1]
+     \State \textbf{function} recognitionSample$(Q, x, \mathcal{D}, \tau)$:
+     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, type $\tau$
+     \State \textbf{Output:} a program whose type unifies with $\tau$
+     \State \textbf{return } $\text{recognitionSample}'(Q,x,\text{start},1,\mathcal{D},\varnothing,\tau)$
+     \Statex
+
+     \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
+     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
+     \State \textbf{Output:} a program whose type unifies with $\tau$
+     \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
+     \State var $\gets$ an unused variable name
+     \State body $\sim$ recognitionSample$'(Q,x,\text{parent},\text{argumentIndex},\mathcal{D},\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$%\Comment{Recursively sample function body}
+     \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
+     %\Endif
+     \Else\Comment{Build an application to give something w/ type $\tau$}
+     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
+     \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
+     \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
+       Q_{\text{parent},e,\text{argumentIndex}}(x)&\text{ if }e\in \mathcal{D}\\
+Q_{\text{parent},\text{var},\text{argumentIndex}}(x)/|\text{variables}|&\text{ if }e\in \mathcal{E}
+       \end{cases}$
+     \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
+     \State newParent$\gets\begin{cases}
+     e &\text{ if }e\in \mathcal{D}\\
+     \text{var}&\text{ if }e\in \mathcal{E}\end{cases}$
+     \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
+     %   \State unify$(\tau,\beta)$
+     \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
+     \State $a_k\sim\text{recognitionSample}'(Q,x,\text{newParent},k,\mathcal{D},\mathcal{E},\alpha_k)$
+     \EndFor
+     \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
+     \EndIf
+     %% \Statex
+     %% \Statex\textbf{where:}
+     %% \State yield$(\tau) = \begin{cases}
+     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+     %%   \tau   &\text{ otherwise.}
+     %% \end{cases}$ \Comment{Final return type of $\tau$}
+     %% \State  args$(\tau) = \begin{cases}
+     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
+     %%   []   &\text{ otherwise.}
+     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
+   \end{algorithmic}
+ \end{algorithm}
+
+ \textbf{Symmetry breaking analysis.}
+
+ \begin{theorem}
+   Let $\mu(\cdot )$ be a distribution over tasks and let $Q^*(\cdot |\cdot )$ be a task-conditional distribution over programs satisfying
+   $$
+   Q^* = \argmax_Q\;\expect_\mu\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log  Q(p|x) \right]
+   $$
+   where $(\mathcal{D},\theta)$ is a generative model over programs. Pick a task $x$ where $\mu(x) > 0$.
+Partition $\Lambda$ into expressions that are observationally equivalent under $x$:
+   $$
+\Lambda = \bigcup_i \mathcal{E}^x_i\text{ where for any }p_1\in \mathcal{E}^x_i\text{ and }p_2\in \mathcal{E}^x_j\text{: }\probability[x|p_1] = \probability[x|p_2] \iff i = j
+$$
+Then there exists an equivalence class $\mathcal{E}^x_i$
+that gets all the probability mass of $Q^*$ -- e.g.,
+$Q^*(p|x) = 0$ whenever $p\not\in \mathcal{E}^x_i$ --
+and there exists a program in that equivalence class
+which gets all of the probability mass assigned by $Q^*(\cdot |x)$ -- e.g.,  there is a $p\in \mathcal{E}^x_i$
+such that $Q^*(p|x) = 1$ -- and that program maximizes $\probability[\cdot |x,\mathcal{D},\theta ]$.
+ \end{theorem}
+ \begin{proof}
+   We proceed by defining the set of ``best programs'' -- programs maximizing the posterior $\probability[\cdot |x,\mathcal{D},\theta]$ -- and then showing that a best program satisfies $Q^*(p|x) = 1$.
+   Define the set of best programs $\mathcal{B}_x$ for the task $x$ by
+   $$
+\mathcal{B}_x = \left\{p\;|\;\probability[p|x,\mathcal{D},\theta] = \max_{p'\in \Lambda}\probability[p'|x,\mathcal{D},\theta] \right\}
+$$
+For convenience define
+$$
+f(Q) = \expect_\mu\left[\max_{p\in \mathcal{B}_x}\log Q(p|x) \right]
+$$
+and observe that $Q^* = \argmax_Q f(Q)$.
+
+Suppose by way of contradiction that there is a $q\not\in \mathcal{B}_x $
+where $Q^*(q|x) = \epsilon > 0$.
+Let $p^*= \argmax_{p\in \mathcal{B}_x}\log Q^*(p|x)$.
+Define
+$$
+Q'(p|x) = \begin{cases}
+  0&\text{ if }p = q\\
+  Q^*(p|x) + \epsilon&\text{ if }p = p^*\\
+  Q^*(p|x)&\text{ otherwise.}
+\end{cases}
+$$
+Then
+\begin{align*}
+  f(Q') - f(Q^*) &= \mu(x) \left(\max_{\substack{p\in \mathcal{B}_x}}\log  Q'(p|x) - \max_{\substack{p\in \mathcal{B}_x}}\log  Q^*(p|x) \right) = \mu(x)\left(\log \left(Q^*(p^*|x) + \epsilon \right) - \log Q^*(p^*|x) \right) > 0
+\end{align*}
+which contradicts the assumption that $Q^*$ maximizes $f(\cdot )$.
+Therefore
+for any $p\not\in \mathcal{B}_x$ we have
+$Q^*(p|x) = 0$.
+
+Suppose by way of contradiction that there are two distinct programs, $q$ and $r$,
+both members of $\mathcal{B}_x$,
+where $Q^*(q|x) = \alpha > 0$
+and $Q^*(r|x) = \beta > 0$.
+Let $p^*= \argmax_{p\in \mathcal{B}_x}\log Q^*(p|x)$.
+If $p^*\not\in \left\{q,r \right\}$ then define
+$$
+Q'(p|x) = \begin{cases}
+  0&\text{ if }p\in \left\{q,r \right\}\\
+  Q^*(p|x) + \alpha + \beta&\text{ if }p = p^*\\
+  Q^*(p|x)&\text{ otherwise.}
+\end{cases}
+$$
+Then
+\begin{align*}
+  f(Q') - f(Q^*) &= \mu(x) \left(\max_{\substack{p\in \mathcal{B}_x}}\log  Q'(p|x) - \max_{\substack{p\in \mathcal{B}_x}}\log  Q^*(p|x) \right)\\ &= \mu(x)\left(\log \left(Q^*(p^*|x) + \alpha + \beta \right) - \log Q^*(p^*|x) \right) > 0  \end{align*}
+which contradicts the assumption that $Q^*$ maximizes $f(\cdot )$.
+Otherwise assume $p^*\in \left\{q,r \right\}$.
+Without loss of generality let $p^* = q$.
+Define
+$$
+Q'(p|x) = \begin{cases}
+  0&\text{ if }p = r\\
+  Q^*(p|x) + \beta&\text{ if }p = p^*\\
+  Q^*(p|x)&\text{ otherwise.}
+\end{cases}
+$$
+Then
+\begin{align*}
+  f(Q') - f(Q^*) &= \mu(x) \left(\max_{\substack{p\in \mathcal{B}_x}}\log  Q'(p|x) - \max_{\substack{p\in \mathcal{B}_x}}\log  Q^*(p|x) \right) = \mu(x)\left(\log \left(Q^*(p^*|x) + \beta \right) - \log Q^*(p^*|x) \right) > 0  \end{align*}
+ which contradicts the assumption that $Q^*$ maximizes $f(\cdot )$.
+ Therefore $Q^*(p|x) > 0$ for at most one $p\in \mathcal{B}_x$.
+ But we already know that $Q^*(p|x) = 0$ for any $p\not\in \mathcal{B}_x$,
+ so it must be the case that $Q^*(\cdot |x)$
+ places all of its probability mass on exactly one $p\in \mathcal{B}_x$.
+ Call that program $p^*$.
+
+ Because the equivalence classes $\left\{\mathcal{E}_i^x \right\}$
+ form a partition of $\Lambda$
+ we know that $p^*$ is a member of exactly one equivalence class; call it $\mathcal{E}_*^x$.
+ This accomplishes one of our goals:
+ we have already shown that $Q^*(p^*|x) = 1$.
+
+ Let $q\in \mathcal{E}_i^x\not=\mathcal{E}_*^x$.
+ Then because the equivalence classes form a partition we know that
+ $q\not= p^*$
+ and so $Q^*(q|x) = 0$,
+ which is what was to be shown:
+ \emph{any}
+ program not in $\mathcal{E}_*^x$
+ gets no probability mass.
+
+ \end{proof}
+
+ Consider an agent tasked with writing programs built from addition and the constants zero and one.
+A bigram parameterization of $Q$ allows
+it to represent the fact that it should never add zero ($Q_{\code{+},\code{0},0} = Q_{\code{+},\code{0},1} = 0$)
+%or never multiply by  1 ($Q_{\code{*},\code{1},0} = Q_{\code{*},\code{1},1} = 0$), or
+or that addition should always associate to the right
+($Q_{\code{+},\code{+},0} = 0$).
+The $\mathcal{L}^{\text{MAP}}$ training objective encourages
+learning these canonical forms.
+Consider two recognition models, $Q_1$ and $Q_2$,
+and two programs in frontier $\mathcal{F}_x$,
+$p_1=\code{(+ (+ 1 1) 1)}$ and $p_2=\code{(+ 1 (+ 1 1))}$,
+where
+\begin{align*}
+  Q_1(p_1|x) = \frac{\epsilon}{2}&\qquad Q_1(p_2|x) = \frac{\epsilon}{2}\\
+  Q_2(p_1|x) = 0&\qquad  Q_2(p_2|x) = \epsilon
+\end{align*}
+i.e., $Q_2$ breaks a symmetry by forcing right associative addition,
+but $Q_1$ does not, instead splitting its probability mass equally between $p_1$ and $p_2$.
+Now because $\probability[p_1|\mathcal{D},\theta] = \probability[p_2|\mathcal{D},\theta]$
+(Algorithm~\ref{sampleProgram}),
+we have
+\begin{align*}
+  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_1)& = \frac{\probability[p_1|\mathcal{D},\theta]\log \frac{\epsilon}{2} + \probability[p_2|\mathcal{D},\theta]\log \frac{\epsilon}{2}}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = \log \frac{\epsilon}{2}\\
+  \mathcal{L}^{\text{posterior}}_{\text{real}}(Q_2)& = \frac{\probability[p_1|\mathcal{D},\theta]\log 0 + \probability[p_2|\mathcal{D},\theta]\log \epsilon}{\probability[p_1|\mathcal{D},\theta] + \probability[p_2|\mathcal{D},\theta]} = +\infty\\
+  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_1)& = \log Q_1(p_1)           = \log Q_1(p_2)   = \log \frac{\epsilon}{2}\\
+  \mathcal{L}^{\text{MAP}}_{\text{real}}(Q_2)& = \log Q_2(p_2) \phantom{ = \log Q_1(p_2) } = \log \epsilon\\
+\end{align*}
+So $\mathcal{L}^{\text{MAP}}$ prefers $Q_2$ (the symmetry breaking
+recognition model), while $\mathcal{L}^{\text{posterior}}$ reverses
+this preference.
+
+To be clear,
+our recognition model does not learn to break \emph{every} possible symmetry in every possible DSL.
+But in practice we found that a bigrams combined with $\mathcal{L}^{\text{MAP}}$
+works well,
+and we use with this combination throughout the paper.
+
 \subsection{Full set of LOGO  tasks}

### 2018-12-30 archive.tex
more writing

@@
 (Figure~\ref{initialExampleDSL}).  This incremental, modular learning
-of deep hierarchies of DSL components is characteristic of \system
+of deep hierarchies of DSL components occurs
 because of the alternation between code writing and code refactoring.
@@
 
-\subsection{Programs that make plans and take actions}
+\subsection{Programs from visual input}
+We consider three domains where the agent must infer a program from an
+image (Figure~\ref{visualSpecs}).
+\subsubsection{Programs that make plans and take actions}
 
@@
 
-\begin{figure}
+\begin{figure}\centering
   \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
   \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
-  \caption{LOGO graphics tasks. The agent must learn to write programs that draw these 90 pictures.}\label{everyLogo}
+  \includegraphics[width = 0.3\textwidth]{figures/sr16.png}
+  \caption{Three domains where the agent infers a program from visual input.
+\textbf{Left}: 16 (out of 145) LOGO graphics tasks.
+Agent writes a program controlling a `pen' that draws the target picture.
+\textbf{Middle}: 9 (out of 112) tower building tasks.
+Agent writes a program controlling a `hand' that builds the target tower.
+\textbf{Right}: 16 (out of 200) symbolic regression tasks.
+Agent writes a program containing continuous real numbers that fits the points along the curve.}\label{visualSpecs}
 \end{figure}
@@
     \begin{tabular}{c}
-      Repeat and rotate subprogram:\\
-          \includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_18.png}
+      Rotational Symmetry:\\
+          \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
       \end{tabular}
@@
 These towers are built from Lego-style blocks
-that snap together on a discrete grid (e.g., we do not model gravity
-and discretize the coordinates of the blocks).
+that snap together on a discrete grid.%%  (e.g., we do not model gravity
+%% and discretize the coordinates of the 
+%% blocks).
 
@@
 12.5\% of the copy tasks.
-With successive wake/sleep cycles
-we assemble a DSL containing
-parametric option-like subroutines (Figure~\ref{towerPrimitives}),
-allowing the agent to solve 90\% of the tasks.
 \begin{figure}
-  \caption{Learned subroutines for building towers out of Lego-style blocks when trained on tasks in Figure~\ref{towerTasks}. These subroutines act like parametric options~\cite{options},
+  \begin{tabular}{cc}
+    \toprule Example learned DSL components&Samples from DSL\\\midrule
+    \begin{tabular}{rl}
+      Brickwall&\begin{tabular}{l}
+        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
+      \end{tabular}\\
+      Bridge&\begin{tabular}{l}
+        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
+      \end{tabular}\\
+      Staircase&\begin{tabular}{l}
+        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
+        \end{tabular}
+    \end{tabular}&
+    \begin{tabular}{l}
+      \includegraphics[width = 5cm]{figures/tower/dreams.png}
+      \end{tabular}
+    \\\bottomrule 
+    \end{tabular}
+  \caption{Learned subroutines for building towers out of Lego-style blocks. These subroutines act like parametric options~\cite{options},
     giving higher-level building blocks that the agent can use to plan.}
@@
 
-\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
-We apply \system
-to symbolic regression problems.  Here, the
-agent observes points along the curve of a function, and must write a
-program that fits those points.  We initially equip our learner with
-addition, multiplication, and division, and task it with solving
-100 %% symbolic regression
-problems, each either a polynomial or rational function.  The recognition model is a
-convnet that observes an image of the target function's
-graph (Fig.~\ref{functions}) --- visually, different kinds of
-polynomials and rational functions produce different kinds of graphs,
-and so the convnet can  look at a graph and predict
-what kind of function best explains it.
-A key difficulty, however, is
-that these problems are best solved with programs containing real
-numbers.  Our solution to this difficulty is to enumerate
- programs with real-valued parameters, and then fit those
+\subsubsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
+Here, the agent observes points along the curve of a function, and
+must write a program that fits those points.  We initially equip our
+learner with addition, multiplication, and division, and task it with
+solving 200 symbolic regression problems, each either a polynomial or
+rational function.  The recognition model is a convnet that observes
+an image of the target function's graph (Fig.~\ref{visualSpecs},
+rightmost) --- visually, different kinds of polynomials and rational
+functions produce different kinds of graphs, and so the convnet can
+look at a graph and predict what kind of function best explains it.  A
+key difficulty, however, is that these problems are best solved with
+programs containing real numbers.  Our solution to this difficulty is
+to enumerate programs with real-valued parameters, and then fit those
 parameters by automatically differentiating through the programs the
-system writes and use gradient descent to fit the parameters.
-We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
-and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
-\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
-  \begin{minipage}[c]{0.3\columnwidth}%\vspace{-0.3cm}
-  \includegraphics[width = \functionSize]{figures/functions/6.png}
-  \includegraphics[width = \functionSize]{figures/functions/48.png}\\
-  \includegraphics[width = \functionSize]{figures/functions/102.png}
-  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
-  \vspace{1pt}%\includegraphics[width = \functionSize]{figures/functions/181.png}
-  %\includegraphics[width = \functionSize]{figures/functions/160.png}
-  \includegraphics[width = \functionSize]{figures/functions/160.png}
-    \includegraphics[width = \functionSize]{figures/functions/149.png}
-  \end{minipage}
-  \begin{minipage}[c]{0.69\columnwidth}    
-    \caption{Recognition model input for symbolic regression.
-DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}\vspace{-1cm}
-        \end{minipage}
-\end{figure}
-We learn a DSL containing 13 new functions,
-mainly templates for polynomials of different orders or ratios of polynomials.
-The model also learns to find programs minimizing the number of continuous parameters ---
-learning to represent linear functions with 
-\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
-which has five continuous parameters.
-This phenomenon arises from our Bayesian framing:
-both the generative model's bias toward shorter programs,
-and the likelihood model's BIC penalty.
-
-\subsection{Learning from Scratch}
-A long-standing dream within the program induction community
-is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
-and then learning to solve a wide swath of
-induction problems~\cite{solomonoff1964formal,schmidhuber2004optimal,hutter2004universal,solomonoff1989system}.
-All existing systems,
-including ours,
-fall far short of this dream,
-and it is unclear (and we believe unlikely)
-that this dream could ever be fully realized.
-How far can we push in this direction?
-``Learning from scratch'' is subjective, but a reasonable
-starting point is the set of primitives provided in 1959
-Lisp~\cite{mccarthy1960recursive}: these include
-conditionals, recursion, arithmetic, and the 
-list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
-A  basic first goal is to start with
-these primitives,
-and then recover a DSL that
-more closely resembles modern functional languages like Haskell and OCaml.
-Recall (Sec.~\ref{sequences})
-that we initially provided our system with functional programming routines like
-\code{map} and \code{fold}.
-
-We ran the following experiment: \system was given a subset of
-the 1959 Lisp primitives, and tasked with solving 22
-programming exercises. A key difference between
-this setup and our previous experiments is that,
-for this experiment,
-the system is
-given primitive recursion,
-whereas previously we had sequestered recursion within
-higher-order functions like \code{map}, \code{fold}, and \code{unfold}.
-
-After running for 93 hours on 64 CPUs, our
-algorithm solves these 22 exercises, along the way assembling a DSL
-with a modern repertoire of
-functional programming idioms and subroutines, including \code{map},
-\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
-and  arithmetic operations like 
-building lists of natural numbers between an interval (see Figure~\ref{learningFromScratch}).
-We did not use the recognition model for this experiment:
-a bottom-up pattern recognizer is of little use
-for acquiring this abstract knowledge
-from less than two dozen problems.
-
-\begin{figure*}  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{cc}\toprule
-    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule
-
-    \begin{tabular}{l}
-      \code{[1\, 9]}$\to $\code{2}\\
-      \code{[5\, 3\, 8]$\to$3}\\
-    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
-      \code{[true\, false]}$\to $\code{2}\\
-      \code{[false\, false\, false]$\to$3}\\
-    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
-
-\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
-%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
-\code{[9\, 8]$\to$[9\, 8\, 0]}\\
-\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\
-
-\code{[2\, 1\, 4]$\to$[2\, 1]}\\
-%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
-\code{[9\, 8]$\to$[9]}\\
-\blueCode{$f(\ell) = $($f_0$ ($\lambda$ (z) (empty? (cdr z))) car cdr $\ell$)}\\\\
-
-\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
-\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
-\blueCode{$f(\ell) = $($f_2$ + $\ell$ 0)}\\\\
-
-\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
-%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
-\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
-\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
-\code{[4\, 2\, 6\, 4]$\to$[-4\, -2\, -6\, -4]}\\
-%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
-\code{[2\, 3\, 0\, 7]$\to$[-2\, -3\, -0\, -7]}\\
-\blueCode{$f(\ell) = $($f_3$ (- 0) $\ell$)}\\\\
-\code{[4\, 2\, 6\, 4]$\to$[5\, 3\, 7\, 5]}\\
-%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
-\code{[2\, 3\, 0\, 7]$\to$[3\, 4\, 1\, 8]}\\
-\blueCode{$f(\ell) = $($f_3$ (+ 1) $\ell$)}\\\\
-
-      \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
-      \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
-      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
-      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
-      \\\\
-
-      \code{3$\to $[0\, 1\, 2]}\\
-      \code{2$\to $[0\, 1]}\\
-      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 $n$)}\\\\
-
-      \code{3$\to $[0\, 1\, 2\, 3]}\\
-      \code{2$\to $[0\, 1\, 2]}\\
-      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 (+ 1 $n$))}\\\\
-
-      \end{tabular}&
-
-%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
-    \begin{tabular}{l}
-        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
-        \hspace{\helpSize}($f_0$: \emph{unfold})\\
-        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
-        \hspace{\helpSize}($f_1$: \emph{index})\\
-        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
-      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
-        \hspace{\helpSize}($f_2$: \emph{fold})\\
-        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
-        \hspace{\helpSize}($f_3$: \emph{map})\\
-        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
-        \hspace{\helpSize}($f_4$: \emph{length})\\
-        \greenCode{$f_5(\code{f},\code{m},\code{n})$\,=\,($f_0$ (= m) f (+ 1) n)}\\
-        \hspace{\helpSize}($f_5$: \emph{generalization of range})
-
-%        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
-$0 $3)))))))
-
-      \end{tabular}
-    %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
-    %% (lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) ($1 ($3 $0)) ($4 $0) ($5 $0)))))))))
-    %% (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (#(lambda (lambda (fix1 $0 (lambda (lambda (if (empty?
-$0) empty (cons ($3 (car $0)) ($1 (cdr $0))))))))) (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
-$0 $3)))))))
-    \\\bottomrule 
-    \end{tabular}
-  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
-  \end{figure*}
-
-We believe that program learners should \emph{not}
-start from scratch,
-but instead should start from
-a rich, domain-agnostic
-basis like those embodied in the standard libraries of modern  languages.
-What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
-and can in principle recover many of the amenities of modern programming systems,
-provided it is given enough computational power and a suitable
-spectrum of tasks.% programming exercises.
-
-\subsection{Learning Generative Models}
-
-We apply \system to learning generative models for images and text (Fig.~\ref{geomCompiled}-\ref{regularExpressions}).
-For images, we learn programs controlling a simulated ``pen,''
-and the task is to look at an image and explain it in terms of a graphics program.
-For text, we learn probabilistic regular expressions -- a simple probabilistic program for which inference is always tractable -- and the task is to
-infer a regex from a collection of strings.
-
+system writes and use gradient descent to fit the parameters.  We
+define the likelihood model, $\probability[x|p]$, by assuming a
+Gaussian noise model for the input/output examples, and penalize the
+use of real-valued parameters using the
+BIC~\cite{Bishop:2006:PRM:1162264}.
+
+We learn a DSL containing 13 new functions, mainly templates for
+different pieces of polynomials or ratios of polynomials.  The model
+also learns to find programs minimizing the number of continuous
+parameters --- for example, learning to represent linear functions
+with \code{(* real (+ x real))}.  This phenomenon arises from our
+Bayesian framing: both the generative model's bias toward shorter
+programs, and the likelihood model's BIC penalty.
+
+h
 \subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
@@
 \begin{figure}
-  \includegraphics[width = 5cm]{figures/logoLearningCurve.png}
-  \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}
-  \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}
-  \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}
+  \begin{tabular}{ccc}
+    \includegraphics[width = 5cm]{figures/logoLearningCurve.png}&
+    \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
+    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\
+    \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
+    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batched.png}&
+  \end{tabular}
+  \caption{}\label{learningCurves}
 \end{figure}
 
-\begin{figure*}\centering
-  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
-  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
-  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
-\vspace{-0.4cm}  \caption{Learning curves for \system both with (\orange{in orange}) and without
-    (\teal{in teal}) the recognition model.
-Solid lines: \% holdout testing tasks solved w/ 10m timeout.
-Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}\vspace{-0.5cm}
-\end{figure*}
-
 \section{Discussion}
 
-%\paragraph{Outlook.}
-We contribute an algorithm, \systemEnding, that learns to program by
-bootstrapping a DSL with new domain-specific primitives that the algorithm
-itself discovers, together with a neural recognition model that learns to
-efficiently deploy the DSL on new tasks. %% We believe this integration of top-down
-%% symbolic representations and bottom-up neural nets --- both of them learned
-%% --- can help make program induction systems more generally useful for AI. 
-%Many
-%directions remain open.
-%\paragraph{Future.}
-Two immediate future goals are to integrate more sophisticated neural recognition
-models~\cite{devlin2017robustfill} and program
-synthesizers~\cite{solar2008program}, which may improve performance in some
-domains over the generic methods used here.
-Another direction is DSL meta-learning: can we find a
-\emph{single} universal primitive set that could bootstrap DSLs for
-new domains, including the domains considered here,  but also many others?
+\subsection{Learning from Scratch}
+A long-standing dream within the program induction community
+is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
+and then learning to solve a wide swath of
+induction problems~\cite{solomonoff1964formal,schmidhuber2004optimal,hutter2004universal,solomonoff1989system}.
+All existing systems,
+including ours,
+fall far short of this dream,
+and it is unclear (and we believe unlikely)
+that this dream could ever be fully realized.
+How far can we push in this direction?
+``Learning from scratch'' is subjective, but a reasonable
+starting point is the set of primitives provided in 1959
+Lisp~\cite{mccarthy1960recursive}: these include
+conditionals, recursion, arithmetic, and the 
+list operators \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
+A  basic first goal is to start with
+these primitives,
+and then recover a DSL that
+more closely resembles modern functional languages like Haskell and OCaml.
+Recall (Sec.~\ref{sequences})
+that we initially provided our system with functional programming routines like
+\code{map} and \code{fold}.
+
+We ran the following experiment: \system was given a subset of the
+1959 Lisp primitives, and tasked with solving 22 programming
+exercises. A key difference between this setup and our previous
+experiments is that, for this experiment, the system is given
+primitive recursion, whereas previously we had sequestered recursion
+within higher-order functions like \code{map}, \code{fold}, and
+\code{unfold}.
+
+After running for 93 hours on 64 CPUs, our
+algorithm solves these 22 exercises, along the way assembling a DSL
+with a modern repertoire of
+functional programming idioms and subroutines, including \code{map},
+\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
+and  arithmetic operations like 
+building lists of natural numbers between an interval (see  Appendix~\ref{appendixMcCarthy}).
+
+We believe that program learners should \emph{not}
+start from scratch,
+but instead should start from
+a rich, domain-agnostic
+basis like those embodied in the standard libraries of modern  languages.
+What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
+and can in principle recover many of the amenities of modern programming systems,
+provided it is given enough computational power and a suitable
+spectrum of tasks.
+
+\subsection{Program Induction as part of the generic AI toolkit}
+
+Our aim with \system is both to show how learning can enable better
+program synthesis, and to show how better program synthesis can be
+more broadly useful for AI. How could a program induction approach to
+AI become broadly useful across a diverse spectrum of AI problems ---
+planning, natural language understanding, machine vision, causal
+inference, and more?
 
@@
      \EndIf
-     %% \Statex
-     %% \Statex\textbf{where:}
-     %% \State yield$(\tau) = \begin{cases}
-     %%   \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-     %%   \tau   &\text{ otherwise.}
-     %% \end{cases}$ \Comment{Final return type of $\tau$}
-     %% \State  args$(\tau) = \begin{cases}
-     %%   [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
-     %%   []   &\text{ otherwise.}
-     %% \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
    \end{algorithmic}
@@
 
- \textbf{Symmetry breaking analysis.}
-
- \begin{theorem}
+ \noindent \textbf{Symmetry breaking.}  Why does the combination of
+ $\mathcal{L}^{\text{MAP}}$ and the bigram parameterization lead to
+ symmetry breaking?  The reason is twofold: (1) the objective
+ $\mathcal{L}^{\text{MAP}}$ prefers symmetry breaking recognition
+ models; and (2) the bigram parameterization permits certain kinds of
+ symmetry breaking.
+ To sharpen these intuitions,
+ we prove (Theorem~\ref{symmetricproof})
+ that any global optimizer of $\mathcal{L}^{\text{MAP}}$
+ breaks symmetries,
+ and then give a concrete worked out example
+ contrasting the behavior of $\mathcal{L}^{\text{MAP}}$ and $\mathcal{L}^{\text{posterior}}$.
+
+ \pagebreak\begin{theorem}\label{symmetricproof}
    Let $\mu(\cdot )$ be a distribution over tasks and let $Q^*(\cdot |\cdot )$ be a task-conditional distribution over programs satisfying
@@
  form a partition of $\Lambda$
- we know that $p^*$ is a member of exactly one equivalence class; call it $\mathcal{E}_*^x$.
- This accomplishes one of our goals:
- we have already shown that $Q^*(p^*|x) = 1$.
-
- Let $q\in \mathcal{E}_i^x\not=\mathcal{E}_*^x$.
+ we know that $p^*$ is a member of exactly one equivalence class; call it $\mathcal{E}_i^x$.
+ Let $q\in \mathcal{E}_j^x\not=\mathcal{E}_i^x$.
  Then because the equivalence classes form a partition we know that
@@
  and so $Q^*(q|x) = 0$,
- which is what was to be shown:
+ which was our first goal:
  \emph{any}
- program not in $\mathcal{E}_*^x$
+ program not in $\mathcal{E}_i^x$
  gets no probability mass.
 
+ Our second goal --- that there is a member of $\mathcal{E}_i^x$ which gets all the probability mass assigned by $Q^*(\cdot |x)$ --- is immediate from $Q^*(p^*|x) = 1$.
+
+ Our final goal --- that $p^*$ maximizes $\probability[\cdot |x,\mathcal{D},\theta]$
+ --- follows from the fact that $p^*\in \mathcal{B}_x$.
  \end{proof}
 
- Consider an agent tasked with writing programs built from addition and the constants zero and one.
+ Notice that Theorem~\ref{symmetricproof} makes no guarantees as to the
+ cross-task systematicity of the symmetry breaking; for example, an
+ optimal recognition model could associate addition to the right for
+ one task and associate addition to the left on another task. \emph{Systematic} breaking of symmetries
+ must arise only as a consequence as
+ the network architecture (i.e., it is more parsimonious to break symmetries the same way for every task than it is to break them differently for each task).
+
+As a concrete example of symmetry breaking, consider an agent tasked with writing programs built from addition and the constants zero and one.
 A bigram parameterization of $Q$ allows
@@
 
+How would this example work out if we did not have a bigram
+parameterization of $Q$?  With a unigram parameterization, $Q_2$ would
+be impossible to express, because it depends on local context within
+the syntax tree of a program. So even though the objective function would prefer symmetry breaking,
+a simple unigram model lacks the expressive power to
+encode it.
+
 To be clear,
@@
 \end{figure}
+
+\subsection{Learning from Scratch: Tasks and DSL}\label{appendixMcCarthy}
+
+We gave our system the following primitives: \code{if}, \code{=},
+\code{>}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
+\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}, all of which
+are present in some form in McCarthy's 1959
+Lisp~\cite{mccarthy1960recursive}.\footnote{McCarthy's first version
+  of Lisp used \code{cond} instead of \code{if}. Because we are using
+  a typed language, we instead used \code{if}, because Lisp-style
+  \code{cond} is unwieldy to express as a function in typed
+  languages.}  We furthermore allowed functions to call themselves,
+which we modeled using the Y combinator.  We did not use the
+recognition model for this experiment: a bottom-up pattern recognizer
+is of little use for acquiring this abstract knowledge from less than
+two dozen problems.
+
+Figure~\ref{learningFromScratch} shows the full set of tasks and the learned DSL.
+\begin{figure*}  \newcommand{\helpSize}{0.25cm}
+  \begin{tabular}{cc}\toprule
+    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule
+
+    \begin{tabular}{l}
+      \code{[1\, 9]}$\to $\code{2}\\
+      \code{[5\, 3\, 8]$\to$3}\\
+    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
+      \code{[true\, false]}$\to $\code{2}\\
+      \code{[false\, false\, false]$\to$3}\\
+    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
+
+\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
+%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
+\code{[9\, 8]$\to$[9\, 8\, 0]}\\
+\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\
+
+\code{[2\, 1\, 4]$\to$[2\, 1]}\\
+%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
+\code{[9\, 8]$\to$[9]}\\
+\blueCode{$f(\ell) = $($f_0$ ($\lambda$ (z) (empty? (cdr z))) car cdr $\ell$)}\\\\
+
+\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
+\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
+\blueCode{$f(\ell) = $($f_2$ + $\ell$ 0)}\\\\
+
+\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
+%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
+\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
+\code{[4\, 2\, 6\, 4]$\to$[-4\, -2\, -6\, -4]}\\
+%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+\code{[2\, 3\, 0\, 7]$\to$[-2\, -3\, -0\, -7]}\\
+\blueCode{$f(\ell) = $($f_3$ (- 0) $\ell$)}\\\\
+\code{[4\, 2\, 6\, 4]$\to$[5\, 3\, 7\, 5]}\\
+%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
+\code{[2\, 3\, 0\, 7]$\to$[3\, 4\, 1\, 8]}\\
+\blueCode{$f(\ell) = $($f_3$ (+ 1) $\ell$)}\\\\
+
+      \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
+      \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
+      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
+      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
+      \\\\
+
+      \code{3$\to $[0\, 1\, 2]}\\
+      \code{2$\to $[0\, 1]}\\
+      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 $n$)}\\\\
+
+      \code{3$\to $[0\, 1\, 2\, 3]}\\
+      \code{2$\to $[0\, 1\, 2]}\\
+      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 (+ 1 $n$))}\\\\
+
+      \end{tabular}&
+
+%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+    \begin{tabular}{l}
+        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
+      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
+        \hspace{\helpSize}($f_0$: \emph{unfold})\\
+        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
+      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
+        \hspace{\helpSize}($f_1$: \emph{index})\\
+        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
+      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
+        \hspace{\helpSize}($f_2$: \emph{fold})\\
+        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
+        \hspace{\helpSize}($f_3$: \emph{map})\\
+        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
+        \hspace{\helpSize}($f_4$: \emph{length})\\
+        \greenCode{$f_5(\code{f},\code{m},\code{n})$\,=\,($f_0$ (= m) f (+ 1) n)}\\
+        \hspace{\helpSize}($f_5$: \emph{generalization of range})
+
+%        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
+$0 $3)))))))
+
+      \end{tabular}
+    %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
+    %% (lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) ($1 ($3 $0)) ($4 $0) ($5 $0)))))))))
+    %% (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (#(lambda (lambda (fix1 $0 (lambda (lambda (if (empty?
+$0) empty (cons ($3 (car $0)) ($1 (cdr $0))))))))) (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
+$0 $3)))))))
+    \\\bottomrule 
+    \end{tabular}
+  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
+  \end{figure*}
 

### 2019-01-01 archive.tex
new learning curves

@@
   \includegraphics[width = \textwidth]{figures/DeepLogo.pdf}
+  \caption{\textbf{Bottom}: Learned graphics programing DSL.
+Each circle encloses a different learned component; arrows point from a component to other components that use it.
+\textbf{Top}: 3 tasks, and below them the programs that solve them.
+Yellow/Blue/Green highlighting indicates which DSL components are used to solve which tasks.}
   \end{figure}
@@
   \begin{tabular}{ccc}
-    \includegraphics[width = 5cm]{figures/logoLearningCurve.png}&
+    \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
     \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
-    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\
+    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
     \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
-    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batched.png}&
+    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
   \end{tabular}

### 2019-01-02 archive.tex
improvements to experiment section

@@
 \begin{figure}
-\centering  \begin{tabular}{ccc}
+  \begin{minipage}[c]{0.3\textwidth}
+    \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
+    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
+    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
+    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
+    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
+    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
+    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};
+
+    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
+    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
+    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
+    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
+    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
+\end{tikzpicture}
+    \end{minipage}\hfill%
+  \begin{tabular}{cll}
      \toprule
-&     Unigram&Bigram \\\midrule
+&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
+     $\mathcal{L}^{\text{posterior}}$&
+     \begin{tabular}{l}
+       \emph{Three samples:}\\
+       \code{(+ 1 0)}\\
+\code{(+ (+ 0 0)}\\\code{\phantom{(+ }(+ 1 0))}\\
+\code{(+ 1 1)}
+\\
+63.0\% right-associative\\37.4\% \code{+0}'s
+       \end{tabular}
+     &
+     \begin{tabular}{l}
+       \emph{Three samples:}\\
+       \code{0}\\
+\code{(+ (+ (+ 0 0)}\\\code{\phantom{(+ (+} (+ 0 1)) 1)}\\
+\code{1}\\
+55.8\% right-associative\\31.9\% \code{+0}'s
+       \end{tabular}
+     \\\\
+     $\mathcal{L}^{\text{MAP}}$&
+     \begin{tabular}{l}
+       \emph{Three samples:}\\
+       \code{1}\\
+       \code{(+ 1 (+ 1 (+ (+ 1 }\\\code{\phantom{(+ }(+ 1 1)) 1)))}\\
+       \code{(+ (+ 1 1) 1)}
+\\48.6\% right-associative\\
+       0.5\% \code{+0}'s       
+       \end{tabular}
+     &
+     \begin{tabular}{l}
+       \emph{Three Samples:}\\
+\code{(+ 1 (+ 1 (+ 1}\\\code{\phantom{(+ 1} (+ 1 (+ 1 1)))))}\\
+\code{0}\\
+\code{(+ 1 (+ 1 (+ 1 1)))}
+       \\
+       \textbf{97.9\% right-associative}\\
+       2.5\% \code{+0}'s       
+       \end{tabular}
+\\    \bottomrule 
+  \end{tabular}
+  \caption{\textbf{Left:} Bigram parameterization of distribution over programs predicted by recognition model.
+  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
+Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task.
+\textbf{Right:} Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero.
+\% right-associative calculated by drawing 500 samples from $Q$.
+$\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition.
+Tasked with building programs from \code{+}, \code{1}, and \code{0}.
+}\label{symmetry}
+\end{figure}
+
+\begin{figure}
+\centering  \begin{tabular}{cll}
+     \toprule
+&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
      $\mathcal{L}^{\text{posterior}}$&
@@
 %% blocks).
-
-The system starts out with four domain-specific primitives:
-\code{left} and \code{right}, which take an integer $n$ as input and
-move the hand left/right $n$ Lego grid spaces; along with
-\code{horizontal} and \code{vertical}, which put down
-horizontal/vertical Lego blocks.  As with LOGO graphics we also
-include our two standard planning primitives (\code{for} and
-\code{get/set}).
-Starting with this basis
-the system can only solve
-12.5\% of the copy tasks.
+The system starts with the same control flow primitives as with
+LOGO graphics,
+and learns parametric `options' for building blocks towers (Figure~\ref{tower}),
+inferring concepts like arches, staircases, bridges and brick walls.
 \begin{figure}
@@
     \end{tabular}
-  \caption{Learned subroutines for building towers out of Lego-style blocks. These subroutines act like parametric options~\cite{options},
-    giving higher-level building blocks that the agent can use to plan.}
+  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{options},
+    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', from learned DSL.}\label{tower}
 \end{figure}
 
-\subsubsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
+\subsubsection{Symbolic Regression}\label{regressionSection}
 Here, the agent observes points along the curve of a function, and
@@
 
-h
 \subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}

### 2019-01-02 archive.tex
improvements to dreaming part of big figure

@@
     \begin{scope}[shift={(9.5,-4.5)}]
-      \node(dreaming) at (2,-1) {\underline{Fantasies}};
-      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
-      \node at ([xshift=1.5cm]d.east) (p1){prog.};
-      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
-      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
+      \node(dreaming) at (1,-1) {\underline{Fantasies}};
+      \node[anchor=center] at ([yshift=-0.5cm]dreaming.south) (d){\textbf{DSL}};
+      \node at ([yshift=-1.75cm]d.south) (p1){program};
+      \draw[squiggle,-> ] (d.south) -- node[sloped,above]{\small sample} (p1.north);
+
+      \node(replay) at ([xshift=2cm]dreaming.east) {\underline{Replays}};
+      \node[anchor=center,align=center] at ([yshift=-0.5cm]replay.south) (d){\textbf{progs. for task}};
+      \node at ([yshift=-1.75cm]d.south) (p1){program};
+      \draw[squiggle,-> ] (d.south) -- node[sloped,above]{\small sample} (p1.north);
+
+      \node(p1) at (1.5,-6) {program};      
+      \node at ([xshift = 2.0cm]p1.east) (t1){ task};
       \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
@@
 
-      \node(replay) at (2,-5) {\underline{Replays}};
-      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
-      \node at ([xshift=1.5cm]d.east) (p1){prog.};
-      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
-      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
-      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
-      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
-        \NeuralNetwork{0.17}};
-      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
-      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
+      \node at ($(-0.25,0.5) + (p1.north)!0.5!(t1.north)$) {\underline{Train recognition model}};
+
+      %% \node at ([xshift = 1.5cm]p1.east) (t1){ task};
+      %% \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
+      %% \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
+      %%   \NeuralNetwork{0.17}};
+      %% \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
+      %% \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);
 

### 2019-01-02 archive.tex
example DSL 's

@@
       \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
-      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
+      \node(n) at ([yshift=-1.2cm,xshift=1.25cm]p1.south) {
         \NeuralNetwork{0.17}};
@@
 
+\subsection{Learned DSLs}
+Here we present representative DSLs learned by our model. DSL primitives
+discovered by the algorithm are prefixed with \lstinline!#!.
+Variables are prefixed with \lstinline!$!, and we adopt De Bruijn indices to
+model bound variables~\cite{pierce}.
+
+\lstset{
+  basicstyle=\footnotesize,
+  escapechar=@,
+  breaklines=true,
+  breakatwhitespace=true,
+  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
+}
+
+\subsubsection{List processing}
+\begin{lstlisting}
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range ($0 (+ 1))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0))))))
+#(+ 1 (+ 1 1))
+#(@$\lambda$@ (@$\lambda$@ (fold $1 (cons $0 empty) (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(+ 1 #(+ 1 (+ 1 1)))
+#(@$\lambda$@ (map (@$\lambda$@ (if ($1 $0) (+ $0 1) 0))))
+#(@$\lambda$@ (cdr (cdr $0)))
+#(@$\lambda$@ (@$\lambda$@ (fold (#(@$\lambda$@ (cdr (cdr $0))) (#(@$\lambda$@ (cdr (cdr $0))) $1)) $0 (@$\lambda$@ (@$\lambda$@ (cdr (#(@$\lambda$@ (@$\lambda$@ (fold $1 (cons $0 empty) (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 (car $0))))))))
+#(@$\lambda$@ (car (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $0 (@$\lambda$@ (empty?
+(#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (gt?
+$0 $1))))))))
+#(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (eq? $0 $1))))))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (is-prime (+ $1 (mod $0 #(+ 1 #(+ 1 (+ 1 1))))))))))
+#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (map (@$\lambda$@ (mod $0 $1))))
+#(@$\lambda$@ (map (@$\lambda$@ (eq? (mod $0 $1) 0))))
+#(@$\lambda$@ (gt? (mod $0 #(+ 1 (+ 1 1))) 0))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (car (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $0 (@$\lambda$@ (empty?
+(#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (gt?
+$0 $1)))))))) (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (gt?
+$1 (length (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $2 (@$\lambda$@ (gt?
+$1 $0))))))))))
+\end{lstlisting}
+
+\subsubsection{Text editing}
+\begin{lstlisting}
+#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1))))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0))))
+#(+ 1)
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq? $2 $1) 0 $0))))))))
+#(@$\lambda$@ (@$\lambda$@ (cons (car $0) (#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1)))) (cons '.' empty) (car $1)))))
+#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (if (char-eq? $2 $0) $1 $0)))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons LPAREN $0) (cons RPAREN empty)))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $0 (length (cdr (cdr $1))))))
+#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+$2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0)))))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $0 (length (cdr (cdr $1)))))) (cdr (cdr $0))))
+#(#(+ 1) 1)
+#(@$\lambda$@ (@$\lambda$@ (cons (car $1) (cons $0 empty))))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+$2 $1) 0 $0)))))))) (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+$2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0))))))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons LPAREN $0) (cons RPAREN empty))) $1) $0) $0)))
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (cons $0 $2)))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 STRING))
+#(#(+ 1) #(#(+ 1) 1))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+$2 $1) 0 $0)))))))) (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) STRING $0) SPACE))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (cons $0 $2))))) (cons $0 $1))))
+#(#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) STRING)
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (cons (car $0) (#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1)))) (cons '.' empty) (car $1))))) (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+$2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0))))))) $1 $0) $1)))
+#(#(+ 1) #(#(+ 1) #(#(+ 1) 1)))
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+$2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0))))))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 STRING)) (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $2 #(#(+ 1) #(#(+ 1) #(#(+ 1) 1))))) $0)))))
+\end{lstlisting}
+
+\subsubsection{Graphics}
+\begin{lstlisting}
+  #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0)))))))
+#(logo_DIVA logo_UA 4)
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY)
+#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0)))
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_FWRT (logo_MULL $2 $1) $4 $0)))))))
+#(@$\lambda$@ (logo_forLoop 7 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 7 logo_epsA $2 $0)))))
+#(@$\lambda$@ (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_SUBA logo_UA logo_epsA) (logo_MULL logo_epsL $0)))
+#(@$\lambda$@ (logo_forLoop logo_IFTY (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 logo_epsA (logo_FWRT logo_epsL logo_epsA $0))))))
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 4 #(logo_DIVA logo_UA 4))
+#(logo_DIVA logo_UA)
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_GETSET $2 (logo_FWRT logo_ZL $4 $0))))))))
+#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_DIVA logo_epsA 2) logo_epsL)
+#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) logo_epsA logo_epsL)
+#(logo_forLoop 3 (@$\lambda$@ (@$\lambda$@ (#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_DIVA logo_epsA 2) logo_epsL) (logo_FWRT logo_ZL #(logo_DIVA logo_UA 4) $0)))))
+#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_GETSET $2 (logo_FWRT logo_ZL $4 $0)))))))) (#(logo_DIVA logo_UA) $0) $0))
+#(@$\lambda$@ (#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0))) (#(@$\lambda$@ (logo_forLoop logo_IFTY (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 logo_epsA (logo_FWRT logo_epsL logo_epsA $0)))))) logo_epsL $0)))
+#(@$\lambda$@ (#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0))) (logo_FWRT logo_UL #(logo_DIVA logo_UA 4) $0)))
+#(@$\lambda$@ (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) logo_epsA (logo_MULL logo_epsL $0)))
+#(logo_FWRT logo_UL logo_UA)
+#(@$\lambda$@ (#(@$\lambda$@ (logo_forLoop 7 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 7 logo_epsA $2 $0))))) (logo_MULL logo_epsL $0)))
+\end{lstlisting}
+
+\subsubsection{Towers}
+\begin{lstlisting}
+  #(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0))))))
+#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0))))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 4 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 8 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (left 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))))))))
+#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0)))))))))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 8 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (left 2 (3x1 $0))))))))
+#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0))))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (1x3 $0)))))
+#(tower_loopM 4 (@$\lambda$@ (@$\lambda$@ (right 6 (3x1 $0)))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0)))))) $0)))))
+#(@$\lambda$@ (right 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))))
+#(@$\lambda$@ (right 2 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))
+#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0)))))) $0))))) $0 (right 2 (#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 4 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))))) $0 $1)))))
+#(tower_loopM 5 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (right 2 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))) $0))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (tower_embed (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 3 6 3 $0)) $0)))))
+#(@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (tower_embed (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 4 5 $4 $0)) $0))))))
+#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 4 5)
+#(@$\lambda$@ (@$\lambda$@ (right 6 (tower_embed $0 $1))))
+#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (right 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))) $0)))))
+\end{lstlisting}
+\subsubsection{Symbolic regression}
+\begin{lstlisting}
+  #(@$\lambda$@ (+. $0 REAL))
+#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL)))
+#(/. REAL)
+#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0)))
+#(@$\lambda$@ (/. (#(/. REAL) $0) $0))
+#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+.
+$0 REAL)) (*.
+$0 REAL))) (*.
+$0 (#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+.
+$0 REAL)) (*.
+$0 REAL))) (*.
+(#(@$\lambda$@ (+.
+$0 REAL)) $0) $0))) $0))))
+#(@$\lambda$@ (#(/. REAL) (#(@$\lambda$@ (+. $0 REAL)) $0)))
+#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. (#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0))) $0) $0)))
+\end{lstlisting}
 \end{document}

### 2019-01-02 archive.tex
more writing

@@
 recursive functions, and the ability to create new functions.
+We `batch' the tasks by randomly shuffling the training set
+and playing small minibatches
+of tasks to the agent during each wake cycle.
 
@@
 \section{Appendix}
+
+\subsection{\system pseudocode}\label{systemPseudocode}
+
+Algorithm~\ref{theWholeThing} specifies how we integrate wake and sleep cycles.
+\begin{algorithm}
+  \caption{Full \system algorithm}\label{theWholeThing}
+  \begin{algorithmic}[1]
+    \State \textbf{function} \systemEnding$(\mathcal{D}, X)$:
+    \State {\bfseries Input:} Initial DSL $\mathcal{D}$, tasks $X$
+    \State \textbf{Output:} Infinite stream of DSLs, recognition models, and frontiers
+    \State \textbf{Hyperparameters:} Batch size $B$, enumeration timeout $T$, maximum frontier size $F$
+    \State $\theta\gets $uniform distribution
+    \State $\mathcal{F}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize frontiers to be empty}
+    \While{true}\Comment{Loop over epochs}
+    \State shuffle $\gets$ random permutation of $X$\Comment{Randomize minibatches}
+    \While{shuffle is not empty}\Comment{Loop over minibatches}
+    \State batch $\gets$ first $B$ elements of shuffle\Comment{Next minibatch of tasks}
+    \State shuffle $\gets$ shuffle with first $B$ elements removed
+    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\mathcal{F}_x\cup \left\{p\;|\;p\in \text{enumerate}(\probability[\cdot |\mathcal{D},\theta],T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
+    \State Train $Q(\cdot |\cdot )$ to minimize $\mathcal{L}^{\text{MAP}}$ across all $\left\{\mathcal{F}_x \right\}_{x\in X}$\Comment{Dream Sleep}
+    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\mathcal{F}_x\cup \left\{p\;|\;p\in \text{enumerate}(Q(\cdot |x),T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
+    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\text{ top }F\text{ elements of }\mathcal{F}_x\text{ as measured by }\probability[\cdot |x,\mathcal{D},\theta]$\Comment{Keep top $F$ programs}
+    \State $\mathcal{D},\theta,\left\{\mathcal{F}_x \right\}_{x\in X}\gets $\textsc{Consolidate}$(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\}_{x\in X})$\Comment{Consolidation Sleep}
+    \State \textbf{yield} $(\mathcal{D},\theta)$, $Q$, $\left\{\mathcal{F}_x \right\}_{x\in X}$ \Comment{Yield the updated DSL, recognition model, and solutions found to tasks}
+    \EndWhile
+    \EndWhile
+
+  \end{algorithmic}
+\end{algorithm}
 

### 2019-01-07 archive.tex
hyper parameters and training details

@@
     \renewcommand{\baselinestretch}{0.5}
-    \renewcommand{\inventionColor}{orange}
+    \newcommand{\inventionColor}{orange}
     \newcommand{\name}[2]{
@@
 
-    \renewcommand{\horizontalSpacing}{3.8cm}
-    \renewcommand{\verticalSpacing}{2cm}
+    \newcommand{\horizontalSpacing}{3.8cm}
+    \newcommand{\verticalSpacing}{2cm}
 
@@
     \renewcommand{\baselinestretch}{0.5}
-    \renewcommand{\inventionColor}{orange}
+    \newcommand{\inventionColor}{orange}
 
@@
     \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
-    \renewcommand{\horizontalSpacing}{4.2cm}
-    \renewcommand{\verticalSpacing}{2cm}
+    \newcommand{\horizontalSpacing}{4.2cm}
+    \newcommand{\verticalSpacing}{2cm}
 
@@
     }
-    \renewcommand{\task}[4]{
+    \newcommand{\task}[4]{
       \node[align=left,anchor=center](t#1) at (#2) {#3};
@@
     }
-    \renewcommand{\nextTask}[4]{
+    \newcommand{\nextTask}[4]{
       \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
@@
       %% defined routines for creating fragmented syntax trees
-      \renewcommand{\syntaxOne}[1]{
+      \newcommand{\syntaxOne}[1]{
         \begin{tikzpicture}[scale=#1,line width=0.35mm]          
@@
 
-\subsection{Consolidation-Sleep: Growing a Domain Specific Language}
+\subsection{Consolidation-Sleep: Growing a Domain Specific Language}\label{consolidationSection}
 
@@
 The second approach is to have $Q$ predict a fixed-dimensional weight
-vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ec2}
+vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ecc}
 or sampler~\cite{menon2013machine}.  This approach can enjoy strong
@@
 take the form of (parent, child, argument index).
-Figure~\ref{bg} diagrams this generative process and Algorithm~\ref{recognitionSample}
+Figure~\ref{symmetry} (left) diagrams this generative process and Algorithm~\ref{recognitionSample}
 specifies a sampling procedure for $Q(\cdot |x)$.
@@
 described next.
-\begin{figure}
-\centering  \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
-    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
-    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
-    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
-    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
-    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
-    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};
-
-    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
-    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
-    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
-    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
-    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
-\end{tikzpicture}
-\caption{Parameterization of distribution over programs predicted by recognition model.
-  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
-Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task.}\label{bg}
-\end{figure}
 
@@
   \begin{minipage}[c]{0.3\textwidth}
-    \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
+    \begin{tikzpicture}[scale=1]%[every node/.style={inner sep=0,outer sep=0}]
     \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
@@
     \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
-    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
-    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
-    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
-    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
+    \draw[->] (k.south)--node[sloped, above]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
+    \draw[->] (k.south)--node[sloped, above]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
+    \draw[->] (m.south)--node[sloped, above]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
+    \draw[->] (m.south)--node[sloped, above]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
 \end{tikzpicture}
@@
 &     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
-     $\mathcal{L}^{\text{posterior}}$&
+     \rotatebox[origin=c]{90}{$\mathcal{L}^{\text{posterior}}$}&
      \begin{tabular}{l}
@@
      \\\\
-     $\mathcal{L}^{\text{MAP}}$&
+     \rotatebox[origin=c]{90}{$\mathcal{L}^{\text{MAP}}$}&
      \begin{tabular}{l}
@@
 \subsubsection{List Processing}\label{listSection}
-We took 218 list manipulation tasks from our previous work~\cite{ec2},
+We took 218 list manipulation tasks from our previous work~\cite{ecc},
 each with 15 input/output examples.  In solving these tasks, the
@@
     \end{tabular}
-  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{options},
+  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{stolle2002learning},
     giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', from learned DSL.}\label{tower}
@@
   \end{tabular}
-  \caption{}\label{learningCurves}
+  \caption{Test set performance across wake/sleep iterations.
+An Epoch is a full pass over the set of training tasks.
+Orange: \systemEnding.
+Teal: No recognition model.
+Purple: No DSL learning.
+Solid lines: \% tasks solved within 10 minute timeout.
+Dashed lines: average search time.
+Shaded region is $\pm 1$ stddev search time.}\label{learningCurves}
 \end{figure}
@@
 program synthesis, and to show how better program synthesis can be
-more broadly useful for AI. How could a program induction approach to
+more broadly useful for AI.
+Many challenges remain.
+
+How could a program induction approach to
 AI become broadly useful across a diverse spectrum of AI problems ---
@@
     \Else\Comment{Build an application to give something w/ type $\tau$}
-    \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+    \State $\text{primitives} \gets \{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$
     $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
@@
 which adds $k$ to all of the free variables in a $\lambda$-expression or version space,
-and we have defined a new operator, $\downshift$, whose purpose is to
-undo the action of $\shift$.
+and we have defined a new operator, $\downshift{}$, whose purpose is to
+undo the action of $\shift{}$.
 We have written definitions recursively,
@@
      \Else\Comment{Build an application to give something w/ type $\tau$}
-     \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
+     \State $\text{primitives} \gets \{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$
      $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
@@
 Figure~\ref{learningFromScratch} shows the full set of tasks and the learned DSL.
-\begin{figure*}  \newcommand{\helpSize}{0.25cm}
-  \begin{tabular}{cc}\toprule
-    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule
-
+\begin{figure}  \newcommand{\helpSize}{0.25cm}
+  \begin{tabular}{ll}\toprule
+    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&%&\normalsize \popp{DSL}\\\midrule
+
+    \begin{tabular}{ll}      
       \begin{tabular}{l}
@@
         \code{[5\, 3\, 8]$\to$3}\\
-    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
-      \code{[true\, false]}$\to $\code{2}\\
-      \code{[false\, false\, false]$\to$3}\\
-    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
-
+        \blueCode{$f(\ell) = $($f_5$ $\ell$)}
+      \end{tabular}&
+      \begin{tabular}{l}
+        \code{[[2 1]\, []]}$\to $\code{[1 0]}\\
+        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
+        \blueCode{$f(\ell) = $($f_2$ $f_4$ $\ell$)}
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
+        \code{[0 1 1 0 0]$\to$[1 1]}\\
+        \code{[9 0 8]$\to$[9\, 8]}\\
+        \blueCode{$f(\ell) = $($f_3$ (eq? 0) $\ell$)}        
+      \end{tabular}&
+      \begin{tabular}{l}
+        \code{[1 -1 0 2]$\to$[1 2]}\\
+        \code{[9 -5 5 0 8]$\to$[9 5 8]}\\
+        \blueCode{$f(\ell) = $($f_3$ (gt? 1) $\ell$)}        
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
         \code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
-%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
         \code{[9\, 8]$\to$[9\, 8\, 0]}\\
-\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\
-
+        \blueCode{$f(\ell) = $($f_0$ cons $\ell$ (cons 0 nil))}
+      \end{tabular}&
+
+      \begin{tabular}{l}
         \code{[2\, 1\, 4]$\to$[2\, 1]}\\
-%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
         \code{[9\, 8]$\to$[9]}\\
-\blueCode{$f(\ell) = $($f_0$ ($\lambda$ (z) (empty? (cdr z))) car cdr $\ell$)}\\\\
-
+        \blueCode{$f(\ell) = $($f_6$ cdr ($\lambda$ (z) (empty? (cdr z))) $\ell$)}
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
         \code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
         \code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
-\blueCode{$f(\ell) = $($f_2$ + $\ell$ 0)}\\\\
-
+        \blueCode{$f(\ell) = $($f_0$ + $\ell$ 0)}
+      \end{tabular}&
+
+      \begin{tabular}{l}
         \code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
-%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
         \code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
-\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
+        \blueCode{$f(\ell) = $($f_2$ ($\lambda$ (x) (+ x x)) $\ell$)}
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
         \code{[4\, 2\, 6\, 4]$\to$[-4\, -2\, -6\, -4]}\\
-%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
         \code{[2\, 3\, 0\, 7]$\to$[-2\, -3\, -0\, -7]}\\
-\blueCode{$f(\ell) = $($f_3$ (- 0) $\ell$)}\\\\
+        \blueCode{$f(\ell) = $($f_2$ (- 0) $\ell$)}
+      \end{tabular}&
+      \begin{tabular}{l}
         \code{[4\, 2\, 6\, 4]$\to$[5\, 3\, 7\, 5]}\\
-%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
         \code{[2\, 3\, 0\, 7]$\to$[3\, 4\, 1\, 8]}\\
-\blueCode{$f(\ell) = $($f_3$ (+ 1) $\ell$)}\\\\
-
+        \blueCode{$f(\ell) = $($f_2$ (+ 1) $\ell$)}
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
         \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
         \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
-      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
-      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
-      \\\\
-
+        \blueCode{$f(\ell) = $($f_6$ ($\lambda$ (l) (cdr (cdr l))) empty? $\ell$}
+      \end{tabular}&
+
+      \begin{tabular}{l}
         \code{3$\to $[0\, 1\, 2]}\\
         \code{2$\to $[0\, 1]}\\
-      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 $n$)}\\\\
-
+        \blueCode{$f(n) = $($f_9$ $n$)}
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
         \code{3$\to $[0\, 1\, 2\, 3]}\\
         \code{2$\to $[0\, 1\, 2]}\\
-      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 (+ 1 $n$))}\\\\
-
+        \blueCode{$f(n) = $($f_9$ (+ 1 $n$))}
       \end{tabular}&
-
-%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
-    \begin{tabular}{l}
-        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
-        \hspace{\helpSize}($f_0$: \emph{unfold})\\
-        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
-      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
-        \hspace{\helpSize}($f_1$: \emph{index})\\
-        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
-      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
-        \hspace{\helpSize}($f_2$: \emph{fold})\\
-        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
-        \hspace{\helpSize}($f_3$: \emph{map})\\
-        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
-        \hspace{\helpSize}($f_4$: \emph{length})\\
-        \greenCode{$f_5(\code{f},\code{m},\code{n})$\,=\,($f_0$ (= m) f (+ 1) n)}\\
-        \hspace{\helpSize}($f_5$: \emph{generalization of range})
-
-%        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
+      \begin{tabular}{l}
+        \code{[9 2]$\to $[9 9 2 2]}\\
+        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
+        \blueCode{$f(l) = $($f_0$ ($\lambda$ (a x) (cons x (cons x a))) l nil)}
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
+        \code{0, [9 2 3]$\to $9}\\
+        \code{3, [0 2 8 4 5 6]$\to $4}\\
+        \blueCode{$f($n$,$l$) = $($f_{10}$ l n)}
+      \end{tabular}&
+      \begin{tabular}{l}
+        \code{1, [9 2 3]$\to $9}\\
+        \code{4, [0 2 8 1 5 6]$\to $1}\\
+        \blueCode{$f($n$,$l$) = $($f_{10}$ l (+ 1 n))}
+      \end{tabular}\\\\
+      \begin{tabular}{l}
+        \code{3$\to $[-3 -2 -1]}\\
+        \code{4$\to $[-4 -3 -2 -1]}\\
+        \blueCode{$f($n$) = $($f_{8}$ 0 n)}
+      \end{tabular}&
+      \begin{tabular}{l}
+        \code{2$\to $[2 1]}\\
+        \code{4$\to $[4 3 2 1]}\\
+        \blueCode{$f($n$) = $($f_{8}$ 0 (- 0 n))}
+      \end{tabular}      
+    \end{tabular}
+    \\\midrule 
+    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+    \begin{tabular}{ll}
+      \begin{tabular}{l}
+        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
+        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
+          \hspace{\helpSize}($f_0$: \emph{fold})
+      \end{tabular}&
+      \begin{tabular}{l}
+        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
+        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
+          \hspace{\helpSize}($f_1$: \emph{unfold})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
+        \hspace{\helpSize}($f_2$: \emph{map})
+      \end{tabular}&
+      \begin{tabular}{l}
+                  \greenCode{$f_3($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (if (f x) a (cons x a))))}\\
+          \hspace{\helpSize}($f_3$: \emph{filter})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_4($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
+        \hspace{\helpSize}($f_4$: \emph{generalization of range})
+      \end{tabular}&
+      \begin{tabular}{l}
+        \greenCode{$f_5($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
+        \hspace{\helpSize}($f_5$: \emph{length})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_6($n$,$p$,$l$)\,=\,$($f_1$ p n car l)}\\
+        \hspace{\helpSize}($f_6$: \emph{specialization of unfold})
+      \end{tabular}&
+      \begin{tabular}{l}
+        \greenCode{$f_7($p$)\,=\,$($f_4$ ($\lambda$ (x) x) p 0)}\\
+        \hspace{\helpSize}($f_7$: \emph{another generalization of range})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_8($n$,$m$)\,=\,$($f_4$ ($\lambda$ (x) (- n x)) (eq? 0) m)}\\
+        \hspace{\helpSize}($f_7$: \emph{count downwards})
+      \end{tabular}&
+      \begin{tabular}{l}
+        \greenCode{$f_9($n$)\,=\,$($f_7$ (eq? n))}\\
+        \hspace{\helpSize}($f_9$: \emph{range})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_{10}($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_9$ n) l))}\\
+        \hspace{\helpSize}($f_{10}$: \emph{index})
+      \end{tabular}&
+
+    \end{tabular}
+%%     \begin{tabular}{l}
+%%         \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
+%%       \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
+%%         \hspace{\helpSize}($f_1$: \emph{index})\\
+
+%%         \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
+%%         \hspace{\helpSize}($f_4$: \emph{length})\\
+%%         \greenCode{$f_5(\code{f},\code{m},\code{n})$\,=\,($f_0$ (= m) f (+ 1) n)}\\
+%%         \hspace{\helpSize}($f_5$: \emph{generalization of range})
+
+%% %        -0.222777	int -> int -> list(int)	#(lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (map (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq?
 $0 $3)))))))
 
-      \end{tabular}
+%%       \end{tabular}
     %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
@@
     \end{tabular}
-  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
-  \end{figure*}
+  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.
+Complete set of tasks and learned DSL shown.
+Learned DSL components are numbered in the order that they are learned.}\label{learningFromScratch}
+  \end{figure}
 
@@
 \end{lstlisting}
+
+\subsection{Hyperparameters and training details}
+
+\noindent\textbf{Neural net architecture} The recognition model for domains with sequential structure (list processing, text editing, regular expressions) is a
+recurrent neural network.
+We use a bidirectional GRU~\cite{cho2014learning} with 64 hidden units that reads each input/output pair; we concatenate the input and output along with a special delimiter
+symbol between them.
+We use a 64-dimensional vectors to embed symbols in the input/output.
+We MaxPool the final hidden unit activations in the GRU along both passes of the bidirectional GRU.
+
+The recognition model for domains with 2D visual structure (LOGO
+graphics, tower building, and symbolic regression) is a convolutional
+neural network.
+We take our convolutional architecture from~\cite{snell2017prototypical}.
+
+We follow the RNN/CNN by an MLP with 128 hidden units and a ReLU activation which then outputs the $Q_{ijk}$ matrix described in~\ref{recognitionAppendix}.
+
+\noindent\textbf{Neural net training} We train our recognition models using Adam~\cite{kingma2014adam} with a learning rate of 0.001.
+
+\noindent\textbf{\system Hyperparameters}
+Due to the high computational cost we
+performed only an informal coarse
+hyperparameter search.
+The most important parameter is
+the enumeration timeout during the wake phase;
+domains that present more challenging program synthesis
+problems require either longer timeouts, more CPUs, or both.
+\begin{center}
+  \begin{tabular}{ccccccc}
+    \toprule
+    Domain&Timeout&CPUs&Batch size&$\lambda$ (Sec.~\ref{consolidationSection})&$\alpha$ (\ref{mapAppendix})&Max frontier (\ref{systemPseudocode})\\\midrule
+    Lists&7m&64&10&1&30&5\\
+    Text&7m&64&10&1&30&5\\
+    Graphics&1h&128&40&1&30&5\\
+    Symbolic regression& & 40&10 & 1&30&5\\
+    Towers& & 64&10 & 1&30&5\\
+    Regexes& & 64&10 & 1&30&5\\
+    \bottomrule
+
+  \end{tabular}
+\end{center}
+
 \end{document}

### 2019-01-08 archive.tex
deep program learning paragraph in introduction

@@
 cycles, both of which are loosely biologically inspired by actual
-human sleep.  The first sleep cycle, which we refer to as
+sleep.  The first sleep cycle, which we refer to as
 \textbf{consolidation}, grows the DSL by replying experiences from
@@
 
+\begin{comment}
 This dream-learning architecture brings together two lines of prior work,
@@
 navigate a new domain of problems.
+\end{comment}
+
+Each wake/sleep cycle creates new DSL components that build on
+components learned in earlier sleep cycles, growing a DSL with nested 
+hierarchies of code.  We identify this cumulative nesting of
+abstractions as a kind of deep representation
+learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
+subset of these learned networks (the DSL). For example, the system
+learns to sort sequences of numbers by invoking a DSL component 4
+layers deep, or draws the leftmost pairs of images in
+Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
+refer to \system as an instance of `deep program learning'.
 
@@
     \bottomrule     
-
   \end{tabular}

### 2019-01-08 archive.tex
random/cherry split for dreams

@@
 omniglot~\cite{lake2015human} are prominent examples.  Here we take
-inspiration from LOGO Turtle graphics~\cite{turtle}, and give our
-agent the ability to control a pen, along with arithmetic operations
-on angles and distances and simple control flow primitives.  We task
-it with drawing a small corpus of images starting from this basis.
+inspiration from LOGO Turtle graphics~\cite{turtle}, tasking it with
+drawing a modest corpus of images while equipping it with control over
+a `pen', along with arithmetic operations on angles and distances.
 
@@
 \begin{figure}
-\centering  \begin{tabular}{c|c}
-      \includegraphics[width = 0.5\textwidth]{figures/initialDreams/montage.png}&
-  \includegraphics[width = 0.5\textwidth]{figures/finalDreams/montage.png}
+\centering  \begin{tabular}{c|cc}
+  \begin{tabular}{c}
+    \includegraphics[width = 0.45\textwidth]{figures/initialDreams/montage.png}
+    \end{tabular}&
+      \begin{tabular}{c}
+        \includegraphics[width = 0.45\textwidth]{figures/finalDreams/cherry_random_montage.png}
+        \end{tabular}&
+      \rotatebox[origin=c]{90}{Cherry picked\hspace{3cm}Random}
     \end{tabular}
-  \caption{Dreams, or samples, from the DSL before (left) and after (right) training on tasks in Figure~\ref{everyLogo}.
+  \caption{Sixteen dreams, or samples, from the DSL before (left) and after (right) training on tasks in Figure~\ref{everyLogo}.
 Blue: where the agent started drawing.

### 2019-01-09 archive.tex
more progress on archive

@@
 
-\section{Wake/Sleep Program Induction}\label{overviewSection}
+\section{Deep Wake/Sleep Program Induction}\label{overviewSection}
 
@@
 
-Viewed as an inference problem, our goal is to start with a collection
-of tasks, written $X$, and infer both a program for each task, as well
+We view these wake sleep/cycles as an approximate inference algorithm that starts with a collection of tasks, written $X$, and infers both a program for each task, as well
 as a distribution over programs encoded by a DSL, written
@@
 $\theta$, and together $(\mathcal{D},\theta)$ define a generative
-model over programs (see Appendix~\ref{sampleProgram}).  We perform
-maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given
+model over programs (Appendix~\ref{sampleProgram}).  We perform
+maximum a posteriori (MAP) inference of the DSL and programs given
 $X$.  Writing $J$ for the joint probability of $(\mathcal{D},\theta)$
-and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
+and $X$, we seek the generative model $\mathcal{D}^*,\theta^*$ solving:
 \begin{align}\label{intractableObjectives}
@@
 \end{align}
-where $\probability[x|p]$ scores the likelihood of a task
-$x\in X$ given a program $p$.\footnote{For example, for list
-  processing, the likelihood is 1 if the program predicts the observed
-  outputs on the observed inputs, and 0 otherwise; when learning a
-  generative model or probabilistic program, the likelihood is the
-  probability of the program sampling the observation.}
-The above equations summarize the problem from the point of view of an ideal Bayesian learner.
-However, Eq.~\ref{intractableObjectives}
-is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
-summing over the  infinite set of all programs.
-In practice we will only ever be able to sum over a finite set of programs.
-So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
-
-\begin{definition}
-A \textbf{frontier of task $x$}, written $\mathcal{F}_x$,
-is a finite set of programs s.t. $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.
-\end{definition}
-\noindent Using the frontiers we  define the following intuitive lower bound on the joint probability, called $\lowerBound$:
+%% where $\probability[x|p]$ scores the likelihood of a task
+%% $x\in X$ given a program $p$.\footnote{For example, for list
+%%   processing, the likelihood is 1 if the program predicts the observed
+%%   outputs on the observed inputs, and 0 otherwise; when learning a
+%%   generative model or probabilistic program, the likelihood is the
+%%   probability of the program sampling the observation.}
+Evaluating Eq.~\ref{intractableObjectives}
+entails marginalizing over the infinite set of all programs -- which is impossible.
+We make a particle-based approximation to Eq.~\ref{intractableObjectives}
+and instead marginalize over a finite \textbf{beam} of programs,
+with one beam per task, collectively written $\left\{\mathcal{B}_x \right\}_{x\in X}$.
+This particle-based approximation is written $\lowerBound (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$
+and acts as a lower bound on the joint:
 \begin{align}
- J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
+ J(\mathcal{D},\theta)\geq \lowerBound  (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, where $|\mathcal{B}_x|$ is small}
 \end{align}
-Our wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$  (\textbf{Wake})
+Wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{B}_x \right\}_{x\in X}$  (\textbf{Wake})
 and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
-\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
-want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
+\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the beams.} Here $(\mathcal{D},\theta)$ is fixed and we
+want to find new programs to add to  the beams so that $\lowerBound$ increases the most.
 $\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
@@
 %Here we use a simple and generic enumerative program synthesis algorithm.
-%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
+%% which we can accomplish by adding new programs to the beams means searching for new programs $p$ for task $x$
 %% where  is large.
-\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
+\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
 Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
-and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.
+and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.
 
@@
 \noindent\textbf{Sleep (Dreaming): tractably maxing $\lowerBound$ w.r.t. the
-  frontiers.}  Here we train %% a neural network, $q$, to predict a
+  beams.}  Here we train %% a neural network, $q$, to predict a
 %% distribution over programs conditioned on a task. The objective of $q$
@@
 $\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
-into the frontiers will most increase $\lowerBound$.
+into the beams will most increase $\lowerBound$.
 
@@
 check if a program $p$ assigns positive probability to a task
-($\probability[x|p] > 0$); if so, we incorporate $p$ into the frontier
-$\mathcal{F}_x$ (Figure~\ref{neuralPipeline}).  We represent programs
-as polymorphicly-typed $\lambda$-calculus expressions, a
-representation closely resembling Lisp and functional languages like
-Haskell and OCaml, including variables, conditionals, higher-order
-recursive functions, and the ability to create new functions.
-We `batch' the tasks by randomly shuffling the training set
-and playing small minibatches
-of tasks to the agent during each wake cycle.
+($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
+$\mathcal{B}_x$. We represent programs as polymorphicly-typed
+$\lambda$-calculus expressions, a representation closely resembling
+Lisp and functional languages like Haskell and OCaml, including
+variables, conditionals, higher-order recursive functions, and the
+ability to create new functions.  We `batch' the tasks by randomly
+shuffling the training set and playing small minibatches of tasks to
+the agent during each wake cycle.
 
@@
 the space of  programs.
+\begin{comment}
   \begin{figure}
@@
       \node[draw,align = center](s) at ([xshift=2cm]e.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
-    \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Frontier\\$\mathcal{F}_x$};
+      \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Beam\\$\mathcal{B}_x$};
 
@@
 Recognition model outputs distribution over program $Q(p|x)$.
-Program output by enumerative search incorporated into frontier if likelihood $\probability[x|p] > 0$}
+Program output by enumerative search incorporated into beam if likelihood $\probability[x|p] > 0$}
     \label{neuralPipeline}
   \end{figure}
+\end{comment}
 
@@
 finding the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
-We replace this marginal with an AIC approximation
+We replace this marginal with an AIC approximation~\cite{akaike1998information}
 and minimize the following expression,
@@
   X}\underbrace{-\log \sum_{p\in
-  \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
+  \mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
 \label{AIC}
@@
 whenever we add a new expression $e$ to the DSL,
-the programs found during waking ($\left\{\mathcal{F}_x \right\}$)
+the programs found during waking ($\left\{\mathcal{B}_x \right\}$)
 are not written in terms of $e$.
@@
         -\log \probability[\mathcal{D}] + 
-\min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{F}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
+\min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{B}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
 \label{factorObjective}
@@
 a version space with $10^6$ nodes, which encodes $10^{14}$
-refactorings.
+refactorings. Appendix~\ref{appendixCompression} specifies how we combine this probabilistic and symbolic machinery to update the DSL.
 
@@
 %% \begin{equation}
-%%         \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
+%%         \log \probability[\mathcal{D}] + \argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 +\sum_{x\in X}\log \sum_{p\in \mathcal{B}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
 %% \label{limitedObjective}
@@
 
-With this machinery in hand, we now have all the pieces needed to
-learn a DSL (Algorithm~\ref{grammarInductionAlgorithm}).
-The functions $I\beta(\cdot )$ and \textsc{refactor} construct
-a version space from a program and extract the shortest program from a version space, respectively (Algorithm~\ref{grammarInductionAlgorithm}, lines 5-6, 14; see Appendix~\ref{appendixVersion}).
-To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
-  and $\lambda$ controls how strongly we regularize the size of the DSL.
-  We place a symmetric Dirichlet prior over the weight vector $\theta$.
-
-To appropriately score each proposed $\mathcal{D}$ we must reestimate
- the weight vector $\theta$ (Algorithm~\ref{grammarInductionAlgorithm}, line 7).
-Although this  may seem 
-very similar to estimating the parameters of a probabilistic context free grammar,
-for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
-our DSLs are context-sensitive due to the presence of variables
-in the programs and also due to the polymorphic typing system.
-Appendix~\ref{mapAppendix} derives a tractable MAP estimator for $\theta$.
-
-\begin{algorithm}%[H]
-  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
-  \caption{DSL Induction Algorithm}
-  \label{grammarInductionAlgorithm}
-  \begin{algorithmic}[1]
-    \State {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
-    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
-    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
-    \While{true}         
-    \State $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
-    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
-    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
-    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
-    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
-    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
-    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
-    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
-    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
-    \State $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
-    \EndWhile
-  \end{algorithmic}
-\end{algorithm}
-
 \begin{figure*}
@@
   \footnotesize
-  \begin{scope}[shift={(2,-11)}]  
+  \begin{scope}[shift={(-1,-12)}]  
     \node[draw, rounded corners](u1) at (0,0) {union};
@@
 
-    \node[anchor=north](p) at ($(-6,0) + (u2)$) {\normalsize $\underbrace{\texttt{(+ 1 1)}}_{\text{\normalsize  Program}}$};
-    \draw[ultra thick,->] ($(0.5,0) + (p.east)$) -- ($(1.5,0) + (p.east)$);
-    \node at ($(1,0.25) + (p.east)$) {Refactors};
+    \node[anchor=north](p) at ($(0,2.5) + (u1)$) {\normalsize Program: \texttt{(+ 1 1)}};
+    \draw[ultra thick,->] ($(0,-0.1) + (p.south)$) --node[sloped, above, inner sep=5]{Refactors} ($(0,0.1) + (u1.north)$);
     \node(g)[anchor=left] at ($(7,0) + (u12.east)$) {\includegraphics[width = 6cm]{figures/vs.eps}};
@@
   \mathcal{L}^{\text{MAP}} &= \mathcal{L}_{\text{Replay}}^{\text{MAP}} + \mathcal{L}_{\text{Fantasy}}^{\text{MAP}}\\
-  \mathcal{L}_{\text{Replay}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
-  \mathcal{L}_{\text{Replay}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
+  \mathcal{L}_{\text{Replay}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{B}_x}
+    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{B}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
+  \mathcal{L}_{\text{Replay}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{B}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
   \mathcal{L}_{\text{Fantasy}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
@@
 omniglot~\cite{lake2015human} are prominent examples.  Here we take
-inspiration from LOGO Turtle graphics~\cite{turtle}, tasking it with
-drawing a modest corpus of images while equipping it with control over
-a `pen', along with arithmetic operations on angles and distances.
+inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
+with drawing a modest corpus of images while equipping it with control
+over a `pen', along with arithmetic operations on angles and
+distances.
 
@@
 %% \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-%% only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL.
+%% only add \textbf{S}ub\textbf{E}xpressions of programs in the beams to the DSL.
 This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
@@
 %% but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-%% \\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
+%% \\\noindent \textbf{Enum}, which enumerates a beam without any learning --- equivalently, our first search step.
 %For each domain,
@@
     \State {\bfseries Input:} Initial DSL $\mathcal{D}$, tasks $X$
-    \State \textbf{Output:} Infinite stream of DSLs, recognition models, and frontiers
-    \State \textbf{Hyperparameters:} Batch size $B$, enumeration timeout $T$, maximum frontier size $F$
+    \State \textbf{Output:} Infinite stream of DSLs, recognition models, and beams
+    \State \textbf{Hyperparameters:} Batch size $B$, enumeration timeout $T$, maximum beam size $F$
     \State $\theta\gets $uniform distribution
-    \State $\mathcal{F}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize frontiers to be empty}
+    \State $\mathcal{F}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize beams to be empty}
     \While{true}\Comment{Loop over epochs}
@@
 
- \subsection{Refactoring code with version spaces}\label{appendixVersion}
+ \subsection{Details of DSL learning}\label{appendixCompression}
+
+ Algorithm~\ref{grammarInductionAlgorithm} specifies our DSL learning procedure.
+ This integrates two toolkits:
+ the machinery of version spaces (Appendix~\ref{appendixVersion})
+ and a probabilistic objective favoring compressive DSLs.
+ The functions $I\beta(\cdot )$ and \textsc{refactor} construct
+a version space from a program and extract the shortest program from a version space, respectively (Algorithm~\ref{grammarInductionAlgorithm}, lines 5-6, 14; Appendix~\ref{appendixVersion}).
+To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
+  and $\lambda$ controls how strongly we regularize the size of the DSL.
+  We place a symmetric Dirichlet prior over the weight vector $\theta$.
+
+To appropriately score each proposed $\mathcal{D}$ we must reestimate
+ the weight vector $\theta$ (Algorithm~\ref{grammarInductionAlgorithm}, line 7).
+Although this  may seem 
+very similar to estimating the parameters of a probabilistic context free grammar,
+for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
+our DSLs are context-sensitive due to the presence of variables
+in the programs and also due to the polymorphic typing system.
+Appendix~\ref{mapAppendix} derives a tractable MAP estimator for $\theta$.
+
+\begin{algorithm}%[H]
+  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
+  \caption{DSL Induction Algorithm}
+  \label{grammarInductionAlgorithm}
+  \begin{algorithmic}[1]
+    \State {\bfseries Input:} Set of beams $\{\mathcal{B}_x\}$
+    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
+    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{B}_x\}$
+    \While{true}         
+    \State $\forall p\in \bigcup_{x}\mathcal{B}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
+    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(v_p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
+    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
+    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*(\mathcal{D})) - \|\theta^*(\mathcal{D})\|_0$ \Comment{objective function}
+    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{B}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
+    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
+    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
+    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
+    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
+    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new DSL}
+    \EndWhile
+  \end{algorithmic}
+\end{algorithm}
+
+ \subsubsection{Refactoring code with version spaces}\label{appendixVersion}
  Formally, a version space is either:
@@
 version spaces takes time linear in the number of programs (written
-$P$) in the frontiers (Algorithm~\ref{grammarInductionAlgorithm}, line
+$P$) in the beams (Algorithm~\ref{grammarInductionAlgorithm}, line
 5), and, in the worst case, exponential time as a function of the
@@
 
-\subsection{Estimating the continuous parameters $\theta$ of a DSL}\label{mapAppendix}
+\subsubsection{Estimating the continuous parameters $\theta$ of a DSL}\label{mapAppendix}
 We use an EM algorithm to estimate the continuous parameters of the DSL, i.e. $\theta$.
@@
     \toprule
-    Domain&Timeout&CPUs&Batch size&$\lambda$ (Sec.~\ref{consolidationSection})&$\alpha$ (\ref{mapAppendix})&Max frontier (\ref{systemPseudocode})\\\midrule
+    Domain&Timeout&CPUs&Batch size&$\lambda$ (\ref{appendixCompression})&$\alpha$ (\ref{mapAppendix})&Max frontier (\ref{systemPseudocode})\\\midrule
     Lists&7m&64&10&1&30&5\\

### 2019-01-09 archive.tex
added some dreams to the deep logo figure

@@
     \draw[->] (p2.south) -- (pu.east);
-    \draw[->] (p3.south) to[out=320,in=45] (line.east);
+    \draw[->] (p3.south) to[out=190,in=120] (line.west);
     \draw[->] (p3.south) to[out=200,in=120] (snow.west);
@@
     \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};
+
+    \node[align=center,rotate=90] at ([xshift=1cm]snow.east) {$\underbrace{\hspace{6cm}}_{\text{\normalsize \textbf{\textsc{Dreams}: Samples from DSL}}}$};
+
+    \node at ([xshift=3cm]semi.east) {\includegraphics[width = 1.5cm]{figures/finalDreams/63smooth_pretty}};
+    \node at ([xshift=3cm]snow.east) {\includegraphics[width = 1.5cm]{figures/finalDreams/400smooth_pretty}};
+    \node at ([xshift=3cm]line.east) {\includegraphics[width = 1.5cm]{figures/finalDreams/399smooth_pretty}};
   \end{tikzpicture}

### 2019-01-09 archive.tex
improvements to deep logo figure

--- 2019-01-09 archive.tex
@@
     \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};
+
+    \node(d1) at ([xshift=3.5cm,yshift=-0.5cm]semi.east) {\includegraphics[width = 1.5cm]{figures/finalDreams/63smooth_pretty}};
+    \node(d2) at ([yshift=-1cm]d1.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/400smooth_pretty}};
+    \node(d3) at ([yshift=-1cm]d2.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/399smooth_pretty}};
+
+    \node(d4) at ([yshift=-1cm]d3.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/425smooth_pretty}};
+    \node(dreamPanel)[draw,fit = (d1) (d2) (d3) (d4)] {};
+    \node[anchor=south,align=center] at (dreamPanel.north) {\textbf{\textsc{Dreams}}\\(\footnotesize Samples from DSL)};
+
+    \node(curly)[align=center,rotate=90] at ([xshift=1cm]snow.east) {$\underbrace{\hspace{6cm}}_{}$};
+    \draw[squiggle,->] ($(0,-3.3) + (curly.east)$) -- node[along, sloped, above]{sample}(dreamPanel.west);
+
   \end{tikzpicture}

### 2019-01-13 archive.tex
more writing

@@
 This particle-based approximation is written $\lowerBound (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$
-and acts as a lower bound on the joint:
+and acts as a lower bound on the joint density:
 \begin{align}
@@
 $\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
-is large (i.e., programs with high posterior probability).
-%Here we use a simple and generic enumerative program synthesis algorithm.
-%% which we can accomplish by adding new programs to the beams means searching for new programs $p$ for task $x$
-%% where  is large.
-\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$.
-Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
-and then updating $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.
-
-Searching for programs is hard because
-of the large combinatorial search space. We ease this difficulty by training a neural recognition model, $Q(\cdot |\cdot )$,
-during the \textbf{Dreaming} phase: $Q$ is trained to approximate the
-posterior over programs, $Q(p|x)\approx \probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
-  Thus training the neural network amortizes the cost of finding programs with high posterior probability.
+is large, i.e., programs with high posterior probability,
+or programs which are a priori likely and which solve the task.
+We find programs solving a task by
+enumerating programs
+from the DSL in decreasing order of their probability
+under the recognition model,
+and then check if a program $p$ assigns positive probability to a task
+($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
+$\mathcal{B}_x$. We represent programs as polymorphicly typed $\lambda$-calculus expressions,
+an expressive formalism
+including conditionals,
+variables,
+higher-order recursive functions,
+and the ability to define new functions.
+\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed and the problem is to search the discrete space of DSLs and find one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
+and then update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.
+
+Finding programs solving tasks is difficult because of the infinitely
+large, combinatorial search landscape. We ease this difficulty by
+training a neural recognition model, $Q(p |x )$, during the
+\textbf{Dreaming} phase: $Q$ is trained to approximate the posterior
+over programs, $Q(p|x)\approx
+\probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
+Thus training the neural network amortizes the cost of finding
+programs with high posterior probability.
 
@@
 
-Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others.
-As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
-the Consolidation cycle has more data from which to learn the DSL; and, because the recognition model is trained on both
-samples from the DSL and programs found during waking,
-the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
+This 3-phase inference procedure works through two distinct kinds of bootstrapping.
+During each sleep cycle the next DSL
+bootstraps off the primitives learned
+during earlier cycles,
+growing an increasingly deep learned program representation (`deep program learning').
+In tandem the DSL and recognition model
+bootstrap each other:
+a more finely tuned
+DSL yields
+richer dreams for the recognition model to learn from,
+while a more accurate recognition model
+solves more tasks during waking which then feed into the next DSL.
 \begin{figure}
@@
 \end{figure}
-
-\subsection{Wake: Solving tasks}\label{explorationSection}
-
-During waking we enumerate programs from the DSL in decreasing order
-of their probability according to the recognition model, and then
-check if a program $p$ assigns positive probability to a task
-($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
-$\mathcal{B}_x$. We represent programs as polymorphicly-typed
-$\lambda$-calculus expressions, a representation closely resembling
-Lisp and functional languages like Haskell and OCaml, including
-variables, conditionals, higher-order recursive functions, and the
-ability to create new functions.  We `batch' the tasks by randomly
-shuffling the training set and playing small minibatches of tasks to
-the agent during each wake cycle.
 
@@
 
+\begin{comment}
+  \subsection{Wake: Solving tasks}\label{explorationSection}
+
+During waking we enumerate programs from the DSL in decreasing order
+of their probability according to the recognition model, and then
+check if a program $p$ assigns positive probability to a task
+($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
+$\mathcal{B}_x$. We represent programs as polymorphicly-typed
+$\lambda$-calculus expressions, a representation closely resembling
+Lisp and functional languages like Haskell and OCaml, including
+variables, conditionals, higher-order recursive functions, and the
+ability to create new functions.  We `batch' the tasks by randomly
+shuffling the training set and playing small minibatches of tasks to
+the agent during each wake cycle.
+
 Why enumerate, when the program synthesis community has invented many
@@
 the space of  programs.
+\end{comment}
 \begin{comment}
@@
 algorithm for automatically refractoring programs, where this
-refactoring exposes common reused subexpressions found across the
+refactoring exposes common reused subexpressions across the
 programs found during waking.  Second, we use this automatic refactoring
@@
 We replace this marginal with an AIC approximation~\cite{akaike1998information}
-and minimize the following expression,
+and marginalize over refactorings of programs in the beams, minimizing the following expression,
 which can be interpreted as a kind of compression:
@@
 \probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
-  X}\underbrace{-\log \sum_{p\in
-  \mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
+  X}\underbrace{-\log \sum_{\substack{p:\\
+      \exists p'\in
+  \mathcal{B}_x: p\text{ refactors into }p'}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
 \label{AIC}
 \end{equation}
-At a high level, our approach is to search locally through the space 
-of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
-fails to decrease.  These small changes consist of introducing new
-candidate $\lambda$-expressions into the DSL.
-
-However, there is a snag with this simple approach:
-whenever we add a new expression $e$ to the DSL,
-the programs found during waking ($\left\{\mathcal{B}_x \right\}$)
-are not written in terms of $e$.
-Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
-after having found the programs
-\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
-As human programmers,
-we can look at these pieces of code and recognize that,
-if we define a new procedure called \code{double},
-defined as \code{($\lambda$ (x) (+ x x))},
-then we can rewrite the original programs as
-\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
-This process is a kind of refactoring
-where a new subroutine is defined (\code{double})
-and the old programs rewritten
-in terms of the new subroutine.
-Figure~\ref{mapFactor}A
-diagrams this refactoring
-process for a more complicated
-setting,
-where the agent must rediscover the higher-order function \code{map}
-starting from the basics of Lisp and the Y-combinator.
-
-We refine our objective to compress
-\emph{refactorings} of the programs found during waking, minimizing
-\begin{equation}
-        -\log \probability[\mathcal{D}] + 
-\min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{B}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
-\label{factorObjective}
-\end{equation}
-where $p\manyReduce p'$ is the standard notation for ``expression $p$
-evaluates to $p'$ by the rules of $\lambda$-calculus''~\cite{pierce}.
-Equation~\ref{factorObjective} captures the idea that we want to add
-new components to the DSL while jointly refactoring our old programs
-in terms of these new components.  But this joint optimization is
-intractable, because there are infinitely many ways of refactoring a
-program.  To make refactoring tractable we first limit the degree to
-which a piece of code can be refactored: rather than consider every
-refactoring, we bound the number of $\lambda$-calculus evaluation
-steps separating a refactoring from its original program.  Now the
-number of refactoring this is finite but astronomically large: for the
-example in Figure~\ref{mapFactor} there are approximately $10^{14}$
-possible refactorings -- a quantity that grows exponentially both as a
-function of program size and a function of the bound on evaluation
-steps. How can we tame this combinatorial explosion?
+But a program has infinitely many possible refactorings, rendering
+Eq.~\ref{AIC} intractable.  Rather than consider every refactoring we
+bound the number of $\lambda$-calculus evaluation steps separating a
+program from its refactoring.  Now the number of refactorings is
+finite but astronomically large: Figure~\ref{mapFactor}A diagrams a
+problem where the agent rediscovers the higher-order function
+\code{map} starting from the basics of Lisp and the Y-combinator, but
+where there are approximately $10^{14}$ possible refactorings -- a quantity
+that grows exponentially both as a function of program size and a
+function of the bound on evaluation steps. How can we tame this
+combinatorial explosion?
 
@@
 Appendix~\ref{appendixVersion}, we give a dynamic program that takes
-as input a program $p$ and then outputs a version space containing the
-refactorings of $p$. Figure~\ref{mapFactor}B diagrams a subtree of a
+as input a program and then outputs a version space containing its
+refactorings. Figure~\ref{mapFactor}B diagrams a subtree of a
 version space containing refactorings of a small program. Our
@@
 refactorings. Appendix~\ref{appendixCompression} specifies how we combine this probabilistic and symbolic machinery to update the DSL.
+At a high level, our approach is to search locally through the space 
+of DSLs, proposing small changes until Eq.~\ref{AIC}
+fails to decrease.
+
+%% However, there is a snag with this simple approach:
+%% whenever we add a new expression $e$ to the DSL,
+%% the programs found during waking
+%% are not written in terms of $e$ ---
+%% and so we must \emph{refactor}
+%% the programs in terms of the new DSL component.
+%% Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
+%% after having found the programs
+%% \code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
+%% As human programmers,
+%% we can look at these pieces of code and recognize that,
+%% if we define a new procedure called \code{double},
+%% defined as \code{($\lambda$ (x) (+ x x))},
+%% then we can rewrite the original programs as
+%% \code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
+%% This process is a kind of refactoring
+%% where a new subroutine is defined (\code{double})
+%% and the old programs rewritten
+%% in terms of the new subroutine.
+%% Figure~\ref{mapFactor}A diagrams a refactoring process where the agent
+%% must rediscover the higher-order function \code{map} starting from the
+%% basics of Lisp and the Y-combinator.
+
+%% We refine our objective to compress
+%% \emph{refactorings} of the programs found during waking, minimizing
+%% \begin{equation}
+%%         -\log \probability[\mathcal{D}] + 
+%% \min_{\theta}\Bigg(-\log \probability[\theta|\mathcal{D}] + \|\theta\|_0 +\sum_{x\in X} - \log \underbrace{\sum_{\substack{p:\\\exists p'\in \mathcal{B}_x: p\manyReduce p'}}}_{\text{Refactors }\mathcal{B}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
+%% \label{factorObjective}
+%% \end{equation}
+%% where $p\manyReduce p'$ is the standard notation for ``expression $p$
+%% evaluates to $p'$ by the rules of $\lambda$-calculus''~\cite{pierce}.
+%% Equation~\ref{factorObjective} captures the idea that we want to add
+%% new components to the DSL while jointly refactoring our old programs
+%% in terms of these new components.  But this joint optimization is
+%% intractable, because there are infinitely many ways of refactoring a
+%% program.  To make refactoring tractable we first limit the degree to
+%% which a piece of code can be refactored: rather than consider every
+%% refactoring, we bound the number of $\lambda$-calculus evaluation
+%% steps separating a refactoring from its original program.  Now the
+%% number of refactoring this is finite but astronomically large: for the
+%% example in Figure~\ref{mapFactor} there are approximately $10^{14}$
+%% possible refactorings -- a quantity that grows exponentially both as a
+%% function of program size and a function of the bound on evaluation
+%% steps. How can we tame this combinatorial explosion?
 
@@
   \caption{DSL learning as code refactoring.
-\textbf{Panel A:} For each task we discover programs during waking, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}).
+\textbf{Panel A:} During waking we discover programs for each task, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}).
 Common subprograms are incorporated into the DSL when they increase a Bayesian objective.
@@
 will show that $\mathcal{L}^{\text{MAP}}$ forces the recognition model
-to break symmetries in the space of programs
-(Sec.~\ref{symmetricSection}).
+to break symmetries in the space of programs.
+
+\noindent\textbf{Parameterizing $Q$.} The recognition model predicts a
+fixed-dimensional tensor encoding a distribution over subroutines in
+the DSL, conditioned on the local context in the syntax tree of the
+program. This local context consists of the parent node in the syntax
+tree, as well as which argument is being generated, functioning as a
+kind of `bigram' model over trees. Figure~\ref{symmetry} (left)
+diagrams this generative process.  This parameterization confers three
+main advantages: (1) it supports fast enumeration and sampling of
+programs, because the recognition model only runs once per
+task, like in~\cite{balog2016deepcoder,ecc,menon2013machine} -- thus
+we can fall back on fast enumeration if the target program is unlike
+the training programs; (2) the recognition model provides fine-grained
+information about the structure of the target program, similar
+to~\cite{devlin2017robustfill,zavershynskyi2018naps}; and (3) in
+conjunction with $\mathcal{L}^{\text{MAP}}$ the recognition model
+learns to break symmetries in the space of programs.
+
+\noindent\textbf{Symmetry breaking.} A good DSL not only exposes high-level building blocks, but also
+carefully restricts the ways in which those building blocks are
+allowed to compose.  For example,
+%% a DSL for list manipulation should
+%% contain both the empty list and a routine for appending lists, but
+%% should not allow appending the empty list.  Similarly 
+a DSL for arithmetic should disallow adding zero, or force
+right-associative addition.  These restrictions break symmetries in
+the space of programs.  A bigram parameterization of the recognition
+model, combined with the $\mathcal{L}^{\text{MAP}}$ training
+objective, interact in a way that breaks symmetries in the program
+space, allowing the agent to more efficiently explore the space of
+programs.  This interaction occurs because the bigram parameterization
+can disallow DSL primitives depending on their local syntactic
+context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
+probability mass onto a single member of a set of syntactically
+distinct but semantically equivalent expressions
+(Appendix~\ref{recognitionAppendix}).
+We experimentally confirm this symmetry-breaking
+behavior by training recognition models that minimize either
+$\mathcal{L}^{\text{MAP}}$/$\mathcal{L}^\text{posterior}$ and which
+use either a bigram parameterization/unigram\footnote{In the unigram variant $Q$ predicts a $|\mathcal{D}| + 1$-dimensional vector: $Q(p|x) = \probability[p|\mathcal{D},\theta_i = Q_i(x)]$,
+  and was used in our prior work~\cite{ecc}} parameterization.
+Figure~\ref{symmetry} (right) shows the result of training $Q$ in these four regimes
+and then sampling programs.
+On this particular run,
+the combination of
+bigrams and $\mathcal{L}^{\text{MAP}}$ learns to
+avoid adding zero and associate addition to the right ---
+different random initializations
+lead to either right or left association.
+
+\begin{comment}
 
@@
 This parameterization
-has three main advantages:
+confers three main advantages:
 (1) it supports fast enumeration
@@
 lead to either right or left association.
-
+  \end{comment}
 \begin{figure}
@@
 
-\begin{figure}
-\centering  \begin{tabular}{cll}
-     \toprule
-&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
-     $\mathcal{L}^{\text{posterior}}$&
-     \begin{tabular}{l}
-       \emph{Three samples:}\\
-       \code{(+ 1 0)}\\
-\code{(+ (+ 0 0) (+ 1 0))}\\
-\code{(+ 1 1)}
-\\
-63.0\% right-associative; 37.4\% \code{+0}'s
-       \end{tabular}
-     &
-     \begin{tabular}{l}
-       \emph{Three samples:}\\
-       \code{0}\\
-\code{(+ (+ (+ 0 0) (+ 0 1)) 1)}\\
-\code{1}\\
-55.8\% right-associative; 31.9\% \code{+0}'s
-       \end{tabular}
-     \\\\
-     $\mathcal{L}^{\text{MAP}}$&
-     \begin{tabular}{l}
-       \emph{Three samples:}\\
-       \code{1}\\
-       \code{(+ 1 (+ 1 (+ (+ 1 (+ 1 1)) 1)))}\\
-       \code{(+ (+ 1 1) 1)}
-\\48.6\% right-associative;
-       0.5\% \code{+0}'s       
-       \end{tabular}
-     &
-     \begin{tabular}{l}
-       \emph{Three Samples:}\\
-\code{(+ 1 (+ 1 (+ 1 (+ 1 (+ 1 1)))))}\\
-\code{0}\\
-\code{(+ 1 (+ 1 (+ 1 1)))}
-       \\
-       \textbf{97.9\% right-associative};
-       2.5\% \code{+0}'s       
-       \end{tabular}
-\\    \bottomrule 
-  \end{tabular}
-  \caption{Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero.
+%% \begin{figure}
+%% \centering  \begin{tabular}{cll}
+%%      \toprule
+%% &     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
+%%      $\mathcal{L}^{\text{posterior}}$&
+%%      \begin{tabular}{l}
+%%        \emph{Three samples:}\\
+%%        \code{(+ 1 0)}\\
+%% \code{(+ (+ 0 0) (+ 1 0))}\\
+%% \code{(+ 1 1)}
+%% \\
+%% 63.0\% right-associative; 37.4\% \code{+0}'s
+%%        \end{tabular}
+%%      &
+%%      \begin{tabular}{l}
+%%        \emph{Three samples:}\\
+%%        \code{0}\\
+%% \code{(+ (+ (+ 0 0) (+ 0 1)) 1)}\\
+%% \code{1}\\
+%% 55.8\% right-associative; 31.9\% \code{+0}'s
+%%        \end{tabular}
+%%      \\\\
+%%      $\mathcal{L}^{\text{MAP}}$&
+%%      \begin{tabular}{l}
+%%        \emph{Three samples:}\\
+%%        \code{1}\\
+%%        \code{(+ 1 (+ 1 (+ (+ 1 (+ 1 1)) 1)))}\\
+%%        \code{(+ (+ 1 1) 1)}
+%% \\48.6\% right-associative;
+%%        0.5\% \code{+0}'s       
+%%        \end{tabular}
+%%      &
+%%      \begin{tabular}{l}
+%%        \emph{Three Samples:}\\
+%% \code{(+ 1 (+ 1 (+ 1 (+ 1 (+ 1 1)))))}\\
+%% \code{0}\\
+%% \code{(+ 1 (+ 1 (+ 1 1)))}
+%%        \\
+%%        \textbf{97.9\% right-associative};
+%%        2.5\% \code{+0}'s       
+%%        \end{tabular}
+%% \\    \bottomrule 
+%%   \end{tabular}
+%%   \caption{Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero.
 \% right-associative calculated by drawing 500 samples from $Q$.
@@
 }\label{symmetry}
-  \end{figure}
+%%   \end{figure}
 
@@
 processing and text editing. In both cases we solve tasks specified by
-a few input/output examples, using a recurrent neural network for the
-recognition model and starting with a generic functional programming
+a input/output examples, starting with a generic functional programming
 basis: \code{foldr}, \code{unfold}, \code{if}, \code{map},
@@
 system composed 16 new DSL components, and discovered multiple
-higher-order functions, such as \code{filter} ($f_0$ in
-Figure~\ref{initialExampleDSL}).  Each round of memory consolidation
-built on components discovered in earlier sleep cycles --- for example
-the agent first learns \code{filter}, uses filter to learn to take the
-maximum element of a list, then uses that routine to learn a new
-component for extracting the $n^{\text{th}}$ largest element of a
-list, which it finally uses to learn how to sort a list of numbers
-(Figure~\ref{initialExampleDSL}).  This incremental, modular learning
-of deep hierarchies of DSL components occurs
-because of the alternation between code writing and code refactoring.
+higher-order functions. Each round of memory consolidation built on
+components discovered in earlier sleep cycles --- for example the
+agent first learns the higher-order function \code{filter}, uses
+\code{filter} to learn to take the maximum element of a list, then
+uses that routine to learn a new component for extracting the
+$n^{\text{th}}$ largest element of a list, which it finally uses to
+learn how to sort a list of numbers (Figure~\ref{initialExampleDSL}).
+This incremental, modular learning of deep hierarchies of DSL
+components occurs because of the alternation between code writing
+(during waking) and code refactoring (during the consolidation phase
+of sleep).
 
@@
 image (Figure~\ref{visualSpecs}).
-\subsubsection{Programs that make plans and take actions}
-
-We apply \system to two domains where the agent plans a series of actions
-in order to draw a picture (Sec.~\ref{logoSection}) or build a tower out of blocks (Sec.~\ref{towerSection}).
-For drawing pictures,
-the agent controls a `pen' that it can pick up or place down while moving across a canvas.
-For building towers,
-the agent controls a `hand' that it moves through a simulated world while
-placing down differently sized blocks.
-For each of these domains,
-the agent observes a target picture or block tower,
-and must either draw the picture or build the tower.
-The recognition model is a CNN that
-observes an image of the target picture or tower.
-The system is initially provided with two control flow operators:
-\code{loop} (a `for' loop) and \code{get/set},
-which saves and then restores the current state of the pen or hand.
+First we consider programs that make plans and take actions:  drawing pictures and building towers out of blocks (Sec.~\ref{logoSection}-\ref{towerSection}).
 
@@
 training data, like polygons, circles, and spirals
-(Figure~\ref{logoPrimitives}, left). It additionally learns more
-abstract visual relationships, like rotational symmetry, which it
-models by incorporating a higher-order function into its DSL
-(Figure~\ref{logoPrimitives}, right).  This abstraction comes from
-jointly compressing programs for all of the training images into its
-DSL.
+(Figure~\ref{logoPrimitives}, left) -- without supervision the agent
+has learned the basic types of objects in its visual world. It
+additionally learns more abstract visual relationships, like
+rotational symmetry, which it models by incorporating a higher-order
+function into its DSL (Figure~\ref{logoPrimitives}, right).  This
+abstraction comes from jointly compressing programs for all of the
+training images into its DSL.
 
@@
 building blocks and motifs acquired from the training data. This
-offers a window into how the generative model bootstraps recognition
+offers a visual window into how the generative model bootstraps recognition
 model training: as the DSL grows more finely tuned to the domain, the
@@
 These towers are built from Lego-style blocks
-that snap together on a discrete grid.%%  (e.g., we do not model gravity
-%% and discretize the coordinates of the 
-%% blocks).
+that snap together on a discrete grid.
 The system starts with the same control flow primitives as with
@@
 \subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
-We evaluate  on held-out testing tasks,
-measuring how many
-tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
-Prior to any learning,
-the system cannot find solutions for most of the tasks,
-and those it does solve take a long time;
-with more wake/sleep iterations,
-we converge upon DSLs and recognition models 
-more closely matching the domain.
+To evaluate the relative importance of DSL learning and recognition
+model training, we evaluate on held-out testing tasks for each of our
+domains, measuring both how many tasks are solved and how long it
+takes to solve them across successive wake/sleep iterations
+(Fig.~\ref{learningCurves}).  We always solve more held-out tasks --
+and generally solve them in less time -- with both components
+combined.  Why?  One hypothesis is that some tasks are best solved by
+DSL learning and others by a neural network, and so including both of
+our sleep cycles  takes the union of these sets of tasks.  Another
+hypothesis is that the DSL and neural network interact
+synergistically, bootstrapping off each other to solve tasks for which
+neither alone suffice. We evaluate the relative weight of these interactions by
+computing the ratio of the tasks solved uniquely by the full model to
 
@@
     \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
+    \includegraphics[width = 5cm]{figures/regexLearningCurve_batch.png}
   \end{tabular}
   \caption{Test set performance across wake/sleep iterations.
-An Epoch is a full pass over the set of training tasks.
 Orange: \systemEnding.
@@
 We ran the following experiment: \system was given a subset of the
-1959 Lisp primitives, and tasked with solving 22 programming
+1959 Lisp primitives, and tasked with solving 18 programming
 exercises. A key difference between this setup and our previous
@@
 
-After running for 93 hours on 64 CPUs, our
-algorithm solves these 22 exercises, along the way assembling a DSL
+After running for 93 hours on 48 CPUs, our
+algorithm solves these 18 exercises, along the way assembling a DSL
 with a modern repertoire of
 functional programming idioms and subroutines, including \code{map},
-\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
+\code{fold}, \code{unfold}, \code{index}, \code{length},
 and  arithmetic operations like 
@@
 
-\subsection{Program Induction as part of the generic AI toolkit}
-
-Our aim with \system is both to show how learning can enable better
-program synthesis, and to show how better program synthesis can be
-more broadly useful for AI.
-Many challenges remain.
-
-How could a program induction approach to
-AI become broadly useful across a diverse spectrum of AI problems ---
-planning, natural language understanding, machine vision, causal
-inference, and more?
+\subsection{Acquiring Domain Expertise}
+
+One interpretation of our system is as a model of the acquisition of
+domain expertise. Humans can acquire expertise across many domains --
+cooking, coding, music, architecture, painting, tennis, or calculus,
+to name a handful of examples, and every child develops expertise in
+natural language, intuitive physics, motor control, kinship
+relationships, and more. Domain experts learn domain-specific
+abstractions, similar to a DSL: for example, a expert chef knows what
+combinations of seasonings go together, or an expert mathematician
+knows a wide set of useful theorems and lemmas.  Jointly, experts
+learn how to recognize when to use these abstractions to solve
+problems.  Becoming an expert involves a learning trajectory that
+unfolds over relatively long time scales, but has modest data
+requirements relative to the dominant machine learning paradigms. For
+example, basic competency in cooking or coding might require on the
+order of learning a hundred recipes or solving a hundred programming
+exercises.  It is this learning regime that we have targeted with
+\systemEnding: Learning about a domain from at most several hundred
+tasks, but where the learning unfolds over many wake/sleep cycles.
+
+\subsection{Prospects for program Induction as part of the generic AI toolkit}
+
+Our aim with \system is to chart a path by which program induction can
+become more broadly useful for AI.  This means viewing the AI
+landscape through the lens of program learning, including the terrain
+considered here --- simple kinds of generative modeling, inverse
+graphics, planning, and programming by example --- but also many
+others like reinforcement learning, commonsense reasoning, natural
+language understanding, and causal inference. Can program induction
+rise to the challenge?  We believe it can, provided we push jointly
+along  many different axes of AI research; and provided we continue
+to integrate learning algorithms -- both symbolic and neural, both
+top-down and bottom-up -- into our artificial agents.
+
+Our system, with its learned DSL and neural recognition model, is one
+embodiment of this hybrid symbolic/neural approach, and enjoys some
+success across small-scale problems in the different domains considered
+here. Scaling to larger problems,
+such as inferring 3-D object models (vs LOGO/Turtle),
+learning natural image grammars (vs ),
 
@@
     \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
-    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new DSL}
+    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(v_p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new DSL}
     \EndWhile
@@
 Recall that our goal is to define an operator over version spaces, $I\beta_n$,
-which calculates the set of $n$-step refactorings.
+which calculates the set of $n$-step refactorings of a program $p$,
+e.g.,
+the set of all programs $p'$ where
+$p'\underbrace{\reduce q\reduce\cdots\reduce}_{\text{$\leq n$ times}} p$,
+where $a \reduce b$
+is the standard notation
+for $a$ rewriting to $b$ according to the standard rewrite rules of $\lambda$-calculus~\cite{pierce}.
+
 We define this operator in terms of another operator, $I\beta'$, which performs a single step of refactoring:
@@
 Complete set of tasks and learned DSL shown.
-Learned DSL components are numbered in the order that they are learned.}\label{learningFromScratch}
+Learned DSL components are numbered in the order that they are learned, \emph{i.e.}, the agent first learns fold, then unfold, then uses fold to define map, etc.}\label{learningFromScratch}
   \end{figure}
@@
     Text&7m&64&10&1&30&5\\
-    Graphics&1h&128&40&1&30&5\\
-    Symbolic regression& & 40&10 & 1&30&5\\
-    Towers& & 64&10 & 1&30&5\\
-    Regexes& & 64&10 & 1&30&5\\
+    Graphics&1h&128&40&1.5&30&5\\
+    Symbolic regression&2m & 40&10 & 1&30&5\\
+    Towers& 2m& 64&10 & 1&30&5\\
+    Regexes&30m & 64&40 & 1&30&5\\
     \bottomrule     

### 2019-02-15 archive.tex
changed color scheme in cycle figure
better figures for learning curves

@@
 
-An age-old dream within AI is a machine that learns and reasons by
+An old dream within AI is a machine that learns and reasons by
 writing its own programs.  This vision stretches back to the
@@
 hierarchies of code.  We identify this cumulative nesting of
-abstractions as a kind of deep representation
+abstractions as a variety of deep representation
 learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
@@
 
-Because any learning problem can in principle be cast as program
-induction, it is important to delimit our focus.  In contrast to
-computer assisted programming~\cite{solar2008program} or genetic
-programming~\cite{DBLP:books/daglib/0070933}, our goal is not to
-automate software engineering, or to synthesize large bodies of code
-starting from scratch.  Ours is a basic AI goal: capturing the human
-ability to learn to think flexibly and efficiently in new domains ---
-to learn what you need to know about a domain so you don't have to
-solve new problems starting from scratch.  We are focused on problems
-that people solve relatively quickly, once they acquire the relevant
-domain expertise.  These correspond to tasks solved by short programs
---- if you have an expressive DSL.
+Our goal with \system is to engineer a system that develops domain
+expertise in humanlike ways. This involves learning both declarative and
+procedural knowledge, like a DSL and a recognition model, but includes
+other features of the development of expertise:
+humans can become experts in many fields,
+and so we evaluate our algorithm across six different problem domains;
+a human expert doesn't become an expert overnight,
+and needs more than a handful of example problems to learn from,
+but doesn't need millions of examples;
+similarly our algorithm's
+learning trajectory unfolds over a series of wake/sleep cycles
+requiring around a hundred problems per domain.
 
@@
 alternatingly finds programs that solve tasks (Wake --
-Figure~\ref{threeCycles} top); improves its DSL by discovering and
-reusing domain-specific subroutines (Consolidation --
-Figure~\ref{threeCycles} left); and trains a neural network that
-efficiently guides search for programs in the DSL (Dreaming --
-Figure~\ref{threeCycles} right).  The learned DSL acts as a a prior on
-programs likely to solve tasks in the domain, while the neural net
-looks at a specific task and produces a ``posterior'' for programs
-likely to solve that specific task (Figure~\ref{threeCycles} middle).
-The neural network thus functions as a \textbf{recognition model}
-supporting a form of approximate Bayesian program induction, jointly
-trained with a \textbf{generative model} for programs encoded in the
-DSL, in the spirit of the Helmholtz machine~\cite{hinton1995wake}. The
-recognition model ensures that searching for programs remains
-tractable even as the DSL (and hence the search space for programs)
-expands.  The generative model, or DSL, distills out common
-abstractions across programs discovered during waking, growing a
-network of increasingly deep and specialized domain-specific concepts
-(Figure~\ref{exampleDSL}, bottom rows).
-
-We view these wake sleep/cycles as an approximate inference algorithm that starts with a collection of tasks, written $X$, and infers both a program for each task, as well
-as a distribution over programs encoded by a DSL, written
-$\mathcal{D}$.  We equip $\mathcal{D}$ with a learned weight vector
-$\theta$, and together $(\mathcal{D},\theta)$ define a generative
-model over programs (Appendix~\ref{sampleProgram}).  We perform
-maximum a posteriori (MAP) inference of the DSL and programs given
-$X$.  Writing $J$ for the joint probability of $(\mathcal{D},\theta)$
-and $X$, we seek the generative model $\mathcal{D}^*,\theta^*$ solving:
-\begin{align}\label{intractableObjectives}
-\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
-  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta &\qquad
-  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
-\end{align}
-%% where $\probability[x|p]$ scores the likelihood of a task
-%% $x\in X$ given a program $p$.\footnote{For example, for list
-%%   processing, the likelihood is 1 if the program predicts the observed
-%%   outputs on the observed inputs, and 0 otherwise; when learning a
-%%   generative model or probabilistic program, the likelihood is the
-%%   probability of the program sampling the observation.}
-Evaluating Eq.~\ref{intractableObjectives}
-entails marginalizing over the infinite set of all programs -- which is impossible.
-We make a particle-based approximation to Eq.~\ref{intractableObjectives}
-and instead marginalize over a finite \textbf{beam} of programs,
-with one beam per task, collectively written $\left\{\mathcal{B}_x \right\}_{x\in X}$.
-This particle-based approximation is written $\lowerBound (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$
-and acts as a lower bound on the joint density:
-\begin{align}
- J(\mathcal{D},\theta)\geq \lowerBound  (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, where $|\mathcal{B}_x|$ is small}
-\end{align}
-Wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{B}_x \right\}_{x\in X}$  (\textbf{Wake})
-and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
-\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the beams.} Here $(\mathcal{D},\theta)$ is fixed and we
-want to find new programs to add to  the beams so that $\lowerBound$ increases the most.
-$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
-is large, i.e., programs with high posterior probability,
-or programs which are a priori likely and which solve the task.
-We find programs solving a task by
-enumerating programs
-from the DSL in decreasing order of their probability
-under the recognition model,
-and then check if a program $p$ assigns positive probability to a task
-($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
-$\mathcal{B}_x$. We represent programs as polymorphicly typed $\lambda$-calculus expressions,
-an expressive formalism
-including conditionals,
-variables,
-higher-order recursive functions,
-and the ability to define new functions.
-\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed and the problem is to search the discrete space of DSLs and find one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
-and then update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.
-
-Finding programs solving tasks is difficult because of the infinitely
-large, combinatorial search landscape. We ease this difficulty by
-training a neural recognition model, $Q(p |x )$, during the
-\textbf{Dreaming} phase: $Q$ is trained to approximate the posterior
-over programs, $Q(p|x)\approx
-\probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
-Thus training the neural network amortizes the cost of finding
-programs with high posterior probability.
-
-\noindent\textbf{Sleep (Dreaming): tractably maxing $\lowerBound$ w.r.t. the
-  beams.}  Here we train %% a neural network, $q$, to predict a
-%% distribution over programs conditioned on a task. The objective of $q$
-%% is
-$Q(p|x)$ to assign high probability to programs $p$ where
-$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
-into the beams will most increase $\lowerBound$.
+Figure~\ref{threeCycles} top); improves its DSL by analyzing programs
+found during waking (Consolidation -- Figure~\ref{threeCycles} left);
+and trains a neural network that efficiently guides search for
+programs in the DSL (Dreaming -- Figure~\ref{threeCycles} right).  The
+learned DSL acts as a a prior on programs likely to solve tasks in the
+domain, while the neural net looks at a specific task and produces a
+``posterior'' for programs likely to solve that specific task
+(Figure~\ref{threeCycles} middle).  The neural network thus functions
+as a \textbf{recognition model} supporting a form of approximate
+Bayesian program induction, jointly trained with a \textbf{generative
+  model} for programs encoded in the DSL, in the spirit of the
+Helmholtz machine~\cite{hinton1995wake}. The recognition model ensures
+that searching for programs remains tractable even as the DSL (and
+hence the search space for programs) expands.  The generative model,
+or DSL, distills out common abstractions across programs discovered
+during waking, growing a network of increasingly deep and specialized
+domain-specific concepts (Figure~\ref{exampleDSL}, bottom rows).
+
+%% Formally, $EC^2$ takes as input a collection of problems, written $X$, which can be solved by programs, and infers a domain specific language -- written $\mathcal{D}$ -- which functions as a generative model over programs likely to solve problems in the domain, and which we identify as an expert's a priori beliefs about what good solutions look like.
+Alongside this generative model it trains a neural recognition model, in the spirit of the Helmholtz machine or wake-sleep algorithm~\cite{dayan1995helmholtz}.
+The recognition model takes as input a specific problem and quickly predicts an approximate posterior over programs likely to solve that problem.
+Writing $p_x$ for the program solving problem $x\in X$ and $Q(p|x)$ for the distribution predicted by the recognition model, $EC^2$ iteratively (and approximately) solves for
+%% \begin{align}
+%% p_x& = \argmax_p \probability[x|p]\probability[p|\mathcal{D}]\label{wakingObjective}\\
+%% \mathcal{D}& = \argmax_\mathcal{D}\probability\left[\mathcal{D} \right]\prod_{x\in X}\sum_{\substack{p:\\p\text{ found during Explore}}} \probability[p|\mathcal{D}]\label{sleepg}\\
+%% Q(p|x)&\approx \probability[p|x,\mathcal{D}]\label{dreamingObjective}
+%% \end{align}
+
+These wake sleep/cycles function as an approximate inference algorithm
+that observes a collection of tasks, written $X$, and infers both a
+program solving each task, as well as a distribution over programs
+encoded by a DSL, written $\mathcal{D}$.  We equip $\mathcal{D}$ with
+a learned weight vector $\theta$, and together $(\mathcal{D},\theta)$
+define a generative model over programs
+(Appendix~\ref{sampleProgram}).  Writing $Q(p|x)$ for the approximate
+posterior predicted by the recognition model, we iteratively (and
+approximately) solve for
+
+\begin{array}{lr}
+  \begin{array}{l}  p_x = \argmax_p \probability[x|p]\probability[p|\mathcal{D}^*,\theta^*]\end{array}
+    &\text{\emph{Wake}}\\\\
+\begin{array}{rl}  
+\mathcal{D}^* = \argmax_\mathcal{D}&\int\probability\left[\mathcal{D},\phantom{^*}\theta \right]\prod_{x\in X}\sum_{\substack{p\text{ found during waking}}} \probability[x|p]\probability[p|\mathcal{D},\phantom{^*}\theta]\;\text{d}\theta\\
+\theta^* = \argmax_\theta&\phantom{\int}\probability\left[\mathcal{D}^*,\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D}^*,\theta]
+ \end{array}
+&\text{\emph{Consolidation sleep}}\\\\
+\begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
+  &\text{\emph{Dream sleep}}
+\end{array}
+
+\noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).
 
@@
 during earlier cycles,
-growing an increasingly deep learned program representation (`deep program learning').
+growing an increasingly deep learned program representation.
 In tandem the DSL and recognition model
@@
 solves more tasks during waking which then feed into the next DSL.
+
+Waking consists of searching for task-specific programs with high
+posterior probability, or programs which are a priori likely and which
+solve a task.  We find programs solving a task by enumerating
+programs from the DSL in decreasing order of their probability under
+the recognition model, and then check if a program $p$ assigns
+positive probability to a task ($\probability[x|p] > 0$). We represent
+programs as polymorphicly typed $\lambda$-calculus expressions, an
+expressive formalism including conditionals, variables, higher-order
+recursive functions, and the ability to define new functions.
+
 \begin{figure}
   \begin{tikzpicture}[line width=0.4mm]
-    \draw[fill=white!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
+    \draw[fill=teal!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
     \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
@@
     \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
-    \draw[fill=white!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
+    \draw[fill=teal!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
     \node at (12,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
-    \draw[fill=white!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);
+    \draw[fill=teal!5!white] (15,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (15,-13) -- (15,-5);
 
@@
 Mathematically this compression takes the form of
-finding the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
+finding the DSL maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
 We replace this marginal with an AIC approximation~\cite{akaike1998information}
-and marginalize over refactorings of programs in the beams, minimizing the following expression,
+and marginalize over refactorings of programs found during waking, minimizing the following expression,
 which can be interpreted as a kind of compression:
 \begin{equation}
-\underbrace{-\log \probability[\mathcal{D}] + \min_{\theta}\Bigg(-\log
+\underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
 \probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
-  X}\underbrace{-\log \sum_{\substack{p:\\
-      \exists p'\in
-  \mathcal{B}_x: p\text{ refactors into }p'}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
+  X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
 \label{AIC}
@@
 
-A structure called a \textbf{version space
-  algebra}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
-neatly resolves this exponential growth.  A version space is a
+To resolve this exponential growth we introduce a new data structure
+combining ideas from version space
+algebras~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}
+and equivalence graphs~\cite{tate2009equality}. A version space is a
 tree-shaped data structure that compactly represents a large set of
 programs and supports efficient set operations like union,
-intersection, and membership checking.  In
-Appendix~\ref{appendixVersion}, we give a dynamic program that takes
-as input a program and then outputs a version space containing its
-refactorings. Figure~\ref{mapFactor}B diagrams a subtree of a
+intersection, and membership checking, while equivalence graphs are
+data structures that track semantic equivalences between program
+subexpressions.  In Appendix~\ref{appendixVersion}, we give a dynamic
+program that takes as input a program and then outputs a version space
+containing its refactorings while tracking semantically equivalent
+subexpressions. Figure~\ref{mapFactor}B diagrams a subtree of a
 version space containing refactorings of a small program. Our
@@
 a version space with $10^6$ nodes, which encodes $10^{14}$
-refactorings. Appendix~\ref{appendixCompression} specifies how we combine this probabilistic and symbolic machinery to update the DSL.
+refactorings. Appendix~\ref{appendixCompression} specifies how we
+combine this probabilistic and symbolic machinery to update the DSL.
 At a high level, our approach is to search locally through the space
-of DSLs, proposing small changes until Eq.~\ref{AIC}
-fails to decrease.
+of DSLs, proposing small changes until Eq.~\ref{AIC} fails to
+decrease.
 
@@
 a DSL for arithmetic should disallow adding zero, or force
-right-associative addition.  These restrictions break symmetries in
-the space of programs.  A bigram parameterization of the recognition
+right-associative addition. A bigram parameterization of the recognition
 model, combined with the $\mathcal{L}^{\text{MAP}}$ training
-objective, interact in a way that breaks symmetries in the program
-space, allowing the agent to more efficiently explore the space of
+objective, interact in a way that breaks symmetries like these, allowing the agent to more efficiently explore the space of
 programs.  This interaction occurs because the bigram parameterization
@@
 $n^{\text{th}}$ largest element of a list, which it finally uses to
-learn how to sort a list of numbers (Figure~\ref{initialExampleDSL}).
+solve a task involving sorting a list of numbers (Figure~\ref{initialExampleDSL}).
 This incremental, modular learning of deep hierarchies of DSL
@@
 rotational symmetry, which it models by incorporating a higher-order
-function into its DSL (Figure~\ref{logoPrimitives}, right).  This
-abstraction comes from jointly compressing programs for all of the
-training images into its DSL.
+function into its DSL (Figure~\ref{logoPrimitives}, right).  %% This
+%% abstraction comes from jointly compressing programs for all of the
+%% training images into its DSL.
 
@@
   \caption{Three domains where the agent infers a program from visual input.
-\textbf{Left}: 16 (out of 145) LOGO graphics tasks.
+\textbf{Left}: 16 (out of 160) LOGO graphics tasks.
 Agent writes a program controlling a `pen' that draws the target picture.
@@
   \begin{tabular}{ccc}
-    \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
-    \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
-    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
-    \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
-    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
-    \includegraphics[width = 5cm]{figures/regexLearningCurve_batch.png}
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/text_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/text_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/logo_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/logo_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/rational_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/rational_time.png}
+    \end{tabular}\\\\\\
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/list_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/list_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/tower_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/tower_time.png}
+    \end{tabular}
   \end{tabular}
   \caption{Test set performance across wake/sleep iterations.
+Each curve is a run with a different random seed.
+Teal: Full model.
+Orange: Dreaming only (no DSL learning).
+Purple: Consolidation only (no recognition model).
+Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks).}\label{learningCurves}
+\end{figure}
+
+%% \begin{figure}
+%%   \begin{tabular}{ccc}
+%%     \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
+%%     \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
+%%     \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
+%%     \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
+%%     \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
+%%     \includegraphics[width = 5cm]{figures/regexLearningCurve_batch.png}
+%%   \end{tabular}
+%%   \caption{Test set performance across wake/sleep iterations.
 Orange: \systemEnding.
@@
 Shaded region is $\pm 1$ stddev search time.}\label{learningCurves}
-\end{figure}
+%% \end{figure}
 
@@
 
+\subsection{Probabilistic Formulation of \system}\label{probabilisticAppendix}
+Our objective is to infer the maximum a posteriori DSL $\mathcal{D}$ and
+parameters $\theta$. Writing $J$ for the joint probability of $(\mathcal{D},\theta)$,
+this corresponds to solving
+\begin{align}\label{intractableObjectives}
+\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]&\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
+  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta& \qquad
+  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
+\end{align}
+where $\probability[x|p]$ scores the likelihood of a task
+$x\in X$ given a program $p$.\footnote{For example, for list
+  processing, the likelihood is 1 if the program predicts the observed
+  outputs on the observed inputs, and 0 otherwise; when learning a
+  generative model or probabilistic program, the likelihood is the
+  probability of the program sampling the observation.}
+
+Evaluating Eq.~\ref{intractableObjectives}
+entails marginalizing over the infinite set of all programs -- which is impossible.
+We make a particle-based approximation to Eq.~\ref{intractableObjectives}
+and instead marginalize over a finite \textbf{beam} of programs,
+with one beam per task, collectively written $\left\{\mathcal{B}_x \right\}_{x\in X}$.
+This particle-based approximation is written $\lowerBound (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$
+and acts as a lower bound on the joint density:
+\begin{align}
+ J(\mathcal{D},\theta)\geq \lowerBound  (\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, where $|\mathcal{B}_x|$ is small}
+\end{align}
+In all of our experiments we set the maximum beam size $|\mathcal{B}_x|$ to 5.
+
+Wake and sleep cycles correspond to alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{B}_x \right\}_{x\in X}$  (\textbf{Wake})
+and $(\mathcal{D},\theta)$ (\textbf{Consolidation}):
+\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the beams.} Here $(\mathcal{D},\theta)$ is fixed and we
+want to find new programs to add to  the beams so that $\lowerBound$ increases the most.
+$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]\propto\probability[p|x,\mathcal{D},\theta]$ 
+is large, i.e., programs with high posterior probability,
+which is the search objective during waking.
+
+%% or programs which are a priori likely and which solve the task.
+%% We find programs solving a task by
+%% enumerating programs
+%% from the DSL in decreasing order of their probability
+%% under the recognition model,
+%% and then check if a program $p$ assigns positive probability to a task
+%% ($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
+%% $\mathcal{B}_x$. We represent programs as polymorphicly typed $\lambda$-calculus expressions,
+%% an expressive formalism
+%% including conditionals,
+%% variables,
+%% higher-order recursive functions,
+%% and the ability to define new functions.
+\\\noindent \textbf{Sleep (Consolidation): Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{B}_x \right\}_{x\in X}$ is held fixed and the problem is to search the discrete space of DSLs and find one maximizing $\int \lowerBound \;\mathrm{d}\theta$,
+and then update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\})$.
+
+Finding programs solving tasks is difficult because of the infinitely
+large, combinatorial search landscape. We ease this difficulty by
+training a neural recognition model, $Q(p |x )$, during the
+\textbf{Dreaming} phase: $Q$ is trained to approximate the posterior
+over programs, $Q(p|x)\approx
+\probability[p|x,\mathcal{D}]\propto\probability[x|p]\probability[p|\mathcal{D}]$.
+Thus training the neural network amortizes the cost of finding
+programs with high posterior probability.
+
+\noindent\textbf{Sleep (Dreaming): tractably maxing $\lowerBound$ w.r.t. the
+  beams.}  Here we train %% a neural network, $q$, to predict a
+%% distribution over programs conditioned on a task. The objective of $q$
+%% is
+$Q(p|x)$ to assign high probability to programs $p$ where
+$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
+into the beams will most increase $\lowerBound$.
+
 \subsection{\system pseudocode}\label{systemPseudocode}
 
-Algorithm~\ref{theWholeThing} specifies how we integrate wake and sleep cycles.
+Algorithm~\ref{theWholeThing} specifies how we integrate wake and
+sleep cycles.
 \begin{algorithm}
@@
  This integrates two toolkits:
- the machinery of version spaces (Appendix~\ref{appendixVersion})
- and a probabilistic objective favoring compressive DSLs.
+ the machinery of version spaces and equivalence graphs (Appendix~\ref{appendixVersion})
+ along with a probabilistic objective favoring compressive DSLs.
  The functions $I\beta(\cdot )$ and \textsc{refactor} construct
@@
 
+\subsubsection{Tracking equivalences}
+
 %% First observe that

### 2019-02-21 archive.tex
updated learning curves

@@
 
-An old dream within AI is a machine that learns and reasons by
-writing its own programs.  This vision stretches back to the
-1960's~\cite{solomonoff1964formal} and and is central to much of
-what it would take to build machines that learn and think like humans.
+An old dream within AI is a machine that learns and reasons by writing
+its own programs.  This vision stretches back to the
+1960's~\cite{solomonoff1964formal} and, if fully realized, could bring
+us much closer to machines that learn and think like humans.
 Computational models of cognition often explain the flexibility and
@@
 text~\cite{gulwani2011automating} and numerical
-sequences~\cite{balog2016deepcoder}, and reasoning over common sense
+sequences~\cite{balog2016deepcoder}, robot motion
+planning~\cite{devlin2017neural}, and reasoning over common sense
 knowledge bases~\cite{muggleton2015meta}.  These systems work in
@@
 concepts.  This joint learning problem models two complementary
-notions of domain expertise: (1) domain experts have at their disposal
+notions of domain expertise: domain experts have at their disposal
 a powerful, yet specialized repertoire of concepts and abstractions
-(analogous to the DSL) while also (2) having accurate intuitions about
+(analogous to the DSL) while also having accurate intuitions about
 when and how to use those concepts to solve new problems.
@@
 refer to as \textbf{dreaming}, improves the agents knowledge of how to
-write programs by training a neural network to help search for
+write code by training a neural network to help search for
 programs. The neural net is trained on replayed experiences as well as
@@
     \begin{tabular}{c}
-      \includegraphics[width = 5cm]{figures/learningCurves/list_hits.png}\\
-      \includegraphics[width = 5cm]{figures/learningCurves/list_time.png}
+      \includegraphics[width = 5cm]{figures/learningCurves/list_hard_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/list_hard_time.png}
     \end{tabular}&
@@
       \includegraphics[width = 5cm]{figures/learningCurves/tower_time.png}
-    \end{tabular}
+    \end{tabular}&
+
+    \includegraphics[width = 5cm]{figures/learningCurves/regular}
   \end{tabular}
@@
 spectrum of tasks.
+
+\subsection{\system and the Exploration-Compression family of algorithms}
+
+Our work sits within the Exploration-Compression (EC) family of
+algorithms.  EC~\cite{Dechter:2013:BLV:2540128.2540316} is a program
+induction framework where an agent alternates between searching, or
+`exploring', the space of programs, and then updating its search
+procedure by compressing programs found during exploration. \system
+grew directly out of research on EC-style systems and both of our
+sleep phases can be interpreted as a kind of compression:
+consolidation aims to compactly refactor code, while the recognition
+model aims to encode a program in as few bits as possible, conditioned
+on a task. Other offshoots of EC include our previous work,
+EC$^2$~\cite{ecc}, the direct predecessor of \systemEnding; and the
+neurosymbolic framework in~\cite{lazaro2019beyond}. Closely aligned ideas
+go much further back~\cite{solomonoff1989system,schmidhuber2004optimal}.
 

### 2019-02-27 archive.tex
added a bunch of logo figures

@@
 model aims to encode a program in as few bits as possible, conditioned
-on a task. Other offshoots of EC include our previous work,
-EC$^2$~\cite{ecc}, the direct predecessor of \systemEnding; and the
-neurosymbolic framework in~\cite{lazaro2019beyond}. Closely aligned ideas
-go much further back~\cite{solomonoff1989system,schmidhuber2004optimal}.
+on a task. Our previous work, EC$^2$~\cite{ecc}, introduced
+neurally-guided search into the EC framework and served as the
+starting point for \systemEnding.  Our work here extends EC$^2$ by (1)
+introducing a new refactoring compression algorithm, (2) enriching the
+neural network with a new bigram-over-trees parameterization and loss
+function that together learn to break symmetries, accelerating program
+search during waking, and (3) demonstrating how this family of
+approaches can be applied to planning and generative modeling
+problems.  Other offshoots of EC include the neurosymbolic framework
+in~\cite{lazaro2019beyond} and the hierarchical Bayesian program
+learner in~\cite{DBLP:conf/icml/LiangJK10}. Closely aligned ideas go
+even further back~\cite{solomonoff1989system,schmidhuber2004optimal}.
 

### 2019-03-17 archive.tex
regex

@@
 
+\usepackage{longtable}
 \usepackage[section]{placeins}
@@
 
-\usepackage{verbatim}
+%\usepackage{verbatim}
  \usepackage{booktabs}
@@
 
+Human knowledge, both in children and in adults, seems to be organized
+into domains of expertise.  Some of us develop competence in cooking,
+calculus, tennis, drawing pictures, or writing software, and virtually
+every child masters natural language, motor control, and intuitive
+theories of kinship, taxonomy, and physics. Over short timescales,
+people can solve new problems within their domain of expertise, and
+over longer time scales, they can grow into domain experts for 
+new classes of problems.  A long-term goal for machine intelligence is
+to build an AI that captures these two human abilities: both solving
+new problems, and expanding the domains of problems that can be solved in
+the first place.
+
+We take the stance that an artificial agent with these humanlike
+abilities needs at least two ingredients.  First it must acquire the
+underlying system of domain-specific concepts.  For drawing pictures
+these concepts include polygons, spirals, arcs, and symmetries; for
+physics these concepts include dot products, vector fields, and
+conservation laws.  Second it must learn to quickly deploy these
+concepts efficiently on new problems: at a glance, human domain
+experts can intuit which compositions of concepts are likely to solve
+the task at hand, even before they begin searching for a solution.
+
+We introduce a computational model that jointly acquires these two
+kinds of domain knowledge.  Because solutions to many kinds of
+problems can often be described as some kind of
+program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
+our system solves a specific problem by searching for a program.  It
+gradually grows its domain-specific knowledge for a class of problems
+by assembling a library of code containing concepts useful for the
+domain, and by training a neural network to quickly infer, for a
+specific problem, what code concepts from its library are likely to
+solve it.  We first investigate our model within classic program
+synthesis domains for manipulating sequences of numbers and text, and
+then consider visual and creative programs for drawing pictures and
+building towers out of toy blocks, altogether considering both
+deterministic and probabilistic programs that act both generatively
+(e.g., producing an artifact like an image or plan) and conditionally
+(e.g., mapping inputs to outputs).
+
+We call our system `DreamCoder' because it works through a novel kind
+of wake/sleep or `dream' learning~\cite{hinton1995wake}, iterating
+through a wake cycle -- where it solves problems by writing programs
+-- and a pair of sleep cycles, both of which are loosely biologically
+inspired. The first sleep cycle, which we refer to as
+\textbf{consolidation}, grows its library of code by replying
+experiences from waking and consolidating them into new code
+abstractions.  This cycle is inspired by the formation of abstractions
+during sleep memory consolidation~\cite{DUDAI201520}.  The second
+sleep cycle, which we refer to as \textbf{dreaming}, improves the
+agents knowledge of how to write code by training a neural network to
+help search for programs. The neural net is trained on replayed
+experiences as well as `fantasies', or samples, from the learned
+library.  These two kinds of dreams are inspired by the distinct
+episodic replay and hallucination components of dream
+sleep~\cite{fosse2003dreaming}.
+
+\system builds on several generations of research in AI, program
+synthesis, and cognitive science, with program induction being one of
+the oldest theoretical frameworks for concept learning within
+artificial intelligence~\cite{solomonoff1964formal}, and the
+conceptually allied `Language of Thought Hypothesis' being almost as
+old~\cite{fodor1975language}. Recent neural program synthesis systems
+pair a fixed programming language to a learned neural network that
+guides program
+search~\cite{spiral,balog2016deepcoder,devlin2017neural}, while recent
+symbolic AI research has developed frameworks for inferring 
+reusable pieces of code shared in a multitask
+setting~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
+A main goal of this work is to combine and refine these techniques
+with the intention of building agents that, like humans, develop
+domain expertise for new classes of problems, motivated by models from
+cognitive science that explain the flexibility and generality of human
+thinking in terms of program-like mental representations, referred to
+as a `language of thought'~\cite{piantadosi2011learning}.
+
+Most prior program synthesis work assumes a fixed library of code
+components~\cite{jha2010oracle} or a hand-engineered domain-specific
+language (a `DSL'~\cite{polozov2015flashmeta,gulwani2011automating}).
+In the vocabulary of program synthesis, our approach is to learn an
+increasingly finely-tuned DSL (during consolidation) while jointly
+learning to deploy it efficiently on new problems (during dreaming).
+Each wake/sleep cycle creates new DSL components that build on
+components learned in earlier sleep cycles, growing a DSL with nested
+hierarchies of code.  We identify this cumulative nesting of
+abstractions as a variety of deep representation
+learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
+subset of these learned networks (the DSL). For example, the system
+learns to sort sequences of numbers by invoking a DSL component 4
+layers deep, or draws the leftmost pairs of images in
+Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
+refer to \system as an instance of `deep program learning'.
+
+\section{Old introduction}
+
 An old dream within AI is a machine that learns and reasons by writing
@@
     \newcommand{\invention}[4]{
-      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,draw,rounded corners,align=center] at (#2) {
+      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,ellipse,draw,rounded corners,align=center] at (#2) {
         \footnotesize #3
@@
 programs, and the likelihood model's BIC penalty.
+
+\subsubsection{Generative Modeling of Text}
+We investigate few-shot learning of generative models by tasking our
+agent with inferring a probabilistic regular expression from a small
+number (5) of strings - for example, observing the strings \$1.20 and \$9.42
+and predicting the regex \$d
+
+\begin{tabular}{ccc}
+  \toprule   Observed text&Inferred regex&Held out testing text\\\midrule
+
+  \end{tabular}
 
@@
 
+\subsection{Results on held out text patterns}
+%\makeatletter
+\renewcommand{\verbatim@font}{\ttfamily\tiny} 
+\input{regex.tex}
+
 \end{document}

### 2019-04-07 archive.tex
updated some learning curves

--- 2019-02-27 archive.tex
@@
 \begin{document}
-
-\begin{figure}
-  \includegraphics[width = \textwidth]{figures/DeepLogo.pdf}
-  \caption{\textbf{Bottom}: Learned graphics programing DSL.
-Each circle encloses a different learned component; arrows point from a component to other components that use it.
-\textbf{Top}: 3 tasks, and below them the programs that solve them.
-Yellow/Blue/Green highlighting indicates which DSL components are used to solve which tasks.}
-  \end{figure}
-
-\begin{figure}
-  \begin{tikzpicture}
-    \renewcommand{\baselinestretch}{0.5}
-    \newcommand{\inventionColor}{orange}
-    \newcommand{\name}[2]{
-      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
-    }
-    \newcommand{\noArguments}[1]{
-      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
-    }
-
-    \newcommand{\horizontalSpacing}{3.8cm}
-    \newcommand{\verticalSpacing}{2cm}
-
-%    \draw (0,0) rectangle (10,5);
-    \node(t1)[align=left] at (1,-1) {
-      \footnotesize +106 769-43$\to $106.769.43\\
-      \footnotesize +83 973-831$\to $83.973.831\\
-      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{\inventionColor}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
-      \footnotesize \hspace{1.25cm}\mcode{(${\color{\inventionColor}f_0}$ \codechar{.} \codechar{ }}\\
-      \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
-    };
-    \node(t2)[align=left,anchor=center] at ([xshift=\horizontalSpacing]t1.center) {
-      \footnotesize Dream Coder$\to$Dr. Dream\\
-      \footnotesize Andy Cencici$\to$Dr. Andy\\
-      \footnotesize Jan Kotas$\to$Dr. Jan\\
-      \footnotesize Nancy Jones$\to$Dr. Nancy\\
-      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_3}$ (${\color{\inventionColor}f_4}$ \codechar{ } s))}
-    };
-    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
-      \footnotesize Dream (Coder)$\to$Coder\\
-      \footnotesize 1 (2) 3 4$\to$2\\
-      \footnotesize the (dog) ran$\to$ran\\
-      \footnotesize +817(33)99$\to$33\\
-      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_6}$ \codechar{(} \codechar{)} s)}
-    };
-
-    \draw (-2,-2.5) -- (15,-2.5);
-    \node[align=left,rotate=90] at (-1.5,-1.3) {Tasks and\\\\\texttt{Programs}};
-
-    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)\\\\System discovers subroutines in {\color{\inventionColor}orange}};
-
-    \newcommand{\invention}[4]{
-      \node(#1)[inner sep=1,anchor=center,text width=2cm,draw,rounded corners,align=center] at (#2) {
-        \footnotesize #3
-      };
-      \node[anchor=north] at ([yshift=3.5pt]#1.south) {#4};
-    }
-    \newcommand{\vinvention}[4]{
-      \invention{#1}{[yshift=-\verticalSpacing]#2.center}{#3}{#4}
-    }
-    \newcommand{\hinvention}[4]{
-      \invention{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
-    }
-
-    \invention{substitution}{1,-4}{
-      ${\color{\inventionColor}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
-       \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
-    }{substitute chars}
-    \vinvention{cat}{substitution}{
-      ${\color{\inventionColor}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
-      \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
-    }{concat strings}
-    \vinvention{looping}{cat}{
-       ${\color{\inventionColor}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
-      \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
-    }{loop until char}
-
-    \hinvention{constant}{substitution}{
-      ${\color{\inventionColor}f_3}($\mcode{s}$) \,=\, $\\
-      \mcode{(${\color{\inventionColor}f_1}$ s string)}
-    }{prepend constant}
-    \vinvention{drop}{constant}{
-      ${\color{\inventionColor}f_4}($\mcode{c}$) \,=\, $\\
-      \mcode{(${\color{\inventionColor}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
-    }{prefix}
-    \vinvention{suffix}{drop}{
-      ${\color{\inventionColor}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
-      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
-    }{suffix}
-
-    \hinvention{delimited}{constant}{
-      \mcode{${\color{\inventionColor}f_6}($x$,$y$,$s$)\,=\,$}\\
-      \mcode{(${\color{\inventionColor}f_4}$ x (${\color{\inventionColor}f_5}$ y c))}
-    }{delimited substring}
-    \draw[->] (delimited.west) -- (drop.east);
-    \draw[->] (delimited.west) -- (suffix.east);
-
-    \vinvention{appendSingle}{delimited}{
-      \mcode{${\color{\inventionColor}f_7}($c$,$s$)\,=\,$}\\
-      \mcode{(${\color{\inventionColor}f_1}$ (cons c nil) s)}}{append character}
-
-    \vinvention{surround}{appendSingle}{
-      \mcode{${\color{\inventionColor}f_8}($s$)\,=\,$}\\
-      \mcode{(cons \codechar{(} (${\color{\inventionColor}f_7}$ \codechar{)} s))}}{surround in parens}
-    \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
-    \draw[->] (appendSingle.west) to[in=40,out=120] ([yshift=0.25cm]cat.east);
-
-    \draw[->] (drop.west) -- (looping.east);
-    \draw[->] (constant.west) -- (cat.east);
-
-    \draw[->] ([xshift=-0.5cm]t2.south) -- (constant.north);
-    \draw[->] ([xshift=0.3cm]t2.south) to[in=30,out=-30] ([xshift=1cm]drop.north);
-
-    \draw[->] ([yshift=0.2cm,xshift=-0.2cm]t1.south) to[in=100,out=200] (substitution.north);
-    \draw[->] ([yshift=0.5cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
-    \draw[->] ([xshift=-0.5cm]t3.south) -- (delimited.north);
-
-    \begin{scope}[shift={(0,-15)}]
-      \invention{filter}{0,0}{
-        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
-        }{higher-order `filter'}
-
-      \vinvention{three}{filter}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
-      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
-      \hinvention{single}{filter}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
-      \vinvention{for}{single}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
-      \vinvention{Max}{for}{\name{5}{l}\\\mcode{%
-          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
-      \hinvention{largest}{single}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
-      \vinvention{count}{largest}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
-      \vinvention{weird}{count}{\name{8}{p$,$l}\mcode{
-          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
-                 {Higher-order incrementer}
-
-      \draw[->] (for.west) -- (three.east);
-      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
-      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
-      \draw[->] (Max.west) -- (filter.east);
-      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);
-
-      \node(t1)[align=left,anchor=center] at (0,3) {
-        \mcode{[1 4 5 8]$\to$[5 8]}\\
-        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
-        \mcode{[3 4 5 6]$\to$[5 6]}\\
-        \mcode{[6 5 4 3]$\to$[6 5]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}
-      };
-      \node(t2)[align=left] at ([xshift=\horizontalSpacing]t1.center) {
-        \mcode{[9 2 4]$\to$[4 2 9]}\\
-        \mcode{[4 3]$\to$[3 4]}\\
-        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
-        \mcode{[]$\to$[]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}
-      };
-      \node(t3)[align=left] at ([xshift=\horizontalSpacing]t2.center) {
-        \mcode{[9 2 3]$\to$[2 3 9]}\\
-        \mcode{[2 1]$\to$[1 2]}\\
-        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
-        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}
-      };
-
-      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]t1.south) to[out=250] (filter.north);
-      \draw[->] ([yshift=0.0cm,xshift=0.3cm]t1.south) to[out=350,in=100] (for.west);
-      \draw[->] ([xshift=-0.7cm]t2.south) -- (single.north);
-      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]t3.south) to[out=240] (largest.north);
-
-    \end{scope}
-
-    \end{tikzpicture}
-\end{figure}
 
@@
 
-An old dream within AI is a machine that learns and reasons by writing
-its own programs.  This vision stretches back to the
-1960's~\cite{solomonoff1964formal} and, if fully realized, could bring
-us much closer to machines that learn and think like humans.
-Computational models of cognition often explain the flexibility and
-richness of human thinking in terms of program learning: from everyday
-thinking and problem solving (motor program induction as an account of
-recognition and generation of handwriting and
-speech~\cite{lake2015human}; functional programs as a model of natural
-language semantics~\cite{SOMETHING}) to learning problems that unfold
-over longer developmental time scales: the child's acquisition of
-intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
-natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
-just a few.  An outstanding challenge, however, is to engineer
-program-learners that display the same level of domain-generality as
-the humans they are meant to model.
-
-Recent program-learning systems developed within the AI and machine
-learning community are impressive along many dimensions, authoring
-programs for problem domains like drawing
-pictures~\cite{spiral,ellis2017learning}, transforming
-text~\cite{gulwani2011automating} and numerical
-sequences~\cite{balog2016deepcoder}, robot motion
-planning~\cite{devlin2017neural}, and reasoning over common sense
-knowledge bases~\cite{muggleton2015meta}.  These systems work in
-different ways, but typically hinge upon a carefully hand-engineered
-Domain Specific Language (DSL).  The DSL restricts the space of
-programs to contain the kinds of concepts needed for one specific
-domain.  For example, a picture-drawing DSL could include concepts
-like circles and spirals, and a DSL for numerical sequences could
-include sorting and reversing lists of numbers.  Modern systems also
-learn how to efficiently deploy the DSL on new
-problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
-unlike human learners -- do not discover the underlying system of
-concepts needed to navigate the domain.
-
-We contribute a program-induction system that learns the
-domain-specific concepts (DSL) while jointly learning how to use those
-concepts.  This joint learning problem models two complementary
-notions of domain expertise: domain experts have at their disposal
-a powerful, yet specialized repertoire of concepts and abstractions
-(analogous to the DSL) while also having accurate intuitions about
-when and how to use those concepts to solve new problems.
-Representative domains,
-along with DSLs we learn for them,
-are shown in Figure~\ref{exampleDSL}.
-
-We call our system `DreamCoder' because it acquires these two kinds of
-expertise through a novel kind of wake/sleep or `dream'
-learning~\cite{hinton1995wake}, iterating through a wake cycle --
-where it solves problems by writing programs -- and a pair of sleep
-cycles, both of which are loosely biologically inspired by actual
-sleep.  The first sleep cycle, which we refer to as
-\textbf{consolidation}, grows the DSL by replying experiences from
-waking and consolidating them into new code abstractions.  This cycle
-is inspired by the formation of abstractions during sleep memory
-consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
-refer to as \textbf{dreaming}, improves the agents knowledge of how to
-write code by training a neural network to help search for
-programs. The neural net is trained on replayed experiences as well as
-`fantasies', or samples, from the DSL.  These two kinds of dreams are
-inspired by the distinct episodic replay and hallucination components
-of dream sleep~\cite{fosse2003dreaming}.
-
-\begin{comment}
-This dream-learning architecture brings together two lines of prior work,
-both of which have been separately influential within
-artificial intelligence.
-One line of work
-considers the problem of learning new concepts, abstractions, or
-`options' from
-experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
-while the other line of work considers the problem of learning how to
-deploy those concepts
-efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
-goal with DreamCoder is to show that the combination of these ideas is
-uniquely powerful, and pushes us toward program-writing systems that,
-like human learners, autonomously acquire the expertise needed to
-navigate a new domain of problems.
-\end{comment}
-
-Each wake/sleep cycle creates new DSL components that build on
-components learned in earlier sleep cycles, growing a DSL with nested 
-hierarchies of code.  We identify this cumulative nesting of
+What would it take to build a machine that can learn everything a
+person does over their lifetime?
+Although this goal will remain distant for the foreseeable future,
+we know that a general learning system like this would need to be able to acquire many different kinds of expertise.
+Virtually every child becomes an expert in natural language, motor control, and social interaction, while learning rich
+intuitive theories of kinship, taxonomy and physics. Many of those
+children will grow into adults with competence in cooking, calculus, tennis, drawing
+pictures, or writing software.
+Expertise -- the ability to quickly solve new problems in a domain -- is crucial.
+Human learning goes far beyond just memorizing a large set of facts or routines,
+but involves higher-order skills,
+like learning to learn new concepts or
+learning to solve new types of problems more quickly and effectively.
+Despite great advances in machine learning,
+we are still far from
+an AI with these abilities.
+The modern AI toolkit has given us machines that learn to play challenging games at superhuman levels,
+but cannot quickly transfer to similar games like a person would;
+or, which can learn to generate convincing English prose,
+but which do not then learn to analyze many different languages,
+like a field linguist would.
+Here, we contribute a computational model that takes a small step toward the goal of building machines that
+grow into domain experts,
+looking at much simpler kinds of problems that humans can learn to solve and indeed become experts in.
+
+Our model is structured
+around the hypothesis that
+domain expertise
+consists primarily of two ingredients.
+First, domain experts have 
+explicit, declarative concepts that are powerful yet finely-tuned.
+A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
+a physicist has concepts like dot products, vector fields, and conservation laws;
+and an architect has concepts like arches, supports, and bridges.
+Second, experts have implicit procedural skill in deploying those concepts
+to quickly solve new problems:
+at a glance, human domain
+experts can intuit which compositions of concepts are likely to solve
+the task at hand, even before searching for a solution.
+Studies of human expertise
+find that,
+compared to novices,
+experts can quickly categorize and classify problems based on the ``deep structure''
+of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
+In short, we take the stance that expertise means both
+having the right concepts, and being able to quickly see how those concepts can be composed into a solution.
+
+Because solutions to many kinds of
+problems can often be described as some kind of
+program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
+our model solves a specific problem by synthesizing a program solving it. It
+gradually grows its domain-specific knowledge for a class of problems
+by assembling a library of code containing concepts useful for the
+domain, and by training a neural network to quickly infer, for a
+specific problem, what code in its library is likely to
+solve it. We think of these two learned components as analogous
+to the explicit concepts and
+implicit procedural skill that human experts develop.
+We will first investigate our model within classic program
+synthesis domains for manipulating sequences of numbers and text, and
+then consider visual and creative programs for drawing pictures and
+building towers out of toy blocks, and finally consider programs for basic kinds of equation discovery and programming language design, altogether considering both
+deterministic and probabilistic programs that act both generatively
+(e.g., producing an artifact like an image or plan) and conditionally
+(e.g., mapping inputs to outputs).
+
+Our model iteratively creates new library routines that build on
+concepts learned earlier in its learning trajectory, growing a library with nested
+hierarchies of code.  We think of this cumulative nesting of
 abstractions as a variety of deep representation
 learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
-subset of these learned networks (the DSL). For example, the system
-learns to sort sequences of numbers by invoking a DSL component 4
+subset of these learned networks (the library). For example, the model
+learns to sort sequences of numbers by invoking a library component 4
 layers deep, or draws the leftmost pairs of images in
 Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
-refer to \system as an instance of `deep program learning'.
-
-Our goal with \system is to engineer a system that develops domain
-expertise in humanlike ways. This involves learning both declarative and
-procedural knowledge, like a DSL and a recognition model, but includes
-other features of the development of expertise:
-humans can become experts in many fields,
-and so we evaluate our algorithm across six different problem domains;
-a human expert doesn't become an expert overnight,
-and needs more than a handful of example problems to learn from,
-but doesn't need millions of examples;
-similarly our algorithm's
-learning trajectory unfolds over a series of wake/sleep cycles
-requiring around a hundred problems per domain.
-
-\begin{figure}\centering
-  \begin{tikzpicture}
-    \renewcommand{\baselinestretch}{0.5}
-    \newcommand{\inventionColor}{orange}
-
-    \newcommand{\name}[2]{
-      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
-    }
-    \newcommand{\noArguments}[1]{
-      \mcode{{\color{\inventionColor}$f_{#1}$}$\,=\,$}
-    }
-    \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
-    \newcommand{\horizontalSpacing}{4.2cm}
-    \newcommand{\verticalSpacing}{2cm}
-
-    \newcommand{\invention}[4]{
-      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,draw,rounded corners,align=center] at (#2) {
-        \footnotesize #3
-      };
-      \node[anchor=north,align=center] at ([yshift=3.5pt]#1.south) {#4};
-    }
-    \newcommand{\vinvention}[4]{
-      \invention{#1}{[yshift=-\verticalSpacing]#2.south}{#3}{#4}
-    }
-    \newcommand{\hinvention}[4]{
-      \invention{#1}{[xshift=\horizontalSpacing]#2.south}{#3}{#4}
-    }
-    \newcommand{\exampleFunction}[2]{
-      \includegraphics[width = 0.5cm]{#1}, \includegraphics[width = 0.5cm]{#2}
-    }
-    \newcommand{\task}[4]{
-      \node[align=left,anchor=center](t#1) at (#2) {#3};
-      \node[align=left](p#1) at ([yshift=-0.75cm]t#1.south) {#4};
-    }
-    \newcommand{\nextTask}[4]{
-      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
-    }
-
-    \begin{scope}[shift={(0,8)}]
-      \task{1}{2,3}{
-        \textsc{Reverse}\\
-        \mcode{[9 2 4]$\to$[4 2 9]}\\
-        \mcode{[4 3]$\to$[3 4]}\\
-        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
-        \mcode{[]$\to$[]}
-      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}}
-      \nextTask{2}{t1}{
-        \textsc{Take if $\geq 5$}\\
-        \mcode{[1 4 5 8]$\to$[5 8]}\\
-        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
-        \mcode{[3 4 5 6]$\to$[5 6]}\\
-        \mcode{[6 5 4 3]$\to$[6 5]}
-      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}}
-
-      \nextTask{3}{t2}{
-        \textsc{Sort}\\
-        \mcode{[9 2 3]$\to$[2 3 9]}\\
-        \mcode{[2 1]$\to$[1 2]}\\
-        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
-        \mcode{[9 8 9]$\to$[8 9 9]}
-      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}}
-
-      %% \vinvention{filter}{p1}{
-      %%   \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
-      %% }{higher-order `filter'}
-      \vinvention{single}{p1}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
-
-      \vinvention{three}{single}{\footnotesize \noArguments{1}\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
-      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
-      \hinvention{filter}{single}{
-        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
-      }{higher-order `filter'}
-      \vinvention{for}{filter}{\noArguments{4}\\\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
-      \vinvention{weird}{for}{\name{8}{p$,$l}\mcode{
-          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0)))}}{Higher-order incrementer}                 
-      \hinvention{largest}{filter}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
-      \vinvention{Max}{largest}{\name{5}{l}\\\mcode{%
-          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
-      \vinvention{count}{Max}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
-
-      \node[rotate=90,align=center] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
-      \node(programLabel)[rotate=90,align=center] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
-      \node[rotate=90,align=center] at ([xshift=-2cm]three.center) {\textbf{\textsc{Domain Specific Language}}\\
-        \small Learned routines in {\color{orange}orange}};
-      \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: List Processing}}};
-
-      \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
-    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};
-
-      \draw[->] (for.west) -- (three.east);
-      \draw[->] (largest.east) to[out=0,in=0] (Max.east);
-      \draw[->] ([yshift=-0.cm]largest.west) -- ([yshift=0.2cm]filter.east);
-      \draw[->] (Max.west) -- (filter.east);
-      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);
-
-      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]p2.south) to[out=250] (filter.north);
-      \draw[->] ([yshift=0.0cm,xshift=0.3cm]p2.south) to[out=350,in=0] (for.east);
-      \draw[->] ([xshift=-0.7cm]p1.south) -- (single.north);
-      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]p3.south) to[out=240] (largest.north);
-
-    \end{scope}
-
-    \renewcommand{\task}[4]{
-      \node(t#1) at (#2) {\includegraphics[width = 2cm]{#3}};
-      \node(p#1) at ([yshift=-0.75cm]t#1.south) {#4};
-    }
-    \newcommand{\nextTask}[4]{
-      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
-    }
-
-    \task{1}{2,0}{figures/logoTasks/logo144.png}{\mcode{(for (i 4) (\use{3} i))}}
-    \nextTask{2}{t1}{figures/logoTasks/logo121.png}{\mcode{(\use{7} 6 (\use{1}; (\use{0} 1)))}}
-    \nextTask{3}{t2}{figures/logoTasks/logo133.png}{\mcode{(\use{7} 7 (\use{8}; \use{6}))}}
-
-    \vinvention{circle}{p1}{
-      \name{0}{r}\mcode{(for (i $\infty$) (move r $\epsilon$) (move $\epsilon$ $\epsilon$))}
-    }{circle\\\exampleFunction{figures/logoTasks/logo51}{figures/logoTasks/logo63_dilated}}
-    \vinvention{pu}{circle}{
-      \noArguments{1}\mcode{(pen-up (move 1 0))}
-    }{pick up pen\\and move}
-    \vinvention{repeated}{pu}{
-      \name{2}{n$,$d$,$a}\mcode{(for (i n) (move d a))}
-    }{repeated segment\\\exampleFunction{figures/logoTasks/logoDemo1_dilated}{figures/logoTasks/logoDemo0_dilated.png}}
-    \hinvention{square}{circle}{
-      \name{3}{d}\mcode{(\use{2} 4 d (/ $2\pi$ 4))}
-    }{square\\\exampleFunction{figures/logoTasks/logo139_dilated}{figures/logoTasks/logo142_dilated}}
-    \vinvention{smooth}{square}{
-      \name{4}{d$,$a}\mcode{(\use{2} $\infty$ d a)}
-    }{Smooth curve\\\exampleFunction{figures/logoTasks/logoDemo2_dilated}{figures/logoTasks/logoDemo3_dilated}}
-    \vinvention{angle}{smooth}{
-      \name{5}{n}\mcode{(/ $2\pi$ n)}
-    }{$\frac{2\pi}{n}$}
-    \hinvention{semi}{square}{
-      \noArguments{6}\mcode{(\use{4} $\epsilon$ $\epsilon$)}
-    }{semicircle\\\exampleFunction{figures/logoTasks/logo55_dilated}{figures/logoTasks/logo62_dilated}}
-    \vinvention{snow}{semi}{
-      \name{7}{n$,$f}\mcode{(for (i n) \\(move 0 (\use{5} n))\\ (get/set f))}
-    }{rotational symmetry\\\exampleFunction{figures/logoTasks/logo119_dilated}{figures/logoTasks/logo68_dilated}}
-    \vinvention{line}{snow}{
-      \noArguments{8}\mcode{(move 1 0)}
-    }{line segment}
-
-    \draw[->] (p1.south) -- (square.north);
-    \draw[->] (p2.south) -- (circle.north);
-    \draw[->] (p3.south) -- (semi.north);
-    \draw[->] (p2.south) -- (snow.west);
-    \draw[->] (p2.south) -- (pu.east);
-    \draw[->] (p3.south) to[out=190,in=120] (line.west);
-    \draw[->] (p3.south) to[out=200,in=120] (snow.west);
-
-    \draw[->,orange] (snow.west) -- (angle.east);
-    \draw[->,orange] (smooth.west) -- (repeated.east);
-    \draw[->,orange] (semi.west) -- (smooth.east);
-    \draw[->,orange] (square.west) -- (repeated.east);
-
-    \node[align=center,rotate=90] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
-    \node(programLabel)[align=center,rotate=90] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
-    \node[rotate=90,align=center] at ([xshift=-2cm]pu.center) {\textbf{\textsc{Domain Specific Language}}\\
-      \small Learned routines in {\color{orange}orange}};
-          \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: Graphics Programming}}};
-
-    \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
-    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};
-
-    \node(d1) at ([xshift=3.5cm,yshift=-0.5cm]semi.east) {\includegraphics[width = 1.5cm]{figures/finalDreams/63smooth_pretty}};
-    \node(d2) at ([yshift=-1cm]d1.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/400smooth_pretty}};
-    \node(d3) at ([yshift=-1cm]d2.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/399smooth_pretty}};
-
-    \node(d4) at ([yshift=-1cm]d3.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/425smooth_pretty}};
-    \node(dreamPanel)[draw,fit = (d1) (d2) (d3) (d4)] {};
-    \node[anchor=south,align=center] at (dreamPanel.north) {\textbf{\textsc{Dreams}}\\(\footnotesize Samples from DSL)};
-
-    \node(curly)[align=center,rotate=90] at ([xshift=1cm]snow.east) {$\underbrace{\hspace{6cm}}_{}$};
-    \draw[squiggle,->] ($(0,-3.3) + (curly.east)$) -- node[along, sloped, above]{sample}(dreamPanel.west);
-
-  \end{tikzpicture}
-  \caption{Two of the six domains we apply our system to.
-Agent observes tasks (top rows) which it solves by writing programs (middle rows) while jointly growing a library (DSL; bottom rows).
-Learned DSLs rediscover multiple higher-order functions (\code{filter} for list functions and rotational symmetry for generative graphics).
-Learned DSL components call each other (arrows).}  \label{exampleDSL}
+refer to our approach as an instance of `deep program learning'.
+
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/dpl/list}
+  \includegraphics[width = \textwidth]{figures/dpl/logo}
+  \caption{}\label{exampleDSL}
 \end{figure}
 
+%% \section{Old introduction}
+
+%% An old dream within AI is a machine that learns and reasons by writing
+%% its own programs.  This vision stretches back to the
+%% 1960's~\cite{solomonoff1964formal} and, if fully realized, could bring
+%% us much closer to machines that learn and think like humans.
+%% Computational models of cognition often explain the flexibility and
+%% richness of human thinking in terms of program learning: from everyday
+%% thinking and problem solving (motor program induction as an account of
+%% recognition and generation of handwriting and
+%% speech~\cite{lake2015human}; functional programs as a model of natural
+%% language semantics~\cite{SOMETHING}) to learning problems that unfold
+%% over longer developmental time scales: the child's acquisition of
+%% intuitive theories (of kinship, taxonomy, etc.)~\cite{Ullman2012} and
+%% natural language grammar~\cite{DBLP:journals/cogsr/SchmidK11}, to name
+%% just a few.  An outstanding challenge, however, is to engineer
+%% program-learners that display the same level of domain-generality as
+%% the humans they are meant to model.
+
+%% Recent program-learning systems developed within the AI and machine
+%% learning community are impressive along many dimensions, authoring
+%% programs for problem domains like drawing
+%% pictures~\cite{spiral,ellis2017learning}, transforming
+%% text~\cite{gulwani2011automating} and numerical
+%% sequences~\cite{balog2016deepcoder}, robot motion
+%% planning~\cite{devlin2017neural}, and reasoning over common sense
+%% knowledge bases~\cite{muggleton2015meta}.  These systems work in
+%% different ways, but typically hinge upon a carefully hand-engineered
+%% Domain Specific Language (DSL).  The DSL restricts the space of
+%% programs to contain the kinds of concepts needed for one specific
+%% domain.  For example, a picture-drawing DSL could include concepts
+%% like circles and spirals, and a DSL for numerical sequences could
+%% include sorting and reversing lists of numbers.  Modern systems also
+%% learn how to efficiently deploy the DSL on new
+%% problems~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}, but --
+%% unlike human learners -- do not discover the underlying system of
+%% concepts needed to navigate the domain.
+
+%% We contribute a program-induction system that learns the
+%% domain-specific concepts (DSL) while jointly learning how to use those
+%% concepts.  This joint learning problem models two complementary
+%% notions of domain expertise: domain experts have at their disposal
+%% a powerful, yet specialized repertoire of concepts and abstractions
+%% (analogous to the DSL) while also having accurate intuitions about
+%% when and how to use those concepts to solve new problems.
+%% Representative domains,
+%% along with DSLs we learn for them,
+%% are shown in Figure~\ref{exampleDSL}.
+
+%% We call our system `DreamCoder' because it acquires these two kinds of
+%% expertise through a novel kind of wake/sleep or `dream'
+%% learning~\cite{hinton1995wake}, iterating through a wake cycle --
+%% where it solves problems by writing programs -- and a pair of sleep
+%% cycles, both of which are loosely biologically inspired by actual
+%% sleep.  The first sleep cycle, which we refer to as
+%% \textbf{consolidation}, grows the DSL by replying experiences from
+%% waking and consolidating them into new code abstractions.  This cycle
+%% is inspired by the formation of abstractions during sleep memory
+%% consolidation~\cite{DUDAI201520}.  The second sleep cycle, which we
+%% refer to as \textbf{dreaming}, improves the agents knowledge of how to
+%% write code by training a neural network to help search for
+%% programs. The neural net is trained on replayed experiences as well as
+%% `fantasies', or samples, from the DSL.  These two kinds of dreams are
+%% inspired by the distinct episodic replay and hallucination components
+%% of dream sleep~\cite{fosse2003dreaming}.
+
+%% This dream-learning architecture brings together two lines of prior work,
+%% both of which have been separately influential within
+%% artificial intelligence.
+%% One line of work
+%% considers the problem of learning new concepts, abstractions, or
+%% `options' from
+%% experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
+%% while the other line of work considers the problem of learning how to
+%% deploy those concepts
+%% efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
+%% goal with DreamCoder is to show that the combination of these ideas is
+%% uniquely powerful, and pushes us toward program-writing systems that,
+%% like human learners, autonomously acquire the expertise needed to
+%% navigate a new domain of problems.
+
+%% Each wake/sleep cycle creates new DSL components that build on
+%% components learned in earlier sleep cycles, growing a DSL with nested 
+%% hierarchies of code.  We identify this cumulative nesting of
+%% abstractions as a variety of deep representation
+%% learning~\cite{lecun2015deep}.  Figure~\ref{exampleDSL} diagrams a
+%% subset of these learned networks (the DSL). For example, the system
+%% learns to sort sequences of numbers by invoking a DSL component 4
+%% layers deep, or draws the leftmost pairs of images in
+%% Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
+%% refer to \system as an instance of `deep program learning'.
+
+%% Our goal with \system is to engineer a system that develops domain
+%% expertise in humanlike ways. This involves learning both declarative and
+%% procedural knowledge, like a DSL and a recognition model, but includes
+%% other features of the development of expertise:
+%% humans can become experts in many fields,
+%% and so we evaluate our algorithm across six different problem domains;
+%% a human expert doesn't become an expert overnight,
+%% and needs more than a handful of example problems to learn from,
+%% but doesn't need millions of examples;
+%% similarly our algorithm's
+%% learning trajectory unfolds over a series of wake/sleep cycles
+%% requiring around a hundred problems per domain.
+
+%\input{dpl1.tex} % old latex deep programming figures
+
 \section{Deep Wake/Sleep Program Induction}\label{overviewSection}
 
-\system~takes as its goal to acquire domain-specific expertise, which
-it learns by solving a collection of programming \textbf{tasks}.  It
-alternatingly finds programs that solve tasks (Wake --
-Figure~\ref{threeCycles} top); improves its DSL by analyzing programs
-found during waking (Consolidation -- Figure~\ref{threeCycles} left);
-and trains a neural network that efficiently guides search for
-programs in the DSL (Dreaming -- Figure~\ref{threeCycles} right).  The
-learned DSL acts as a a prior on programs likely to solve tasks in the
-domain, while the neural net looks at a specific task and produces a
-``posterior'' for programs likely to solve that specific task
-(Figure~\ref{threeCycles} middle).  The neural network thus functions
-as a \textbf{recognition model} supporting a form of approximate
-Bayesian program induction, jointly trained with a \textbf{generative
-  model} for programs encoded in the DSL, in the spirit of the
-Helmholtz machine~\cite{hinton1995wake}. The recognition model ensures
-that searching for programs remains tractable even as the DSL (and
-hence the search space for programs) expands.  The generative model,
-or DSL, distills out common abstractions across programs discovered
-during waking, growing a network of increasingly deep and specialized
-domain-specific concepts (Figure~\ref{exampleDSL}, bottom rows).
+Our model, called \systemEnding, works through a novel kind
+of wake/sleep or `dream' learning~\cite{hinton1995wake}, iterating
+through a wake cycle -- where it solves \textbf{tasks} by writing programs (Figure~\ref{threeCycles} top)
+-- and a pair of sleep cycles, both of which are loosely biologically
+inspired. The first sleep cycle, which we refer to as
+\textbf{consolidation}, grows its library of code by replying
+experiences from waking and consolidating them into new code
+abstractions (Figure~\ref{threeCycles} left).  This cycle is inspired by the formation of abstractions
+during sleep memory consolidation~\cite{DUDAI201520}.  The second
+sleep cycle, which we refer to as \textbf{dreaming}, improves the
+agents knowledge of how to write code by training a neural network to
+help quickly search for programs. The neural net is trained on replayed
+experiences as well as `fantasies', or samples, from the learned
+library (Figure~\ref{threeCycles} right).  These two kinds of dreams are inspired by the distinct
+episodic replay and hallucination components of dream
+sleep~\cite{fosse2003dreaming}.
+
+Viewed as a probabilistic inference problem, \system observes a set of
+tasks, written $X$, and infers both a program solving each task, as
+well as a prior distribution over programs likely to solve tasks in
+the domain (Figure~\ref{threeCycles} middle). This prior is encoded by
+a library, written $\mathcal{D}$, which combined with a
+learned weight vector $\theta$
+defines a generative model over programs
+(Appendix~\ref{sampleProgram}).  The neural network learns to invert
+this generative model by predicting, conditioned on a task, a
+``posterior'' distribution over programs likely to solve that specific
+task,
+thus functioning as a \textbf{recognition model} and trained jointly with the generative model, in the spirit of the Helmholtz machine~\cite{hinton1995wake}.
+Writing $Q(p|x)$ for the approximate
+posterior predicted by the recognition model, wake/sleep cycles correspond to iteratively (and
+approximately) solving for
+
+\begin{array}{lr}
+  \begin{array}{l}  p_x = \argmax_p \probability[x|p]\probability[p|\mathcal{D}^*,\theta^*]\text{, for each task $x\in X$}\end{array}
+    &\text{\emph{Wake}}\\\\
+\begin{array}{rl}  
+\mathcal{D}^* = \argmax_\mathcal{D}&\int\probability\left[\mathcal{D},\phantom{^*}\theta \right]\prod_{x\in X}\sum_{\substack{p\text{ found during waking}}} \probability[x|p]\probability[p|\mathcal{D},\phantom{^*}\theta]\;\text{d}\theta\\
+\theta^* = \argmax_\theta&\phantom{\int}\probability\left[\mathcal{D}^*,\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D}^*,\theta]
+ \end{array}
+&\text{\emph{Consolidation sleep}}\\\\
+\begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
+  &\text{\emph{Dream sleep}}
+\end{array}
+
+\noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).
+
+%% The model inputs a collection of programming \textbf{tasks}, and then
+%% alternatingly finds programs that solve tasks (Wake --
+%% Figure~\ref{threeCycles} top); improves its library by analyzing programs
+%% found during waking (Consolidation -- Figure~\ref{threeCycles} left);
+%% and trains a neural network that efficiently guides search for
+%% programs written using the library (Dreaming -- Figure~\ref{threeCycles} right).  The
+%% learned library acts as a a prior on programs likely to solve tasks in the
+%% domain, while the neural net looks at a specific task and produces a
+%% ``posterior'' for programs likely to solve that specific task
+%% (Figure~\ref{threeCycles} middle).  The neural network thus functions
+%% as a \textbf{recognition model} supporting a form of approximate
+%% Bayesian program induction, jointly trained with a \textbf{generative
+%%   model} for programs encoded in the DSL, in the spirit of the
+%% Helmholtz machine~\cite{hinton1995wake}. The recognition model ensures
+%% that searching for programs remains tractable even as the library (and
+%% hence the search space for programs) expands.  The generative model,
+%% or library, distills out common abstractions across programs discovered
+%% during waking, growing a network of increasingly deep and specialized
+%% domain-specific concepts (Figure~\ref{exampleDSL}, bottom rows).
 
@@
 
-These wake sleep/cycles function as an approximate inference algorithm
-that observes a collection of tasks, written $X$, and infers both a
-program solving each task, as well as a distribution over programs
-encoded by a DSL, written $\mathcal{D}$.  We equip $\mathcal{D}$ with
-a learned weight vector $\theta$, and together $(\mathcal{D},\theta)$
-define a generative model over programs
-(Appendix~\ref{sampleProgram}).  Writing $Q(p|x)$ for the approximate
-posterior predicted by the recognition model, we iteratively (and
-approximately) solve for
-
-\begin{array}{lr}
-  \begin{array}{l}  p_x = \argmax_p \probability[x|p]\probability[p|\mathcal{D}^*,\theta^*]\end{array}
-    &\text{\emph{Wake}}\\\\
-\begin{array}{rl}  
-\mathcal{D}^* = \argmax_\mathcal{D}&\int\probability\left[\mathcal{D},\phantom{^*}\theta \right]\prod_{x\in X}\sum_{\substack{p\text{ found during waking}}} \probability[x|p]\probability[p|\mathcal{D},\phantom{^*}\theta]\;\text{d}\theta\\
-\theta^* = \argmax_\theta&\phantom{\int}\probability\left[\mathcal{D}^*,\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D}^*,\theta]
- \end{array}
-&\text{\emph{Consolidation sleep}}\\\\
-\begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
-  &\text{\emph{Dream sleep}}
-\end{array}
-
-\noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).
-
-This 3-phase inference procedure works through two distinct kinds of bootstrapping.
-During each sleep cycle the next DSL
-bootstraps off the primitives learned
-during earlier cycles,
-growing an increasingly deep learned program representation.
-In tandem the DSL and recognition model
-bootstrap each other:
-a more finely tuned
-DSL yields
-richer dreams for the recognition model to learn from,
-while a more accurate recognition model
-solves more tasks during waking which then feed into the next DSL.
+This 3-phase inference procedure works through two distinct kinds of
+bootstrapping.  During each sleep cycle the next library bootstraps
+off the concepts learned during earlier cycles, growing an
+increasingly deep learned program representation.  In tandem the
+generative and recognition models bootstrap each other: a more finely
+tuned library of concepts yields richer dreams for the recognition
+model to learn from, while a more accurate recognition model solves
+more tasks during waking which then feed into the next library.
 
@@
 posterior probability, or programs which are a priori likely and which
-solve a task.  We find programs solving a task by enumerating
-programs from the DSL in decreasing order of their probability under
-the recognition model, and then check if a program $p$ assigns
+solve a task. During a Wake cycle we sample a minibatch of tasks and find programs solving specific task by enumerating
+programs in decreasing order of their probability under
+the recognition model,  then checking if a program $p$ assigns
 positive probability to a task ($\probability[x|p] > 0$). We represent
@@
 
+\section{Related work}
+
+We build on several generations of research in AI, program synthesis,
+and cognitive science, with program induction being one of the oldest
+theoretical frameworks for concept learning within artificial
+intelligence~\cite{solomonoff1964formal}, and the conceptually allied
+`Language of Thought Hypothesis' being almost as
+old~\cite{fodor1975language}. Recent neural program synthesis systems
+pair a fixed domain-specific programming language (a `DSL') to a
+learned neural network that guides program
+search~\cite{spiral,balog2016deepcoder,devlin2017neural}, while recent
+symbolic AI research has developed frameworks for learning the DSL by
+inferring reusable pieces of
+code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
+A main goal of this work is to combine and refine these techniques
+with the intention of building agents that, like humans, develop
+domain expertise for new classes of problems.  Work in cognitive
+modeling has often exploited program-like representations for both
+concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
+learning~\cite{logical,Ullman2012},
+and we view these models of concept learning as
+a kind of program induction,
+and see theory learning as similar to inducing a DSL or library.
+
+Deep learning is a dominant machine intelligence paradigm that
+has had many successes
+at creating machines with impressive expertise at
+specific tasks in vision, natural language, gameplaying, and many other areas.
+ideas: we want to become an expert from modest amounts of data; we want to be able to transfer to different types of problems, and to do this zero shot;
+
 \begin{figure}
@@
     \begin{scope}[shift={(0.5,0.5)}]
-      \node[align=right] at (0,-0.5) (d){\textbf{DSL}\\
+      \node[align=right] at (0,-0.5) (d){\textbf{Library}\\
         $f_1$, $f_2$, ...};
@@
     \begin{scope}[shift={(9.4,-3.5)},scale=0.6,line width=0.05mm]
-      \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
+      \node[obs,scale=0.7] at (3.5,3) (dx){Library};
       \node[latent,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
@@
       \node(dreaming) at (1,-1) {\underline{Fantasies}};
-      \node[anchor=center] at ([yshift=-0.5cm]dreaming.south) (d){\textbf{DSL}};
+      \node[anchor=center] at ([yshift=-0.5cm]dreaming.south) (d){\textbf{Library}};
       \node at ([yshift=-1.75cm]d.south) (p1){program};
@@
       \begin{scope}[shift={(-3.8,-8)},scale=0.6,line width=0.05mm]
-        \node[obs,scale=0.7] at (3.5,3) (dx){DSL};
+        \node[obs,scale=0.7] at (3.5,3) (dx){Library};
         \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
@@
         \begin{tikzpicture}[scale=0.9,line width=0.5mm]
-          \node[align=center] at (0,0) {\textbf{new DSL} w/ \texttt{(+ x 1)}:};
+          \node[align=center] at (0,0) {\textbf{new Library} w/ \texttt{(+ x 1)}:};
           \begin{scope}[shift={(0.6,-0.5)}]
@@
       \begin{scope}[shift={(4,-8)},scale=0.6,line width=0.05mm]
-        \node[latent,scale=0.7] at (3.5,3) (dx){DSL};
+        \node[latent,scale=0.7] at (3.5,3) (dx){Library};
         \node[obs,scale=0.7] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
@@
       \spiral{(3.5,1)}{3.5}
-      \node[latent,scale=1] at (3.5,3) (dx){DSL};
+      \node[latent,scale=1] at (3.5,3) (dx){Library};
       \node[latent,scale=1] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
@@
 
-\begin{comment}
-  \subsection{Wake: Solving tasks}\label{explorationSection}
-
-During waking we enumerate programs from the DSL in decreasing order
-of their probability according to the recognition model, and then
-check if a program $p$ assigns positive probability to a task
-($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
-$\mathcal{B}_x$. We represent programs as polymorphicly-typed
-$\lambda$-calculus expressions, a representation closely resembling
-Lisp and functional languages like Haskell and OCaml, including
-variables, conditionals, higher-order recursive functions, and the
-ability to create new functions.  We `batch' the tasks by randomly
-shuffling the training set and playing small minibatches of tasks to
-the agent during each wake cycle.
-
-Why enumerate, when the program synthesis community has invented many
-sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
-We have two reasons:
-(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
-(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
-the space of  programs.
-\end{comment}
-\begin{comment}
-  \begin{figure}
-    \begin{tikzpicture}[every node/.style={rounded corners,thick}]
-      \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
-      \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
-        e.g., RNN or CNN};
-      %    \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
-      \node[draw,align = center](s) at ([xshift=2cm]e.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
-      \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Beam\\$\mathcal{B}_x$};
-
-      \draw[->] (x.east) -- (e.west);
-      \draw[->] (e.east) -- (s.west);
-      %    \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
-      \draw[->] (s.east) -- (f.west);
-      \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
-      \node[align=center] at ($(0,-1) + (e.center)$) {$\underbrace{\hspace{4cm}}_{\normalsize\text{\normalsize Recognition model $Q(p|x)$}}$};
-      \node[align=center] at ($(s.west)!0.5!(f.east) + (0,-1)$) {$\underbrace{\hspace{5cm}}_{\text{\normalsize Search}}$};
-    \end{tikzpicture}
-    \caption{Neurally-guided program inference pipeline.
+%% \begin{comment}
+
+%%   \subsection{Wake: Solving tasks}\label{explorationSection}
+
+%% During waking we enumerate programs from the DSL in decreasing order
+%% of their probability according to the recognition model, and then
+%% check if a program $p$ assigns positive probability to a task
+%% ($\probability[x|p] > 0$); if so, we incorporate $p$ into the beam
+%% $\mathcal{B}_x$. We represent programs as polymorphicly-typed
+%% $\lambda$-calculus expressions, a representation closely resembling
+%% Lisp and functional languages like Haskell and OCaml, including
+%% variables, conditionals, higher-order recursive functions, and the
+%% ability to create new functions.  We `batch' the tasks by randomly
+%% shuffling the training set and playing small minibatches of tasks to
+%% the agent during each wake cycle.
+
+%% Why enumerate, when the program synthesis community has invented many
+%% sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
+%% We have two reasons:
+%% (1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
+%% (2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
+%% the space of  programs.
+%% \end{comment}
+%% \begin{comment}
+%%   \begin{figure}
+%%     \begin{tikzpicture}[every node/.style={rounded corners,thick}]
+%%       \node(x)[draw,align=center] at (0,0) {Task $x$\\e.g., input/outputs};
+%%       \node(e)[draw, align=center] at ([xshift=3cm]x.east) {Domain-specific encoder\\
+%%         e.g., RNN or CNN};
+%%       %    \node(m)[draw,align = center] at ([xshift=1.1cm]e.east) {MLP};%\\Output: $(|\mathcal{D}| + 2)\times(|\mathcal{D}| + 1)\times A$};
+%%       \node[draw,align = center](s) at ([xshift=2cm]e.east) {Enumerate\\(Algorithm~\ref{recognitionSample})};
+%%       \node(f)[draw,align = center] at ([xshift=2cm]s.east) {Beam\\$\mathcal{B}_x$};
+
+%%       \draw[->] (x.east) -- (e.west);
+%%       \draw[->] (e.east) -- (s.west);
+%%       %    \draw[->] (m.east) -- node(p)[fill=white,align=center,midway,inner sep=0,outer sep=0,rotate=90]{\small $Q_{ijk}(x)$}(s.west);
+%%       \draw[->] (s.east) -- (f.west);
+%%       \draw[->] (x.north) -- ([yshift=15]x.north) -- node[above = 0pt,midway]{\small likelihood $\probability[x|p]$}([yshift=15,xshift=40]s.north) -- ($(s.east)!0.5!(f.west)$);
+%%       \node[align=center] at ($(0,-1) + (e.center)$) {$\underbrace{\hspace{4cm}}_{\normalsize\text{\normalsize Recognition model $Q(p|x)$}}$};
+%%       \node[align=center] at ($(s.west)!0.5!(f.east) + (0,-1)$) {$\underbrace{\hspace{5cm}}_{\text{\normalsize Search}}$};
+%%     \end{tikzpicture}
+%%     \caption{Neurally-guided program inference pipeline.
 Recognition model outputs distribution over program $Q(p|x)$.
 Program output by enumerative search incorporated into beam if likelihood $\probability[x|p] > 0$}
-    \label{neuralPipeline}
-  \end{figure}
-\end{comment}
+%%     \label{neuralPipeline}
+%%   \end{figure}
+%% \end{comment}
 
@@
 
-\begin{comment}
-
-\subsubsection{Parameterizing $Q$}\label{recognitionParameterization}
-
-Broadly the literature contains two different approaches to
-parameterizing conditional distributions over programs.  The first
-approach~\cite{devlin2017robustfill,zavershynskyi2018naps} is to use a
-recurrent network to predict the entire program token-by-token, which
-has the advantage that, if the network is sufficiently powerful, it
-can completely solve the synthesis problem.  The disadvantage is that
-these models can perform poorly at out-of-sample
-generalization~\cite{}, which is critical for our setting, as the
-agent may need to solve new tasks that are qualitatively different
-from the tasks it has solved so far.
-
-%% Second, a powerful deep
-%% recurrent network may be costly to sample or enumerate from --- so if
-%% the network cannot easily solve a task, we cannot compensate with
-%% rapid sampling or enumeration.  In contrast, state-of-the-art
-%% enumerative program synthesizers evaluate millions of programs per
-%% second~\cite{feser2015synthesizing}.
-
-The second approach is to have $Q$ predict a fixed-dimensional weight
-vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ecc}
-or sampler~\cite{menon2013machine}.  This approach can enjoy strong
-out-of-sample generalization, because it can fall back on enumeration
-or sampling when the target program is unlike the training programs.
-A main drawback is that the neural net is deliberately handicapped,
-and can only send so much information about the target program.
-
-We adopt a middle ground between these two extremes.  Our recognition
-model predicts a distribution over primitives in the DSL,
-conditioned on the local context in the syntax tree of the
-program. When predicting the next node to add to the syntax tree of a program,
-the recognition model
-conditions on the parent node, as well as
-which argument is being generated.
-This is a kind of `bigram' model over trees,
-where the bigrams
-take the form of (parent, child, argument index).
-Figure~\ref{symmetry} (left) diagrams this generative process and Algorithm~\ref{recognitionSample}
-specifies a sampling procedure for $Q(\cdot |x)$.
-This parameterization
-confers three main advantages:
-(1) it supports fast enumeration
-and sampling of programs,
-because the
-recognition model
-only needs to run once for each task;
-(2) it allows the recognition model to provide fine-grained
-information about the structure of the target program;
-and (3)
-training this recognition model
-causes it to learn to break symmetries in the space of programs,
-described next.
-
-\subsubsection{Learning to break symmetries in program space}\label{symmetricSection}
-
-A good DSL not only exposes high-level building blocks, but also
-carefully restricts the ways in which those building blocks are
-allowed to compose.  For example,
-%% a DSL for list manipulation should
-%% contain both the empty list and a routine for appending lists, but
-%% should not allow appending the empty list.  Similarly 
-a DSL for arithmetic should contain both addition and the number zero
-but disallow adding zero.  These restrictions break symmetries in the
-space of programs.  A bigram parameterization of the recognition
-model, combined with the $\mathcal{L}^{\text{MAP}}$ training
-objective, interact in a way that breaks symmetries in the program
-space, allowing the agent to more efficiently explore the space of
-programs.  This interaction occurs because the bigram parameterization
-can disallow DSL primitives depending on their local syntactic
-context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
-probability mass onto a single member of a set of syntactically
-distinct but semantically equivalent expressions
-(Appendix~\ref{recognitionAppendix}).
-
-We experimentally confirm this symmetry-breaking
-behavior by training recognition models that minimize either
-$\mathcal{L}^{\text{MAP}}$/$\mathcal{L}^\text{posterior}$ and which
-use either a bigram parameterization/unigram\footnote{In the unigram variant $Q$ predicts a $|\mathcal{D}| + 1$-dimensional vector: $Q(p|x) = \probability[p|\mathcal{D},\theta_i = Q_i(x)]$,
-  and was used in our prior work~\cite{ecc}} parameterization.
-Figure~\ref{symmetry} shows the result of training $Q$ in these four regimes for a DSL containing \code{+}, \code{0}, and \code{1}
-and then sampling programs.
-On this particular run,
-the combination of
-bigrams and $\mathcal{L}^{\text{MAP}}$ learns to
-avoid adding zero and associate addition to the right ---
-different random initializations
-lead to either right or left association.
-  \end{comment}
+%% \subsubsection{Parameterizing $Q$}\label{recognitionParameterization}
+
+%% Broadly the literature contains two different approaches to
+%% parameterizing conditional distributions over programs.  The first
+%% approach~\cite{devlin2017robustfill,zavershynskyi2018naps} is to use a
+%% recurrent network to predict the entire program token-by-token, which
+%% has the advantage that, if the network is sufficiently powerful, it
+%% can completely solve the synthesis problem.  The disadvantage is that
+%% these models can perform poorly at out-of-sample
+%% generalization~\cite{}, which is critical for our setting, as the
+%% agent may need to solve new tasks that are qualitatively different
+%% from the tasks it has solved so far.
+
+%% %% Second, a powerful deep
+%% %% recurrent network may be costly to sample or enumerate from --- so if
+%% %% the network cannot easily solve a task, we cannot compensate with
+%% %% rapid sampling or enumeration.  In contrast, state-of-the-art
+%% %% enumerative program synthesizers evaluate millions of programs per
+%% %% second~\cite{feser2015synthesizing}.
+
+%% The second approach is to have $Q$ predict a fixed-dimensional weight
+%% vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ecc}
+%% or sampler~\cite{menon2013machine}.  This approach can enjoy strong
+%% out-of-sample generalization, because it can fall back on enumeration
+%% or sampling when the target program is unlike the training programs.
+%% A main drawback is that the neural net is deliberately handicapped,
+%% and can only send so much information about the target program.
+
+%% We adopt a middle ground between these two extremes.  Our recognition
+%% model predicts a distribution over primitives in the DSL,
+%% conditioned on the local context in the syntax tree of the
+%% program. When predicting the next node to add to the syntax tree of a program,
+%% the recognition model
+%% conditions on the parent node, as well as
+%% which argument is being generated.
+%% This is a kind of `bigram' model over trees,
+%% where the bigrams
+%% take the form of (parent, child, argument index).
+%% Figure~\ref{symmetry} (left) diagrams this generative process and Algorithm~\ref{recognitionSample}
+%% specifies a sampling procedure for $Q(\cdot |x)$.
+%% This parameterization
+%% confers three main advantages:
+%% (1) it supports fast enumeration
+%% and sampling of programs,
+%% because the
+%% recognition model
+%% only needs to run once for each task;
+%% (2) it allows the recognition model to provide fine-grained
+%% information about the structure of the target program;
+%% and (3)
+%% training this recognition model
+%% causes it to learn to break symmetries in the space of programs,
+%% described next.
+
+%% \subsubsection{Learning to break symmetries in program space}\label{symmetricSection}
+
+%% A good DSL not only exposes high-level building blocks, but also
+%% carefully restricts the ways in which those building blocks are
+%% allowed to compose.  For example,
+%% %% a DSL for list manipulation should
+%% %% contain both the empty list and a routine for appending lists, but
+%% %% should not allow appending the empty list.  Similarly 
+%% a DSL for arithmetic should contain both addition and the number zero
+%% but disallow adding zero.  These restrictions break symmetries in the
+%% space of programs.  A bigram parameterization of the recognition
+%% model, combined with the $\mathcal{L}^{\text{MAP}}$ training
+%% objective, interact in a way that breaks symmetries in the program
+%% space, allowing the agent to more efficiently explore the space of
+%% programs.  This interaction occurs because the bigram parameterization
+%% can disallow DSL primitives depending on their local syntactic
+%% context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
+%% probability mass onto a single member of a set of syntactically
+%% distinct but semantically equivalent expressions
+%% (Appendix~\ref{recognitionAppendix}).
+
+%% We experimentally confirm this symmetry-breaking
+%% behavior by training recognition models that minimize either
+%% $\mathcal{L}^{\text{MAP}}$/$\mathcal{L}^\text{posterior}$ and which
+%% use either a bigram parameterization/unigram\footnote{In the unigram variant $Q$ predicts a $|\mathcal{D}| + 1$-dimensional vector: $Q(p|x) = \probability[p|\mathcal{D},\theta_i = Q_i(x)]$,
+%%   and was used in our prior work~\cite{ecc}} parameterization.
+%% Figure~\ref{symmetry} shows the result of training $Q$ in these four regimes for a DSL containing \code{+}, \code{0}, and \code{1}
+%% and then sampling programs.
+%% On this particular run,
+%% the combination of
+%% bigrams and $\mathcal{L}^{\text{MAP}}$ learns to
+%% avoid adding zero and associate addition to the right ---
+%% different random initializations
+%% lead to either right or left association.
+
 \begin{figure}
@@
 
+%\subsection{Results on held out text patterns}
+%\makeatletter
+%\renewcommand{\verbatim@font}{\ttfamily\tiny} 
+%\input{regex.tex}
+
 \end{document}

### 2019-04-07 dpl1.tex
updated some learning curves

\begin{figure}
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \newcommand{\inventionColor}{orange}
    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
    }

    \newcommand{\horizontalSpacing}{3.8cm}
    \newcommand{\verticalSpacing}{2cm}

%    \draw (0,0) rectangle (10,5);
    \node(t1)[align=left] at (1,-1) {
      \footnotesize +106 769-43$\to $106.769.43\\
      \footnotesize +83 973-831$\to $83.973.831\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(}\subnode{f01}{${\color{\inventionColor}f_0}$}\mcode{  \codechar{.} \codechar{-}}\\
      \footnotesize \hspace{1.25cm}\mcode{(${\color{\inventionColor}f_0}$ \codechar{.} \codechar{ }}\\
      \footnotesize \hspace{1.4cm}\mcode{(cdr s)))}
    };
    \node(t2)[align=left,anchor=center] at ([xshift=\horizontalSpacing]t1.center) {
      \footnotesize Dream Coder$\to$Dr. Dream\\
      \footnotesize Andy Cencici$\to$Dr. Andy\\
      \footnotesize Jan Kotas$\to$Dr. Jan\\
      \footnotesize Nancy Jones$\to$Dr. Nancy\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_3}$ (${\color{\inventionColor}f_4}$ \codechar{ } s))}
    };
    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
      \footnotesize Dream (Coder)$\to$Coder\\
      \footnotesize 1 (2) 3 4$\to$2\\
      \footnotesize the (dog) ran$\to$ran\\
      \footnotesize +817(33)99$\to$33\\
      \footnotesize $f(\text{\mcode{s}}) = $\mcode{(${\color{\inventionColor}f_6}$ \codechar{(} \codechar{)} s)}
    };

    \draw (-2,-2.5) -- (15,-2.5);
    \node[align=left,rotate=90] at (-1.5,-1.3) {Tasks and\\\\\texttt{Programs}};

    \node[align=left,rotate=90] at (-1.5,-6.3) {Domain Specific Language (DSL)\\\\System discovers subroutines in {\color{\inventionColor}orange}};

    \newcommand{\invention}[4]{
      \node(#1)[inner sep=1,anchor=center,text width=2cm,draw,rounded corners,align=center] at (#2) {
        \footnotesize #3
      };
      \node[anchor=north] at ([yshift=3.5pt]#1.south) {#4};
    }
    \newcommand{\vinvention}[4]{
      \invention{#1}{[yshift=-\verticalSpacing]#2.center}{#3}{#4}
    }
    \newcommand{\hinvention}[4]{
      \invention{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

    \invention{substitution}{1,-4}{
      ${\color{\inventionColor}f_0}($\mcode{s}$,$\mcode{a}$,$\mcode{b}$) \,=\, $\\
       \mcode{(map ($\lambda$ (x) (if (= x a)\\ b x)) s)}
    }{substitute chars}
    \vinvention{cat}{substitution}{
      ${\color{\inventionColor}f_1}($\mcode{a}$,$\mcode{b}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y)\\ (cons x y))\\ a b))}
    }{concat strings}
    \vinvention{looping}{cat}{
       ${\color{\inventionColor}f_2}($\mcode{c}$,$\mcode{p}$,$\mcode{s}$) \,=\, $\\
      \mcode{(unfold l ($\lambda$ (x) (= c (car x))) p ($\lambda$ (z) (cdr z)))}
    }{loop until char}

    \hinvention{constant}{substitution}{
      ${\color{\inventionColor}f_3}($\mcode{s}$) \,=\, $\\
      \mcode{(${\color{\inventionColor}f_1}$ s string)}
    }{prepend constant}
    \vinvention{drop}{constant}{
      ${\color{\inventionColor}f_4}($\mcode{c}$) \,=\, $\\
      \mcode{(${\color{\inventionColor}f_2}$ c ($\lambda$ }\\\mcode{(z) (car z)))}
    }{prefix}
    \vinvention{suffix}{drop}{
      ${\color{\inventionColor}f_5}($\mcode{s}$,$\mcode{c}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
    }{suffix}

    \hinvention{delimited}{constant}{
      \mcode{${\color{\inventionColor}f_6}($x$,$y$,$s$)\,=\,$}\\
      \mcode{(${\color{\inventionColor}f_4}$ x (${\color{\inventionColor}f_5}$ y c))}
    }{delimited substring}
    \draw[->] (delimited.west) -- (drop.east);
    \draw[->] (delimited.west) -- (suffix.east);

    \vinvention{appendSingle}{delimited}{
      \mcode{${\color{\inventionColor}f_7}($c$,$s$)\,=\,$}\\
      \mcode{(${\color{\inventionColor}f_1}$ (cons c nil) s)}}{append character}

    \vinvention{surround}{appendSingle}{
      \mcode{${\color{\inventionColor}f_8}($s$)\,=\,$}\\
      \mcode{(cons \codechar{(} (${\color{\inventionColor}f_7}$ \codechar{)} s))}}{surround in parens}
    \draw[->] (surround.west) to[in=200,out=160] ([yshift=-0.25cm]appendSingle.west);
    \draw[->] (appendSingle.west) to[in=40,out=120] ([yshift=0.25cm]cat.east);

    \draw[->] (drop.west) -- (looping.east);
    \draw[->] (constant.west) -- (cat.east);

    \draw[->] ([xshift=-0.5cm]t2.south) -- (constant.north);
    \draw[->] ([xshift=0.3cm]t2.south) to[in=30,out=-30] ([xshift=1cm]drop.north);

    \draw[->] ([yshift=0.2cm,xshift=-0.2cm]t1.south) to[in=100,out=200] (substitution.north);
    \draw[->] ([yshift=0.5cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
    \draw[->] ([xshift=-0.5cm]t3.south) -- (delimited.north);

    \begin{scope}[shift={(0,-15)}]
      \invention{filter}{0,0}{
        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
        }{higher-order `filter'}

      \vinvention{three}{filter}{\footnotesize ${\color{\inventionColor}f_1} = $\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
      \hinvention{single}{filter}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}
      \vinvention{for}{single}{\footnotesize ${\color{\inventionColor}f_4} = $\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
      \vinvention{Max}{for}{\name{5}{l}\\\mcode{%
          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
      \hinvention{largest}{single}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
      \vinvention{count}{largest}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}
      \vinvention{weird}{count}{\name{8}{p$,$l}\mcode{
          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0))))}}
                 {Higher-order incrementer}

      \draw[->] (for.west) -- (three.east);
      \draw[->] ([yshift=-0.0cm]largest.west) -- (Max.east);
      \draw[->] ([yshift=-0.cm]largest.west) to[out=240,in=330] ([yshift=-0.1cm]filter.east);
      \draw[->] (Max.west) -- (filter.east);
      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);

      \node(t1)[align=left,anchor=center] at (0,3) {
        \mcode{[1 4 5 8]$\to$[5 8]}\\
        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
        \mcode{[3 4 5 6]$\to$[5 6]}\\
        \mcode{[6 5 4 3]$\to$[6 5]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}
      };
      \node(t2)[align=left] at ([xshift=\horizontalSpacing]t1.center) {
        \mcode{[9 2 4]$\to$[4 2 9]}\\
        \mcode{[4 3]$\to$[3 4]}\\
        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
        \mcode{[]$\to$[]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}
      };
      \node(t3)[align=left] at ([xshift=\horizontalSpacing]t2.center) {
        \mcode{[9 2 3]$\to$[2 3 9]}\\
        \mcode{[2 1]$\to$[1 2]}\\
        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
        \footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}
      };

      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]t1.south) to[out=250] (filter.north);
      \draw[->] ([yshift=0.0cm,xshift=0.3cm]t1.south) to[out=350,in=100] (for.west);
      \draw[->] ([xshift=-0.7cm]t2.south) -- (single.north);
      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]t3.south) to[out=240] (largest.north);

    \end{scope}

    \end{tikzpicture}
\end{figure}

\begin{figure}\centering
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \newcommand{\inventionColor}{orange}

    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$\,=\,$}
    }
    \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
    \newcommand{\horizontalSpacing}{4.2cm}
    \newcommand{\verticalSpacing}{2cm}

    \newcommand{\invention}[4]{
      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,ellipse,draw,rounded corners,align=center] at (#2) {
        \footnotesize #3
      };
      \node[anchor=north,align=center] at ([yshift=3.5pt]#1.south) {#4};
    }
    \newcommand{\vinvention}[4]{
      \invention{#1}{[yshift=-\verticalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\hinvention}[4]{
      \invention{#1}{[xshift=\horizontalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\exampleFunction}[2]{
      \includegraphics[width = 0.5cm]{#1}, \includegraphics[width = 0.5cm]{#2}
    }
    \newcommand{\task}[4]{
      \node[align=left,anchor=center](t#1) at (#2) {#3};
      \node[align=left](p#1) at ([yshift=-0.75cm]t#1.south) {#4};
    }
    \newcommand{\nextTask}[4]{
      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

    \begin{scope}[shift={(0,8)}]
      \task{1}{2,3}{
        \textsc{Reverse}\\
        \mcode{[9 2 4]$\to$[4 2 9]}\\
        \mcode{[4 3]$\to$[3 4]}\\
        \mcode{[3 4 5 6]$\to$[6 5 4 3]}\\
        \mcode{[]$\to$[]}
      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(fold ($\lambda$ (x a)}\\\hspace{0.25cm}\mcode{ (${\color{\inventionColor}f_3}$ x a)) nil z)}}
      \nextTask{2}{t1}{
        \textsc{Take if $\geq 5$}\\
        \mcode{[1 4 5 8]$\to$[5 8]}\\
        \mcode{[9 8 3 8]$\to$[9 8 8]}\\
        \mcode{[3 4 5 6]$\to$[5 6]}\\
        \mcode{[6 5 4 3]$\to$[6 5]}
      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(${\color{\inventionColor}f_0}$ ($\lambda$ (x)}\\\hspace{0.5cm}\mcode{ (> x ${\color{\inventionColor}f_4}$)) z)}}

      \nextTask{3}{t2}{
        \textsc{Sort}\\
        \mcode{[9 2 3]$\to$[2 3 9]}\\
        \mcode{[2 1]$\to$[1 2]}\\
        \mcode{[8 7 9 2 5]$\to$[2 5 7 8 9]}\\
        \mcode{[9 8 9]$\to$[8 9 9]}
      }{\footnotesize $f(\text{\mcode{z}}) = $\mcode{(map ($\lambda$ (i)}\\\hspace{0.5cm}\mcode{(${\color{\inventionColor}f_6}$ z i)) }\\\hspace{0.5cm}\mcode{(+ 1 (range }\\\hspace{0.5cm}\mcode{(length z)))}}

      %% \vinvention{filter}{p1}{
      %%   \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
      %% }{higher-order `filter'}
      \vinvention{single}{p1}{\name{3}{z$,$l}\\\mcode{(fold ($\lambda$ (x y)\\(cons x y))\\ z l)}}{Append element}

      \vinvention{three}{single}{\footnotesize \noArguments{1}\\\mcode{(+ 1 (+ 1 1))}}{The number `3'}
      \vinvention{prefix}{three}{\name{2}{n$,$z}\\\mcode{(map ($\lambda$ (i) (index i z)) (range (+ 1 n)))}}{First $n$ elements}
      \hinvention{filter}{single}{
        \name{0}{p$,$z}\mcode{\\(fold ($\lambda$ (x a)\\ (if (p x) (cons x a) a))\\ nil z)} 
      }{higher-order `filter'}
      \vinvention{for}{filter}{\noArguments{4}\\\mcode{(+ 1 ${\color{\inventionColor}f_1}$)}}{The number `4'}
      \vinvention{weird}{for}{\name{8}{p$,$l}\mcode{
          (map ($\lambda$ (x) (if (p x) (+ 1 x) 0)))}}{Higher-order incrementer}                 
      \hinvention{largest}{filter}{\name{6}{l$,$n}\\\mcode{(${\color{\inventionColor}f_5}$ (${\color{\inventionColor}f_0}$ ($\lambda$ (x) \\(> n (length (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (> x y)) l)))) l))}}{$n^{\text{th}}$ largest element}
      \vinvention{Max}{largest}{\name{5}{l}\\\mcode{%
          (car (${\color{\inventionColor}f_0}$ ($\lambda$ (x) (nil? (${\color{\inventionColor}f_0}$ l ($\lambda$ (z) (> z x))))) l))}}{max of list}
      \vinvention{count}{Max}{\name{7}{p$,$l}\\\mcode{(length\\ (${\color{\inventionColor}f_0}$ ($\lambda$ (y) (= x y)) l))}}{count}

      \node[rotate=90,align=center] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
      \node(programLabel)[rotate=90,align=center] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
      \node[rotate=90,align=center] at ([xshift=-2cm]three.center) {\textbf{\textsc{Domain Specific Language}}\\
        \small Learned routines in {\color{orange}orange}};
      \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: List Processing}}};

      \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};

      \draw[->] (for.west) -- (three.east);
      \draw[->] (largest.east) to[out=0,in=0] (Max.east);
      \draw[->] ([yshift=-0.cm]largest.west) -- ([yshift=0.2cm]filter.east);
      \draw[->] (Max.west) -- (filter.east);
      \draw[->] ([yshift=0.1cm]count.west) -- ([yshift=-0.3cm]filter.east);

      \draw[->] ([yshift=0.35cm,xshift=-0.3cm]p2.south) to[out=250] (filter.north);
      \draw[->] ([yshift=0.0cm,xshift=0.3cm]p2.south) to[out=350,in=0] (for.east);
      \draw[->] ([xshift=-0.7cm]p1.south) -- (single.north);
      \draw[->] ([xshift=-1.3cm,yshift=0.6cm]p3.south) to[out=240] (largest.north);

    \end{scope}

    \renewcommand{\task}[4]{
      \node(t#1) at (#2) {\includegraphics[width = 2cm]{#3}};
      \node(p#1) at ([yshift=-0.75cm]t#1.south) {#4};
    }
    \newcommand{\nextTask}[4]{
      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

    \task{1}{2,0}{figures/logoTasks/logo144.png}{\mcode{(for (i 4) (\use{3} i))}}
    \nextTask{2}{t1}{figures/logoTasks/logo121.png}{\mcode{(\use{7} 6 (\use{1}; (\use{0} 1)))}}
    \nextTask{3}{t2}{figures/logoTasks/logo133.png}{\mcode{(\use{7} 7 (\use{8}; \use{6}))}}

    \vinvention{circle}{p1}{
      \name{0}{r}\mcode{(for (i $\infty$) (move r $\epsilon$) (move $\epsilon$ $\epsilon$))}
    }{circle\\\exampleFunction{figures/logoTasks/logo51}{figures/logoTasks/logo63_dilated}}
    \vinvention{pu}{circle}{
      \noArguments{1}\mcode{(pen-up (move 1 0))}
    }{pick up pen\\and move}
    \vinvention{repeated}{pu}{
      \name{2}{n$,$d$,$a}\mcode{(for (i n) (move d a))}
    }{repeated segment\\\exampleFunction{figures/logoTasks/logoDemo1_dilated}{figures/logoTasks/logoDemo0_dilated.png}}
    \hinvention{square}{circle}{
      \name{3}{d}\mcode{(\use{2} 4 d (/ $2\pi$ 4))}
    }{square\\\exampleFunction{figures/logoTasks/logo139_dilated}{figures/logoTasks/logo142_dilated}}
    \vinvention{smooth}{square}{
      \name{4}{d$,$a}\mcode{(\use{2} $\infty$ d a)}
    }{Smooth curve\\\exampleFunction{figures/logoTasks/logoDemo2_dilated}{figures/logoTasks/logoDemo3_dilated}}
    \vinvention{angle}{smooth}{
      \name{5}{n}\mcode{(/ $2\pi$ n)}
    }{$\frac{2\pi}{n}$}
    \hinvention{semi}{square}{
      \noArguments{6}\mcode{(\use{4} $\epsilon$ $\epsilon$)}
    }{semicircle\\\exampleFunction{figures/logoTasks/logo55_dilated}{figures/logoTasks/logo62_dilated}}
    \vinvention{snow}{semi}{
      \name{7}{n$,$f}\mcode{(for (i n) \\(move 0 (\use{5} n))\\ (get/set f))}
    }{rotational symmetry\\\exampleFunction{figures/logoTasks/logo119_dilated}{figures/logoTasks/logo68_dilated}}
    \vinvention{line}{snow}{
      \noArguments{8}\mcode{(move 1 0)}
    }{line segment}

    \draw[->] (p1.south) -- (square.north);
    \draw[->] (p2.south) -- (circle.north);
    \draw[->] (p3.south) -- (semi.north);
    \draw[->] (p2.south) -- (snow.west);
    \draw[->] (p2.south) -- (pu.east);
    \draw[->] (p3.south) to[out=190,in=120] (line.west);
    \draw[->] (p3.south) to[out=200,in=120] (snow.west);

    \draw[->,orange] (snow.west) -- (angle.east);
    \draw[->,orange] (smooth.west) -- (repeated.east);
    \draw[->,orange] (semi.west) -- (smooth.east);
    \draw[->,orange] (square.west) -- (repeated.east);

    \node[align=center,rotate=90] at ([xshift=-2cm]t1.center) {\textbf{\textsc{Tasks}}\\};
    \node(programLabel)[align=center,rotate=90] at ([xshift=-2cm]p1.center) {\texttt{\textbf{Progs.}}\\};
    \node[rotate=90,align=center] at ([xshift=-2cm]pu.center) {\textbf{\textsc{Domain Specific Language}}\\
      \small Learned routines in {\color{orange}orange}};
          \node at ([yshift=0.5cm]t2.north) {\textbf{\Large \textsc{Domain: Graphics Programming}}};

    \draw [dashed] ([xshift=0.5cm,yshift=-0.25\verticalSpacing]programLabel.east) -- ([xshift=15cm,yshift=-0.25\verticalSpacing]programLabel.east);
    \node[align=left,anchor=center] at ([xshift=14cm,yshift=-0.3\verticalSpacing]programLabel.east) {\large $\uparrow$\;system inputs\\\\\\\large $\downarrow$\;system outputs};

    \node(d1) at ([xshift=3.5cm,yshift=-0.5cm]semi.east) {\includegraphics[width = 1.5cm]{figures/finalDreams/63smooth_pretty}};
    \node(d2) at ([yshift=-1cm]d1.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/400smooth_pretty}};
    \node(d3) at ([yshift=-1cm]d2.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/399smooth_pretty}};

    \node(d4) at ([yshift=-1cm]d3.center) {\includegraphics[width = 1.5cm]{figures/finalDreams/425smooth_pretty}};
    \node(dreamPanel)[draw,fit = (d1) (d2) (d3) (d4)] {};
    \node[anchor=south,align=center] at (dreamPanel.north) {\textbf{\textsc{Dreams}}\\(\footnotesize Samples from DSL)};

    \node(curly)[align=center,rotate=90] at ([xshift=1cm]snow.east) {$\underbrace{\hspace{6cm}}_{}$};
    \draw[squiggle,->] ($(0,-3.3) + (curly.east)$) -- node[along, sloped, above]{sample}(dreamPanel.west);

  \end{tikzpicture}
  \caption{Two of the six domains we apply our system to. Agent observes tasks (top rows) which it solves by writing programs (middle rows) while jointly growing a library (DSL; bottom rows). Learned DSLs rediscover multiple higher-order functions (\code{filter} for list functions and rotational symmetry for generative graphics). Learned DSL components call each other (arrows).}  \label{exampleDSL}
\end{figure}

### 2019-04-08 archive.tex
discussion progress

@@
 and an architect has concepts like arches, supports, and bridges.
-Second, experts have implicit procedural skill in deploying those concepts
+Second, experts have implicit, procedural skill in deploying those concepts
 to quickly solve new problems:
@@
 In short, we take the stance that expertise means both
-having the right concepts, and being able to quickly see how those concepts can be composed into a solution.
-
-Because solutions to many kinds of
-problems can often be described as some kind of
+having the right explicit concepts, and being able to quickly see how those concepts can be composed into a solution.
+
+Because solutions to many kinds of problems can often be described as
+some kind of
 program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
-our model solves a specific problem by synthesizing a program solving it. It
-gradually grows its domain-specific knowledge for a class of problems
-by assembling a library of code containing concepts useful for the
-domain, and by training a neural network to quickly infer, for a
-specific problem, what code in its library is likely to
-solve it. We think of these two learned components as analogous
-to the explicit concepts and
-implicit procedural skill that human experts develop.
+our model solves a specific problem by synthesizing a program solving
+it. It gradually grows its explicit domain-specific knowledge for a
+class of problems by assembling a library of code containing concepts
+useful for the domain, and by training a neural network to quickly
+infer, for a specific problem, what code in its library is likely to
+solve it. We think of these two learned components as analogous to the
+explicit concepts and implicit procedural skill that human experts
+develop.  The model is an instance of what in the machine learning
+community is called a wake-sleep algorithm, where for us, `waking'
+corresponds to solving problems, while a pair of `sleep' cycles
+correspond to improving the library of explicit concepts and training
+the neural net to improve implicit procedural knowledge of how to
+write code.
+
 We will first investigate our model within classic program
@@
 \section{Discussion}
+
+\emph{(what we found)} Our goal was to build a model capturing
+humanlike aspects of learning to be an expert in a new domain of
+problems.  Two main findings of our work are that, first, a single
+program-induction system can learn to solve large sets of problems
+from many qualitatively different domains; and second, that fully
+acquiring expertise in these domains hinges both upon explicit
+declarative knowledge, and implicit procedural skill.  We have
+deliberately placed the interplay of declarative and procedural
+knowledge at the center of our model, and believe that both these
+kinds of learning are crucial for building agents that, like humans,
+autonomously learn to navigate a new domain of problems.
+
+Our model's domain knowledge grows over a series of wake/sleep cycles,
+with the solutions to easier problems bootstrapping
+solving of harder tasks.
+This dynamic is related to, but distinct from, curriculum learning.
+In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead sees randomly sampled batches of tasks, effectively self-pacing its way through an implicit curriculum.
+But humans select their tasks in much richer ways,
+and can even generate their own tasks to solve,
+either as stepping stones toward hard, unsolved problems or motivated by curiosity and aesthetics.
+Building agents that generate their own problems in these humanlike ways is a necessary next step
+if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
+both of which go far beyond the relatively small-scale demonstrations in this work.
+
+A basic representational challenge is to
+learn concepts fundamentally inexpressible in terms of the starting basis: \system invents procedures for
+adding vectors, taking dot products, and calculating forces,
+but can not make the jump needed to define new terms like `acceleration' and `coordinate system,'
+concepts inexpressible as a library of procedures.
+We believe approaches based either on term rewriting systems or defining new generalized algebraic datatypes could give traction to fundamental problems like these,
+strategies which are closely allied with conceptual role semantics.
+Indeed, humans often open up whole new fields of inquiry by introducing concepts not
+definable in terms of prior abstractions,
+like the meaning of `allele' in genetics
+or `force' in physics.
+
+related work
+
+limitations
+
+how will go beyond? generate own tasks; real DSL learning (all expressible in base language); lifelong learning (forgetting?)
 
@@
     \toprule
-    Domain&Timeout&CPUs&Batch size&$\lambda$ (\ref{appendixCompression})&$\alpha$ (\ref{mapAppendix})&Max frontier (\ref{systemPseudocode})\\\midrule
-    Lists&7m&64&10&1&30&5\\
-    Text&7m&64&10&1&30&5\\
-    Graphics&1h&128&40&1.5&30&5\\
+    Domain&Timeout&CPUs&Batch size&$\lambda$ (\ref{appendixCompression})&$\alpha$ (\ref{mapAppendix})&Max beam size (\ref{systemPseudocode})\\\midrule
+    Lists&7m&64&10&1.5&30&5\\
+    Text&7m&64&10&1.5&30&5\\
+    Graphics&1h&96&50&1.5&30&5\\
     Symbolic regression&2m & 40&10 & 1&30&5\\
-    Towers& 2m& 64&10 & 1&30&5\\
-    Regexes&30m & 64&40 & 1&30&5\\
+    Towers& 1h& 64&50 & 1.5&30&5\\
+    Regexes&30m & 64&40 & 1.5&30&5\\
     \bottomrule     

### 2019-04-09 archive.tex
Add TSNE section to appendix

@@
 \section{Appendix}
+
+\subsection{Neural Initial Problem Representations}
+
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/all_tsne.png}
+  \caption{}\label{allTSNE}
+\end{figure}
+
+How do \systemEnding `s representations of problems themselves change, as the model acquires expertise?
+Within cognitive science, a rich body of evidence suggests that human experts learn to `see` problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}.
+Novices notice a problem’s `surface features`, the literal objects and details made obvious by the task description.
+Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures`, the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.
+
+\systemEnding`s neural recognition model can be seen as forming analogous `initial` problem representations, implicitly encoded within the activations of the neural network.
+Intuitively, these initial neural representations map problem features to a `gist` over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves.
+To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding`s learning trajectory.
+Figure \ref{allTSNE} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm.
+We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories.
+Rather, as \systemEnding acquires domain expertise, the neural network adapts how it transforms each task`s surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.
 

### 2019-04-10 archive.tex
regex text from Max

@@
 and predicting the regex \$d
+
+For this task, we used the dataset from [cite Luke], consisting of a
+large corpus of string concepts. Each string concept has 5 ``train"
+strings and 5 ``test" strings.
+
+We trained \system using 128 string concepts, randomly sampled from
+the corpus, and tested on another 128 randomly sampled string
+concepts.
+
+Our initial DSL consisted of all printable characters, four special
+character classes (lowercase letters $\\l$, uppercase letters $\\u$,
+digits $\\d$, and '.', which represents all possible characters), and
+3 functions: \textbf{Kleene Star}, \textbf{Maybe}, \textbf{Or}. We
+also included a task-dependent 'constant' token, which prints the
+largest constant string shared across the 5 training tasks.
+Concatenation was done via continuation-style, as above in the logo
+domain ... [idk how you wanted this to be mentioned?]
+
+Because this domain has a relatively large number of parameters, we
+used a modified version of the recognition model, where only unigram
+parameters were inferred. We found that this recognition model trained
+quicker and more had more stable convergence.
+
+Probabilistic regexes have no ground truth notion of ``solved," so we
+evaluate differently from other domains. Our main evaluation metric is
+to report a lower bound on the posterior predictive of each task given
+the learned grammar:
+
+\begin{align} 
+\Probablility[x_{test}|x_{train}, \mathcal D , \theta ] &= \integral
+dp \Probability[x_{test}| p] \Probablility[p| x_{train}, \mathcal D ,
+\theta]\\
+&= \integral dp \Probability[x_{test}| p] \Probablility[x_{train}|p]
+\Probablility[p|\mathcal D , \theta]\\
+\geq& \sum_{p\in F} \Probability[x_{test}| p] \Probablility[x_
+{train}|p] \Probablility[p|\mathcal D , \theta]
+\end{align}
+
+%\begin{align} 
+%\Probablility[x|\mathcal D , \theta ] &= \integral dp \Probablility
+[x|p] \Probablility[p|\mathcal D , \theta]\\
+%\geq& \sum_{p\in F} \Probablility[x|p] \Probablility[p|\mathcal D ,
+\theta]
+%\end{align}
+
+where $x_{train}$ are the training strings for task $x$, and $x_
+{test}$ are the held-out test strings, not seen during search.
+
+For this domain we use an additional post-processing step, in order to
+generate realistic samples from the probabilistic programs. For each
+task and each found probabilistic regex, we separately optimize the
+continuous regex production parameters in order to maximize the
+likelihood on the training strings for that task. We apply additive
+smoothing to ensure optimization does not over-fit to the training
+tasks.
 

### 2019-04-11 archive.tex
moved expertise content into main paper

@@
 a input/output examples, starting with a generic functional programming
-basis: \code{foldr}, \code{unfold}, \code{if}, \code{map},
-\code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
-\code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
-\code{is-nil}.
+basis, including routines like \code{map}, \code{fold}, \code{cons}, \code{car}, \code{cdr}, etc.
+%% : \code{foldr}, \code{unfold}, \code{if}, \code{map},
+%% \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
+%% \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
+%% \code{is-nil}.
 
@@
 each with 15 input/output examples.  In solving these tasks, the
-system composed 16 new DSL components, and discovered multiple
+system typically composes around 20 new DSL components, and discovered multiple
 higher-order functions. Each round of memory consolidation built on
@@
 prior work.
+
+How do \systemEnding's representations of these problems themselves change, as the model acquires expertise?
+Within cognitive science, a rich body of evidence suggests that human experts learn to `see' problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}.
+Novices notice a problem’s `surface features', the literal objects and details made obvious by the task description.
+Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures', the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.
+
+\systemEnding's neural recognition model can be seen as forming analogous `initial' problem representations, implicitly encoded within the activations of the neural network.
+Intuitively, these initial neural representations map problem features to a `gist' over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves.
+To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding's learning trajectory.
+Figure \ref{textExpertiseCluster} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm.
+We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories.
+Rather, as \system acquires domain expertise, the neural network adapts how it transforms each task's surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.
+
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/text_tsne.png}
+  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of text tasks (neural net activations), after first wake/sleep cycle.
+Middle: at-a-glance organization after 20 wake/sleep cycles.
+Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution.
+More examples in Fig.~\ref{allTSNE}}\label{textExpertiseCluster}
+\end{figure}
 
@@
 
-\subsection{Acquiring Domain Expertise}
-
-One interpretation of our system is as a model of the acquisition of
-domain expertise. Humans can acquire expertise across many domains --
-cooking, coding, music, architecture, painting, tennis, or calculus,
-to name a handful of examples, and every child develops expertise in
-natural language, intuitive physics, motor control, kinship
-relationships, and more. Domain experts learn domain-specific
-abstractions, similar to a DSL: for example, a expert chef knows what
-combinations of seasonings go together, or an expert mathematician
-knows a wide set of useful theorems and lemmas.  Jointly, experts
-learn how to recognize when to use these abstractions to solve
-problems.  Becoming an expert involves a learning trajectory that
-unfolds over relatively long time scales, but has modest data
-requirements relative to the dominant machine learning paradigms. For
-example, basic competency in cooking or coding might require on the
-order of learning a hundred recipes or solving a hundred programming
-exercises.  It is this learning regime that we have targeted with
-\systemEnding: Learning about a domain from at most several hundred
-tasks, but where the learning unfolds over many wake/sleep cycles.
-
 \subsection{Prospects for program Induction as part of the generic AI toolkit}
@@
 \section{Appendix}
-
-\subsection{Neural Initial Problem Representations}
-
-\begin{figure}
-  \includegraphics[width = \textwidth]{figures/all_tsne.png}
-  \caption{}\label{allTSNE}
-\end{figure}
-
-How do \systemEnding `s representations of problems themselves change, as the model acquires expertise?
-Within cognitive science, a rich body of evidence suggests that human experts learn to `see` problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}.
-Novices notice a problem’s `surface features`, the literal objects and details made obvious by the task description.
-Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures`, the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.
-
-\systemEnding`s neural recognition model can be seen as forming analogous `initial` problem representations, implicitly encoded within the activations of the neural network.
-Intuitively, these initial neural representations map problem features to a `gist` over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves.
-To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding`s learning trajectory.
-Figure \ref{allTSNE} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm.
-We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories.
-Rather, as \systemEnding acquires domain expertise, the neural network adapts how it transforms each task`s surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.
 
@@
 
+\subsection{Neural Problem Representations}
+
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/all_tsne.png}
+  \caption{}\label{allTSNE}
+\end{figure}
+
 \subsection{Full set of LOGO  tasks}

### 2019-04-14 archive.tex
updated latex to new local figure
a little more writing

@@
   \includegraphics[width = \textwidth]{figures/dpl/list}
-  \includegraphics[width = \textwidth]{figures/dpl/logo}
+  \includegraphics[width = \textwidth]{figures/DeepLogo_4_11}
   \caption{}\label{exampleDSL}
@@
 inspired. The first sleep cycle, which we refer to as
-\textbf{consolidation}, grows its library of code by replying
+\textbf{abstraction}, grows its library of code by replying
 experiences from waking and consolidating them into new code
@@
  \end{array}
-&\text{\emph{Consolidation sleep}}\\\\
+&\text{\emph{Abstraction (sleep)}}\\\\
 \begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
-  &\text{\emph{Dream sleep}}
+  &\text{\emph{Dreaming (sleep)}}
 \end{array}
@@
 recursive functions, and the ability to define new functions.
-
-\section{Related work}
-
-We build on several generations of research in AI, program synthesis,
-and cognitive science, with program induction being one of the oldest
-theoretical frameworks for concept learning within artificial
-intelligence~\cite{solomonoff1964formal}, and the conceptually allied
-`Language of Thought Hypothesis' being almost as
-old~\cite{fodor1975language}. Recent neural program synthesis systems
-pair a fixed domain-specific programming language (a `DSL') to a
-learned neural network that guides program
-search~\cite{spiral,balog2016deepcoder,devlin2017neural}, while recent
-symbolic AI research has developed frameworks for learning the DSL by
-inferring reusable pieces of
-code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
-A main goal of this work is to combine and refine these techniques
-with the intention of building agents that, like humans, develop
-domain expertise for new classes of problems.  Work in cognitive
-modeling has often exploited program-like representations for both
-concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
-learning~\cite{logical,Ullman2012},
-and we view these models of concept learning as
-a kind of program induction,
-and see theory learning as similar to inducing a DSL or library.
-
-Deep learning is a dominant machine intelligence paradigm that
-has had many successes
-at creating machines with impressive expertise at
-specific tasks in vision, natural language, gameplaying, and many other areas.
-ideas: we want to become an expert from modest amounts of data; we want to be able to transfer to different types of problems, and to do this zero shot;
 
@@
       \node[latent,scale=1] at (3.5,3) (dx){Library};
-      \node[latent,scale=1] at ([yshift=-1.7cm,xshift=0cm]dx) (zp){prog};
+      \node[latent,scale=1] at ([yshift=-2cm,xshift=0cm]dx) (zp){prog};
       \node[obs,scale=1] at ([yshift=-1.45cm]zp) (xp) {task};
@@
   \caption{\textbf{Middle:} \system as a graphical model.
-Agent observes programming tasks (e.g., input/outputs for list processing or images for graphics programs), which it explains with latent programs, while jointly inferring a latent Domain Specific Language (DSL) capturing cross-program regularities.
+Agent observes programming tasks (e.g., input/outputs for list processing or images for graphics programs), which it explains with latent programs, while jointly inferring a latent library capturing cross-program regularities.
 A neural network, called the \emph{recognition model} (red arrows) is trained to quickly infer programs with high posterior probability.
-\textbf{Top}: Wake phase infers programs while holding the DSL and recognition model fixed.
-\textbf{Left}: Sleep (Consolidation) phase infers DSL while holding the programs fixed by refactoring programs found during waking and extracting common components.
+\textbf{Top}: Wake phase infers programs while holding the library and recognition model fixed.
+\textbf{Left}: Sleep (Consolidation) phase updates library while holding the programs fixed by refactoring programs found during waking and extracting common components.
 \textbf{Right}: Sleep (Dreaming) phase trains recognition model to predict approximate posterior over programs conditioned on task.
-Trained on `Fantasies' (programs sampled from DSL) \& `Replays' (programs found during waking).}\label{threeCycles}
+Trained on `Fantasies' (programs sampled from library) \& `Replays' (programs found during waking).}\label{threeCycles}
 \end{figure}
@@
 
-\subsection{Consolidation-Sleep: Growing a Domain Specific Language}\label{consolidationSection}
-
-The DSL offers a set of abstractions that allow an agent to concisely
+\subsection{Abstraction Sleep: Growing a Library of Concepts}\label{consolidationSection}
+
+The library offers a set of abstractions that allow an agent to concisely
 express solutions to the tasks at hand. We automatically discover
@@
 programs found during waking.  Second, we use this automatic refactoring
-process to search for DSLs that maximally compress these programs by
+process to search for libraries that maximally compress these programs by
 incorporating reused subexpressions into the DSL.
@@
 Mathematically this compression takes the form of
-finding the DSL maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
+finding the library maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
 We replace this marginal with an AIC approximation~\cite{akaike1998information}
@@
 refactorings. Appendix~\ref{appendixCompression} specifies how we
-combine this probabilistic and symbolic machinery to update the DSL.
+combine this probabilistic and symbolic machinery to update the library.
 At a high level, our approach is to search locally through the space
-of DSLs, proposing small changes until Eq.~\ref{AIC} fails to
+of libraries, proposing small changes until Eq.~\ref{AIC} fails to
 decrease.
@@
   \end{tikzpicture}
-  \caption{DSL learning as code refactoring.
+  \caption{Library learning as code refactoring.
 \textbf{Panel A:} During waking we discover programs for each task, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}).
-Common subprograms are incorporated into the DSL when they increase a Bayesian objective.
-Intuitively, these new DSL components best compress the programs found during waking.
+Common subprograms are incorporated into the library when they increase a Bayesian objective.
+Intuitively, these new components best compress the programs found during waking.
 \textbf{Panel B:} \# of possible refactorings grows exponentially with program size, so we represent refactorings using version spaces, which augment syntax trees with a \emph{union} operator whose children are themselves version spaces.
@@
 sources of self-supervised data: \emph{replays} of programs discovered
-during waking, and \emph{fantasies}, or programs drawn from the DSL.
+during waking, and \emph{fantasies}, or programs drawn from $(\mathcal{D},\theta)$.
 Replays ensure that the recognition model is trained on the actual
@@
 \noindent\textbf{Parameterizing $Q$.} The recognition model predicts a
-fixed-dimensional tensor encoding a distribution over subroutines in
-the DSL, conditioned on the local context in the syntax tree of the
+fixed-dimensional tensor encoding a distribution over routines in
+the library, conditioned on the local context in the syntax tree of the
 program. This local context consists of the parent node in the syntax
@@
 
-\noindent\textbf{Symmetry breaking.} A good DSL not only exposes high-level building blocks, but also
-carefully restricts the ways in which those building blocks are
+\noindent\textbf{Symmetry breaking.} Effective domain-specific representations not only exposes high-level building blocks, but also
+carefully restrict the ways in which those building blocks are
 allowed to compose.  For example,
@@
 %% should not allow appending the empty list.  Similarly 
-a DSL for arithmetic should disallow adding zero, or force
+when searching over arithmetic expressions, one could disallow adding zero, and force
 right-associative addition. A bigram parameterization of the recognition
@@
 programs.  This interaction occurs because the bigram parameterization
-can disallow DSL primitives depending on their local syntactic
+can disallow library routines depending on their local syntactic
 context, while the $\mathcal{L}^{\text{MAP}}$ objective forces all
@@
 each with 15 input/output examples.  In solving these tasks, the
-system typically composes around 20 new DSL components, and discovered multiple
+system typically composes around 20 new library routines, and discovered multiple
 higher-order functions. Each round of memory consolidation built on
@@
 solve a task involving sorting a list of numbers (Figure~\ref{initialExampleDSL}).
-This incremental, modular learning of deep hierarchies of DSL
+This incremental, modular learning of deep hierarchies of library
 components occurs because of the alternation between code writing
-(during waking) and code refactoring (during the consolidation phase
+(during waking) and code refactoring (during the abstraction phase
 of sleep).
-
-%% \begin{figure}[b]\centering
-%% \vspace{-0.5cm}  \begin{tabular}{lll}
-%%     \toprule
-%%     Name & Input & Output \\\midrule
-%%     repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
-%%     drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
-%%     rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
-%%     count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
-%%     keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
-%%     product & [7\, 1\, 6\, 2] & 84 \\
-%%     \bottomrule
-%%   \end{tabular}
-%%   \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
-%% \end{figure}
 
@@
 Excel~\cite{gulwani2011automatin}.  This prior work uses
-hand-engineered DSLs and hand-engineered search strategies.  Here, we
+hand-engineered libraries of primitives and hand-engineered search strategies.  Here, we
 will show that we can jointly learn both these ingredients and surpass
@@
 inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
-with drawing a modest corpus of images while equipping it with control
+with drawing a corpus of 160 images while equipping it with control
 over a `pen', along with arithmetic operations on angles and
@@
 
-Inside its learned DSL we find interpretable parametric drawing
+Inside its learned library we find interpretable parametric drawing
 routines corresponding to the families of visual objects in its
@@
 additionally learns more abstract visual relationships, like
-rotational symmetry, which it models by incorporating a higher-order
-function into its DSL (Figure~\ref{logoPrimitives}, right).  %% This
-%% abstraction comes from jointly compressing programs for all of the
-%% training images into its DSL.
-
-What does \system dream of?  Prior to learning samples from the DSL
+rotational symmetry, which it models by abstracting out a new higher-order
+function into its library (Figure~\ref{logoPrimitives}, right).
+
+What does \system dream of?  Prior to learning random programs written using the library
 are simple and largely unstructured (Figure~\ref{logoDreams}, left).
@@
 offers a visual window into how the generative model bootstraps recognition
-model training: as the DSL grows more finely tuned to the domain, the
+model training: as the library grows more finely tuned to the domain, the
 neural net gets richer and more highly varied training data.
@@
   \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{stolle2002learning},
-    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', from learned DSL.}\label{tower}
+    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
 \end{figure}
@@
 \subsubsection{Generative Modeling of Text}
-We investigate few-shot learning of generative models by tasking our
+Few-shot learning of simple generative concepts comes naturally to
+humans, from learning new rules in natural language~\cite{marcus1999rule},
+to learning routines for symbols and signs
+\cite{lake2015human},
+to learning new motor routines for producing words.
+We investigate few-shot generative modeling by tasking our
 agent with inferring a probabilistic regular expression from a small
-number (5) of strings - for example, observing the strings \$1.20 and \$9.42
-and predicting the regex \$d
-
-For this task, we used the dataset from [cite Luke], consisting of a
-large corpus of string concepts. Each string concept has 5 ``train"
-strings and 5 ``test" strings.
-
-We trained \system using 128 string concepts, randomly sampled from
-the corpus, and tested on another 128 randomly sampled string
-concepts.
-
-Our initial DSL consisted of all printable characters, four special
-character classes (lowercase letters $\\l$, uppercase letters $\\u$,
-digits $\\d$, and '.', which represents all possible characters), and
-3 functions: \textbf{Kleene Star}, \textbf{Maybe}, \textbf{Or}. We
-also included a task-dependent 'constant' token, which prints the
-largest constant string shared across the 5 training tasks.
-Concatenation was done via continuation-style, as above in the logo
-domain ... [idk how you wanted this to be mentioned?]
-
-Because this domain has a relatively large number of parameters, we
-used a modified version of the recognition model, where only unigram
-parameters were inferred. We found that this recognition model trained
-quicker and more had more stable convergence.
-
-Probabilistic regexes have no ground truth notion of ``solved," so we
-evaluate differently from other domains. Our main evaluation metric is
-to report a lower bound on the posterior predictive of each task given
-the learned grammar:
-
-\begin{align} 
-\Probablility[x_{test}|x_{train}, \mathcal D , \theta ] &= \integral
-dp \Probability[x_{test}| p] \Probablility[p| x_{train}, \mathcal D ,
-\theta]\\
-&= \integral dp \Probability[x_{test}| p] \Probablility[x_{train}|p]
-\Probablility[p|\mathcal D , \theta]\\
-\geq& \sum_{p\in F} \Probability[x_{test}| p] \Probablility[x_
-{train}|p] \Probablility[p|\mathcal D , \theta]
-\end{align}
-
-%\begin{align} 
-%\Probablility[x|\mathcal D , \theta ] &= \integral dp \Probablility
-[x|p] \Probablility[p|\mathcal D , \theta]\\
-%\geq& \sum_{p\in F} \Probablility[x|p] \Probablility[p|\mathcal D ,
-\theta]
-%\end{align}
-
-where $x_{train}$ are the training strings for task $x$, and $x_
-{test}$ are the held-out test strings, not seen during search.
-
-For this domain we use an additional post-processing step, in order to
-generate realistic samples from the probabilistic programs. For each
-task and each found probabilistic regex, we separately optimize the
-continuous regex production parameters in order to maximize the
-likelihood on the training strings for that task. We apply additive
-smoothing to ensure optimization does not over-fit to the training
-tasks.
-
-\begin{tabular}{ccc}
-  \toprule   Observed text&Inferred regex&Held out testing text\\\midrule
-
-  \end{tabular}
-
-\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
-To evaluate the relative importance of DSL learning and recognition
-model training, we evaluate on held-out testing tasks for each of our
-domains, measuring both how many tasks are solved and how long it
-takes to solve them across successive wake/sleep iterations
-(Fig.~\ref{learningCurves}).  We always solve more held-out tasks --
-and generally solve them in less time -- with both components
-combined.  Why?  One hypothesis is that some tasks are best solved by
-DSL learning and others by a neural network, and so including both of
-our sleep cycles  takes the union of these sets of tasks.  Another
-hypothesis is that the DSL and neural network interact
-synergistically, bootstrapping off each other to solve tasks for which
-neither alone suffice. We evaluate the relative weight of these interactions by
-computing the ratio of the tasks solved uniquely by the full model to
-
-%% We compare with ablations of our model on held out tasks.
-%% The purpose of this ablation study is 
-%% both to examine the role of each component of \systemEnding,
-%% as well as to compare with
-%% prior approaches in the literature:
-%% a head-to-head
-%% comparison of program synthesizers is complicated by the fact that
-%% each system, including ours, makes idiosyncratic 
-%% assumptions about the space of programs and the statement of tasks.
-
-%% Nevertheless, much prior work can be modeled within our setup. 
-%% We compare with the following ablations (Tbl~\ref{baselineComparisons};
-%% Fig~\ref{learningCurves}):
-%% \\\noindent \textbf{No NN:} lesions the recognition model.
-%% \\\noindent \textbf{NPS}, which does not learn the DSL,
-%% instead learning the recognition model
-%% from samples drawn from the fixed DSL.
-%% We call this NPS (Neural Program Synthesis)
-%% because this is closest to how
-%% RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
-%% \\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
-%% only add \textbf{S}ub\textbf{E}xpressions of programs in the beams to the DSL.
-This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
-%% \\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
-%% but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
-%% \\\noindent \textbf{Enum}, which enumerates a beam without any learning --- equivalently, our first search step.
-%For each domain,
-%% We are interested both in how many tasks the
-%% agent can solve and how quickly it can find those solutions.
-%% Tbl.~\ref{baselineComparisons}
-%% compares our model against these alternatives.
-%% We consistently
-%% improve on the baselines,
-%% and find that lesioning the recognition model
-%% and lesioning it also slows down the convergence of the algorithm,
-%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
-%% This supports a view of the recognition model as a way of amortizing the cost of search.
-
+number of strings, where these strings are drawn from CSV columns crawled from the web~\cite{LUKE}. Across wake/sleep cycles, our agent learns to learn
+regular expressions that describe the structure of a typically occurring text concepts,
+like numbers or dates.
+Figure~\ref{qualitativeRegex} shows example model outputs on held out generative modeling tasks,
+contrasting the full model with ablations that lesion either the library learning or recognition model training.
 \begin{figure}
-  \begin{tabular}{ccc}
-    \begin{tabular}{c}
-      \includegraphics[width = 5cm]{figures/learningCurves/text_hits.png}\\
-      \includegraphics[width = 5cm]{figures/learningCurves/text_time.png}
-    \end{tabular}&
-
-    \begin{tabular}{c}
-      \includegraphics[width = 5cm]{figures/learningCurves/logo_hits.png}\\
-      \includegraphics[width = 5cm]{figures/learningCurves/logo_time.png}
-    \end{tabular}&
-
-    \begin{tabular}{c}
-      \includegraphics[width = 5cm]{figures/learningCurves/rational_hits.png}\\
-      \includegraphics[width = 5cm]{figures/learningCurves/rational_time.png}
-    \end{tabular}\\\\\\
-
-    \begin{tabular}{c}
-      \includegraphics[width = 5cm]{figures/learningCurves/list_hard_hits.png}\\
-      \includegraphics[width = 5cm]{figures/learningCurves/list_hard_time.png}
-    \end{tabular}&
-
-    \begin{tabular}{c}
-      \includegraphics[width = 5cm]{figures/learningCurves/tower_hits.png}\\
-      \includegraphics[width = 5cm]{figures/learningCurves/tower_time.png}
-    \end{tabular}&
-
-    \includegraphics[width = 5cm]{figures/learningCurves/regular}
-  \end{tabular}
-  \caption{Test set performance across wake/sleep iterations.
-Each curve is a run with a different random seed.
-Teal: Full model.
-Orange: Dreaming only (no DSL learning).
-Purple: Consolidation only (no recognition model).
-Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks).}\label{learningCurves}
+  \includegraphics[width = \textwidth]{figures/regexQualitative.pdf}
+  \caption{Qualitative results on probabilistic generative text modeling.
+Agent observes 5 strings (`Input' column) and infers a probabilistic regex (MAP program--regex primitives highlighted in red), from which it can imagine more examples of the latent text concept.
+No Library: Dreaming only (ablates library learning).
+No Rec: Consolidation only (ablates recognition model)}\label{qualitativeRegex}
 \end{figure}
 
-%% \begin{figure}
-%%   \begin{tabular}{ccc}
-%%     \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
-%%     \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
-%%     \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
-%%     \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
-%%     \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
-%%     \includegraphics[width = 5cm]{figures/regexLearningCurve_batch.png}
-%%   \end{tabular}
-%%   \caption{Test set performance across wake/sleep iterations.
-Orange: \systemEnding.
-Teal: No recognition model.
-Purple: No DSL learning.
-Solid lines: \% tasks solved within 10 minute timeout.
-Dashed lines: average search time.
-Shaded region is $\pm 1$ stddev search time.}\label{learningCurves}
-%% \end{figure}
-
-\section{Discussion}
-
-\emph{(what we found)} Our goal was to build a model capturing
-humanlike aspects of learning to be an expert in a new domain of
-problems.  Two main findings of our work are that, first, a single
-program-induction system can learn to solve large sets of problems
-from many qualitatively different domains; and second, that fully
-acquiring expertise in these domains hinges both upon explicit
-declarative knowledge, and implicit procedural skill.  We have
-deliberately placed the interplay of declarative and procedural
-knowledge at the center of our model, and believe that both these
-kinds of learning are crucial for building agents that, like humans,
-autonomously learn to navigate a new domain of problems.
-
-Our model's domain knowledge grows over a series of wake/sleep cycles,
-with the solutions to easier problems bootstrapping
-solving of harder tasks.
-This dynamic is related to, but distinct from, curriculum learning.
-In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead sees randomly sampled batches of tasks, effectively self-pacing its way through an implicit curriculum.
-But humans select their tasks in much richer ways,
-and can even generate their own tasks to solve,
-either as stepping stones toward hard, unsolved problems or motivated by curiosity and aesthetics.
-Building agents that generate their own problems in these humanlike ways is a necessary next step
-if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
-both of which go far beyond the relatively small-scale demonstrations in this work.
-
-A basic representational challenge is to
-learn concepts fundamentally inexpressible in terms of the starting basis: \system invents procedures for
-adding vectors, taking dot products, and calculating forces,
-but can not make the jump needed to define new terms like `acceleration' and `coordinate system,'
-concepts inexpressible as a library of procedures.
-We believe approaches based either on term rewriting systems or defining new generalized algebraic datatypes could give traction to fundamental problems like these,
-strategies which are closely allied with conceptual role semantics.
-Indeed, humans often open up whole new fields of inquiry by introducing concepts not
-definable in terms of prior abstractions,
-like the meaning of `allele' in genetics
-or `force' in physics.
-
-related work
-
-limitations
-
-how will go beyond? generate own tasks; real DSL learning (all expressible in base language); lifelong learning (forgetting?)
-
-\subsection{Learning from Scratch}
-A long-standing dream within the program induction community
-is ``learning from scratch'': starting with a \emph{minimal} Turing-complete programming language,
+\subsection{From libraries to languages}
+
+Rather than start with a domain-specific basis and then enrich it over successive wake/sleep cycles -- as these prior experiments have done --
+could we instead start with a highly generic basis,
+and then learn a library encoding something closer to the basic domain language?
+We first consider learning a basic language for equations describing physical laws,
+starting \emph{without} concepts like vectors or forces
+and instead providing sequences of real numbers and Lisp-style sequence manipulation.
+We collected 57 equations for physical laws from AP and MCAT physics `cheat sheets' --
+forms like Coulomb's law and expressions for the center of mass of a set of objects --
+along with
+reference vector algebra identities from machine learning textbooks,
+and tasked our agent with inferring programs calculating these input/output mappings.
+In solving these tasks, \system first discovers the building blocks of vector algebra,
+like inner products, vector sums, and norms,
+and then uses this vocabulary to infer library routines for common
+physical relationships like inverse square laws and integrating second derivatives over time.
+
+Could we also learn the basics of programming, including these foundational sequence manipulation primitives?
+This goal is relevant to a long-standing dream within the program induction community to ``learn from scratch'': starting with a \emph{minimal} Turing-complete programming language,
 and then learning to solve a wide swath of
@@
 
-\subsection{\system and the Exploration-Compression family of algorithms}
+\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
+To evaluate the relative importance of library learning and recognition
+model training, we evaluate on held-out testing tasks for each of our
+domains, measuring both how many tasks are solved and how long it
+takes to solve them across successive wake/sleep iterations
+(Fig.~\ref{learningCurves}).  We always solve more held-out tasks --
+and generally solve them in less time -- with both components
+combined.  Why?  One hypothesis is that some tasks are best solved by
+DSL learning and others by a neural network, and so including both of
+our sleep cycles  takes the union of these sets of tasks.  Another
+hypothesis is that the DSL and neural network interact
+synergistically, bootstrapping off each other to solve tasks for which
+neither alone suffice. We evaluate the relative weight of these interactions by
+computing the ratio of the tasks solved uniquely by the full model to
+
+\begin{figure}
+  \begin{tabular}{ccc}
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/text_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/text_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/logo_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/logo_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/rational_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/rational_time.png}
+    \end{tabular}\\\\\\
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/list_hard_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/list_hard_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/tower_hits.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/tower_time.png}
+    \end{tabular}&
+
+    \begin{tabular}{c}
+      \includegraphics[width = 5cm]{figures/learningCurves/regex_marginal_test_unigram_gen.png}\\
+      \includegraphics[width = 5cm]{figures/learningCurves/regex_task_train_unigram_gen.png}
+      \end{tabular}
+  \end{tabular}
+  \caption{Test set performance across wake/sleep iterations.
+Each curve is a run with a different random seed.
+Teal: Full model.
+Orange: Dreaming only (no DSL learning).
+Purple: Consolidation only (no recognition model).
+Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks).
+Generative text modeling plots show (top) posterior predictive likelihood of held-out strings on held out tasks and (bottom) lower bound on marginal likelihood of held out tasks, both of which are averaged per character.}\label{learningCurves}
+\end{figure}
+
+\section{Discussion}
+
+\emph{(what we found)} Our goal was to build a model capturing
+humanlike aspects of learning to be an expert in a new domain of
+problems.  Two main findings of our work are that, first, a single
+program-induction system can learn to solve large sets of problems
+from many qualitatively different domains; and second, that fully
+acquiring expertise in these domains hinges both upon explicit
+declarative knowledge, and implicit procedural skill.  We have
+deliberately placed the interplay of declarative and procedural
+knowledge at the center of our model, and believe that both these
+kinds of learning are crucial for building agents that, like humans,
+autonomously learn to navigate a new domain of problems.
+
+\emph{(related work)} \system builds on several generations of research in AI, program synthesis,
+and cognitive science, with program induction being one of the oldest
+theoretical frameworks for concept learning within artificial
+intelligence~\cite{solomonoff1964formal}, and the conceptually allied
+`Language of Thought Hypothesis' being almost as
+old~\cite{fodor1975language}. Recent neural program synthesis systems
+pair a fixed domain-specific programming language (a `DSL') to a
+learned neural network that guides program
+search~\cite{spiral,balog2016deepcoder,devlin2017neural}, while recent
+symbolic AI research has developed frameworks for learning the DSL by
+inferring reusable pieces of
+code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
+A main goal of this work is to combine and refine these techniques
+with the intention of building agents that, like humans, develop
+domain expertise for new classes of problems.  Work in cognitive
+modeling has often exploited program-like representations for both
+concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
+learning~\cite{logical,Ullman2012},
+and we view these models of concept learning as
+a kind of program induction,
+and see theory learning as similar to inducing a DSL or library.
 
 Our work sits within the Exploration-Compression (EC) family of
-algorithms.  EC~\cite{Dechter:2013:BLV:2540128.2540316} is a program
-induction framework where an agent alternates between searching, or
-`exploring', the space of programs, and then updating its search
-procedure by compressing programs found during exploration. \system
+algorithms.  EC~\cite{Dechter:2013:BLV:2540128.2540316} algorithms alternate between searching, or
+`exploring', the space of programs, and then updating their search
+procedure by compressing solutions found during exploration. \system
 grew directly out of research on EC-style systems and both of our
@@
 model aims to encode a program in as few bits as possible, conditioned
-on a task. Our previous work, EC$^2$~\cite{ecc}, introduced
-neurally-guided search into the EC framework and served as the
-starting point for \systemEnding.  Our work here extends EC$^2$ by (1)
-introducing a new refactoring compression algorithm, (2) enriching the
-neural network with a new bigram-over-trees parameterization and loss
-function that together learn to break symmetries, accelerating program
-search during waking, and (3) demonstrating how this family of
-approaches can be applied to planning and generative modeling
-problems.  Other offshoots of EC include the neurosymbolic framework
+on a task. %%  Our previous work, EC$^2$~\cite{ecc}, introduced
+%% neurally-guided search into the EC framework and served as the
+%% starting point for \systemEnding.  Our work here extends EC$^2$ by (1)
+%% introducing a new refactoring compression algorithm, (2) enriching the
+%% neural network with a new bigram-over-trees parameterization and loss
+%% function that together learn to break symmetries, accelerating program
+%% search during waking, and (3) demonstrating how this family of
+%% approaches can be applied to planning and generative modeling
+%% problems. 
+Other offshoots of EC include our prior work, EC$^2$~\cite{ecc}, the neurosymbolic framework
 in~\cite{lazaro2019beyond} and the hierarchical Bayesian program
@@
 
-\subsection{Prospects for program Induction as part of the generic AI toolkit}
-
-Our aim with \system is to chart a path by which program induction can
-become more broadly useful for AI.  This means viewing the AI
-landscape through the lens of program learning, including the terrain
-considered here --- simple kinds of generative modeling, inverse
-graphics, planning, and programming by example --- but also many
-others like reinforcement learning, commonsense reasoning, natural
-language understanding, and causal inference. Can program induction
-rise to the challenge?  We believe it can, provided we push jointly
-along  many different axes of AI research; and provided we continue
-to integrate learning algorithms -- both symbolic and neural, both
-top-down and bottom-up -- into our artificial agents.
-
-Our system, with its learned DSL and neural recognition model, is one
-embodiment of this hybrid symbolic/neural approach, and enjoys some
-success across small-scale problems in the different domains considered
-here. Scaling to larger problems,
-such as inferring 3-D object models (vs LOGO/Turtle),
-learning natural image grammars (vs ),
+\emph{(limitations - and how to go beyond them)} Our model's domain knowledge grows over a series of wake/sleep cycles,
+with the solutions to easier problems bootstrapping
+solving of harder tasks.
+This dynamic is related to, but distinct from, curriculum learning.
+In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
+But humans select their tasks in much richer ways,
+and can even generate their own tasks to solve,
+either as stepping stones toward hard, unsolved problems or motivated by considerations like curiosity and aesthetics.
+Building agents that generate their own problems in these humanlike ways is a necessary next step
+if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
+both of which go far beyond the relatively small-scale demonstrations in this work.
+
+\emph{(I feel the lifelong learning setting is less important of a point to make - we might remove this)} Over a lifetime humans develop expertise in many domains --
+unlike our model,
+which learned to solve many classes of problems,
+but only class one at a time.
+Scaling approaches like ours to this multidomain, lifelong setting
+could be enabled by a hierarchical Bayesian model of DSLs for many domains,
+capturing  shared conceptual structure between similar domains  (e.g., Parcheesi and chess)
+and separating knowledge between dissimilar domains (e.g., Parcheesi and cheese making).
+
+A basic representational challenge is to
+learn concepts fundamentally inexpressible in terms of the starting basis. \system invents procedures for
+adding vectors, taking dot products, and calculating forces,
+but can not make the jump needed to define new terms like `acceleration' and `coordinate system,'
+concepts inexpressible as a library of procedures.
+We believe approaches based either on term rewriting systems or defining new generalized algebraic datatypes could give traction to fundamental problems like these,
+and we note that these strategies are similar in spirit to conceptual role models of semantics.
+Indeed, humans often open up whole new fields of inquiry by introducing concepts not
+definable in terms of prior abstractions,
+like the meaning of `allele' in genetics
+or `force' in physics.
+
+Programs, as a knowledge representation,
+hold promise because of their universal expressiveness and
+compositional systemacity.
+We are optimistic that programs can become a more standard part of the AI toolkit,
+and that a key enabling insight will be combination of
+explicit declarative concepts and implicit procedural skill, both jointly learned.
+
+%% \subsection{Prospects for program Induction as part of the generic AI toolkit}
+
+%% Our aim with \system is to chart a path by which program induction can
+%% become more broadly useful for AI.  This means viewing the AI
+%% landscape through the lens of program learning, including the terrain
+%% considered here --- simple kinds of generative modeling, inverse
+%% graphics, planning, and programming by example --- but also many
+%% others like reinforcement learning, commonsense reasoning, natural
+%% language understanding, and causal inference. Can program induction
+%% rise to the challenge?  We believe it can, provided we push jointly
+%% along  many different axes of AI research; and provided we continue
+%% to integrate learning algorithms -- both symbolic and neural, both
+%% top-down and bottom-up -- into our artificial agents.
+
+%% Our system, with its learned DSL and neural recognition model, is one
+%% embodiment of this hybrid symbolic/neural approach, and enjoys some
+%% success across small-scale problems in the different domains considered
+%% here. Scaling to larger problems,
+%% such as inferring 3-D object models (vs LOGO/Turtle),
+%% learning natural image grammars (vs ),
 

### 2019-04-14 archive.tex
deep program learning experiment analysis

@@
 
-\section{Deep Wake/Sleep Program Induction}\label{overviewSection}
+\section{A Wake/Sleep Algorithm for Program Induction}\label{overviewSection}
 
@@
 -- and a pair of sleep cycles, both of which are loosely biologically
-inspired. The first sleep cycle, which we refer to as
-\textbf{abstraction}, grows its library of code by replying
+inspired by the different memory consolidation processes that occur during sleep~\cite{rasch2013sleep}. The first sleep cycle, which we refer to as
+\textbf{abstraction}, grows the library of code by replying
 experiences from waking and consolidating them into new code
-abstractions (Figure~\ref{threeCycles} left).  This cycle is inspired by the formation of abstractions
-during sleep memory consolidation~\cite{DUDAI201520}.  The second
+abstractions (Figure~\ref{threeCycles} left).  This cycle is loosely inspired by the formation of declarative abstractions
+during slow-wave sleep memory consolidation~\cite{DUDAI201520}.  The second
 sleep cycle, which we refer to as \textbf{dreaming}, improves the
-agents knowledge of how to write code by training a neural network to
+agents procedural knowledge of how to write code by training a neural network to
 help quickly search for programs. The neural net is trained on replayed
-experiences as well as `fantasies', or samples, from the learned
+experiences as well as `fantasies', or sampled programs, built from the learned
 library (Figure~\ref{threeCycles} right).  These two kinds of dreams are inspired by the distinct
@@
 process to search for libraries that maximally compress these programs by
-incorporating reused subexpressions into the DSL.
+incorporating reused subexpressions into the library.
 
@@
 $n^{\text{th}}$ largest element of a list, which it finally uses to
-solve a task involving sorting a list of numbers (Figure~\ref{initialExampleDSL}).
+solve a task involving sorting a list of numbers (Figure~\ref{exampleDSL}).
 This incremental, modular learning of deep hierarchies of library
@@
 
-Procedural visual concepts are studied across AI and cognitive science
---- Bongard problems~\cite{Moscow}, Raven's progressive
-matrices~\cite{raven2003raven}, and Lake et al.'s BPL model of
-omniglot~\cite{lake2015human} are prominent examples.  Here we take
+Procedural or generative visual concepts 
+--- from Bongard problems~\cite{Moscow}, to handwritten characters~\cite{lake2015human,hofstadter1993letter}, to Raven's progressive
+matrices~\cite{raven2003raven} --- are studied across AI and cognitive science.  Here we take
 inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
@@
 \begin{figure}
-  \begin{tabular}{cc}
-    \toprule Example learned DSL components&Samples from DSL\\\midrule
+\centering  \begin{tabular}{cc}
+    \toprule Example learned DSL components&Samples written using final library\\\midrule
     \begin{tabular}{rl}
@@
 and generally solve them in less time -- with both components
-combined.  Why?  One hypothesis is that some tasks are best solved by
-DSL learning and others by a neural network, and so including both of
-our sleep cycles  takes the union of these sets of tasks.  Another
-hypothesis is that the DSL and neural network interact
-synergistically, bootstrapping off each other to solve tasks for which
-neither alone suffice. We evaluate the relative weight of these interactions by
-computing the ratio of the tasks solved uniquely by the full model to
-
+combined, suggesting a synergy between library learning and
+recognition model training. Examining the libraries learned with and without
+recognition models reveals a difference in the typical \emph{depth}
+of learned subroutines, and this average depth correlates
+with the number of solved testing tasks ($r=0.71$, $p < 10^{-8}$).
+In contrast, the correlation between the \emph{number} of learned subroutines and number of solved tasks is weaker and not statistically significant ($r=0.20$, $p > 0.05$),
+suggesting that the recognition model bootstraps
+``better'' libraries, where ``better'' is at least partly explained by the depth
+of the learned representation. Figure~\ref{depthMontage} shows typical `deep' libraries learned by
+\system and contrasts this depth with \% solved tasks, both with and without the recognition model.
 \begin{figure}
@@
 
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/dpl/montage.pdf}
+  \caption{Top: Learned libraries diagrammed as networks.
+Arrows point from a learned concept to all other concepts that use it.
+Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (right).
+Full model (circles on scatterplots) learns deeper libraries than no-recognition-model ablation (crosses on scatterplots).
+Each scatter plot point run with a different random seed.}\label{depthMontage}
+  \end{figure}
+
 \section{Discussion}
@@
 humanlike aspects of learning to be an expert in a new domain of
-problems.  Two main findings of our work are that, first, a single
+problems. Our model works through a wake/sleep algorithm that grows an increasingly deep
+representation of a domain's conceptual structure,
+while learning to quickly combine these concepts into programs
+solving specific problems within the domain.
+Two main findings of our work are that, first, a single
 program-induction system can learn to solve large sets of problems
@@
 
-\subsection{Full set of LOGO  tasks}
+\subsection{LOGO Graphics Experiment}
+
+We apply our system to 160 LOGO graphics tasks, split 50/50 test/train (Figure~\ref{everyLogoTask}).
+For each task of the agent must write a program that moves a simulated pen across the canvas,
+with the goal of drawing the target image, as measured by pixel level similarity.
+
+We initially provide our agent with two control flow primitives: \code{for} (a `for` loop), and \code{get/set}, which gets and saves the current state of the agent's pen, executes a block of code, and then restores the state to its previous value.
+We provide two different drawing primitives: \code{move}, which takes as input both a distance and angle, and moves the pen forward by that distance and rotates by that angle, and \code{pen-up}, which lifts up the pen and then executes a block of code \emph{without} placing down ink on the canvas.
+
+These programs are imperative (produce a sequence of actions) rather than purely functional (calculate a stateless mapping).
+The \system infrastructure only supports purely functional $\lambda$-calculus programs.
+To embed imperative routines within this framework,
+we model imperative routines using a state monad~\cite{wadler1990comprehending} and encode each imperative action in a continuation passing style where each imperative primitive takes as input the actions to execute afterwards (i.e., the continuation),
+and the program as a whole takes as input a final no-op continuation.
+This continuation passing set up allows the program synthesizer to not need the monadic `bind' operator to sequence chains of commands.
+To be concrete, our imperative primitives have the following types:
+
+\begin{tabular}{l}
+  \code{move} : \code{distance}$\to$\code{angle}$\to$\code{State ()}$\to$\code{State ()}\\
+  \code{pen-up} : \code{(State ()$\to$State ())}$\to$\code{State ()}$\to$\code{State ()}\\
+  \code{for} : \code{int$\to$(State ()$\to$State ())}$\to$\code{State ()}$\to$\code{State ()}\\
+  \code{get/set} : \code{(State ()$\to$State ())}$\to$\code{State ()}$\to$\code{State ()}
+\end{tabular}
+
+and each synthesized program has type \code{State ()}$\to$\code{State ()}.
+We additionally provide the agent with natural numbers (1 through 9), arithmetic operations (addition/subtraction/multiplication/division), unit distances and angles (1 meter and $2\pi$ radians), and special values $\infty$ (approximated as $20$) and $\epsilon$ (approximated as $1/20$).
+We distinguish distances and angles in the type system, only allowing a distance/angle to be added/subtracted with another object that is also a distance/angle, and only allowing multiplication/division by unitless scalars.
+This trick requires no modification of the underlying \system software, and is a standard way of embedding units and dimensions within a type system~\cite{fsunits}.
 
@@
   \includegraphics[width = \textwidth]{figures/fullLogo.png}
-  \caption{Full set of LOGO graphics tasks that we apply our system to}
+  \caption{Full set of LOGO graphics tasks that we apply our system to}\label{everyLogoTask}
 \end{figure}
 
-\subsection{Full set of tower tasks}
+\subsection{Tower Building Experiment}
+We apply our system to 112 tower building tasks (Figure~\ref{everyTowerTask}), split 50/50 test/train.
+We use the same monadic, continuation-passing encoding of imperative programs as we used for LOGO graphics,
+and include the exact same control flow primitives.
+Rather than moving a pen over a canvas, the agent here moves a simulated hand left/right over a 2D tower building stage, and has at its disposal an unlimited supply of horizontal and vertical blocks.
+All coordinates are discretized.
+Horizontal blocks have size 6x2 and vertical blocks have size 2x6.
+The state of the simulated hand maintains both its position (a single discrete scalar) and its orientation (a single boolean, either facing left are facing right).
+We include two domain specific primitives for adjusting the position of the hand: \code{move}, which takes as input a scalar $d$ and moves the hand forward  by distance $d$,
+and \code{reverse},
+which flips the orientation of the hand.
+The agent has two additional domain specific primitives for dropping a horizontal/vertical block at the current position of the hand.
+
 \begin{figure}
   \includegraphics[width = \textwidth]{figures/fullTower.png}
-  \caption{Full set of tower building tasks that we apply our system to}
+  \caption{Full set of tower building tasks that we apply our system to}\label{everyTowerTask}
 \end{figure}

### 2019-04-15 archive.tex
Update archive.tex
libraries to language figure starting

@@
 \begin{figure}
+  \includegraphics[width = 1.2\textwidth]{figures/DeepLogo_4_11}
   \includegraphics[width = \textwidth]{figures/dpl/list}
-  \includegraphics[width = \textwidth]{figures/DeepLogo_4_11}
   \caption{}\label{exampleDSL}
@@
 starting \emph{without} concepts like vectors or forces
-and instead providing sequences of real numbers and Lisp-style sequence manipulation.
+and instead providing Lisp-style sequence manipulation primitives like
+\code{map} and \code{fold}.
 We collected 57 equations for physical laws from AP and MCAT physics `cheat sheets' --
@@
 physical relationships like inverse square laws and integrating second derivatives over time.
+
+\begin{figure}
+  \centering
+  \begin{tabular}{lll}\toprule
+    &\multicolumn{1}{c}{Physical Laws}&\multicolumn{1}{c}{Recursive Functions}\\\midrule
+    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&
+    % physical laws programs and tasks
+    \begin{tabular}{l}
+      \begin{tabular}{l}
+        Center-of-mass:\\
+      \end{tabular}\\\\
+      \begin{tabular}{l}
+        Coulomb's law:\\
+      \end{tabular}\\\\
+      \begin{tabular}{l}
+        Integrating position over time:\\
+        \end{tabular}
+      \end{tabular}
+    &
+    % recursive functions programs and tasks
+    \begin{tabular}{l}
+      \begin{tabular}{l}
+        Duplicate each list element:\\
+        \code{[9 2]$\to $[9 9 2 2]}\\
+        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
+        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
+        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
+      \end{tabular}\\\\
+
+      \begin{tabular}{l}
+        Lengths of list-of-lists:\\
+        \code{[[2 1] [1 2 9]\, []]}$\to $\code{[2 3 0]}\\
+        \code{[[5 2 9 3 7 5 2]]}$\to $\code{[7]}\\
+        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
+        \blueCode{$f(\ell) = $($f_2$ ($\lambda$ (z) ($f_4$ z)) $\ell$)}
+      \end{tabular}\\\\
+      \begin{tabular}{l}
+        1-index into list:\\
+        \code{1, [9 2 3]$\to $9}\\
+        \code{2, [1 5 8]$\to $5}\\
+        \code{4, [0 2 8 1 5 6]$\to $1}\\
+        \blueCode{$f($n$,$l$) = $($f_7$ l (+ 1 n))}
+      \end{tabular}\\\\
+
+      \end{tabular}
+    \\\midrule
+    \rotatebox[origin=c]{90}{\popp{Library}}&
+    % physical laws library
+    &
+    % recursive functions programs and tasks
+    \begin{tabular}{l}
+      \begin{tabular}{l}
+        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
+        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
+          \hspace{\helpSize}($f_0$: \emph{fold})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
+        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
+          \hspace{\helpSize}($f_1$: \emph{unfold})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
+        \hspace{\helpSize}($f_2$: \emph{map})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_3($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
+        \hspace{\helpSize}($f_3$: \emph{count upward until predicate holds})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_4($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
+        \hspace{\helpSize}($f_4$: \emph{length})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_5($p$)\,=\,$($f_3$ ($\lambda$ (x) x) p 0)}\\
+        \hspace{\helpSize}($f_5$: \emph{count upward from 0 until predicate holds})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_6($n$)\,=\,$($f_5$ (eq? n))}\\
+        \hspace{\helpSize}($f_6$: \emph{range})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_7($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
+        \hspace{\helpSize}($f_7$: \emph{index})
+      \end{tabular}
+
+    \end{tabular}
+
+    \\\bottomrule
+
+  \end{tabular}
+  \end{figure}
 
@@
       \begin{tabular}{l}
-        \code{[[2 1]\, []]}$\to $\code{[1 0]}\\
+        \code{[[2 1]\, []]}$\to $\code{[2 0]}\\
         \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
-        \blueCode{$f(\ell) = $($f_2$ $f_4$ $\ell$)}
+        \blueCode{$f(\ell) = $($f_2$ $f_5$ $\ell$)}
       \end{tabular}\\\\
@@
         \greenCode{$f_4($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
-        \hspace{\helpSize}($f_4$: \emph{generalization of range})
+        \hspace{\helpSize}($f_4$: \emph{count upward until predicate holds})
       \end{tabular}&
@@
         \greenCode{$f_7($p$)\,=\,$($f_4$ ($\lambda$ (x) x) p 0)}\\
-        \hspace{\helpSize}($f_7$: \emph{another generalization of range})
+        \hspace{\helpSize}($f_7$: \emph{count upward from 0 until predicate holds})
       \end{tabular}\\

### 2019-04-15 archive.tex
improvement to waking sleeping figure but still in progress

@@
     \begin{scope}[shift={(0.5,0.5)}]
-      \node[align=right] at (0,-0.5) (d){\textbf{Library}\\
-        $f_1$, $f_2$, ...};
+      \node[align=center] at (0,-0.5) (d){\textbf{Library}\\
+        \code{$f_1($x$)=$(+ x 1)}\\
+        \code{$f_2($z$)=$(fold cons (cons z nil))}\\
+         ...};
       \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
-        \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[7\, 3]}         \\
-        \footnotesize    \code{[4\, 3\, 2\, 1]}$\to$\code{[4\, 3]} };
+        \footnotesize            \code{[7\, 2\, 3]}$\to$\code{[4 3 8]}         \\
+        \footnotesize    \code{[3\, 8]}$\to$\code{[9 4]}\\
+        \footnotesize    \code{[4\, 3\, 2]}$\to$\code{[3 4 5]} };
 
@@
       \node[align=left] at ([xshift=3cm]s.east) (f) {\textbf{Programs:}\\
-        \small    \code{($f_1$ $\ell$ ($\lambda$ (x) (> x 2)))}\\
+        \small    \code{(map $f_1$ (fold $f_2$ nil x))}\\
         \small $\cdots\cdots\cdots$};
@@
 
-    \node at (0,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
+    \node at (0,-4.75) {\textbf{\textsc{Sleep: Abstraction}}};
     \draw[fill=teal!5!white] (-3,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-3,-13) -- (-3,-5);
@@
 \textbf{Top}: Wake phase infers programs while holding the library and recognition model fixed.
-\textbf{Left}: Sleep (Consolidation) phase updates library while holding the programs fixed by refactoring programs found during waking and extracting common components.
+\textbf{Left}: Sleep (Abstraction) phase updates library while holding the programs fixed by refactoring programs found during waking and abstracting out common components (highlighted in orange).
 \textbf{Right}: Sleep (Dreaming) phase trains recognition model to predict approximate posterior over programs conditioned on task.

### 2019-04-17 archive.tex
improved week/sleep diagram
more progress on physical laws

@@
   \begin{tikzpicture}[line width=0.4mm]
-    \draw[fill=teal!5!white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
-    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};
+    \draw[fill=teal!5!white] (-1.25,1.25) -- (13.25,1.25) -- (13.25,-4) -- (-1.25,-4) -- (-1.25,1.25);
+    \node at (5.5,1.5) {\textsc{\textbf{Wake}}};
 
     \begin{scope}[shift={(0.5,0.5)}]
-      \node[align=center] at (0,-0.5) (d){\textbf{Library}\\
-        \code{$f_1($x$)=$(+ x 1)}\\
-        \code{$f_2($z$)=$(fold cons (cons z nil))}\\
-         ...};
+      \node[align=center] at (0,-0.5) (d){
+        \begin{tabular}{l}
+          \multicolumn{1}{c}{\textbf{Library}}\\
+        \small\code{$f_1($x$)=$(+ x 1)}\\
+        \small\code{$f_2($z$)=$(fold cons}\\
+        \small\phantom{\code{$f_2($z$)$}}\code{(cons z nil))}\\
+\small $\cdots\cdots\cdots$
+                \end{tabular}};
       \node[align=center] at ([yshift = -2cm]d) (t){\textbf{Task}\\
@@
     \end{scope}
+
     \begin{scope}[shift={(9.4,-3.5)},scale=0.6,line width=0.05mm]
@@
 and then learn a library encoding something closer to the basic domain language?
-We first consider learning a basic language for equations describing physical laws,
-starting \emph{without} concepts like vectors or forces
-and instead providing Lisp-style sequence manipulation primitives like
+We first consider learning a basic language for physical laws,
+starting with generic sequence-manipulation primitives,
+and then investigate how these fundamental sequence-manipulating routines could themselves be learned.
+
+We first task \system with inferring 57 equations describing physical laws,
+like closed forms for ballistic motion, inverse square laws,
+expressions for calculating centers of mass, etc.,
+drawing verbatim from AP and MCAT physics `cheat sheets'
+along with equation guides from machine learning textbooks.
+These relationships are easy to express
+given concepts like vectors or forces.
+Rather than `bake in' abstractions like these we
+initialize the system with Lisp-style sequence manipulation primitives like
 \code{map} and \code{fold}.
-We collected 57 equations for physical laws from AP and MCAT physics `cheat sheets' --
-forms like Coulomb's law and expressions for the center of mass of a set of objects --
-along with
-reference vector algebra identities from machine learning textbooks,
-and tasked our agent with inferring programs calculating these input/output mappings.
+We automatically convert each task into a
+regression problem by randomly generating real or vector valued input/outputs,
+and encode vectors as lists of reals.
+
 In solving these tasks, \system first discovers the building blocks of vector algebra,
-like inner products, vector sums, and norms,
-and then uses this vocabulary to infer library routines for common
-physical relationships like inverse square laws and integrating second derivatives over time.
+like inner products, vector sums, and norms.
+It then uses this vocabulary to infer library routines for common
+physical relationships like the vector form inverse square laws
+and integrating second derivatives over time,
+effectively undergoing a `change of basis' from the initial Lisp-style primitives into a
+physics-style basis. Figure~\ref{librariesToLanguages}~(left) illustrates a subset
+of the provided tasks and learned library for this domain.
 
@@
         Center-of-mass:\\
+        $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
+        \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
+        \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
       \end{tabular}\\\\
@@
         Coulomb's law:\\
+        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
+        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
+        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        
       \end{tabular}\\\\
@@
         Integrating position over time:\\
+        $x_0$, $\dot{x}$, $\ddot{x}$, $t\mapsto \frac{1}{2}\ddot{x}t^2 + \dot{x}t + x_0$\\
+
         \end{tabular}
@@
     % physical laws library
+    \begin{tabular}{l}
+      \begin{tabular}{l}
+        \greenCode{$f_0($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ (* x x) y)) v)}\\
+        \hspace{\helpSize}($f_0$: \emph{L-2 norm})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_1($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ x y)) v)}\\
+        \hspace{\helpSize}($f_1$: \emph{Sum vector components})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_2($v,u$)\,=\,$($f_1$ (zip v u ($\lambda$ (x y) (* x y))))}\\
+        \hspace{\helpSize}($f_2$: \emph{Dot product})
+      \end{tabular}\\
+      %(lambda (fold $0 (map (lambda (*. $0 0.)) (car $0)) (lambda (lambda (zip $0 $1 (lambda (lambda (+. $0 $1)))))))) 
+      \begin{tabular}{l}
+        \greenCode{$f_3($vs$)\,=\,$(fold vs (map ($\lambda$ (x) (* x 0))}\\
+        \phantom{\greenCode{$f_3($vs$)\,=\,$(fold vs (}}\greenCode{(car vs))}\\
+        \phantom{\greenCode{$f_3($vs$)\,=\,$(}}\greenCode{($\lambda$ (v u) (zip v u ($\lambda$ (x y) (+ x y)))))}\\
+        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_4($x$)\,=\,$(/ x (+ 1 1))}\\
+        \hspace{\helpSize}($f_4$: \emph{Divide by two})
+      \end{tabular}\\
+      %      %#(lambda (lambda (lambda (lambda (+. $2 (*. $0 (+. (*. (#(lambda (/. $0 (+. 1. 1.))) $0) $1) $3))))))) $2 $3 $1 $0)))))
+      \begin{tabular}{l}
+        \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
+        \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_6($v$)\,=\,$(power ($f_0$ v) ($f_4$ 1))}\\
+        \hspace{\helpSize}($f_6$: \emph{n-dimensional vector length})
+        \end{tabular}
+      \end{tabular}
     &
@@
   \end{tabular}
+  \caption{\textbf{Left:} Learning a language for physical laws starting with recursive list routines like \code{map} and \code{fold}.
+Agent observes input/outputs of latent functions encoding physical laws (top left, 20 input/outputs per task).
+Vectors are represented as lists of real numbers.
+Physical constants expressed in Planck units.
+\textbf{Right:} Learning a language for recursive list routines starting with only recursion and primitives found in 1959 Lisp.
+Agent rediscovers \code{map}, \code{fold}, \code{unfold}, etc.
+For full tasks and learned libraries see Appendix~\ref{appendixMcCarthy}}\label{librariesToLanguages}
   \end{figure}
 
-Could we also learn the basics of programming, including these foundational sequence manipulation primitives?
+Could we also learn the basics of programming, including these recursive sequence manipulation primitives?
 This goal is relevant to a long-standing dream within the program induction community to ``learn from scratch'': starting with a \emph{minimal} Turing-complete programming language,
@@
 within higher-order functions like \code{map}, \code{fold}, and
-\code{unfold}.
-
-After running for 93 hours on 48 CPUs, our
-algorithm solves these 18 exercises, along the way assembling a DSL
-with a modern repertoire of
-functional programming idioms and subroutines, including \code{map},
-\code{fold}, \code{unfold}, \code{index}, \code{length},
-and  arithmetic operations like 
-building lists of natural numbers between an interval (see  Appendix~\ref{appendixMcCarthy}).
+\code{unfold}.  In solving these 18 exercises, the algorithm assembles
+a DSL containing a modern repertoire of functional programming idioms
+and subroutines, including \code{map}, \code{fold}, \code{unfold},
+\code{index}, \code{length}, and arithmetic operations like building
+lists of natural numbers between an interval (see
+Figure~\ref{librariesToLanguages}, right).
 
@@
 basis like those embodied in the standard libraries of modern  languages.
-What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
+What these experiments reveal is that \system doesn't \emph{need} to start from a rich basis,
 and can in principle recover many of the amenities of modern programming systems,
@@
 
-\subsection{Learning from Scratch: Tasks and DSL}\label{appendixMcCarthy}
+\subsection{Learning from Scratch: Tasks and Library}\label{appendixMcCarthy}
 
@@
 
-Figure~\ref{learningFromScratch} shows the full set of tasks and the learned DSL.
+Figure~\ref{learningFromScratch} shows the full set of tasks and the learned library.
 \begin{figure}  \newcommand{\helpSize}{0.25cm}
@@
     \\\midrule 
-    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
+    \rotatebox[origin=c]{90}{\normalsize \popp{Library}}&
     \begin{tabular}{ll}

### 2019-04-18 archive.tex
scientific laws
added caption and latest deep program learning figures

@@
 \begin{figure}
-  \includegraphics[width = 1.2\textwidth]{figures/DeepLogo_4_11}
-  \includegraphics[width = \textwidth]{figures/dpl/list}
-  \caption{}\label{exampleDSL}
+  \includegraphics[width = \textwidth]{figures/DeepLogo_4_15}
+  \includegraphics[width = \textwidth]{figures/DeepList_4_17}
+  \caption{Model solves tasks (left; top, task is to draw an image.
+bottom, task is input/output mapping) by writing programs (right column) where these programs invoke explicit declarative knowledge in the form of library routines (middle) jointly learned w/ the solutions.
+Library code can call each other (arrows), forming a layered network of domain-specific concepts.
+Fantasies (below library) are tasks sampled by randomly combining code from library.
+Fantasies are used to train a neural net to help search for programs.}\label{exampleDSL}
 \end{figure}
@@
       \begin{tabular}{l}
-        Center-of-mass:\\
-        $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
-        \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
-        \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
+        Kinetic energy:\\
+        $m$, $\vec{v}\mapsto \frac{1}{2}m|\vec{v}|^2$\\
+        \code{1, [0 2 -1]$\to$2.5}\\
+        \code{2, [2 1 3]$\to$14}\\
+        \blueCode{$f(m$, $\vec{v}) = $($f_4$ (* $m$ ($f_0$ $\vec{v}$)))}
+
+        %% Center-of-mass:\\
+        %% $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
+        %% \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
+        %% \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
       \end{tabular}\\\\
@@
         \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
-        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        
+        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        \\
+        \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
+        \hspace{0.2cm}\blueCode{($f_7$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
       \end{tabular}\\\\
       \begin{tabular}{l}
-        Integrating position over time:\\
-        $x_0$, $\dot{x}$, $\ddot{x}$, $t\mapsto \frac{1}{2}\ddot{x}t^2 + \dot{x}t + x_0$\\
-
+        Integrating angle over time:\\
+        $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
+        \code{1, 3, 0, 2$\to$7}\\
+        \code{-5, 0, 10, 1$\to$0}\\
+        \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
         \end{tabular}
@@
         \hspace{\helpSize}($f_6$: \emph{n-dimensional vector length})
+      \end{tabular}\\
+      %(lambda (lambda (lambda (map (lambda (*.
+(/.
+(*.
+(/.
+$0 (#(lambda (fold $0 0.
+(lambda (lambda (+.
+(*.
+$1 $1) $0))))) $1)) $3) (#(lambda (power (#(lambda (fold $0 0.
+(lambda (lambda (+.
+(*.
+$1 $1) $0))))) $0) (#(lambda (/.
+$0 (+.
+1.
+1.))) 1.))) $1)) $2)) $0)
+    %(lambda (lambda (lambda (map ($\lambda$ (x) (* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)
+    \begin{tabular}{l}
+      \greenCode{$f_7($a,b,v$)\,=\,$(map ($\lambda$ (x)}\\
+      \greenCode{\phantom{$f_7($a,b}(* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)}\\
+      \hspace{\helpSize}($f_7$: Inverse square law)
       \end{tabular}
     \end{tabular}
+
     &
@@
 and see theory learning as similar to inducing a DSL or library.
+
+Our model alternatingly searches for solutions to tasks and then improves its search strategy;
+within program induction, these principles were first elucidated by the Exploration-Compression (EC) algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
 

### 2019-04-18 archive.tex
better deep programming montage

@@
 \begin{figure}
-  \includegraphics[width = \textwidth]{figures/dpl/montage.pdf}
+  \includegraphics[width = \textwidth]{deepMontage.pdf}
   \caption{Top: Learned libraries diagrammed as networks.
 Arrows point from a learned concept to all other concepts that use it.
-Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (right).
+Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (middle), while total library size is more weakly associated with solving tasks (right).
 Full model (circles on scatterplots) learns deeper libraries than no-recognition-model ablation (crosses on scatterplots).

### 2019-04-18 deepMontage.tex
better deep programming montage

\documentclass{standalone}
\usepackage{tikz}
\usepackage{anyfontsize}
% all other packages and stuff you need for the picture
%
\begin{document}
\begin{tikzpicture}
  \node(logo)[anchor=east,align=center] at (0,0) {{\Huge \textbf{LOGO Graphics}}\\\includegraphics[scale=0.5]{figures/dpl/logo.pdf}};
  \node(tower)[anchor=west,align=center] at (logo.east) {{\Huge \textbf{Tower Building}}\\\includegraphics[scale=0.5]{figures/dpl/tower.pdf}};
  \node(list)[anchor=north,align=center] at (logo.south) {{\Huge \textbf{List Processing}}\\\includegraphics[scale=0.5]{figures/dpl/listGraph.pdf}};
  \node(text)[anchor=west,align=center] at (list.east) {{\Huge \textbf{Text Editing}}\\\includegraphics[scale=0.5]{figures/dpl/text.pdf}};
  \node(regex)[anchor=north,align=center] at (list.south) {{\Huge \textbf{Generative Text Patterns}}\\\includegraphics[scale=0.5]{figures/dpl/regex.pdf}};
  \node(scientific)[anchor=north,align=center] at (text.south) {{\Huge \textbf{Scientific Laws}}\\\includegraphics[scale=0.5]{figures/dpl/scientificLaws.pdf}};
  \node(McCarthy)[anchor=north,align=center] at (scientific.south) {{\Huge \textbf{Basic Recursive Functions}}\\\includegraphics[scale=0.5]{figures/dpl/McCarthy.pdf}};
  \node(rational)[anchor=west,align=center] at (tower.east) {{\Huge \textbf{Symbolic Regression}}\\\includegraphics[scale=0.5]{figures/dpl/rational.pdf}};

  \node(g1)[anchor=north] at (regex.south) {\includegraphics[scale=1.5]{figures/dpl/depthVersusAccuracy_MEAN.eps}};
  \node(g2)[anchor=west] at (g1.east) {\includegraphics[scale=1.5]{figures/dpl/depthVersusAccuracy_MAX.eps}};
  \node(g3)[anchor=west] at (g2.east) {\includegraphics[scale=1.5]{figures/dpl/depthVersusAccuracy_SIZE.eps}};
\end{tikzpicture}
\end{document}

### 2019-04-18 archive.tex
almost done except for the conclusion!

@@
 
-\subsection{Programs that manipulate sequences}\label{sequences}
+\subsection{Conditional programs that manipulate sequences}\label{sequences}
 We first apply \system to two classic benchmark domains: list
 processing and text editing. In both cases we solve tasks specified by
-a input/output examples, starting with a generic functional programming
-basis, including routines like \code{map}, \code{fold}, \code{cons}, \code{car}, \code{cdr}, etc.
+a conditional mapping (i.e., input/output examples), starting with a
+generic functional programming basis, including routines like
+\code{map}, \code{fold}, \code{cons}, \code{car}, \code{cdr}, etc.
 %% : \code{foldr}, \code{unfold}, \code{if}, \code{map},
@@
 system typically composes around 20 new library routines, and discovered multiple
-higher-order functions. Each round of memory consolidation built on
-components discovered in earlier sleep cycles --- for example the
+higher-order functions. Each round of abstraction built on
+concepts discovered in earlier sleep cycles --- for example the
 agent first learns the higher-order function \code{filter}, uses
@@
 
-How do \systemEnding's representations of these problems themselves change, as the model acquires expertise?
-Within cognitive science, a rich body of evidence suggests that human experts learn to `see' problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}.
-Novices notice a problem’s `surface features', the literal objects and details made obvious by the task description.
-Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures', the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.
-
-\systemEnding's neural recognition model can be seen as forming analogous `initial' problem representations, implicitly encoded within the activations of the neural network.
-Intuitively, these initial neural representations map problem features to a `gist' over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves.
-To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding's learning trajectory.
-Figure \ref{textExpertiseCluster} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm.
-We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories.
-Rather, as \system acquires domain expertise, the neural network adapts how it transforms each task's surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.
-
+\subsection{Generative \& Procedural Programs}
+%% We consider three domains where the agent must infer a program from an
+%% image (Figure~\ref{visualSpecs}).
+%% First we consider programs that make plans and take actions:  drawing pictures and building towers out of blocks (Sec.~\ref{logoSection}-\ref{towerSection}).
+
+\subsubsection{Programs for Smooth Trajectories}\label{regressionSection}
+Many procedures that people learn, like as motor routines, can be
+thought of as some kind of smooth trajectory. Modeling these
+procedures as programs entails jointly inferring their combinatorial
+structure and their continuous parameters.  We consider simple
+programs generating trajectories described by polynomials (up to
+degree 4) and rational functions, and equip our learner with a
+primitive for introducing a continuous parameter into the discrete
+structure of its programs, which an inner loop optimizes via gradient
+descent. We penalize the use of real-valued parameters using a
+BIC~\cite{Bishop:2006:PRM:1162264} likelihood model.  We use a convolutional neural
+network as a `visual' recognition model, which observes a graph of the
+trajectory over time, allowing the agent to `eyeball' the trajectory
+prior to writing code that describes it
+(Figure~\ref{symbolicRegressionExamples}).  In solving 100 such tasks,
+the model learns to find programs minimizing the number of continuous
+parameters --- this phenomenon arises from our Bayesian framing: both
+the generative model's bias toward shorter programs, and the likelihood
+model's BIC penalty.
 \begin{figure}
-  \includegraphics[width = \textwidth]{figures/text_tsne.png}
-  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of text tasks (neural net activations), after first wake/sleep cycle.
-Middle: at-a-glance organization after 20 wake/sleep cycles.
-Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution.
-More examples in Fig.~\ref{allTSNE}}\label{textExpertiseCluster}
+    \includegraphics[width = \textwidth]{figures/sr16_horizontal.png}
+  \caption{Sixteen smooth-trajectory tasks.
+Agent writes a program containing continuous real numbers that fits the points along the curve.
+Recognition model is a CNN that takes as input a graph of the trajectory over time (shown above).
+Task is `solved' if program matches points on curve.}\label{symbolicRegressionExamples}
   \end{figure}
-
-\subsection{Programs from visual input}
-We consider three domains where the agent must infer a program from an
-image (Figure~\ref{visualSpecs}).
-First we consider programs that make plans and take actions:  drawing pictures and building towers out of blocks (Sec.~\ref{logoSection}-\ref{towerSection}).
-
-\subsubsection{Programs that draw pictures}\label{logoSection}
+\subsubsection{Programs that generate images}\label{logoSection}
 
@@
 inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
-with drawing a corpus of 160 images while equipping it with control
+with drawing a corpus of 160 images (Figure~\ref{logoMaster}, top) while equipping it with control
 over a `pen', along with arithmetic operations on angles and
@@
 training data, like polygons, circles, and spirals
-(Figure~\ref{logoPrimitives}, left) -- without supervision the agent
+(Figure~\ref{logoPrimitives}, middle) -- without supervision the agent
 has learned the basic types of objects in its visual world. It
@@
 rotational symmetry, which it models by abstracting out a new higher-order
-function into its library (Figure~\ref{logoPrimitives}, right).
+function into its library.
 
 What does \system dream of?  Prior to learning random programs written using the library
-are simple and largely unstructured (Figure~\ref{logoDreams}, left).
+are simple and largely unstructured (Figure~\ref{logoMaster}, bottom left).
 After training the samples become richly structured
-(Figure~\ref{logoDreams}, right), compositionally recombining latent
+(Figure~\ref{logoMaster}, bottom right), compositionally recombining latent
 building blocks and motifs acquired from the training data. This
@@
 neural net gets richer and more highly varied training data.
-
-%% has primitives for moving the pen forward and rotating it (\code{move}),
-%% picking the pen up and then putting it down (\code{pen-up}),
-%% arithmetic operations on angles and distances,
-%% and constants ($2\pi$, 0 through 9, and $\infty$\footnote{To ensure that programs terminate, we set $\infty = 20$}).
-
-\begin{figure}\centering
-  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
-  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
-  \includegraphics[width = 0.3\textwidth]{figures/sr16.png}
-  \caption{Three domains where the agent infers a program from visual input.
-\textbf{Left}: 16 (out of 160) LOGO graphics tasks.
-Agent writes a program controlling a `pen' that draws the target picture.
-\textbf{Middle}: 9 (out of 112) tower building tasks.
-Agent writes a program controlling a `hand' that builds the target tower.
-\textbf{Right}: 16 (out of 200) symbolic regression tasks.
-Agent writes a program containing continuous real numbers that fits the points along the curve.}\label{visualSpecs}
-\end{figure}
 \begin{figure}
-\centering  \begin{tabular}{cc}
-    \toprule
-    Parametric drawing routines&Higher-order drawing routine\\\midrule
+  \centering
+  \begin{tabular}{cc}
+\toprule     \multicolumn{2}{c}{    Subset of 160 tasks}\\\midrule
+    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/horizontalLogoTasks.png}}\\\\
+
+    \midrule  Parametric drawing routines in library&Higher-order drawing routine in library\\\midrule
     \begin{tabular}{rl}
@@
     \begin{tabular}{c}
-      Rotational Symmetry:\\
+      Radial Symmetry:\\
           \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
@@
 
-    \\\bottomrule 
+    \\\bottomrule \\\\
+    \multicolumn{2}{c}{
+      \begin{tabular}{ccc}
+        Dreams before learning&Dreams after learning&\\
+        \begin{tabular}{c}
+          \includegraphics[width = 0.4\textwidth]{figures/initialDreams/montage.png}
+        \end{tabular}&
+        \begin{tabular}{c}
+          \includegraphics[width = 0.4\textwidth]{figures/finalDreams/cherry_random_montage.png}
+        \end{tabular}&
+        \rotatebox[origin=c]{90}{Cherry picked\hspace{2.5cm}Random}
       \end{tabular}
-  \caption{Example primitives learned by \system when trained on tasks in Figure~\ref{everyLogo}.
+    }
+  \end{tabular}
+  \caption{\textbf{Top}: 48 (out of 160) LOGO graphics tasks.
+Agent writes a program controlling a `pen' that draws the target picture.
+\textbf{Middle}: Example learned library routines.
 Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right).
 Each row of images on the left is the same code executed with different parameters.
-Each image on the right is the same code executed with different parameters and with a different subprogram provided as input.}\label{logoPrimitives}
-  \end{figure}
-\begin{figure}
-\centering  \begin{tabular}{c|cc}
-  \begin{tabular}{c}
-    \includegraphics[width = 0.45\textwidth]{figures/initialDreams/montage.png}
-    \end{tabular}&
-      \begin{tabular}{c}
-        \includegraphics[width = 0.45\textwidth]{figures/finalDreams/cherry_random_montage.png}
-        \end{tabular}&
-      \rotatebox[origin=c]{90}{Cherry picked\hspace{3cm}Random}
-    \end{tabular}
-  \caption{Sixteen dreams, or samples, from the DSL before (left) and after (right) training on tasks in Figure~\ref{everyLogo}.
+Each image on the right is the same code executed with different parameters and with a different subprogram provided as input.
+\textbf{Bottom}: Sixteen dreams, or sampled programs, from library before (left) and after (right) learning.
 Blue: where the agent started drawing.
-Pink: where the agent ended drawing.}\label{logoDreams}
+Pink: where the agent ended drawing.}\label{logoMaster}
   \end{figure}
@@
 tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
-(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
+(Figure~\ref{tower}, top).  Here the agent observes both an image of a
 tower and the locations of each of its blocks, and must write a
@@
 LOGO graphics,
-and learns parametric `options' for building blocks towers (Figure~\ref{tower}),
+and learns parametric `options' for building blocks towers (Figure~\ref{tower}, bottom left),
 inferring concepts like arches, staircases, bridges and brick walls.
@@
   \centering  \begin{tabular}{cc}
+    \toprule
+    \multicolumn{2}{c}{Subset of 112 Tasks}\\\midrule
+    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower12.png}}\\\midrule \\
     \toprule Example learned DSL components&Samples written using final library\\\midrule
@@
       Brickwall&\begin{tabular}{l}
-        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
+        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bricks.png}
       \end{tabular}\\
       Bridge&\begin{tabular}{l}
-        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
+        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bridge.png}
       \end{tabular}\\
       Staircase&\begin{tabular}{l}
-        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
+        \includegraphics[width = 6cm]{figures/tower/tower_dsl_staircase.png}
         \end{tabular}
@@
     \begin{tabular}{l}
-      \includegraphics[width = 5cm]{figures/tower/dreams.png}
+      \includegraphics[width = 6cm]{figures/tower/dreams.png}
       \end{tabular}
@@
     \end{tabular}
-  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{stolle2002learning},
-    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
+  \caption{\textbf{Top}: Twelve (out of 112) tower building tasks.
+Agent writes a program controlling a `hand' that builds the target tower.
+\textbf{Bottom left}: Three learned library routines.
+These components act like parametric options~\cite{stolle2002learning},
+    giving higher-level building blocks that the agent can use to plan. \textbf{Bottom Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
 \end{figure}
 
-\subsubsection{Symbolic Regression}\label{regressionSection}
-Here, the agent observes points along the curve of a function, and
-must write a program that fits those points.  We initially equip our
-learner with addition, multiplication, and division, and task it with
-solving 200 symbolic regression problems, each either a polynomial or
-rational function.  The recognition model is a convnet that observes
-an image of the target function's graph (Fig.~\ref{visualSpecs},
-rightmost) --- visually, different kinds of polynomials and rational
-functions produce different kinds of graphs, and so the convnet can
-look at a graph and predict what kind of function best explains it.  A
-key difficulty, however, is that these problems are best solved with
-programs containing real numbers.  Our solution to this difficulty is
-to enumerate programs with real-valued parameters, and then fit those
-parameters by automatically differentiating through the programs the
-system writes and use gradient descent to fit the parameters.  We
-define the likelihood model, $\probability[x|p]$, by assuming a
-Gaussian noise model for the input/output examples, and penalize the
-use of real-valued parameters using the
-BIC~\cite{Bishop:2006:PRM:1162264}.
-
-We learn a DSL containing 13 new functions, mainly templates for
-different pieces of polynomials or ratios of polynomials.  The model
-also learns to find programs minimizing the number of continuous
-parameters --- for example, learning to represent linear functions
-with \code{(* real (+ x real))}.  This phenomenon arises from our
-Bayesian framing: both the generative model's bias toward shorter
-programs, and the likelihood model's BIC penalty.
-
-\subsubsection{Generative Modeling of Text}
+\subsubsection{Probabilistic Generative Modeling of Text}
 Few-shot learning of simple generative concepts comes naturally to
@@
 
-\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
+\subsection{Analyzing \system across domains}\label{quantitative}
 To evaluate the relative importance of library learning and recognition
@@
 Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (middle), while total library size is more weakly associated with solving tasks (right).
-Full model (circles on scatterplots) learns deeper libraries than no-recognition-model ablation (crosses on scatterplots).
+Full model (stars on scatterplots) learns deeper libraries than no-recognition-model ablation (triangles on scatterplots).
 Each scatter plot point run with a different random seed.}\label{depthMontage}
+\end{figure}
+
+How do \systemEnding's representations of these tasks themselves change, as the model acquires expertise?
+Within cognitive science, a rich body of evidence suggests that human experts learn to `see' problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}.
+Novices notice a problem’s `surface features', the literal objects and details made obvious by the task description.
+Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures', the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.
+
+\systemEnding's neural recognition model can be seen as forming analogous `initial' problem representations, implicitly encoded within the activations of the neural network.
+Intuitively, these initial neural representations map problem features to a `gist' over the model’s current domain-specific concepts -- they guide and constrain how the model searches for problem solutions, but are not symbolically structured solutions themselves.
+To probe how these representations change as the model gains a richer conceptual vocabulary within each domain, we compare how they cluster problems within that domain over the course of \systemEnding's learning trajectory.
+Figure \ref{textExpertiseCluster} depicts TSNE visualizations of these neural problem representations at the first and last iteration of the algorithm.
+We color-code each problem using semantic, human-interpretable categories determined apriori by the task designers, but critically, the model itself never has access to these categories.
+Rather, as \system acquires domain expertise, the neural network adapts how it transforms each task's surface features, restructuring its initial problem representation space to draw together problems that share deeper, more abstract conceptual similarities.
+
+\begin{figure}
+  \includegraphics[width = \textwidth]{figures/textLogo_tsne.png}
+  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of tasks (neural net activations), after first wake/sleep cycle.
+Middle: at-a-glance organization after last wake/sleep cycle.
+Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution.
+We convert each program into a feature vector by counting the \# times each library routine is used, and then embed those feature vectors.
+More examples in Fig.~\ref{allTSNE}}\label{textExpertiseCluster}
 \end{figure}

### 2019-04-19 archive.tex
even more writing

@@
 program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
-our model solves a specific problem by synthesizing a program solving
+our model approaches a specific problem by searching for a program solving
 it. It gradually grows its explicit domain-specific knowledge for a
@@
 Our model iteratively creates new library routines that build on
-concepts learned earlier in its learning trajectory, growing a library with nested
+concepts acquired earlier in its learning trajectory, growing a library with nested
 hierarchies of code.  We think of this cumulative nesting of
@@
 more closely resembles modern functional languages like Haskell and OCaml.
-Recall (Sec.~\ref{sequences})
-that we initially provided our system with functional programming routines like
-\code{map} and \code{fold}.
 
@@
 and generally solve them in less time -- with both components
-combined, suggesting a synergy between library learning and
-recognition model training. Examining the libraries learned with and without
+combined, suggesting a synergy between these forms of declarative and procedural knowledge.
+Examining the libraries learned with and without
 recognition models reveals a difference in the typical \emph{depth}

### 2019-04-19 archive.tex
rough draft

@@
 SyGuS additionally comes with a
-different hand-engineered DSL \emph{for each text editing problem}.
+different hand-engineered libraries of primitives \emph{for each text editing problem}.
 %\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task.
 For these experiments, our system did not use this assistance.}
-Here  we learned a single DSL
+Here  we learned a single library of text-editing concepts
 that applied generically to
@@
 model's BIC penalty.
-\begin{figure}
+\begin{figure}[b]
     \includegraphics[width = \textwidth]{figures/sr16_horizontal.png}
@@
 training data, like polygons, circles, and spirals
-(Figure~\ref{logoPrimitives}, middle) -- without supervision the agent
+(Figure~\ref{logoMaster}, middle) -- without supervision the agent
 has learned the basic types of objects in its visual world. It
@@
 LOGO graphics,
-and learns parametric `options' for building blocks towers (Figure~\ref{tower}, bottom left),
-inferring concepts like arches, staircases, bridges and brick walls.
+and inside its learned library we find parametric `options' for building blocks towers (Figure~\ref{tower}, bottom left),
+including latent concepts like arches, staircases, bridges and brick walls.
 \begin{figure}
@@
     \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower12.png}}\\\midrule \\
-    \toprule Example learned DSL components&Samples written using final library\\\midrule
+    \toprule Example learned Library components&Samples written using final library\\\midrule
     \begin{tabular}{rl}
@@
 agent with inferring a probabilistic regular expression from a small
-number of strings, where these strings are drawn from CSV columns crawled from the web~\cite{LUKE}. Across wake/sleep cycles, our agent learns to learn
+number of strings, where these strings are drawn from CSV columns crawled from the web~\cite{mws}. Across wake/sleep cycles, our agent learns to learn
 regular expressions that describe the structure of a typically occurring text concepts,
@@
 
-We first task \system with inferring 57 equations describing physical laws,
+We first task \system with inferring 60 equations describing physical laws,
 like closed forms for ballistic motion, inverse square laws,
@@
 these primitives,
-and then recover a DSL that
+and then recover a representation that
 more closely resembles modern functional languages like Haskell and OCaml.
@@
 \code{unfold}.  In solving these 18 exercises, the algorithm assembles
-a DSL containing a modern repertoire of functional programming idioms
+a library containing a modern repertoire of functional programming idioms
 and subroutines, including \code{map}, \code{fold}, \code{unfold},
@@
 Teal: Full model.
-Orange: Dreaming only (no DSL learning).
+Orange: Dreaming only (no library learning).
 Purple: Consolidation only (no recognition model).
@@
 learned neural network that guides program
-search~\cite{spiral,balog2016deepcoder,devlin2017neural}, while recent
+search~\cite{DBLP:journals/corr/abs-1804-01118,balog2016deepcoder,devlin2017neural}, while recent
 symbolic AI research has developed frameworks for learning the DSL by
@@
 
-Our model alternatingly searches for solutions to tasks and then improves its search strategy;
-within program induction, these principles were first elucidated by the Exploration-Compression (EC) algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
-
-Our work  sits within the Exploration-Compression (EC) family of
-algorithms.  EC~\cite{Dechter:2013:BLV:2540128.2540316} algorithms alternate between searching, or
-`exploring', the space of programs, and then updating their search
-procedure by compressing solutions found during exploration. \system
-grew directly out of research on EC-style systems and both of our
-sleep phases can be interpreted as a kind of compression:
-consolidation aims to compactly refactor code, while the recognition
-model aims to encode a program in as few bits as possible, conditioned
-on a task. %%  Our previous work, EC$^2$~\cite{ecc}, introduced
-%% neurally-guided search into the EC framework and served as the
-%% starting point for \systemEnding.  Our work here extends EC$^2$ by (1)
-%% introducing a new refactoring compression algorithm, (2) enriching the
-%% neural network with a new bigram-over-trees parameterization and loss
-%% function that together learn to break symmetries, accelerating program
-%% search during waking, and (3) demonstrating how this family of
-%% approaches can be applied to planning and generative modeling
-%% problems. 
-Other offshoots of EC include our prior work, EC$^2$~\cite{ecc}, the neurosymbolic framework
-in~\cite{lazaro2019beyond} and the hierarchical Bayesian program
-learner in~\cite{DBLP:conf/icml/LiangJK10}. Closely aligned ideas go
-even further back~\cite{solomonoff1989system,schmidhuber2004optimal}.
+Our model draws on ideas first elucidated by the
+Exploration-Compression family of algorithms~\cite{Dechter:2013:BLV:2540128.2540316,ecc,lazaro2019beyond,DBLP:conf/icml/LiangJK10,solomonoff1989system,schmidhuber2004optimal,ozkural2011teraflop},
+which alternate
+between searching, or `exploring', a space of solutions, and then
+updating their search procedure by compressing solutions found during
+exploration.  Both of our sleep phases can be interpreted as a kind of
+compression: abstraction aims to compactly refactor code, while the
+recognition model aims to encode a program in as few bits as possible,
+conditioned on a task. What \system contributes here is the joint capture of
+declarative aspects of domain knowledge -- considered in prior EC works -- and
+the equally crucial procedural knowledge,
+both of which human experts seem to need.
+
+%% Our model alternatingly searches for solutions to tasks and then improves its search strategy;
+%% within program induction, these principles were first elucidated by the Exploration-Compression (EC) algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
+
+%% Our work  sits within the Exploration-Compression (EC) family of
+%% algorithms.%%  Our previous work, EC$^2$~\cite{ecc}, introduced
+%% %% neurally-guided search into the EC framework and served as the
+%% %% starting point for \systemEnding.  Our work here extends EC$^2$ by (1)
+%% %% introducing a new refactoring compression algorithm, (2) enriching the
+%% %% neural network with a new bigram-over-trees parameterization and loss
+%% %% function that together learn to break symmetries, accelerating program
+%% %% search during waking, and (3) demonstrating how this family of
+%% %% approaches can be applied to planning and generative modeling
+%% %% problems. 
+%% Other offshoots of EC include our prior work, EC$^2$~\cite{ecc}, the neurosymbolic framework
+%% in~\cite{lazaro2019beyond} and the hierarchical Bayesian program
+%% learner in~\cite{DBLP:conf/icml/LiangJK10}. Closely aligned ideas go
+%% even further back~\cite{solomonoff1989system,schmidhuber2004optimal}.
 
@@
 explicit declarative concepts and implicit procedural skill, both jointly learned.
+Looking forward we anticipate that making machines which can \emph{simultaneously}
+develop expertise for many different classes of problems, as humans do,
+could be enabled by metalearning a library, DSL, or `language-of-thought':
+learning a single generic programmatic substrate,
+which can then differentiate itself into domain-specific
+representations for many different classes of problems.
 

### 2019-04-21 archive.tex
improved physics results

@@
 additionally learns more abstract visual relationships, like
-rotational symmetry, which it models by abstracting out a new higher-order
+radial symmetry, which it models by abstracting out a new higher-order
 function into its library.
@@
 expressions for calculating centers of mass, etc.,
-drawing verbatim from AP and MCAT physics `cheat sheets'
+which we took from AP and MCAT physics `cheat sheets'
 along with equation guides from machine learning textbooks.
@@
   \centering
+  \setlength{\tabcolsep}{2pt} % default is 6pt
   \begin{tabular}{lll}\toprule
@@
         \code{2, [2 1 3]$\to$14}\\
-        \blueCode{$f(m$, $\vec{v}) = $($f_4$ (* $m$ ($f_0$ $\vec{v}$)))}
+        %#(lambda (lambda (*. (/. $0 (+. $1 $1))))) 1. $1 (#(lambda (fold $0 0. (lambda (lambda (+. $0 (*. $1 $1)))))) $0))))
+        \blueCode{$f(m$, $\vec{v}) = $($f_5$ 1 $m$ ($f_0$ $\vec{v}$))}
 
@@
         \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
-        \hspace{0.2cm}\blueCode{($f_7$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
+        \hspace{0.2cm}\blueCode{($f_{10}$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
       \end{tabular}\\\\
       \begin{tabular}{l}
-        Integrating angle over time:\\
-        $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
-        \code{1, 3, 0, 2$\to$7}\\
-        \code{-5, 0, 10, 1$\to$0}\\
-        \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
+        %% Integrating angle over time:\\
+        %% $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
+        %% \code{1, 3, 0, 2$\to$7}\\
+        %% \code{-5, 0, 10, 1$\to$0}\\
+        %% \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
+        Newton's Second Law:\\
+        $m$, $\{\vec{F}_i \}\mapsto \frac{1}{m}\sum_i\vec{F}_i = a$\\
+        \code{2, [[2 4 1]]$\to$[1 2 0.5]}\\
+        \code{3, [[0 0 2] [3 0 1]]$\to$[1 0 0.333]}\\
+        \blueCode{$f(m$, forces$) = $($f_4$ ($f_7$ m) ($f_3$ forces))}
         \end{tabular}
@@
         \code{4, [0 2 8 1 5 6]$\to $1}\\
-        \blueCode{$f($n$,$l$) = $($f_7$ l (+ 1 n))}
+        \blueCode{$f($n$,$l$) = $($f_8$ l (+ 1 n))}
       \end{tabular}\\\\
@@
         \greenCode{$f_0($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ (* x x) y)) v)}\\
-        \hspace{\helpSize}($f_0$: \emph{L-2 norm})
+        \hspace{\helpSize}($f_0$: \emph{L-2 norm}) % still here
       \end{tabular}\\
@@
         \greenCode{$f_1($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ x y)) v)}\\
-        \hspace{\helpSize}($f_1$: \emph{Sum vector components})
+        \hspace{\helpSize}($f_1$: \emph{Sum vector components}) % still here
       \end{tabular}\\
@@
         \greenCode{$f_2($v,u$)\,=\,$($f_1$ (zip v u ($\lambda$ (x y) (* x y))))}\\
-        \hspace{\helpSize}($f_2$: \emph{Dot product})
+        \hspace{\helpSize}($f_2$: \emph{Dot product}) % still here
       \end{tabular}\\ 
-      %(lambda (fold $0 (map (lambda (*. $0 0.)) (car $0)) (lambda (lambda (zip $0 $1 (lambda (lambda (+. $0 $1)))))))) 
       \begin{tabular}{l}
@@
         \phantom{\greenCode{$f_3($vs$)\,=\,$(}}\greenCode{($\lambda$ (v u) (zip v u ($\lambda$ (x y) (+ x y)))))}\\
-        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors})
+        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors}) % still here
       \end{tabular}\\
       \begin{tabular}{l}
-        \greenCode{$f_4($x$)\,=\,$(/ x (+ 1 1))}\\
-        \hspace{\helpSize}($f_4$: \emph{Divide by two})
+        \greenCode{$f_4($x,v$)\,=\,$(map ($\lambda$ (y) (* x y)) v)}\\
+        \hspace{\helpSize}($f_4$: \emph{Multiply vector by scalar})
+        \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_5($x,y,z$)\,=\,$(* (/ y (+ x x)) z)}
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_6($x,y$)\,=\,$(* ($f_5$ 1 x y) x)}\hspace{1cm}\greenCode{$f_7($x$)\,=\,$(/ 1 x)}
       \end{tabular}\\
       %      %#(lambda (lambda (lambda (lambda (+. $2 (*. $0 (+. (*. (#(lambda (/. $0 (+. 1. 1.))) $0) $1) $3))))))) $2 $3 $1 $0)))))
-      \begin{tabular}{l}
-        \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
-        \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
-      \end{tabular}\\
-      \begin{tabular}{l}
-        \greenCode{$f_6($v$)\,=\,$(power ($f_0$ v) ($f_4$ 1))}\\
-        \hspace{\helpSize}($f_6$: \emph{n-dimensional vector length})
+      %% \begin{tabular}{l}
+      %%   \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
+      %%   \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
+      %% \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_8($v$)\,=\,$(power ($f_0$ v) ($f_6$ 1 1))}\\
+        \hspace{\helpSize}($f_8$: \emph{vector length})
       \end{tabular}\\
@@
       \begin{tabular}{l}
-      \greenCode{$f_7($a,b,v$)\,=\,$(map ($\lambda$ (x)}\\
-      \greenCode{\phantom{$f_7($a,b}(* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)}\\
-      \hspace{\helpSize}($f_7$: Inverse square law)
+        \greenCode{$f_9($x,y,z$)\,=\,$(* (/ x ($f_0$ z)) y)}\\
+        \hspace{\helpSize}($f_9$: \emph{scalar form of inverse square law})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_{10}($x,y,v$)\,=\,$($f_4$ ($f_9$ y (/ x ($f_7$ v)) v) v)}\\
+        \hspace{\helpSize}($f_{10}$: \emph{vector form of inverse square law})
         \end{tabular}
@@
       \begin{tabular}{l}
-        \greenCode{$f_7($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
-        \hspace{\helpSize}($f_7$: \emph{index})
+        \greenCode{$f_7($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (if (f x) a (cons x a))))}\\
+          \hspace{\helpSize}($f_7$: \emph{filter})
+      \end{tabular}\\
+      \begin{tabular}{l}
+        \greenCode{$f_8($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
+        \hspace{\helpSize}($f_8$: \emph{index})
       \end{tabular}
@@
   \end{tabular}
+  \setlength{\tabcolsep}{6pt} % default is 6pt
   \caption{\textbf{Left:} Learning a language for physical laws starting with recursive list routines like \code{map} and \code{fold}.
@@
 
-\emph{(what we found)} Our goal was to build a model capturing
+Our goal was to build a model capturing
 humanlike aspects of learning to be an expert in a new domain of
@@
 
-\emph{(related work)} \system builds on several generations of research in AI, program synthesis,
+\system builds on several generations of research in AI, program synthesis,
 and cognitive science, with program induction being one of the oldest
@@
 
-\emph{(limitations - and how to go beyond them)} Our model's domain knowledge grows over a series of wake/sleep cycles,
+Our model's domain knowledge grows over a series of wake/sleep cycles,
 with the solutions to easier problems bootstrapping
@@
 
-\emph{(I feel the lifelong learning setting is less important of a point to make - we might remove this)} Over a lifetime humans develop expertise in many domains --
-unlike our model,
-which learned to solve many classes of problems,
-but only class one at a time.
-Scaling approaches like ours to this multidomain, lifelong setting
-could be enabled by a hierarchical Bayesian model of DSLs for many domains,
-capturing  shared conceptual structure between similar domains  (e.g., Parcheesi and chess)
-and separating knowledge between dissimilar domains (e.g., Parcheesi and cheese making).
+%% Over a lifetime humans develop expertise in many domains --
+%% unlike our model,
+%% which learned to solve many classes of problems,
+%% but only class one at a time.
+%% Scaling approaches like ours to this multidomain, lifelong setting
+%% could be enabled by a hierarchical Bayesian model of DSLs for many domains,
+%% capturing  shared conceptual structure between similar domains  (e.g., Parcheesi and chess)
+%% and separating knowledge between dissimilar domains (e.g., Parcheesi and cheese making).
 

### 2019-04-30 archive.tex
better tower image
Towers copy, citation

@@
 
-Inspired by the classic AI `copy task' --- where an agent must look at
+Inspired by the classic AI `copy demo' --- where an agent must look at
 an image of a tower made of toy blocks and re-create the
@@
     \multicolumn{2}{c}{Subset of 112 Tasks}\\\midrule
-    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower12.png}}\\\midrule \\
+    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower_montage_10.png}}\\\midrule \\
     \toprule Example learned Library components&Samples written using final library\\\midrule
@@
     \end{tabular}
-  \caption{\textbf{Top}: Twelve (out of 112) tower building tasks.
+  \caption{\textbf{Top}: Ten (out of 112) tower building tasks.
 Agent writes a program controlling a `hand' that builds the target tower.

### 2019-05-01 archive.tex
better dreams for towers

@@
 compared to novices,
-experts can quickly categorize and classify problems based on the ``deep structure''
+experts can, at a glance, categorize and classify problems based on the ``deep structure''
 of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
@@
 our model approaches a specific problem by searching for a program solving
-it. It gradually grows its explicit domain-specific knowledge for a
+it. It gradually grows its domain-specific knowledge for a
 class of problems by assembling a library of code containing concepts
-useful for the domain, and by training a neural network to quickly
+useful for the domain, while jointly training a neural network to quickly
 infer, for a specific problem, what code in its library is likely to
@@
 learns to sort sequences of numbers by invoking a library component 4
-layers deep, or draws the leftmost pairs of images in
+layers deep, or draws the topmost image in
 Figure~\ref{exampleDSL} using a depth-3 component.  For this reason we
@@
 
-\begin{figure}
-  \includegraphics[width = \textwidth]{figures/DeepLogo_4_15}
-  \includegraphics[width = \textwidth]{figures/DeepList_4_17}
+\begin{figure}\centering
+  \includegraphics[width = \textwidth]{figures/deepLogo_may1}
+  \includegraphics[width = \textwidth]{figures/deepList_may1}
   \caption{Model solves tasks (left; top, task is to draw an image.
@@
 Library code can call each other (arrows), forming a layered network of domain-specific concepts.
-Fantasies (below library) are tasks sampled by randomly combining code from library.
-Fantasies are used to train a neural net to help search for programs.}\label{exampleDSL}
+%% Fantasies (below library) are tasks sampled by randomly combining code from library.
+Fantasies are used to train a neural net to help search for programs.
+  }\label{exampleDSL}
 \end{figure}
@@
 finding the library maximizing $\int \probability[\mathcal{D},\theta]\probability[X|\mathcal{D},\theta]\;\mathrm{d}\theta$ (Sec.~\ref{overviewSection}).
-We replace this marginal with an AIC approximation~\cite{akaike1998information}
-and marginalize over refactorings of programs found during waking, minimizing the following expression,
+We %% replace this marginal with an AIC approximation~\cite{akaike1998information}
+%% and
+marginalize over refactorings of programs found during waking, seeking to minimize the following expression,
 which can be interpreted as a kind of compression:
+%% \begin{equation}
+%% \underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
+%% \probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
+%%   X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
+%% \label{AIC}
+%% \end{equation}
 \begin{equation}
-\underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
-\probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
-  X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
+\underbrace{  -\log \probability[\mathcal{D}]_{\substack{\phantom{p}\\\\\\\\\\}}}_{\text{Description length of library}}\underbrace{-\log \int \probability[\theta|\mathcal{D}]\prod_{x\in X}\sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]\;\mathrm{d}\theta}_{\text{Description length of tasks given library}}
+%% \underbrace{-\log \probability[\mathcal{D}] + \min_{\substack{\theta\\\\\\\\}}\Bigg(-\log
+%% \probability[\theta|\mathcal{D}] + \|\theta\|_0}_{\text{Description length of }(\mathcal{D},\theta)} +\sum_{x\in
+%%   X}\underbrace{-\log \sum_{\substack{p \text{ a refactoring of } p'\\ p'\text{ found during waking}}}\probability[x|p]\probability[p|\mathcal{D},\theta]}_{\text{Description length of programs for task }x}\Bigg)
 \label{AIC}
@@
 each with 15 input/output examples.  In solving these tasks, the
-system typically composes around 20 new library routines, and discovered multiple
-higher-order functions. Each round of abstraction built on
+system typically composes around 20 new library routines, and rediscovers
+higher-order functions such as \code{filter}. Each round of abstraction built on
 concepts discovered in earlier sleep cycles --- for example the
-agent first learns the higher-order function \code{filter}, uses
-\code{filter} to learn to take the maximum element of a list, then
-uses that routine to learn a new component for extracting the
+agent first learns \code{filter}, then uses
+it to learn to take the maximum element of a list, then
+uses that routine to learn a new library routine for extracting the
 $n^{\text{th}}$ largest element of a list, which it finally uses to
@@
 and algorithms that synthesize text editing programs ship in Microsoft
-Excel~\cite{gulwani2011automatin}.  This prior work uses
+Excel~\cite{gulwani2011automating}.  This prior work uses
 hand-engineered libraries of primitives and hand-engineered search strategies.  Here, we
-will show that we can jointly learn both these ingredients and surpass
+will show that we can jointly learn both these ingredients and perform comparably to
 the state-of-the-art domain-general program synthesizers on a standard
@@
 
-%% Because our enumerative search procedure cannot generate string %
-%% constants, we instead enumerate programs with string-valued
-%% parameters.  For example, to learn a program that prepends ``Dr.'', we
-%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
-%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
-%% define $\probability[x|p]$ by approximately marginalizing out the
-%% string parameters via a simple dynamic program.
-%% In Sec.~\ref{regressionSection}, we will use a similar trick to
-%% synthesize programs containing real numbers, but using gradient
-%% descent instead of dynamic programming.
-
 We trained our system on 128 automatically-generated text editing tasks, with 4 input/output examples each.
-We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
+We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition.
+Figure~\ref{textEditExamples} shows representative text-editing problems.
+Before any learning,
 \system solves 3.7\% of the problems within 10 minutes with an average search time of 235 seconds.
@@
 prior work.
+\begin{figure}[h]\centering
+  \begin{tabular}{ccc}
+    \fbox{\begin{tabular}{c}
+        \;\,Dream Coder$\to$(Coder)\\
+        Domain Expertise$\to$(Expertise)
+    \end{tabular}}&
+    \fbox{\begin{tabular}{c}
+        Temple Anna H$\to $Dr. TAH\\
+        Lara Gregori$\to $Dr. LG
+    \end{tabular}}&
+    \fbox{\begin{tabular}{c}
+      this test-split here.$\to$split here\\
+      another-text.test for you$\to$text
+    \end{tabular}}
+  \end{tabular}
+  \caption{Three example text editing tasks.
+Each task is specified by input/output pairs (notated above as input$\to$output), and the goal is to infer a program producing the input/output mapping.}\label{texEditProblems}
+\end{figure}
 
@@
       \begin{tabular}{ccc}
-        Dreams before learning&Dreams after learning&\\
-        \begin{tabular}{c}
-          \includegraphics[width = 0.4\textwidth]{figures/initialDreams/montage.png}
-        \end{tabular}&
-        \begin{tabular}{c}
-          \includegraphics[width = 0.4\textwidth]{figures/finalDreams/cherry_random_montage.png}
-        \end{tabular}&
-        \rotatebox[origin=c]{90}{Cherry picked\hspace{2.5cm}Random}
+        Dreams before learning&Dreams after learning&Dreams after learning (cherry picked)\\
+          \includegraphics[width = 0.3\textwidth]{figures/dreams/beforeLearning25}&
+          \includegraphics[width = 0.3\textwidth]{figures/dreams/final25}&
+          \includegraphics[width = 0.3\textwidth]{figures/dreams/cherry_picked/montage}
       \end{tabular}
@@
 Each image on the right is the same code executed with different parameters and with a different subprogram provided as input.
-\textbf{Bottom}: Sixteen dreams, or sampled programs, from library before (left) and after (right) learning.
+\textbf{Bottom}: 25 dreams, or sampled programs, from library before (left column) and after (middle \& right) learning.
 Blue: where the agent started drawing.
@@
     \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower_montage_10.png}}\\\midrule \\
-    \toprule Example learned Library components&Samples written using final library\\\midrule
+    \toprule Example learned library components&Dreams from learned library\\\midrule
     \begin{tabular}{rl}
@@
     \begin{tabular}{l}
-      \includegraphics[width = 6cm]{figures/tower/dreams.png}
+      \includegraphics[width = 8cm]{figures/tower/dreams/cherry_montage.png}
       \end{tabular}
@@
 It then uses this vocabulary to infer library routines for common
-physical relationships like the vector form inverse square laws
+physical relationships like the vector form of inverse square laws
 and integrating second derivatives over time,
@@
         Coulomb's law:\\
-        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
+        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
         \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
@@
 Orange: Dreaming only (no library learning).
-Purple: Consolidation only (no recognition model).
+Purple: Abstraction only (no recognition model).
 Search time plots show solid lines (time averaged over all tasks) and dotted lines (time averaged over solved tasks).
@@
 
-Our goal was to build a model capturing
-humanlike aspects of learning to be an expert in a new domain of
+Our goal was to build a computational model capturing
+aspects of learning to be an expert in a new domain of
 problems. Our model works through a wake/sleep algorithm that grows an increasingly deep
@@
 kinds of learning are crucial for building agents that, like humans,
-autonomously learn to navigate a new domain of problems.
+autonomously learn to solve a new domain of problems.
 
@@
 A main goal of this work is to combine and refine these techniques
-with the intention of building agents that, like humans, develop
+with the intention of building agents that develop
 domain expertise for new classes of problems.  Work in cognitive
@@
 This dynamic is related to, but distinct from, curriculum learning.
-In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
+In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks, but here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
 But humans select their tasks in much richer ways,

### 2019-05-02 archive.tex
deep program learning
title

@@
 
+\title{\systemEnding: A Wake/Sleep Algorithm for Developing Domain Expertise}
+\date{}
 \begin{document}
+\maketitle{}
 
@@
 \begin{figure}\centering
-  \includegraphics[width = \textwidth]{figures/deepLogo_may1}
-  \includegraphics[width = \textwidth]{figures/deepList_may1}
+  \includegraphics[width = \textwidth]{figures/deepLogo_may2.pdf}\vspace{2.5cm}
+  \includegraphics[width = \textwidth]{figures/deepList_may2.pdf}
   \caption{Model solves tasks (left; top, task is to draw an image.
@@
 with the number of solved testing tasks ($r=0.71$, $p < 10^{-8}$).
-In contrast, the correlation between the \emph{number} of learned subroutines and number of solved tasks is weaker and not statistically significant ($r=0.20$, $p > 0.05$),
-suggesting that the recognition model bootstraps
-``better'' libraries, where ``better'' is at least partly explained by the depth
+In contrast, the correlation between the \emph{number} of learned subroutines and number of solved tasks is weaker ($r=0.20$, $p > 0.05$), although, within each domain, the recognition model still leads to bigger libraries.
+These results suggest that the recognition model bootstraps
+``better'' libraries, where ``better'' is largely explained by the depth
 of the learned representation. Figure~\ref{depthMontage} shows typical `deep' libraries learned by
@@
 Arrows point from a learned concept to all other concepts that use it.
-Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (middle), while total library size is more weakly associated with solving tasks (right).
-Full model (stars on scatterplots) learns deeper libraries than no-recognition-model ablation (triangles on scatterplots).
+Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (middle).
+Total library size is more also associated with solving tasks (right), but across domains the association is weaker.
+Full model (stars on scatterplots) learns bigger and deeper libraries than no-recognition-model ablation (triangles on scatterplots).
 Each scatter plot point run with a different random seed.}\label{depthMontage}

### 2019-05-06 archive.tex
stream of consciousness based on Josh conversation

@@
 
-\title{\systemEnding: A Wake/Sleep Algorithm for Developing Domain Expertise}
+\title{\systemEnding: Building Deep Domain Expertise with Wake/Sleep Bayesian Program Learning}
 \date{}
@@
 \maketitle{}
+
+\vspace{-1cm}
 
@@
 like a field linguist would.
-Here, we contribute a computational model that takes a small step toward the goal of building machines that
+Here, we present a computational model that takes a step toward the goal of building machines that
 grow into domain experts,
-looking at much simpler kinds of problems that humans can learn to solve and indeed become experts in.
-
+looking at a number of problem domains that humans can learn and indeed become experts in.
+
+we draw inspiration from cognitive science.
+this inspiration says that there are two things.
+we do both these things.
+hooray for us.
 Our model is structured
@@
 program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
-our model approaches a specific problem by searching for a program solving
-it. It gradually grows its domain-specific knowledge for a
+our model approaches a specific problem by searching for a program
+solving it. It gradually grows its domain-specific knowledge for a
 class of problems by assembling a library of code containing concepts
-useful for the domain, while jointly training a neural network to quickly
-infer, for a specific problem, what code in its library is likely to
-solve it. We think of these two learned components as analogous to the
-explicit concepts and implicit procedural skill that human experts
-develop.  The model is an instance of what in the machine learning
-community is called a wake-sleep algorithm, where for us, `waking'
-corresponds to solving problems, while a pair of `sleep' cycles
-correspond to improving the library of explicit concepts and training
-the neural net to improve implicit procedural knowledge of how to
-write code.
+useful for the domain, while jointly training a neural network to
+quickly infer, for a specific problem, what code in its library is
+likely to solve it. We think of these two learned components as
+analogous to the explicit concepts and implicit procedural skill that
+human experts develop.  The model, called \systemEnding, is an
+instance of what in the machine learning community is called a
+wake-sleep algorithm, where for us, `waking' corresponds to solving
+problems, while a pair of `sleep' cycles correspond to improving the
+library of explicit concepts and training the neural net to improve
+implicit procedural knowledge.
 
@@
 
-Our model, called \systemEnding, works through a novel kind
+\system works through a novel kind
 of wake/sleep or `dream' learning~\cite{hinton1995wake}, iterating
@@
 approximately) solving for
-
-\begin{array}{lr}
-  \begin{array}{l}  p_x = \argmax_p \probability[x|p]\probability[p|\mathcal{D}^*,\theta^*]\text{, for each task $x\in X$}\end{array}
-    &\text{\emph{Wake}}\\\\
-\begin{array}{rl}  
-\mathcal{D}^* = \argmax_\mathcal{D}&\int\probability\left[\mathcal{D},\phantom{^*}\theta \right]\prod_{x\in X}\sum_{\substack{p\text{ found during waking}}} \probability[x|p]\probability[p|\mathcal{D},\phantom{^*}\theta]\;\text{d}\theta\\
-\theta^* = \argmax_\theta&\phantom{\int}\probability\left[\mathcal{D}^*,\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D}^*,\theta]
- \end{array}
-&\text{\emph{Abstraction (sleep)}}\\\\
-\begin{array}{l} Q(p|x)\approx \probability[p|x,\mathcal{D}^*,\theta^*]\end{array}
-  &\text{\emph{Dreaming (sleep)}}
-\end{array}
-
+\begin{align}
+  p_x &= \argmax_{\substack{p\\Q(p|x)\text{ is large}}} \probability[x|p]\probability[p|\mathcal{D},\theta]\text{, for each task $x\in X$}&&\text{\emph{Wake}}\\
+  \mathcal{D},\theta &= \argmax_{\mathcal{D},\theta}\probability\left[\mathcal{D},\theta \right]\prod_{x\in X}\sum_{p\text{ found during waking}} \probability[x|p]\probability[p|\mathcal{D},\theta]&&\text{\emph{Sleep: Abstraction}}\\
+  Q(p|x)&\approx \probability[p|x,\mathcal{D},\theta]&&\text{\emph{Sleep: Dreaming}}  
+\end{align}
 \noindent which serves to maximize a lower bound on the posterior over $(\mathcal{D},\theta)$ given $X$ (Appendix~\ref{probabilisticAppendix}).
@@
       \begin{tabular}{ccc}
-        Dreams before learning&Dreams after learning&Dreams after learning (cherry picked)\\
-          \includegraphics[width = 0.3\textwidth]{figures/dreams/beforeLearning25}&
-          \includegraphics[width = 0.3\textwidth]{figures/dreams/final25}&
-          \includegraphics[width = 0.3\textwidth]{figures/dreams/cherry_picked/montage}
+        Dreams before learning&Dreams after learning (random)&Dreams after learning (selected)\\
+          \fbox{\includegraphics[width = 0.3\textwidth]{figures/dreams/beforeLearning25}}&
+          \fbox{\includegraphics[width = 0.3\textwidth]{figures/dreams/final25}}&
+          \fbox{\includegraphics[width = 0.3\textwidth]{figures/dreams/cherry_picked/montage}}
       \end{tabular}
@@
 Agent rediscovers \code{map}, \code{fold}, \code{unfold}, etc.
-For full tasks and learned libraries see Appendix~\ref{appendixMcCarthy}}\label{librariesToLanguages}
+For full tasks and learned libraries see Appendix~\ref{physicsAppendix,appendixMcCarthy}}\label{librariesToLanguages}
   \end{figure}
@@
 updating their search procedure by compressing solutions found during
-exploration.  Both of our sleep phases can be interpreted as a kind of
+exploration.
+EC : not only one difference. we actually incorporate several improvements.
+for example our compression is very sophisticated.
+it is a fancy kind of re- factoring.
+exploring is analogous to waking except that we also are better here because we can break symmetries and have a deep network.
+Both of our sleep phases can be interpreted as a kind of
 compression: abstraction aims to compactly refactor code, while the
@@
 both of which human experts seem to need.
+also it turns out that computers need these two too!!!
 
@@
     \State $\theta\gets $uniform distribution
-    \State $\mathcal{F}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize beams to be empty}
+    \State $\mathcal{B}_x\gets\varnothing $, $\forall x\in X$\Comment{Initialize beams to be empty}
     \While{true}\Comment{Loop over epochs}
@@
     \State shuffle $\gets$ shuffle with first $B$ elements removed
-    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\mathcal{F}_x\cup \left\{p\;|\;p\in \text{enumerate}(\probability[\cdot |\mathcal{D},\theta],T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
-    \State Train $Q(\cdot |\cdot )$ to minimize $\mathcal{L}^{\text{MAP}}$ across all $\left\{\mathcal{F}_x \right\}_{x\in X}$\Comment{Dream Sleep}
-    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\mathcal{F}_x\cup \left\{p\;|\;p\in \text{enumerate}(Q(\cdot |x),T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
-    \State $\forall x\in \text{batch}$: $\mathcal{F}_x\gets\text{ top }F\text{ elements of }\mathcal{F}_x\text{ as measured by }\probability[\cdot |x,\mathcal{D},\theta]$\Comment{Keep top $F$ programs}
-    \State $\mathcal{D},\theta,\left\{\mathcal{F}_x \right\}_{x\in X}\gets $\textsc{Consolidate}$(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\}_{x\in X})$\Comment{Consolidation Sleep}
-    \State \textbf{yield} $(\mathcal{D},\theta)$, $Q$, $\left\{\mathcal{F}_x \right\}_{x\in X}$ \Comment{Yield the updated DSL, recognition model, and solutions found to tasks}
+    \State $\forall x\in \text{batch}$: $\mathcal{B}_x\gets\mathcal{B}_x\cup \left\{p\;|\;p\in \text{enumerate}(\probability[\cdot |\mathcal{D},\theta],T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
+    \State Train $Q(\cdot |\cdot )$ to minimize $\mathcal{L}^{\text{MAP}}$ across all $\left\{\mathcal{B}_x \right\}_{x\in X}$\Comment{Dream Sleep}
+    \State $\forall x\in \text{batch}$: $\mathcal{B}_x\gets\mathcal{B}_x\cup \left\{p\;|\;p\in \text{enumerate}(Q(\cdot |x),T)\text{ if }\probability[x|p] > 0 \right\}$\Comment{Wake}
+    \State $\forall x\in \text{batch}$: $\mathcal{B}_x\gets\text{ top }F\text{ elements of }\mathcal{B}_x\text{ as measured by }\probability[\cdot |x,\mathcal{D},\theta]$\Comment{Keep top $F$ programs}
+    \State $\mathcal{D},\theta,\left\{\mathcal{B}_x \right\}_{x\in X}\gets $\textsc{Consolidate}$(\mathcal{D},\theta,\left\{\mathcal{B}_x \right\}_{x\in X})$\Comment{Consolidation Sleep}
+    \State \textbf{yield} $(\mathcal{D},\theta)$, $Q$, $\left\{\mathcal{B}_x \right\}_{x\in X}$ \Comment{Yield the updated DSL, recognition model, and solutions found to tasks}
     \EndWhile
@@
   %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
-  \caption{DSL Induction Algorithm}
+  \caption{Library Induction Algorithm}
   \label{grammarInductionAlgorithm}
@@
     %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
-    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
+    \State \textbf{Output:} Library $\mathcal{D}$, weight vector $\theta$
     \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{B}_x\}$
@@
     \State $\forall p\in \bigcup_{x}\mathcal{B}_x: $ $v_p\gets I\beta(p)$ \Comment{Construct a version space for each program}
-    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(v_p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
+    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{B}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(v_p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood under $(\mathcal{D}',\theta)$}
     \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
     \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*(\mathcal{D})) - \|\theta^*(\mathcal{D})\|_0$ \Comment{objective function}
-    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{B}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
-    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
-    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
-    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
-    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
-    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(v_p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new DSL}
+    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{B}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new components}
+    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new libraries}
+    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new library}
+    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to library led to a better score}
+    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better library. Update library.}
+    \State $\forall x\;:\;\mathcal{B}_x\gets\left\{\text{\textsc{refactor}}(v_p|\mathcal{D})\;:\; p\in \mathcal{B}_x\right\}$\Comment{Refactor beams in terms of new library}
     \EndWhile
@@
 Important for our purposes, it is also efficient to refactor the
-members of a version space's extension in terms of a new DSL.
+members of a version space's extension in terms of a new library.
 We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
@@
   \end{align*}
-where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
+where $\text{size}(e|\mathcal{D})$ for program $e$ and library $\mathcal{D}$
 is the size of the syntax tree of $e$,
@@
 
-We now formally prove that $I\beta$ exhaustively enumerates the space of possible refactorings.
+We now formally prove that $I\beta_n$ exhaustively enumerates the space of possible refactorings.
 Our approach is to first prove that $S_k$ exhaustively enumerates
@@
 
-From these results the consistency and completeness of $I\beta$ follows:
+From these results the consistency and completeness of $I\beta_n$ follows:
 \begin{theorem}
@@
 
+Having introduced the version-space machinery, we now are in a position to define
+how these version spaces aggregate into a single data structure, one for each program, tracking every equivalence revealed by
+$I\beta_n$.
+Observe that every time we calculate $I\beta_n(p)$, we obtain a version space containing expressions semantically equivalent to program $p$.
+In line with prior work in program analysis~\cite{tate2009equality},
+we track these equivalences,
+and finally return a single structure for each program compiling all of these equivalences.
+Concretely,
+Algorithm~\ref{grammarInductionAlgorithm} calculates (line 5), for each program $p$,
+a version space $I\beta(p)$ defined as
+\begin{equation*}
+  I\beta(p) = I\beta_n(p)\uplus\begin{cases}
+  \text{if $p = (f\;x)$: }I\beta(f)\;I\beta(x)\\
+  \text{if $p = \lambda b$: }\lambda I\beta(b)\\
+  \text{if $p$ is an index or primitive: }\varnothing 
+  \end{cases}
+\end{equation*}
+where $n$, the amount of refactoring, is a hyper parameter.
+We set $n$ to 3 for all experiments with the exception of learning recursive functions from basic Lisp primitives (Appendix~\ref{appendixMcCarthy}),
+where more involved refactoring was required; for that experiment we set $n = 4$.
+
 %% First observe that
@@
 
-\subsubsection{Computational complexity of DSL learning}
-
-How long does each update to the DSL in
+\subsubsection{Computational complexity of library learning}
+
+How long does each update to the library in
 Algorithm~\ref{grammarInductionAlgorithm} take?  Constructing the
@@
 spaces, so is $O(V)$.  For each proposal we have to refactor every
-program (line 6), so this means we spend $O(V^2) = O(P^22^n)$ per DSL
+program (line 6), so this means we spend $O(V^2) = O(P^22^n)$ per library
 update.  In practice this quadratic dependence on $P$ (the number of
@@
 For each version space $v$ we calculate a \emph{beam}, which is a
-function from a DSL $\mathcal{D}$ to a shortest program in
+function from a library $\mathcal{D}$ to a shortest program in
 $\denotation{v}$ using primitives in $\mathcal{D}$.  Our strategy will
@@
 the set `proposals' is defined in line 10 of Algorithm~\ref{grammarInductionAlgorithm},
-and $\mathcal{D}$ is the current DSL:
+and $\mathcal{D}$ is the current library:
 \begin{align*}
@@
 After performing this beam search,
-we take only the top $I$ proposals as measured by $-\sum_x\min_{p\in \mathcal{F}_x}\text{beam}_{v_p}(\mathcal{D}')$.
+we take only the top $I$ proposals as measured by $-\sum_x\min_{p\in \mathcal{B}_x}\text{beam}_{v_p}(\mathcal{D}')$.
 We set $I = 300$ in all of our experiments,
@@
 Next,
-we describe our MAP estimator for the continuous parameters ($\theta$) of the DSL.
-
-\subsubsection{Estimating the continuous parameters $\theta$ of a DSL}\label{mapAppendix}
-We use an EM algorithm to estimate the continuous parameters of the DSL, i.e. $\theta$.
+we describe our MAP estimator for the continuous parameters ($\theta$) of the generative model.
+
+\subsubsection{Estimating the continuous parameters $\theta$ of the generative model}\label{mapAppendix}
+We use an EM algorithm to estimate the continuous parameters of the generative model, i.e. $\theta$.
 Suppressing dependencies on $\mathcal{D}$, the EM updates are
@@
 \label{maximizeStep}  \theta& = \argmax_\theta \log P(\theta) + \sum_x \expect_{q_x}\left[\log \probability\left[p|\theta \right] \right]\\
-  q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{F}_x \right]
+  q_x(p)&\propto \probability[x|p]\probability[p|\theta]\indicator\left[p\in \mathcal{B}_x \right]
 \end{align}
@@
 This lower bound is tight whenever all
-of the types of the expressions in the DSL are not polymorphic, in which case our DSL is equivalent to a PCFG
+of the types of the expressions in the library are not polymorphic, in which case our library is equivalent to a PCFG
 and this estimator is equivalent to the inside/outside algorithm.
-Polymorphism introduces context-sensitivity to the DSL,
+Polymorphism introduces context-sensitivity to the library,
 and exactly maximizing the likelihood with respect to $\theta$
@@
   \mathcal{L}^{\text{MAP}} &= \mathcal{L}_{\text{Replay}}^{\text{MAP}} + \mathcal{L}_{\text{Fantasy}}^{\text{MAP}}\\
-  \mathcal{L}_{\text{Replay}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
-    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
-\mathcal{L}_{\text{Replay}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
+  \mathcal{L}_{\text{Replay}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{B}_x}
+    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{B}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
+\mathcal{L}_{\text{Replay}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{B}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
   \mathcal{L}_{\text{Fantasy}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
@@
 But, evaluating $\mathcal{L}_{\text{Fantasy}}$ involves drawing programs from
-the current DSL, running them to get their outputs,
+the current library, running them to get their outputs,
 and then training $Q$ to regress from the input/outputs to the program.
@@
 The $\mathcal{L}_{\text{Fantasy}}^{\text{MAP}}$ objective involves
-finding the MAP program solving a task drawn from the DSL.  To make
+finding the MAP program solving a task drawn from the library.  To make
 this tractable, rather than \emph{sample} programs as training data
@@
 $Q$ predicts a $(|\mathcal{D}|+2)\times (|\mathcal{D}|+1)\times A$-dimensional tensor,
-where $A$ is the maximum arity\footnote{The arity of a function is the number of arguments that it takes as input.} of any primitive in the DSL.
+where $A$ is the maximum arity\footnote{The arity of a function is the number of arguments that it takes as input.} of any primitive in the library.
 Slightly abusing notation, we write this tensor as $Q_{ijk}(x)$,
@@
      \State \textbf{function} recognitionSample$(Q, x, \mathcal{D}, \tau)$:
-     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, type $\tau$
+     \State {\bfseries Input:} recognition model $Q$, task $x$, library $\mathcal{D}$, type $\tau$
      \State \textbf{Output:} a program whose type unifies with $\tau$
@@
      \State \textbf{function} recognitionSample$'(Q, x, \text{parent}, \text{argumentIndex}, \mathcal{D}, \mathcal{E}, \tau)$:
-     \State {\bfseries Input:} recognition model $Q$, task $x$, DSL $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
+     \State {\bfseries Input:} recognition model $Q$, task $x$, library $\mathcal{D}$, parent $\in \mathcal{D}\cup\left\{\text{start},\text{var} \right\}$, argumentIndex $\in \mathbb{N}$, environment $\mathcal{E}$, type $\tau$
      \State \textbf{Output:} a program whose type unifies with $\tau$
@@
 Consider two recognition models, $Q_1$ and $Q_2$,
-and two programs in frontier $\mathcal{F}_x$,
+and two programs in beam $\mathcal{B}_x$,
 $p_1=\code{(+ (+ 1 1) 1)}$ and $p_2=\code{(+ 1 (+ 1 1))}$,
@@
 To be clear,
-our recognition model does not learn to break \emph{every} possible symmetry in every possible DSL.
+our recognition model does not learn to break \emph{every} possible symmetry in every possible library or DSL.
 But in practice we found that a bigrams combined with $\mathcal{L}^{\text{MAP}}$
@@
 \subsection{Neural Problem Representations}
-
+Figure~\ref{allTSNE} shows how \system organizes and represents its tasks over its learning trajectory for several domains.
+We visualize the inferred program representations (left columns) by
+calculating a feature vector for each task $x$, written $\phi(x)$,
+whose $i^{\text{th}}$ component is the expected number of times that the $i^\text{th}$  library component was used when solving the task:
+\[
+\phi(x)_i = \expect_{\probability[p |\mathcal{D},\theta,x]}\left[ \text{number of times }i\in \mathcal{D} \text{ occurs in }p\right]
+\]
+and then embed those feature vectors in a two dimensional space using tSNE~\cite{maaten2008visualizing}.
+We visualize the recognition model's task representations by
+flattening the neural output $Q_{ijk}(x)$ (Appendix~\ref{recognitionAppendix}) to
+obtain a feature vector for each task, which we visualize also using tSNE both after the first wake/sleep cycle (left column) and after the final wake/sleep cycle (middle column).
 \begin{figure}
   \includegraphics[width = \textwidth]{figures/all_tsne.png}
-  \caption{}\label{allTSNE}
+  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of tasks (neural net activations), after first wake/sleep cycle.
+Middle: at-a-glance organization after last wake/sleep cycle.
+Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution.
+We convert each program into a feature vector by counting the \# times each library routine is used, and then embed those feature vectors.}\label{allTSNE}
 \end{figure}
@@
 
-\subsection{Learning from Scratch: Tasks and Library}\label{appendixMcCarthy}
+\subsection{Learning a language for physical laws}\label{physicsAppendix}
+
+We took equations from MCAT as well as AP Physics B \& C cheat sheets (\url{http://mcat.prep101.com/wp-content/uploads/ES_MCATPhysics.pdf}; \url{https://secure-media.collegeboard.org/digitalServices/pdf/ap/ap-physics-1-equations-table.pdf}; \url{https://secure-media.collegeboard.org/digitalServices/pdf/ap/physics-c-tables-and-equations-list.pdf}),
+choosing representative laws,
+but excluding expressions with trigonometric functions, multivariate differentials, and integrals.
+These equations included  free fall velocity ($\sqrt{2gh}$),
+velocity under ballistic motion ($v^2 = v_0^2 + 2a(x - x_0)$),
+velocity magnitude ($v = \sqrt{v_x^2 + v_y^2}$),
+angular acceleration ($a_r = v^2/R$),
+mass energy equivalence ($e = mc^2$),
+expressions for center of mass ($\sum_i m_ix_i/\sum_im_i$ and $(m_1\vec{x_1} + m_2\vec{x_2})/(m_1 + m_2)$),
+density ($\pho = m/v$), pressure ($F/A$), power ($P = I^2R$ and $P = V^2/R$),
+RMS voltage ($V/\sqrt{2}$), energy stored in a capacitor ($U = 1/2CV^2$, $1/2QV$, and $1/2Q^2/C$),
+optical power ($P = 1/f$),
+
+Each equation was converted to a symbolic regression problem by sampling 10 random input/output examples,
+where the inputs correspond to free variables in the target expression.
+Physical constants were expressed in Planck units, which generally drops certain special coefficients from
+these expressions (e.g., the gravitational constant $G = 1$).
+We gave the system the following primitives: \code{0}, \code{1}, $\pi$, \code{power}, \code{+}, \code{*}, \code{/}, \code{-}, \code{map}, \code{fold}, \code{zip}, \code{cons}, \code{car}, \code{cdr}, and \code{nil}.
+
+At each wake cycle, rather than sample a random minibatch of tasks,
+we gave the agent all tasks it had not solved in the past two wake cycles.
+This had the effect of the agent solving the easiest tasks first,
+and then reallocating its compute time to a smaller set of harder, unsolved tasks.
+
+\subsection{Learning a language of recursive functions}\label{appendixMcCarthy}
 
@@
   \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.
-Complete set of tasks and learned DSL shown.
-Learned DSL components are numbered in the order that they are learned, \emph{i.e.}, the agent first learns fold, then unfold, then uses fold to define map, etc.}\label{learningFromScratch}
+Complete set of tasks and learned library shown.
+Learned library routines are numbered in the order that they are learned, \emph{i.e.}, the agent first learns fold, then unfold, then uses fold to define map, etc.}\label{learningFromScratch}
   \end{figure}
@@
 
-\subsection{Learned DSLs}
-Here we present representative DSLs learned by our model. DSL primitives
-discovered by the algorithm are prefixed with \lstinline!#!.
-Variables are prefixed with \lstinline!$!, and we adopt De Bruijn indices to
-model bound variables~\cite{pierce}.
-
-\lstset{
-  basicstyle=\footnotesize,
-  escapechar=@,
-  breaklines=true,
-  breakatwhitespace=true,
-  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
-}
-
-\subsubsection{List processing}
-\begin{lstlisting}
-#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range ($0 (+ 1))))))
-#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0))))))
-#(+ 1 (+ 1 1))
-#(@$\lambda$@ (@$\lambda$@ (fold $1 (cons $0 empty) (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
-#(+ 1 #(+ 1 (+ 1 1)))
-#(@$\lambda$@ (map (@$\lambda$@ (if ($1 $0) (+ $0 1) 0))))
-#(@$\lambda$@ (cdr (cdr $0)))
-#(@$\lambda$@ (@$\lambda$@ (fold (#(@$\lambda$@ (cdr (cdr $0))) (#(@$\lambda$@ (cdr (cdr $0))) $1)) $0 (@$\lambda$@ (@$\lambda$@ (cdr (#(@$\lambda$@ (@$\lambda$@ (fold $1 (cons $0 empty) (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 (car $0))))))))
-#(@$\lambda$@ (car (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $0 (@$\lambda$@ (empty?
+%% \subsection{Learned DSLs}
+%% Here we present representative DSLs learned by our model. DSL primitives
+%% discovered by the algorithm are prefixed with \lstinline!#!.
+%% Variables are prefixed with \lstinline!$!, and we adopt De Bruijn indices to
+%% model bound variables~\cite{pierce}.
+
+%% \lstset{
+%%   basicstyle=\footnotesize,
+%%   escapechar=@,
+%%   breaklines=true,
+%%   breakatwhitespace=true,
+%%   postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
+%% }
+
+%% \subsubsection{List processing}
+%% \begin{lstlisting}
+%% #(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range ($0 (+ 1))))))
+%% #(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0))))))
+%% #(+ 1 (+ 1 1))
+%% #(@$\lambda$@ (@$\lambda$@ (fold $1 (cons $0 empty) (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+%% #(+ 1 #(+ 1 (+ 1 1)))
+%% #(@$\lambda$@ (map (@$\lambda$@ (if ($1 $0) (+ $0 1) 0))))
+%% #(@$\lambda$@ (cdr (cdr $0)))
+%% #(@$\lambda$@ (@$\lambda$@ (fold (#(@$\lambda$@ (cdr (cdr $0))) (#(@$\lambda$@ (cdr (cdr $0))) $1)) $0 (@$\lambda$@ (@$\lambda$@ (cdr (#(@$\lambda$@ (@$\lambda$@ (fold $1 (cons $0 empty) (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 (car $0))))))))
+%% #(@$\lambda$@ (car (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $0 (@$\lambda$@ (empty?
 (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (gt?
 $0 $1))))))))
-#(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (eq? $0 $1))))))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (is-prime (+ $1 (mod $0 #(+ 1 #(+ 1 (+ 1 1))))))))))
-#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
-#(@$\lambda$@ (map (@$\lambda$@ (mod $0 $1))))
-#(@$\lambda$@ (map (@$\lambda$@ (eq? (mod $0 $1) 0))))
-#(@$\lambda$@ (gt? (mod $0 #(+ 1 (+ 1 1))) 0))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (car (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $0 (@$\lambda$@ (empty?
+%% #(@$\lambda$@ (@$\lambda$@ (length (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (eq? $0 $1))))))
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (is-prime (+ $1 (mod $0 #(+ 1 #(+ 1 (+ 1 1))))))))))
+%% #(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+%% #(@$\lambda$@ (map (@$\lambda$@ (mod $0 $1))))
+%% #(@$\lambda$@ (map (@$\lambda$@ (eq? (mod $0 $1) 0))))
+%% #(@$\lambda$@ (gt? (mod $0 #(+ 1 (+ 1 1))) 0))
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (car (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $0 (@$\lambda$@ (empty?
 (#(@$\lambda$@ (@$\lambda$@ (fold $1 empty (@$\lambda$@ (@$\lambda$@ (if ($2 $1) (cons $1 $0) $0)))))) $1 (@$\lambda$@ (gt?
@@
 $1 $0))))))))))
-\end{lstlisting}
-
-\subsubsection{Text editing}
-\begin{lstlisting}
-#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
-#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1))))
-#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0))))
-#(+ 1)
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq? $2 $1) 0 $0))))))))
-#(@$\lambda$@ (@$\lambda$@ (cons (car $0) (#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1)))) (cons '.' empty) (car $1)))))
-#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (if (char-eq? $2 $0) $1 $0)))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons LPAREN $0) (cons RPAREN empty)))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $0 (length (cdr (cdr $1))))))
-#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+%% \end{lstlisting}
+
+%% \subsubsection{Text editing}
+%% \begin{lstlisting}
+%% #(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0))))))
+%% #(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1))))
+%% #(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0))))
+%% #(+ 1)
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq? $2 $1) 0 $0))))))))
+%% #(@$\lambda$@ (@$\lambda$@ (cons (car $0) (#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1)))) (cons '.' empty) (car $1)))))
+%% #(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (if (char-eq? $2 $0) $1 $0)))))
+%% #(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons LPAREN $0) (cons RPAREN empty)))
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $0 (length (cdr (cdr $1))))))
+%% #(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
 $2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0)))))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $0 (length (cdr (cdr $1)))))) (cdr (cdr $0))))
-#(#(+ 1) 1)
-#(@$\lambda$@ (@$\lambda$@ (cons (car $1) (cons $0 empty))))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+%% #(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $0 (length (cdr (cdr $1)))))) (cdr (cdr $0))))
+%% #(#(+ 1) 1)
+%% #(@$\lambda$@ (@$\lambda$@ (cons (car $1) (cons $0 empty))))
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
 $2 $1) 0 $0)))))))) (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
 $2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0))))))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) (cons LPAREN $0) (cons RPAREN empty))) $1) $0) $0)))
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (cons $0 $2)))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 STRING))
-#(#(+ 1) #(#(+ 1) 1))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+%% #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (cons $0 $2)))))
+%% #(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 STRING))
+%% #(#(+ 1) #(#(+ 1) 1))
+%% #(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
 $2 $1) 0 $0)))))))) (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) STRING $0) SPACE))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (cons $0 $2))))) (cons $0 $1))))
-#(#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) STRING)
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (cons (car $0) (#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1)))) (cons '.' empty) (car $1))))) (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (cons $0 $2))))) (cons $0 $1))))
+%% #(#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) STRING)
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (cons (car $0) (#(@$\lambda$@ (@$\lambda$@ (cons '.' (cons $0 $1)))) (cons '.' empty) (car $1))))) (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
 $2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0))))))) $1 $0) $1)))
-#(#(+ 1) #(#(+ 1) #(#(+ 1) 1)))
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
+%% #(#(+ 1) #(#(+ 1) #(#(+ 1) 1)))
+%% #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $1 (#(@$\lambda$@ (@$\lambda$@ (cdr (fold (#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $1 (fold (cdr $1) 0 (@$\lambda$@ (@$\lambda$@ (#(+ 1) (if (char-eq?
 $2 $1) 0 $0)))))))) $1 $0) $1 (@$\lambda$@ (@$\lambda$@ (cdr $0))))))) (#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (fold $1 $0 (@$\lambda$@ (@$\lambda$@ (cons $1 $0)))))) $0 STRING)) (#(@$\lambda$@ (@$\lambda$@ (map (@$\lambda$@ (index $0 $2)) (range $0)))) $2 #(#(+ 1) #(#(+ 1) #(#(+ 1) 1))))) $0)))))
-\end{lstlisting}
-
-\subsubsection{Graphics}
-\begin{lstlisting}
-  #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0)))))))
-#(logo_DIVA logo_UA 4)
-#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY)
-#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0)))
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_FWRT (logo_MULL $2 $1) $4 $0)))))))
-#(@$\lambda$@ (logo_forLoop 7 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 7 logo_epsA $2 $0)))))
-#(@$\lambda$@ (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_SUBA logo_UA logo_epsA) (logo_MULL logo_epsL $0)))
-#(@$\lambda$@ (logo_forLoop logo_IFTY (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 logo_epsA (logo_FWRT logo_epsL logo_epsA $0))))))
-#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 4 #(logo_DIVA logo_UA 4))
-#(logo_DIVA logo_UA)
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_GETSET $2 (logo_FWRT logo_ZL $4 $0))))))))
-#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_DIVA logo_epsA 2) logo_epsL)
-#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) logo_epsA logo_epsL)
-#(logo_forLoop 3 (@$\lambda$@ (@$\lambda$@ (#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_DIVA logo_epsA 2) logo_epsL) (logo_FWRT logo_ZL #(logo_DIVA logo_UA 4) $0)))))
-#(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_GETSET $2 (logo_FWRT logo_ZL $4 $0)))))))) (#(logo_DIVA logo_UA) $0) $0))
-#(@$\lambda$@ (#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0))) (#(@$\lambda$@ (logo_forLoop logo_IFTY (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 logo_epsA (logo_FWRT logo_epsL logo_epsA $0)))))) logo_epsL $0)))
-#(@$\lambda$@ (#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0))) (logo_FWRT logo_UL #(logo_DIVA logo_UA 4) $0)))
-#(@$\lambda$@ (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) logo_epsA (logo_MULL logo_epsL $0)))
-#(logo_FWRT logo_UL logo_UA)
-#(@$\lambda$@ (#(@$\lambda$@ (logo_forLoop 7 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 7 logo_epsA $2 $0))))) (logo_MULL logo_epsL $0)))
-\end{lstlisting}
-
-\subsubsection{Towers}
-\begin{lstlisting}
-  #(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0))))))
-#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0))))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 4 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 8 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (left 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))))))))
-#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0)))))))))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 8 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (left 2 (3x1 $0))))))))
-#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0))))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (1x3 $0)))))
-#(tower_loopM 4 (@$\lambda$@ (@$\lambda$@ (right 6 (3x1 $0)))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0)))))) $0)))))
-#(@$\lambda$@ (right 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))))
-#(@$\lambda$@ (right 2 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))
-#(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0)))))) $0))))) $0 (right 2 (#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 4 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))))) $0 $1)))))
-#(tower_loopM 5 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (right 2 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))) $0))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (tower_embed (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 3 6 3 $0)) $0)))))
-#(@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (tower_embed (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 4 5 $4 $0)) $0))))))
-#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 4 5)
-#(@$\lambda$@ (@$\lambda$@ (right 6 (tower_embed $0 $1))))
-#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (right 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))) $0)))))
-\end{lstlisting}
-\subsubsection{Symbolic regression}
-\begin{lstlisting}
-  #(@$\lambda$@ (+. $0 REAL))
-#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL)))
-#(/. REAL)
-#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0)))
-#(@$\lambda$@ (/. (#(/. REAL) $0) $0))
-#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+.
+%% \end{lstlisting}
+
+%% \subsubsection{Graphics}
+%% \begin{lstlisting}
+%%   #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0)))))))
+%% #(logo_DIVA logo_UA 4)
+%% #(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY)
+%% #(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0)))
+%% #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_FWRT (logo_MULL $2 $1) $4 $0)))))))
+%% #(@$\lambda$@ (logo_forLoop 7 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 7 logo_epsA $2 $0)))))
+%% #(@$\lambda$@ (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_SUBA logo_UA logo_epsA) (logo_MULL logo_epsL $0)))
+%% #(@$\lambda$@ (logo_forLoop logo_IFTY (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 logo_epsA (logo_FWRT logo_epsL logo_epsA $0))))))
+%% #(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 4 #(logo_DIVA logo_UA 4))
+%% #(logo_DIVA logo_UA)
+%% #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_GETSET $2 (logo_FWRT logo_ZL $4 $0))))))))
+%% #(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_DIVA logo_epsA 2) logo_epsL)
+%% #(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) logo_epsA logo_epsL)
+%% #(logo_forLoop 3 (@$\lambda$@ (@$\lambda$@ (#(#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) (logo_DIVA logo_epsA 2) logo_epsL) (logo_FWRT logo_ZL #(logo_DIVA logo_UA 4) $0)))))
+%% #(@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $1 (@$\lambda$@ (@$\lambda$@ (logo_GETSET $2 (logo_FWRT logo_ZL $4 $0)))))))) (#(logo_DIVA logo_UA) $0) $0))
+%% #(@$\lambda$@ (#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0))) (#(@$\lambda$@ (logo_forLoop logo_IFTY (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 logo_epsA (logo_FWRT logo_epsL logo_epsA $0)))))) logo_epsL $0)))
+%% #(@$\lambda$@ (#(logo_PT (@$\lambda$@ (logo_FWRT logo_UL logo_ZA $0))) (logo_FWRT logo_UL #(logo_DIVA logo_UA 4) $0)))
+%% #(@$\lambda$@ (#(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) logo_IFTY) logo_epsA (logo_MULL logo_epsL $0)))
+%% #(logo_FWRT logo_UL logo_UA)
+%% #(@$\lambda$@ (#(@$\lambda$@ (logo_forLoop 7 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (logo_forLoop $2 (@$\lambda$@ (@$\lambda$@ (logo_FWRT $2 $3 $0))))))) 7 logo_epsA $2 $0))))) (logo_MULL logo_epsL $0)))
+%% \end{lstlisting}
+
+%% \subsubsection{Towers}
+%% \begin{lstlisting}
+%%   #(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0))))))
+%% #(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0))))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 4 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 8 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (left 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))))))))
+%% #(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0)))))))))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 8 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (left 2 (3x1 $0))))))))
+%% #(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0))))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (1x3 $0)))))
+%% #(tower_loopM 4 (@$\lambda$@ (@$\lambda$@ (right 6 (3x1 $0)))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0)))))) $0)))))
+%% #(@$\lambda$@ (right 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))))
+%% #(@$\lambda$@ (right 2 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))
+%% #(@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (1x3 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (1x3 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) (right 4 $0)))))) $0))))) $0 (right 2 (#(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left 4 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))))) $0 $1)))))
+%% #(tower_loopM 5 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (right 2 (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0))) $0))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (tower_embed (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 3 6 3 $0)) $0)))))
+%% #(@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (tower_embed (@$\lambda$@ (#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 4 5 $4 $0)) $0))))))
+%% #(#(@$\lambda$@ (@$\lambda$@ (@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (left $3 (3x1 (right 3 (3x1 (left $4 $0))))))))))) 4 5)
+%% #(@$\lambda$@ (@$\lambda$@ (right 6 (tower_embed $0 $1))))
+%% #(@$\lambda$@ (tower_loopM $0 (@$\lambda$@ (@$\lambda$@ (#(@$\lambda$@ (right 4 (#(@$\lambda$@ (1x3 (1x3 (right 4 (1x3 (1x3 $0)))))) (#(@$\lambda$@ (1x3 (left 4 (1x3 (right 2 (3x1 $0)))))) $0)))) $0)))))
+%% \end{lstlisting}
+%% \subsubsection{Symbolic regression}
+%% \begin{lstlisting}
+%%   #(@$\lambda$@ (+. $0 REAL))
+%% #(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL)))
+%% #(/. REAL)
+%% #(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0)))
+%% #(@$\lambda$@ (/. (#(/. REAL) $0) $0))
+%% #(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+.
 $0 REAL)) (*.
@@
 $0 REAL)) $0) $0))) $0))))
-#(@$\lambda$@ (#(/. REAL) (#(@$\lambda$@ (+. $0 REAL)) $0)))
-#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. (#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0))) $0) $0)))
-\end{lstlisting}
+%% #(@$\lambda$@ (#(/. REAL) (#(@$\lambda$@ (+. $0 REAL)) $0)))
+%% #(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. (#(@$\lambda$@ (#(@$\lambda$@ (#(@$\lambda$@ (+. $0 REAL)) (*. $0 REAL))) (*. (#(@$\lambda$@ (+. $0 REAL)) $0) $0))) $0) $0)))
+%% \end{lstlisting}
 
@@
 
-\noindent\textbf{\system Hyperparameters}
+\noindent\textbf{Hyperparameters}
 Due to the high computational cost we
@@
     Domain&Timeout&CPUs&Batch size&$\lambda$ (\ref{appendixCompression})&$\alpha$ (\ref{mapAppendix})&Max beam size (\ref{systemPseudocode})\\\midrule
+    Symbolic regression&2m & 40&10 & 1&30&5\\
     Lists&7m&64&10&1.5&30&5\\
@@
     Graphics&1h&96&50&1.5&30&5\\
-    Symbolic regression&2m & 40&10 & 1&30&5\\
     Towers& 1h& 64&50 & 1.5&30&5\\
     Regexes&30m & 64&40 & 1.5&30&5\\
+    Physical laws&4h&64&See~\ref{physicsAppendix}&1 & 30&5\\
+    Recursive functions&4h&64&See~\ref{appendixMcCarthy}&1 & 30&5\\
     \bottomrule     
